### 2024-07-17 install_packages.Rmd
add renv info after installing packages

---
title: "packages"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file installs packages.

There is no need to do this after the first installation, as it should all be 
controlled by renv().

This file is just for Sven's benefit, so that he can see how he installed them.

# install the libraries that we will be using #

## install ##

As I first created a public repo on Github (sven-panis/Tutorial_Event_History_Analysis) which I then cloned to a new R project (RStudo: File->New Project->Version Control->Git), I need to install and initialize renv myself.

```{r}
install.packages("renv")
renv::init()
```

I then closed the project, reopened it, and added the following two lines to .Rprofile (so that git will not ignore the renv.lock file, and renv::snapshot() will recored everything):
options(renv.settings.vcs.ignore.local = "FALSE") 
renv::settings$snapshot.type("all")

Next, general packages for data wrangling and visualisation used in various scripts.

```{r}
install.packages(c("tidyverse", "RColorBrewer", "patchwork", "ggridges"))
```

Next, packages for model building in brms used mainly in models.Rmd and effects.Rmd
files.

```{r}
install.packages("remotes")
remotes::install_github("stan-dev/cmdstanr")
library(cmdstanr)
install_cmdstan(cores=11) 
check_cmdstan_toolchain() 
## in the above, set cores correctly for your machine and without a folder location
## specified it just uses the default folder location. The default is fine for most
## setups though. To set the folder, include: dir = "/path/.cmdstan"

## you may also need to configure the toolchain, if you use a mac like I do:
## https://github.com/stan-dev/rstan/wiki/Configuring-C---Toolchain-for-Mac
## https://mac.thecoatlessprofessor.com/macrtools/

install.packages("devtools")
devtools::install_github("jmgirard/standist")
# 
install.packages(c("brms", "tidybayes", "bayesplot", "future"))
```

To build regression models in other packages such as lme4:

```{r}
#install.packages("lme4", "nlme")
```

Next, packages for papaja and manuscript writing

```{r install-pkg}
# install.packages(c("devtools", "remotes"))
# 
devtools::install_github("crsh/papaja")
devtools::install_github("crsh/citr")
# 
install.packages("tinytex")
tinytex::install_tinytex()
# 
# # and now some useful addins for writing .Rmd from R Studio word count
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# 
# # remedy for shortcuts in markdown writing
remotes::install_github("ThinkR-open/remedy")
# 
# # grammar checking - cool...
devtools::install_github("ropenscilabs/gramr")
```

Finally, use renv() to take a snapshot of installed packages and update the lock.file

```{r snapshot-renv}
# take a snapshot and update the lock.file
renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

### 2024-07-18 manuscript/manuscript.Rmd
add manuscript and bibtex files

---
title             : "A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data"
shorttitle        : "Tutorial on hazard modeling"

author: 
  - name          : "Sven Panis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "ETH GLC, room G16.2, Gloriastrasse 37/39, 8006 Zürich"
    email         : "sven.panis@hest.ethz.ch"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Richard Ramsey"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "ETH Zürich"

authornote: |
  Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).

  <!-- Enter author note here. -->

abstract: |
  Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology. To move beyond mean performance measures, various distributional analyses have been proposed. Here we focus on one particular distributional analysis known as discrete-time event history analysis, a.k.a. hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis. Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4. We discuss how to manage inter-individual differences, implications for experimental design, and various options when analysing time-to-event data using discrete-time survival analysis.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "response times, event history analysis, Bayesian regression models"
wordcount         : "X"

bibliography      : ["r-references.bib", "ownreferences.bib"]

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
editor_options: 
  chunk_output_type: console
header-includes:
  - \raggedbottom
---

```{r setup, include = FALSE}
pkg <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")

lapply(pkg, library, character.only = TRUE)

r_refs(file = "r-references.bib")
my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r plot-settings}
## theme settings for ggplot
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18), 
          title = element_text(size = 18),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

```{r global-chunk-settings}
## set the figure options
knitr::opts_chunk$set(fig.pos='H', out.extra = '', out.height = "67%",
                      fig.align = "center") # initial version
```

# Introduction

In experimental psychology, it is standard practice to analyse response times (RTs) using analysis-of-variance. However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events. Such information is useful not only for interpretation, but also for cognitive psychophysiology and model selection [@panisAnalyzingResponseTimes2020].

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->

# Results

We used `r my_r_citation$r` for creating.

`r my_r_citation$pkgs` <!-- This creates the footnote -->

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2024-07-18 manuscript/manuscript.tex
add manuscript and bibtex files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{response times, event history analysis, Bayesian regression models\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data},
  pdfauthor={Sven Panis1 \& Richard Ramsey1},
  pdflang={en-EN},
  pdfkeywords={response times, event history analysis, Bayesian regression models},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data}
\author{Sven Panis\textsuperscript{1} \& Richard Ramsey\textsuperscript{1}}
\date{}

\shorttitle{Tutorial on hazard modeling}

\authornote{

Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).

The authors made the following contributions. Sven Panis: Writing - Original Draft Preparation, Writing - Review \& Editing; Richard Ramsey: Conceptualization, Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Sven Panis, ETH GLC, room G16.2, Gloriastrasse 37/39, 8006 Zürich. E-mail: \href{mailto:sven.panis@hest.ethz.ch}{\nolinkurl{sven.panis@hest.ethz.ch}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} ETH Zürich}

\abstract{%
Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology. To move beyond mean performance measures, various distributional analyses have been proposed. Here we focus one particular distributional analysis known as discrete-time event history analysis, a.k.a. hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis. Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4. We discuss how to manage inter-individual differences, implications for experimental design, and various options when analysing time-to-event data using discrete-time survival analysis.
}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

In experimental psychology, it is standard practice to analyse response times (RTs) using analysis-of-variance. However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events. Such information is useful not only for interpretation, but also for cognitive psychophysiology and model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).

\section{Methods}\label{methods}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\subsection{Participants}\label{participants}

\subsection{Material}\label{material}

\subsection{Procedure}\label{procedure}

\subsection{Data analysis}\label{data-analysis}

\section{Results}\label{results}

We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{citr} (Version 0.3.2; Aust, 2019), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for creating.

\section{Discussion}\label{discussion}

\newpage

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-R-citr}
Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://github.com/crsh/citr}

\bibitem[\citeproctext]{ref-R-papaja}
Aust, F., \& Barth, M. (2023). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}

\bibitem[\citeproctext]{ref-R-tinylabels}
Barth, M. (2023). \emph{{tinylabels}: Lightweight variable labels}. Retrieved from \url{https://cran.r-project.org/package=tinylabels}

\bibitem[\citeproctext]{ref-R-lubridate}
Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. Retrieved from \url{https://www.jstatsoft.org/v40/i03/}

\bibitem[\citeproctext]{ref-R-tibble}
Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}

\bibitem[\citeproctext]{ref-R-RColorBrewer}
Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}

\bibitem[\citeproctext]{ref-panisAnalyzingResponseTimes2020}
Panis, S., Schmidt, F., Wolkersdorfer, M. P., \& Schmidt, T. (2020). Analyzing {Response Times} and {Other Types} of {Time-to-Event Data Using Event History Analysis}: {A Tool} for {Mental Chronometry} and {Cognitive Psychophysiology}. \emph{I-Perception}, \emph{11}(6), 2041669520978673. \url{https://doi.org/10.1177/2041669520978673}

\bibitem[\citeproctext]{ref-R-patchwork}
Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://CRAN.R-project.org/package=patchwork}

\bibitem[\citeproctext]{ref-R-base}
R Core Team. (2024). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-R-ggplot2}
Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}

\bibitem[\citeproctext]{ref-R-forcats}
Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}

\bibitem[\citeproctext]{ref-R-stringr}
Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}

\bibitem[\citeproctext]{ref-R-tidyverse}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., \ldots{} Yutani, H. (2019). Welcome to the {tidyverse}. \emph{Journal of Open Source Software}, \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\bibitem[\citeproctext]{ref-R-dplyr}
Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}

\bibitem[\citeproctext]{ref-R-purrr}
Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}

\bibitem[\citeproctext]{ref-R-readr}
Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}

\bibitem[\citeproctext]{ref-R-tidyr}
Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}

\end{CSLReferences}

\end{document}

### 2024-07-19 manuscript/manuscript.Rmd
update Introduction

@@
     role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
+      - "Conceptualization"
       - "Writing - Original Draft Preparation"
@@
 However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
-Such information is useful not only for interpretation, but also for cognitive psychophysiology and model selection [@panisAnalyzingResponseTimes2020].
+Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+
+# Tutorial 1: Calculating descriptive statistics using a life table
 

### 2024-07-22 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
Update readme

---
title: "Tutorial_1"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.

# Load the libraries that we will be using.

```{r load-pkg}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

# Read in the raw data from Panis and Schmidt (2016).

```{r read-raw-data}
raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T)
head(raw)
tail(raw)
glimpse(raw)
```

# Do some wrangling to get a simpler data set.

```{r wrangle-raw-data}
# remove practice trials
data <- raw %>% filter(practice==0) # 9240 trials: 6 participants x 70 blocks x 22 trials-per-block

# remove unnecessary columns
data <- data %>% select(-c(ss,practice,rd,form,task,primeduration,
                                             blankduration,maskduration,targetduration,
                                             void,target_dir,corresp,resperr))
# remove trials with no response info
sel = which(data$resp==0)
data <- data[-c(sel),] # 9189 rows left

# keep only the no-mask trials
data <- data %>% filter(mask_type == 0) %>%
                 select(-mask_type) # 2737 rows left, and 8 variables

summary(data)
# vp = Versuchsperson or subject
# bl = block number
# tr = trial number within block
# prime_type: no prime (0), consistent prime (1), inconsistent prime (2)
# resp = response left (1) or right (2)
# respac = accuracy
# rt = response time
# TrialNr = trial number within experiment
```

```{r save-data-file}
# save wrangled data file
write_csv(data, "Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
```

# Read in simpler data file, change variable names, and create factors.

```{r read-wrangled-data}
data_wr <- read_csv("Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
head(data_wr)

# change variable names (to pid, trial, condition, rt, and acc)
colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
# pid to indicate column with person IDs
# trial must be a unique trial number per subject
# condition must be a factor indicating the experimental levels of the single independent variable (1,2,...)
# rt to indicate column with response times
# acc to indicate column with accuracies

# create factor for condition (prime type)
data_wr <- data_wr %>% mutate(condition = condition + 1,
                              condition = factor(condition, levels = c(1,2,3), labels = c("neutral","congruent","incongruent")))
summary(data_wr)
```

# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).

Load the functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 205) as above (line 83).

```{r load-functions}
# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
censor <- function(df, timeout, bin_width){
  if(!(timeout %% bin_width == 0)){
    return("The censoring time must be a multiple of the bin width!")
  }
  if(timeout < 0 | bin_width < 0){
    return("Both timeout and bin_width must be larger than 0!")
  }
  df %>% mutate(right_censored = 0,
                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
                drt = ceiling(rtc/bin_width), # drt = discrete response time
                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
}

# function ptb creates a person-trial-bin oriented data set
ptb <- function(df){
  df %>% uncount(weights = drt) %>% # create drt rows per trial
         group_by(trial) %>% 
         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
         ungroup()
}

# function setup_lt sets up a life table for each level of condition (1 independent variable)
setup_lt <- function(ptb){
  ptb %>% mutate(event = str_c("event", event)) %>%
          group_by(condition,period) %>% 
          count(event) %>% 
          ungroup() %>% 
          pivot_wider(names_from = event,
                      values_from = n) %>% 
          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
                 event1 = ifelse(is.na(event1),0,event1),
                 risk_set = event0 + event1) %>% # define the risk set
          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
          group_by(condition) %>%
          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
          ungroup() 
}

# function calc_ca calculates the conditional accuracies 
calc_ca <- function(df){
  df %>% filter(right_censored==0) %>%
         group_by(condition,drt,cens_time,bin_width) %>%
         summarize(ca = mean(acc) %>% round(digits = 2),
                   n = n(),
                   .groups = 'drop') %>%
         ungroup() %>%
         mutate(period = drt,
                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
         select(-drt)
}

# function join_lt_ca joins the conditional accuracies to the life tables
join_lt_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}

# function extract_median is used to extract the S(t).50 quantiles (estimated median RTs)
extract_median <- function(df){
  above_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival > .5) %>% 
      slice(n()) # take last row
  below_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival < .5) %>% 
      slice(1) # take first row
  # pull period above
  period_above <- pull(above_pct50, period)
  # pull survivor function values
  survival_above <- pull(above_pct50, survival)
  survival_below <- pull(below_pct50, survival)
  # estimate median by interpolation
  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
}

# function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects
  library(patchwork)
  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
  median_period <- extract_median(df)
  n_conditions <- nlevels(df$condition)
  data_median <- c()
  for(i in 1:n_conditions){
    data_median <- append(data_median, c(median_period[i], median_period[i]))
  }

  data_medians <- tibble(period= data_median,
                         survival = rep(c(.5, 0),n_conditions),
                         condition = rep(1:n_conditions, each=2))
# plot the hazard functions
p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=hazard)) +
  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits = c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,haz_yaxis)) +
  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top")
# plot the survivor functions
p2 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
  # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("neutral","congruent","incongruent"))),
            data = data_medians, 
            linetype = 3, show.legend = F) +

  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="S(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())
# plot the conditional accuracy functions
p3 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p1/p2/p3
}
```

Apply the functions

```{r apply-functions}
# create a dataframe with two columns (pid and data); the latter contains a list of nested data frames
data_nested <- data_wr %>% group_nest(pid)
# apply the functions using map() and map2()
data_final <- data_nested %>% 
                     mutate(censored  = map(data, censor, 600, 40)) %>%   # user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
```

# Extract info, save plot for each subject, create person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.

```{r extract-info}
# view original data (column 2) for subject 1:
pluck(data_final, 2, 1) # pluck() ignores the round(digits=2) function
print(pluck(data_final, 2, 1), n=20)

# view person-trial-bin data set (column 4) for subject 1:
pluck(data_final,4,1)
print(pluck(data_final,4,1),n=30)

# view lifetable (column 7) for each condition for subject 1
pluck(data_final, 7, 1) %>% filter(condition=="neutral")
pluck(data_final, 7, 1) %>% filter(condition=="congruent")
pluck(data_final, 7, 1) %>% filter(condition=="incongruent")

# view plot (column 8) for subject 1. 
pluck(data_final, 8, 1)
```

```{r save-plots}

```

```{r create-file-hazard-modeling}

```

```{r save-file-hazard-modeling}

```

# Optional: Create single plot for data aggregated across subjects

```{r aggregated-data-plot}

```

### 2024-07-22 manuscript/manuscript.Rmd
update manuscript folder

@@
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and various options when analysing time-to-event data using discrete-time survival analysis.
+Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+We discuss how to manage inter-individual differences, implications for experimental design, and select among various options when analysing time-to-event data using discrete-time survival analysis.
 
@@
 
-In experimental psychology, it is standard practice to analyse response times (RTs) using analysis-of-variance.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations using analysis-of-variance.
+However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+In this tutorial we focus on a distributional method for analyzing time-to-event data that is known as discrete-time event history analysis (EHA), a.k.a.
+survival, hazard, duration, failure-time, and transition analysis [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
+Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+
+To apply EHA, one must be able to define the event of interest (e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+
+The shape of a distribution of waiting times can be described in multiple ways (Luce)
+Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
+The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+
+Similarly, after dividing time in discrete, contiguous time bins indexed by t, discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT $>$ t) and the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t), and not on the probability mass function and the cumulative distribution function.
+The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons. First, ...
+Second...
+
+We focus on factorial within-subject designs in which a large number of measurements are taken for a relatively small set of participants. Such so-called small-*N* designs are ... (Smith and little)
+
+We used `r my_r_citation$r` for all reported analyses. Web links are printed in bold.
+
+`r my_r_citation$pkgs` <!-- This creates the footnote -->
 
 # Tutorial 1: Calculating descriptive statistics using a life table
+
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016, available on [**ResearchGate**](https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect).
+In their first experiment, @panisWhatShapingRT2016 presented a double arrow for xx ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
+Participants had to indicate the direction in which the arrow pointed using their corresponding index finger.
+Response time and accuracy were recorded on each trial.
+Prime type (neutral, congruent, incongruent) and mask type were manipulated.
+Here we focus on the subset of trials in which no mask was presented.
+The xx-ms prime stimulus was a double arrow with onset at -xx ms for the congruent and incongruent prime conditions.
+
+# Tutorial 2: Fitting Bayesian hazard models
+
+# Tutorial 3: Fitting Frequentist hazard models
 
@@
 
-We used `r my_r_citation$r` for creating.
-
-`r my_r_citation$pkgs` <!-- This creates the footnote -->
-
 # Discussion

### 2024-07-22 manuscript/manuscript.tex
update manuscript folder

@@
 The authors made the following contributions.
-Sven Panis: Writing - Original Draft Preparation, Writing - Review \& Editing; Richard Ramsey: Conceptualization, Writing - Review \& Editing, Supervision.
+Sven Panis: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Richard Ramsey: Conceptualization, Writing - Review \& Editing, Supervision.
 
@@
 To move beyond mean performance measures, various distributional analyses have been proposed.
-Here we focus one particular distributional analysis known as discrete-time event history analysis, a.k.a.
+Here we focus on one particular distributional analysis known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and various options when analysing time-to-event data using discrete-time survival analysis.
+Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+We discuss how to manage inter-individual differences, implications for experimental design, and select among various options when analysing time-to-event data using discrete-time survival analysis.
 }
@@
 
-In experimental psychology, it is standard practice to analyse response times (RTs) using analysis-of-variance.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
-Such information is useful not only for interpretation, but also for cognitive psychophysiology and model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations using analysis-of-variance.
+However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
+Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+In this tutorial we focus on a distributional method for analyzing time-to-event data that is known as discrete-time event history analysis (EHA), a.k.a.
+survival, hazard, duration, failure-time, and transition analysis (Allison, 1982, 2010; Singer \& Willett, 2003).
+Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+
+To apply EHA, one must be able to define the event of interest (e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+
+The shape of a distribution of waiting times can be described in multiple ways (Luce)
+Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
+The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+
+Similarly, after dividing time in discrete, contiguous time bins indexed by t, discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT \(>\) t) and the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t), and not on the probability mass function and the cumulative distribution function.
+The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons. First, \ldots{}
+Second\ldots{}
+
+We focus on factorial within-subject designs in which a large number of measurements are taken for a relatively small set of participants.
+Such so-called small-\emph{N} designs are \ldots{} (Smith and little)
+
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{citr} (Version 0.3.2; Aust, 2019), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+Web links are printed in bold.
+
+\section{Tutorial 1: Calculating descriptive statistics using a life table}\label{tutorial-1-calculating-descriptive-statistics-using-a-life-table}
+
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016), available on \href{https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect}{\textbf{ResearchGate}}.
+In their first experiment, Panis and Schmidt (2016) presented a double arrow for xx ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
+Participants had to indicate the direction in which the arrow pointed using their corresponding index finger.
+Response time and accuracy were recorded on each trial.
+Prime type (neutral, congruent, incongruent) and mask type were manipulated.
+Here we focus on the subset of trials in which no mask was presented.
+The xx-ms prime stimulus was a double arrow with onset at -xx ms for the congruent and incongruent prime conditions.
+
+\section{Tutorial 2: Fitting Bayesian hazard models}\label{tutorial-2-fitting-bayesian-hazard-models}
+
+\section{Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{citr} (Version 0.3.2; Aust, 2019), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for creating.
-
 \section{Discussion}\label{discussion}
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-allisonDiscreteTimeMethodsAnalysis1982}
+Allison, P. D. (1982). Discrete-{Time Methods} for the {Analysis} of {Event Histories}. \emph{Sociological Methodology}, \emph{13}, 61. \url{https://doi.org/10.2307/270718}
+
+\bibitem[\citeproctext]{ref-allisonSurvivalAnalysisUsing2010}
+Allison, P. D. (2010). \emph{Survival analysis using {SAS}: A practical guide} (2. ed). Cary, NC: SAS Press.
+
 \bibitem[\citeproctext]{ref-R-citr}
@@
 
+\bibitem[\citeproctext]{ref-panisWhatShapingRT2016}
+Panis, S., \& Schmidt, T.
+(2016).
+What {Is Shaping RT} and {Accuracy Distributions}?
+{Active} and {Selective Response Inhibition Causes} the {Negative Compatibility Effect}.
+\emph{Journal of Cognitive Neuroscience}, \emph{28}(11), 1651--1671.
+\url{https://doi.org/10.1162/jocn_a_00998}
+
 \bibitem[\citeproctext]{ref-R-patchwork}
@@
 R Core Team. (2024). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+
+\bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
+Singer, J. D., \& Willett, J. B. (2003). \emph{Applied longitudinal data analysis: Modeling change and event occurrence}. Oxford ; New York: Oxford University Press.
 

### 2024-07-22 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
Finish Tutorial 1

@@
 
-This file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
+This Tutorial_1 file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
 
@@
 
-# change variable names (to pid, trial, condition, rt, and acc)
+# change relevant variable names (to pid, trial, condition, rt, and acc)
 colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
-# pid to indicate column with person IDs
+# pid to indicate column with unique person IDs
 # trial must be a unique trial number per subject
-# condition must be a factor indicating the experimental levels of the single independent variable (1,2,...)
-# rt to indicate column with response times
-# acc to indicate column with accuracies
+# condition must be a factor indicating the levels (1,2,...) and labels of the single independent variable 
+# rt to indicate column with response times (in ms)
+# acc to indicate column with accuracies (1/0)
 
@@
 
-# function extract_median is used to extract the S(t).50 quantiles (estimated median RTs)
+# function extract_median is used to extract the S(t).50 quantiles (i.e., the estimated median RTs)
 extract_median <- function(df){
@@
 # function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
-plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects
+plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects (see line 244)
   library(patchwork)
@@
 data_final <- data_nested %>% 
-                     mutate(censored  = map(data, censor, 600, 40)) %>%   # user input: censoring time, and bin width
+                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
                      mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
@@
 
-# Extract info, save plot for each subject, create person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.
+# Extract info, save plot for each subject, save a lifetable, create and save the person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.
 
@@
 pluck(data_final, 8, 1)
+
+# view the median RTs in a tibble (median_tb)
+median_bin <- map(data_final$lifetable_ca, extract_median)
+binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
+median_rt  <- modify_depth(median_bin,1,~.x*binsize)
+median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
+median_tb
 ```
@@
 ```{r save-plots}
+# save the plot for each subject
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 6, height = 8, dpi = 600)
+```
 
+```{r create-single-lifetable}
+# get the binsize in ms
+binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
+# extract the lifetable (column 7 in data_final) for condition 1 (neutral) for subject 6
+lifetable_neutral_s6 <- data_final[[7]][[6]] %>% 
+                        filter(condition=="neutral") %>% 
+                        mutate(bin = period*binsize,
+                               events = event1) %>%
+                        select(bin,risk_set,events,hazard,se_haz,survival,se_surv,ca,se_ca) 
+# get riskset for first bin
+riskset <- lifetable_neutral_s6 %>% select(risk_set) %>% slice(1) %>% pull()
+
+# create a row for time point 0
+row0 <- c(bin=0,risk_set=riskset,events=0,hazard=0,se_haz=0,survival=1,se_surv=0,ca=NA,se_ca=NA)
+# add new row
+lifetable_neutral_s6_complete <- lifetable_neutral_s6 %>%
+                                 add_row(!!! row0) %>%
+                                 arrange(bin)
+lifetable_neutral_s6_complete
+```
+
+```{r save-single-lifetable}
+write_csv(lifetable_neutral_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
 ```
@@
 ```{r create-file-hazard-modeling}
-
+# select the person-trial-bin data set for each subject, unnest them, and select relevant variables.
+input_modeling_hazard <- data_final %>% 
+                         select(pid, ptb_data) %>%  unnest(ptb_data) %>%
+                         select(pid,bl,tr,trial,condition,period,event)
 ```
@@
 ```{r save-file-hazard-modeling}
-
+write_csv(input_modeling_hazard, file="Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
 ```
 
-# Optional: Create single plot for data aggregated across subjects
+```{r create-file-ca-modeling}
+# select the censored data set for each subject, unnest them, discard trials that are censored (RT > 600 or no response detected before the deadline), and select relevant variables. 
+input_modeling_ca <- data_final %>%
+                     select(pid, censored) %>% unnest(censored) %>%
+                     filter(right_censored == 0) %>% # only keep uncensored trials
+                     select(pid,bl,tr,trial,condition,acc,drt)
+```
+
+```{r save-file-ca-modeling}
+write_csv(input_modeling_ca, file="Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
+```
+
+# Optional: Create single plot for data aggregated across subjects. This is only recommended when all subjects show the same qualitative patterns.
 
 ```{r aggregated-data-plot}
-
+# create single subject number, and unique trial number
+data_agg <- data_wr %>% mutate(oldtrial=trial,
+                               oldpid=pid,
+                               trial = 1:n(),
+                               pid = 1)
+# apply functions
+data_agg_nested <- data_agg %>% group_nest(pid)
+data_agg_final <- data_agg_nested %>% 
+                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
+                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
+                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
+                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
+                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
+                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+# view plot
+pluck(data_final, 8, 1)
 ```

### 2024-07-22 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
change neutral to blank

@@
 data_wr <- data_wr %>% mutate(condition = condition + 1,
-                              condition = factor(condition, levels = c(1,2,3), labels = c("neutral","congruent","incongruent")))
+                              condition = factor(condition, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
 summary(data_wr)
@@
   # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
-  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("neutral","congruent","incongruent"))),
+  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
             data = data_medians, 
@@
 # view lifetable (column 7) for each condition for subject 1
-pluck(data_final, 7, 1) %>% filter(condition=="neutral")
+pluck(data_final, 7, 1) %>% filter(condition=="blank")
 pluck(data_final, 7, 1) %>% filter(condition=="congruent")

### 2024-07-22 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
change neutral to blank

@@
 # extract the lifetable (column 7 in data_final) for condition 1 (neutral) for subject 6
-lifetable_neutral_s6 <- data_final[[7]][[6]] %>% 
-                        filter(condition=="neutral") %>% 
+lifetable_blank_s6 <- data_final[[7]][[6]] %>% 
+                        filter(condition=="blank") %>% 
                         mutate(bin = period*binsize,
@@
 # get riskset for first bin
-riskset <- lifetable_neutral_s6 %>% select(risk_set) %>% slice(1) %>% pull()
+riskset <- lifetable_blank_s6 %>% select(risk_set) %>% slice(1) %>% pull()
 
@@
 # add new row
-lifetable_neutral_s6_complete <- lifetable_neutral_s6 %>%
+lifetable_blank_s6_complete <- lifetable_blank_s6 %>%
                                  add_row(!!! row0) %>%
                                  arrange(bin)
-lifetable_neutral_s6_complete
+lifetable_blank_s6_complete
 ```
@@
 ```{r save-single-lifetable}
-write_csv(lifetable_neutral_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
+write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
 ```

### 2024-07-23 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
correct row0 info

@@
 
-# Read in the raw data from Panis and Schmidt (2016).
+# Read in the raw data from Panis and Schmidt (2016) - can be skipped.
 
@@
 
-# Do some wrangling to get a simpler data set.
+# Do some wrangling to get a simpler data set - can be skipped.
 
@@
 # create a row for time point 0
-row0 <- c(bin=0,risk_set=riskset,events=0,hazard=0,se_haz=0,survival=1,se_surv=0,ca=NA,se_ca=NA)
+row0 <- c(bin=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,survival=1,se_surv=0,ca=NA,se_ca=NA)
 # add new row

### 2024-07-23 manuscript/manuscript.Rmd
update manuscript

--- 2024-07-19 manuscript/manuscript.Rmd
@@
 Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and various options when analysing time-to-event data using discrete-time survival analysis.
+We discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
 
@@
 
-bibliography      : ["r-references.bib", "ownreferences.bib"]
+bibliography      : ["r-references.bib", "extrareferences.bib"]
 
@@
 knitr::opts_chunk$set(fig.pos='H', out.extra = '', out.height = "67%",
-                      fig.align = "center") # initial version
+                      fig.align = "center") 
 ```
@@
 
-In experimental psychology, it is standard practice to analyse response times (RTs) using analysis-of-variance.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations using analysis-of-variance.
+However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+In this tutorial we focus on a distributional method for analyzing time-to-event data that is known as discrete-time event history analysis (EHA), a.k.a.
+survival, hazard, duration, failure-time, and transition analysis [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
+Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+
+To apply EHA, one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+
+The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].
+Let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
+The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+
+Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+Discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT $>$ t) and the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t), and not on the probability mass function and the cumulative distribution function.
+The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+This latter function is also known as the micro-level speed-accuracy tradeoff function.
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by @holdenDispersionResponseTimes2009, “probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions” (p.
+331).
+Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if F~A~(t) > F~B~(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
+Townsend (1990) showed that a complete ordering on the hazard functions —$\lambda$~A~(t) > $\lambda$~B~(t) for all t— implies a complete ordering on both the cumulative distribution and survivor functions —F~A~(t) > F~B~(t) and S~A~(t) < S~B~(t)— which in turn implies an ordering on the mean latencies —mean A < mean B.
+In contrast, an ordering on two means does not imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does not imply a complete ordering on the corresponding hazard functions.
+This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period so that we only know that the RT must be larger than some value.
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010) which is useful for cognitive psychophysiology [@meyerModernMentalChronometry1988].
+Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+For example, fitting parametric functions or computational models to data without studying the shape of the h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+
+We focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-*N* designs).
+This  approach  emphasizes  the  precision  and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
+In contrast to the large-*N* design that averages across many participants without being able to scrutinize individual data patterns,  small-*N*  designs  retain  crucial  information  about  the  data  patterns  of  individual  observers.
+This  is  of  great  advantage whenever participants differ systematically in their strategies or in the time-courses of their effects, so that blindly averaging them would lead to misleading data patterns.
+Indeed, @smithSmallBeautifulDefense2018 argue that, “if psychology is to be a mature quantitative science, then its primary theoretical aim should be to investigate systematic functional relationships as they are manifested at the individual participant level” (p.
+2083).
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can have excellent power [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
+
+We used `r my_r_citation$r` for all reported analyses. Web links are printed in bold.
+
+`r my_r_citation$pkgs` <!-- This creates the footnote -->
 
 # Tutorial 1: Calculating descriptive statistics using a life table
+
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016, available on [**ResearchGate**](https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect).
+In their first experiment, @panisWhatShapingRT2016 presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
+Participants had to indicate the direction in which the double arrow pointed using their corresponding index finger, within 800 ms after target onset.
+Response time and accuracy were recorded on each trial.
+Prime type (blank, congruent, incongruent) and mask type were manipulated.
+Here we focus on the subset of trials in which no mask was presented.
+The 13-ms prime stimulus was a double arrow with onset at -187 ms for the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
+
+After loading in the data file, one has to (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
+The required column names are as follows:
+
+* "pid", indicating unique participant IDs;
+* "trial", indicating each unique trial per participant;
+* "condition", a factor indicating the levels of the independent variable (1, 2, ...) and the corresponding labels;
+* "rt", indicating the response times in ms;
+* "acc", inicating the accuracies (1/0).
+
+In the code of Tutorial 1, this is accomplished as follows.
+
+\scriptsize
+```{r setup-data-tut1, echo=TRUE}
+data_wr <- read_csv("../Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
+colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
+data_wr <- data_wr %>% 
+  mutate(condition = condition + 1, # original levels were 0, 1, 2.
+         condition = factor(condition, levels=c(1,2,3), labels=c("blank","congruent","incongruent")))
+```
+\normalsize
+
+```{r load-functions, echo=F}
+# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
+censor <- function(df, timeout, bin_width){
+  if(!(timeout %% bin_width == 0)){
+    return("The censoring time must be a multiple of the bin width!")
+  }
+  if(timeout < 0 | bin_width < 0){
+    return("Both timeout and bin_width must be larger than 0!")
+  }
+  df %>% mutate(right_censored = 0,
+                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
+                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
+                drt = ceiling(rtc/bin_width), # drt = discrete response time
+                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
+}
+
+# function ptb creates a person-trial-bin oriented data set
+ptb <- function(df){
+  df %>% uncount(weights = drt) %>% # create drt rows per trial
+         group_by(trial) %>% 
+         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
+         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
+         ungroup()
+}
+
+# function setup_lt sets up a life table for each level of condition (1 independent variable)
+setup_lt <- function(ptb){
+  ptb %>% mutate(event = str_c("event", event)) %>%
+          group_by(condition,period) %>% 
+          count(event) %>% 
+          ungroup() %>% 
+          pivot_wider(names_from = event,
+                      values_from = n) %>% 
+          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+                 event1 = ifelse(is.na(event1),0,event1),
+                 risk_set = event0 + event1) %>% # define the risk set
+          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
+          group_by(condition) %>%
+          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
+                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
+                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+          ungroup() 
+}
+
+# function calc_ca calculates the conditional accuracies 
+calc_ca <- function(df){
+  df %>% filter(right_censored==0) %>%
+         group_by(condition,drt,cens_time,bin_width) %>%
+         summarize(ca = mean(acc) %>% round(digits = 2),
+                   n = n(),
+                   .groups = 'drop') %>%
+         ungroup() %>%
+         mutate(period = drt,
+                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
+         select(-drt)
+}
+
+# function join_lt_ca joins the conditional accuracies to the life tables
+join_lt_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}
+
+# function extract_median is used to extract the S(t).50 quantiles (i.e., the estimated median RTs)
+extract_median <- function(df){
+  above_pct50 <- df %>% 
+      group_by(condition) %>%
+      filter(survival > .5) %>% 
+      slice(n()) # take last row
+  below_pct50 <- df %>% 
+      group_by(condition) %>%
+      filter(survival < .5) %>% 
+      slice(1) # take first row
+  # pull period above
+  period_above <- pull(above_pct50, period)
+  # pull survivor function values
+  survival_above <- pull(above_pct50, survival)
+  survival_below <- pull(below_pct50, survival)
+  # estimate median by interpolation
+  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
+}
+
+# function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
+plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects (see line 244)
+  library(patchwork)
+  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
+  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
+  median_period <- extract_median(df)
+  n_conditions <- nlevels(df$condition)
+  data_median <- c()
+  for(i in 1:n_conditions){
+    data_median <- append(data_median, c(median_period[i], median_period[i]))
+  }
+
+  data_medians <- tibble(period= data_median,
+                         survival = rep(c(.5, 0),n_conditions),
+                         condition = rep(1:n_conditions, each=2))
+# plot the hazard functions
+p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=hazard)) +
+  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
+  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits = c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,haz_yaxis)) +
+  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
+  theme(legend.background = element_rect(fill = "transparent"),
+        panel.grid = element_blank(),
+        legend.position = "top")
+# plot the survivor functions
+p2 <-df %>%
+  ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=survival), show.legend = F) +
+  geom_point(aes(y=survival), size=1, show.legend = F) +
+  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
+  # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
+  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
+            data = data_medians, 
+            linetype = 3, show.legend = F) +
+
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits=c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,1)) +
+  labs(x="", y="S(t)",
+       colour="Condition") +
+  theme(panel.grid = element_blank())
+# plot the conditional accuracy functions
+p3 <-df %>%
+  ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=ca), show.legend = F) +
+  geom_point(aes(y=ca), size=1, show.legend = F) +
+  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits=c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,1)) +
+  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
+       colour="Condition") +
+  theme(panel.grid = element_blank())
+
+p1/p2/p3
+}
+```
+
+To set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t) using functional programming, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our function "censor()", as follows.
+
+\scriptsize
+```{r nest-apply-functions-tut1, echo = TRUE}
+data_nested <- data_wr %>% group_nest(pid)
+data_final <- data_nested %>% 
+  mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
+  mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
+  mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca(t)
+  mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca(t)
+  mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca(t)
+  mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+```
+\normalsize
+
+Note that the censoring time should be a multiple of the bin width (both in ms).
+The censoring time should be a time point after which no informative responses are expected anymore.
+In experiments that implement a response deadline in each trial the censoring time can equal that deadline time point.
+Trials with a RT larger than the censoring time, or trials in which no response is emitted during the data collection period, are treated as right-censored observations in EHA.
+In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
+Removing such trials before calculating the mean event time can introduce a sampling bias.
+The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
+The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
+One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
+
+Table 1 shows the life table for condition "blank" (no prime stimulus presented) - compare to Figure 1.
+A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
+At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
+
+Figure 1 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
+By using discrete-time h(t) functions of event occurrence - in combination with ca(t) functions for two-choice tasks - one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+
+For example, for participant 6, the estimated hazard values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, when the waiting time has increased until *240 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
+
+(ref:life-table-caption) The life table for the blank prime condition of participant 6.
+
+(ref:life-note-caption) The column named "bin" indicates the endpoint of each time bin (in ms), and includes time point zero.
+For example the first bin is (0,40] with the starting point excluded and the endpoint included.
+se = standard error.
+ca = conditional accuracy.
+
+```{r life-table}
+life_tab <- read_csv("../Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
+
+apa_table(
+  life_tab,
+  caption = "(ref:life-table-caption)",
+  note = "(ref:life-note-caption)",
+  placement = "H"
+)
+```
+
+(ref:descr-fig-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
+
+```{r eha-plot, fig.cap = "(ref:descr-fig-caption)"}
+knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt.png")
+```
+
+Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+
+  However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+
+  These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+Qualitatively similar results were obtained for the other five participants.
+Also, in their second Experiment, @panisWhatShapingRT2016 showed that the negative compatibility effect in the mask-present conditions is time-locked to mask onset.
+This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022; @schmidtResponseInhibitionNegative2022].
+In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and response accuracy (1/0) in each bin of a selected time range, respectively.
+
+# Tutorial 2: Fitting Bayesian hazard models
+
+# Tutorial 3: Fitting Frequentist hazard models
 

### 2024-07-23 manuscript/manuscript.tex
update manuscript

@@
 \usepackage{xcolor}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and select among various options when analysing time-to-event data using discrete-time survival analysis.
+Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+We discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
 }
@@
 survival, hazard, duration, failure-time, and transition analysis (Allison, 1982, 2010; Singer \& Willett, 2003).
-Across three tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
 
-To apply EHA, one must be able to define the event of interest (e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+To apply EHA, one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
 
-The shape of a distribution of waiting times can be described in multiple ways (Luce)
+The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991).
+Let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
 Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
@@
 
-Similarly, after dividing time in discrete, contiguous time bins indexed by t, discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT \(>\) t) and the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t), and not on the probability mass function and the cumulative distribution function.
+Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+Discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT \(>\) t) and the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t), and not on the probability mass function and the cumulative distribution function.
 The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+This latter function is also known as the micro-level speed-accuracy tradeoff function.
 
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons. First, \ldots{}
-Second\ldots{}
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions'' (p.~331).
+Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
+Townsend (1990) showed that a complete ordering on the hazard functions ---\(\lambda\)\textsubscript{A}(t) \textgreater{} \(\lambda\)\textsubscript{B}(t) for all t--- implies a complete ordering on both the cumulative distribution and survivor functions ---F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) and S\textsubscript{A}(t) \textless{} S\textsubscript{B}(t)--- which in turn implies an ordering on the mean latencies ---mean A \textless{} mean B.
+In contrast, an ordering on two means does not imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does not imply a complete ordering on the corresponding hazard functions.
+This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period so that we only know that the RT must be larger than some value.
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010) which is useful for cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+For example, fitting parametric functions or computational models to data without studying the shape of the h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
 
-We focus on factorial within-subject designs in which a large number of measurements are taken for a relatively small set of participants.
-Such so-called small-\emph{N} designs are \ldots{} (Smith and little)
+We focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-\emph{N} designs).
+This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
+In contrast to the large-\emph{N} design that averages across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
+This is of great advantage whenever participants differ systematically in their strategies or in the time-courses of their effects, so that blindly averaging them would lead to misleading data patterns.
+Indeed, Smith and Little (2018) argue that, ``if psychology is to be a mature quantitative science, then its primary theoretical aim should be to investigate systematic functional relationships as they are manifested at the individual participant level'' (p.~2083).
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can have excellent power (Baker et al., 2021; Smith \& Little, 2018).
 
@@
 To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016), available on \href{https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect}{\textbf{ResearchGate}}.
-In their first experiment, Panis and Schmidt (2016) presented a double arrow for xx ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
-Participants had to indicate the direction in which the arrow pointed using their corresponding index finger.
+In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
+Participants had to indicate the direction in which the double arrow pointed using their corresponding index finger, within 800 ms after target onset.
 Response time and accuracy were recorded on each trial.
-Prime type (neutral, congruent, incongruent) and mask type were manipulated.
+Prime type (blank, congruent, incongruent) and mask type were manipulated.
 Here we focus on the subset of trials in which no mask was presented.
-The xx-ms prime stimulus was a double arrow with onset at -xx ms for the congruent and incongruent prime conditions.
+The 13-ms prime stimulus was a double arrow with onset at -187 ms for the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
+
+After loading in the data file, one has to (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
+The required column names are as follows:
+
+\begin{itemize}
+\tightlist
+\item
+  ``pid'', indicating unique participant IDs;
+\item
+  ``trial'', indicating each unique trial per participant;
+\item
+  ``condition'', a factor indicating the levels of the independent variable (1, 2, \ldots) and the corresponding labels;
+\item
+  ``rt'', indicating the response times in ms;
+\item
+  ``acc'', inicating the accuracies (1/0).
+\end{itemize}
+
+In the code of Tutorial 1, this is accomplished as follows.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{data\_wr }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/DataExp1\_6subjects\_wrangled.csv"}\NormalTok{)}
+\FunctionTok{colnames}\NormalTok{(data\_wr) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"pid"}\NormalTok{,}\StringTok{"bl"}\NormalTok{,}\StringTok{"tr"}\NormalTok{,}\StringTok{"condition"}\NormalTok{,}\StringTok{"resp"}\NormalTok{,}\StringTok{"acc"}\NormalTok{,}\StringTok{"rt"}\NormalTok{,}\StringTok{"trial"}\NormalTok{) }
+\NormalTok{data\_wr }\OtherTok{\textless{}{-}}\NormalTok{ data\_wr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =}\NormalTok{ condition }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, }\CommentTok{\# original levels were 0, 1, 2.}
+         \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{), }\AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{,}\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+To set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t) using functional programming, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our function ``censor()'', as follows.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{data\_nested }\OtherTok{\textless{}{-}}\NormalTok{ data\_wr }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{group\_nest}\NormalTok{(pid)}
+\NormalTok{data\_final }\OtherTok{\textless{}{-}}\NormalTok{ data\_nested }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{censored  =} \FunctionTok{map}\NormalTok{(data, censor, }\DecValTok{600}\NormalTok{, }\DecValTok{40}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}   \CommentTok{\# !
+user input: censoring time, and bin width}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data  =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}           \CommentTok{\# create person{-}trial{-}bin dataset}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}      \CommentTok{\# create life tables without ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condacc   =} \FunctionTok{map}\NormalTok{(censored, calc\_ca)) }\SpecialCharTok{\%\textgreater{}\%}       \CommentTok{\# calculate ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable\_ca =} \FunctionTok{map2}\NormalTok{(lifetable, condacc, join\_lt\_ca)) }\SpecialCharTok{\%\textgreater{}\%}    \CommentTok{\# create life tables with ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{plot      =} \FunctionTok{map2}\NormalTok{(}\AttributeTok{.x =}\NormalTok{ lifetable\_ca, }\AttributeTok{.y =}\NormalTok{ pid, plot\_eha,}\DecValTok{1}\NormalTok{))  }\CommentTok{\# create plots }
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Note that the censoring time should be a multiple of the bin width (both in ms).
+The censoring time should be a time point after which no informative responses are expected anymore.
+In experiments that implement a response deadline in each trial the censoring time can equal that deadline time point.
+Trials with a RT larger than the censoring time, or trials in which no response is emitted during the data collection period, are treated as right-censored observations in EHA.
+In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
+Removing such trials before calculating the mean event time can introduce a sampling bias.
+The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
+The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
+One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
+
+Table 1 shows the life table for condition ``blank'' (no prime stimulus presented) - compare to Figure 1.
+A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
+At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
+
+Figure 1 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
+By using discrete-time h(t) functions of event occurrence - in combination with ca(t) functions for two-choice tasks - one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+
+For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.22 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:life-table}This is the life table for condition blank of participant 6.}
+
+\begin{tabular}{lllllllll}
+\toprule
+bin & \multicolumn{1}{c}{risk\_set} & \multicolumn{1}{c}{events} & \multicolumn{1}{c}{hazard} & \multicolumn{1}{c}{se\_haz} & \multicolumn{1}{c}{survival} & \multicolumn{1}{c}{se\_surv} & \multicolumn{1}{c}{ca} & \multicolumn{1}{c}{se\_ca}\\
+\midrule
+0.00 & 219.00 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
+40.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+80.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+120.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+160.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+200.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+240.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+280.00 & 219.00 & 7.00 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
+320.00 & 212.00 & 13.00 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
+360.00 & 199.00 & 26.00 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
+400.00 & 173.00 & 40.00 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
+440.00 & 133.00 & 48.00 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
+480.00 & 85.00 & 37.00 & 0.44 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
+520.00 & 48.00 & 32.00 & 0.67 & 0.07 & 0.07 & 0.02 & 1.00 & 0.00\\
+560.00 & 16.00 & 9.00 & 0.56 & 0.12 & 0.03 & 0.01 & 1.00 & 0.00\\
+600.00 & 7.00 & 4.00 & 0.57 & 0.19 & 0.01 & 0.01 & 1.00 & 0.00\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} The column named ``bin'' indicates the endpoint of each time bin (in ms), and includes time point zero. se = standard error. ca = conditional accuracy.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=12in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+
+}
+
+\caption{Discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
+\end{figure}
+
+Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+
+However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.07 for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
+
+These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+Qualitatively similar results were obtained for the other five participants.
+Also, in their second Experiment, Panis and Schmidt (2016) showed that the negative compatibility effect in the mask-present conditions is time-locked to mask onset.
+This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Schmidt, Panis, Wolkersdorfer, \& Vorberg, 2022).
+In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and response accuracy (1/0) in each bin of a selected time range, respectively.
 
@@
 
+\bibitem[\citeproctext]{ref-bakerPowerContoursOptimising2021}
+Baker, D.
+H., Vilidaite, G., Lygo, F.
+A., Smith, A.
+K., Flack, T.
+R., Gouws, A.
+D., \& Andrews, T.
+J.
+(2021).
+Power contours: {Optimising} sample size and precision in experimental psychology and human neuroscience.
+\emph{Psychological Methods}, \emph{26}(3), 295--314.
+\url{https://doi.org/10.1037/met0000337}
+
 \bibitem[\citeproctext]{ref-R-tinylabels}
@@
 
+\bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
+Holden, J.
+G., Van Orden, G.
+C., \& Turvey, M.
+T.
+(2009).
+Dispersion of response times reveals cognitive dynamics.
+\emph{Psychological Review}, \emph{116}(2), 318--342.
+\url{https://doi.org/10.1037/a0014849}
+
+\bibitem[\citeproctext]{ref-kantowitzInterpretationReactionTime2021}
+Kantowitz, B.
+H., \& Pachella, R.
+G.
+(2021).
+The {Interpretation} of {Reaction Time} in {Information-Processing Research} 1.
+\emph{Human Information Processing}, 41--82.
+\url{https://doi.org/10.4324/9781003176688-2}
+
+\bibitem[\citeproctext]{ref-kelsoOutlineGeneralTheory2013}
+Kelso, J.
+A.
+S., Dumas, G., \& Tognoli, E.
+(2013).
+Outline of a general theory of behavior and brain coordination.
+\emph{Neural Networks: The Official Journal of the International Neural Network Society}, \emph{37}, 120--131.
+\url{https://doi.org/10.1016/j.neunet.2012.09.003}
+
+\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
+Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
+\bibitem[\citeproctext]{ref-meyerModernMentalChronometry1988}
+Meyer, D. E., Osman, A. M., Irwin, D. E., \& Yantis, S. (1988). Modern mental chronometry. \emph{Biological Psychology}, \emph{26}(1-3), 3--67. \url{https://doi.org/10.1016/0301-0511(88)90013-0}
+
 \bibitem[\citeproctext]{ref-R-tibble}
@@
 Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}
+
+\bibitem[\citeproctext]{ref-panisHowCanWe2020}
+Panis, S.
+(2020).
+How can we learn what attention is?
+{Response} gating via multiple direct routes kept in check by inhibitory control processes.
+\emph{Open Psychology}, \emph{2}(1), 238--279.
+\url{https://doi.org/10.1515/psych-2020-0107}
+
+\bibitem[\citeproctext]{ref-panisStudyingDynamicsVisual2020}
+Panis, S., Moran, R., Wolkersdorfer, M.
+P., \& Schmidt, T.
+(2020).
+Studying the dynamics of visual search behavior using {RT} hazard and micro-level speed--accuracy tradeoff functions: {A} role for recurrent object recognition and cognitive control processes.
+\emph{Attention, Perception, \& Psychophysics}, \emph{82}(2), 689--714.
+\url{https://doi.org/10.3758/s13414-019-01897-z}
 
@@
 
+\bibitem[\citeproctext]{ref-panisWhenDoesInhibition2022}
+Panis, S., \& Schmidt, T.
+(2022).
+When does {``inhibition of return''} occur in spatial cueing tasks?
+{Temporally} disentangling multiple cue-triggered effects using response history and conditional accuracy analyses.
+\emph{Open Psychology}, \emph{4}(1), 84--114.
+\url{https://doi.org/10.1515/psych-2022-0005}
+
+\bibitem[\citeproctext]{ref-panisNeuropsychologicalEvidenceTemporal2017}
+Panis, S., Torfs, K., Gillebert, C.
+R., Wagemans, J., \& Humphreys, G.
+W.
+(2017).
+Neuropsychological evidence for the temporal dynamics of category-specific naming.
+\emph{Visual Cognition}, \emph{25}(1-3), 79--99.
+\url{https://doi.org/10.1080/13506285.2017.1330790}
+
+\bibitem[\citeproctext]{ref-panisTimecourseContingenciesPerceptual2009}
+Panis, S., \& Wagemans, J.
+(2009).
+Time-course contingencies in perceptual organization and identification of fragmented object outlines.
+\emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{35}(3), 661--687.
+\url{https://doi.org/10.1037/a0013547}
+
 \bibitem[\citeproctext]{ref-R-patchwork}
@@
 
+\bibitem[\citeproctext]{ref-schmidtResponseInhibitionNegative2022}
+Schmidt, T., Panis, S., Wolkersdorfer, M.
+P., \& Vorberg, D.
+(2022).
+Response inhibition in the {Negative Compatibility Effect} in the absence of inhibitory stimulus features.
+\emph{Open Psychology}, \emph{4}(1), 219--230.
+\url{https://doi.org/10.1515/psych-2022-0012}
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
 Singer, J. D., \& Willett, J. B. (2003). \emph{Applied longitudinal data analysis: Modeling change and event occurrence}. Oxford ; New York: Oxford University Press.
+
+\bibitem[\citeproctext]{ref-smithSmallBeautifulDefense2018}
+Smith, P.
+L., \& Little, D.
+R.
+(2018).
+Small is beautiful: {In} defense of the small-{N} design.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(6), 2083--2101.
+\url{https://doi.org/10.3758/s13423-018-1451-8}
+
+\bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}
+Townsend, J.
+T.
+(1990).
+Truth and consequences of ordinal differences in statistical distributions: {Toward} a theory of hierarchical inference.
+\emph{Psychological Bulletin}, \emph{108}(3), 551--567.
+\url{https://doi.org/10.1037/0033-2909.108.3.551}
+
+\bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}
+Wickelgren, W. A. (1977). Speed-accuracy tradeoff and information processing dynamics. \emph{Acta Psychologica}, \emph{41}(1), 67--85. \url{https://doi.org/10.1016/0001-6918(77)90012-9}
 

### 2024-07-24 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
correct error with resp==0

@@
 ```{r read-raw-data}
-raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T)
+raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) # 9372 rows
 head(raw)
@@
 # remove practice trials
-data <- raw %>% filter(practice==0) # 9240 trials: 6 participants x 70 blocks x 22 trials-per-block
+data <- raw %>% filter(practice==0) # 9240 experimental trials: 6 participants x 70 blocks x 22 trials-per-block
 
@@
                                              void,target_dir,corresp,resperr))
-# remove trials with no response info
-sel = which(data$resp==0)
-data <- data[-c(sel),] # 9189 rows left
+# look at trials where resp==0
+data %>% filter(resp==0) 
+data %>% filter(resp==0) %>% print(n=51)
+
+# remove trials with no response info if rt < 800 (800 was the response deadline in each trial):
+# the 46 trials where rt >= 800 and resp==0 are trials in which the subject did not respond; these should not be deleted because they will be treated as right-censored observations in EHA!
+# the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
+sel = which(data$resp==0 & data$rt < 800)
+data <- data[-c(sel),] # 9235 rows
 
@@
 data <- data %>% filter(mask_type == 0) %>%
-                 select(-mask_type) # 2737 rows left, and 8 variables
-
+                 select(-mask_type) # 2757 rows and 8 variables
 summary(data)
@@
 # tr = trial number within block
-# prime_type: no prime (0), consistent prime (1), inconsistent prime (2)
-# resp = response left (1) or right (2)
+# prime_type: no prime or blank (0), consistent prime (1), inconsistent prime (2)
+# resp = no button (0), button left (1), button right (2) 
 # respac = accuracy
@@
 
-# change relevant variable names (to pid, trial, condition, rt, and acc)
+# change some variable names to required names (pid, trial, condition, rt, and acc)
 colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
@@
 
-Load the functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 205) as above (line 83).
+Load the functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 210) as above (line 88).
 
@@
                          select(pid,bl,tr,trial,condition,period,event)
+summary(input_modeling_hazard)
 ```

### 2024-07-24 Tutorial_2_Bayesian/Tutorial_2.Rmd
add tutorial 2

---
title: "Tutorial_2"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build a Bayesian hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions.

# Load the libraries that we will be using

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

# Set options.

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()
detectCores()
```

```{r check-info}
packageVersion("cmdstanr")
devtools::session_info("rstan")
```

# Load the person-trial-bin data set that we saved in Tutorial 1.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, 3 conditions, 15 periods, and event indicator (0/1)
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
ptb_data <- ptb_data %>% filter(period > 5)

# create factor condition, with "blank" as the reference level
ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
summary(ptb_data)

# center discrete TIME (period) on bin 9, and trial on trial 1000
ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
                                trial_c = (trial - 1000)/1000)

# create dummy variables for each time period (to illustrate how to fit a general specification of TIME)
ptb_data <- ptb_data %>% mutate(d6 = if_else(period == 6, 1, 0),
                                d7 = if_else(period == 7, 1, 0),
                                d8 = if_else(period == 8, 1, 0),
                                d9 = if_else(period == 9, 1, 0),
                                d10 = if_else(period == 10, 1, 0),
                                d11 = if_else(period == 11, 1, 0),
                                d12 = if_else(period == 12, 1, 0),
                                d13 = if_else(period == 13, 1, 0),
                                d14 = if_else(period == 14, 1, 0),
                                d15 = if_else(period == 15, 1, 0))

# remove unnecessary columns before fitting a model
ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period,d6,d7,d8,d9,d10,d11,d12,d13,d14,d15)) # 12840 obs. of 5 variables
head(ptb_data,n=17)
summary(ptb_data)
```

# Set up priors, and fit the model with full random effects structure.

```{r priors}
priors <- c(
  set_prior("normal(0, 1)", class = "b"), # for beta parameters 
  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale
  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
  set_prior("lkj(2)", class = "cor") # for correlations between RE
)
```

```{r full-random-effects-model}
plan(multicore)

model_full_RE <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9*trial_c + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           (1 + condition*period_9*trial_c +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_full_RE")

# the formula above comes down to the following effects:
# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+ ##effect of incongruent interacts cubicly with TIME
# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
```

Model_full_RE took 3 hours to run.

```{r check-model-full-RE}
summary(model_full_RE)
```

```{r full-random-effects-model}
plan(multicore)

model_cubic_RE <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_cubic_RE")

# the formula above comes down to the following effects:
# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3) ##effect of incongruent interacts cubicly with TIME
```

```{r check-model-full-RE}
summary(model_cubic_RE)
```

# plot hazard functions

```{r}

```

# visualize posterior distributions of the effects of congruent and incongruent primes on hazard relative to blank prime, for each time bin in trial 1000, and for each time bin in trial 100.

```{r load-model}
model_cubic_RE <- readRDS("Tutorial_2_Bayesian/models/model_cubic_RE.rds") 
```

```{r make-plot-effects}
post <-as_draws_df(model_cubic_RE) %>% # 8000 draws x 299 variables
   select(starts_with("b_")) %>%       # 8000 x 18
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial100 = -900/1000)  
# effects for trial 1000
p1congruent <- post %>% 
  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

p2incongruent <- post %>% 
  mutate(postdistr = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  labs(x = "time bin", y = "effect of incongruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

p1congruent + p2incongruent
#(p1+p2+p3)/(p4+p5+p6)/(p7+p8+p9)/(p10+p11+p12)/(p13+p14+p15)
```

```{r save-plot-effects}
ggsave("figures/effects_knot_1.png", width = 12, height = 16, dpi = 600)

```

### 2024-07-24 manuscript/manuscript.Rmd
update manuscript

@@
 ---
-title             : "A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data"
-shorttitle        : "Tutorial on hazard modeling"
+title             : "A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data"
+shorttitle        : "Tutorial on hazard analysis"
 
@@
 Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
+Along the way, we discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
 
@@
 The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
+When creating the plots using our function plot_eha(), some warning messages will likely be generated, like these:
+
+* Removed 2 rows containing missing values or values outside the scale range (`geom_line()`). 
+* Removed 2 rows containing missing values or values outside the scale range (`geom_point()`). 
+* Removed 2 rows containing missing values or values outside the scale range (`geom_segment()`).
+
+The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
@@
 In other words, when the waiting time has increased until *240 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
+
+Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+
+  However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 
-Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-
-  However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-
   These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
@@
 
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the data set used in the first tutorial.
+
+model formula 
+analytic decisions: specification of the effect of TIME,
+
+assumptions:
+- linearity: continuous variables (TIME, trial number) 
+- additivity:
+- proportionality:
+
 # Tutorial 3: Fitting Frequentist hazard models
 
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+In this third tutorial we illustrate how to fit a frequentist hazard regression model for the data set used in the first tutorial.
 
-## Participants
+# Tutorial 4: Calculating descriptive statistics when there are two independent variables
 
-## Material
-
-## Procedure
-
-## Data analysis
-<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->
-
-# Results
+In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016.
 
 # Discussion
+
+## Individual differences
+
+- role of response deadlines, low-level vs. higher-level processes, 
+- clustering algorithms based on h(t) and ca(t) data
+-
+
+## Cognitive psychophysiology and computational model selection
+
+## Power analysis
+
+- example repo on github
+
+## Preregistration
+
+- example preregistration for knot data
+
+# Conclusions
 

### 2024-07-24 manuscript/manuscript.tex
update manuscript

@@
 \hypersetup{
-  pdftitle={A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data},
+  pdftitle={A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data},
   pdfauthor={Sven Panis1 \& Richard Ramsey1},
@@
 
-\title{A tutorial on Bayesian and Frequentist Event History Analysis for psychological time-to-event data}
+\title{A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data}
 \author{Sven Panis\textsuperscript{1} \& Richard Ramsey\textsuperscript{1}}
@@
 
-\shorttitle{Tutorial on hazard modeling}
+\shorttitle{Tutorial on hazard analysis}
 
@@
 Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-We discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
+Along the way, we discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
 }
@@
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
+When plotting the data, some warning messages will be generated, like these:
+
+\begin{itemize}
+\tightlist
+\item
+  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_line()}).
+\item
+  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_point()}).
+\item
+  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_segment()}).
+\end{itemize}
+
+The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
@@
 
+Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+
+However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.07 for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
+
 \begin{table}[H]
@@
 
-\caption{\label{tab:life-table}This is the life table for condition blank of participant 6.}
+\caption{\label{tab:life-table}The life table for the blank prime condition of participant 6.}
 
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} The column named ``bin'' indicates the endpoint of each time bin (in ms), and includes time point zero. se = standard error. ca = conditional accuracy.}
+\normalsize{\textit{Note.} The column named ``bin'' indicates the endpoint of each time bin (in ms), and includes time point zero.
+For example the first bin is (0,40{]} with the starting point excluded and the endpoint included.
+se = standard error.
+ca = conditional accuracy.}
 \end{tablenotes}
@@
 
-\caption{Discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
+\caption{Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
 \end{figure}
-
-Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-
-However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.07 for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the data set used in the first tutorial.
+
 \section{Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
 
-\section{Methods}\label{methods}
+In this third tutorial we illustrate how to fit a frequentist hazard regression model for the data set used in the first tutorial.
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+\section{Tutorial 4: Calculating descriptive statistics when there are two independent variables}\label{tutorial-4-calculating-descriptive-statistics-when-there-are-two-independent-variables}
 
-\subsection{Participants}\label{participants}
-
-\subsection{Material}\label{material}
-
-\subsection{Procedure}\label{procedure}
-
-\subsection{Data analysis}\label{data-analysis}
-
-\section{Results}\label{results}
+In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016).
 
 \section{Discussion}\label{discussion}
+
+\subsection{Individual differences}\label{individual-differences}
+
+\begin{itemize}
+\tightlist
+\item
+  role of response deadlines, low-level vs.~higher-level processes,
+\item
+  clustering algorithms based on h(t) and ca(t) data
+\item
+\end{itemize}
+
+\subsection{Cognitive psychophysiology and computational model selection}\label{cognitive-psychophysiology-and-computational-model-selection}
+
+\subsection{Power analysis}\label{power-analysis}
+
+\begin{itemize}
+\tightlist
+\item
+  example repo on github
+\end{itemize}
+
+\subsection{Preregistration}\label{preregistration}
+
+\begin{itemize}
+\tightlist
+\item
+  example preregistration for knot data
+\end{itemize}
+
+\section{Conclusions}\label{conclusions}
 

### 2024-07-24 Tutorial_2_Bayesian/Tutorial_2.Rmd
update Tutorial 2

@@
 print(ptb_data,n=30)
-summary(ptb_data) # 26602 rows: 6 participants, 3 conditions, 15 periods, and event indicator (0/1)
+summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
 ```
@@
 
-# create dummy variables for each time period (to illustrate how to fit a general specification of TIME)
-ptb_data <- ptb_data %>% mutate(d6 = if_else(period == 6, 1, 0),
-                                d7 = if_else(period == 7, 1, 0),
-                                d8 = if_else(period == 8, 1, 0),
-                                d9 = if_else(period == 9, 1, 0),
-                                d10 = if_else(period == 10, 1, 0),
-                                d11 = if_else(period == 11, 1, 0),
-                                d12 = if_else(period == 12, 1, 0),
-                                d13 = if_else(period == 13, 1, 0),
-                                d14 = if_else(period == 14, 1, 0),
-                                d15 = if_else(period == 15, 1, 0))
-
 # remove unnecessary columns before fitting a model
-ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period,d6,d7,d8,d9,d10,d11,d12,d13,d14,d15)) # 12840 obs. of 5 variables
+ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
 head(ptb_data,n=17)
 summary(ptb_data)
+# 6 subjects
+# condition: blank (6401), congruent (2642), incongruent (3797)
+# period_9: -3 to 6
+# trial_c: -0.999 to 0.540
 ```

### 2024-08-06 manuscript/manuscript.Rmd
I made some edits to the start of the intro

@@
 shorttitle        : "Tutorial on hazard analysis"
-
 author: 
@@
       - "Supervision"
-
 affiliation:
@@
     institution   : "ETH Zürich"
-
 authornote: |
   Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).
-
   <!-- Enter author note here. -->
-
 abstract: |
@@
   <!-- https://tinyurl.com/ybremelq -->
-
 keywords          : "response times, event history analysis, Bayesian regression models"
 wordcount         : "X"
-
 bibliography      : ["r-references.bib", "extrareferences.bib"]
-
 floatsintext      : yes
@@
 mask              : no
-
 figurelist        : no
@@
 footnotelist      : no
-
 classoption       : "man"
@@
 
-In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations using analysis-of-variance.
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating mean average performance across a series of trials.
 However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
@@
 survival, hazard, duration, failure-time, and transition analysis [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+Across four tutorials that we make publicly available on Github and the Open Science Framework (OSF), we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
 

### 2024-08-06 notes/manuscript_notes.Rmd
I added some notes/feedback on the draft manuscript

---
title: "manuscript_notes"
author: "Rich"
date: "2024-08-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is a way to provide feedback on draft versions of the manuscript. 
I can edit the manuscript directly of course and I have done in a few places (which git should track), but this is a space to provide comments.

I'm not sure of this is the best way to go. But so be it.

At the end we can add this folder to .gitignore, so we don't need to post this online.

# Aug 6th 2024 #

## Abstract ##

You say: To move beyond mean performance measures, various distributional analyses have been proposed.

This makes complete sense. But I would say it differently. I would emphasise in a sentence or two why mean performance measures are limited, and thus highlight the benefit of non-mean measures. And maybe try to find simpler language for distributional measures. Again, it is completely precise and accurate, but I worry that most folks will read it and think "that aint anything I care about" when in actual fact it might be extremely relevant. So, in other words, I think the abstract needs a clearer hook for the generalist experimental psych reader. Something like ... mean avg measures (across trials) hide/obscure a lot of information about psychological processes, such as when psychological states occur in time. In contrast, blah blah approaches reveal how psych processes occur in specific time windows (etc. etc.). And this changes key conclusions about psych processes. Here we provide a tutorial on how to do this [[and then return to the details of the tutorial as you already state]]. Even though I have written a lot here, I expect the changes to the abstract to be small. I just wanted to try to make my point clear here to you. I hope that makes sense.

## Introduction ##

I would slightly re-structure the introduction. I would use some subsections in the intro maybe also to organise ideas. The main change that I propose is to have a first section that provides the motivation in simple terms (non-modelling terms), as well as a roadmap for what is to come next in the rest of the paper, so that folks know what to expect. This is especially important for reviews/tutorials etc., since it doesn't follow the standard empirical paper approach i.e., intro, methods, results ... Then I would move all of the modelling stuff that is currently in the intro into an entirely new section. Here are some ideas on structure with example sections numbered...

1. motivation = means vs distribution/timewindow approach. Benefits etc. I would have a plot in this part of the intro. Just real or simulated raw data - like we discussed in our meeting on Monday for your future talk - bar plots of means between conditions juxtaposed with a distributional analysis across time. I think the visually hits home the basic point about how timing is important. Then all the modelling and lifetable stuff  that comes later is important as a method, but we get in there nice and early with a clear message on why it is so ubiquitous in exp psych (e.g., almost everyone plots mean RT, for example). So we make the benefit/value really clear, really early.

2. Roadmap of the paper, which outlines what to expect. e.g., We first provide a brief overview of hazard models to orient the reader to the basic concepts and ideas that we will use through the paper. However, this will remain relatively short, and for detailed treatment, see blah blah REFs. And we *must* provide some figures here. Figures really help with stats stuff, because it brings more people in. Can you think of some figures to illustrate key modelling concepts? Or do you think the mean vs distribution figure above in point (1) will be enough? We then provide 4 different tutorials, each of which is written is R code and publicly available. The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can also analyse time to event data.... Tutorial 1 blah. 2 blah. 3 blah. 4 blah. Provide a sentence or two on each tutorial.

## Background on EHA analyses ##

Here is where you move the modelling stuff into a separate section. 
This is where you get into the more technical stuff.
keep this short and refer to other, more detailed reviews (including your own papers).
The reason for keeping this section short is that we want to focus on a worked examples of how to do the analyses, rather than justify why you should use these techniques. Loads of papers have focussed on why before. And let's say this explicitly at the start of this section, so that if anyone complains that we need more background info., we have already explained why we kept it short.

## Tutorial 1 ##

## Tutorial 2 ##

...

### 2024-08-07 manuscript/manuscript.Rmd
update manuscript

--- 2024-07-24 manuscript/manuscript.Rmd
@@
 ```{r setup, include = FALSE}
-pkg <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")
+# Tutorial 1 and 4
+pkg1 <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")
 
-lapply(pkg, library, character.only = TRUE)
+lapply(pkg1, library, character.only = TRUE)
 
+# Tutorial 2
+pkg2 <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
+         "brms", "tidybayes", "bayesplot", "future", "parallel")
+
+lapply(pkg2, library, character.only = TRUE)
+
+options(brms.backend = "cmdstanr",
+        mc.cores = parallel::detectCores(),
+        future.fork.enable = TRUE,
+        future.rng.onMisuse = "ignore") ## automatically set in RStudio
+
+supportsMulticore()
+detectCores()
+
+# Tutorial 3
+
+# R references
 r_refs(file = "r-references.bib")
@@
 Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT $>$ t) and the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t), and not on the probability mass function and the cumulative distribution function.
+Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function p(t) = h(t).S(t-1) and the cumulative distribution function F(t) = 1-S(t).
 The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the hazard function, which assesses the unique risk associated with each time bin, the survivor function cumulates the bin-by-bin risks of event *non*occurrence.
 For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
@@
 
-We used `r my_r_citation$r` for all reported analyses. Web links are printed in bold.
+We used `r my_r_citation$r` for all reported analyses.
+Web links are printed in bold.
+The content of this tutorial is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 
-  However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-
 (ref:life-table-caption) The life table for the blank prime condition of participant 6.
@@
 
+   However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+
   These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the data set used in the first tutorial.
+When you want to study how hazard depends on various predictors, you can fit regression models to the data [@singerAppliedLongitudinalData2003].
+There are two analytic decisions one has to make.
+First, one has to select an analysis time range, i.e., a contiguous set of bins for which there is enough data for each participant.
+Second, one can choose the logit link function which transforms a (hazard) probability into the log of the odds ratio, or the complementary log-log (cloglog) link function, which yields the logarithm of the negated logarithm of the probability of event *non*occurrence.
+An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
+The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
+Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
+One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one.
 
-model formula 
-analytic decisions: specification of the effect of TIME,
+An example discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
 
-assumptions:
-- linearity: continuous variables (TIME, trial number) 
-- additivity:
-- proportionality:
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-1) + $\alpha$~3~(TIME-1)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-1)].
+
+The main predictor variable TIME is the time bin index t that is centered on value 1 in this example.
+The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor.
+Predictors can be discrete, continuous, and time-varying or time-invariant.
+For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
+However, if the predictor interacts linearly with time (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 1 by $\beta$~2~ cloglog-hazard units (when TIME-1 = 0), in bin 2 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-1 = 1), and so forth.
+To interpret the effects of the predictors, the parameter estimates are exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+
+In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (see Tutorial 1); @allisonSurvivalAnalysisUsing2010).
+When there is clustering in the data, as in the case of a small-*N* design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models [@allisonSurvivalAnalysisUsing2010].
+
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the masked response priming data set used in the first tutorial. 
+In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
+
+First, we select the analysis range (200,600] and the cloglog link, and use a polynomial to specify the effect of TIME in the "blank" prime condition.
+Second, based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] and because cognition is likely the behavior of a non-linear dynamical system [ref], we relax all three assumptions, as follows:
+
+\scriptsize
+```{r prepare-data, echo=T}
+# load data
+ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
+# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
+ptb_data <- ptb_data %>% filter(period > 5)
+# create factor condition, with "blank" as the reference level
+ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+# center discrete TIME (period) on bin 9, and trial on trial 1000
+ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
+                                trial_c = (trial - 1000)/1000)
+# remove unnecessary columns before fitting a model
+ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
+```
+
+```{r set-priors, echo=T}
+priors <- c(
+  set_prior("normal(0, 1)", class = "b"), # for beta parameters 
+  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale
+  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
+  set_prior("lkj(2)", class = "cor") # for correlations between RE
+)
+```
+
+```{r fit-model, echo=T}
+plan(multicore)
+model_full_RE <-
+   brm(data = ptb_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + 
+                           condition*period_9*trial_c + 
+                           condition*I(period_9^2) + 
+                           condition*I(period_9^3) +
+                           (1 + condition*period_9*trial_c +
+                           condition*I(period_9^2) +
+                           condition*I(period_9^3) | pid),
+       prior = priors,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "../Tutorial_2_Bayesian/models/model_full_RE")
+```
+\normalsize
+
+To test whether (centered) trial number affects behavior, we fit a model without the variable trial_c.
+
+Use WAIC to compare models.
+
+Plot the effects of congruent and incongruent for each time bin for the selected model.
+
+Plot the model-based hazard and survivor functions.
 

### 2024-08-07 manuscript/manuscript.tex
update manuscript files

@@
 Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time survivor function S(t) = P(RT \(>\) t) and the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t), and not on the probability mass function and the cumulative distribution function.
+Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t) and the discrete-time survivor function S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]}, and not on the probability mass function p(t) = h(t).S(t-1) and the cumulative distribution function F(t) = 1-S(t).
 The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the hazard function, which assesses the unique risk associated with each time bin, the survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence.
 For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{citr} (Version 0.3.2; Aust, 2019), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
+The content of this tutorial is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
 
@@
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-When plotting the data, some warning messages will be generated, like these:
+When creating the plots using our function plot\_eha(), some warning messages will likely be generated, like these:
 
@@
 
-For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.22 for the blank, congruent, and incongruent prime conditions, respectively.
+For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
 In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
@@
 In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-
-However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.07 for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 \midrule
-0.00 & 219.00 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
-40.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-80.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-120.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-160.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-200.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-240.00 & 219.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-280.00 & 219.00 & 7.00 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
-320.00 & 212.00 & 13.00 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
-360.00 & 199.00 & 26.00 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
-400.00 & 173.00 & 40.00 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
-440.00 & 133.00 & 48.00 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
-480.00 & 85.00 & 37.00 & 0.44 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
-520.00 & 48.00 & 32.00 & 0.67 & 0.07 & 0.07 & 0.02 & 1.00 & 0.00\\
-560.00 & 16.00 & 9.00 & 0.56 & 0.12 & 0.03 & 0.01 & 1.00 & 0.00\\
-600.00 & 7.00 & 4.00 & 0.57 & 0.19 & 0.01 & 0.01 & 1.00 & 0.00\\
+0.00 & 220.00 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
+40.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+80.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+120.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+160.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+200.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+240.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+280.00 & 220.00 & 7.00 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
+320.00 & 213.00 & 13.00 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
+360.00 & 200.00 & 26.00 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
+400.00 & 174.00 & 40.00 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
+440.00 & 134.00 & 48.00 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
+480.00 & 86.00 & 37.00 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
+520.00 & 49.00 & 32.00 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00\\
+560.00 & 17.00 & 9.00 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00\\
+600.00 & 8.00 & 4.00 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00\\
 \bottomrule
@@
 
+However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.06 for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
+
 These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the data set used in the first tutorial.
+When you want to study how hazard depends on various predictors, you can fit regression models to the data (Singer \& Willett, 2003).
+There are two analytic decisions one has to make.
+First, one has to select an analysis time range, i.e., a contiguous set of bins for which there is enough data for each participant.
+Second, one can choose the logit link function which transforms a (hazard) probability into the log of the odds ratio, or the complementary log-log (cloglog) link function, which yields the logarithm of the negated logarithm of the probability of event \emph{non}occurrence.
+An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
+The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
+Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
+One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one.
+
+An example discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
+
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-1) + \(\alpha\)\textsubscript{3}(TIME-1)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-1){]}.
+
+The main predictor variable TIME is the time bin index t that is centered on value 1 in this example.
+The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor.
+Predictors can be discrete, continuous, and time-varying or time-invariant.
+For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units.
+However, if the predictor interacts linearly with time (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 1 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-1 = 0), in bin 2 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-1 = 1), and so forth.
+To interpret the effects of the predictors, the parameter estimates are exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+
+In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (see Tutorial 1); Allison (2010)).
+When there is clustering in the data, as in the case of a small-\emph{N} design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models (Allison, 2010).
+
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the masked response priming data set used in the first tutorial.
+In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
+
+First, we select the analysis range (200,600{]} and the cloglog link, and use a polynomial to specify the effect of TIME in the ``blank'' prime condition.
+Second, based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) and because cognition is likely the behavior of a non-linear dynamical system {[}ref{]}, we relax all three assumptions, as follows:
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# load data}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
+\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{)}
+\CommentTok{\# create factor condition, with "blank" as the reference level}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
+\CommentTok{\# center discrete TIME (period) on bin 9, and trial on trial 1000}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{period\_9 =}\NormalTok{ period }\SpecialCharTok{{-}} \DecValTok{9}\NormalTok{,}
+                                \AttributeTok{trial\_c =}\NormalTok{ (trial }\SpecialCharTok{{-}} \DecValTok{1000}\NormalTok{)}\SpecialCharTok{/}\DecValTok{1000}\NormalTok{)}
+\CommentTok{\# remove unnecessary columns before fitting a model}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period)) }\CommentTok{\# 12840 obs.
+of 5 variables}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{priors }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{), }\CommentTok{\# for beta parameters }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"student\_t(7.61, 0, 1.57)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{), }\CommentTok{\# flat prior for intercept on hazard scale}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{), }\CommentTok{\# for standard deviation of RE}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{) }\CommentTok{\# for correlations between RE}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plan}\NormalTok{(multicore)}
+\NormalTok{model\_full\_RE }\OtherTok{\textless{}{-}}
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ ptb\_data,}
+       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
+       \AttributeTok{file =} \StringTok{"../Tutorial\_2\_Bayesian/models/model\_full\_RE"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+To test whether (centered) trial number affects behavior, we fit a model without the variable trial\_c.
+
+Use WAIC to compare models.
+
+Plot the effects of congruent and incongruent for each time bin for the selected model.
+
+Plot the model-based hazard and survivor functions.
 
@@
 
+\bibitem[\citeproctext]{ref-R-RJ-2021-048}
+Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+
+\bibitem[\citeproctext]{ref-R-brms_a}
+Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\bibitem[\citeproctext]{ref-R-brms_b}
+Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411. \url{https://doi.org/10.32614/RJ-2018-017}
+
+\bibitem[\citeproctext]{ref-R-brms_c}
+Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with {brms} and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}
+
+\bibitem[\citeproctext]{ref-R-Rcpp_b}
+Eddelbuettel, D., \& Balamuta, J.
+J.
+(2018).
+{Extending {R} with {C++}: A Brief Introduction to {Rcpp}}.
+\emph{The American Statistician}, \emph{72}(1), 28--36.
+\url{https://doi.org/10.1080/00031305.2017.1375990}
+
+\bibitem[\citeproctext]{ref-R-Rcpp_a}
+Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. \url{https://doi.org/10.18637/jss.v040.i08}
+
+\bibitem[\citeproctext]{ref-R-cmdstanr}
+Gabry, J., Češnovar, R., Johnson, A., \& Bronder, S. (2024). \emph{Cmdstanr: R interface to 'CmdStan'}. Retrieved from \url{https://github.com/stan-dev/cmdstanr}
+
+\bibitem[\citeproctext]{ref-R-bayesplot}
+Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in bayesian workflow. \emph{J. R. Stat. Soc. A}, \emph{182}, 389--402. \url{https://doi.org/10.1111/rssa.12378}
+
+\bibitem[\citeproctext]{ref-R-standist}
+Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
+
 \bibitem[\citeproctext]{ref-R-lubridate}
@@
 
+\bibitem[\citeproctext]{ref-R-tidybayes}
+Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
+
 \bibitem[\citeproctext]{ref-kelsoOutlineGeneralTheory2013}
@@
 
+\bibitem[\citeproctext]{ref-kurzAppliedLongitudinalDataAnalysis2023}
+Kurz, A. S. (2023a). \emph{Applied longitudinal data analysis in brms and the tidyverse} (version 0.0.3). Retrieved from \url{https://bookdown.org/content/4253/}
+
+\bibitem[\citeproctext]{ref-kurzStatisticalRethinkingSecondEd2023}
+Kurz, A. S. (2023b). \emph{Statistical rethinking with brms, ggplot2, and the tidyverse: {Second} edition} (version 0.4.0). Retrieved from \url{https://bookdown.org/content/4857/}
+
 \bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
 Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
+\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2018}
+McElreath, R. (2018). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {Stan}} (1st ed.). {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9781315372495}
 
@@
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
-Singer, J. D., \& Willett, J. B. (2003). \emph{Applied longitudinal data analysis: Modeling change and event occurrence}. Oxford ; New York: Oxford University Press.
+Singer, J. D., \& Willett, J. B. (2003). \emph{Applied {Longitudinal Data Analysis}: {Modeling Change} and {Event Occurrence}}. Oxford, New York: Oxford University Press.
 

### 2024-08-07 Tutorial_2_Bayesian/Tutorial_2.Rmd
add models

@@
 priors <- c(
-  set_prior("normal(0, 1)", class = "b"), # for beta parameters 
-  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale
+  set_prior("normal(0, 1)", class = "b"), # for beta parameters when using logit link
+
+  #set_prior("skew_normal(-0.5, .95,-4)", class = "b"), # for beta parameters when using cloglog link
+
+  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale when using logit link;
+
+  #set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),   # weakly regularizing prior for intercept when using cloglog link (hazard values below .5 more likely than values above .5)
   set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
@@
 )
+
+inv_cloglog <- function(c){
+  1-exp(-exp(c))
+}
+cloglog <- function(p){
+  log(-log(1-p))
+}
+
+dat <- 
+  tibble(c1 = rnorm(1e5,0,1), # usa as weakly reguralizing prior for beta parameters (when using logit link)
+         c2 = rt(1e5,7.61)* 1.57) %>% # use as weakly reguralizing prior for interceptparameter (when using logit 
+  mutate(p1 = inv_logit_scaled(c1),
+         p2 = inv_logit_scaled(c2))
+
+plot(density(dat$c1))
+plot(density(dat$c2))
+plot(density(dat$p1))
+plot(density(dat$p2))
+
+# cloglog
+dat2 <- tibble(c1 = rnorm(1e5,-1,1.1), # cloglog intercept
+              c2 = rskew_normal(1e5, -0.5, sigma = .95, alpha = -4)) %>% # beta toward .6
+
+  mutate(p1 = inv_cloglog(c1),
+         p2 = inv_cloglog(c2))
+
+plot(density(dat2$c1))
+plot(density(dat2$c2))
+plot(density(dat2$p1))
+plot(density(dat2$p2))
+
 ```
@@
 
-model_full_RE <-
+model_full_RE_newprior <-
    brm(data = ptb_data,
@@
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_full_RE")
+       file = "Tutorial_2_Bayesian/models/model_full_RE_newprior")
 
@@
 
-Model_full_RE took 3 hours to run.
+Model_full_RE took between 3 and 4 hours to run.
 
@@
 
-```{r full-random-effects-model}
+Fit a model that relaxes proportionality assumption.
+
+```{r}
+# remove unnecessary columns before fitting a model
+ptb_data <- ptb_data %>% select(-c(trial_c)) # 12840 obs. of 4 variables
+head(ptb_data,n=17)
+summary(ptb_data)
+```
+
+```{r cubic-random-effects-model}
 plan(multicore)
@@
 
-```{r check-model-full-RE}
+```{r check-model-cubic-RE}
 summary(model_cubic_RE)
+```
+
+#### relax all ass
+
+```{r cubictime-quadtrial-model}
+plan(multicore)
+
+model_cubictime_quadtrial <-
+   brm(data = ptb_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + 
+                           condition*period_9*trial_c + 
+                           condition*period_9*I(trial_c^2) + 
+                           condition*I(period_9^2)*trial_c + 
+                           condition*I(period_9^2)*I(trial_c^2) +
+                           condition*I(period_9^3) +
+                      (1 + condition*period_9*trial_c +
+                           condition*period_9*I(trial_c^2) +
+                           condition*I(period_9^2)*trial_c +
+                           condition*I(period_9^2)*I(trial_c^2) +
+                           condition*I(period_9^3) | pid),
+       prior = priors,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_cubictime_quadtrial")
+
+# the formula above comes down to the following effects:
+# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
+# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
+# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
+# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+ ##effect of incongruent interacts cubicly with TIME
+# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
+# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
+# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
 ```
@@
 ```{r}
-
+model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
+model_full_RE_newprior <- readRDS("Tutorial_2_Bayesian/models/model_full_RE_newprior.rds") 
+fixef(model_full_RE)
+fixef(model_full_RE_newprior)
 ```
@@
 ```{r load-model}
-model_cubic_RE <- readRDS("Tutorial_2_Bayesian/models/model_cubic_RE.rds") 
+model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
 ```
@@
 ```{r make-plot-effects}
-post <-as_draws_df(model_cubic_RE) %>% # 8000 draws x 299 variables
+post <-as_draws_df(model_full_RE) %>% # 8000 draws x 299 variables
    select(starts_with("b_")) %>%       # 8000 x 18
@@
                      limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
   labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
@@
                      limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
   labs(x = "time bin", y = "effect of incongruent") +

### 2024-08-27 manuscript/manuscript.Rmd
comment out model with wrong priors temporarily

@@
 ```{r fit-model, echo=T}
-plan(multicore)
-model_full_RE <-
-   brm(data = ptb_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + 
-                           condition*period_9*trial_c + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           (1 + condition*period_9*trial_c +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) | pid),
-       prior = priors,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "../Tutorial_2_Bayesian/models/model_full_RE")
+#plan(multicore)
+#model_full_RE <-
+#   brm(data = ptb_data,
+#       family = binomial(link="cloglog"),
+#       event | trials(1) ~ 0 + Intercept + 
+#                           condition*period_9*trial_c + 
+#                           condition*I(period_9^2) + 
+#                           condition*I(period_9^3) +
+#                           (1 + condition*period_9*trial_c +
+#                           condition*I(period_9^2) +
+#                           condition*I(period_9^3) | pid),
+#       prior = priors,
+#       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+#       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+#       seed = 12, init = "0",
+#       file = "../Tutorial_2_Bayesian/models/model_full_RE")
 ```

### 2024-08-27 manuscript/manuscript.tex
update refs and manuscript

@@
 survival, hazard, duration, failure-time, and transition analysis (Allison, 1982, 2010; Singer \& Willett, 2003).
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+Across four tutorials that we make publicly available on Github and the Open Science Framework (OSF), we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 \begin{Highlighting}[]
-\FunctionTok{plan}\NormalTok{(multicore)}
-\NormalTok{model\_full\_RE }\OtherTok{\textless{}{-}}
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ ptb\_data,}
-       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
-       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"../Tutorial\_2\_Bayesian/models/model\_full\_RE"}\NormalTok{)}
+\CommentTok{\#plan(multicore)}
+\CommentTok{\#model\_full\_RE \textless{}{-}}
+\CommentTok{\#   brm(data = ptb\_data,}
+\CommentTok{\#       family = binomial(link="cloglog"),}
+\CommentTok{\#       event | trials(1) \textasciitilde{} 0 + Intercept + }
+\CommentTok{\#                           condition*period\_9*trial\_c + }
+\CommentTok{\#                           condition*I(period\_9\^{}2) + }
+\CommentTok{\#                           condition*I(period\_9\^{}3) +}
+\CommentTok{\#                           (1 + condition*period\_9*trial\_c +}
+\CommentTok{\#                           condition*I(period\_9\^{}2) +}
+\CommentTok{\#                           condition*I(period\_9\^{}3) | pid),}
+\CommentTok{\#       prior = priors,}
+\CommentTok{\#       chains = 4, cores = 4, iter = 3000, warmup = 1000,}
+\CommentTok{\#       control = list(adapt\_delta = 0.999, step\_size = 0.04, max\_treedepth = 12),}
+\CommentTok{\#       seed = 12, init = "0",}
+\CommentTok{\#       file = "../Tutorial\_2\_Bayesian/models/model\_full\_RE")}
 \end{Highlighting}

### 2024-08-27 Tutorial_2_Bayesian/Tutorial_2.Rmd
add figure with link functions

@@
 ```
+
+# Plot the logit and complementary log-log (cloglog) link functions
+
+```{r plot-links}
+probability <- (1:99999)/100000
+logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
+logit    <- function(x) { return( log(x/(1-x)) )}
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+cloglog    <- function(x) { return( log(-1*log(1-x)) )}
+
+cloglog_prob <- cloglog(probability)
+logit_prob <- logit(probability)
+dataplot <- cbind(probability,cloglog_prob,logit_prob)
+
+ggplot() +
+  geom_hline(yintercept=0, color="white") +
+  geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
+  geom_line(data=dataplot,aes(y=cloglog_prob,x=probability,colour="cloglog"),linewidth=1) +
+  scale_color_manual(name = "Link function:", values = c("logit" = "darkblue", "cloglog" = "red")) +
+  geom_vline(xintercept = logistic(0), linetype="dotted", linewidth = 0.3) +   
+  geom_vline(xintercept = inverse_cloglog(0), linetype="dotted", linewidth = 0.3) +
+  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90) +
+  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90) +
+  scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
+  labs(x = "Probability",
+        y = "logit or cloglog scale") +
+  theme(panel.grid = element_blank(),
+        axis.text=element_text(size=12),
+        axis.title=element_text(size=14),
+      #  legend.position="top",
+        legend.text = element_text(size=14),
+      legend.position = "inside",
+        legend.position.inside=c(.2,.75),
+      legend.background = element_rect(fill="white") )
+```
+
+```{r save-plot-links}
+ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 12, height = 10, dpi = 600)
+```
+
+# Visualize different prior distributions on the logit and cloglog scales
 

### 2024-08-27 Tutorial_2_Bayesian/Tutorial_2.Rmd
update prior distributions

@@
 # Visualize different prior distributions on the logit and cloglog scales
+
+To gain a sense of what prior logit values would approximate a uniform distriubution on the probability scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
+Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function).
+
+## Simulate, convert, and fit
+
+## Visualize prior distributions
+
+```{r}
+logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
+logit    <- function(x) { return( log(x/(1-x)) )}
+invcloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+cloglog    <- function(x) { return( log(-1*log(1-x)) )}
+
+pr1 <- tibble(prior = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  mutate(p = logistic(prior)) %>% 
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-13, y=60000, label="N(0,4)",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l1 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("logistic") +
+  theme(panel.grid = element_blank())
+
+c1 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 4)) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("inverse cloglog")+
+  theme(panel.grid = element_blank())
+
+pr2 <- tibble(prior = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-6, y=60000, label="N(0,2)",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l2 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c2 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 2)) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# flat prior for logistic
+pr3 <- tibble(prior = rt(1e6, df = 7.61)* 1.57) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-4, y=100000, label="t(7.61)*1.57",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l3 <- tibble(log_odds = rt(1e6, df = 7.61)* 1.57) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c3 <- tibble(cloglog_prob = rt(1e6, df = 7.61)* 1.57) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  #geom_histogram(bins = 50) +
+  #scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# prior to .5
+pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3, y=60000, label="N(0,1)",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l4 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 1))%>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  geom_vline(xintercept=logistic(0), color="red") +
+  theme(panel.grid = element_blank())
+
+c4 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 1)) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+ # geom_histogram(bins = 50) +
+#  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# flat on cloglog
+pr5 <- tibble(prior = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3, y=60000, label="skew_N(-0.59,1.26,-4.20)",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l5 <- tibble(log_odds = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20))%>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+#  geom_histogram(bins = 50) +
+#  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c5 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# regu to invcloglog(0)
+pr6 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .7, alpha = -2.20)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-1, y=60000, label="skew_N(-0.2,.7,-2.20)",
+              color="red", size=6) +
+  theme(panel.grid = element_blank())
+
+l6 <- tibble(log_odds = rskew_normal(1e6, mu=-0.2, sigma = .7, alpha = -2.20))%>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+ # geom_histogram(bins = 50) +
+#  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c6 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .7, alpha = -2.20)) %>% 
+  mutate(p = invcloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  geom_vline(xintercept=invcloglog(0), color = "red") +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6)
+
+```
+
+```{r}
+ggsave("figures/plot_of_priors.png", width=9, height=10,dpi=300)
+```
+
+The log-odds Normal(0,1) gently regularizes p towards .5, but still allows for stronger values. This might be a good prior to use for the beta parameters.
+
+N(0,4) make no sense for logit-hazard models.
+
+Compare three standard deviations.
+
+The log-odds Normal(0,1) gently regularizes p towards .5, but still allows for stronger values. This might be a good prior to use for the beta parameters.
+
+The alpha parameters tend to drift toward the lower end of the probability range in general (i.e., they are typically below .5 in hazard space, also for RT studies).
+The log-odds Normal(0, 1.5) prior is nearly flat in probability space, but it does still push the mass away from the boundaries.
+
+```{r}
+set.seed(11)
+
+# generate draws from U(0,1)
+dat <- 
+  tibble(p = runif(1e5, 0, 1)) %>% 
+  mutate(g = logit(p),
+         c = cloglog(p)) 
+# display
+dat %>%   
+  ggplot(aes(x = c)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+plot(density(dat$c))
+
+# fit models
+fit11.11 <-
+  brm(data = dat,
+      family = student,
+      g ~ 1,
+      chains = 4, cores = 4,
+      file = "models/fit11.11")
+
+fit11.11b <-
+  brm(data = dat,
+      family = skew_normal(),
+      c ~ 1,
+      chains = 4, cores = 4,
+      file = "models/fit11.11b")
+
+fit11.11b2 <-
+  brm(data = dat,
+      family = student(),
+      c ~ 1,
+      chains = 4, cores = 4,
+      file = "models/fit11.11b2")
+
+```
+
+```{r}
+print(fit11.11)
+print(fit11.11b)
+print(fit11.11b2)
+
+pp_check(fit11.11b)
+```
+
+Now we can reverse the process. Here’s what it would look like if we simulated from the Student  
+t-distribution based on those posterior means and then converted the results into the probability metric.
+
+```{r}
+set.seed(11)
+
+# cloglog
+#library(sn)
+
+tibble(c = rskew_normal(1e8, mu=-0.59, sigma = 1.26, alpha = -4.20) ) %>% # simulate from skew-normal #distribution !!!!!!!! for intercept 
+#  !!!!!!!!!
+
+#tibble(c = rsn(1e6, xi=-1, omega = 1, alpha = -2) ) %>% # simulate from skew-normal distribution
+
+   mutate(p = invcloglog(c)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+```
 

### 2024-08-30 Tutorial_2_Bayesian/Tutorial_2.Rmd
update Tutorial 2

--- 2024-08-27 Tutorial_2_Bayesian/Tutorial_2.Rmd
@@
 
-In this file we build a Bayesian hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions.
+In this file we build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials.
 
@@
 ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-summary(ptb_data)
 
-# center discrete TIME (period) on bin 9, and trial on trial 1000
-ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
-                                trial_c = (trial - 1000)/1000)
+# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.
+ptb_data <- ptb_data %>% 
+        mutate(period_9 = period - 9,
+               trial_c = (trial - 1000)/1000,
+               d6  = if_else(period == 6, 1, 0),
+               d7  = if_else(period == 7, 1, 0),
+               d8  = if_else(period == 8, 1, 0),
+               d9  = if_else(period == 9, 1, 0),
+               d10 = if_else(period == 10, 1, 0),
+               d11 = if_else(period == 11, 1, 0),
+               d12 = if_else(period == 12, 1, 0),
+               d13 = if_else(period == 13, 1, 0),
+               d14 = if_else(period == 14, 1, 0),
+               d15 = if_else(period == 15, 1, 0))
 
-# remove unnecessary columns before fitting a model
-ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
 head(ptb_data,n=17)
@@
 
-# Set up priors, and fit the model with full random effects structure.
+To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
+Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.
 
-```{r priors}
-priors <- c(
-  set_prior("normal(0, 1)", class = "b"), # for beta parameters when using logit link
+## Simulate, convert, and fit
 
-  #set_prior("skew_normal(-0.5, .95,-4)", class = "b"), # for beta parameters when using cloglog link
+```{r simulate-convert}
+set.seed(11)
 
-  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale when using logit link;
+logit    <- function(x) { return( log(x/(1-x)) )}
+cloglog  <- function(x) { return( log(-1*log(1-x)) )}
 
-  #set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),   # weakly regularizing prior for intercept when using cloglog link (hazard values below .5 more likely than values above .5)
-  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
-  set_prior("lkj(2)", class = "cor") # for correlations between RE
-)
+# generate draws from U(0,1) and convert
+dat <- 
+  tibble(p = runif(1e6, 0, 1)) %>% 
+  mutate(g = logit(p),
+         c = cloglog(p)) 
+# display
+dat %>%   
+  ggplot(aes(x = c)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
 
-inv_cloglog <- function(c){
-  1-exp(-exp(c))
-}
-cloglog <- function(p){
-  log(-log(1-p))
-}
-
-dat <- 
-  tibble(c1 = rnorm(1e5,0,1), # usa as weakly reguralizing prior for beta parameters (when using logit link)
-         c2 = rt(1e5,7.61)* 1.57) %>% # use as weakly reguralizing prior for interceptparameter (when using logit 
-  mutate(p1 = inv_logit_scaled(c1),
-         p2 = inv_logit_scaled(c2))
-
-plot(density(dat$c1))
-plot(density(dat$c2))
-plot(density(dat$p1))
-plot(density(dat$p2))
-
-# cloglog
-dat2 <- tibble(c1 = rnorm(1e5,-1,1.1), # cloglog intercept
-              c2 = rskew_normal(1e5, -0.5, sigma = .95, alpha = -4)) %>% # beta toward .6
-
-  mutate(p1 = inv_cloglog(c1),
-         p2 = inv_cloglog(c2))
-
-plot(density(dat2$c1))
-plot(density(dat2$c2))
-plot(density(dat2$p1))
-plot(density(dat2$p2))
-
+plot(density(dat$c))
 ```
 
-```{r full-random-effects-model}
+```{r fit-skew-normal}
+# fit models
+fit_skewN <-
+  brm(data = dat,
+      family = skew_normal(),
+      c ~ 1,
+      chains = 4, cores = 4,
+      file = "Tutorial_2_Bayesian/models/fit_skewN")
+```
+
+```{r}
+summary(fit_skewN) 
+```
+
+Now we can reverse the process.
+We simulate from the skew-Normal distribution based on the posterior means for mu, sigma, and alpha, and then convert the results into the probability (i.e., discrete-time hazard) metric.
+
+```{r check-results}
+set.seed(11)
+
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+
+tibble(c = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22) ) %>% 
+  mutate(p = inverse_cloglog(c)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+```
+
+## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale
+
+```{r plot-priors}
+logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
+logit    <- function(x) { return( log(x/(1-x)) )}
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+cloglog    <- function(x) { return( log(-1*log(1-x)) )}
+
+set.seed(23)
+
+# A N(0,4) prior on the logit and cloglog scales pushes mass to probabilities of 0 and 1
+pr1 <- tibble(prior = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-13, y=60000, label="N(0,4)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l1 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("logistic(prior)") +
+  theme(panel.grid = element_blank())
+
+c1 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 4)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("inverse_cloglog(prior)")+
+  theme(panel.grid = element_blank())
+
+# A N(0,2) prior on the logit and cloglog scales pushes mass to probabilities of 0 and/or 1
+pr2 <- tibble(prior = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-6, y=60000, label="N(0,2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l2 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c2 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# A student-t(df=7.61) prior with scale 1.57 on the logit scale approximates a uniform distribution on the probability scale.
+This might be a good prior to use for the alpha parameters or Intercept in a logit-hazard model.
+pr3 <- tibble(prior = rt(1e6, df = 7.61)* 1.57) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-12, y=140000, label="t(7.61)*1.57",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l3 <- tibble(log_odds = rt(1e6, df = 7.61)* 1.57) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c3 <- ggplot()
+
+# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. This might be a good prior to use for the beta parameters in a logit-hazard model.
+pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3, y=60000, label="N(0,1)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'D', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l4 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 1))%>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  geom_vline(xintercept=logistic(0), color="red") +
+  theme(panel.grid = element_blank())
+
+c4 <- ggplot()
+
+# A skew_Normal(-0.59,1.26,-4.22) prior on the cloglog scale approxiates a uniform distr.
+on the hazard scale.
+This uninformative prior might be good for the alpha parameters or Intercept in a cloglog-hazard model.
+pr5 <- tibble(prior = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-5.5, y=55000, label="skew_N(-0.59,1.26,-4.22)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'E', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l5 <- ggplot()
+
+c5 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# The skew_Normal(-1,1,-2) on the cloglog scale is a weakly informative prior for the alpha parameters or Intercept in a cloglog-hazard model because hazard values below .5 more likely than values above .5 in general.
+pr6 <- tibble(prior = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-5, y=55000, label="skew_N(-1,1,-2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'F', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l6 <- ggplot()
+
+c6 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# A skew_Normal(-0.2,0.71,-2.2) on the cloglog scale gently regularizes p towards .6321. This might be a good prior to use for the beta parameters in a cloglog-hazard model.
+pr7 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3.2, y=55000, label="skew_N(-0.2,.71,-2.2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'G', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l7 <- ggplot()
+
+c7 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  geom_vline(xintercept=invcloglog(0), color = "red") +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) / (l7 + pr7 + c7)
+```
+
+```{r save-plot-priors}
+ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
+```
+
+# Fit 4 cloglog-hazard models.
+
+## Model M1: general specification of TIME in the baseline condition (blank prime), and main effects of prime type, and trial number.
+
+### Prepare data for M1
+
+```{r data-M1}
+# remove unnecessary columns before fitting a model
+M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9)) # 12840 obs. of 14 variables
+head(M1_data)
+summary(M1_data)
+```
+
+### Set up priors for M1
+
+```{r priors-cloglog-M1}
+priors_M1 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"), # weakly informative prior for alpha parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d9"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"), 
+  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
+)
+```
+
+### Fit model M1
+
+```{r fit-model-M1}
 plan(multicore)
 
-model_full_RE_newprior <-
-   brm(data = ptb_data,
+model_M1 <-
+   brm(data = M1_data,
        family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + 
-                           condition*period_9*trial_c + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           (1 + condition*period_9*trial_c +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) | pid),
-       prior = priors,
+       event | trials(1) ~ 0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c +
+                          (0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+       prior = priors_M1,
        chains = 4, cores = 4, iter = 3000, warmup = 1000,
@@
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_full_RE_newprior")
-
-# the formula above comes down to the following effects:
-# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
-# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
-# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
-# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+ ##effect of incongruent interacts cubicly with TIME
-# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
-# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
-# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
+       file = "Tutorial_2_Bayesian/models/model_M1")
 ```
 
-Model_full_RE took between 3 and 4 hours to run.
+Model_M1 took about 96 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
-```{r check-model-full-RE}
-summary(model_full_RE)
+```{r check-model-M1}
+summary(model_M1)
 ```
+
+## Model M2: third-order polynomical specification of TIME in the baseline condition (blank prime), and main effects of prime types, and trial number.
+
+### Prepare data for M2
+
+```{r data-M2}
+# remove unnecessary columns before fitting a model
+M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
+head(M2_data)
+```
+
+### Set up priors for M2
+
+```{r priors-cloglog-M2}
+priors_M2 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
+)
+```
+
+### Fit model M2
+
+```{r fit-model-M2}
+plan(multicore)
+
+model_M2 <-
+   brm(data = M2_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c +
+                          (1 + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c | pid),
+       prior = priors_M2,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M2")
+```
+
+Model_M2 took about .... to run.
+
+```{r check-model-M2}
+summary(model_M2)
+```
+
+## Model M3: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax proportionality assumption for prime types, and trial number.
+
+### Prepare data for M3
+
+```{r data-M3}
+# remove unnecessary columns before fitting a model
+M3_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
+head(M3_data)
+```
+
+### Set up priors for M3
+
+```{r priors-cloglog-M3}
+priors_M3 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
+)
+```
+
+### Fit model M3
+
+```{r fit-model-M3}
+plan(multicore)
+
+model_M3 <- 
+   brm(data = M3_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
+                           condition*period_9 + 
+                           condition*I(period_9^2) + 
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) +
+                           (1 + condition*period_9 +
+                           condition*I(period_9^2) +
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) | pid),
+       prior = priors_M3,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M3")
+```
+
+Model_M3 took about .... to run.
+
+```{r check-model-M3}
+summary(model_M3)
+```
+
+## Model M4: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.
+
+# Compare 4 models
+
+# Plot effects for selected model
+
+##########
 

### 2024-09-10 Tutorial_2_Bayesian/Tutorial_2.Rmd
update Tutorial 2

@@
   geom_histogram(bins = 50) +
-  geom_vline(xintercept=invcloglog(0), color = "red") +
+  geom_vline(xintercept=inverse_cloglog(0), color = "red") +
   scale_y_continuous(NULL, breaks = NULL) +
@@
 # remove unnecessary columns before fitting a model
-M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9)) # 12840 obs. of 14 variables
+M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. of 14 variables
 head(M1_data)
 summary(M1_data)
+
+#M1_data <- within(M1_data, condition <- relevel(condition, ref = "blank"))
+
 ```
@@
   set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"), # weakly informative prior for alpha parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"),# weakly informative prior for alpha parameters when using cloglog link 
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d9"), 
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
@@
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
   set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
@@
        family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c +
-                          (0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+       event | trials(1) ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
+         condition + trial_c +
+
+                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
        prior = priors_M1,
@@
 
-Model_M1 took about 96 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+Model_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
 ```{r check-model-M1}
+
 summary(model_M1)
+
+model_M1c <- readRDS("Tutorial_2_Bayesian/models/model_M1c.rds")
+
+fixef(model_M1)
+
+```
+
+```{r data-M1b}
+# remove unnecessary columns before fitting a model
+M1b_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. of 14 variables
+head(M1b_data)
+summary(M1b_data)
+
+#M1_data <- within(M1_data, condition <- relevel(condition, ref = "blank"))
+
+```
+
+```{r priors-cloglog-M1}
+priors_M1b <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  #set_prior("skew_normal(-1,1,-2)", class = "Intercept"), # weakly informative prior for alpha parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
+  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
+)
+```
+
+```{r fit-model-M1}
+plan(multicore)
+
+model_M1b <-
+   brm(data = M1b_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + 
+         condition + trial_c +
+
+                   (0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+       prior = priors_M1b,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M1b")
+```
+
+117 minutes
+
+```{r}
+summary(model_M1)
+
+fixef(model_M1b)
+
+```
+
+```{r fit-model-M1c}
+plan(multicore)
+
+model_M1c <-
+   brm(data = M1b_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + 
+         condition + trial_c +
+
+                   (1 + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+       prior = priors_M1b,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M1c")
 ```
@@
   set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
   set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
@@
        family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c +
-                          (1 + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c | pid),
+       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c +
+                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c | pid),
        prior = priors_M2,
@@
 
-Model_M2 took about .... to run.
+Model_M2 took about 144 minutes to run.
 
@@
 summary(model_M2)
+model_M2b <- readRDS("Tutorial_2_Bayesian/models/model_M2b.rds")
+fixef(model_M2b)
+```
+
+```{r priors-cloglog-M2b}
+priors_M2b <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
+)
+```
+
+```{r fit-model-M2}
+plan(multicore)
+
+model_M2b <-
+   brm(data = M2_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c +
+                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c | pid),
+       prior = priors_M2,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M2b")
+```
+
+```{r}
+fixef(model_M2b)
 ```
@@
   set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using cloglog link
   set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
@@
 
-Model_M3 took about .... to run.
+Model_M3 took about 268 minutes to run.
 
@@
 
+### Prepare data for M4
+
+```{r data-M4}
+# remove unnecessary columns before fitting a model
+M4_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
+head(M4_data)
+```
+
+### Set up priors for M4
+
+```{r priors-cloglog-M4}
+priors_M4 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using cloglog link
+  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
+  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
+)
+```
+
+### Fit model M4
+
+```{r fit-model-M4}
+plan(multicore)
+
+model_M4 <- 
+   brm(data = M4_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
+                           condition*period_9 + 
+                           condition*I(period_9^2) + 
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) +
+                           I(trial_c^2)*period_9 + 
+                           I(trial_c^2)*I(period_9^2) + 
+                           condition*period_9*trial_c + 
+                           condition*I(period_9^2)*trial_c +
+                           condition*period_9*I(trial_c^2) + 
+                           condition*I(period_9^2)*I(trial_c^2) +
+                           (1 + condition*period_9 +
+                           condition*I(period_9^2) +
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) +
+                           I(trial_c^2)*period_9 + 
+                           I(trial_c^2)*I(period_9^2) + 
+                           condition*period_9*trial_c + 
+                           condition*I(period_9^2)*trial_c +
+                           condition*period_9*I(trial_c^2) + 
+                           condition*I(period_9^2)*I(trial_c^2)  | pid),
+       prior = priors_M4,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M4")
+```
+
+Model_M4 took about  8 hours to run.
+
+```{r check-model-M4}
+summary(model_M4)
+
+fixef(model_M4)[c("conditioncongruent", "conditionincongruent"), -2] %>% exp()
+
+```
+
 # Compare 4 models
 
-# Plot effects for selected model
+```{r}
+model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
+model_M2 <- readRDS("Tutorial_2_Bayesian/models/model_M2.rds")
+model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
+model_M4 <- readRDS("Tutorial_2_Bayesian/models/model_M4.rds")
+```
+
+Using WAIC and LOO for comparing nonnested models.
+
+```{r}
+model_M1  <- add_criterion(model_M1, c("loo", "waic"))
+model_M2  <- add_criterion(model_M2, c("loo", "waic"))
+model_M3  <- add_criterion(model_M3, c("loo", "waic"))
+model_M4  <- add_criterion(model_M4, c("loo", "waic"))
+```
+
+Compare all three models:
+
+```{r}
+loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F)
+loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F)
+```
+
+```{r}
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
+
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
+
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "stacking") %>% round(digits = 3)
+```
+
+# Plot effects for selected model M4
+
+Plot Pareto k estimates
+
+```{r}
+loo(model_M4)$diagnostics %>% 
+  data.frame() %>% 
+  # attach the `id` values
+  bind_cols(M4_data) %>% 
+  mutate(id = 1:n()) %>%
+
+  ggplot(aes(x = id, y = pareto_k)) +
+  geom_point(alpha = 3/4) + 
+  geom_text(data = . %>% filter(pareto_k > .2),
+            aes(x = id + 2, label = id),
+            size = 3, hjust = 0) +
+  theme(panel.grid = element_blank())
+```
+
+# visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trial 1000.
+
+```{r make-plot-effects}
+post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
+   select(starts_with("b_")) %>%       # 8000 x 31
+   expand_grid(period_9 = -3:6) %>%   
+   mutate(period_9sq = period_9^2,
+          period_9cu = period_9^3,
+          trial100 = -900/1000)  
+# effects for trial 1000
+p1congruent <- post %>% 
+  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
+  group_by(period_9) %>%
+
+  ggplot(aes(x = period_9, y = postdistr)) +
+  stat_lineribbon(show.legend=T) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=0, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
+  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90))
+
+p2incongruent <- post %>% 
+  mutate(postdistr = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
+  group_by(period_9) %>%
+
+  ggplot(aes(x = period_9, y = postdistr)) +
+  stat_lineribbon(show.legend=F) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=0, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
+  labs(x = "time bin", y = "effect of incongruent") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90))
+
+p1congruent + p2incongruent
+```
+
+```{r save-plot-effects}
+ggsave("figures/effects_con_incon_trial1000.png", width = 12, height = 16, dpi = 600)
+```
+
+```{r}
+post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
+   select(starts_with("b_")) %>%       # 8000 x 31
+   expand_grid(period_9 = -3:6) %>%   
+   mutate(period_9sq = period_9^2,
+          period_9cu = period_9^3,
+          trial100 = -900/1000)  
+# effects for trial 1000
+plot_facet <- post %>% 
+  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
+         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
+
+  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
+  select(condition, period_9, value) %>%
+ # mutate(value_hr = exp(value)) %>%
+  group_by(period_9) %>%
+
+   ggplot(aes(x = period_9, y = value)) +
+  stat_lineribbon(show.legend=T) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=0, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-8,-6,-4,-2,0,2,4,6,8), limits = c(-8,8)) +
+  labs(x = "time bin", y = "cloglog-hazard") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90),
+        panel.background = element_rect(fill="white", color="black"),
+        legend.position = "top") +
+  facet_wrap(~ condition)
+
+plot_facet
+```
+
+# plot effects on hazard ratio scale
+
+```{r}
+plot_facet_hr <- post %>% 
+  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
+         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
+
+  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
+  select(condition, period_9, value) %>%
+  mutate(value_hr = exp(value)) %>%
+  group_by(period_9) %>%
+
+   ggplot(aes(x = period_9, y = value_hr)) +
+  stat_lineribbon(show.legend=T) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=1, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,7)) +
+  scale_y_continuous( limits = c(0,40)) +
+  labs(x = "time bin", y = "hazard ratio") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top") +
+  facet_wrap(~ condition)
+
+plot_facet_hr
+
+```
+
+# Plot model-based predicted cloglog-hazard for subject 6
+
+```{r}
+make_fitted <- function(fit, scale, ...) {
+
+  fitted(fit,
+         newdata = nd,
+         scale = scale,
+         ...) %>% 
+    data.frame() %>% 
+    bind_cols(nd)
+}
+```
+
+# model_M4: Plot predicted cloglog(hazard) for single subject using fitted() in make_fitted().
+
+```{r}
+# define the `newdata`
+nd <- tibble(pid = 1:6) %>%
+  expand_grid(incon = 0:1,
+                con  = 0:1,
+                trial_c = c(-0.5,0,0.5),
+                period_9 = -3:6) %>%
+  filter(!(con ==1 & incon == 1)) %>%
+  mutate(Iperiod_9E2 = period_9^2,
+         Iperiod_9E3 = period_9^3,
+         Itrial_cE2 = trial_c^2,
+         condition = ifelse(con==1 & incon==0,"congruent",
+                     ifelse(con==0 & incon==1,"incongruent", "blank")),
+         condition = factor(condition, levels=c("blank" ,"congruent","incongruent"))) 
+```
+
+```{r}
+selected_subject = 6 # select pid here (1:6) manually
+
+get_variables(model_M4)
+```
+
+First, on the cloglog-scale:
+
+```{r}
+make_fitted(model_M4, scale = "linear") %>% 
+  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
+         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
+         period = period_9+9) %>% 
+  filter(pid==selected_subject) %>% 
+  # plot!
+  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
+             fill = condition, color = condition)) +
+  geom_ribbon(alpha = 1/5, linewidth = 0) +
+  geom_line() +
+  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
+  ylab("fitted cloglog(hazard)") +
+  coord_cartesian(ylim = c(-5, 1)) +
+  theme(panel.grid = element_blank(),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top",
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~ trial)
+```
+
+Same on hazard scale
+
+```{r}
+make_fitted(model_M4, scale = "response") %>% 
+  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
+         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
+         period = period_9+9) %>% 
+  filter(pid==selected_subject) %>% 
+  # plot!
+  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
+             fill = condition, color = condition)) +
+  geom_ribbon(alpha = 1/5, linewidth = 0) +
+  geom_line() +
+  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
+  ylab("fitted hazard") +
+  coord_cartesian(ylim = c(0, 1)) +
+  theme(panel.grid = element_blank(),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top",
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~ trial)
+```
 

### 2024-09-12 Tutorial_2_Bayesian/Tutorial_2.Rmd
update tutorial 3

@@
 
-In this file we build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials.
+In this file we plot priors for the logit and cloglog link functions, build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials for the selected model.
 
@@
 # remove unnecessary columns before fitting a model
-M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. of 14 variables
+M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. 
 head(M1_data)
 summary(M1_data)
-
-#M1_data <- within(M1_data, condition <- relevel(condition, ref = "blank"))
-
 ```
@@
 ```{r check-model-M1}
-
 summary(model_M1)
-
-model_M1c <- readRDS("Tutorial_2_Bayesian/models/model_M1c.rds")
-
 fixef(model_M1)
-
-```
-
-```{r data-M1b}
-# remove unnecessary columns before fitting a model
-M1b_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. of 14 variables
-head(M1b_data)
-summary(M1b_data)
-
-#M1_data <- within(M1_data, condition <- relevel(condition, ref = "blank"))
-
-```
-
-```{r priors-cloglog-M1}
-priors_M1b <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  #set_prior("skew_normal(-1,1,-2)", class = "Intercept"), # weakly informative prior for alpha parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
-  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
-)
-```
-
-```{r fit-model-M1}
-plan(multicore)
-
-model_M1b <-
-   brm(data = M1b_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + 
-         condition + trial_c +
-
-                   (0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
-       prior = priors_M1b,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M1b")
-```
-
-117 minutes
-
-```{r}
-summary(model_M1)
-
-fixef(model_M1b)
-
-```
-
-```{r fit-model-M1c}
-plan(multicore)
-
-model_M1c <-
-   brm(data = M1b_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + 
-         condition + trial_c +
-
-                   (1 + d6 + d7 + d8 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
-       prior = priors_M1b,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M1c")
 ```
@@
 summary(model_M2)
-model_M2b <- readRDS("Tutorial_2_Bayesian/models/model_M2b.rds")
-fixef(model_M2b)
-```
-
-```{r priors-cloglog-M2b}
-priors_M2b <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
-  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
-)
-```
-
-```{r fit-model-M2}
-plan(multicore)
-
-model_M2b <-
-   brm(data = M2_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c +
-                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c | pid),
-       prior = priors_M2,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M2b")
-```
-
-```{r}
-fixef(model_M2b)
+fixef(model_M2)
 ```
@@
        event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
-                           condition*period_9 + 
-                           condition*I(period_9^2) + 
+                          condition*period_9*trial_c + 
+                          condition*period_9*I(trial_c^2) + 
+                          condition*I(period_9^2)*trial_c +
+                          condition*I(period_9^2)*I(trial_c^2) +
                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
-                           I(trial_c^2)*period_9 + 
-                           I(trial_c^2)*I(period_9^2) + 
-                           condition*period_9*trial_c + 
+                          (1 +  condition*period_9*trial_c +
+                                condition*period_9*I(trial_c^2) + 
                                 condition*I(period_9^2)*trial_c +
-                           condition*period_9*I(trial_c^2) + 
-                           condition*I(period_9^2)*I(trial_c^2) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
+                                condition*I(period_9^2)*I(trial_c^2)  
                                 condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           I(trial_c^2)*period_9 + 
-                           I(trial_c^2)*I(period_9^2) + 
-                           condition*period_9*trial_c + 
-                           condition*I(period_9^2)*trial_c +
-                           condition*period_9*I(trial_c^2) + 
-                           condition*I(period_9^2)*I(trial_c^2)  | pid),
+                                trial_c*I(period_9^3) | pid),
        prior = priors_M4,
@@
 
+# the formula above comes down to the following effects:
+# cloglog[h(t)]=Intercept+     ## the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
+# period_9+I(period_9^2)+I(period_9^3)+    ##a cubic specification of TIME for reference condition "blank"
+# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+      ##effect of congruent interacts cubicly with TIME
+# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+     ##effect of incongruent interacts cubicly with TIME
+# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
+# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
+# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
 ```
@@
 
-Plot Pareto k estimates
+## Pareto k estimates
 
@@
 
-# visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trial 1000.
+## Visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trial 1000.
 
@@
           trial100 = -900/1000)
+
 # effects for trial 1000
@@
 
-# Plot model-based predicted cloglog-hazard for subject 6
+# Plot model-based predicted cloglog-hazard and hazard functions, for a single subject
 
-```{r}
+```{r select-pid}
+selected_subject = 6 # select pid here (1:6) manually
+```
+
+```{r help-function}
 make_fitted <- function(fit, scale, ...) {
@@
 
-# model_M4: Plot predicted cloglog(hazard) for single subject using fitted() in make_fitted().
+# Set up new data
 
@@
 
-```{r}
-selected_subject = 6 # select pid here (1:6) manually
-
-get_variables(model_M4)
-```
-
-First, on the cloglog-scale:
+# First, on the cloglog-scale:
 
@@
   filter(pid==selected_subject) %>% 
-  # plot!
+  # plot
   ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
@@
 
-Same on hazard scale
+# Second, on hazard scale:
 
@@
 ```
-
-##########
-
-Fit a model that relaxes proportionality assumption.
-
-```{r}
-# remove unnecessary columns before fitting a model
-ptb_data <- ptb_data %>% select(-c(trial_c)) # 12840 obs. of 4 variables
-head(ptb_data,n=17)
-summary(ptb_data)
-```
-
-```{r cubic-random-effects-model}
-plan(multicore)
-
-model_cubic_RE <-
-   brm(data = ptb_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + 
-                           condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) | pid),
-       prior = priors,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_cubic_RE")
-
-# the formula above comes down to the following effects:
-# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
-# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
-# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
-# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3) ##effect of incongruent interacts cubicly with TIME
-```
-
-```{r check-model-cubic-RE}
-summary(model_cubic_RE)
-```
-
-#### relax all ass
-
-```{r cubictime-quadtrial-model}
-plan(multicore)
-
-model_cubictime_quadtrial <-
-   brm(data = ptb_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + 
-                           condition*period_9*trial_c + 
-                           condition*period_9*I(trial_c^2) + 
-                           condition*I(period_9^2)*trial_c + 
-                           condition*I(period_9^2)*I(trial_c^2) +
-                           condition*I(period_9^3) +
-                      (1 + condition*period_9*trial_c +
-                           condition*period_9*I(trial_c^2) +
-                           condition*I(period_9^2)*trial_c +
-                           condition*I(period_9^2)*I(trial_c^2) +
-                           condition*I(period_9^3) | pid),
-       prior = priors,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_cubictime_quadtrial")
-
-# the formula above comes down to the following effects:
-# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
-# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
-# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
-# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+ ##effect of incongruent interacts cubicly with TIME
-# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
-# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
-# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
-```
-
-# plot hazard functions
-
-```{r}
-model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
-model_full_RE_newprior <- readRDS("Tutorial_2_Bayesian/models/model_full_RE_newprior.rds") 
-fixef(model_full_RE)
-fixef(model_full_RE_newprior)
-```
-
-# visualize posterior distributions of the effects of congruent and incongruent primes on hazard relative to blank prime, for each time bin in trial 1000, and for each time bin in trial 100.
-
-```{r load-model}
-model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
-```
-
-```{r make-plot-effects}
-post <-as_draws_df(model_full_RE) %>% # 8000 draws x 299 variables
-   select(starts_with("b_")) %>%       # 8000 x 18
-   expand_grid(period_9 = -3:6) %>%   
-   mutate(period_9sq = period_9^2,
-          period_9cu = period_9^3,
-          trial100 = -900/1000)  
-# effects for trial 1000
-p1congruent <- post %>% 
-  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
-  group_by(period_9) %>%
-
-  ggplot(aes(x = period_9, y = postdistr)) +
-  stat_lineribbon(show.legend=F) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
-  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90))
-
-p2incongruent <- post %>% 
-  mutate(postdistr = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
-  group_by(period_9) %>%
-
-  ggplot(aes(x = period_9, y = postdistr)) +
-  stat_lineribbon(show.legend=F) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
-  labs(x = "time bin", y = "effect of incongruent") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90))
-
-p1congruent + p2incongruent
-#(p1+p2+p3)/(p4+p5+p6)/(p7+p8+p9)/(p10+p11+p12)/(p13+p14+p15)
-```
-
-```{r save-plot-effects}
-ggsave("figures/effects_knot_1.png", width = 12, height = 16, dpi = 600)
-
-```

### 2024-09-12 Tutorial_3.Rmd
update tutorial 3

---
title: "Tutorial_3"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

# Load the libraries that we will be using

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")

lapply(pkg, library, character.only = TRUE)
```

# Load the person-trial-bin data set that we saved in Tutorial 1.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
ptb_data <- ptb_data %>% filter(period > 5)

# create factor condition, with "blank" as the reference level
ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ptb_data <- ptb_data %>% 
        mutate(period_9 = period - 9,
               trial_c = (trial - 1000)/1000)

head(ptb_data,n=17)
summary(ptb_data)
# 6 subjects
# condition: blank (6401), congruent (2642), incongruent (3797)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

```{r get-data}
M3_data <- ptb_data %>% select(pid, period_9, trial_c, condition, event)
```

```{r fit-model}
model_M3_freq <- glmer(event ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("bobyqa", "Nelder_Mead"),
                                     optCtrl = list(maxfun=1000000)), 
              data=M3_data, 
              family=binomial(link="cloglog"))
```

3 hours...

```{r inspect-model}
summary(model_M3_freq)

model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")

fixef(model_M3_freq)

coef(model_M3_f)

model_M3_f@beta

summary(model_M3_freq)

model_M3_freq@theta
model_M3_freq@u

model_M3_freq@flist

save(model_M3_freq, file="Tutorial_3_Frequentist/model_M3_freq.RData")

plot(1:16,fixef(model_M3_f),  ylim=c(-2,1), xaxis=NULL)
axis(1,at=1:16,labels=names(fixef(model_M3_f)),)
points(fixef(model_M3)[,1], col="red")
points(fixef(model_M3_fr), col="green")
points(fixef(model_M3_freq), col="blue")

length(names(fixef(model_M3_f)))
```

```{r fit-model}
model_M3_fr <- glmer(event ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=1000000)), 
              data=M3_data, 
              family=binomial(link="cloglog"))
```

```{r save-model}
save(model_M3_fr, file="Tutorial_3_Frequentist/model_M3_fr.RData")
```

```{r fit-model}
model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data=M3_data, 
              family=binomial(link="cloglog"))
```

```{r save-model}
save(model_M3_f, file="Tutorial_3_Frequentist/model_M3_f.RData")
```

### 2024-09-17 Tutorial_4.Rmd
add Tutorial_4

---
title: "Tutorial_4"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tutorial_4 illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).

# Load the libraries that we will be using.

```{r load-pkg}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

# Read in the raw data from Panis and Schmidt (2016).

```{r read-raw-data}
raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) # 9372 rows
head(raw)
tail(raw)
glimpse(raw)
```

# Do some wrangling.

```{r wrangle-raw-data}
# remove practice trials
data_wr <- raw %>% filter(practice==0) # 9240 experimental trials: 6 participants x 70 blocks x 22 trials-per-block

# remove unnecessary columns
data_wr <- data_wr %>% select(-c(ss,practice,rd,form,task,primeduration,
                                             blankduration,maskduration,targetduration,
                                             void,target_dir,corresp,resperr))
# look at trials where resp==0
data_wr %>% filter(resp==0) 
data_wr %>% filter(resp==0) %>% print(n=51)

# remove trials with no response info if rt < 800 (800 was the response deadline in each trial):
# the 46 trials where rt >= 800 and resp==0 are trials in which the subject did not respond; these should not be deleted because they will be treated as right-censored observations in EHA!
# the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
sel = which(data_wr$resp==0 & data_wr$rt < 800)
data_wr <- data_wr[-c(sel),] # 9235 rows

summary(data_wr)
# vp = Versuchsperson or subject
# bl = block number
# tr = trial number within block
# mask_type: no mask (0), relevant mask (1), irrelevant mask (2), lines mask (3)
# prime_type: no prime or blank (0), consistent prime (1), inconsistent prime (2)
# resp = no button (0), button left (1), button right (2) 
# respac = accuracy
# rt = response time
# TrialNr = trial number within experiment
```

# Change variable names, and create factors.

```{r create-names-and-factors}
# change some variable names to required names (pid, trial, condition1, condition2, rt, and acc)
colnames(data_wr) <- c("pid","bl","tr","condition1", "condition2","resp","acc","rt","trial") 
# pid to indicate column with unique person IDs
# trial must be a unique trial number per subject
# condition1 must be a factor indicating the levels (1,2,...) and labels of the first independent variable (here: mask_type)
# condition2 must be a factor indicating the levels (1,2,...) and labels of the second independent variable (here: prime_type)
# rt to indicate column with response times (in ms)
# acc to indicate column with accuracies (1/0)

# create factors for conditions (mask type and prime type): the levels of condition1 are displayed across columns; the levels of condition2 are displayed within each plot.
data_wr <- data_wr %>% mutate(condition1 = condition1 + 1, # avoid level "0"
                              condition1 = factor(condition1, levels = c(1,2,3,4), labels = c("nomask","relevant","irrelevant","lines")),
                              condition2 = condition2 + 1, # avoid level "0"
                              condition2 = factor(condition2, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
summary(data_wr)
```

# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).

Load the functions. Make sure you rename, and specify the same levels and labels for factors condition1 and condition2 in the function plot_eha_2IV as above.

```{r load-functions}
# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
censor <- function(df, timeout, bin_width){
  if(!(timeout %% bin_width == 0)){
    return("The censoring time must be a multiple of the bin width!")
  }
  if(timeout < 0 | bin_width < 0){
    return("Both timeout and bin_width must be larger than 0!")
  }
  df %>% mutate(right_censored = 0,
                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
                drt = ceiling(rtc/bin_width), # drt = discrete response time
                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
}

# function ptb creates a person-trial-bin oriented data set
ptb <- function(df){
  df %>% uncount(weights = drt) %>% # create drt rows per trial
         group_by(trial) %>% 
         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
         ungroup()
}

# function setup_lt sets up a life table for each level 2 independent variables
setup_lt_2IV <- function(ptb){
  ptb %>% mutate(event = str_c("event", event)) %>%
          group_by(condition1,condition2,period) %>% 
          count(event) %>% 
          ungroup() %>% 
          pivot_wider(names_from = event,
                      values_from = n) %>% 
          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
                 event1 = ifelse(is.na(event1),0,event1),
                 risk_set = event0 + event1) %>% # define the risk set
          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
          group_by(condition1,condition2) %>%
          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
          ungroup() 
}

# function calc_ca calculates the conditional accuracies 
calc_ca_2IV <- function(df){
  df %>% filter(right_censored==0) %>%
         group_by(condition1,condition2,drt,cens_time,bin_width) %>%
         summarize(ca = mean(acc) %>% round(digits = 2),
                   n = n(),
                   .groups = 'drop') %>%
         ungroup() %>%
         mutate(period = drt,
                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
         select(-drt)
}

# function join_lt_ca_2IV joins the conditional accuracies to the life tables when there are two independent variables
join_lt_ca_2IV <- function(df1,df2){df1 %>% left_join(df2, join_by(condition1,condition2,period))}

# function extract_median_2IV is used to extract the S(t).50 quantiles when there are two independent variables
extract_median_2IV <- function(df){
  above_pct50 <- df %>% 
      group_by(condition1,condition2) %>%
      filter(survival >= .5) %>% 
      slice(n()) # take last row
  below_pct50 <- df %>% 
      group_by(condition1,condition2) %>%
      filter(survival < .5) %>% 
      slice(1) # take first row
  # pull period above
  period_above <- pull(above_pct50, period)
  # pull survivor function values
  survival_above <- pull(above_pct50, survival)
  survival_below <- pull(below_pct50, survival)
  # estimate median by interpolation
  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
}

# function plot_eha_2IV plots the hazard, survivor, and ca(t) functions, when there are two independent variables
plot_eha_2IV <- function(df,subj,haz_yaxis){ 
  library(patchwork)
  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
  median_period <- extract_median_2IV(df)
  n_iv1 <- nlevels(df$condition1) 
  n_iv2 <- nlevels(df$condition2) 
  n_conditions <- n_iv1*n_iv2
  data_median <- c()
  for(i in 1:n_conditions){
    data_median <- append(data_median, c(median_period[i], median_period[i]))
  }
  # Create correctly named factors (here: prime and mask) and specify the same levels and labels as above
  data_medians <- tibble(period= data_median,
                         survival = rep(c(.5, 0),n_conditions),
                         prime =  factor(rep(rep(1:n_iv2, each=2),n_iv1),levels=c(1:3),labels=c("blank","congruent","incongruent")),
                         mask = factor(rep(1:n_iv1, each=2*n_iv2),levels=c(1:4),labels=c("nomask", "relevant", "irrelevant","lines")))

# plot the hazard functions
p1 <-df %>% 
  mutate(mask = condition1,       # Create correctly named factor
         prime = condition2) %>%  # Create correctly named factor
  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
  geom_line(aes(y=hazard)) +
  geom_point(aes(y=hazard), size=1) + labs(color="Prime") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
                     limits = c(1,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top",
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor

# plot the survivor functions
p2 <-df %>%
  mutate(mask = condition1,      # Create correctly named factor
         prime = condition2) %>% # Create correctly named factor
  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
  # add vertical lines to indicate the estimated median RTs with geom_path()
  geom_path(aes(x=period, y=survival),
            data = data_medians, 
            linetype = 3, show.legend = F) +
  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
                     limits=c(1,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="S(t)") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor

# plot the conditional accuracy functions
p3 <-df %>%
  mutate(mask = condition1,       # Create correctly named factor
         prime = condition2) %>%  # Create correctly named factor
  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
                     limits=c(1,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin t's endpoint (ms)", y="ca(t)") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor

p1/p2/p3
}
```

Apply the functions

```{r apply-functions}
# create a data frame with two columns (pid and data); the latter contains a list of nested data frames
data_nested <- data_wr %>% group_nest(pid)
# apply the functions using map() and map2()
data_final <- data_nested %>% 
                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, ptb)) %>%           
                     mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%      # create life tables without ca for 2 independent variables
                     mutate(condacc   = map(censored, calc_ca_2IV)) %>%       # calculate conditional accuracy for 2 independent variables
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  # add ca to lifetables for 2 independent variables
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV,1))  # create plots for 2 independent variables
```

# Extract info, and save plot for each subject

```{r extract-info, warning=F}
# view original data (column 2 in data_final) for subject 1:
pluck(data_final, 2, 1) %>% print(n=20)

# view person-trial-bin data set (column 4) for subject 1:
pluck(data_final,4,1) %>% print(n=30)

# view lifetable (column 7) for a specific combination of condition levels for subject 1:
pluck(data_final, 7, 1) %>% filter(condition1=="nomask" & condition2 == "blank")

# view plot (column 8) for subject 1:
pluck(data_final, 8, 1)

# view the median RTs in a tibble:
median_bin <- map(data_final$lifetable_ca, extract_median_2IV)
binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
median_rt  <- modify_depth(median_bin,1,~.x*binsize)
median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
median_tb %>% print(n=6*12)
```

```{r save-plots, warning=F}
# save the plot for each subject
map2(paste0("Tutorial_4_descriptive_2IV/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 8, dpi = 600)
```

### 2024-09-17 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
change filter condition

@@
       group_by(condition) %>%
-      filter(survival > .5) %>% 
+      filter(survival >= .5) %>% 
       slice(n()) # take last row

### 2024-09-17 Tutorial_3.Rmd
add code to compare model outputs

@@
 ```
+
+# Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3
+
+```{r load-models}
+model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
+load("Tutorial_3_Frequentist/model_M3_f.RData")
+```
+
+```{r extract-estimates}
+est <- tibble(brm = fixef(model_M3)[,1],
+              glmer = fixef(model_M3_f),
+              effects = names(fixef(model_M3_f))) 
+```
+
+```{r plot-estimates}
+cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
+shapes  <- c("s1" = 16, "s2" = 17)
+
+p1 <- ggplot(data = est, aes(x = effects))
+p1 <- p1 + geom_point(aes( y = brm,  color = "c1", shape = "s1"))
+p1 <- p1 + geom_point(aes( y = glmer, color = "c2", shape = "s2")) 
+p1 <- p1 + labs( x = "Effects", y = "cloglog-hazard" )
+p1 <- p1 + scale_color_manual(name = "Model", 
+                                breaks = c("c1", "c2"), 
+                                values = cols,
+                                labels = c("brm", "glmer"))
+p1 <- p1 + scale_shape_manual(name = "Model", 
+                              breaks = c("s1", "s2"),
+                              values = shapes,
+                              labels = c("brm", "glmer")) +
+   theme(panel.grid = element_blank(),
+         panel.background = element_rect(fill="white",color="black"),
+         legend.position = "top",
+         axis.text.x = element_text(angle=90)) 
+p1
+```

### 2024-09-17 Tutorial_3.Rmd
remove redundant models

@@
 ```{r fit-model}
-model_M3_freq <- glmer(event ~ 1 + condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
-
-              # control parameters, data set, and complementary log-log link function
-              control = glmerControl(optimizer = c("bobyqa", "Nelder_Mead"),
-                                     optCtrl = list(maxfun=1000000)), 
-              data=M3_data, 
-              family=binomial(link="cloglog"))
-```
-
-3 hours...
-
-```{r inspect-model}
-summary(model_M3_freq)
-
-model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
-
-fixef(model_M3_freq)
-
-coef(model_M3_f)
-
-model_M3_f@beta
-
-summary(model_M3_freq)
-
-model_M3_freq@theta
-model_M3_freq@u
-
-model_M3_freq@flist
-
-save(model_M3_freq, file="Tutorial_3_Frequentist/model_M3_freq.RData")
-
-plot(1:16,fixef(model_M3_f),  ylim=c(-2,1), xaxis=NULL)
-axis(1,at=1:16,labels=names(fixef(model_M3_f)),)
-points(fixef(model_M3)[,1], col="red")
-points(fixef(model_M3_fr), col="green")
-points(fixef(model_M3_freq), col="blue")
-
-length(names(fixef(model_M3_f)))
-```
-
-```{r fit-model}
-model_M3_fr <- glmer(event ~ 1 + condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
-
-              # control parameters, data set, and complementary log-log link function
-              control = glmerControl(optimizer = c("nlminbwrap"),
-                                     optCtrl = list(maxfun=1000000)), 
-              data=M3_data, 
-              family=binomial(link="cloglog"))
-```
-
-```{r save-model}
-save(model_M3_fr, file="Tutorial_3_Frequentist/model_M3_fr.RData")
-```
-
-```{r fit-model}
 model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
@@
 ```
+
+Model_M3_f took about 3 hours.
 

### 2024-09-17 Tutorial_1.Rmd
add function for aggregated data

--- 2024-09-17 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
@@
 
-```{r load-pkg}
+```{r load-pkg, results='hide'}
 pkg <- c("tidyverse", "patchwork")
@@
 head(raw)
-tail(raw)
-glimpse(raw)
+#tail(raw)
+#glimpse(raw)
 ```
@@
 
-```{r extract-info}
+```{r extract-info, warning=F}
 # view original data (column 2) for subject 1:
-pluck(data_final, 2, 1) # pluck() ignores the round(digits=2) function
-print(pluck(data_final, 2, 1), n=20)
+pluck(data_final, 2, 1) %>% print(n=20)
 
 # view person-trial-bin data set (column 4) for subject 1:
-pluck(data_final,4,1)
-print(pluck(data_final,4,1),n=30)
+pluck(data_final,4,1) %>% print(n=30)
 
@@
                                pid = 1)
+
+# function plot_eha_agg plots the hazard, survivor, and ca(t) functions for aggregated data.
+plot_eha_agg <- function(df,subj,haz_yaxis){ 
+  library(patchwork)
+  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
+  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
+  median_period <- extract_median(df)
+  n_conditions <- nlevels(df$condition)
+  data_median <- c()
+  for(i in 1:n_conditions){
+    data_median <- append(data_median, c(median_period[i], median_period[i]))
+  }
+
+  data_medians <- tibble(period= data_median,
+                         survival = rep(c(.5, 0),n_conditions),
+                         condition = rep(1:n_conditions, each=2))
+# plot the hazard functions
+p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=hazard)) +
+  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
+  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits = c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,haz_yaxis)) +
+  labs(x="", y="h(t)", title = "Aggregated data (N=6)") +           # New info in title
+  theme(legend.background = element_rect(fill = "transparent"),
+        panel.grid = element_blank(),
+        legend.position = "top")
+# plot the survivor functions
+p2 <-df %>%
+  ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=survival), show.legend = F) +
+  geom_point(aes(y=survival), size=1, show.legend = F) +
+  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
+  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
+            data = data_medians, 
+            linetype = 3, show.legend = F) +
+
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits=c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,1)) +
+  labs(x="", y="S(t)",
+       colour="Condition") +
+  theme(panel.grid = element_blank())
+# plot the conditional accuracy functions
+p3 <-df %>%
+  ggplot(aes(x=period, color=condition, group=condition)) +
+  geom_line(aes(y=ca), show.legend = F) +
+  geom_point(aes(y=ca), size=1, show.legend = F) +
+  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
+  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
+                     limits=c(0,cutoff/binsize)) +
+  scale_y_continuous(limits = c(0,1)) +
+  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
+       colour="Condition") +
+  theme(panel.grid = element_blank())
+
+p1/p2/p3
+}
+
 # apply functions
@@
                      mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_agg,1))  # create plots 
 # view plot
-pluck(data_final, 8, 1)
+pluck(data_agg_final, 8, 1)
 ```

### 2024-09-17 Tutorial_3.Rmd
add eval=F

@@
 
-```{r fit-model}
+```{r fit-model, eval=F}
 model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
@@
 
-```{r save-model}
+```{r save-model, eval=F}
 save(model_M3_f, file="Tutorial_3_Frequentist/model_M3_f.RData")

### 2024-09-17 Tutorial_2_Bayesian/Tutorial_2.Rmd
add eval=F

@@
 
-```{r save-plot-priors}
+```{r save-plot-priors, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
@@
 
-```{r fit-model-M1}
+```{r fit-model-M1, eval=F}
 plan(multicore)
@@
 
-```{r fit-model-M2}
+```{r fit-model-M2, eval=F}
 plan(multicore)
@@
 
-```{r fit-model-M3}
+```{r fit-model-M3, eval=F}
 plan(multicore)
@@
 
-```{r fit-model-M4}
+```{r fit-model-M4, eval=F}
 plan(multicore)
@@
 Model_M4 took about  8 hours to run.
-
-```{r check-model-M4}
-summary(model_M4)
-
-fixef(model_M4)[c("conditioncongruent", "conditionincongruent"), -2] %>% exp()
-
-# the formula above comes down to the following effects:
-# cloglog[h(t)]=Intercept+     ## the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
-# period_9+I(period_9^2)+I(period_9^3)+    ##a cubic specification of TIME for reference condition "blank"
-# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+      ##effect of congruent interacts cubicly with TIME
-# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+     ##effect of incongruent interacts cubicly with TIME
-# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
-# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
-# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
-```
 
@@
 
-```{r save-plot-effects}
+```{r save-plot-effects, eval=F}
 ggsave("figures/effects_con_incon_trial1000.png", width = 12, height = 16, dpi = 600)

### 2024-09-18 Tutorial_1.Rmd
add plot for 2 subjects and change orientation x-axis labels

@@
         panel.grid = element_blank(),
-        legend.position = "top")
+        legend.position = "top",
+        axis.text.x = element_text(angle=90))
 # plot the survivor functions
@@
        colour="Condition") +
-  theme(panel.grid = element_blank())
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90))
 # plot the conditional accuracy functions
@@
        colour="Condition") +
-  theme(panel.grid = element_blank())
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90))
 
@@
 
+Combine plots of 2 subjects.
+
+```{r plot-two-subjects}
+pa <- pluck(data_final, 8, 3)
+pb <- pluck(data_final, 8, 6)
+
+pa | pb 
+```
+
+```{r save-plot-two-subjects}
+ggsave("Tutorial_1_descriptive_stats/figures/Plot_subjects3and6_PanisSchmidt.png", width = 8, height=8, dpi=600)
+```
+
+Create a life table.
+
 ```{r create-single-lifetable}
@@
 
+Create input file for hazard models.
+
 ```{r create-file-hazard-modeling}
@@
 ```
+
+Create input file for ca(t) models.
 

### 2024-09-18 manuscript/manuscript.Rmd
update manuscript

@@
   Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-To move beyond mean performance measures, various distributional analyses have been proposed.
-Here we focus on one particular distributional analysis known as discrete-time event history analysis, a.k.a.
+The orthodox method for analysing such data - comparing means with analysis-of-variance - is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
+Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.
+  Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-Along the way, we discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
+We illustrate how one can calculate the descriptive statistics, and how one can implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
+We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
 
@@
 ```{r setup, include = FALSE}
-# Tutorial 1 and 4
+# Tutorials 1 and 4
 pkg1 <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")
@@
 # Tutorial 3
+pkg3 <- c("lme4", "nlme")
+
+lapply(pkg3, library, character.only = TRUE)
 
@@
 
-In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating mean average performance across a series of trials.
-However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
+## Means versus distributional shapes
+
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+However, differences in means conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
-In this tutorial we focus on a distributional method for analyzing time-to-event data that is known as discrete-time event history analysis (EHA), a.k.a.
-survival, hazard, duration, failure-time, and transition analysis [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
-Across four tutorials that we make publicly available on Github and the Open Science Framework (OSF), we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
 
-To apply EHA, one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+(ref:descr-fig1-caption) Means versus distributional shapes.
+
+```{r plot1, fig.cap = "(ref:descr-fig1-caption)", out.width="80%"}
+knitr::include_graphics("../Figures/rt_haz_combi_2.jpeg")
+```
+
+In each example, the mean RT is lower in condition 2 compared to condition 1.
+However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
+In the second example, the distributional effect starts around 400 ms and is gone by 800 ms.
+And in the third example, the distributional effect reverses around 550 ms.
+
+## Outline of the paper
+
+In this paper we focus on a distributional method known as discrete-time event history analysis, a.k.a.
+hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
+We first provide a brief overview of hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
+However, this will remain relatively short, and for detailed treatment, see @singerAppliedLongitudinalData2003, @allisonDiscreteTimeMethodsAnalysis1982, and @allisonSurvivalAnalysisUsing2010.
+
+We then provide four different tutorials, each of which is written in R code and publicly available on Github and the Open Science Framework (OSF).
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
+In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published data set when there is one independent variable.
+The descriptive statistics are plotted, and we comment on their interpretation.
+In Tutorial 2 we illustrate how one can fit Bayesian hazard models to the data.
+After selecting the best of four models, we plot the model-based effects and the model fits for a few subjects.
+In Tutorial 3 we illustrate how to fit hazard models in a frequentist framework.
+We compare the model-based effects between Bayesian and frequentist approaches.
+In Tutorial 4 we illustrate how to calculate the descriptive statistics when there are two independent variables.
+
+# Overview of hazard analysis
+
+To apply event history analysis (EHA), one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, a saccade onset, a fixation offset, etc.), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+Both the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
 
@@
 The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the hazard function, which assesses the unique risk associated with each time bin, the survivor function cumulates the bin-by-bin risks of event *non*occurrence.
-For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+Unlike the discrete-time hazard function, which assesses the unique risk associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence.
+
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 This latter function is also known as the micro-level speed-accuracy tradeoff function.
@@
 Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-For example, fitting parametric functions or computational models to data without studying the shape of the h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+For example, fitting parametric functions or computational models to data without studying the shape of the discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
 
@@
 Web links are printed in bold.
-The content of this tutorial is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
+The content of the tutorials is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and response accuracy (1/0) in each bin of a selected time range, respectively.
+As we will show in Tutorials 2 and 3, statistical models for h(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
 

### 2024-09-18 manuscript/manuscript.tex
update manuscript

@@
 Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-To move beyond mean performance measures, various distributional analyses have been proposed.
-Here we focus on one particular distributional analysis known as discrete-time event history analysis, a.k.a.
+The orthodox method for analysing such data - comparing means with analysis-of-variance - is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
+Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.\\
+Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-Across four tutorials that we make publicly available on Github and OSF, we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-Along the way, we discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time survival analysis.
+We illustrate how one can calculate the descriptive statistics, and how one can implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
+We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
 }
@@
 
-In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating mean average performance across a series of trials.
-However, differences in means conceal when an experimental effect starts, how long it lasts, and whether its onset is time-locked to other events.
+\subsection{Means versus distributional shapes}\label{means-versus-distributional-shapes}
+
+In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+However, differences in means conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
-In this tutorial we focus on a distributional method for analyzing time-to-event data that is known as discrete-time event history analysis (EHA), a.k.a.
-survival, hazard, duration, failure-time, and transition analysis (Allison, 1982, 2010; Singer \& Willett, 2003).
-Across four tutorials that we make publicly available on Github and the Open Science Framework (OSF), we illustrate how to calculate and interpret descriptive statistics, and how to implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
+As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
 
-To apply EHA, one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, saccade onset, fixation offset), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Figures/rt_haz_combi_2}
+
+}
+
+\caption{Means versus distributional shapes.}\label{fig:plot1}
+\end{figure}
+
+In each example, the mean RT is lower in condition 2 compared to condition 1.
+However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
+In the second example, the distributional effect starts around 400 ms and is gone by 800 ms.
+And in the third example, the distributional effect reverses around 550 ms.
+
+\subsection{Outline of the paper}\label{outline-of-the-paper}
+
+In this paper we focus on a particular distributional method known as discrete-time event history analysis, a.k.a.
+hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
+We first provide a brief overview of discrete-time hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
+However, this will remain relatively short, and for detailed treatment, see Singer and Willett (2003), Allison (1982), and Allison (2010).
+
+We then provide four different tutorials, each of which is written in R code and publicly available on Github and the Open Science Framework (OSF).
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
+In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published data set when there is one independent variable.
+The descriptive statistics are plotted, and we comment on their interpretation.
+In Tutorial 2 we illustrate how one can fit Bayesian hazard models to the data.
+After selecting the best of four models, we plot the model-based effects and the model fits for a few subjects.
+In Tutorial 3 we illustrate how to fit hazard models in a frequentist framework.
+We compare the model-based effects between Bayesian and frequentist approaches.
+In Tutorial 4 we illustrate how to calculate the descriptive statistics when there are two independent variables.
+
+\section{Overview of hazard analysis}\label{overview-of-hazard-analysis}
+
+To apply event history analysis (EHA), one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, a saccade onset, a fixation offset, etc.), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
 
@@
 Unlike the hazard function, which assesses the unique risk associated with each time bin, the survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence.
+
 For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
-The content of this tutorial is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
+The content of the tutorials is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
 
@@
 
+\bibitem[\citeproctext]{ref-R-lme4}
+Bates, D., Mächler, M., Bolker, B., \& Walker, S.
+(2015).
+Fitting linear mixed-effects models using {lme4}.
+\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
+\url{https://doi.org/10.18637/jss.v067.i01}
+
+\bibitem[\citeproctext]{ref-R-Matrix}
+Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://CRAN.R-project.org/package=Matrix}
+
 \bibitem[\citeproctext]{ref-R-RJ-2021-048}
@@
 
+\bibitem[\citeproctext]{ref-R-nlme}
+Pinheiro, J. C., \& Bates, D. M. (2000). \emph{Mixed-effects models in s and s-PLUS}. New York: Springer. \url{https://doi.org/10.1007/b98882}
+
 \bibitem[\citeproctext]{ref-R-base}

### 2024-09-18 notes/manuscript_notes.Rmd
update manuscript

@@
 
-This makes complete sense.
-But I would say it differently.
-I would emphasise in a sentence or two why mean performance measures are limited, and thus highlight the benefit of non-mean measures.
+This makes complete sense. But I would say it differently. I would emphasise in a sentence or two why mean performance measures are limited, and thus highlight the benefit of non-mean measures.
+
 And maybe try to find simpler language for distributional measures.
 Again, it is completely precise and accurate, but I worry that most folks will read it and think "that aint anything I care about" when in actual fact it might be extremely relevant.
+
 So, in other words, I think the abstract needs a clearer hook for the generalist experimental psych reader.
@@
 And this changes key conclusions about psych processes.
+
 Here we provide a tutorial on how to do this [[and then return to the details of the tutorial as you already state]].
@@
 
-I would slightly re-structure the introduction.
-I would use some subsections in the intro maybe also to organise ideas.
-The main change that I propose is to have a first section that provides the motivation in simple terms (non-modelling terms), as well as a roadmap for what is to come next in the rest of the paper, so that folks know what to expect.
+I would slightly re-structure the introduction. I would use some subsections in the intro maybe also to organise ideas.
+
+The main change that I propose is to have a first section that provides the motivation in simple terms (non-modelling terms),
+
+as well as a roadmap for what is to come next in the rest of the paper, so that folks know what to expect.
 This is especially important for reviews/tutorials etc., since it doesn't follow the standard empirical paper approach i.e., intro, methods, results ...
-Then I would move all of the modelling stuff that is currently in the intro into an entirely new section.
-Here are some ideas on structure with example sections numbered...
 
-1.
-motivation = means vs distribution/timewindow approach.
-Benefits etc.
+Then I would move all of the modelling stuff that is currently in the intro into an entirely new section. Here are some ideas on structure with example sections numbered...
+
+1. motivation = means vs distribution/timewindow approach. Benefits etc.
+
 I would have a plot in this part of the intro.
 Just real or simulated raw data - like we discussed in our meeting on Monday for your future talk - bar plots of means between conditions juxtaposed with a distributional analysis across time.
+
 I think the visually hits home the basic point about how timing is important.
@@
 However, this will remain relatively short, and for detailed treatment, see blah blah REFs.
+
 And we *must* provide some figures here.
@@
 Or do you think the mean vs distribution figure above in point (1) will be enough?
+
 We then provide 4 different tutorials, each of which is written is R code and publicly available.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can also analyse time to event data....
-Tutorial 1 blah.
-2 blah.
-3 blah.
-4 blah.
-Provide a sentence or two on each tutorial.
+Tutorial 1 blah. 2 blah. 3 blah. 4 blah. Provide a sentence or two on each tutorial.
 

### 2024-09-18 manuscript/manuscript.Rmd
update manuscript

@@
   Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data - comparing means with analysis-of-variance - is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
+The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
 Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.
@@
 In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
-As we will show in Tutorials 2 and 3, statistical models for h(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
+As we will show in Tutorials 2 and 3, statistical models for h(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
 

### 2024-09-18 manuscript/manuscript.tex
update manuscript

@@
 Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data - comparing means with analysis-of-variance - is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
+The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
 Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.\\
@@
 
-In this paper we focus on a particular distributional method known as discrete-time event history analysis, a.k.a.
+In this paper we focus on a distributional method known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of discrete-time hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
+We first provide a brief overview of hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
 However, this will remain relatively short, and for detailed treatment, see Singer and Willett (2003), Allison (1982), and Allison (2010).
@@
 To apply event history analysis (EHA), one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, a saccade onset, a fixation offset, etc.), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+Both the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
 
@@
 The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the hazard function, which assesses the unique risk associated with each time bin, the survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence.
+Unlike the discrete-time hazard function, which assesses the unique risk associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence.
 
-For two-choice RT data, the discrete-time hazard function can be extended with the conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 This latter function is also known as the micro-level speed-accuracy tradeoff function.
@@
 Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-For example, fitting parametric functions or computational models to data without studying the shape of the h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
+For example, fitting parametric functions or computational models to data without studying the shape of the discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can each be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and response accuracy (1/0) in each bin of a selected time range, respectively.
+As we will show in Tutorials 2 and 3, statistical models for h(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
 

### 2024-09-18 Tutorial_2.Rmd
update chunk options

--- 2024-09-17 Tutorial_2_Bayesian/Tutorial_2.Rmd
@@
 
-```{r load-pkg}
+```{r load-pkg, results='hide'}
 pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
@@
 
-```{r set-options}
+```{r set-options, results='hide'}
 options(brms.backend = "cmdstanr",
@@
 
-```{r check-info}
+```{r check-info, results='hide'}
 packageVersion("cmdstanr")
@@
 
-```{r fit-skew-normal}
+```{r fit-skew-normal, eval=F}
 # fit models
@@
 ```{r}
+fit_skewN <- readRDS("Tutorial_2_Bayesian/models/fit_skewN.rds")
 summary(fit_skewN) 
@@
 
-```{r check-model-M1}
+```{r check-model-M1, eval=F}
+model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
 summary(model_M1)
@@
 
-```{r check-model-M2}
+```{r check-model-M2, eval=F}
+model_M2 <- readRDS("Tutorial_2_Bayesian/models/model_M2.rds")
 summary(model_M2)
@@
 
-```{r check-model-M3}
+```{r check-model-M3, eval=F}
+model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
+
 summary(model_M3)
@@
 
+```{r check-model-M4, eval=F}
+model_M4 <- readRDS("Tutorial_2_Bayesian/models/model_M4.rds")
+summary(model_M4)
+```
+
 # Compare 4 models
 
-```{r}
+```{r compare-models}
 model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
@@
 
-```{r}
+```{r add-criterion, eval=F}
 model_M1  <- add_criterion(model_M1, c("loo", "waic"))
@@
 
-```{r}
+```{r loo-compare}
 loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F)
@@
 
-```{r}
+```{r model-weights}
 model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
@@
 
-```{r}
+```{r pareto}
 loo(model_M4)$diagnostics %>% 
@@
 
-```{r make-plot-effects}
-post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
-   select(starts_with("b_")) %>%       # 8000 x 31
-   expand_grid(period_9 = -3:6) %>%   
-   mutate(period_9sq = period_9^2,
-          period_9cu = period_9^3,
-          trial100 = -900/1000)  
-# effects for trial 1000
-p1congruent <- post %>% 
-  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
-  group_by(period_9) %>%
-
-  ggplot(aes(x = period_9, y = postdistr)) +
-  stat_lineribbon(show.legend=T) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
-  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90))
-
-p2incongruent <- post %>% 
-  mutate(postdistr = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
-  group_by(period_9) %>%
-
-  ggplot(aes(x = period_9, y = postdistr)) +
-  stat_lineribbon(show.legend=F) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
-  labs(x = "time bin", y = "effect of incongruent") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90))
-
-p1congruent + p2incongruent
-```
-
-```{r save-plot-effects, eval=F}
-ggsave("figures/effects_con_incon_trial1000.png", width = 12, height = 16, dpi = 600)
-```
-
-```{r}
+```{r post-distr}
 post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
@@
 
+```{r save-plot-effects, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png", width = 12, height = 12, dpi = 600)
+```
+
 # plot effects on hazard ratio scale
 
-```{r}
+```{r haz-ratio}
 plot_facet_hr <- post %>% 
@@
 
-```{r}
+```{r new-data}
 # define the `newdata`
@@
 
-```{r}
+```{r subject-cloglog}
 make_fitted(model_M4, scale = "linear") %>% 
@@
 
-```{r}
+```{r subject-hazard}
 make_fitted(model_M4, scale = "response") %>% 

### 2024-09-18 manuscript/manuscript.Rmd
update manuscript

@@
 The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
+We  discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
 
@@
 
-In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016.
+In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
 
@@
 
+We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analysing time-to-event data.
+By providing this tutorial, we hope that researchers will start using hazard analysis more often.
+While we focused on within-subject factorial small-*N* designs, event history analysis has been developed to also encompass other designs.
+
+## Advantages of hazard functions
+
 ## Individual differences
 
-- role of response deadlines, low-level vs. higher-level processes, 
+- role of response deadlines, low-level vs. higher-level processes, overlap between distributions
 - clustering algorithms based on h(t) and ca(t) data

### 2024-09-18 manuscript/manuscript.tex
update manuscript

@@
 The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
+We discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
 }
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 
-In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016).
+In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
 
 \section{Discussion}\label{discussion}
+
+We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analysing time-to-event data.
+By providing this tutorial, we hope that researchers will start using hazard analysis more often.
+While we focused on within-subject factorial small-\emph{N} designs, event history analysis has been developed to also encompass other designs.
+
+\subsection{Advantages of hazard functions}\label{advantages-of-hazard-functions}
 
@@
 \item
-  role of response deadlines, low-level vs.~higher-level processes,
+  role of response deadlines, low-level vs.~higher-level processes, overlap between distributions
 \item

### 2024-09-19 manuscript/manuscript.Rmd
update manuscript

--- 2024-09-18 manuscript/manuscript.Rmd
@@
   Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
-Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.
+The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and how they evolve with increasing waiting time.
+Such information can change key conclusions about psychological processes and can be revealed by using distributional measures that portray the detailed shape of time-to-event distributions.
   Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
@@
 The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
+We  discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, the advantages of a hazard analysis over other distributional methods available in the literature, limitations, and extensions.
+Our ultimate goal is to convince readers to start using hazard analysis more often when dealing with time-to-event data.
 
@@
 In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in means conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
-Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
+Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
@@
 However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
-In the second example, the distributional effect starts around 400 ms and is gone by 800 ms.
-And in the third example, the distributional effect reverses around 550 ms.
+In the second example, the effect starts around 400 ms and is gone by 800 ms.
+And in the third example, the effect reverses around 550 ms.
 
@@
 This latter function is also known as the micro-level speed-accuracy tradeoff function.
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by @holdenDispersionResponseTimes2009, “probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions” (p.
-331).
-Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if F~A~(t) > F~B~(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
-Townsend (1990) showed that a complete ordering on the hazard functions —$\lambda$~A~(t) > $\lambda$~B~(t) for all t— implies a complete ordering on both the cumulative distribution and survivor functions —F~A~(t) > F~B~(t) and S~A~(t) < S~B~(t)— which in turn implies an ordering on the mean latencies —mean A < mean B.
-In contrast, an ordering on two means does not imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does not imply a complete ordering on the corresponding hazard functions.
-This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
-For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period so that we only know that the RT must be larger than some value.
-This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010) which is useful for cognitive psychophysiology [@meyerModernMentalChronometry1988].
-Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-For example, fitting parametric functions or computational models to data without studying the shape of the discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
 
@@
 
-Figure 1 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
+Figure 2 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
 By using discrete-time h(t) functions of event occurrence - in combination with ca(t) functions for two-choice tasks - one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
@@
 
-(ref:descr-fig-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
+(ref:descr-fig2-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
 
-```{r eha-plot, fig.cap = "(ref:descr-fig-caption)"}
+```{r eha-plot, fig.cap = "(ref:descr-fig2-caption)"}
 knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt.png")
@@
 Qualitatively similar results were obtained for the other five participants.
-Also, in their second Experiment, @panisWhatShapingRT2016 showed that the negative compatibility effect in the mask-present conditions is time-locked to mask onset.
+These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+
+  Also, in their second Experiment, @panisWhatShapingRT2016 showed that the negative compatibility effect in the mask-present conditions (see Tutorial 4) is time-locked to mask onset.
 This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022; @schmidtResponseInhibitionNegative2022].
@@
 
-When you want to study how hazard depends on various predictors, you can fit regression models to the data [@singerAppliedLongitudinalData2003].
-There are two analytic decisions one has to make.
-First, one has to select an analysis time range, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, one can choose the logit link function which transforms a (hazard) probability into the log of the odds ratio, or the complementary log-log (cloglog) link function, which yields the logarithm of the negated logarithm of the probability of event *non*occurrence.
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model to the masked response priming data set used in the first tutorial.
+Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
+
+There are several analytic decisions one has to make.
+First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
+Second, given that the dependent variable is binary, one has to select a link function (see Figure 3).
+
+(ref:descr-fig3-caption) The logit and cloglog link functions.
+
+```{r plot-link-functions, fig.cap = "(ref:descr-fig3-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
+```
+
+The symmetric logit link function transforms a (hazard) probability into the log of the odds ratio.
+The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event *non*occurrence.
 An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
@@
 Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
-One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one.
+One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
 
-An example discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
 
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-1) + $\alpha$~3~(TIME-1)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-1)].
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)].
 
-The main predictor variable TIME is the time bin index t that is centered on value 1 in this example.
+The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
 The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor.
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
 For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
-However, if the predictor interacts linearly with time (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 1 by $\beta$~2~ cloglog-hazard units (when TIME-1 = 0), in bin 2 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-1 = 1), and so forth.
-To interpret the effects of the predictors, the parameter estimates are exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
+To interpret the effects of a predictor,its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+When using the logit link, exponentiating a $\beta$ parameter results in an odds ratio.
 
@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the masked response priming data set used in the first tutorial. 
 In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
 
-First, we select the analysis range (200,600] and the cloglog link, and use a polynomial to specify the effect of TIME in the "blank" prime condition.
-Second, based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] and because cognition is likely the behavior of a non-linear dynamical system [ref], we relax all three assumptions, as follows:
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the data.
+We select the analysis range (200,600] and the cloglog link.
+The data is prepared as follows.
 
@@
 ```{r prepare-data, echo=T}
-# load data
-ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
-ptb_data <- ptb_data %>% filter(period > 5)
+# load person-trial-bin oriented data set
+#ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
+# select analysis time range: (200,600] with 10 bins (time bin ranks t=6 to t=15) of 40 ms
+#ptb_data <- ptb_data %>% filter(period > 5)
 # create factor condition, with "blank" as the reference level
-ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+#ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
 # center discrete TIME (period) on bin 9, and trial on trial 1000
-ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
-                                trial_c = (trial - 1000)/1000)
+#ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
+  #                              trial_c = (trial - 1000)/1000)
 # remove unnecessary columns before fitting a model
-ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
-```
-
-```{r set-priors, echo=T}
-priors <- c(
-  set_prior("normal(0, 1)", class = "b"), # for beta parameters 
-  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef = "Intercept"), # flat prior for intercept on hazard scale
-  set_prior("normal(0, 1)", class = "sd"), # for standard deviation of RE
-  set_prior("lkj(2)", class = "cor") # for correlations between RE
-)
-```
-
-```{r fit-model, echo=T}
-#plan(multicore)
-#model_full_RE <-
-#   brm(data = ptb_data,
-#       family = binomial(link="cloglog"),
-#       event | trials(1) ~ 0 + Intercept + 
-#                           condition*period_9*trial_c + 
-#                           condition*I(period_9^2) + 
-#                           condition*I(period_9^3) +
-#                           (1 + condition*period_9*trial_c +
-#                           condition*I(period_9^2) +
-#                           condition*I(period_9^3) | pid),
-#       prior = priors,
-#       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-#       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-#       seed = 12, init = "0",
-#       file = "../Tutorial_2_Bayesian/models/model_full_RE")
+#ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
 ```
@@
 
-To test whether (centered) trial number affects behavior, we fit a model without the variable trial_c.
+## Model 1: A general specification of TIME, and main effects of congruency and trial number
 
-Use WAIC to compare models.
+For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+
+## Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
+, and use a polynomial to specify the effect of TIME in the "blank" prime condition.
+
+## Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
+
+## Model 4: A polynomial specification of TIME, and relaxing all three assumptions
+
+Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] and because cognition is likely the behavior of a non-linear dynamical system [ref], we relax all three assumptions in model 4.
+
+## Compare the models.
+
+Use WAIC to compare models (ref McElreath, Kurz)
+
+## Plot congruency effects and subject-specific fits for the selected model
 
@@
 
-In this third tutorial we illustrate how to fit a frequentist hazard regression model for the data set used in the first tutorial.
+In this third tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in the first tutorial.
 
@@
 
+We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
+By providing this tutorial, we hope that researchers will start using hazard analysis more often, due to the many advantages described below.
+While we focus here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+
+## Advantages of hazard functions
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by @holdenDispersionResponseTimes2009, “probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions” (p.
+331; see also Figure 1 in ....) when one is interested in studying the detailed shape of a RT distribution.
+
+Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if F~A~(t) > F~B~(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
+Townsend (1990) showed that a complete ordering on the hazard functions —$\lambda$~A~(t) > $\lambda$~B~(t) for all t— implies a complete ordering on both the cumulative distribution and survivor functions —F~A~(t) > F~B~(t) and S~A~(t) < S~B~(t)— which in turn implies an ordering on the mean latencies —mean A < mean B.
+In contrast, an ordering on two means does *not* imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does *not* imply a complete ordering on the corresponding hazard functions.
+This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+Thus, because the discrete-time hazard function identifies unique information about event occurrence in each bin, while F(t) accumulates the complements of the hazard estimates from the current and previous bins -- S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)] --, instead of using delta-plots -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions (see ref).
+
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (i.e., the response deadline).
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
+As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
+This orthodox procedure can lead to a sampling bias, however, which results in underestimation of the mean.
+By introducting a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects with behavioral effeccts when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
+
+Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the same detailed shape of the empirical RT distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+
 ## Individual differences
 
-- role of response deadlines, low-level vs. higher-level processes, 
-- clustering algorithms based on h(t) and ca(t) data
--
+One important issue is that of possible individual differences in the overall location of the distribution, and the time course of psychological effects.
+For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
+These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 600 ms for simple detection tasks, 1000 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
+Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response), introducing a response deadline is recommended when designing RT experiments.
+Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
+But even when using a response deadline, participants can differ qualitatively in the effects they display (refs).
+One way to deal with this is to run a clustering algorithm on the individual hazard estimates across all conditions.
+The obtained dendrogram can then be used to identify (hopefully big) clusters of participants that behave similarly, and to identify (hopefully small) clusters of participants with outlying behavioral patterns.
+One might then exclude the outlying participants before fitting a hazard model.
 
-## Cognitive psychophysiology and computational model selection
+## Limitations
 
-## Power analysis
+One important limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+There is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
+We therefore recommend to design as many trials as possible per experimental condition given the available resources.
+Another limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
 
-- example repo on github
+## Extensions
 
-## Preregistration
-
-- example preregistration for knot data
+The hazard models in this tutorial assume that there is one event of interest.
+However, in certain situations, more than one event of interest might exist.
+For example, ...
+- multi-state hazard models refs
 
 # Conclusions
+see paper...
 

### 2024-09-19 manuscript/manuscript.tex
update manuscript

--- 2024-09-18 manuscript/manuscript.tex
@@
 Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and whether they are time-locked to stimuli.
-Such information can change key conclusions about psychological processes and can be revealed by using distributional measures.\\
+The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and how they evolve with increasing waiting time.
+Such information can change key conclusions about psychological processes and can be revealed by using distributional measures that portray the detailed shape of time-to-event distributions.\\
 Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
@@
 The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We further discuss how to manage inter-individual differences, implications for experimental design, and how to select among various options when analysing time-to-event data using discrete-time hazard analysis.
+We discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, and the advantages of a hazard analysis over other distributional methods available in the literature.
+Our ultimate goal is to convince readers to start using hazard analysis more often when dealing with time-to-event data.
 }
@@
 
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331).
-Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
-Townsend (1990) showed that a complete ordering on the hazard functions ---\(\lambda\)\textsubscript{A}(t) \textgreater{} \(\lambda\)\textsubscript{B}(t) for all t--- implies a complete ordering on both the cumulative distribution and survivor functions ---F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) and S\textsubscript{A}(t) \textless{} S\textsubscript{B}(t)--- which in turn implies an ordering on the mean latencies ---mean A \textless{} mean B.
-In contrast, an ordering on two means does not imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does not imply a complete ordering on the corresponding hazard functions.
-This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
-For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period so that we only know that the RT must be larger than some value.
-This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010) which is useful for cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
-Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-For example, fitting parametric functions or computational models to data without studying the shape of the discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
-
 We focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-\emph{N} designs).
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 Also, in their second Experiment, Panis and Schmidt (2016) showed that the negative compatibility effect in the mask-present conditions is time-locked to mask onset.
-This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Schmidt, Panis, Wolkersdorfer, \& Vorberg, 2022).
+This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Schmidt, Panis, Wolkersdorfer, \& Vorberg, 2022).
 In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
@@
 
-When you want to study how hazard depends on various predictors, you can fit regression models to the data (Singer \& Willett, 2003).
-There are two analytic decisions one has to make.
-First, one has to select an analysis time range, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, one can choose the logit link function which transforms a (hazard) probability into the log of the odds ratio, or the complementary log-log (cloglog) link function, which yields the logarithm of the negated logarithm of the probability of event \emph{non}occurrence.
+In this second tutorial we illustrate how to fit a Bayesian hazard regression model to the masked response priming data set used in the first tutorial.
+Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
+
+There are several analytic decisions one has to make.
+First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
+Second, given that the dependent variable is binary, one has to select a link function (see Figure 3).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}
+
+}
+
+\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
+\end{figure}
+
+The symmetric logit link function transforms a (hazard) probability into the log of the odds ratio.
+The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event \emph{non}occurrence.
 An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
@@
 Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
-One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one.
+One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
 
-An example discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
 
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-1) + \(\alpha\)\textsubscript{3}(TIME-1)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-1){]}.
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.
 
-The main predictor variable TIME is the time bin index t that is centered on value 1 in this example.
+The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
 The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor.
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
 For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units.
-However, if the predictor interacts linearly with time (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 1 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-1 = 0), in bin 2 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-1 = 1), and so forth.
-To interpret the effects of the predictors, the parameter estimates are exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth.
+To interpret the effects of a predictor,its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.
 
@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model for the masked response priming data set used in the first tutorial.
 In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
 
-First, we select the analysis range (200,600{]} and the cloglog link, and use a polynomial to specify the effect of TIME in the ``blank'' prime condition.
-Second, based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) and because cognition is likely the behavior of a non-linear dynamical system {[}ref{]}, we relax all three assumptions, as follows:
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the data.
+We select the analysis range (200,600{]} and the cloglog link.
+The data is prepared as follows.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# load data}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
-\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{)}
+\CommentTok{\# load person{-}trial{-}bin oriented data set}
+\CommentTok{\#ptb\_data \textless{}{-} read\_csv("Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv")}
+\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks t=6 to t=15) of 40 ms}
+\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% filter(period \textgreater{} 5)}
 \CommentTok{\# create factor condition, with "blank" as the reference level}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
+\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))}
 \CommentTok{\# center discrete TIME (period) on bin 9, and trial on trial 1000}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{period\_9 =}\NormalTok{ period }\SpecialCharTok{{-}} \DecValTok{9}\NormalTok{,}
-                                \AttributeTok{trial\_c =}\NormalTok{ (trial }\SpecialCharTok{{-}} \DecValTok{1000}\NormalTok{)}\SpecialCharTok{/}\DecValTok{1000}\NormalTok{)}
+\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% mutate(period\_9 = period {-} 9,}
+  \CommentTok{\#                              trial\_c = (trial {-} 1000)/1000)}
 \CommentTok{\# remove unnecessary columns before fitting a model}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period)) }\CommentTok{\# 12840 obs.
-of 5 variables}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{priors }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{), }\CommentTok{\# for beta parameters }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"student\_t(7.61, 0, 1.57)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{), }\CommentTok{\# flat prior for intercept on hazard scale}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{), }\CommentTok{\# for standard deviation of RE}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{) }\CommentTok{\# for correlations between RE}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#plan(multicore)}
-\CommentTok{\#model\_full\_RE \textless{}{-}}
-\CommentTok{\#   brm(data = ptb\_data,}
-\CommentTok{\#       family = binomial(link="cloglog"),}
-\CommentTok{\#       event | trials(1) \textasciitilde{} 0 + Intercept + }
-\CommentTok{\#                           condition*period\_9*trial\_c + }
-\CommentTok{\#                           condition*I(period\_9\^{}2) + }
-\CommentTok{\#                           condition*I(period\_9\^{}3) +}
-\CommentTok{\#                           (1 + condition*period\_9*trial\_c +}
-\CommentTok{\#                           condition*I(period\_9\^{}2) +}
-\CommentTok{\#                           condition*I(period\_9\^{}3) | pid),}
-\CommentTok{\#       prior = priors,}
-\CommentTok{\#       chains = 4, cores = 4, iter = 3000, warmup = 1000,}
-\CommentTok{\#       control = list(adapt\_delta = 0.999, step\_size = 0.04, max\_treedepth = 12),}
-\CommentTok{\#       seed = 12, init = "0",}
-\CommentTok{\#       file = "../Tutorial\_2\_Bayesian/models/model\_full\_RE")}
+\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% select({-}c(bl,tr,trial,period)) \# 12840 obs. of 5 variables}
 \end{Highlighting}
@@
 
-To test whether (centered) trial number affects behavior, we fit a model without the variable trial\_c.
+\subsection{Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
 
-Use WAIC to compare models.
+For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+
+\subsection{Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+
+, and use a polynomial to specify the effect of TIME in the ``blank'' prime condition.
+
+\subsection{Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
+
+\subsection{Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
+
+Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) and because cognition is likely the behavior of a non-linear dynamical system {[}ref{]}, we relax all three assumptions in model 4.
+
+\subsection{Compare the models.}\label{compare-the-models.}
+
+Use WAIC to compare models (ref McElreath, Kurz)
+
+\subsection{Plot congruency effects and subject-specific fits for the selected model}\label{plot-congruency-effects-and-subject-specific-fits-for-the-selected-model}
 
@@
 
-In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016).
+In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
 
@@
 
+We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
+By providing this tutorial, we hope that researchers will start using hazard analysis more often, due to the many advantages described below.
+While we focus here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+
+\subsection{Advantages of hazard functions}\label{advantages-of-hazard-functions}
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions'' (p.~331; see also Figure 1 in \ldots.) when one is interested in studying the detailed shape of a RT distribution.
+
+Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
+Townsend (1990) showed that a complete ordering on the hazard functions ---\(\lambda\)\textsubscript{A}(t) \textgreater{} \(\lambda\)\textsubscript{B}(t) for all t--- implies a complete ordering on both the cumulative distribution and survivor functions ---F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) and S\textsubscript{A}(t) \textless{} S\textsubscript{B}(t)--- which in turn implies an ordering on the mean latencies ---mean A \textless{} mean B.
+In contrast, an ordering on two means does \emph{not} imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does \emph{not} imply a complete ordering on the corresponding hazard functions.
+This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
+Thus, because the discrete-time hazard function identifies unique information about event occurrence in each bin, while F(t) accumulates the complements of the hazard estimates from the current and previous bins -- S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} --, instead of using delta-plots -- differences in quantiles from F(t)\(^-1\) -- one can simply plot delta-h(t) functions (see ref).
+
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (i.e., the response deadline).
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
+As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
+This orthodox procedure can lead to a sampling bias, however, which results in underestimation of the mean.
+By introducting a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects with behavioral effeccts when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+
+Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the same detailed shape of the empirical RT distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
+
 \subsection{Individual differences}\label{individual-differences}
 
-\begin{itemize}
-\tightlist
-\item
-  role of response deadlines, low-level vs.~higher-level processes,
-\item
-  clustering algorithms based on h(t) and ca(t) data
-\item
-\end{itemize}
+One important issue is that of possible individual differences in the overall location of the distribution, and the time course of psychological effects.
+For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
+These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 600 ms for simple detection tasks, 1000 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
+Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response), introducing a response deadline is recommended when designing RT experiments.
+Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
+But even when using a response deadline, participants can differ qualitatively in the effects they display (refs).
+One way to deal with this is to run a clustering algorithm on the individual hazard estimates across all conditions.
+The obtained dendrogram can then be used to identify (hopefully big) clusters of participants that behave similarly, and to identify (hopefully small) clusters of participants with outlying behavioral patterns.
+One might then exclude the outlying participants before fitting a hazard model.
 
-\subsection{Cognitive psychophysiology and computational model selection}\label{cognitive-psychophysiology-and-computational-model-selection}
+\subsection{Limitations}\label{limitations}
 
-\subsection{Power analysis}\label{power-analysis}
+One important limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+There is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
+We therefore recommend to design as many trials as possible per experimental condition given the available resources.
+Another limitation of multilevel hazard modeling is that it might take some time to fit the model.
 
-\begin{itemize}
-\tightlist
-\item
-  example repo on github
-\end{itemize}
+\subsection{Extensions}\label{extensions}
 
-\subsection{Preregistration}\label{preregistration}
-
-\begin{itemize}
-\tightlist
-\item
-  example preregistration for knot data
-\end{itemize}
+The hazard models in this tutorial assume that there is one event of interest.
+However, in certain situations, more than one event of interest might exist.
+For example, \ldots{} - multi-state hazard models refs
 
 \section{Conclusions}\label{conclusions}
+
+see paper\ldots{}
 

### 2024-09-21 manuscript/manuscript.Rmd
update manuscript

@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the data.
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1.
 We select the analysis range (200,600] and the cloglog link.
@@
 # load person-trial-bin oriented data set
-#ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-# select analysis time range: (200,600] with 10 bins (time bin ranks t=6 to t=15) of 40 ms
-#ptb_data <- ptb_data %>% filter(period > 5)
+ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
+
+# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
+ptb_data <- ptb_data %>% filter(period > 5)
+
 # create factor condition, with "blank" as the reference level
-#ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-# center discrete TIME (period) on bin 9, and trial on trial 1000
-#ptb_data <- ptb_data %>% mutate(period_9 = period - 9,
-  #                              trial_c = (trial - 1000)/1000)
-# remove unnecessary columns before fitting a model
-#ptb_data <- ptb_data %>% select(-c(bl,tr,trial,period)) # 12840 obs. of 5 variables
+ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+
+# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.
+ptb_data <- ptb_data %>% 
+        mutate(period_9 = period - 9,
+               trial_c = (trial - 1000)/1000,
+               d6  = if_else(period == 6, 1, 0),
+               d7  = if_else(period == 7, 1, 0),
+               d8  = if_else(period == 8, 1, 0),
+               d9  = if_else(period == 9, 1, 0),
+               d10 = if_else(period == 10, 1, 0),
+               d11 = if_else(period == 11, 1, 0),
+               d12 = if_else(period == 12, 1, 0),
+               d13 = if_else(period == 13, 1, 0),
+               d14 = if_else(period == 14, 1, 0),
+               d15 = if_else(period == 15, 1, 0))
 ```
 \normalsize
+
+## Prior distributions
+
+To get the posterior distribution of the parameters given the data, we need to specify the prior distribution of the parameters.
+The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+
+(ref:descr-fig4-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+
+```{r plot-priors, fig.cap = "(ref:descr-fig4-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")
+```
+
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
+
+To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the beta parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+
+To gain a sense of what prior *cloglog* values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
+Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the alpha parameters or intercept(s) in a cloglog-hazard model.
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the beta parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
 
@@
 For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2, this is accomplished as follows.
+
+\scriptsize
+```{r data-M1, echo=T}
+# remove unnecessary columns before fitting a model
+M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9))
+
+# Specify priors
+priors_M1 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
+  set_prior("normal(0, 1)", class = "sd"),        
+  set_prior("lkj(2)", class = "cor")            
+)
+```
+\normalsize
+
+We can now estimate our first Bayesion regression model, as follows.
+
+\scriptsize
+```{r fit-model-M1, echo=T, eval=F}
+plan(multicore)
+
+model_M1 <-
+   brm(data = M1_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
+         condition + trial_c +
+
+                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+       prior = priors_M1,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M1")
+```
+\normalsize
+
+Estimating model_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
 ## Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
-, and use a polynomial to specify the effect of TIME in the "blank" prime condition.
+
+For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
 
@@
 
-## Advantages of hazard functions
+## Advantages of hazard analysis
 
@@
 Hazard functions are thus more diagnostic than density functions” (p.
-331; see also Figure 1 in ....) when one is interested in studying the detailed shape of a RT distribution.
+331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in ....).
 
@@
 But even when using a response deadline, participants can differ qualitatively in the effects they display (refs).
-One way to deal with this is to run a clustering algorithm on the individual hazard estimates across all conditions.
+One way to deal with this is to describe and interpret the different patterns.
+Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
 The obtained dendrogram can then be used to identify (hopefully big) clusters of participants that behave similarly, and to identify (hopefully small) clusters of participants with outlying behavioral patterns.

### 2024-09-21 manuscript/manuscript.tex
update manuscript

@@
 The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, and the advantages of a hazard analysis over other distributional methods available in the literature.
+We discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, the advantages of a hazard analysis over other distributional methods available in the literature, limitations, and extensions.
 Our ultimate goal is to convince readers to start using hazard analysis more often when dealing with time-to-event data.
@@
 In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in means conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
-Such information is useful not only for interpretation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
+Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 Qualitatively similar results were obtained for the other five participants.
-Also, in their second Experiment, Panis and Schmidt (2016) showed that the negative compatibility effect in the mask-present conditions is time-locked to mask onset.
+These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+
+Also, in their second Experiment, Panis and Schmidt (2016) showed that the negative compatibility effect in the mask-present conditions (see Tutorial 4) is time-locked to mask onset.
 This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Schmidt, Panis, Wolkersdorfer, \& Vorberg, 2022).
@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the data.
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1.
 We select the analysis range (200,600{]} and the cloglog link.
@@
 \CommentTok{\# load person{-}trial{-}bin oriented data set}
-\CommentTok{\#ptb\_data \textless{}{-} read\_csv("Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv")}
-\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks t=6 to t=15) of 40 ms}
-\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% filter(period \textgreater{} 5)}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
+
+\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{)}
+
 \CommentTok{\# create factor condition, with "blank" as the reference level}
-\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))}
-\CommentTok{\# center discrete TIME (period) on bin 9, and trial on trial 1000}
-\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% mutate(period\_9 = period {-} 9,}
-  \CommentTok{\#                              trial\_c = (trial {-} 1000)/1000)}
-\CommentTok{\# remove unnecessary columns before fitting a model}
-\CommentTok{\#ptb\_data \textless{}{-} ptb\_data \%\textgreater{}\% select({-}c(bl,tr,trial,period)) \# 12840 obs. of 5 variables}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
+
+\CommentTok{\# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} 
+        \FunctionTok{mutate}\NormalTok{(}\AttributeTok{period\_9 =}\NormalTok{ period }\SpecialCharTok{{-}} \DecValTok{9}\NormalTok{,}
+               \AttributeTok{trial\_c =}\NormalTok{ (trial }\SpecialCharTok{{-}} \DecValTok{1000}\NormalTok{)}\SpecialCharTok{/}\DecValTok{1000}\NormalTok{,}
+               \AttributeTok{d6  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d7  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{7}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d8  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{8}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d9  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{9}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d10 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{10}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d11 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{11}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d12 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{12}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d13 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{13}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d14 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{14}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+               \AttributeTok{d15 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{15}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{))}
 \end{Highlighting}
@@
 
+\subsection{Prior distributions}\label{prior-distributions}
+
+To get the posterior distribution of the parameters given the data, we need to specify the prior distribution of the parameters.
+The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}
+
+}
+
+\caption{Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
+\end{figure}
+
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
+
+To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the beta parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+
+To gain a sense of what prior \emph{cloglog} values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
+Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the alpha parameters or intercept(s) in a cloglog-hazard model.
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the beta parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
+
 \subsection{Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
@@
 For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2, this is accomplished as follows.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# remove unnecessary columns before fitting a model}
+\NormalTok{M1\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period, period\_9,d9))  }
+
+\CommentTok{\# Specify priors}
+\NormalTok{priors\_M1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}0.2,0.71,{-}2.2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),       }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d6"}\NormalTok{),}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d7"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d8"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d10"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d11"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d12"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d13"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d14"}\NormalTok{), }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d15"}\NormalTok{),}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{),}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{),        }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{)            }
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+We can now estimate our first Bayesion regression model, as follows.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plan}\NormalTok{(multicore)}
+
+\NormalTok{model\_M1 }\OtherTok{\textless{}{-}}
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M1\_data,}
+       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}
+\NormalTok{         condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
+
+\NormalTok{                   (d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors\_M1,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
+       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M1"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Estimating model\_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
@@
 
-, and use a polynomial to specify the effect of TIME in the ``blank'' prime condition.
+For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
 
@@
 
-\subsection{Advantages of hazard functions}\label{advantages-of-hazard-functions}
+\subsection{Advantages of hazard analysis}\label{advantages-of-hazard-analysis}
 
@@
 First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331; see also Figure 1 in \ldots.) when one is interested in studying the detailed shape of a RT distribution.
+Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in \ldots.).
 
@@
 We therefore recommend to design as many trials as possible per experimental condition given the available resources.
-Another limitation of multilevel hazard modeling is that it might take some time to fit the model.
+Another limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
 

### 2024-09-22 manuscript/manuscript.Rmd
update manuscript

@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model to the masked response priming data set used in the first tutorial.
+In this second tutorial we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
 
-There are several analytic decisions one has to make.
+There are several analytic decisions one has to make when fitting a hazard model.
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
@@
 
-In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (see Tutorial 1); @allisonSurvivalAnalysisUsing2010).
+In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set; @allisonSurvivalAnalysisUsing2010).
 When there is clustering in the data, as in the case of a small-*N* design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models [@allisonSurvivalAnalysisUsing2010].
@@
 
-To get the posterior distribution of the parameters given the data, we need to specify the prior distribution of the parameters.
+To get the posterior distribution of each parameters given the data, we need to specify the prior distribution of each parameter.
 The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
@@
 Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
-However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the alpha parameters or intercept(s) in a cloglog-hazard model.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
 A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the beta parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
@@
 \scriptsize
-```{r data-M1, echo=T}
+```{r data-priors-M1, echo=T}
 # remove unnecessary columns before fitting a model
@@
 
-Estimating model_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
@@
 
-For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+We first remove unncessary columns and specify our priors.
+
+\scriptsize
+```{r data-priors-M2, echo=T}
+# remove unnecessary columns
+M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15))
+
+# Specify priors
+priors_M2 <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),    
+  set_prior("normal(0, 1)", class = "sd"),            
+  set_prior("lkj(2)", class = "cor")                 
+)
+```
+\normalsize
+
+Now we can fit model 2.
+
+\scriptsize
+```{r fit-model-M2, eval=F, echo=T}
+plan(multicore)
+
+model_M2 <-
+   brm(data = M2_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c +
+                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
+                          condition + trial_c | pid),
+       prior = priors_M2,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M2")
+```
+\normalsize
+
+Estimating model M2 took about 144 minutes.
 
 ## Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
+
+For the third model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables congruency (variable "condition") and trial number (variable "trial_c").
+We use the same data set and priors as for model 2.
+
+\scriptsize
+```{r fit-model-M3, eval=F, echo=T}
+M3_data <- M2_data
+priors_M3 <- priors_M2
+plan(multicore)
+
+model_M3 <- 
+   brm(data = M3_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
+                           condition*period_9 + 
+                           condition*I(period_9^2) + 
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) +
+                           (1 + condition*period_9 +
+                           condition*I(period_9^2) +
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) | pid),
+       prior = priors_M3,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M3")
+```
+\normalsize
+
+Estimating model M3 took about 268 minutes.
 
@@
 Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] and because cognition is likely the behavior of a non-linear dynamical system [ref], we relax all three assumptions in model 4.
+We use the same data set and priors as for model 2.
+
+\scriptsize 
+```{r fit-model-M4, eval=F, echo=T}
+M4_data <- M2_data
+priors_M4 <- priors_M2
+plan(multicore)
+
+model_M4 <- 
+   brm(data = M4_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
+                          condition*period_9*trial_c + 
+                          condition*period_9*I(trial_c^2) + 
+                          condition*I(period_9^2)*trial_c +
+                          condition*I(period_9^2)*I(trial_c^2) +
+                          condition*I(period_9^3) +
+                          trial_c*I(period_9^3) +
+                          (1 +  condition*period_9*trial_c +
+                                condition*period_9*I(trial_c^2) + 
+                                condition*I(period_9^2)*trial_c +
+                                condition*I(period_9^2)*I(trial_c^2)  
+                                condition*I(period_9^3) +
+                                trial_c*I(period_9^3) | pid),
+       prior = priors_M4,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M4")
+```
+\normalsize
+
+Estimating model M4 took about 8 hours.
 
@@
 
-Use WAIC to compare models (ref McElreath, Kurz)
+We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
 
-## Plot congruency effects and subject-specific fits for the selected model
+\scriptsize 
+```{r compare-models, echo=T}
+model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds")
+model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds")
+model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds")
+model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds")
+
+#model_M1  <- add_criterion(model_M1, c("loo", "waic"))
+#model_M2  <- add_criterion(model_M2, c("loo", "waic"))
+#model_M3  <- add_criterion(model_M3, c("loo", "waic"))
+#model_M4  <- add_criterion(model_M4, c("loo", "waic"))
+
+loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F)
+loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F)
+
+# model weights
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
+
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
+```
+\normalsize
+
+Clearly, both weighting schemes prefer model M4.
+
+## Plot congruency effects and subject-specific fits for model M4.
 
@@
 
+Figure 5
+
 Plot the model-based hazard and survivor functions.
+
+Figure 6
 
@@
 
+model and comparison and note on fitting
+
+Figure 7
+
 # Tutorial 4: Calculating descriptive statistics when there are two independent variables
@@
 In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
+Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+(ref:descr-fig8-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
+
+```{r plot-2ivs, fig.cap = "(ref:descr-fig8-caption)", out.width='95%'}
+knitr::include_graphics("../Tutorial_4_descriptive_2IV/Plot_2IV_for_subject6_PanisSchmidt.png")
+```
 
@@
 Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the same detailed shape of the empirical RT distributions.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].

### 2024-09-22 manuscript/manuscript.tex
update manuscript

@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 
-In this second tutorial we illustrate how to fit a Bayesian hazard regression model to the masked response priming data set used in the first tutorial.
+In this second tutorial we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
 
-There are several analytic decisions one has to make.
+There are several analytic decisions one has to make when fitting a hazard model.
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
@@
 
-In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (see Tutorial 1); Allison (2010)).
+In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set; Allison (2010)).
 When there is clustering in the data, as in the case of a small-\emph{N} design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models (Allison, 2010).
@@
 
-Estimating model\_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
@@
 
-For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+We first remove unncessary columns and specify our priors.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# remove unnecessary columns}
+\NormalTok{M2\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) }
+
+\CommentTok{\# Specify priors}
+\NormalTok{priors\_M2 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}0.2,0.71,{-}2.2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),      }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{),    }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{),            }
+  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{)                 }
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Now we can fit model 2.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plan}\NormalTok{(multicore)}
+
+\NormalTok{model\_M2 }\OtherTok{\textless{}{-}}
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M2\_data,}
+       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
+\NormalTok{                          (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors\_M2,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
+       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M2"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Estimating model M2 took about 144 minutes.
 
@@
 
+For the third model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables congruency (variable ``condition'') and trial number (variable ``trial\_c'').
+We use the same data set and priors as for model 2.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{M3\_data }\OtherTok{\textless{}{-}}\NormalTok{ M2\_data}
+\NormalTok{priors\_M3 }\OtherTok{\textless{}{-}}\NormalTok{ priors\_M2}
+\FunctionTok{plan}\NormalTok{(multicore)}
+
+\NormalTok{model\_M3 }\OtherTok{\textless{}{-}} 
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M3\_data,}
+       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+} \CommentTok{\# Note that duplicate terms in the model formula are ignored}
+\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors\_M3,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
+       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M3"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Estimating model M3 took about 268 minutes.
+
 \subsection{Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
@@
 Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) and because cognition is likely the behavior of a non-linear dynamical system {[}ref{]}, we relax all three assumptions in model 4.
+We use the same data set and priors as for model 2.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{M4\_data }\OtherTok{\textless{}{-}}\NormalTok{ M2\_data}
+\NormalTok{priors\_M4 }\OtherTok{\textless{}{-}}\NormalTok{ priors\_M2}
+\FunctionTok{plan}\NormalTok{(multicore)}
+
+\NormalTok{model\_M4 }\OtherTok{\textless{}{-}} 
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M4\_data,}
+       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+} \CommentTok{\# Note that duplicate terms in the model formula are ignored}
+\NormalTok{                          condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
+\NormalTok{                          condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                          (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{  condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                                condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)  }
+\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                                trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors\_M4,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
+       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M4"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Estimating model M4 took about 8 hours.
 
@@
 
-Use WAIC to compare models (ref McElreath, Kurz)
-
-\subsection{Plot congruency effects and subject-specific fits for the selected model}\label{plot-congruency-effects-and-subject-specific-fits-for-the-selected-model}
+We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights (Kurz, 2023a; McElreath, 2018).
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_M1 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M1.rds"}\NormalTok{)}
+\NormalTok{model\_M2 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M2.rds"}\NormalTok{)}
+\NormalTok{model\_M3 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M3.rds"}\NormalTok{)}
+\NormalTok{model\_M4 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M4.rds"}\NormalTok{)}
+
+\CommentTok{\#model\_M1  \textless{}{-} add\_criterion(model\_M1, c("loo", "waic"))}
+\CommentTok{\#model\_M2  \textless{}{-} add\_criterion(model\_M2, c("loo", "waic"))}
+\CommentTok{\#model\_M3  \textless{}{-} add\_criterion(model\_M3, c("loo", "waic"))}
+\CommentTok{\#model\_M4  \textless{}{-} add\_criterion(model\_M4, c("loo", "waic"))}
+
+\FunctionTok{loo\_compare}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{criterion =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{print}\NormalTok{(}\AttributeTok{simplify =}\NormalTok{ F)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##          elpd_diff se_diff elpd_loo se_elpd_loo p_loo   se_p_loo looic  
+## model_M4     0.0       0.0 -5094.4     62.9       127.6     4.2  10188.9
+## model_M3   -21.0      10.1 -5115.4     62.5        77.3     2.5  10230.8
+## model_M1  -254.9      24.4 -5349.3     64.8        72.7     1.7  10698.6
+## model_M2  -256.8      23.6 -5351.2     64.8        39.7     1.0  10702.5
+##          se_looic
+## model_M4   125.8 
+## model_M3   125.1 
+## model_M1   129.6 
+## model_M2   129.7
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{loo\_compare}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{criterion =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{print}\NormalTok{(}\AttributeTok{simplify =}\NormalTok{ F)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##          elpd_diff se_diff elpd_waic se_elpd_waic p_waic  se_p_waic waic   
+## model_M4     0.0       0.0 -5092.8      62.9        125.9     4.0   10185.5
+## model_M3   -22.1      10.0 -5114.9      62.5         76.8     2.4   10229.7
+## model_M1  -256.3      24.3 -5349.1      64.8         72.5     1.6   10698.2
+## model_M2  -258.4      23.6 -5351.2      64.8         39.6     1.0   10702.4
+##          se_waic
+## model_M4   125.8
+## model_M3   125.1
+## model_M1   129.6
+## model_M2   129.7
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# model weights}
+\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## model_M1 model_M2 model_M3 model_M4 
+##        0        0        0        1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## model_M1 model_M2 model_M3 model_M4 
+##        0        0        0        1
+\end{verbatim}
+
+\normalsize
+
+Clearly, both weighting schemes prefer model M4.
+
+\subsection{Plot congruency effects and subject-specific fits for model M4.}\label{plot-congruency-effects-and-subject-specific-fits-for-model-m4.}
 
@@
 
+Figure 5
+
 Plot the model-based hazard and survivor functions.
 
+Figure 6
+
 \section{Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
@@
 
+model and comparison and note on fitting
+
+Figure 7
+
 \section{Tutorial 4: Calculating descriptive statistics when there are two independent variables}\label{tutorial-4-calculating-descriptive-statistics-when-there-are-two-independent-variables}
@@
 In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
+Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_4_descriptive_2IV/Plot_2IV_for_subject6_PanisSchmidt}
+
+}
+
+\caption{Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
+\end{figure}
 
@@
 Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the same detailed shape of the empirical RT distributions.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).

### 2024-09-22 Tutorial_2.Rmd
update Tutorial 2

@@
 ```{r save-plot-effects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png", width = 12, height = 12, dpi = 600)
+ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png", width = 12, height = 12, dpi = 300)
 ```
@@
 ```
+
+```{r save-plot-fit-hazard, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M4effects_subject6.png", width = 10, height = 10, dpi = 300)
+```

### 2024-09-22 manuscript/manuscript.Rmd
update manuscript

@@
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+
 As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
+In each example, the mean RT is lower in condition 2 compared to condition 1.
+However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
+In the second example, the effect starts around 400 ms and is gone by 800 ms.
+And in the third example, the effect reverses around 550 ms.
 
@@
 ```
-
-In each example, the mean RT is lower in condition 2 compared to condition 1.
-However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
-In the second example, the effect starts around 400 ms and is gone by 800 ms.
-And in the third example, the effect reverses around 550 ms.
 
@@
 In Tutorial 2 we illustrate how one can fit Bayesian hazard models to the data.
-After selecting the best of four models, we plot the model-based effects and the model fits for a few subjects.
+We discuss possible link functions, and plot the model-based effects of our predictors of interest.
 In Tutorial 3 we illustrate how to fit hazard models in a frequentist framework.
-We compare the model-based effects between Bayesian and frequentist approaches.
+Even though both frameworks generate similar parameter estimates, we note that model convergence is often not obtained in the frequentist framework.
 In Tutorial 4 we illustrate how to calculate the descriptive statistics when there are two independent variables.
@@
 
-Plot the effects of congruent and incongruent for each time bin for the selected model.
+Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial number 1000 for the selected model.
 
-Figure 5
+(ref:descr-fig5-caption) 50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial number 1000.
 
-Plot the model-based hazard and survivor functions.
+```{r plot-prime-effects, fig.cap = "(ref:descr-fig5-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png")
+```
 
-Figure 6
+Figure 6 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+
+(ref:descr-fig6-caption) Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.
+
+```{r plot-hazard-subject6, fig.cap = "(ref:descr-fig6-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_subject6.png")
+```
 
@@
 In this third tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in the first tutorial.
+For illustration purposes, we only fitted model M3 using the function glmer() from the package lme4.
 
-model and comparison and note on fitting
+\scriptsize 
+```{r fit-model, eval=F, echo=T}
+model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
+                           condition*I(period_9^2) + 
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) +
+                           (1 + condition*period_9 +
+                           condition*I(period_9^2) +
+                           condition*I(period_9^3) +
+                           trial_c*period_9 + 
+                           trial_c*I(period_9^2) + 
+                           trial_c*I(period_9^3) | pid),
 
-Figure 7
+              # control parameters, data set, and complementary log-log link function
+              control = glmerControl(optimizer = c("nlminbwrap"),
+                                     optCtrl = list(maxfun=10000000)), 
+              data=M3_data, 
+              family=binomial(link="cloglog"))
+```
+\normalsize
+
+In Figure 7 we compere the parameter estimates of model M3 from brms() with those of glmer().
+
+(ref:descr-fig7-caption) Parameter estimates for model M3 from brms() and glmer().
+
+```{r plot-comparison, fig.cap = "(ref:descr-fig7-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png")
+```
+
+Figure 7 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+This is of course one of the reasons why Bayesian modeling has become so popular in the last years.
 
@@
 
-One important limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
-There is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
-We therefore recommend to design as many trials as possible per experimental condition given the available resources.
-Another limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
+Compared to the orthodox method -- comparing means with ANOVA --, the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
+Another issue that might initially look as a limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+However, as nature does not reveal itself easily, obtaining insight into behavioral dynamics simply requires more data per condition than is usually collected under the orthodox method.
+In general, there is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function; We therefore recommend to design as many trials as possible per experimental condition given the available resources.
 
@@
 The hazard models in this tutorial assume that there is one event of interest.
+For RT data, this event consitutes a single transition between an "idle" state and a "responded" state.
 However, in certain situations, more than one event of interest might exist.
-For example, ...
-- multi-state hazard models refs
+For example, an individual might transition back and forth between a "healthy" state and a "depression" state, before being absorbed in a final "death" state.
+When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (refs).
+Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
+Another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
 

### 2024-09-22 manuscript/manuscript.tex
update manuscript

@@
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+
 As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Figures/rt_haz_combi_2}
-
-}
-
-\caption{Means versus distributional shapes.}\label{fig:plot1}
-\end{figure}
-
 In each example, the mean RT is lower in condition 2 compared to condition 1.
@@
 And in the third example, the effect reverses around 550 ms.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Figures/rt_haz_combi_2}
+
+}
+
+\caption{Means versus distributional shapes.}\label{fig:plot1}
+\end{figure}
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 
-Plot the effects of congruent and incongruent for each time bin for the selected model.
-
-Figure 5
-
-Plot the model-based hazard and survivor functions.
-
-Figure 6
+Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial number 1000 for the selected model.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000}
+
+}
+
+\caption{50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial number 1000.}\label{fig:plot-prime-effects}
+\end{figure}
+
+Figure 6 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_subject6}
+
+}
+
+\caption{Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.}\label{fig:plot-hazard-subject6}
+\end{figure}
 
@@
 In this third tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in the first tutorial.
-
-model and comparison and note on fitting
-
-Figure 7
+For illustration purposes, we only fitted model M3 using the function glmer() from the package lme4.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_M3\_f }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(event }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
+
+              \CommentTok{\# control parameters, data set, and complementary log{-}log link function}
+              \AttributeTok{control =} \FunctionTok{glmerControl}\NormalTok{(}\AttributeTok{optimizer =} \FunctionTok{c}\NormalTok{(}\StringTok{"nlminbwrap"}\NormalTok{),}
+                                     \AttributeTok{optCtrl =} \FunctionTok{list}\NormalTok{(}\AttributeTok{maxfun=}\DecValTok{10000000}\NormalTok{)), }
+              \AttributeTok{data=}\NormalTok{M3\_data, }
+              \AttributeTok{family=}\FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+In Figure 7 we compere the parameter estimates of model M3 from brms() with those of glmer().
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_3_Frequentist/comparison}
+
+}
+
+\caption{Parameter estimates for model M3 from brms() and glmer().}\label{fig:plot-comparison}
+\end{figure}
+
+Figure 7 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+This is of course one of the reasons why Bayesian modeling has become so popular in the last years.
 
@@
 The hazard models in this tutorial assume that there is one event of interest.
+For RT data, this event consitutes a single transition between an ``idle'' state and a ``responded'' state.
 However, in certain situations, more than one event of interest might exist.
-For example, \ldots{} - multi-state hazard models refs
+For example, an individual might transition back and forth between a ``healthy'' state and a ``depression'' state, before being absorbed in a final ``death'' state.
+When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (refs).
+Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
+Another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
 

### 2024-09-22 Tutorial_2.Rmd
increase figure text size

@@
   geom_vline(xintercept = inverse_cloglog(0), linetype="dotted", linewidth = 0.3) +
-  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90) +
-  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90) +
+  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90, size=4) +
+  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90,size=4) +
   scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
@@
   theme(panel.grid = element_blank(),
-        axis.text=element_text(size=12),
-        axis.title=element_text(size=14),
+        axis.text=element_text(size=16),
+        axis.title=element_text(size=18),
       #  legend.position="top",
-        legend.text = element_text(size=14),
+        legend.text = element_text(size=18),
       legend.position = "inside",
@@
 ```{r save-plot-links}
-ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 12, height = 10, dpi = 600)
+ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 300)
 ```

### 2024-09-22 manuscript/manuscript.Rmd
update manuscript

@@
 In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
@@
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
+We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, and for detailed treatment, see @singerAppliedLongitudinalData2003, @allisonDiscreteTimeMethodsAnalysis1982, and @allisonSurvivalAnalysisUsing2010.
@@
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
-In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published data set when there is one independent variable.
+In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published RT data set when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
@@
 Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function p(t) = h(t).S(t-1) and the cumulative distribution function F(t) = 1-S(t).
-The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence.
+Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
 
@@
 
-We focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-*N* designs).
+Here we focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-*N* designs).
 This  approach  emphasizes  the  precision  and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
@@
 
-`r my_r_citation$pkgs` <!-- This creates the footnote -->
+`r my_r_citation$pkgs`
+
+<!-- This creates the footnote -->
 
@@
 The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event *non*occurrence.
-An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
+An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption.
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
 
@@
 
-In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set; @allisonSurvivalAnalysisUsing2010).
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X~1~) can be written as follows:
+
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~].
+
+In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set [@allisonSurvivalAnalysisUsing2010].
 When there is clustering in the data, as in the case of a small-*N* design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models [@allisonSurvivalAnalysisUsing2010].
@@
 
-To get the posterior distribution of each parameters given the data, we need to specify the prior distribution of each parameter.
+To get the posterior distribution of each parameters given the data, we need to specify a prior distribution for each parameter.
 The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
-
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
 Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the beta parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
 
@@
 However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
-A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the beta parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
 
@@
 
-Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] and because cognition is likely the behavior of a non-linear dynamical system [ref], we relax all three assumptions in model 4.
+Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022], we relax all three assumptions in model 4.
 We use the same data set and priors as for model 2.
@@
 
-In Figure 7 we compere the parameter estimates of model M3 from brms() with those of glmer().
+In Figure 7 we compere the parameter estimates of model M3 from brm() with those of glmer().
 
-(ref:descr-fig7-caption) Parameter estimates for model M3 from brms() and glmer().
+(ref:descr-fig7-caption) Parameter estimates for model M3 from brm() and glmer().
 
@@
 
-(ref:descr-fig8-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
+(ref:descr-fig8-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
 
@@
 We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
-By providing this tutorial, we hope that researchers will start using hazard analysis more often, due to the many advantages described below.
-While we focus here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+By providing this tutorial, we hope that researchers will start using descriptive and inferential hazard analysis more often, due to the many advantages described below.
+While we focused here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+We also discuss the issue of individual differences, limitations, and extensions.
 
@@
 Hazard functions are thus more diagnostic than density functions” (p.
-331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in ....).
+331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
 
@@
 For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
-Thus, because the discrete-time hazard function identifies unique information about event occurrence in each bin, while F(t) accumulates the complements of the hazard estimates from the current and previous bins -- S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)] --, instead of using delta-plots -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions (see ref).
+As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions [see @panisHowCanWe2020].
 
@@
 Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
@@
 Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
-But even when using a response deadline, participants can differ qualitatively in the effects they display (refs).
+But even when using a response deadline, participants can differ qualitatively in the effects they display [see @panisHowCanWe2020].
 One way to deal with this is to describe and interpret the different patterns.
 Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
-The obtained dendrogram can then be used to identify (hopefully big) clusters of participants that behave similarly, and to identify (hopefully small) clusters of participants with outlying behavioral patterns.
+The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with outlying behavioral patterns.
 One might then exclude the outlying participants before fitting a hazard model.
 
-## Limitations
+## Limitation(s)
 
@@
 For example, an individual might transition back and forth between a "healthy" state and a "depression" state, before being absorbed in a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (refs).
+When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states [@SteeleF2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
 
 # Conclusions
-see paper...
+
+RT and accuracy distributions are a rich source of information on the time course of cognitive processing.
+The changing effects of our experimental manipulations with increases in waiting time become strikingly clear when looking at response hazards and microlevel speed-accuracy trade-off functions.
+We hope that experimental psychologists and cognitive neuroscientists are less reluctant to embrace event history analysis after reading this tutorial.
 

### 2024-09-22 manuscript/manuscript.tex
update manuscript

@@
 In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves over increasing waiting time, and whether its onset is time-locked to other events.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events.
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
@@
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of hazard analysis to orient the reader to the basic concepts and ideas that we will use throughout the paper.
+We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, and for detailed treatment, see Singer and Willett (2003), Allison (1982), and Allison (2010).
@@
 Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t) and the discrete-time survivor function S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]}, and not on the probability mass function p(t) = h(t).S(t-1) and the cumulative distribution function F(t) = 1-S(t).
-The discrete-time hazard probability function gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence.
+Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t) and the discrete-time survivor function S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]}, and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
 
@@
 
-We focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-\emph{N} designs).
+Here we focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-\emph{N} designs).
 This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event \emph{non}occurrence.
-An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption (see below).
+An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption.
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X1, X2) and the cloglog link function can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
 
@@
 
-In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software (after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set; Allison (2010)).
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
+
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.
+
+In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (Allison, 2010).
 When there is clustering in the data, as in the case of a small-\emph{N} design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models (Allison, 2010).
@@
 
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
-
 To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
 Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the beta parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
 
@@
 
-Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) and because cognition is likely the behavior of a non-linear dynamical system {[}ref{]}, we relax all three assumptions in model 4.
+Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
 We use the same data set and priors as for model 2.
@@
 
-In Figure 7 we compere the parameter estimates of model M3 from brms() with those of glmer().
+In Figure 7 we compere the parameter estimates of model M3 from brm() with those of glmer().
 
@@
 
-\caption{Parameter estimates for model M3 from brms() and glmer().}\label{fig:plot-comparison}
+\caption{Parameter estimates for model M3 from brm() and glmer().}\label{fig:plot-comparison}
 \end{figure}
@@
 
-\caption{Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
+\caption{Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
 \end{figure}
@@
 We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
-By providing this tutorial, we hope that researchers will start using hazard analysis more often, due to the many advantages described below.
-While we focus here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+By providing this tutorial, we hope that researchers will start using descriptive and inferential hazard analysis more often, due to the many advantages described below.
+While we focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+We also discuss the issue of individual differences, limitations, and extensions.
 
@@
 First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in \ldots.).
+Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
 
@@
 For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
-Thus, because the discrete-time hazard function identifies unique information about event occurrence in each bin, while F(t) accumulates the complements of the hazard estimates from the current and previous bins -- S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} --, instead of using delta-plots -- differences in quantiles from F(t)\(^-1\) -- one can simply plot delta-h(t) functions (see ref).
+As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)\(^-1\) -- one can simply plot delta-h(t) functions (see Panis, 2020).
 
@@
 Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
@@
 Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
-But even when using a response deadline, participants can differ qualitatively in the effects they display (refs).
+But even when using a response deadline, participants can differ qualitatively in the effects they display (see Panis, 2020).
 One way to deal with this is to describe and interpret the different patterns.
 Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
-The obtained dendrogram can then be used to identify (hopefully big) clusters of participants that behave similarly, and to identify (hopefully small) clusters of participants with outlying behavioral patterns.
+The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with outlying behavioral patterns.
 One might then exclude the outlying participants before fitting a hazard model.
 
-\subsection{Limitations}\label{limitations}
+\subsection{Limitation(s)}\label{limitations}
 
@@
 For example, an individual might transition back and forth between a ``healthy'' state and a ``depression'' state, before being absorbed in a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (refs).
+When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
 
@@
 
-see paper\ldots{}
+RT and accuracy distributions are a rich source of information on the time course of cognitive processing.
+The changing effects of our experimental manipulations with increases in waiting time become strikingly clear when looking at response hazards and microlevel speed-accuracy trade-off functions.
+We hope that experimental psychologists and cognitive neuroscientists are less reluctant to embrace event history analysis after reading this tutorial.
 
@@
 
+\bibitem[\citeproctext]{ref-SteeleF2004}
+Steele, F., Goldstein, H., \& Browne, W.
+(2004).
+A general multilevel multistate competing risks model for event history data, with an application to a study of contraceptive use dynamics.
+\emph{Statistical Modelling}, \emph{4}(2), 145--159.
+\url{https://doi.org/10.1191/1471082X04st069oa}
+
 \bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}

### 2024-09-22 manuscript/manuscript.tex
update manuscript

@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.

### 2024-09-25 sims/sims.Rmd
added a folder and file to simulate hazard data and make some illustrative plots for Figure 1

---
title: "sims"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file simulates the data used in Figure 1 of the manuscript and then creates some plots.

# load the libraries that we will be using #

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "truncnorm", "faux", "ggpmisc")

lapply(pkg, library, character.only = TRUE)
```

## plot settings ##

theme settings for ggplot

```{r plot-settings}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# section 1 #

## create a baseline hazard function ##

set up and plot a user-defined discrete-time hazard function.
This hazard function takes on a reasonable shape for speeded rt tasks.
The shape serves as a reference point that other effects simulated conditions can be compared against.

```{r}
## define some variables
followup <- 1000 # length of data collection period in ms for each trial
nr_bins  <- 10
bin_width <- followup/nr_bins
bin_endpoints <- (1:nr_bins)*bin_width

## define the hazard function at each of 10 time bins
hazard_cond1 <- c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39)

## create a df
data_h <- tibble(
  hazard = hazard_cond1, 
  time = bin_endpoints
  )

## a quick plot to take a look
p1.1 <- ggplot(data_h, aes(x = time, y = hazard)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = bin_endpoints) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("endpoints of discrete time bins (ms)") +
  ylab("discrete-time hazard probability")
p1.1
```

# section 2 #

## create a function that can simulate hazrad functions between conditions ##

this is to simplify as much as possible and to use it to illustrate one key message. e.g., mean conceals what time reveals

Just two conditions.

### first, let's do it step-by-step (without a function) to check the process ###

define parameters

```{r}
subj_n = 10
rep_n = 200
cond_n = 2
haz_n = 10
bin_n = 10
hazard_cond1 <- c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39)
hazard_cond2 <- hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2)
hazard <- c(hazard_cond1, hazard_cond2)
```

setup the data

```{r}
# make it reproducible
set.seed(1)

# set up data structure
dat_test <- add_random(subj = subj_n, rep = rep_n) %>% 
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
  add_within("subj", bin = 1:bin_n) %>% 
  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
         event = rbinom(n(), 1, hazf)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()

head(dat_test)
str(dat_test)
summary(dat_test)
```

take a look

data check

```{r}
dat_test %>%
  distinct(outcomel)

dat_test %>%
  tally(outcomel)

dat_test %>%
  distinct(outcomen)
```

plot

```{r}
p2.1 <- ggplot(dat_test, aes(x = bin, y = hazf, colour = condition)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bins") +
  ylab("discrete-time hazard probability")
p2.1
```

create summary data

at the pid level

```{r}
pid_dat_test <- dat_test %>% 
  group_by(subj, condition, bin) %>% 
  summarise(outcomen = mean(outcomen, na.rm=TRUE),
            n=n())
pid_dat_test
```

at the group level

```{r}
group_dat_test <- pid_dat_test %>% 
  group_by(condition, bin) %>% 
  summarise(group_outcomen = mean(outcomen, na.rm=TRUE),
            n = n(),
            group_sd = sd(outcomen, na.rm = TRUE),
            sem = (group_sd/sqrt(n)))
group_dat_test
```

plot pid and group summary data together

```{r}
p2.2 <- ggplot(group_dat_test, aes(x = bin, y = group_outcomen, 
                   colour = condition, fill=condition)) +
  geom_line(linewidth=1) + 
  geom_point(size=3) +
  geom_jitter(data=pid_dat_test, aes(y=outcomen), height=0, width = 0.1, alpha = 0.5) +
  geom_errorbar(aes(ymin = group_outcomen-sem, ymax = group_outcomen+sem),
                width=.2, linewidth=1) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x="timebin",
       y="probability")
p2.2
```

### now write a function to simulate hazard data ###

```{r}
sim_haz <- function(subj_n = 10, rep_n = 200, bin_n = 10, cond_n = 2,  # these can be changed when calling the f
               hazard_cond1 = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
               hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2),
               hazard = c(hazard_cond1, hazard_cond2),   # 
                ... # helps the function work with pmap() below
                ) {

  # set up data structure
  data <- add_random(subj = subj_n, rep = rep_n) %>% 
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
  add_within("subj", bin = 1:bin_n) %>% 
  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
         event = rbinom(n(), 1, hazf)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()

  # glimpse(data) # only use this when testing the code
}
```

Here’s a quick example of how our function works. You can change these parameters
and create some different data.

```{r}
sim_haz(subj_n = 20, rep_n = 200) # if you uncomment glimpse above,
# it will let you glimpse the data that's generated. this is useful for checking / testing code purposes.
```

test it by varying the subj_n

```{r}
data_n10 <- sim_haz(subj_n = 10, rep_n = 200)
data_n20 <- sim_haz(subj_n = 20, rep_n = 200)
```

data check

```{r}
data_n10 %>%
  distinct(subj, rep, condition)
```

a quick plot to check

```{r}
p2.3 <- ggplot(data_n10, aes(x = bin, y = hazf, colour = condition)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bins") +
  ylab("discrete-time hazard probability")
p2.3
```

ok, this seems to work, as we would expect.

# section 3 #

## convert hazard data into rt data ##

rt at the trial level

use a truncated normal distribution to privude upper and lower bounds

```{r}
rt_dat_test <- data_n10 %>%
  select(-contrast, -event, -cumsum, -outcomel) %>% 
  mutate(u = bin*100,
         l = u-100,
         m = (l + u)/2) %>% 
  filter(outcomen > 0) %>% 
  group_by(subj, rep, condition, bin) %>% 
  mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
head(rt_dat_test)
summary(rt_dat_test)
```

data check

```{r}
rt_dat_test %>% 
  group_by(condition, bin) %>% 
  summarise(min = min(rt),
            max=max(rt),
            mean=mean(rt))

rt_dat_test %>% 
  group_by(condition, bin) %>% 
  tally()
```

this looks sensible.

plot

```{r}
ggplot(rt_dat_test, aes(x=rt, colour = condition)) +
   geom_density()

rt_dat_test %>% 
  mutate(bin = factor(bin)) %>% 
  ggplot(aes(x=rt, colour = bin)) +
  geom_density() +
  facet_wrap(~condition)
```

create summary data per condition at the pid and group level for plotting mean comparisons

at the pid level

```{r}
rt_dat_pid <- rt_dat_test %>%
  group_by(subj, condition) %>% 
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # 
            sem = (sd/sqrt(n)))
rt_dat_pid
```

at the group level

```{r}
rt_dat_group <- rt_dat_test %>%
  group_by(condition) %>% 
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = length(unique(subj)), # 
            sem = (sd/sqrt(n)))
rt_dat_group
```

violin plot

```{r}
p3.1 <- ggplot(rt_dat_pid, aes(x=condition, y = mean, 
                             fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = rt_dat_group, 
             aes(y = mean), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = rt_dat_group,
                aes(y = mean, ymin = mean-sem, ymax = mean +sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = rt_dat_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   ggtitle("mean rt by condition") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   ylab("mean rt (ms)")
p3.1
```

plot together

```{r}
p3.2 <- p3.1 / p2.2
p3.2
```

# section 4 #

## create functions to automate repeat processes ##

These can then be used below to simulate, wrangle and plot the data

### create a function to summarise hazard data ###

at the pid level

```{r}
haz_pid <- function(df) {
  df %>% 
    group_by(subj, condition, bin) %>% 
    summarise(outcomen = mean(outcomen, na.rm=TRUE),
              n=n(),
              .groups = "drop")
}
```

at the group level

```{r}
haz_group <- function(df) {
  df %>% 
    group_by(condition, bin) %>% 
    summarise(group_outcomen = mean(outcomen, na.rm=TRUE),
              n = n(),
              group_sd = sd(outcomen, na.rm = TRUE),
              sem = (group_sd/sqrt(n)),
              .groups = "drop")
}
```

test them

```{r}
haz_pid_ftest <- haz_pid(data_n10)
haz_pid_ftest

haz_group_ftest <- haz_group(haz_pid_ftest)
haz_group_ftest
```

### create a function to convert hazards into rt data ###

```{r}
rt_dat <- function(df) {
  df %>%
    select(-contrast, -event, -cumsum, -outcomel) %>% 
    mutate(u = bin*100,
         l = u-100,
         m = (l + u)/2) %>% 
    filter(outcomen > 0) %>% 
    group_by(subj, rep, condition, bin) %>% 
    mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u)) %>% 
    ungroup()
}
```

test it

```{r}
rt_dat_ftest <- rt_dat(data_n10)
head(rt_dat_ftest)
glimpse(rt_dat_ftest)
```

### create a function to summarise rt data ###

at the pid level

```{r}
rt_pid <- function(df) {
  df %>% 
    group_by(subj, condition) %>% 
    summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # 
            sem = (sd/sqrt(n)),
            .groups = "drop")
}
```

at the group level

```{r}
rt_group <- function(df) {
  df %>% 
    group_by(condition) %>% 
    summarise(mean = mean(rt),
              sd = sd(rt), 
              n = length(unique(subj)),
              sem = (sd/sqrt(n)),
              .groups="drop")
}
```

test them

```{r}
rt_pid_ftest <- rt_pid(rt_dat_ftest)
rt_pid_ftest

rt_group_ftest <- rt_group(rt_dat_ftest)
rt_group_ftest
```

### create a function to plot the hazard data ###

hazard plot function

```{r}
haz_plot <- function(df_pid, df_group) {
  ggplot(df_group, aes(x = bin, y = group_outcomen, 
                   colour = condition, fill=condition)) +
  geom_line(linewidth=1) + 
  geom_point(size=3) +
  geom_jitter(data=df_pid, aes(y=outcomen), height=0, width = 0.1, alpha = 0.5) +
  geom_errorbar(aes(ymin = group_outcomen-sem, ymax = group_outcomen+sem),
                width=.2, linewidth=1) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x="timebin",
       y="probability")
}
```

test it

```{r}
p_haz_test <- haz_plot(haz_pid_ftest, haz_group_ftest)
p_haz_test
```

### create a function to plot the rt data ###

mean rt plot function

```{r}
rt_plot <- function(df_pid, df_group) {
   ggplot(df_pid, aes(x=condition, y = mean, 
                      fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = df_group, 
             aes(y = mean), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = df_group,
                aes(y = mean, ymin = mean-sem, ymax = mean+sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = df_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   ylab("mean rt (ms)")
}
```

test it

```{r}
p_rt_test <- rt_plot(rt_pid_ftest, rt_group_ftest)
p_rt_test
```

# section 5 #

## use the function to simulate A=B for the mean, but then some different distributions ##

simulate hazard data

```{r}
## make it reproducible
set.seed(123)

## mean rt = same, haz = same
haz1 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1))

## mean rt = same, haz = later difference, B > A
haz2 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 1.8, 1.8, 1.5, 1))

## mean rt = same, haz = later difference, A > B
haz3 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 0.2, 0.2, 0.5, 1))
```

create summary data

```{r}
haz1_pid <- haz_pid(haz1) %>% 
  mutate(example = "example 1")
haz1_pid

haz1_group <- haz_group(haz1_pid) %>% 
  mutate(example = "example 1")
haz1_group

haz2_pid <- haz_pid(haz2) %>% 
  mutate(example = "example 2")
haz2_pid

haz2_group <- haz_group(haz2_pid) %>% 
  mutate(example = "example 2")
haz2_group

haz3_pid <- haz_pid(haz3) %>% 
  mutate(example = "example 3")
haz3_pid

haz3_group <- haz_group(haz3_pid) %>% 
  mutate(example = "example 3")
haz3_group
```

plot the hazard data

```{r}
p5.1 <- haz_plot(haz1_pid, haz1_group)
p5.1

p5.2 <- haz_plot(haz2_pid, haz2_group)
p5.2

p5.3 <- haz_plot(haz3_pid, haz3_group)
p5.3
```

convert to rt

```{r}
## make it reproducible
set.seed(123)

rt1 <- rt_dat(haz1)
head(rt1)

rt2 <- rt_dat(haz2)
head(rt2)

rt3 <- rt_dat(haz3)
head(rt3)
```

summarise the rt data

```{r}
rt1_pid <- summary_pid(rt1) %>%
  mutate(example = "example 1")

rt1_group <- summary_group(rt1) %>% 
  mutate(example = "example 1")

rt2_pid <- summary_pid(rt2) %>% 
  mutate(example = "example 2")

rt2_group <- summary_group(rt2) %>% 
  mutate(example = "example 2")

rt3_pid <- summary_pid(rt3) %>% 
  mutate(example = "example 3")

rt3_group <- summary_group(rt3) %>% 
  mutate(example = "example 3")
```

plot the rt data

```{r}
p5.4 <- rt_plot(rt1_pid, rt1_group)
p5.4

p5.5 <- rt_plot(rt2_pid, rt2_group)
p5.5

p5.6 <- rt_plot(rt3_pid, rt3_group)
p5.6
```

## combine the data and plot together ##

```{r}
## hazard
haz_pid_123 <- bind_rows(haz1_pid, haz2_pid, haz3_pid) %>% 
  mutate(example=factor(example))
head(haz_pid_123)

haz_group_123 <- bind_rows(haz1_group, haz2_group, haz3_group) %>% 
  mutate(example=factor(example))
head(haz_group_123)

## rt
rt_pid_123 <- bind_rows(rt1_pid, rt2_pid, rt3_pid) %>% 
  mutate(example=factor(example))
head(rt_pid_123)

rt_group_123 <- bind_rows(rt1_group, rt2_group, rt3_group) %>% 
  mutate(example=factor(example))
head(rt_group_123)
```

plot

hazard

```{r}
p5.7 <- haz_plot(haz_pid_123, haz_group_123) +
  facet_wrap(~example)
p5.7
```

rt

```{r}
p5.8 <- rt_plot(rt_pid_123, rt_group_123) +
  facet_wrap(~example)
p5.8
```

combine plots

```{r}
p5.9 <- p5.8 / p5.7
p5.9
```

# section 6 #

## use the function to simulate A>B for the mean, but then some different distributions ##

simulate hazard data

```{r}
## make it reproducible
set.seed(456)

## mean rt A>B, haz = early diff
haz4 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 3, 3, 3, 2, 1, 1, 1, 1, 1))

## mean rt A>B, haz = middle diff
haz5 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 2.5, 1.7, 1.7, 1.2, 1, 1, 1))

## mean rt A>B, haz = early diff and later reversal
haz6 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 3, 3, 3, 2, 0.5, 0.4, 0.3, 0.2, 0.2))
```

create summary data

```{r}
haz4_pid <- haz_pid(haz4) %>% 
  mutate(example = "example 4")
haz4_pid

haz4_group <- haz_group(haz4_pid) %>% 
  mutate(example = "example 4")
haz4_group

haz5_pid <- haz_pid(haz5) %>% 
  mutate(example = "example 5")
haz5_pid

haz5_group <- haz_group(haz5_pid) %>% 
  mutate(example = "example 5")
haz5_group

haz6_pid <- haz_pid(haz6) %>% 
  mutate(example = "example 6")
haz6_pid

haz6_group <- haz_group(haz6_pid) %>% 
  mutate(example = "example 6")
haz6_group
```

plot the hazard data

```{r}
p6.1 <- haz_plot(haz4_pid, haz4_group)
p6.1

p6.2 <- haz_plot(haz5_pid, haz5_group)
p6.2

p6.3 <- haz_plot(haz6_pid, haz6_group)
p6.3
```

convert to rt

```{r}
## make it reproducible
set.seed(456)

rt4 <- rt_dat(haz4)
head(rt4)

rt5 <- rt_dat(haz5)
head(rt5)

rt6 <- rt_dat(haz6)
head(rt6)
```

summarise the rt data

```{r}
rt4_pid <- summary_pid(rt4) %>%
  mutate(example = "example 4")

rt4_group <- summary_group(rt4) %>% 
  mutate(example = "example 4")

rt5_pid <- summary_pid(rt5) %>% 
  mutate(example = "example 5")

rt5_group <- summary_group(rt5) %>% 
  mutate(example = "example 5")

rt6_pid <- summary_pid(rt6) %>% 
  mutate(example = "example 6")

rt6_group <- summary_group(rt6) %>% 
  mutate(example = "example 6")
```

plot the rt data

```{r}
p6.4 <- rt_plot(rt4_pid, rt4_group)
p6.4

p6.5 <- rt_plot(rt5_pid, rt5_group)
p6.5

p6.6 <- rt_plot(rt6_pid, rt6_group)
p6.6
```

## combine the data and plot together ##

```{r}
## hazard
haz_pid_456 <- bind_rows(haz4_pid, haz5_pid, haz6_pid) %>% 
  mutate(example=factor(example))
head(haz_pid_456)

haz_group_456 <- bind_rows(haz4_group, haz5_group, haz6_group) %>% 
  mutate(example=factor(example))
head(haz_group_456)

## rt
rt_pid_456 <- bind_rows(rt4_pid, rt5_pid, rt6_pid) %>% 
  mutate(example=factor(example))
head(rt_pid_456)

rt_group_456 <- bind_rows(rt4_group, rt5_group, rt6_group) %>% 
  mutate(example=factor(example))
head(rt_group_456)
```

plot

hazard

```{r}
p6.7 <- haz_plot(haz_pid_456, haz_group_456) +
  facet_wrap(~example)
p6.7
```

rt

```{r}
p6.8 <- rt_plot(rt_pid_456, rt_group_456) +
  facet_wrap(~example)
p6.8
```

combine plots

```{r}
p6.9 <- p6.8 / p6.7
p6.9
```

big plot

```{r}
# p6.10 <- (p5.9 | p6.9) +
#   plot_layout(axes = 'collect')
# p6.10
```

# section 7 #

## combine all examples into one df and plot ##

### wrangle ###

```{r}
# haz
haz_pid_all <- bind_rows(haz_pid_123, haz_pid_456) %>% 
  mutate(example=factor(example))
head(haz_pid_all)

haz_group_all <- bind_rows(haz_group_123, haz_group_456) %>% 
  mutate(example=factor(example))
head(haz_group_all)

## rt
rt_pid_all <- bind_rows(rt_pid_123, rt_pid_456) %>% 
  mutate(example=factor(example))
head(rt_pid_all)

rt_group_all <- bind_rows(rt_group_123, rt_group_456) %>% 
  mutate(example=factor(example))
head(rt_group_all)
```

### plot ###

hazard

```{r}
p7.1 <- haz_plot(haz_pid_all, haz_group_all) +
  facet_wrap(~example, nrow = 1)
p7.1
```

rt

```{r}
p7.2 <- rt_plot(rt_pid_all, rt_group_all) +
  facet_wrap(~example, nrow = 1)
p7.2
```

combine plots

```{r}
p7.3 <- (p7.2 / p7.1) +
  plot_annotation(tag_levels = 'A')
p7.3

ggsave("sims/figures/rt_haz.jpeg",
       width = 12, height = 8, dpi=800)
```

## use patchwork to combine prior plots ##

```{r}
# p7.4 <- (p5.9 / p6.9) 
#   # plot_annotation(tag_levels = 'A')
# p7.4
# 
# ggsave("sims/figures/rt_haz_2.jpeg",
#        width = 12, height = 8, dpi=800)
```

# section 8 #

## try to inset rt within the hazard plot panels ##

create a haz plot with two rows

```{r}
p8.1 <- haz_plot(haz_pid_all, haz_group_all) +
  facet_wrap(~example, nrow = 2)
p8.1
```

create a rt plot two rows

```{r}
p8.2 <- rt_plot(rt_pid_all, rt_group_all) +
  facet_wrap(~example, nrow = 2)
p8.2
```

now let's try using ggpmisc to inset rt within the haz plots

## let's first try a simpler example ##

the rt plot should be simpler and have smaller text, if it is going to be an inset plot

let's create a couple of mean rt plots. and then facet by hazard.

### create a function for simpler rt plots ###

a simpler mean rt plot function

```{r}
rt_plot_inset <- function(df_pid, df_group) {
   ggplot(df_pid, aes(x=condition, y = mean, 
                      fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = df_group, 
             aes(y = mean), size = 2, position=pd2, colour="black") +
   geom_errorbar(data = df_group,
                aes(y = mean, ymin = mean-sem, ymax = mean+sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = df_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   labs(y="",
        x="") +
   theme_bw() +
   theme(legend.position = "none",
         text = element_text(face = "bold"),
         panel.background = element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank())
}
```

test it

```{r}
p_rt_inset_test <- rt_plot_inset(rt_pid_ftest, rt_group_ftest)
p_rt_inset_test
```

now create a df with the ggplot object in there, so it can be used later

```{r}
inset <- tibble(x = 0, y = 1,
             plot = list(p_rt_inset_test))
```

combine it with a hazard plot

using geom_plot

```{r}
p_haz_inset <- p2.2 +
  geom_plot(data = inset, 
            vp.width = 1/4,
            aes(x, y, label = plot))
p_haz_inset

# ggsave("sims/figures/haz_inset.jpeg",
#        width = 8, height = 8, dpi=800)
```

### now incorporate a facet ###

create six examples of the simpler rt plots

```{r}
p8.1s <- rt_plot_inset(rt1_pid, rt1_group)
p8.1s

p8.2s <- rt_plot_inset(rt2_pid, rt2_group)
p8.2s

p8.3s <- rt_plot_inset(rt3_pid, rt3_group)
p8.3s

p8.4s <- rt_plot_inset(rt4_pid, rt4_group)
p8.4s

p8.5s <- rt_plot_inset(rt5_pid, rt5_group)
p8.5s

p8.6s <- rt_plot_inset(rt6_pid, rt6_group)
p8.6s
```

create the inset data

```{r}
inset_facet <- tibble(x = rep(0, times = 6),
                      y = rep(1, times = 6), 
                      example = factor(c("example 1", "example 2", "example 3",
                                     "example 4", "example 5", "example 6")),
                      plot = list(p8.1s, p8.2s, p8.3s, p8.4s, p8.5s, p8.6s))
```

add the inset to the main plot

```{r}
p_haz_inset_facet <- p8.1 +
  geom_plot(data = inset_facet,
            vp.width = 1/3,
            aes(x, y, label = plot))
p_haz_inset_facet

ggsave("sims/figures/haz_inset_facet.jpeg",
       width = 14, height = 10, dpi=800)
```

### 2024-10-01 sims/sims.Rmd
added code to simulate data and create figures

@@
 
-## create a function that can simulate hazrad functions between conditions ##
+## create a function that can simulate hazard functions between conditions ##
 
@@
 ```{r}
-rt1_pid <- summary_pid(rt1) %>%
+rt1_pid <- rt_pid(rt1) %>%
   mutate(example = "example 1")
 
-rt1_group <- summary_group(rt1) %>% 
+rt1_group <- rt_group(rt1) %>% 
   mutate(example = "example 1")
 
-rt2_pid <- summary_pid(rt2) %>% 
+rt2_pid <- rt_pid(rt2) %>% 
   mutate(example = "example 2")
 
-rt2_group <- summary_group(rt2) %>% 
+rt2_group <- rt_group(rt2) %>% 
   mutate(example = "example 2")
 
-rt3_pid <- summary_pid(rt3) %>% 
+rt3_pid <- rt_pid(rt3) %>% 
   mutate(example = "example 3")
 
-rt3_group <- summary_group(rt3) %>% 
+rt3_group <- rt_group(rt3) %>% 
   mutate(example = "example 3")
@@
 ```{r}
-rt4_pid <- summary_pid(rt4) %>%
+rt4_pid <- rt_pid(rt4) %>%
   mutate(example = "example 4")
 
-rt4_group <- summary_group(rt4) %>% 
+rt4_group <- rt_group(rt4) %>% 
   mutate(example = "example 4")
 
-rt5_pid <- summary_pid(rt5) %>% 
+rt5_pid <- rt_pid(rt5) %>% 
   mutate(example = "example 5")
 
-rt5_group <- summary_group(rt5) %>% 
+rt5_group <- rt_group(rt5) %>% 
   mutate(example = "example 5")
 
-rt6_pid <- summary_pid(rt6) %>% 
+rt6_pid <- rt_pid(rt6) %>% 
   mutate(example = "example 6")
 
-rt6_group <- summary_group(rt6) %>% 
+rt6_group <- rt_group(rt6) %>% 
   mutate(example = "example 6")
@@
 ```{r}
-# p6.10 <- (p5.9 | p6.9) +
-#   plot_layout(axes = 'collect')
+# p6.10 <- (p5.9 / p6.9) +
+#   plot_layout(axes = 'collect',
+#               guides = 'collect')
 # p6.10
@@
 ggsave("sims/figures/haz_inset_facet.jpeg",
-       width = 14, height = 10, dpi=800)
-```
+       width = 14, height = 10, dpi=1000)
+```
+
+# section 9 #
+
+## create a figure that illustrates hazard plus conditional accuracy ##
+
+### create the hazard data ###
+
+create hazard data
+
+```{r}
+## make it reproducible
+set.seed(7)
+
+## create the data
+haz7 <- sim_haz(subj_n = 10, rep_n = 400, 
+                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2))
+```
+
+calculate summary hazard data
+
+```{r}
+haz7_pid <- haz_pid(haz7)
+haz7_pid
+
+haz7_group <- haz_group(haz7_pid)
+haz7_group
+```
+
+plot hazard
+
+```{r}
+p9.1 <- haz_plot(haz7_pid, haz7_group) +
+  labs(title = "hazard") +
+  theme(legend.position = "none")
+p9.1
+
+## if you want to remove individual pid points for simplicity
+# p9.1 <- p9.1 %>%  
+#   remove_geom('point')
+# p9.1
+```
+
+### convert to rt data ###
+
+convert to rt data
+
+```{r}
+## make it reproducible
+set.seed(7)
+
+rt7 <- rt_dat(haz7)
+head(rt7)
+```
+
+calculate summary rt data
+
+```{r}
+rt7_pid <- rt_pid(rt7)
+rt7_pid
+
+rt7_group <- rt_group(rt7)
+rt7_group
+```
+
+simple plot rt
+
+```{r}
+# p9.2 <- rt_plot(rt7_pid, rt7_group)
+# p9.2
+
+p9.2 <- rt_plot_inset(rt7_pid, rt7_group)
+p9.2
+```
+
+### inset rt within the hazard plot ###
+
+create the inset data
+
+```{r}
+inset_hazrt <- tibble(
+  x = 1,
+  y = 0.9, 
+  plot = list(p9.2)
+)
+```
+
+and plot
+
+```{r}
+p9.3 <- p9.1 +
+  geom_plot(data = inset_hazrt,
+            vp.width = 1/3,
+            aes(x, y, label = plot))
+p9.3
+```
+
+### accuracy ###
+
+create some data
+
+```{r}
+acc <- tibble(
+  time = seq(1,10,1),
+  cond1 = rep(1,length = 10),
+  cond2 = c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
+) %>% 
+  pivot_longer(-time,
+               names_to = "condition",
+               values_to = "accuracy") %>% 
+  mutate(condition = factor(condition))
+head(acc)
+```
+
+create group level summary data
+
+```{r}
+acc_group <- acc %>% 
+  group_by(condition) %>% 
+  summarise(group_accuracy = mean(accuracy, na.rm = TRUE),
+            n=n(),
+            group_sd=sd(accuracy, na.rm = TRUE),
+            sem = (group_sd/sqrt(n)),
+            .groups = "drop")
+acc_group
+```
+
+line plot per bin
+
+```{r}
+p9.4 <-ggplot(acc, aes(x = time, y = accuracy,
+                         colour = condition)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  xlab("timebin") +
+  ylab("probability") +
+  ggtitle("accuracy") 
+p9.4
+```
+
+mean average plot
+
+```{r}
+p9.5 <- ggplot(acc, aes(x=condition, y = accuracy, 
+                      fill = condition, colour = condition)) +
+   geom_jitter(position=position_jitterdodge(dodge.width =1),
+               alpha = 0.5) +
+   geom_violin(alpha = 0.7) +
+   geom_point(data = acc_group, 
+             aes(y = group_accuracy), size = 2, position=pd2, colour = "black") +
+   geom_errorbar(data = acc_group,
+                aes(y = group_accuracy, ymin = group_accuracy-sem, ymax = group_accuracy+sem),
+                width=.2, position=pd2, colour = "black") +
+   geom_line(data = acc_group, 
+            aes(y = group_accuracy, group=1),
+            position=pd2,  colour = "black") +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   scale_y_continuous(limits = c(0, 1), breaks = seq(0,1,0.25)) +
+   labs(y="",
+        x="") +
+   theme_bw() +
+   theme(legend.position = "none",
+         text = element_text(face = "bold"),
+         panel.background = element_blank(),
+         panel.grid.major = element_blank(),
+         panel.grid.minor = element_blank())
+p9.5
+```
+
+inset mean acc within the acc plot
+
+create the inset data
+
+```{r}
+inset_acc <- tibble(
+  x = 1,
+  y = 0.9, 
+  plot = list(p9.5)
+)
+```
+
+and plot
+
+```{r}
+p9.6 <- p9.4 +
+  geom_plot(data = inset_acc,
+            vp.width = 1/3,
+            aes(x, y, label = plot))
+p9.6
+```
+
+## plot together ##
+
+```{r}
+p9.7 <- (p9.3 / p9.6) +
+  plot_layout(guides = 'collect',
+              axes = 'collect')
+p9.7
+
+## without inset
+# ggsave("sims/figures/haz_acc.jpeg",
+#        width = 8, height = 8, dpi=800)
+
+## with inset
+# ggsave("sims/figures/haz_acc_inset.jpeg",
+#        width = 8, height = 8, dpi=800)
+```
+
+It doesn't make sense to create the inset with this data because the bins (1:10) are just to illustrate the point rather than being tied to actual RTs, like in a real study where bin 3 may represent 250-300ms say.
+
+## create a figure that illustrates hazard and survivor functions ##
+
+create the data
+
+```{r}
+data_hs <- tibble(
+  timebin = seq(1,10,1),
+  hazard = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
+  survivor = c(1, 0.97, 0.91, 0.79, 0.61, 0.39, 0.22, 0.08, 0.04, 0.02)
+) %>% 
+  pivot_longer(-timebin,
+               names_to = "dv",
+               values_to = "probability") %>% 
+  mutate(dv = factor(dv))
+head(data_hs)
+```
+
+plot
+
+```{r}
+p9.7 <-ggplot(data_hs, aes(x = timebin, y = probability,
+                         colour = dv)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  ggtitle("hazard and survivor functions") 
+p9.7
+
+ggsave("sims/figures/haz_surv.jpeg",
+       width = 7, height = 5, dpi=800)
+```

### 2024-10-01 notes/manuscript_notes.Rmd
added notes on the manuscript

@@
 ...
+
+# September 2024 #
+
+I received your draft and made a bunch of edits and the comments are below.
+I added my edits to git and made a commit, then pushed them back.
+
+## general stuff ##
+
+### functions ###
+
+maybe a function directory in supps?
+Each of your user-defined functions can be listed with a description of what it does. Like in an R package?
+
+### references ###
+
+I know why you have the issue with Bengtsson 2021 reference.
+For some reason, it is simply in your .bib file more than once.
+So there is some typo/bug that needs fixing so that the ref is only listed once.
+
+At least most of the the code works fine, it is just your .bib file that needs fixing.
+
+More generally, I'm not sure whether it is better to have the R package REFs in a footnote or in the main text as a paragraph. We can decide later on what works best.
+
+### terms ###
+
+we use the terms hazard and EHA interchangeably. We should pick one, define it and then stick to it.
+
+For example, the first line of the section "overview of hazard analysis" starts with "To apply event history analysis"
+
+### formulas ###
+
+I think we should go "formula free" in the main text, and instead report formulas in supplementary materials. The reason is as I now make clear at the start of sectino 2:
+
+>For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs).
+Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
+Our focus here is not on providing a detailed account of the underlying regression equations, since this has been done many times before.
+Instead, we want to provide an intuition to how EHA works in general and in the context of experimental psychology.
+As such, we supply relevant equations in supplementary materials and refer to them in the text whenever relevant.
+
+I would also add as a general commwent that whenever we do insert formulas in text (in the future or in supplementary material here) we should number them and we should clearly define every part of the equation.
+
+### big computational chunks in the manuscript ###
+
+I know one of the benefits of using Rmarkdown and papaja is that you can have seamless integration of code chunks and prose.
+However, I also value some efficiency and simplicity.
+So, for me, if a code chunk is simple and short e.g., simulate some easy data and plot, then I can see why it is nice to have it in.
+But when the code chunks get really big like the 'load-functions' chunk around line 256, I don't think it works as nicely.
+The reason I say this is at least twofold.
+1) since we are providing the tutorials as separate .Rmd files and .html files, then the underlying code chunks and functions are of course available to everyone.
+So the main text can be the front-end window dressing to get folks to want to jump into and use the code.
+That means, we can just insert images from the tutorial folders, rather than have massive code chunks, which then make the manuscript.Rmd file unnecessarily long.
+2) the second reason is that unless we use 'cache' function or eval=F, then it slows down the knit process considerably as a whole bunch of computations have to be performed every time we generate the pdf.
+And as a process of editing and revision, I like to be able to see how my edits change the output and I like to read the pdf as a doc to see how it flows.
+So, for that, I'd like to keep the manuscript file lightweight and easy to render.
+That means we keep all code chunks that actually compute stuff in the tutorials and then mainly just refer to .jpegs in the main text.
+
+Does that make sense?
+
+### tables ###
+
+tables 1 and 2, which I added, would be better and easier to read if they were either placed side-by-side or incorporated into one table. I'll leave that for you to figure out. You're welcome!!
+
+### summary of the steps involved ###
+
+for each tutorial, I think it would be nice to have a summary of the steps involved, as a roadmap.
+e.g., a list of the steps that need to be taken or a flowchart.
+I think it would be nice to see which chunky bits of stuff need to happen.
+Just an idea to ease the burden of information and give people a guide.
+
+## abstract ##
+
+I need to shorten it.
+
+## Tutorial 1 ##
+
+no need to cite research gate, just cite the paper and put the data on our GitHub project / OSF, so that it is available.
+
+### figure 4 - h(t), S(t), ca(t) ###
+
+the figure that shows all three panels.
+
+Can we add shading that highlights particular time-windows that we refer to in the text?
+I think that would help.
+
+And as we disucssed before, we need to be mindful that folks know exactly what eaxh panels add on its own and/or in combination with the other panels.
+
+### unpack some statements a little more ###
+
+e.g., you say
+
+>Qualitatively similar results were obtained for the other five participants.
+These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+
+Extend this paragraph to make it crystal clear what we mean. Can we unpack the theoretical relevance and bring it back into how EHA is revealing?
+
+And why don't we show the group average data, rather than just PID 6? It seems odd to me without an explanation/justification.
+
+### I removed the following text from this section ###
+
+>Also, in their second Experiment, @panisWhatShapingRT2016 showed that the negative compatibility effect in the mask-present conditions (see Tutorial 4) is time-locked to mask onset.
+This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022; @schmidtResponseInhibitionNegative2022].
+In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+
+This is really interesting, but it doesn't seem to fit in this section. In this section, we just want to tell people how to wrangle and look at the data.
+
+Maybe in a later section we can add in this text to explain the significance or implications.
+
+It also read a little bit like we were making inferential claims from descriptive stats, so it shoudl wait until we get to a later tutorial maybe.
+
+### link function plot ###
+
+this can go in supps, as we are not writing a regression textbook.
+
+### this losing people and we need a way to simplify ###
+
+You say this...
+
+>An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
+
+>cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)].
+
+>The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
+The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
+Predictors can be discrete, continuous, and time-varying or time-invariant.
+For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
+To interpret the effects of a predictor,its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+When using the logit link, exponentiating a $\beta$ parameter results in an odds ratio.
+
+>An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X~1~) can be written as follows:
+
+>cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~].
+
+But we need to bring people along, as that is our stated aim here, so this needs to go in supps.
+Then we can keep the main text accessible, without compromising on the underlying regression equations.
+But these have all been stated in introductory texts so it should not be front and centre for us.
+But it should be there for the interested reader and hence why supps is perfects.
+
+I also suggest we discuss how to interpret the results when we get to the results.
+
+### prior plot ###
+
+again, move this plot to supps.
+It is a beautiful plot and hard-won by you to get this right.
+But move it to supps and let people get on with their lives!
+If they want to see it, then they can.
+But we need to keep the flow and momentum in the main text.
+
+>To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+
+>To gain a sense of what prior *cloglog* values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
+Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
+
+### no need to output code for priors and such like ###
+
+people are not learning anything by you showing them the code that lists priors.
+Instead, give them an intuition about the general approach and then they can see the implementation in the raw tutorial code.
+e.g., explain the justification and approach to priors in simple, overview terms, and then let them read the code if they want to know how to set them.
+
+### avoid repetition in the model building steps ###
+
+for model 1 in Tutorial 2 - the first inferential model - we should outline the key steps.
+e.g., read in the data, set priors, build the model, evaluate the model via model comparison and parameter estimates.
+But, as I said in the previous subsection, there is no need to show code chunks for most fo these steps.
+Instead, I would focus on the general steps and procedures and let the tutorial .Rmd and .html files present the code.
+
+Then, as we move to Model 2, we only mention what differs between models 1 and 2. And the same for the rest. This keeps it short.
+
+And when we move on to the frequentist models with lme4, we should only say something like... the general process is similar, except there are no priors to set.
+
+Again, in the main text, we do not want to drowned people in implementation detail when we have a set of tutorials to do exactly that.
+
+### model building sections ###
+
+rather than output the code, which I've removed for the majority of instances, give some intuition on why we would want to specify formulas and models in this way.
+
+e.g., why have time in a certain form and why build from model 1 through to 4. When you add or change something, say why that might be useful to explore. But don't make it overly complicated!
+
+e.g., if time effects are not linear, which seems likely, then one may want to try different parameterisations of time...
+
+### model comparison ###
+
+you say:
+
+>Clearly, both weighting schemes prefer model M4.
+
+How is this clear? I don't think there are any plots or tables in the main text.
+
+### parameter plots ###
+
+let's have a paragraph on how we might interpret these plots.
+Can we have a table with summary data and quantiles, just to show how they might be interpreted and reported?
+
+### frequentist versus Bayesian plot ###
+
+I like the idea of comparing. That's nice. But let's improve the figure and include intervals as well as points.
+
+## Tutorial 4 ##
+
+>Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+This needs explaining or removing.
+
+## Disucssion ##
+
+[[This para needs to be way shorter and easier to read or we get rid of it]]
+
+>Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if F~A~(t) > F~B~(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
+Townsend (1990) showed that a complete ordering on the hazard functions —$\lambda$~A~(t) > $\lambda$~B~(t) for all t— implies a complete ordering on both the cumulative distribution and survivor functions —F~A~(t) > F~B~(t) and S~A~(t) < S~B~(t)— which in turn implies an ordering on the mean latencies —mean A < mean B.
+In contrast, an ordering on two means does *not* imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does *not* imply a complete ordering on the corresponding hazard functions.
+This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions [see @panisHowCanWe2020].

### 2024-10-01 manuscript/manuscript.Rmd
I updated the manuscript via direct edits

@@
 ---
-title             : "A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data"
-shorttitle        : "Tutorial on hazard analysis"
+title             : "Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows"
+shorttitle        : "A tutorial on hazard analysis"
 author: 
@@
   Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).
+  Social Brain Sciences lab, Department of Humanities, Social and Political Sciences (D-GESS).
   <!-- Enter author note here. -->
 abstract: |
-  Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and how they evolve with increasing waiting time.
-Such information can change key conclusions about psychological processes and can be revealed by using distributional measures that portray the detailed shape of time-to-event distributions.
-  Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
-hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We illustrate how one can calculate the descriptive statistics, and how one can implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We  discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, the advantages of a hazard analysis over other distributional methods available in the literature, limitations, and extensions.
-Our ultimate goal is to convince readers to start using hazard analysis more often when dealing with time-to-event data.
+  Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
+However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+The ability to reveal finer-grained, "temporal states" of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
+Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
+To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis.
+We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferntial statistics via Bayesian and frequentist multi-level regression modelling.
+We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
+Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
 
   <!-- https://tinyurl.com/ybremelq -->
-keywords          : "response times, event history analysis, Bayesian regression models"
+keywords          : "response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology"
 wordcount         : "X"
-bibliography      : ["r-references.bib", "extrareferences.bib"]
+bibliography      : ["extrareferences.bib", "r-references.bib"]
 floatsintext      : yes
@@
 r_refs(file = "r-references.bib")
-my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)
+my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) ## 
 ```
@@
 
-# Introduction
+# 1. Introduction
 
-## Means versus distributional shapes
+## 1.1 Motivation and background context: Comparing means versus distributional shapes
 
-In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events.
+In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (insert REF).
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 
-As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
-In each example, the mean RT is lower in condition 2 compared to condition 1.
-However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
-In the second example, the effect starts around 400 ms and is gone by 800 ms.
-And in the third example, the effect reverses around 550 ms.
+As a simple illustration, Figure 1 shows the results of several simulated RT datasets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distribution differs (Figure 1, top row).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the rt distribution differs in each case (Figure 1, bottom row).
+Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distribution would lead to a different conclusion in every case.
 
-(ref:descr-fig1-caption) Means versus distributional shapes.
+(ref:descr-fig1-caption) Means versus distributional shapes for six different simulated dataset examples.
+For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
+Insets show mean reaction time per condition.
 
-```{r plot1, fig.cap = "(ref:descr-fig1-caption)", out.width="80%"}
-knitr::include_graphics("../Figures/rt_haz_combi_2.jpeg")
+```{r plot1, fig.cap = "(ref:descr-fig1-caption)", out.width="90%"}
+knitr::include_graphics("../sims/figures/haz_inset_facet.jpeg")
 ```
 
-## Outline of the paper
+Why does this matter for research in psychology?
+Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the timecourse of psychological states.
+As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
+For example, the distributions in Example 4 show that the effect starts around 200 ms and is gone by 600 ms.
+In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms.
+And in the Example 6, the effect reverses around between 500 and 600 ms.
+What kind of theory or set of theories could account for such effects?
+Are there new auxiliary assumptions that theories need to adopt?
+And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
+As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to theory building toolkit.
 
-In this paper we focus on a distributional method known as discrete-time event history analysis, a.k.a.
+From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict when events will occur is not a new endeavour.
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
+The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
+Here, in the paper, we hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+
+## 1.2 Aims and structure of the paper
+
+In this paper, we focus on a distributional method known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
 We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, and for detailed treatment, see @singerAppliedLongitudinalData2003, @allisonDiscreteTimeMethodsAnalysis1982, and @allisonSurvivalAnalysisUsing2010.
+However, this will remain relatively short, as this has been covered in detail before @singerAppliedLongitudinalData2003, @allisonDiscreteTimeMethodsAnalysis1982, and @allisonSurvivalAnalysisUsing2010, and our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you should do them.
 
-We then provide four different tutorials, each of which is written in R code and publicly available on Github and the Open Science Framework (OSF).
+We then provide four different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
-In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published RT data set when there is one independent variable.
+Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
+Additionally, each tutorial is made available as .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
+
+In Tutorial 1, we illustrate how to process or "wrangle" a previously published RT dataset to calculate descriptive statistics when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 2 we illustrate how one can fit Bayesian hazard models to the data.
+In Tutorial 2, we illustrate how one can fit Bayesian multi-level regression hazard models to the data using the R package brms.
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 3 we illustrate how to fit hazard models in a frequentist framework.
-Even though both frameworks generate similar parameter estimates, we note that model convergence is often not obtained in the frequentist framework.
-In Tutorial 4 we illustrate how to calculate the descriptive statistics when there are two independent variables.
+In Tutorial 3, we illustrate how to fit the same type of regression hazard models in a frequentist framework using the R package lme4.
+We then briefly compare and contrast these inferential frameworks when applied to EHA.
+In Tutorial 4, we provide a generalisation of the approach to illustrate one might descriptive statistics when using a more complex design, such as when there are two independent variables.
 
-# Overview of hazard analysis
+In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (REFs) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
 
-To apply event history analysis (EHA), one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, a saccade onset, a fixation offset, etc.), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
-Both the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+# 2. A brief introduction to hazard analysis
 
-The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].
-Let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
-Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
-The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs).
+Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
+Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
+Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
+As such, we only supply regression equations in supplementary materials and then refer to them in the text whenever relevant.
 
-Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+## 2.1 Basic features of hazard analysis
+
+To apply event history analysis (EHA), one must be able to:
+
+1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.)
+
+2. define time point zero (e.g., target stimulus onset, fixation onset)
+
+3. measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+
+The definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+Since our focus here is on hazard models that use discrete time units, we describe that approach.
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the first response could occur in 550 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+
+Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X).
+The equations that define both of these functions are reported in supplementary materials (Supp XX).
 The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs after bin t.
+In other words, the survivor function reflects the likelihood that the event occurs in a subsequent timebin.
 
-For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-This latter function is also known as the micro-level speed-accuracy tradeoff function.
+The survivor function can help to qualify or provide context to the interpretation of the hazard function.
+For example, it can give a sense of how many trials may contribute to that part of the distribution.
+If each participant completes 100 trials in an experiment, and the survivor function prob of 0.03, then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
+Therefore, the error bars in this part of the distribution would be wider and less precise compared to other parts.
 
-Here we focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-*N* designs).
+(ref:descr-fig2-1-caption) Hazard and survivor functions
+
+```{r plot2-1, fig.cap = "(ref:descr-fig2-1-caption)", out.width="80%"}
+knitr::include_graphics("../sims/figures/haz_surv.jpeg")
+```
+
+## 2.2 Hazard analysis in the context of experimental psychology
+
+### 2.2.1 A worked example
+
+In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer.
+For example, a task may involve choosing between two response options with only one of them being correct.
+For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
+X in Supps), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+
+Integrating results between hazard and conditional accuracy functions can be informative for understanding psychological processes.
+To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
+In a standard stimulus-response-compatibility paradigm, there is a prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
+The prime can then be congruent or incongruent with the target.
+Taken together, the results show that for early responses (< bin 6), responses always follow the prime (and not the target, as instructed).
+And then for later responses, the prime can be overridden, as both conditions are now always correct.
+This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
+And such a conclusion is not supporting when the effects are explored over a timeline.
+Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
+
+(ref:descr-fig2-2-caption) Hazard and conditional accuracy
+
+```{r plot2-2, fig.cap = "(ref:descr-fig2-2-caption)", out.width="80%"}
+knitr::include_graphics("../sims/figures/haz_acc.jpeg")
+```
+
+Unlocking the temporal states of cognitive processes can be revealing in and of itself for theory development and the understanding of basic psychological processes.
+Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
+Do all participants show similar temporal states or are there individual differences?
+Do such individual differences extend to those individuals that have been diagnosed with some form of psychopathology?
+How do temporal states relate to brain-based mechanisms that might be studied using other methods from cognitive neuroscience?
+And how much of theory in cognitive psychology would be in need of revision if mean-average comparisons were supplemented with a temporal states approach?
+
+### 2.2.2 Implications for designing experiments
+
+Performing hazard analyses in experimental psychology has implications for how experiments are designed.
+Indeed, if trials are categorised as a function of when they occur, then each timebin will only include a subset of the total number of trials.
+For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
+Those 100 trials must be distributed in some manner across the chosen number of bins.
+
+In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+Accordingly, experiment designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-*N* designs).
 This approach  emphasizes  the  precision  and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
-In contrast to the large-*N* design that averages across many participants without being able to scrutinize individual data patterns,  small-*N*  designs  retain  crucial  information  about  the  data  patterns  of  individual  observers.
-This  is  of  great  advantage whenever participants differ systematically in their strategies or in the time-courses of their effects, so that blindly averaging them would lead to misleading data patterns.
-Indeed, @smithSmallBeautifulDefense2018 argue that, “if psychology is to be a mature quantitative science, then its primary theoretical aim should be to investigate systematic functional relationships as they are manifested at the individual participant level” (p.
-2083).
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can have excellent power [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
+
+In contrast to the large-*N* design that typically average across many participants without being able to scrutinize individual data patterns,  small-*N*  designs  retain  crucial  information  about  the  data  patterns  of  individual  observers.
+This can be advantageous whenever participants differ systematically in their strategies or in the time-courses of their effects, so that averaging them would lead to misleading data patterns.
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
 
@@
 
-# Tutorial 1: Calculating descriptive statistics using a life table
+# 3. An overview of the general analytical workflow
 
-To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016, available on [**ResearchGate**](https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect).
+Although the focus is on EHA, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
+
+## 3.1 Data science workflow and descriptive statistics
+
+Descriptive, data science workflow.
+Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
+Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
+[[more here, as necessary]].
+
+## 3.2 Inferential statistical approach
+
+Our lab adopts a estimation approach to multi-level regression (Kruschke & Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?).
+We also use a "keep it maximal" approach to specifying varying (or random) effects (Barr et al., 2013).
+This means that wherever possible we include varying intercepts and slopes per pid.
+To make inferences, we use two main approaches.
+We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy.
+We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
+
+# 4. Tutorials
+
+[[provide a short summary of the aims and scope of each tutorial, as well as the links between them]].
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own user-defined functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+
+Then a list of tutorials:
+
+1a. Wrangle raw data and descriptive stats (T1).
+1b.
+
+2a.
+2b.
+
+3a.
+3b.
+
+Inferential stats (T2 and T3).
+
+Generalisation (T4).
+Should this be online in Supps??
+It would make the main text shorter and simpler, but make it no less available.
+We could just have a sentence at the end of T1, which says that we provide a generalisation and extension in T4, which is in Supps.
+
+Plannng (T5) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
+
+## 4.1 Tutorial 1: Calculating descriptive statistics using a life table
+
+### 4.1.1 Data wrangling aims
+
+Our data wrangling procedures serve two related purposes.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions.
+Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches.
+The two types of data structure we label as 'person-trial' data (Table 1) and 'person-trial-bin' data (Table 2).
+The 'person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
+In contrast, the 'person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
+Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
+Likewise, the event could occur in bin 1 there would only be 1 row of data for that trial.
+
+```{r}
+## make it reproducible
+set.seed(123)
+
+## create the ca data
+ca <- tibble(
+  pid = as.integer(rep(1, times=10)),
+  trial = 1:10,
+  condition = sample(rep(c("congruent", "incongruent"), each = 5)),
+  rt = rnorm(10, 500, 100),
+  accuracy = rbinom(10,1,0.7)
+)
+
+## create the ha data
+ha <- tibble(
+  pid = c(rep(1, times=9)),
+  trial = c(rep(1, times=4), rep(2,times=5)),
+  condition = c(rep("congruent", times=4), rep("incongruent", times=5)),
+  timebin = c(seq(1,4), seq(1,5)),
+  event = c(0, 0, 0, 1, 0, 0, 0, 0, 1)
+) %>% 
+  mutate_if(is.numeric, as.integer)
+```
+
+(ref:ca-data-table-caption) Data structure for 'person-trial' data
+
+(ref:ca-data-note-caption) The first 10 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.
+
+```{r ca-data-table}
+apa_table(
+  ca,
+  caption = "(ref:ca-data-table-caption)",
+  note = "(ref:ca-data-note-caption)",
+  placement = "H"
+)
+```
+
+(ref:ha-data-table-caption) Data structure for 'person-trial-bin' data
+
+(ref:ha-data-note-caption) The first 2 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.
+
+```{r ha-data-table}
+apa_table(
+  ha,
+  caption = "(ref:ha-data-table-caption)",
+  note = "(ref:ha-data-note-caption)",
+  placement = "H"
+)
+```
+
+### 4.1.2 A real data wrangling example
+
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016.
 In their first experiment, @panisWhatShapingRT2016 presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 
-After loading in the data file, one has to (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
+There are several data wrangling steps to be taken. First, we need to load the data before (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
+
 The required column names are as follows:
@@
 * "rt", indicating the response times in ms;
-* "acc", inicating the accuracies (1/0).
+* "acc", indicating the accuracies (1/0).
 
@@
 
-To set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t) using functional programming, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our function "censor()", as follows.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t).
+To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our function "censor()", as follows.
 
@@
 In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
-Removing such trials before calculating the mean event time can introduce a sampling bias.
+Removing such trials before calculating the mean event time can introduce a sampling bias (REFs).
 The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
 The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-When creating the plots using our function plot_eha(), some warning messages will likely be generated, like these:
+
+The next step is to plot the data using our custom made plotting tool plot_eha(). When creating the plots, some warning messages will likely be generated, like these:
 
@@
 
-Table 1 shows the life table for condition "blank" (no prime stimulus presented) - compare to Figure 1.
+Table 3 shows the life table for condition "blank" (no prime stimulus presented) - compare to Figure 1.
 A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
@@
 
-Figure 2 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
+Figure 4 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
 By using discrete-time h(t) functions of event occurrence - in combination with ca(t) functions for two-choice tasks - one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
@@
 
-  These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+  These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
@@
 
-  Also, in their second Experiment, @panisWhatShapingRT2016 showed that the negative compatibility effect in the mask-present conditions (see Tutorial 4) is time-locked to mask onset.
-This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022; @schmidtResponseInhibitionNegative2022].
-In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
 As we will show in Tutorials 2 and 3, statistical models for h(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
+As suchm multi-level regression is what we tuen to in the next tutorials.
 
-# Tutorial 2: Fitting Bayesian hazard models
+## 4.2 Tutorial 2: Fitting Bayesian hazard models
 
-In this second tutorial we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+In this second tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
+
+### 4.2.1 Hazard model considerations
 
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see Figure 3).
-
-(ref:descr-fig3-caption) The logit and cloglog link functions.
-
-```{r plot-link-functions, fig.cap = "(ref:descr-fig3-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
-```
-
-The symmetric logit link function transforms a (hazard) probability into the log of the odds ratio.
-The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event *non*occurrence.
-An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption.
+Second, given that the dependent variable is binary, one has to select a link function (see Supps).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-
-An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
-
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)].
-
-The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
-Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
-To interpret the effects of a predictor,its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
-When using the logit link, exponentiating a $\beta$ parameter results in an odds ratio.
-
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X~1~) can be written as follows:
-
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~].
+We provide relevant example regression formulas in supplementary materials.
 
@@
 
-## Prior distributions
+### 4.2.2 Prior distributions
 
 To get the posterior distribution of each parameters given the data, we need to specify a prior distribution for each parameter.
-The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+The middle column of Supplementary Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
 
@@
 
-(ref:descr-fig4-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+### 4.2.3 Model 1: A general specification of TIME, and main effects of congruency and trial number
 
-```{r plot-priors, fig.cap = "(ref:descr-fig4-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")
-```
-
-To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
-
-To gain a sense of what prior *cloglog* values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
-Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
-However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
-A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
-
-## Model 1: A general specification of TIME, and main effects of congruency and trial number
+[[Here let's give some intuition on why we would want to setup the model like this]]
 
@@
 
-## Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
+### 4.2.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
+
+[[Here let's give some intuition on why we would want to modify the formula and model features]]
 
@@
 
-\scriptsize
-```{r data-priors-M2, echo=T}
-# remove unnecessary columns
-M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15))
+<!-- \scriptsize -->
+<!-- ```{r data-priors-M2, echo=T} -->
+<!-- # remove unnecessary columns -->
+<!-- M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15))  -->
 
-# Specify priors
-priors_M2 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),    
-  set_prior("normal(0, 1)", class = "sd"),            
-  set_prior("lkj(2)", class = "cor")                 
-)
-```
-\normalsize
+<!-- # Specify priors -->
+<!-- priors_M2 <- c( -->
+<!--   set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       -->
+<!--   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),     -->
+<!--   set_prior("normal(0, 1)", class = "sd"),             -->
+<!--   set_prior("lkj(2)", class = "cor")                  -->
+<!-- ) -->
+<!-- ``` -->
+<!-- \normalsize -->
 
-Now we can fit model 2.
+<!-- Now we can fit model 2. -->
 
-\scriptsize
-```{r fit-model-M2, eval=F, echo=T}
-plan(multicore)
+<!-- \scriptsize -->
+<!-- ```{r fit-model-M2, eval=F, echo=T} -->
+<!-- plan(multicore) -->
 
-model_M2 <-
-   brm(data = M2_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c +
-                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c | pid),
-       prior = priors_M2,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M2")
-```
-\normalsize
+<!-- model_M2 <- -->
+<!--    brm(data = M2_data, -->
+<!--        family = binomial(link="cloglog"), -->
+<!--        event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) +  -->
+<!--                           condition + trial_c + -->
+<!--                           (1 + period_9 + I(period_9^2) + I(period_9^3) +  -->
+<!--                           condition + trial_c | pid), -->
+<!--        prior = priors_M2, -->
+<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
+<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
+<!--        seed = 12, init = "0", -->
+<!--        file = "Tutorial_2_Bayesian/models/model_M2") -->
+<!-- ``` -->
+<!-- \normalsize -->
 
@@
 
-## Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
+### 4.2.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
+
+[[Here let's give some intuition on why we would want to modify the formula and model features]]
 
@@
 
-\scriptsize
-```{r fit-model-M3, eval=F, echo=T}
-M3_data <- M2_data
-priors_M3 <- priors_M2
-plan(multicore)
+<!-- \scriptsize -->
+<!-- ```{r fit-model-M3, eval=F, echo=T} -->
+<!-- M3_data <- M2_data -->
+<!-- priors_M3 <- priors_M2 -->
+<!-- plan(multicore) -->
 
-model_M3 <- 
-   brm(data = M3_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
-                           condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
-       prior = priors_M3,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M3")
-```
-\normalsize
+<!-- model_M3 <-  -->
+<!--    brm(data = M3_data, -->
+<!--        family = binomial(link="cloglog"), -->
+<!--        event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored -->
+<!--                            condition*period_9 +  -->
+<!--                            condition*I(period_9^2) +  -->
+<!--                            condition*I(period_9^3) + -->
+<!--                            trial_c*period_9 +  -->
+<!--                            trial_c*I(period_9^2) +  -->
+<!--                            trial_c*I(period_9^3) + -->
+<!--                            (1 + condition*period_9 + -->
+<!--                            condition*I(period_9^2) + -->
+<!--                            condition*I(period_9^3) + -->
+<!--                            trial_c*period_9 +  -->
+<!--                            trial_c*I(period_9^2) +  -->
+<!--                            trial_c*I(period_9^3) | pid), -->
+<!--        prior = priors_M3, -->
+<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
+<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
+<!--        seed = 12, init = "0", -->
+<!--        file = "Tutorial_2_Bayesian/models/model_M3") -->
+<!-- ``` -->
+<!-- \normalsize -->
 
@@
 
-## Model 4: A polynomial specification of TIME, and relaxing all three assumptions
+### 4.2.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions
 
@@
 
-\scriptsize 
-```{r fit-model-M4, eval=F, echo=T}
-M4_data <- M2_data
-priors_M4 <- priors_M2
-plan(multicore)
+<!-- \scriptsize  -->
+<!-- ```{r fit-model-M4, eval=F, echo=T} -->
+<!-- M4_data <- M2_data -->
+<!-- priors_M4 <- priors_M2 -->
+<!-- plan(multicore) -->
 
-model_M4 <- 
-   brm(data = M4_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
-                          condition*period_9*trial_c + 
-                          condition*period_9*I(trial_c^2) + 
-                          condition*I(period_9^2)*trial_c +
-                          condition*I(period_9^2)*I(trial_c^2) +
-                          condition*I(period_9^3) +
-                          trial_c*I(period_9^3) +
-                          (1 +  condition*period_9*trial_c +
-                                condition*period_9*I(trial_c^2) + 
-                                condition*I(period_9^2)*trial_c +
-                                condition*I(period_9^2)*I(trial_c^2)  
-                                condition*I(period_9^3) +
-                                trial_c*I(period_9^3) | pid),
-       prior = priors_M4,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M4")
-```
-\normalsize
+<!-- model_M4 <-  -->
+<!--    brm(data = M4_data, -->
+<!--        family = binomial(link="cloglog"), -->
+<!--        event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored -->
+<!--                           condition*period_9*trial_c +  -->
+<!--                           condition*period_9*I(trial_c^2) +  -->
+<!--                           condition*I(period_9^2)*trial_c + -->
+<!--                           condition*I(period_9^2)*I(trial_c^2) + -->
+<!--                           condition*I(period_9^3) + -->
+<!--                           trial_c*I(period_9^3) + -->
+<!--                           (1 +  condition*period_9*trial_c + -->
+<!--                                 condition*period_9*I(trial_c^2) +  -->
+<!--                                 condition*I(period_9^2)*trial_c + -->
+<!--                                 condition*I(period_9^2)*I(trial_c^2)   -->
+<!--                                 condition*I(period_9^3) + -->
+<!--                                 trial_c*I(period_9^3) | pid), -->
+<!--        prior = priors_M4, -->
+<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
+<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
+<!--        seed = 12, init = "0", -->
+<!--        file = "Tutorial_2_Bayesian/models/model_M4") -->
+<!-- ``` -->
+<!-- \normalsize -->
 
@@
 
-## Compare the models.
+### 4.2.7 Compare the models.
 
@@
 
-\scriptsize 
-```{r compare-models, echo=T}
-model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds")
-model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds")
-model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds")
-model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds")
+<!-- \scriptsize  -->
+<!-- ```{r compare-models, echo=T, eval=FALSE} -->
+<!-- model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds") -->
+<!-- model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds") -->
+<!-- model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds") -->
+<!-- model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds") -->
 
-#model_M1  <- add_criterion(model_M1, c("loo", "waic"))
-#model_M2  <- add_criterion(model_M2, c("loo", "waic"))
-#model_M3  <- add_criterion(model_M3, c("loo", "waic"))
-#model_M4  <- add_criterion(model_M4, c("loo", "waic"))
+<!-- #model_M1  <- add_criterion(model_M1, c("loo", "waic")) -->
+<!-- #model_M2  <- add_criterion(model_M2, c("loo", "waic")) -->
+<!-- #model_M3  <- add_criterion(model_M3, c("loo", "waic")) -->
+<!-- #model_M4  <- add_criterion(model_M4, c("loo", "waic")) -->
 
-loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F)
-loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F)
+<!-- loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F) -->
+<!-- loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F) -->
 
-# model weights
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
+<!-- # model weights -->
+<!-- model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3) -->
 
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
-```
-\normalsize
+<!-- model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3) -->
+<!-- ``` -->
+<!-- \normalsize -->
 
@@
 
-## Plot congruency effects and subject-specific fits for model M4.
+### 4.2.8 Evaluate parameter estimates
 
@@
 
-# Tutorial 3: Fitting Frequentist hazard models
+[[let's have a paragraph on how we might interpret these plots.]]
+
+## 4.3 Tutorial 3: Fitting Frequentist hazard models
 
@@
 
-\scriptsize 
-```{r fit-model, eval=F, echo=T}
-model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
+<!-- \scriptsize  -->
+<!-- ```{r fit-model, eval=F, echo=T} -->
+<!-- model_M3_f <- glmer(event ~ 1 + condition*period_9 +  -->
+<!--                            condition*I(period_9^2) +  -->
+<!--                            condition*I(period_9^3) + -->
+<!--                            trial_c*period_9 +  -->
+<!--                            trial_c*I(period_9^2) +  -->
+<!--                            trial_c*I(period_9^3) + -->
+<!--                            (1 + condition*period_9 + -->
+<!--                            condition*I(period_9^2) + -->
+<!--                            condition*I(period_9^3) + -->
+<!--                            trial_c*period_9 +  -->
+<!--                            trial_c*I(period_9^2) +  -->
+<!--                            trial_c*I(period_9^3) | pid), -->
 
-              # control parameters, data set, and complementary log-log link function
-              control = glmerControl(optimizer = c("nlminbwrap"),
-                                     optCtrl = list(maxfun=10000000)), 
-              data=M3_data, 
-              family=binomial(link="cloglog"))
-```
-\normalsize
+<!--               # control parameters, data set, and complementary log-log link function -->
+<!--               control = glmerControl(optimizer = c("nlminbwrap"), -->
+<!--                                      optCtrl = list(maxfun=10000000)),  -->
+<!--               data=M3_data,  -->
+<!--               family=binomial(link="cloglog")) -->
+<!-- ``` -->
+<!-- \normalsize -->
 
-In Figure 7 we compere the parameter estimates of model M3 from brm() with those of glmer().
+In Figure 7 we compare the parameter estimates of model M3 from brm() with those of glmer().
 
@@
 However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
-This is of course one of the reasons why Bayesian modeling has become so popular in the last years.
+This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
+But the price you pay for being able to fit more complex models in a Bayesian framework is computation time.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorial 2 took several hours to build.
 
-# Tutorial 4: Calculating descriptive statistics when there are two independent variables
+## 4.4 Tutorial 4: Generalising to a more complex design
 
-In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
+So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
+But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
+The purpose of Tutorial 4, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
+We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
+
+To this end, Tutorial 4 illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in Tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
 Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
@@
 
-# Discussion
+# 5. Discussion
 
-We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
-By providing this tutorial, we hope that researchers will start using descriptive and inferential hazard analysis more often, due to the many advantages described below.
-While we focused here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
-We also discuss the issue of individual differences, limitations, and extensions.
+This main motivation for writing this paper is the observation that event history analysis remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA provides compared to more conventional analyses.
+By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA in the future.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA, which is the increase in analytical complexity compared to mean-average comparisons.
+While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
+In the following, we discuss issues relating to individual differences, limitations of the approach, and future extensions.
 
-## Advantages of hazard analysis
+## 5.1 Advantages of hazard analysis
 
@@
 331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
+Therefore, [[why should people care?
+What is the functional relevance for exp psych and researchers?]]
 
+[[This para needs to be way shorter and easier to read or we get rid of it]]
 Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
@@
 
-## Individual differences
+## 5.2 Individual differences
 
@@
 Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
+
 But even when using a response deadline, participants can differ qualitatively in the effects they display [see @panisHowCanWe2020].
@@
 
-## Limitation(s)
+## 5.3 Limitation(s)
 
-Compared to the orthodox method -- comparing means with ANOVA --, the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
-Another issue that might initially look as a limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
-However, as nature does not reveal itself easily, obtaining insight into behavioral dynamics simply requires more data per condition than is usually collected under the orthodox method.
-In general, there is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function; We therefore recommend to design as many trials as possible per experimental condition given the available resources.
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
+Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
+For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect?
+After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
+Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
 
-## Extensions
+## 5.4 Extensions
 
 The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this event consitutes a single transition between an "idle" state and a "responded" state.
+For RT data, this event constitutes a single transition between an "idle" state and a "responded" state.
 However, in certain situations, more than one event of interest might exist.
-For example, an individual might transition back and forth between a "healthy" state and a "depression" state, before being absorbed in a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states [@SteeleF2004].
+For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state.
+When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states [@SteeleF2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (REF).
+[[give a concrete example for this latter point]]
 
-# Conclusions
+# 6. Conclusions
 
-RT and accuracy distributions are a rich source of information on the time course of cognitive processing.
-The changing effects of our experimental manipulations with increases in waiting time become strikingly clear when looking at response hazards and microlevel speed-accuracy trade-off functions.
-We hope that experimental psychologists and cognitive neuroscientists are less reluctant to embrace event history analysis after reading this tutorial.
+RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
+We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.
+On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.
 

### 2024-10-01 manuscript/manuscript.tex
I updated the manuscript via direct edits

--- 2024-09-22 manuscript/manuscript.tex
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 % definitions for citeproc citations
@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian regression models\newline\indent Word count: X}
+\keywords{response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology\newline\indent Word count: X}
 \usepackage{lineno}
@@
 \hypersetup{
-  pdftitle={A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data},
+  pdftitle={Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows},
   pdfauthor={Sven Panis1 \& Richard Ramsey1},
   pdflang={en-EN},
-  pdfkeywords={response times, event history analysis, Bayesian regression models},
+  pdfkeywords={response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology},
   hidelinks,
@@
 
-\title{A tutorial on Bayesian and Frequentist Event History Analyses for psychological time-to-event data}
+\title{Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows}
 \author{Sven Panis\textsuperscript{1} \& Richard Ramsey\textsuperscript{1}}
@@
 
-\shorttitle{Tutorial on hazard analysis}
+\shorttitle{A tutorial on hazard analysis}
 
@@
 Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).
+Social Brain Sciences lab, Department of Humanities, Social and Political Sciences (D-GESS).
 
@@
 \abstract{%
-Time-to-event data such as response times, saccade latencies, and fixation durations are ubiquitous in experimental psychology.
-The orthodox method for analysing such data -- comparing means with analysis-of-variance -- is actually hiding a lot of information about psychological effects, such as their onset time and duration, and how they evolve with increasing waiting time.
-Such information can change key conclusions about psychological processes and can be revealed by using distributional measures that portray the detailed shape of time-to-event distributions.\\
-Here we provide a set of tutorials on how to implement one particular distributional method known as discrete-time event history analysis, a.k.a.
-hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We illustrate how one can calculate the descriptive statistics, and how one can implement Bayesian and frequentist regression models, using the R packages tidyverse, brms, and lme4.
-The R code is publicly available on Github and OSF, and can easily be adapted for other data sets.
-We discuss possible link functions and prior distributions, how to manage inter-individual differences, implications for experimental design, the advantages of a hazard analysis over other distributional methods available in the literature, limitations, and extensions.
-Our ultimate goal is to convince readers to start using hazard analysis more often when dealing with time-to-event data.
+Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
+However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
+Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
+To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis.
+We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferntial statistics via Bayesian and frequentist multi-level regression modelling.
+We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
+Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
 }
@@
 
-\section{Introduction}\label{introduction}
-
-\subsection{Means versus distributional shapes}\label{means-versus-distributional-shapes}
-
-In experimental psychology, it is still standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events.
+\section{1. Introduction}\label{introduction}
+
+\subsection{1.1 Motivation and background context: Comparing means versus distributional shapes}\label{motivation-and-background-context-comparing-means-versus-distributional-shapes}
+
+In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (insert REF).
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 
-As a simple illustration, Figure 1 shows three examples of how an observed difference in mean response times (RTs) between two experimental conditions conceals differences in the shapes of the underlying RT distributions.
-In each example, the mean RT is lower in condition 2 compared to condition 1.
-However, the distributions in the first example show that the effect starts around 200 ms and is gone by 600 ms.
-In the second example, the effect starts around 400 ms and is gone by 800 ms.
-And in the third example, the effect reverses around 550 ms.
+As a simple illustration, Figure 1 shows the results of several simulated RT datasets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distribution differs (Figure 1, top row).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the rt distribution differs in each case (Figure 1, bottom row).
+Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distribution would lead to a different conclusion in every case.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Figures/rt_haz_combi_2}
+{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../sims/figures/haz_inset_facet}
 
@@
 
-\caption{Means versus distributional shapes.}\label{fig:plot1}
+\caption{Means versus distributional shapes for six different simulated dataset examples.
+For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
+Insets show mean reaction time per condition.}\label{fig:plot1}
 \end{figure}
 
-\subsection{Outline of the paper}\label{outline-of-the-paper}
-
-In this paper we focus on a distributional method known as discrete-time event history analysis, a.k.a.
+Why does this matter for research in psychology?
+Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the timecourse of psychological states.
+As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
+For example, the distributions in Example 4 show that the effect starts around 200 ms and is gone by 600 ms.
+In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms.
+And in the Example 6, the effect reverses around between 500 and 600 ms.
+What kind of theory or set of theories could account for such effects?
+Are there new auxiliary assumptions that theories need to adopt?
+And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
+As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to theory building toolkit.
+
+From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict when events will occur is not a new endeavour.
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
+The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
+Here, in the paper, we hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+
+\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
+
+In this paper, we focus on a distributional method known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
 We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, and for detailed treatment, see Singer and Willett (2003), Allison (1982), and Allison (2010).
-
-We then provide four different tutorials, each of which is written in R code and publicly available on Github and the Open Science Framework (OSF).
+However, this will remain relatively short, as this has been covered in detail before Singer and Willett (2003), Allison (1982), and Allison (2010), and our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you should do them.
+
+We then provide four different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
-In Tutorial 1 we illustrate how to calculate the descriptive statistics for a published RT data set when there is one independent variable.
+Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
+Additionally, each tutorial is made available as .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
+
+In Tutorial 1, we illustrate how to process or ``wrangle'' a previously published RT dataset to calculate descriptive statistics when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 2 we illustrate how one can fit Bayesian hazard models to the data.
+In Tutorial 2, we illustrate how one can fit Bayesian multi-level regression hazard models to the data using the R package brms.
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 3 we illustrate how to fit hazard models in a frequentist framework.
-Even though both frameworks generate similar parameter estimates, we note that model convergence is often not obtained in the frequentist framework.
-In Tutorial 4 we illustrate how to calculate the descriptive statistics when there are two independent variables.
-
-\section{Overview of hazard analysis}\label{overview-of-hazard-analysis}
-
-To apply event history analysis (EHA), one must be able to define the event of interest (any qualitative change that can be situated in time, e.g., a button press, a saccade onset, a fixation offset, etc.), time point zero (e.g., target stimulus onset, fixation onset), and measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
-Both the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
-
-The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991).
-Let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
-Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
-The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
-
-Similarly, after dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t\textbar{} RT \(\geq\) t) and the discrete-time survivor function S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]}, and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+In Tutorial 3, we illustrate how to fit the same type of regression hazard models in a frequentist framework using the R package lme4.
+We then briefly compare and contrast these inferential frameworks when applied to EHA.
+In Tutorial 4, we provide a generalisation of the approach to illustrate one might descriptive statistics when using a more complex design, such as when there are two independent variables.
+
+In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (REFs) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
+
+\section{2. A brief introduction to hazard analysis}\label{a-brief-introduction-to-hazard-analysis}
+
+For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs).
+Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
+Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
+Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
+As such, we only supply regression equations in supplementary materials and then refer to them in the text whenever relevant.
+
+\subsection{2.1 Basic features of hazard analysis}\label{basic-features-of-hazard-analysis}
+
+To apply event history analysis (EHA), one must be able to:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.)
+\item
+  define time point zero (e.g., target stimulus onset, fixation onset)
+\item
+  measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+\end{enumerate}
+
+The definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+Since our focus here is on hazard models that use discrete time units, we describe that approach.
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the first response could occur in 550 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+
+Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X).
+The equations that define both of these functions are reported in supplementary materials (Supp XX).
 The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
-
-For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct \textbar{} RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-This latter function is also known as the micro-level speed-accuracy tradeoff function.
-
-Here we focus on factorial within-subject designs in which a large number of observations are made on a relatively small number of participants (small-\emph{N} designs).
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t.
+In other words, the survivor function reflects the likelihood that the event occurs in a subsequent timebin.
+
+The survivor function can help to qualify or provide context to the interpretation of the hazard function.
+For example, it can give a sense of how many trials may contribute to that part of the distribution.
+If each participant completes 100 trials in an experiment, and the survivor function prob of 0.03, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
+Therefore, the error bars in this part of the distribution would be wider and less precise compared to other parts.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_surv}
+
+}
+
+\caption{Hazard and survivor functions}\label{fig:plot2-1}
+\end{figure}
+
+\subsection{2.2 Hazard analysis in the context of experimental psychology}\label{hazard-analysis-in-the-context-of-experimental-psychology}
+
+\subsubsection{2.2.1 A worked example}\label{a-worked-example}
+
+In the context of experimental psychology, it is common for participants to be presented with a task that has a right and wrong answer.
+For example, a task may involve choosing between two response options with only one of them being correct.
+For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
+X in Supps), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+
+Integrating results between hazard and conditional accuracy functions can be informative for understanding psychological processes.
+To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer.
+In a standard stimulus-response-compatibility paradigm, there is prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
+The prime can then be congruent or incongruent with the target.
+Taken together, the results show that for early responses (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
+And then for later responses, the prime can be overridden, as both conditions are now always correct.
+This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
+And such a conclusion is not supporting when the effects are explored over a timeline.
+Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_acc}
+
+}
+
+\caption{Hazard and conditional accuracy}\label{fig:plot2-2}
+\end{figure}
+
+Unlocking the temporal states of cognitive processes can be revealing in and of itself for theory development and the understanding of basic psychological processes.
+Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
+Do all participants show similar temporal states or are there individual differences?
+Do such individual differences extend to those individuals that have been diagnosed with some form of psychopathology?
+How do temporal states relate to brain-based mechanisms that might be studied using other methods from cognitive neuroscience?
+And how much of theory in cognitive psychology would be in need of revision if mean-average comparisons were supplemented with a temporal states approach?
+
+\subsubsection{2.2.2 Implications for designing experiments}\label{implications-for-designing-experiments}
+
+Performing hazard analyses in experimental psychology has implications for how experiments are designed.
+Indeed, if trials are categorised as a function of when they occur, then each timebin will only include a subset of the total number of trials.
+For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
+Those 100 trials must be distributed in some manner across the chosen number of bins.
+
+In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+Accordingly, experiment designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
 This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
-In contrast to the large-\emph{N} design that averages across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
-This is of great advantage whenever participants differ systematically in their strategies or in the time-courses of their effects, so that blindly averaging them would lead to misleading data patterns.
-Indeed, Smith and Little (2018) argue that, ``if psychology is to be a mature quantitative science, then its primary theoretical aim should be to investigate systematic functional relationships as they are manifested at the individual participant level'' (p.~2083).
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can have excellent power (Baker et al., 2021; Smith \& Little, 2018).
-
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+
+In contrast to the large-\emph{N} design that typically average across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
+This can be advantageous whenever participants differ systematically in their strategies or in the time-courses of their effects, so that averaging them would lead to misleading data patterns.
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
+
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.164; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 
-\section{Tutorial 1: Calculating descriptive statistics using a life table}\label{tutorial-1-calculating-descriptive-statistics-using-a-life-table}
-
-To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016), available on \href{https://www.researchgate.net/publication/304069212_What_Is_Shaping_RT_and_Accuracy_Distributions_Active_and_Selective_Response_Inhibition_Causes_the_Negative_Compatibility_Effect}{\textbf{ResearchGate}}.
+\section{3. An overview of the general analytical workflow}\label{an-overview-of-the-general-analytical-workflow}
+
+Although the focus is on EHA, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
+
+\subsection{3.1 Data science workflow and descriptive statistics}\label{data-science-workflow-and-descriptive-statistics}
+
+Descriptive, data science workflow.
+Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
+Functional programming basically means you don't write your own loops but instead use functions that
+
+\subsection{3.2 Inferential statistical approach}\label{inferential-statistical-approach}
+
+Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?).
+We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013).
+This means that wherever possible we include varying intercepts and slopes per pid.
+To make inferences, we use two main approaches.
+We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy.
+We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
+
+\section{4. Tutorials}\label{tutorials}
+
+{[}{[}provide a short summary of the aims and scope of each tutorial, as well as the links between them{]}{]}.
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own user-defined functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+
+Then a list of tutorials:
+
+1a. Wrangle raw data and descriptive stats (T1).
+1b.
+
+2a.
+2b.
+
+3a.
+3b.
+
+Inferential stats (T2 and T3).
+
+Generalisation (T4).
+Should this be online in Supps??
+It would make the main text shorter and simpler, but make it no less available.
+We could just have a sentence at the end of T1, which says that we provide a generalisation and extension in T4, which is in Supps.
+
+Plannng (T5) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
+
+\subsection{4.1 Tutorial 1: Calculating descriptive statistics using a life table}\label{tutorial-1-calculating-descriptive-statistics-using-a-life-table}
+
+\subsubsection{4.1.1 Data wrangling aims}\label{data-wrangling-aims}
+
+{[}{[}here we might have a figure that visulises the different strucutre of data for hazard analysis and conditional accuracy analysis.
+it is just easier to see visually.
+e.g., conditional acc is how it comes e.g., an RT per trial, whereas hazard needs to be wrangled so that the structure changes.
+I think that is worth visualising.
+e.g., two columns side-by-side.
+top XX rows visualised.
+And if we give them labels, such as `person-trial' data and `person-trial-bin' data, it will helps folsk to follow along.
+I include some text below as an example.{]}{]}
+
+Our data wrangling procedures serve two related purposes.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions.
+Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches.
+The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2).
+The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
+In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
+Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs when it generates a 1.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
+Likewise, the event could occur in bin 1 there would only be 1 row of data for that trial.
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:ca-data-table}Data structure for `person-trial' data}
+
+\begin{tabular}{lllll}
+\toprule
+pid & \multicolumn{1}{c}{trial} & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{rt} & \multicolumn{1}{c}{accuracy}\\
+\midrule
+1 & 1 & congruent & 373.49 & 1\\
+1 & 2 & incongruent & 431.31 & 1\\
+1 & 3 & congruent & 455.43 & 0\\
+1 & 4 & incongruent & 622.41 & 1\\
+1 & 5 & incongruent & 535.98 & 1\\
+1 & 6 & incongruent & 540.08 & 1\\
+1 & 7 & congruent & 511.07 & 1\\
+1 & 8 & incongruent & 444.42 & 1\\
+1 & 9 & congruent & 678.69 & 1\\
+1 & 10 & congruent & 549.79 & 1\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} The first 10 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:ha-data-table}Data structure for `person-trial-bin' data}
+
+\begin{tabular}{lllll}
+\toprule
+pid & \multicolumn{1}{c}{trial} & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{event}\\
+\midrule
+1 & 1 & congruent & 1 & 0\\
+1 & 1 & congruent & 2 & 0\\
+1 & 1 & congruent & 3 & 0\\
+1 & 1 & congruent & 4 & 1\\
+1 & 2 & incongruent & 1 & 0\\
+1 & 2 & incongruent & 2 & 0\\
+1 & 2 & incongruent & 3 & 0\\
+1 & 2 & incongruent & 4 & 0\\
+1 & 2 & incongruent & 5 & 1\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} The first 2 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{4.1.2 A real data wrangling example}\label{a-real-data-wrangling-example}
+
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016).
 In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 \item
-  ``acc'', inicating the accuracies (1/0).
+  ``acc'', indicating the accuracies (1/0).
 \end{itemize}
@@
 
-To set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t) using functional programming, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our function ``censor()'', as follows.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t).
+To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our function ``censor()'', as follows.
 
@@
 In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
-Removing such trials before calculating the mean event time can introduce a sampling bias.
+Removing such trials before calculating the mean event time can introduce a sampling bias (REFs).
 The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-When creating the plots using our function plot\_eha(), some warning messages will likely be generated, like these:
+
+The next step is to plot the data using our custom made plotting tool plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:
 
@@
 
-Table 1 shows the life table for condition ``blank'' (no prime stimulus presented) - compare to Figure 1.
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) - compare to Figure 1.
 A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
@@
 
-{\centering \includegraphics[width=12in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 
-These results show that this participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
@@
 
-Also, in their second Experiment, Panis and Schmidt (2016) showed that the negative compatibility effect in the mask-present conditions (see Tutorial 4) is time-locked to mask onset.
-This example shows that a simple difference between two means fails to reveal the dynamic behavior people display in many experimental paradigms (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Schmidt, Panis, Wolkersdorfer, \& Vorberg, 2022).
-In other words, statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
 As we will show in Tutorials 2 and 3, statistical models for h(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
-
-\section{Tutorial 2: Fitting Bayesian hazard models}\label{tutorial-2-fitting-bayesian-hazard-models}
-
-In this second tutorial we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+As suchm multi-level regression is what we tuen to in the next tutorials.
+
+\subsection{4.2 Tutorial 2: Fitting Bayesian hazard models}\label{tutorial-2-fitting-bayesian-hazard-models}
+
+In this second tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
+
+\subsubsection{4.2.1 Hazard model considerations}\label{hazard-model-considerations}
 
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see Figure 3).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}
-
-}
-
-\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
-\end{figure}
-
-The symmetric logit link function transforms a (hazard) probability into the log of the odds ratio.
-The asymmetric complementary log-log (cloglog) link function transforms hazard into the logarithm of the negated logarithm of the probability of event \emph{non}occurrence.
-An important difference between these two link functions is that cloglog provides a discrete-time hazard model that has a built-in proportional hazards assumption, while logit provides a proportional odds assumption.
+Second, given that the dependent variable is binary, one has to select a link function (see Supps).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-
-An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
-
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.
-
-The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
-Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth.
-To interpret the effects of a predictor,its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
-When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.
-
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
-
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.
+We provide relevant example regression formulas in supplementary materials.
 
@@
 
-\subsection{Prior distributions}\label{prior-distributions}
+\subsubsection{4.2.2 Prior distributions}\label{prior-distributions}
 
 To get the posterior distribution of each parameters given the data, we need to specify a prior distribution for each parameter.
-The middle column of Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+The middle column of Supplementary Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}
-
-}
-
-\caption{Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
-\end{figure}
-
-To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
-
-To gain a sense of what prior \emph{cloglog} values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
-Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
-However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
-A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
-
-\subsection{Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+\subsubsection{4.2.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+
+{[}{[}Here let's give some intuition on why we would want to setup the model like this{]}{]}
 
@@
 
-\subsection{Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+\subsubsection{4.2.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+
+{[}{[}Here let's give some intuition on why we would want to modify the formula and model features{]}{]}
 
@@
 
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# remove unnecessary columns}
-\NormalTok{M2\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) }
-
-\CommentTok{\# Specify priors}
-\NormalTok{priors\_M2 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}0.2,0.71,{-}2.2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),      }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{),    }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{),            }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{)                 }
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-Now we can fit model 2.
-
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{plan}\NormalTok{(multicore)}
-
-\NormalTok{model\_M2 }\OtherTok{\textless{}{-}}
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M2\_data,}
-       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
-\NormalTok{                          (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors\_M2,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
-       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M2"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
 Estimating model M2 took about 144 minutes.
 
-\subsection{Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
+\subsubsection{4.2.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
+
+{[}{[}Here let's give some intuition on why we would want to modify the formula and model features{]}{]}
 
@@
 
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{M3\_data }\OtherTok{\textless{}{-}}\NormalTok{ M2\_data}
-\NormalTok{priors\_M3 }\OtherTok{\textless{}{-}}\NormalTok{ priors\_M2}
-\FunctionTok{plan}\NormalTok{(multicore)}
-
-\NormalTok{model\_M3 }\OtherTok{\textless{}{-}} 
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M3\_data,}
-       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+} \CommentTok{\# Note that duplicate terms in the model formula are ignored}
-\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors\_M3,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
-       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M3"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
 Estimating model M3 took about 268 minutes.
 
-\subsection{Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
-
-Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
+\subsubsection{4.2.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
+
+Based on previous work (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
 We use the same data set and priors as for model 2.
 
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{M4\_data }\OtherTok{\textless{}{-}}\NormalTok{ M2\_data}
-\NormalTok{priors\_M4 }\OtherTok{\textless{}{-}}\NormalTok{ priors\_M2}
-\FunctionTok{plan}\NormalTok{(multicore)}
-
-\NormalTok{model\_M4 }\OtherTok{\textless{}{-}} 
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M4\_data,}
-       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+} \CommentTok{\# Note that duplicate terms in the model formula are ignored}
-\NormalTok{                          condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
-\NormalTok{                          condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                          (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{  condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                                condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)  }
-\NormalTok{                                condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                                trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors\_M4,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
-       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M4"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
 Estimating model M4 took about 8 hours.
 
-\subsection{Compare the models.}\label{compare-the-models.}
+\subsubsection{4.2.7 Compare the models.}\label{compare-the-models.}
 
@@
 
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_M1 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M1.rds"}\NormalTok{)}
-\NormalTok{model\_M2 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M2.rds"}\NormalTok{)}
-\NormalTok{model\_M3 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M3.rds"}\NormalTok{)}
-\NormalTok{model\_M4 }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\StringTok{"../Tutorial\_2\_Bayesian/models/model\_M4.rds"}\NormalTok{)}
-
-\CommentTok{\#model\_M1  \textless{}{-} add\_criterion(model\_M1, c("loo", "waic"))}
-\CommentTok{\#model\_M2  \textless{}{-} add\_criterion(model\_M2, c("loo", "waic"))}
-\CommentTok{\#model\_M3  \textless{}{-} add\_criterion(model\_M3, c("loo", "waic"))}
-\CommentTok{\#model\_M4  \textless{}{-} add\_criterion(model\_M4, c("loo", "waic"))}
-
-\FunctionTok{loo\_compare}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{criterion =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{print}\NormalTok{(}\AttributeTok{simplify =}\NormalTok{ F)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##          elpd_diff se_diff elpd_loo se_elpd_loo p_loo   se_p_loo looic  
-## model_M4     0.0       0.0 -5094.4     62.9       127.6     4.2  10188.9
-## model_M3   -21.0      10.1 -5115.4     62.5        77.3     2.5  10230.8
-## model_M1  -254.9      24.4 -5349.3     64.8        72.7     1.7  10698.6
-## model_M2  -256.8      23.6 -5351.2     64.8        39.7     1.0  10702.5
-##          se_looic
-## model_M4   125.8 
-## model_M3   125.1 
-## model_M1   129.6 
-## model_M2   129.7
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{loo\_compare}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{criterion =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{print}\NormalTok{(}\AttributeTok{simplify =}\NormalTok{ F)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##          elpd_diff se_diff elpd_waic se_elpd_waic p_waic  se_p_waic waic   
-## model_M4     0.0       0.0 -5092.8      62.9        125.9     4.0   10185.5
-## model_M3   -22.1      10.0 -5114.9      62.5         76.8     2.4   10229.7
-## model_M1  -256.3      24.3 -5349.1      64.8         72.5     1.6   10698.2
-## model_M2  -258.4      23.6 -5351.2      64.8         39.6     1.0   10702.4
-##          se_waic
-## model_M4   125.8
-## model_M3   125.1
-## model_M1   129.6
-## model_M2   129.7
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# model weights}
-\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M1 model_M2 model_M3 model_M4 
-##        0        0        0        1
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M1 model_M2 model_M3 model_M4 
-##        0        0        0        1
-\end{verbatim}
-
-\normalsize
-
 Clearly, both weighting schemes prefer model M4.
 
-\subsection{Plot congruency effects and subject-specific fits for model M4.}\label{plot-congruency-effects-and-subject-specific-fits-for-model-m4.}
+\subsubsection{4.2.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}
 
@@
 
-\section{Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
+{[}{[}let's have a paragraph on how we might interpret these plots.{]}{]}
+
+\subsection{4.3 Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
 
@@
 
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_M3\_f }\OtherTok{\textless{}{-}} \FunctionTok{glmer}\NormalTok{(event }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid),}
-
-              \CommentTok{\# control parameters, data set, and complementary log{-}log link function}
-              \AttributeTok{control =} \FunctionTok{glmerControl}\NormalTok{(}\AttributeTok{optimizer =} \FunctionTok{c}\NormalTok{(}\StringTok{"nlminbwrap"}\NormalTok{),}
-                                     \AttributeTok{optCtrl =} \FunctionTok{list}\NormalTok{(}\AttributeTok{maxfun=}\DecValTok{10000000}\NormalTok{)), }
-              \AttributeTok{data=}\NormalTok{M3\_data, }
-              \AttributeTok{family=}\FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-In Figure 7 we compere the parameter estimates of model M3 from brm() with those of glmer().
+In Figure 7 we compare the parameter estimates of model M3 from brm() with those of glmer().
 
@@
 However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
-This is of course one of the reasons why Bayesian modeling has become so popular in the last years.
-
-\section{Tutorial 4: Calculating descriptive statistics when there are two independent variables}\label{tutorial-4-calculating-descriptive-statistics-when-there-are-two-independent-variables}
-
-In this final tutorial we illustrate how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
+This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
+But the price you pay for being able to fit more complex models in a Bayesian framework is computation time.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorial 2 took several hours to build.
+
+\subsection{4.4 Tutorial 4: Generalising to a more complex design}\label{tutorial-4-generalising-to-a-more-complex-design}
+
+So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
+But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
+The purpose of Tutorial 4, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
+We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
+
+To this end, Tutorial 4 illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in Tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
 Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
@@
 
-\section{Discussion}\label{discussion}
-
-We noticed that many researchers are still reluctant to abandon analysis-of-variance and switch to event history analysis when analyzing time-to-event data.
-By providing this tutorial, we hope that researchers will start using descriptive and inferential hazard analysis more often, due to the many advantages described below.
-While we focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
-We also discuss the issue of individual differences, limitations, and extensions.
-
-\subsection{Advantages of hazard analysis}\label{advantages-of-hazard-analysis}
+\section{5. Discussion}\label{discussion}
+
+This main motivation for writing this paper is the observation that event history analysis remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA provides compared to more conventional analyses.
+By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA in the future.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA, which is the increase in analytical complexity compared to mean-average comparisons.
+While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
+In the following, we discuss issues relating to individual differences, limitations of the approach, and future extensions.
+
+\subsection{5.1 Advantages of hazard analysis}\label{advantages-of-hazard-analysis}
 
@@
 
-\subsection{Individual differences}\label{individual-differences}
+\subsection{5.2 Individual differences}\label{individual-differences}
 
@@
 
-\subsection{Limitation(s)}\label{limitations}
-
-Compared to the orthodox method -- comparing means with ANOVA --, the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters.
-Another issue that might initially look as a limitation is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
-However, as nature does not reveal itself easily, obtaining insight into behavioral dynamics simply requires more data per condition than is usually collected under the orthodox method.
-In general, there is a tradeoff between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function; We therefore recommend to design as many trials as possible per experimental condition given the available resources.
-
-\subsection{Extensions}\label{extensions}
+\subsection{5.3 Limitation(s)}\label{limitations}
+
+Compared to the orthodox method -- comparing mean-averages between conditions --, the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
+We, therefore, recommend to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
+For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect?
+It be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
+Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
+
+\subsection{5.4 Extensions}\label{extensions}
 
 The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this event consitutes a single transition between an ``idle'' state and a ``responded'' state.
+For RT data, this event constitutes a single transition between an ``idle'' state and a ``responded'' state.
 However, in certain situations, more than one event of interest might exist.
-For example, an individual might transition back and forth between a ``healthy'' state and a ``depression'' state, before being absorbed in a final ``death'' state.
+For example, in a more medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depression'' state, before being absorbed in a final ``death'' state.
 When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
@@
 Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
-
-\section{Conclusions}\label{conclusions}
-
-RT and accuracy distributions are a rich source of information on the time course of cognitive processing.
-The changing effects of our experimental manipulations with increases in waiting time become strikingly clear when looking at response hazards and microlevel speed-accuracy trade-off functions.
-We hope that experimental psychologists and cognitive neuroscientists are less reluctant to embrace event history analysis after reading this tutorial.
+{[}{[}give a concrete example for this latter point{]}{]}
+
+\section{6. Conclusions}\label{conclusions}
+
+RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
+We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built functions, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.
+On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.
 
@@
 
-\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
-Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
-
 \bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2018}
@@
 
-\bibitem[\citeproctext]{ref-schmidtResponseInhibitionNegative2022}
-Schmidt, T., Panis, S., Wolkersdorfer, M.
-P., \& Vorberg, D.
-(2022).
-Response inhibition in the {Negative Compatibility Effect} in the absence of inhibitory stimulus features.
-\emph{Open Psychology}, \emph{4}(1), 219--230.
-\url{https://doi.org/10.1515/psych-2022-0012}
-
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}

### 2024-10-11 manuscript/manuscript.Rmd
change cite R options

@@
 # R references
-r_refs(file = "r-references.bib")
+r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation...
 my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) ## 
@@
 To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
-In a standard stimulus-response-compatibility paradigm, there is a prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
+In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
 The prime can then be congruent or incongruent with the target.
+Figure 3 shows that the early upswing in hazard is equal for both conditions, and that early responses are always correct in ..
+and always incorrect in the incongruent condtion.
 Taken together, the results show that for early responses (< bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later responses, the prime can be overridden, as both conditions are now always correct.
+And then for later bins, response hazard is lower in incongruent compared to congruent trials, as .....the prime can be overridden, as both conditions are now always correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-And such a conclusion is not supporting when the effects are explored over a timeline.
+And such a conclusion is not supported when the effects are explored over a timeline.
 Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.

### 2024-10-11 manuscript/manuscript.tex
change cite R options

@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 % definitions for citeproc citations
@@
 
-In the context of experimental psychology, it is common for participants to be presented with a task that has a right and wrong answer.
+In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer.
 For example, a task may involve choosing between two response options with only one of them being correct.
@@
 Integrating results between hazard and conditional accuracy functions can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer.
-In a standard stimulus-response-compatibility paradigm, there is prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
+To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
+In a standard stimulus-response-compatibility paradigm, there is a prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
 The prime can then be congruent or incongruent with the target.
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.164; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 Web links are printed in bold.
@@
 Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
-Functional programming basically means you don't write your own loops but instead use functions that
+Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
+{[}{[}more here, as necessary{]}{]}.
 
@@
 
-{[}{[}here we might have a figure that visulises the different strucutre of data for hazard analysis and conditional accuracy analysis.
-it is just easier to see visually.
-e.g., conditional acc is how it comes e.g., an RT per trial, whereas hazard needs to be wrangled so that the structure changes.
-I think that is worth visualising.
-e.g., two columns side-by-side.
-top XX rows visualised.
-And if we give them labels, such as `person-trial' data and `person-trial-bin' data, it will helps folsk to follow along.
-I include some text below as an example.{]}{]}
-
 Our data wrangling procedures serve two related purposes.
@@
 In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs when it generates a 1.
+Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
@@
 
-{\centering \includegraphics[height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[width=12in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-
+Therefore, {[}{[}why should people care?
+What is the functional relevance for exp psych and researchers?{]}{]}
+
+{[}{[}This para needs to be way shorter and easier to read or we get rid of it{]}{]}
 Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
@@
 Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
+
 But even when using a response deadline, participants can differ qualitatively in the effects they display (see Panis, 2020).
@@
 
-Compared to the orthodox method -- comparing mean-averages between conditions --, the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
 Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
-We, therefore, recommend to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
+Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
 For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect?
-It be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
+After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
 Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
@@
 However, in certain situations, more than one event of interest might exist.
-For example, in a more medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depression'' state, before being absorbed in a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state models which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
+For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before a final ``death'' state.
+When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial.
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (REF).
 {[}{[}give a concrete example for this latter point{]}{]}
@@
 RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
-We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built functions, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.
+We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.
 On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.

### 2024-10-11 manuscript/manuscript.Rmd
add supplementary Material file

@@
 classoption       : "man"
+appendix          : SupplementaryMaterial.Rmd
 output            : papaja::apa6_pdf
+
 # output            : papaja::apa6_word
@@
 :::
+
+\newpage
+
+```{r child = "SupplementaryMaterial.Rmd"}
+```

### 2024-10-11 manuscript/manuscript.tex
add supplementary Material file

@@
 To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
-In a standard stimulus-response-compatibility paradigm, there is a prime stimulus (e.g., a double arrow) followed by a target stimulus (single arrow).
+In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
 The prime can then be congruent or incongruent with the target.
+Figure 3 shows that the early upswing in hazard is equal for both conditions, and that early responses are always correct in ..
+and always incorrect in the incongruent condtion.
 Taken together, the results show that for early responses (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later responses, the prime can be overridden, as both conditions are now always correct.
+And then for later bins, response hazard is lower in incongruent compared to congruent trials, as \ldots..the prime can be overridden, as both conditions are now always correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-And such a conclusion is not supporting when the effects are explored over a timeline.
+And such a conclusion is not supported when the effects are explored over a timeline.
 Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
@@
 
+\newpage
+
+\section{Supplementary Material}\label{supplementary-material}
+
+here is the supp
+
 \end{document}

### 2024-10-16 manuscript/SupplementaryMaterial.Rmd
add info supp

--- 2024-09-18 notes/manuscript_notes.Rmd
@@
----
-title: "manuscript_notes"
-author: "Rich"
-date: "2024-08-06"
-output: html_document
-editor_options: 
-  chunk_output_type: console
----
+# Supplementary material
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+## A. Definitions of discrete-time hazard, surivor, and conditional accuracy functions
 
-This file is a way to provide feedback on draft versions of the manuscript. 
-I can edit the manuscript directly of course and I have done in a few places (which git should track), but this is a space to provide comments.
+cfr paper meemory hazard function
 
-I'm not sure of this is the best way to go. But so be it.
+## B. Custom functions for descriptive discrete-time hazard analysis
 
-At the end we can add this folder to .gitignore, so we don't need to post this online.
+We defined a few custom functions that we list here.
 
-# Aug 6th 2024 #
+### Identifying right-censored data and calculating the time bin index
 
-## Abstract ##
+###
 
-You say: To move beyond mean performance measures, various distributional analyses have been proposed.
+## C. Link functions
 
-This makes complete sense. But I would say it differently. I would emphasise in a sentence or two why mean performance measures are limited, and thus highlight the benefit of non-mean measures.
+## D. Regression equations
 
-And maybe try to find simpler language for distributional measures.
-Again, it is completely precise and accurate, but I worry that most folks will read it and think "that aint anything I care about" when in actual fact it might be extremely relevant.
+An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
 
-So, in other words, I think the abstract needs a clearer hook for the generalist experimental psych reader.
-Something like ...
-mean avg measures (across trials) hide/obscure a lot of information about psychological processes, such as when psychological states occur in time.
-In contrast, blah blah approaches reveal how psych processes occur in specific time windows (etc.
-etc.).
-And this changes key conclusions about psych processes.
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)].
 
-Here we provide a tutorial on how to do this [[and then return to the details of the tutorial as you already state]].
-Even though I have written a lot here, I expect the changes to the abstract to be small.
-I just wanted to try to make my point clear here to you.
-I hope that makes sense.
+The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
+The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
+Predictors can be discrete, continuous, and time-varying or time-invariant.
+For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
+To interpret the effects of a predictor,its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+When using the logit link, exponentiating a $\beta$ parameter results in an odds ratio.
 
-## Introduction ##
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X~1~) can be written as follows:
 
-I would slightly re-structure the introduction. I would use some subsections in the intro maybe also to organise ideas.
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~].
 
-The main change that I propose is to have a first section that provides the motivation in simple terms (non-modelling terms),
+## E. Prior distributions
 
-as well as a roadmap for what is to come next in the rest of the paper, so that folks know what to expect.
-This is especially important for reviews/tutorials etc., since it doesn't follow the standard empirical paper approach i.e., intro, methods, results ...
+To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
 
-Then I would move all of the modelling stuff that is currently in the intro into an entirely new section. Here are some ideas on structure with example sections numbered...
-
-1. motivation = means vs distribution/timewindow approach. Benefits etc.
-
-I would have a plot in this part of the intro.
-Just real or simulated raw data - like we discussed in our meeting on Monday for your future talk - bar plots of means between conditions juxtaposed with a distributional analysis across time.
-
-I think the visually hits home the basic point about how timing is important.
-Then all the modelling and lifetable stuff  that comes later is important as a method, but we get in there nice and early with a clear message on why it is so ubiquitous in exp psych (e.g., almost everyone plots mean RT, for example).
-So we make the benefit/value really clear, really early.
-
-2.
-Roadmap of the paper, which outlines what to expect.
-e.g., We first provide a brief overview of hazard models to orient the reader to the basic concepts and ideas that we will use through the paper.
-However, this will remain relatively short, and for detailed treatment, see blah blah REFs.
-
-And we *must* provide some figures here.
-Figures really help with stats stuff, because it brings more people in.
-Can you think of some figures to illustrate key modelling concepts?
-Or do you think the mean vs distribution figure above in point (1) will be enough?
-
-We then provide 4 different tutorials, each of which is written is R code and publicly available.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can also analyse time to event data....
-Tutorial 1 blah. 2 blah. 3 blah. 4 blah. Provide a sentence or two on each tutorial.
-
-## Background on EHA analyses ##
-
-Here is where you move the modelling stuff into a separate section. 
-This is where you get into the more technical stuff.
-keep this short and refer to other, more detailed reviews (including your own papers).
-The reason for keeping this section short is that we want to focus on a worked examples of how to do the analyses, rather than justify why you should use these techniques.
-Loads of papers have focussed on why before.
-And let's say this explicitly at the start of this section, so that if anyone complains that we need more background info., we have already explained why we kept it short.
-
-## Tutorial 1 ##
-
-## Tutorial 2 ##
-
-...
+To gain a sense of what prior *cloglog* values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
+Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).

### 2024-10-16 manuscript/manuscript.Rmd
add info supp

@@
 After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response could occur in 550 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+For example, the first response could occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+Continuous RT data is treated here as interval-censored data.
 
 Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X).
-The equations that define both of these functions are reported in supplementary materials (Supp XX).
-The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+The equations that define both of these functions are reported in supplementary materials (part A).
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occured in the past, i.e., in one of the prior bins.
+This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs after bin t.
-In other words, the survivor function reflects the likelihood that the event occurs in a subsequent timebin.
+In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
 
@@
 Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+
 Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
@@
 Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
+
+We did not discuss continuous-time hazard analysis.
+As indicated by --allison -- learning discrete-time methods first will help in learning continuous-time methods.
+Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
+However, they require much more data to estimate the continuous-time hazard (rate) function well.
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than ~200 trials per condition per experiment.
 

### 2024-10-16 Tutorial_1a.Rmd
add Tutorial 1a and custom functions file

--- 2024-09-17 Tutorial_1_descriptive_stats/Tutorial_1.Rmd
@@
 
-Load the functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 210) as above (line 88).
+Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 118 in custom_functions.R) as above (line 88 in Tutorial_1a).
 
 ```{r load-functions}
-# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
-censor <- function(df, timeout, bin_width){
-  if(!(timeout %% bin_width == 0)){
-    return("The censoring time must be a multiple of the bin width!")
-  }
-  if(timeout < 0 | bin_width < 0){
-    return("Both timeout and bin_width must be larger than 0!")
-  }
-  df %>% mutate(right_censored = 0,
-                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
-                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
-                drt = ceiling(rtc/bin_width), # drt = discrete response time
-                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
-}
-
-# function ptb creates a person-trial-bin oriented data set
-ptb <- function(df){
-  df %>% uncount(weights = drt) %>% # create drt rows per trial
-         group_by(trial) %>% 
-         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
-         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
-         ungroup()
-}
-
-# function setup_lt sets up a life table for each level of condition (1 independent variable)
-setup_lt <- function(ptb){
-  ptb %>% mutate(event = str_c("event", event)) %>%
-          group_by(condition,period) %>% 
-          count(event) %>% 
-          ungroup() %>% 
-          pivot_wider(names_from = event,
-                      values_from = n) %>% 
-          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
-                 event1 = ifelse(is.na(event1),0,event1),
-                 risk_set = event0 + event1) %>% # define the risk set
-          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
-          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
-          group_by(condition) %>%
-          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
-                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
-                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
-          ungroup() 
-}
-
-# function calc_ca calculates the conditional accuracies 
-calc_ca <- function(df){
-  df %>% filter(right_censored==0) %>%
-         group_by(condition,drt,cens_time,bin_width) %>%
-         summarize(ca = mean(acc) %>% round(digits = 2),
-                   n = n(),
-                   .groups = 'drop') %>%
-         ungroup() %>%
-         mutate(period = drt,
-                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
-         select(-drt)
-}
-
-# function join_lt_ca joins the conditional accuracies to the life tables
-join_lt_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}
-
-# function extract_median is used to extract the S(t).50 quantiles (i.e., the estimated median RTs)
-extract_median <- function(df){
-  above_pct50 <- df %>% 
-      group_by(condition) %>%
-      filter(survival >= .5) %>% 
-      slice(n()) # take last row
-  below_pct50 <- df %>% 
-      group_by(condition) %>%
-      filter(survival < .5) %>% 
-      slice(1) # take first row
-  # pull period above
-  period_above <- pull(above_pct50, period)
-  # pull survivor function values
-  survival_above <- pull(above_pct50, survival)
-  survival_below <- pull(below_pct50, survival)
-  # estimate median by interpolation
-  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
-}
-
-# function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
-plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects (see line 244)
-  library(patchwork)
-  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
-  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
-  median_period <- extract_median(df)
-  n_conditions <- nlevels(df$condition)
-  data_median <- c()
-  for(i in 1:n_conditions){
-    data_median <- append(data_median, c(median_period[i], median_period[i]))
-  }
-
-  data_medians <- tibble(period= data_median,
-                         survival = rep(c(.5, 0),n_conditions),
-                         condition = rep(1:n_conditions, each=2))
-# plot the hazard functions
-p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=hazard)) +
-  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
-  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits = c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,haz_yaxis)) +
-  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
-  theme(legend.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        legend.position = "top")
-# plot the survivor functions
-p2 <-df %>%
-  ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=survival), show.legend = F) +
-  geom_point(aes(y=survival), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
-  # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
-  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
-            data = data_medians, 
-            linetype = 3, show.legend = F) +
-
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits=c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="S(t)",
-       colour="Condition") +
-  theme(panel.grid = element_blank())
-# plot the conditional accuracy functions
-p3 <-df %>%
-  ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=ca), show.legend = F) +
-  geom_point(aes(y=ca), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits=c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
-       colour="Condition") +
-  theme(panel.grid = element_blank())
-
-p1/p2/p3
-}
+source("custom_functions.R")
 ```
 
-Apply the functions
+Apply the custom functions
 
@@
 # view plot (column 8) for subject 1. 
-pluck(data_final, 8, 1)
+pluck(data_final, 8, 6)
 
@@
 ```
+
+Combine plots of 2 subjects.
+
+```{r plot-two-subjects}
+pa <- pluck(data_final, 8, 3)
+pb <- pluck(data_final, 8, 6)
+
+pa | pb 
+```
+
+```{r save-plot-two-subjects}
+ggsave("Tutorial_1_descriptive_stats/figures/Plot_subjects3and6_PanisSchmidt.png", width = 8, height=8, dpi=600)
+```
+
+Create a life table.
 
@@
 
-```{r aggregated-data-plot}
+```{r agg-data}
 # create single subject number, and unique trial number
@@
                                pid = 1)
+```
+
+```{r agg-descr}
 # apply functions
@@
                      mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_agg,1))  # create plots 
+```
+
+```{r agg-plot}
 # view plot
-pluck(data_final, 8, 1)
+pluck(data_agg_final, 8, 1)
 ```

### 2024-10-16 custom_functions.R

@@ -1,0 +1,237 @@
+# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
+censor <- function(df, timeout, bin_width){
+  if(!(timeout %% bin_width == 0)){
+    return("The censoring time must be a multiple of the bin width!")
+  }
+  if(timeout < 0 | bin_width < 0){
+    return("Both timeout and bin_width must be larger than 0!")
+  }
+  df %>% mutate(right_censored = 0,
+                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
+                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
+                drt = ceiling(rtc/bin_width), # drt = discrete RT or time bin rank
+                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
+}
+
+# function ptb creates a person-trial-bin oriented data set
+ptb <- function(df){
+  df %>% uncount(weights = drt) %>% # create drt rows per trial
+    group_by(trial) %>% 
+    mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
+    mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
+    ungroup()
+}
+
+# function setup_lt sets up a life table for each level of condition (1 independent variable)
+setup_lt <- function(ptb){
+  ptb %>% mutate(event = str_c("event", event)) %>%
+    group_by(condition,period) %>% 
+    count(event) %>% 
+    ungroup() %>% 
+    pivot_wider(names_from = event,
+                values_from = n) %>% 
+    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+           event1 = ifelse(is.na(event1),0,event1),
+           risk_set = event0 + event1) %>% # define the risk set
+    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
+           pmass = event1/max(risk_set),
[…]

### 2024-10-16 Tutorial_1a.Rmd
add new Tutorial 1a

@@
 ---
-title: "Tutorial_1"
+title: "Tutorial_1a"
 author: "Sven Panis"
@@
 
-This Tutorial_1 file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
+This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
 
@@
 
-```{r save-plots}
+```{r save-plots, eval=F}
 # save the plot for each subject
 map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 6, height = 8, dpi = 600)
-```
-
-Combine plots of 2 subjects.
-
-```{r plot-two-subjects}
-pa <- pluck(data_final, 8, 3)
-pb <- pluck(data_final, 8, 6)
-
-pa | pb 
-```
-
-```{r save-plot-two-subjects}
-ggsave("Tutorial_1_descriptive_stats/figures/Plot_subjects3and6_PanisSchmidt.png", width = 8, height=8, dpi=600)
 ```
@@
 
-```{r save-single-lifetable}
+```{r save-single-lifetable, eval=F}
 write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
@@
 
-```{r save-file-hazard-modeling}
+```{r save-file-hazard-modeling, eval=F}
 write_csv(input_modeling_hazard, file="Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
@@
 
-```{r save-file-ca-modeling}
+```{r save-file-ca-modeling, eval=F}
 write_csv(input_modeling_ca, file="Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
@@
 
-```{r agg-plot}
+```{r agg-plot, warning=F}
 # view plot

### 2024-10-16 manuscript/SupplementaryMaterial.Rmd
update supp

@@
 
-cfr paper meemory hazard function
+The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].
+
+After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+This latter function is also known as the micro-level speed-accuracy tradeoff function.
+
+When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
+The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
 
@@
 ## C. Link functions
+
+(ref:descr-fig3-caption) The logit and cloglog link functions.
+
+```{r plot-link-functions, fig.cap = "(ref:descr-fig3-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
+```
 
@@
 
+(ref:descr-fig4-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+
+```{r plot-priors, fig.cap = "(ref:descr-fig4-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")
+```
+
 To gain a sense of what prior *cloglog* values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).

### 2024-10-16 Tutorial_1b.Rmd
add tut1b

--- 2024-09-17 Tutorial_4.Rmd
@@
 ---
-title: "Tutorial_4"
+title: "Tutorial_1b"
 author: "Sven Panis"
@@
 
-Tutorial_4 illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
+Tutorial_1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
 
@@
 head(raw)
-tail(raw)
-glimpse(raw)
 ```
@@
 
-# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).
+# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), P(t), and ca(t)).
 
@@
 ```{r load-functions}
-# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
-censor <- function(df, timeout, bin_width){
-  if(!(timeout %% bin_width == 0)){
-    return("The censoring time must be a multiple of the bin width!")
-  }
-  if(timeout < 0 | bin_width < 0){
-    return("Both timeout and bin_width must be larger than 0!")
-  }
-  df %>% mutate(right_censored = 0,
-                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
-                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
-                drt = ceiling(rtc/bin_width), # drt = discrete response time
-                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
-}
-
-# function ptb creates a person-trial-bin oriented data set
-ptb <- function(df){
-  df %>% uncount(weights = drt) %>% # create drt rows per trial
-         group_by(trial) %>% 
-         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
-         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
-         ungroup()
-}
-
-# function setup_lt sets up a life table for each level 2 independent variables
-setup_lt_2IV <- function(ptb){
-  ptb %>% mutate(event = str_c("event", event)) %>%
-          group_by(condition1,condition2,period) %>% 
-          count(event) %>% 
-          ungroup() %>% 
-          pivot_wider(names_from = event,
-                      values_from = n) %>% 
-          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
-                 event1 = ifelse(is.na(event1),0,event1),
-                 risk_set = event0 + event1) %>% # define the risk set
-          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
-          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
-          group_by(condition1,condition2) %>%
-          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
-                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
-                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
-          ungroup() 
-}
-
-# function calc_ca calculates the conditional accuracies 
-calc_ca_2IV <- function(df){
-  df %>% filter(right_censored==0) %>%
-         group_by(condition1,condition2,drt,cens_time,bin_width) %>%
-         summarize(ca = mean(acc) %>% round(digits = 2),
-                   n = n(),
-                   .groups = 'drop') %>%
-         ungroup() %>%
-         mutate(period = drt,
-                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
-         select(-drt)
-}
-
-# function join_lt_ca_2IV joins the conditional accuracies to the life tables when there are two independent variables
-join_lt_ca_2IV <- function(df1,df2){df1 %>% left_join(df2, join_by(condition1,condition2,period))}
-
-# function extract_median_2IV is used to extract the S(t).50 quantiles when there are two independent variables
-extract_median_2IV <- function(df){
-  above_pct50 <- df %>% 
-      group_by(condition1,condition2) %>%
-      filter(survival >= .5) %>% 
-      slice(n()) # take last row
-  below_pct50 <- df %>% 
-      group_by(condition1,condition2) %>%
-      filter(survival < .5) %>% 
-      slice(1) # take first row
-  # pull period above
-  period_above <- pull(above_pct50, period)
-  # pull survivor function values
-  survival_above <- pull(above_pct50, survival)
-  survival_below <- pull(below_pct50, survival)
-  # estimate median by interpolation
-  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
-}
-
-# function plot_eha_2IV plots the hazard, survivor, and ca(t) functions, when there are two independent variables
-plot_eha_2IV <- function(df,subj,haz_yaxis){ 
-  library(patchwork)
-  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
-  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
-  median_period <- extract_median_2IV(df)
-  n_iv1 <- nlevels(df$condition1) 
-  n_iv2 <- nlevels(df$condition2) 
-  n_conditions <- n_iv1*n_iv2
-  data_median <- c()
-  for(i in 1:n_conditions){
-    data_median <- append(data_median, c(median_period[i], median_period[i]))
-  }
-  # Create correctly named factors (here: prime and mask) and specify the same levels and labels as above
-  data_medians <- tibble(period= data_median,
-                         survival = rep(c(.5, 0),n_conditions),
-                         prime =  factor(rep(rep(1:n_iv2, each=2),n_iv1),levels=c(1:3),labels=c("blank","congruent","incongruent")),
-                         mask = factor(rep(1:n_iv1, each=2*n_iv2),levels=c(1:4),labels=c("nomask", "relevant", "irrelevant","lines")))
-
-# plot the hazard functions
-p1 <-df %>% 
-  mutate(mask = condition1,       # Create correctly named factor
-         prime = condition2) %>%  # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=hazard)) +
-  geom_point(aes(y=hazard), size=1) + labs(color="Prime") +
-  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits = c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
-  theme(legend.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-# plot the survivor functions
-p2 <-df %>%
-  mutate(mask = condition1,      # Create correctly named factor
-         prime = condition2) %>% # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=survival), show.legend = F) +
-  geom_point(aes(y=survival), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
-  # add vertical lines to indicate the estimated median RTs with geom_path()
-  geom_path(aes(x=period, y=survival),
-            data = data_medians, 
-            linetype = 3, show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits=c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="S(t)") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-# plot the conditional accuracy functions
-p3 <-df %>%
-  mutate(mask = condition1,       # Create correctly named factor
-         prime = condition2) %>%  # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=ca), show.legend = F) +
-  geom_point(aes(y=ca), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits=c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="Time bin t's endpoint (ms)", y="ca(t)") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-p1/p2/p3
-}
+source("custom_functions.R")
 ```
@@
 
-```{r save-plots, warning=F}
+```{r save-plots, warning=F, eval=F}
 # save the plot for each subject
-map2(paste0("Tutorial_4_descriptive_2IV/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 8, dpi = 600)
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 10, dpi = 300)
 ```

### 2024-10-16 custom_functions.R
update custom functions file

@@ -230,8 +230,155 @@ plot_eha_agg <- function(df,subj,haz_yaxis){
     theme(panel.grid = element_blank(),
           axis.text.x = element_text(angle=90))
   
-  
   (p1|p3) / (p2|p4)
+}
+
+
+# function setup_lt_2IV sets up life tables for 2 independent variables
+setup_lt_2IV <- function(ptb){
+  ptb %>% mutate(event = str_c("event", event)) %>%
+    group_by(condition1,condition2,period) %>% 
+    count(event) %>% 
+    ungroup() %>% 
+    pivot_wider(names_from = event,
+                values_from = n) %>% 
+    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+           event1 = ifelse(is.na(event1),0,event1),
+           risk_set = event0 + event1) %>% # define the risk set
+    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4),
+           pmass = event1/max(risk_set),
+           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
+    group_by(condition1,condition2) %>%
+    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
+           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
+           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+    ungroup() 
+}
+
+# function calc_ca_2IV calculates the conditional accuracies for 2 IVs
+calc_ca_2IV <- function(df){
+  df %>% filter(right_censored==0) %>%
+    group_by(condition1,condition2,drt,cens_time,bin_width) %>%
+    summarize(ca = mean(acc) %>% round(digits = 2),
+              n = n(),
+              .groups = 'drop') %>%
+    ungroup() %>%
+    mutate(period = drt,
+           se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
+    select(-drt)
+}
+
+# function join_lt_ca_2IV joins the conditional accuracies to the life tables when there are two independent variables
+join_lt_ca_2IV <- function(df1,df2){df1 %>% left_join(df2, join_by(condition1,condition2,period))}
+
[…]

### 2024-10-16 manuscript/SupplementaryMaterial.Rmd
update supp

@@
 The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].
+After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+Discrete-time EHA focuses on the discrete-time hazard function
 
-After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t).
+\noindent h(t) = P(RT = t| RT $\geq$ t)   \hfill  (1)
+
+\noindent and the discrete-time survivor function
+
+\noindent S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)]  \hfill  (2)
+
+\noindent and not on the probability mass function
+
+\noindent P(t) = P(RT = t) = h(t).S(t-1)  \hfill  (3)
+
+\noindent nor the cumulative distribution function
+
+\noindent F(t) = 1-S(t) = P(RT $\leq$ t).   \hfill  (4)
+
 The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
@@
 
-For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
+
+\noindent ca(t) = P(correct | RT = t)   \hfill  (5)
+
+which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 This latter function is also known as the micro-level speed-accuracy tradeoff function.
@@
 
-We defined a few custom functions that we list here.
+We defined 13 custom functions that we list here.
 
-### Identifying right-censored data and calculating the time bin index
-
-###
+* censor(df,timeout,bin_width)
+* ptb(df)
+* setup_lt(ptb)
+* setup_lt_2IV(ptb)
+* calc_ca(df)
+* calc_ca_2IV(df)
+* join_lt_ca(df1,df2)
+* join_lt_ca_2IV(df1,df2)
+* extract_median(df)
+* extract_median_2IV(df)
+* plot_eha(df,subj,haz_yaxis)
+* plot_eha_2IV(df,subj,haz_yaxis)
+* plot_eha_agg(df,subj,haz_yaxis)
 

### 2024-10-16 manuscript/manuscript.Rmd
update ms.

@@
 For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distribution differs (Figure 1, top row).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the rt distribution differs in each case (Figure 1, bottom row).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distribution differs in each case (Figure 1, bottom row).
 Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distribution would lead to a different conclusion in every case.
@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to theory building toolkit.
+As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to our theory building toolkit.
 
@@
 
-In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (REFs) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials [@allisonSurvivalAnalysisUsing2010] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
 Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
@@
 
-For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs).
+For a comprehensive background context to hazard analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003].
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
@@
 We also use a "keep it maximal" approach to specifying varying (or random) effects (Barr et al., 2013).
-This means that wherever possible we include varying intercepts and slopes per pid.
+This means that wherever possible we include varying intercepts and slopes per participant
 To make inferences, we use two main approaches.
@@
 
-[[provide a short summary of the aims and scope of each tutorial, as well as the links between them]].
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own user-defined functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+Tutorials 1a and 1b show how to calculate and plot the descriptive statistics when there are one and two independent variables, respectively.
+Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
+Tutorials 3a and 3b show how to implement multilevel models for hazard and conditional accuracy in the frequentist framework, respectively.
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
 
@@
 
-1a. Wrangle raw data and descriptive stats (T1).
-1b.
-
-2a.
-2b.
-
-3a.
-3b.
-
-Inferential stats (T2 and T3).
+1a. Wrangle raw data and descriptive stats for one independent variable.
+1b. Wrangle raw data and descriptive stats for two independent variables.
+
+2a. Bayesian multilevel modeling for h(t)
+2b. Bayesian multilevel modeling for ca(t)
+
+3a. Frequentist multilevel modeling for h(t)
+3b. Frequentist multilevel modeling for ca(t)
 
@@
 
-## 4.1 Tutorial 1: Calculating descriptive statistics using a life table
+## 4.1 Tutorial 1a: Calculating descriptive statistics using a life table
 
@@
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
-Likewise, the event could occur in bin 1 there would only be 1 row of data for that trial.
+Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.
 
@@
 
-There are several data wrangling steps to be taken. First, we need to load the data before (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
+There are several data wrangling steps to be taken. First, we need to load the data before we (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
 
@@
 
-In the code of Tutorial 1, this is accomplished as follows.
+In the code of Tutorial 1a, this is accomplished as follows.
 
@@
 
-```{r load-functions, echo=F}
-# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
-censor <- function(df, timeout, bin_width){
-  if(!(timeout %% bin_width == 0)){
-    return("The censoring time must be a multiple of the bin width!")
-  }
-  if(timeout < 0 | bin_width < 0){
-    return("Both timeout and bin_width must be larger than 0!")
-  }
-  df %>% mutate(right_censored = 0,
-                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
-                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
-                drt = ceiling(rtc/bin_width), # drt = discrete response time
-                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
-}
-
-# function ptb creates a person-trial-bin oriented data set
-ptb <- function(df){
-  df %>% uncount(weights = drt) %>% # create drt rows per trial
-         group_by(trial) %>% 
-         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
-         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
-         ungroup()
-}
-
-# function setup_lt sets up a life table for each level of condition (1 independent variable)
-setup_lt <- function(ptb){
-  ptb %>% mutate(event = str_c("event", event)) %>%
-          group_by(condition,period) %>% 
-          count(event) %>% 
-          ungroup() %>% 
-          pivot_wider(names_from = event,
-                      values_from = n) %>% 
-          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
-                 event1 = ifelse(is.na(event1),0,event1),
-                 risk_set = event0 + event1) %>% # define the risk set
-          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
-          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
-          group_by(condition) %>%
-          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
-                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
-                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
-          ungroup() 
-}
-
-# function calc_ca calculates the conditional accuracies 
-calc_ca <- function(df){
-  df %>% filter(right_censored==0) %>%
-         group_by(condition,drt,cens_time,bin_width) %>%
-         summarize(ca = mean(acc) %>% round(digits = 2),
-                   n = n(),
-                   .groups = 'drop') %>%
-         ungroup() %>%
-         mutate(period = drt,
-                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
-         select(-drt)
-}
-
-# function join_lt_ca joins the conditional accuracies to the life tables
-join_lt_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}
-
-# function extract_median is used to extract the S(t).50 quantiles (i.e., the estimated median RTs)
-extract_median <- function(df){
-  above_pct50 <- df %>% 
-      group_by(condition) %>%
-      filter(survival > .5) %>% 
-      slice(n()) # take last row
-  below_pct50 <- df %>% 
-      group_by(condition) %>%
-      filter(survival < .5) %>% 
-      slice(1) # take first row
-  # pull period above
-  period_above <- pull(above_pct50, period)
-  # pull survivor function values
-  survival_above <- pull(above_pct50, survival)
-  survival_below <- pull(below_pct50, survival)
-  # estimate median by interpolation
-  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
-}
-
-# function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
-plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects (see line 244)
-  library(patchwork)
-  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
-  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
-  median_period <- extract_median(df)
-  n_conditions <- nlevels(df$condition)
-  data_median <- c()
-  for(i in 1:n_conditions){
-    data_median <- append(data_median, c(median_period[i], median_period[i]))
-  }
-
-  data_medians <- tibble(period= data_median,
-                         survival = rep(c(.5, 0),n_conditions),
-                         condition = rep(1:n_conditions, each=2))
-# plot the hazard functions
-p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=hazard)) +
-  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
-  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits = c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,haz_yaxis)) +
-  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
-  theme(legend.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        legend.position = "top")
-# plot the survivor functions
-p2 <-df %>%
-  ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=survival), show.legend = F) +
-  geom_point(aes(y=survival), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
-  # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
-  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
-            data = data_medians, 
-            linetype = 3, show.legend = F) +
-
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits=c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="S(t)",
-       colour="Condition") +
-  theme(panel.grid = element_blank())
-# plot the conditional accuracy functions
-p3 <-df %>%
-  ggplot(aes(x=period, color=condition, group=condition)) +
-  geom_line(aes(y=ca), show.legend = F) +
-  geom_point(aes(y=ca), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
-  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
-                     limits=c(0,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
-       colour="Condition") +
-  theme(panel.grid = element_blank())
-
-p1/p2/p3
-}
-```
-
 Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t).
@@
 \scriptsize
-```{r nest-apply-functions-tut1, echo = TRUE}
+```{r nest-apply-functions-tut1, echo = TRUE, eval=F}
 data_nested <- data_wr %>% group_nest(pid)
@@
 
+```{r load-data}
+load("../Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata") # loads data_final
+```
+
 For example, for participant 6, the estimated hazard values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
@@
 
-## 4.2 Tutorial 2: Fitting Bayesian hazard models
+## 4.2 Tutorial 1b: Generalising to a more complex design
+
+So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
+But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
+The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
+We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
+
+To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
+Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+(ref:descr-fig8-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
+
+```{r plot-2ivs, fig.cap = "(ref:descr-fig8-caption)", out.width='95%'}
+knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt.png")
+```
+
+## 4.3 Tutorial 2a: Fitting Bayesian hazard models
 
@@
 
-### 4.2.1 Hazard model considerations
+### 4.3.1 Hazard model considerations
 
@@
 
-### 4.2.2 Prior distributions
+### 4.3.2 Prior distributions
 
@@
 
-### 4.2.3 Model 1: A general specification of TIME, and main effects of congruency and trial number
+### 4.3.3 Model 1: A general specification of TIME, and main effects of congruency and trial number
 
@@
 
-### 4.2.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
+### 4.3.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
 
@@
 
-### 4.2.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
+### 4.3.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
 
@@
 
-### 4.2.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions
+### 4.3.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions
 
@@
 
-### 4.2.7 Compare the models.
+### 4.3.7 Compare the models.
 
@@
 
-### 4.2.8 Evaluate parameter estimates
+### 4.3.8 Evaluate parameter estimates
 
@@
 
-## 4.3 Tutorial 3: Fitting Frequentist hazard models
+## 4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
+
+## 4.5 Tutorial 3a: Fitting Frequentist hazard models
 
@@
 
-## 4.4 Tutorial 4: Generalising to a more complex design
-
-So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
-But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
-The purpose of Tutorial 4, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
-We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
-
-To this end, Tutorial 4 illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
-Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
-
-(ref:descr-fig8-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
-
-```{r plot-2ivs, fig.cap = "(ref:descr-fig8-caption)", out.width='95%'}
-knitr::include_graphics("../Tutorial_4_descriptive_2IV/Plot_2IV_for_subject6_PanisSchmidt.png")
-```
+## 4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models
 

### 2024-10-16 manuscript/manuscript.tex
update ms.

@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to theory building toolkit.
+As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to our theory building toolkit.
 
@@
 
-In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (REFs) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
 Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
@@
 
-For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (REFs).
+For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (Singer \& Willett, 2003).
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
@@
 After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response could occur in 550 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+For example, the first response could occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+Continuous RT data is treated here as interval-censored data.
 
 Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X).
-The equations that define both of these functions are reported in supplementary materials (Supp XX).
-The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+The equations that define both of these functions are reported in supplementary materials (part A).
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occured in the past, i.e., in one of the prior bins.
+This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t.
-In other words, the survivor function reflects the likelihood that the event occurs in a subsequent timebin.
+In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
 
@@
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013).
-This means that wherever possible we include varying intercepts and slopes per pid.
+This means that wherever possible we include varying intercepts and slopes per participant
 To make inferences, we use two main approaches.
@@
 
-{[}{[}provide a short summary of the aims and scope of each tutorial, as well as the links between them{]}{]}.
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own user-defined functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+Tutorials 1a and 1b show how to calculate and plot the descriptive statistics when there are one and two independent variables, respectively.
+Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
+Tutorials 3a and 3b show how to implement multilevel models for hazard and conditional accuracy in the frequentist framework, respectively.
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
 
@@
 
-\subsection{4.1 Tutorial 1: Calculating descriptive statistics using a life table}\label{tutorial-1-calculating-descriptive-statistics-using-a-life-table}
+\subsection{4.1 Tutorial 1a: Calculating descriptive statistics using a life table}\label{tutorial-1a-calculating-descriptive-statistics-using-a-life-table}
 
@@
 
-\subsection{4.2 Tutorial 2: Fitting Bayesian hazard models}\label{tutorial-2-fitting-bayesian-hazard-models}
+\subsection{4.2 Tutorial 1b: Generalising to a more complex design}\label{tutorial-1b-generalising-to-a-more-complex-design}
+
+So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
+But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
+The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
+We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
+
+To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
+Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt}
+
+}
+
+\caption{Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
+\end{figure}
+
+\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models}\label{tutorial-2a-fitting-bayesian-hazard-models}
 
@@
 
-\subsubsection{4.2.1 Hazard model considerations}\label{hazard-model-considerations}
+\subsubsection{4.3.1 Hazard model considerations}\label{hazard-model-considerations}
 
@@
 
-\subsubsection{4.2.2 Prior distributions}\label{prior-distributions}
+\subsubsection{4.3.2 Prior distributions}\label{prior-distributions}
 
@@
 
-\subsubsection{4.2.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+\subsubsection{4.3.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
 
@@
 
-\subsubsection{4.2.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
+\subsubsection{4.3.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
 
@@
 
-\subsubsection{4.2.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
+\subsubsection{4.3.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
 
@@
 
-\subsubsection{4.2.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
+\subsubsection{4.3.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
 
@@
 
-\subsubsection{4.2.7 Compare the models.}\label{compare-the-models.}
+\subsubsection{4.3.7 Compare the models.}\label{compare-the-models.}
 
@@
 
-\subsubsection{4.2.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}
+\subsubsection{4.3.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}
 
@@
 
-\subsection{4.3 Tutorial 3: Fitting Frequentist hazard models}\label{tutorial-3-fitting-frequentist-hazard-models}
+\subsection{4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}
+
+\subsection{4.5 Tutorial 3a: Fitting Frequentist hazard models}\label{tutorial-3a-fitting-frequentist-hazard-models}
 
@@
 
-\subsection{4.4 Tutorial 4: Generalising to a more complex design}\label{tutorial-4-generalising-to-a-more-complex-design}
-
-So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
-But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
-The purpose of Tutorial 4, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
-We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
-
-To this end, Tutorial 4 illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1, we simply present the sample-based functions for participant 6 in Figure 8.
-Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_4_descriptive_2IV/Plot_2IV_for_subject6_PanisSchmidt}
-
-}
-
-\caption{Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
-\end{figure}
+\subsection{4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models}\label{tutorial-3b-fitting-frequentist-conditional-accuracy-models}
 
@@
 Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
+
+We did not discuss continuous-time hazard analysis.
+As indicated by --allison -- learning discrete-time methods first will help in learning continuous-time methods.
+Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
+However, they require much more data to estimate the continuous-time hazard (rate) function well.
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.
 
@@
 
+\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
+Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
 \bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2018}
@@
 
-\section{Supplementary Material}\label{supplementary-material}
-
-here is the supp
+\section{Supplementary material}\label{supplementary-material}
+
+\subsection{A. Definitions of discrete-time hazard, surivor, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-and-conditional-accuracy-functions}
+
+The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991).
+After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+Discrete-time EHA focuses on the discrete-time hazard function
+
+\noindent h(t) = P(RT = t\textbar{} RT \(\geq\) t) \hfill  (1)
+
+\noindent and the discrete-time survivor function
+
+\noindent S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} \hfill  (2)
+
+\noindent and not on the probability mass function
+
+\noindent P(t) = P(RT = t) = h(t).S(t-1) \hfill  (3)
+
+\noindent nor the cumulative distribution function
+
+\noindent F(t) = 1-S(t) = P(RT \(\leq\) t). \hfill  (4)
+
+The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
+
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
+
+\noindent ca(t) = P(correct \textbar{} RT = t) \hfill  (5)
+
+which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+This latter function is also known as the micro-level speed-accuracy tradeoff function.
+
+When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
+The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+
+\subsection{B. Custom functions for descriptive discrete-time hazard analysis}\label{b.-custom-functions-for-descriptive-discrete-time-hazard-analysis}
+
+We defined 13 custom functions that we list here.
+
+\begin{itemize}
+\tightlist
+\item
+  censor(df,timeout,bin\_width)
+\item
+  ptb(df)
+\item
+  setup\_lt(ptb)
+\item
+  setup\_lt\_2IV(ptb)
+\item
+  calc\_ca(df)
+\item
+  calc\_ca\_2IV(df)
+\item
+  join\_lt\_ca(df1,df2)
+\item
+  join\_lt\_ca\_2IV(df1,df2)
+\item
+  extract\_median(df)
+\item
+  extract\_median\_2IV(df)
+\item
+  plot\_eha(df,subj,haz\_yaxis)
+\item
+  plot\_eha\_2IV(df,subj,haz\_yaxis)
+\item
+  plot\_eha\_agg(df,subj,haz\_yaxis)
+\end{itemize}
+
+\subsection{C. Link functions}\label{c.-link-functions}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}
+
+}
+
+\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
+\end{figure}
+
+\subsection{D. Regression equations}\label{d.-regression-equations}
+
+An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
+
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.
+
+The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
+The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
+The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
+Predictors can be discrete, continuous, and time-varying or time-invariant.
+For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth.
+To interpret the effects of a predictor,its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.
+
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
+
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.
+
+\subsection{E. Prior distributions}\label{e.-prior-distributions}
+
+To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}
+
+}
+
+\caption{Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
+\end{figure}
+
+To gain a sense of what prior \emph{cloglog} values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
+Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
+However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
+A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
 

### 2024-10-16 Tutorial_2a.Rmd
rename 2 tuts

--- 2024-09-22 Tutorial_2.Rmd
@@
 ---
-title: "Tutorial_2"
+title: "Tutorial_2a"
 author: "Sven Panis"
@@
 
-In this file we plot priors for the logit and cloglog link functions, build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials for the selected model.
+In this Tutorial 2a we plot priors for the logit and cloglog link functions, build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials for the selected model.
 

### 2024-10-16 Tutorial_3a.Rmd
rename 2 tuts

--- 2024-09-17 Tutorial_3.Rmd
@@
 ---
-title: "Tutorial_3"
+title: "Tutorial_3a"
 author: "Sven Panis"
@@
 
-```{r load-pkg}
+```{r load-pkg, results='hide'}
 pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")
@@
 ```
+
+```{r save-plot-estimates, eval=F}
+ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=8,dpi=300)
+```

### 2024-10-18 Tutorial_2a.Rmd
update tutorial 2a

@@
 
-# Load the person-trial-bin data set that we saved in Tutorial 1.
+# Load the person-trial-bin data set that we saved in Tutorial 1a.
 
@@
                                 condition*I(period_9^2)*trial_c +
-                                condition*I(period_9^2)*I(trial_c^2)  
+                                condition*I(period_9^2)*I(trial_c^2)  +
                                 condition*I(period_9^3) +

### 2024-10-18 Tutorial_2b.Rmd
add tutorials 2b and 3b

---
title: "Tutorial_2b"
author: "sven panis"
date: "2024-10-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data.

# Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

# Set options.

```{r set-options, results='hide'}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()
detectCores()
```

```{r check-info, results='hide'}
packageVersion("cmdstanr")
devtools::session_info("rstan")
```

# Load the person-trial data set that we saved in Tutorial 1a.

```{r load-data}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
print(ca_data,n=30)
summary(ca_data) # 2683 rows: 6 participants, trial, 3 conditions, 15 periods, and acc
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% filter(drt > 5)

# create factor condition, with "blank" as the reference level
ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ca_data <- ca_data %>% 
        mutate(period_9 = drt - 9,
               trial_c = (trial - 1000)/1000)

head(ca_data,n=17)
summary(ca_data)
# 6 subjects
# condition: blank (1303), congruent (668), incongruent (626)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit model  M4_ca: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.

## Prepare data for M4_ca

```{r data-M4-ca}
# remove unnecessary columns before fitting a model
M4_ca_data <- ca_data %>% select(-c(bl,tr,trial,drt)) # 2597 rows
head(M4_ca_data)
```

### Set up priors for M4_ca

```{r priors-cloglog-M4-ca}
priors_M4_ca <- c(
  set_prior("normal(0,1)", class = "b"),       # weakly informative prior for beta parameters when using logit link
  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using logit link
  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
)
```

### Fit model M4_ca

```{r fit-model-M4-ca, eval=F}
plan(multicore)

model_M4_ca <- 
   brm(data = M4_ca_data,
       family = binomial(link="logit"),
       acc | trials(1) ~ 0 + Intercept +
                          condition*period_9*trial_c + 
                          condition*period_9*I(trial_c^2) + 
                          condition*I(period_9^2)*trial_c +
                          condition*I(period_9^2)*I(trial_c^2) +
                          condition*I(period_9^3) +
                          trial_c*I(period_9^3) +
                          (1 +  condition*period_9*trial_c +
                                condition*period_9*I(trial_c^2) + 
                                condition*I(period_9^2)*trial_c +
                                condition*I(period_9^2)*I(trial_c^2) +
                                condition*I(period_9^3) +
                                trial_c*I(period_9^3) | pid),
       prior = priors_M4_ca,
       chains = 4, cores = 4, iter = 3000, warmup = 1000, 
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 14),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M4_ca")
```

Model_M4_ca took about 7 hours to run.

```{r check-model-M4-ca, eval=F}
model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
summary(model_M4_ca)
```

# Display effects

### 2024-10-18 Tutorial_3b.Rmd
add tutorials 2b and 3b

---
title: "Tutorial_3b"
author: "sven panis"
date: "2024-10-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

# Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")

lapply(pkg, library, character.only = TRUE)
```

# Load the data set for ca(t) modeling that we saved in Tutorial 1a.

```{r load-data-ca}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
print(ca_data,n=30)
summary(ca_data) # 2683 rows
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% filter(drt > 5)

# create factor condition, with "blank" as the reference level
ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ca_data <- ca_data %>% 
        mutate(period_9 = drt - 9,
               trial_c = (trial - 1000)/1000)

head(ca_data,n=17)
summary(ca_data)
# 6 subjects
# condition: blank (1303), congruent (668), incongruent (626)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

```{r get-data}
M3_ca_data <- ca_data %>% select(pid, period_9, trial_c, condition, acc)
```

```{r fit-model, eval=F}
model_M3_ca_f <- glmer(acc ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data=M3_ca_data, 
              family=binomial(link="logit"))
```

Model_M3_ca_f took about 15 minutes to fit.

```{r save-model, eval=F}
save(model_M3_ca_f, file="Tutorial_3_Frequentist/model_M3_ca_f.RData")
```

### 2024-10-18 manuscript/SupplementaryMaterial.Rmd
update ms.

@@
 
-\noindent F(t) = 1-S(t) = P(RT $\leq$ t).   \hfill  (4)
+\noindent F(t) = P(RT $\leq$ t) = 1-S(t)   \hfill  (4)
 
-The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+The discrete-time hazard function of event occurrence gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1.
+From equation 3 we find that hazard in bin t is equal to P(t)/S(t-1).
 
@@
 
-which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-This latter function is also known as the micro-level speed-accuracy tradeoff function.
+which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT > t-1) = P(RT >= t) tells you on how many percent of the trials the estimate h(t) = ..
+is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) tells you on how many percent of the trials the estimate ca(t) is based.
 
@@
 Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
-The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+The hazard rate function gives you the instantaneous *rate* of event occurrence at time point t, given that the event has not occurred yet.
 
@@
 
-* censor(df,timeout,bin_width)
-* ptb(df)
-* setup_lt(ptb)
-* setup_lt_2IV(ptb)
-* calc_ca(df)
-* calc_ca_2IV(df)
-* join_lt_ca(df1,df2)
-* join_lt_ca_2IV(df1,df2)
-* extract_median(df)
-* extract_median_2IV(df)
-* plot_eha(df,subj,haz_yaxis)
-* plot_eha_2IV(df,subj,haz_yaxis)
-* plot_eha_agg(df,subj,haz_yaxis)
+* censor(df,timeout,bin_width) : divide the time segment (0,timeout] in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
+* ptb(df) : transform the person-trial data set to the person-trial-bin data set
+* setup_lt(ptb) : set up a life table for each level of 1 independent variable
+* setup_lt_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
+* calc_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
+* calc_ca_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
+* join_lt_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
+* join_lt_ca_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
+* extract_median(df) : estimate quantiles S(t).50 (1 independent variable)
+* extract_median_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
+* plot_eha(df,subj,haz_yaxis) : create plots of the discrete-time functions (1 independent variable)
+* plot_eha_2IV(df,subj,haz_yaxis) : create plots of the discrete-time functions (2 independent variables)
+* plot_eha_agg(df,subj,haz_yaxis) : create 1 plot for aggregated data (1 independent variable)
 
@@
 
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)].
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)]
 
@@
 
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~].
+cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~]
 

### 2024-10-18 manuscript/manuscript.Rmd
update ms.

@@
 The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occured in the past, i.e., in one of the prior bins.
+In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
 This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
@@
 For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
-X in Supps), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+X in Supps), which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 
@@
 We used `r my_r_citation$r` for all reported analyses.
-Web links are printed in bold.
 The content of the tutorials is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
@@
 Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
-Tutorials 3a and 3b show how to implement multilevel models for hazard and conditional accuracy in the frequentist framework, respectively.
+Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
 Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
 
-Then a list of tutorials:
+Our list of tutorials is as follows:
 
@@
 Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the timewindow of interest.
 Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
+It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individulas,  and/or different groups with qualitatively different behavior.
 
-Table 3 shows the life table for condition "blank" (no prime stimulus presented) - compare to Figure 1.
-A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
+Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6 - compare to Figure 1.
+A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
 At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
@@
 
+When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
+
   These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
@@
 At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
-As we will show in Tutorials 2 and 3, statistical models for h(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) in each bin of a selected time range.
-As suchm multi-level regression is what we tuen to in the next tutorials.
+As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected timewindow for analysis.
+As such, multi-level regression is what we turn to in those tutorials.
+But first we consider the descriptive stats for two independent variables.
 
@@
 As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
-Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+Compared to the no-mask condition (column 1 in Figure 8), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+T...
 
@@
 
-In this second tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
@@
 
-In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
+In general, there are three assumptions one can make or relax when adding experimental predictor variables and other covariates: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1.
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We select the analysis range (200,600] and the cloglog link.

### 2024-10-19 Tutorial_2a.Rmd
add effects tut2a

@@
 
-## Visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trial 1000.
+## Visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trials 500, 1000, and 1500.
 
 ```{r post-distr}
+#get_variables(model_M4)[1:31]
+
 post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
@@
           period_9cu = period_9^3,
-          trial100 = -900/1000)
+          trial500 = -500/1000,
+          trial500sq = trial500^2,
+          trial1500 = 500/1000,
+          trial1500sq = trial1500^2)
 
-# effects for trial 1000
-plot_facet <- post %>% 
-  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
-         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
+# effects for trials 500, 1000, and 1500
+effects <- post %>% 
+  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
+            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
+            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
+            period_9sq * trial1500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
+            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
+            period_9sq * trial1500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,
 
-  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
+         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
+            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
+            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
+            period_9sq * trial1500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
+            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
+            period_9sq * trial1500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,
+
+         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
+            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
+            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
+            period_9sq * trial500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
+            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
+            period_9sq * trial500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,
+
+         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
+            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
+            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
+            period_9sq * trial500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
+            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
+            period_9sq * trial500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,
+
+         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
+
+         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
+```
+
+```{r plot-effects}
+plot_facet <- effects %>%
+  pivot_longer(cols=`congruent trial 1500`:`incongruent trial 1000`, names_to = "condition" ) %>%
   select(condition, period_9, value) %>%
- # mutate(value_hr = exp(value)) %>%
   group_by(period_9) %>%
@@
                      limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-8,-6,-4,-2,0,2,4,6,8), limits = c(-8,8)) +
+  scale_y_continuous(breaks = c(-12,-10,-8,-6,-4,-2,0,2,4,6,8,10,12), limits = c(-12,12)) +
   labs(x = "time bin", y = "cloglog-hazard") +
@@
         legend.position = "top") +
-  facet_wrap(~ condition)
+  facet_wrap(~ factor(condition,levels=c("congruent trial 500","congruent trial 1000","congruent trial 1500","incongruent trial 500","incongruent trial 1000","incongruent trial 1500")))
 
@@
 ```{r save-plot-effects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png", width = 12, height = 12, dpi = 300)
+ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
+```
+
+# calculate point and interval estimates, and hazard ratios
+
+```{r table-interval-estimates}
+int_c_500 <- effects %>%
+  select(`congruent trial 500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`congruent trial 500`),
+         mean = `congruent trial 500`,
+         condition = "c500") %>%
+  select(bin_endpoint,condition, mean, .lower, .upper, .width, `hazard ratio`)
+
+int_i_500 <- effects %>%
+  select(`incongruent trial 500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`incongruent trial 500`),
+         mean = `incongruent trial 500`,
+         condition = "i500") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width, `hazard ratio`)
+
+int_c_1000 <- effects %>%
+  select(`congruent trial 1000`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`congruent trial 1000`),
+         mean = `congruent trial 1000`,
+         condition = "c1000" ) %>%
+  select(bin_endpoint,condition,mean, .lower, .upper, .width, `hazard ratio`)
+
+int_i_1000 <- effects %>%
+  select(`incongruent trial 1000`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`incongruent trial 1000`),
+         mean = `incongruent trial 1000`,
+         condition = "i1000") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width,`hazard ratio`)
+
+int_c_1500 <- effects %>%
+  select(`congruent trial 1500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`congruent trial 1500`),
+         mean = `congruent trial 1500`,
+         condition = "c1500" ) %>%
+  select(bin_endpoint, condition, mean, .lower, .upper, .width, `hazard ratio`)
+
+int_i_1500 <- effects %>%
+  select(`incongruent trial 1500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `hazard ratio` = exp(`incongruent trial 1500`),
+         mean = `incongruent trial 1500`,
+         condition = "i1500") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width,  `hazard ratio`)
+
+test <- rbind(int_c_500, int_c_1000, int_c_1500,int_i_500,int_i_1000, int_i_1500)
+test %>% print(n=60)
+```
+
+```{r save-table-interval-estimates, eval=F}
+write_csv(test, file="Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
 ```

### 2024-10-19 Tutorial_2b.Rmd
add effects tut2b

@@
 # Display effects
+
+```{r}
+model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
+```
+
+## Pareto k estimates
+
+```{r pareto}
+loo(model_M4_ca)$diagnostics %>% 
+  data.frame() %>% 
+  # attach the `id` values
+  bind_cols(M4_ca_data) %>% 
+  mutate(id = 1:n()) %>%
+
+  ggplot(aes(x = id, y = pareto_k)) +
+  geom_point(alpha = 3/4) + 
+  geom_text(data = . %>% filter(pareto_k > .2),
+            aes(x = id + 2, label = id),
+            size = 3, hjust = 0) +
+  theme(panel.grid = element_blank())
+```
+
+## Visualize posterior distributions of the effects of congruent and incongruent primes on logit-ca(t) relative to blank prime, for each time bin in trials 500, 1000, and 1500.
+
+```{r post-distr}
+get_variables(model_M4_ca)[1:31]
+
+post <-as_draws_df(model_M4_ca) %>% # 8000 draws x 715 variables
+   select(starts_with("b_")) %>%       # 8000 x 31
+   expand_grid(period_9 = -3:6) %>%   
+   mutate(period_9sq = period_9^2,
+          period_9cu = period_9^3,
+          trial500 = -500/1000,
+          trial500sq = trial500^2,
+          trial1500 = 500/1000,
+          trial1500sq = trial1500^2)
+
+# effects for trials 500, 1000, and 1500
+effects <- post %>% 
+  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
+            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
+            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
+            period_9sq * trial1500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
+            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
+            period_9sq * trial1500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`,
+
+         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
+            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
+            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
+            period_9sq * trial1500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
+            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
+            period_9sq * trial1500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`,
+
+         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
+            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
+            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
+            period_9sq * trial500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
+            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
+            period_9sq * trial500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`,
+
+         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
+            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
+            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
+            period_9sq * trial500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
+            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
+            period_9sq * trial500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`,
+
+         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
+            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
+
+         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
+            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
+```
+
+```{r plot-effects}
+plot_facet <- effects %>%
+  pivot_longer(cols=`congruent trial 1500`:`incongruent trial 1000`, names_to = "condition" ) %>%
+  select(condition, period_9, value) %>%
+  group_by(period_9) %>%
+
+  ggplot(aes(x = period_9, y = value)) +
+  stat_lineribbon(show.legend=T) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=0, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,6)) +
+  scale_y_continuous(breaks = c(-40,-30,-20,-10,0,10,20,30,40), limits = c(-40,40)) +
+  labs(x = "time bin", y = "logit-ca") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90),
+        panel.background = element_rect(fill="white", color="black"),
+        legend.position = "top") +
+  facet_wrap(~ factor(condition,levels=c("congruent trial 500","congruent trial 1000","congruent trial 1500","incongruent trial 500","incongruent trial 1000","incongruent trial 1500")))
+
+plot_facet
+```
+
+```{r save-plot-effects, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
+```
+
+# calculate point and interval estimates, and hazard ratios
+
+```{r table-interval-estimates}
+int_c_500 <- effects %>%
+  select(`congruent trial 500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`congruent trial 500`),
+         mean = `congruent trial 500`,
+         condition = "c500") %>%
+  select(bin_endpoint,condition, mean, .lower, .upper, .width, `odds ratio`)
+
+int_i_500 <- effects %>%
+  select(`incongruent trial 500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`incongruent trial 500`),
+         mean = `incongruent trial 500`,
+         condition = "i500") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width, `odds ratio`)
+
+int_c_1000 <- effects %>%
+  select(`congruent trial 1000`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`congruent trial 1000`),
+         mean = `congruent trial 1000`,
+         condition = "c1000" ) %>%
+  select(bin_endpoint,condition,mean, .lower, .upper, .width, `odds ratio`)
+
+int_i_1000 <- effects %>%
+  select(`incongruent trial 1000`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`incongruent trial 1000`),
+         mean = `incongruent trial 1000`,
+         condition = "i1000") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width,`odds ratio`)
+
+int_c_1500 <- effects %>%
+  select(`congruent trial 1500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`congruent trial 1500`),
+         mean = `congruent trial 1500`,
+         condition = "c1500" ) %>%
+  select(bin_endpoint, condition, mean, .lower, .upper, .width, `odds ratio`)
+
+int_i_1500 <- effects %>%
+  select(`incongruent trial 1500`, period_9) %>%
+  group_by(period_9) %>%
+  mean_hdi() %>%
+  mutate(bin_endpoint = (period_9 + 9)*40,
+        `odds ratio` = exp(`incongruent trial 1500`),
+         mean = `incongruent trial 1500`,
+         condition = "i1500") %>%
+  select(bin_endpoint, condition,mean, .lower, .upper, .width,  `odds ratio`)
+
+test <- rbind(int_c_500, int_c_1000, int_c_1500,int_i_500,int_i_1000, int_i_1500)
+test %>% print(n=60)
+```
+
+```{r save-table-interval-estimates, eval=F}
+write_csv(test, file="Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
+```
+
+# plot effects on hazard ratio scale
+
+```{r haz-ratio}
+plot_facet_hr <- post %>% 
+  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
+         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
+
+  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
+  select(condition, period_9, value) %>%
+  mutate(value_hr = exp(value)) %>%
+  group_by(period_9) %>%
+
+   ggplot(aes(x = period_9, y = value_hr)) +
+  stat_lineribbon(show.legend=T) +
+  scale_fill_brewer() +
+  geom_hline(yintercept=1, linetype="dashed", color = "red") +
+  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
+                     limits = c(-3,7)) +
+  scale_y_continuous( limits = c(0,40)) +
+  labs(x = "time bin", y = "hazard ratio") +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle=90),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top") +
+  facet_wrap(~ condition)
+
+plot_facet_hr
+
+```
+
+# Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject
+
+```{r select-pid}
+selected_subject = 6 # select pid here (1:6) manually
+```
+
+```{r help-function}
+make_fitted <- function(fit, scale, ...) {
+
+  fitted(fit,
+         newdata = nd,
+         scale = scale,
+         ...) %>% 
+    data.frame() %>% 
+    bind_cols(nd)
+}
+```
+
+# Set up new data
+
+```{r new-data}
+# define the `newdata`
+nd <- tibble(pid = 1:6) %>%
+  expand_grid(incon = 0:1,
+                con  = 0:1,
+                trial_c = c(-0.5,0,0.5),
+                period_9 = -3:6) %>%
+  filter(!(con ==1 & incon == 1)) %>%
+  mutate(Iperiod_9E2 = period_9^2,
+         Iperiod_9E3 = period_9^3,
+         Itrial_cE2 = trial_c^2,
+         condition = ifelse(con==1 & incon==0,"congruent",
+                     ifelse(con==0 & incon==1,"incongruent", "blank")),
+         condition = factor(condition, levels=c("blank" ,"congruent","incongruent"))) 
+```
+
+# First, on the logit-scale:
+
+```{r subject-logit}
+make_fitted(model_M4_ca, scale = "linear") %>% 
+  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
+         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
+         period = period_9+9) %>% 
+  filter(pid==selected_subject) %>% 
+  # plot
+  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
+             fill = condition, color = condition)) +
+  geom_ribbon(alpha = 1/5, linewidth = 0) +
+  geom_line() +
+  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
+  ylab("fitted logit(ca)") +
+  coord_cartesian(ylim = c(-20, 20)) +
+  theme(panel.grid = element_blank(),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top",
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~ trial)
+```
+
+# Second, on ca(t) scale:
+
+```{r subject-hazard}
+make_fitted(model_M4_ca, scale = "response") %>% 
+  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
+         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
+         period = period_9+9) %>% 
+  filter(pid==selected_subject) %>% 
+  # plot
+  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
+             fill = condition, color = condition)) +
+  geom_ribbon(alpha = 1/5, linewidth = 0) +
+  geom_line() +
+  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
+  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
+  ylab("fitted ca(t)") +
+  coord_cartesian(ylim = c(0, 1)) +
+  theme(panel.grid = element_blank(),
+        panel.background = element_rect(fill="white",color="black"),
+        legend.position = "top",
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~ trial)
+```
+
+```{r save-plot-fit-hazard, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M4effects_subject6.png", width = 10, height = 10, dpi = 300)
+```

### 2024-10-19 manuscript/manuscript.Rmd
update ms.

@@
 
-Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial number 1000 for the selected model.
+Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500 for the selected model.
 
@@
 ```{r plot-prime-effects, fig.cap = "(ref:descr-fig5-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png")
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png")
+```
+
+Tabel X shows the mean point estimates and the upper and lower limits of the 95% highest density intervals for each bin
+
+```{r}
+int_tab <- read_csv("../Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
+```
+
+(ref:int-table-caption) Point and interval estimates.
+
+(ref:int-table-note-caption) The posterior distributions of the effects of congruent and incongruent primes relative to no primes are summarized by point estimates and the lower and upper ...
+
+```{r int-table}
+apa_table(
+  int_tab,
+  caption = "(ref:int-table-caption)",
+  note = "(ref:int-table-note-caption)",
+  placement = "H",
+  font_size = "scriptsize",
+  longtable = TRUE,
+  repeat_header = TRUE
+)
 ```

### 2024-10-19 manuscript/manuscript.tex
update ms.

@@
 For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
-X in Supps), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+X in Supps), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 
@@
 We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-Web links are printed in bold.
 The content of the tutorials is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
@@
 Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the timewindow of interest.
 Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) - compare to Figure 1.
-A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se).
+It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individulas, and/or different groups with qualitatively different behavior.
+
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6 - compare to Figure 1.
+A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
 At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
@@
 
+When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
+
 These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
@@
 As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
-Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+
+Compared to the no-mask condition (column 1 in Figure 8), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+T\ldots{}
 
@@
 
-In this second tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
@@
 
-In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
-
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1.
+In general, there are three assumptions one can make or relax when adding experimental predictor variables and other covariates: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
+
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We select the analysis range (200,600{]} and the cloglog link.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials}
 
@@
 \end{figure}
+
+Tabel X shows the mean point estimates and the upper and lower limits of the 95\% highest density intervals for each bin
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} The posterior distributions of the effects of congruent and incongruent primes relative to no primes are summarized by point estimates and the lower and upper \ldots{}}
+\end{TableNotes}
+
+\scriptsize{
+
+\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:int-table}Point and interval estimates.}\\
+\toprule
+bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:int-table} continued}}\\
+\toprule
+bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
+\midrule
+\endhead
+240.00 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82\\
+280.00 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03\\
+320.00 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45\\
+360.00 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91\\
+400.00 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.71\\
+440.00 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.64\\
+480.00 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62\\
+520.00 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.62\\
+560.00 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.60\\
+600.00 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.52\\
+240.00 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63\\
+280.00 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.13\\
+320.00 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90\\
+360.00 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99\\
+400.00 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68\\
+440.00 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58\\
+480.00 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58\\
+520.00 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.64\\
+560.00 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71\\
+600.00 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.78\\
+240.00 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02\\
+280.00 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.66\\
+320.00 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.52\\
+360.00 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68\\
+400.00 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.40\\
+440.00 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28\\
+480.00 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23\\
+520.00 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20\\
+560.00 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18\\
+600.00 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15\\
+240.00 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10\\
+280.00 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72\\
+320.00 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58\\
+360.00 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.34\\
+400.00 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.30\\
+440.00 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.34\\
+480.00 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44\\
+520.00 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.58\\
+560.00 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65\\
+600.00 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.56\\
+240.00 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68\\
+280.00 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88\\
+320.00 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78\\
+360.00 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.40\\
+400.00 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.33\\
+440.00 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.39\\
+480.00 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.56\\
+520.00 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87\\
+560.00 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27\\
+600.00 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52\\
+240.00 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07\\
+280.00 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.94\\
+320.00 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.72\\
+360.00 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.35\\
+400.00 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28\\
+440.00 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.35\\
+480.00 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55\\
+520.00 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00\\
+560.00 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.79\\
+600.00 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
 
@@
 
-\noindent F(t) = 1-S(t) = P(RT \(\leq\) t). \hfill  (4)
-
-The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
-Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
+\noindent F(t) = P(RT \(\leq\) t) = 1-S(t) \hfill  (4)
+
+The discrete-time hazard function of event occurrence gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
+The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1.
+From equation 3 we find that hazard in bin t is equal to P(t)/S(t-1).
 
@@
 
-which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-This latter function is also known as the micro-level speed-accuracy tradeoff function.
+which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \textgreater= t) tells you on how many percent of the trials the estimate h(t) = ..
+is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) tells you on how many percent of the trials the estimate ca(t) is based.
 
@@
 Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
-The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.
+The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.
 
@@
 \item
-  censor(df,timeout,bin\_width)
-\item
-  ptb(df)
-\item
-  setup\_lt(ptb)
-\item
-  setup\_lt\_2IV(ptb)
-\item
-  calc\_ca(df)
-\item
-  calc\_ca\_2IV(df)
-\item
-  join\_lt\_ca(df1,df2)
-\item
-  join\_lt\_ca\_2IV(df1,df2)
-\item
-  extract\_median(df)
-\item
-  extract\_median\_2IV(df)
-\item
-  plot\_eha(df,subj,haz\_yaxis)
-\item
-  plot\_eha\_2IV(df,subj,haz\_yaxis)
-\item
-  plot\_eha\_agg(df,subj,haz\_yaxis)
+  censor(df,timeout,bin\_width) : divide the time segment (0,timeout{]} in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
+\item
+  ptb(df) : transform the person-trial data set to the person-trial-bin data set
+\item
+  setup\_lt(ptb) : set up a life table for each level of 1 independent variable
+\item
+  setup\_lt\_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
+\item
+  calc\_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
+\item
+  calc\_ca\_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
+\item
+  join\_lt\_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
+\item
+  join\_lt\_ca\_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
+\item
+  extract\_median(df) : estimate quantiles S(t).50 (1 independent variable)
+\item
+  extract\_median\_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
+\item
+  plot\_eha(df,subj,haz\_yaxis) : create plots of the discrete-time functions (1 independent variable)
+\item
+  plot\_eha\_2IV(df,subj,haz\_yaxis) : create plots of the discrete-time functions (2 independent variables)
+\item
+  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for aggregated data (1 independent variable)
 \end{itemize}
@@
 
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}
 
@@
 
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.
+cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}
 

### 2024-10-19 Tutorial_2b.Rmd
update tut2b

@@
 ```{r save-plot-effects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
+ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
 ```
@@
 ```{r save-table-interval-estimates, eval=F}
-write_csv(test, file="Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
+write_csv(test, file="Tutorial_2_Bayesian/tables/ca_effects_intervals_table.csv")
 ```
@@
 ```{r save-plot-fit-hazard, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_subject6.png", width = 10, height = 10, dpi = 300)
+ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_subject6.png", width = 10, height = 10, dpi = 300)
 ```

### 2024-10-19 Tutorial_2b.Rmd
update tutorials 2a and 2b

@@
 
-```{r}
+```{r load-model-M4-ca}
 model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
@@
 
-# plot effects on hazard ratio scale
+# plot effects on odds ratio scale
 
-```{r haz-ratio}
+```{r odds-ratio}
 plot_facet_hr <- post %>% 
@@
 
-```{r subject-hazard}
+```{r subject-ca}
 make_fitted(model_M4_ca, scale = "response") %>% 

### 2024-10-20 Tutorial_2a.Rmd
update tutorials 2a and 2b
update stat_lineribbon in 2a

@@
 
-```{r save-plot-links}
+```{r save-plot-links, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 300)
@@
   ggplot(aes(x = period_9, y = value)) +
-  stat_lineribbon(show.legend=T) +
+  stat_lineribbon(point_interval = mean_hdi,.width = c(0.5, 0.8, 0.95),show.legend=T) +
   scale_fill_brewer() +
@@
   scale_y_continuous(breaks = c(-12,-10,-8,-6,-4,-2,0,2,4,6,8,10,12), limits = c(-12,12)) +
-  labs(x = "time bin", y = "cloglog-hazard") +
+  labs(x = "time bin endpoint", y = "cloglog-hazard") +
   theme(panel.grid = element_blank(),

### 2024-10-20 Tutorial_2b.Rmd
update stat_lineribbon 2b

@@
                      limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-40,-30,-20,-10,0,10,20,30,40), limits = c(-40,40)) +
+  scale_y_continuous(breaks = c(-50,-40,-30,-20,-10,0,10,20,30,40,50), limits = c(-50,50)) +
   labs(x = "time bin", y = "logit-ca") +

### 2024-10-20 manuscript/SupplementaryMaterial.Rmd
update suppl. material

@@
 
-which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+\noindent which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 
-The survivor function provides a context for the hazard function, as S(t-1) = P(RT > t-1) = P(RT >= t) tells you on how many percent of the trials the estimate h(t) = ..
-is based.
-The probability mass function provides a context for the conditional accuracy function, as P(t) tells you on how many percent of the trials the estimate ca(t) is based.
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT > t-1) = P(RT $\geq$ t) tells you on how many percent of the trials the estimate h(t) = P(RT = t| RT $\geq$ t) is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct | RT = t) is based.
 
@@
 
+When you want to analyse simple RT data from a detection experiment with one independent variable, the functions calc_ca() and join_lt_ca() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot_eha().
+When you want to analyse simple RT data from a detection experiment with two independent variables, the functions calc_ca_2IV() and join_lt_ca_2IV() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot_eha_2IV().
+
 ## C. Link functions
 
-(ref:descr-fig3-caption) The logit and cloglog link functions.
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 11.
 
-```{r plot-link-functions, fig.cap = "(ref:descr-fig3-caption)", out.width='80%'}
+(ref:descr-plot11-caption) The logit and cloglog link functions.
+
+```{r plot-link-functions, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
@@
 
-(ref:descr-fig4-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+(ref:descr-plot12-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
 
-```{r plot-priors, fig.cap = "(ref:descr-fig4-caption)", out.width='80%'}
+```{r plot-priors, fig.cap = "(ref:descr-plot12-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")

### 2024-10-20 manuscript/manuscript.Rmd
update ms.

--- 2024-10-18 manuscript/manuscript.Rmd
@@
 Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (insert REF).
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022].
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to our theory building toolkit.
+As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of waiting time -- to our theory building toolkit.
 
-From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict when events will occur is not a new endeavour.
+From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
 Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
@@
 
-In this paper, we focus on a distributional method known as discrete-time event history analysis, a.k.a.
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
 We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before @singerAppliedLongitudinalData2003, @allisonDiscreteTimeMethodsAnalysis1982, and @allisonSurvivalAnalysisUsing2010, and our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you should do them.
+However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010], and our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you should do them.
 
-We then provide four different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+We then provide six different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
@@
 
-In Tutorial 1, we illustrate how to process or "wrangle" a previously published RT dataset to calculate descriptive statistics when there is one independent variable.
+In Tutorial 1a, we illustrate how to process or "wrangle" a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 2, we illustrate how one can fit Bayesian multi-level regression hazard models to the data using the R package brms.
+In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
+In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 3, we illustrate how to fit the same type of regression hazard models in a frequentist framework using the R package lme4.
+In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis.
+In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to EHA.
-In Tutorial 4, we provide a generalisation of the approach to illustrate one might descriptive statistics when using a more complex design, such as when there are two independent variables.
+In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
 
 In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials [@allisonSurvivalAnalysisUsing2010] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time hazard analysis on time-to-event data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
 
@@
 Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
-As such, we only supply regression equations in supplementary materials and then refer to them in the text whenever relevant.
+As such, we only supply regression equations in the supplementary material and then refer to them in the text whenever relevant.
 
@@
 After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response could occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
 Continuous RT data is treated here as interval-censored data.
 
-Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X).
-The equations that define both of these functions are reported in supplementary materials (part A).
+Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure 2).
+The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A).
 The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
@@
 The survivor function can help to qualify or provide context to the interpretation of the hazard function.
-For example, it can give a sense of how many trials may contribute to that part of the distribution.
-If each participant completes 100 trials in an experiment, and the survivor function prob of 0.03, then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in this part of the distribution would be wider and less precise compared to other parts.
+For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
+If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500], then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
+Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.
 
-(ref:descr-fig2-1-caption) Hazard and survivor functions
+(ref:descr-plot2-caption) Hazard and survivor functions
 
-```{r plot2-1, fig.cap = "(ref:descr-fig2-1-caption)", out.width="80%"}
+```{r plot2, fig.cap = "(ref:descr-plot2-caption)", out.width="80%"}
 knitr::include_graphics("../sims/figures/haz_surv.jpeg")
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
-X in Supps), which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+For such two-choice RT data, the discrete-time hazard analysis can be extended with a discrete-time SAT analyis.
+Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 
@@
 The prime can then be congruent or incongruent with the target.
-Figure 3 shows that the early upswing in hazard is equal for both conditions, and that early responses are always correct in ..
-and always incorrect in the incongruent condtion.
-Taken together, the results show that for early responses (< bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later bins, response hazard is lower in incongruent compared to congruent trials, as .....the prime can be overridden, as both conditions are now always correct.
+Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+These results show that for early responses (< bin 6), responses always follow the prime (and not the target, as instructed).
+And then for later bins, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 
-(ref:descr-fig2-2-caption) Hazard and conditional accuracy
+(ref:descr-plot3-caption) Hazard and conditional accuracy
 
-```{r plot2-2, fig.cap = "(ref:descr-fig2-2-caption)", out.width="80%"}
+```{r plot3, fig.cap = "(ref:descr-plot3-caption)", out.width="80%"}
 knitr::include_graphics("../sims/figures/haz_acc.jpeg")
@@
 In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
-Accordingly, experiment designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-*N* designs).
+Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-*N* designs).
 This approach  emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
 
@@
 1b. Wrangle raw data and descriptive stats for two independent variables.
-
 2a. Bayesian multilevel modeling for h(t)
 2b. Bayesian multilevel modeling for ca(t)
-
 3a. Frequentist multilevel modeling for h(t)
@@
 
-Generalisation (T4).
-Should this be online in Supps??
-It would make the main text shorter and simpler, but make it no less available.
-We could just have a sentence at the end of T1, which says that we provide a generalisation and extension in T4, which is in Supps.
-
-Plannng (T5) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
+Plannng (T4) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
 
@@
 In contrast, the 'person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the timewindow of interest.
-Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.
+Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
+Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t).
-To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our function "censor()", as follows.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material.
+To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
 
@@
 
-The next step is to plot the data using our custom made plotting tool plot_eha(). When creating the plots, some warning messages will likely be generated, like these:
+The next step is to plot the data using our custom function plot_eha(). When creating the plots, some warning messages will likely be generated, like these:
 
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individulas,  and/or different groups with qualitatively different behavior.
+It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
-Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6 - compare to Figure 1.
+Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6 -- compare to Figure 1.
 A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
@@
 
-Figure 4 displays the discrete-time hazard, survivor, and conditional accuracy functions for each prime condition for participant 6.
-By using discrete-time h(t) functions of event occurrence - in combination with ca(t) functions for two-choice tasks - one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
+By using discrete-time h(t) functions of event occurrence -- in combination with ca(t) functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
 
@@
 
-(ref:descr-fig2-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
+(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
 
-```{r eha-plot, fig.cap = "(ref:descr-fig2-caption)"}
+```{r eha-plot, fig.cap = "(ref:descr-plot4-caption)"}
 knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt.png")
@@
 Qualitatively similar results were obtained for the other five participants.
+
   These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
 At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected timewindow for analysis.
+As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected time window for analysis.
 As such, multi-level regression is what we turn to in those tutorials.
-But first we consider the descriptive stats for two independent variables.
+But first we consider calculating the descriptive statistics for two independent variables.
 
@@
 
-So far in this paper, we have use a simple experimental design, which involved one condition with two levels.
-But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than two levels.
+So far in this paper, we have use a simple experimental design, which involved one condition with three levels.
+But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than three levels.
 The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 5.
 
-Compared to the no-mask condition (column 1 in Figure 8), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
-T...
+Compared to the no-mask condition (column 1 in Figure 5), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
 
-(ref:descr-fig8-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
+(ref:descr-plot5-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
 
-```{r plot-2ivs, fig.cap = "(ref:descr-fig8-caption)", out.width='95%'}
+```{r plot-2ivs, fig.cap = "(ref:descr-plot5-caption)", out.width='95%'}
 knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt.png")
@@
 
-## 4.3 Tutorial 2a: Fitting Bayesian hazard models
+## 4.3 Tutorial 2a: Fitting Bayesian hazard models to time-to-event data
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see Supps).
+Second, given that the dependent variable is binary, one has to select a link function (see part C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-We provide relevant example regression formulas in supplementary materials.
+We provide relevant example regression formulas in part D of the supplementary material.
 
@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1a.
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We select the analysis range (200,600] and the cloglog link.
@@
 
-To get the posterior distribution of each parameters given the data, we need to specify a prior distribution for each parameter.
-The middle column of Supplementary Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
+To get the posterior distribution of each model parameter given the data, we need to specify a prior distribution for each parameter.
+The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
 
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The other rows in Figure 9 show prior distributions on the logit and cloglog scale that we use instead.
 
@@
 
-[[Here let's give some intuition on why we would want to setup the model like this]]
-
-For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2, this is accomplished as follows.
-
-\scriptsize
-```{r data-priors-M1, echo=T}
-# remove unnecessary columns before fitting a model
-M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9))
-
-# Specify priors
-priors_M1 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"),
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
-  set_prior("normal(0, 1)", class = "sd"),        
-  set_prior("lkj(2)", class = "cor")            
-)
-```
-\normalsize
-
-We can now estimate our first Bayesion regression model, as follows.
+When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, you can use a general specification of TIME, i.e., one intercept per time bin.
+In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M1 is specified as follows.
 
@@
                            condition + trial_c +
-
                            (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
@@
 
+After selecting the binomial family and the cloglog link, the model formula is specified.
+The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of priming condition and centered trial number.
+Each of these effects is allowed to vary across individuals.
 Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
@@
 
-[[Here let's give some intuition on why we would want to modify the formula and model features]]
+When the shape of the hazard function is rather smooth, one can fit a more parsimonious model by using a polynomial specification of TIME.
+For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
+The model formula for model M2 looks as follows.
 
-For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-We first remove unncessary columns and specify our priors.
+\scriptsize 
+```{r formula-model-M2, eval=F, echo=T}
+event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) +
+                           condition + trial_c +
+                           (1 + period_9 + I(period_9^2) + I(period_9^3) +
+                           condition + trial_c | pid),
+```
+\normalsize
 
-<!-- \scriptsize -->
-<!-- ```{r data-priors-M2, echo=T} -->
-<!-- # remove unnecessary columns -->
-<!-- M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15))  -->
-
-<!-- # Specify priors -->
-<!-- priors_M2 <- c( -->
-<!--   set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       -->
-<!--   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),     -->
-<!--   set_prior("normal(0, 1)", class = "sd"),             -->
-<!--   set_prior("lkj(2)", class = "cor")                  -->
-<!-- ) -->
-<!-- ``` -->
-<!-- \normalsize -->
-
-<!-- Now we can fit model 2. -->
-
-<!-- \scriptsize -->
-<!-- ```{r fit-model-M2, eval=F, echo=T} -->
-<!-- plan(multicore) -->
-
-<!-- model_M2 <- -->
-<!--    brm(data = M2_data, -->
-<!--        family = binomial(link="cloglog"), -->
-<!--        event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) +  -->
-<!--                           condition + trial_c + -->
-<!--                           (1 + period_9 + I(period_9^2) + I(period_9^3) +  -->
-<!--                           condition + trial_c | pid), -->
-<!--        prior = priors_M2, -->
-<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
-<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
-<!--        seed = 12, init = "0", -->
-<!--        file = "Tutorial_2_Bayesian/models/model_M2") -->
-<!-- ``` -->
-<!-- \normalsize -->
-
+Because TIME is centered on bin 9, and trial number on trial 1000, the Intercept represents the cloglog-hazard in bin 9 for the blank prime condition in trial 1000.
 Estimating model M2 took about 144 minutes.
@@
 
-[[Here let's give some intuition on why we would want to modify the formula and model features]]
+So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
+However, the descriptive plots suggest that the effect of condition can vary across time bins.
+For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable "condition") and centered trial number (variable "trial_c").
 
-For the third model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables congruency (variable "condition") and trial number (variable "trial_c").
-We use the same data set and priors as for model 2.
+\scriptsize
+```{r formula-model-M3, eval=F, echo=T}
+        event | trials(1) ~ 0 + Intercept + 
+                            condition*period_9 +  
+                            condition*I(period_9^2) +  
+                            condition*I(period_9^3) + 
+                            trial_c*period_9 +  
+                            trial_c*I(period_9^2) +  
+                            trial_c*I(period_9^3) + 
+                            (1 + condition*period_9 + 
+                            condition*I(period_9^2) + 
+                            condition*I(period_9^3) + 
+                            trial_c*period_9 +  
+                            trial_c*I(period_9^2) +  
+                            trial_c*I(period_9^3) | pid), 
+```
+\normalsize
 
-<!-- \scriptsize -->
-<!-- ```{r fit-model-M3, eval=F, echo=T} -->
-<!-- M3_data <- M2_data -->
-<!-- priors_M3 <- priors_M2 -->
-<!-- plan(multicore) -->
-
-<!-- model_M3 <-  -->
-<!--    brm(data = M3_data, -->
-<!--        family = binomial(link="cloglog"), -->
-<!--        event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored -->
-<!--                            condition*period_9 +  -->
-<!--                            condition*I(period_9^2) +  -->
-<!--                            condition*I(period_9^3) + -->
-<!--                            trial_c*period_9 +  -->
-<!--                            trial_c*I(period_9^2) +  -->
-<!--                            trial_c*I(period_9^3) + -->
-<!--                            (1 + condition*period_9 + -->
-<!--                            condition*I(period_9^2) + -->
-<!--                            condition*I(period_9^3) + -->
-<!--                            trial_c*period_9 +  -->
-<!--                            trial_c*I(period_9^2) +  -->
-<!--                            trial_c*I(period_9^3) | pid), -->
-<!--        prior = priors_M3, -->
-<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
-<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
-<!--        seed = 12, init = "0", -->
-<!--        file = "Tutorial_2_Bayesian/models/model_M3") -->
-<!-- ``` -->
-<!-- \normalsize -->
-
-Estimating model M3 took about 268 minutes.
+Note that duplicate terms in the model formula are ignored. Estimating model M3 took about 268 minutes.
 
@@
 Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022], we relax all three assumptions in model 4.
-We use the same data set and priors as for model 2.
+We thus add a squared term for the continuous predictor centered trial number -- I(trial_c^2) -- and include interaction terms.
 
-<!-- \scriptsize  -->
-<!-- ```{r fit-model-M4, eval=F, echo=T} -->
-<!-- M4_data <- M2_data -->
-<!-- priors_M4 <- priors_M2 -->
-<!-- plan(multicore) -->
+\scriptsize  
+```{r fit-model-M4, eval=F, echo=T} 
+event | trials(1) ~ 0 + Intercept + 
+                           condition*period_9*trial_c +  
+                           condition*period_9*I(trial_c^2) +
+                           condition*I(period_9^2)*trial_c + 
+                           condition*I(period_9^2)*I(trial_c^2) + 
+                           condition*I(period_9^3) + 
+                           trial_c*I(period_9^3) + 
+                           (1 +  condition*period_9*trial_c + 
+                                 condition*period_9*I(trial_c^2) +  
+                                 condition*I(period_9^2)*trial_c + 
+                                 condition*I(period_9^2)*I(trial_c^2)   
+                                 condition*I(period_9^3) + 
+                                 trial_c*I(period_9^3) | pid)
+```
+\normalsize
 
-<!-- model_M4 <-  -->
-<!--    brm(data = M4_data, -->
-<!--        family = binomial(link="cloglog"), -->
-<!--        event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored -->
-<!--                           condition*period_9*trial_c +  -->
-<!--                           condition*period_9*I(trial_c^2) +  -->
-<!--                           condition*I(period_9^2)*trial_c + -->
-<!--                           condition*I(period_9^2)*I(trial_c^2) + -->
-<!--                           condition*I(period_9^3) + -->
-<!--                           trial_c*I(period_9^3) + -->
-<!--                           (1 +  condition*period_9*trial_c + -->
-<!--                                 condition*period_9*I(trial_c^2) +  -->
-<!--                                 condition*I(period_9^2)*trial_c + -->
-<!--                                 condition*I(period_9^2)*I(trial_c^2)   -->
-<!--                                 condition*I(period_9^3) + -->
-<!--                                 trial_c*I(period_9^3) | pid), -->
-<!--        prior = priors_M4, -->
-<!--        chains = 4, cores = 4, iter = 3000, warmup = 1000, -->
-<!--        control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12), -->
-<!--        seed = 12, init = "0", -->
-<!--        file = "Tutorial_2_Bayesian/models/model_M4") -->
-<!-- ``` -->
-<!-- \normalsize -->
-
-Estimating model M4 took about 8 hours.
+Again, duplicate terms in the model formula are ignored. Estimating model M4 took about 8 hours.
 
@@
 
-We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
+We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
 
-<!-- \scriptsize  -->
-<!-- ```{r compare-models, echo=T, eval=FALSE} -->
-<!-- model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds") -->
-<!-- model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds") -->
-<!-- model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds") -->
-<!-- model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds") -->
+```{r compare-models, echo=F, eval=T} 
+model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds") 
+model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds") 
+model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds") 
+model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds") 
+```
 
-<!-- #model_M1  <- add_criterion(model_M1, c("loo", "waic")) -->
-<!-- #model_M2  <- add_criterion(model_M2, c("loo", "waic")) -->
-<!-- #model_M3  <- add_criterion(model_M3, c("loo", "waic")) -->
-<!-- #model_M4  <- add_criterion(model_M4, c("loo", "waic")) -->
+\scriptsize
+```{r show-weights, echo=T, eval=T}
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
+model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
+```
+\normalsize
 
-<!-- loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F) -->
-<!-- loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F) -->
-
-<!-- # model weights -->
-<!-- model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3) -->
-
-<!-- model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3) -->
-<!-- ``` -->
-<!-- \normalsize -->
-
-Clearly, both weighting schemes prefer model M4.
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M4, and a weight of 0 to the other three simpler models.
 
@@
 
-Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial number 1000 for the selected model.
+To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
-(ref:descr-fig5-caption) 50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial number 1000.
+Figure 6 shows one point (mean) and three highest density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
 
-```{r plot-prime-effects, fig.cap = "(ref:descr-fig5-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png")
+(ref:descr-plot6-caption) Means and 50/80/95% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
+
+```{r plot-prime-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png")
 ```
 
-Figure 6 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest density interval.
+Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
 
-(ref:descr-fig6-caption) Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.
+```{r}
+int_tab <- read_csv("../Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
+```
 
-```{r plot-hazard-subject6, fig.cap = "(ref:descr-fig6-caption)", out.width='80%'}
+(ref:int-table-caption) Point and interval estimates, and hazard ratios.
+
+(ref:int-table-note-caption) c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.
+
+```{r int-table}
+apa_table(
+  int_tab,
+  caption = "(ref:int-table-caption)",
+  note = "(ref:int-table-note-caption)",
+  placement = "H",
+  font_size = "scriptsize",
+  longtable = TRUE,
+  repeat_header = TRUE
+)
+```
+
+Interpretation...
+
+To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+
+(ref:descr-plot7-caption) Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.
+
+```{r plot-hazard-subject6, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_subject6.png")
@@
 
-[[let's have a paragraph on how we might interpret these plots.]]
+## 4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
 
-## 4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
+In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).
+
+To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+
+Figure 8 shows one point (mean) and three highest density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+
+(ref:descr-plot8-caption) Means and 50/80/95% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
+
+```{r plot-prime-ca-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials.png")
+```
+
+Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95% highest density interval.
+Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
+
+```{r}
+int_ca_tab <- read_csv("../Tutorial_2_Bayesian/tables/ca_effects_intervals_table.csv")
+```
+
+(ref:int-ca-table-caption) Point and interval estimates, and odds ratios.
+
+(ref:int-ca-table-note-caption) c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.
+
+```{r int-ca-table}
+apa_table(
+  int_ca_tab,
+  caption = "(ref:int-ca-table-caption)",
+  note = "(ref:int-ca-table-note-caption)",
+  placement = "H",
+  font_size = "scriptsize",
+  longtable = TRUE,
+  repeat_header = TRUE
+)
+```
+
+Interpretation...
+
+plot 9 ca(t)
 
@@
 
-In this third tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in the first tutorial.
-For illustration purposes, we only fitted model M3 using the function glmer() from the package lme4.
+In this fifth tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Alternatively, one could also use the function glmmPQL() from the R package MASS.
+The resulting hazard model is called M3_f.
 
-<!-- \scriptsize  -->
-<!-- ```{r fit-model, eval=F, echo=T} -->
-<!-- model_M3_f <- glmer(event ~ 1 + condition*period_9 +  -->
-<!--                            condition*I(period_9^2) +  -->
-<!--                            condition*I(period_9^3) + -->
-<!--                            trial_c*period_9 +  -->
-<!--                            trial_c*I(period_9^2) +  -->
-<!--                            trial_c*I(period_9^3) + -->
-<!--                            (1 + condition*period_9 + -->
-<!--                            condition*I(period_9^2) + -->
-<!--                            condition*I(period_9^3) + -->
-<!--                            trial_c*period_9 +  -->
-<!--                            trial_c*I(period_9^2) +  -->
-<!--                            trial_c*I(period_9^3) | pid), -->
+In Figure 10 we compare the parameter estimates of model M3 from brm() with those of model M3_f from glmer().
 
-<!--               # control parameters, data set, and complementary log-log link function -->
-<!--               control = glmerControl(optimizer = c("nlminbwrap"), -->
-<!--                                      optCtrl = list(maxfun=10000000)),  -->
-<!--               data=M3_data,  -->
-<!--               family=binomial(link="cloglog")) -->
-<!-- ``` -->
-<!-- \normalsize -->
+(ref:descr-plot10-caption) Parameter estimates for model M3 from brm() and model M3_f from glmer().
 
-In Figure 7 we compare the parameter estimates of model M3 from brm() with those of glmer().
-
-(ref:descr-fig7-caption) Parameter estimates for model M3 from brm() and glmer().
-
-```{r plot-comparison, fig.cap = "(ref:descr-fig7-caption)", out.width='80%'}
+```{r plot-comparison, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png")
@@
 
-Figure 7 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
 However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
@@
 But the price you pay for being able to fit more complex models in a Bayesian framework is computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorial 2 took several hours to build.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a and 2b took several hours to build.
 
 ## 4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models
+
+In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Again, the resulting conditional accuracy model M3_ca_f did not converge and resulted in a singular fit.
 
@@
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss issues relating to individual differences, limitations of the approach, and future extensions.
+In the following, we discuss issues relating to model complexity versus interpretability, individual differences, limitations of the approach, and future extensions.
 
@@
 
-## 5.2 Individual differences
+## 5.2 Model complexity versus interpretability
+
+Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to many higher-order interactions. 
+For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period_9), and the passage of time on the across-trial or within-experiment time scale (variable trial_c).
+When trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
+From a theoretical perspective, adding more than 1 time scale is important to capture plasticity (e.g., proactive control) and other learning effects that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
+From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental variables, because adding time scales quickly increases model complexity.
+
+## 5.3 Individual differences
 
@@
 
-## 5.3 Limitation(s)
+Another approach: fit models to individual subjects and describe prevalence...
+
+## 5.4 Limitation(s)
 
@@
 
-## 5.4 Extensions
+## 5.5 Extensions
 

### 2024-10-20 manuscript/manuscript.tex
update ms.

@@
 Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (insert REF).
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009).
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension to our theory building toolkit.
-
-From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict when events will occur is not a new endeavour.
+As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of waiting time -- to our theory building toolkit.
+
+From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
 Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
@@
 
-In this paper, we focus on a distributional method known as discrete-time event history analysis, a.k.a.
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time event history analysis, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
 We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before Singer and Willett (2003), Allison (1982), and Allison (2010), and our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you should do them.
-
-We then provide four different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003), and our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you should do them.
+
+We then provide six different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
@@
 
-In Tutorial 1, we illustrate how to process or ``wrangle'' a previously published RT dataset to calculate descriptive statistics when there is one independent variable.
+In Tutorial 1a, we illustrate how to process or ``wrangle'' a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 2, we illustrate how one can fit Bayesian multi-level regression hazard models to the data using the R package brms.
+In Tutorial 1b we provide a generalisation of this approach to illustrate how one might descriptive statistics when using a more complex design, such as when there are two independent variables.
+In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT-hazard data using the R package brms.
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 3, we illustrate how to fit the same type of regression hazard models in a frequentist framework using the R package lme4.
+In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis.
+In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to EHA.
-In Tutorial 4, we provide a generalisation of the approach to illustrate one might descriptive statistics when using a more complex design, such as when there are two independent variables.
+In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
 
 In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform hazard analysis.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time hazard analysis on time-to-event data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
 
@@
 The survivor function can help to qualify or provide context to the interpretation of the hazard function.
-For example, it can give a sense of how many trials may contribute to that part of the distribution.
-If each participant completes 100 trials in an experiment, and the survivor function prob of 0.03, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in this part of the distribution would be wider and less precise compared to other parts.
+For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
+If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500{]}, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
+Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.
 
@@
 
-\caption{Hazard and survivor functions}\label{fig:plot2-1}
+\caption{Hazard and survivor functions}\label{fig:plot2}
 \end{figure}
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ.
-X in Supps), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 
@@
 The prime can then be congruent or incongruent with the target.
-Figure 3 shows that the early upswing in hazard is equal for both conditions, and that early responses are always correct in ..
-and always incorrect in the incongruent condtion.
-Taken together, the results show that for early responses (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later bins, response hazard is lower in incongruent compared to congruent trials, as \ldots..the prime can be overridden, as both conditions are now always correct.
+Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+These results show that for early responses (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
+And then for later bins, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 
-\caption{Hazard and conditional accuracy}\label{fig:plot2-2}
+\caption{Hazard and conditional accuracy}\label{fig:plot3}
 \end{figure}
@@
 In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
-Accordingly, experiment designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
+Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
 This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions.
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
 
@@
 In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the timewindow of interest.
-Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.
+Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
+Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t) and ca(t).
-To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our function ``censor()'', as follows.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material.
+To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
 
@@
 
-The next step is to plot the data using our custom made plotting tool plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:
+The next step is to plot the data using our custom function plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:
 
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individulas, and/or different groups with qualitatively different behavior.
-
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6 - compare to Figure 1.
+It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6 -- compare to Figure 1.
 A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
@@
 Qualitatively similar results were obtained for the other five participants.
+
 These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
 At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected timewindow for analysis.
+As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected time window for analysis.
 As such, multi-level regression is what we turn to in those tutorials.
-But first we consider the descriptive stats for two independent variables.
+But first we consider calculating the descriptive statistics for two independent variables.
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.
-
-Compared to the no-mask condition (column 1 in Figure 8), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
-T\ldots{}
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 5.
+
+Compared to the no-mask condition (column 1 in Figure 5), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
 
@@
 
-\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models}\label{tutorial-2a-fitting-bayesian-hazard-models}
+\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models to time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-time-to-event-data}
 
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see Supps).
+Second, given that the dependent variable is binary, one has to select a link function (see part C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 
-To get the posterior distribution of each parameters given the data, we need to specify a prior distribution for each parameter.
-The middle column of Supplementary Figure 4 shows seven examples of prior distributions on the logit and/or cloglog scales.
-
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 3 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability (i.e., discrete-time hazard) scale, as most mass is pushed to probabilities of 0 and 1.
+To get the posterior distribution of each model parameter given the data, we need to specify a prior distribution for each parameter.
+The middle column of Figure 9 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
+
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 9 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The other rows in Figure 9 show prior distributions on the logit and cloglog scale that we use instead.
 
@@
 
-{[}{[}Here let's give some intuition on why we would want to setup the model like this{]}{]}
-
-For the first model, we use a general specification of TIME (i.e., one intercept per time bin) for the baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2, this is accomplished as follows.
-
-\scriptsize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# remove unnecessary columns before fitting a model}
-\NormalTok{M1\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(bl,tr,trial,period, period\_9,d9))  }
-
-\CommentTok{\# Specify priors}
-\NormalTok{priors\_M1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}0.2,0.71,{-}2.2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),       }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d6"}\NormalTok{),}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d7"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d8"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d10"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d11"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d12"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d13"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d14"}\NormalTok{), }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"d15"}\NormalTok{),}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"skew\_normal({-}1,1,{-}2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef =} \StringTok{"Intercept"}\NormalTok{),}
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{),        }
-  \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"lkj(2)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"cor"}\NormalTok{)            }
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-We can now estimate our first Bayesion regression model, as follows.
+When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, you can use a general specification of TIME, i.e., one intercept per time bin.
+In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
+Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M1 is specified as follows.
 
@@
 \NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
-
 \NormalTok{                           (d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
@@
 
-{[}{[}Here let's give some intuition on why we would want to modify the formula and model features{]}{]}
-
-For the second model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-We first remove unncessary columns and specify our priors.
-
+When the shape of the hazard function is rather smooth, one can fit a more parsemoneous model by using a polynomial specification of TIME.
+For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
+The model formula for model M2 looks as follows.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
+\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Because TIME is centered on bin 9, and trial number on trial 1000, the Intercept represents the cloglog-hazard in bin 9 for the blank prime condition in trial 1000.
 Estimating model M2 took about 144 minutes.
@@
 
-{[}{[}Here let's give some intuition on why we would want to modify the formula and model features{]}{]}
-
-For the third model, we use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables congruency (variable ``condition'') and trial number (variable ``trial\_c'').
-We use the same data set and priors as for model 2.
-
-Estimating model M3 took about 268 minutes.
+So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
+However, the descriptive plots suggest that the effect of condition can vary across time bins.
+For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and trial number (variable ``trial\_c'').
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{        event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
+\NormalTok{                            condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
+\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
+\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                            (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
+\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
+\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid), }
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Note that duplicate terms in the model formula are ignored. Estimating model M3 took about 268 minutes.
 
@@
 
-Based on previous work (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
-We use the same data set and priors as for model 2.
-
-Estimating model M4 took about 8 hours.
+Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
+We thus add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}  
+\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{  condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
+\NormalTok{                                 condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
+\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)   }
+\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
+\NormalTok{                                 trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid)}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+Again, duplicate terms in the model formula are ignored. Estimating model M4 took about 8 hours.
 
@@
 
-We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights (Kurz, 2023a; McElreath, 2018).
-
-Clearly, both weighting schemes prefer model M4.
+We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2018).
+
+\scriptsize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## model_M1 model_M2 model_M3 model_M4 
+##        0        0        0        1
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## model_M1 model_M2 model_M3 model_M4 
+##        0        0        0        1
+\end{verbatim}
+
+\normalsize
+
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M4, and a weight of 0 to the other three simpler models.
 
@@
 
-Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500 for the selected model.
+To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+
+Figure 6 shows the point and effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-Tabel X shows the mean point estimates and the upper and lower limits of the 95\% highest density intervals for each bin
+Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95\% highest density interval.
+Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
 
@@
 \begin{TableNotes}[para]
-\normalsize{\textit{Note.} The posterior distributions of the effects of congruent and incongruent primes relative to no primes are summarized by point estimates and the lower and upper \ldots{}}
+\normalsize{\textit{Note.} c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.}
 \end{TableNotes}
@@
 \begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:int-table}Point and interval estimates.}\\
+\caption{\label{tab:int-table}Point and interval estimates, and hazard ratios.}\\
 \toprule
@@
 
-Figure 6 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+Interpretation\ldots{}
+
+To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
 
@@
 
-{[}{[}let's have a paragraph on how we might interpret these plots.{]}{]}
-
 \subsection{4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}
 
+In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).
+
+To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+
+Figure 8 shows the point and effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials}
+
+}
+
+\caption{50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
+\end{figure}
+
+Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95\% highest density interval.
+Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.}
+\end{TableNotes}
+
+\scriptsize{
+
+\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:int-ca-table}Point and interval estimates, and odds ratios.}\\
+\toprule
+bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{odds ratio}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:int-ca-table} continued}}\\
+\toprule
+bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{odds ratio}\\
+\midrule
+\endhead
+240.00 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24,618,458.61\\
+280.00 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4,846.09\\
+320.00 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.79\\
+360.00 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.89\\
+400.00 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.19\\
+440.00 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.31\\
+480.00 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.41\\
+520.00 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.40\\
+560.00 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.12\\
+600.00 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.00\\
+240.00 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1,899,836.02\\
+280.00 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1,961.83\\
+320.00 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.87\\
+360.00 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.57\\
+400.00 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.14\\
+440.00 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.06\\
+480.00 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.10\\
+520.00 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.00\\
+560.00 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.00\\
+600.00 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.00\\
+240.00 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10,001,085.39\\
+280.00 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5,124.44\\
+320.00 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.12\\
+360.00 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.82\\
+400.00 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.58\\
+440.00 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.82\\
+480.00 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.08\\
+520.00 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.00\\
+560.00 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.00\\
+600.00 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.00\\
+240.00 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.00\\
+280.00 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.00\\
+320.00 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.02\\
+360.00 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.38\\
+400.00 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.59\\
+440.00 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.59\\
+480.00 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.30\\
+520.00 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.04\\
+560.00 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3,282,435.93\\
+600.00 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46,319,712,352,328.76\\
+240.00 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.00\\
+280.00 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.00\\
+320.00 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.04\\
+360.00 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.59\\
+400.00 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.10\\
+440.00 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.68\\
+480.00 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.20\\
+520.00 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3,307.44\\
+560.00 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164,758,701.84\\
+600.00 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10,165,856,639,901,592.00\\
+240.00 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.00\\
+280.00 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.00\\
+320.00 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.06\\
+360.00 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.87\\
+400.00 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.57\\
+440.00 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.21\\
+480.00 & i1500 & 5.52 & -13.87 & 26.57 & 0.95 & 249.51\\
+520.00 & i1500 & 12.67 & -31.46 & 58.41 & 0.95 & 318,500.40\\
+560.00 & i1500 & 25.50 & -53.08 & 115.21 & 0.95 & 119,299,568,240.94\\
+600.00 & i1500 & 45.85 & -86.60 & 200.06 & 0.95 & 81,670,189,671,651,033,088.00\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
+
+Interpretation\ldots{}
+
+plot 9 ca(t)
+
 \subsection{4.5 Tutorial 3a: Fitting Frequentist hazard models}\label{tutorial-3a-fitting-frequentist-hazard-models}
 
-In this third tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in the first tutorial.
-For illustration purposes, we only fitted model M3 using the function glmer() from the package lme4.
-
-In Figure 7 we compare the parameter estimates of model M3 from brm() with those of glmer().
+In this fifth tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Alternatively, one could also use the function glmmPQL() from the R package MASS.
+The resulting hazard model is called M3\_f.
+
+In Figure 10 we compare the parameter estimates of model M3 from brm() with those of model M3\_f from glmer().
 
@@
 
-\caption{Parameter estimates for model M3 from brm() and glmer().}\label{fig:plot-comparison}
+\caption{Parameter estimates for model M3 from brm() and model M3\_f from glmer().}\label{fig:plot-comparison}
 \end{figure}
 
-Figure 7 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
 However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
@@
 But the price you pay for being able to fit more complex models in a Bayesian framework is computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorial 2 took several hours to build.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a and 2b took several hours to build.
 
 \subsection{4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models}\label{tutorial-3b-fitting-frequentist-conditional-accuracy-models}
+
+In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
+For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Again, the resulting conditional accuracy model M3\_ca\_f did not converge and resulted in a singular fit.
 
@@
 
-which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+\noindent which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 
-The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \textgreater= t) tells you on how many percent of the trials the estimate h(t) = ..
-is based.
-The probability mass function provides a context for the conditional accuracy function, as P(t) tells you on how many percent of the trials the estimate ca(t) is based.
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \(\geq\) t) tells you on how many percent of the trials the estimate h(t) = P(RT = t\textbar{} RT \(\geq\) t) is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct \textbar{} RT = t) is based.
 
@@
 
+When you want to analyse simple RT data from a detection experiment with one independent variable, the functions calc\_ca() and join\_lt\_ca() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot\_eha().
+When you want to analyse simple RT data from a detection experiment with two independent variables, the functions calc\_ca\_2IV() and join\_lt\_ca\_2IV() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot\_eha\_2IV().
+
 \subsection{C. Link functions}\label{c.-link-functions}
+
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 11.
 

### 2024-10-20 manuscript/manuscript.Rmd
update ms.

@@
 
-1a. Wrangle raw data and descriptive stats for one independent variable.
-1b. Wrangle raw data and descriptive stats for two independent variables.
-2a. Bayesian multilevel modeling for h(t)
-2b. Bayesian multilevel modeling for ca(t)
-3a. Frequentist multilevel modeling for h(t)
-3b. Frequentist multilevel modeling for ca(t)
+* 1a. Wrangle raw data and descriptive stats for one independent variable.
+* 1b. Wrangle raw data and descriptive stats for two independent variables.
+* 2a. Bayesian multilevel modeling for h(t)
+* 2b. Bayesian multilevel modeling for ca(t)
+* 3a. Frequentist multilevel modeling for h(t)
+* 3b. Frequentist multilevel modeling for ca(t)
 
@@
 For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period_9), and the passage of time on the across-trial or within-experiment time scale (variable trial_c).
-When trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
 From a theoretical perspective, adding more than 1 time scale is important to capture plasticity (e.g., proactive control) and other learning effects that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
@@
 
-Another approach: fit models to individual subjects and describe prevalence...
+Another approach: fit models to individual subjects and describe prevalence...REF
 
@@
 We did not discuss continuous-time hazard analysis.
-As indicated by --allison -- learning discrete-time methods first will help in learning continuous-time methods.
+As indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time methods first will help in learning continuous-time methods.
 Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
@@
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (REF).
-[[give a concrete example for this latter point]]
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@@allisonSurvivalAnalysisUsing2010].
+For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
 

### 2024-10-20 manuscript/manuscript.tex
update ms.

@@
 The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 1b we provide a generalisation of this approach to illustrate how one might descriptive statistics when using a more complex design, such as when there are two independent variables.
-In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT-hazard data using the R package brms.
+In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
+In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+For such two-choice RT data, the discrete-time hazard analysis can be extended with a discrete-time SAT analyis.
+Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to autmate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
 
@@
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial.
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1a.
+In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We select the analysis range (200,600{]} and the cloglog link.
@@
 To get the posterior distribution of each model parameter given the data, we need to specify a prior distribution for each parameter.
-The middle column of Figure 9 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
-
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 9 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
+
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 The other rows in Figure 9 show prior distributions on the logit and cloglog scale that we use instead.
@@
 
-When the shape of the hazard function is rather smooth, one can fit a more parsemoneous model by using a polynomial specification of TIME.
+When the shape of the hazard function is rather smooth, one can fit a more parsimonious model by using a polynomial specification of TIME.
 For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
@@
 However, the descriptive plots suggest that the effect of condition can vary across time bins.
-For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and trial number (variable ``trial\_c'').
+For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and centered trial number (variable ``trial\_c'').
 
@@
 
-Figure 6 shows the point and effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 6 shows one point (mean) and three highest density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-\caption{50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial number 1000.}\label{fig:plot-prime-effects}
+\caption{Means and 50/80/95\% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-effects}
 \end{figure}
 
-Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95\% highest density interval.
+Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
@@
 
-Figure 8 shows the point and effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 8 shows one point (mean) and three highest density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-\caption{50/80/95 percentile intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
+\caption{Means and 50/80/95\% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
 \end{figure}
@@
 
-\subsection{5.2 Individual differences}\label{individual-differences}
+\subsection{5.2 Model complexity versus interpretability}\label{model-complexity-versus-interpretability}
+
+Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to many higher-order interactions.
+For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period\_9), and the passage of time on the across-trial or within-experiment time scale (variable trial\_c).
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
+From a theoretical perspective, adding more than 1 time scale is important to capture plasticity (e.g., proactive control) and other learning effects that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
+From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental variables, because adding time scales quickly increases model complexity.
+
+\subsection{5.3 Individual differences}\label{individual-differences}
 
@@
 
-\subsection{5.3 Limitation(s)}\label{limitations}
+Another approach: fit models to individual subjects and describe prevalence\ldots REF
+
+\subsection{5.4 Limitation(s)}\label{limitations}
 
@@
 We did not discuss continuous-time hazard analysis.
-As indicated by --allison -- learning discrete-time methods first will help in learning continuous-time methods.
+As indicated by Allison (2010), learning discrete-time methods first will help in learning continuous-time methods.
 Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
@@
 
-\subsection{5.4 Extensions}\label{extensions}
+\subsection{5.5 Extensions}\label{extensions}
 
@@
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (REF).
-{[}{[}give a concrete example for this latter point{]}{]}
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (@ Allison, 2010).
+For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
 

### 2024-10-21 notes/manuscript_notes.Rmd
update ms.
respond to notes from Rich

@@
 
+-> Our custom functions are now located in the file custom_functions.R.
+The functions are also listed in the supplementary material.
+
 ### references ###
@@
 
+-> The .bib file is generated automatically. I have solved the problem by adding the argument "append=F" to the function r_refs().
+
 More generally, I'm not sure whether it is better to have the R package REFs in a footnote or in the main text as a paragraph. We can decide later on what works best.
+
+-> ok.
 
@@
 we use the terms hazard and EHA interchangeably. We should pick one, define it and then stick to it. 
+For example, the first line of the section "overview of hazard analysis" starts with "To apply event history analysis"
 
-For example, the first line of the section "overview of hazard analysis" starts with "To apply event history analysis"
+-> I agree, but we are also discussing SAT analyses. I had a reviewer once who insisted that SAT is different from EHA. So we could talk about EHA/SAT analyses perhaps?
 
@@
 I would also add as a general commwent that whenever we do insert formulas in text (in the future or in supplementary material here) we should number them and we should clearly define every part of the equation.
+
+-> all equations and formulas are moved to the supplementary material. Equations are numbered.
 
@@
 
+-> Yes, the big code chunks are removed.
+
 ### tables ###
@@
 tables 1 and 2, which I added, would be better and easier to read if they were either placed side-by-side or incorporated into one table. I'll leave that for you to figure out. You're welcome!!
+
+-> still to do.
 
@@
 Just an idea to ease the burden of information and give people a guide.
+
+-> still to do.
 
@@
 
+-> ok.
+
 ### figure 4 - h(t), S(t), ca(t) ###
@@
 the figure that shows all three panels.
+
+-> I added the probability mass function P(t) for each condition.
 
@@
 
+-> I added the following text:
+These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
+
 And why don't we show the group average data, rather than just PID 6? It seems odd to me without an explanation/justification.
+
+-> I added the following text:
+It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+
+When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
 
@@
 
+-> ok.
+
 Maybe in a later section we can add in this text to explain the significance or implications.
 
+-> I added the following text to the conclusion:
+Statistically controlling for the passage of time during data analysis is thus equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+
 It also read a little bit like we were making inferential claims from descriptive stats, so it shoudl wait until we get to a later tutorial maybe.
+
+-> I agree.
 
@@
 this can go in supps, as we are not writing a regression textbook.
+
+-> Done.
 
@@
 
+-> The example regression equations are moved to the supplementary material.
+
 I also suggest we discuss how to interpret the results when we get to the results.
+
+-> Done.
 
@@
 But we need to keep the flow and momentum in the main text.
+
+-> Done.
 
@@
 
+-> I removed the code that lists priors. I kept the model code for M1, and the model formula for models M2 to M4.
+
 ### avoid repetition in the model building steps ###
@@
 
+-> I agree.
+
 And when we move on to the frequentist models with lme4, we should only say something like... the general process is similar, except there are no priors to set.
+
+-> Done.
 
@@
 rather than output the code, which I've removed for the majority of instances, give some intuition on why we would want to specify formulas and models in this way.
+
+-> Done.
 
@@
 you say:
-
 >Clearly, both weighting schemes prefer model M4.
@@
 How is this clear? I don't think there are any plots or tables in the main text.
+
+-> I added the output of the model_weights() functions to section 4.3.7.
 
@@
 
+-> I added that table in Tutorials 2a and 2b.
+
 ### frequentist versus Bayesian plot ###
@@
 I like the idea of comparing. That's nice. But let's improve the figure and include intervals as well as points.
+
+-> to do.
 
@@
 This needs explaining or removing.
+
+-> I added the following text:
+In the no-mask condition (column 1 in Figure 5), we observe a positive compatibility effect in the hazard and ca(t) functions, as congruent primes temporarily generate higher values for hazard and conditional accuracy compared to incongruent primes.
+However, when a (relevant, irrelevant, or lines) mask is present (columns 2-4), there is a negative compatibility effect in the hazard and conditional accuracy functions, as congruent primes temporarily generate *lower* values for hazard and conditional accuracy compared to incongruent primes.
 
@@
 As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions [see @panisHowCanWe2020].
+
+-> I simplified the text to this:
+Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if h~A~(t) > h~B~(t) for all t, then both hazard functions are said to show a complete ordering.
+Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 Performing hazard analyses in experimental psychology has implications for how experiments are designed.
-Indeed, if trials are categorised as a function of when they occur, then each timebin will only include a subset of the total number of trials.
+Indeed, if trials are categorised as a function of when responses occur, then each timebin will only include a subset of the total number of trials.
 For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
@@
 
-Compared to the no-mask condition (column 1 in Figure 5), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+In the no-mask condition (column 1 in Figure 5), we observe a positive compatibility effect in the hazard and ca(t) functions, as congruent primes temporarily generate higher values for hazard and conditional accuracy compared to incongruent primes.
+However, when a (relevant, irrelevant, or lines) mask is present (columns 2-4), there is a negative compatibility effect in the hazard and conditional accuracy functions, as congruent primes temporarily generate *lower* values for hazard and conditional accuracy compared to incongruent primes.
 
@@
 
-When the shape of the hazard function is rather smooth, one can fit a more parsimonious model by using a polynomial specification of TIME.
+When the shape of the hazard function is rather smooth, as it is here for RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
 For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
@@
 However, the descriptive plots suggest that the effect of condition can vary across time bins.
+Previous research has indeed shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
 For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable "condition") and centered trial number (variable "trial_c").
@@
 
-Figure 6 shows one point (mean) and three highest density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 6 shows one point (mean) and three highest posterior density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
 
-(ref:descr-plot6-caption) Means and 50/80/95% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
+(ref:descr-plot6-caption) Means and 50/80/95% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
 
@@
 
-Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest density interval.
+Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
@@
 
-(ref:int-table-caption) Point and interval estimates, and hazard ratios.
+(ref:int-table-caption) Point and 95% highest posterior density interval estimates, and hazard ratios.
 
@@
 
-Figure 8 shows one point (mean) and three highest density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 8 shows one point (mean) and three highest posteriror density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
 
-(ref:descr-plot8-caption) Means and 50/80/95% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
+(ref:descr-plot8-caption) Means and 50/80/95% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
 
@@
 
-Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95% highest density interval.
+Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
@@
 
-(ref:int-ca-table-caption) Point and interval estimates, and odds ratios.
+(ref:int-ca-table-caption) Point and 95% highest posterior density interval estimates, and odds ratios.
 
@@
 In this fifth tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in Tutorial 1a.
+The general process is similar to that in Tutorial 2a, except that there are no priors to set.
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
@@
 Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
-However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+However, the random effects structure of model M3_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
@@
 331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
-Therefore, [[why should people care?
-What is the functional relevance for exp psych and researchers?]]
+Therefore, when the goal is to study how psychological effects change over time, hazard functions are preferred.
 
-[[This para needs to be way shorter and easier to read or we get rid of it]]
 Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if F~A~(t) > F~B~(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
-Townsend (1990) showed that a complete ordering on the hazard functions —$\lambda$~A~(t) > $\lambda$~B~(t) for all t— implies a complete ordering on both the cumulative distribution and survivor functions —F~A~(t) > F~B~(t) and S~A~(t) < S~B~(t)— which in turn implies an ordering on the mean latencies —mean A < mean B.
-In contrast, an ordering on two means does *not* imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does *not* imply a complete ordering on the corresponding hazard functions.
-This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, if h~A~(t) > h~B~(t) for all t, then both hazard functions are said to show a complete ordering.
+Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
 For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
-As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)$^-1$ -- one can simply plot delta-h(t) functions [see @panisHowCanWe2020].
 
@@
 This orthodox procedure can lead to a sampling bias, however, which results in underestimation of the mean.
-By introducting a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
 
@@
 
-Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to many higher-order interactions. 
+Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions. 
 For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period_9), and the passage of time on the across-trial or within-experiment time scale (variable trial_c).
@@
 For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
+Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.
 
@@
 RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
+Statistically controlling for the passage of time during data analysis is thus equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
 We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
-As such, we only supply regression equations in the supplementary material and then refer to them in the text whenever relevant.
+As such, we only supply regression equations in the supplementary material (part D) and then refer to them in the text whenever relevant.
 
@@
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions.
-Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches.
-The two types of data structure we label as 'person-trial' data (Table 1) and 'person-trial-bin' data (Table 2).
-The 'person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
-In contrast, the 'person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
-Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the timecourse of psychological processes.
+Second, we want to produce two different data sets that can each be submitted to different types of inferential modelling approaches.
+The two types of data structure we label as 'person-trial' data and 'person-trial-bin' data.
+The 'person-trial' data (Table 1) will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
+This data set is used when fitting conditional accuracy models.
 
@@
 
+In contrast, the 'person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
+Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+This data set is used when fitting hazard models.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
+Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
+
 (ref:ha-data-table-caption) Data structure for 'person-trial-bin' data
 
-(ref:ha-data-note-caption) The first 2 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.
+(ref:ha-data-note-caption) The first 2 trials for participant 1 from Table 1 are shown. The width of the time bins is 100 ms. These data are simulated and for illustrative purposes only.
 

### 2024-10-21 manuscript/manuscript.tex
update ms.

@@
 
+\begin{itemize}
+\tightlist
+\item
   1a. Wrangle raw data and descriptive stats for one independent variable.
+\item
   1b. Wrangle raw data and descriptive stats for two independent variables.
+\item
   2a. Bayesian multilevel modeling for h(t)
+\item
   2b. Bayesian multilevel modeling for ca(t)
+\item
   3a. Frequentist multilevel modeling for h(t)
+\item
   3b. Frequentist multilevel modeling for ca(t)
+\end{itemize}
 
@@
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions.
-Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches.
-The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2).
-The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
-In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
-Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the timecourse of psychological processes.
+Second, we want to produce two different data sets that can each be submitted to different types of inferential modelling approaches.
+The two types of data structure we label as `person-trial' data and `person-trial-bin' data.
+The `person-trial' data (Table 1) will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
+This data set is used when fitting conditional accuracy models.
 
@@
 
+In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
+Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+This data set is used when fitting hazard models.
+It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
+Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
+
 \begin{table}[H]
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} The first 2 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.}
+\normalsize{\textit{Note.} The first 2 trials for participant 1 from Table 1 are shown. The width of the time bins is 100 ms. These data are simulated and for illustrative purposes only.}
 \end{tablenotes}
@@
 
-Compared to the no-mask condition (column 1 in Figure 5), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
+In the no-mask condition (column 1 in Figure 5), we observe a positive compatibility effect in the hazard and ca(t) functions, as congruent primes temporarily generate higher values for hazard and conditional accuracy compared to incongruent primes.
+However, when a (relevant, irrelevant, or lines) mask is present (columns 2-4), there is a negative compatibility effect in the hazard and conditional accuracy functions, as congruent primes temporarily generate \emph{lower} values for hazard and conditional accuracy compared to incongruent primes.
 
@@
 However, the descriptive plots suggest that the effect of condition can vary across time bins.
+Previous research has indeed shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
 For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and centered trial number (variable ``trial\_c'').
@@
 
-Figure 6 shows one point (mean) and three highest density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 6 shows one point (mean) and three highest posterior density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-\caption{Means and 50/80/95\% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-effects}
+\caption{Means and 50/80/95\% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-effects}
 \end{figure}
 
-Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest density interval.
+Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
@@
 \begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:int-table}Point and interval estimates, and hazard ratios.}\\
+\caption{\label{tab:int-table}Point and 95\% highest posterior density interval estimates, and hazard ratios.}\\
 \toprule
@@
 
-Figure 8 shows one point (mean) and three highest density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 8 shows one point (mean) and three highest posteriror density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-\caption{Means and 50/80/95\% highest density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
+\caption{Means and 50/80/95\% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
 \end{figure}
 
-Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower limits of the 95\% highest density interval.
+Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
@@
 \begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:int-ca-table}Point and interval estimates, and odds ratios.}\\
+\caption{\label{tab:int-ca-table}Point and 95\% highest posterior density interval estimates, and odds ratios.}\\
 \toprule
@@
 Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
-However, the random effects structure of model M3 was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+However, the random effects structure of model M3\_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
@@
 Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-Therefore, {[}{[}why should people care?
-What is the functional relevance for exp psych and researchers?{]}{]}
-
-{[}{[}This para needs to be way shorter and easier to read or we get rid of it{]}{]}
+Therefore, when the goal is to study how psychological effects change over time, hazard functions are preferred.
+
 Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) for all t, then both cumulative distribution functions are said to show a complete ordering.
-Townsend (1990) showed that a complete ordering on the hazard functions ---\(\lambda\)\textsubscript{A}(t) \textgreater{} \(\lambda\)\textsubscript{B}(t) for all t--- implies a complete ordering on both the cumulative distribution and survivor functions ---F\textsubscript{A}(t) \textgreater{} F\textsubscript{B}(t) and S\textsubscript{A}(t) \textless{} S\textsubscript{B}(t)--- which in turn implies an ordering on the mean latencies ---mean A \textless{} mean B.
-In contrast, an ordering on two means does \emph{not} imply a complete ordering on the corresponding F(t) and S(t) functions, and a complete ordering on these latter functions does \emph{not} imply a complete ordering on the corresponding hazard functions.
-This means that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, if h\textsubscript{A}(t) \textgreater{} h\textsubscript{B}(t) for all t, then both hazard functions are said to show a complete ordering.
+Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
 For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
-As a result, instead of using delta-plots for RT -- differences in quantiles from F(t)\(^-1\) -- one can simply plot delta-h(t) functions (see Panis, 2020).
 
@@
 
-Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to many higher-order interactions.
+Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions.
 For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period\_9), and the passage of time on the across-trial or within-experiment time scale (variable trial\_c).

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@@allisonSurvivalAnalysisUsing2010].
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@allisonSurvivalAnalysisUsing2010].
 For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.

### 2024-10-21 manuscript/manuscript.tex
update ms.

@@
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (@ Allison, 2010).
+Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (Allison, 2010).
 For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.

### 2024-10-21 manuscript/manuscript.Rmd
add interpretation to ms.

@@
 
-Interpretation...
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-estimate in the baseline condition (no prime; red striped lines in Figure 6).
+For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240] of trial 500.
+Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360], (360,400], and (400,440].
+For example, in bin (320,360], the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
+While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and target-triggered response (e.g., right)
+
+In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240] and (240,280], while incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440].
+Due to practice, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+
+Towards the end of the experiment (trial 1500), both congruent and incongruent primes have positive and negative effects.
+Positive effects are present in bins (200,240] and (240,280].
+Incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440], and congruent primes now also have negative effects in bins (360,400] and (400,440].
+These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period_9), but also on the across-trial/within-experiment time scale (variable trial_c).
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of two separate cognitive processes.
+
+@panisWhatShapingRT2016 distinguished between response competition effects and cognitive control effects.
+While response competition is present in the incongruent trials throughout the experiment, @panisWhatShapingRT2016 argued that a cognitive control process is also involved, such as active top-down inhibition of the prime-triggered response.
+In other words, people learn that the premature prime-triggered response has to be actively inhibited in order to respond to the target stimulus.
+This inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the inhibitory effect of response competition in the incongruent trials.
 
@@
 
-Interpretation...
+Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines in Figure 8).
+For example, the odds ratio for congruent primes in bin (320,360] in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
+Incongruent primes have a negative effect on logit-ca(t) in time bins (200,240], (240,280], and (280,320] throughout the experiment, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines).
 
-plot 9 ca(t)
+To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+
+(ref:descr-plot9-caption) Model-based ca(t) functions for participant 6 in trial 500, 1000, and 1500.
+
+```{r plot-ca-subject6, fig.cap = "(ref:descr-plot9-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4_ca_effects_subject6.png")
+```
 

### 2024-10-21 manuscript/manuscript.tex
add interpretation to ms.

@@
 
-Interpretation\ldots{}
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-estimate in the baseline condition (no prime; red striped lines in Figure 6).
+For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240{]} of trial 500.
+Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360{]}, (360,400{]}, and (400,440{]}.
+For example, in bin (320,360{]}, the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
+While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and target-triggered response (e.g., right)
+
+In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}.
+Due to practice, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+
+Towards the end of the experiment (trial 1500), both congruent and incongruent primes have positive and negative effects.
+Positive effects are present in bins (200,240{]} and (240,280{]}.
+Incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}, and congruent primes now also have negative effects in bins (360,400{]} and (400,440{]}.
+These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period\_9), but also on the across-trial/within-experiment time scale (variable trial\_c).
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of two separate cognitive processes.
+
+Panis and Schmidt (2016) distinguished between response competition effects and cognitive control effects.
+While response competition is present in the incongruent trials throughout the experiment, Panis and Schmidt (2016) argued that a cognitive control process is also involved, such as active top-down inhibition of the prime-triggered response.
+In other words, people learn that the premature prime-triggered response has to be actively inhibited in order to respond to the target stimulus.
+This inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the inhibitory effect of response competition in the incongruent trials.
 
@@
 
-Interpretation\ldots{}
-
-plot 9 ca(t)
+Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines in Figure 8).
+For example, the odds ratio for congruent primes in bin (320,360{]} in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
+Incongruent primes have a negative effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, and (280,320{]} throughout the experiment, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines).
+
+To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4_ca_effects_subject6}
+
+}
+
+\caption{Model-based ca(t) functions for participant 6 in trial 500, 1000, and 1500.}\label{fig:plot-ca-subject6}
+\end{figure}
 

### 2024-10-21 manuscript/manuscript.Rmd
correct small error

@@
 
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-estimate in the baseline condition (no prime; red striped lines in Figure 6).
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240] of trial 500.

### 2024-10-21 manuscript/manuscript.Rmd
update interpretation hazard results

@@
 These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period_9), but also on the across-trial/within-experiment time scale (variable trial_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of two separate cognitive processes.
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of separate cognitive processes.
 
-@panisWhatShapingRT2016 distinguished between response competition effects and cognitive control effects.
-While response competition is present in the incongruent trials throughout the experiment, @panisWhatShapingRT2016 argued that a cognitive control process is also involved, such as active top-down inhibition of the prime-triggered response.
-In other words, people learn that the premature prime-triggered response has to be actively inhibited in order to respond to the target stimulus.
-This inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the inhibitory effect of response competition in the incongruent trials.
+@panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
+While response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent prime trials.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
+This global inhibitory effect might becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
 

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 @panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent prime trials.
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent prime trials.
 In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
-This global inhibitory effect might becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (here the target stimulus), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
 

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 @panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent prime trials.
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
 In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
-This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+This globa inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
 Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (here the target stimulus), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

--- 2024-10-21 manuscript/manuscript.Rmd
@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis.
+In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis for choice RT data.
 In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
@@
 
-Integrating results between hazard and conditional accuracy functions can be informative for understanding psychological processes.
+Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
 To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
@@
 Incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440], and congruent primes now also have negative effects in bins (360,400] and (400,440].
+
 These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period_9), but also on the across-trial/within-experiment time scale (variable trial_c).
@@
 @panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent prime trials.
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
 In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
 This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
-Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (here the target stimulus), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
+Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
 
@@
 In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, and (b) we use the logit link function.
 For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).

### 2024-10-21 Tutorial_2a.Rmd
delete old code

@@
 model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
-
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "stacking") %>% round(digits = 3)
 ```
@@
 
-# calculate point and interval estimates, and hazard ratios
+## calculate point and interval estimates, and hazard ratios
 
@@
 
-# plot effects on hazard ratio scale
-
-```{r haz-ratio}
-plot_facet_hr <- post %>% 
-  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
-         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
-
-  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
-  select(condition, period_9, value) %>%
-  mutate(value_hr = exp(value)) %>%
-  group_by(period_9) %>%
-
-   ggplot(aes(x = period_9, y = value_hr)) +
-  stat_lineribbon(show.legend=T) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=1, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,7)) +
-  scale_y_continuous( limits = c(0,40)) +
-  labs(x = "time bin", y = "hazard ratio") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top") +
-  facet_wrap(~ condition)
-
-plot_facet_hr
-
-```
-
-# Plot model-based predicted cloglog-hazard and hazard functions, for a single subject
+## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject
 
@@
 
-# Set up new data
+### Set up new data
 
@@
 
-# First, on the cloglog-scale:
+### First, on the cloglog-scale:
 
@@
 
-# Second, on hazard scale:
+### Second, on hazard scale:
 

### 2024-10-21 Tutorial_2b.Rmd
delete old code

@@
 
-# plot effects on odds ratio scale
-
-```{r odds-ratio}
-plot_facet_hr <- post %>% 
-  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
-         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
-
-  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
-  select(condition, period_9, value) %>%
-  mutate(value_hr = exp(value)) %>%
-  group_by(period_9) %>%
-
-   ggplot(aes(x = period_9, y = value_hr)) +
-  stat_lineribbon(show.legend=T) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=1, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,7)) +
-  scale_y_continuous( limits = c(0,40)) +
-  labs(x = "time bin", y = "hazard ratio") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top") +
-  facet_wrap(~ condition)
-
-plot_facet_hr
-
-```
-
-# Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject
+## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject
 
@@
 
-# First, on the logit-scale:
+### First, on the logit-scale:
 
@@
 
-# Second, on ca(t) scale:
+### Second, on ca(t) scale:
 

### 2024-10-21 manuscript/manuscript.tex
update ms.

@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis.
+In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis for choice RT data.
 In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
@@
 
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-estimate in the baseline condition (no prime; red striped lines in Figure 6).
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240{]} of trial 500.
@@
 Incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}, and congruent primes now also have negative effects in bins (360,400{]} and (400,440{]}.
+
 These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period\_9), but also on the across-trial/within-experiment time scale (variable trial\_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of two separate cognitive processes.
-
-Panis and Schmidt (2016) distinguished between response competition effects and cognitive control effects.
-While response competition is present in the incongruent trials throughout the experiment, Panis and Schmidt (2016) argued that a cognitive control process is also involved, such as active top-down inhibition of the prime-triggered response.
-In other words, people learn that the premature prime-triggered response has to be actively inhibited in order to respond to the target stimulus.
-This inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the inhibitory effect of response competition in the incongruent trials.
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of separate cognitive processes.
+
+Panis and Schmidt (2016) distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
+This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+Interestingly, while Panis and Schmidt (2016) did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
 
@@
 In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
 For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).

### 2024-10-21 manuscript/SupplementaryMaterial.Rmd
update notes and supps
update suppl. mat.

@@
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 12 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

@@
 While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
-The other rows in Figure 9 show prior distributions on the logit and cloglog scale that we use instead.
+The other rows in Figure 12 show prior distributions on the logit and cloglog scale that we use instead.
 
@@
 In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, and (b) we use the logit link function.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
 For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).

### 2024-10-21 manuscript/manuscript.tex
update ms.

@@
 To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 4 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 12 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).

### 2024-10-23 manuscript/manuscript.Rmd
update ms. and notes

@@
 To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis.
-We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferntial statistics via Bayesian and frequentist multi-level regression modelling.
+We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
 We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
@@
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
-Here, in the paper, we hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+Here, in the paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
 
@@
 
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time event history analysis, a.k.a.
+In this paper, we focus on a distributional method for time-to-event data known as *discrete-time event history analysis*, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
+We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010], and our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you should do them.
@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to EHA.
+In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
+In Tutorial 3a, we illustrate how to fit the same type of multilevel regression models for RT data in a frequentist framework using the R package lme4.
+We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
 In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
 
-In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials [@allisonSurvivalAnalysisUsing2010] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time hazard analysis on time-to-event data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials [@allisonSurvivalAnalysisUsing2010] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
 
-# 2. A brief introduction to hazard analysis
+# 2. A brief introduction to event history analysis
 
-For a comprehensive background context to hazard analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003].
+For a comprehensive background context to event history analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003].
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
-Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
+Instead, we want to provide an intuition to how event history analysis works in general as well as in the context of experimental psychology.
 As such, we only supply regression equations in the supplementary material (part D) and then refer to them in the text whenever relevant.
 
-## 2.1 Basic features of hazard analysis
+## 2.1 Basic features of event history analysis
 
-To apply event history analysis (EHA), one must be able to:
+To apply event history analysis (EHA), a.k.a. hazard analysis, one must be able to:
 
@@
 
-Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure 2).
+Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
 The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A).
-The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+The discrete-time hazard function gives you for each bin the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
@@
 
+[[this paragraph is not so clear yet... Perhaps add error bars to the functions in Figure 2?]]
 The survivor function can help to qualify or provide context to the interpretation of the hazard function.
@@
 
-## 2.2 Hazard analysis in the context of experimental psychology
+## 2.2 Event history analysis in the context of experimental psychology
 
@@
 
-In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer.
+In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer.
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time hazard analysis can be extended with a discrete-time SAT analyis.
+For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+We refer to this extended analysis for choice RT data as EHA/SAT.
 
@@
 Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
-These results show that for early responses (< bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later bins, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
+And then for longer waiting times, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 
-Performing hazard analyses in experimental psychology has implications for how experiments are designed.
+Performing event history analyses in experimental psychology has implications for how experiments are designed.
 Indeed, if trials are categorised as a function of when responses occur, then each timebin will only include a subset of the total number of trials.
@@
 
-Although the focus is on EHA, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
+Although the focus is on EHA/SAT, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
 
@@
 
-Tutorials 1a and 1b show how to calculate and plot the descriptive statistics when there are one and two independent variables, respectively.
+Tutorials 1a and 1b show how to calculate and plot the descriptive statistics of EHA/SAT when there are one and two independent variables, respectively.
 Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
@@
 
-* 1a. Wrangle raw data and descriptive stats for one independent variable.
-* 1b. Wrangle raw data and descriptive stats for two independent variables.
+* 1a. Wrangle raw data and calculate descriptive stats for one independent variable.
+* 1b. Wrangle raw data and calculate descriptive stats for two independent variables.
 * 2a. Bayesian multilevel modeling for h(t)
@@
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the timecourse of psychological processes.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes.
 Second, we want to produce two different data sets that can each be submitted to different types of inferential modelling approaches.
@@
 The 'person-trial' data (Table 1) will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
-This data set is used when fitting conditional accuracy models.
+This data set is used when fitting conditional accuracy models (Tutorials 2b and 3b).
 
@@
 Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-This data set is used when fitting hazard models.
+This data set is used when fitting hazard models (Tutorials 2a and 3a).
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
@@
 Removing such trials before calculating the mean event time can introduce a sampling bias (REFs).
-The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
+
+The person-trial-bin oriented dataset is created by our custom function ptb(), and it has one row for each time bin of each trial that is at risk for event occurrence.
 The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
 
-The next step is to plot the data using our custom function plot_eha(). When creating the plots, some warning messages will likely be generated, like these:
+The next step is to set up the life table -- custom function setup_lt() --, calculate the conditional accuracies -- custom function calc_ca() --, and then plot the descriptive statistics using our custom function plot_eha().
+When creating the plots, some warning messages will likely be generated, like these:
 
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
-Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6 -- compare to Figure 1.
-A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
+In general, it is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+
+Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
 At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
@@
 Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
-By using discrete-time h(t) functions of event occurrence -- in combination with ca(t) functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
 
@@
 
-  These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower response are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
 
-  These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+In general, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
 Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
-At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected time window for analysis.
-As such, multi-level regression is what we turn to in those tutorials.
+At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA/SAT.
+As we will show in later Tutorials, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
 But first we consider calculating the descriptive statistics for two independent variables.
@@
 
-(ref:descr-plot5-caption) Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
+(ref:descr-plot5-caption) Sample-based discrete-time hazard, survivor, conditional accuracy, and probability mass functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
 
@@
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in Tutorial 1a.
-Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
 
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see part C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see part C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 
-When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, you can use a general specification of TIME, i.e., one intercept per time bin.
+When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., one intercept per time bin.
 In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
@@
                            condition + trial_c +
-                           (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
+                           (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + 
+                           condition + trial_c | pid),
        prior = priors_M1,
@@
 The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of priming condition and centered trial number.
-Each of these effects is allowed to vary across individuals.
+Each of these effects is allowed to vary across individuals (variable pid).
 Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
@@
 
-When the shape of the hazard function is rather smooth, as it is here for RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
+When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
 For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
@@
 So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
-However, the descriptive plots suggest that the effect of condition can vary across time bins.
-Previous research has indeed shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
+However, the descriptive plots suggest that the effect of prime-target congruency varies across time bins.
+Previous research has shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
 For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable "condition") and centered trial number (variable "trial_c").
@@
 
-To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
@@
 Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
-For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240] of trial 500.
+For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240] of trial 500.
 Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360], (360,400], and (400,440].
 For example, in bin (320,360], the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
-While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and target-triggered response (e.g., right)
+While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and the target-triggered response (e.g., right)
 
@@
 @panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
-This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the response to the target stimulus.
+This global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
 Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).
+For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a) in the conditional accuracy model called M4_ca.
 
-To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Alternatively, one could also use the function glmmPQL() from the R package MASS (REF).
 The resulting hazard model is called M3_f.
@@
 
-This main motivation for writing this paper is the observation that event history analysis remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA provides compared to more conventional analyses.
-By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA, which is the increase in analytical complexity compared to mean-average comparisons.
-While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that event history analysis can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+This main motivation for writing this paper is the observation that event history and SAT analyses remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
+While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that EHAS/SAT can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
@@
 331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
-Therefore, when the goal is to study how psychological effects change over time, hazard functions are preferred.
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
 
@@
 
-Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
 EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
@@
 Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response), introducing a response deadline is recommended when designing RT experiments.
-Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard models.
+Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard and conditional accuracy models.
 
@@
 
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 
@@
 
-We did not discuss continuous-time hazard analysis.
-As indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time methods first will help in learning continuous-time methods.
+We did not discuss continuous-time EHA, nor continuous-time SAT analysis.
+As indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods.
 Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
@@
 For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states [@SteeleF2004].
+When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states [@SteeleF2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
@@
 RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
-Statistically controlling for the passage of time during data analysis is thus equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+Statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
 We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.

### 2024-10-23 manuscript/manuscript.tex
update ms. and notes

@@
 
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time event history analysis, a.k.a.
+In this paper, we focus on a distributional method for time-to-event data known as \emph{discrete-time event history analysis}, a.k.a.
 hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of hazard analysis to orient the reader to the basic concepts that we will use throughout the paper.
+We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003), and our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you should do them.
@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the hazard analysis for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of hazard regression models in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to EHA.
+In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
+In Tutorial 3a, we illustrate how to fit the same type of multilevel regression models for RT data in a frequentist framework using the R package lme4.
+We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
 In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
 
-In summary, even though event history analyses is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using hazard analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time hazard analysis on time-to-event data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
-
-\section{2. A brief introduction to hazard analysis}\label{a-brief-introduction-to-hazard-analysis}
-
-For a comprehensive background context to hazard analysis, we recommend several excellent textbooks (Singer \& Willett, 2003).
+In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+
+\section{2. A brief introduction to event history analysis}\label{a-brief-introduction-to-event-history-analysis}
+
+For a comprehensive background context to event history analysis, we recommend several excellent textbooks (Singer \& Willett, 2003).
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
-Instead, we want to provide an intuition to how EHA works in general as well as in the context of experimental psychology.
+Instead, we want to provide an intuition to how event history analysis works in general as well as in the context of experimental psychology.
 As such, we only supply regression equations in the supplementary material (part D) and then refer to them in the text whenever relevant.
 
-\subsection{2.1 Basic features of hazard analysis}\label{basic-features-of-hazard-analysis}
-
-To apply event history analysis (EHA), one must be able to:
+\subsection{2.1 Basic features of event history analysis}\label{basic-features-of-event-history-analysis}
+
+To apply event history analysis (EHA), a.k.a. hazard analysis, one must be able to:
 
@@
 
+{[}{[}this paragraph is not so clear yet\ldots{} Perhaps add error bars to the functions in Figure 2?{]}{]}
 The survivor function can help to qualify or provide context to the interpretation of the hazard function.
@@
 
-\subsection{2.2 Hazard analysis in the context of experimental psychology}\label{hazard-analysis-in-the-context-of-experimental-psychology}
+\subsection{2.2 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
 
@@
 
-In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer.
+In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer.
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time hazard analysis can be extended with a discrete-time SAT analyis.
+For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+We refer to this extended analysis for choice RT data as EHA/SAT.
 
@@
 Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
-These results show that for early responses (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-And then for later bins, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
+And then for longer waiting times, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 \item
-  1a. Wrangle raw data and descriptive stats for one independent variable.
-\item
-  1b. Wrangle raw data and descriptive stats for two independent variables.
+  1a. Wrangle raw data and calculate descriptive stats for one independent variable.
+\item
+  1b. Wrangle raw data and calculate descriptive stats for two independent variables.
 \item
@@
 The `person-trial' data (Table 1) will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
-This data set is used when fitting conditional accuracy models.
+This data set is used when fitting conditional accuracy models (Tutorials 2b and 3b).
 
@@
 Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-This data set is used when fitting hazard models.
+This data set is used when fitting hazard models (Tutorials 2a and 3a).
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
@@
 Removing such trials before calculating the mean event time can introduce a sampling bias (REFs).
-The person-trial-bin oriented dataset has one row for each time bin of each trial that is at risk for event occurrence.
+
+The person-trial-bin oriented dataset is created by our custom function ptb(), and it has one row for each time bin of each trial that is at risk for event occurrence.
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
 
-The next step is to plot the data using our custom function plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:
+The next step is to set up the life table -- custom function setup\_lt() --, calculate the conditional accuracies -- custom function calc\_ca() --, and then plot the descriptive statistics using our custom function plot\_eha().
+When creating the plots, some warning messages will likely be generated, like these:
 
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
-
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6 -- compare to Figure 1.
-A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se).
+
+In general, it is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
 At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
@@
 
-These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only later response are fully controlled by the target stimulus.
+These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower response are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
 
-These results go against the (often implicit) assumption that all observed responses are primed responses to the target stimulus.
+In general, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
 Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
-At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA.
-As we will show in Tutorials 2 and 3, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and accuracy (1/0) in each bin of a selected time window for analysis.
-As such, multi-level regression is what we turn to in those tutorials.
+At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA/SAT.
+As we will show in later Tutorials, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
 But first we consider calculating the descriptive statistics for two independent variables.
@@
 
-\caption{Sample-based discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
+\caption{Sample-based discrete-time hazard, survivor, conditional accuracy, and probability mass functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
 \end{figure}
@@
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in Tutorial 1a.
-Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
+In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
 
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable is binary, one has to select a link function (see part C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see part C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 \NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
-\NormalTok{                           (d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
+\NormalTok{                           (d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}
+\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
        \AttributeTok{prior =}\NormalTok{ priors\_M1,}
@@
 So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
-However, the descriptive plots suggest that the effect of condition can vary across time bins.
-Previous research has indeed shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
+However, the descriptive plots suggest that the effect of prime-target congruency varies across time bins.
+Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
 For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and centered trial number (variable ``trial\_c'').
@@
 Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
-For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials, in bin (200,240{]} of trial 500.
+For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240{]} of trial 500.
 Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360{]}, (360,400{]}, and (400,440{]}.
 For example, in bin (320,360{]}, the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
-While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and target-triggered response (e.g., right)
+While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and the target-triggered response (e.g., right)
 
@@
 Panis and Schmidt (2016) distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the target stimulus.
-This global inhibitory effect might become visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the response to the target stimulus.
+This global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
 Interestingly, while Panis and Schmidt (2016) did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a).
-
-To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effect of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a) in the conditional accuracy model called M4\_ca.
+
+To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Alternatively, one could also use the function glmmPQL() from the R package MASS (REF).
 The resulting hazard model is called M3\_f.
@@
 
-This main motivation for writing this paper is the observation that event history analysis remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA provides compared to more conventional analyses.
-By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA, which is the increase in analytical complexity compared to mean-average comparisons.
-While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that event history analysis can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+This main motivation for writing this paper is the observation that event history and SAT analyses remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
+While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHAS/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
@@
 Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-Therefore, when the goal is to study how psychological effects change over time, hazard functions are preferred.
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
 
@@
 
-Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and ca(t) functions) in order to know what to derive on the microscopic level.
+Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
 EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
@@
 
-We did not discuss continuous-time hazard analysis.
-As indicated by Allison (2010), learning discrete-time methods first will help in learning continuous-time methods.
+We did not discuss continuous-time EHA, nor continuous-time SAT analysis.
+As indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods.
 Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
@@
 For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
+When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.

### 2024-10-23 notes/manuscript_notes.Rmd
update notes and supps
update ms. and notes

@@
 
--> still to do.
+-> Still to do... It does not seem straightforward to place apa tables side by side. For now, I kept them separate and placed text between them.
 
@@
 
--> I removed the code that lists priors. I kept the model code for M1, and the model formula for models M2 to M4.
+-> I removed the code that lists priors. I kept the model code for M1, and the model formulas for models M2 to M4.
 
@@
 
--> to do.
+-> to do. Which intervals should we plot for the frequentist approach?
 
@@
 For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+
+# 23 oktober 2024
+
+Still to do:
+- combine tables showing person-trial and person-trial-period data sets.
+
+- REF for time-to-death (line 127 in ms.)
+
+- REF for reviews and tutorials on EHA (line 140)
+
+- REF for regression (line 145)
+
+- change paragraph (delete it and refer to suppl. material?), and add error bars to Figure 2? (line 162)
+
+- update section 3 and add REFs (line 200-215)
+
+- REF sampling bias (line 332)
+
+- REF for package MASS (line 665)
+
+- REF for prevalence (line 710)
+
+-frequentist versus Bayesian plot: which intervals should we plot for the frequentist model?
+
+- summary of steps involved in tutorials

### 2024-10-24 Tutorial_1a.Rmd
update tut1a
add content list to tutorials

@@
 
-# Load the libraries that we will be using.
+Content:
+1. Load libraries
+2. Read and wrangle data
+3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
+4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models
+5. Optional: set up life tables and make plots for data aggregated across participants
+
+# 1. Load the libraries that we will be using.
 
@@
 
-# Read in the raw data from Panis and Schmidt (2016) - can be skipped.
+# 2. Read and wrangle data
+
+Read in the raw data from Panis and Schmidt (2016) - can be skipped.
 
@@
 
-# Do some wrangling to get a simpler data set - can be skipped.
+Do some wrangling to get a simpler data set - can be skipped.
 
@@
 
-# Read in simpler data file, change variable names, and create factors.
+Read in simpler data file, change variable names, and create factors.
 
@@
 
-# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).
+# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
 
@@
 
-# Extract info, save plot for each subject, save a lifetable, create and save the person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.
+Save the calculations.
+
+```{r save-descr, eval=F}
+save(data_final, file="Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata")
+```
+
+# 4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models
 
@@
 ```
+
+Save the plots.
 
@@
 
-Create input file for hazard models.
+Create input file for fitting hazard models.
 
@@
 
-Create input file for ca(t) models.
+Create input file for fitting ca(t) models.
 
@@
 
-# Optional: Create single plot for data aggregated across subjects. This is only recommended when all subjects show the same qualitative patterns.
+# 5. Optional: set up life tables and make plots for data aggregated across participants
+
+Note: This is only recommended when all subjects show the same qualitative patterns.
 

### 2024-10-24 Tutorial_1b.Rmd
add content list to tutorials

@@
 
-Tutorial_1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
+Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
 
-# Load the libraries that we will be using.
+Content:
+1. Load libraries
+2. Read and wrangle data
+3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
+4. Extract info (life table, plots, ...)
+
+# 1. Load the libraries that we will be using.
 
@@
 
-# Read in the raw data from Panis and Schmidt (2016).
+# 2. Read and wrangle data
+
+Read in the raw data from Panis and Schmidt (2016).
 
@@
 
-# Do some wrangling.
+Do some wrangling.
 
@@
 
-# Change variable names, and create factors.
+Change variable names, and create factors.
 
@@
 
-# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), P(t), and ca(t)).
+# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
 
@@
 
-# Extract info, and save plot for each subject
+# 4. Extract info (life table, plots, ...)
 
@@
 
+Save the plots.
+
 ```{r save-plots, warning=F, eval=F}

### 2024-10-24 Tutorial_2a.Rmd
add content list to tutorials

@@
 
-# Load the libraries that we will be using
+Content:
+1. Load libraries.
+2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
+3. Plot the logit and complementary log-log (cloglog) link functions.
+4. Visualize different prior distributions on the logit and cloglog scales.
+5. Fit hazard models.
+6. Compare models using loo and waic.
+7. Display effects for the selected model.
+
+# 1. Load the libraries that we will be using.
 
@@
 
-# Set options.
+Set options.
 
@@
 
-# Load the person-trial-bin data set that we saved in Tutorial 1a.
+# 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
 
@@
 
-# Prepare the data set.
+Prepare the data set.
 
@@
 
-# Plot the logit and complementary log-log (cloglog) link functions
+# 3. Plot the logit and complementary log-log (cloglog) link functions.
 
@@
 
-# Visualize different prior distributions on the logit and cloglog scales
+# 4. Visualize different prior distributions on the logit and cloglog scales.
 
@@
 
-## Simulate, convert, and fit
+## Simulate, convert, and fit.
 
@@
   theme(panel.grid = element_blank())
-
-plot(density(dat$c))
 ```
@@
 ```{r fit-skew-normal, eval=F}
-# fit models
+# fit model
 fit_skewN <-
@@
 
-## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale
+## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale.
 
@@
 
-# Fit 4 cloglog-hazard models.
+# 5. Fit hazard models.
 
@@
 
-### Fit model M1
+### Fit model M1.
 
@@
 
-### Prepare data for M2
+### Prepare data for M2.
 
@@
 
-### Set up priors for M2
+### Set up priors for M2.
 
@@
 
-### Fit model M2
+### Fit model M2.
 
@@
 
-### Prepare data for M3
+### Prepare data for M3.
 
@@
 
-### Set up priors for M3
+### Set up priors for M3.
 
@@
 
-### Fit model M3
+### Fit model M3.
 
@@
 
-### Prepare data for M4
+### Prepare data for M4.
 
@@
 
-### Set up priors for M4
+### Set up priors for M4.
 
@@
 
-### Fit model M4
+### Fit model M4.
 
@@
 
-# Compare 4 models
+# 6. Compare models using loo and waic.
 
@@
 
-Compare all three models:
+Compare all three models.
 
@@
 
-# Plot effects for selected model M4
+# 7. Display effects for the selected model (model_M4).
 
-## Pareto k estimates
+## Pareto k estimates.
 
@@
 
-## calculate point and interval estimates, and hazard ratios
+## calculate point and interval estimates, and hazard ratios.
 
@@
 
-## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject
+## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject.
 
@@
 
-### Set up new data
+### Set up new data.
 
@@
 
-### First, on the cloglog-scale:
+### First, on the cloglog-scale.
 
@@
 
-### Second, on hazard scale:
+### Second, on hazard scale.
 

### 2024-10-24 Tutorial_2b.Rmd
add content list to tutorials

@@
 
-# Load the libraries that we will be using
+Content:
+1. Load libraries.
+2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
+3. Fit a conditional accuracy model.
+4. Display effects for the conditional accuracy model.
+
+# 1. Load the libraries that we will be using.
 
@@
 
-# Load the person-trial data set that we saved in Tutorial 1a.
+# 2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
 
@@
 
-# Fit model  M4_ca: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.
+# 3. Fit a conditional accuracy model.
 
-## Prepare data for M4_ca
+We fit a single model "M4_ca": We use a third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.
+
+## Prepare data for M4_ca.
 
@@
 
-### Set up priors for M4_ca
+### Set up priors for M4_ca.
 
@@
 
-### Fit model M4_ca
+### Fit model M4_ca.
 
@@
 
-# Display effects
+# 4. Display effects for the conditional accuracy model.
 
@@
 
-## calculate point and interval estimates, and hazard ratios
+## calculate point and interval estimates, and odds ratios.
 
@@
 
-## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject
+## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject.
 
@@
 
-### First, on the logit-scale:
+### First, on the logit-ca(t) scale.
 
@@
 
-### Second, on ca(t) scale:
+### Second, on the ca(t) scale.
 

### 2024-10-24 Tutorial_3a.Rmd
add fixed effects code
add content list to tutorials

@@
 
-# Load the libraries that we will be using
+Content:
+1. Load libraries.
+2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
+3. Fit hazard model in the frequentist framework (model_M3_f).
+4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.
+
+# 1. Load the libraries that we will be using
 
 ```{r load-pkg, results='hide'}
-pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")
+pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")
 
@@
 
-# Load the person-trial-bin data set that we saved in Tutorial 1.
+# 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
 
@@
 
-# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)
+# 3. 3. Fit hazard model in the frequentist framework (model_M3_f)
 
@@
 ```
+
+Model_M3_f contains the same effects as Model_M3 in Tutorial 2a.
 
@@
 
-# Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3
+```{r effects-M3-f}
+load("Tutorial_3_Frequentist/model_M3_f.RData")
+summary(model_M3_f)
+```
+
+# 4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.
 
@@
 ```{r extract-estimates}
-est <- tibble(brm = fixef(model_M3)[,1],
-              glmer = fixef(model_M3_f),
-              effects = names(fixef(model_M3_f))) 
+est <- tibble(brm = fixef(model_M3)[,c(1,3,4)],
+              glmer = tidy(model_M3_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
+              effectname = names(fixef(model_M3_f)),
+              position = 1:16) 
 ```
@@
 
-p1 <- ggplot(data = est, aes(x = effects))
-p1 <- p1 + geom_point(aes( y = brm,  color = "c1", shape = "s1"))
-p1 <- p1 + geom_point(aes( y = glmer, color = "c2", shape = "s2")) 
-p1 <- p1 + labs( x = "Effects", y = "cloglog-hazard" )
-p1 <- p1 + scale_color_manual(name = "Model", 
+p1 <- ggplot(data = est, aes(x = position)) +
+ geom_point(aes( y = brm[,"Estimate"],  color = "c1", shape = "s1")) +
+ geom_point(aes( y = glmer$estimate, color = "c2", shape = "s2"), position = position_nudge(x = .1))  +
+ geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], color = "c1"), show.legend = T) +
+ geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, color = "c2"), show.legend = T, position = position_nudge(x = .1)) +
+ geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 1:16,labels = est$effectname) +
+ labs( x = "Effects in model M3", y = "cloglog-hazard" ) +
+ scale_color_manual(name = "Model", 
                                 breaks = c("c1", "c2"), 
                                 values = cols,
-                                labels = c("brm", "glmer"))
-p1 <- p1 + scale_shape_manual(name = "Model", 
+                                labels = c("brm", "glmer")) +
+ scale_shape_manual(name = "Model", 
                               breaks = c("s1", "s2"),
@@
          legend.position = "top",
-         axis.text.x = element_text(angle=90)) 
+         axis.text.x = element_text(size=11,angle=90),
+         axis.text.y = element_text(size=13),
+         axis.title.x = element_text(size = 14),
+         axis.title.y = element_text(size = 14),
+         legend.text=element_text(size=13)) 
 p1

### 2024-10-24 Tutorial_3b.Rmd
add fixed effects code
add content list to tutorials

@@
 
-# Load the libraries that we will be using
+Content:
+1. Load libraries.
+2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
+3. Fit a conditional accuracy model in the frequentist framework (model M3_ca_f).
+
+# 1. Load the libraries that we will be using.
 
@@
 
-# Load the data set for ca(t) modeling that we saved in Tutorial 1a.
+# 2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
 
@@
 
-# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)
+# 3. Fit a conditional accuracy model in the frequentist framework (model M3_ca_f).
 
@@
 ```
+
+Model_M3_ca_f contains the same effects as Model_M3 in Tutorial 2a.
 
@@
 ```
+
+```{r effects-M3-f}
+load("Tutorial_3_Frequentist/model_M3_ca_f.RData")
+summary(model_M3_ca_f)
+```

### 2024-10-24 manuscript/SupplementaryMaterial.Rmd
update suppl. material

@@
 * join_lt_ca_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
-* extract_median(df) : estimate quantiles S(t).50 (1 independent variable)
-* extract_median_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
+* extract_median(df) : estimate quantiles S(t)~.50~ (1 independent variable)
+* extract_median_2IV(df) : estimate quantiles S(t)~.50~ (2 independent variables)
 * plot_eha(df,subj,haz_yaxis) : create plots of the discrete-time functions (1 independent variable)

### 2024-10-24 notes/manuscript_notes.Rmd
update notes todo

--- 2024-10-21 notes/manuscript_notes.Rmd
@@
 For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+
+# 23 oktober 2024
+
+Still to do:
+- combine tables showing person-trial and person-trial-period data sets
+- REF for regression (line 145) 
+- update section 3 and add REFs (line 200-215)
+
+- REF for time-to-death (line 127 in ms.) : Done
+- REF for reviews and tutorials on EHA (line 140) : Done
+- change paragraph (delete it and refer to suppl. material?), and add error bars to Figure 2? (line 162): Done (no error bars added, but I refer to suppl. material)
+- REF sampling bias (line 332): Instead of adding a REF, I rephrased the sentence.
+- REF for package MASS (line 665) : Done
+- REF for prevalence (line 710) : Done
+-frequentist versus Bayesian plot with intervals: Done
+- summary of steps involved in tutorials: Done.
+- check odds ratio rounding : Done 
+(added to apa_table(): format.args = list(
+       digits = c(0, 2,2,2,2,2,5),
+        decimal.mark = ".", big.mark = ""))

### 2024-10-24 manuscript/manuscript.Rmd
update ms. (add refs and format.args)

@@
 wordcount         : "X"
-bibliography      : ["extrareferences.bib", "r-references.bib"]
+bibliography      : ["extrareferences2.bib", "r-references.bib"]
 floatsintext      : yes
@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death [@williammatthewmakehamLawMortalityConstruction1860].
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
@@
 
-In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials [@allisonSurvivalAnalysisUsing2010] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
 Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
@@
 In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
+As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
 
-[[this paragraph is not so clear yet... Perhaps add error bars to the functions in Figure 2?]]
-The survivor function can help to qualify or provide context to the interpretation of the hazard function.
+<!-- The survivor function can help to qualify or provide context to the interpretation of the hazard function.
 For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
 If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500], then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.
+Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.-->
 
@@
 We refer to this extended analysis for choice RT data as EHA/SAT.
+As explained in part A of the supplementary material, the probability mass function provides a context for interpreting the conditional accuracy function.
 
@@
 In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
-Removing such trials before calculating the mean event time can introduce a sampling bias (REFs).
+Removing such trials before calculating the mean event time will result in underestimation of the true mean.
 
@@
 
-The next step is to set up the life table -- custom function setup_lt() --, calculate the conditional accuracies -- custom function calc_ca() --, and then plot the descriptive statistics using our custom function plot_eha().
+The next step is to set up the life table using our custom function setup_lt(), calculate the conditional accuracies using our custom function calc_ca(), add the ca(t) estimates to the life table using our custom function join_lt_ca(), and then plot the descriptive statistics using our custom function plot_eha().
 When creating the plots, some warning messages will likely be generated, like these:
@@
 
-Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022], we relax all three assumptions in model 4.
-We thus add a squared term for the continuous predictor centered trial number -- I(trial_c^2) -- and include interaction terms.
+Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] we expect nonlinear effects of trial number.
+We thus relax all three assumptions in model 4.
+We add a squared term for the continuous predictor centered trial number -- I(trial_c^2) -- and include interaction terms.
 
@@
   longtable = TRUE,
-  repeat_header = TRUE
+  repeat_header = TRUE,
+  format.args = list(
+       digits = c(0, 2,2,2,2,2,5),
+        decimal.mark = ".", big.mark = "")
 )
@@
 In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240] and (240,280], while incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440].
-Due to practice, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+Due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
 
@@
   longtable = TRUE,
-  repeat_header = TRUE
+  repeat_header = TRUE,
+  format.args = list(
+       digits = c(0, 2,2,2,2,2,13),
+        decimal.mark = ".", big.mark = "")
 )
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (REF).
+Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
 The resulting hazard model is called M3_f.
@@
 
-(ref:descr-plot10-caption) Parameter estimates for model M3 from brm() and model M3_f from glmer().
+(ref:descr-plot10-caption) Parameter estimates for model M3 from brm() -- means and 95% quantile intervals -- and model M3_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
 
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
 Again, the resulting conditional accuracy model M3_ca_f did not converge and resulted in a singular fit.
@@
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
-From a theoretical perspective, adding more than 1 time scale is important to capture plasticity (e.g., proactive control) and other learning effects that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
-From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental variables, because adding time scales quickly increases model complexity.
+From a theoretical perspective, adding more than 1 time scale -- and their interactions -- is important to capture plasticity and other learning effects (e.g., proactive control) that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
+From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental predictor variables, because adding time scales quickly increases model complexity.
 
@@
 
-Another approach: fit models to individual subjects and describe prevalence...REF
+Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021].
+This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
+This approach allows one to quantify how typical or uncommon an observed effect is in the population, and the uncertainty around this estimate.
 
@@
 For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states [@SteeleF2004].
+When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.

### 2024-10-24 manuscript/manuscript.tex
update ms. (add refs and format.args)

@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death (William Matthew Makeham, 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
@@
 
-In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
 Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
@@
 In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
-
-{[}{[}this paragraph is not so clear yet\ldots{} Perhaps add error bars to the functions in Figure 2?{]}{]}
-The survivor function can help to qualify or provide context to the interpretation of the hazard function.
-For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
-If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500{]}, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.
+As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
 
@@
 
-The next step is to set up the life table -- custom function setup\_lt() --, calculate the conditional accuracies -- custom function calc\_ca() --, and then plot the descriptive statistics using our custom function plot\_eha().
+The next step is to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha().
 When creating the plots, some warning messages will likely be generated, like these:
@@
 
-Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4.
-We thus add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.
+Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) we expect nonlinear effects of trial number.
+We thus relax all three assumptions in model 4.
+We add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.
 
@@
 \endhead
-240.00 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82\\
-280.00 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03\\
-320.00 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45\\
-360.00 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91\\
-400.00 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.71\\
-440.00 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.64\\
-480.00 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62\\
-520.00 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.62\\
-560.00 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.60\\
-600.00 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.52\\
-240.00 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63\\
-280.00 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.13\\
-320.00 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90\\
-360.00 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99\\
-400.00 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68\\
-440.00 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58\\
-480.00 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58\\
-520.00 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.64\\
-560.00 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71\\
-600.00 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.78\\
-240.00 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02\\
-280.00 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.66\\
-320.00 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.52\\
-360.00 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68\\
-400.00 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.40\\
-440.00 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28\\
-480.00 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23\\
-520.00 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20\\
-560.00 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18\\
-600.00 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15\\
-240.00 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10\\
-280.00 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72\\
-320.00 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58\\
-360.00 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.34\\
-400.00 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.30\\
-440.00 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.34\\
-480.00 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44\\
-520.00 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.58\\
-560.00 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65\\
-600.00 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.56\\
-240.00 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68\\
-280.00 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88\\
-320.00 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78\\
-360.00 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.40\\
-400.00 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.33\\
-440.00 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.39\\
-480.00 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.56\\
-520.00 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87\\
-560.00 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27\\
-600.00 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52\\
-240.00 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07\\
-280.00 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.94\\
-320.00 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.72\\
-360.00 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.35\\
-400.00 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28\\
-440.00 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.35\\
-480.00 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55\\
-520.00 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00\\
-560.00 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.79\\
-600.00 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75\\
+240 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82080\\
+280 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03199\\
+320 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45383\\
+360 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91397\\
+400 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.70802\\
+440 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.63522\\
+480 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62035\\
+520 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.61980\\
+560 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.59543\\
+600 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.51694\\
+240 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63183\\
+280 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.12611\\
+320 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90342\\
+360 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99277\\
+400 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68359\\
+440 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58403\\
+480 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58190\\
+520 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.63546\\
+560 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71487\\
+600 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.77863\\
+240 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02227\\
+280 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.65584\\
+320 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.51502\\
+360 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68077\\
+400 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.39703\\
+440 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28245\\
+480 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23037\\
+520 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20247\\
+560 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18021\\
+600 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15267\\
+240 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10461\\
+280 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72468\\
+320 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58233\\
+360 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.33902\\
+400 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.29661\\
+440 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.33991\\
+480 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44474\\
+520 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.57904\\
+560 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65388\\
+600 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.55820\\
+240 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68463\\
+280 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88377\\
+320 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78490\\
+360 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.39866\\
+400 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.32935\\
+440 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.38574\\
+480 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.55825\\
+520 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87013\\
+560 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27313\\
+600 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52411\\
+240 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07329\\
+280 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.93821\\
+320 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.71847\\
+360 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.34502\\
+400 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28360\\
+440 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.34778\\
+480 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55459\\
+520 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00234\\
+560 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.78962\\
+600 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75123\\
 \bottomrule
@@
 In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}.
-Due to practice, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+Due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
 
@@
 \endhead
-240.00 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24,618,458.61\\
-280.00 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4,846.09\\
-320.00 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.79\\
-360.00 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.89\\
-400.00 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.19\\
-440.00 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.31\\
-480.00 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.41\\
-520.00 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.40\\
-560.00 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.12\\
-600.00 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.00\\
-240.00 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1,899,836.02\\
-280.00 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1,961.83\\
-320.00 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.87\\
-360.00 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.57\\
-400.00 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.14\\
-440.00 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.06\\
-480.00 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.10\\
-520.00 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.00\\
-560.00 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.00\\
-600.00 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.00\\
-240.00 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10,001,085.39\\
-280.00 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5,124.44\\
-320.00 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.12\\
-360.00 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.82\\
-400.00 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.58\\
-440.00 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.82\\
-480.00 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.08\\
-520.00 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.00\\
-560.00 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.00\\
-600.00 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.00\\
-240.00 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.00\\
-280.00 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.00\\
-320.00 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.02\\
-360.00 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.38\\
-400.00 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.59\\
-440.00 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.59\\
-480.00 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.30\\
-520.00 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.04\\
-560.00 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3,282,435.93\\
-600.00 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46,319,712,352,328.76\\
-240.00 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.00\\
-280.00 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.00\\
-320.00 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.04\\
-360.00 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.59\\
-400.00 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.10\\
-440.00 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.68\\
-480.00 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.20\\
-520.00 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3,307.44\\
-560.00 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164,758,701.84\\
-600.00 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10,165,856,639,901,592.00\\
-240.00 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.00\\
-280.00 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.00\\
-320.00 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.06\\
-360.00 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.87\\
-400.00 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.57\\
-440.00 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.21\\
-480.00 & i1500 & 5.52 & -13.87 & 26.57 & 0.95 & 249.51\\
-520.00 & i1500 & 12.67 & -31.46 & 58.41 & 0.95 & 318,500.40\\
-560.00 & i1500 & 25.50 & -53.08 & 115.21 & 0.95 & 119,299,568,240.94\\
-600.00 & i1500 & 45.85 & -86.60 & 200.06 & 0.95 & 81,670,189,671,651,033,088.00\\
+240 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24618458.6117376089096\\
+280 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4846.0911175084821\\
+320 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.7913398591541\\
+360 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.8918515266144\\
+400 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.1907589464532\\
+440 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.3094780953354\\
+480 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.4104953301397\\
+520 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.3973471906322\\
+560 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.1244195286526\\
+600 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.0000573483764\\
+240 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1899836.0176862408407\\
+280 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1961.8278183493471\\
+320 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.8743225515544\\
+360 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.5713234449885\\
+400 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.1378140505531\\
+440 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.0636745027736\\
+480 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.0982112115060\\
+520 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.0008224865296\\
+560 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.0000002080447\\
+600 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.0000000000005\\
+240 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10001085.3946626689285\\
+280 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5124.4351045810508\\
+320 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.1208056719069\\
+360 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.8233561607888\\
+400 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.5848218057173\\
+440 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.8181575226781\\
+480 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.0826145785621\\
+520 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.0008862056864\\
+560 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.0000003362909\\
+600 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.0000000000015\\
+240 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.0000000000730\\
+280 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.0000261643171\\
+320 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.0246001196011\\
+360 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.3780876098423\\
+400 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.5916606671310\\
+440 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.5871858652992\\
+480 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.3019651321687\\
+520 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.0443659277399\\
+560 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3282435.9279020344839\\
+600 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46319712352328.7578125000000\\
+240 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.0000000003243\\
+280 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.0000632158160\\
+320 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.0419655563481\\
+360 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.5909004105316\\
+400 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.0992267336787\\
+440 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.6827111411806\\
+480 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.2031868705690\\
+520 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3307.4439707159477\\
+560 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164758701.8390493392944\\
+600 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10165856639901592.0000000000000\\
+240 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.0000000012362\\
+280 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.0001189822174\\
+320 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.0575421441866\\
+360 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.8709638702927\\
+400 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.5699339941536\\
+440 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.2076572160001\\
+480 & i1500 & 5.52 & -13.87 & 26.57 & 0.95 & 249.5050127299390\\
+520 & i1500 & 12.67 & -31.46 & 58.41 & 0.95 & 318500.3986836019321\\
+560 & i1500 & 25.50 & -53.08 & 115.21 & 0.95 & 119299568240.9411773681641\\
+600 & i1500 & 45.85 & -86.60 & 200.06 & 0.95 & 81670189671651033088.0000000000000\\
 \bottomrule
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (REF).
+Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
 The resulting hazard model is called M3\_f.
@@
 
-\caption{Parameter estimates for model M3 from brm() and model M3\_f from glmer().}\label{fig:plot-comparison}
+\caption{Parameter estimates for model M3 from brm() -- means and 95\% quantile intervals -- and model M3\_f from glmer() -- maximum likelihood estimates and 95\% confidence intervals.}\label{fig:plot-comparison}
 \end{figure}
@@
 For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS.
+Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
 Again, the resulting conditional accuracy model M3\_ca\_f did not converge and resulted in a singular fit.
@@
 
-Another approach: fit models to individual subjects and describe prevalence\ldots REF
+Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021).
+This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
+This approach allows one to quantify how typical or uncommon an observed effect is in the population, and the uncertainty around this estimate.
 
@@
 
+\bibitem[\citeproctext]{ref-blossfeldTechniquesEventHistory2002}
+Blossfeld, H.-P., \& Rohwer, G. (2002). \emph{Techniques of event history modeling: {New} approaches to causal analysis, 2nd ed} (pp. x, 310). Mahwah, NJ, US: Lawrence Erlbaum Associates Publishers.
+
+\bibitem[\citeproctext]{ref-box-steffensmeierEventHistoryModeling2004}
+Box-Steffensmeier, J. M. (2004). Event history modeling: A guide for social scientists. Cambridge: University Press.
+
 \bibitem[\citeproctext]{ref-R-brms_a}
@@
 
+\bibitem[\citeproctext]{ref-hosmerAppliedSurvivalAnalysis2011}
+Hosmer, D. W., Lemeshow, S., \& May, S. (2011). \emph{Applied {Survival Analysis}: {Regression Modeling} of {Time} to {Event Data}} (2nd ed). Hoboken: John Wiley \& Sons.
+
+\bibitem[\citeproctext]{ref-inceBayesianInferencePopulation2021}
+Ince, R. A., Paton, A. T., Kay, J. W., \& Schyns, P. G. (2021). Bayesian inference of population prevalence. \emph{eLife}, \emph{10}, e62461. \url{https://doi.org/10.7554/eLife.62461}
+
 \bibitem[\citeproctext]{ref-kantowitzInterpretationReactionTime2021}
@@
 
+\bibitem[\citeproctext]{ref-landesIntroductionEventHistory2020}
+Landes, J., Engelhardt, S. C., \& Pelletier, F. (2020). An introduction to event history analyses for ecologists. \emph{Ecosphere}, \emph{11}(10), e03238. \url{https://doi.org/10.1002/ecs2.3238}
+
 \bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
@@
 
+\bibitem[\citeproctext]{ref-ripleyMASSSupportFunctions2024}
+Ripley, B., Venables, B., Bates, D.
+M., ca 1998), K.
+H.
+(partial.
+port, ca 1998), A.
+G.
+(partial.
+port, \& polr), D.
+F.
+(support.
+functions for.
+(2024).
+\emph{{MASS}: {Support Functions} and {Datasets} for {Venables} and {Ripley}'s {MASS}}.
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
@@
 
-\bibitem[\citeproctext]{ref-SteeleF2004}
+\bibitem[\citeproctext]{ref-steeleGeneralMultilevelMultistate2004}
 Steele, F., Goldstein, H., \& Browne, W.
@@
 \url{https://doi.org/10.1191/1471082X04st069oa}
+
+\bibitem[\citeproctext]{ref-teachmanAnalyzingSocialProcesses1983}
+Teachman, J.
+D.
+(1983).
+Analyzing social processes: {Life} tables and proportional hazards models.
+\emph{Social Science Research}, \emph{12}(3), 263--301.
+\url{https://doi.org/10.1016/0049-089X(83)90015-7}
 
@@
 Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
+
+\bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
+William Matthew Makeham. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
 

### 2024-10-25 manuscript/manuscript.Rmd
update ms. and extrareferences2.bib

@@
 Descriptive, data science workflow.
-Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
+Data wrangling via tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
 Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke & Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?).
-We also use a "keep it maximal" approach to specifying varying (or random) effects (Barr et al., 2013).
+Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018] ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018] ; Kurz, 202?).
+We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
 This means that wherever possible we include varying intercepts and slopes per participant

### 2024-10-25 manuscript/manuscript.tex
update ms. and extrareferences2.bib

@@
 Descriptive, data science workflow.
-Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
+Data wrangling via tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
 Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?).
-We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013).
+Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018) ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2018) ; Kurz, 202?).
+We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
 This means that wherever possible we include varying intercepts and slopes per participant
@@
 
+\bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
+Barr, D.
+J., Levy, R., Scheepers, C., \& Tily, H.
+J.
+(2013).
+Random effects structure for confirmatory hypothesis testing: {Keep} it maximal.
+\emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001.
+\url{https://doi.org/10.1016/j.jml.2012.11.001}
+
 \bibitem[\citeproctext]{ref-R-tinylabels}
@@
 \url{https://doi.org/10.1016/j.neunet.2012.09.003}
+
+\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
+Kruschke, J.
+K., \& Liddell, T.
+M.
+(2018).
+The {Bayesian New Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power analysis from a {Bayesian} perspective.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 178--206.
+\url{https://doi.org/10.3758/s13423-016-1221-4}
 
@@
 
+\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
+Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G.
+(2023).
+\emph{R for data science: Import, tidy, transform, visualize, and model data} (2nd edition).
+Beijing Boston Farnham Sebastopol Tokyo: O'Reilly.
+
 \bibitem[\citeproctext]{ref-R-dplyr}
@@
 \item
-  extract\_median(df) : estimate quantiles S(t).50 (1 independent variable)
-\item
-  extract\_median\_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
+  extract\_median(df) : estimate quantiles S(t)\textsubscript{.50} (1 independent variable)
+\item
+  extract\_median\_2IV(df) : estimate quantiles S(t)\textsubscript{.50} (2 independent variables)
 \item

### 2024-10-27 Tutorial_2a.Rmd
change label in plot priors

@@
   scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-12, y=140000, label="t(7.61)*1.57",
+  annotate(geom="text", x=-12, y=140000, label="t(7.61,0,1.57)",
               color="red", size=6) +

### 2024-10-27 manuscript/manuscript.Rmd
update ms.

@@
 ---
-title             : "Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows"
+title             : "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows"
 shorttitle        : "A tutorial on hazard analysis"
@@
 In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022].
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development and our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 
-As a simple illustration, Figure 1 shows the results of several simulated RT datasets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distribution differs (Figure 1, top row).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distribution differs in each case (Figure 1, bottom row).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distribution would lead to a different conclusion in every case.
+As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributios differ (Figure 1, top row).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, bottom row).
+Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
 
-(ref:descr-fig1-caption) Means versus distributional shapes for six different simulated dataset examples.
+(ref:descr-fig1-caption) Means versus distributional shapes for six different simulated data set examples.
+The first second after stimulus onset is divided in ten bins of 100 ms.
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
@@
 Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the timecourse of psychological states.
+Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of waiting time -- to our theory building toolkit.
+As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of time -- to our theory building toolkit.
 
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death [@williammatthewmakehamLawMortalityConstruction1860].
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death [e.g., @williammatthewmakehamLawMortalityConstruction1860].
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
@@
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
-Additionally, each tutorial is made available as .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
+Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
 
@@
 In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of multilevel regression models for RT data in a frequentist framework using the R package lme4.
+In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
@@
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs after bin t.
-In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
 As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
@@
 
-(ref:descr-plot2-caption) Hazard and survivor functions
+(ref:descr-plot2-caption) Discrete-time hazard and survivor functions.
 
@@
 Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
+To illustrate, we consider a hypothetical choice RT example that is inspired by real data [@panisWhatShapingRT2016], but simplified to make the main point clearer (Figure 3).
 In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
@@
 
-(ref:descr-plot3-caption) Hazard and conditional accuracy
+(ref:descr-plot3-caption) Discrete-time hazard and conditional accuracy functions.
 
@@
 In contrast to the large-*N* design that typically average across many participants without being able to scrutinize individual data patterns,  small-*N*  designs  retain  crucial  information  about  the  data  patterns of individual observers.
-This can be advantageous whenever participants differ systematically in their strategies or in the time-courses of their effects, so that averaging them would lead to misleading data patterns.
+This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
@@
 Descriptive, data science workflow.
-Data wrangling via tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
+We perform data wrangling following tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
 Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
@@
 
-Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018] ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018] ; Kurz, 202?).
+Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018] ; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].
 We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
@@
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy.
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2018].
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 In contrast, the 'person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the 'person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+Therefore, the 'person-trial-bin' data set generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
 This data set is used when fitting hazard models (Tutorials 2a and 3a).
@@
   mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
-  mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
+  mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin data set
   mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca(t)
@@
 
-The person-trial-bin oriented dataset is created by our custom function ptb(), and it has one row for each time bin of each trial that is at risk for event occurrence.
+The person-trial-bin oriented data set is created by our custom function ptb(), and it has one row for each time bin (of each trial) that is at risk for event occurrence.
 The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
@@
 
-In general, it is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+In general, it is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was always guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
@@
   note = "(ref:life-note-caption)",
-  placement = "H"
+  placement = "H",
+  format.args = list(
+    digits = c(0,0,0,2,2,2,2,2,2),
+    decimal.mark = ".", big.mark = "")
 )
@@
 
-(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
+(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
 
@@
 
-When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
+When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
 
-These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower response are fully controlled by the target stimulus.
+These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
@@
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multi-level regression models to the RT data of the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
-Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
+Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t) in a selected baseline condition.
 One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-We provide relevant example regression formulas in part D of the supplementary material.
+Relevant example regression formulas are provided in equations 6 and 7 in part D of the supplementary material.
 
-In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set [@allisonSurvivalAnalysisUsing2010].
+In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set [@allisonSurvivalAnalysisUsing2010].
 When there is clustering in the data, as in the case of a small-*N* design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models [@allisonSurvivalAnalysisUsing2010].
@@
 In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We select the analysis range (200,600] and the cloglog link.
+We decided to select the analysis time window (200,600] and the cloglog link.
 The data is prepared as follows.
@@
 ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-
-# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
+# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)
 ptb_data <- ptb_data %>% filter(period > 5)
-
 # create factor condition, with "blank" as the reference level
 ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-
-# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.
+# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.
 ptb_data <- ptb_data %>% 
@@
 
-When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., one intercept per time bin.
+When you do not want to make assumptions about the shape of the hazard function in the selected baseline condition, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one intercept per time bin.
 In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
@@
 ```{r fit-model-M1, echo=T, eval=F}
-plan(multicore)
-
 model_M1 <-
@@
 After selecting the binomial family and the cloglog link, the model formula is specified.
-The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of priming condition and centered trial number.
+The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of prime-target congruency (variable condition) and centered trial number (variable trial_c).
 Each of these effects is allowed to vary across individuals (variable pid).
@@
 When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
-For our second example model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
+For our second example model, we thus use a third-order polynomial specification of TIME for the selected baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
 The model formula for model M2 looks as follows.
@@
 Because TIME is centered on bin 9, and trial number on trial 1000, the Intercept represents the cloglog-hazard in bin 9 for the blank prime condition in trial 1000.
-Estimating model M2 took about 144 minutes.
+Estimating model M2 took about 2.5 hours.
 
@@
 
-So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
-However, the descriptive plots suggest that the effect of prime-target congruency varies across time bins.
+So far, we assumed that the effect of our predictors prime-target congruecy and centered trial number are the same in each time bin.
+However, the descriptive plots (e.g., Figure 4) suggest that the effect of prime-target congruency varies across time bins.
 Previous research has shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
-For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable "condition") and centered trial number (variable "trial_c").
+For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable condition) and centered trial number (variable trial_c).
 
@@
 
-Note that duplicate terms in the model formula are ignored. Estimating model M3 took about 268 minutes.
+Note that duplicate terms in the model formula are ignored. Estimating model M3 took about 4.5 hours.
 
@@
 
-Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] we expect nonlinear effects of trial number.
+Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] we expect nonlinear effects of trial number on the hazard of response occurrence.
 We thus relax all three assumptions in model 4.
@@
 
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240] of trial 500.
@@
 
-In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240] and (240,280], while incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440].
-Due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+In the middle of the experiment (trial 1000), both congruent and incongruent primes have positive effects in bins (200,240] and (240,280], while incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440].
+Probably due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms early in a trial (compared to 40 ms at the beginning of the experiment) compared to the blank prime condition.
 
@@
 
-@panisWhatShapingRT2016 distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
+@panisWhatShapingRT2016 distinguished between automatic response competition (bottom-up lateral inhibition between response channels), active and global inhibition (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
 While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the response to the target stimulus.
-This global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
-Interestingly, while @panisWhatShapingRT2016 did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the correct response to the target stimulus.
+Thus, it seems that this global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+Interestingly, while @panisWhatShapingRT2016 did not test interactions between prime-target congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which nicely coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
 
@@
 
-In this fourth tutorial, we illustrate how to fit a Bayesian regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+In this fourth tutorial, we illustrate how to fit a Bayesian multi-level regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
@@
 
-To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes on logit-ca relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
@@
 
-Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines in Figure 8).
+Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (blank prime; red striped lines in Figure 8).
 For example, the odds ratio for congruent primes in bin (320,360] in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
@@
 
-In this fifth tutorial we illustrate how to fit a multilevel hazard regression model in the frequentist framework, for the data set used in Tutorial 1a.
+In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data set used in Tutorial 1a.
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
@@
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
-But the price you pay for being able to fit more complex models in a Bayesian framework is computation time.
+But the price you pay for being able to fit more complex random effects models in a Bayesian framework is computation time.
 In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a and 2b took several hours to build.
@@
 
-This main motivation for writing this paper is the observation that event history and SAT analyses remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research, which means the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
 Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
-While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that EHAS/SAT can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
+While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by @holdenDispersionResponseTimes2009, “probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+First, as discussed by @holdenDispersionResponseTimes2009, “probability density [and mass] functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
 Hazard functions are thus more diagnostic than density functions” (p.
@@
 
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (i.e., the response deadline).
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
 This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
@@
 As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
-This orthodox procedure can lead to a sampling bias, however, which results in underestimation of the mean.
+This orthodox procedure leads to underestimation of the true mean.
 By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
@@
 (Allison, 2010).
-This is useful for linking physiological effects with behavioral effeccts when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
 
@@
 
-## 5.4 Limitation(s)
+## 5.4 Limitations
 
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multi-level hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 
-Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
 Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
@@
 However, in certain situations, more than one event of interest might exist.
-For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
+For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before being absorbed into a final "death" state.
+When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize event history analysis to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.

### 2024-10-27 manuscript/manuscript.tex
update ms.

@@
 \hypersetup{
-  pdftitle={Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows},
+  pdftitle={Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows},
   pdfauthor={Sven Panis1 \& Richard Ramsey1},
@@
 
-\title{Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows}
+\title{Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows}
 \author{Sven Panis\textsuperscript{1} \& Richard Ramsey\textsuperscript{1}}
@@
 In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact of theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009).
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development and our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
 Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 
-As a simple illustration, Figure 1 shows the results of several simulated RT datasets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distribution differs (Figure 1, top row).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distribution differs in each case (Figure 1, bottom row).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distribution would lead to a different conclusion in every case.
+As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributios differ (Figure 1, top row).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, bottom row).
+Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
 
@@
 
-\caption{Means versus distributional shapes for six different simulated dataset examples.
+\caption{Means versus distributional shapes for six different simulated data set examples.
+The first second after stimulus onset is divided in ten bins of 100 ms.
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
@@
 And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of waiting time -- to our theory building toolkit.
+As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of time -- to our theory building toolkit.
 
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death (William Matthew Makeham, 1860).
+Indeed, hundreds of years ago, analytical methods were developed to predict time to death (e.g., William Matthew Makeham, 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
@@
 In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of multilevel regression models for RT data in a frequentist framework using the R package lme4.
+In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
@@
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t.
-In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
 As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
@@
 
-\caption{Hazard and survivor functions}\label{fig:plot2}
+\caption{Discrete-time hazard and survivor functions.}\label{fig:plot2}
 \end{figure}
@@
 Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical example that is inspired by real data (Panis et al., 2016), but simplified to make the main point clearer (Figure 3).
+To illustrate, we consider a hypothetical choice RT example that is inspired by real data (Panis \& Schmidt, 2016), but simplified to make the main point clearer (Figure 3).
 In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
@@
 
-\caption{Hazard and conditional accuracy}\label{fig:plot3}
+\caption{Discrete-time hazard and conditional accuracy functions.}\label{fig:plot3}
 \end{figure}
@@
 In contrast to the large-\emph{N} design that typically average across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
-This can be advantageous whenever participants differ systematically in their strategies or in the time-courses of their effects, so that averaging them would lead to misleading data patterns.
+This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
@@
 Descriptive, data science workflow.
-Data wrangling via tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
+We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
 Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018) ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2018) ; Kurz, 202?).
+Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018) ; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2018).
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
@@
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy.
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2018).
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+Therefore, the `person-trial-bin' data set generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
 This data set is used when fitting hazard models (Tutorials 2a and 3a).
@@
 user input: censoring time, and bin width}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data  =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}           \CommentTok{\# create person{-}trial{-}bin dataset}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data  =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}           \CommentTok{\# create person{-}trial{-}bin data set}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}      \CommentTok{\# create life tables without ca(t)}
@@
 
-The person-trial-bin oriented dataset is created by our custom function ptb(), and it has one row for each time bin of each trial that is at risk for event occurrence.
+The person-trial-bin oriented data set is created by our custom function ptb(), and it has one row for each time bin (of each trial) that is at risk for event occurrence.
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
@@
 \midrule
-0.00 & 220.00 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
-40.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-80.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-120.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-160.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-200.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-240.00 & 220.00 & 0.00 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-280.00 & 220.00 & 7.00 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
-320.00 & 213.00 & 13.00 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
-360.00 & 200.00 & 26.00 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
-400.00 & 174.00 & 40.00 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
-440.00 & 134.00 & 48.00 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
-480.00 & 86.00 & 37.00 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
-520.00 & 49.00 & 32.00 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00\\
-560.00 & 17.00 & 9.00 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00\\
-600.00 & 8.00 & 4.00 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00\\
+0 & 220 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
+40 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+80 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+120 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+160 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+200 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+240 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
+280 & 220 & 7 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
+320 & 213 & 13 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
+360 & 200 & 26 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
+400 & 174 & 40 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
+440 & 134 & 48 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
+480 & 86 & 37 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
+520 & 49 & 32 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00\\
+560 & 17 & 9 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00\\
+600 & 8 & 4 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00\\
 \bottomrule
@@
 
-\caption{Estimated discrete-time hazard, survivor, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
+\caption{Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
 \end{figure}
@@
 
-When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
-
-These results suggest that the participant is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower response are fully controlled by the target stimulus.
+When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
+
+These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
@@
 
-In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multi-level regression models to the RT data of the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t).
+Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t) in a selected baseline condition.
 One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-We provide relevant example regression formulas in part D of the supplementary material.
-
-In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial-oriented data set into a person-trial-bin-oriented data set (Allison, 2010).
+Relevant example regression formulas are provided in equations 6 and 7 in part D of the supplementary material.
+
+In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set (Allison, 2010).
 When there is clustering in the data, as in the case of a small-\emph{N} design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models (Allison, 2010).
@@
 In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We select the analysis range (200,600{]} and the cloglog link.
+We decided to select the analysis time window (200,600{]} and the cloglog link.
 The data is prepared as follows.
@@
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
-
-\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
+\CommentTok{\# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)}
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{)}
-
 \CommentTok{\# create factor condition, with "blank" as the reference level}
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
-
-\CommentTok{\# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.}
+\CommentTok{\# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.}
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} 
@@
 \begin{Highlighting}[]
-\FunctionTok{plan}\NormalTok{(multicore)}
-
 \NormalTok{model\_M1 }\OtherTok{\textless{}{-}}
@@
 After selecting the binomial family and the cloglog link, the model formula is specified.
-The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of priming condition and centered trial number.
+The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of prime-target congruency (variable condition) and centered trial number (variable trial\_c).
 Each of these effects is allowed to vary across individuals (variable pid).
@@
 
-So far, we assumed that the effect of our predictors condition and centered trial number are the same in each time bin.
-However, the descriptive plots suggest that the effect of prime-target congruency varies across time bins.
+So far, we assumed that the effect of our predictors prime-target congruecy and centered trial number are the same in each time bin.
+However, the descriptive plots (e.g., Figure 4) suggest that the effect of prime-target congruency varies across time bins.
 Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
-For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable ``condition'') and centered trial number (variable ``trial\_c'').
+For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable condition) and centered trial number (variable trial\_c).
 
@@
 
-Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) we expect nonlinear effects of trial number.
+Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) we expect nonlinear effects of trial number on the hazard of response occurrence.
 We thus relax all three assumptions in model 4.
@@
 
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hzard estimate in the baseline condition (no prime; red striped lines in Figure 6).
+Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240{]} of trial 500.
@@
 
-In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}.
-Due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).
+In the middle of the experiment (trial 1000), both congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}.
+Probably due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms early in a trial (compared to 40 ms at the beginning of the experiment) compared to the blank prime condition.
 
@@
 
-Panis and Schmidt (2016) distinguished between automatic response competition effects (bottom-up lateral inhibition between response channels), active and global inhibition effects (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
+Panis and Schmidt (2016) distinguished between automatic response competition (bottom-up lateral inhibition between response channels), active and global inhibition (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
 While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the response to the target stimulus.
-This global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
-Interestingly, while Panis and Schmidt (2016) did not test interactions between congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
+In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the correct response to the target stimulus.
+Thus, it seems that this global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
+Interestingly, while Panis and Schmidt (2016) did not test interactions between prime-target congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which nicely coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
 
@@
 
-To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
+To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes on logit-ca relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
 
@@
 
-Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines in Figure 8).
+Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (blank prime; red striped lines in Figure 8).
 For example, the odds ratio for congruent primes in bin (320,360{]} in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
@@
 
-This main motivation for writing this paper is the observation that event history and SAT analyses remains under-used in psychological research, which means the field of research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research, which means the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
 Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
-While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHAS/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
+While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density {[}and mass{]} functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
 Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
@@
 (Allison, 2010).
-This is useful for linking physiological effects with behavioral effeccts when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
 
@@
 
-\subsection{5.4 Limitation(s)}\label{limitations}
-
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
-
-Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution.
+\subsection{5.4 Limitations}\label{limitations}
+
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multi-level hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+
+Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
 Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
@@
 However, in certain situations, more than one event of interest might exist.
-For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state models, which generalize event history analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
+For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before being absorbed into a final ``death'' state.
+When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize event history analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
@@
 
+\bibitem[\citeproctext]{ref-wolkersdorferTemporalDynamicsSequential2020}
+Wolkersdorfer, M.
+P., Panis, S., \& Schmidt, T.
+(2020).
+Temporal dynamics of sequential motor activation in a dual-prime paradigm: {Insights} from conditional accuracy and hazard functions.
+\emph{Attention, Perception, \& Psychophysics}, \emph{82}(5), 2581--2602.
+\url{https://doi.org/10.3758/s13414-020-02010-5}
+
 \end{CSLReferences}
@@
 
-\subsection{A. Definitions of discrete-time hazard, surivor, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-and-conditional-accuracy-functions}
+\subsection{A.
+Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-probability-mass-and-conditional-accuracy-functions}
 
@@
 After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Discrete-time EHA focuses on the discrete-time hazard function
+Because waiting times can only increase, discrete-time EHA focuses on the discrete-time hazard function
 
@@
 
-The discrete-time hazard function of event occurrence gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
@@
 
-\noindent which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
-Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
+Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
 The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.
@@
 \item
-  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for aggregated data (1 independent variable)
+  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for data aggregated across participants (1 independent variable)
 \end{itemize}
@@
 
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}
+\noindent cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\beta\)\textsubscript{0}ONE + \(\beta\)\textsubscript{1}(TIME-9) + \(\beta\)\textsubscript{2}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{3}X\textsubscript{1} + \(\beta\)\textsubscript{4}X\textsubscript{2} + \(\beta\)\textsubscript{5}X\textsubscript{2}(TIME-9){]} \hfill  (6)
 
 The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
+The first set of terms within brackets, the parameters \(\beta\)\textsubscript{0} to \(\beta\)\textsubscript{2} multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
+The second set of terms (the beta parameters \(\beta\)\textsubscript{3} to \(\beta\)\textsubscript{5}) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
@@
 However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth.
-To interpret the effects of a predictor,its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+To interpret the effects of a predictor, its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
 When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.
@@
 
-cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}
+\noindent cloglog{[}h(t){]} = {[}\(\beta\)\textsubscript{0}D1 + \(\beta\)\textsubscript{1}D2 + \(\beta\)\textsubscript{2}D3 + \(\beta\)\textsubscript{3}D4 + \(\beta\)\textsubscript{4}D5 + \(\beta\)\textsubscript{5}D6{]} + {[}\(\beta\)\textsubscript{6}X\textsubscript{1}{]} \hfill  (7)
 

### 2024-10-27 manuscript/manuscript.Rmd
add and delete REFs

@@
 
-For a comprehensive background context to event history analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003].
+For a comprehensive background context to event history analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
@@
 
-Based on previous work [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022] we expect nonlinear effects of trial number on the hazard of response occurrence.
+Based on previous work [e.g., @panisHowCanWe2020] we expect nonlinear effects of trial number on the hazard of response occurrence.
 We thus relax all three assumptions in model 4.
 We add a squared term for the continuous predictor centered trial number -- I(trial_c^2) -- and include interaction terms.
+For example, how the effect of congruent primes changes across time bins within a trial might change across the trials within an experiment.
 

### 2024-10-27 manuscript/manuscript.tex
add and delete REFs

@@
 
-For a comprehensive background context to event history analysis, we recommend several excellent textbooks (Singer \& Willett, 2003).
+For a comprehensive background context to event history analysis, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
 Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
@@
 
-Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) we expect nonlinear effects of trial number on the hazard of response occurrence.
+Based on previous work (e.g., Panis, 2020) we expect nonlinear effects of trial number on the hazard of response occurrence.
 We thus relax all three assumptions in model 4.
 We add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.
+For example, how the effect of congruent primes changes across time bins within a trial might change across the trials within an experiment.
 

### 2024-10-28 manuscript/manuscript.Rmd
update ms.

@@
 
-To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trials 500, 1000, and 1500.
 
-(ref:descr-plot7-caption) Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.
+(ref:descr-plot7-caption) Model-based hazard functions for each prime type for participant 6 in trials 500, 1000, and 1500.
 
@@
 
-To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trials 500, 1000, and 1500.
 
-(ref:descr-plot9-caption) Model-based ca(t) functions for participant 6 in trial 500, 1000, and 1500.
+(ref:descr-plot9-caption) Model-based ca(t) functions for each prime type for participant 6 in trials 500, 1000, and 1500.
 

### 2024-10-28 Tutorial_1a.Rmd

@@
 
-This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
+This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
 
 Content:
+
 1. Load libraries

### 2024-10-28 manuscript/SupplementaryMaterial.Rmd

--- 2024-10-21 manuscript/SupplementaryMaterial.Rmd
@@
 
-## A. Definitions of discrete-time hazard, surivor, and conditional accuracy functions
+## A. Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions
 
@@
 
-The discrete-time hazard function of event occurrence gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins.
+The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
@@
 
-\noindent which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
-Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
+Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
 The hazard rate function gives you the instantaneous *rate* of event occurrence at time point t, given that the event has not occurred yet.
@@
 * join_lt_ca_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
-* extract_median(df) : estimate quantiles S(t).50 (1 independent variable)
-* extract_median_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
+* extract_median(df) : estimate quantiles S(t)~.50~ (1 independent variable)
+* extract_median_2IV(df) : estimate quantiles S(t)~.50~ (2 independent variables)
 * plot_eha(df,subj,haz_yaxis) : create plots of the discrete-time functions (1 independent variable)
 * plot_eha_2IV(df,subj,haz_yaxis) : create plots of the discrete-time functions (2 independent variables)
-* plot_eha_agg(df,subj,haz_yaxis) : create 1 plot for aggregated data (1 independent variable)
+* plot_eha_agg(df,subj,haz_yaxis) : create 1 plot for data aggregated across participants (1 independent variable)
 
@@
 
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~ONE + $\alpha$~2~(TIME-9) + $\alpha$~3~(TIME-9)$^2$] +                                    [$\beta$~1~X~1~ + $\beta$~2~X~2~ + $\beta$~3~X~2~(TIME-9)]
+\noindent cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\beta$~0~ONE + $\beta$~1~(TIME-9) + $\beta$~2~(TIME-9)$^2$] +                                    [$\beta$~3~X~1~ + $\beta$~4~X~2~ + $\beta$~5~X~2~(TIME-9)]  \hfill  (6)
 
 The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the alpha parameters multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
-The second set of terms (the beta parameters) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
+The first set of terms within brackets, the parameters $\beta$~0~ to $\beta$~2~ multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
+The second set of terms (the beta parameters $\beta$~3~ to $\beta$~5~) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
@@
 However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
-To interpret the effects of a predictor,its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
+To interpret the effects of a predictor, its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
 When using the logit link, exponentiating a $\beta$ parameter results in an odds ratio.
@@
 
-cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\alpha$~1~D1 + $\alpha$~2~D2 + $\alpha$~3~D3 + $\alpha$~4~D4 + $\alpha$~5~D5 + $\alpha$~6~D6] + [$\beta$~1~X~1~]
+\noindent cloglog[h(t)] =  [$\beta$~0~D1 + $\beta$~1~D2 + $\beta$~2~D3 + $\beta$~3~D4 + $\beta$~4~D5 + $\beta$~5~D6] + [$\beta$~6~X~1~]  \hfill  (7)
 

### 2024-10-28 manuscript/manuscript.tex

@@
 
-To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trials 500, 1000, and 1500.
 
@@
 
-\caption{Model-based hazard functions for participant 6 in trial 500, 1000, and 1500.}\label{fig:plot-hazard-subject6}
+\caption{Model-based hazard functions for each prime type for participant 6 in trials 500, 1000, and 1500.}\label{fig:plot-hazard-subject6}
 \end{figure}
@@
 
-To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trial 500, 1000, and 1500.
+To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trials 500, 1000, and 1500.
 
@@
 
-\caption{Model-based ca(t) functions for participant 6 in trial 500, 1000, and 1500.}\label{fig:plot-ca-subject6}
+\caption{Model-based ca(t) functions for each prime type for participant 6 in trials 500, 1000, and 1500.}\label{fig:plot-ca-subject6}
 \end{figure}

### 2024-10-30 manuscript/SupplementaryMaterial.Rmd
Added a section to Supps

@@
 A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
+
+## F. Advantages of hazard analysis
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by @holdenDispersionResponseTimes2009, “probability density [and mass] functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions” (p.
+331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
+
+Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if h~A~(t) > h~B~(t) for all t, then both hazard functions are said to show a complete ordering.
+Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
+
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
+As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
+This orthodox procedure leads to underestimation of the true mean.
+By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
+
+Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].

### 2024-10-30 manuscript/manuscript.Rmd
edits to the manuscript

@@
 We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
-We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
+Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
@@
 footnotelist      : no
-classoption       : "man"
+classoption       : "man, donotrepeattitle"
 appendix          : SupplementaryMaterial.Rmd
@@
 
-In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
-Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+In experimental psychology, it is standard practice to analyse reaction times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
+Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 
 As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributios differ (Figure 1, top row).
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, top row).
 In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, bottom row).
@@
 In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms.
-And in the Example 6, the effect reverses around between 500 and 600 ms.
-What kind of theory or set of theories could account for such effects?
+And in the Example 6, the effect reverses between 500 and 600 ms.
+What kind of theory or theories could account for such effects?
 Are there new auxiliary assumptions that theories need to adopt?
-And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of time -- to our theory building toolkit.
+And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
+As we show later using published examples, for many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
 
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death [e.g., @williammatthewmakehamLawMortalityConstruction1860].
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @williammatthewmakehamLawMortalityConstruction1860].
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
-Here, in the paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
+Here, in the current paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
 
@@
 We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010], and our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you should do them.
+However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
+Indeed, our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
 
-We then provide six different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
@@
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
+Finally, the manuscript itself is written in R using the papaja() package [[insert REF for papaja() here]], which makes it computationally reproducible, in terms of the undeerlying data and figures.
 
@@
 In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
+
 In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
@@
 
+In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
+We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
+[[more to come here, once we have written the tutorial]].
+
 In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
@@
 For a comprehensive background context to event history analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
-Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
-Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
-Instead, we want to provide an intuition to how event history analysis works in general as well as in the context of experimental psychology.
-As such, we only supply regression equations in the supplementary material (part D) and then refer to them in the text whenever relevant.
+Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
+Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
+Instead, we want to provide an intuition regarding how event history analysis works in general, as well as in the context of experimental psychology.
+As such, we only supply regression equations in supplementary material and then refer to them in the text whenever relevant.
 
@@
 
-To apply event history analysis (EHA), a.k.a. hazard analysis, one must be able to:
+To apply event history analysis (EHA), one must be able to:
 
@@
 The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A).
-The discrete-time hazard function gives you for each bin the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
@@
 In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
-As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
 
-<!-- The survivor function can help to qualify or provide context to the interpretation of the hazard function.
+The survivor function can help to qualify or provide context to the interpretation of the hazard function.
 For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
 If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500], then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.-->
+Therefore, the error bars in later parts of the hazard function would typically be wider and less precise compared to earlier parts.
 
@@
 
-## 2.2 Event history analysis in the context of experimental psychology
+## 2.2 Benefits of event history analysis
 
-### 2.2.1 A worked example
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
+Instead, we briefly summarise list the benefits below, and refer the reader to section F of Supplementary Materials for more detailed coverage of the benefits.
+A summary of the benefits are as follows:
+
+1. Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
+
+2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences than mean-average comaprisons may conceal (Townsend, 1990).
+
+3.
+EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
+Trials with longer RTs are not discarded, but instead controbute to the risk set in each time bin.
+
+4.
+Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
+
+5.
+EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
+
+## 2.3 Event history analysis in the context of experimental psychology
+
+To make EHA more relevant to researchers studying cognitive psychology and cognitive neuroscience, in this section we provide a relevant worked example and consider implications that are relevant to that domain of research.
+
+### 2.3.1 A worked example
 
@@
 For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 We refer to this extended analysis for choice RT data as EHA/SAT.
-As explained in part A of the supplementary material, the probability mass function provides a context for interpreting the conditional accuracy function.
 
@@
 The prime can then be congruent or incongruent with the target.
+
 Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
-And then for longer waiting times, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
-This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-And such a conclusion is not supported when the effects are explored over a timeline.
+And then for longer waiting times, the response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+
+This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT across trials.
+Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
+For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials.
+But such a conclusion is not supported when the effects are explored over a timeline.
 Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
@@
 
-Unlocking the temporal states of cognitive processes can be revealing in and of itself for theory development and the understanding of basic psychological processes.
+Unlocking the temporal states of cognitive processes can be revealing for theory development and the understanding of basic psychological processes.
 Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
@@
 
-### 2.2.2 Implications for designing experiments
+### 2.3.2 Implications for designing experiments
 
@@
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
-
-We used `r my_r_citation$r` for all reported analyses.
-The content of the tutorials is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
-
-`r my_r_citation$pkgs`
-
-<!-- This creates the footnote -->
 
@@
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
+
+## 3.3 Implementation
+
+We used `r my_r_citation$r` for all reported analyses.
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
+
+`r my_r_citation$pkgs`
+
+<!-- This creates the footnote -->
 
@@
 
-In general, it is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was always guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone may be guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
@@
 
-When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
-
 These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
+When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 5.
-
-In the no-mask condition (column 1 in Figure 5), we observe a positive compatibility effect in the hazard and ca(t) functions, as congruent primes temporarily generate higher values for hazard and conditional accuracy compared to incongruent primes.
-However, when a (relevant, irrelevant, or lines) mask is present (columns 2-4), there is a negative compatibility effect in the hazard and conditional accuracy functions, as congruent primes temporarily generate *lower* values for hazard and conditional accuracy compared to incongruent primes.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 as part of Tutorial 1b for those that are interested.
 
@@
 ```{r plot-2ivs, fig.cap = "(ref:descr-plot5-caption)", out.width='95%'}
-knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt.png")
+# knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt.png")
 ```
@@
 Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
+For simplicity and ease of presentation, we only tabulate data for a subset of the design in the main text (trial 500 for congruent and incongruent conditions).
+For the full table, see Supplementary materials or Tutorual X.
 
@@
 ```{r int-table}
-apa_table(
-  int_tab,
-  caption = "(ref:int-table-caption)",
-  note = "(ref:int-table-note-caption)",
-  placement = "H",
-  font_size = "scriptsize",
-  longtable = TRUE,
-  repeat_header = TRUE,
-  format.args = list(
-       digits = c(0, 2,2,2,2,2,5),
-        decimal.mark = ".", big.mark = "")
-)
+# apa_table(
+#   int_tab,
+#   caption = "(ref:int-table-caption)",
+#   note = "(ref:int-table-note-caption)",
+#   placement = "H",
+#   font_size = "scriptsize",
+#   longtable = TRUE,
+#   repeat_header = TRUE,
+#   format.args = list(
+#        digits = c(0, 2,2,2,2,2,5),
+#         decimal.mark = ".", big.mark = "")
+# )
 ```
 
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
+Based on Figure 6 and Supplementary Table XX, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240] of trial 500.
@@
 These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period_9), but also on the across-trial/within-experiment time scale (variable trial_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of separate cognitive processes.
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, suggests the involvement of separate cognitive processes.
 
@@
 
-Figure 8 shows one point (mean) and three highest posteriror density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 8 shows one point (mean) and three highest posterior density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
+Supplementary Table XX shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
@@
 ```{r int-ca-table}
-apa_table(
-  int_ca_tab,
-  caption = "(ref:int-ca-table-caption)",
-  note = "(ref:int-ca-table-note-caption)",
-  placement = "H",
-  font_size = "scriptsize",
-  longtable = TRUE,
-  repeat_header = TRUE,
-  format.args = list(
-       digits = c(0, 2,2,2,2,2,13),
-        decimal.mark = ".", big.mark = "")
-)
+# apa_table(
+#   int_ca_tab,
+#   caption = "(ref:int-ca-table-caption)",
+#   note = "(ref:int-ca-table-note-caption)",
+#   placement = "H",
+#   font_size = "scriptsize",
+#   longtable = TRUE,
+#   repeat_header = TRUE,
+#   format.args = list(
+#        digits = c(0, 2,2,2,2,2,13),
+#         decimal.mark = ".", big.mark = "")
+# )
 ```
 
-Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (blank prime; red striped lines in Figure 8).
+Based on Figure 8 and Supplementary Table XX, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (blank prime; red dashed lines in Figure 8).
 For example, the odds ratio for congruent primes in bin (320,360] in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
@@
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar,, which makes sense given the close similarity in model structure.
 However, the random effects structure of model M3_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
@@
 
+## 4.4 Tutorial 4: Planning
+
+This needs adding... RR to do this.
+
 # 5. Discussion
 
-This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research, which means the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research.
+As a consequence, the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to mean-average comparisons.
 While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss issues relating to model complexity versus interpretability, individual differences, limitations of the approach, and future extensions.
+In the following, we discuss issues relating to model complexity and interpretability, individual differences, as well as limitations of the approach and future extensions.
 
-## 5.1 Advantages of hazard analysis
+## 5.1 What are the main use-cases of EHA for understanding cognition and brain function?
 
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by @holdenDispersionResponseTimes2009, “probability density [and mass] functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions” (p.
-331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
-Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
+For those researchers, like ourselves, who are primarily interested in understanding human cognitive and brain systems, we consider two broadly-defined, main use-cases of EHA.
+First, as we hope to have made clear by this point, EHA is one way to investigating a "temporal states" approach to cognitive processes.
+EHA provides one way to uncover when cognitive states may start and stop, as well as what they may be tied to or interact with.
+Therefore, if your research questions concern **when** and **for how long** psychological states occur, our EHA tutorials could be useful tools for you to use.
 
-Second, because RT distributions may differ from one another in multiple ways, @townsendTruthConsequencesOrdinal1990 developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if h~A~(t) > h~B~(t) for all t, then both hazard functions are said to show a complete ordering.
-Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
-For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.
-
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
-This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
-In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
-As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
-This orthodox procedure leads to underestimation of the true mean.
-By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
-
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
-
-Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+Second, even if you are not primarily interested in studying the temporal states of cognition, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on mean-average comparisons.
+Given that distinctly different inferences can be made from the same data based on whether one computes a mean-average across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement mean-average comparisons with EHA.
+One could envisage scenarios where the implicit assumption of an effect manifesting across all of the time bins measured would not be supported by EHA.
+Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain aspects of time.
 
@@
 
-Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions. 
-For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period_9), and the passage of time on the across-trial or within-experiment time scale (variable trial_c).
-However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
-From a theoretical perspective, adding more than 1 time scale -- and their interactions -- is important to capture plasticity and other learning effects (e.g., proactive control) that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
-From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental predictor variables, because adding time scales quickly increases model complexity.
+EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions. 
+For example, model M4 contains two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
+From a theoretical perspective, adding more than 1 time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
+From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
+As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
+This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices.
 
@@
 Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
-The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with outlying behavioral patterns.
-One might then exclude the outlying participants before fitting a hazard model.
+The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
+One might then exclude the smaller sub-group of participants before fitting a hazard model or consider the possibility that different cognitive processes may be at play during task performance across the different sub-groups.
 
-Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021].
+Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021], which is a from of Small-N approach (Smith & Little, 2018).
 This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
@@
 
-RT and accuracy distributions are a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
+Estimating the temporal distributions of RT and accuracy provide a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
 Statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.

### 2024-10-30 manuscript/manuscript.tex
edits to the manuscript

@@
 \documentclass[
-  man,floatsintext]{apa6}
+  man, donotrepeattitle,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 
-In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development and our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
-Such information is useful not only for interpretation of the effects, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+In experimental psychology, it is standard practice to analyse reaction times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
+Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 
 As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributios differ (Figure 1, top row).
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, top row).
 In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, bottom row).
@@
 In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms.
-And in the Example 6, the effect reverses around between 500 and 600 ms.
-What kind of theory or set of theories could account for such effects?
+And in the Example 6, the effect reverses between 500 and 600 ms.
+What kind of theory or theories could account for such effects?
 Are there new auxiliary assumptions that theories need to adopt?
-And are there new experiments that need to be run to test the novel predictions that follow from these analyses?
-As we show later using concrete examples from past experimental data, for many psychological questions this ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes as well as adding a relatively under-used dimension -- the passage of time -- to our theory building toolkit.
+And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
+As we show later using published examples, for many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
 
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict time to death (e.g., William Matthew Makeham, 1860).
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., William Matthew Makeham, 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
-Here, in the paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
+Here, in the current paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
 
@@
 We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003), and our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you should do them.
-
-We then provide six different tutorials, each of which is written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003).
+Indeed, our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
+
+We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
@@
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
+Finally, the manuscript itself is written in R using the papaja() package {[}{[}insert REF for papaja() here{]}{]}, which makes it computationally reproducible, in terms of the undeerlying data and figures.
 
@@
 
+In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
+We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
+{[}{[}more to come here, once we have written the tutorial{]}{]}.
+
 In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
@@
 For a comprehensive background context to event history analysis, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
-Likewise, for general introduction to understanding regression equations, we recommend several introductory level textbooks (REFs).
-Our focus here is not on providing a detailed account of the underlying regression equations, since this topics has been comprehensively covered many times before.
-Instead, we want to provide an intuition to how event history analysis works in general as well as in the context of experimental psychology.
-As such, we only supply regression equations in the supplementary material (part D) and then refer to them in the text whenever relevant.
+Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
+Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
+Instead, we want to provide an intuition regarding how event history analysis works in general, as well as in the context of experimental psychology.
+As such, we only supply regression equations in supplementary material and then refer to them in the text whenever relevant.
 
@@
 In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
-As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
+
+The survivor function can help to qualify or provide context to the interpretation of the hazard function.
+For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
+If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500{]}, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
+Therefore, the error bars in later parts of the hazard function would typically be wider and less precise compared to earlier parts.
 
@@
 
-\subsection{2.2 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
-
-\subsubsection{2.2.1 A worked example}\label{a-worked-example}
+\subsection{2.2 Benefits of event history analysis}\label{benefits-of-event-history-analysis}
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
+Instead, we briefly summarise list the benefits below, and refer the reader to section F of Supplementary Materials for more detailed coverage of the benefits.
+A summary of the benefits are as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
+\item
+  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences than mean-average comaprisons may conceal (Townsend, 1990).
+\item
+  EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
+Trials with longer RTs are not discarded, but instead controbute to the risk set in each time bin.
+\item
+  Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+\item
+  EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
+\end{enumerate}
+
+\subsection{2.3 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
+
+To make EHA more relevant to researchers studying cognitive psychology and cognitive neuroscience, in this section we provide a relevant worked example and consider implications that are relevant to that domain of research.
+
+\subsubsection{2.3.1 A worked example}\label{a-worked-example}
 
@@
 For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 We refer to this extended analysis for choice RT data as EHA/SAT.
-As explained in part A of the supplementary material, the probability mass function provides a context for interpreting the conditional accuracy function.
 
@@
 The prime can then be congruent or incongruent with the target.
+
 Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-And then for longer waiting times, response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
-This is interesting because mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-And such a conclusion is not supported when the effects are explored over a timeline.
+And then for longer waiting times, the response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+
+This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT across trials.
+Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
+For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials.
+But such a conclusion is not supported when the effects are explored over a timeline.
 Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
@@
 
-\subsubsection{2.2.2 Implications for designing experiments}\label{implications-for-designing-experiments}
+\subsubsection{2.3.2 Implications for designing experiments}\label{implications-for-designing-experiments}
 
@@
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
-
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-The content of the tutorials is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
 
@@
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
+
+\subsection{3.3 Implementation}\label{implementation}
+
+We used R (Version 4.4.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
 
@@
 
-When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
-
 These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
+When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 5.
-
-In the no-mask condition (column 1 in Figure 5), we observe a positive compatibility effect in the hazard and ca(t) functions, as congruent primes temporarily generate higher values for hazard and conditional accuracy compared to incongruent primes.
-However, when a (relevant, irrelevant, or lines) mask is present (columns 2-4), there is a negative compatibility effect in the hazard and conditional accuracy functions, as congruent primes temporarily generate \emph{lower} values for hazard and conditional accuracy compared to incongruent primes.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.95\linewidth,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt}
-
-}
-
-\caption{Sample-based discrete-time hazard, survivor, conditional accuracy, and probability mass functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).}\label{fig:plot-2ivs}
-\end{figure}
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 as part of Tutorial 1b for those that are interested.
 
@@
 Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
-
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{TableNotes}[para]
-\normalsize{\textit{Note.} c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.}
-\end{TableNotes}
-
-\scriptsize{
-
-\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:int-table}Point and 95\% highest posterior density interval estimates, and hazard ratios.}\\
-\toprule
-bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:int-table} continued}}\\
-\toprule
-bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
-\midrule
-\endhead
-240 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82080\\
-280 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03199\\
-320 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45383\\
-360 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91397\\
-400 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.70802\\
-440 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.63522\\
-480 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62035\\
-520 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.61980\\
-560 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.59543\\
-600 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.51694\\
-240 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63183\\
-280 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.12611\\
-320 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90342\\
-360 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99277\\
-400 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68359\\
-440 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58403\\
-480 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58190\\
-520 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.63546\\
-560 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71487\\
-600 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.77863\\
-240 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02227\\
-280 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.65584\\
-320 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.51502\\
-360 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68077\\
-400 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.39703\\
-440 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28245\\
-480 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23037\\
-520 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20247\\
-560 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18021\\
-600 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15267\\
-240 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10461\\
-280 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72468\\
-320 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58233\\
-360 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.33902\\
-400 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.29661\\
-440 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.33991\\
-480 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44474\\
-520 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.57904\\
-560 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65388\\
-600 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.55820\\
-240 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68463\\
-280 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88377\\
-320 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78490\\
-360 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.39866\\
-400 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.32935\\
-440 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.38574\\
-480 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.55825\\
-520 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87013\\
-560 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27313\\
-600 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52411\\
-240 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07329\\
-280 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.93821\\
-320 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.71847\\
-360 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.34502\\
-400 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28360\\
-440 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.34778\\
-480 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55459\\
-520 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00234\\
-560 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.78962\\
-600 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75123\\
-\bottomrule
-\addlinespace
-\insertTableNotes
-\end{longtable}
-
-}
-
-\end{ThreePartTable}
-\end{center}
-
-Based on Figure 6 and Table 4, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
+For simplicity and ease of presentation, we only tabulate data for a subset of the design in the main text (trial 500 for congruent and incongruent conditions).
+For the full table, see Supplementary materials or Tutorual X.
+
+Based on Figure 6 and Supplementary Table XX, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
 For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240{]} of trial 500.
@@
 These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period\_9), but also on the across-trial/within-experiment time scale (variable trial\_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, strongly suggests the involvement of separate cognitive processes.
+The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, suggests the involvement of separate cognitive processes.
 
@@
 
-Figure 8 shows one point (mean) and three highest posteriror density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+Figure 8 shows one point (mean) and three highest posterior density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
 
@@
 
-Table 5 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
+Supplementary Table XX shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
 Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
 
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{TableNotes}[para]
-\normalsize{\textit{Note.} c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.}
-\end{TableNotes}
-
-\scriptsize{
-
-\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:int-ca-table}Point and 95\% highest posterior density interval estimates, and odds ratios.}\\
-\toprule
-bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{odds ratio}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:int-ca-table} continued}}\\
-\toprule
-bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{odds ratio}\\
-\midrule
-\endhead
-240 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24618458.6117376089096\\
-280 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4846.0911175084821\\
-320 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.7913398591541\\
-360 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.8918515266144\\
-400 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.1907589464532\\
-440 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.3094780953354\\
-480 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.4104953301397\\
-520 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.3973471906322\\
-560 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.1244195286526\\
-600 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.0000573483764\\
-240 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1899836.0176862408407\\
-280 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1961.8278183493471\\
-320 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.8743225515544\\
-360 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.5713234449885\\
-400 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.1378140505531\\
-440 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.0636745027736\\
-480 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.0982112115060\\
-520 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.0008224865296\\
-560 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.0000002080447\\
-600 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.0000000000005\\
-240 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10001085.3946626689285\\
-280 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5124.4351045810508\\
-320 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.1208056719069\\
-360 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.8233561607888\\
-400 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.5848218057173\\
-440 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.8181575226781\\
-480 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.0826145785621\\
-520 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.0008862056864\\
-560 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.0000003362909\\
-600 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.0000000000015\\
-240 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.0000000000730\\
-280 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.0000261643171\\
-320 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.0246001196011\\
-360 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.3780876098423\\
-400 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.5916606671310\\
-440 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.5871858652992\\
-480 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.3019651321687\\
-520 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.0443659277399\\
-560 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3282435.9279020344839\\
-600 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46319712352328.7578125000000\\
-240 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.0000000003243\\
-280 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.0000632158160\\
-320 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.0419655563481\\
-360 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.5909004105316\\
-400 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.0992267336787\\
-440 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.6827111411806\\
-480 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.2031868705690\\
-520 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3307.4439707159477\\
-560 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164758701.8390493392944\\
-600 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10165856639901592.0000000000000\\
-240 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.0000000012362\\
-280 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.0001189822174\\
-320 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.0575421441866\\
-360 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.8709638702927\\
-400 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.5699339941536\\
-440 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.2076572160001\\
-480 & i1500 & 5.52 & -13.87 & 26.57 & 0.95 & 249.5050127299390\\
-520 & i1500 & 12.67 & -31.46 & 58.41 & 0.95 & 318500.3986836019321\\
-560 & i1500 & 25.50 & -53.08 & 115.21 & 0.95 & 119299568240.9411773681641\\
-600 & i1500 & 45.85 & -86.60 & 200.06 & 0.95 & 81670189671651033088.0000000000000\\
-\bottomrule
-\addlinespace
-\insertTableNotes
-\end{longtable}
-
-}
-
-\end{ThreePartTable}
-\end{center}
-
-Based on Figure 8 and Table 5, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (blank prime; red striped lines in Figure 8).
+Based on Figure 8 and Supplementary Table XX, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (blank prime; red dashed lines in Figure 8).
 For example, the odds ratio for congruent primes in bin (320,360{]} in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
@@
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar,, which makes sense given the close similarity in model structure.
 However, the random effects structure of model M3\_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
@@
 
+\subsection{4.4 Tutorial 4: Planning}\label{tutorial-4-planning}
+
+This needs adding\ldots{} RR to do this.
+
 \section{5. Discussion}\label{discussion}
 
-This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research, which means the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
+This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research.
+As a consequence, the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is the increase in analytical complexity compared to mean-average comparisons.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to mean-average comparisons.
 While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss issues relating to model complexity versus interpretability, individual differences, limitations of the approach, and future extensions.
-
-\subsection{5.1 Advantages of hazard analysis}\label{advantages-of-hazard-analysis}
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density {[}and mass{]} functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
-
-Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if h\textsubscript{A}(t) \textgreater{} h\textsubscript{B}(t) for all t, then both hazard functions are said to show a complete ordering.
-Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
-For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
-
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
-This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
-In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
-As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
-This orthodox procedure leads to underestimation of the true mean.
-By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
-
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
-
-Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
+In the following, we discuss issues relating to model complexity and interpretability, individual differences, as well as limitations of the approach and future extensions.
+
+\subsection{5.1 What are the main use-cases of EHA for understanding cognition and brain function?}\label{what-are-the-main-use-cases-of-eha-for-understanding-cognition-and-brain-function}
+
+For those researchers, like ourselves, who are primarily interested in understanding human cognitive and brain systems, we consider two broadly-defined, main use-cases of EHA.
+First, as we hope to have made clear by this point, EHA is one way to investigating a ``temporal states'' approach to cognitive processes.
+EHA provides one way to uncover when cognitive states may start and stop, as well as what they may be tied to or interact with.
+Therefore, if your research questions concern \textbf{when} and \textbf{for how long} psychological states occur, our EHA tutorials could be useful tools for you to use.
+
+Second, even if you are not primarily interested in studying the temporal states of cognition, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on mean-average comparisons.
+Given that distinctly different inferences can be made from the same data based on whether one computes a mean-average across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement mean-average comparisons with EHA.
+One could envisage scenarios where the implicit assumption of an effect manifesting across all of the time bins measured would not be supported by EHA.
+Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain aspects of time.
 
@@
 
-Models for discrete-time h(t) and ca(t) can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions.
-For example, model M4 contains two time scales as covariates: the passage of time on the across-bin or within-trial time scale (variable period\_9), and the passage of time on the across-trial or within-experiment time scale (variable trial\_c).
-However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (across-bin or within-trial, across-trial or within-block, across-block or within-session, and across-session or within-experiment).
-From a theoretical perspective, adding more than 1 time scale -- and their interactions -- is important to capture plasticity and other learning effects (e.g., proactive control) that play out on such longer time scales (across-trials, across-blocks, across-sessions), and that are probably present in each experiment in general.
-From a practical perspective, therefore, it might be interesting for interpretational purposes to limit the number of experimental predictor variables, because adding time scales quickly increases model complexity.
+EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions.
+For example, model M4 contains two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
+From a theoretical perspective, adding more than 1 time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
+From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
+As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
+This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices.
 
@@
 Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
-The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with outlying behavioral patterns.
-One might then exclude the outlying participants before fitting a hazard model.
-
-Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021).
+The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
+One might then exclude the smaller sub-group of participants before fitting a hazard model or consider the possibility that different cognitive processes may be at play during task performance across the different sub-groups.
+
+Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021), which is a from of Small-N approach (Smith \& Little, 2018).
 This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
@@
 \bibitem[\citeproctext]{ref-R-forcats}
-Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}
+Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://CRAN.R-project.org/package=forcats}
 
 \bibitem[\citeproctext]{ref-R-stringr}
-Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}
+Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://CRAN.R-project.org/package=stringr}
 
@@
 \bibitem[\citeproctext]{ref-R-dplyr}
-Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
+Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}
 
 \bibitem[\citeproctext]{ref-R-purrr}
-Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
+Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
 
 \bibitem[\citeproctext]{ref-R-readr}
-Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
+Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
 
 \bibitem[\citeproctext]{ref-R-tidyr}
-Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
+Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
 
@@
 
+\subsection{F. Advantages of hazard analysis}\label{f.-advantages-of-hazard-analysis}
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density {[}and mass{]} functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
+Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
+
+Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
+For example, if h\textsubscript{A}(t) \textgreater{} h\textsubscript{B}(t) for all t, then both hazard functions are said to show a complete ordering.
+Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
+
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
+This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
+In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
+As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
+This orthodox procedure leads to underestimation of the true mean.
+By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+
+Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer et al., 1988).
+
+Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
+
 \end{document}

### 2024-10-30 notes/manuscript_notes.Rmd
update of notes.
notes/feedback on the manuscript

@@
 
--> still to do.
+-> I added a content list at the start of each Tutorial_x.Rmd file.
 
@@
         decimal.mark = ".", big.mark = ""))
+
+# October 29, 2024 #
+
+## minor comments ##
+
+- I removed the repetition of the title on the intro page. This always annoys me!
+
+- Minor point. Response time vs reaction time? Line 1 of the intro, we say response. 
+But don't we want to say reaction time, as most people use that term?
+
+- Makeham reference. Shouldn't it just be Makeham, 1860, rather than first and middle names?
+
+- I would explicitly insert a reference in the main text for papaja().
+See the citation subsection of this webpage: https://github.com/crsh/papaja.
+And see page 6 in the manuscript for the text that I added.
+I feel it is worth highlighting / advertising since the developers can get more recognition and maybe others will start using it too.
+
+- I would remove reference to specific equations throughout the manuscript.
+Instead, I would just say it once, like you do at the start of section 2.
+e.g., we only supply regression equations in supps.
+then don't mention it again because it disrupts the flow in my view.
+Folks who want more can easily flick to the supps and read all they want.
+
+- section 2.1.
+we use an abbreviation for EHA.
+and we also define it as hazard.
+Since we have already given the other terms for EHA, let's just use EHA.
+I hope that makes sense.
+And shouldn't we abbreviate to EHA the first time we mention it, rather than in section 2.1?
+
+- section 2.1. You say...
+
+>Continuous RT data is treated here as interval-censored data.
+
+But in the context of this paragraph, this makes no sense and just seems tacked on and disconnected. Can we explain or remove?
+
+- Also in section 2.1, you say in the middle of a paragraph...
+
+>This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
+
+Two things here. 1) At this stage in proceedings, I'm not even sure we need to give the benefits of using a hazard, but instead, just give the definition. So we might just remove it.
+2) If we do include it, then this sentence does not communicate how or why it is more diagnostic, just that it is more diagnostic.
+And we provide no support for this claim.
+Later on in the discussion (I think) there is more detail regarding the greater diagnosticity.
+But here, we get caught between two posts and don't provide sufficient detail.
+So, I would just remove it from here for now, to make things shorter and simpler, at least for the time being.
+
+- Also in section 2.1, you say ...
+
+>As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.
+
+Again, I'm not sure we can leave people hanging like this.
+Instead, I would maybe wrap some of this contextualisation into the figure legend of Figure 2.
+And delete this sentence.
+Or include 1 or 2 sentences that explain how the survivor can help contextualise the hazard, but make it clear and relevant.
+
+[[As I went back to the manuscript, I saw an old paragrpah and re-inlcuded it, as I think it is relevant.
+e.g., if we make the hazard and survivor functions a big deal by inlcuding them in a figure, then we really should explain what to make of them and why they are important.]]
+
+- section 4.1.2. I don't understand why it is -187ms. This needs explaining because it is not obvious.
+
+- I really like it when you mention the use of "our" custom functions on page 19.
+That seems exactly what we should be doing.
+i.e., present a workflow and functions that make life easier for others to do it.
+
+- what does the following mean?
+
+>duplicate terms in the model formula are ignored
+
+This should be clearly explained or removed.
+
+- I like the walkthrough of the parameter estimates in section 4.3.8. That's nice and clear.
+
+- can we say why we have the following sentence?
+
+>For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+
+Were they too complex to fit or something else?
+
+## regressions references ##
+
+Gelman, A., Hill, J., & Vehtari, A. (2021). Regression and Other Stories. Cambridge University Press.
+
+Winter, B. (2019). Statistics for Linguists: An Introduction Using R. Routledge.
+
+## table 1 and table 2 ##
+
+Personally, I think having Tables 1 and 2 separately is fine. Let's leave it for now.
+
+## figure 1 ##
+
+I changed the layout of figure 1 as I thought it was easier to see.
+I also changed the way RT data were generated to a uniform distribution.
+
+## figure 3 ##
+
+I tried to insert a little more explanation as to why the joint consideration of hazard plus ca is psychologically interesting.
+One problem I had, however, is that if I calculate mean average RT from these hazards, then there is no difference in RT.
+I guess if we want to visually illustrate the differences it might be nice to have an insert plot showing the typical interferece measure of mean RT and then the distribution plots.
+However, it might also be better just to keep this super simple and make the point that we do at the moment, which is that both together help guide inferences.
+And let's worry less about plotting mean RT in this figure.
+
+## life tables ##
+
+section 4.1.2
+We mention life tables, but have we defined them before this point? If not, we need to define them first. And mention why they are important. If not, they just come out of nowhere.
+
+indeed, maybe the below quoted text should come earlier to explain life tables
+
+>Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
+
+## figure 5 ##
+
+Move this to the Tutorial. It is way too complicated for an already long paper.
+Folks who are interested can look at the tutorial.
+
+## code can be more compact ##
+
+e.g., this
+
+```{r}
+# load person-trial-bin oriented data set
+ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
+# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)
+ptb_data <- ptb_data %>% filter(period > 5)
+# create factor condition, with "blank" as the reference level
+ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.
+ptb_data <- ptb_data %>% 
+        mutate(period_9 = period - 9,
+               trial_c = (trial - 1000)/1000,
+               d6  = if_else(period == 6, 1, 0),
+               d7  = if_else(period == 7, 1, 0),
+               d8  = if_else(period == 8, 1, 0),
+               d9  = if_else(period == 9, 1, 0),
+               d10 = if_else(period == 10, 1, 0),
+               d11 = if_else(period == 11, 1, 0),
+               d12 = if_else(period == 12, 1, 0),
+               d13 = if_else(period == 13, 1, 0),
+               d14 = if_else(period == 14, 1, 0),
+               d15 = if_else(period == 15, 1, 0))
+```
+
+can instead be...
+
+```{r}
+# load person-trial-bin oriented data set
+ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv") %>% 
+# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)
+filter(period > 5) %>% 
+# create factor condition, with "blank" as the reference level
+mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent"))) %>% 
+# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.
+mutate(period_9 = period - 9,
+       trial_c = (trial - 1000)/1000,
+       d6  = if_else(period == 6, 1, 0),
+       d7  = if_else(period == 7, 1, 0),
+       d8  = if_else(period == 8, 1, 0),
+       d9  = if_else(period == 9, 1, 0),
+       d10 = if_else(period == 10, 1, 0),
+       d11 = if_else(period == 11, 1, 0),
+       d12 = if_else(period == 12, 1, 0),
+       d13 = if_else(period == 13, 1, 0),
+       d14 = if_else(period == 14, 1, 0),
+       d15 = if_else(period == 15, 1, 0))
+```
+
+And I also wonder. Should this even be in the main text, given that it is in the tutorial?
+I'm not sure.
+But it doesn't strike me as overly important to have the repetition or redundancy, and we have a long paper already.
+So folks might just get tired of having to figure out more stuff when they don't need to if they only want to read along.
+What do you think?
+
+## model building ##
+
+adapt_delta seems high. I rarely go above 0.99.
+And why not init_r = 0.1 or similar instead init = aexactly zero every time.
+0.1 means that the starting point varies a bit between -0.1 and 0.1 (i think)
+
+## table 4 and 5 ##
+
+These are way too long. 
+My suggestions as follows:
+
+1. just show con and inc 500 for the purposes of the manuscript. then in the main text and legend we can explain that it is just a selection of data for the purposes of understanding.
+The full tables can be in SUpps if you feel it is necessary.
+
+2. remove the .width column. just say in the legend 95% quantiles.
+
+3. split c500 and i500 into two columns. e.g, trial =500, condition = con or inc but in two separate columns.
+
+4. Round to two dp.
+
+5.
+And what's going on with some of those hazard values and odds ratio values?
+Is it just because one value could be so small that the ratio can be absolutely massive?
+It just seems odd without any explanation?
+e.g., take a look at the odds ratio column in table 5.
+
+For now, I just removed them to make the doc easier to manage.
+
+## model formulas ##
+
+some things are not clear, such as:
+
+- model_M1. 
+--The use of 0 and Intercept needs explaining.
+--adapt_delta is high. I rarely have to go above 0.99. Does the model not look right if this is set lower? This could be a sign that the model is not a suitable structure.
+--step_size - I have never set this. Can you send me some info on this parameter?
+-- init set at exactly zero.
+This should be avoided whenever possible.
+Try init_r = 0.1.
+This means that the start points will vary between -0.1 and 0.1, rather than always being the same value (zero).
+-- Although condition is defined earlier. I would make it clear in this section what condition means, as it is super important.
+
+- model_M2
+-- clearly define what period_9 means.
+-- It seems to me that you have quadratic and 3rd order terms, but you only mention 3rd order in the text.
+
+## loo plots ##
+
+They seem to be missing from section 4.3.7. 
+They would be useful in combinartion with model weights.
+
+## psychological / cognitive interpretation ##
+
+You say ...
+
+>The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, suggests the involvement of separate cognitive processes.
+
+Note - I removed the word "strongly" from here, as I'm just not sure it is appropriate and it adds nothing anyway.
+
+And my question: We argue the above, but couldn't an equally plausible account be that it reflects the exact same cognitive process, but for different durations?
+i.e., different onsets and offsets?
+Is there some aspect to the argument that I am missing?
+If so, we need to provide those details or give the alternative interpretation.
+It is still interesting, but I think we need to be clearer here and a little more careful.
+
+And a more general comment on this section, which I don't, as yet, have a perfect solution to.
+On page 37, the paragraph starts with Panis and Schmidt (2016) distinguished between automatic ...
+I like the aim of this paragraph because it brings folks back to cognitive mechanisms and cog psych, which many readers may be interested in, since that is who the article is aimed at.
+In other words, it aims to show how all this fancy modelling and effort can bring folks to a meaningfully different interpretation of the operation of the underlying system.
+With that said, after reading quite a bit of stats stuff, modelling stuff and code chunks, parameters etc., to them dump on top a bunch of new processes and terms that are not previously outlined, just feels too much for readers.
+Like they are punch-drunk at this point and we can tell them anything because they stopped listening or caring.
+
+So, what to do? I outline some options below.
+
+1.
+create a new subsection 4.3.9 Relevance for theory (or a similar title) where we place the paragraph and explain why and how such results matter for theory development.
+By doing so, we accept that it is long and complicated, but we think it is important enough to give it the space it needs.
+
+2.
+We largely remove the paragraph and instead point readers towards Panis and Schmidt 2016.
+e.g., To see how such results matter for theory development and the distinction between competing theories, see Panis and Schmidt 2016.
+
+## posterior predictions? ##
+
+You say...
+
+>To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trials 500, 1000, and 1500.
+
+Are these posterior predictions? If so, they need a little explanation.
+
+## figure settings ##
+
+The below figure settings are in the manuscript file.
+
+```{r plot-settings}
+## theme settings for ggplot
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 18), 
+          title = element_text(size = 18),
+          legend.position = "bottom")
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+This was confusing initially, because your figures definitely do not follow these settings.
+
+Then I realised that you are just taking figures from files. 
+So, the files used to generate the figures need this section adding to the top of the script, so that they are generated in a consistent manner.
+
+Also, I would be explicit with sizing when saving figures, if you don't so this already.
+If not, the size of the saved figure is determined by the plot window in RStudio and that can change depending on the screen and size of RStudio at the time.
+And also set the figure dpi higher for a crisper resolution e.g., dpi=800..
+
+https://ggplot2.tidyverse.org/reference/ggsave.html
+
+e.g.,
+
+```{r}
+ggsave("sims/figures/haz_surv.jpeg",
+       width = 7, height = 5, dpi=800)
+```
+
+One more important thing - colours/aesthetics etc. 
+I think we can improve our data vis game. 
+e.g., Figure 6 and 8 look great, but Figure 7 and 9 need some work.
+I think we can pick better colour palettes e.g., https://r-graph-gallery.com/38-rcolorbrewers-palettes.html
+
+I often go with the following palette ("Dark2"), see text and chunk below as an example.
+use colour and fill depending on whether you want the particular palette applied to colour or fill aesthetics or both.
+
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_fill_brewer(palette = "Dark2") +
+
+```{r}
+p9.7 <-ggplot(data_hs, aes(x = timebin, y = probability,
+                         colour = dv)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  ggtitle("hazard and survivor functions") 
+p9.7
+```
+
+This is important for several reasons.
+1. Data vis is just important and can really make a paper stand out.
+2. We need to be consistent throughout and we don't want some tables/figures looking great and others looking half-finished or an after thought.
+
+## figure legends ##
+
+in some cases, much more information is needed. e.g., Figure 7 and 9. Folks need a guide.
+
+## figure layout ##
+
+have you considered 1 column instead of 3 for figure 7/9?
+does it look better/easier to understand? Maybe try it and see?
+
+Figure 9 looks odd.
+Can we re-szie and re-colour as sggested above.
+and maybe try one column for the facet rather than 3.
+To set this, just insert "ncol=1" into the facet argument.
+e.g., facet_wrap(~condition, ncol=1).
+
+Figure 10 can be flipped such that the axes are opposite. e.g., flip x and y. then the text will be easier to read and plots easier to understand, in terms of numerical values.
+
+## Discussion ##
+
+I really liked reading the discussion. 
+For the most part, it was compact and efficient and covered some big-ticket items.
+The one exception was the 5.1 advantages section.
+It was all very clear, but it read like a background to the topic, which should come in the intro, rather than at the end.
+e.g., we've just been through all of this text, code, figures and model building and then we decide to say there are advantages??
+It just seems out of sync.
+So, I moved the text to Supplementary Material and created a new sub-section in supps.
+There are several benefits of doing so.
+
+1) we can keep the discussion focussed on things that are relevant to exp psych folks, rather than more general stats topics, which would be relevant for textbooks and the like.
+
+2) we can link to these benefits earlier on when we give a background to EHA in section 2.1.
+I've created a new sub-section in sectino 2 and added a sumamry paragraph that points the reader to the supplementary materials for the details.
+
+3) We can add something else into the discussion that is more relevant to our aim here.
+I have tried to add in a section on common use-cases for exp psych folks.
+It might not quite work or fit in yet, but the motivation was something like this.
+You may want to use it if you are interested in the timeline of effects.
+But you may also need to use it from a counfound/technical/method perspective because it might qualify inferences made from mean-average comparisons.
+The reason is just to make it clear that you need not be interested in temporal states to find this useful on a methodological/analytical perspective.

### 2024-10-30 sims/sims.Rmd
typo fixed
updates to the sim files and figs

@@
 
-use a truncated normal distribution to privude upper and lower bounds
+use a uniform distribution to provide upper and lower bounds
 
@@
   group_by(subj, rep, condition, bin) %>% 
-  mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
+  # mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
+  mutate(rt = runif(n=1, min=l, max=u))
 head(rt_dat_test)
@@
     group_by(subj, rep, condition, bin) %>% 
-    mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u)) %>% 
+    mutate(rt = runif(n=1, min=l, max=u)) %>% 
     ungroup()
@@
 
+re-plot with 2 columns, rather than 2 rows, and re-order the facets so that 1-3 are in col1
+
+```{r}
+p8.1b <- haz_plot(haz_pid_all, haz_group_all) +
+  facet_wrap(~fct_relevel(example,'example 1','example 4','example 2','example 5',
+                          'example 3', 'example 6'), ncol = 2)
+p8.1b
+```
+
 create a rt plot two rows
@@
 p8.2
+```
+
+re-plot rt with 2 columns, rather than 2 rows
+
+```{r}
+p8.2b <- rt_plot(rt_pid_all, rt_group_all) +
+  facet_wrap(~example, ncol = 2)
+p8.2b
 ```
@@
 
-```{r}
-p_haz_inset_facet <- p8.1 +
+with two rows
+
+```{r}
+# p_haz_inset_facet <- p8.1 +
+#   geom_plot(data = inset_facet,
+#             vp.width = 1/3,
+#             aes(x, y, label = plot))
+# p_haz_inset_facet
+# 
+# ggsave("sims/figures/haz_inset_facet.jpeg",
+#        width = 14, height = 10, dpi=1000)
+```
+
+with two columns
+
+```{r}
+p_haz_inset_facet <- p8.1b +
   geom_plot(data = inset_facet,
@@
 ggsave("sims/figures/haz_inset_facet.jpeg",
-       width = 14, height = 10, dpi=1000)
+       width = 10, height = 12, dpi=1000)
 ```
@@
 
+# section 10 #
+
 ## create a figure that illustrates hazard and survivor functions ##

### 2024-10-31 Tutorial_1a.Rmd
add table of content

@@
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 

### 2024-11-05 install_packages.Rmd
install new packages for Tutorial 3
add broom.mixed package
add packages for sims.Rmd

@@
 
+Next, packages for simulation (Figures 1, 2, and 3)
+
+```{r}
+install.packages(c("faux","ggpmisc","truncnorm"))
+```
+
 Next, packages for model building in brms used mainly in models.Rmd and effects.Rmd
@@
 install.packages(c("brms", "tidybayes", "bayesplot", "future"))
+
+install.packages("broom.mixed")
 ```
@@
 ```{r}
-#install.packages("lme4", "nlme")
+install.packages("lme4", "nlme")
 ```

### 2024-11-05 sims/sims.Rmd
update sims.Rmd to create 2 new Figures (2 and 3 in ms.)

@@
 ```
+
+# Section 11: Simulate hazard data for 1 subject, 1 condition, 200 trials, and plot empirical hazard and survivor functions with error bars (Figure 2)
+
+## Simulate hazard data based on user-defined hazard function
+
+```{r}
+## make it reproducible
+set.seed(7)
+
+## apply function sim_haz
+data_single <- sim_haz(subj_n = 1, rep_n = 200) %>%
+               filter(condition == "cond1") %>%
+               mutate(period = bin,
+                      trial = str_sub(rep, start=6, end=6)) %>%
+               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
+head(data_single, n=20)
+```
+
+## set up life table
+
+```{r}
+lifetable <- data_single %>%
+    mutate(event = str_c("event", event)) %>%
+    group_by(condition,period) %>% 
+    count(event) %>% 
+    ungroup() %>% 
+    pivot_wider(names_from = event,
+                values_from = n) %>% 
+    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+           event1 = ifelse(is.na(event1),0,event1),
+           risk_set = event0 + event1) %>% # define the risk set
+    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
+           pmass = event1/max(risk_set),
+           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
+    group_by(condition) %>%
+    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
+           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
+           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+    ungroup() 
+```
+
+## wrangle life table for plotting
+
+```{r}
+minimallt <- lifetable %>% select(period,condition, hazard, se_haz, survival, se_surv) %>%
+  pivot_longer(cols = c(hazard,survival),
+               names_to = "dv",
+               values_to = "probability") %>%
+  mutate(se = ifelse(dv=="hazard", se_haz, se_surv))
+```
+
+## plot hazard and survivor functions
+
+```{r}
+p11.1 <-ggplot(minimallt, aes(x = period, y = probability,
+                         colour = dv)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  geom_linerange(aes(ymin=probability-se, ymax=probability+se), show.legend = F) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  ggtitle("hazard and survivor functions") +
+  xlab("timebin")
+p11.1
+
+ggsave("sims/figures/haz_surv_single.jpeg",
+       width = 7, height = 5, dpi=800)
+```
+
+# Section 12 (Figure 3): plot of hazard and ca(t) for 2 conditions with error bars
+
+## hazard
+
+### generate data set
+
+```{r}
+## make it reproducible
+set.seed(8)
+
+## apply function sim_haz
+data_single2 <- sim_haz(subj_n = 1, rep_n = 2000) %>%
+               mutate(period = bin,
+                      trial = str_sub(rep, start=6, end=6)) %>%
+               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
+head(data_single2, n=20)
+```
+
+### set up life table
+
+```{r}
+lifetable2 <- data_single2 %>%
+    mutate(event = str_c("event", event)) %>%
+    group_by(condition,period) %>% 
+    count(event) %>% 
+    ungroup() %>% 
+    pivot_wider(names_from = event,
+                values_from = n) %>% 
+    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+           event1 = ifelse(is.na(event1),0,event1),
+           risk_set = event0 + event1) %>% # define the risk set
+    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
+           pmass = event1/max(risk_set),
+           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
+    group_by(condition) %>%
+    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
+           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
+           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+    ungroup() 
+```
+
+### plot hazard functions
+
+```{r}
+p12.1 <-ggplot(lifetable2, aes(x = period, y = hazard,
+                         colour = condition)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  ggtitle("hazard")  +
+  xlab("timebin") + ylab("probability")
+p12.1
+```
+
+## accuracy
+
+create some data
+
+```{r}
+acc <- tibble(
+  time = seq(1,10,1),
+  cond1 = rep(1,length = 10),
+  cond2 = c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
+) %>% 
+  pivot_longer(-time,
+               names_to = "condition",
+               values_to = "accuracy") %>% 
+  mutate(condition = factor(condition))
+head(acc,n=20)
+```
+
+line plot per bin
+
+```{r}
+p12.2 <-ggplot(acc, aes(x = time, y = accuracy,
+                         colour = condition)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10) +
+  scale_y_continuous(limits = c(0,1)) +
+  xlab("timebin") +
+  ylab("probability") +
+  ggtitle("conditional accuracy") 
+p12.2
+```
+
+Combine plots
+
+```{r}
+p12.3 <- (p12.1 / p12.2) +
+  plot_layout(guides = 'collect',
+              axes = 'collect')
+p12.3
+
+ggsave("sims/figures/haz_acc_single.jpeg",
+       width = 7, height = 5, dpi=800)
+
+```

### 2024-11-06 manuscript/SupplementaryMaterial.Rmd
update ms. and suppl. material.

@@
 The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
+This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
@@
 
-When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y]: x < RT $\leq$ y.
+If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
 Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).

### 2024-11-06 manuscript/manuscript.Rmd
update ms. and suppl. material.

@@
 The ability to reveal finer-grained, "temporal states" of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
-One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
-To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis.
+Moreover, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
+To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
 We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
@@
 Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @williammatthewmakehamLawMortalityConstruction1860].
-The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
-Here, in the current paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
+The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020].
+Here, in the current paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
+We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
 
@@
 
-In this paper, we focus on a distributional method for time-to-event data known as *discrete-time event history analysis*, a.k.a.
-hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
+In this paper, we focus on discrete-time EHA.
+We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
@@
 We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja() package [[insert REF for papaja() here]], which makes it computationally reproducible, in terms of the undeerlying data and figures.
+Finally, the manuscript itself is written in R using the papaja() package [@pap2024], which makes it computationally reproducible, in terms of the underlying data and figures.
 
@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
+In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
 In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
+We then briefly compare and contrast these inferential frameworks when applied to EHA.
 In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
@@
 
-In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
 
@@
 
-For a comprehensive background context to event history analysis, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
+For a comprehensive background context to EHA, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
 Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
-Instead, we want to provide an intuition regarding how event history analysis works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in supplementary material and then refer to them in the text whenever relevant.
+Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
+As such, we only supply regression equations in the supplementary material.
 
@@
 
-To apply event history analysis (EHA), one must be able to:
+To apply EHA, one must be able to:
 
@@
 
-The definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+In EHA, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
 Since our focus here is on hazard models that use discrete time units, we describe that approach.
@@
 For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
-Continuous RT data is treated here as interval-censored data.
 
 Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A).
+The equations that define both of these functions are reported in part A of the supplementary material.
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs after bin t.
@@
 
-The survivor function can help to qualify or provide context to the interpretation of the hazard function.
-For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
-If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500], then only 3% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in later parts of the hazard function would typically be wider and less precise compared to earlier parts.
-
 (ref:descr-plot2-caption) Discrete-time hazard and survivor functions.
+While the hazard function is the vehicle for inferring the time course of cognitive processes, the survivor function at S(t-1) can help to qualify or provide context to the interpretation of the hazard function at h(t).
+For example, the high hazard of h(t=5) = .xx is experienced only by S(t-1=4) = .yy percent of the trials.
+Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will typically be wider and less precise compared to earlier parts.
 
 ```{r plot2, fig.cap = "(ref:descr-plot2-caption)", out.width="80%"}
-knitr::include_graphics("../sims/figures/haz_surv.jpeg")
+knitr::include_graphics("../sims/figures/haz_surv_single.jpeg")
 ```
@@
 
-2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences than mean-average comaprisons may conceal (Townsend, 1990).
+2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences that mean-average comparisons may conceal (Townsend, 1990).
 
@@
 EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with longer RTs are not discarded, but instead controbute to the risk set in each time bin.
+Trials with longer RTs are not discarded, but instead contribute to the risk set in each time bin.
 
@@
 Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
+[@allisonSurvivalAnalysisUsing2010].
 This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
+For such two-choice RT data, the discrete-time EHA of the RT data can be extended with a discrete-time SAT analysis of the timed accuracy data.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
@@
 ```{r plot3, fig.cap = "(ref:descr-plot3-caption)", out.width="80%"}
-knitr::include_graphics("../sims/figures/haz_acc.jpeg")
+knitr::include_graphics("../sims/figures/haz_acc_single.jpeg")
 ```
@@
 
-Performing event history analyses in experimental psychology has implications for how experiments are designed.
+Performing EHA in experimental psychology has implications for how experiments are designed.
 Indeed, if trials are categorised as a function of when responses occur, then each timebin will only include a subset of the total number of trials.
@@
 Here we focus on the subset of trials in which no mask was presented.
-The 13-ms prime stimulus was a double arrow with onset at -187 ms for the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
+The 13-ms prime stimulus was a double arrow presented 187 ms before target onset in the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
 
@@
 Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t) in a selected baseline condition.
-One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below).
-Relevant example regression formulas are provided in equations 6 and 7 in part D of the supplementary material.
+One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below; see also part D of the supplementary material).
 
@@
 
-This main motivation for writing this paper is the observation that event history and SAT analyses remain under-used in psychological research.
+This main motivation for writing this paper is the observation that EHA and SAT analysis remain under-used in psychological research.
 As a consequence, the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
@@
 For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before being absorbed into a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize event history analysis to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
+When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize EHA to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
@@
 Statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
-We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing event history analysis and investigating the temporal profile of cognitive states.
+We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing EHA and investigating the temporal profile of cognitive states.
 On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.

### 2024-11-06 manuscript/manuscript.tex
update ms. and suppl. material.

@@
 The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
-One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
-To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis.
+Moreover, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
+To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
 We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., William Matthew Makeham, 1860).
-The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
-Here, in the current paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Makeham, William M., 1860).
+The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
+Here, in the current paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
+We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
 
@@
 
-In this paper, we focus on a distributional method for time-to-event data known as \emph{discrete-time event history analysis}, a.k.a.
-hazard analysis, duration analysis, failure-time analysis, survival analysis, and transition analysis.
-We first provide a brief overview of event history analysis to orient the reader to the basic concepts that we will use throughout the paper.
+In this paper, we focus on discrete-time EHA.
+We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003).
@@
 We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply the analyses to their own time-to-event data sets.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja() package {[}{[}insert REF for papaja() here{]}{]}, which makes it computationally reproducible, in terms of the undeerlying data and figures.
+Finally, the manuscript itself is written in R using the papaja() package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.
 
@@
 We discuss possible link functions, and plot the model-based effects of our predictors of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the event history analysis of RT data for choice RT data.
+In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
 In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to event history analysis.
+We then briefly compare and contrast these inferential frameworks when applied to EHA.
 In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
@@
 
-In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
-Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
 
@@
 
-For a comprehensive background context to event history analysis, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
+For a comprehensive background context to EHA, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
 Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
-Instead, we want to provide an intuition regarding how event history analysis works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in supplementary material and then refer to them in the text whenever relevant.
+Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
+As such, we only supply regression equations in the supplementary material.
 
@@
 For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
-Continuous RT data is treated here as interval-censored data.
 
 Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A).
+The equations that define both of these functions are reported in part A of the supplementary material.
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t.
@@
 
-The survivor function can help to qualify or provide context to the interpretation of the hazard function.
-For example, it can give a sense of how many trials contribute to each part of the hazard distribution.
-If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500{]}, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials.
-Therefore, the error bars in later parts of the hazard function would typically be wider and less precise compared to earlier parts.
-
 \begin{figure}[H]
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_surv}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_surv_single}
 
@@
 
-\caption{Discrete-time hazard and survivor functions.}\label{fig:plot2}
+\caption{Discrete-time hazard and survivor functions.
+While the hazard function is the vehicle for inferring the time course of cognitive processes, the survivor function at S(t-1) can help to qualify or provide context to the interpretation of the hazard function at h(t).
+For example, the high hazard of h(t=5) = .xx is experienced only by S(t-1=4) = .yy percent of the trials.
+Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will typically be wider and less precise compared to earlier parts.}\label{fig:plot2}
 \end{figure}
@@
 \item
-  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences than mean-average comaprisons may conceal (Townsend, 1990).
+  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences that mean-average comparisons may conceal (Townsend, 1990).
 \item
   EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with longer RTs are not discarded, but instead controbute to the risk set in each time bin.
+Trials with longer RTs are not discarded, but instead contribute to the risk set in each time bin.
 \item
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis.
+For such two-choice RT data, the discrete-time EHA of the RT data can be extended with a discrete-time SAT analysis of the timed accuracy data.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_acc}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_acc_single}
 
@@
 
-We used R (Version 4.4.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
@@
 For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before being absorbed into a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize event history analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
+When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize EHA to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
 Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
@@
 Aust, F., \& Barth, M. (2023). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}
+
+\bibitem[\citeproctext]{ref-pap2024}
+Aust, F., \& Barth, M. (2024). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
 
@@
 Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
+\bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
+Makeham, William M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
 
@@
 \bibitem[\citeproctext]{ref-R-forcats}
-Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://CRAN.R-project.org/package=forcats}
+Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}
 
 \bibitem[\citeproctext]{ref-R-stringr}
-Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://CRAN.R-project.org/package=stringr}
+Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}
 
@@
 \bibitem[\citeproctext]{ref-R-dplyr}
-Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}
+Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
 
 \bibitem[\citeproctext]{ref-R-purrr}
-Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
 
 \bibitem[\citeproctext]{ref-R-readr}
-Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
+Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
 
 \bibitem[\citeproctext]{ref-R-tidyr}
-Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
-
-\bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-William Matthew Makeham. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
+Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
 
@@
 The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
+This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
@@
 
-When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y.
+If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
 Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).

### 2024-11-06 notes/manuscript_notes.Rmd
update ms. and suppl. material.

@@
 
+-> I replaced each reaction with response, as reaction is often used for detection tasks only.
+
 - Makeham reference. Shouldn't it just be Makeham, 1860, rather than first and middle names?
+
+-> Yes, I will check the bib file... correct version is in extrareferences.bib (backup for extrareferences2.bib)
 
@@
 
+-> ref papaja is in extrareferences.bib
+
 - I would remove reference to specific equations throughout the manuscript.
@@
 Folks who want more can easily flick to the supps and read all they want.
+
+-> removed equation numbers in sections 2 and 2..
 
@@
 
+-> yes, done.
+
 - section 2.1. You say...
@@
 But in the context of this paragraph, this makes no sense and just seems tacked on and disconnected. Can we explain or remove?
+
+-> I removed it, and added it with explanation in the supplementary material - part A.
 
@@
 So, I would just remove it from here for now, to make things shorter and simpler, at least for the time being.
+
+-> Deleted and moved to supplementary material.
 
@@
 
+-> Sentence deleted, but wrapped with old paragraph in figure legend of Figure 2.
+
 - section 4.1.2. I don't understand why it is -187ms. This needs explaining because it is not obvious.
+
+-> The prime is presented *before* the target. I removed the minus and made it explicit.
 
@@
 i.e., present a workflow and functions that make life easier for others to do it.
+
+-> great.
 
@@
 
+-> I clearly explained it..... todo
+
 - I like the walkthrough of the parameter estimates in section 4.3.8. That's nice and clear.
+
+-> ok.
 
@@
 
+-> I expected problems with fitting the very complex model M4. But even M3 has problems.
+
 ## regressions references ##
@@
 Winter, B. (2019). Statistics for Linguists: An Introduction Using R. Routledge.
+
+-> added in
 
@@
 
+-> I prefer the last option. I created a new figure 3.
+
 ## life tables ##
@@
 We mention life tables, but have we defined them before this point? If not, we need to define them first. And mention why they are important. If not, they just come out of nowhere.
-
 indeed, maybe the below quoted text should come earlier to explain life tables
@@
 
+-> Done.  todo...
+
 ## figure 5 ##
@@
 Folks who are interested can look at the tutorial.
+
+-> Done.  todo....
 
@@
 
+-> I think it is important for naive readers to get an idea of how the variables in the models presented later are calculated.
+
 ## model building ##
@@
 0.1 means that the starting point varies a bit between -0.1 and 0.1 (i think)
+
+-> adapt-delta of 0.999 is not unusual for generalized linear mixed models.
 
@@
 
+-> Done todo
+
 3. split c500 and i500 into two columns. e.g, trial =500, condition = con or inc but in two separate columns.
+
+-> done todo
 
@@
 
+-> Explanation for large values is given in the table captions.  todo
+
 ## model formulas ##
@@
 --The use of 0 and Intercept needs explaining.
+
+-> done.
+
 --adapt_delta is high. I rarely have to go above 0.99. Does the model not look right if this is set lower? This could be a sign that the model is not a suitable structure.
+
 --step_size - I have never set this. Can you send me some info on this parameter?
@@
 This means that the start points will vary between -0.1 and 0.1, rather than always being the same value (zero).
+
 -- Although condition is defined earlier. I would make it clear in this section what condition means, as it is super important.
+
+-> Done... todo
 
@@
 -- clearly define what period_9 means.
+
 -- It seems to me that you have quadratic and 3rd order terms, but you only mention 3rd order in the text.
+
+-> I thought a "third order polynomial" implies a constant, linear, quadratic and cubic component...
 
@@
 They would be useful in combinartion with model weights.
+
+-> todo.  Done....
 
@@
 e.g., To see how such results matter for theory development and the distinction between competing theories, see Panis and Schmidt 2016.
+
+-> I prefere option 1, and added a new subsection... todo
 

### 2024-11-06 manuscript/manuscript.Rmd
update ms.

@@
 The ability to reveal finer-grained, "temporal states" of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Moreover, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states.
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
 To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
-We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
+We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multilevel regression modelling.
 Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
-Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
+Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
 
@@
 As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, top row).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, bottom row).
+For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
+In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, right).
 Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
@@
 The first second after stimulus onset is divided in ten bins of 100 ms.
+Timebin indicates the bin rank.
+The first bin is (0,100], the last bin is (900,1000].
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
@@
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in Example 4 show that the effect starts around 200 ms and is gone by 600 ms.
-In contrast, in Example 5, the effect starts around 400 ms and is gone by 800 ms.
+For example, the distributions in Example 4 show that the effect starts between 100 and 200 ms (in timebin 2) and is gone when the waiting time reaches 500 ms or more.
+In contrast, in Example 5, the effect starts around 300 ms and is gone by 700 ms.
 And in the Example 6, the effect reverses between 500 and 600 ms.
@@
 For a comprehensive background context to EHA, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
-Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
+Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks [@gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
@@
 
-1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.)
+1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
 
-2. define time point zero (e.g., target stimulus onset, fixation onset)
+2. define time point zero (e.g., target stimulus onset, fixation onset);
 
@@
 After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600).
+For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600 ms).
 
@@
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs after bin t.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
 In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
@@
 (ref:descr-plot2-caption) Discrete-time hazard and survivor functions.
-While the hazard function is the vehicle for inferring the time course of cognitive processes, the survivor function at S(t-1) can help to qualify or provide context to the interpretation of the hazard function at h(t).
-For example, the high hazard of h(t=5) = .xx is experienced only by S(t-1=4) = .yy percent of the trials.
-Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will typically be wider and less precise compared to earlier parts.
+Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
+While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
+For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t-1=5) = .44.
+Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.
 
@@
 
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer.
+In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task.
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data can be extended with a discrete-time SAT analysis of the timed accuracy data.
+For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-We refer to this extended analysis for choice RT data as EHA/SAT.
+We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
 
@@
 To illustrate, we consider a hypothetical choice RT example that is inspired by real data [@panisWhatShapingRT2016], but simplified to make the main point clearer (Figure 3).
-In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
+In a standard priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
 The prime can then be congruent or incongruent with the target.
 
-Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
-And then for longer waiting times, the response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+Between 500 and 600 ms the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 
@@
 Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials.
+For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials, and are indexed by each recorded response.
 But such a conclusion is not supported when the effects are explored over a timeline.
@@
 (ref:descr-plot3-caption) Discrete-time hazard and conditional accuracy functions.
+Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
 
@@
 This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
 
@@
 
-Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018] ; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].
+Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].
 We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
@@
 
-Note that duplicate terms in the model formula are ignored. Estimating model M3 took about 4.5 hours.
+Note that duplicate terms across the interaction terms in the model formula (e.g., condition) are ignored. Estimating model M3 took about 4.5 hours.
 
@@
 
-Again, duplicate terms in the model formula are ignored. Estimating model M4 took about 8 hours.
+Again, duplicate terms across the interaction terms in the model formula are ignored. Estimating model M4 took about 8 hours.
 
@@
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Because we expected that the random effects structure of model M4 would be too complex for a freqentist approach, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
@@
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar,, which makes sense given the close similarity in model structure.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
 However, the random effects structure of model M3_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.

### 2024-11-06 manuscript/manuscript.tex
update ms.

@@
 The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Moreover, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states.
+Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states.
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
 To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
-We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling.
+We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multilevel regression modelling.
 Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
-Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
+Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
 }
@@
 The first second after stimulus onset is divided in ten bins of 100 ms.
+Timebin indicates the bin rank.
+The first bin is (0,100{]}, the last bin is (900,1000{]}.
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
@@
 For a comprehensive background context to EHA, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
-Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Sven - See my comments in the notes file for relevant references here).
+Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
@@
 \item
-  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.)
-\item
-  define time point zero (e.g., target stimulus onset, fixation onset)
+  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+\item
+  define time point zero (e.g., target stimulus onset, fixation onset);
 \item
@@
 In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
 In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
@@
 \caption{Discrete-time hazard and survivor functions.
-While the hazard function is the vehicle for inferring the time course of cognitive processes, the survivor function at S(t-1) can help to qualify or provide context to the interpretation of the hazard function at h(t).
-For example, the high hazard of h(t=5) = .xx is experienced only by S(t-1=4) = .yy percent of the trials.
-Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will typically be wider and less precise compared to earlier parts.}\label{fig:plot2}
+Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
+While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
+For example, the high hazard of .60 = h(t=6) is experienced only by .44 = S(t-1=5) percent of the trials.
+Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.}\label{fig:plot2}
 \end{figure}
@@
 
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer.
+In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task.
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data can be extended with a discrete-time SAT analysis of the timed accuracy data.
+For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
 Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-We refer to this extended analysis for choice RT data as EHA/SAT.
+We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
 
@@
 To illustrate, we consider a hypothetical choice RT example that is inspired by real data (Panis \& Schmidt, 2016), but simplified to make the main point clearer (Figure 3).
-In a standard response priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
+In a standard priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
 The prime can then be congruent or incongruent with the target.
 
-Figure 3 shows that the early upswing in hazard is equal for both prime conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-And then for longer waiting times, the response hazard is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
+Between 500 and 600 ms the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
 
@@
 
-\caption{Discrete-time hazard and conditional accuracy functions.}\label{fig:plot3}
+\caption{Discrete-time hazard and conditional accuracy functions.
+Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).}\label{fig:plot3}
 \end{figure}
@@
 This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
 
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018) ; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2018).
+Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2018).
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
@@
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+Because we expected that the random effects structure of model M4 would be too complex for a freqentist approach, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
@@
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar,, which makes sense given the close similarity in model structure.
+Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
 However, the random effects structure of model M3\_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
@@
 Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in bayesian workflow. \emph{J. R. Stat. Soc. A}, \emph{182}, 389--402. \url{https://doi.org/10.1111/rssa.12378}
+
+\bibitem[\citeproctext]{ref-gelmanRegressionOtherStories2020}
+Gelman, A., Hill, J., \& Vehtari, A.
+(2020).
+Regression and {Other Stories}.
+https://www.cambridge.org/highereducation/books/regression-and-other-stories/DD20DD6C9057118581076E54E40C372C; Cambridge University Press.
+\url{https://doi.org/10.1017/9781139161879}
 
@@
 
+\bibitem[\citeproctext]{ref-winterStatisticsLinguistsIntroduction2019}
+Winter, B. (2019). \emph{Statistics for {Linguists}: {An Introduction Using R}}. New York: Routledge. \url{https://doi.org/10.4324/9781315165547}
+
 \bibitem[\citeproctext]{ref-wolkersdorferTemporalDynamicsSequential2020}

### 2024-11-06 notes/manuscript_notes.Rmd
update notes

@@
 
--> I clearly explained it..... todo
+-> changed to: Note that duplicate terms across the interaction terms in the model formula (e.g., condition) are ignored.
 
@@
 
--> I expected problems with fitting the very complex model M4. But even M3 has problems.
+-> I expected problems with fitting the very complex model M4. I added this info.
 
@@
 
--> added in
+-> done.
 
@@
 
+-> ok.
+
 ## figure 1 ##
@@
 I also changed the way RT data were generated to a uniform distribution.
+
+-> ok.
 
@@
 
--> I prefer the last option. I created a new figure 3.
+-> I prefer the last option. I created a new figure 3 and added info to its caption.
 
@@
 
--> I think it is important for naive readers to get an idea of how the variables in the models presented later are calculated.
+-> I think it is important for naive readers to get an idea of how the variables in the model formulas presented later are calculated.
 
@@
 
--> Done todo
+->  todo
 
@@
 
--> done todo
+->  todo
 
 4. Round to two dp.
+
+-> todo
 
@@
 
--> done.
+-> todo
 
@@
 
--> Done... todo
+-> todo
 
@@
 -- clearly define what period_9 means.
+
+-> todo
 
@@
 
--> todo.  Done....
+-> todo.
 
@@
 The reason is just to make it clear that you need not be interested in temporal states to find this useful on a methodological/analytical perspective.
+
+# comments Sven 6 nov
+
+1. Figure 1: connect the points per subject?
+
+2.

### 2024-11-06 sims/sims.Rmd
update sims

@@
 ## apply function sim_haz
-data_single2 <- sim_haz(subj_n = 1, rep_n = 2000) %>%
+data_single2 <- sim_haz(subj_n = 1, rep_n = 1000) %>%
                mutate(period = bin,
-                      trial = str_sub(rep, start=6, end=6)) %>%
-               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
+                      trial = as.integer(str_sub(rep, start=4, end=7)),
+                      condition = factor(condition, labels=c("congruent","incongruent"))) %>%
+               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep"))  %>%
+               mutate(acc = ifelse(period<6 & condition=="incongruent",0,1),
+                      acc = ifelse(period==6 & condition=="incongruent" & trial < 500,0,acc)) 
 head(data_single2, n=20)
+summary(data_single2)
+
 ```
@@
 
+## add ca(t)
+
+```{r}
+ca <- data_single2 %>% 
+    group_by(condition,period) %>%
+    summarize(ca = mean(acc) %>% round(digits = 2),
+              n = n(),
+              .groups = 'drop') %>%
+    ungroup() %>%
+    mutate(se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) 
+ca
+```
+
 ### plot hazard functions
@@
   ggtitle("hazard")  +
-  xlab("timebin") + ylab("probability")
+  xlab("timebin") + ylab("probability") + guides(color=guide_legend(title="prime type"))
 p12.1
@@
   time = seq(1,10,1),
-  cond1 = rep(1,length = 10),
-  cond2 = c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
+  congruent = rep(1,length = 10),
+  incongruent = c(0, 0, 0, 0, 0, .5, 1, 1, 1, 1)
 ) %>% 
@@
 ```{r}
-p12.2 <-ggplot(acc, aes(x = time, y = accuracy,
+p12.2 <-ggplot(ca, aes(x = period, y = ca,
                          colour = condition)) +
@@
   geom_point(size=3) +
+  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
   scale_colour_brewer(palette = "Dark2") + 
@@
   ylab("probability") +
-  ggtitle("conditional accuracy") 
+  ggtitle("conditional accuracy") + 
+  guides(color=guide_legend(title="prime type"))
 p12.2
@@
 ggsave("sims/figures/haz_acc_single.jpeg",
-       width = 7, height = 5, dpi=800)
-
-```
+       width = 7, height = 6, dpi=800)
+
+```

### 2024-11-07 Tutorial_4.Rmd
tutorial 4 initial commit

@@
 title: "Tutorial_4"
-author: "Sven Panis"
+author: "Rich"
 date: "`r Sys.Date()`"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-Tutorial_4 illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
+Tutorial 4 shows one way to plan the design of a future experiment using EHA.
 
-# Load the libraries that we will be using.
+# 1. Load the libraries that we will be using #
 
-```{r load-pkg}
-pkg <- c("tidyverse", "patchwork")
+```{r load-pkg, results='hide'}
+pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
+         "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux")
 
@@
 
-# Read in the raw data from Panis and Schmidt (2016).
+Set options.
 
-```{r read-raw-data}
-raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) # 9372 rows
-head(raw)
-tail(raw)
-glimpse(raw)
+```{r set-options, results='hide'}
+options(brms.backend = "cmdstanr",
+        mc.cores = parallel::detectCores(),
+        future.fork.enable = TRUE,
+        future.rng.onMisuse = "ignore") ## automatically set in RStudio
+
+supportsMulticore()
+detectCores()
 ```
 
-# Do some wrangling.
-
-```{r wrangle-raw-data}
-# remove practice trials
-data_wr <- raw %>% filter(practice==0) # 9240 experimental trials: 6 participants x 70 blocks x 22 trials-per-block
-
-# remove unnecessary columns
-data_wr <- data_wr %>% select(-c(ss,practice,rd,form,task,primeduration,
-                                             blankduration,maskduration,targetduration,
-                                             void,target_dir,corresp,resperr))
-# look at trials where resp==0
-data_wr %>% filter(resp==0) 
-data_wr %>% filter(resp==0) %>% print(n=51)
-
-# remove trials with no response info if rt < 800 (800 was the response deadline in each trial):
-# the 46 trials where rt >= 800 and resp==0 are trials in which the subject did not respond; these should not be deleted because they will be treated as right-censored observations in EHA!
-# the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
-sel = which(data_wr$resp==0 & data_wr$rt < 800)
-data_wr <- data_wr[-c(sel),] # 9235 rows
-
-summary(data_wr)
-# vp = Versuchsperson or subject
-# bl = block number
-# tr = trial number within block
-# mask_type: no mask (0), relevant mask (1), irrelevant mask (2), lines mask (3)
-# prime_type: no prime or blank (0), consistent prime (1), inconsistent prime (2)
-# resp = no button (0), button left (1), button right (2) 
-# respac = accuracy
-# rt = response time
-# TrialNr = trial number within experiment
+```{r check-info, results='hide'}
+# packageVersion("cmdstanr")
+# devtools::session_info("rstan")
 ```
 
-# Change variable names, and create factors.
+theme settings for ggplot
 
-```{r create-names-and-factors}
-# change some variable names to required names (pid, trial, condition1, condition2, rt, and acc)
-colnames(data_wr) <- c("pid","bl","tr","condition1", "condition2","resp","acc","rt","trial") 
-# pid to indicate column with unique person IDs
-# trial must be a unique trial number per subject
-# condition1 must be a factor indicating the levels (1,2,...) and labels of the first independent variable (here: mask_type)
-# condition2 must be a factor indicating the levels (1,2,...) and labels of the second independent variable (here: prime_type)
-# rt to indicate column with response times (in ms)
-# acc to indicate column with accuracies (1/0)
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"),
+          legend.position = "bottom")
+)
 
-# create factors for conditions (mask type and prime type): the levels of condition1 are displayed across columns; the levels of condition2 are displayed within each plot.
-data_wr <- data_wr %>% mutate(condition1 = condition1 + 1, # avoid level "0"
-                              condition1 = factor(condition1, levels = c(1,2,3,4), labels = c("nomask","relevant","irrelevant","lines")),
-                              condition2 = condition2 + 1, # avoid level "0"
-                              condition2 = factor(condition2, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
-summary(data_wr)
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
 ```
 
-# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).
+# 2. A brief outline of the workflow #
 
-Load the functions. Make sure you rename, and specify the same levels and labels for factors condition1 and condition2 in the function plot_eha_2IV as above.
+## 2.1 - Main background sources ##
 
-```{r load-functions}
-# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
-censor <- function(df, timeout, bin_width){
-  if(!(timeout %% bin_width == 0)){
-    return("The censoring time must be a multiple of the bin width!")
-  }
-  if(timeout < 0 | bin_width < 0){
-    return("Both timeout and bin_width must be larger than 0!")
-  }
-  df %>% mutate(right_censored = 0,
-                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
-                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
-                drt = ceiling(rtc/bin_width), # drt = discrete response time
-                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
-}
+The basic approach is based upon two main sources:
 
-# function ptb creates a person-trial-bin oriented data set
-ptb <- function(df){
-  df %>% uncount(weights = drt) %>% # create drt rows per trial
-         group_by(trial) %>% 
-         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
-         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
+The basic structure and code follows the examples outlined by Solomon Kurz in his
+'power' blog posts and Lisa Debruine's faux{} package.
+
+For Solomon's blog posts, see here: [https://solomonkurz.netlify.app/tags/power/](https://solomonkurz.netlify.app/tags/power/)
+
+For Lisa's faux package, see here: [https://debruine.github.io/faux/](https://debruine.github.io/faux/).
+The vignettes provided by the faux package are particularly helpful and worth following before you take a deep dive in the below code.
+
+And for some simpler worked examples of the general workflow (i.e., simpler than using EHA), then take a look at these examples by Rich, which show how you might simulate multi-level data for a range of factorial designs: [https://github.com/rich-ramsey/sim_demo](https://github.com/rich-ramsey/sim_demo)
+
+## 2.2 - Basic workflow ##
+
+The basic workflow is as follows:
+
+1. Fit a regression model to an existing dataset.
+2. Use the regression model parameters to simulate a one new dataset.
+3. Create a function to iterate over many datatsets and vary parameters of interest (e.g., sample size, trial count, effect size).
+4. Summarise the simulated data to estimate likely power or precision of the research design options.
+
+## 2.3 - Exceptions to the normal workflow ##
+
+The basic workflow above is a modified version of that recommended by Solomon and Lisa.
+In a typical workflow, once the datasets are simulated, a regression model would be fit to each datatset.
+Then, the parameters from each model would be summarised to make a judgment about the likely precision or power of the desing.
+However, in the case of Model_4 in this tutorial, which took ~8 hours to build, we are not able to easily build 1000 models per variation in sample size and/or trial count, as that would take an unreasonably long time.
+Instead, we choose to take a simpler approach and summarise the data per bin and condition, rather than the parameter estimates from the model.
+Although this is not as satisfying, we feel it can still be a useful guide to set expectations. 
+To further save time and computational resources, we also choose to use a much simpler model than Model_4, in order to outline the basic approach to planning via data simulation. 
+The details on this simpler model structure are provided below.
+
+# 3. Read in the data #
+
+Here, we again use data form Panis and Schmidt (2016), but to make things simpler, we focus on only 6 timebins (timebins 4-9) and two conditions (con and inc).
+
+## read in and wrangle the raw data ##
+
+```{r load-data}
+data <- read_csv("Tutorial_4_planning/data/tidyppp_Exp1_n6.csv") %>%
+  ## no mask condition only, neutral prime removed, timebins 4-9 only
+  filter(no_ma == 1, no_pr < 1, timebin %in% c(4:9)) %>% 
+  ## remove unnecessary columns
+  select(-c(bl, target_dir, corresp, resp, respac, rel_ma, irr_ma, ran_ma, no_ma)) %>%
+  ## rename the error column
+  rename(error = resperr) %>% 
+  ## create a new variable and some factors
+  mutate(prime = if_else(con_pr == 1, "con", "inc"),
+         prime = factor(prime,
+                        levels = c("con", "inc")),
+         timebin = factor(timebin)) %>% 
+  ## simplify the data by selecting a few variables only
+  select(pid, timebin, prime, outcome)
+head(data)
+str(data)
+```
+
+data check
+
+```{r check-conds}
+data %>% 
+  distinct(timebin, prime)
+```
+
+## create summary data ##
+
+summary data
+
+```{r calculate-summary-data}
+data_pid <- data %>% 
+  group_by(pid, timebin, prime) %>% 
+  summarise(outcome = mean(outcome))
+head(data_pid)
+
+data_group <- data %>% 
+  group_by(timebin, prime) %>% 
+  summarise(outcome = mean(outcome))
+head(data_group)
+```
+
+quick plot
+
+```{r quick-summary-plot}
+p3.1 <- ggplot(data_group, aes(x=timebin, y = outcome, colour = prime)) +
+   geom_line(aes(group = prime)) + geom_point() +
+   geom_jitter(data=data_pid, alpha = 0.5, width = 0.1, height = 0) +
+   scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+   scale_x_discrete(labels = c(1:6)) +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   labs(title = "Panis & Schmidt (2016) (N=6)",
+        x = "time bins")
+p3.1
+```
+
+check the number of trials
+
+```{r check-trials}
+tally <- data %>% 
+  group_by(pid, prime, timebin) %>%
+  summarise(n=n(),
+            sum=sum(outcome),
+            .groups = "drop")
+tally
+```
+
+# 4. Build an initial model #
+
+In general, it can sometimes be a lot simpler to use an index coding approach to build regression models.
+
+see this link for some background on an index coding approach:
+https://bookdown.org/content/4857/the-many-variables-the-spurious-waffles.html#many-categories.
+
+and then maybe specify the interaction term like this...
+https://discourse.mc-stan.org/t/specifying-formulates-for-interaction-between-categorical-variables-with-the-index-coding-approach-in-brms/29449/3
+
+And in the context of data simulation, index coding can be particularly useful, because it may be often easier and more intuitive to set condition means, rather than more complicated regression parameters, such as interaction terms.
+
+So, the approach here is to build an initial model using an index coding approach, but with a much simpler dataset than has been used in prior tutorials.
+
+Because we are building Bayesian regression models, which generate a posterior distribution per condition of interest in our index coding approach, we can simply calculate contrasts of interests and associated prevision or interval estimates (95% quantile intervals, for example) to address any questions we may have about comparisons between conditions, such as inc vs con in a particular timebin.
+
+## formula ##
+
+with varying intercept and slope for prime by pid
+
+```{r model-formula}
+formula = bf(outcome ~ 0 + timebin:prime +
+               (0 + timebin:prime | pid))
+```
+
+## check the priors available ##
+
+```{r priors}
+get_prior(formula,
+          data = data, family = bernoulli(link = "cloglog"))
+```
+
+## visualise priors ##
+
+For an index coding model, priors will be set differently to how they were set in the earlier tutorials, because the earlier tutorials did not use an index coding approach.
+Using an index coding approach, each condition has a prior for the mean of that condition.
+And as we know from plot 3.1 above, the condition averages across the six time bins selected range from near zero on the hazard probability scale to 0.50. 
+And in principle, the more general case could involve any hazard value from zero to 1.
+Therefore, if we try to stick to a weakly informative approach to priors [see here for more details](https://github.com/stan-dev/stan/wiki/prior-choice-recommendations), then priors for the cloglog values should cover a fairly broad range.
+
+For a look at values across the cloglog scale and how it compares to the logit scale, see [here](https://towardsdatascience.com/a-gentle-introduction-to-complementary-log-log-regression-8ac3c5c1cd83).
+
+Ok, so let's take a look at some cloglog values at the lower and upper end of the probability scale.
+
+```{r cloglog-vals}
+cloglog_vals = clogloglink(c(0.01,0.99))
+cloglog_vals
+
+# -4.600149  1.527180
+```
+
+and now take a look at some distributions
+
+```{r visualise-priors}
+visualize("normal(0, 2)", "normal(0, 1)", "normal(0, 0.5)",
+          xlim = c(-4, 4))
+```
+
+(0,2) looks like it covers all of our likely values, even if it is too broad.
+
+## set priors ##
+
+We set the prior for the mean as (normal(0,2)) and then we use some default recommendations from McElreath 2020 for the rest.
+
+```{r set-priors}
+priors <- c(
+  set_prior("normal(0, 2)", class = "b"),
+  set_prior("normal(0, 1)", class = "sd"),
+  set_prior("lkj(2)", class = "cor")
+)
+```
+
+## run the model ##
+
+This model takes ~17 minutes to build on a 2020 MacBook Pro (2 GHz Quad-Core Intel Core i5).
+For this tutorial, this chunk is skipped to save time. 
+Change eval=TRUE in the code chunk to run this model.
+
+```{r build-model, eval=FALSE}
+plan(multicore)
+bi <- brm(formula = formula,
+        data = data, family = bernoulli(link = "cloglog"),
+        prior = priors,
+        iter = 2000, warmup = 1000, cores = 8, chains = 4,
+        control = list(adapt_delta = 0.95),
+        save_pars = save_pars(all=TRUE),
+        seed = 123,
+        init = 0.01,
+        file = "Tutorial_4_planning/models/bi")
+summary(bi)
+```
+
+The model built without any concerns, errors or warnings.
+
+At this point, you would normally do the following:
+
+- model checking by looking at the convergence of chains and model diagnostics. 
+We'll skip this step for now, just to move on to the main focus of planning and data simulation.
+
+- Summarise and visualise the parameter estimates and/or posterior predictions.
+Below, we quickly summarise fixed effects from the model in cloglog values, and we also convert them to hazard values, which tend to be far easier to interpret.
+
+## visualise fixed effect parameters from model bi ##
+
+read in the existing model object (if you did not build the model yourself)
+
+```{r load-model-object}
+bi <- readRDS("Tutorial_4_planning/models/bi.rds")
+summary(bi)
+```
+
+wrangle the posterior distribution
+
+```{r wrangle-posterior}
+## take the posterior draws
+post <- as_draws_df(bi) %>% 
+  select(-lp__) %>% 
+  as_tibble()
+
+## create a summary, and use medians with robust = TRUE
+post_summary <- posterior_summary(bi, robust = TRUE)
+
+## just look at the fixed effects
+post_qi_b <- post %>%
+  select(starts_with("b_")) %>%
+  pivot_longer(everything()) %>%
+  group_by(name) %>% 
+  median_qi(value)
+head(post_qi_b)
+```
+
+a quick plot using posterior samples and tidybayes
+
+```{r plot-fixed}
+tidy_fixed <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% # select and rename in simpler labels.
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) %>% # move from wide to long
+  mutate(key = factor(name, levels=unique(name)),
+         bin = rep(1:6, length.out = 48000),
+         prime = rep(c("con", "inc"), each = 6, length.out = 48000),
+         bin=factor(bin),
+         prime=factor(prime))
+head(tidy_fixed)
+tail(tidy_fixed)
+
+check.labels <- tidy_fixed %>% 
+  distinct(key, bin, prime) 
+check.labels
+
+# plot
+p4.1 <- ggplot(tidy_fixed, aes(x = bin, y = value, fill=prime)) +  
+  stat_halfeye(alpha=0.7) +
+  labs(title = "Model bi cloglog",
+       x = "timebin", y = "cloglog") +
+  scale_fill_brewer(palette = "Dark2") 
+p4.1
+
+ggsave ("Tutorial_4_planning/figures/index_cloglog.jpeg",
+        width = 6, height = 4)
+```
+
+calculate hazards per condition per bin in the posterior dist
+
+```{r calculate-bi-hazard}
+tidy_haz <- tidy_fixed %>%
+  select(chain, iter, draw, bin, prime, value) %>% 
+  mutate(hazard = exp(value))
+head(tidy_haz)
+```
+
+plot hazards
+
+```{r plot-bi-hazards}
+p4.2 <- ggplot(tidy_haz, aes(x = bin, y = hazard,
+                                    fill=prime)) +  
+  stat_halfeye(alpha=0.7) +
+  labs(title = "hazard estimates",
+       x = "timebin", y = "hazard") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) 
+p4.2
+
+ggsave ("Tutorial_4_planning/figures/index_hazard.jpeg",
+        width = 6, height = 4)
+```
+
+warning about removing rows.
+
+# 5. Simulate a single dataset #
+
+Note: For multi-level data, data simulation can be quite a convoluted process, and we therefore strongly recommend reading the power blogs by Solomon Kurz and working through the faux package vignettes, before getting stuck into the below code.
+
+read in a prior model, if not already loaded.
+
+```{r load-model-object-sim, eval=FALSE}
+bi <- readRDS("Tutorial_4_planning/models/bi.rds")
+summary(bi)
+```
+
+Define some parameters
+
+```{r define-sim-parameters}
+# define parameters
+# specify some features of the design
+subj_n = 10  # number of subjects
+rep_n = 200 # number of trial repeats 
+cond_n = 2 # number of conditions
+bin_n = 6 # number of time bins
+
+## set fixed effects
+## as a reminder
+fixef <- as_tibble(fixef(bi), rownames = "term") %>%
+  mutate(across(where(is.double), \(x) round(x, 2)))
+fixef
+
+## b1 to b6 = timebin 1 to 6.
+## c = con
+b1c = -4.66
+b2c = -3.24
+b3c = -2.32
+b4c = -1.49
+b5c = -1.21
+b6c = -0.60
+
+## i = inc
+b1i = -4.37
+b2i = -3.21
+b3i = -2.37
+b4i = -2.27
+b5i = -3.04
+b6i = -2.45
+
+## set varying effects by pid
+## as a reminder
+varcor <- VarCorr(bi)
+glimpse(varcor)
+str(varcor)
+
+## extract sd
+sd <- as_tibble(varcor$pid$sd, rownames = "term") %>%
+  mutate(across(where(is.double), \(x) round(x, 2)))
+sd
+
+u1c_sd = 1.06   # 
+u2c_sd = 1.34   # 
+u3c_sd = 0.60   # 
+u4c_sd = 0.47   # 
+u5c_sd = 0.58   #
+u6c_sd = 0.72   #
+
+u1i_sd = 1.17   # 
+u2i_sd = 1.05   # 
+u3i_sd = 0.91   # 
+u4i_sd = 1.47   # 
+u5i_sd = 1.34   #
+u6i_sd = 0.62   # 
+```
+
+make a correlation matrix
+
+```{r define-cor-mat}
+# varcor <- VarCorr(b3c_out)
+# varcor
+# str(varcor)
+
+## take each one
+sd <- varcor$pid$sd
+cor <- varcor$pid$cor
+cov <- varcor$pid$cov
+
+## make it tidy
+cor_mat_tidy <- as_tibble(cor, rownames = "term") %>% ## 
+  rename_with(~str_replace_all(.x, '\\.', '_')) %>% 
+  pivot_longer(-term,
+               values_to = "value",
+               names_to = c("parameter", "term2"),
+               names_pattern = "([a-zA-Z_\\d?]+)_([a-zA-Z:\\d?]+)") %>% ## pivot longer and use names_pattern to select the parts of the column names to separate on.
+the above is regex for each bit between the separator ("-").
+e.g., ([a-zA-Z]+\\d+).
+this looks for letters and digits before the first "_".
+This (\\d?) looks for the possibility of digits.
+  ## the rest of this code just turns the variables into what we want
+  filter(parameter == "Estimate") %>%
+  select(-parameter) %>% 
+  pivot_wider(names_from = "term2",
+              values_from = "value") %>% 
+  select(-term)
+head(cor_mat_tidy)
+str(cor_mat_tidy)
+
+## check with the model summary
+# summary(bi)
+
+## remove names and make a matrix as faux{} likes it like that
+## full cor mat
+cor_mat <- unname(as.matrix(cor_mat_tidy))
+cor_mat
+```
+
+setup the data structure
+
+```{r setup-sim-d1}
+# make it reproducible
+set.seed(1)
+
+d1 <- add_random(subj = subj_n, rep = rep_n) %>%
+  add_within("subj", condition = c("cond1", "cond2")) %>%
+  add_contrast("condition", "treatment", add_cols = TRUE, 
+               colnames = c("cond")) %>%
+  add_within("rep", bin = 1:bin_n) %>%
+  ### create new conditions here that respect the twelve levels b1c, b1i etc.
+  mutate(con = if_else(condition == "cond1", 1, 0),
+         inc = if_else(condition == "cond2", 1, 0)) %>% 
+  mutate(bin1 = if_else(bin == 1, 1, 0),
+         bin2 = if_else(bin == 2, 1, 0),
+         bin3 = if_else(bin == 3, 1, 0),
+         bin4 = if_else(bin == 4, 1, 0),
+         bin5 = if_else(bin == 5, 1, 0),
+         bin6 = if_else(bin == 6, 1, 0)) %>% 
+  # add random effects 
+  add_ranef("subj", u1c = u1c_sd, u2c = u2c_sd, u3c = u3c_sd, u4c = u4c_sd,
+            u5c = u5c_sd, u6c = u6c_sd,
+            u1i = u1i_sd, u2i = u2i_sd, u3i = u3i_sd, u4i = u4i_sd,
+            u5i = u5i_sd, u6i = u6i_sd, 
+            .cors = cor_mat) %>% 
+  # calculate logit
+  mutate(cloglog = (b1c + u1c) * con * bin1 + (b2c + u2c) * con * bin2 + 
+           (b3c + u3c) * con * bin3 + (b4c + u4c) * con * bin4  + 
+           (b5c + u5c) * con * bin5 + (b6c + u6c) * con * bin6 +
+           (b1i + u1i) * inc * bin1 + (b2i + u2i) * inc * bin2 + 
+           (b3i + u3i) * inc * bin3 + (b4i + u4i) * inc * bin4  + 
+           (b5i + u5i) * inc * bin5 + (b6i + u6i) * inc * bin6) %>% 
+  # calculate inverse cloglog
+  mutate(prob = clogloglink(cloglog, inverse = TRUE)) %>%  
+  # calculate event
+  mutate(event = rbinom(n(), 1, prob)) 
+head(d1)
+str(d1)
+summary(d1)
+glimpse(d1)
+
+## save initial data
+# write_csv(d1, "Tutorial_4_planning/data/d1.csv") #
+```
+
+summarise the data produced so far
+
+```{r summarise-sim-d1}
+d1_pid <- d1 %>%
+  group_by(subj, condition, bin) %>%
+  summarise(
+    across(cloglog:event, \(x) mean(x, na.rm = TRUE)),
+    .groups = "drop"
+    )
+d1_pid
+
+d1_group <- d1 %>%
+  group_by(condition, bin) %>%
+  summarise(
+    across(cloglog:event, \(x) mean(x, na.rm = TRUE)),
+    .groups = "drop"
+    )
+d1_group
+```
+
+a quick plot
+
+cloglog
+
+```{r plot-d1-clog}
+p5.1 <- ggplot(d1_group, aes(x=bin, y = cloglog, colour = condition)) +
+   geom_line(aes(group = condition)) + 
+   geom_point() +
+   geom_jitter(data=d1_pid, alpha = 0.5, width = 0.1, height = 0) +
+   # scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+   scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   labs(title = "cloglog - simulation d1 (N=10)",
+        x = "time bins")
+p5.1
+```
+
+event prob
+
+```{r plot-d1-prob}
+p5.2 <- ggplot(d1_group, aes(x=bin, y = event, colour = condition)) +
+   geom_line(aes(group = condition)) + 
+   geom_point() +
+   geom_jitter(data=d1_pid, alpha = 0.5, width = 0.1, height = 0) +
+   scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+   scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   labs(title = "event prob - simulation d1 (N=10)",
+        x = "time bins")
+p5.2
+```
+
+convert the data into outcome data (this next chunk can be incorporated into the data structure chunk above, I'm just doing separately here, so that you can see what is initially generated)
+
+```{r calculate-outcome-d1}
+d1_out <- d1 %>% 
+  group_by(subj, rep, condition) %>%
+  mutate(cumsum = cumsum(event),
+         outcomel = event==1 & cumsum(event) < 2,
+         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
+                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
+  drop_na(outcomen) %>% 
   ungroup()
-}
+head(d1_out)
 
-# function setup_lt sets up a life table for each level 2 independent variables
-setup_lt_2IV <- function(ptb){
-  ptb %>% mutate(event = str_c("event", event)) %>%
-          group_by(condition1,condition2,period) %>% 
-          count(event) %>% 
-          ungroup() %>% 
-          pivot_wider(names_from = event,
-                      values_from = n) %>% 
-          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
-                 event1 = ifelse(is.na(event1),0,event1),
-                 risk_set = event0 + event1) %>% # define the risk set
-          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
-          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
-          group_by(condition1,condition2) %>%
-          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
-                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
-                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+## save outcome data
+# write_csv(d1_out, "Tutorial_4_planning/data/d1_out.csv")
+```
+
+summarise the outcome data, which is produced based on the prob value, right?
+
+```{r summary-d1-out}
+d1_out_pid <- d1_out %>% 
+  group_by(subj, condition, bin) %>% 
+  summarise(outcome = mean(outcomen)) %>% 
+  mutate(prime = if_else(condition == "cond1", "con", "inc"),
+         prime = factor(prime))
+d1_out_pid
+
+d1_out_group <- d1_out %>% 
+  group_by(condition, bin) %>% 
+  summarise(outcome = mean(outcomen)) %>% 
+  mutate(prime = if_else(condition == "cond1", "con", "inc"),
+         prime = factor(prime))
+d1_out_group
+```
+
+plot outcome using the prime variable to make it comparable to the raw data
+
+```{r plot-d1-outcome}
+p5.3 <- ggplot(d1_out_group, aes(x = bin, y = outcome, colour = prime)) +
+  geom_line(aes(group = prime)) + 
+  geom_point() +
+  geom_jitter(data=d1_out_pid, alpha=0.5, width = 0.1, height = 0) +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+  scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
+  scale_colour_brewer(palette = "Dark2") +
+  labs(title = "simulation d1 (N=10)",
+       x = "time bins",
+       y = "outcome")
+p5.3
+```
+
+compare the raw data to the simulated data for a single simulated dataset
+
+```{r plot-raw-plus-d1}
+p5.4 <- (p3.1 / p5.3) +
+   plot_layout(guides = 'collect',
+               axes = 'collect')
+p5.4
+
+ggsave ("Tutorial_4_planning/figures/raw_vs_d1.jpeg",
+        width = 8, height = 8)
+```
+
+ok, this looks reasonably sensible.
+
+# 6. Think about effect sizes
+
+take a look at avg hazards and hazard ratios at the group level in the original data
+
+```{r plot-raw}
+p3.1
+```
+
+this is the descriptive data per bin and condition for hazards
+probability of a response/event, given that it has not happened yet
+
+this is the group level summary data plotted in dat_plot
+
+```{r plot-raw-summary}
+data_group
+```
+
+calculate the hazard ratio
+
+```{r calculate-hazard}
+hazard <- data_group %>% 
+  pivot_wider(names_from = "prime",
+              values_from = "outcome") %>% 
+  mutate(diff = con-inc,
+         ratio = inc/con)
+hazard
+
+# timebin     con    inc     diff ratio
+#   <fct>     <dbl>  <dbl>    <dbl> <dbl>
+# 1 4       0.00975 0.0139 -0.00416 1.43 
+# 2 5       0.0478  0.0423  0.00551 0.885
+# 3 6       0.0901  0.0913 -0.00121 1.01 
+# 4 7       0.188   0.115   0.0732  0.611
+# 5 8       0.252   0.0495  0.203   0.196
+# 6 9       0.401   0.0848  0.316   0.211
+```
+
+key hazard and hazard ratio values of interest are as follows....
+
+```{r key-hazards, eval=FALSE}
+## bin4
+## 0.61 - 39% reduction in hazard from con to inc
+## 0.115/0.188
+
+## bin5
+## 0.20 - 80% reduction in hazard from con to inc
+## 0.0495/0.252
+
+## bin6
+## 0.21 - 79% reduction in hazard from con to inc
+## 0.401/0.0848
+```
+
+Ok, if we wanted to simulate a bunch of effect sizes, we could think about the following.
+
+Let's just focus on bin 6 to keep things simple.
+And keep in mind that we do need to think about the absolute hazard values, as well as the ratio.
+
+So con is close to 0.1, so let's fix it there as a round number and calculate some deviations from it.
+
+Given that the effect of interest was ~80% reduction in Sven's 2016 paper, let's play around with a range of effect sizes from e.g., 0.25, 0.5, 0.75.
+These would translate to a 75% reduction, 50% reduction and 25% reduction from con to inc.
+
+```{r sim1-bin6-effect-sizes}
+bin6 <- tibble(
+  effect_id = 1:3,
+  inc = rep(0.1, times=3),
+  ratio = c(0.25, 0.5, 0.75),
+  con = inc/ratio,
+  inc_clog = clogloglink(inc),
+  con_clog = clogloglink(con)
+)
+bin6
+
+#  effect_id   inc ratio   con inc_clog con_clog
+#       <int> <dbl> <dbl> <dbl>    <dbl>    <dbl>
+# 1         1   0.1  0.25 0.4      -2.25   -0.672
+# 2         2   0.1  0.5  0.2      -2.25   -1.50 
+# 3         3   0.1  0.75 0.133    -2.25   -1.94 
+```
+
+# 7. Create a function to iterate through and simulate many datatsets
+
+create a function to simulate data
+These values are from the initial model and data from Sven's paper, but they can be modified when running the function, which we do below.
+
+```{r index-sim-function}
+index_sim <- function(subj_n = 10, rep_n = 200, bin_n = 6,  # these can be changed when calling the function
+                b1c = -4.66, b2c = -3.24, b3c = -2.32, b4c = -1.49, b5c = -1.21, b6c = -0.6, 
+                b1i = -4.37, b2i = -3.21, b3i = -2.37, b4i = -2.27, b5i = -3.04, b6i = -2.45, # fixed effects
+                u1c_sd = 1.06, u2c_sd = 1.34, u3c_sd = 0.60, u4c_sd = 0.47, u5c_sd = 0.58, u6c_sd = 0.72, 
+                u1i_sd = 1.17, u2i_sd = 1.05, u3i_sd = 0.91, u4i_sd = 1.47, u5i_sd = 1.34, u6i_sd = 0.62, # varying effects
+                cors = cor_mat,   # correlations between effects
+                ... # helps the function work with pmap() below
+                ) {
+  # set up data structure
+  data <- add_random(subj = subj_n, rep = rep_n) %>%
+  add_within("subj", condition = c("cond1", "cond2")) %>%
+  add_contrast("condition", "treatment", add_cols = TRUE, 
+               colnames = c("cond")) %>%
+  add_within("rep", bin = 1:bin_n) %>%
+  ### create new conditions here that respect the twelve levels b1c, b1i etc.
+  mutate(con = if_else(condition == "cond1", 1, 0),
+         inc = if_else(condition == "cond2", 1, 0)) %>% 
+  mutate(bin1 = if_else(bin == 1, 1, 0),
+         bin2 = if_else(bin == 2, 1, 0),
+         bin3 = if_else(bin == 3, 1, 0),
+         bin4 = if_else(bin == 4, 1, 0),
+         bin5 = if_else(bin == 5, 1, 0),
+         bin6 = if_else(bin == 6, 1, 0)) %>% 
+  # add random effects 
+  add_ranef("subj", u1c = u1c_sd, u2c = u2c_sd, u3c = u3c_sd, u4c = u4c_sd,
+            u5c = u5c_sd, u6c = u6c_sd,
+            u1i = u1i_sd, u2i = u2i_sd, u3i = u3i_sd, u4i = u4i_sd,
+            u5i = u5i_sd, u6i = u6i_sd, 
+            .cors = cor_mat) %>% 
+  # calculate logit
+  mutate(cloglog = (b1c + u1c) * con * bin1 + (b2c + u2c) * con * bin2 + 
+           (b3c + u3c) * con * bin3 + (b4c + u4c) * con * bin4  + 
+           (b5c + u5c) * con * bin5 + (b6c + u6c) * con * bin6 +
+           (b1i + u1i) * inc * bin1 + (b2i + u2i) * inc * bin2 + 
+           (b3i + u3i) * inc * bin3 + (b4i + u4i) * inc * bin4  + 
+           (b5i + u5i) * inc * bin5 + (b6i + u6i) * inc * bin6) %>% 
+  # calculate inverse cloglog
+  mutate(prob = clogloglink(cloglog, inverse = TRUE)) %>%  
+  # calculate event
+  mutate(event = rbinom(n(), 1, prob)) %>% 
+  group_by(subj, rep, condition) %>%
+  mutate(cumsum = cumsum(event),
+         outcomel = event==1 & cumsum(event) < 2,
+         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
+                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
+  drop_na(outcomen) %>%
   ungroup()
-}
 
-# function calc_ca calculates the conditional accuracies 
-calc_ca_2IV <- function(df){
-  df %>% filter(right_censored==0) %>%
-         group_by(condition1,condition2,drt,cens_time,bin_width) %>%
-         summarize(ca = mean(acc) %>% round(digits = 2),
-                   n = n(),
-                   .groups = 'drop') %>%
-         ungroup() %>%
-         mutate(period = drt,
-                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
-         select(-drt)
-}
-
-# function join_lt_ca_2IV joins the conditional accuracies to the life tables when there are two independent variables
-join_lt_ca_2IV <- function(df1,df2){df1 %>% left_join(df2, join_by(condition1,condition2,period))}
-
-# function extract_median_2IV is used to extract the S(t).50 quantiles when there are two independent variables
-extract_median_2IV <- function(df){
-  above_pct50 <- df %>% 
-      group_by(condition1,condition2) %>%
-      filter(survival >= .5) %>% 
-      slice(n()) # take last row
-  below_pct50 <- df %>% 
-      group_by(condition1,condition2) %>%
-      filter(survival < .5) %>% 
-      slice(1) # take first row
-  # pull period above
-  period_above <- pull(above_pct50, period)
-  # pull survivor function values
-  survival_above <- pull(above_pct50, survival)
-  survival_below <- pull(below_pct50, survival)
-  # estimate median by interpolation
-  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
-}
-
-# function plot_eha_2IV plots the hazard, survivor, and ca(t) functions, when there are two independent variables
-plot_eha_2IV <- function(df,subj,haz_yaxis){ 
-  library(patchwork)
-  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
-  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
-  median_period <- extract_median_2IV(df)
-  n_iv1 <- nlevels(df$condition1) 
-  n_iv2 <- nlevels(df$condition2) 
-  n_conditions <- n_iv1*n_iv2
-  data_median <- c()
-  for(i in 1:n_conditions){
-    data_median <- append(data_median, c(median_period[i], median_period[i]))
-  }
-  # Create correctly named factors (here: prime and mask) and specify the same levels and labels as above
-  data_medians <- tibble(period= data_median,
-                         survival = rep(c(.5, 0),n_conditions),
-                         prime =  factor(rep(rep(1:n_iv2, each=2),n_iv1),levels=c(1:3),labels=c("blank","congruent","incongruent")),
-                         mask = factor(rep(1:n_iv1, each=2*n_iv2),levels=c(1:4),labels=c("nomask", "relevant", "irrelevant","lines")))
-
-# plot the hazard functions
-p1 <-df %>% 
-  mutate(mask = condition1,       # Create correctly named factor
-         prime = condition2) %>%  # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=hazard)) +
-  geom_point(aes(y=hazard), size=1) + labs(color="Prime") +
-  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits = c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
-  theme(legend.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-# plot the survivor functions
-p2 <-df %>%
-  mutate(mask = condition1,      # Create correctly named factor
-         prime = condition2) %>% # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=survival), show.legend = F) +
-  geom_point(aes(y=survival), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
-  # add vertical lines to indicate the estimated median RTs with geom_path()
-  geom_path(aes(x=period, y=survival),
-            data = data_medians, 
-            linetype = 3, show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits=c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="", y="S(t)") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-# plot the conditional accuracy functions
-p3 <-df %>%
-  mutate(mask = condition1,       # Create correctly named factor
-         prime = condition2) %>%  # Create correctly named factor
-  ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
-  geom_line(aes(y=ca), show.legend = F) +
-  geom_point(aes(y=ca), size=1, show.legend = F) +
-  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
-  scale_x_continuous(breaks = c(1:(cutoff/binsize)), labels=c(1:(cutoff/binsize)*binsize),
-                     limits=c(1,cutoff/binsize)) +
-  scale_y_continuous(limits = c(0,1)) +
-  labs(x="Time bin t's endpoint (ms)", y="ca(t)") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor
-
-p1/p2/p3
+  # glimpse(data) # only use this when testing the code
 }
@@
 
-Apply the functions
+Here’s a quick example of how our function works. You can change these parameters
+and create some different data.
 
-```{r apply-functions}
-# create a data frame with two columns (pid and data); the latter contains a list of nested data frames
-data_nested <- data_wr %>% group_nest(pid)
-# apply the functions using map() and map2()
-data_final <- data_nested %>% 
-                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
-                     mutate(ptb_data  = map(censored, ptb)) %>%           
-                     mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%      # create life tables without ca for 2 independent variables
-                     mutate(condacc   = map(censored, calc_ca_2IV)) %>%       # calculate conditional accuracy for 2 independent variables
-                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  # add ca to lifetables for 2 independent variables
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV,1))  # create plots for 2 independent variables
+```{r test-index-sim}
+test_sim <- index_sim(subj_n = 10, rep_n = 200) # if you uncomment glimpse above,
+# it will let you glimpse the data that's generated. this is useful for checking / testing code purposes.
 ```
 
-# Extract info, and save plot for each subject
+create a function to summarise the simulated data
 
-```{r extract-info, warning=F}
-# view original data (column 2 in data_final) for subject 1:
-pluck(data_final, 2, 1) %>% print(n=20)
-
-# view person-trial-bin data set (column 4) for subject 1:
-pluck(data_final,4,1) %>% print(n=30)
-
-# view lifetable (column 7) for a specific combination of condition levels for subject 1:
-pluck(data_final, 7, 1) %>% filter(condition1=="nomask" & condition2 == "blank")
-
-# view plot (column 8) for subject 1:
-pluck(data_final, 8, 1)
-
-# view the median RTs in a tibble:
-median_bin <- map(data_final$lifetable_ca, extract_median_2IV)
-binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
-median_rt  <- modify_depth(median_bin,1,~.x*binsize)
-median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
-median_tb %>% print(n=6*12)
+```{r index-summary-function}
+index_summary <- function(df) {
+  df %>% 
+  group_by(subj, condition, bin) %>% 
+  summarise(n = n(), 
+            sum = sum(outcomen),
+            mean_outcome = mean(outcomen, na.rm = TRUE),
+            sd_outcome = sd(outcomen, na.rm = TRUE),
+            sem = (sd_outcome/sqrt(length(unique(rep)))),
+            .groups="drop") 
+}
 ```
 
-```{r save-plots, warning=F}
-# save the plot for each subject
-map2(paste0("Tutorial_4_descriptive_2IV/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 8, dpi = 600)
+test it
+
+```{r test-index-summary}
+test_summary <- index_summary(test_sim)
+head(test_summary)
 ```
+
+# 8. Simulate data for a range of parameters
+
+vary trial count per condition and effect size, keep pid fixed at 10.
+This will take some time, probably hours...go get a beer/coffee.
+For the purposes of this tutorial, eval is set to false in the below chunks, to save time.
+
+```{r run-sim1, eval=FALSE}
+plan(multicore)
+x1 <- crossing(
+  exp = 1:1000, # number of experiment replicates
+  subj_n = 10, # range of subject N
+  rep_n = c(100, 200, 400),
+  b6i = -2.25,
+  b6c = c(-1.94, -1.50, -0.67)
+) %>%
+  mutate(d = pmap(., index_sim)) %>% 
+  mutate(s = map(d, index_summary)) %>% 
+  select(-d)
+```
+
+# 9. Summarise the simulated data
+
+```{r summarise-sim1, eval=FALSE}
+sx1 <- x1 %>%
+  unnest(s) %>%
+  mutate(exp=factor(exp),
+         subj_n=factor(subj_n),
+         rep_n=factor(rep_n),
+         b6c = factor(b6c,
+                      levels = c("-1.94", "-1.5", "-0.67"),
+                      labels = c("25%", "50%", "75%")),
+         subj=factor(subj),
+         bin=factor(bin))
+sx1
+```
+
+take a look
+
+```{r sim1-quick-look, eval=FALSE}
+head(sx1)
+tail(sx1)
+
+sx1 %>% 
+  distinct(b6c)
+```
+
+save out a file
+
+```{r save-sim1, eval=FALSE}
+## save the first simulation
+write_csv(sx1, "Tutorial_4_planning/data/sim1/sim1_data.csv")
+```
+
+## calculate summary data ##
+
+read in the file, if it is already computed
+
+```{r load-sim1}
+sx1 <- read_csv("Tutorial_4_planning/data/sim1/sim1_data.csv") %>%
+  mutate(exp=factor(exp),
+         subj_n=factor(subj_n),
+         rep_n=factor(rep_n),
+         b6c = factor(b6c,
+                      levels = c("-1.94", "-1.5", "-0.67"),
+                      labels = c("25%", "50%", "75%")),
+         subj=factor(subj),
+         bin=factor(bin))
+sx1
+```
+
+at the exp level
+
+```{r exp-summary-sim1}
+sx1_exp <- sx1 %>% 
+  group_by(exp, subj_n, rep_n, b6c, condition, bin) %>% 
+  summarise(exp_mean_outcome = mean(mean_outcome, na.rm = TRUE),
+            exp_mean_sum=mean(sum, na.rm = TRUE),
+            exp_sum=sum(sum),
+            n = length(unique(subj)), 
+            sd=sd(mean_outcome, na.rm = TRUE),
+            sem = (sd/sqrt(length(unique(subj)))),
+            .groups = "drop")
+sx1_exp
+```
+
+at the sim level
+
+```{r sim-summary-sim1}
+sx1_sim <- sx1_exp %>% 
+  group_by(subj_n, rep_n, b6c, condition, bin) %>% 
+  summarise(sim_mean_outcome = mean(exp_mean_outcome, na.rm = TRUE),
+            sim_mean_sum = mean(exp_mean_sum, na.rm = TRUE),
+            sim_sum = mean(exp_sum, na.rm = TRUE),
+            n = length(unique(exp)), 
+            sd=sd(exp_mean_outcome, na.rm = TRUE),
+            sem = (sd/sqrt(length(unique(exp)))),
+            .groups = "drop")
+sx1_sim
+```
+
+a quick plot of the data
+
+```{r sim1-violin}
+p9.1 <- ggplot(sx1_sim, 
+                aes(x = bin, y = sim_mean_outcome, 
+                    colour = condition)) +
+  geom_jitter(data=sx1_exp, aes(y = exp_mean_outcome),
+              alpha=0.5, width = 0.1, height = 0) +
+  geom_line(aes(group = condition)) + 
+  geom_point(size=2, colour = "black") +
+  geom_errorbar(aes(ymin = sim_mean_outcome-sem*1.96, ymax = sim_mean_outcome+sem*1.96),
+                width=.2, colour = "black") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+  labs(x="timebin",
+       y="outcome") + 
+  facet_grid(fct_rev(rep_n)~b6c)
+p9.1
+
+ggsave ("Tutorial_4_planning/figures/sim1/sim_violin.jpeg",
+        width = 10, height = 8, dpi=800)
+```
+
+ok, look at some values in bin 6
+
+at the exp level
+
+```{r sim1-bin6-exp-values}
+sx1_exp_check <- sx1_exp %>%
+  filter(bin %in% c(6)) %>% 
+  select(exp, rep_n, b6c, condition, bin, exp_mean_outcome) %>%
+  pivot_wider(names_from = "condition",
+              values_from = "exp_mean_outcome") %>% 
+  group_by(exp, rep_n, b6c, bin) %>% 
+  mutate(ratio = cond2/cond1) 
+head(sx1_exp_check)
+```
+
+and at the sim level
+
+```{r sim1-bin6-sim-values}
+sx1_sim_check <- sx1_exp_check %>%
+  group_by(rep_n, b6c, bin) %>% 
+  summarise(sim_ratio = mean(ratio),
+            sim_sd=sd(ratio),
+            sim_sem=sim_sd/sqrt(1000),
+            .groups = "drop")
+head(sx1_sim_check)
+```
+
+plot ratio values
+
+```{r plot-ratio-values}
+p9.2 <- ggplot(sx1_sim_check, 
+                aes(x = b6c, y = sim_ratio)) +
+  geom_point(size=1.5) +
+  geom_errorbar(aes(ymin = sim_ratio-sim_sem*1.96, ymax = sim_ratio+sim_sem*1.96)) + 
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "red") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+  labs(x="timebin",
+       y="outcome") + 
+  facet_wrap(~rep_n)
+p9.2
+```
+
+# 10. calculate power / precision #
+
+basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound quantiles exclude zero?
+This of course is not a model-based estimate.
+But I think it should give a good guide to what we might expect.
+
+## calculate difference scores ##
+
+at the exp level
+
+```{r sim1-calculate-diff-scores}
+sx1_diff_exp <- sx1 %>%
+  pivot_wider(id_cols = c(exp, subj_n, rep_n, b6c, subj, bin),
+              names_from = "condition",
+              values_from = "mean_outcome") %>% 
+  mutate(diff = cond1 - cond2) %>%
+  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
+  summarise(exp_mean_diff = mean(diff, na.rm = TRUE),
+            exp_sd = sd(diff, na.rm = TRUE),
+            n = n(), # n here is the total trials per condition per pid
+            exp_sem = (exp_sd/sqrt(n)),
+            exp_ci = 1.96*exp_sem,
+            exp_dz = exp_mean_diff/exp_sd,
+            .groups = "drop") 
+head(sx1_diff_exp)
+tail(sx1_diff_exp)
+```
+
+at the group/sim level
+
+```{r sim1-summary-diff-scores}
+sx1_diff_sim <- sx1_diff_exp %>% 
+  group_by(subj_n, rep_n, b6c, bin) %>% 
+  summarise(mean_diff = mean(exp_mean_diff, na.rm = TRUE),
+            sd = sd(exp_mean_diff, na.rm = TRUE),
+            n=n(),
+            sem = (sd/sqrt((n))),
+            ci = 1.96*sem) 
+sx1_diff_sim
+```
+
+## plot ##
+
+violin
+
+diff in original units
+
+```{r sim1-plot-sim-diffs}
+p10.1 <- ggplot(sx1_diff_exp, aes(x=bin, y = exp_mean_diff,
+                                colour = bin, fill = bin)) +
+   geom_jitter(alpha = 0.5, width = 0.1) +
+   geom_violin(alpha = 0.7) +
+   geom_point(data = sx1_diff_sim, 
+             aes(y = mean_diff), size = 3, position=pd2, colour="black") +
+   geom_errorbar(data = sx1_diff_sim,
+                aes(y = mean_diff, ymin = mean_diff-ci, ymax = mean_diff+ci),
+                width=.2, position=pd2, colour = "black") +
+   geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   theme(legend.position = "none") +
+   ylab("mean outcome") +
+   ggtitle("difference score (cond1-cond2)") +
+   facet_grid(fct_rev(rep_n)~b6c)
+p10.1
+
+ggsave("Tutorial_4_planning/figures/sim1/sim_diffs.jpeg",
+       width = 10, height = 8, dpi=800)
+```
+
+plot each exp's difference score and associated 95% interval
+
+create some factors and new variables
+
+```{r sim1-wrangle-diffs}
+sx1_diff_exp_2 <- sx1_diff_exp %>%
+  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
+  mutate(lower = exp_mean_diff-exp_ci,
+         upper = exp_mean_diff+exp_ci,
+         above_zero = if_else(lower > 0, "yes", "no"), 
+         above_zero = factor(above_zero, levels = c("no", "yes")))
+sx1_diff_exp_2
+```
+
+calculate "power" in a quick and dirty way based on 95% CI
+
+```{r sim1-calculate-power}
+power_x1 <- sx1_diff_exp_2 %>%
+  group_by(subj_n, rep_n, b6c, bin) %>%
+  mutate(check = ifelse(lower > 0, 1, 0)) %>%
+  summarise(power = mean(check, na.rm = TRUE)) %>% 
+  filter(bin %in% c(6))
+power_x1
+```
+
+plot power
+
+```{r sim1-power-tile}
+p10.2 <- ggplot(power_x1, aes(x = b6c, y=rep_n, fill = power)) +
+  geom_tile() +
+  geom_text(aes(label = sprintf("%.3f", power)), color = "white", size = 10) +
+  scale_fill_viridis_c(limits = c(0, 1)) +
+  facet_wrap(~bin) +
+  labs(y="rep_n", x="b6c")
+p10.2
+#
+ggsave ("Tutorial_4_planning/figures/sim1/power.jpeg",
+        width = 10, height = 6, dpi=800)
+```
+
+join power to the df
+
+```{r sim1-join-dfs}
+sx1_diff_power <- sx1_diff_exp_2 %>%
+  filter(bin %in% c(6)) %>% 
+  inner_join(power_x1, by = c("subj_n", "rep_n", "b6c", "bin")) %>% 
+  mutate(power = round(power * 100, 2)) 
+head(sx1_diff_power)
+```
+
+plot
+
+```{r sim1-plot-pointrange}
+p10.3 <- sx1_diff_power %>%
+  ggplot(aes(x = exp, y = exp_mean_diff, ymin = lower, ymax = upper)) +
+  geom_pointrange(fatten = 1/2, aes(colour=above_zero)) +
+  geom_hline(yintercept = 0, colour = "red") +
+  scale_colour_manual(values=c("darkgrey","black")) +
+  geom_text(aes(x=700, y=-0.35, label = sprintf("%.1f%s", power, "% power")), 
+            color = "darkgrey", size = 5) +
+  theme(legend.position = "none") +
+  labs(x = "sim # (i.e., simulation index)",
+       y = "hazard difference") +
+  scale_x_discrete(breaks = c(250,500,750,1000)) +
+  facet_grid(fct_rev(rep_n)~b6c)
+p10.3
+
+ggsave ("Tutorial_4_planning/figures/sim1/bin6_diffs.jpeg",
+        width = 10, height = 8, dpi=800)
+```
+
+plot power as a bar plot
+
+```{r sim1-power-col}
+p10.4 <- ggplot(power_x1, aes(x=rep_n, y=power,
+                           colour = b6c, fill = b6c)) +
+  geom_col(alpha = 0.5) +
+  geom_hline(yintercept = 0.8, colour = "red", linetype = "dashed") +
+  geom_hline(yintercept = 0.9, colour = "black", linetype = "dashed") +
+  geom_hline(yintercept = 0.95, colour = "blue", linetype = "dashed") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_fill_brewer(palette = "Dark2") +
+  theme(legend.position = "none") +
+  facet_wrap(~b6c, nrow=1)
+p10.4
+
+ggsave ("Tutorial_4_planning/figures/sim1/bin6_power_col.jpeg",
+        width = 8, height = 6, dpi=800)
+```
+
+ok, on the basis of this analysis, we might make the following conclusions:
+
+- effects of con vs inc that are as large as previously reported by Panis et al., 2016 would be easy to detect with 100 trials and 10 pids. And probably less data than that also.
+
+- Across all effect sizes, there is no benefit of adding more than 200 trials per condition.
+
+- Within the context of these absolute hazard values, somewhere between a 50% reduction and a 75% reduction would give you 80% power to detect an effect, should it exist.
+e.g., a 75% reduction is nearly 100% power.
+A 50% reduction is 70% power.
+So maybe a 60% reduction would give 80% power, with N=10 and trial =200.
+We could run more sims to figure that out.
+
+- Other analyses to run - how many individual pids per exp show a clear effect of con vs inc in bin 6?
+This would be useful to know if we used the small-N design logic and the individual as the unit of replication.
+
+# 11. Run some more follow-up sims #
+
+Let's try a 50%, 60% and 70% reduction across N=10,15 and 20.
+
+let's check effect sizes quickly.
+
+```{r sim2-bin6-effect-sizes}
+bin6_v2 <- tibble(
+  effect_id = 1:3,
+  inc = rep(0.1, times=3),
+  ratio = c(0.3, 0.4, 0.5),
+  con = inc/ratio,
+  inc_clog = clogloglink(inc),
+  con_clog = clogloglink(con)
+)
+bin6_v2
+
+# effect_id   inc ratio   con inc_clog con_clog
+#       <int> <dbl> <dbl> <dbl>    <dbl>    <dbl>
+# 1         1   0.1   0.3 0.333    -2.25   -0.903
+# 2         2   0.1   0.4 0.25     -2.25   -1.25 
+# 3         3   0.1   0.5 0.2      -2.25   -1.50 
+```
+
+## run the sim ##
+
+vary effect size and pid.
+Again, as before, this will take some time, probably hours...go get a beer/coffee.
+And again, eval is set to FALSE, to save time for the tutorial.
+
+```{r run-sim2, eval=FALSE}
+plan(multicore)
+x2 <- crossing(
+  exp = 1:1000, # number of experiment replicates
+  subj_n = c(10,15,20), # range of subject N
+  rep_n = 200,
+  b6i = -2.25,
+  b6c = c(-0.90, -1.25, -1.50)
+) %>%
+  mutate(d = pmap(., index_sim)) %>% 
+  mutate(s = map(d, index_summary)) %>% 
+  select(-d)
+```
+
+## summarise the simulated data ##
+
+```{r summary-sim2, eval=FALSE}
+sx2 <- x2 %>%
+  unnest(s) %>%
+  mutate(exp=factor(exp),
+         subj_n=factor(subj_n),
+         rep_n=factor(rep_n),
+         b6c = factor(b6c,
+                      levels = c("-1.5", "-1.25", "-0.9"),
+                      labels = c("50%", "60%", "70%")),
+         subj=factor(subj),
+         bin=factor(bin))
+sx2
+```
+
+take a look
+
+```{r sim2-quick-look, eval=FALSE}
+head(sx2)
+tail(sx2)
+
+sx2 %>% 
+  distinct(b6c)
+```
+
+save out a file
+
+```{r save-sim2, eval=FALSE}
+## save the first simulation
+write_csv(sx2, "Tutorial_4_planning/data/sim2/sim2_data.csv")
+```
+
+## calculate summary data ##
+
+read in the file, if it is already computed
+
+```{r load-sim2}
+sx2 <- read_csv("Tutorial_4_planning/data/sim2/sim2_data.csv") %>%
+  mutate(exp=factor(exp),
+         subj_n=factor(subj_n),
+         rep_n=factor(rep_n),
+         b6c = factor(b6c,
+                      levels = c("-1.5", "-1.25", "-0.9"),
+                      labels = c("50%", "60%", "70%")),
+         subj=factor(subj),
+         bin=factor(bin))
+sx2
+```
+
+at the exp level
+
+```{r sim2-exp-summary}
+sx2_exp <- sx2 %>% 
+  group_by(exp, subj_n, rep_n, b6c, condition, bin) %>% 
+  summarise(exp_mean_outcome = mean(mean_outcome, na.rm = TRUE),
+            exp_mean_sum=mean(sum, na.rm = TRUE),
+            exp_sum=sum(sum),
+            n = length(unique(subj)), 
+            sd=sd(mean_outcome, na.rm = TRUE),
+            sem = (sd/sqrt(length(unique(subj)))),
+            .groups = "drop")
+sx2_exp
+```
+
+at the sim level
+
+```{r sim2-sim-summary}
+sx2_sim <- sx2_exp %>% 
+  group_by(subj_n, rep_n, b6c, condition, bin) %>% 
+  summarise(sim_mean_outcome = mean(exp_mean_outcome, na.rm = TRUE),
+            sim_mean_sum = mean(exp_mean_sum, na.rm = TRUE),
+            sim_sum = mean(exp_sum, na.rm = TRUE),
+            n = length(unique(exp)), 
+            sd=sd(exp_mean_outcome, na.rm = TRUE),
+            sem = (sd/sqrt(length(unique(exp)))),
+            .groups = "drop")
+sx2_sim
+```
+
+a quick plot of the data
+
+```{r sim2-violin}
+p11.1 <- ggplot(sx2_sim, 
+                aes(x = bin, y = sim_mean_outcome, 
+                    colour = condition)) +
+  geom_jitter(data=sx2_exp, aes(y = exp_mean_outcome),
+              alpha=0.5, width = 0.1, height = 0) +
+  geom_line(aes(group = condition)) + 
+  geom_point(size=2, colour = "black") +
+  geom_errorbar(aes(ymin = sim_mean_outcome-sem*1.96, ymax = sim_mean_outcome+sem*1.96),
+                width=.2, colour = "black") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+  labs(x="timebin",
+       y="outcome") + 
+  facet_grid(fct_rev(subj_n)~b6c)
+p11.1
+
+ggsave ("Tutorial_4_planning/figures/sim2/sim_violin.jpeg",
+        width = 10, height = 8, dpi=800)
+```
+
+ok, look at some values in bin 6
+
+at the exp level
+
+```{r sim2-bin6-exp-values}
+sx2_exp_check <- sx2_exp %>%
+  filter(bin %in% c(6)) %>% 
+  select(exp, subj_n, b6c, condition, bin, exp_mean_outcome) %>%
+  pivot_wider(names_from = "condition",
+              values_from = "exp_mean_outcome") %>% 
+  group_by(exp, subj_n, b6c, bin) %>% 
+  mutate(ratio = cond2/cond1) 
+head(sx2_exp_check)
+```
+
+and at the sim level
+
+```{r sim2-bin6-sim-values}
+sx2_sim_check <- sx2_exp_check %>%
+  group_by(subj_n, b6c, bin) %>% 
+  summarise(sim_ratio = mean(ratio),
+            sim_sd=sd(ratio),
+            sim_sem=sim_sd/sqrt(1000),
+            .groups = "drop")
+head(sx2_sim_check)
+```
+
+plot ratio values
+
+```{r sim2-ratio-values}
+p11.2 <- ggplot(sx2_sim_check, 
+                aes(x = b6c, y = sim_ratio)) +
+  geom_point(size=1.5) +
+  geom_errorbar(aes(ymin = sim_ratio-sim_sem*1.96, ymax = sim_ratio+sim_sem*1.96)) + 
+  geom_hline(yintercept = c(0.3, 0.4, 0.5), colour = "red") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
+  labs(x="timebin",
+       y="outcome") + 
+  facet_wrap(~subj_n)
+p11.2
+```
+
+## calculate power / precision ##
+
+basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound quantiles exclude zero?
+This of course is not a model-based estimate.
+But I think it should give a good guide to what we might expect.
+
+## calculate difference scores ##
+
+at the exp level
+
+```{r sim2-calculate-diff-scores}
+sx2_diff_exp <- sx2 %>%
+  pivot_wider(id_cols = c(exp, subj_n, rep_n, b6c, subj, bin),
+              names_from = "condition",
+              values_from = "mean_outcome") %>% 
+  mutate(diff = cond1 - cond2) %>%
+  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
+  summarise(exp_mean_diff = mean(diff, na.rm = TRUE),
+            exp_sd = sd(diff, na.rm = TRUE),
+            n = n(), # n here is the total trials per condition per pid
+            exp_sem = (exp_sd/sqrt(n)),
+            exp_ci = 1.96*exp_sem,
+            exp_dz = exp_mean_diff/exp_sd,
+            .groups = "drop") 
+head(sx2_diff_exp)
+tail(sx2_diff_exp)
+```
+
+at the group/sim level
+
+```{r sim2-summary-diff-scores}
+sx2_diff_sim <- sx2_diff_exp %>% 
+  group_by(subj_n, rep_n, b6c, bin) %>% 
+  summarise(mean_diff = mean(exp_mean_diff, na.rm = TRUE),
+            sd = sd(exp_mean_diff, na.rm = TRUE),
+            n=n(),
+            sem = (sd/sqrt((n))),
+            ci = 1.96*sem) 
+sx2_diff_sim
+```
+
+## plot ##
+
+violin
+
+diff in original units
+
+```{r sim2-plot-sim-diffs}
+p11.3 <- ggplot(sx2_diff_exp, aes(x=bin, y = exp_mean_diff,
+                                colour = bin, fill = bin)) +
+   geom_jitter(alpha = 0.5, width = 0.1) +
+   geom_violin(alpha = 0.7) +
+   geom_point(data = sx2_diff_sim, 
+             aes(y = mean_diff), size = 3, position=pd2, colour="black") +
+   geom_errorbar(data = sx2_diff_sim,
+                aes(y = mean_diff, ymin = mean_diff-ci, ymax = mean_diff+ci),
+                width=.2, position=pd2, colour = "black") +
+   geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+   theme(legend.position = "none") +
+   ylab("mean outcome") +
+   ggtitle("difference score (cond1-cond2)") +
+   facet_grid(fct_rev(subj_n)~b6c)
+p11.3
+
+ggsave("Tutorial_4_planning/figures/sim2/sim_diffs.jpeg",
+       width = 10, height = 8, dpi=800)
+```
+
+plot each exp's difference score and associated 95% interval
+
+create some factors and new variables
+
+```{r sim2-wrangle-diffs}
+sx2_diff_exp_2 <- sx2_diff_exp %>%
+  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
+  mutate(lower = exp_mean_diff-exp_ci,
+         upper = exp_mean_diff+exp_ci,
+         above_zero = if_else(lower > 0, "yes", "no"), 
+         above_zero = factor(above_zero, levels = c("no", "yes")))
+sx2_diff_exp_2
+```
+
+calculate "power" in a quick and dirty way based on 95% CI
+
+```{r sim2-calculate-power}
+power_x2 <- sx2_diff_exp_2 %>%
+  group_by(subj_n, rep_n, b6c, bin) %>%
+  mutate(check = ifelse(lower > 0, 1, 0)) %>%
+  summarise(power = mean(check, na.rm = TRUE)) %>% 
+  filter(bin %in% c(6))
+power_x2
+```
+
+plot power
+
+```{r sim2-power-tile}
+p11.4 <- ggplot(power_x2, aes(x = b6c, y=subj_n, fill = power)) +
+  geom_tile() +
+  geom_text(aes(label = sprintf("%.3f", power)), color = "white", size = 10) +
+  scale_fill_viridis_c(limits = c(0, 1)) +
+  facet_wrap(~bin) +
+  labs(y="subj_n", x="b6c")
+p11.4
+#
+ggsave ("Tutorial_4_planning/figures/sim2/power.jpeg",
+        width = 10, height = 6, dpi=800)
+```
+
+join power to the df
+
+```{r sim2-join-dfs}
+sx2_diff_power <- sx2_diff_exp_2 %>%
+  filter(bin %in% c(6)) %>% 
+  inner_join(power_x2, by = c("subj_n", "rep_n", "b6c", "bin")) %>% 
+  mutate(power = round(power * 100, 2)) 
+head(sx2_diff_power)
+```
+
+plot
+
+```{r sim2-plot-pointrange}
+p11.5 <- sx2_diff_power %>%
+  ggplot(aes(x = exp, y = exp_mean_diff, ymin = lower, ymax = upper)) +
+  geom_pointrange(fatten = 1/2, aes(colour=above_zero)) +
+  geom_hline(yintercept = 0, colour = "red") +
+  scale_colour_manual(values=c("darkgrey","black")) +
+  geom_text(aes(x=700, y=-0.35, label = sprintf("%.1f%s", power, "% power")), 
+            color = "darkgrey", size = 5) +
+  theme(legend.position = "none") +
+  labs(x = "sim # (i.e., simulation index)",
+       y = "hazard difference") +
+  scale_x_discrete(breaks = c(250,500,750,1000)) +
+  facet_grid(fct_rev(subj_n)~b6c)
+p11.5
+
+ggsave ("Tutorial_4_planning/figures/sim2/bin6_diffs.jpeg",
+        width = 10, height = 8, dpi=800)
+```
+
+plot power as a bar plot
+
+```{r sim2-power-col}
+p11.6 <- ggplot(power_x2, aes(x=subj_n, y=power,
+                           colour = b6c, fill = b6c)) +
+  geom_col(alpha = 0.5) +
+  geom_hline(yintercept = 0.8, colour = "red", linetype = "dashed") +
+  geom_hline(yintercept = 0.9, colour = "black", linetype = "dashed") +
+  geom_hline(yintercept = 0.95, colour = "blue", linetype = "dashed") +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_fill_brewer(palette = "Dark2") +
+  theme(legend.position = "none") +
+  facet_wrap(~b6c, nrow=1)
+p11.6
+
+ggsave ("Tutorial_4_planning/figures/sim2/bin6_power_col.jpeg",
+        width = 8, height = 6, dpi=800)
+```
+
+A summary of these power calculations might be as follows:
+
+For a 70% reduction (0.3 hazard ratio), N=10, trial=200 per condition, would give nearly 100% power.
+For a 60% reduction (0.4 hazard ratio), N=10, trial=200 per condition, would give nearly 90% power.
+For a 50% reduction (0.5 hazard ratio), N=15, trial=200 per condition, would give over 80% power.
+
+And a conclusion / judgment / decision process might be:
+
+Well, like almost always, it depends on your objectives. Some considerations might be...
+
+How much power or precision are you looking to obtain in this particular study?
+Are you running multiple studies that have some form of replication built in?
+What resources do you have at your disposal, such as time, money and personnel?
+How easy or difficult is it to obtain the specific type of sample?
+
+My thoughts for studies in my lab might be something like this...
+
+Pick 0.4 or 0.5 as a target effect size since this is much smaller than that observed in published studies, then pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80% power mark.
+
+**But**, and this is an important "but", do not solely rely on one study.
+Run a follow-up experiment that replicates and extends the initial result.
+By doing so, you avoid the Cult of the Isolated Single Study, and it reduces the reliance on any one type of power analysis.
+Instead, you are aiming for common patterns across two or more experiments, rather than trying to make the case that a single study has sufficient evidential value to hit some criterion mark.

### 2024-11-07 notes/model_notes.Rmd
added model notes and tutorial notes

---
title: "model_notes"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is a way to provide feedback on draft versions of the models that were built as part of the tutorials.

# November 4th 2024 #

## a few minor points and/or questions ##

1. Model formula.

I get the 0 + ... logic in general, when you do not want an overall intercept and/or when you want to manually set the intercept somewhere. 
But I'm curious about the setup that you have when you use 0 + Intercept, when "Intercept" is a column in a dataframe with all values being "1". Isn't that just the same as setting 1 + ... whatever else ?  
Do you see what I mean? As I understand it (but I may be confused), you are saying zero but then setting the intercept at 1 anyway, which just seems odd to me. Why not just set it at 1, like normal?

## interpreting parameter estimates and interval values ##

- I understand, as it is currently written, how to interpret and use the information in the model comparison section.
e.g., out of 4 models, model 4 is more accurate at making out-of-sample predictions. 
This means that whatever model 4 brings in addition to the other models, it helps make out-of-sample predictions more accurate.
In this case, model 4 adds a dimension of experiment time interacting with condition (and maybe other stuff also).
So this means that condition effects appear to vary by the time of the experiment or testing session.

Ok, so far so good.

What I am finding a little harder to understand in a quantitative sense is the size and precision of key effects of interest, which relate to condition * timebin * trial time interactions.
For a comparison, if we had a simple gaussian model with a factorial design (one condition, with two levels: inc vs con), we could estimate the size, direction and interval width of key parameters of interest. e.g., the factor condition  (inc > con) leads to an increase in reaction time by 30ms [20,40] with 95% quantiles in square brackets. This would be our key effect size. And we could report it and make an inference about it. Others could power new studies based on these reported effect sizes etc.

Of course, when model parameters are reported in different units to those that might be intuitive as effect sizes, such as cloglog values, one needs to transform them first to hazard values or ratios. But, when models get really complex (like in EHA), then one single parameter (even after transformation) is hard to interpret on its own, because the effect size of interest is a product of many different and possibly interacting parameters.

And, as I understand things, this is where posterior predictions comes in. 
Posterior predictions are generated from all of the parameters in the model.
And as they are generated for every draw in the posterior distribution, you get a distribution of predicted values, which can be summarised using tidybayes etc., to produce point and interval estimates, as well as plots of the posterior predictions distributions.

I say all of this to provide background context for understanding section 4.3.8.
The top row of Figure 5 shows that the middle timebins within a trial (400, 440, say) become more negative (compared to neutral) later in the experiment (trial 1500).
Statements like that reflect a 3-way interaction: condition * timebin * trial number.
But to show that quantitatively we would need the difference between these 3 factors to be displayed and interval estimates provided. 
Does this make sense?
We would need the difference scores between panel 1500 computed from panel 500 and the difference of panel 1000 computed from 500.
Otherwise, as I understand it, we are just showing numerical differences at each trial numbers (500, 1000 and 1500) rather than computing the differences between those trial numbers.

Now, if I return to your model specification, you have several condition * timebin * trial number interactions that reflect linear, quadratic and cubic treatments of time. So, focussing on any one of these could be misleading to the overall picture. And hence why posterior predictions are useful because they use all of the parameters.

So, at this stage, my question and comment is as follows: in these situations it feels like posterior predictions PLUS paired contrasts between conditions/columns in the posterior would be really useful and possibly even essential to make an inference of the type we want to. e.g., we need to quantify how much and with what level of precision (interval width) the difference between con and blank changes over the experiment.

If we do not do this, then I worry that we are falling into the trap of the below paper:

https://www.tandfonline.com/doi/abs/10.1198/000313006X152649

Now, given how tremendously complicated these models are, I may be confused on many levels.

However, what is absolutely certain is that we need to be clear on how one would make an inference about the effect of condition over two time scales (within trial and across trials). 
And at the moment, we have not quantified this in terms of an effect size and an associated interval width. 
Instead, we have done model comparison and plotted some parameter estimates per levels of the design, which produce interval widths for condition vs neutral across timbin. But we have not produced relevant comparisons across trial number.

I hope this makes some sense. 
Of course, we can discuss this also, if that's easier.

As an illustration of what I think I mean (!), if we take Figure 6 (Model-based hazard functions).
If this figure was for the group average (rather than only pid 6).
Then we could compute a difference score for each timebin of con compared to neutral, as well as inc compared to neutral.
That's all good.
But we would then need to calculate the difference of these differences between 1000 and 500 and between 1500 and 500.
At least we would need to compute these if we wanted to quantify how much trial number impacts the size of the condition differences when compared to neutral.

## different ways to calculate posterior predicitons / effects of interest ##

Ok, I hope that the prior comment above makes sense.

And to follow-up on this prior comment, I have now done some more digging and made some calculations using model_M4 and the tidybayes add_epred_draws() function.

And in addition, I looked at your code for calculating estimates in Tutorial_2a.

I have a bunch of code that I will share with you, but here I will just provide a summary. I think I have come to an interim conclusion, which is as follows...

the below code chunk is your code from Tutorial_2a for calculating effects by manually adding parameters from the model together.

```{r post-distr}
#get_variables(model_M4)[1:31]

post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
   select(starts_with("b_")) %>%       # 8000 x 31
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial500 = -500/1000,
          trial500sq = trial500^2,
          trial1500 = 500/1000,
          trial1500sq = trial1500^2)

# effects for trials 500, 1000, and 1500
effects <- post %>% 
  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
            period_9sq * trial1500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
            period_9sq * trial1500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,

         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
            period_9sq * trial1500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
            period_9sq * trial1500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,

         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
            period_9sq * trial500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
            period_9sq * trial500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,

         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
            period_9sq * trial500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
            period_9sq * trial500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,

         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,

         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
```

And I *think* this is fine, although it is very hard to check and VERY VERY easy to make mistakes, given the long-winded nature of the term names and the complexity of the model.

As I understand things, another way to go would be as follows:

note: dat_M4 is just the data that you used in model_M4
e.g.,

```{r}
# dat_M4 <- as_tibble(model_M4$data)
```

and now use add_epred_draws()

```{r}
epreds <- dat_M4 %>% 
  data_grid(pid, condition,
            period_9 = seq(from = -3, to = 6, by = 1),
            trial_c = c(-0.5, 0, 0.5)) %>%
  add_epred_draws(model_M4) %>% 
  ungroup()
head(epreds)
```

This just uses add_epred_draws instead of calculating by hand.
To my mind, this is far safer and more reliable/robust and a hell of a lot more compact.

Once you have these predictions, you can wrangle and compare contrasts, plot, and summarise as you normally would using tidybayes functions.

I will share the code so that you can take a look.

### 2024-11-07 notes/tutorial_notes.Rmd
added model notes and tutorial notes

---
title: "tutorial_notes"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is a way to provide feedback on draft versions of the tutorials. 
I can edit the tutorials directly of course (which git should track), but this is a space to provide comments.

# October 31st 2024 #

The basic workflow of the tutorials seems pretty good to me.
I can do some more fine-tuning of the code after the next revision of the manuscript,
given that some stuff will change.

## table of contents, aesthetics and apperance ##

I think we can easily improve the visual appearance of the tutorials in .html format by using a theme,
as well as their ease of navigation by setting a table of contents function.
To do so, include the following in th YAML header.

output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united

This is what we did in Andrew Wildman's psychopy tutorial. 
There are other themes, but this works for me.

### 2024-11-18 Tutorial_2a.Rmd
update html and Rmd files
update Tutorial 2a and ms.

@@
 date: "`r Sys.Date()`"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-In this Tutorial 2a we plot priors for the logit and cloglog link functions, build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials for the selected model.
-
-Content:
-1. Load libraries.
-2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
-3. Plot the logit and complementary log-log (cloglog) link functions.
-4. Visualize different prior distributions on the logit and cloglog scales.
-5. Fit hazard models.
-6. Compare models using loo and waic.
-7. Display effects for the selected model.
+Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data.
+The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (prime = blank, congruent, or incongruent).
+We compare models using WAIC and LOO, and interpret parameter estimates for one model.
+We also plot the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
 
@@
 
+theme settings for ggplot
+
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"),
+          legend.position = "bottom")
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
 # 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
@@
 print(ptb_data,n=30)
-summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
+summary(ptb_data) # 26602 rows: 6 participants, trial (experiment-wise), 3 conditions, 15 periods, and event indicator (0/1)
 ```
 
-Prepare the data set.
+Wrangle the data set.
 
-```{r prepare-data}
+```{r wrangle-data}
+ptb_data <- ptb_data %>% 
 # select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
-ptb_data <- ptb_data %>% filter(period > 5)
+  filter(period > 5) %>%
 
-# create factor condition, with "blank" as the reference level
-ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-
-# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.
-ptb_data <- ptb_data %>% 
-        mutate(period_9 = period - 9,
-               trial_c = (trial - 1000)/1000,
+# create categorical predictor for TIME named "timebin" with index coding
+  mutate(timebin = factor(period,levels=c(6:15)),
+# create continuous predictor for TIME named "period_9", centered on bin 9,
+         period_9 = period - 9,
+# create binary variables "dummys" to indicate each bin in the analysis
          d6  = if_else(period == 6, 1, 0),
@@
          d14 = if_else(period == 14, 1, 0),
-               d15 = if_else(period == 15, 1, 0))
+         d15 = if_else(period == 15, 1, 0),
+
+# create continuous predictor for trial number named "trial_c", centered on bin 1000 and rescaled
+         trial_c = (trial - 1000)/1000,
+# create categorical predictor for trial number named "stage" (early,middle,late) with index coding
+         stage = ifelse(trial <= 500, 1, ifelse(trial > 1000, 3, 2)),
+         stage = factor(stage, levels=c(1,2,3)),
+
+# create factor "condition", with "blank" as the reference level
+         condition = factor(condition, labels = c("blank", "congruent","incongruent")),
+# create categorical predictor "prime" with index-coding
+         prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+         prime = factor(prime,levels=c(1,2,3))) %>%
+  select(pid,trial,trial_c,stage,condition,prime,period,period_9,timebin,d6:d15)
 
@@
 summary(ptb_data)
-# 6 subjects
-# condition: blank (6401), congruent (2642), incongruent (3797)
-# period_9: -3 to 6
-# trial_c: -0.999 to 0.540
 ```
 
-# 3. Plot the logit and complementary log-log (cloglog) link functions.
+# 3. Fit hazard models.
 
-```{r plot-links}
-probability <- (1:99999)/100000
-logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
-logit    <- function(x) { return( log(x/(1-x)) )}
-inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
-cloglog         <- function(x) { return( log(-1*log(1-x)) )}
+## Model M0: A null model without experimental predictors
 
-cloglog_prob <- cloglog(probability)
-logit_prob <- logit(probability)
-dataplot <- cbind(probability,cloglog_prob,logit_prob)
+The first model we fit is a "random intercepts" model, where we fit a single intercept for each timebin and allow each intercept to vary between participants.
+This is a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial). 
+There are two ways to implement such a model. First, we can use the index-coding approach, as .. .by McElreath...
 
-ggplot() +
-  geom_hline(yintercept=0, color="white") +
-  geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
-  geom_line(data=dataplot,aes(y=cloglog_prob,x=probability,colour="cloglog"),linewidth=1) +
-  scale_color_manual(name = "Link function:", values = c("logit" = "darkblue", "cloglog" = "red")) +
-  geom_vline(xintercept = logistic(0), linetype="dotted", linewidth = 0.3) +   
-  geom_vline(xintercept = inverse_cloglog(0), linetype="dotted", linewidth = 0.3) +
-  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90, size=4) +
-  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90,size=4) +
-  scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
-  labs(x = "Probability",
-        y = "logit or cloglog scale") +
-  theme(panel.grid = element_blank(),
-        axis.text=element_text(size=16),
-        axis.title=element_text(size=18),
-      #  legend.position="top",
-        legend.text = element_text(size=18),
-      legend.position = "inside",
-        legend.position.inside=c(.2,.75),
-      legend.background = element_rect(fill="white") )
+Prepare the data file, and specify priors.
+
+```{r data-priors-M0}
+# remove unnecessary columns before fitting a model
+M0_data <- ptb_data %>% select(pid, event, timebin)
+
+priors_M0 <- c(
+  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), # flat prior for p
+  set_prior("normal(0, 1)", class = "sd"),                    
+  set_prior("lkj(2)", class = "cor")                        
+)
 ```
 
-```{r save-plot-links, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 300)
+Fit the model (M0). old model M0d
+
+```{r fit-model-M0, eval=F}
+plan(multicore)
+
+model_M0 <-
+   brm(data = M0_data,
+       family = binomial(link="cloglog"),
+       event|trials(1) ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "models/model_M0")
 ```
 
-# 4. Visualize different prior distributions on the logit and cloglog scales.
+2582.7 seconds.
 
-To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
-Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.
+Second, we can fit the same model using a dummy coding approach, as follows.
 
-## Simulate, convert, and fit.
+Prepare the data file, and specify priors.
 
-```{r simulate-convert}
-set.seed(11)
+Fit the model with dummy coding (M0d).
 
-logit    <- function(x) { return( log(x/(1-x)) )}
-cloglog  <- function(x) { return( log(-1*log(1-x)) )}
+### Make assumptions to simplify the model
 
-# generate draws from U(0,1) and convert
-dat <- 
-  tibble(p = runif(1e6, 0, 1)) %>% 
-  mutate(g = logit(p),
-         c = cloglog(p)) 
-# display
-dat %>%   
-  ggplot(aes(x = c)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-```
+Instead of using a general specification of TIME, we can treat TIME as a continuous variable and make assumptions about how hazard changes over time within a trial.
+For example, we might assume that hazard changes in a linear + quadratic fashion over timebins within a trial.
 
-```{r fit-skew-normal, eval=F}
-# fit model
-fit_skewN <-
-  brm(data = dat,
-      family = skew_normal(),
-      c ~ 1,
-      chains = 4, cores = 4,
-      file = "Tutorial_2_Bayesian/models/fit_skewN")
-```
+Prepare the data file, and specify priors.
 
-```{r}
-fit_skewN <- readRDS("Tutorial_2_Bayesian/models/fit_skewN.rds")
-summary(fit_skewN) 
-```
+Fit the model with a continuous predictor (M0c).
 
-Now we can reverse the process.
-We simulate from the skew-Normal distribution based on the posterior means for mu, sigma, and alpha, and then convert the results into the probability (i.e., discrete-time hazard) metric.
+## Model M1.
 
-```{r check-results}
-set.seed(11)
+The next model we fit also includes our predictor variable prime-target congruency.
+First, we can use index coding for the categorical predictor.
 
-inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+Second, if we do not want to use index coding, assume that hazard can vary in a linear+quadratic fashion over time, and relax the proportionality assumption, then we can fit the following model.
 
-tibble(c = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22) ) %>% 
-  mutate(p = inverse_cloglog(c)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-```
+## Model M2.
 
-## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale.
+The next model we fit also includes our predictor trial number.
 
-```{r plot-priors}
-logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
-logit    <- function(x) { return( log(x/(1-x)) )}
-inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
-cloglog    <- function(x) { return( log(-1*log(1-x)) )}
+First, we can categorise the continuous predictor trial number by grouping them in one of three "stages" of the experiment (stage 1 = trials 1 to 500; stage 2 = trials 501 to 1000; stage 3 = trials 1001 and later) and use index coding.
 
-set.seed(23)
+Second, we can treat trial number and bin number as continuous variables, and center them.
 
-# A N(0,4) prior on the logit and cloglog scales pushes mass to probabilities of 0 and 1
-pr1 <- tibble(prior = rnorm(1e6, mean = 0, sd = 4)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-13, y=60000, label="N(0,4)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
+# Model comparisons.
 
-l1 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 4)) %>%  
-  mutate(p = logistic(log_odds)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  ggtitle("logistic(prior)") +
-  theme(panel.grid = element_blank())
-
-c1 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 4)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  ggtitle("inverse_cloglog(prior)")+
-  theme(panel.grid = element_blank())
-
-# A N(0,2) prior on the logit and cloglog scales pushes mass to probabilities of 0 and/or 1
-pr2 <- tibble(prior = rnorm(1e6, mean = 0, sd = 2)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-6, y=60000, label="N(0,2)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l2 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 2)) %>%  
-  mutate(p = logistic(log_odds)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-c2 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 2)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-# A student-t(df=7.61) prior with scale 1.57 on the logit scale approximates a uniform distribution on the probability scale.
-This might be a good prior to use for the alpha parameters or Intercept in a logit-hazard model.
-pr3 <- tibble(prior = rt(1e6, df = 7.61)* 1.57) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-12, y=140000, label="t(7.61,0,1.57)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l3 <- tibble(log_odds = rt(1e6, df = 7.61)* 1.57) %>%  
-  mutate(p = logistic(log_odds)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-c3 <- ggplot()
-
-# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. This might be a good prior to use for the beta parameters in a logit-hazard model.
-pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-3, y=60000, label="N(0,1)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'D', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l4 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 1))%>%  
-  mutate(p = logistic(log_odds)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  geom_vline(xintercept=logistic(0), color="red") +
-  theme(panel.grid = element_blank())
-
-c4 <- ggplot()
-
-# A skew_Normal(-0.59,1.26,-4.22) prior on the cloglog scale approxiates a uniform distr.
-on the hazard scale.
-This uninformative prior might be good for the alpha parameters or Intercept in a cloglog-hazard model.
-pr5 <- tibble(prior = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-5.5, y=55000, label="skew_N(-0.59,1.26,-4.22)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'E', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l5 <- ggplot()
-
-c5 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-# The skew_Normal(-1,1,-2) on the cloglog scale is a weakly informative prior for the alpha parameters or Intercept in a cloglog-hazard model because hazard values below .5 more likely than values above .5 in general.
-pr6 <- tibble(prior = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-5, y=55000, label="skew_N(-1,1,-2)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'F', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l6 <- ggplot()
-
-c6 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-# A skew_Normal(-0.2,0.71,-2.2) on the cloglog scale gently regularizes p towards .6321. This might be a good prior to use for the beta parameters in a cloglog-hazard model.
-pr7 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-3.2, y=55000, label="skew_N(-0.2,.71,-2.2)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'G', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l7 <- ggplot()
-
-c7 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  geom_vline(xintercept=inverse_cloglog(0), color = "red") +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) / (l7 + pr7 + c7)
-```
-
-```{r save-plot-priors, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
-```
-
-# 5. Fit hazard models.
-
-## Model M1: general specification of TIME in the baseline condition (blank prime), and main effects of prime type, and trial number.
+# Model inference.
 
@@
 ```
+
+# 3. Plot the logit and complementary log-log (cloglog) link functions.
+
+```{r plot-links}
+probability <- (1:99999)/100000
+logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
+logit    <- function(x) { return( log(x/(1-x)) )}
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+cloglog         <- function(x) { return( log(-1*log(1-x)) )}
+
+cloglog_prob <- cloglog(probability)
+logit_prob <- logit(probability)
+dataplot <- cbind(probability,cloglog_prob,logit_prob)
+
+ggplot() +
+  geom_hline(yintercept=0, color="white") +
+  geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
+  geom_line(data=dataplot,aes(y=cloglog_prob,x=probability,colour="cloglog"),linewidth=1) +
+  scale_color_manual(name = "Link function:", values = c("logit" = "darkblue", "cloglog" = "red")) +
+  geom_vline(xintercept = logistic(0), linetype="dotted", linewidth = 0.3) +   
+  geom_vline(xintercept = inverse_cloglog(0), linetype="dotted", linewidth = 0.3) +
+  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90, size=4) +
+  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90,size=4) +
+  scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
+  labs(x = "Probability",
+        y = "logit or cloglog scale") +
+  theme(panel.grid = element_blank(),
+        axis.text=element_text(size=16),
+        axis.title=element_text(size=18),
+      #  legend.position="top",
+        legend.text = element_text(size=18),
+      legend.position = "inside",
+        legend.position.inside=c(.2,.75),
+      legend.background = element_rect(fill="white") )
+```
+
+```{r save-plot-links, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 300)
+```
+
+# 4. Visualize different prior distributions on the logit and cloglog scales.
+
+To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
+Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.
+
+## Simulate, convert, and fit.
+
+```{r simulate-convert}
+set.seed(11)
+
+logit    <- function(x) { return( log(x/(1-x)) )}
+cloglog  <- function(x) { return( log(-1*log(1-x)) )}
+
+# generate draws from U(0,1) and convert
+dat <- 
+  tibble(p = runif(1e6, 0, 1)) %>% 
+  mutate(g = logit(p),
+         c = cloglog(p)) 
+# display
+dat %>%   
+  ggplot(aes(x = c)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+```
+
+```{r fit-skew-normal, eval=F}
+# fit model
+fit_skewN <-
+  brm(data = dat,
+      family = skew_normal(),
+      c ~ 1,
+      chains = 4, cores = 4,
+      file = "Tutorial_2_Bayesian/models/fit_skewN")
+```
+
+```{r}
+fit_skewN <- readRDS("Tutorial_2_Bayesian/models/fit_skewN.rds")
+summary(fit_skewN) 
+```
+
+Now we can reverse the process.
+We simulate from the skew-Normal distribution based on the posterior means for mu, sigma, and alpha, and then convert the results into the probability (i.e., discrete-time hazard) metric.
+
+```{r check-results}
+set.seed(11)
+
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+
+tibble(c = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22) ) %>% 
+  mutate(p = inverse_cloglog(c)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+```
+
+## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale.
+
+```{r plot-priors}
+logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
+logit    <- function(x) { return( log(x/(1-x)) )}
+inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
+cloglog    <- function(x) { return( log(-1*log(1-x)) )}
+
+set.seed(23)
+
+# A N(0,4) prior on the logit and cloglog scales pushes mass to probabilities of 0 and 1
+pr1 <- tibble(prior = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-13, y=60000, label="N(0,4)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l1 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 4)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("logistic(prior)") +
+  theme(panel.grid = element_blank())
+
+c1 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 4)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  ggtitle("inverse_cloglog(prior)")+
+  theme(panel.grid = element_blank())
+
+# A N(0,2) prior on the logit and cloglog scales pushes mass to probabilities of 0 and/or 1
+pr2 <- tibble(prior = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-6, y=60000, label="N(0,2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l2 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 2)) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c2 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# A student-t(df=7.61) prior with scale 1.57 on the logit scale approximates a uniform distribution on the probability scale.
+This might be a good prior to use for the alpha parameters or Intercept in a logit-hazard model.
+pr3 <- tibble(prior = rt(1e6, df = 7.61)* 1.57) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-12, y=140000, label="t(7.61,0,1.57)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l3 <- tibble(log_odds = rt(1e6, df = 7.61)* 1.57) %>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+c3 <- ggplot()
+
+# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. This might be a good prior to use for the beta parameters in a logit-hazard model.
+pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3, y=60000, label="N(0,1)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'D', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l4 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 1))%>%  
+  mutate(p = logistic(log_odds)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  geom_vline(xintercept=logistic(0), color="red") +
+  theme(panel.grid = element_blank())
+
+c4 <- ggplot()
+
+# A skew_Normal(-0.59,1.26,-4.22) prior on the cloglog scale approxiates a uniform distr.
+on the hazard scale.
+This uninformative prior might be good for the alpha parameters or Intercept in a cloglog-hazard model.
+pr5 <- tibble(prior = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-5.5, y=55000, label="skew_N(-0.59,1.26,-4.22)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'E', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l5 <- ggplot()
+
+c5 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# The skew_Normal(-1,1,-2) on the cloglog scale is a weakly informative prior for the alpha parameters or Intercept in a cloglog-hazard model because hazard values below .5 more likely than values above .5 in general.
+pr6 <- tibble(prior = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-5, y=55000, label="skew_N(-1,1,-2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'F', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l6 <- ggplot()
+
+c6 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+# A skew_Normal(-0.2,0.71,-2.2) on the cloglog scale gently regularizes p towards .6321. This might be a good prior to use for the beta parameters in a cloglog-hazard model.
+pr7 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>%  
+  ggplot(aes(x = prior)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  annotate(geom="text", x=-3.2, y=55000, label="skew_N(-0.2,.71,-2.2)",
+              color="red", size=6) +
+  annotate(geom = 'text', label = 'G', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
+  theme(panel.grid = element_blank())
+
+l7 <- ggplot()
+
+c7 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  geom_vline(xintercept=inverse_cloglog(0), color = "red") +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
+
+(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) / (l7 + pr7 + c7)
+```
+
+```{r save-plot-priors, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
+```

### 2024-11-20 Tutorial_2a.Rmd
add new models info

@@
 The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (prime = blank, congruent, or incongruent).
-We compare models using WAIC and LOO, and interpret parameter estimates for one model.
+We then compare models using WAIC and LOO, and interpret parameter estimates for one model.
 We also plot the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
@@
          prime = factor(prime,levels=c(1,2,3))) %>%
-  select(pid,trial,trial_c,stage,condition,prime,period,period_9,timebin,d6:d15)
+  select(pid,event,trial,trial_c,stage,condition,prime,period,period_9,timebin,d6:d15)
 
@@
 This is a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial). 
-There are two ways to implement such a model. First, we can use the index-coding approach, as .. .by McElreath...
+There are two ways to implement such a model. First, we can use the index-coding approach, which supplies an intercept for each level of TIME.
 
@@
 ```{r data-priors-M0}
-# remove unnecessary columns before fitting a model
 M0_data <- ptb_data %>% select(pid, event, timebin)
@@
 priors_M0 <- c(
-  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), # flat prior for p
+  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), # See suppl. material 
   set_prior("normal(0, 1)", class = "sd"),                    
@@
 
-Fit the model (M0). old model M0d
+Fit model M0.
 
@@
 
-model_M0 <-
+model_M0 <-                     # wrong data file name
    brm(data = M0_data,
@@
        seed = 12, init = "0",
-       file = "models/model_M0")
-```
-
-2582.7 seconds.
+       file = "Tutorial_2_Bayesian/models/model_M0")
+```
+
+This took 43 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+
+```{r}
+model_M0 <- readRDS("Tutorial_2_Bayesian/models/model_M0.rds")
+summary(model_M0)
+formula(model_M0)
+plot(model_M0)
+```
 
@@
 
+```{r data-priors-M0d}
+M0d_data <- ptb_data %>% select(pid, event, d6:d8, d10:d15)
+
+priors_M0d <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
+  set_prior("normal(0, 1)", class = "sd"),  
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
 Fit the model with dummy coding (M0d).
 
+```{r fit-model-M0d, eval=F}
+plan(multicore)
+
+model_M0d <-            # bernouilli.  :(
+   brm(data = M0d_data,
+       family = bernoulli(link="cloglog"),
+       event ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
+                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15  | pid),
+       prior = priors_M0d,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M0d")
+```
+
+M0d took 77 minutes.
+
+```{r}
+model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
+summary(model_M0d)
+formula(model_M0d)
+```
+
 ### Make assumptions to simplify the model
@@
 Instead of using a general specification of TIME, we can treat TIME as a continuous variable and make assumptions about how hazard changes over time within a trial.
-For example, we might assume that hazard changes in a linear + quadratic fashion over timebins within a trial.
+For example, we might assume that hazard changes in a linear + quadratic fashion over time bins within a trial.
 
@@
 
+```{r data-priors-M0c}
+M0c_data <- ptb_data %>% select(pid, event, period_9) %>%
+  mutate(period_9_sq = period_9^2,
+         period_9_cu = period_9^3)
+
+priors_M0c <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0, 1)", class = "sd"),  
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
 Fit the model with a continuous predictor (M0c).
 
-## Model M1.
-
-The next model we fit also includes our predictor variable prime-target congruency.
+```{r fit-model-M0c, eval=F}
+plan(multicore)
+
+model_M0c <-
+   brm(data = M0c_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq + period_9_cu + 
+                          (0 + Intercept + period_9 + period_9_sq + period_9_cu  | pid),
+       prior = priors_M0c,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M0c")
+```
+
+Model_M0c took about 144 minutes to run.
+
+```{r}
+model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
+summary(model_M0c)
+formula(model_M0c)
+```
+
+## Model M1. Adding our experimental manipulation.
+
+The next model we fit also includes our predictor variable prime type.
 First, we can use index coding for the categorical predictor.
 
-Second, if we do not want to use index coding, assume that hazard can vary in a linear+quadratic fashion over time, and relax the proportionality assumption, then we can fit the following model.
-
-## Model M2.
-
-The next model we fit also includes our predictor trial number.
-
-First, we can categorise the continuous predictor trial number by grouping them in one of three "stages" of the experiment (stage 1 = trials 1 to 500; stage 2 = trials 501 to 1000; stage 3 = trials 1001 and later) and use index coding.
-
-Second, we can treat trial number and bin number as continuous variables, and center them.
-
-# Model comparisons.
-
-# Model inference.
+Prepare the data file, and specify priors.
+
+```{r data-priors-M1}
+M1_data <- ptb_data %>% select(pid, event, timebin, prime)
+
+priors_M1 <- c(
+  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), 
+  set_prior("normal(0, 1)", class = "sd"),                    
+  set_prior("lkj(2)", class = "cor")                        
+)
+```
+
+Fit model M1.
+
+```{r fit-model-M1, eval=F}
+plan(multicore)
+
+model_M1 <-
+   brm(data = M1_data,
+       family = binomial(link="cloglog"),
+       event|trials(1) ~ 0 + timebin:prime  +
+                        (0 + timebin:prime | pid),
+       prior = priors_M1,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M1")
+```
+
+Model M1 took about 126 minutes to fit.
+
+```{r}
+model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
+summary(model_M1)
+formula(model_M1)
+```
+
+Second, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach and treat TIME as a continuous variable.
+For example, we may assume that hazard can change in a linear + quadratic + cubic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic + cubic fashion, and fit the following model (M1d).
+
+```{r data-priors-M1d}
+M1d_data <- ptb_data %>% select(pid, event, period_9, condition) %>%
+
+priors_M1d <- c(
+  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"), 
+  set_prior("normal(0, 1)", class = "sd"),            
+  set_prior("lkj(2)", class = "cor")                  
+)
+```
+
+```{r fit-model-M1d, eval=F}
+plan(multicore)
+
+model_M1d <- 
+   brm(data = M1d_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + 
+                           condition*period_9 + 
+                           condition*period_9_sq + 
+                           condition*period_9_cu +
+                           (1 + condition*period_9 +
+                           condition*period_9_sq +
+                           condition*period_9_cu | pid),
+       prior = priors_M1d,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M1d")
+```
+
+Model_M1d took about 268 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
+
+```{r}
+model_M1d <- readRDS("Tutorial_2_Bayesian/models/model_M1d.rds")
+summary(model_M1d)
+formula(model_M1d)
+```
+
+One could also fit a model with prime and period_9.
+However, to include interactions between a index coded categorical variable and a continuous variable in brm(), one has to switch to the non-linear syntax, as illustrated in the following model formula.
+
+```{r}
+bf(event|trials(1) ~ 0 + a + b * period_9 + c * period_9_sq,
+                 a ~ 0 + prime + (0 + prime |i| pid),
+                 b ~ 0 + prime + (0 + prime |i| pid),
+                 c ~ 0 + prime + (0 + prime |i| pid),
+                 nl = TRUE)
+```
+
+The priors can be set as follows.
+
+```{r priors-nonlinearsyntax}
+priors <- c(
+  prior(skew_normal(-0.59,1.26,-4.22), class = b, nlpar = a), 
+  prior(skew_normal(-0.2,.71,-2.2), class = b, nlpar = b), 
+  prior(skew_normal(-0.2,.71,-2.2), class = b, nlpar = c), 
+  prior(exponential(1), class = sd, group = pid, nlpar = a),
+  prior(exponential(1), class = sd, group = pid, nlpar = b),
+  prior(lkj(2), class = cor, group = pid)
+)
+```
+
+## Model M2. Including effects of trial number.
+
+Up until now, we have been working with one time scale, TIME or bin rank within a trial.
+While many cognitive processes play out on this short time scale (milliseconds to seconds), some play out on longer time scales (minutes to hours), e.g., learning processes.
+
+When we are interested in studying how the hazard of response occurence in our priming experiment also changes on a longer time scale, we can add trial number into the model formula.
+Here we simply illustrate some possibilities with index and reference coding.
+
+First, we can categorize the predictor trial number by grouping trials in one of three "stages" of the experiment (stage 1 = trials 1 to 500; stage 2 = trials 501 to 1000; stage 3 = trials 1001 and later) and use index coding (variable "stage" in ptb_data).
+
+```{r}
+event|trials(1) ~ 0 + timebin:prime:stage  + (0 + timebin:prime:stage | pid)
+```
+
+Second, we can treat trial number as a continuous variable, and make assumptions about the way hazard changes with trial number (e.g., linear, quadratic, etc.) for each combination of timebin and prime.
+
+```{r}
+bf(event|trials(1) ~ 0 + a + b * trial_c + c * trial_c_sq,
+                 a ~ 0 + timebin:prime + (0 + timebin:prime |i| pid),
+                 b ~ 0 + timebin:prime + (0 + timebin:prime |i| pid),
+                 c ~ 0 + timebin:prime + (0 + timebin:prime |i| pid),
+                 nl = TRUE)
+```
+
+Third, one can use dummy coding for prime type (variable "condition" in ptb_data), and treat TIME and trial number as continous variables.
+
+```{r}
+ bf(event | trials(1) ~ 0 + Intercept + 
+                          condition*period_9*trial_c + 
+                          condition*period_9*trial_c_sq + 
+                          condition*period_9_sq*trial_c +
+                          condition*period_9_sq*trial_c_sq +
+                          (1 +  condition*period_9*trial_c +
+                                condition*period_9*trial_c^2) + 
+                                condition*period_9_sq*trial_c +
+                                condition*period_9_sq*trial_c_sq  | pid)
+```
+
+# Compare models using loo and waic.
+
+The predictive accuracy of a set of models can be compared using WAIC and LOO.
+
+```{r compare-models}
+model_M0  <- readRDS("Tutorial_2_Bayesian/models/model_M0.rds")
+model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
+model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
+model_M1  <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
+model_M1d <- readRDS("Tutorial_2_Bayesian/models/model_M1d.rds")
+```
+
+Using WAIC and LOO for comparing nonnested models.
+
+```{r add-criterion, eval=F}
+model_M0  <- add_criterion(model_M0, c("loo", "waic"))
+model_M0d <- add_criterion(model_M0d, c("loo", "waic"))
+model_M0c <- add_criterion(model_M0c, c("loo", "waic"))
+model_M1  <- add_criterion(model_M1, c("loo", "waic"))
+model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
+```
+
+Compare all three models.
+
+```{r loo-compare}
+loo_compare(model_M0, model_M0d, model_M0c, model_M1, model_M1d, criterion = "loo") %>% print(simplify = F)
+loo_compare(model_M0, model_M0d, model_M0c, model_M1, model_M1d, criterion = "waic") %>% print(simplify = F)
+```
+
+```{r model-weights}
+model_weights(model_M0, model_M0d, model_M0c, model_M1, model_M1d, weights = "loo") %>% round(digits = 3)
+
+model_weights(model_M0, model_M0d, model_M0c, model_M1, model_M1d, weights = "waic") %>% round(digits = 3)
+```
+
+# Model inference for model M1.
+
+## Diagnostics
+
+The Pareto k estimates can be displayed as follows.
+
+```{r pareto}
+loo(model_M1)$diagnostics %>% 
+  data.frame() %>% 
+  # attach the `id` values
+  bind_cols(M1_data) %>% 
+  mutate(id = 1:n()) %>%
+
+  ggplot(aes(x = id, y = pareto_k)) +
+  geom_point(alpha = 3/4) + 
+  geom_text(data = . %>% filter(pareto_k > .2),
+            aes(x = id + 2, label = id),
+            size = 3, hjust = 0) +
+  theme(panel.grid = element_blank())
+```
+
+## Posterior distributions of the parameters
+
+## Expected value of the posterior predictive distribution
+
+We can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, and for each participant in the data set.
+
+Figure X
+
+### Grand average marginal effects
+
+### Subject-specific average marginal effects
+
+And at the group level:
+
+## Example conclusions for model M1.
+
+What can we conclude from model M1 about our research question, i.e., what is the temporal dynamics of the effect of prime-target congruency on RT?
+In other words, in which time bins does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of xxx ms)?
 
@@
 
-c3 <- ggplot()
+c3 <- tibble(cloglog_prob = rt(1e6, df = 7.61)* 1.57) %>% 
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+  theme(panel.grid = element_blank())
 
@@
 
-c4 <- ggplot()
+c4 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 1))%>%  
+  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
+  ggplot(aes(x = p)) +
+  geom_histogram(bins = 50) +
+  scale_y_continuous(NULL, breaks = NULL) +
+ theme(panel.grid = element_blank())
 

### 2024-11-20 Tutorial_2a.Rmd
update 2a

@@
 pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
-         "brms", "tidybayes", "bayesplot", "future", "parallel")
+         "brms", "tidybayes", "bayesplot", "future", "parallel", "modelr")
 
@@
 
-## Posterior distributions of the parameters
+## Posterior distributions of the population-level parameters of M1
+
+Wrangle the posterior draws.
+
+```{r}
+post <- as_draws_df(model_M1) %>% 
+  select(-lp__) %>% 
+  as_tibble()
+
+post_summary <- posterior_summary(model_M1, robust = TRUE)
+post_summary[1:30,]
+
+post_qi_b <- post %>%
+  select(starts_with("b_")) %>%
+  pivot_longer(everything()) %>%
+  group_by(name) %>% 
+  median_qi(value) %>% 
+  arrange(name)
+head(post_qi_b) # 30 "fixed" effects 
+```
+
+Visualise fixed effects.
+
+```{r}
+tidy_fixed <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) %>% 
+  mutate(timebin = str_sub(name,10,11),
+         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
+         condition = str_sub(name,17,18),
+         condition = factor(str_remove(condition,"e"),
+                            levels=c(1,2,3),
+                            labels=c("blank","congruent","incongruent")))
+head(tidy_fixed)
+tail(tidy_fixed)
+
+# plot
+p_tidy_fixed <- ggplot(tidy_fixed, aes(x = timebin, y = value, fill=condition)) +  
+  stat_halfeye(alpha=0.7) +
+  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1',
+       x = "time bin", y = "cloglog-hazard") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+   theme(axis.text.x = element_text(angle=90)) +
+  facet_wrap(~condition)
+p_tidy_fixed
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_postdistr.png", width = 10, height = 8, dpi = 800)
+```
 
@@
 
-We can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, and for each participant in the data set.
-
-Figure X
+We can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, for each participant in the data set, and for a brand new hypothetical participant.
+
+For the average participant, using add_epred_draws().
+
+```{r}
+dat_M1 <- as_tibble(model_M1$data)
+
+epreds_grand <- dat_M1 %>% 
+  data_grid(timebin, prime)  %>% 
+  add_epred_draws(model_M1, re_formula = NA) %>% 
+  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
+```
+
+Summarize and plot predicted hazard values.
+
+```{r}
+epreds_grand %>% group_by(timebin,prime) %>% mean_qi()
+```
+
+```{r}
+ ggplot(epreds_grand, aes(x=timebin, y=.epred, 
+                       fill=prime, color=prime)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,.6)) +
+    labs(y = "predicted hazard") +
+    ggtitle("Average participant")
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_pred_grand.png", width = 10, height = 8, dpi = 800)
+```
+
+Again, for the average participant, but using posterior draws.
+
+```{r}
+tidy_fixed %>%
+  mutate(haz = 1-exp((-1)*exp(value))) %>% # inverse cloglog
+
+ggplot(aes(x=timebin, y=haz, 
+                       fill=condition, color=condition)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,.6)) +
+    labs(y = "predicted hazard") +
+    ggtitle("Average participant")
+```
+
+Second, for each participant in the data set.
+
+```{r}
+epreds_pid <- dat_M1 %>% 
+  data_grid(pid,timebin, prime)  %>% 
+  add_epred_draws(model_M1, re_formula = NULL) %>% 
+  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
+```
+
+Summarize and plot predicted hazard values.
+
+```{r}
+epreds_pid %>% group_by(pid,timebin,prime) %>% mean_qi()
+```
+
+```{r}
+ ggplot(epreds_pid, aes(x=timebin, y=.epred, 
+                       fill=prime, color=prime)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,1)) +
+    labs(y = "predicted hazard") +
+    ggtitle("Each participant (N = 6)") +
+    facet_wrap(~pid, ncol=3)
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_pred_pid.png", width = 10, height = 8, dpi = 800)
+```
+
+Third, for a completely new hypothetical participant.
+
+```{r}
+epreds_pid_new <- 
+  posterior_epred(model_M1, 
+                  newdata = expand_grid(timebin=6:15,prime=1:3,pid=7), 
+                  re_formula = NULL, 
+                  allow_new_levels=T, 
+                  sample_new_levels = "gaussian") %>% 
+  as_tibble %>%
+  mutate(draw = 1:n()) %>%
+  pivot_longer(V1:V30,values_to = ".epred", names_to = "condition") %>%
+
+  mutate(prime = factor(rep(c(1,2,3), 10*8000), levels=c(1,2,3),labels=c("blank","congruent","incongruent")),
+         timebin = factor(rep(rep(c(6:15),each=3),8000),levels=c(6:15)  )) %>%
+  ungroup()
+```
+
+```{r}
+epreds_pid_new <- dat_M1 %>% 
+  data_grid(timebin, prime)  %>%
+  mutate(pid = 7) %>%
+  add_epred_draws(model_M1, re_formula = NULL, allow_new_levels=T,    sample_new_levels = "gaussian") %>% 
+  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
+```
+
+Summarize and plot predicted hazard values.
+
+```{r}
+epreds_pid_new %>% group_by(pid,timebin,prime) %>% mean_qi()
+```
+
+```{r}
+ ggplot(epreds_pid_new, aes(x=timebin, y=.epred, 
+                       fill=prime, color=prime)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,1)) +
+    labs(y = "predicted hazard") +
+    ggtitle("A brand new hypothetical participant") 
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_pred_pid_new.png", width = 10, height = 8, dpi = 800)
+```
 
@@
 
+We are actually interested in the difference in predicted hazard between congruent and blank primes on the one hand, and between incongruent and blank primes on the other hand, for each time bin.
+When based on the grand mean, this is known as the grand average marginal effect (AME).
+
+```{r}
+epreds_grand_diffs <- epreds_grand %>%
+  pivot_wider(id_cols = c(timebin, .draw),
+              names_from = "prime",
+              values_from = ".epred") %>% #80000 x 5
+  mutate(`congruent minus blank` = congruent - blank,
+         `incongruent minus blank` = incongruent -  blank) %>% 
+  select(-c(blank,congruent,incongruent)) %>% 
+  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
+               names_to = "contrast",
+               values_to = "pred_haz")
+
+epreds_grand_diffs # 160 000 rows 
+```
+
+Summarize and plot.
+
+```{r}
+epreds_grand_diffs %>%
+  group_by(contrast,timebin) %>%
+    mean_qi(.width = c(.95)) %>%
+  arrange(contrast,timebin,.width)
+```
+
+!!!!!  Make a table ...
+
+```{r}
+ ggplot(epreds_grand_diffs, aes(x=timebin, y=pred_haz, 
+                       fill=contrast)) +
+  stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.8,.95)) +
+ # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
+  geom_hline(yintercept = 0, color = "red", lty = 3) +
+  scale_y_continuous(limits=c(-0.5,0.3)) +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(y = "difference in predicted hazard") +
+  ggtitle("Grand average marginal effect") +
+  facet_wrap(~contrast)
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_grand_AME.png", width = 10, height = 8, dpi = 800)
+```
+
+Note that these grand AMEs can also be calculated based on the posterior draws.
+
+```{r}
+post %>% # 8000 x 30 = 240000
+  mutate_all(function(x){1-exp((-1)*exp(x))}) %>%
+  mutate(bin6_cb = `b_timebin6:prime2` - `b_timebin6:prime1`,
+         bin7_cb = `b_timebin7:prime2` - `b_timebin7:prime1`,
+         bin8_cb = `b_timebin8:prime2` - `b_timebin8:prime1`,
+         bin9_cb = `b_timebin9:prime2` - `b_timebin9:prime1`,
+         bin10_cb = `b_timebin10:prime2` - `b_timebin10:prime1`,
+         bin11_cb = `b_timebin11:prime2` - `b_timebin11:prime1`,
+         bin12_cb = `b_timebin12:prime2` - `b_timebin12:prime1`,
+         bin13_cb = `b_timebin13:prime2` - `b_timebin13:prime1`,
+         bin14_cb = `b_timebin14:prime2` - `b_timebin14:prime1`,
+         bin15_cb = `b_timebin15:prime2` - `b_timebin15:prime1`,
+         bin6_ib = `b_timebin6:prime3` - `b_timebin6:prime1`,
+         bin7_ib = `b_timebin7:prime3` - `b_timebin7:prime1`,
+         bin8_ib = `b_timebin8:prime3` - `b_timebin8:prime1`,
+         bin9_ib = `b_timebin9:prime3` - `b_timebin9:prime1`,
+         bin10_ib = `b_timebin10:prime3` - `b_timebin10:prime1`,
+         bin11_ib = `b_timebin11:prime3` - `b_timebin11:prime1`,
+         bin12_ib = `b_timebin12:prime3` - `b_timebin12:prime1`,
+         bin13_ib = `b_timebin13:prime3` - `b_timebin13:prime1`,
+         bin14_ib = `b_timebin14:prime3` - `b_timebin14:prime1`,
+         bin15_ib = `b_timebin15:prime3` - `b_timebin15:prime1`) %>%
+  select(starts_with("bin"))  %>% # 8000 x 20
+
+  pivot_longer(cols = bin6_cb:bin15_ib, names_to = "condition", values_to = "haz_diff") %>%
+  mutate(bin = str_sub(condition,4,5),
+         bin = str_remove(bin,"_"),
+         bin = factor(bin, levels=c(6:15)),
+         comp = str_sub(condition,6,8),
+         comp = str_remove(comp,"_"),
+         contrast = ifelse(comp == "cb", "congruent minus blank", "incongruent minus blank")) %>%
+
+  ggplot(aes(x=bin, y = haz_diff, fill = contrast)) +
+    stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
+    geom_hline(yintercept = 0, color = "red", lty = 3) +
+    scale_y_continuous(limits=c(-0.5,0.3)) +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    theme(axis.text.x = element_text(angle=90)) +
+    labs(y = "difference in predicted hazard", x = "timebin") +
+    ggtitle("Grand average marginal effect") +
+    facet_wrap(~contrast)
+```
+
 ### Subject-specific average marginal effects
 
-And at the group level:
+To calculate the subject-specific AMEs for each time bin, we create contrasts between the conditional means.
+
+```{r}
+epreds_pid_diffs <- epreds_pid %>%
+  pivot_wider(id_cols = c(pid,timebin, .draw),
+              names_from = "prime",
+              values_from = ".epred") %>% #80000 x 5
+  mutate(`congruent minus blank` = congruent - blank,
+         `incongruent minus blank` = incongruent -  blank) %>% 
+  select(-c(blank,congruent,incongruent)) %>% 
+  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
+               names_to = "contrast",
+               values_to = "pred_haz") 
+epreds_pid_diffs # 160 000 rows 
+```
+
+Summarize and plot the contrasts in predicted hazard values.
+
+```{r}
+epreds_pid_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
+```
+
+```{r}
+ ggplot(epreds_pid_diffs, aes(x=timebin, y=pred_haz, 
+                       fill=contrast, color=contrast)) +
+    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
+    geom_hline(yintercept = 0, color = "red", lty = 3) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(-0.6,0.6)) +
+    labs(y = "difference in predicted hazard") +
+    ggtitle("Subject-specific AMEs") +
+    facet_wrap(~pid, ncol=3)
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_pid_AMEs.png", width = 10, height = 8, dpi = 800)
+```
+
+[[remove next section....]]
+And check whether an average of subject-specific AMEs is equal to the grand AME... NO !!!
+
+```{r}
+epreds_pid_diffs_av <- epreds_pid_diffs %>% group_by(timebin,contrast,.draw) %>%
+  mutate(average_condmeans = mean(pred_haz))
+```
+
+```{r}
+ ggplot(epreds_pid_diffs_av, aes(x=timebin, y=pred_haz, 
+                       fill=contrast)) +
+  stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.8,.95)) +
+ # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
+  geom_hline(yintercept = 0, color = "red", lty = 3) +
+  scale_y_continuous(limits=c(-0.5,0.3)) +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(y = "difference in predicted hazard") +
+  ggtitle("Average of subject-specific AMEs") +
+  facet_wrap(~contrast)
+```
+
+### AMEs for a brand new hypothetical participant
+
+```{r}
+epreds_pid_new_diffs <- epreds_pid_new %>%
+  pivot_wider(id_cols = c(pid,timebin, .draw),
+              names_from = "prime",
+              values_from = ".epred") %>% #80000 x 5
+  mutate(`congruent minus blank` = congruent - blank,
+         `incongruent minus blank` = incongruent -  blank) %>% 
+  select(-c(blank,congruent,incongruent)) %>% 
+  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
+               names_to = "contrast",
+               values_to = "pred_haz") 
+epreds_pid_new_diffs # 160 000 rows 
+```
+
+Summarize and plot the contrasts in predicted hazard values.
+
+```{r}
+epreds_pid_new_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
+```
+
+```{r}
+ ggplot(epreds_pid_new_diffs, aes(x=timebin, y=pred_haz, 
+                       fill=contrast, color=contrast)) +
+    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
+    geom_hline(yintercept = 0, color = "red", lty = 3) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(-0.7,0.7)) +
+    labs(y = "difference in predicted hazard") +
+    ggtitle("AMEs for a brand new\nhypothetical participant") +
+    facet_wrap(~contrast)
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1_pid_new_AMEs.png", width = 10, height = 8, dpi = 800)
+```
 
@@
 What can we conclude from model M1 about our research question, i.e., what is the temporal dynamics of the effect of prime-target congruency on RT?
-In other words, in which time bins does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of xxx ms)?
+In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of xxx ms)?
+
+We can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
+The contrast "congruent minus blank" was estimated to be 0.0925 hazard units in bin 6 (95% CrI = [0.0411, 0.183]), and 0.152 hazard units in bin 7 (95% CrI = [0.0489, 0.262]).
+For the other bins, the 95% credible interval contained zero.
+The contrast "incongruent minus blank" was estimated to be 0.0943 hazard units in bin 6 (95% CrI = [0.0322, 0.208]), -0.204 hazard units in bin 9 (95% CrI = [-0.331, -0.0769]), -0.275 hazard units in bin 10 (95% CrI = [-0.45, -0.0513]), and -0.213 hazard units in bin 11 (95% CrI = [-0.377, -0.0137]).
+For the other bins, the 95% credible interval contained zero.
+
+There are thus two phases of performance for the average subject between 200 and 600 ms after target onset.
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240].
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
+
+The posterior distribution of each contrast can also be summarized by considering its proportion below or above some value, like zero.
+For example, here are the proportions that each contrast is larger and smaller than 0:
+
+```{r}
+pabove <- epreds_grand_diffs %>% group_by(timebin,contrast) %>%
+    summarize(prop_above = mean(pred_haz > 0)) %>% arrange(contrast,timebin)
+pbelow <- epreds_grand_diffs %>% group_by(timebin,contrast) %>%
+    summarize(prop_below = mean(pred_haz < 0)) %>% arrange(contrast,timebin)
+pabove %>% inner_join(pbelow, by = c("timebin","contrast"))
+```
+
+Thus the probability that the contrast "congruent minus blank" is larger than 0, is larger than .80 in bins 6 to 9.
+And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .80 in bins 8 to 13.
+
+#####################
 

### 2024-11-20 Tutorial_2a.Rmd
add code to save table, and upload csv file

@@
 ```{r}
-epreds_grand_diffs %>%
+table <- epreds_grand_diffs %>%
   group_by(contrast,timebin) %>%
     mean_qi(.width = c(.95)) %>%
-  arrange(contrast,timebin,.width)
-```
-
-!!!!!  Make a table ...
+  arrange(contrast,timebin,.width) %>%
+  select(-c(.width,.point,.interval))
+```
+
+```{r save-table-grand-AMEs, eval=F}
+write_csv(table, file="Tutorial_2_Bayesian/tables/grand_AMEs.csv")
+```
 
@@
 In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
+The effect of an incongruent prime thus depends on how much waiting time has passed since target onset.
 

### 2024-11-21 Tutorial_2a.Rmd
add conclusions for M1 to 2a

@@
 
-We can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
+If we want to study the average effect of prime type on hazard, uncontaminated by interindividual differences, we can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
 The contrast "congruent minus blank" was estimated to be 0.0925 hazard units in bin 6 (95% CrI = [0.0411, 0.183]), and 0.152 hazard units in bin 7 (95% CrI = [0.0489, 0.262]).
@@
 For the other bins, the 95% credible interval contained zero.
+Note that we could also have calculated hazard ratios instead of hazard differences.
 
@@
 In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
-The effect of an incongruent prime thus depends on how much waiting time has passed since target onset.
+The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
 
@@
 And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .80 in bins 8 to 13.
+
+Plot of posterior distributions of sigma(timebinX:primeY) : check size differences between timebins and prime types?
+
+If we want to focus more on interindividual differences, we can study the subject-specific AMEs in Figure XX.
+While most participants show the grand AMEs, three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] - subject 2 also in bin (320,360].
+This dip is also present in the AMEs for a brand new hypothetical participant.
+
+Future studies could (a) increase the number of participants to estimate the porportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
+For example, Panis and Schmidt (ref) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here:the target).
+Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the experiment-wide time scale, next to the trial-wide time scale (TIME).
+In Tutorial_2a.Rmd we provide a number of model formula that should get you going.
 

### 2024-11-23 Tutorial_2a.Rmd
update tut 2a + add png

@@
 
+Perform a prior predictive check.
+
+```{r}
+plan(multicore)
+
+model_M0_prior <-                     # wrong data file name
+   brm(data = M0_data,
+       family = bernoulli(link="cloglog"),
+       event ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = 'only',
+       file = "Tutorial_2_Bayesian/models/model_M0_prior")
+```
+
+```{r}
+
+#pp_check(model_M0_prior, stat="mean")
+
+#http://bruno.nicenboim.me/bayescogsci/
+
+# prior predictive check for 1 time bin
+cloglog_predictive_distribution <-
+  function(samples, N_obs) {
+    # empty data frame with headers:
+    df_pred <- tibble(trial = numeric(0),
+                      y_pred = numeric(0),
+                      iter = numeric(0))
+    # i iterates from 1 to the length of samples,
+      for (i in seq_along(samples)) {
+        cloglog_haz <- samples[i]
+        df_pred <- bind_rows(df_pred,
+                             tibble(trial = 1:N_obs,
+                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
+                                    iter = i))
+    }
+    df_pred # return
+  }
+# simulate prior hazard values for 1 bin
+N_samples <- 3000
+N_obs <- 100 #nrow(M0_data)
+#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
+cloglog_samples <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept
+
+prior_pred <- cloglog_predictive_distribution(cloglog_samples,N_obs)
+
+prior_pred_haz <- prior_pred %>% group_by(iter) %>% summarise(pred_haz = mean(y_pred))
+
+# plot 
+ggplot(prior_pred_haz, aes(x=pred_haz)) +
+  geom_histogram(binwidth=.01) +
+  scale_x_continuous(limits = c(0,1))
+```
+
+```{r}
+# Generate prior predictive samples for 10 bins 
+bins_cloglog_predictive_distribution <-
+  function(samples_bin, N_obs, N_bins) {
+    # empty data frame with headers:
+    df_pred <- tibble(trial = numeric(0),
+                      y_pred = numeric(0),
+                      iter = numeric(0),
+                      bin = numeric(0))
+    # i iterates from 1 to the length of samples,
+    # b interates from 1 to N_bins
+    for( b in 1:N_bins){
+      for (i in 1:dim(samples_bin)[2]) {
+        cloglog_haz <- samples_bin[b,i]
+        df_pred <- bind_rows(df_pred,
+                             tibble(trial = 1:N_obs,
+                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
+                                    iter = i,
+                                    bin = b))
+      }
+    }
+    df_pred # return
+  }
+
+# simulate prior hazard functions for 10 bin
+N_samples <- 40
+N_obs <- 100
+N_bins <- 10
+#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
+cloglog_samples_bin <- data.frame()
+for(b in 1:N_bins){
+  cloglog_samples_bin[b,1:N_samples] <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept
+}
+
+prior_pred <- bins_cloglog_predictive_distribution(cloglog_samples_bin,N_obs,N_bins)
+
+prior_pred_haz <- prior_pred %>% group_by(iter,bin) %>% summarise(pred_haz = mean(y_pred))
+
+# plot 
+ggplot(prior_pred_haz, aes(x=bin, y=pred_haz, group=iter)) +
+  #geom_histogram(binwidth=.01) +
+  geom_line(color="black") +
+  geom_line(data = prior_pred_haz %>% filter(iter==4), color="red", linewidth = 2) +
+  geom_line(data = prior_pred_haz %>% filter(iter==18), color="blue", linewidth = 2) +
+
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(1,10), breaks = c(1:10))
+
+```
+
+```{r}
+# Generate prior predictive hazard functions for 6 participants
+newdata_prior = tibble(timebin = c(6:15)) %>% 
+                            expand_grid(pid = 1:6)
+N_obs = 100
+
+# prepare columns of data set
+df_pred <- tibble(timebin = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
+  prior_pred <- prior_pred %>% mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+head(df_pred)
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% group_by(pid,timebin,.draw) %>% summarise(pred_haz = mean(.prediction))
+summary(df_pred_haz)
+
+# plot 40 draws per participant
+ggplot(df_pred_haz %>% filter(.draw < 21), aes(x=timebin, y=pred_haz, group=.draw)) +
+
+   geom_line(color="black") +
+   geom_line(data=df_pred_haz %>% filter(.draw == 20), color="red", linewidth=2) +
+
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
+  facet_wrap(~pid)
+
+```
+
 Fit model M0.
@@
 formula(model_M0c)
+```
+
+## Intermezzo: Prior predictive checks
+
+```{r fit-model-M0c, eval=F}
+plan(multicore)
+
+model_M0c_prior <-
+   brm(data = M0c_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq + period_9_cu + 
+                          (0 + Intercept + period_9 + period_9_sq + period_9_cu  | pid),
+       prior = priors_M0c,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior")
+```
+
+```{r}
+# Generate prior predictive hazard functions for 6 participants
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2,
+         period_9_cu = period_9^3) %>%
+  expand_grid(pid = 1:6)
+
+N_obs = 100
+
+# prepare columns of data set
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = double(0),
+                  period_9_cu = double(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0c_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
+  prior_pred <- prior_pred %>% mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+head(df_pred)
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% group_by(pid,period_9,.draw) %>% summarise(pred_haz = mean(.prediction))
+summary(df_pred_haz)
+
+# plot 40 draws per participant
+ggplot(df_pred_haz %>% filter(.draw < 2), aes(x=period_9, y=pred_haz, group=.draw)) +
+
+   geom_line(color="black") +
+   geom_line(data=df_pred_haz %>% filter(.draw == 17), color="red", linewidth=2) +
+
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+```
+
+##
+
+```{r}
+dat <- tibble(x = -3:6,
+              x2 = x^2,
+              x3 = x^3,
+              cloglog1 = -1.8   + 0.3*x   - 0.1*x2   + 0.009*x3,
+              cloglog2 = -1     + 0.5*x   - 0.1*x2   + 0.004*x3,
+              cloglog3 = -3     - 0.1*x   + 0.12*x2  - 0.002*x3,
+              cloglog4 = -0.40  + 0.2*x   - 0.04*x2  + 0.003*x3,
+              haz1 = 1 - exp(-1*exp(cloglog1)),
+              haz2 = 1 - exp(-1*exp(cloglog2)),
+              haz3 = 1 - exp(-1*exp(cloglog3)),
+              haz4 = 1 - exp(-1*exp(cloglog4)))
+
+p1<-ggplot(dat, aes(x=x)) +  
+      geom_line(aes(y=cloglog1), color="black") +
+      geom_line(aes(y=cloglog2), color="green") +
+      geom_line(aes(y=cloglog3), color="red") +
+      geom_line(aes(y=cloglog4), color="blue") +
+      scale_y_continuous(limits=c(-6,2)) +
+      scale_x_continuous(breaks=c(-3:6))
+
+p2<-ggplot(dat, aes(x=x))+
+      geom_line(aes(y=haz1), color="black") + 
+      geom_line(aes(y=haz2), color="green") +
+      geom_line(aes(y=haz3), color="red") +
+      geom_line(aes(y=haz4), color="blue") +
+      scale_y_continuous(limits=c(0,1)) +
+      scale_x_continuous(breaks=c(-3:6))
+
+p1|p2
+```
+
+```{r}
+tibble(cloglog = rnorm(1000, .1, .1)) %>%
+  ggplot(aes(x=cloglog)) +
+  geom_density()
+```
+
+```{r data-priors-M0c}
+M0c_data <- ptb_data %>% select(pid, event, period_9) %>%
+  mutate(period_9_sq = period_9^2)
+
+priors_M0c_prior <- c(
+  #set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
+ set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
+
+#set_prior("normal(0.5,.5)",class="b",coef="period_9"),
+#set_prior("normal(-.1,.2)",class="b",coef="period_9_sq"),   # Np3
+
+set_prior("normal(0.1,.08)",class="b",coef="period_9"), # ok !!!
+set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
+
+#set_prior("skew_normal(0.3,0.1,2)",class="b",coef="period_9"), #skNp3
+#set_prior("skew_normal(0,0.1,0)",class="b",coef="period_9_sq"),
+
+ set_prior("normal(0, .01)", class = "sd"),  
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c, eval=F}
+plan(multicore)
+
+model_M0_lin_prior448 <-
+   brm(data = M0c_data,
+       family = binomial(link="cloglog"),
+       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq +
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_prior,
+       chains = 4, cores = 4, iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0_lin_prior448")
+```
+
+```{r}
+post <- as_draws_df(model_M0_lin_prior448) %>% 
+  select(-lp__) %>% 
+  as_tibble()
+
+tidy_fixed <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) 
+head(tidy_fixed)
+tail(tidy_fixed)
+
+# plot
+p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = name)) +  
+  stat_halfeye(alpha=0.7,point_interval = "median_qi",.width=c(.8,.99) ) 
+p_tidy_fixed
+
+```
+
+```{r}
+#model_M0_lin_prior14 <-readRDS("Tutorial_2_Bayesian/models/model_M0c_lin_prior14.rds")               
+# Generate prior predictive hazard functions for 6 participants
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+N_obs = 150 # increase for smoother results
+
+# prepare columns of data set
+df_pred <- tibble(period_9 = integer(0),
+                 period_9_sq = integer(0),
+                 pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+set.seed(1)
+
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0_lin_prior448, 
+                                    newdata = newdata_prior, 
+                                    summary=F, 
+                                    ndraws=NULL)
+  prior_pred <- prior_pred %>% 
+    mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+head(df_pred)
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction)) 
+summary(df_pred_haz)
+
+# plot 40 draws per participant
+sel = sample(1:8000,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <7639), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   geom_line(data=df_pred_haz %>% filter(.draw ==7418), 
+             color="red", linewidth=2) +
+  geom_line(data=df_pred_haz %>% filter(.draw == 6001), 
+            color="green", linewidth=2) +
+  geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
 ```
@@
   prior(exponential(1), class = sd, group = pid, nlpar = b),
+  prior(exponential(1), class = sd, group = pid, nlpar = c),
   prior(lkj(2), class = cor, group = pid)
@@
 
-For the average participant, using add_epred_draws().
+First, for the average participant, using add_epred_draws().
 
@@
 
-Plot of posterior distributions of sigma(timebinX:primeY) : check size differences between timebins and prime types?
+[[todo or not todo]]
+Plot of posterior distributions of sigma(timebinX:primeY) : check size differences between timebins and prime types? -> these are also population-level effects
 
@@
 Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the experiment-wide time scale, next to the trial-wide time scale (TIME).
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale (TIME).
 In Tutorial_2a.Rmd we provide a number of model formula that should get you going.

### 2024-11-23 Tutorial_2a.Rmd
update 2a cosmetics

@@
                                     summary=F, 
-                                    ndraws=NULL)
-  prior_pred <- prior_pred %>% 
+                                    ndraws=NULL) %>% 
                 mutate(obs = i)
@@
   data_grid(timebin, prime) %>% 
-  add_epred_draws(model_M1, re_formula = NA) %>% 
-  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  add_epred_draws(model_M1, 
+                  re_formula = NA) %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
   ungroup()
@@
 ```{r}
-epreds_grand %>% group_by(timebin,prime) %>% mean_qi()
+epreds_grand %>% 
+  group_by(timebin,prime) %>% 
+  mean_qi()
 ```
@@
   data_grid(pid,timebin, prime) %>% 
-  add_epred_draws(model_M1, re_formula = NULL) %>% 
-  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  add_epred_draws(model_M1, 
+                  re_formula = NULL) %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
   ungroup()
@@
 ```{r}
-epreds_pid %>% group_by(pid,timebin,prime) %>% mean_qi()
+epreds_pid %>% 
+  group_by(pid,timebin,prime) %>% 
+  mean_qi()
 ```
@@
   posterior_epred(model_M1, 
-                  newdata = expand_grid(timebin=6:15,prime=1:3,pid=7), 
+                  newdata = expand_grid(timebin=6:15,
+                                        prime=1:3,
+                                        pid=7), 
                   re_formula = NULL, 
@@
   mutate(draw = 1:n()) %>%
-  pivot_longer(V1:V30,values_to = ".epred", names_to = "condition") %>%
-
-  mutate(prime = factor(rep(c(1,2,3), 10*8000), levels=c(1,2,3),labels=c("blank","congruent","incongruent")),
-         timebin = factor(rep(rep(c(6:15),each=3),8000),levels=c(6:15)  )) %>%
+  pivot_longer(V1:V30,
+               values_to = ".epred", 
+               names_to = "condition") %>%
+  mutate(prime = factor(rep(c(1,2,3), 10*8000), 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent")),
+         timebin = factor(rep(rep(c(6:15),each=3),8000),
+                          levels=c(6:15))) %>%
   ungroup()
@@
   mutate(pid = 7) %>%
-  add_epred_draws(model_M1, re_formula = NULL, allow_new_levels=T,    sample_new_levels = "gaussian") %>% 
-  mutate(prime = factor(prime, levels=c(1,2,3),labels=c("blank","congruent","incongruent"))) %>%
+  add_epred_draws(model_M1, 
+                  re_formula = NULL, 
+                  allow_new_levels=T,    
+                  sample_new_levels = "gaussian") %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
   ungroup()
@@
 ```{r}
-epreds_pid_new %>% group_by(pid,timebin,prime) %>% mean_qi()
+epreds_pid_new %>% 
+  group_by(pid,timebin,prime) %>% 
+  mean_qi()
 ```
@@
                        fill=contrast)) +
-  stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.8,.95)) +
+  stat_lineribbon(alpha = 0.5, 
+                  point_interval = "mean_qi",
+                  .width=c(.8,.95)) +
  # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
-  geom_hline(yintercept = 0, color = "red", lty = 3) +
+  geom_hline(yintercept = 0, 
+             color = "red", 
+             lty = 3) +
   scale_y_continuous(limits=c(-0.5,0.3)) +
@@
 
-  pivot_longer(cols = bin6_cb:bin15_ib, names_to = "condition", values_to = "haz_diff") %>%
+  pivot_longer(cols = bin6_cb:bin15_ib, 
+               names_to = "condition", 
+               values_to = "haz_diff") %>%
   mutate(bin = str_sub(condition,4,5),
@@
          comp = str_remove(comp,"_"),
-         contrast = ifelse(comp == "cb", "congruent minus blank", "incongruent minus blank")) %>%
-
-  ggplot(aes(x=bin, y = haz_diff, fill = contrast)) +
-    stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
-    geom_hline(yintercept = 0, color = "red", lty = 3) +
+         contrast = ifelse(comp == "cb", 
+         "congruent minus blank", "incongruent minus blank")) %>%
+
+  ggplot(aes(x=bin, y=haz_diff, fill=contrast)) +
+    stat_halfeye(point_interval = "mean_qi",
+                 .width=c(.8,.95)) +
+    geom_hline(yintercept = 0, 
+               color = "red", 
+               lty = 3) +
     scale_y_continuous(limits=c(-0.5,0.3)) +
@@
 epreds_pid_new_diffs <- epreds_pid_new %>%
-  pivot_wider(id_cols = c(pid,timebin, .draw),
+  pivot_wider(id_cols = c(pid,timebin,.draw),
               names_from = "prime",
-              values_from = ".epred") %>% #80000 x 5
+              values_from = ".epred") %>% 
   mutate(`congruent minus blank` = congruent - blank,
@@
                values_to = "pred_haz") 
-epreds_pid_new_diffs # 160 000 rows 
+epreds_pid_new_diffs 
 ```
@@
                        fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    stat_lineribbon(alpha = 0.5, 
+                    point_interval = "mean_qi",
+                    .width=c(.95)) +
     #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
@@
 
-What can we conclude from model M1 about our research question, i.e., what is the temporal dynamics of the effect of prime-target congruency on RT?
+What can we conclude from model M1 about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
 In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of xxx ms)?
 
-If we want to study the average effect of prime type on hazard, uncontaminated by interindividual differences, we can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
 The contrast "congruent minus blank" was estimated to be 0.0925 hazard units in bin 6 (95% CrI = [0.0411, 0.183]), and 0.152 hazard units in bin 7 (95% CrI = [0.0489, 0.262]).
@@
 ```{r}
-pabove <- epreds_grand_diffs %>% group_by(timebin,contrast) %>%
-    summarize(prop_above = mean(pred_haz > 0)) %>% arrange(contrast,timebin)
-pbelow <- epreds_grand_diffs %>% group_by(timebin,contrast) %>%
-    summarize(prop_below = mean(pred_haz < 0)) %>% arrange(contrast,timebin)
+pabove <- epreds_grand_diffs %>% 
+  group_by(timebin,contrast) %>%
+  summarize(prop_above = mean(pred_haz > 0)) %>% 
+  arrange(contrast,timebin)
+
+pbelow <- epreds_grand_diffs %>% 
+  group_by(timebin,contrast) %>%
+  summarize(prop_below = mean(pred_haz < 0)) %>% 
+  arrange(contrast,timebin)
+
 pabove %>% inner_join(pbelow, by = c("timebin","contrast"))
@@
 
-If we want to focus more on interindividual differences, we can study the subject-specific AMEs in Figure XX.
+If we want to focus more on inter-individual differences, we can study the subject-specific AMEs in Figure XX.
 While most participants show the grand AMEs, three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] - subject 2 also in bin (320,360].
@@
 
-Future studies could (a) increase the number of participants to estimate the porportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
+Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
 For example, Panis and Schmidt (ref) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here:the target).

### 2024-11-24 Tutorial_1a.Rmd
update tutorials 1a and 1b

--- 2024-10-24 Tutorial_1a.Rmd
@@
 ---
-title: "Tutorial_1a"
+title: "Tutorial 1a"
 author: "Sven Panis"
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.
-
-Content:
-1. Load libraries
-2. Read and wrangle data
-3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
-4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models
-5. Optional: set up life tables and make plots for data aggregated across participants
+This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to calculate descriptive statistics.
+It uses functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
+To set up a life table, the user has to select a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (i.e., a cutoff time, or timeout).
+The goal is to find the optimal bin width and censoring time (both in ms) to portray the effects of interest.
+Note that all response times must be positive.
 
@@
 
-# 2. Read and wrangle data
+Theme settings for ggplot.
 
-Read in the raw data from Panis and Schmidt (2016) - can be skipped.
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"))
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+# 2. Read and wrangle data.
+
+Read in the raw data file from Panis and Schmidt (2016) - can be skipped.
 
 ```{r read-raw-data}
-raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) # 9372 rows
-head(raw)
-#tail(raw)
-#glimpse(raw)
+raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", 
+                  delim="\t",
+                  col_names=T)
 ```
@@
 # remove practice trials
-data <- raw %>% filter(practice==0) # 9240 experimental trials: 6 participants x 70 blocks x 22 trials-per-block
+data <- raw %>% filter(practice==0) # 9240 experimental trials: 6 subj x 70 blocks x 22 trials-per-block
 
@@
                            void,target_dir,corresp,resperr))
+
 # look at trials where resp==0
-data %>% filter(resp==0) 
 data %>% filter(resp==0) %>% print(n=51)
@@
 sel = which(data$resp==0 & data$rt < 800)
-data <- data[-c(sel),] # 9235 rows
+data <- data[-c(sel),]
 
@@
 data <- data %>% filter(mask_type == 0) %>%
-                 select(-mask_type) # 2757 rows and 8 variables
+                 select(-mask_type)
+
 summary(data)
@@
 
-# change some variable names to required names (pid, trial, condition, rt, and acc)
+# change some variable names to required names: pid, trial, condition, rt, and acc.
 colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial")
+
 # pid to indicate column with unique person IDs
@@
 data_wr <- data_wr %>% mutate(condition = condition + 1,
-                              condition = factor(condition, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
+                              condition = factor(condition, 
+                                                 levels = c(1,2,3), 
+                                                 labels = c("blank","congruent","incongruent")))
 summary(data_wr)
@@
 
-Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 118 in custom_functions.R) as above (line 88 in Tutorial_1a).
+Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 134-136 in custom_functions.R) as above (lines 96-99).
 
@@
 ```{r apply-functions}
-# create a dataframe with two columns (pid and data); the latter contains a list of nested data frames
+# create a dataframe with two columns (pid and data); 
+# the latter contains a list of nested data frames
 data_nested <- data_wr %>% group_nest(pid)
+
 # apply the functions using map() and map2()
 data_final <- data_nested %>% 
-                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
-                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
-                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
-                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
-                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+    # ! user input: censoring time, and bin width
+    mutate(censored = map(data, censor, 600, 40)) %>%
+    # create person-trial-bin data set
+    mutate(ptb_data = map(censored, ptb)) %>%
+    # create life tables without ca
+    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
+    # calculate ca
+    mutate(condacc = map(censored, calc_ca)) %>%     
+    # create life tables with ca
+    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
+    # create plots, and set haz_yaxis to 1, and first_bin_shown to 2
+    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha, 1, 2))  
+    # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
+    #            Recommended to set this to 1 when plotting the data of individual subjects.
+    # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
+    #            you can set first_bin_shown to another value, such as 5 (i.e., do not show the 
+    #            first 4 bins in the plots).
 ```
@@
 
-# 4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models
+# 4. Extract info (life table, plots, ...) and create files for fitting hazard and conditional accuracy models
 
@@
 # view original data (column 2) for subject 1:
-pluck(data_final, 2, 1) %>% print(n=20)
+pluck(data_final, 2, 1)
 
@@
 # save the plot for each subject
-map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 6, height = 8, dpi = 600)
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 8, height = 8, dpi = 800)
 ```
@@
 # get the binsize in ms
-binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
+binsize    <- data_final$censored[[1]] %>% 
+                pull(bin_width) %>% 
+                max(na.rm=T)
+
 # extract the lifetable (column 7 in data_final) for condition 1 (neutral) for subject 6
@@
                                events = event1) %>%
-                        select(bin,risk_set,events,hazard,se_haz,survival,se_surv,ca,se_ca) 
+                        select(bin,risk_set,events,hazard,se_haz,
+                              survival,se_surv,ca,se_ca)
+
 # get riskset for first bin
-riskset <- lifetable_blank_s6 %>% select(risk_set) %>% slice(1) %>% pull()
+riskset <- lifetable_blank_s6 %>% 
+  select(risk_set) %>% 
+  slice(1) %>% 
+  pull()
 
 # create a row for time point 0
-row0 <- c(bin=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,survival=1,se_surv=0,ca=NA,se_ca=NA)
+row0 <- c(bin=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,
+          survival=1,se_surv=0,ca=NA,se_ca=NA)
+
 # add new row
@@
 input_modeling_hazard <- data_final %>% 
-                         select(pid, ptb_data) %>%  unnest(ptb_data) %>%
+                           select(pid, ptb_data) %>%  
+                           unnest(ptb_data) %>%
                            select(pid,bl,tr,trial,condition,period,event)
@@
 ```{r create-file-ca-modeling}
-# select the censored data set for each subject, unnest them, discard trials that are censored (RT > 600 or no response detected before the deadline), and select relevant variables. 
+# select the censored data set for each subject, unnest them, discard trials that are 
+# censored (RT > 600 or no response detected before the deadline), and select relevant variables. 
 input_modeling_ca <- data_final %>%
-                     select(pid, censored) %>% unnest(censored) %>%
+                       select(pid, censored) %>% 
+                       unnest(censored) %>%
                        filter(right_censored == 0) %>% # only keep uncensored trials
@@
 data_agg_final <- data_agg_nested %>% 
-                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
-                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
-                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
-                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
-                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_agg,1))  # create plots 
+    mutate(censored  = map(data, censor, 600, 40)) %>%
+    mutate(ptb_data  = map(censored, ptb)) %>%          
+    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
+    mutate(condacc   = map(censored, calc_ca)) %>%      
+    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
+    # set haz_yaxis to .7, first_bin_shown to 1, aggregated_data to TRUE, and Nsubj to 6 
+    mutate(plot = map2(.x=lifetable_ca, .y=pid, plot_eha, .7, 3, T, 6)) 
+  # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
+  #            Recommended to set this to 1 when plotting the data of individual subjects.
+  # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
+  #.           you can set first_bin_shown to another value, such as 5 (i.e., do not show the 
+  #            first 4 bins in the plots).
+  # aggregated_data: For a dataset with data aggregated across individuals, set to TRUE. 
+  #            This only controls the title of the plots.
+  # Nsubj: The number of individuals before aggregation. Only used to set the title of the plot.
 ```

### 2024-11-24 Tutorial_1b.Rmd
update tutorials 1a and 1b

@@
 
-Content:
-1. Load libraries
-2. Read and wrangle data
-3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
-4. Extract info (life table, plots, ...)
-
 # 1. Load the libraries that we will be using.
@@
 
-# 2. Read and wrangle data
+Theme settings for ggplot.
+
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 14, face = "bold"), 
+          title = element_text(size = 14, face = "bold"))
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+# 2. Read and wrangle data.
 
@@
 ```{r read-raw-data}
-raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) # 9372 rows
+raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) 
 head(raw)
@@
 # remove practice trials
-data_wr <- raw %>% filter(practice==0) # 9240 experimental trials: 6 participants x 70 blocks x 22 trials-per-block
+data_wr <- raw %>% filter(practice==0)
 
@@
 # look at trials where resp==0
-data_wr %>% filter(resp==0) 
-data_wr %>% filter(resp==0) %>% print(n=51)
+data_wr %>% 
+  filter(resp==0) %>% 
+  print(n=51)
 
@@
 sel = which(data_wr$resp==0 & data_wr$rt < 800)
-data_wr <- data_wr[-c(sel),] # 9235 rows
+data_wr <- data_wr[-c(sel),]
 
@@
 colnames(data_wr) <- c("pid","bl","tr","condition1", "condition2","resp","acc","rt","trial")
+
 # pid to indicate column with unique person IDs
@@
 
-# create factors for conditions (mask type and prime type): the levels of condition1 are displayed across columns; the levels of condition2 are displayed within each plot.
-data_wr <- data_wr %>% mutate(condition1 = condition1 + 1, # avoid level "0"
-                              condition1 = factor(condition1, levels = c(1,2,3,4), labels = c("nomask","relevant","irrelevant","lines")),
-                              condition2 = condition2 + 1, # avoid level "0"
-                              condition2 = factor(condition2, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
+# create factors for conditions (mask type and prime type): 
+# the levels of condition1 are displayed across columns; 
+# the levels of condition2 are displayed within each plot.
+data_wr <- data_wr %>% 
+  mutate(condition1 = condition1 + 1, 
+         condition1 = factor(condition1, 
+                             levels = c(1,2,3,4), 
+                             labels = c("nomask","relevant","irrelevant","lines")),
+         condition2 = condition2 + 1, 
+         condition2 = factor(condition2, 
+                             levels = c(1,2,3), 
+                             labels = c("blank","congruent","incongruent")))
 summary(data_wr)
@@
 data_final <- data_nested %>% 
-                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
+  # ! user input: censoring time, and bin width
+  mutate(censored  = map(data, censor, 600, 40)) %>%   
   mutate(ptb_data  = map(censored, ptb)) %>%           
-                     mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%      # create life tables without ca for 2 independent variables
-                     mutate(condacc   = map(censored, calc_ca_2IV)) %>%       # calculate conditional accuracy for 2 independent variables
-                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  # add ca to lifetables for 2 independent variables
-                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV,1))  # create plots for 2 independent variables
+  # create life tables without ca for 2 independent variables
+  mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%    
+  # calculate conditional accuracy for 2 independent variables
+  mutate(condacc   = map(censored, calc_ca_2IV)) %>%       
+  # add ca to lifetables for 2 independent variables
+  mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  
+  # create plots for 2 independent variables: set haz_yaxis to 1, and first_bin_shown to 2
+  mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV,1,2))  
 ```
@@
 # view original data (column 2 in data_final) for subject 1:
-pluck(data_final, 2, 1) %>% print(n=20)
+pluck(data_final, 2, 1) %>% 
+  print(n=20)
 
 # view person-trial-bin data set (column 4) for subject 1:
-pluck(data_final,4,1) %>% print(n=30)
+pluck(data_final,4,1) %>% 
+  print(n=30)
 
 # view lifetable (column 7) for a specific combination of condition levels for subject 1:
-pluck(data_final, 7, 1) %>% filter(condition1=="nomask" & condition2 == "blank")
+pluck(data_final, 7, 1) %>% 
+  filter(condition1=="nomask" & condition2 == "blank")
 
@@
 # save the plot for each subject
-map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 10, dpi = 300)
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 10, dpi = 800)
 ```

### 2024-11-24 custom_functions.R
update custom_functions.R and .Rdata for tut 1a

@@ -1,5 +1,7 @@
-# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
+# function censor creates censored observations and discrete response times (drt), 
+# given a user-defined censoring time (timeout) and bin width in ms
 censor <- function(df, timeout, bin_width){
+
   if(!(timeout %% bin_width == 0)){
     return("The censoring time must be a multiple of the bin width!")
   }
@@ -7,10 +9,15 @@ censor <- function(df, timeout, bin_width){
     return("Both timeout and bin_width must be larger than 0!")
   }
   df %>% mutate(right_censored = 0,
-                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
-                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
-                drt = ceiling(rtc/bin_width), # drt = discrete RT or time bin rank
-                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
+                # censored response times
+                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2),
+                # 1 = right censored observation, 0 = observed rt
+                right_censored = ifelse(rtc == timeout,1,right_censored), 
+                # drt = discrete RT or time bin rank
+                drt = ceiling(rtc/bin_width), 
+                # save both user-defined parameters for plotting
+                cens_time = timeout, 
+                bin_width = bin_width) 
 }
 
 # function ptb creates a person-trial-bin oriented data set
@@ -18,7 +25,8 @@ ptb <- function(df){
   df %>% uncount(weights = drt) %>% # create drt rows per trial
     group_by(trial) %>% 
     mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
-    mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
+    mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% 
+    # event = 1 indicates response occurrence
     ungroup()
 }
 
@@ -30,18 +38,23 @@ setup_lt <- function(ptb){
     ungroup() %>% 
[…]

### 2024-11-24 Tutorial_2a.Rmd
update tut 2a

@@
 The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (prime = blank, congruent, or incongruent).
-We then compare models using WAIC and LOO, and interpret parameter estimates for one model.
-We also plot the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
+We perform prior predictive checks to infer the prior distributions we will employ.
+We compare models using WAIC and LOO, and interpret parameter estimates for one model.
+We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
 
@@
 ```{r data-priors-M0}
-M0_data <- ptb_data %>% select(pid, event, timebin)
-
-priors_M0 <- c(
-  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), # See suppl. material 
+data_M0i <- ptb_data %>% select(pid, event, timebin)
+
+priors_M0i <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b"), 
   set_prior("normal(0, 1)", class = "sd"),                    
@@
 
-Fit model M0.
-
-```{r fit-model-M0, eval=F} 
+Fit model M0i.
+
+```{r fit-model-M0i, eval=F} 
 plan(multicore)
 
-model_M0 <-                     # wrong data file name
-   brm(data = M0_data,
-       family = binomial(link="cloglog"),
-       event|trials(1) ~ 0 + timebin + (0 + timebin | pid),
-       prior = priors_M0,
+model_M0i <-                    
+   brm(data = data_M0i,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0i,
        chains = 4, cores = 4, iter = 3000, warmup = 1000,
@@
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M0")
-```
-
-This took 43 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
-
-```{r}
-model_M0 <- readRDS("Tutorial_2_Bayesian/models/model_M0.rds")
-summary(model_M0)
-formula(model_M0)
-plot(model_M0)
+       file = "Tutorial_2_Bayesian/models/model_M0i")
+```
+
+This took 28 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+
+```{r}
+model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
+summary(model_M0i)
+formula(model_M0i)
+plot(model_M0i)
 ```
@@
 ```{r data-priors-M0d}
-M0d_data <- ptb_data %>% select(pid, event, d6:d8, d10:d15)
+data_M0d <- ptb_data %>% select(pid, event, d6:d8, d10:d15)
 
@@
 
-model_M0d <-            # bernouilli.  :(
-   brm(data = M0d_data,
+model_M0d <-      
+   brm(data = data_M0d,
        family = bernoulli(link="cloglog"),
@@
 
-M0d took 77 minutes.
+M0d took about 77 minutes.
 
@@
 model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
-summary(model_M0d)
 formula(model_M0d)
+fixef(model_M0d)
 ```
@@
 ```{r data-priors-M0c}
-M0c_data <- ptb_data %>% select(pid, event, period_9) %>%
-  mutate(period_9_sq = period_9^2,
-         period_9_cu = period_9^3)
+data_M0c <- ptb_data %>% 
+  select(pid, event, period_9) %>%
+  mutate(period_9_sq = period_9^2)
 
 priors_M0c <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
   set_prior("normal(0, 1)", class = "sd"),  
@@
 model_M0c <-
-   brm(data = M0c_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq + period_9_cu + 
-                          (0 + Intercept + period_9 + period_9_sq + period_9_cu  | pid),
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq + 
+              (0 + Intercept + period_9 + period_9_sq | pid),
        prior = priors_M0c,
@@
 
-Model_M0c took about 144 minutes to run.
+Model_M0c took about 71 minutes to run.
 
@@
               cloglog1 = -1.8   + 0.3*x   - 0.1*x2   + 0.009*x3,
-              cloglog2 = -1     + 0.5*x   - 0.1*x2   + 0.004*x3,
+              cloglog2 = -1     + 0.4*x   - 0.1*x2   + 0.007*x3,
               cloglog3 = -3     - 0.1*x   + 0.12*x2  - 0.002*x3,
@@
 
-set_prior("normal(0.1,.08)",class="b",coef="period_9"), # ok !!!
+#set_prior("normal(0.1,.08)",class="b",coef="period_9"), # ok !!!
+#set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
+
+set_prior("normal(0.1,.1)",class="b",coef="period_9"), # ok !!!
 set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
+#set_prior("normal(0,.003)",class="b",coef="period_9_cu"),
 
@@
 
-model_M0_lin_prior448 <-
+model_M0_lin_prior454 <-
    brm(data = M0c_data,
@@
        sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0_lin_prior448")
-```
-
-```{r}
-post <- as_draws_df(model_M0_lin_prior448) %>% 
+       file = "Tutorial_2_Bayesian/models/model_M0_lin_prior454")
+```
+
+```{r}
+#model_M0_lin_prior448 <- readRDS("Tutorial_2_Bayesian/models/model_M0_lin_prior448.rds")
+
+post <- as_draws_df(model_M0_lin_prior454) %>% 
   select(-lp__) %>% 
@@
 # plot
-p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = name)) +  
-  stat_halfeye(alpha=0.7,point_interval = "median_qi",.width=c(.8,.99) ) 
+p_tidy_fixed <- ggplot(tidy_fixed, aes(x = name, y = value, fill = name)) +  
+  stat_halfeye(alpha=0.7, point_interval = "median_qi",.width=c(.8,.99) ) 
 p_tidy_fixed
@@
 
-N_obs = 150 # increase for smoother results
+N_obs = 100 # increase for smoother results
 
@@
                  period_9_sq = integer(0),
+                # period_9_ce = integer(0),
                  pid = integer(0),
@@
 for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0_lin_prior448, 
+  prior_pred <- add_predicted_draws(model_M0_lin_prior454, 
                                     newdata = newdata_prior, 
@@
 ggplot(df_pred_haz %>% 
-   filter(.draw <7639), aes(x=period_9, y=pred_haz, group=.draw)) +
+   filter(.draw <639), aes(x=period_9, y=pred_haz, group=.draw)) +
    geom_line(color="black") +
-   geom_line(data=df_pred_haz %>% filter(.draw ==7418), 
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
              color="red", linewidth=2) +
-  geom_line(data=df_pred_haz %>% filter(.draw == 6001), 
+  geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
             color="green", linewidth=2) +

### 2024-11-24 Tutorial_2a.Rmd
update tuts

@@
 Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data.
-The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (prime = blank, congruent, or incongruent).
-We perform prior predictive checks to infer the prior distributions we will employ.
+The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent).
+We perform some prior predictive checks to infer the prior distributions we will employ.
 We compare models using WAIC and LOO, and interpret parameter estimates for one model.
@@
 ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-print(ptb_data,n=30)
-summary(ptb_data) # 26602 rows: 6 participants, trial (experiment-wise), 3 conditions, 15 periods, and event indicator (0/1)
+head(ptb_data)
 ```
@@
          period_9 = period - 9,
-# create binary variables "dummys" to indicate each bin in the analysis
+# create binary variables to indicate each bin
          d6  = if_else(period == 6, 1, 0),
@@
 
-# create continuous predictor for trial number named "trial_c", centered on bin 1000 and rescaled
+# create continuous predictor for trial number named "trial_c", centered on bin 1000 and rescale
          trial_c = (trial - 1000)/1000,
@@
 head(ptb_data,n=17)
-summary(ptb_data)
 ```
@@
 
-The first model we fit is a "random intercepts" model, where we fit a single intercept for each timebin and allow each intercept to vary between participants.
-This is a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial). 
-There are two ways to implement such a model. First, we can use the index-coding approach, which supplies an intercept for each level of TIME.
-
-Prepare the data file, and specify priors.
-
-```{r data-priors-M0}
+The first model we fit is a "random intercepts" model, where we fit a single grand intercept for each timebin and add random intercepts that vary between participants.
+This constitutes a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial).
+
+There are two ways to implement such a model. First, we can use the index-coding approach, which provides an intercept for each level of TIME (variable timebin in ptb_data).
+
+Prepare the data file, and specify priors. The skew_normal prior is set for each grand intercept on the cloglog scale, and should reflect our prior beliefs.
+
+```{r data-priors-M0i}
 data_M0i <- ptb_data %>% select(pid, event, timebin)
@@
 
-Perform a prior predictive check.
+Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflects our prior beliefs.
+First, sample the prior distributions using sample_prior="only".
 
@@
 
-model_M0_prior <-                     # wrong data file name
-   brm(data = M0_data,
+model_M0i_prior <-                  
+   brm(data = data_M0i,
        family = bernoulli(link="cloglog"),
-       event ~ 0 + timebin + (0 + timebin | pid),
-       prior = priors_M0,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       formula = event ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0i,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
-       sample_prior = 'only',
-       file = "Tutorial_2_Bayesian/models/model_M0_prior")
-```
-
-```{r}
-
-#pp_check(model_M0_prior, stat="mean")
-
-#http://bruno.nicenboim.me/bayescogsci/
-
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0i_prior")
+```
+
+Next, use them to predict prior data. To better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.
+
+```{r ppc-manual}
 # prior predictive check for 1 time bin
-cloglog_predictive_distribution <-
-  function(samples, N_obs) {
-    # empty data frame with headers:
+# see http://bruno.nicenboim.me/bayescogsci/
+
+# cloglog and inverse-cloglog link functions
+inverse_cloglog <- function(x) {return(1-(exp(-1*exp(x))))}
+cloglog <- function(x) {return(log(-1*log(1-x)))}
+
+# Specify
+N_samples <- 1000 # the number of samples from the skew_normal
+N_obs <- 100      # the number of simulated observations per sample
+set.seed(1)
+
+# function to generate binary observations from samples
+cloglog_predictive_distribution <- function(samples, Nobs) {
+
+    # empty data frame with headers
     df_pred <- tibble(trial = numeric(0),
@@
                       iter = numeric(0))
-    # i iterates from 1 to the length of samples,
+    # i iterates from 1 to the length of samples
     for (i in seq_along(samples)) {
@@
           tibble(trial = 1:N_obs,
-                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
+                 y_pred = rbinom(Nobs, 1, inverse_cloglog(cloglog_haz)),
                  iter = i))
     }
-    df_pred # return
+    df_pred 
  }
-# simulate prior hazard values for 1 bin
-N_samples <- 3000
-N_obs <- 100 #nrow(M0_data)
-#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
-cloglog_samples <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept
-
+
+# sample prior cloglog-hazard values for 1 bin
+cloglog_samples <- rskew_normal(N_samples,-1,1,-2)
+
+# apply the function
 prior_pred <- cloglog_predictive_distribution(cloglog_samples,N_obs)
 
-prior_pred_haz <- prior_pred %>% group_by(iter) %>% summarise(pred_haz = mean(y_pred))
+# calculate hazard as the mean over binary observations
+prior_pred_haz <- prior_pred %>% 
+  group_by(iter) %>% 
+  summarise(pred_haz = mean(y_pred))
 
@@
   geom_histogram(binwidth=.01) +
-  scale_x_continuous(limits = c(0,1))
-```
-
-```{r}
-# Generate prior predictive samples for 10 bins 
-bins_cloglog_predictive_distribution <-
-  function(samples_bin, N_obs, N_bins) {
-    # empty data frame with headers:
-    df_pred <- tibble(trial = numeric(0),
-                      y_pred = numeric(0),
-                      iter = numeric(0),
-                      bin = numeric(0))
-    # i iterates from 1 to the length of samples,
-    # b interates from 1 to N_bins
-    for( b in 1:N_bins){
-      for (i in 1:dim(samples_bin)[2]) {
-        cloglog_haz <- samples_bin[b,i]
-        df_pred <- bind_rows(df_pred,
-                             tibble(trial = 1:N_obs,
-                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
-                                    iter = i,
-                                    bin = b))
-      }
-    }
-    df_pred # return
-  }
-
-# simulate prior hazard functions for 10 bin
-N_samples <- 40
-N_obs <- 100
-N_bins <- 10
-#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
-cloglog_samples_bin <- data.frame()
-for(b in 1:N_bins){
-  cloglog_samples_bin[b,1:N_samples] <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept
-}
-
-prior_pred <- bins_cloglog_predictive_distribution(cloglog_samples_bin,N_obs,N_bins)
-
-prior_pred_haz <- prior_pred %>% group_by(iter,bin) %>% summarise(pred_haz = mean(y_pred))
-
-# plot 
-ggplot(prior_pred_haz, aes(x=bin, y=pred_haz, group=iter)) +
-  #geom_histogram(binwidth=.01) +
-  geom_line(color="black") +
-  geom_line(data = prior_pred_haz %>% filter(iter==4), color="red", linewidth = 2) +
-  geom_line(data = prior_pred_haz %>% filter(iter==18), color="blue", linewidth = 2) +
-
-  scale_y_continuous(limits = c(0,1)) +
-  scale_x_continuous(limits = c(1,10), breaks = c(1:10))
-
-```
-
-```{r}
+  scale_x_continuous(limits = c(0,1)) +
+  labs(x = "simulated hazard",
+       title = "Prior predictive distribution",
+       subtitle = str_c(N_samples," samples; ",N_obs," observations per sample"))
+```
+
+The shape of the prior predictive distribution looks as expected (compare with Figure X in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
+
+Now use add_predicted_draws() to simulate prior-based data for 10 bins and 6 subjects, while take into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
+
+```{r ppc-M0i-prior}
 # Generate prior predictive hazard functions for 6 participants
-newdata_prior = tibble(timebin = c(6:15)) %>% 
+newdata_prior = tibble(timebin = 6:15) %>% 
            expand_grid(pid = 1:6)
+
+# Specify
 N_obs = 100
+set.seed(2)
 
@@
 
+# Call add_predicted_draws() for each simulated observation
 for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
-  prior_pred <- prior_pred %>% mutate(obs = i)
+  prior_pred <- add_predicted_draws(model_M0i_prior, 
+                                    newdata=newdata_prior, 
+                                    summary=F, 
+                                    ndraws=NULL) %>% 
+                mutate(obs = i)
   df_pred <- bind_rows(df_pred, prior_pred)
 }
-head(df_pred)
-
-# calculate hazard per draw (average across obs)
-df_pred_haz <- df_pred %>% group_by(pid,timebin,.draw) %>% summarise(pred_haz = mean(.prediction))
-summary(df_pred_haz)
+
+# calculate hazard per draw (average across observations)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,timebin,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
 
 # plot 40 draws per participant
-ggplot(df_pred_haz %>% filter(.draw < 21), aes(x=timebin, y=pred_haz, group=.draw)) +
-
+ggplot(data = df_pred_haz %>% filter(.draw < 41), 
+       aes(x=timebin, y=pred_haz, group=.draw)) +
    geom_line(color="black") +
-   geom_line(data=df_pred_haz %>% filter(.draw == 20), color="red", linewidth=2) +
-
+   geom_line(data=df_pred_haz %>% filter(.draw == 20), 
+             color="red", 
+             linewidth=2) +
   scale_y_continuous(limits = c(0,1)) +
   scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
+  labs(x = "time bin",
+       y = "simulated hazard",
+       title = "Prior predictive distributions",
+       subtitle = str_c("8000 samples; ",N_obs," observations per sample")) +
   facet_wrap(~pid)
-
-```
-
-Fit model M0i.
+```
+
+Now we can fit model M0i to get the posterior distributions.
 
@@
        prior = priors_M0i,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
@@
 
-```{r}
+```{r insppect-M0i}
 model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
@@
 ```{r data-priors-M0d}
+# select bin indicator variables
 data_M0d <- ptb_data %>% select(pid, event, d6:d8, d10:d15)
@@
        family = bernoulli(link="cloglog"),
-       event ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
+       formula = event ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
                    (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15  | pid),
        prior = priors_M0d,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
@@
 
-```{r}
+```{r inpect-M0d}
 model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
@@
 
-### Make assumptions to simplify the model
-
-Instead of using a general specification of TIME, we can treat TIME as a continuous variable and make assumptions about how hazard changes over time within a trial.
-For example, we might assume that hazard changes in a linear + quadratic fashion over time bins within a trial.
+Third, we can make assumptions to simplify the model.
+Instead of using a general specification of TIME (bin rank), we can treat TIME as a continuous variable and make assumptions about how cloglog-hazard changes over time within a trial.
+For example, we might assume that cloglog-hazard changes in a linear ("period_9") + quadratic ("period_9_sq") fashion over time bins within a trial.
 
@@
 
-Fit the model with a continuous predictor (M0c).
-
-```{r fit-model-M0c, eval=F}
+Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflect our prior beliefs. 
+The following code allows to change the coefficients and generate (cloglog-)hazard functions.
+Once you have found the coefficients for period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to estimate a mean and standard deviation for each coefficient.
+
+```{r search-coef}
+dat <- tibble(x = -3:6,
+              x2 = x^2,
+              cloglog1 = -1.8   + 0.3*x   - 0.1*x2,
+              cloglog2 = -1     + 0.4*x   - 0.1*x2,
+              cloglog3 = -3     - 0.1*x   + 0.12*x2,
+              cloglog4 = -0.40  + 0.2*x   - 0.04*x2,
+              haz1 = 1 - exp(-1*exp(cloglog1)),
+              haz2 = 1 - exp(-1*exp(cloglog2)),
+              haz3 = 1 - exp(-1*exp(cloglog3)),
+              haz4 = 1 - exp(-1*exp(cloglog4)))
+
+p1<-ggplot(dat, aes(x=x)) +  
+      geom_line(aes(y=cloglog1), color="black") +
+      geom_line(aes(y=cloglog2), color="green") +
+      geom_line(aes(y=cloglog3), color="red") +
+      geom_line(aes(y=cloglog4), color="blue") +
+      scale_y_continuous(limits=c(-6,2)) +
+      scale_x_continuous(breaks=c(-3:6)) +
+      labs(x = "timebin", y = "cloglog-hazard")
+
+p2<-ggplot(dat, aes(x=x))+
+      geom_line(aes(y=haz1), color="black") + 
+      geom_line(aes(y=haz2), color="green") +
+      geom_line(aes(y=haz3), color="red") +
+      geom_line(aes(y=haz4), color="blue") +
+      scale_y_continuous(limits=c(0,1)) +
+      scale_x_continuous(breaks=c(-3:6)) +
+      labs(x = "timebin", y = "hazard")
+
+p1|p2
+```
+
+First, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random variables while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
+
+```{r priors-M0c-ppc}
+priors_M0c_ppc <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .1)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior, eval=F}
 plan(multicore)
 
-model_M0c <-
+model_M0c_prior <-
    brm(data = data_M0c,
@@
                           (0 + Intercept + period_9 + period_9_sq | pid),
-       prior = priors_M0c,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M0c")
-```
-
-Model_M0c took about 71 minutes to run.
-
-```{r}
-model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
-summary(model_M0c)
-formula(model_M0c)
-```
-
-## Intermezzo: Prior predictive checks
-
-```{r fit-model-M0c, eval=F}
-plan(multicore)
-
-model_M0c_prior <-
-   brm(data = M0c_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq + period_9_cu + 
-                          (0 + Intercept + period_9 + period_9_sq + period_9_cu  | pid),
-       prior = priors_M0c,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       prior = priors_M0c_ppc,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
@@
 
-```{r}
+Check the prior distributions.
+
+```{r prior-distr}
+# extract prior draws
+post <- as_draws_df(model_M0c_prior) %>% 
+  select(-lp__) %>% 
+  as_tibble()
+
+tidy_prior <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter))
+
+# plot
+ggplot(tidy_prior, aes(x = name, y = value, fill = name)) +  
+  stat_halfeye(alpha=0.7, 
+               point_interval = "median_qi",
+               .width=c(.8,.99), 
+               show.legend = F) +
+  labs(x = "parameter",
+       title = "Samples from the prior distributions")
+```
+
+Generate prior predictive check for each participant.
+
+```{r ppc-model-M0c-prior}
 # Generate prior predictive hazard functions for 6 participants
+# Specify
+N_obs = 100 # increase for smoother results
+set.seed(1)
+
+# make new data
 newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2,
-         period_9_cu = period_9^3) %>%
+  mutate(period_9_sq = period_9^2) %>%
   expand_grid(pid = 1:6)
 
-N_obs = 100
-
-# prepare columns of data set
+# prepare columns of data set with predictions
 df_pred <- tibble(period_9 = integer(0),
-                  period_9_sq = double(0),
-                  period_9_cu = double(0),
+                  period_9_sq = integer(0),
                   pid = integer(0),
@@
 
+# extract predictions
 for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0c_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
-  prior_pred <- prior_pred %>% mutate(obs = i)
-  df_pred <- bind_rows(df_pred, prior_pred)
-}
-head(df_pred)
-
-# calculate hazard per draw (average across obs)
-df_pred_haz <- df_pred %>% group_by(pid,period_9,.draw) %>% summarise(pred_haz = mean(.prediction))
-summary(df_pred_haz)
-
-# plot 40 draws per participant
-ggplot(df_pred_haz %>% filter(.draw < 2), aes(x=period_9, y=pred_haz, group=.draw)) +
-
-   geom_line(color="black") +
-   geom_line(data=df_pred_haz %>% filter(.draw == 17), color="red", linewidth=2) +
-
-  scale_y_continuous(limits = c(0,1)) +
-  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
-  facet_wrap(~pid)
-```
-
-##
-
-```{r}
-dat <- tibble(x = -3:6,
-              x2 = x^2,
-              x3 = x^3,
-              cloglog1 = -1.8   + 0.3*x   - 0.1*x2   + 0.009*x3,
-              cloglog2 = -1     + 0.4*x   - 0.1*x2   + 0.007*x3,
-              cloglog3 = -3     - 0.1*x   + 0.12*x2  - 0.002*x3,
-              cloglog4 = -0.40  + 0.2*x   - 0.04*x2  + 0.003*x3,
-              haz1 = 1 - exp(-1*exp(cloglog1)),
-              haz2 = 1 - exp(-1*exp(cloglog2)),
-              haz3 = 1 - exp(-1*exp(cloglog3)),
-              haz4 = 1 - exp(-1*exp(cloglog4)))
-
-p1<-ggplot(dat, aes(x=x)) +  
-      geom_line(aes(y=cloglog1), color="black") +
-      geom_line(aes(y=cloglog2), color="green") +
-      geom_line(aes(y=cloglog3), color="red") +
-      geom_line(aes(y=cloglog4), color="blue") +
-      scale_y_continuous(limits=c(-6,2)) +
-      scale_x_continuous(breaks=c(-3:6))
-
-p2<-ggplot(dat, aes(x=x))+
-      geom_line(aes(y=haz1), color="black") + 
-      geom_line(aes(y=haz2), color="green") +
-      geom_line(aes(y=haz3), color="red") +
-      geom_line(aes(y=haz4), color="blue") +
-      scale_y_continuous(limits=c(0,1)) +
-      scale_x_continuous(breaks=c(-3:6))
-
-p1|p2
-```
-
-```{r}
-tibble(cloglog = rnorm(1000, .1, .1)) %>%
-  ggplot(aes(x=cloglog)) +
-  geom_density()
-```
-
-```{r data-priors-M0c}
-M0c_data <- ptb_data %>% select(pid, event, period_9) %>%
-  mutate(period_9_sq = period_9^2)
-
-priors_M0c_prior <- c(
-  #set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
- set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
-
-#set_prior("normal(0.5,.5)",class="b",coef="period_9"),
-#set_prior("normal(-.1,.2)",class="b",coef="period_9_sq"),   # Np3
-
-#set_prior("normal(0.1,.08)",class="b",coef="period_9"), # ok !!!
-#set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
-
-set_prior("normal(0.1,.1)",class="b",coef="period_9"), # ok !!!
-set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
-#set_prior("normal(0,.003)",class="b",coef="period_9_cu"),
-
-#set_prior("skew_normal(0.3,0.1,2)",class="b",coef="period_9"), #skNp3
-#set_prior("skew_normal(0,0.1,0)",class="b",coef="period_9_sq"),
-
- set_prior("normal(0, .01)", class = "sd"),  
-  set_prior("lkj(2)", class = "cor")  
-)
-```
-
-```{r fit-model-M0c, eval=F}
-plan(multicore)
-
-model_M0_lin_prior454 <-
-   brm(data = M0c_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq  +
-                          (0 + Intercept + period_9 + period_9_sq  | pid),
-       prior = priors_M0c_prior,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0_lin_prior454")
-```
-
-```{r}
-#model_M0_lin_prior448 <- readRDS("Tutorial_2_Bayesian/models/model_M0_lin_prior448.rds")
-
-post <- as_draws_df(model_M0_lin_prior454) %>% 
-  select(-lp__) %>% 
-  as_tibble()
-
-tidy_fixed <- post %>% 
-  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
-  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
-  pivot_longer(-c(chain, draw, iter)) 
-head(tidy_fixed)
-tail(tidy_fixed)
-
-# plot
-p_tidy_fixed <- ggplot(tidy_fixed, aes(x = name, y = value, fill = name)) +  
-  stat_halfeye(alpha=0.7, point_interval = "median_qi",.width=c(.8,.99) ) 
-p_tidy_fixed
-
-```
-
-```{r}
-#model_M0_lin_prior14 <-readRDS("Tutorial_2_Bayesian/models/model_M0c_lin_prior14.rds")               
-# Generate prior predictive hazard functions for 6 participants
-newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2) %>%
-  expand_grid(pid = 1:6)
-
-N_obs = 100 # increase for smoother results
-
-# prepare columns of data set
-df_pred <- tibble(period_9 = integer(0),
-                 period_9_sq = integer(0),
-                # period_9_ce = integer(0),
-                 pid = integer(0),
-                  .row = integer(0),
-                  .chain = integer(0),
-                  .iteration = integer(0),
-                  .draw = integer(0),
-                  .prediction = integer(0),
-                  obs = integer(0))
-
-set.seed(1)
-
-for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0_lin_prior454, 
+  prior_pred <- add_predicted_draws(model_M0c_prior, 
                                     newdata = newdata_prior, 
@@
 
-head(df_pred)
-
 # calculate hazard per draw (average across obs)
@@
   summarise(pred_haz = mean(.prediction))
-summary(df_pred_haz)
-
-# plot 40 draws per participant
-sel = sample(1:8000,3,replace=F)
+
+# plot 200 draws per participant, highlighting three
+sel = sample(1:200,3,replace=F)
 
 ggplot(df_pred_haz %>% 
-   filter(.draw <639), aes(x=period_9, y=pred_haz, group=.draw)) +
+   filter(.draw <200), aes(x=period_9, y=pred_haz, group=.draw)) +
    geom_line(color="black") +
+   # highlight
    geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
@@
 
+A wide range of differently shaped hazard functions can be observed, and early bins are unlikely to have high hazard values, consistent with our prior beliefs.
+Note that some simulated hazard functions start off with a high hazard in bin -3.
+These reflect a combination of negative slopes and a positive coefficient for period_9_sq.
+These possibilities are required to encompass hazard functions that initially decrease and then increase, as observed in the empirical hazard functions for incongruent primes.
+
+Now increase the variance again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.
+
+```{r fit-model-M0c, eval=F}
+plan(multicore)
+
+model_M0c <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq + 
+              (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       file = "Tutorial_2_Bayesian/models/model_M0c")
+```
+
+Model_M0c took about 71 minutes to run.
+
+```{r inspect-M0c}
+model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
+summary(model_M0c)
+formula(model_M0c)
+```
+
 ## Model M1. Adding our experimental manipulation.
 
-The next model we fit also includes our predictor variable prime type.
-First, we can use index coding for the categorical predictor.
+The next models we fit also include our predictor variable prime type.
+First, we can use index coding for this categorical predictor.
 
@@
 ```{r data-priors-M1}
-M1_data <- ptb_data %>% select(pid, event, timebin, prime)
-
-priors_M1 <- c(
-  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), 
+data_M1i <- ptb_data %>% select(pid, event, timebin, prime)
+
+priors_M1i <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b"), 
   set_prior("normal(0, 1)", class = "sd"),                    
@@
 
-Fit model M1.
-
-```{r fit-model-M1, eval=F}
+Fit model M1i. The interaction between timebin and prime will specify 30 grand intercepts.
+
+```{r fit-model-M1i, eval=F}
 plan(multicore)
 
-model_M1 <-
-   brm(data = M1_data,
-       family = binomial(link="cloglog"),
-       event|trials(1) ~ 0 + timebin:prime  +
+model_M1i <-
+   brm(data = data_M1i,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + timebin:prime  +
                         (0 + timebin:prime | pid),
-       prior = priors_M1,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       prior = priors_M1i,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M1")
-```
-
-Model M1 took about 126 minutes to fit.
-
-```{r}
-model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
-summary(model_M1)
-formula(model_M1)
+       file = "Tutorial_2_Bayesian/models/model_M1i")
+```
+
+Model M1i took about 124 minutes to fit.
+
+```{r inspect-M1i}
+model_M1i <- readRDS("Tutorial_2_Bayesian/models/model_M1i.rds")
+fixef(model_M1i)
+formula(model_M1i)
 ```
@@
 Second, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach and treat TIME as a continuous variable.
-For example, we may assume that hazard can change in a linear + quadratic + cubic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic + cubic fashion, and fit the following model (M1d).
+For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the following model (M1d).
 
 ```{r data-priors-M1d}
-M1d_data <- ptb_data %>% select(pid, event, period_9, condition) %>%
+data_M1d <- ptb_data %>% select(pid, event, period_9, condition) %>%
+  mutate(period_9_sq = period_9^2)
+
+# check which priors to set for model_M1d
+check_priors_M1d <- get_prior(formula = event ~ 0 + Intercept + 
+                                        condition*period_9 + 
+                                        condition*period_9_sq + 
+                                        (1 + condition*period_9 +
+                                        condition*period_9_sq | pid),
+                              data = data_M1d,
+                              familiy = bernoulli(link="cloglog"))
 
 priors_M1d <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"), 
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"), 
+  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef = "period_9_sq"),  
+  set_prior("normal(0,.4)", class = "b", coef = "conditioncongruent"),
+  set_prior("normal(0,.4)", class = "b", coef = "conditionincongruent"),
   set_prior("normal(0, 1)", class = "sd"),            
@@
 model_M1d <- 
-   brm(data = M1d_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + 
+   brm(data = data_M1d,
+       family = bernoulli(link="cloglog"),
+       event ~ 0 + Intercept + 
                    condition*period_9 + 
                    condition*period_9_sq + 
-                           condition*period_9_cu +
               (1 + condition*period_9 +
-                           condition*period_9_sq +
-                           condition*period_9_cu | pid),
+                   condition*period_9_sq | pid),
        prior = priors_M1d,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
@@
 
-Model_M1d took about 268 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Model_M1d took about 145 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
 
@@
 
-One could also fit a model with prime and period_9.
-However, to include interactions between a index coded categorical variable and a continuous variable in brm(), one has to switch to the non-linear syntax, as illustrated in the following model formula.
-
-```{r}
-bf(event|trials(1) ~ 0 + a + b * period_9 + c * period_9_sq,
+One could also fit a model with the variables prime and period_9.
+However, to include interactions between an index-coded categorical variable and a continuous variable in brm(), one has to switch to the non-linear syntax, as illustrated in the following model formula.
+
+```{r}
+formula = bf(event ~ 0 + a + b * period_9 + c * period_9_sq,
                  a ~ 0 + prime + (0 + prime |i| pid),
@@
 
-The priors can be set as follows.
+The priors could be set as follows.
 
 ```{r priors-nonlinearsyntax}
+# check which priors to set
+get_prior(formula=bf(event ~ 0 + a + b * period_9 + c * period_9_sq,
+                 a ~ 0 + prime + (0 + prime |i| pid),
+                 b ~ 0 + prime + (0 + prime |i| pid),
+                 c ~ 0 + prime + (0 + prime |i| pid),
+                 nl = TRUE),
+          data = ptb_data %>% 
+                   select(pid, event, period_9, prime) %>%
+                   mutate(period_9_sq = period_9^2),
+          family = bernoulli(link="cloglog"))
+
+# set priors
 priors <- c(
-  prior(skew_normal(-0.59,1.26,-4.22), class = b, nlpar = a), 
-  prior(skew_normal(-0.2,.71,-2.2), class = b, nlpar = b), 
-  prior(skew_normal(-0.2,.71,-2.2), class = b, nlpar = c), 
+  prior(skew_normal(-1,1,-2), class = b, nlpar = a), 
+  prior(normal(0.1,.2), class = b, nlpar = b), 
+  prior(normal(0,.7), class = b, nlpar = c), 
   prior(exponential(1), class = sd, group = pid, nlpar = a),
@@
 ```{r}
-event|trials(1) ~ 0 + timebin:prime:stage  + (0 + timebin:prime:stage | pid)
+event ~ 0 + timebin:prime:stage  + (0 + timebin:prime:stage | pid)
 ```
@@
 ```{r}
-bf(event|trials(1) ~ 0 + a + b * trial_c + c * trial_c_sq,
+bf(event ~ 0 + a + b * trial_c + c * trial_c_sq,
        a ~ 0 + timebin:prime + (0 + timebin:prime |i| pid),
@@
 
-Third, one can use dummy coding for prime type (variable "condition" in ptb_data), and treat TIME and trial number as continous variables.
-
-```{r}
- bf(event | trials(1) ~ 0 + Intercept + 
+Third, one can use dummy coding for prime type (variable "condition" in ptb_data), and treat TIME ("period_9") and trial number ("trial_c") as continous variables.
+
+```{r}
+ bf(event ~ 0 + Intercept + 
             condition*period_9*trial_c + 
@@
                   condition*period_9_sq*trial_c +
-                                condition*period_9_sq*trial_c_sq  | pid)
-```
-
-# Compare models using loo and waic.
+                  condition*period_9_sq*trial_c_sq|pid)
+```
+
+# 4. Compare models using loo and waic.
 
@@
 ```{r compare-models}
-model_M0  <- readRDS("Tutorial_2_Bayesian/models/model_M0.rds")
+model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
 model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
 model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
-model_M1  <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
+model_M1i <- readRDS("Tutorial_2_Bayesian/models/model_M1i.rds")
 model_M1d <- readRDS("Tutorial_2_Bayesian/models/model_M1d.rds")
@@
 ```{r add-criterion, eval=F}
-model_M0  <- add_criterion(model_M0, c("loo", "waic"))
+model_M0i <- add_criterion(model_M0i, c("loo", "waic"))
 model_M0d <- add_criterion(model_M0d, c("loo", "waic"))
 model_M0c <- add_criterion(model_M0c, c("loo", "waic"))
-model_M1  <- add_criterion(model_M1, c("loo", "waic"))
+model_M1i <- add_criterion(model_M1i, c("loo", "waic"))
 model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
@@
 ```{r loo-compare}
-loo_compare(model_M0, model_M0d, model_M0c, model_M1, model_M1d, criterion = "loo") %>% print(simplify = F)
-loo_compare(model_M0, model_M0d, model_M0c, model_M1, model_M1d, criterion = "waic") %>% print(simplify = F)
+loo_compare(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, criterion = "loo") %>% print(simplify = F)
+loo_compare(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, criterion = "waic") %>% print(simplify = F)
 ```
@@
 ```{r model-weights}
-model_weights(model_M0, model_M0d, model_M0c, model_M1, model_M1d, weights = "loo") %>% round(digits = 3)
-
-model_weights(model_M0, model_M0d, model_M0c, model_M1, model_M1d, weights = "waic") %>% round(digits = 3)
-```
-
-# Model inference for model M1.
+model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "loo") %>% round(digits = 3)
+
+model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "waic") %>% round(digits = 3)
+```
+
+# 5. Model inference for model M1i.
 
@@
 ```{r pareto}
-loo(model_M1)$diagnostics %>% 
+loo(model_M1i)$diagnostics %>% 
   data.frame() %>% 
   # attach the `id` values
-  bind_cols(M1_data) %>% 
+  bind_cols(data_M0i) %>% 
   mutate(id = 1:n()) %>%
@@
 ```{r}
-post <- as_draws_df(model_M1) %>% 
+post <- as_draws_df(model_M1i) %>% 
   select(-lp__) %>% 
@@
 
-post_summary <- posterior_summary(model_M1, robust = TRUE)
+post_summary <- posterior_summary(model_M1i, robust = TRUE)
 post_summary[1:30,]
@@
   stat_halfeye(alpha=0.7) +
-  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1',
+  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1i',
        x = "time bin", y = "cloglog-hazard") +
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_postdistr.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_postdistr.png", width = 10, height = 8, dpi = 800)
 ```
@@
 ```{r}
-dat_M1 <- as_tibble(model_M1$data)
-
-epreds_grand <- dat_M1 %>% 
+dat_M1i <- as_tibble(model_M1i$data)
+
+epreds_grand <- dat_M1i %>% 
   data_grid(timebin, prime) %>% 
-  add_epred_draws(model_M1, 
+  add_epred_draws(model_M1i, 
                   re_formula = NA) %>% 
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_pred_grand.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_pred_grand.png", width = 10, height = 8, dpi = 800)
 ```
@@
 ```{r}
-epreds_pid <- dat_M1 %>% 
+epreds_pid <- dat_M1i %>% 
   data_grid(pid,timebin, prime) %>% 
-  add_epred_draws(model_M1, 
+  add_epred_draws(model_M1i, 
                   re_formula = NULL) %>% 
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_pred_pid.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid.png", width = 10, height = 8, dpi = 800)
 ```
@@
 epreds_pid_new <- 
-  posterior_epred(model_M1, 
+  posterior_epred(model_M1i, 
                   newdata = expand_grid(timebin=6:15,
@@
                   sample_new_levels = "gaussian") %>% 
-  as_tibble %>%
+  as_tibble() %>%
   mutate(draw = 1:n()) %>%
@@
 ```{r}
-epreds_pid_new <- dat_M1 %>% 
+epreds_pid_new <- dat_M1i %>% 
   data_grid(timebin, prime) %>%
   mutate(pid = 7) %>%
-  add_epred_draws(model_M1, 
+  add_epred_draws(model_M1i, 
                   re_formula = NULL, 
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_pred_pid_new.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
 ```
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_grand_AME.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_grand_AME.png", width = 10, height = 8, dpi = 800)
 ```
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_pid_AMEs.png", width = 10, height = 8, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_pid_AMEs.png", width = 10, height = 8, dpi = 800)
 ```
@@
 ```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1_pid_new_AMEs.png", width = 10, height = 8, dpi = 800)
-```
-
-## Example conclusions for model M1.
+ggsave("Tutorial_2_Bayesian/figures/M1i_pid_new_AMEs.png", width = 10, height = 8, dpi = 800)
+```
+
+## Example conclusions for model M1i.
 
@@
 
-Thus the probability that the contrast "congruent minus blank" is larger than 0, is larger than .80 in bins 6 to 9.
-And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .80 in bins 8 to 13.
+Thus the probability that the contrast "congruent minus blank" is larger than 0, is larger than .80 in bins 6 to 8.
+And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .80 in bins 8 to 12.
 
@@
 While most participants show the grand AMEs, three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] - subject 2 also in bin (320,360].
-This dip is also present in the AMEs for a brand new hypothetical participant.
 
@@
 
-#####################
-
-### Prepare data for M1
-
-```{r data-M1}
-# remove unnecessary columns before fitting a model
-M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9,d9)) # 12840 obs. 
-head(M1_data)
-summary(M1_data)
-```
-
-### Set up priors for M1
-
-```{r priors-cloglog-M1}
-priors_M1 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"),# weakly informative prior for alpha parameters when using cloglog link 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"),
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),
-  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
-)
-```
-
-### Fit model M1.
-
-```{r fit-model-M1, eval=F}
-plan(multicore)
-
-model_M1 <-
-   brm(data = M1_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
-         condition + trial_c +
-
-                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
-       prior = priors_M1,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M1")
-```
-
-Model_M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
-
-```{r check-model-M1, eval=F}
-model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
-summary(model_M1)
-fixef(model_M1)
-```
-
-## Model M2: third-order polynomical specification of TIME in the baseline condition (blank prime), and main effects of prime types, and trial number.
-
-### Prepare data for M2.
-
-```{r data-M2}
-# remove unnecessary columns before fitting a model
-M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
-head(M2_data)
-```
-
-### Set up priors for M2.
-
-```{r priors-cloglog-M2}
-priors_M2 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
-  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
-)
-```
-
-### Fit model M2.
-
-```{r fit-model-M2, eval=F}
-plan(multicore)
-
-model_M2 <-
-   brm(data = M2_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c +
-                          (1 + period_9 + I(period_9^2) + I(period_9^3) + 
-                          condition + trial_c | pid),
-       prior = priors_M2,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M2")
-```
-
-Model_M2 took about 144 minutes to run.
-
-```{r check-model-M2, eval=F}
-model_M2 <- readRDS("Tutorial_2_Bayesian/models/model_M2.rds")
-summary(model_M2)
-fixef(model_M2)
-```
-
-## Model M3: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax proportionality assumption for prime types, and trial number.
-
-### Prepare data for M3.
-
-```{r data-M3}
-# remove unnecessary columns before fitting a model
-M3_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
-head(M3_data)
-```
-
-### Set up priors for M3.
-
-```{r priors-cloglog-M3}
-priors_M3 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using cloglog link
-  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
-)
-```
-
-### Fit model M3.
-
-```{r fit-model-M3, eval=F}
-plan(multicore)
-
-model_M3 <- 
-   brm(data = M3_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
-                           condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
-       prior = priors_M3,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M3")
-```
-
-Model_M3 took about 268 minutes to run.
-
-```{r check-model-M3, eval=F}
-model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
-
-summary(model_M3)
-```
-
-## Model M4: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.
-
-### Prepare data for M4.
-
-```{r data-M4}
-# remove unnecessary columns before fitting a model
-M4_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
-head(M4_data)
-```
-
-### Set up priors for M4.
-
-```{r priors-cloglog-M4}
-priors_M4 <- c(
-  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using cloglog link
-  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
-)
-```
-
-### Fit model M4.
-
-```{r fit-model-M4, eval=F}
-plan(multicore)
-
-model_M4 <- 
-   brm(data = M4_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
-                          condition*period_9*trial_c + 
-                          condition*period_9*I(trial_c^2) + 
-                          condition*I(period_9^2)*trial_c +
-                          condition*I(period_9^2)*I(trial_c^2) +
-                          condition*I(period_9^3) +
-                          trial_c*I(period_9^3) +
-                          (1 +  condition*period_9*trial_c +
-                                condition*period_9*I(trial_c^2) + 
-                                condition*I(period_9^2)*trial_c +
-                                condition*I(period_9^2)*I(trial_c^2)  +
-                                condition*I(period_9^3) +
-                                trial_c*I(period_9^3) | pid),
-       prior = priors_M4,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
-       seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M4")
-```
-
-Model_M4 took about 8 hours to run.
-
-```{r check-model-M4, eval=F}
-model_M4 <- readRDS("Tutorial_2_Bayesian/models/model_M4.rds")
-summary(model_M4)
-```
-
-# 6. Compare models using loo and waic.
-
-```{r compare-models}
-model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
-model_M2 <- readRDS("Tutorial_2_Bayesian/models/model_M2.rds")
-model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
-model_M4 <- readRDS("Tutorial_2_Bayesian/models/model_M4.rds")
-```
-
-Using WAIC and LOO for comparing nonnested models.
-
-```{r add-criterion, eval=F}
-model_M1  <- add_criterion(model_M1, c("loo", "waic"))
-model_M2  <- add_criterion(model_M2, c("loo", "waic"))
-model_M3  <- add_criterion(model_M3, c("loo", "waic"))
-model_M4  <- add_criterion(model_M4, c("loo", "waic"))
-```
-
-Compare all three models.
-
-```{r loo-compare}
-loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "loo") %>% print(simplify = F)
-loo_compare(model_M1, model_M2, model_M3, model_M4, criterion = "waic") %>% print(simplify = F)
-```
-
-```{r model-weights}
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
-
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
-```
-
-# 7. Display effects for the selected model (model_M4).
-
-## Pareto k estimates.
-
-```{r pareto}
-loo(model_M4)$diagnostics %>% 
-  data.frame() %>% 
-  # attach the `id` values
-  bind_cols(M4_data) %>% 
-  mutate(id = 1:n()) %>%
-
-  ggplot(aes(x = id, y = pareto_k)) +
-  geom_point(alpha = 3/4) + 
-  geom_text(data = . %>% filter(pareto_k > .2),
-            aes(x = id + 2, label = id),
-            size = 3, hjust = 0) +
-  theme(panel.grid = element_blank())
-```
-
-## Visualize posterior distributions of the effects of congruent and incongruent primes on cloglog-hazard relative to blank prime, for each time bin in trials 500, 1000, and 1500.
-
-```{r post-distr}
-#get_variables(model_M4)[1:31]
-
-post <-as_draws_df(model_M4) %>% # 8000 draws x 715 variables
-   select(starts_with("b_")) %>%       # 8000 x 31
-   expand_grid(period_9 = -3:6) %>%   
-   mutate(period_9sq = period_9^2,
-          period_9cu = period_9^3,
-          trial500 = -500/1000,
-          trial500sq = trial500^2,
-          trial1500 = 500/1000,
-          trial1500sq = trial1500^2)
-
-# effects for trials 500, 1000, and 1500
-effects <- post %>% 
-  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
-            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
-            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
-            period_9sq * trial1500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
-            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
-            period_9sq * trial1500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,
-
-         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
-            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
-            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
-            period_9sq * trial1500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
-            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
-            period_9sq * trial1500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,
-
-         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
-            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
-            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
-            period_9sq * trial500 * `b_conditioncongruent:Iperiod_9E2:trial_c` +
-            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
-            period_9sq * trial500sq * `b_conditioncongruent:Iperiod_9E2:Itrial_cE2`,
-
-         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
-            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
-            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
-            period_9sq * trial500 * `b_conditionincongruent:Iperiod_9E2:trial_c` +
-            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
-            period_9sq * trial500sq * `b_conditionincongruent:Iperiod_9E2:Itrial_cE2`,
-
-         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
-
-         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
-```
-
-```{r plot-effects}
-plot_facet <- effects %>%
-  pivot_longer(cols=`congruent trial 1500`:`incongruent trial 1000`, names_to = "condition" ) %>%
-  select(condition, period_9, value) %>%
-  group_by(period_9) %>%
-
-  ggplot(aes(x = period_9, y = value)) +
-  stat_lineribbon(point_interval = mean_hdi,.width = c(0.5, 0.8, 0.95),show.legend=T) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-12,-10,-8,-6,-4,-2,0,2,4,6,8,10,12), limits = c(-12,12)) +
-  labs(x = "time bin endpoint", y = "cloglog-hazard") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90),
-        panel.background = element_rect(fill="white", color="black"),
-        legend.position = "top") +
-  facet_wrap(~ factor(condition,levels=c("congruent trial 500","congruent trial 1000","congruent trial 1500","incongruent trial 500","incongruent trial 1000","incongruent trial 1500")))
-
-plot_facet
-```
-
-```{r save-plot-effects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
-```
-
-## calculate point and interval estimates, and hazard ratios.
-
-```{r table-interval-estimates}
-int_c_500 <- effects %>%
-  select(`congruent trial 500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`congruent trial 500`),
-         mean = `congruent trial 500`,
-         condition = "c500") %>%
-  select(bin_endpoint,condition, mean, .lower, .upper, .width, `hazard ratio`)
-
-int_i_500 <- effects %>%
-  select(`incongruent trial 500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`incongruent trial 500`),
-         mean = `incongruent trial 500`,
-         condition = "i500") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width, `hazard ratio`)
-
-int_c_1000 <- effects %>%
-  select(`congruent trial 1000`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`congruent trial 1000`),
-         mean = `congruent trial 1000`,
-         condition = "c1000" ) %>%
-  select(bin_endpoint,condition,mean, .lower, .upper, .width, `hazard ratio`)
-
-int_i_1000 <- effects %>%
-  select(`incongruent trial 1000`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`incongruent trial 1000`),
-         mean = `incongruent trial 1000`,
-         condition = "i1000") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width,`hazard ratio`)
-
-int_c_1500 <- effects %>%
-  select(`congruent trial 1500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`congruent trial 1500`),
-         mean = `congruent trial 1500`,
-         condition = "c1500" ) %>%
-  select(bin_endpoint, condition, mean, .lower, .upper, .width, `hazard ratio`)
-
-int_i_1500 <- effects %>%
-  select(`incongruent trial 1500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `hazard ratio` = exp(`incongruent trial 1500`),
-         mean = `incongruent trial 1500`,
-         condition = "i1500") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width,  `hazard ratio`)
-
-test <- rbind(int_c_500, int_c_1000, int_c_1500,int_i_500,int_i_1000, int_i_1500)
-test %>% print(n=60)
-```
-
-```{r save-table-interval-estimates, eval=F}
-write_csv(test, file="Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
-```
-
-## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject.
-
-```{r select-pid}
-selected_subject = 6 # select pid here (1:6) manually
-```
-
-```{r help-function}
-make_fitted <- function(fit, scale, ...) {
-
-  fitted(fit,
-         newdata = nd,
-         scale = scale,
-         ...) %>% 
-    data.frame() %>% 
-    bind_cols(nd)
-}
-```
-
-### Set up new data.
-
-```{r new-data}
-# define the `newdata`
-nd <- tibble(pid = 1:6) %>%
-  expand_grid(incon = 0:1,
-                con  = 0:1,
-                trial_c = c(-0.5,0,0.5),
-                period_9 = -3:6) %>%
-  filter(!(con ==1 & incon == 1)) %>%
-  mutate(Iperiod_9E2 = period_9^2,
-         Iperiod_9E3 = period_9^3,
-         Itrial_cE2 = trial_c^2,
-         condition = ifelse(con==1 & incon==0,"congruent",
-                     ifelse(con==0 & incon==1,"incongruent", "blank")),
-         condition = factor(condition, levels=c("blank" ,"congruent","incongruent"))) 
-```
-
-### First, on the cloglog-scale.
-
-```{r subject-cloglog}
-make_fitted(model_M4, scale = "linear") %>% 
-  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
-         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
-         period = period_9+9) %>% 
-  filter(pid==selected_subject) %>% 
-  # plot
-  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
-             fill = condition, color = condition)) +
-  geom_ribbon(alpha = 1/5, linewidth = 0) +
-  geom_line() +
-  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
-  ylab("fitted cloglog(hazard)") +
-  coord_cartesian(ylim = c(-5, 1)) +
-  theme(panel.grid = element_blank(),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~ trial)
-```
-
-### Second, on hazard scale.
-
-```{r subject-hazard}
-make_fitted(model_M4, scale = "response") %>% 
-  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
-         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
-         period = period_9+9) %>% 
-  filter(pid==selected_subject) %>% 
-  # plot
-  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
-             fill = condition, color = condition)) +
-  geom_ribbon(alpha = 1/5, linewidth = 0) +
-  geom_line() +
-  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
-  ylab("fitted hazard") +
-  coord_cartesian(ylim = c(0, 1)) +
-  theme(panel.grid = element_blank(),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~ trial)
-```
-
-```{r save-plot-fit-hazard, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4effects_subject6.png", width = 10, height = 10, dpi = 300)
-```
-
-# 3. Plot the logit and complementary log-log (cloglog) link functions.
+# 6. Plot the logit and complementary log-log (cloglog) link functions.
 
@@
 ```{r save-plot-links, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 300)
-```
-
-# 4. Visualize different prior distributions on the logit and cloglog scales.
+ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 8, height = 8, dpi = 800)
+```
+
+# 7. Visualize different prior distributions on the logit and cloglog scales.
 
@@
   annotate(geom="text", x=-13, y=60000, label="N(0,4)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
   annotate(geom="text", x=-6, y=60000, label="N(0,2)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
   annotate(geom="text", x=-12, y=140000, label="t(7.61,0,1.57)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
   annotate(geom="text", x=-3, y=60000, label="N(0,1)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'D', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
   scale_y_continuous(NULL, breaks = NULL) +
-  geom_vline(xintercept=logistic(0), color="red") +
+  #geom_vline(xintercept=logistic(0), color="red") +
   theme(panel.grid = element_blank())
@@
   annotate(geom="text", x=-5.5, y=55000, label="skew_N(-0.59,1.26,-4.22)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'E', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
   annotate(geom="text", x=-5, y=55000, label="skew_N(-1,1,-2)",
-              color="red", size=6) +
+              color="red", size=4) +
   annotate(geom = 'text', label = 'F', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
@@
 
-# A skew_Normal(-0.2,0.71,-2.2) on the cloglog scale gently regularizes p towards .6321. This might be a good prior to use for the beta parameters in a cloglog-hazard model.
-pr7 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>%  
-  ggplot(aes(x = prior)) +
-  geom_histogram(bins = 50) +
-  scale_y_continuous(NULL, breaks = NULL) +
-  annotate(geom="text", x=-3.2, y=55000, label="skew_N(-0.2,.71,-2.2)",
-              color="red", size=6) +
-  annotate(geom = 'text', label = 'G', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
-  theme(panel.grid = element_blank())
-
-l7 <- ggplot()
-
-c7 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>% 
-  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
-  ggplot(aes(x = p)) +
-  geom_histogram(bins = 50) +
-  geom_vline(xintercept=inverse_cloglog(0), color = "red") +
-  scale_y_continuous(NULL, breaks = NULL) +
-  theme(panel.grid = element_blank())
-
-(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) / (l7 + pr7 + c7)
+((l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) ) &  
+  theme(text = element_text(size = 10, face = "bold"), 
+        title = element_text(size = 10, face = "bold"))
 ```
@@
 ```{r save-plot-priors, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
-```
+ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=800)
+```

### 2024-11-24 Tutorial_2a.Rmd
update tuts 2a and 2b

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, results='hide'}
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-```{r}
+```{r M0i-prior, eval=F}
 plan(multicore)
@@
        file = "Tutorial_2_Bayesian/models/model_M0i_prior")
+```
+
+```{r inspect-M0i-prior}
+model_M0i_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0i_prior.rds")
 ```
@@
 
-```{r ppc-M0i-prior}
+```{r ppc-M0i-prior, eval=F}
 # Generate prior predictive hazard functions for 6 participants
@@
 
-```{r insppect-M0i}
+```{r inspect-M0i}
 model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
@@
 
+```{r inspect-M0c-prior}
+model_M0c_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior.rds")
+```
+
 Check the prior distributions.
@@
 
-```{r ppc-model-M0c-prior}
+```{r ppc-model-M0c-prior, eval=F}
 # Generate prior predictive hazard functions for 6 participants
@@
 
-```{r}
+```{r post-M0i}
 post <- as_draws_df(model_M1i) %>% 
@@
 
-```{r}
+```{r plot-post-M0i}
 tidy_fixed <- post %>% 
@@
 
-```{r save-post, eval=F}
+```{r save-pred-grand, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_grand.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r save-post, eval=F}
+```{r save-pred-pid, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r save-post, eval=F}
+```{r save-pred-pid-new, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r save-post, eval=F}
+```{r save-grand-AMEs, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_grand_AME.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r save-post, eval=F}
+```{r save-pid-AMEs, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pid_AMEs.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r save-post, eval=F}
+```{r save-pid-new-AMEs, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pid_new_AMEs.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r}
+```{r props-contrasts}
 pabove <- epreds_grand_diffs %>% 

### 2024-11-24 Tutorial_2b.Rmd
update html and Rmd files
update tuts 2a and 2b

@@
 date: "2024-10-18"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-Content:
-1. Load libraries.
-2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
-3. Fit a conditional accuracy model.
-4. Display effects for the conditional accuracy model.
-
 # 1. Load the libraries that we will be using.
@@
 
-# 2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
+theme settings for ggplot
 
-```{r load-data}
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"),
+          legend.position = "bottom")
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+# 2. Load and wrangle the person-trial data set.
+
+```{r load-data, results='hide'}
 ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
-print(ca_data,n=30)
-summary(ca_data) # 2683 rows: 6 participants, trial, 3 conditions, 15 periods, and acc
+head(ca_data)
 ```
@@
 # select analysis time range: (200,600]
-ca_data <- ca_data %>% filter(drt > 5)
+ca_data <- ca_data %>% filter(drt > 5) %>%
 
 # create factor condition, with "blank" as the reference level
-ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-
-# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
-ca_data <- ca_data %>% 
-        mutate(period_9 = drt - 9,
+mutate(condition = factor(condition, 
+                          labels = c("blank", "congruent","incongruent")),
+       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+      prime = factor(prime,levels=c(1,2,3)),
+       timebin = factor(drt, levels=c(6:15)),
+       period_9 = drt - 9,
        trial_c = (trial - 1000)/1000)
 
-head(ca_data,n=17)
-summary(ca_data)
-# 6 subjects
-# condition: blank (1303), congruent (668), incongruent (626)
-# period_9: -3 to 6
-# trial_c: -0.999 to 0.540
+head(ca_data)
 ```
@@
 
-We fit a single model "M4_ca": We use a third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.
+We fit a single model "M1i_ca", using index coding for both TIME and condition (variables "timebin" and "prime" in ca_data).
 
-## Prepare data for M4_ca.
+## Prepare data for M1i_ca.
 
-```{r data-M4-ca}
+```{r data-M1i-ca}
 # remove unnecessary columns before fitting a model
-M4_ca_data <- ca_data %>% select(-c(bl,tr,trial,drt)) # 2597 rows
-head(M4_ca_data)
+data_M1i_ca <- ca_data %>% select(pid,acc,timebin,prime)
+head(data_M1i_ca)
 ```
 
-### Set up priors for M4_ca.
+## Set up priors for M1i_ca.
 
-```{r priors-cloglog-M4-ca}
-priors_M4_ca <- c(
-  set_prior("normal(0,1)", class = "b"),       # weakly informative prior for beta parameters when using logit link
-  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using logit link
-  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
-  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
+```{r priors-cloglog-M1i-ca}
+priors_M1i_ca <- c(
+  set_prior("student_t(7.61, 0, 1.57)", class = "b"), # use flat prior on logit scale for the proportion ca 
+  set_prior("normal(0, 1)", class = "sd"),                    
+  set_prior("lkj(2)", class = "cor")                        
 )
@@
 
-### Fit model M4_ca.
+## Fit model M1i_ca.
 
-```{r fit-model-M4-ca, eval=F}
+```{r fit-model-M1i-ca, eval=F}
 plan(multicore)
 
-model_M4_ca <- 
-   brm(data = M4_ca_data,
-       family = binomial(link="logit"),
-       acc | trials(1) ~ 0 + Intercept +
-                          condition*period_9*trial_c + 
-                          condition*period_9*I(trial_c^2) + 
-                          condition*I(period_9^2)*trial_c +
-                          condition*I(period_9^2)*I(trial_c^2) +
-                          condition*I(period_9^3) +
-                          trial_c*I(period_9^3) +
-                          (1 +  condition*period_9*trial_c +
-                                condition*period_9*I(trial_c^2) + 
-                                condition*I(period_9^2)*trial_c +
-                                condition*I(period_9^2)*I(trial_c^2) +
-                                condition*I(period_9^3) +
-                                trial_c*I(period_9^3) | pid),
-       prior = priors_M4_ca,
+model_M1i_ca <- 
+   brm(data = data_M1i_ca,
+       family = bernoulli(link="logit"),  # use logit link for conditional accuracy
+       formula = acc ~ 0 + timebin:prime  +
+                        (0 + timebin:prime | pid),
+       prior = priors_M1i_ca,
        chains = 4, cores = 4, iter = 3000, warmup = 1000, 
@@
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M4_ca")
+       file = "Tutorial_2_Bayesian/models/model_M1i_ca")
 ```
 
-Model_M4_ca took about 7 hours to run.
+Model_M1i_ca took about 11 minutes to run.
 
-```{r check-model-M4-ca, eval=F}
-model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
-summary(model_M4_ca)
+```{r inspect-M1i-ca}
+model_M1i_ca <- readRDS("Tutorial_2_Bayesian/models/model_M1i_ca.rds")
+fixef(model_M1i_ca)
 ```
@@
 
-```{r load-model-M4-ca}
-model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
-```
-
 ## Pareto k estimates
@@
 ```{r pareto}
-loo(model_M4_ca)$diagnostics %>% 
+loo(model_M1i_ca)$diagnostics %>% 
   data.frame() %>% 
   # attach the `id` values
-  bind_cols(M4_ca_data) %>% 
+  bind_cols(data_M1i_ca) %>% 
   mutate(id = 1:n()) %>%
-
   ggplot(aes(x = id, y = pareto_k)) +
@@
 
-## Visualize posterior distributions of the effects of congruent and incongruent primes on logit-ca(t) relative to blank prime, for each time bin in trials 500, 1000, and 1500.
+## Plot posterior distributions for M1i_ca.
 
-```{r post-distr}
-get_variables(model_M4_ca)[1:31]
+Wrangle the posterior draws.
 
-post <-as_draws_df(model_M4_ca) %>% # 8000 draws x 715 variables
-   select(starts_with("b_")) %>%       # 8000 x 31
-   expand_grid(period_9 = -3:6) %>%   
-   mutate(period_9sq = period_9^2,
-          period_9cu = period_9^3,
-          trial500 = -500/1000,
-          trial500sq = trial500^2,
-          trial1500 = 500/1000,
-          trial1500sq = trial1500^2)
+```{r post-M0i}
+post_ca <- as_draws_df(model_M1i_ca) %>% 
+  select(-lp__) %>% 
+  as_tibble()
 
-# effects for trials 500, 1000, and 1500
-effects <- post %>% 
-  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
-            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
-            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
-            period_9sq * trial1500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
-            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
-            period_9sq * trial1500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`,
+post_summary <- posterior_summary(model_M1i_ca, robust = TRUE)
+post_summary[1:30,]
 
-         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
-            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
-            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
-            period_9sq * trial1500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
-            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
-            period_9sq * trial1500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`,
-
-         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
-            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
-            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
-            period_9sq * trial500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
-            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
-            period_9sq * trial500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`,
-
-         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
-            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
-            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
-            period_9sq * trial500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
-            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
-            period_9sq * trial500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`,
-
-         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
-            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
-
-         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
-            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
+post_qi_b <- post_ca %>%
+  select(starts_with("b_")) %>%
+  pivot_longer(everything()) %>%
+  group_by(name) %>% 
+  median_qi(value) %>% 
+  arrange(name)
+head(post_qi_b) # 30 "fixed" effects 
 ```
 
-```{r plot-effects}
-plot_facet <- effects %>%
-  pivot_longer(cols=`congruent trial 1500`:`incongruent trial 1000`, names_to = "condition" ) %>%
-  select(condition, period_9, value) %>%
-  group_by(period_9) %>%
+Visualise fixed effects.
 
-  ggplot(aes(x = period_9, y = value)) +
-  stat_lineribbon(point_interval = mean_hdi,.width = c(0.5, 0.8, 0.95),show.legend=T) +
-  scale_fill_brewer() +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
-                     limits = c(-3,6)) +
-  scale_y_continuous(breaks = c(-50,-40,-30,-20,-10,0,10,20,30,40,50), limits = c(-50,50)) +
-  labs(x = "time bin", y = "logit-ca") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle=90),
-        panel.background = element_rect(fill="white", color="black"),
-        legend.position = "top") +
-  facet_wrap(~ factor(condition,levels=c("congruent trial 500","congruent trial 1000","congruent trial 1500","incongruent trial 500","incongruent trial 1000","incongruent trial 1500")))
+```{r plot-post-M0i-ca}
+tidy_fixed_ca <- post_ca %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) %>% 
+  mutate(timebin = str_sub(name,10,11),
+         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
+         condition = str_sub(name,17,18),
+         condition = factor(str_remove(condition,"e"),
+                            levels=c(1,2,3),
+                            labels=c("blank","congruent","incongruent")))
+head(tidy_fixed_ca)
+tail(tidy_fixed_ca)
 
-plot_facet
+# plot
+p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
+  stat_halfeye(alpha=0.7) +
+  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1i_ca',
+       x = "time bin", 
+       y = "logit-ca") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_y_continuous(limits=c(-10,10)) +
+  theme(axis.text.x = element_text(angle=90)) +
+  facet_wrap(~condition)
+p_tidy_fixed_ca
 ```
 
-```{r save-plot-effects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png", width = 10, height = 8, dpi = 800)
 ```
 
-## calculate point and interval estimates, and odds ratios.
+## Plot predicted conditional accuracy functions
 
-```{r table-interval-estimates}
-int_c_500 <- effects %>%
-  select(`congruent trial 500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`congruent trial 500`),
-         mean = `congruent trial 500`,
-         condition = "c500") %>%
-  select(bin_endpoint,condition, mean, .lower, .upper, .width, `odds ratio`)
+### Grand mean
 
-int_i_500 <- effects %>%
-  select(`incongruent trial 500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`incongruent trial 500`),
-         mean = `incongruent trial 500`,
-         condition = "i500") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width, `odds ratio`)
+```{r epred-grand}
+dat_M1i_ca <- as_tibble(model_M1i_ca$data)
 
-int_c_1000 <- effects %>%
-  select(`congruent trial 1000`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`congruent trial 1000`),
-         mean = `congruent trial 1000`,
-         condition = "c1000" ) %>%
-  select(bin_endpoint,condition,mean, .lower, .upper, .width, `odds ratio`)
-
-int_i_1000 <- effects %>%
-  select(`incongruent trial 1000`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`incongruent trial 1000`),
-         mean = `incongruent trial 1000`,
-         condition = "i1000") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width,`odds ratio`)
-
-int_c_1500 <- effects %>%
-  select(`congruent trial 1500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`congruent trial 1500`),
-         mean = `congruent trial 1500`,
-         condition = "c1500" ) %>%
-  select(bin_endpoint, condition, mean, .lower, .upper, .width, `odds ratio`)
-
-int_i_1500 <- effects %>%
-  select(`incongruent trial 1500`, period_9) %>%
-  group_by(period_9) %>%
-  mean_hdi() %>%
-  mutate(bin_endpoint = (period_9 + 9)*40,
-        `odds ratio` = exp(`incongruent trial 1500`),
-         mean = `incongruent trial 1500`,
-         condition = "i1500") %>%
-  select(bin_endpoint, condition,mean, .lower, .upper, .width,  `odds ratio`)
-
-test <- rbind(int_c_500, int_c_1000, int_c_1500,int_i_500,int_i_1000, int_i_1500)
-test %>% print(n=60)
+epreds_grand <- dat_M1i_ca %>% 
+  data_grid(timebin, prime) %>% 
+  add_epred_draws(model_M1i_ca, 
+                  re_formula = NA) %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
 ```
 
-```{r save-table-interval-estimates, eval=F}
-write_csv(test, file="Tutorial_2_Bayesian/tables/ca_effects_intervals_table.csv")
+Summarize and plot predicted hazard values.
+
+```{r epred-grand-summary}
+epreds_grand %>% 
+  group_by(timebin,prime) %>% 
+  mean_qi()
 ```
 
-## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject.
-
-```{r select-pid}
-selected_subject = 6 # select pid here (1:6) manually
+```{r epred-grand-plot}
+ ggplot(epreds_grand, aes(x=timebin, y=.epred, 
+                          fill=prime, color=prime)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,1)) +
+    labs(y = "predicted conditional accuracy") +
+    ggtitle("Average participant")
 ```
 
-```{r help-function}
-make_fitted <- function(fit, scale, ...) {
-
-  fitted(fit,
-         newdata = nd,
-         scale = scale,
-         ...) %>% 
-    data.frame() %>% 
-    bind_cols(nd)
-}
+```{r save-epred-grand, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_grand.png", width = 10, height = 8, dpi = 800)
 ```
 
-### Set up new data.
+### Subject-specific
 
-```{r new-data}
-# define the `newdata`
-nd <- tibble(pid = 1:6) %>%
-  expand_grid(incon = 0:1,
-                con  = 0:1,
-                trial_c = c(-0.5,0,0.5),
-                period_9 = -3:6) %>%
-  filter(!(con ==1 & incon == 1)) %>%
-  mutate(Iperiod_9E2 = period_9^2,
-         Iperiod_9E3 = period_9^3,
-         Itrial_cE2 = trial_c^2,
-         condition = ifelse(con==1 & incon==0,"congruent",
-                     ifelse(con==0 & incon==1,"incongruent", "blank")),
-         condition = factor(condition, levels=c("blank" ,"congruent","incongruent"))) 
+```{r epred-pid}
+epreds_pid <- dat_M1i_ca %>% 
+  data_grid(pid,timebin, prime) %>% 
+  add_epred_draws(model_M1i_ca, 
+                  re_formula = NULL) %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
 ```
 
-### First, on the logit-ca(t) scale.
+Summarize and plot predicted hazard values.
 
-```{r subject-logit}
-make_fitted(model_M4_ca, scale = "linear") %>% 
-  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
-         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
-         period = period_9+9) %>% 
-  filter(pid==selected_subject) %>% 
-  # plot
-  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
-             fill = condition, color = condition)) +
-  geom_ribbon(alpha = 1/5, linewidth = 0) +
-  geom_line() +
-  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
-  ylab("fitted logit(ca)") +
-  coord_cartesian(ylim = c(-20, 20)) +
-  theme(panel.grid = element_blank(),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~ trial)
+```{r epred-pid-summary}
+epreds_pid %>% 
+  group_by(pid,timebin,prime) %>% 
+  mean_qi()
 ```
 
-### Second, on the ca(t) scale.
-
-```{r subject-ca}
-make_fitted(model_M4_ca, scale = "response") %>% 
-  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
-         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
-         period = period_9+9) %>% 
-  filter(pid==selected_subject) %>% 
-  # plot
-  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
-             fill = condition, color = condition)) +
-  geom_ribbon(alpha = 1/5, linewidth = 0) +
-  geom_line() +
-  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
-  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
-  ylab("fitted ca(t)") +
-  coord_cartesian(ylim = c(0, 1)) +
-  theme(panel.grid = element_blank(),
-        panel.background = element_rect(fill="white",color="black"),
-        legend.position = "top",
-        axis.text.x = element_text(angle=90)) +
-  facet_wrap(~ trial)
+```{r plot-epred-pid}
+ ggplot(epreds_pid, aes(x=timebin, y=.epred, 
+                       fill=prime, color=prime)) +
+    stat_lineribbon(alpha = 0.5, .width = .95) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,1)) +
+    labs(y = "predicted conditional accuracy") +
+    ggtitle("Each participant (N = 6)") +
+    facet_wrap(~pid, ncol=3)
 ```
 
-```{r save-plot-fit-hazard, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_subject6.png", width = 10, height = 10, dpi = 300)
+```{r save-epred-pid, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png", width = 10, height = 8, dpi = 800)
 ```
+
+## Plot average marginal effects (AMEs)
+
+### Grand mean AMEs
+
+```{r grand-ame}
+epreds_grand_diffs <- epreds_grand %>%
+  pivot_wider(id_cols = c(timebin, .draw),
+              names_from = "prime",
+              values_from = ".epred") %>% #80000 x 5
+  mutate(`congruent minus blank` = congruent - blank,
+         `incongruent minus blank` = incongruent -  blank) %>% 
+  select(-c(blank,congruent,incongruent)) %>% 
+  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
+               names_to = "contrast",
+               values_to = "pred_ca")
+
+epreds_grand_diffs # 160 000 rows 
+```
+
+Summarize and plot.
+
+```{r grand-ame-summary}
+table <- epreds_grand_diffs %>%
+  group_by(contrast,timebin) %>%
+  mean_qi(.width = c(.95)) %>%
+  arrange(contrast,timebin,.width) %>%
+  select(-c(.width,.point,.interval))
+```
+
+```{r save-table-grand-AMEs, eval=F}
+write_csv(table, file="Tutorial_2_Bayesian/tables/ca_grand_AMEs.csv")
+```
+
+```{r grand-ame-plot}
+ ggplot(epreds_grand_diffs, aes(x=timebin, y=pred_ca, 
+                       fill=contrast)) +
+  stat_lineribbon(alpha = 0.5, 
+                  point_interval = "mean_qi",
+                  .width=c(.8,.95)) +
+ # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
+  geom_hline(yintercept = 0, 
+             color = "red", 
+             lty = 3) +
+  scale_y_continuous(limits=c(-1,1)) +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(y = "difference in predicted ca") +
+  ggtitle("Grand average marginal effect") +
+  facet_wrap(~contrast)
+```
+
+```{r save-grand-AMEs, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png", width = 10, height = 8, dpi = 800)
+```
+
+### Subject-specific AMEs
+
+To calculate the subject-specific AMEs for each time bin, we create contrasts between the conditional means.
+
+```{r epred-pid-ame}
+epreds_pid_diffs <- epreds_pid %>%
+  pivot_wider(id_cols = c(pid,timebin, .draw),
+              names_from = "prime",
+              values_from = ".epred") %>% #80000 x 5
+  mutate(`congruent minus blank` = congruent - blank,
+         `incongruent minus blank` = incongruent -  blank) %>% 
+  select(-c(blank,congruent,incongruent)) %>% 
+  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
+               names_to = "contrast",
+               values_to = "pred_ca") 
+epreds_pid_diffs # 160 000 rows 
+```
+
+Summarize and plot the contrasts in predicted hazard values.
+
+```{r epred-pid-ame-summary}
+epreds_pid_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
+```
+
+```{r epred-pid-ame-plot}
+ ggplot(epreds_pid_diffs, aes(x=timebin, y=pred_ca, 
+                       fill=contrast, color=contrast)) +
+    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
+    geom_hline(yintercept = 0, color = "red", lty = 3) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(-1,1)) +
+    labs(y = "difference in predicted ca") +
+    ggtitle("Subject-specific AMEs") +
+    facet_wrap(~pid, ncol=3)
+```
+
+```{r save-pid-AMEs, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pid_AMEs.png", width = 10, height = 8, dpi = 800)
+```

### 2024-11-26 Tutorial_2a.Rmd
update tuts 2a, 2b, 3a, and 3b

@@
 
-```{r compare-models}
+```{r load-models}
 model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
@@
 
-```{r pareto}
+```{r pareto-M1i}
 loo(model_M1i)$diagnostics %>% 
@@
 
-## Posterior distributions of the population-level parameters of M1
+## Posterior distributions of the population-level parameters of M1i
 
@@
 
-```{r post-M0i}
+```{r post-M1i}
 post <- as_draws_df(model_M1i) %>% 
@@
 
-```{r plot-post-M0i}
+```{r plot-post-M1i}
 tidy_fixed <- post %>% 
@@
 p_tidy_fixed <- ggplot(tidy_fixed, aes(x = timebin, y = value, fill=condition)) +  
-  stat_halfeye(alpha=0.7) +
-  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1i',
+  stat_halfeye(point_interval="median_qi", 
+               .width = c(0.80,0.95),
+               alpha=0.7) +
+  labs(title = 'Posterior distributions for population-level\neffects in Model M1i',
        x = "time bin", y = "cloglog-hazard") +
@@
   group_by(timebin,prime) %>% 
-  mean_qi()
+  median_qi(.width=0.95)
 ```
@@
                           fill=prime, color=prime)) +
-    stat_lineribbon(alpha = 0.5, .width = .95) +
+    stat_lineribbon(point_interval="median_qi",.width=.95,alpha = 0.5) +
     scale_fill_brewer(palette = "Dark2") +
@@
 
-Again, for the average participant, but using posterior draws.
+Again, for the average participant, but now using posterior draws.
 
@@
            fill=condition, color=condition)) +
-    stat_lineribbon(alpha = 0.5, .width = .95) +
+    stat_lineribbon(point_interval="median_qi",
+                    .width=.95,
+                    alpha = 0.5) +
     scale_fill_brewer(palette = "Dark2") +
@@
   group_by(pid,timebin,prime) %>% 
-  mean_qi()
+  median_qi(.width=0.95)
 ```
@@
                        fill=prime, color=prime)) +
-    stat_lineribbon(alpha = 0.5, .width = .95) +
+    stat_lineribbon(point_interval="median_qi",
+                    .width=.95,
+                    alpha = 0.5) +
     scale_fill_brewer(palette = "Dark2") +
@@
 Third, for a completely new hypothetical participant.
-
-```{r}
-epreds_pid_new <- 
-  posterior_epred(model_M1i, 
-                  newdata = expand_grid(timebin=6:15,
-                                        prime=1:3,
-                                        pid=7), 
-                  re_formula = NULL, 
-                  allow_new_levels=T, 
-                  sample_new_levels = "gaussian") %>% 
-  as_tibble() %>%
-  mutate(draw = 1:n()) %>%
-  pivot_longer(V1:V30,
-               values_to = ".epred", 
-               names_to = "condition") %>%
-  mutate(prime = factor(rep(c(1,2,3), 10*8000), 
-                        levels=c(1,2,3),
-                        labels=c("blank","congruent","incongruent")),
-         timebin = factor(rep(rep(c(6:15),each=3),8000),
-                          levels=c(6:15))) %>%
-  ungroup()
-```
 
@@
   group_by(pid,timebin,prime) %>% 
-  mean_qi()
+  median_qi(.width=.95)
 ```
@@
 
-### Grand average marginal effects
+## Average marginal effects (AMEs)
+
+### Grand AMEs
 
@@
                names_to = "contrast",
-               values_to = "pred_haz")
+               values_to = "diff")
 
@@
 ```{r}
- ggplot(epreds_grand_diffs, aes(x=timebin, y=pred_haz, 
+ ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
                        fill=contrast)) +
@@
                   .width=c(.8,.95)) +
- # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
   geom_hline(yintercept = 0, 
@@
                names_to = "condition", 
-               values_to = "haz_diff") %>%
+               values_to = "diff") %>%
   mutate(bin = str_sub(condition,4,5),
@@
 
-  ggplot(aes(x=bin, y=haz_diff, fill=contrast)) +
-    stat_halfeye(point_interval = "mean_qi",
+  ggplot(aes(x=bin, y=diff, fill=contrast)) +
+     stat_lineribbon(alpha = 0.5, 
+                  point_interval = "mean_qi",
                   .width=c(.8,.95)) +
@@
                names_to = "contrast",
-               values_to = "pred_haz") 
+               values_to = "diff") 
 epreds_pid_diffs # 160 000 rows 
@@
 
-Summarize and plot the contrasts in predicted hazard values.
-
-```{r}
-epreds_pid_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
-```
-
-```{r}
- ggplot(epreds_pid_diffs, aes(x=timebin, y=pred_haz, 
+Summarize the contrasts in predicted hazard values.
+
+```{r}
+table <- epreds_pid_diffs %>% 
+  group_by(pid,timebin,contrast) %>% 
+  mean_qi(.width = .95) %>%
+  arrange(contrast,timebin,.width) %>%
+  select(-c(.width,.point,.interval))
+```
+
+```{r save-table-pid-AMEs, eval=F}
+write_csv(table, file="Tutorial_2_Bayesian/tables/pid_AMEs.csv")
+```
+
+Plot the contrasts in predicted hazard values.
+
+```{r}
+ ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
                               fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    stat_lineribbon(alpha = 0.5, 
+                    point_interval = "mean_qi",
+                    .width=c(.8,.95)) +
     #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
@@
 
-[[remove next section....]]
+[[remove next section and 2 code chunks....]]
 And check whether an average of subject-specific AMEs is equal to the grand AME... NO !!!
@@
                names_to = "contrast",
-               values_to = "pred_haz") 
+               values_to = "diff") 
 epreds_pid_new_diffs 
@@
 ```{r}
-epreds_pid_new_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
+epreds_pid_new_diffs %>% 
+  group_by(pid,timebin,contrast) %>% 
+  median_qi(.width = .95) %>%
+  arrange(contrast,timebin,.width) %>%
+  select(-c(.width,.point,.interval))
 ```
@@
     theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(-0.7,0.7)) +
+    scale_y_continuous(limits=c(-1,1)) +
     labs(y = "difference in predicted hazard") +
@@
 
-What can we conclude from model M1 about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
-In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of xxx ms)?
-
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on the grand AMEs of congruent and incongruent primes in each bin.
-The contrast "congruent minus blank" was estimated to be 0.0925 hazard units in bin 6 (95% CrI = [0.0411, 0.183]), and 0.152 hazard units in bin 7 (95% CrI = [0.0489, 0.262]).
+What can we conclude from model M1i about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
+In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of 187 ms)?
+
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
+The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 7 (95% CrI = [0.04, 0.25]).
 For the other bins, the 95% credible interval contained zero.
-The contrast "incongruent minus blank" was estimated to be 0.0943 hazard units in bin 6 (95% CrI = [0.0322, 0.208]), -0.204 hazard units in bin 9 (95% CrI = [-0.331, -0.0769]), -0.275 hazard units in bin 10 (95% CrI = [-0.45, -0.0513]), and -0.213 hazard units in bin 11 (95% CrI = [-0.377, -0.0137]).
+The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin 9 (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin 10 (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin 11 (95% CrI = [-0.40, -0.03]).
 For the other bins, the 95% credible interval contained zero.
@@
 
-There are thus two phases of performance for the average subject between 200 and 600 ms after target onset.
+There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
 In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240].
@@
   group_by(timebin,contrast) %>%
-  summarize(prop_above = mean(pred_haz > 0)) %>% 
+  summarize(prop_above = mean(diff > 0)) %>% 
   arrange(contrast,timebin)
@@
   group_by(timebin,contrast) %>%
-  summarize(prop_below = mean(pred_haz < 0)) %>% 
+  summarize(prop_below = mean(diff < 0)) %>% 
   arrange(contrast,timebin)
 
-pabove %>% inner_join(pbelow, by = c("timebin","contrast"))
-```
-
-Thus the probability that the contrast "congruent minus blank" is larger than 0, is larger than .80 in bins 6 to 8.
-And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .80 in bins 8 to 12.
-
-[[todo or not todo]]
-Plot of posterior distributions of sigma(timebinX:primeY) : check size differences between timebins and prime types? -> these are also population-level effects
-
-If we want to focus more on inter-individual differences, we can study the subject-specific AMEs in Figure XX.
-While most participants show the grand AMEs, three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] - subject 2 also in bin (320,360].
+table_prop <- pabove %>% inner_join(pbelow, by = c("timebin","contrast"))
+```
+
+```{r save-table-props, eval=F}
+write_csv(table_prop, file="Tutorial_2_Bayesian/tables/contrasts_props.csv")
+```
+
+Thus, the probability that the contrast "congruent minus blank" is larger than 0, is larger than .9 in bins 6 to 8.
+And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .9 in bins 9 to 12.
+
+If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
 
 Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
-For example, Panis and Schmidt (ref) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here:the target).
+For example, @panisWhatShapingRT2016 concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
 Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale (TIME).
+This effect might be masked for incongruent primes by the response competition effect.
+
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 9).
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
 In Tutorial_2a.Rmd we provide a number of model formula that should get you going.
@@
 ggplot() +
-  geom_hline(yintercept=0, color="white") +
+  geom_hline(yintercept=0, color="grey") +
   geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
@@
 
-# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. This might be a good prior to use for the beta parameters in a logit-hazard model.
+# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. 
 pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  

### 2024-11-26 Tutorial_2b.Rmd
update tuts 2a, 2b, 3a, and 3b

@@
 p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
-  stat_halfeye(alpha=0.7) +
-  labs(title = 'Posterior distributions for "fixed" effects\nin Model M1i_ca',
+  stat_halfeye(point_interval = "median_qi",
+               .width = c(.80,.95),
+               alpha=0.7) +
+  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
        x = "time bin", 
@@
   group_by(timebin,prime) %>% 
-  mean_qi()
+  median_qi()
 ```
@@
                           fill=prime, color=prime)) +
-    stat_lineribbon(alpha = 0.5, .width = .95) +
+    stat_lineribbon(point_interval="median_qi",
+                    alpha = 0.5, 
+                    .width = .95) +
     scale_fill_brewer(palette = "Dark2") +
@@
   group_by(pid,timebin,prime) %>% 
-  mean_qi()
+  median_qi()
 ```
@@
                        fill=prime, color=prime)) +
-    stat_lineribbon(alpha = 0.5, .width = .95) +
+    stat_lineribbon(point_interval="median_qi",
+                    alpha = 0.5, 
+                    .width = .95) +
     scale_fill_brewer(palette = "Dark2") +
@@
                names_to = "contrast",
-               values_to = "pred_ca")
+               values_to = "diff_ca")
 
@@
 ```{r grand-ame-plot}
- ggplot(epreds_grand_diffs, aes(x=timebin, y=pred_ca, 
+ ggplot(epreds_grand_diffs, aes(x=timebin, y=diff_ca, 
                        fill=contrast)) +
@@
                names_to = "contrast",
-               values_to = "pred_ca") 
+               values_to = "diff_ca") 
 epreds_pid_diffs # 160 000 rows 
@@
 ```{r epred-pid-ame-summary}
-epreds_pid_diffs %>% group_by(pid,timebin,contrast) %>% mean_qi()
+epreds_pid_diffs %>% 
+  group_by(pid,timebin,contrast) %>% 
+  mean_qi()
 ```
@@
 ```{r epred-pid-ame-plot}
- ggplot(epreds_pid_diffs, aes(x=timebin, y=pred_ca, 
+ ggplot(epreds_pid_diffs, aes(x=timebin, y=diff_ca, 
                        fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.95)) +
+    stat_lineribbon(alpha = 0.5, 
+                    point_interval = "mean_qi",
+                    .width=c(.95)) +
     #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +

### 2024-11-26 Tutorial_3a.Rmd
update html and Rmd files
update tuts 2a, 2b, 3a, and 3b

--- 2024-09-17 Tutorial_3.Rmd
@@
 ---
-title: "Tutorial_3"
+title: "Tutorial_3a"
 author: "Sven Panis"
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-# Load the libraries that we will be using
+# 1. Load the libraries that we will be using
 
-```{r load-pkg}
-pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")
+```{r load-pkg, results='hide'}
+pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")
 
@@
 
-# Load the person-trial-bin data set that we saved in Tutorial 1.
+Theme settings for ggplot.
+
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"))
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+# 2. Load and wrangle the person-trial-bin data set.
 
@@
 ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-print(ptb_data,n=30)
-summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
 ```
 
-# Prepare the data set.
+Prepare the data set.
 
 ```{r prepare-data}
-# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
-ptb_data <- ptb_data %>% filter(period > 5)
+# select analysis time range
+ptb_data <- ptb_data %>% filter(period > 5) %>%
 
 # create factor condition, with "blank" as the reference level
-ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent"))) %>%
 
-# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
-ptb_data <- ptb_data %>% 
+# center TIME (period) on bin 9, center trial on trial 1000 and rescale, 
+# create variables with index coding (timebin, prime).
 mutate(period_9 = period - 9,
-               trial_c = (trial - 1000)/1000)
+       trial_c = (trial - 1000)/1000,
+       timebin = factor(period, levels=c(6:15)),
+       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+       prime = factor(prime, levels=c(1,2,3)))
 
 head(ptb_data,n=17)
-summary(ptb_data)
-# 6 subjects
-# condition: blank (6401), congruent (2642), incongruent (3797)
-# period_9: -3 to 6
-# trial_c: -0.999 to 0.540
 ```
 
-# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)
+# 3. 3. Fit hazard model in the frequentist framework (model_M1i_f)
 
 ```{r get-data}
-M3_data <- ptb_data %>% select(pid, period_9, trial_c, condition, event)
+data_M1i_f <- ptb_data %>% select(pid, event, timebin, prime)
 ```
 
+Model_M1i_f contains the same effects as Model_M1i in Tutorial 2a.
+
 ```{r fit-model, eval=F}
-model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
-
+model_M1i_f <- glmer(event ~ 0 + timebin:prime  +
+                            (0 + timebin:prime | pid),
           # control parameters, data set, and complementary log-log link function
@@
                                  optCtrl = list(maxfun=10000000)), 
-              data=M3_data, 
-              family=binomial(link="cloglog"))
+          data = data_M1i_f, 
+          family = binomial(link="cloglog"))
 ```
 
-Model_M3_f took about 3 hours.
+Model_M1i_f took about 3 hours.
 
-```{r save-model, eval=F}
-save(model_M3_f, file="Tutorial_3_Frequentist/model_M3_f.RData")
+```{r save-M1i-f, eval=F}
+save(model_M1i_f, file="Tutorial_3_Frequentist/model_M1i_f.RData")
 ```
 
-# Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3
+```{r inspect-M1i-f}
+load("Tutorial_3_Frequentist/model_M1i_f.RData")
+summary(model_M1i_f)
+```
+
+# 4. Compare MLE estimates from model_M1i_f with the posterior mean estimates from model_M1i.
 
 ```{r load-models}
-model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
-load("Tutorial_3_Frequentist/model_M3_f.RData")
+model_M1i <- readRDS("Tutorial_2_Bayesian/models/model_M1i.rds")
+load("Tutorial_3_Frequentist/model_M1i_f.RData")
 ```
@@
 ```{r extract-estimates}
-est <- tibble(brm = fixef(model_M3)[,1],
-              glmer = fixef(model_M3_f),
-              effects = names(fixef(model_M3_f))) 
+est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
+              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
+              effectname = names(fixef(model_M1i_f)),
+              position = 1:30)
+
+colnames(est)
 ```
@@
 
-p1 <- ggplot(data = est, aes(x = effects))
-p1 <- p1 + geom_point(aes( y = brm,  color = "c1", shape = "s1"))
-p1 <- p1 + geom_point(aes( y = glmer, color = "c2", shape = "s2")) 
-p1 <- p1 + labs( x = "Effects", y = "cloglog-hazard" )
-p1 <- p1 + scale_color_manual(name = "Model", 
+p1 <- ggplot(data = est, aes(x = position)) +
+ geom_point(aes( y = brm[,"Estimate"],  
+                 color = "c1", 
+                 shape = "s1"),
+                 size = 1.5) +
+ geom_point(aes( y = glmer$estimate, 
+                 color = "c2", 
+                 shape = "s2"),
+                 size = 1.5, 
+                 position = position_nudge(x = .2))  +
+ geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
+                    color = "c1"), 
+                show.legend = F) +
+ geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
+                    color = "c2"), 
+                show.legend = F, 
+                position = position_nudge(x = .2)) +
+ geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 1:30,labels = est$effectname) +
+ scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +
+
+ labs( x = "Effects in model M1i", y = "cloglog-hazard" ) +
+ scale_color_manual(name = "Model", 
                                 breaks = c("c1", "c2"), 
                                 values = cols,
-                                labels = c("brm", "glmer"))
-p1 <- p1 + scale_shape_manual(name = "Model", 
+                                labels = c("brm", "glmer")) +
+ scale_shape_manual(name = "Model", 
                               breaks = c("s1", "s2"),
@@
                               labels = c("brm", "glmer")) +
-   theme(panel.grid = element_blank(),
-         panel.background = element_rect(fill="white",color="black"),
-         legend.position = "top",
-         axis.text.x = element_text(angle=90)) 
+  coord_flip() #+
+ #theme(panel.grid = element_blank())
+         #legend.position = "top",
+         #axis.text.x = element_text(size=11,angle=90),
+         #axis.text.y = element_text(size=13),
+         #axis.title.x = element_text(size = 14),
+         #axis.title.y = element_text(size = 14),
+         #legend.text=element_text(size=13)) 
 p1
 ```
+
+```{r save-plot-comparison, eval=F}
+ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=8,dpi=800)
+```

### 2024-11-26 Tutorial_3b.Rmd
update html and Rmd files
update tuts 2a, 2b, 3a, and 3b

--- 2024-10-18 Tutorial_3b.Rmd
@@
 date: "2024-10-18"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    theme: united
 editor_options: 
@@
 
-# Load the libraries that we will be using
+# 1. Load the libraries that we will be using.
 
@@
 
-# Load the data set for ca(t) modeling that we saved in Tutorial 1a.
+Theme settings for ggplot.
+
+```{r plot-settings, results='hide'}
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 22, face = "bold"), 
+          title = element_text(size = 22, face = "bold"))
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+# 2. Load and wrangle the person-trial data set.
 
@@
 ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
-print(ca_data,n=30)
-summary(ca_data) # 2683 rows
 ```
 
-# Prepare the data set.
+Prepare the data set.
 
@@
 # select analysis time range: (200,600]
-ca_data <- ca_data %>% filter(drt > 5)
+ca_data <- ca_data %>% filter(drt > 5) %>%
 
-# create factor condition, with "blank" as the reference level
-ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
+# create factor condition, with "blank" as the reference level; center TIME; center and rescale trial; create index-coded variables timebin and prime
+mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")),
+       period_9 = drt - 9,
+       trial_c = (trial - 1000)/1000,
+       timebin = factor(drt, levels=c(6:15)),
+       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+       prime = factor(prime, levels=c(1,2,3)))
 
-# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
-ca_data <- ca_data %>% 
-        mutate(period_9 = drt - 9,
-               trial_c = (trial - 1000)/1000)
-
-head(ca_data,n=17)
-summary(ca_data)
-# 6 subjects
-# condition: blank (1303), congruent (668), incongruent (626)
-# period_9: -3 to 6
-# trial_c: -0.999 to 0.540
+head(ca_data)
 ```
 
-# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)
+# 3. Fit a conditional accuracy model in the frequentist framework (model M1i_ca_f).
 
 ```{r get-data}
-M3_ca_data <- ca_data %>% select(pid, period_9, trial_c, condition, acc)
+data_M1i_ca_f <- ca_data %>% select(pid, acc, timebin, prime)
 ```
 
+Model_M1i_ca_f contains the same effects as Model_M1i_ca in Tutorial 2b.
+
 ```{r fit-model, eval=F}
-model_M3_ca_f <- glmer(acc ~ 1 + condition*period_9 + 
-                           condition*I(period_9^2) + 
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) +
-                           (1 + condition*period_9 +
-                           condition*I(period_9^2) +
-                           condition*I(period_9^3) +
-                           trial_c*period_9 + 
-                           trial_c*I(period_9^2) + 
-                           trial_c*I(period_9^3) | pid),
+model_M1i_ca_f <- glmer(acc ~ 0 + timebin:prime  +
+                        (0 + timebin:prime | pid),
 
@@
                                      optCtrl = list(maxfun=10000000)), 
-              data=M3_ca_data, 
-              family=binomial(link="logit"))
+              data = data_M1i_ca_f, 
+              family = binomial(link="logit"))
 ```
 
-Model_M3_ca_f took about 15 minutes to fit.
+Model_M1i_ca_f took about 90 minutes to fit.
 
 ```{r save-model, eval=F}
-save(model_M3_ca_f, file="Tutorial_3_Frequentist/model_M3_ca_f.RData")
+save(model_M1i_ca_f, file="Tutorial_3_Frequentist/model_M1i_ca_f.RData")
 ```
+
+```{r effects-M1i-ca-f}
+load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
+fixef(model_M1i_ca_f)
+summary(model_M1i_ca_f)
+```
+
+# 4. Compare MLE estimates from model_M1i_ca_f with the posterior mean estimates from model_M1i_ca.
+
+```{r load-models}
+model_M1i_ca <- readRDS("Tutorial_2_Bayesian/models/model_M1i_ca.rds")
+load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
+```
+
+```{r extract-estimates}
+est <- tibble(brm = fixef(model_M1i_ca)[,c(1,3,4)],
+              glmer = tidy(model_M1i_ca_f, 
+                           effects = c("fixed"), 
+                           conf.int=T, 
+                           conf.level = 0.95)[,c(2,3,7,8)], 
+              effectname = names(fixef(model_M1i_ca_f)),
+              position = 1:30)
+
+summary(est)
+head(est)
+```
+
+```{r plot-estimates}
+cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
+shapes  <- c("s1" = 16, "s2" = 17)
+
+p1 <- ggplot(data = est, aes(x = position)) +
+ geom_point(aes( y = brm[,"Estimate"],  
+                 color = "c1", 
+                 shape = "s1"),
+                 size = 1.5) +
+ geom_point(aes( y = glmer$estimate, 
+                 color = "c2", 
+                 shape = "s2"),
+                 size = 1.5, 
+                 position = position_nudge(x = .2))  +
+ geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
+                    color = "c1"), 
+                show.legend = F) +
+ geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
+                    color = "c2"), 
+                show.legend = F, 
+                position = position_nudge(x = .2)) +
+ geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 1:30,labels = est$effectname) +
+ scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +
+
+ labs( x = "Effects in model M1i_ca", y = "logit-ca" ) +
+ scale_color_manual(name = "Model", 
+                                breaks = c("c1", "c2"), 
+                                values = cols,
+                                labels = c("brm", "glmer")) +
+ scale_shape_manual(name = "Model", 
+                              breaks = c("s1", "s2"),
+                              values = shapes,
+                              labels = c("brm", "glmer")) +
+  coord_flip() #+
+ #theme(panel.grid = element_blank())
+         #legend.position = "top",
+         #axis.text.x = element_text(size=11,angle=90),
+         #axis.text.y = element_text(size=13),
+         #axis.title.x = element_text(size = 14),
+         #axis.title.y = element_text(size = 14),
+         #legend.text=element_text(size=13)) 
+p1
+```
+
+```{r save-plot-comparison, eval=F}
+ggsave("Tutorial_3_Frequentist/Comparison_ca.png", height=8, width=8,dpi=800)
+```

### 2024-11-26 manuscript/manuscript.Rmd
update ms.

@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
 
@@
 
-* 1a. Wrangle raw data and calculate descriptive stats for one independent variable.
-* 1b. Wrangle raw data and calculate descriptive stats for two independent variables.
+* 1a. Wrangle raw data and calculate descriptive stats for one independent variable
+* 1b. Wrangle raw data and calculate descriptive stats for two independent variables
 * 2a. Bayesian multilevel modeling for h(t)
@@
 * 3b. Frequentist multilevel modeling for ca(t)
-
-Plannng (T4) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
+* 4.  Simulation and power analysis for planning experiments
 
@@
 
-In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone may be guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 as part of Tutorial 1b for those that are interested.
-
-(ref:descr-plot5-caption) Sample-based discrete-time hazard, survivor, conditional accuracy, and probability mass functions for participant 6, as a function of prime type (blank, congruent, incongruent) and mask type (no mask, relevant, irrelevant, lines).
-
-```{r plot-2ivs, fig.cap = "(ref:descr-plot5-caption)", out.width='95%'}
-# knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject6_PanisSchmidt.png")
-```
-
-## 4.3 Tutorial 2a: Fitting Bayesian hazard models to time-to-event data
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial 1b for those that are interested.
+
+## 4.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data
 
@@
 
-There are several analytic decisions one has to make when fitting a hazard model.
+There are several analytic decisions one has to make when fitting a discrete-time hazard model.
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
-Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t) in a selected baseline condition.
-One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below; see also part D of the supplementary material).
-
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
+And when you treat time as a discrete predictor, you can choose between reference coding and index coding.
 In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set [@allisonSurvivalAnalysisUsing2010].
@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
+In tutorial_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We decided to select the analysis time window (200,600] and the cloglog link.
-The data is prepared as follows.
+Below, we shortly discuss three of these models.
+The person-trial-bin data set is prepared as follows.
 
 \scriptsize
-```{r prepare-data, echo=T}
-# load person-trial-bin oriented data set
-ptb_data <- read_csv("../Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
-# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)
-ptb_data <- ptb_data %>% filter(period > 5)
-# create factor condition, with "blank" as the reference level
-ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))
-# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.
+```{r wrangle-data, eval=F, echo=T}
+# read in the file we saved in tutorial 1a
+ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
+
 ptb_data <- ptb_data %>% 
+# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
+filter(period > 5) %>%
+       # continuous predictor for TIME named "period_9", centered on bin 9
 mutate(period_9 = period - 9,
-               trial_c = (trial - 1000)/1000,
-               d6  = if_else(period == 6, 1, 0),
-               d7  = if_else(period == 7, 1, 0),
-               d8  = if_else(period == 8, 1, 0),
-               d9  = if_else(period == 9, 1, 0),
-               d10 = if_else(period == 10, 1, 0),
-               d11 = if_else(period == 11, 1, 0),
-               d12 = if_else(period == 12, 1, 0),
-               d13 = if_else(period == 13, 1, 0),
-               d14 = if_else(period == 14, 1, 0),
-               d15 = if_else(period == 15, 1, 0))
+       # categorical predictor for TIME named "timebin" with index coding
+       timebin = factor(period,levels=c(6:15)),
+       # factor "condition" using reference coding, with "blank" as the reference level
+       condition = factor(condition, labels = c("blank", "congruent","incongruent")),
+       # categorical predictor "prime" with index coding
+       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+       prime = factor(prime,levels=c(1,2,3)))
 ```
@@
 
-To get the posterior distribution of each model parameter given the data, we need to specify a prior distribution for each parameter.
-The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
-
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
-The other rows in Figure 12 show prior distributions on the logit and cloglog scale that we use instead.
-
-### 4.3.3 Model 1: A general specification of TIME, and main effects of congruency and trial number
-
-When you do not want to make assumptions about the shape of the hazard function in the selected baseline condition, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one intercept per time bin.
-In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M1 is specified as follows.
-
-\scriptsize
-```{r fit-model-M1, echo=T, eval=F}
-model_M1 <-
-   brm(data = M1_data,
-       family = binomial(link="cloglog"),
-       event | trials(1) ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
-                           condition + trial_c +
-                           (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15 + 
-                           condition + trial_c | pid),
-       prior = priors_M1,
-       chains = 4, cores = 4, iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
+To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
+In Tutorial_2a.Rmd we perform some prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
+
+The middle column of Figure 16 in part E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in supplementary Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+
+### 4.3.3 Model M0i: A null model with index coding
+
+When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
+In this first model, we use a general specification of TIME using index coding, and do not include experimental predictors.
+We call this model "M0i".
+
+Before we fit model M0i, we select the necessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M0i is specified as follows.
+
+```{r fit-model-M0i, echo=T, eval=F} 
+model_M0i <-                    
+   brm(data = data_M0i,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0i,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
        seed = 12, init = "0",
-       file = "Tutorial_2_Bayesian/models/model_M1")
-```
-\normalsize
-
-After selecting the binomial family and the cloglog link, the model formula is specified.
-The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of prime-target congruency (variable condition) and centered trial number (variable trial_c).
-Each of these effects is allowed to vary across individuals (variable pid).
-Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
-
-### 4.3.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number
-
-When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
-For our second example model, we thus use a third-order polynomial specification of TIME for the selected baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
-The model formula for model M2 looks as follows.
-
-\scriptsize 
-```{r formula-model-M2, eval=F, echo=T}
-event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) +
-                           condition + trial_c +
-                           (1 + period_9 + I(period_9^2) + I(period_9^3) +
-                           condition + trial_c | pid),
-```
-\normalsize
-
-Because TIME is centered on bin 9, and trial number on trial 1000, the Intercept represents the cloglog-hazard in bin 9 for the blank prime condition in trial 1000.
-Estimating model M2 took about 2.5 hours.
-
-### 4.3.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption
-
-So far, we assumed that the effect of our predictors prime-target congruecy and centered trial number are the same in each time bin.
-However, the descriptive plots (e.g., Figure 4) suggest that the effect of prime-target congruency varies across time bins.
+       file = "Tutorial_2_Bayesian/models/model_M0i")
+```
+
+After selecting the bernoulli family and the cloglog link, the model formula is specified.
+The specification "0 + ..." removes the default intercept in brm().
+The fixed effects include an intercept for each level of timebin.
+Each of these intercepts is allowed to vary across individuals (variable pid).
+Estimating model M0i took about 30 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+
+### 4.3.4 Model M1i: Adding the effects of prime-target congruency
+
 Previous research has shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
-For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable condition) and centered trial number (variable trial_c).
-
-\scriptsize
-```{r formula-model-M3, eval=F, echo=T}
-        event | trials(1) ~ 0 + Intercept + 
-                            condition*period_9 +  
-                            condition*I(period_9^2) +  
-                            condition*I(period_9^3) + 
-                            trial_c*period_9 +  
-                            trial_c*I(period_9^2) +  
-                            trial_c*I(period_9^3) + 
-                            (1 + condition*period_9 + 
-                            condition*I(period_9^2) + 
-                            condition*I(period_9^3) + 
-                            trial_c*period_9 +  
-                            trial_c*I(period_9^2) +  
-                            trial_c*I(period_9^3) | pid), 
-```
-\normalsize
-
-Note that duplicate terms across the interaction terms in the model formula (e.g., condition) are ignored. Estimating model M3 took about 4.5 hours.
-
-### 4.3.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions
-
-Based on previous work [e.g., @panisHowCanWe2020] we expect nonlinear effects of trial number on the hazard of response occurrence.
-We thus relax all three assumptions in model 4.
-We add a squared term for the continuous predictor centered trial number -- I(trial_c^2) -- and include interaction terms.
-For example, how the effect of congruent primes changes across time bins within a trial might change across the trials within an experiment.
-
-\scriptsize  
-```{r fit-model-M4, eval=F, echo=T} 
-event | trials(1) ~ 0 + Intercept + 
-                           condition*period_9*trial_c +  
-                           condition*period_9*I(trial_c^2) +
-                           condition*I(period_9^2)*trial_c + 
-                           condition*I(period_9^2)*I(trial_c^2) + 
-                           condition*I(period_9^3) + 
-                           trial_c*I(period_9^3) + 
-                           (1 +  condition*period_9*trial_c + 
-                                 condition*period_9*I(trial_c^2) +  
-                                 condition*I(period_9^2)*trial_c + 
-                                 condition*I(period_9^2)*I(trial_c^2)   
-                                 condition*I(period_9^3) + 
-                                 trial_c*I(period_9^3) | pid)
-```
-\normalsize
-
-Again, duplicate terms across the interaction terms in the model formula are ignored. Estimating model M4 took about 8 hours.
-
-### 4.3.7 Compare the models.
-
-We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
-
-```{r compare-models, echo=F, eval=T} 
-model_M1 <- readRDS("../Tutorial_2_Bayesian/models/model_M1.rds") 
-model_M2 <- readRDS("../Tutorial_2_Bayesian/models/model_M2.rds") 
-model_M3 <- readRDS("../Tutorial_2_Bayesian/models/model_M3.rds") 
-model_M4 <- readRDS("../Tutorial_2_Bayesian/models/model_M4.rds") 
-```
-
-\scriptsize
+In the next model, we use index coding for both TIME (variable "timebin") and the categorical predictor prime-target-congruency (variable "prime"), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
+Here is the model formula of this model that we call "M1i".
+
+```{r fit-model-M1i, eval=F, echo=T}
+event ~ 0 + timebin:prime + (0 + timebin:prime | pid)
+```
+
+Estimating model M1i took about 124 minutes.
+
+### 4.3.4 Model M1d: A more parsimonious model
+
+When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
+Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").
+
+For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called "M1d".
+Here is its model formula.
+
+```{r fit-model-M1d, eval=F, echo=T}
+event ~ 0 + Intercept + condition*period_9 + condition*period_9_sq + 
+                   (1 + condition*period_9 + condition*period_9_sq | pid)
+```
+
+The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
+The variable period_9_sq is a squared version of period_9.
+Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 145 minutes to run.
+
+### 4.3.5 Compare the models.
+
+We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
+
+```{r load-models, echo=F, eval=T, echo=F} 
+model_M0i <- readRDS("../Tutorial_2_Bayesian/models/model_M0i.rds") 
+model_M1i <- readRDS("../Tutorial_2_Bayesian/models/model_M1i.rds") 
+model_M1d <- readRDS("../Tutorial_2_Bayesian/models/model_M1d.rds")
+
+model_M0i <- add_criterion(model_M0i, c("loo", "waic"))
+model_M1i <- add_criterion(model_M1i, c("loo", "waic"))
+model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
+```
+
 ```{r show-weights, echo=T, eval=T}
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "loo") %>% round(digits = 3)
-model_weights(model_M1, model_M2, model_M3, model_M4, weights = "waic") %>% round(digits = 3)
-```
-\normalsize
-
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M4, and a weight of 0 to the other three simpler models.
-
-### 4.3.8 Evaluate parameter estimates
-
-To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
-
-Figure 6 shows one point (mean) and three highest posterior density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
-
-(ref:descr-plot6-caption) Means and 50/80/95% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
-
-```{r plot-prime-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png")
-```
-
-Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
-Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
-For simplicity and ease of presentation, we only tabulate data for a subset of the design in the main text (trial 500 for congruent and incongruent conditions).
-For the full table, see Supplementary materials or Tutorual X.
-
-```{r}
-int_tab <- read_csv("../Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
-```
-
-(ref:int-table-caption) Point and 95% highest posterior density interval estimates, and hazard ratios.
-
-(ref:int-table-note-caption) c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.
-
-```{r int-table}
-# apa_table(
-#   int_tab,
-#   caption = "(ref:int-table-caption)",
-#   note = "(ref:int-table-note-caption)",
-#   placement = "H",
+model_weights(model_M0i, model_M1i, model_M1d, weights = "loo") %>% round(digits = 2)
+model_weights(model_M0i, model_M1i, model_M1d, weights = "waic") %>% round(digits = 2) 
+```
+
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other two simpler models.
+
+### 4.3.6 Interpreting model M1i
+
+To make inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80% and 95% credible intervals).
+
+(ref:descr-plot5-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
+
+```{r plot-fixed-effects, fig.cap = "(ref:descr-plot5-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_postdistr.png")
+```
+
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6), and for each participant in the data set (Figure 7).
+
+(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.
+
+```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_grand.png")
+```
+
+(ref:descr-plot7-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.
+
+```{r plot-pred-pid-effects, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_pid.png")
+```
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 8; grand average marginal effect) and for each participant in the data set (Figure 9; subject-specific average marginal effect).
+
+(ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the average participant.
+
+```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_grand_AME.png")
+```
+
+(ref:descr-plot9-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for each participant.
+
+```{r plot-pid-ame-effects, fig.cap = "(ref:descr-plot9-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pid_AMEs.png")
+```
+
+Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.
+
+```{r load-tables, eval=T, echo=F}
+tab_grand_ame <- read_csv("../Tutorial_2_Bayesian/tables/grand_AMEs.csv")
+#tab_pid_ame <- read_csv("../Tutorial_2_Bayesian/tables/pid_AMEs.csv")
+tab_contrast_props <- read_csv("../Tutorial_2_Bayesian/tables/contrasts_props.csv")
+```
+
+(ref:table-grand-ame-caption) Point (mean) and 95% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.
+
+(ref:table-grand-ame-note-caption) diff = difference in predicted hazard.
+
+```{r table-grand-ame}
+ apa_table(
+   tab_grand_ame,
+   caption = "(ref:table-grand-ame-caption)",
+   note = "(ref:table-grand-ame-note-caption)",
+   placement = "H",
 #   font_size = "scriptsize",
-#   longtable = TRUE,
-#   repeat_header = TRUE,
-#   format.args = list(
-#        digits = c(0, 2,2,2,2,2,5),
-#         decimal.mark = ".", big.mark = "")
-# )
-```
-
-Based on Figure 6 and Supplementary Table XX, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240] on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
-For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240] of trial 500.
-Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360], (360,400], and (400,440].
-For example, in bin (320,360], the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
-While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and the target-triggered response (e.g., right)
-
-In the middle of the experiment (trial 1000), both congruent and incongruent primes have positive effects in bins (200,240] and (240,280], while incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440].
-Probably due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms early in a trial (compared to 40 ms at the beginning of the experiment) compared to the blank prime condition.
-
-Towards the end of the experiment (trial 1500), both congruent and incongruent primes have positive and negative effects.
-Positive effects are present in bins (200,240] and (240,280].
-Incongruent primes again have negative effects in bins (320,360], (360,400], and (400,440], and congruent primes now also have negative effects in bins (360,400] and (400,440].
-
-These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period_9), but also on the across-trial/within-experiment time scale (variable trial_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, suggests the involvement of separate cognitive processes.
-
-@panisWhatShapingRT2016 distinguished between automatic response competition (bottom-up lateral inhibition between response channels), active and global inhibition (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the correct response to the target stimulus.
-Thus, it seems that this global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
-Interestingly, while @panisWhatShapingRT2016 did not test interactions between prime-target congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which nicely coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
-
-To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trials 500, 1000, and 1500.
-
-(ref:descr-plot7-caption) Model-based hazard functions for each prime type for participant 6 in trials 500, 1000, and 1500.
-
-```{r plot-hazard-subject6, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_subject6.png")
-```
+   longtable = TRUE,
+   repeat_header = TRUE,
+   format.args = list(
+        digits = c(0,0,2,2,2),
+        decimal.mark = ".", big.mark = "")
+)
+```
+
+#### Example conclusions for M1i
+
+What can we conclude from model M1i about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
+In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of 187 ms)?
+
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
+The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 7 (95% CrI = [0.04, 0.25]).
+For the other bins, the 95% credible interval contained zero.
+The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin 9 (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin 10 (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin 11 (95% CrI = [-0.40, -0.03]).
+For the other bins, the 95% credible interval contained zero.
+Note that we could also have calculated hazard ratios instead of hazard differences.
+
+There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240].
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
+The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
+
+The posterior distribution of each contrast can also be summarized by considering its proportion below or above some value, like zero.
+Table 5 shows the proportion of the posterior distribution below or above zero, for each time bin and contrast.
+
+(ref:table-props-caption) Summarizing the posterior distributions of each contrast by their proportions below and above zero.
+
+(ref:table-props-note-caption) prop_below = proportion below zero: prop_above = proportion above zero.
+
+```{r table-contrasts_props, eval=T}
+ apa_table(
+   tab_contrast_props,
+   caption = "(ref:table-props-caption)",
+   note = "(ref:table-props-note-caption)",
+   placement = "H",
+#   font_size = "scriptsize",
+   longtable = TRUE,
+   repeat_header = TRUE,
+   format.args = list(
+        digits = c(0,0,2,2),
+        decimal.mark = ".", big.mark = "")
+)
+```
+
+Thus, the probability that the contrast "congruent minus blank" is larger than 0, is larger than .9 in bins 6 to 8.
+And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .9 in bins 9 to 12.
+
+If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
+
+Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
+For example, @panisWhatShapingRT2016 concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
+Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
+This effect might be masked for incongruent primes by the response competition effect.
+
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 9).
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
+In Tutorial_2a.Rmd we provide a number of model formula that should get you going.
 
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a) in the conditional accuracy model called M4_ca.
-
-To make inferences from the parameter estimates in model M4_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes on logit-ca relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
-
-Figure 8 shows one point (mean) and three highest posterior density interval (50/80/95%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
-
-(ref:descr-plot8-caption) Means and 50/80/95% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.
-
-```{r plot-prime-ca-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials.png")
-```
-
-Supplementary Table XX shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95% highest posterior density interval.
-Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
-
-```{r}
-int_ca_tab <- read_csv("../Tutorial_2_Bayesian/tables/ca_effects_intervals_table.csv")
-```
-
-(ref:int-ca-table-caption) Point and 95% highest posterior density interval estimates, and odds ratios.
-
-(ref:int-ca-table-note-caption) c500 = effect of congruent primes in trial 500; c1000 = effect of congruent primes in trial 1000; c1500 = effect of congruent primes in trial 1500; i500 = effect of incongruent primes in trial 500; i1000 = effect of incongruent primes in trial 1000; i1500 = effect of incongruent primes in trial 1500.
-
-```{r int-ca-table}
-# apa_table(
-#   int_ca_tab,
-#   caption = "(ref:int-ca-table-caption)",
-#   note = "(ref:int-ca-table-note-caption)",
-#   placement = "H",
+For illustration purposes, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i_ca.
+
+To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 10, together with point (median) and interval estimates (80% and 95% credible intervals).
+
+(ref:descr-plot10-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
+
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
+```
+
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 11), and for each participant in the data set (Figure 12).
+
+(ref:plot11-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.
+
+```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot11-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_grand.png")
+```
+
+(ref:plot12-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.
+
+```{r plot-pred-pid-effects-ca, fig.cap = "(ref:plot12-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png")
+```
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 13; grand average marginal effect).
+
+(ref:descr-plot13-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.
+
+```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot13-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png")
+```
+
+Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.
+
+```{r load-tables-ca, eval=T, echo=F}
+tab_ca_grand_ame <- read_csv("../Tutorial_2_Bayesian/tables/ca_grand_AMEs.csv")
+```
+
+(ref:table-ca-grand-ame-caption) Point (mean) and 95% credible interval summary of estimated differences in conditional accuracy, for each time bin and contrast, in the average participant.
+
+(ref:table-ca-grand-ame-note-caption) diff = difference in predicted conditional accuracy.
+
+```{r table-ca-grand-ame}
+ apa_table(
+   tab_ca_grand_ame,
+   caption = "(ref:table-ca-grand-ame-caption)",
+   note = "(ref:table-ca-grand-ame-note-caption)",
+   placement = "H",
 #   font_size = "scriptsize",
-#   longtable = TRUE,
-#   repeat_header = TRUE,
-#   format.args = list(
-#        digits = c(0, 2,2,2,2,2,13),
-#         decimal.mark = ".", big.mark = "")
-# )
-```
-
-Based on Figure 8 and Supplementary Table XX, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240], (240,280], (280,320], and (320,360], relative to the logit-ca(t) estimates in the baseline condition (blank prime; red dashed lines in Figure 8).
-For example, the odds ratio for congruent primes in bin (320,360] in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
-Incongruent primes have a negative effect on logit-ca(t) in time bins (200,240], (240,280], and (280,320] throughout the experiment, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines).
-
-To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trials 500, 1000, and 1500.
-
-(ref:descr-plot9-caption) Model-based ca(t) functions for each prime type for participant 6 in trials 500, 1000, and 1500.
-
-```{r plot-ca-subject6, fig.cap = "(ref:descr-plot9-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4_ca_effects_subject6.png")
-```
+   longtable = TRUE,
+   repeat_header = TRUE,
+   format.args = list(
+        digits = c(0,0,2,2,2),
+        decimal.mark = ".", big.mark = "")
+)
+```
+
+Based on Figure 13 and Table 6, we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
 
@@
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-Because we expected that the random effects structure of model M4 would be too complex for a freqentist approach, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+
+For illustration purposes, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-The resulting hazard model is called M3_f.
-
-In Figure 10 we compare the parameter estimates of model M3 from brm() with those of model M3_f from glmer().
-
-(ref:descr-plot10-caption) Parameter estimates for model M3 from brm() -- means and 95% quantile intervals -- and model M3_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
-
-```{r plot-comparison, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
+The resulting hazard model is called M1i_f.
+
+In Figure 14 we compare the parameter estimates of model M1i from brm() with those of model M1i_f from glmer().
+
+(ref:descr-plot14-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
+
+```{r plot-comparison, fig.cap = "(ref:descr-plot14-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png")
@@
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
-However, the random effects structure of model M3_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+Figure 14 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
+However, model M1i_f did not converge and resulted in a singular fit.
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
 But the price you pay for being able to fit more complex random effects models in a Bayesian framework is computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a and 2b took several hours to build.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
 
@@
 In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+For illustration purposes, we only fitted the effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-Again, the resulting conditional accuracy model M3_ca_f did not converge and resulted in a singular fit.
+Again, the resulting conditional accuracy model M1i_ca_f did not converge and resulted in a singular fit.
 

### 2024-11-26 manuscript/manuscript.tex
update Tutorial 2a and ms.
update ms.

@@
 While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by .44 = S(t-1=5) percent of the trials.
+For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t-1=5) = .44.
 Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.}\label{fig:plot2}
@@
 \item
-  1a. Wrangle raw data and calculate descriptive stats for one independent variable.
-\item
-  1b. Wrangle raw data and calculate descriptive stats for two independent variables.
+  1a. Wrangle raw data and calculate descriptive stats for one independent variable
+\item
+  1b. Wrangle raw data and calculate descriptive stats for two independent variables
 \item
@@
   3b. Frequentist multilevel modeling for ca(t)
+\item
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \setcounter{enumi}{3}
+  \tightlist
+  \item
+    Simulation and power analysis for planning experiments
+  \end{enumerate}
 \end{itemize}
-
-Plannng (T4) - if we get a simulation and power analysis script working, which we are happy with then we could include it here.
 
@@
 
-{\centering \includegraphics[width=12in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[width=21.33in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 as part of Tutorial 1b for those that are interested.
-
-\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models to time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-time-to-event-data}
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial 1b for those that are interested.
+
+\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-discrete-time-to-event-data}
 
@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose a specification of the effect of discrete TIME (i.e., the time bin index t) in a selected baseline condition.
-One can choose a general specification (one intercept per bin) or a functional specification, such as a polynomial one (compare model 1 with models 2, 3, and 4 below; see also part D of the supplementary material).
+Third, one has to choose whether to treat discrete TIME (i.e., the time bin index t) as a discrete or continuous predictor.
+And when you treat time as a discrete predictor, you can choose between reference coding and index coding.
 
@@
 
-In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
+In tutorial\_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
 We decided to select the analysis time window (200,600{]} and the cloglog link.
-The data is prepared as follows.
+Below, we shortly discuss three of these models.
+The person-trial-bin data set is prepared as follows.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# load person{-}trial{-}bin oriented data set}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
-\CommentTok{\# select analysis time window: (200,600] with 10 bins (time bin ranks 6 to 15)}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{)}
-\CommentTok{\# create factor condition, with "blank" as the reference level}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
-\CommentTok{\# center TIME (variable period) on bin 9, and variable trial on number 1000 and rescale; Add dummy variables for each bin.}
+\CommentTok{\# read in the file we saved in tutorial 1a}
+\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
+
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} 
+\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
+\FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+       \CommentTok{\# continuous predictor for TIME named "period\_9", centered on bin 9}
 \FunctionTok{mutate}\NormalTok{(}\AttributeTok{period\_9 =}\NormalTok{ period }\SpecialCharTok{{-}} \DecValTok{9}\NormalTok{,}
-               \AttributeTok{trial\_c =}\NormalTok{ (trial }\SpecialCharTok{{-}} \DecValTok{1000}\NormalTok{)}\SpecialCharTok{/}\DecValTok{1000}\NormalTok{,}
-               \AttributeTok{d6  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d7  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{7}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d8  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{8}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d9  =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{9}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d10 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{10}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d11 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{11}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d12 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{12}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d13 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{13}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d14 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{14}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-               \AttributeTok{d15 =} \FunctionTok{if\_else}\NormalTok{(period }\SpecialCharTok{==} \DecValTok{15}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{))}
+       \CommentTok{\# categorical predictor for TIME named "timebin" with index coding}
+       \AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period,}\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
+       \CommentTok{\# factor "condition" using reference coding, with "blank" as the reference level}
+       \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)),}
+       \CommentTok{\# categorical predictor "prime" with index coding}
+       \AttributeTok{prime =} \FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"blank"}\NormalTok{,}\DecValTok{1}\NormalTok{, }\FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"congruent"}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)),}
+       \AttributeTok{prime =} \FunctionTok{factor}\NormalTok{(prime,}\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)))}
 \end{Highlighting}
@@
 
-To get the posterior distribution of each model parameter given the data, we need to specify a prior distribution for each parameter.
-The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions on the logit and/or cloglog scales.
-
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
-The other rows in Figure 12 show prior distributions on the logit and cloglog scale that we use instead.
-
-\subsubsection{4.3.3 Model 1: A general specification of TIME, and main effects of congruency and trial number}\label{model-1-a-general-specification-of-time-and-main-effects-of-congruency-and-trial-number}
-
-When you do not want to make assumptions about the shape of the hazard function in the selected baseline condition, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one intercept per time bin.
-In this first model, we use a general specification of TIME for the selected baseline condition (blank prime), and assume that the effects of prime-target congruency and trial number are proportional and additive, and that the effect of trial number is linear.
-Before we fit model 1, we remove unnecessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M1 is specified as follows.
-
-\scriptsize
+To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
+In Tutorial\_2a.Rmd we perform some prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
+
+The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+
+\subsubsection{4.3.3 Model M0i: A general specification of TIME}\label{model-m0i-a-general-specification-of-time}
+
+When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
+In this first model, we use a general specification of TIME using index coding, and call the model ``M0i''.
+
+Before we fit model M0i, we select the necessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M0i is specified as follows.
 
@@
 \begin{Highlighting}[]
-\NormalTok{model\_M1 }\OtherTok{\textless{}{-}}
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ M1\_data,}
-       \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-\NormalTok{       event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}
-\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
-\NormalTok{                           (d6 }\SpecialCharTok{+}\NormalTok{ d7 }\SpecialCharTok{+}\NormalTok{ d8 }\SpecialCharTok{+} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ d10 }\SpecialCharTok{+}\NormalTok{ d11 }\SpecialCharTok{+}\NormalTok{ d12 }\SpecialCharTok{+}\NormalTok{ d13 }\SpecialCharTok{+}\NormalTok{ d14 }\SpecialCharTok{+}\NormalTok{ d15 }\SpecialCharTok{+}
-\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors\_M1,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }\AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }\AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
+\NormalTok{model\_M0i }\OtherTok{\textless{}{-}}                    
+   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ data\_M0i,}
+       \AttributeTok{family =} \FunctionTok{bernoulli}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
+       \AttributeTok{formula =}\NormalTok{ event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin }\SpecialCharTok{|}\NormalTok{ pid),}
+       \AttributeTok{prior =}\NormalTok{ priors\_M0i,}
+       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }
+       \AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }
+                      \AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }
+                      \AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
        \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M1"}\NormalTok{)}
+       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M0i"}\NormalTok{)}
 \end{Highlighting}
@@
 
-\normalsize
-
-After selecting the binomial family and the cloglog link, the model formula is specified.
-The fixed effects include 9 dummy variables, the explicit Intercept variable (which represents bin 9 in this example), and the main effects of prime-target congruency (variable condition) and centered trial number (variable trial\_c).
-Each of these effects is allowed to vary across individuals (variable pid).
-Estimating model M1 took about 70 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
-
-\subsubsection{4.3.4 Model 2: A polynomial specification of TIME, and main effects of congruency and trial number}\label{model-2-a-polynomial-specification-of-time-and-main-effects-of-congruency-and-trial-number}
-
-When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by using a polynomial specification of TIME.
-For our second example model, we thus use a third-order polynomial specification of TIME for the selected baseline condition (blank prime), and again assume that the effects of prime-target congruency and centered trial number are proportional and additive, and that the effect of trial number is linear.
-The model formula for model M2 looks as follows.
-
-\scriptsize
+After selecting the bernoulli family and the cloglog link, the model formula is specified.
+The specification ``0 + \ldots{}'' removes the default intercept in brm().
+The fixed effects include an intercept for each level of timebin.
+Each of these intercepts is allowed to vary across individuals (variable pid).
+Estimating model M0i took about 30 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
+
+\subsubsection{4.3.4 Model M1i: Adding the effects of prime-target congruency}\label{model-m1i-adding-the-effects-of-prime-target-congruency}
+
+Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
+In the next model, we use index coding for both TIME (variable ``timebin'') and the categorical predictor prime-target-congruency (variable ``prime''), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
+Here is the model formula of this model that we call ``M1i''.
 
@@
 \begin{Highlighting}[]
-\NormalTok{event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{+}
-\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ period\_9 }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition }\SpecialCharTok{+}\NormalTok{ trial\_c }\SpecialCharTok{|}\NormalTok{ pid),}
+\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{|}\NormalTok{ pid)}
 \end{Highlighting}
@@
 
-\normalsize
-
-Because TIME is centered on bin 9, and trial number on trial 1000, the Intercept represents the cloglog-hazard in bin 9 for the blank prime condition in trial 1000.
-Estimating model M2 took about 2.5 hours.
-
-\subsubsection{4.3.5 Model 3: A polynomial specification of TIME, and relaxing the proportionality assumption}\label{model-3-a-polynomial-specification-of-time-and-relaxing-the-proportionality-assumption}
-
-So far, we assumed that the effect of our predictors prime-target congruecy and centered trial number are the same in each time bin.
-However, the descriptive plots (e.g., Figure 4) suggest that the effect of prime-target congruency varies across time bins.
-Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
-For the third model, we thus use a third-order polynomial specification of TIME for the baseline condition (blank prime), and relax the proportionality assumption for the predictor variables prime-target congruency (variable condition) and centered trial number (variable trial\_c).
-
-\scriptsize
+Estimating model M1i took about 124 minutes.
+
+\subsubsection{4.3.4 Model M1d: A more parsimonious model}\label{model-m1d-a-more-parsimonious-model}
+
+When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
+Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable ``condition'') and treat TIME as a continuous variable (variable ``period\_9'').
+
+For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called ``M1d''.
+Here is its model formula.
 
@@
 \begin{Highlighting}[]
-\NormalTok{        event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
-\NormalTok{                            condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
-\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
-\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                            (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+} 
-\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                            condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}  
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}  
-\NormalTok{                            trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid), }
+\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9\_sq }\SpecialCharTok{+}
+\NormalTok{                   (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9\_sq }\SpecialCharTok{|}\NormalTok{ pid)}
 \end{Highlighting}
@@
 
-\normalsize
-
-Note that duplicate terms across the interaction terms in the model formula (e.g., condition) are ignored. Estimating model M3 took about 4.5 hours.
-
-\subsubsection{4.3.6 Model 4: A polynomial specification of TIME, and relaxing all three assumptions}\label{model-4-a-polynomial-specification-of-time-and-relaxing-all-three-assumptions}
-
-Based on previous work (e.g., Panis, 2020) we expect nonlinear effects of trial number on the hazard of response occurrence.
-We thus relax all three assumptions in model 4.
-We add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.
-For example, how the effect of congruent primes changes across time bins within a trial might change across the trials within an experiment.
-
-\scriptsize
+The specification ``0 + Intercept + \ldots{}'' removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
+Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition.
+
+\subsubsection{4.3.5 Compare the models.}\label{compare-the-models.}
+
+We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2018).
 
@@
 \begin{Highlighting}[]
-\NormalTok{event }\SpecialCharTok{|} \FunctionTok{trials}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}  
-\NormalTok{                           condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                           condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{  condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+} 
-\NormalTok{                                 condition}\SpecialCharTok{*}\NormalTok{period\_9}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\NormalTok{trial\_c }\SpecialCharTok{+}
-\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(trial\_c}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)   }
-\NormalTok{                                 condition}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{+} 
-\NormalTok{                                 trial\_c}\SpecialCharTok{*}\FunctionTok{I}\NormalTok{(period\_9}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ pid)}
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }
 \end{Highlighting}
@@
 
-\normalsize
-
-Again, duplicate terms across the interaction terms in the model formula are ignored. Estimating model M4 took about 8 hours.
-
-\subsubsection{4.3.7 Compare the models.}\label{compare-the-models.}
-
-We can compare the four models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2018).
-
-\scriptsize
+\begin{verbatim}
+##     model_M0i     model_M1i     model_M1d 
+## 2.462363e-183  1.000000e+00  5.739791e-17
+\end{verbatim}
 
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }
 \end{Highlighting}
@@
 \begin{verbatim}
-## model_M1 model_M2 model_M3 model_M4 
-##        0        0        0        1
+##     model_M0i     model_M1i     model_M1d 
+## 8.545948e-184  1.000000e+00  2.055767e-17
 \end{verbatim}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M1, model\_M2, model\_M3, model\_M4, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M1 model_M2 model_M3 model_M4 
-##        0        0        0        1
-\end{verbatim}
-
-\normalsize
-
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M4, and a weight of 0 to the other three simpler models.
-
-\subsubsection{4.3.8 Evaluate parameter estimates}\label{evaluate-parameter-estimates}
-
-To make inferences from the parameter estimates in model M4, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
-
-Figure 6 shows one point (mean) and three highest posterior density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500.
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other two simpler models.
+
+\subsubsection{4.3.6 Interpreting model M1i}\label{interpreting-model-m1i}
+
+To make inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_postdistr}
 
@@
 
-\caption{Means and 50/80/95\% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-effects}
+\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i.}\label{fig:plot-fixed-effects}
 \end{figure}
 
-Table 4 shows the summaries of the draws from the posterior distributions of the effects of congruent and incongurent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
-Also, by exponentiating the mean we obtain an effect size in terms of a hazard ratio.
-For simplicity and ease of presentation, we only tabulate data for a subset of the design in the main text (trial 500 for congruent and incongruent conditions).
-For the full table, see Supplementary materials or Tutorual X.
-
-Based on Figure 6 and Supplementary Table XX, we see that at the beginning of the experiment (trial 500), congruent and incongruent primes have a positive effect in time bin (200,240{]} on cloglog-hazard, relative to the cloglog-hazard estimate in the baseline condition (no prime; red striped lines in Figure 6).
-For example, the hazard ratio shows that the hazard of response occurrence for congruent primes is estimated to be 8.82 times higher than that for no-prime trials in bin (200,240{]} of trial 500.
-Incongruent primes also have a negative effect on cloglog-hazard in bins (320,360{]}, (360,400{]}, and (400,440{]}.
-For example, in bin (320,360{]}, the hazard ratio shows that the hazard of response occurrence for incongruent prime is estimated to be .34 times smaller than that for no-prime trials.
-While the early positive effects reflect responses to the prime stimulus, the later negative effect for incongruent primes likely reflects response competition between the prime-triggered response (e.g., left) and the target-triggered response (e.g., right)
-
-In the middle of the experiment (trial 1000), both congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}.
-Probably due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms early in a trial (compared to 40 ms at the beginning of the experiment) compared to the blank prime condition.
-
-Towards the end of the experiment (trial 1500), both congruent and incongruent primes have positive and negative effects.
-Positive effects are present in bins (200,240{]} and (240,280{]}.
-Incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}, and congruent primes now also have negative effects in bins (360,400{]} and (400,440{]}.
-
-These results show that the effect of prime-target congruency changes not only on the across-bin/within-trial time scale (variable period\_9), but also on the across-trial/within-experiment time scale (variable trial\_c).
-The fact that congruent primes generate negative effects for 80 ms (compared to no-prime trials) towards the end of the experiment, while incongruent primes generate negative effects for 120 ms throughout the experiment, suggests the involvement of separate cognitive processes.
-
-Panis and Schmidt (2016) distinguished between automatic response competition (bottom-up lateral inhibition between response channels), active and global inhibition (top-down nonselective response inhibition), and active and selective inhibition (top-down selective response inhibition).
-While automatic response competition can be expected to be present in the incongruent trials throughout the experiment, active and global response inhibition effects might be present in both congruent and incongruent (unmasked) prime trials.
-In other words, people learn that the prime-triggered response is premature and that they have to temporarily slow down (increase the global response threshold) in order to allow gating of the correct response to the target stimulus.
-Thus, it seems that this global inhibitory effect becomes visible in the congruent (compared to no-prime) trials towards the end of the experiment, while it might be masked by the automatic inhibitory effect of response competition in the incongruent trials.
-Interestingly, while Panis and Schmidt (2016) did not test interactions between prime-target congruency and trial number, they concluded that active (i.e., top-down) response inhibition starts around 360 ms after the onset of the second stimulus (the target stimulus in no-mask trials), which nicely coincides with the onset of the negative effect of congruent primes observed here in trial 1500.
-
-To conclude this Tutorial 2a, Figure 7 shows the model-based hazard functions for each prime type for participant 6, in trials 500, 1000, and 1500.
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6), and for each participant in the data set (Figure 7).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_subject6}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_grand}
 
@@
 
-\caption{Model-based hazard functions for each prime type for participant 6 in trials 500, 1000, and 1500.}\label{fig:plot-hazard-subject6}
+\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.}\label{fig:plot-pred-grand-effects}
 \end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_pid}
+
+}
+
+\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.}\label{fig:plot-pred-pid-effects}
+\end{figure}
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 8; grand average marginal effect) and for each participant in the data set (Figure 9; subject-specific average marginal effect).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_grand_AME}
+
+}
+
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the average participant.}\label{fig:plot-grand-ame-effects}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pid_AMEs}
+
+}
+
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for each participant.}\label{fig:plot-pid-ame-effects}
+\end{figure}
+
+Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% credible interval, for the average participant.
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} diff = difference in predicted hazard.}
+\end{TableNotes}
+
+\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table-grand-ame}Point (mean) and 95\% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.}\\
+\toprule
+contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table-grand-ame} continued}}\\
+\toprule
+contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
+\midrule
+\endhead
+congruent minus blank & 6 & 0.09 & 0.02 & 0.17\\
+congruent minus blank & 7 & 0.14 & 0.04 & 0.25\\
+congruent minus blank & 8 & 0.11 & -0.02 & 0.24\\
+congruent minus blank & 9 & 0.08 & -0.12 & 0.27\\
+congruent minus blank & 10 & -0.08 & -0.30 & 0.15\\
+congruent minus blank & 11 & -0.10 & -0.31 & 0.11\\
+congruent minus blank & 12 & -0.04 & -0.24 & 0.17\\
+congruent minus blank & 13 & -0.08 & -0.33 & 0.20\\
+congruent minus blank & 14 & -0.10 & -0.33 & 0.15\\
+congruent minus blank & 15 & -0.01 & -0.27 & 0.27\\
+incongruent minus blank & 6 & 0.09 & 0.01 & 0.21\\
+incongruent minus blank & 7 & 0.08 & -0.03 & 0.23\\
+incongruent minus blank & 8 & -0.06 & -0.17 & 0.07\\
+incongruent minus blank & 9 & -0.19 & -0.31 & -0.06\\
+incongruent minus blank & 10 & -0.27 & -0.45 & -0.04\\
+incongruent minus blank & 11 & -0.23 & -0.40 & -0.03\\
+incongruent minus blank & 12 & -0.17 & -0.33 & 0.02\\
+incongruent minus blank & 13 & -0.10 & -0.31 & 0.14\\
+incongruent minus blank & 14 & -0.06 & -0.27 & 0.15\\
+incongruent minus blank & 15 & 0.03 & -0.23 & 0.27\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\paragraph{Example conclusions for M1i}\label{example-conclusions-for-m1i}
+
+What can we conclude from model M1i about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
+In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of 187 ms)?
+
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
+The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin 6 (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin 7 (95\% CrI = {[}0.04, 0.25{]}).
+For the other bins, the 95\% credible interval contained zero.
+The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin 6 (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin 9 (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin 10 (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin 11 (95\% CrI = {[}-0.40, -0.03{]}).
+For the other bins, the 95\% credible interval contained zero.
+Note that we could also have calculated hazard ratios instead of hazard differences.
+
+There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240{]}.
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440{]}.
+The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
+
+The posterior distribution of each contrast can also be summarized by considering its proportion below or above some value, like zero.
+Table 5 shows the proportion of the posterior distribution below or above zero, for each time bin and contrast.
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} prop\_below = proportion below zero: prop\_above = proportion above zero.}
+\end{TableNotes}
+
+\begin{longtable}{llll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table-contrasts_props}Summarizing the posterior distributions of each contrast by their proportions below and above zero.}\\
+\toprule
+timebin & \multicolumn{1}{c}{contrast} & \multicolumn{1}{c}{prop\_above} & \multicolumn{1}{c}{prop\_below}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table-contrasts_props} continued}}\\
+\toprule
+timebin & \multicolumn{1}{c}{contrast} & \multicolumn{1}{c}{prop\_above} & \multicolumn{1}{c}{prop\_below}\\
+\midrule
+\endhead
+6 & congruent minus blank & 0.99 & 0.01\\
+7 & congruent minus blank & 0.99 & 0.01\\
+8 & congruent minus blank & 0.95 & 0.05\\
+9 & congruent minus blank & 0.79 & 0.21\\
+10 & congruent minus blank & 0.24 & 0.76\\
+11 & congruent minus blank & 0.15 & 0.85\\
+12 & congruent minus blank & 0.33 & 0.67\\
+13 & congruent minus blank & 0.28 & 0.72\\
+14 & congruent minus blank & 0.19 & 0.81\\
+15 & congruent minus blank & 0.47 & 0.53\\
+6 & incongruent minus blank & 0.98 & 0.02\\
+7 & incongruent minus blank & 0.92 & 0.08\\
+8 & incongruent minus blank & 0.12 & 0.88\\
+9 & incongruent minus blank & 0.00 & 1.00\\
+10 & incongruent minus blank & 0.01 & 0.99\\
+11 & incongruent minus blank & 0.02 & 0.98\\
+12 & incongruent minus blank & 0.04 & 0.96\\
+13 & incongruent minus blank & 0.20 & 0.80\\
+14 & incongruent minus blank & 0.27 & 0.73\\
+15 & incongruent minus blank & 0.58 & 0.42\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+Thus, the probability that the contrast ``congruent minus blank'' is larger than 0, is larger than .9 in bins 6 to 8.
+And the probability that the contrast ``incongruent minus blank'' is smaller than 0, is larger than .9 in bins 9 to 12.
+
+If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
+
+Future studies could (a) increase the number of participants to estimate the proportion of ``dippers'' in the subject population, and/or (b) try to explain why this dip occurs.
+For example, Panis and Schmidt (2016) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
+Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after \textasciitilde360 ms, and towards the end of the experiment.
+This effect might be masked for incongruent primes by the response competition effect.
+
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 9).
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
+In Tutorial\_2a.Rmd we provide a number of model formula that should get you going.
 
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M4 (see Tutorial 2a) in the conditional accuracy model called M4\_ca.
-
-To make inferences from the parameter estimates in model M4\_ca, we summarize the draws from the posterior distributions of the effects of congruent and incongruent primes on logit-ca relative to the blank prime condition, in each time bin for trial numbers 500, 1000, and 1500, in terms of point and interval estimates.
-
-Figure 8 shows one point (mean) and three highest posterior density interval (50/80/95\%) estimates for the effects of congruent and incongruent primes relative to neutral primes on logit-ca, for each time bin in trial numbers 500, 1000, and 1500.
+For illustration purposes, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i\_ca.
+
+To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 10, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_postdistr}
 
@@
 
-\caption{Means and 50/80/95\% highest posterior density intervals of the draws from the posterior distributions representing the effect of congruent and incongruent primes relative to neutral primes in each bin for trial numbers 500, 1000, and 1500.}\label{fig:plot-prime-ca-effects}
+\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca.}\label{fig:plot-ca-fixed-effects}
 \end{figure}
 
-Supplementary Table XX shows the summaries of the draws from the posterior distributions of the effects of congruent and incongruent primes relative to the blank prime condition in trials 500, 1000, and 1500, in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% highest posterior density interval.
-Also, by exponentiating the mean we obtain an effect size in terms of an odds ratio.
-
-Based on Figure 8 and Supplementary Table XX, we see that throughout the experiment (trials 500, 1000, and 1500), congruent primes have a positive effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the logit-ca(t) estimates in the baseline condition (blank prime; red dashed lines in Figure 8).
-For example, the odds ratio for congruent primes in bin (320,360{]} in trial 500 shows that the odds of a correct response are estimated to be 8.89 times higher than the odds of a correct response when there is no prime.
-Incongruent primes have a negative effect on logit-ca(t) in time bins (200,240{]}, (240,280{]}, and (280,320{]} throughout the experiment, relative to the logit-ca(t) estimates in the baseline condition (no prime; red striped lines).
-
-To conclude this Tutorial 2b, Figure 9 shows the model-based ca(t) functions for each prime type for participant 6, in trials 500, 1000, and 1500.
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 11), and for each participant in the data set (Figure 12).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4_ca_effects_subject6}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_grand}
 
@@
 
-\caption{Model-based ca(t) functions for each prime type for participant 6 in trials 500, 1000, and 1500.}\label{fig:plot-ca-subject6}
+\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.}\label{fig:plot-pred-grand-effects-ca}
 \end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_pid}
+
+}
+
+\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.}\label{fig:plot-pred-pid-effects-ca}
+\end{figure}
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 13; grand average marginal effect).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_grand_AME}
+
+}
+
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.}\label{fig:plot-ca-grand-ame-effects}
+\end{figure}
+
+Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% credible interval, for the average participant.
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} diff = difference in predicted conditional accuracy.}
+\end{TableNotes}
+
+\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table-ca-grand-ame}Point (mean) and 95\% credible interval summary of estimated differences in conditional accuracy, for each time bin and contrast, in the average participant.}\\
+\toprule
+contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff\_ca} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table-ca-grand-ame} continued}}\\
+\toprule
+contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff\_ca} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
+\midrule
+\endhead
+congruent minus blank & 6 & 0.36 & 0.06 & 0.74\\
+congruent minus blank & 7 & 0.32 & 0.13 & 0.57\\
+congruent minus blank & 8 & 0.16 & 0.06 & 0.31\\
+congruent minus blank & 9 & 0.07 & 0.02 & 0.17\\
+congruent minus blank & 10 & 0.00 & -0.03 & 0.03\\
+congruent minus blank & 11 & -0.03 & -0.11 & 0.02\\
+congruent minus blank & 12 & -0.02 & -0.10 & 0.03\\
+congruent minus blank & 13 & -0.04 & -0.20 & 0.06\\
+congruent minus blank & 14 & -0.20 & -0.59 & 0.07\\
+congruent minus blank & 15 & -0.06 & -0.49 & 0.28\\
+incongruent minus blank & 6 & -0.61 & -0.91 & -0.22\\
+incongruent minus blank & 7 & -0.64 & -0.84 & -0.38\\
+incongruent minus blank & 8 & -0.70 & -0.86 & -0.40\\
+incongruent minus blank & 9 & -0.21 & -0.59 & 0.04\\
+incongruent minus blank & 10 & -0.03 & -0.15 & 0.02\\
+incongruent minus blank & 11 & -0.02 & -0.07 & 0.02\\
+incongruent minus blank & 12 & 0.00 & -0.04 & 0.04\\
+incongruent minus blank & 13 & 0.01 & -0.06 & 0.08\\
+incongruent minus blank & 14 & 0.02 & -0.09 & 0.18\\
+incongruent minus blank & 15 & -0.09 & -0.39 & 0.23\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+Based on Figure 13 and Table 6, we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
 
@@
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-Because we expected that the random effects structure of model M4 would be too complex for a freqentist approach, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+
+For illustration purposes, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting hazard model is called M3\_f.
-
-In Figure 10 we compare the parameter estimates of model M3 from brm() with those of model M3\_f from glmer().
+The resulting hazard model is called M1i\_f.
+
+In Figure 14 we compare the parameter estimates of model M1i from brm() with those of model M1i\_f from glmer().
 
@@
 
-\caption{Parameter estimates for model M3 from brm() -- means and 95\% quantile intervals -- and model M3\_f from glmer() -- maximum likelihood estimates and 95\% confidence intervals.}\label{fig:plot-comparison}
+\caption{Parameter estimates for model M1i from brm() -- means and 95\% credible intervals -- and model M1i\_f from glmer() -- maximum likelihood estimates and 95\% confidence intervals.}\label{fig:plot-comparison}
 \end{figure}
 
-Figure 10 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
-However, the random effects structure of model M3\_f was already too complex for the frequentist model as it did not converge and resulted in a singular fit.
+Figure 14 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
+However, model M1i\_f did not converge and resulted in a singular fit.
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
 But the price you pay for being able to fit more complex random effects models in a Bayesian framework is computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a and 2b took several hours to build.
+In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
 
@@
 In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-For illustration purposes, we only fitted the effects from model M3 (see Tutorial 2a) using the function glmer() from the R package lme4.
+For illustration purposes, we only fitted the effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-Again, the resulting conditional accuracy model M3\_ca\_f did not converge and resulted in a singular fit.
+Again, the resulting conditional accuracy model M1i\_ca\_f did not converge and resulted in a singular fit.
 

### 2024-11-26 manuscript/manuscript.Rmd
upudate ms.

@@
 To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
 In their first experiment, @panisWhatShapingRT2016 presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 
-\scriptsize
+\footnotesize
 ```{r setup-data-tut1, echo=TRUE}
-data_wr <- read_csv("../Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
+data_wr<-read_csv("../Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
 colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
@@
   mutate(condition = condition + 1, # original levels were 0, 1, 2.
-         condition = factor(condition, levels=c(1,2,3), labels=c("blank","congruent","incongruent")))
+         condition = factor(condition, 
+                            levels=c(1,2,3), 
+                            labels=c("blank","congruent","incongruent")))
 ```
@@
 
-\scriptsize
+\footnotesize
 ```{r nest-apply-functions-tut1, echo = TRUE, eval=F}
@@
 data_final <- data_nested %>% 
-  mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
-  mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin data set
-  mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca(t)
-  mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca(t)
-  mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca(t)
-  mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
+  # ! user input: censoring time, and bin width
+  mutate(censored = map(data, censor, 600, 40)) %>%   
+  # create person-trial-bin data set
+  mutate(ptb_data = map(censored, ptb)) %>%          
+  # create life tables without ca(t)
+  mutate(lifetable = map(ptb_data, setup_lt)) %>%   
+  # calculate ca(t)
+  mutate(condacc = map(censored, calc_ca)) %>%      
+  # create life tables with ca(t)
+  mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    
+  # create plots 
+  mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  
 ```
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant. 
-
 In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
@@
 
-\scriptsize
+\footnotesize
 ```{r wrangle-data, eval=F, echo=T}
@@
 
+\footnotesize
 ```{r show-weights, echo=T, eval=T}
@@
 ```
+\normalsize
 

### 2024-11-26 manuscript/manuscript.tex
upudate ms.

@@
 To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016).
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
 In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 
-\scriptsize
+\footnotesize
 
@@
 \begin{Highlighting}[]
-\NormalTok{data\_wr }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/DataExp1\_6subjects\_wrangled.csv"}\NormalTok{)}
+\NormalTok{data\_wr}\OtherTok{\textless{}{-}}\FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/DataExp1\_6subjects\_wrangled.csv"}\NormalTok{)}
 \FunctionTok{colnames}\NormalTok{(data\_wr) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"pid"}\NormalTok{,}\StringTok{"bl"}\NormalTok{,}\StringTok{"tr"}\NormalTok{,}\StringTok{"condition"}\NormalTok{,}\StringTok{"resp"}\NormalTok{,}\StringTok{"acc"}\NormalTok{,}\StringTok{"rt"}\NormalTok{,}\StringTok{"trial"}\NormalTok{) }
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =}\NormalTok{ condition }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, }\CommentTok{\# original levels were 0, 1, 2.}
-         \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{), }\AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{,}\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
+         \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }
+                            \AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{), }
+                            \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{,}\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
 \end{Highlighting}
@@
 \NormalTok{data\_final }\OtherTok{\textless{}{-}}\NormalTok{ data\_nested }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{censored  =} \FunctionTok{map}\NormalTok{(data, censor, }\DecValTok{600}\NormalTok{, }\DecValTok{40}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}   \CommentTok{\# !
-user input: censoring time, and bin width}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data  =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}           \CommentTok{\# create person{-}trial{-}bin data set}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}      \CommentTok{\# create life tables without ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condacc   =} \FunctionTok{map}\NormalTok{(censored, calc\_ca)) }\SpecialCharTok{\%\textgreater{}\%}       \CommentTok{\# calculate ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable\_ca =} \FunctionTok{map2}\NormalTok{(lifetable, condacc, join\_lt\_ca)) }\SpecialCharTok{\%\textgreater{}\%}    \CommentTok{\# create life tables with ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{plot      =} \FunctionTok{map2}\NormalTok{(}\AttributeTok{.x =}\NormalTok{ lifetable\_ca, }\AttributeTok{.y =}\NormalTok{ pid, plot\_eha,}\DecValTok{1}\NormalTok{))  }\CommentTok{\# create plots }
+  \CommentTok{\# ! user input: censoring time, and bin width}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{censored =} \FunctionTok{map}\NormalTok{(data, censor, }\DecValTok{600}\NormalTok{, }\DecValTok{40}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# create person{-}trial{-}bin data set}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}          
+  \CommentTok{\# create life tables without ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}   
+  \CommentTok{\# calculate ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condacc =} \FunctionTok{map}\NormalTok{(censored, calc\_ca)) }\SpecialCharTok{\%\textgreater{}\%}      
+  \CommentTok{\# create life tables with ca(t)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable\_ca =} \FunctionTok{map2}\NormalTok{(lifetable, condacc, join\_lt\_ca)) }\SpecialCharTok{\%\textgreater{}\%}    
+  \CommentTok{\# create plots }
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{plot =} \FunctionTok{map2}\NormalTok{(}\AttributeTok{.x =}\NormalTok{ lifetable\_ca, }\AttributeTok{.y =}\NormalTok{ pid, plot\_eha,}\DecValTok{1}\NormalTok{))  }
 \end{Highlighting}
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-
-In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone may be guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
 
@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose whether to treat discrete TIME (i.e., the time bin index t) as a discrete or continuous predictor.
-And when you treat time as a discrete predictor, you can choose between reference coding and index coding.
-
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
+And when you treat time as a discrete predictor, you can choose between reference coding and index coding.\\
 In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set (Allison, 2010).
@@
 
-The middle column of Figure 12 in part E of the supplementary material shows seven examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 12 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
-
-\subsubsection{4.3.3 Model M0i: A general specification of TIME}\label{model-m0i-a-general-specification-of-time}
+The middle column of Figure 16 in part E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in supplementary Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+
+\subsubsection{4.3.3 Model M0i: A null model with index coding}\label{model-m0i-a-null-model-with-index-coding}
 
 When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
-In this first model, we use a general specification of TIME using index coding, and call the model ``M0i''.
+In this first model, we use a general specification of TIME using index coding, and do not include experimental predictors.
+We call this model ``M0i''.
 
@@
 The specification ``0 + Intercept + \ldots{}'' removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
+The variable period\_9\_sq is a squared version of period\_9.
 Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 145 minutes to run.
 
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 ## model_M0i model_M1i model_M1d 
-## 2.462363e-183  1.000000e+00  5.739791e-17
+##         0         1         0
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }
 \end{Highlighting}
@@
 ## model_M0i model_M1i model_M1d 
-## 8.545948e-184  1.000000e+00  2.055767e-17
+##         0         1         0
 \end{verbatim}
@@
 \item
-  plot\_eha(df,subj,haz\_yaxis) : create plots of the discrete-time functions (1 independent variable)
-\item
-  plot\_eha\_2IV(df,subj,haz\_yaxis) : create plots of the discrete-time functions (2 independent variables)
-\item
-  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for data aggregated across participants (1 independent variable)
+  plot\_eha(df,subj,haz\_yaxis=1,first\_bin\_shown=1,aggregated\_data=F,Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+\item
+  plot\_eha\_2IV(df,subj,haz\_yaxis=1,first\_bin\_shown=1,aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
 \end{itemize}
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 11.
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 15.
 
@@
 
-\caption{Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
+\caption{Prior distributions for the Intercept on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
 \end{figure}

### 2024-11-26 notes/manuscript_notes.Rmd
upudate ms.

@@
 
--> I added a content list at the start of each Tutorial_x.Rmd file.
+-> I added a toc for each tutorial.
 
@@
 And as we disucssed before, we need to be mindful that folks know exactly what eaxh panels add on its own and/or in combination with the other panels.
+
+-> I think shading is not necessary.
 
@@
 
--> to do. Which intervals should we plot for the frequentist approach?
+-> Done.
 
@@
 - combine tables showing person-trial and person-trial-period data sets
-- REF for regression (line 145) 
-- update section 3 and add REFs (line 200-215)
-
 - REF for time-to-death (line 127 in ms.) : Done
 - REF for reviews and tutorials on EHA (line 140) : Done
-- change paragraph (delete it and refer to suppl. material?), and add error bars to Figure 2? (line 162): Done (no error bars added, but I refer to suppl. material)
 - REF sampling bias (line 332): Instead of adding a REF, I rephrased the sentence.
@@
 - I removed the repetition of the title on the intro page. This always annoys me!
+
+-> great!
 
@@
 
--> I expected problems with fitting the very complex model M4. I added this info.
+-> I expected problems with fitting the very complex model M4. 
+I want to keep this tutorial short.
 
@@
 
--> Done.  todo...
+-> Done.
 
@@
 
--> Done.  todo....
+-> Done.
 
@@
 
+-> Done.
+
 And I also wonder. Should this even be in the main text, given that it is in the tutorial?
@@
 
+-> init="0" avoids the chain initialization problems I experienced.
+
 ## table 4 and 5 ##
@@
 
-->  todo
+->  Done.
 
@@
 
-->  todo
+->  Table has changed...
 
@@
 
--> todo
+-> Done.
 
@@
 
--> Explanation for large values is given in the table captions.  todo
+-> Tables have changed in the new ms.
 
@@
 - model_M1.
+
 --The use of 0 and Intercept needs explaining.
 
--> todo
+-> Done.
 
 --adapt_delta is high. I rarely have to go above 0.99. Does the model not look right if this is set lower? This could be a sign that the model is not a suitable structure.
+
+-> .999 is not unusual for generalized linear mixed models.
 
@@
 
--> todo.
+-> todo!!!!!!!
 
@@
 
--> I prefere option 1, and added a new subsection... todo
+-> I added a new subsection.
 
@@
 Are these posterior predictions? If so, they need a little explanation.
+
+-> Done.
 
@@
 
+-> Done.
+
 Also, I would be explicit with sizing when saving figures, if you don't so this already.
@@
 ```
+
+-> Done.
 
@@
 
+-> Done.
+
 ## figure legends ##
@@
 in some cases, much more information is needed. e.g., Figure 7 and 9. Folks need a guide.
+
+-> I have to check for all Figures... todo
 
@@
 
+-> Figure 9 has changed.
+
 Figure 10 can be flipped such that the axes are opposite. e.g., flip x and y. then the text will be easier to read and plots easier to understand, in terms of numerical values.
+
+-> Done
 
@@
 
-2.
+# comments Sven 26 november
+
+Still to do:
+
+- add refs for some websites I consulted
+- check the references in the ms to various parts of the supplementary material
+- check legends of all Figures, and add more info if needed
+
+- make loo plots (I did not find code to do that. Do you have a reference for such code?)
+
+- add info on simulation and power analysis (section 4.4)
+
+- use timebin or time bins?
+
+- use multi-level or multilevel ?

### 2024-11-26 Tutorial_2a.Rmd
update tuts and suppl. mat.

@@
 
-cloglog_prob <- cloglog(probability)
-logit_prob <- logit(probability)
-dataplot <- cbind(probability,cloglog_prob,logit_prob)
+cloglog <- cloglog(probability)
+logit <- logit(probability)
+dataplot <- cbind(probability,cloglog,logit) %>%
+  as_tibble() %>%
+  pivot_longer(cloglog:logit, names_to = "link", values_to = "value")
 
@@
   geom_hline(yintercept=0, color="grey") +
-  geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
-  geom_line(data=dataplot,aes(y=cloglog_prob,x=probability,colour="cloglog"),linewidth=1) +
-  scale_color_manual(name = "Link function:", values = c("logit" = "darkblue", "cloglog" = "red")) +
+  geom_line(data=dataplot,aes(y=value,x=probability,colour=link),linewidth=1) +
+  geom_line(data=dataplot,aes(y=value,x=probability,colour=link),linewidth=1) +
+  scale_color_brewer(palette = "Dark2") +
   geom_vline(xintercept = logistic(0), linetype="dotted", linewidth = 0.3) +   
@@
   annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90,size=4) +
-  scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
   labs(x = "Probability",
         y = "logit or cloglog scale") +
-  theme(panel.grid = element_blank(),
-        axis.text=element_text(size=16),
-        axis.title=element_text(size=18),
-      #  legend.position="top",
-        legend.text = element_text(size=18),
-      legend.position = "inside",
-        legend.position.inside=c(.2,.75),
-      legend.background = element_rect(fill="white") )
+  theme(panel.grid = element_blank())
 ```

### 2024-11-26 Tutorial_3a.Rmd
update tuts and suppl. mat.

@@
 est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
-              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
+              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(3,7,8)],
               effectname = names(fixef(model_M1i_f)),
               position = 1:30) 
-
-colnames(est)
+est
 ```
@@
 ```{r plot-estimates}
-cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
+# check hex codes for color paletter
+mypal <- brewer.pal(8,"Dark2")
+
+# specify colors and shapes
+cols    <- c( "c1" = mypal[1], "c2" = mypal[2])
 shapes  <- c("s1" = 16, "s2" = 17)
 
+# plot
 p1 <- ggplot(data = est, aes(x = position)) +
@@
                               labels = c("brm", "glmer")) +
-  coord_flip() #+
- #theme(panel.grid = element_blank())
-         #legend.position = "top",
-         #axis.text.x = element_text(size=11,angle=90),
-         #axis.text.y = element_text(size=13),
-         #axis.title.x = element_text(size = 14),
-         #axis.title.y = element_text(size = 14),
-         #legend.text=element_text(size=13)) 
+  coord_flip() 
 p1

### 2024-11-26 manuscript/SupplementaryMaterial.Rmd
update tuts and suppl. mat.

@@
 * extract_median_2IV(df) : estimate quantiles S(t)~.50~ (2 independent variables)
-* plot_eha(df,subj,haz_yaxis) : create plots of the discrete-time functions (1 independent variable)
-* plot_eha_2IV(df,subj,haz_yaxis) : create plots of the discrete-time functions (2 independent variables)
-* plot_eha_agg(df,subj,haz_yaxis) : create 1 plot for data aggregated across participants (1 independent variable)
+* plot_eha(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+* plot_eha_2IV(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,   Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 11.
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 15.
 
-(ref:descr-plot11-caption) The logit and cloglog link functions.
+(ref:descr-plot15-caption) The logit and cloglog link functions.
 
-```{r plot-link-functions, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'}
+```{r plot-link-functions, fig.cap = "(ref:descr-plot15-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
@@
 
-(ref:descr-plot12-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+(ref:descr-plot16-caption) Prior distributions for the Intercept on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
 
-```{r plot-priors, fig.cap = "(ref:descr-plot12-caption)", out.width='80%'}
+```{r plot-priors, fig.cap = "(ref:descr-plot16-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")

### 2024-11-26 custom_functions.R
update color figures and custum-functions

@@ -111,6 +111,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
   # Nsubj: The number of individuals before aggregation. Only used to set the title of the plot.
   
   library(patchwork) # to combine figures
+  library(RColorBrewer) # color palette 
   cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
   binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
   median_period <- extract_median(df)
@@ -136,6 +137,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                size=1) + 
     geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), 
                    show.legend = F) +
+    scale_color_brewer(palette = "Dark2") +
     scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
@@ -165,6 +167,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
               data = data_medians, 
               linetype = 3, 
               show.legend = F) +
+    scale_color_brewer(palette = "Dark2") +
     scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
@@ -183,9 +186,11 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                size=1) +
     geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), 
                    show.legend = F) +
+    scale_color_brewer(palette = "Dark2") +
     scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
+
     scale_y_continuous(limits = c(0,1)) +
     labs(x = "bin endpoints (ms)", 
          y = "ca(t)",
[…]

### 2024-11-26 Tutorial_2a.Rmd
update html files (and 2 tuts)

@@
 
-```{r}
+```{r inspect-M1d}
 model_M1d <- readRDS("Tutorial_2_Bayesian/models/model_M1d.rds")
@@
 
-```{r}
+```{r nonlinearsyntax}
 formula = bf(event ~ 0 + a + b * period_9 + c * period_9_sq,
@@
 
-```{r}
+```{r model-formula-1}
 event ~ 0 + timebin:prime:stage  + (0 + timebin:prime:stage | pid)
@@
 
-```{r}
+```{r model-formula-2}
 bf(event ~ 0 + a + b * trial_c + c * trial_c_sq,
@@
 
-```{r}
+```{r model-formula-3}
  bf(event ~ 0 + Intercept + 
@@
 
-```{r}
+```{r epred-grand}
 dat_M1i <- as_tibble(model_M1i$data)
@@
 
-```{r}
+```{r epred-grand-summary}
 epreds_grand %>% 
@@
 
-```{r}
+```{r epred-grand-plot}
  ggplot(epreds_grand, aes(x=timebin, y=.epred, 
@@
 
-```{r save-pred-grand, eval=F}
+```{r save-epred-grand, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_grand.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r}
+```{r epred-grand-post}
 tidy_fixed %>%
@@
 
-```{r}
+```{r epred-pid}
 epreds_pid <- dat_M1i %>% 
@@
 
-```{r}
+```{r epred-pid-summary}
 epreds_pid %>% 
@@
 
-```{r}
+```{r epred-pid-plot}
  ggplot(epreds_pid, aes(x=timebin, y=.epred, 
@@
 
-```{r save-pred-pid, eval=F}
+```{r save-epred-pid, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r}
+```{r epred-new-pid}
 epreds_pid_new <- dat_M1i %>% 
@@
 
-```{r}
+```{r epred-new-pid-summary}
 epreds_pid_new %>% 
@@
 
-```{r}
+```{r epred-new-pid-plot}
  ggplot(epreds_pid_new, aes(x=timebin, y=.epred, 
@@
 
-```{r save-pred-pid-new, eval=F}
+```{r save-epred-pid-new, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
@@
 
-```{r}
+```{r epred-grand-diff}
 epreds_grand_diffs <- epreds_grand %>%
@@
 
-```{r}
+```{r epred-grand-diff-summary}
 table <- epreds_grand_diffs %>%
@@
 
-```{r}
+```{r epred-grand-diff-plot}
  ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
@@
 
-```{r}
+```{r epred-grand-ame-post}
 post %>% # 8000 x 30 = 240000
@@
 
-```{r}
+```{r epred-pid-diff}
 epreds_pid_diffs <- epreds_pid %>%
@@
 
-```{r}
+```{r epred-pid-diff-summary}
 table <- epreds_pid_diffs %>% 
@@
 
-```{r}
+```{r epred-pid-diff-plot}
  ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
@@
 epreds_pid_diffs_av <- epreds_pid_diffs %>% group_by(timebin,contrast,.draw) %>%
-  mutate(average_condmeans = mean(pred_haz))
+  mutate(average_condmeans = mean(diff))
 ```
@@
 ```{r}
- ggplot(epreds_pid_diffs_av, aes(x=timebin, y=pred_haz, 
+ ggplot(epreds_pid_diffs_av, aes(x=timebin, y=diff, 
                        fill=contrast)) +
@@
 
-```{r}
+```{r epred-new-pid-diff}
 epreds_pid_new_diffs <- epreds_pid_new %>%
@@
 
-```{r}
+```{r epred-new-pid-diff-summary}
 epreds_pid_new_diffs %>% 
@@
 
-```{r}
- ggplot(epreds_pid_new_diffs, aes(x=timebin, y=pred_haz, 
+```{r epred-new-pid-diff-plot}
+ ggplot(epreds_pid_new_diffs, aes(x=timebin, y=diff, 
                        fill=contrast, color=contrast)) +
@@
 
-```{r}
+```{r read-model-fit-skewN}
 fit_skewN <- readRDS("Tutorial_2_Bayesian/models/fit_skewN.rds")

### 2024-11-26 notes/manuscript_notes.Rmd
add note

@@
 
+- add figures of the prior predictive checks in Tutorial_2a.html
+
 - add refs for some websites I consulted
+
 - check the references in the ms to various parts of the supplementary material
+
 - check legends of all Figures, and add more info if needed

### 2024-11-26 Tutorial_2a.Rmd
add figures of PPC for 2 models

@@
 The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent).
-We perform some prior predictive checks to infer the prior distributions we will employ.
+We perform some prior predictive checks to determine the prior distributions we will employ.
 We compare models using WAIC and LOO, and interpret parameter estimates for one model.
@@
 
-Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflects our prior beliefs.
+Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflects our prior beliefs.
 First, sample the prior distributions using sample_prior="only".
@@
 
-```{r ppc-manual}
+```{r ppc-manual, eval=T}
 # prior predictive check for 1 time bin
@@
 
-The shape of the prior predictive distribution looks as expected (compare with Figure X in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
+```{r save-ppc, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png", width = 10, height = 8, dpi = 800)
+```
+
+The shape of the prior predictive distribution looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
 
@@
 
+```{r save-ppc-ten-bins-six-subjects, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0i_ppc.png", width = 10, height = 8, dpi = 800)
+```
+
 Now we can fit model M0i to get the posterior distributions.
@@
 summary(model_M0i)
-formula(model_M0i)
 plot(model_M0i)
@@
 
-Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflect our prior beliefs. 
+Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs. 
 The following code allows to change the coefficients and generate (cloglog-)hazard functions.
-Once you have found the coefficients for period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to estimate a mean and standard deviation for each coefficient.
+Once you have found the coefficients for the intercept, period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to determine a suited mean and standard deviation for each coefficient.
 
@@
 
-First, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random variables while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
+Next, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
 
@@
 
-```{r inspect-M0c-prior}
+```{r load-M0c-prior}
 model_M0c_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior.rds")
@@
 
-Generate prior predictive check for each participant.
+Next, generate prior predictive check for each participant.
 
@@
 
+```{r save-ppc-M1c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc.png", width = 10, height = 8, dpi = 800)
+```
+
 A wide range of differently shaped hazard functions can be observed, and early bins are unlikely to have high hazard values, consistent with our prior beliefs.
@@
 
-Now increase the variance again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.
+Now increase the standard deviation again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.
 

### 2024-12-01 Tutorial_2a.Rmd
correct small error
update readme and tut 2a

@@
 We compare models using WAIC and LOO, and interpret parameter estimates for one model.
-We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
+We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
 
@@
 
-Prepare the data file, and specify priors. The skew_normal prior is set for each grand intercept on the cloglog scale, and should reflect our prior beliefs.
+Prepare the data file, and specify priors.
+The skew_normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that hazard values above .5 are less likely than values below .5.
 
@@
 
-Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflects our prior beliefs.
+### Intermezzo: Prior Predictive Check for M0i.
+
+Perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
 First, sample the prior distributions using sample_prior="only".
@@
 
-Now we can fit model M0i to get the posterior distributions.
+After our prior predictive check, we can fit model M0i to get the posterior distributions.
 
@@
 
-```{r save-ppc-M1c, eval=F}
+```{r save-ppc-M0c, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M0c_ppc.png", width = 10, height = 8, dpi = 800)
 ```
+
+#
+Suppose we believe apriori that hazard functions are flat.
+We specify our new priors:
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc2 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .001)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior, eval=F}
+plan(multicore)
+
+model_M0c_prior2 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc2,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior2")
+```
+
+```{r ppc-model-M0c-prior, eval=F}
+# Generate prior predictive hazard functions for 6 participants
+# Specify
+N_obs = 100 # increase for smoother results
+set.seed(1)
+
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# prepare columns of data set with predictions
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# extract predictions
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0c_prior2, 
+                                    newdata = newdata_prior, 
+                                    summary=F, 
+                                    ndraws=200) %>% 
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot 200 draws per participant, highlighting three
+sel = sample(1:200,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <200), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   # highlight
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
+             color="red", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
+            color="green", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+```
+
+Or, suppose we believe that hazard increases with time; More specifically, we assume that cloglog-hazard increases linearly over time.
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc3 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.3,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .001)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior, eval=F}
+plan(multicore)
+
+model_M0c_prior3 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc3,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
+```
+
+```{r ppc-model-M0c-prior, eval=F}
+# Generate prior predictive hazard functions for 6 participants
+# Specify
+N_obs = 300 # increase for smoother results
+set.seed(1)
+
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# prepare columns of data set with predictions
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# extract predictions
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0c_prior3, 
+                                    newdata = newdata_prior, 
+                                    summary=F, 
+                                    ndraws=20) %>% 
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot 200 draws per participant, highlighting three
+sel = sample(1:20,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <= 20), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   # highlight
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
+             color="red", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
+            color="green", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+```
+
+Or, suppose we believe that hazard first increases with time, and then decreases;
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc3 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.6,.2)", class = "b", coef = "period_9"), 
+  set_prior("normal(-0,.09)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .1)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior, eval=F}
+plan(multicore)
+
+model_M0c_prior3 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc3,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
+```
+
+```{r ppc-model-M0c-prior, eval=F}
+# Generate prior predictive hazard functions for 6 participants
+# Specify
+N_obs = 100 # increase for smoother results
+N_samples = 8000
+set.seed(1)
+
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# prepare columns of data set with predictions
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# extract predictions
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0c_prior3, 
+                                    newdata = newdata_prior, 
+                                    summary=F, 
+                                    ndraws=N_samples) %>% 
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot 200 draws per participant, highlighting three
+sel = sample(1:N_samples,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <= N_samples), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   # highlight
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
+             color="red", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
+            color="green", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+```
+
+Or, check settings.
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc3 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .1)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior, eval=F}
+plan(multicore)
+
+model_M0c_prior3 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc3,
+       chains = 4, cores = 4, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
+```
+
+```{r ppc-model-M0c-prior, eval=F}
+# Generate prior predictive hazard functions for 6 participants
+# Specify
+N_obs = 100 # increase for smoother results
+N_samples = 8000
+set.seed(1)
+
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# prepare columns of data set with predictions
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# extract predictions
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0c_prior3, 
+                                    newdata = newdata_prior, 
+                                    summary=F, 
+                                    ndraws=N_samples) %>% 
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot 200 draws per participant, highlighting three
+sel = sample(1:N_samples,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <= N_samples), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   # highlight
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
+             color="red", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
+            color="green", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+```
+
+#
 

### 2024-12-01 manuscript/manuscript.Rmd
update ms.

@@
 Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
-We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
 Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
 
-  <!-- https://tinyurl.com/ybremelq -->
 keywords          : "response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology"
@@
 ```{r setup, include = FALSE}
-# Tutorials 1 and 4
 pkg1 <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")
@@
 
-# Tutorial 2
 pkg2 <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
@@
 
-# Tutorial 3
 pkg3 <- c("lme4", "nlme")
@@
 # R references
-r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation...
-my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) ## 
+r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
+my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) 
 ```
@@
 
-```{r plot-settings}
-## theme settings for ggplot
-theme_set(
-  theme_bw() +
-    theme(text = element_text(size = 18), 
-          title = element_text(size = 18),
-          legend.position = "bottom")
-)
-
-## Set the amount of dodge in figures
-pd <- position_dodge(0.7)
-pd2 <- position_dodge(1)
-```
-
 ```{r global-chunk-settings}
@@
 
-In experimental psychology, it is standard practice to analyse reaction times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
+In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
 Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
@@
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
-Insets show mean reaction time per condition.
+Insets show mean response time per condition.
 
@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @williammatthewmakehamLawMortalityConstruction1860].
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @halleyVIEstimateDegrees1997; @williammatthewmakehamLawMortalityConstruction1860].
 The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020].
-Here, in the current paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+
+## 1.2 Aims and structure of the paper
+
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
 survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
 We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
-
-## 1.2 Aims and structure of the paper
-
-In this paper, we focus on discrete-time EHA.
+Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+
 We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
-Indeed, our primary aim here is to introduce a set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
+Indeed, our primary aim here is to introduce the set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
 
@@
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja() package [@pap2024], which makes it computationally reproducible, in terms of the underlying data and figures.
+Finally, the manuscript itself is written in R using the papaja package [@pap2024], which makes it computationally reproducible, in terms of the underlying data and figures.
 
@@
 In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
-We discuss possible link functions, and plot the model-based effects of our predictors of interest.
+We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
 In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
+
+In Tutorial 3a, we shortly illustrate how to fit similar multilevel regression models for RT data in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to EHA.
@@
 
-2. define time point zero (e.g., target stimulus onset, fixation onset);
+2. define time point zero (e.g., target stimulus onset, fixation onset, etc.);
 
@@
 While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t-1=5) = .44.
+For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t=5) = .44.
 Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.
@@
 
-2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences that mean-average comparisons may conceal (Townsend, 1990).
+2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
 
@@
 EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with longer RTs are not discarded, but instead contribute to the risk set in each time bin.
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see below).
 
@@
 
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task.
+In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a discrimination task.
 For example, a task may involve choosing between two response options with only one of them being correct.
@@
 
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
-Between 500 and 600 ms the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
-For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
-
-This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT across trials.
+During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
+
+This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT and accuracy across trials.
 Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Bin width equals 100 ms.
 
@@
 
-Descriptive, data science workflow.
 We perform data wrangling following tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
@@
 We used `r my_r_citation$r` for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
-
-`r my_r_citation$pkgs`
-
-<!-- This creates the footnote -->
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
+
+`r my_r_citation$pkgs` <!-- create footnote -->
 
@@
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes.
+First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes, using a life table.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+
 Second, we want to produce two different data sets that can each be submitted to different types of inferential modelling approaches.
@@
 To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
-At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
 In their first experiment, @panisWhatShapingRT2016 presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 Table 3 shows the life table for condition "blank" (no prime stimulus presented) for participant 6.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
-At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
-
-Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
-By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
-
-```{r load-data}
-load("../Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata") # loads data_final
-```
-
-For example, for participant 6, the estimated hazard values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, when the waiting time has increased until *240 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
-
-Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 For example the first bin is (0,40] with the starting point excluded and the endpoint included.
+At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
 se = standard error.
 ca = conditional accuracy.
+NA = undefined.
 
@@
 
-(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.
+Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
+By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+
+(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6.
+Vertical dotted lines indicate the estimated median RTs.
+Error bars represent +/- 1 Standard Error of the respective proportion.
 
@@
 ```
+
+```{r load-data}
+load("../Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata") # loads data_final
+```
+
+For example, for participant 6, the estimated hazard values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, when the waiting time has increased until *240 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
+
+Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
-And when you treat time as a discrete predictor, you can choose between reference coding and index coding.
+And when you treat a variable as a discrete predictor, you can choose between reference coding and index coding.
+
 In the case of a large-*N* design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set [@allisonSurvivalAnalysisUsing2010].
@@
        # categorical predictor for TIME named "timebin" with index coding
-       timebin = factor(period,levels=c(6:15)),
+       timebin = factor(period, levels = c(6:15)),
        # factor "condition" using reference coding, with "blank" as the reference level
@@
        # categorical predictor "prime" with index coding
-       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
-       prime = factor(prime,levels=c(1,2,3)))
+       prime = ifelse(condition=="blank", 1, ifelse(condition=="congruent", 2, 3)),
+       prime = factor(prime, levels = c(1,2,3)))
 ```
@@
 To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
-In Tutorial_2a.Rmd we perform some prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
+In Tutorial_2a.Rmd we perform a few prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
 
@@
 Previous research has shown that psychological effects typically change over time [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhenDoesInhibition2022].
-In the next model, we use index coding for both TIME (variable "timebin") and the categorical predictor prime-target-congruency (variable "prime"), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
+In the next model, therefore, we use index coding for both TIME (variable "timebin") and the categorical predictor prime-target-congruency (variable "prime"), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
 Here is the model formula of this model that we call "M1i".
@@
 
-(ref:table-props-caption) Summarizing the posterior distributions of each contrast by their proportions below and above zero.
-
-(ref:table-props-note-caption) prop_below = proportion below zero: prop_above = proportion above zero.
+(ref:table-props-caption) Summarizing the posterior distributions of each contrast by their proportion below and above zero.
+
+(ref:table-props-note-caption) prop_below = proportion below zero; prop_above = proportion above zero.
 
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i_ca.
+To keep the tutorial short, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i_ca.
 
@@
 
-For illustration purposes, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
+To keep this tutorial short, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
@@
 In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-For illustration purposes, we only fitted the effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
+To keep it short, we only fitted the effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].

### 2024-12-01 manuscript/manuscript.tex
update ms.

@@
 Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
-We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
 Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
@@
 For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
-Insets show mean reaction time per condition.}\label{fig:plot1}
+Insets show mean response time per condition.}\label{fig:plot1}
 \end{figure}
@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Makeham, William M., 1860).
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1997; Makeham, 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
-Here, in the current paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+
+\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
+
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
 survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
 We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Moreover, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
-
-\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
-
-In this paper, we focus on discrete-time EHA.
+Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+
 We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
 However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003).
-Indeed, our primary aim here is to introduce a set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
+Indeed, our primary aim here is to introduce the set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
 
@@
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja() package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.
+Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.
 
@@
 In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
-We discuss possible link functions, and plot the model-based effects of our predictors of interest.
+We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
 In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
-In Tutorial 3a, we illustrate how to fit the same type of multilevel regression model for RT data in a frequentist framework using the R package lme4.
+
+In Tutorial 3a, we shortly illustrate how to fit similar multilevel regression models for RT data in a frequentist framework using the R package lme4.
 We then briefly compare and contrast these inferential frameworks when applied to EHA.
@@
 \item
-  define time point zero (e.g., target stimulus onset, fixation onset);
+  define time point zero (e.g., target stimulus onset, fixation onset, etc.);
 \item
@@
 While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t-1=5) = .44.
+For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t=5) = .44.
 Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.}\label{fig:plot2}
@@
 \item
-  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences that mean-average comparisons may conceal (Townsend, 1990).
+  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
 \item
   EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with longer RTs are not discarded, but instead contribute to the risk set in each time bin.
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see below).
 \item
@@
 
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition.
 These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-Between 500 and 600 ms the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
-For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these later bins are correct.
-
-This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT across trials.
+During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
+
+This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT and accuracy across trials.
 Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
@@
 \caption{Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).}\label{fig:plot3}
+Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Bin width equals 100 ms.}\label{fig:plot3}
 \end{figure}
@@
 
-Descriptive, data science workflow.
 We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
@@
 We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 
@@
 To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016).
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
-At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
 In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
-At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.
-
-Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
-By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
-
-For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
-
-Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6. At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
 
@@
 For example the first bin is (0,40{]} with the starting point excluded and the endpoint included.
+At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
 se = standard error.
@@
 
+Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
+By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+
 \begin{figure}[H]
@@
 
-\caption{Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6, as a function of the passage of discrete waiting time.}\label{fig:eha-plot}
+\caption{Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6.
+Vertical dotted lines indicate the estimated median RTs.
+Error bars represent 95\% confidence intervals for the respective proportion.}\label{fig:eha-plot}
 \end{figure}
+
+For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
+
+Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 
+\footnotesize
+
 \begin{Shaded}
@@
 \end{verbatim}
+
+\normalsize
 
@@
 \begin{TableNotes}[para]
-\normalsize{\textit{Note.} prop\_below = proportion below zero: prop\_above = proportion above zero.}
+\normalsize{\textit{Note.} prop\_below = proportion below zero; prop\_above = proportion above zero.}
 \end{TableNotes}
@@
 \begin{longtable}{llll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table-contrasts_props}Summarizing the posterior distributions of each contrast by their proportions below and above zero.}\\
+\caption{\label{tab:table-contrasts_props}Summarizing the posterior distributions of each contrast by their proportion below and above zero.}\\
 \toprule
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
@@
 
+\bibitem[\citeproctext]{ref-halleyVIEstimateDegrees1997}
+Halley, E.
+(1997).
+{VI}.
+{An} estimate of the degrees of the mortality of mankind; drawn from curious tables of the births and funerals at the city of {Breslaw}; with an attempt to ascertain the price of annuities upon lives.
+\emph{Philosophical Transactions of the Royal Society of London}, \emph{17}(196), 596--610.
+\url{https://doi.org/10.1098/rstl.1693.0007}
+
+\bibitem[\citeproctext]{ref-heiss2021}
+Heiss, A. (2021, November 10). A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel {Bayesian} Models. \url{https://doi.org/10.59350/wbn93-edb02}
+
 \bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
@@
 \bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-Makeham, William M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
+Makeham. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
 

### 2024-12-02 Tutorial_2a.Rmd
update ms. and tuts

--- 2024-11-26 Tutorial_2a.Rmd
@@
 
-Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflects our prior beliefs.
+### Intermezzo: Prior Predictive Check for M0i.
+
+Following Gelman et al.
+(2020), we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
 First, sample the prior distributions using sample_prior="only".
@@
        prior = priors_M0i,
-       chains = 4, cores = 4, 
+       chains = 1, cores = 1, 
        iter = 3000, warmup = 1000,
@@
 
-Next, use them to predict prior data. To better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.
+Next, use them to predict prior data. But to better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.
 
@@
 # Specify
-N_samples <- 1000 # the number of samples from the skew_normal
+N_samples <- 2000 # the number of samples from the skew_normal
 N_obs <- 100      # the number of simulated observations per sample
@@
 # apply the function
-prior_pred <- cloglog_predictive_distribution(cloglog_samples,N_obs)
+prior_pred <- cloglog_predictive_distribution(cloglog_samples,
+                                              N_obs)
 
@@
        title = "Prior predictive distribution",
-       subtitle = str_c(N_samples," samples; ",N_obs," observations per sample"))
+       subtitle = str_c(N_samples," samples; ",
+                        N_obs," observations per sample"))
 ```
@@
 
-The shape of the prior predictive distribution looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
-
-Now use add_predicted_draws() to simulate prior-based data for 10 bins and 6 subjects, while take into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
+```{r plot-ppc-1-bin, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png")
+```
+
+The shape of the prior predictive distribution for 1 bin looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
+
+Now use add_predicted_draws() to simulate prior-based data for 10 bins and 6 subjects, while taking into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
 
@@
 # plot 40 draws per participant
-ggplot(data = df_pred_haz %>% filter(.draw < 41), 
+ggplot(data = df_pred_haz %>% filter(.draw < 200), 
        aes(x=timebin, y=pred_haz, group=.draw)) +
@@
 
-Now we can fit model M0i to get the posterior distributions.
+```{r plot-ppc-M0i, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc.png")
+```
+
+The red hazard function is one of the 200 functions plotted, for each of six participants. They can take on any shape, with values below .5 being more likely than values above .5.
+
+After our prior predictive check, we can fit model M0i to get the posterior distributions.
 
@@
 model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
-formula(model_M0d)
 fixef(model_M0d)
@@
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
-  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0.3,.25)", class = "b", coef = "period_9"), 
   set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
@@
 
+### Intermezzo. Prior predictive check for model M0c.
+
 Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs.
-The following code allows to change the coefficients and generate (cloglog-)hazard functions.
+
+To get an idea of how the intercept and both slopes control the shape of the hazard function, the following code allows to change the coefficients and generate (cloglog-)hazard functions.
 Once you have found the coefficients for the intercept, period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to determine a suited mean and standard deviation for each coefficient.
@@
               x2 = x^2,
-              cloglog1 = -1.8   + 0.3*x   - 0.1*x2,
-              cloglog2 = -1     + 0.4*x   - 0.1*x2,
-              cloglog3 = -3     - 0.1*x   + 0.12*x2,
-              cloglog4 = -0.40  + 0.2*x   - 0.04*x2,
+              cloglog1 = -1.8   + 0.3*x    - 0.1*x2,  # black
+              cloglog2 = -1     + 0.41*x   - 0.04*x2,  # green
+              cloglog3 = -3     - 0.13*x   + 0.12*x2, # red
+              cloglog4 = -0.40  + 0.6*x    - 0.11*x2, # blue
               haz1 = 1 - exp(-1*exp(cloglog1)),
@@
 
-Next, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
-
-```{r priors-M0c-ppc}
-priors_M0c_ppc <- c(
+For example, the coefficients for x vary from -0.13 to 0.6, and we might specify N(.3,.25) as a prior distribution, Those for x-squared vary from -0.11 to 0.12, and we might specify N(0,.06) as a prior distribution.
+
+Suppose we expect (cloglog-)hazard functions that increase almost linearly with time.
+We also set the standard deviation of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
+
+```{r priors-M0c-ppc1}
+priors_M0c_ppc1 <- c(
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
-  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
-  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
-  set_prior("normal(0, .1)", class = "sd"),  # set standard deviation to a small value for now
+  set_prior("normal(0.3,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .01)", class = "sd"), 
   set_prior("lkj(2)", class = "cor")  
@@
 
-```{r fit-model-M0c-prior, eval=F}
+Sample the prior distributions.
+
+```{r fit-model-M0c-prior1, eval=F}
 plan(multicore)
 
-model_M0c_prior <-
+model_M0c_prior1 <-
    brm(data = data_M0c,
@@
                           (0 + Intercept + period_9 + period_9_sq | pid),
-       prior = priors_M0c_ppc,
-       chains = 4, cores = 4, 
+       prior = priors_M0c_ppc1,
+       chains = 1, cores = 1, 
        iter = 3000, warmup = 1000,
@@
        sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0c_prior")
-```
-
-```{r load-M0c-prior}
-model_M0c_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior.rds")
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior1")
+```
+
+```{r load-M0c-prior1}
+model_M0c_prior1 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior1.rds")
 ```
@@
 
-```{r prior-distr}
+```{r prior1-distr}
 # extract prior draws
-post <- as_draws_df(model_M0c_prior) %>% 
+post <- as_draws_df(model_M0c_prior1) %>% 
   select(-lp__) %>% 
@@
 
-```{r ppc-model-M0c-prior, eval=F}
-# Generate prior predictive hazard functions for 6 participants
-# Specify
-N_obs = 100 # increase for smoother results
+Create a function.
+
+```{r function-plot-ppc, eval=F}
+plot_ppc <- function(fit_prior, N_samples, N_obs, new_data_prior){
+
 set.seed(1)
-
-# make new data
-newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2) %>%
-  expand_grid(pid = 1:6)
 
@@
 for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0c_prior, 
-                                    newdata = newdata_prior, 
+  prior_pred <- add_predicted_draws(fit_prior, 
+                                    newdata = new_data_prior, 
                                     summary=F, 
-                                    ndraws=NULL) %>% 
+                                    ndraws=N_samples) %>% 
                 mutate(obs = i)
@@
 
-# plot 200 draws per participant, highlighting three
-sel = sample(1:200,3,replace=F)
+# plot N_samples draws per participant, highlighting three
+sel = sample(1:N_samples,3,replace=F)
 
 ggplot(df_pred_haz %>% 
-   filter(.draw <200), aes(x=period_9, y=pred_haz, group=.draw)) +
+   filter(.draw <N_samples), aes(x=period_9, y=pred_haz, group=.draw)) +
    geom_line(color="black") +
@@
   facet_wrap(~pid)
-```
-
-```{r save-ppc-M1c, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0c_ppc.png", width = 10, height = 8, dpi = 800)
+}
+```
+
+Apply the function.
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p1 <- plot_ppc(model_M0c_prior1,2000,100,newdata_prior)
+p1
+```
+
+For each of 2000 draws from the prior distributions, a black hazard function is plotted. Three randomly selected hazard functions are highlighted in red, green, and blue.
+
+```{r save-ppc1-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc1.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc1-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc1.png")
+```
+
+Or suppose we believe apriori that the hazard functions are flat.
+We specify our new priors:
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc2 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .001)", class = "sd"), 
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+Sample the prior distributions 2000 times:
+
+```{r fit-model-M0c-prior2, eval=F}
+plan(multicore)
+
+model_M0c_prior2 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc2,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior2")
+```
+
+```{r load-M0c-prior2}
+model_M0c_prior2 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior2.rds")
+```
+
+Apply the function plot_ppc().
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p2 <- plot_ppc(model_M0c_prior2,2000,100,newdata_prior)
+p2
+```
+
+Now the prior predictive hazard functions are all rather flat, with an intercept controlled by the prior skew-normal distribution, i.e., values below .5 are more likely than values above .5.
+
+```{r save-ppc2-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc2.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc2-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc2.png")
+```
+
+Finally, perform a ppc for the selected prior distributions for M0c (see lines 369-374), and increase the standard deviation for the normal prior for class "sd" somewhat.
+
+```{r priors-M0c-ppc3}
+priors_M0c_ppc3 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.3,.25)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .1)", class = "sd"),  
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior3, eval=F}
+plan(multicore)
+
+model_M0c_prior3 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc3,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
+```
+
+```{r load-M0c-prior3}
+model_M0c_prior3 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior3.rds")
+```
+
+Apply the function plot_ppc().
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p3 <- plot_ppc(model_M0c_prior3,2000,100,newdata_prior)
+p3
 ```
@@
 
-Now increase the standard deviation again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.
+```{r save-ppc3-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc3.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc3-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc3.png")
+```
+
+After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions.
 
@@
 
-Model_M0c took about 71 minutes to run.
+Model_M0c took about 19 minutes to run.
 
@@
 ```{r data-priors-M1d}
-data_M1d <- ptb_data %>% select(pid, event, period_9, condition) %>%
+data_M1d <- ptb_data %>% 
+  select(pid, event, period_9, condition) %>%
   mutate(period_9_sq = period_9^2)
@@
   set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"), 
-  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
+  set_prior("normal(0.3,.25)", class = "b", coef = "period_9"), 
   set_prior("normal(0,.06)", class= "b", coef = "period_9_sq"),  
@@
 
-Model_M1d took about 145 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Model_M1d took about 184 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
 
@@
 summary(model_M1d)
-formula(model_M1d)
 ```
@@
   prior(skew_normal(-1,1,-2), class = b, nlpar = a), 
-  prior(normal(0.1,.2), class = b, nlpar = b), 
-  prior(normal(0,.7), class = b, nlpar = c), 
+  prior(normal(0.3,.25), class = b, nlpar = b), 
+  prior(normal(0,.06), class = b, nlpar = c), 
   prior(exponential(1), class = sd, group = pid, nlpar = a),
@@
 ```
+
+Observations with Pareto k values above .7 are problematic.
 
@@
 
-We can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, for each participant in the data set, and for a brand new hypothetical participant.
+Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, for each participant in the data set, and for a brand new hypothetical participant.
 
@@
 
-We are actually interested in the difference in predicted hazard between congruent and blank primes on the one hand, and between incongruent and blank primes on the other hand, for each time bin.
+Following Heiss (2021), we are actually interested in the difference in predicted hazard between congruent and blank primes on the one hand, and between incongruent and blank primes on the other hand, for each time bin.
 When based on the grand mean, this is known as the grand average marginal effect (AME).

### 2024-12-02 manuscript/SupplementaryMaterial.Rmd
update ms. and tuts

@@
 
-We defined 13 custom functions that we list here.
+We defined 12 custom functions that we list here.
 
@@
 * join_lt_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
-* join_lt_ca_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
+* join_lt_ca_2IV(df1, df2) : add the ca(t) estimates to the life tables (2 independent variables)
 * extract_median(df) : estimate quantiles S(t)~.50~ (1 independent variable)
 * extract_median_2IV(df) : estimate quantiles S(t)~.50~ (2 independent variables)
-* plot_eha(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
-* plot_eha_2IV(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,   Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+* plot_eha(df, subj, haz_yaxis=1, first_bin_shown=1, aggregated_data=F, Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+* plot_eha_2IV(df, subj, haz_yaxis=1, first_bin_shown=1, aggregated_data=F,   Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
 
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
 
@@
 Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~1~ cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~2~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~2~ + $\beta$~3~ cloglog-hazard units (when TIME-9 = 1), and so forth.
+For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~3~ cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~4~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~4~ + $\beta$~5~ cloglog-hazard units (when TIME-9 = 1), and so forth.
 To interpret the effects of a predictor, its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
@@
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 12 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 16 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
@@
 However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
-A skew-normal distribution with a mean of -0.2, a standard deviation of 0.71, and a skewness of -2.2 might be a good weakly informative prior for the non-intercept parameters in a cloglog-hazard model as it gently regularizes p towards .6321 (i.e., a zero effect on the cloglog scale).
 
@@
 331) when one is interested in studying the detailed shape of a RT distribution [see also Figure 1 in @panisAnalyzingResponseTimes2020].
-Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are the preferred ways to describe the RT + accuracy data.
 

### 2024-12-02 manuscript/manuscript.Rmd
update ms. and tuts

@@
 Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in the supplementary material.
+As such, we only supply regression equations in part D of the supplementary material.
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial 1b for those that are interested.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial_1b.Rmd for those that are interested.
 
@@
 To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
-In Tutorial_2a.Rmd we perform a few prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
+In Tutorial_2a.Rmd we perform a few prior predictive checks to make sure our selected prior distributions reflect our prior beliefs [@gelmanBayesianWorkflow2020].
 
 The middle column of Figure 16 in part E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in supplementary Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 Each of these intercepts is allowed to vary across individuals (variable pid).
+We request 2000 samples from the posterior distribution for each of four chains.
 Estimating model M0i took about 30 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
@@
 Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 145 minutes to run.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 184 minutes to run.
 
@@
 EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions. 
-For example, model M4 contains two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+For example, some of the models discussed in Tutorial 2a (M2) contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
@@
 As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
-This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices.
+This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4).
 

### 2024-12-02 notes/manuscript_notes.Rmd
update ms. and tuts

@@
 
-- add figures of the prior predictive checks in Tutorial_2a.html
-
-- add refs for some websites I consulted
-
-- check the references in the ms to various parts of the supplementary material
-
 - check legends of all Figures, and add more info if needed
@@
 
-- add info on simulation and power analysis (section 4.4)
+- add info on simulation and power analysis (section 4.4 + summary in section 4)
 

### 2024-12-02 Tutorial_2a.Rmd
add 2a html
update 2a rmd and html

@@
 
-```{r ppc-manual, eval=T}
+```{r ppc-manual, eval=F}
 # prior predictive check for 1 time bin
@@
 summary(model_M0i)
-plot(model_M0i)
+#plot(model_M0i)
 ```

### 2024-12-02 manuscript/manuscript.tex
update ms.

@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1997; Makeham, 1860).
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1997; Makeham, William M., 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
@@
 
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6. At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
+Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
 
@@
 se = standard error.
-ca = conditional accuracy.}
+ca = conditional accuracy.
+NA = undefined.}
 \end{tablenotes}
@@
 Vertical dotted lines indicate the estimated median RTs.
-Error bars represent 95\% confidence intervals for the respective proportion.}\label{fig:eha-plot}
+Error bars represent +/- 1 Standard Error of the respective proportion.}\label{fig:eha-plot}
 \end{figure}
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial 1b for those that are interested.
+As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial\_1b.Rmd for those that are interested.
 
@@
 Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
-And when you treat time as a discrete predictor, you can choose between reference coding and index coding.\\
+And when you treat a variable as a discrete predictor, you can choose between reference coding and index coding.
+
 In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set (Allison, 2010).
@@
        \CommentTok{\# categorical predictor for TIME named "timebin" with index coding}
-       \AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period,}\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
+       \AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
        \CommentTok{\# factor "condition" using reference coding, with "blank" as the reference level}
@@
        \CommentTok{\# categorical predictor "prime" with index coding}
-       \AttributeTok{prime =} \FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"blank"}\NormalTok{,}\DecValTok{1}\NormalTok{, }\FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"congruent"}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)),}
-       \AttributeTok{prime =} \FunctionTok{factor}\NormalTok{(prime,}\AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)))}
+       \AttributeTok{prime =} \FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"blank"}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"congruent"}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{)),}
+       \AttributeTok{prime =} \FunctionTok{factor}\NormalTok{(prime, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)))}
 \end{Highlighting}
@@
 To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
-In Tutorial\_2a.Rmd we perform some prior predictive checks to make sure our selected prior distributions reflect our prior beliefs.
+In Tutorial\_2a.Rmd we perform a few prior predictive checks to make sure our selected prior distributions reflect our prior beliefs (Gelman, Vehtari, et al., 2020).
 
 The middle column of Figure 16 in part E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in supplementary Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
-In the next model, we use index coding for both TIME (variable ``timebin'') and the categorical predictor prime-target-congruency (variable ``prime''), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
+In the next model, therefore, we use index coding for both TIME (variable ``timebin'') and the categorical predictor prime-target-congruency (variable ``prime''), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
 Here is the model formula of this model that we call ``M1i''.
@@
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-For illustration purposes, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i\_ca.
+To keep the tutorial short, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i\_ca.
 
@@
 
-For illustration purposes, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
+To keep this tutorial short, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
@@
 In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-For illustration purposes, we only fitted the effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
+To keep it short, we only fitted the effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
@@
 EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions.
-For example, model M4 contains two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+For example, some of the models discussed in Tutorial 2a (M2) contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
@@
 
+\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., \ldots{} Modrák, M. (2020). \emph{Bayesian {Workflow}}. arXiv. \url{https://doi.org/10.48550/arXiv.2011.01808}
+
 \bibitem[\citeproctext]{ref-R-standist}
@@
 \bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-Makeham. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
+Makeham, William M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
 
@@
 \item
-  join\_lt\_ca\_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
+  join\_lt\_ca\_2IV(df1, df2) : add the ca(t) estimates to the life tables (2 independent variables)
 \item
@@
 \item
-  plot\_eha(df,subj,haz\_yaxis=1,first\_bin\_shown=1,aggregated\_data=F,Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
-\item
-  plot\_eha\_2IV(df,subj,haz\_yaxis=1,first\_bin\_shown=1,aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+  plot\_eha(df, subj, haz\_yaxis=1, first\_bin\_shown=1, aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
+\item
+  plot\_eha\_2IV(df, subj, haz\_yaxis=1, first\_bin\_shown=1, aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
 \end{itemize}
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a third-order polynomial specification for TIME can be written as follows:
+An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
 
@@
 Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{1} cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{2} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{2} + \(\beta\)\textsubscript{3} cloglog-hazard units (when TIME-9 = 1), and so forth.
+For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{3} cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{4} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{4} + \(\beta\)\textsubscript{5} cloglog-hazard units (when TIME-9 = 1), and so forth.
 To interpret the effects of a predictor, its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
@@
 To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 12 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 16 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
@@
 Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are preferred.
+Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are the preferred ways to describe the RT + accuracy data.
 

### 2024-12-02 Tutorial_3b.Rmd
update html files (and 2 tuts)
update html files

@@
 ```{r load-pkg, results='hide'}
-pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")
+pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")
 
@@
 load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
-fixef(model_M1i_ca_f)
 summary(model_M1i_ca_f)
@@
               position = 1:30)
-
-summary(est)
-head(est)
 ```

### 2024-12-02 Tutorial_1a.Rmd
update Rmd and html

@@
     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
-    # set haz_yaxis to .7, first_bin_shown to 1, aggregated_data to TRUE, and Nsubj to 6 
+    # set haz_yaxis to .7, first_bin_shown to 3, aggregated_data to TRUE, and Nsubj to 6 
     mutate(plot = map2(.x=lifetable_ca, .y=pid, plot_eha, .7, 3, T, 6)) 
@@
   # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
-  #.           you can set first_bin_shown to another value, such as 5 (i.e., do not show the 
-  #            first 4 bins in the plots).
+  #.           you can set first_bin_shown to another value, such as 3 (i.e., do not show the 
+  #            first3 bins in the plots).
   # aggregated_data: For a dataset with data aggregated across individuals, set to TRUE. 

### 2024-12-02 Tutorial_2a.Rmd
update Rmd and html

@@
 
-Now use add_predicted_draws() to simulate prior-based data for 10 bins and 6 subjects, while taking into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
+Now use add_predicted_draws() to simulate prior-based data for our 10 bins and 6 subjects, while taking into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
 
@@
                                     summary=F, 
-                                    ndraws=NULL) %>% 
+                                    ndraws=NULL) %>% # all 2000
                 mutate(obs = i)
@@
 
-# plot 40 draws per participant
+# plot 200 draws per participant
 ggplot(data = df_pred_haz %>% filter(.draw < 200), 
@@
        title = "Prior predictive distributions",
-       subtitle = str_c("8000 samples; ",N_obs," observations per sample")) +
+       subtitle = str_c("200 samples; ",N_obs," observations per sample")) +
   facet_wrap(~pid)
@@
 model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
-summary(model_M0i)
+fixef(model_M0i)
 #plot(model_M0i)
@@
 summary(model_M0c)
-formula(model_M0c)
 ```
@@
 fixef(model_M1i)
-formula(model_M1i)
 ```
@@
 model_M1d <- readRDS("Tutorial_2_Bayesian/models/model_M1d.rds")
-summary(model_M1d)
+fixef(model_M1d)
 ```
@@
 
-```{r model-formula-1}
+```{r model-formula-1, eval=F}
 event ~ 0 + timebin:prime:stage  + (0 + timebin:prime:stage | pid)
@@
 
-```{r model-formula-2}
+```{r model-formula-2, eval=F}
 bf(event ~ 0 + a + b * trial_c + c * trial_c_sq,
@@
 
-```{r model-formula-3}
+```{r model-formula-3, eval=F}
  bf(event ~ 0 + Intercept + 

### 2024-12-02 Tutorial_2b.Rmd
update Rmd and html

@@
                     alpha = 0.5, 
-                    .width = .95) +
+                    .width = c(.8,.95)) +
     scale_fill_brewer(palette = "Dark2") +
@@
                     point_interval = "mean_qi",
-                    .width=c(.95)) +
+                    .width=c(.8,.95)) +
     #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +

### 2024-12-03 manuscript/manuscript.Rmd
minor edits

@@
 Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
-
 keywords          : "response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology"
@@
 output            : papaja::apa6_pdf
-
 # output            : papaja::apa6_word
@@
 We perform data wrangling following tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
-Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
-[[more here, as necessary]].
+Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
+In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of EHA a lot quicker and more efficient.
 
@@
 For example, @panisWhatShapingRT2016 concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
-Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
+Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to.
+To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
 This effect might be masked for incongruent primes by the response competition effect.

### 2024-12-03 notes/manuscript_notes.Rmd
update ms. and notes
added manuscript notes

@@
 
+- delete code in Tutorial_2a about brand new hypothetical subject?
+
 - use timebin or time bins?
@@
 - use multi-level or multilevel ?
+
+## December 2nd 2024 ##
+
+I read your comments above.
+
+On first glance, I really like the new plots!
+
+I'll dig in now with some comments.
+
+### intro ###
+
+I think the Halley Ref should be 1693 and not 1997? We have a copy of the original manuscriopt, so it seems reasonable to cite the original?
+
+### section 3.1 ###
+
+maybe add a sentence to highlights how we add our own custom functions to make the wrangling and plotting sections quicker and more efficient.
+e.g., we add to this approach by writing our own custom functions.
+
+### section 3.2 ###
+
+first sentence, the McElreath reference should be 2020 for the textbook and not 2018.
+
+### 4.3.4 Model M1d ###
+
+I'm not convinced we need this model to be reported in the main text. It is complex and deviates from the prior model structure with index coding.
+Folks need to get their head around polynomial specifications of time. And we don't really need them to and we don't really give them much chance to understand it. It is kind of like "trust me".
+Instead, I would have this in the tutorial with a qualifier saying...you can also build models that specify time differently...
+
+Then in the main text just compare M0i and M1i, for the purposes of demonstrating how to do it.
+The basic conclusion being that adding the factor condition improves predictive accuracy compare to timebin only. 
+That's maybe quite basic, but also quite important too for our conclusion - congruency matters.
+
+Or maybe have these models, all with index coding:
+
+- grand intercept only (no predictors)
+- timebin only (M0i)
+- congruency only (without timebin)
+- congruency:timebin (M1i)
+
+Or just stick with two models, as you have above (M0i and M1i).
+
+The main thing we want to communicate is how and why to do model comparison. So let's keep it as simple as possible, at least in the main text.
+
+### section 4.3.3 ###
+
+we need to define index coding.
+
+maybe somewhere earlier we could have a subsection that explains different coding schemes for categorical variables?
+e.g., dummy, reference and contrast it with index.
+Then say we focus on index in the main text, as we think it is easier to interpret means per condition.
+
+### section 4.3.6 ###
+
+Tables 4 and 5 can be moved to the tutorial only (or Supps) whichever makes most sense.
+
+Figure 5 y-axis. Should it just be cloglog? Rather than cloglog-hazard?
+
+Another way to plot Figure 5 would be to avoid a facet and instead fill by prime.
+That way, you might benefit from the principle of putting things close together that you want to compare e.g., the effects of prime should be easier to see if prime comparisons are close together.
+Maybe at least plot it the other way to have a look and see what you think?
+I'd like to see it at least.
+
+There is no need to commit the space needed when they are redundant with the figures.
+e.g., we can just say in the main text, the point estimates and quantile intervals can be reported in a table (see the tutorial for details).
+That saves loads of space and time scanning values when the figures are beautiful and standalone as they are.
+
+also, the title of this sub-section should be about evaluating parameter estimates (rather than interpreting model ...).
+This is because you can use model comparison to interpret the model.
+So we should make the subsection distinctive.
+
+This figures in this section are wonderful. Nice work!
+
+I suggest two changes:
+
+1.
+Combine Figures 6 and 7 into one figure (so that the group average is at the top and the individual PIDs are below.
+Patchwork can do this for you.
+And if you include + plot_annotation(tag_levels = 'A') then you'll get a nice A and B to label the different parts of the plot.)
+
+2. As (1) combine Figures 8 and 9.
+
+on page 38, you say:
+
+>Note that we could also have calculated hazard ratios instead of hazard differences.
+
+This feels a little disjointed from what comes before in the paragraph.
+Either remove it or provide an extra sentence that distinguishes between the absolute hazard difference and the hazard ratio.
+
+Figure 6.
+Change the title from "Average participant" as it is a bit misleading.
+Maybe something like "group average" or "grand average" or even remove the title completely once you combine it with the individual pid plots because it will probably become obvious, especially with a figure legend explaining things.
+
+Figure 9. spell out AMEs in full to avoid jargon.
+
+### section 4.4 ###
+
+Combine Figures 11 and 12 like I suggest above.
+
+Put Table 6 in the tutorial and remove it from the main text.
+
+Figure 14.
+Do a facet_wrap by prime type.
+Put timebin on the x axis and cloglog on the y-axis.
+This makes the orientation similar to the prior plots and will therefore be easier to follow for the reader.

### 2024-12-03 notes/model_notes.Rmd
added some model notes for tutorial 2a

@@
 I will share the code so that you can take a look.
+
+# December 3rd 2024 #
+
+## Tutorial 2a ##
+
+I haven't had time to dig through the prior predictive checks yet, as I wanted to focus on the main plots in the paper.
+
+If I follow correctly, it is re-assuring that wrangling the posterior draws gives the same results as using add_epred_draws(). Am I following correctly though?
+
+when you predict data for a new participant, couldn't you also just supply new data that has no pid variable?
+Instead, it would just have timebin and prime condition and the model would make appropriate estimates??
+I'd be curious to see this also (not that we necessarily need it in the tutorial, I'm just curious.)

### 2024-12-04 Tutorial_4.Rmd
updated Tutorial 4

@@
 1. Fit a regression model to an existing dataset.
-2. Use the regression model parameters to simulate a one new dataset.
-3. Create a function to iterate over many datatsets and vary parameters of interest (e.g., sample size, trial count, effect size).
+2. Use the regression model parameters to simulate one new dataset.
+3. Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
 4. Summarise the simulated data to estimate likely power or precision of the research design options.
@@
 ggsave ("Tutorial_4_planning/figures/raw_vs_d1.jpeg",
-        width = 8, height = 8)
+        width = 8, height = 8, dpi=800)
 ```
@@
   scale_fill_brewer(palette = "Dark2") +
+  labs(title = "Simulation 1") +
   theme(legend.position = "none") +
@@
   scale_fill_brewer(palette = "Dark2") +
+  labs(title = "Simulation 2") +
   theme(legend.position = "none") +
@@
 Instead, you are aiming for common patterns across two or more experiments, rather than trying to make the case that a single study has sufficient evidential value to hit some criterion mark.
+
+# 12. Join some plots together #
+
+Join the power column plots together from sim1 and sim2.
+
+```{r}
+p12.1 <- (p10.4 | p11.6) +
+  plot_annotation(tag_levels = 'A') +
+  plot_layout(axes = 'collect')
+p12.1
+
+ggsave ("Tutorial_4_planning/figures/bin6_power_col.jpeg",
+        width = 12, height = 7, dpi=800)
+```

### 2024-12-04 manuscript/manuscript.Rmd
added text for Tutorial 4 (planning)

@@
 
-## 4.4 Tutorial 4: Planning
-
-This needs adding... RR to do this.
+## 4.7 Tutorial 4: Planning
+
+In the final tutorial, we look at planning a future experiment, which uses EHA.
+
+### 4.7.1 Background
+
+The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman et al., 2020).
+The basic structure and code follows the examples outlined by Solomon Kurz in his
+'power' blog posts (Kurz, 2019) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper (DeBruine & Barr, 2021).
+
+### 4.7.2 Basic workflow
+
+The basic workflow is as follows:
+
+1. Fit a regression model to an existing dataset.
+2. Use the regression model parameters to simulate one new dataset.
+3. Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
+4. Summarise the simulated data to estimate likely power or precision of the research design options.
+
+Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data.
+However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be computationally demanding for desktop machines.
+So, for ease, here we just use the raw simulated datasets to guide future expectations.
+
+In the below, we only provide a high-level summary of the process and let readers dive into the details within the tutorial should they feel so inclined.
+
+### 4.7.3 Fit a regression model and simulate one dataset
+
+We again use the data from Panis & Schmidt (2016) to provide a worked example.
+We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
+We chose to focus on a subsample of the data to ease the computational burden.
+We also used a full varying effects structure, with the model formula, as follows:
+
+```{r sim-model-bi, eval=F, echo=T}
+event ~ 0 + timebin:prime + (0 + timebin:prime | pid)
+```
+
+We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
+The raw data and the simulated data are plotted in Figure 15 and show quite close correspondence, which is re-assuring.
+But, this is only one dataset.
+What we really want to do is simulate many datasets and vary parameters of interest, which is what we turn to in the next section.
+
+(ref:simdat-plot15-caption) Raw data from Panis & Schmidt (2016) and simulated data from 10 participants.
+
+```{r simdat, fig.cap = "(ref:simdat-plot15-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_4_Planning/figures/raw_vs_d1.jpeg")
+```
+
+### 4.7.4 Simulate and summarise data across a range of parameter values
+
+Here we use the same data simulation process as used above, but instead of simulating one dataset, we simulate 1000 datasets per variation in parameter values.
+Specifically, in Simulation 1, we vary the number of trials per condition (100, 200, and 400), as well as the effect size in bin 6. 
+We focus on bin 6 only, in terms of varying the effect size, just to make things simpler and easier to understand.
+The effect size observed in bin 6 in this subsample of data was a 79% reduction in hazard value from the congruent prime (0.401 hazard value) to the incongruent prime condition (0.085 hazard value).
+In other words, a hazard ratio of 0.21 (e.g., 0.085/0.401 = 0.21).
+As a starting point, we chose three effect sizes, which covered a fairly broad range of hazard ratios (0.25, 0.5, 0.75), which correspond to a 75%, 50% and 25% reduction in hazard value as a function of prime condition.
+
+Summary results from Simulation 1 are shown in Figure 16A.
+Figure 16A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
+We are aware that "power" is not part of a Bayesian analytical workflow, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
+
+The results of Simulation 1 show that if we were targeting an effect size similar to the one reported in the original study, then testing 10 participants and collecting 100 trials per condition would be enough to provide over 95% power.
+However, we could not be as confident about smaller effects, such as a hazard ratio of 50% or 25%.
+From this simulation, we can see that somewhere between an effect size of a 50% and 75% reduction in hazard value, power increases to a range that most researchers would consider acceptable (i.e., >95% power).
+To probe this space a little further, we decided to run a second simulation, which varied different parameters
+
+In Simulation 2, we varied the effect size between a different range of values (0.5, 0.4, 0.3), which correspond to a 50%, 60% and 70% reduction in hazard value as a function of prime condition.
+In addition, we varied the number of participants per experiment between 10, 15, and 20 participants.
+Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
+Summary results from Simulation 2 are shown in Figure 16B.
+A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):
+
+- For a 70% reduction (0.3 hazard ratio), N=10 would give nearly 100% power.
+- For a 60% reduction (0.4 hazard ratio), N=10 would give nearly 90% power.
+- For a 50% reduction (0.5 hazard ratio), N=15 would give over 80% power.
+
+(ref:power-plot16-caption) Statistical power across data Simulation 1 (A) and Simulation 2 (B).
+Power was calculated as the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+In Simulation 1, the effect size was varied between a 25%, 50% and 75% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
+In Simulation 2, the effect size was varied between a 50%, 60% and 70% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80% (red), 90% (black) and 95% (blue) power.
+Abbreviations: rep_n = the number of trials per experimental condition; sub_j = the number of participants per simulated experiment.
+
+```{r power, fig.cap = "(ref:power-plot16-caption)", out.width='100%'}
+knitr::include_graphics("../Tutorial_4_Planning/figures/bin6_power_col.jpeg")
+```
+
+### 4.7.5 Planning decisions
+
+Now that we have summarised our simulated data, what planning decisions could we make about a future study?
+How many trials per condition should we collect and how many participants should we test? 
+Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022). 
+There is no straigtforward and clear-cut answer.
+Some considerations might be...
+
+  - How much power or precision are you looking to obtain in this particular study?
+  - Are you running multiple studies that have some form of replication built in?
+  - What resources do you have at your disposal, such as time, money and personnel?
+  - How easy or difficult is it to obtain the specific type of sample?
+
+If we were running this kind of study in our lab, what would we do?
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon (Panis & Schmidt, 2016).
+Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80% power mark.
+If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide >90% power for an effect size of 0.5.
+
+**But**, and this is an important "but", unless there are unavoidable reasons, no matter what planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
+Instead, we would run a follow-up experiment that replicates and extends the initial result. 
+By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis. 
+Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study on its own has sufficient evidential value to hit some criterion mark.
 

### 2024-12-04 manuscript/manuscript.tex
added text for Tutorial 4 (planning)

@@
 We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
-Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
-{[}{[}more here, as necessary{]}{]}.
+Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
+In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of EHA a lot quicker and more efficient.
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, n.d.), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Heiss (2021), Kurz (2023a), and Kurz (2023b).
@@
 
-{\centering \includegraphics[width=21.33in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 For example, Panis and Schmidt (2016) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
-Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to; To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after \textasciitilde360 ms, and towards the end of the experiment.
+Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to.
+To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after \textasciitilde360 ms, and towards the end of the experiment.
 This effect might be masked for incongruent primes by the response competition effect.
@@
 
-\subsection{4.4 Tutorial 4: Planning}\label{tutorial-4-planning}
-
-This needs adding\ldots{} RR to do this.
+\subsection{4.7 Tutorial 4: Planning}\label{tutorial-4-planning}
+
+In the final tutorial, we look at planning a future experiment, which uses EHA.
+
+\subsubsection{4.7.1 Background}\label{background}
+
+The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman et al., 2020).
+The basic structure and code follows the examples outlined by Solomon Kurz in his
+`power' blog posts (Kurz, 2019) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
+
+\subsubsection{4.7.2 Basic workflow}\label{basic-workflow}
+
+The basic workflow is as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Fit a regression model to an existing dataset.
+\item
+  Use the regression model parameters to simulate one new dataset.
+\item
+  Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
+\item
+  Summarise the simulated data to estimate likely power or precision of the research design options.
+\end{enumerate}
+
+Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data.
+However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be too computationally demanding for desktop machines.
+So, for ease, here we just use the raw simulated datasets to set expectations.
+
+In the below, we only provide a high-level summary of the process and let readers dive into the details within the tutorial should they feel so inclined.
+
+\subsubsection{4.7.3 Fit a regression model and simulate one dataset}\label{fit-a-regression-model-and-simulate-one-dataset}
+
+We again use the data from Panis \& Schmidt (2016) to provide a worked example.
+We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
+We chose to focus on a subsample of the data to ease the computational burden.
+We also used a full varying effects structure, with the model formula, as follows:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{|}\NormalTok{ pid)}
+\end{Highlighting}
+\end{Shaded}
+
+We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
+The raw data and the simulated data are plotted in Figure 15 and show quite close correspondence, which is re-assuring.
+But, of course, this is only one dataset.
+What we really want to do is simulate many datasets and vary parameters of interest, which is what we turn to in the next section.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_4_Planning/figures/raw_vs_d1}
+
+}
+
+\caption{Raw data from Panis \& Schmidt (2016) and simulated data from 10 participants.}\label{fig:simdat}
+\end{figure}
+
+\subsubsection{4.7.4 Simulate and summarise data across a range of parameter values}\label{simulate-and-summarise-data-across-a-range-of-parameter-values}
+
+Here we use the same data simulating process as used above, but instead of simulating one dataset, we simulate 1000 datasets per variation in parameter values.
+Specifically, in Simulation 1, we vary the number of trials per condition (100, 200, and 400), as well as the effect size in bin 6.
+We focus on bin 6 only, in terms of varying the effect size, just to make things simpler and easier to understand.
+The effect size observed in bin 6 in this subsample of data was a 79\% reduction in hazard value from the congruent prime (0.401 hazard value) to the incongruent prime condition (0.085 hazard value).
+In order words, a hazard ratio of 0.21.
+As a starting point, we chose three effect sizes, which covered a fairly broad range of hazard ratios (0.25, 0.5, 0.75), which correspond to a 75\%, 50\% and 25\% reduction in hazard as a function of prime condition.
+
+Summary results from Simulation 1 are shown in Figure 16A.
+Figure 16A which depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+We know that ``power'' is not a Bayesian thing, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
+In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
+The results show that if we were targetting an effect size similar to the one reported in the orgininal study, the using 10 participants and 100 trials per condition would be enough to provide over 95\% power.
+However, we could not be as confident about smaller effects, such as a hazard ratio of 50\% or 25\%.
+From this simulation, we can see that somewhere between an effect size of 50\% and 75\%, power increases to a range that most researchers would consider acceptable (i.e., \textgreater95\% power).
+To probe this space a little further, we decided to run a second simulation, which varied different parameters
+
+I Simulation 2, we varied the effect size across different values (0.5, 0.4, 0.3), which correspond to a 50\%, 60\% and 70\% reduction in hazard as a function of prime condition.
+In addition, we varied the number of participants in the data (10, 15, and 20 participants).
+Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
+Summary results from Simulation 2 are shown in Figure 16B.
+A summary of these power calculations might be as follows:
+
+For a 70\% reduction (0.3 hazard ratio), N=10, trial=200 per condition, would give nearly 100\% power.
+For a 60\% reduction (0.4 hazard ratio), N=10, trial=200 per condition, would give nearly 90\% power.
+For a 50\% reduction (0.5 hazard ratio), N=15, trial=200 per condition, would give over 80\% power.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth,height=0.67\textheight,]{../Tutorial_4_Planning/figures/bin6_power_col}
+
+}
+
+\caption{Statistical power across data Simulation 1 (A) and Simulation 2 (B).
+Power was calculated as the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+In Simulation 1, the effect size was varied between a 25\%, 50\% and 75\% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
+In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.}\label{fig:power}
+\end{figure}
+
+\subsubsection{4.7.5 Planning decisions}\label{planning-decisions}
+
+Now that we have summarised our simulated data, what planning decisions do we make about future study?
+How many trials per condition should we collect and how many participants should we test?
+Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022).
+There is no straigtforward and clear-cut answer.
+Some considerations might be\ldots{}
+
+\begin{itemize}
+\tightlist
+\item
+  How much power or precision are you looking to obtain in this particular study?
+\item
+  Are you running multiple studies that have some form of replication built in?
+\item
+  What resources do you have at your disposal, such as time, money and personnel?
+\item
+  How easy or difficult is it to obtain the specific type of sample?
+\end{itemize}
+
+If we were running this kind of study in our lab, what would we do?
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon, then pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80\% power mark.
+If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
+
+\textbf{But}, and this is an important ``but'', unless there are unavoidable reasons, no matter what planning choices we made, we would not solely rely on one study.
+Instead, we would run a follow-up experiment that replicates and extends the initial result.
+By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of power analysis.
+Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study has sufficient evidential value to hit some criterion mark.
 
@@
 \bibitem[\citeproctext]{ref-R-standist}
-Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
+Girard, J. (n.d.). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
 

### 2024-12-04 manuscript/manuscript.tex
minor updates to the manuscript

@@
 Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data.
-However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be too computationally demanding for desktop machines.
-So, for ease, here we just use the raw simulated datasets to set expectations.
+However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be computationally demanding for desktop machines.
+So, for ease, here we just use the raw simulated datasets to guide future expectations.
 
@@
 The raw data and the simulated data are plotted in Figure 15 and show quite close correspondence, which is re-assuring.
-But, of course, this is only one dataset.
+But, this is only one dataset.
 What we really want to do is simulate many datasets and vary parameters of interest, which is what we turn to in the next section.
@@
 
-Here we use the same data simulating process as used above, but instead of simulating one dataset, we simulate 1000 datasets per variation in parameter values.
+Here we use the same data simulation process as used above, but instead of simulating one dataset, we simulate 1000 datasets per variation in parameter values.
 Specifically, in Simulation 1, we vary the number of trials per condition (100, 200, and 400), as well as the effect size in bin 6.
@@
 The effect size observed in bin 6 in this subsample of data was a 79\% reduction in hazard value from the congruent prime (0.401 hazard value) to the incongruent prime condition (0.085 hazard value).
-In order words, a hazard ratio of 0.21.
-As a starting point, we chose three effect sizes, which covered a fairly broad range of hazard ratios (0.25, 0.5, 0.75), which correspond to a 75\%, 50\% and 25\% reduction in hazard as a function of prime condition.
+In other words, a hazard ratio of 0.21 (e.g., 0.085/0.401 = 0.21).
+As a starting point, we chose three effect sizes, which covered a fairly broad range of hazard ratios (0.25, 0.5, 0.75), which correspond to a 75\%, 50\% and 25\% reduction in hazard value as a function of prime condition.
 
 Summary results from Simulation 1 are shown in Figure 16A.
-Figure 16A which depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
-We know that ``power'' is not a Bayesian thing, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
+Figure 16A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
-The results show that if we were targetting an effect size similar to the one reported in the orgininal study, the using 10 participants and 100 trials per condition would be enough to provide over 95\% power.
+We are aware that ``power'' is not part of a Bayesian analytical workflow, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
+
+The results of Simulation 1 show that if we were targeting an effect size similar to the one reported in the original study, then testing 10 participants and collecting 100 trials per condition would be enough to provide over 95\% power.
 However, we could not be as confident about smaller effects, such as a hazard ratio of 50\% or 25\%.
-From this simulation, we can see that somewhere between an effect size of 50\% and 75\%, power increases to a range that most researchers would consider acceptable (i.e., \textgreater95\% power).
+From this simulation, we can see that somewhere between an effect size of a 50\% and 75\% reduction in hazard value, power increases to a range that most researchers would consider acceptable (i.e., \textgreater95\% power).
 To probe this space a little further, we decided to run a second simulation, which varied different parameters
 
-I Simulation 2, we varied the effect size across different values (0.5, 0.4, 0.3), which correspond to a 50\%, 60\% and 70\% reduction in hazard as a function of prime condition.
-In addition, we varied the number of participants in the data (10, 15, and 20 participants).
+In Simulation 2, we varied the effect size between a different range of values (0.5, 0.4, 0.3), which correspond to a 50\%, 60\% and 70\% reduction in hazard value as a function of prime condition.
+In addition, we varied the number of participants per experiment between 10, 15, and 20 participants.
 Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
 Summary results from Simulation 2 are shown in Figure 16B.
-A summary of these power calculations might be as follows:
-
-For a 70\% reduction (0.3 hazard ratio), N=10, trial=200 per condition, would give nearly 100\% power.
-For a 60\% reduction (0.4 hazard ratio), N=10, trial=200 per condition, would give nearly 90\% power.
-For a 50\% reduction (0.5 hazard ratio), N=15, trial=200 per condition, would give over 80\% power.
+A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):
+
+\begin{itemize}
+\tightlist
+\item
+  For a 70\% reduction (0.3 hazard ratio), N=10 would give nearly 100\% power.
+\item
+  For a 60\% reduction (0.4 hazard ratio), N=10 would give nearly 90\% power.
+\item
+  For a 50\% reduction (0.5 hazard ratio), N=15 would give over 80\% power.
+\end{itemize}
 
@@
 In Simulation 1, the effect size was varied between a 25\%, 50\% and 75\% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
-In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.}\label{fig:power}
+In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.
+Abbreviations: rep\_n = the number of trials per experimental condition; sub\_j = the number of participants per simulated experiment.}\label{fig:power}
 \end{figure}
@@
 
-Now that we have summarised our simulated data, what planning decisions do we make about future study?
+Now that we have summarised our simulated data, what planning decisions could we make about a future study?
 How many trials per condition should we collect and how many participants should we test?
@@
 If we were running this kind of study in our lab, what would we do?
-We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon, then pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80\% power mark.
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon (Panis \& Schmidt, 2016).
+Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80\% power mark.
 If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
 
-\textbf{But}, and this is an important ``but'', unless there are unavoidable reasons, no matter what planning choices we made, we would not solely rely on one study.
+\textbf{But}, and this is an important ``but'', unless there are unavoidable reasons, no matter what planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
 Instead, we would run a follow-up experiment that replicates and extends the initial result.
-By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of power analysis.
-Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study has sufficient evidential value to hit some criterion mark.
+By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis.
+Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study on its own has sufficient evidential value to hit some criterion mark.
 

### 2024-12-05 Tutorial_2a.Rmd
added a figure with 6 hypothetical pids

@@
 pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
-         "brms", "tidybayes", "bayesplot", "future", "parallel", "modelr")
+         "brms", "tidybayes", "bayesplot", "future", "parallel", "modelr",
+         "rstan")
 
@@
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
+```
+
+Fourth, six completely new hypothetical participants.
+
+```{r epred-new-pid-6}
+epreds_pid_new6 <- dat_M1i %>% 
+  data_grid(timebin, prime, pid = 7:12) %>%
+  add_epred_draws(model_M1i, 
+                  re_formula = NULL, 
+                  allow_new_levels=T,    
+                  sample_new_levels = "gaussian") %>% 
+  mutate(prime = factor(prime, 
+                        levels=c(1,2,3),
+                        labels=c("blank","congruent","incongruent"))) %>%
+  ungroup()
+```
+
+Summarize and plot predicted hazard values.
+
+```{r epred-new-pid-6-summary}
+epreds_pid_new6 %>% 
+  group_by(pid,timebin,prime) %>% 
+  median_qi(.width=.95)
+```
+
+```{r epred-new-pid-6-plot}
+ ggplot(epreds_pid_new6, aes(x=timebin, y=.epred, 
+                       fill=prime, color=prime)) +
+    stat_lineribbon(point_interval="median_qi",
+                    .width=.95,
+                    alpha = 0.5) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    theme(axis.text.x = element_text(angle=90)) +
+    scale_y_continuous(limits=c(0,1)) +
+    labs(y = "predicted hazard") +
+    ggtitle("Six new hypothetical participants") +
+    facet_wrap(~pid, ncol=3)
+```
+
+```{r save-epred-pid-new, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new_6.png", width = 10, height = 8, dpi = 800)
 ```

### 2024-12-05 notes/tutorial_notes.Rmd
added comments on Tutorial 2a. We've already discussed these in person at this point.

@@
 There are other themes, but this works for me.
+
+# December 4th 2024 #
+
+I took a look at tutorial2a.
+
+## overall comments on structure ##
+
+My feeling is that it is comprehensive and clear. Nice work on the prior predictive checks.
+
+However, it is long. And there is a lot of additional material, which extends beyond what is required.
+This is just an observation, and not necessarily a criticism.
+
+To give you a few examples, the sidenote about prior predictive checks is long.
+And you calculate things twice in several places e.g., from draws manually and using convenience functions.
+Now, this might all add to the richness and depth of understanding that the tutorial provides.
+But, at the same time, it does appear to be going against our stated aim of removing barriers to entry.
+If we consider a more minimal structure, we would supply code in the tutorial to get the job done and nothing more.
+Readers would be able to walk along with the most direct and straightforward link to the main text.
+There could be a case that we are not showing enough skill in selecting what is most important.
+
+Now, at the moment, I'm conflicted.
+The richness and context plus thinking and logic is good.
+But it makes it long with loads of plots.
+My worry is that someone reads the manuscript, flips to the tutorial and then thinks "oh boy, what's all this?" and then gives up.
+
+For now, I say we leave it, but just keep it in mind for future revisions and accessibility, as it may need re-thinking later.
+
+## epred draws ##
+
+These look good in general. 
+But let's have a quick discussion about one of them. Something came up when I was playing around.
+I need to run it by you.

### 2024-12-05 Tutorial_2b.Rmd
minor edits 2b.

@@
 # select analysis time range: (200,600]
-ca_data <- ca_data %>% filter(drt > 5) %>%
-
+ca_data <- ca_data %>% 
+  filter(drt > 5) %>%
   # create factor condition, with "blank" as the reference level
@@
                             labels = c("blank", "congruent","incongruent")),
-       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
+         prime = ifelse(condition=="blank",1, 
+                 ifelse(condition=="congruent",2,3)),
          prime = factor(prime,levels=c(1,2,3)),
@@
          trial_c = (trial - 1000)/1000) 
-
 head(ca_data)
@@
 # remove unnecessary columns before fitting a model
-data_M1i_ca <- ca_data %>% select(pid,acc,timebin,prime)
+data_M1i_ca <- ca_data %>% 
+  select(pid,acc,timebin,prime)
 head(data_M1i_ca)

### 2024-12-05 notes/tutorial_notes.Rmd
comments on Tutorial 2b

@@
 I need to run it by you.
+
+# December 5th 2024 #
+
+I took a look at tutorial2b.
+
+## coding conventions ##
+
+I edited stuff to make each code argument after a pipe a new line. 
+I think it is easier to read this way.
+You can also "re-indent" lines to make the code more consistent in terms of indents also.
+e.g., cmd+I or select the tab "code" then "re-indent lines"
+
+## priors ##
+
+line 101. we don't say anything about how we came to set the priors how we did. 
+Should we say something? Or do we say it elsewhere?
+Maybe a few lines about weakly iformative priors and what that means in this context?
+
+## plots ##
+
+a very minor point.
+for the grand average marginal effects, you facet by condition.
+for the single subject marginal effects, you overlap the lines.
+This is not a big deal, I'm just pointing out the difference.
+e.g., you could plot the grand average in the same way. 
+I'm not saying that would be better. I'm just saying that it would then be consistent.
+It makes no difference overall as they are both very clear as they are.

### 2024-12-05 Tutorial_1a.Rmd
added tidyverse wrangling style and reformatted code chunks

@@
 # remove practice trials
-data <- raw %>% filter(practice==0) # 9240 experimental trials: 6 subj x 70 blocks x 22 trials-per-block
+data <- raw %>% 
+  filter(practice==0) # 9240 experimental trials: 6 subj x 70 blocks x 22 trials-per-block
 
 # remove unnecessary columns
-data <- data %>% select(-c(ss,practice,rd,form,task,primeduration,
+data <- data %>% 
+  select(-c(ss,practice,rd,form,task,primeduration,
             blankduration,maskduration,targetduration,
@@
 # look at trials where resp==0
-data %>% filter(resp==0) %>% print(n=51)
+data %>% 
+  filter(resp==0) %>% 
+  print(n=51)
 
@@
 # the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
-sel = which(data$resp==0 & data$rt < 800)
-data <- data[-c(sel),]
+data <- data %>% 
+  filter(!(resp == 0 & rt < 800))
+
+data %>%
+  distinct(resp)
 
 # keep only the no-mask trials
-data <- data %>% filter(mask_type == 0) %>%
+data <- data %>% 
+  filter(mask_type == 0) %>%
   select(-mask_type)
@@
 # change some variable names to required names: pid, trial, condition, rt, and acc.
-colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial")
+data_wr <- data_wr %>% 
+  rename(pid = vp, condition = prime_type, acc = respac, trial = TrialNr)
 
@@
 # create factor for condition (prime type)
-data_wr <- data_wr %>% mutate(condition = condition + 1,
+data_wr <- data_wr %>% 
+  mutate(condition = condition + 1,
          condition = factor(condition,
@@
 # the latter contains a list of nested data frames
-data_nested <- data_wr %>% group_nest(pid)
+data_nested <- data_wr %>% 
+  group_nest(pid)
 
@@
 # view person-trial-bin data set (column 4) for subject 1:
-pluck(data_final,4,1) %>% print(n=30)
+pluck(data_final,4,1) %>% 
+  print(n=30)
 
 # view lifetable (column 7) for each condition for subject 1
-pluck(data_final, 7, 1) %>% filter(condition=="blank") 
-pluck(data_final, 7, 1) %>% filter(condition=="congruent")
-pluck(data_final, 7, 1) %>% filter(condition=="incongruent")
+pluck(data_final, 7, 1) %>% 
+  filter(condition=="blank") 
+pluck(data_final, 7, 1) %>% 
+  filter(condition=="congruent")
+pluck(data_final, 7, 1) %>% 
+  filter(condition=="incongruent")
 
@@
 median_bin <- map(data_final$lifetable_ca, extract_median)
-binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
+
+binsize    <- data_final$censored[[1]] %>% 
+  pull(bin_width) %>% 
+  max(na.rm=T)
+
 median_rt  <- modify_depth(median_bin,1,~.x*binsize)
-median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
+
+median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% 
+  list_rbind(names_to = "pid")
 median_tb
@@
 # save the plot for each subject
-map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 8, height = 8, dpi = 800)
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), 
+     data_final$plot, ggsave, width = 8, height = 8, dpi = 800)
 ```
@@
 # create single subject number, and unique trial number
-data_agg <- data_wr %>% mutate(oldtrial=trial,
+data_agg <- data_wr %>% 
+  mutate(oldtrial=trial,
          oldpid=pid,
@@
 # apply functions
-data_agg_nested <- data_agg %>% group_nest(pid)
+data_agg_nested <- data_agg %>% 
+  group_nest(pid)
+
 data_agg_final <- data_agg_nested %>% 

### 2024-12-05 Tutorial_1b.Rmd
update tuts
added tidyverse wrangling style and reformatted code chunks

@@
 ```{r read-raw-data}
-raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T) 
+raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", 
+                  delim="\t",
+                  col_names=T) 
 head(raw)
@@
 # remove practice trials
-data_wr <- raw %>% filter(practice==0)
+data_wr <- raw %>% 
+  filter(practice==0)
 
 # remove unnecessary columns
-data_wr <- data_wr %>% select(-c(ss,practice,rd,form,task,primeduration,
+data_wr <- data_wr %>% 
+  select(-c(ss,practice,rd,form,task,primeduration,
             blankduration,maskduration,targetduration,
             void,target_dir,corresp,resperr))
+
 # look at trials where resp==0
@@
 # the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
-sel = which(data_wr$resp==0 & data_wr$rt < 800)
-data_wr <- data_wr[-c(sel),]
+data <- data %>% 
+  filter(!(resp == 0 & rt < 800))
 
@@
 # change some variable names to required names (pid, trial, condition1, condition2, rt, and acc)
-colnames(data_wr) <- c("pid","bl","tr","condition1", "condition2","resp","acc","rt","trial")
+data_wr <- data_wr %>% 
+  rename(pid = vp, condition1 = mask_type, condition2 = prime_type, 
+         acc = respac, trial = TrialNr)
 
@@
 # create a data frame with two columns (pid and data); the latter contains a list of nested data frames
-data_nested <- data_wr %>% group_nest(pid)
+data_nested <- data_wr %>% 
+  group_nest(pid)
+
 # apply the functions using map() and map2()
@@
 median_bin <- map(data_final$lifetable_ca, extract_median_2IV)
-binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
+
+binsize    <- data_final$censored[[1]] %>% 
+  pull(bin_width) %>% 
+  max(na.rm=T)
+
 median_rt  <- modify_depth(median_bin,1,~.x*binsize)
-median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
-median_tb %>% print(n=6*12)
+
+median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% 
+  list_rbind(names_to = "pid")
+
+median_tb %>% 
+  print(n=6*12)
 ```
@@
 # save the plot for each subject
-map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 12, height = 10, dpi = 800)
+map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), 
+     data_final$plot, ggsave, width = 12, height = 10, dpi = 800)
 ```

### 2024-12-08 custom_functions.R
update notes, refs, and custom functions

@@ -327,7 +327,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
                          mask = factor(rep(1:n_iv1, each=2*n_iv2),
                                        levels=c(1:4),
                                        labels=c("nomask", "relevant", "irrelevant","lines")))
-  
+   
   # Remove the first "first_bin_shown - 1" bins from the plots.
   df <- df %>% filter(period >= first_bin_shown)
   
@@ -339,104 +339,104 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
     geom_line(aes(y=hazard)) +
     geom_point(aes(y=hazard), 
                size=1) + 
-    geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), 
-                   show.legend = F) +
+    geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz),
+                  show.legend = F) +
     scale_color_brewer(palette = "Dark2") +
-    scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
-                       labels = c(first_bin_shown:(cutoff/binsize)*binsize),
-                       limits = c(first_bin_shown,cutoff/binsize)) +
+    scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)),
+                      labels = c(first_bin_shown:(cutoff/binsize)*binsize),
+                      limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,haz_yaxis)) +
-    labs(x = "bin endpoints (ms)", 
-         y = "h(t)", 
-         color = "Prime") +
+    labs(x = "bin endpoints (ms)",
+        y = "h(t)",
+        color = "Prime") +
     theme(panel.grid = element_blank(),
-          axis.text.x = element_text(angle=90)) +
-    facet_wrap(~mask,nrow=1,ncol=4) # Apply correctly named factor  
-  
+         axis.text.x = element_text(angle=90)) +
+    facet_wrap(~mask, nrow=1, ncol=4) # Apply correctly named factor
+
   # plot the survivor functions
   p2 <-df %>%
     mutate(mask = condition1,      # Create correctly named factor
            prime = condition2) %>% # Create correctly named factor
     ggplot(aes(x=period, color=prime, group=prime)) + # Apply correctly named factor
[…]

### 2024-12-08 custom_functions.R
update custom plotting functions

@@ -145,7 +145,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
     labs(x = "bin endpoints (ms)", 
          y = "h(t)",
          color = "Condition") +
-    theme(panel.grid = element_blank(),
+    theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
   
   # plot the survivor functions
@@ -175,7 +175,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
     labs(x = "bin endpoints (ms)", 
          y = "S(t)",
          color = "Condition") +
-    theme(panel.grid = element_blank(),
+    theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
   
   # plot the conditional accuracy functions
@@ -195,7 +195,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
     labs(x = "bin endpoints (ms)", 
          y = "ca(t)",
          color = "Condition") +
-    theme(panel.grid = element_blank(),
+    theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
   
   # plot the probability mass functions
@@ -214,7 +214,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
     labs(x = "bin endpoints (ms)", 
          y = "P(t)",
          color = "Condition") +
-    theme(panel.grid = element_blank(),
+    theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
   
   # create title
@@ -231,7 +231,9 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
   
   p1_theme + p2_theme + p3_theme + p4 + 
     plot_annotation(title = title) +
-    plot_layout(guides = "collect") & theme(legend.position = "bottom")
+    plot_layout(guides = "collect", 
+                axes = "collect_x") & 
+    theme(legend.position = "bottom")
  
 }
 
@@ -349,7 +351,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
[…]

### 2024-12-08 Tutorial_1b.Rmd
update tut 1b

@@
 # the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
-data <- data %>% 
+data_wr <- data_wr %>% 
   filter(!(resp == 0 & rt < 800))

### 2024-12-08 manuscript/manuscript.Rmd
update ms. and notes (ms and tut)

@@
 my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)
+
 ```
@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @halleyVIEstimateDegrees1997; @williammatthewmakehamLawMortalityConstruction1860].
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @Halley1693; @williammatthewmakehamLawMortalityConstruction1860].
 The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020].
@@
 Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of EHA a lot quicker and more efficient.
+In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of descriptive EHA a lot quicker and more efficient.
 
@@
 
-Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].
+Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
 We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
@@
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2018].
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2020].
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 We used `r my_r_citation$r` for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2018, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 
-We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2018; @kurzAppliedLongitudinalDataAnalysis2023].
+We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2020; @kurzAppliedLongitudinalDataAnalysis2023].
 

### 2024-12-08 manuscript/manuscript.tex
update ms. and notes (ms and tut)

@@
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1997; Makeham, William M., 1860).
+Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1693; Makeham, 1860).
 The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2018).
+Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
@@
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2018).
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020).
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 
-We used R (Version 4.4.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, n.d.), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Heiss (2021), Kurz (2023a), and Kurz (2023b).
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 
@@
 
-{\centering \includegraphics[height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[width=21.33in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 
-We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2018).
+We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2020).
 
@@
 \bibitem[\citeproctext]{ref-R-standist}
-Girard, J. (n.d.). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
+Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
 
@@
 
-\bibitem[\citeproctext]{ref-halleyVIEstimateDegrees1997}
+\bibitem[\citeproctext]{ref-Halley1693}
 Halley, E.
-(1997).
-{VI}.
-{An} estimate of the degrees of the mortality of mankind; drawn from curious tables of the births and funerals at the city of {Breslaw}; with an attempt to ascertain the price of annuities upon lives.
+(1693).
+VI.
+An estimate of the degrees of the mortality of mankind; drawn from curious tables of the births and funerals at the city of breslaw; with an attempt to ascertain the price of annuities upon lives.
 \emph{Philosophical Transactions of the Royal Society of London}, \emph{17}(196), 596--610.
@@
 \bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-Makeham, William M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
-
-\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2018}
-McElreath, R. (2018). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {Stan}} (1st ed.). {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9781315372495}
+Makeham, W. M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
+
+\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
+McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
 
@@
 \bibitem[\citeproctext]{ref-R-dplyr}
-Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}
+Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
 
 \bibitem[\citeproctext]{ref-R-purrr}
-Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
 
 \bibitem[\citeproctext]{ref-R-readr}
-Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
+Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
 
 \bibitem[\citeproctext]{ref-R-tidyr}
-Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
+Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
 

### 2024-12-08 notes/manuscript_notes.Rmd
update notes, refs, and custom functions
update ms. and notes (ms and tut)

@@
 
+-> Done.
+
 ### section 3.1 ###
@@
 
+-> Done.
+
 ### section 3.2 ###
@@
 first sentence, the McElreath reference should be 2020 for the textbook and not 2018.
+
+-> Done. Here and elsewhere...
+
+Other refs to add:
+
+Lakens, D. (2022). Sample Size Justification. Collabra: Psychology, 8(1), 33267. https://doi.org/10.1525/collabra.33267
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C.
+C., Carpenter, B., Yao, Y., Kennedy, L., Gabry, J., Bürkner, P.-C., & Modrák, M.
+(2020).
+Bayesian Workflow (arXiv:2011.01808).
+arXiv.
+http://arxiv.org/abs/2011.01808
+Kurz, A. S. https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/
+DeBruine, L.
+M., & Barr, D.
+J.
+(2021).
+Understanding Mixed-Effects Models Through Data Simulation.
+Advances in Methods and Practices in Psychological Science, 4(1), 2515245920965119.
+https://doi.org/10.1177/2515245920965119
+
+-> Done.
 
@@
 
+-> Model M1d removed from main text.
+
 ### section 4.3.3 ###
@@
 we need to define index coding.
+
+-> McElreath book... def
 
@@
 
+->
+
 Figure 5 y-axis. Should it just be cloglog? Rather than cloglog-hazard?
@@
 I'd like to see it at least.
+
+->
 
@@
 
+->
+
 on page 38, you say:
@@
 
+-> I will remove it.
+
 Figure 6.
@@
 
+->
+
 Figure 9. spell out AMEs in full to avoid jargon.
+
+->
 

### 2024-12-08 Tutorial_2a.Rmd
update ms and tut 2a

@@
 For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the following model (M1d).
+
+### 4.3.4 Model M1d: A more parsimonious model
+
+##
+When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
+Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").
+
+For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called "M1d".
+Here is its model formula.
+
+```{r fit-model-M1d, eval=F, echo=T}
+event ~ 0 + Intercept + condition*period_9 + condition*period_9_sq + 
+                   (1 + condition*period_9 + condition*period_9_sq | pid)
+```
+
+The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
+The variable period_9_sq is a squared version of period_9.
+Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 184 minutes to run.
+##
 

### 2024-12-08 manuscript/manuscript.Rmd
update ms and tut 2a

@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
-Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
-And when you treat a variable as a discrete predictor, you can choose between reference coding and index coding.
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor.
+And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding.
+With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level.
+Brm() will then construct k-1 indicator variables (see model M1d in Tutorial_2a.Rmd for an example).
+With index coding, one constructs an index variable which contains integers that correspond to different categories (see models M0i and M1i below).
+As explained by @mcelreathStatisticalRethinkingBayesian2020, the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.
 
@@
 We decided to select the analysis time window (200,600] and the cloglog link.
-Below, we shortly discuss three of these models.
+Below, we shortly discuss two of these models.
 The person-trial-bin data set is prepared as follows.
@@
 filter(period > 5) %>%
-       # continuous predictor for TIME named "period_9", centered on bin 9
-mutate(period_9 = period - 9,
-       # categorical predictor for TIME named "timebin" with index coding
-       timebin = factor(period, levels = c(6:15)),
+       # define categorical predictor TIME as index variable named timebin
+mutate(timebin = factor(period, levels = c(6:15)),
        # factor "condition" using reference coding, with "blank" as the reference level
@@
 
-### 4.3.4 Model M1d: A more parsimonious model
-
-When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
-Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").
-
-For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called "M1d".
-Here is its model formula.
-
-```{r fit-model-M1d, eval=F, echo=T}
-event ~ 0 + Intercept + condition*period_9 + condition*period_9_sq + 
-                   (1 + condition*period_9 + condition*period_9_sq | pid)
-```
-
-The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
-The variable period_9_sq is a squared version of period_9.
-Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 184 minutes to run.
-
 ### 4.3.5 Compare the models.
 
-We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2020; @kurzAppliedLongitudinalDataAnalysis2023].
+We can compare the two models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2020; @kurzAppliedLongitudinalDataAnalysis2023].
 
@@
 model_M1i <- readRDS("../Tutorial_2_Bayesian/models/model_M1i.rds") 
-model_M1d <- readRDS("../Tutorial_2_Bayesian/models/model_M1d.rds")
+#model_M1d <- readRDS("../Tutorial_2_Bayesian/models/model_M1d.rds")
 
@@
 model_M1i <- add_criterion(model_M1i, c("loo", "waic"))
-model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
+#model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
 ```
@@
 ```{r show-weights, echo=T, eval=T}
-model_weights(model_M0i, model_M1i, model_M1d, weights = "loo") %>% round(digits = 2)
-model_weights(model_M0i, model_M1i, model_M1d, weights = "waic") %>% round(digits = 2) 
+model_weights(model_M0i, model_M1i, weights = "loo") %>% round(digits = 2)
+model_weights(model_M0i, model_M1i, weights = "waic") %>% round(digits = 2) 
 ```
@@
 
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other two simpler models.
-
-### 4.3.6 Interpreting model M1i
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other simpler model.
+
+### 4.3.6 Evaluating parameter estimates in model M1i
 

### 2024-12-08 manuscript/manuscript.tex
update ms and tut 2a

@@
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a discrete or continuous predictor.
-And when you treat a variable as a discrete predictor, you can choose between reference coding and index coding.
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor.
+And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding.
+With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level.
+Brm() will then construct k-1 indicator variables (see model M1d in Tutorial\_2a.Rmd for an example).
+With index coding, one constructs an index variable which contains integers that correspond to different categories (see models M0i and M1i below).
+As explained by McElreath (2020), the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.
 
@@
 We decided to select the analysis time window (200,600{]} and the cloglog link.
-Below, we shortly discuss three of these models.
+Below, we shortly discuss two of these models.
 The person-trial-bin data set is prepared as follows.
@@
 \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-       \CommentTok{\# continuous predictor for TIME named "period\_9", centered on bin 9}
-\FunctionTok{mutate}\NormalTok{(}\AttributeTok{period\_9 =}\NormalTok{ period }\SpecialCharTok{{-}} \DecValTok{9}\NormalTok{,}
-       \CommentTok{\# categorical predictor for TIME named "timebin" with index coding}
-       \AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
+       \CommentTok{\# define categorical predictor TIME as index variable named timebin}
+\FunctionTok{mutate}\NormalTok{(}\AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
        \CommentTok{\# factor "condition" using reference coding, with "blank" as the reference level}
@@
 
-\subsubsection{4.3.4 Model M1d: A more parsimonious model}\label{model-m1d-a-more-parsimonious-model}
-
-When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
-Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable ``condition'') and treat TIME as a continuous variable (variable ``period\_9'').
-
-For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called ``M1d''.
-Here is its model formula.
+\subsubsection{4.3.5 Compare the models.}\label{compare-the-models.}
+
+We can compare the two models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2020).
+
+\footnotesize
 
@@
 \begin{Highlighting}[]
-\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ Intercept }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9\_sq }\SpecialCharTok{+}
-\NormalTok{                   (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9 }\SpecialCharTok{+}\NormalTok{ condition}\SpecialCharTok{*}\NormalTok{period\_9\_sq }\SpecialCharTok{|}\NormalTok{ pid)}
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 
-The specification ``0 + Intercept + \ldots{}'' removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
-The variable period\_9\_sq is a squared version of period\_9.
-Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 184 minutes to run.
-
-\subsubsection{4.3.5 Compare the models.}\label{compare-the-models.}
-
-We can compare the three models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2020).
-
-\footnotesize
+\begin{verbatim}
+## model_M0i model_M1i 
+##         0         1
+\end{verbatim}
 
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }
 \end{Highlighting}
@@
 \begin{verbatim}
-## model_M0i model_M1i model_M1d 
-##         0         1         0
+## model_M0i model_M1i 
+##         0         1
 \end{verbatim}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, model\_M1d, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M0i model_M1i model_M1d 
-##         0         1         0
-\end{verbatim}
-
 \normalsize
 
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other two simpler models.
-
-\subsubsection{4.3.6 Interpreting model M1i}\label{interpreting-model-m1i}
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other simpler model.
+
+\subsubsection{4.3.6 Evaluating parameter estimates in model M1i}\label{evaluating-parameter-estimates-in-model-m1i}
 

### 2024-12-08 notes/manuscript_notes.Rmd
update ms and tut 2a

@@
 
--> McElreath book... def
+-> Done in section 4.3.1.
 
@@
 
+-> I guess the addition in section 4.3.1 is clear enough now.
+
 ### section 4.3.6 ###
@@
 Figure 5 y-axis. Should it just be cloglog? Rather than cloglog-hazard?
+
+-> cloglog-hazard is better I guess.
 
@@
 
+-> Done. Title changed for 4.3.6.
+
 This figures in this section are wonderful. Nice work!
@@
 on page 38, you say:
-
 >Note that we could also have calculated hazard ratios instead of hazard differences.
-
 This feels a little disjointed from what comes before in the paragraph.
@@
 This makes the orientation similar to the prior plots and will therefore be easier to follow for the reader.
+
+# Sven dec 8
+
+I also moved prior predictive checks to last section of 2a.
+
+Mention 2b ppc...
+
+Section 4...
+
+Tutorial 2a ... M1d...

### 2024-12-09 Tutorial_2a.Rmd
update tut 2a rmd and html

@@
 
-Second, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach and treat TIME as a continuous variable.
-For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the following model (M1d).
-
-### 4.3.4 Model M1d: A more parsimonious model
-
-##
-When the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
+Second, when the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
 Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").
@@
 For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the model called "M1d".
-Here is its model formula.
-
-```{r fit-model-M1d, eval=F, echo=T}
-event ~ 0 + Intercept + condition*period_9 + condition*period_9_sq + 
-                   (1 + condition*period_9 + condition*period_9_sq | pid)
-```
-
-The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
-The variable period_9_sq is a squared version of period_9.
-Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model M1d took about 184 minutes to run.
-##
+
+Get the data, and set priors.
 
@@
 ```
+
+Fit model M1d.
 
@@
 
-Model_M1d took about 184 minutes to run. Note that duplicate terms in the model formula (e.g., condition) are ignored.
+The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves.
+The variable period_9_sq is a squared version of period_9.
+Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition.
+Model_M1d took about 184 minutes to run.
+Note that duplicate terms in the model formula (e.g., condition) are ignored.
 
@@
 
-```{r save-epred-pid-new, eval=F}
+```{r save-epred-pid-6-new, eval=F}
 ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new_6.png", width = 10, height = 8, dpi = 800)
@@
 
+Save as table.
+
 ```{r save-table-grand-AMEs, eval=F}
@@
 ```
+
+Show table.
+
+Plot.
 
@@
 
+Save as table.
+
 ```{r save-table-pid-AMEs, eval=F}

### 2024-12-09 manuscript/manuscript.Rmd
update ms. and ms-notes

@@
 
-The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman et al., 2020).
+The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it [@gelmanBayesianWorkflow2020].
 The basic structure and code follows the examples outlined by Solomon Kurz in his
-'power' blog posts (Kurz, 2019) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper (DeBruine & Barr, 2021).
+'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper [@debruineUnderstandingMixedEffectsModels2021].
 
@@
 
-We again use the data from Panis & Schmidt (2016) to provide a worked example.
+We again use the data from @panisWhatShapingRT2016 to provide a worked example.
 We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
@@
 
-(ref:simdat-plot15-caption) Raw data from Panis & Schmidt (2016) and simulated data from 10 participants.
+(ref:simdat-plot15-caption) Raw data from @panisWhatShapingRT2016 and simulated data from 10 participants.
 
@@
 How many trials per condition should we collect and how many participants should we test? 
-Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022). 
+Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources [@lakensSampleSizeJustification2022]. 
 There is no straigtforward and clear-cut answer.
@@
 If we were running this kind of study in our lab, what would we do?
-We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon (Panis & Schmidt, 2016).
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon [@panisWhatShapingRT2016].
 Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80% power mark.
@@
 Instead, we would run a follow-up experiment that replicates and extends the initial result. 
-By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis. 
+By doing so, we would aim to avoid the Cult of the Isolated Single Study [@nelderStatisticsStatisticalScience1999; @tongStatisticalInferenceEnables2019], and thus reduce the reliance on any one type of planning tool, such as a power analysis.
 Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study on its own has sufficient evidential value to hit some criterion mark.

### 2024-12-09 manuscript/manuscript.tex
update ms. and ms-notes

@@
 
-The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman et al., 2020).
+The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
 The basic structure and code follows the examples outlined by Solomon Kurz in his
-`power' blog posts (Kurz, 2019) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
+`power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
 
@@
 
-We again use the data from Panis \& Schmidt (2016) to provide a worked example.
+We again use the data from Panis and Schmidt (2016) to provide a worked example.
 We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
@@
 
-\caption{Raw data from Panis \& Schmidt (2016) and simulated data from 10 participants.}\label{fig:simdat}
+\caption{Raw data from Panis and Schmidt (2016) and simulated data from 10 participants.}\label{fig:simdat}
 \end{figure}
@@
 Instead, we would run a follow-up experiment that replicates and extends the initial result.
-By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1986; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis.
+By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1999; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis.
 Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study on its own has sufficient evidential value to hit some criterion mark.
@@
 
+\bibitem[\citeproctext]{ref-debruineUnderstandingMixedEffectsModels2021}
+DeBruine, L.
+M., \& Barr, D.
+J.
+(2021).
+Understanding {Mixed-Effects Models Through Data Simulation}.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920965119.
+\url{https://doi.org/10.1177/2515245920965119}
+
 \bibitem[\citeproctext]{ref-R-Rcpp_b}
@@
 
+\bibitem[\citeproctext]{ref-lakensSampleSizeJustification2022}
+Lakens, D. (2022). Sample {Size Justification}. \emph{Collabra: Psychology}, \emph{8}(1), 33267. \url{https://doi.org/10.1525/collabra.33267}
+
 \bibitem[\citeproctext]{ref-landesIntroductionEventHistory2020}
@@
 Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}
+
+\bibitem[\citeproctext]{ref-nelderStatisticsStatisticalScience1999}
+Nelder, J.
+A.
+(1999).
+From {Statistics} to {Statistical Science}.
+\emph{Journal of the Royal Statistical Society.
+Series D (The Statistician)}, \emph{48}(2), 257--269.
+Retrieved from \url{https://www.jstor.org/stable/2681191}
 
@@
 
+\bibitem[\citeproctext]{ref-tongStatisticalInferenceEnables2019}
+Tong, C.
+(2019).
+Statistical {Inference Enables Bad Science}; {Statistical Thinking Enables Good Science}.
+\emph{The American Statistician}, \emph{73}(sup1), 246--261.
+\url{https://doi.org/10.1080/00031305.2018.1518264}
+
 \bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}

### 2024-12-09 Tutorial_2a.Rmd
update 2a

@@
          "brms", "tidybayes", "bayesplot", "future", "parallel", "modelr",
-         "rstan")
+         "rstan", "knitr")
 
@@
 Show table.
+
+```{r}
+kable(table, caption = "Point (mean) and 95% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.")
+```
 

### 2024-12-09 manuscript/manuscript.Rmd
update ms

@@
 
-```{r table-grand-ame}
+```{r table-grand-ame, eval = F}
  apa_table(

### 2024-12-09 manuscript/manuscript.tex
update ms

@@
 
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{TableNotes}[para]
-\normalsize{\textit{Note.} diff = difference in predicted hazard.}
-\end{TableNotes}
-
-\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table-grand-ame}Point (mean) and 95\% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.}\\
-\toprule
-contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:table-grand-ame} continued}}\\
-\toprule
-contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
-\midrule
-\endhead
-congruent minus blank & 6 & 0.09 & 0.02 & 0.17\\
-congruent minus blank & 7 & 0.14 & 0.04 & 0.25\\
-congruent minus blank & 8 & 0.11 & -0.02 & 0.24\\
-congruent minus blank & 9 & 0.08 & -0.12 & 0.27\\
-congruent minus blank & 10 & -0.08 & -0.30 & 0.15\\
-congruent minus blank & 11 & -0.10 & -0.31 & 0.11\\
-congruent minus blank & 12 & -0.04 & -0.24 & 0.17\\
-congruent minus blank & 13 & -0.08 & -0.33 & 0.20\\
-congruent minus blank & 14 & -0.10 & -0.33 & 0.15\\
-congruent minus blank & 15 & -0.01 & -0.27 & 0.27\\
-incongruent minus blank & 6 & 0.09 & 0.01 & 0.21\\
-incongruent minus blank & 7 & 0.08 & -0.03 & 0.23\\
-incongruent minus blank & 8 & -0.06 & -0.17 & 0.07\\
-incongruent minus blank & 9 & -0.19 & -0.31 & -0.06\\
-incongruent minus blank & 10 & -0.27 & -0.45 & -0.04\\
-incongruent minus blank & 11 & -0.23 & -0.40 & -0.03\\
-incongruent minus blank & 12 & -0.17 & -0.33 & 0.02\\
-incongruent minus blank & 13 & -0.10 & -0.31 & 0.14\\
-incongruent minus blank & 14 & -0.06 & -0.27 & 0.15\\
-incongruent minus blank & 15 & 0.03 & -0.23 & 0.27\\
-\bottomrule
-\addlinespace
-\insertTableNotes
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
 \paragraph{Example conclusions for M1i}\label{example-conclusions-for-m1i}

### 2024-12-09 Tutorial_2a.Rmd
update 2a
update tuts

@@
 The skew_normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that hazard values above .5 are less likely than values below .5.
+A prior predictive check is performed in section 8.1.
 
@@
 
-### Intermezzo: Prior Predictive Check for M0i.
-
-Following Gelman et al.
-(2020), we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
-First, sample the prior distributions using sample_prior="only".
-
-```{r M0i-prior, eval=F}
-plan(multicore)
-
-model_M0i_prior <-                  
-   brm(data = data_M0i,
-       family = bernoulli(link="cloglog"),
-       formula = event ~ 0 + timebin + (0 + timebin | pid),
-       prior = priors_M0i,
-       chains = 1, cores = 1, 
-       iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, 
-                      step_size = 0.04, 
-                      max_treedepth = 12),
-       seed = 12, init = "0",
-       sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0i_prior")
-```
-
-```{r inspect-M0i-prior}
-model_M0i_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0i_prior.rds")
-```
-
-Next, use them to predict prior data. But to better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.
-
-```{r ppc-manual, eval=F}
-# prior predictive check for 1 time bin
-# see http://bruno.nicenboim.me/bayescogsci/
-
-# cloglog and inverse-cloglog link functions
-inverse_cloglog <- function(x) {return(1-(exp(-1*exp(x))))}
-cloglog <- function(x) {return(log(-1*log(1-x)))}
-
-# Specify
-N_samples <- 2000 # the number of samples from the skew_normal
-N_obs <- 100      # the number of simulated observations per sample
-set.seed(1)
-
-# function to generate binary observations from samples
-cloglog_predictive_distribution <- function(samples, Nobs) {
-
-    # empty data frame with headers
-    df_pred <- tibble(trial = numeric(0),
-                      y_pred = numeric(0),
-                      iter = numeric(0))
-    # i iterates from 1 to the length of samples
-    for (i in seq_along(samples)) {
-      cloglog_haz <- samples[i]
-      df_pred <- bind_rows(df_pred,
-          tibble(trial = 1:N_obs,
-                 y_pred = rbinom(Nobs, 1, inverse_cloglog(cloglog_haz)),
-                 iter = i))
-    }
-    df_pred 
- }
-
-# sample prior cloglog-hazard values for 1 bin
-cloglog_samples <- rskew_normal(N_samples,-1,1,-2)
-
-# apply the function
-prior_pred <- cloglog_predictive_distribution(cloglog_samples,
-                                              N_obs)
-
-# calculate hazard as the mean over binary observations
-prior_pred_haz <- prior_pred %>% 
-  group_by(iter) %>% 
-  summarise(pred_haz = mean(y_pred))
-
-# plot 
-ggplot(prior_pred_haz, aes(x=pred_haz)) +
-  geom_histogram(binwidth=.01) +
-  scale_x_continuous(limits = c(0,1)) +
-  labs(x = "simulated hazard",
-       title = "Prior predictive distribution",
-       subtitle = str_c(N_samples," samples; ",
-                        N_obs," observations per sample"))
-```
-
-```{r save-ppc, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png", width = 10, height = 8, dpi = 800)
-```
-
-```{r plot-ppc-1-bin, out.width='80%'}
-knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png")
-```
-
-The shape of the prior predictive distribution for 1 bin looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
-
-Now use add_predicted_draws() to simulate prior-based data for our 10 bins and 6 subjects, while taking into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
-
-```{r ppc-M0i-prior, eval=F}
-# Generate prior predictive hazard functions for 6 participants
-newdata_prior = tibble(timebin = 6:15) %>% 
-           expand_grid(pid = 1:6)
-
-# Specify
-N_obs = 100
-set.seed(2)
-
-# prepare columns of data set
-df_pred <- tibble(timebin = integer(0),
-                  pid = integer(0),
-                  .row = integer(0),
-                  .chain = integer(0),
-                  .iteration = integer(0),
-                  .draw = integer(0),
-                  .prediction = integer(0),
-                  obs = integer(0))
-
-# Call add_predicted_draws() for each simulated observation
-for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(model_M0i_prior, 
-                                    newdata=newdata_prior, 
-                                    summary=F, 
-                                    ndraws=NULL) %>% # all 2000
-                mutate(obs = i)
-  df_pred <- bind_rows(df_pred, prior_pred)
-}
-
-# calculate hazard per draw (average across observations)
-df_pred_haz <- df_pred %>% 
-  group_by(pid,timebin,.draw) %>% 
-  summarise(pred_haz = mean(.prediction))
-
-# plot 200 draws per participant
-ggplot(data = df_pred_haz %>% filter(.draw < 200), 
-       aes(x=timebin, y=pred_haz, group=.draw)) +
-   geom_line(color="black") +
-   geom_line(data=df_pred_haz %>% filter(.draw == 20), 
-             color="red", 
-             linewidth=2) +
-  scale_y_continuous(limits = c(0,1)) +
-  scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
-  labs(x = "time bin",
-       y = "simulated hazard",
-       title = "Prior predictive distributions",
-       subtitle = str_c("200 samples; ",N_obs," observations per sample")) +
-  facet_wrap(~pid)
-```
-
-```{r save-ppc-ten-bins-six-subjects, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0i_ppc.png", width = 10, height = 8, dpi = 800)
-```
-
-```{r plot-ppc-M0i, out.width='80%'}
-knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc.png")
-```
-
-The red hazard function is one of the 200 functions plotted, for each of six participants. They can take on any shape, with values below .5 being more likely than values above .5.
-
-After our prior predictive check, we can fit model M0i to get the posterior distributions.
+Fit model M0i.
 
@@
 
-Prepare the data file, and specify priors.
+Prepare the data file, and specify priors. To find the priors, we performed a prior predictive check which is reported in section 8.2.
 
@@
 
-### Intermezzo. Prior predictive check for model M0c.
-
-Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs.
-
-To get an idea of how the intercept and both slopes control the shape of the hazard function, the following code allows to change the coefficients and generate (cloglog-)hazard functions.
-Once you have found the coefficients for the intercept, period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to determine a suited mean and standard deviation for each coefficient.
-
-```{r search-coef}
-dat <- tibble(x = -3:6,
-              x2 = x^2,
-              cloglog1 = -1.8   + 0.3*x    - 0.1*x2,  # black
-              cloglog2 = -1     + 0.41*x   - 0.04*x2,  # green
-              cloglog3 = -3     - 0.13*x   + 0.12*x2, # red
-              cloglog4 = -0.40  + 0.6*x    - 0.11*x2, # blue
-              haz1 = 1 - exp(-1*exp(cloglog1)),
-              haz2 = 1 - exp(-1*exp(cloglog2)),
-              haz3 = 1 - exp(-1*exp(cloglog3)),
-              haz4 = 1 - exp(-1*exp(cloglog4)))
-
-p1<-ggplot(dat, aes(x=x)) +  
-      geom_line(aes(y=cloglog1), color="black") +
-      geom_line(aes(y=cloglog2), color="green") +
-      geom_line(aes(y=cloglog3), color="red") +
-      geom_line(aes(y=cloglog4), color="blue") +
-      scale_y_continuous(limits=c(-6,2)) +
-      scale_x_continuous(breaks=c(-3:6)) +
-      labs(x = "timebin", y = "cloglog-hazard")
-
-p2<-ggplot(dat, aes(x=x))+
-      geom_line(aes(y=haz1), color="black") + 
-      geom_line(aes(y=haz2), color="green") +
-      geom_line(aes(y=haz3), color="red") +
-      geom_line(aes(y=haz4), color="blue") +
-      scale_y_continuous(limits=c(0,1)) +
-      scale_x_continuous(breaks=c(-3:6)) +
-      labs(x = "timebin", y = "hazard")
-
-p1|p2
-```
-
-For example, the coefficients for x vary from -0.13 to 0.6, and we might specify N(.3,.25) as a prior distribution, Those for x-squared vary from -0.11 to 0.12, and we might specify N(0,.06) as a prior distribution.
-
-Suppose we expect (cloglog-)hazard functions that increase almost linearly with time.
-We also set the standard deviation of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
-
-```{r priors-M0c-ppc1}
-priors_M0c_ppc1 <- c(
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
-  set_prior("normal(0.3,.001)", class = "b", coef = "period_9"), 
-  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
-  set_prior("normal(0, .01)", class = "sd"), 
-  set_prior("lkj(2)", class = "cor")  
-)
-```
-
-Sample the prior distributions.
-
-```{r fit-model-M0c-prior1, eval=F}
-plan(multicore)
-
-model_M0c_prior1 <-
-   brm(data = data_M0c,
-       family = bernoulli(link="cloglog"),
-       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
-                          (0 + Intercept + period_9 + period_9_sq | pid),
-       prior = priors_M0c_ppc1,
-       chains = 1, cores = 1, 
-       iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, 
-                      step_size = 0.04, 
-                      max_treedepth = 12),
-       seed = 12, init = "0",
-       sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0c_prior1")
-```
-
-```{r load-M0c-prior1}
-model_M0c_prior1 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior1.rds")
-```
-
-Check the prior distributions.
-
-```{r prior1-distr}
-# extract prior draws
-post <- as_draws_df(model_M0c_prior1) %>% 
-  select(-lp__) %>% 
-  as_tibble()
-
-tidy_prior <- post %>% 
-  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
-  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
-  pivot_longer(-c(chain, draw, iter))
-
-# plot
-ggplot(tidy_prior, aes(x = name, y = value, fill = name)) +  
-  stat_halfeye(alpha=0.7, 
-               point_interval = "median_qi",
-               .width=c(.8,.99), 
-               show.legend = F) +
-  labs(x = "parameter",
-       title = "Samples from the prior distributions")
-```
-
-Next, generate prior predictive check for each participant.
-
-Create a function.
-
-```{r function-plot-ppc, eval=F}
-plot_ppc <- function(fit_prior, N_samples, N_obs, new_data_prior){
-
-set.seed(1)
-
-# prepare columns of data set with predictions
-df_pred <- tibble(period_9 = integer(0),
-                  period_9_sq = integer(0),
-                  pid = integer(0),
-                  .row = integer(0),
-                  .chain = integer(0),
-                  .iteration = integer(0),
-                  .draw = integer(0),
-                  .prediction = integer(0),
-                  obs = integer(0))
-
-# extract predictions
-for(i in 1:N_obs){
-  prior_pred <- add_predicted_draws(fit_prior, 
-                                    newdata = new_data_prior, 
-                                    summary=F, 
-                                    ndraws=N_samples) %>% 
-                mutate(obs = i)
-  df_pred <- bind_rows(df_pred, prior_pred)
-}
-
-# calculate hazard per draw (average across obs)
-df_pred_haz <- df_pred %>% 
-  group_by(pid,period_9,.draw) %>% 
-  summarise(pred_haz = mean(.prediction))
-
-# plot N_samples draws per participant, highlighting three
-sel = sample(1:N_samples,3,replace=F)
-
-ggplot(df_pred_haz %>% 
-   filter(.draw <N_samples), aes(x=period_9, y=pred_haz, group=.draw)) +
-   geom_line(color="black") +
-   # highlight
-   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
-             color="red", linewidth=2) +
-   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
-            color="green", linewidth=2) +
-   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
-          color="blue", linewidth=2) +
-  scale_y_continuous(limits = c(0,1)) +
-  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
-  facet_wrap(~pid)
-}
-```
-
-Apply the function.
-
-```{r eval=F}
-# make new data
-newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2) %>%
-  expand_grid(pid = 1:6)
-
-# make ppc plot
-p1 <- plot_ppc(model_M0c_prior1,2000,100,newdata_prior)
-p1
-```
-
-For each of 2000 draws from the prior distributions, a black hazard function is plotted. Three randomly selected hazard functions are highlighted in red, green, and blue.
-
-```{r save-ppc1-M0c, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0c_ppc1.png", width = 10, height = 8, dpi = 800)
-```
-
-```{r plot-ppc1-M0c, out.width='80%'}
-knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc1.png")
-```
-
-Or suppose we believe apriori that the hazard functions are flat.
-We specify our new priors:
-
-```{r priors-M0c-ppc2}
-priors_M0c_ppc2 <- c(
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
-  set_prior("normal(0,.001)", class = "b", coef = "period_9"), 
-  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
-  set_prior("normal(0, .001)", class = "sd"), 
-  set_prior("lkj(2)", class = "cor")  
-)
-```
-
-Sample the prior distributions 2000 times:
-
-```{r fit-model-M0c-prior2, eval=F}
-plan(multicore)
-
-model_M0c_prior2 <-
-   brm(data = data_M0c,
-       family = bernoulli(link="cloglog"),
-       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
-                          (0 + Intercept + period_9 + period_9_sq | pid),
-       prior = priors_M0c_ppc2,
-       chains = 1, cores = 1, 
-       iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, 
-                      step_size = 0.04, 
-                      max_treedepth = 12),
-       seed = 12, init = "0",
-       sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0c_prior2")
-```
-
-```{r load-M0c-prior2}
-model_M0c_prior2 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior2.rds")
-```
-
-Apply the function plot_ppc().
-
-```{r eval=F}
-# make new data
-newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2) %>%
-  expand_grid(pid = 1:6)
-
-# make ppc plot
-p2 <- plot_ppc(model_M0c_prior2,2000,100,newdata_prior)
-p2
-```
-
-Now the prior predictive hazard functions are all rather flat, with an intercept controlled by the prior skew-normal distribution, i.e., values below .5 are more likely than values above .5.
-
-```{r save-ppc2-M0c, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0c_ppc2.png", width = 10, height = 8, dpi = 800)
-```
-
-```{r plot-ppc2-M0c, out.width='80%'}
-knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc2.png")
-```
-
-Finally, perform a ppc for the selected prior distributions for M0c (see lines 369-374), and increase the standard deviation for the normal prior for class "sd" somewhat.
-
-```{r priors-M0c-ppc3}
-priors_M0c_ppc3 <- c(
-  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
-  set_prior("normal(0.3,.25)", class = "b", coef = "period_9"), 
-  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
-  set_prior("normal(0, .1)", class = "sd"),  
-  set_prior("lkj(2)", class = "cor")  
-)
-```
-
-```{r fit-model-M0c-prior3, eval=F}
-plan(multicore)
-
-model_M0c_prior3 <-
-   brm(data = data_M0c,
-       family = bernoulli(link="cloglog"),
-       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
-                          (0 + Intercept + period_9 + period_9_sq | pid),
-       prior = priors_M0c_ppc3,
-       chains = 1, cores = 1, 
-       iter = 3000, warmup = 1000,
-       control = list(adapt_delta = 0.999, 
-                      step_size = 0.04, 
-                      max_treedepth = 12),
-       seed = 12, init = "0",
-       sample_prior = "only",
-       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
-```
-
-```{r load-M0c-prior3}
-model_M0c_prior3 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior3.rds")
-```
-
-Apply the function plot_ppc().
-
-```{r eval=F}
-# make new data
-newdata_prior = tibble(period_9 = c(-3:6)) %>% 
-  mutate(period_9_sq = period_9^2) %>%
-  expand_grid(pid = 1:6)
-
-# make ppc plot
-p3 <- plot_ppc(model_M0c_prior3,2000,100,newdata_prior)
-p3
-```
-
-A wide range of differently shaped hazard functions can be observed, and early bins are unlikely to have high hazard values, consistent with our prior beliefs. 
-Note that some simulated hazard functions start off with a high hazard in bin -3.
-These reflect a combination of negative slopes and a positive coefficient for period_9_sq.
-These possibilities are required to encompass hazard functions that initially decrease and then increase, as observed in the empirical hazard functions for incongruent primes.
-
-```{r save-ppc3-M0c, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M0c_ppc3.png", width = 10, height = 8, dpi = 800)
-```
-
-```{r plot-ppc3-M0c, out.width='80%'}
-knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc3.png")
-```
-
-After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions.
+Fit model M0c.
 
@@
 
+#
+```{r plot-post-M1i-alt}
+tidy_fixed <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) %>% 
+  mutate(timebin = str_sub(name,10,11),
+         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
+         condition = str_sub(name,17,18),
+         condition = factor(str_remove(condition,"e"),
+                            levels=c(1,2,3),
+                            labels=c("blank","congruent","incongruent")))
+head(tidy_fixed)
+tail(tidy_fixed)
+
+# plot
+p_tidy_fixed <- ggplot(tidy_fixed, aes(x = timebin, y = value, fill=condition)) +  
+  stat_halfeye(point_interval="median_qi", 
+               .width = c(0.80,0.95),
+               alpha=0.6,
+               position=pd) +
+  labs(title = 'Posterior distributions for population-level\neffects in Model M1i',
+       x = "time bin", y = "cloglog-hazard") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+   theme(axis.text.x = element_text(angle=90)) 
+#+
+#  facet_wrap(~condition)
+p_tidy_fixed
+```
+
+```{r save-post, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_postdistr_alt3.png", width = 10, height = 8, dpi = 800)
+```
+
+#
+
 ```{r save-post, eval=F}
@@
 
-Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, for each participant in the data set, and for a brand new hypothetical participant.
+Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, and for each participant in the data set.
 
@@
 ```{r epred-grand-plot}
- ggplot(epreds_grand, aes(x=timebin, y=.epred, 
+ p1 <- ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                           fill=prime, color=prime)) +
@@
     theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(0,.6)) +
-    labs(y = "predicted hazard") +
-    ggtitle("Average participant")
-```
-
-```{r save-epred-grand, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pred_grand.png", width = 10, height = 8, dpi = 800)
+    scale_y_continuous(limits=c(0,.7)) +
+    labs(y = "predicted hazard")
+
+p1
 ```
@@
 ```{r epred-pid-plot}
- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
+p2 <- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                        fill=prime, color=prime)) +
@@
     labs(y = "predicted hazard") +
-    ggtitle("Each participant (N = 6)") +
+   # ggtitle("Each participant (N = 6)") +
     facet_wrap(~pid, ncol=3)
-```
-
-```{r save-epred-pid, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid.png", width = 10, height = 8, dpi = 800)
-```
-
-Third, for a completely new hypothetical participant.
-
-```{r epred-new-pid}
-epreds_pid_new <- dat_M1i %>% 
-  data_grid(timebin, prime) %>%
-  mutate(pid = 7) %>%
-  add_epred_draws(model_M1i, 
-                  re_formula = NULL, 
-                  allow_new_levels=T,    
-                  sample_new_levels = "gaussian") %>% 
-  mutate(prime = factor(prime, 
-                        levels=c(1,2,3),
-                        labels=c("blank","congruent","incongruent"))) %>%
-  ungroup()
-```
-
-Summarize and plot predicted hazard values.
-
-```{r epred-new-pid-summary}
-epreds_pid_new %>% 
-  group_by(pid,timebin,prime) %>% 
-  median_qi(.width=.95)
-```
-
-```{r epred-new-pid-plot}
- ggplot(epreds_pid_new, aes(x=timebin, y=.epred, 
-                       fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",
-                    .width=.95,
-                    alpha = 0.5) +
-    scale_fill_brewer(palette = "Dark2") +
-    scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-    theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted hazard") +
-    ggtitle("A brand new hypothetical participant") 
-```
-
-```{r save-epred-pid-new, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
-```
-
-Fourth, six completely new hypothetical participants.
-
-```{r epred-new-pid-6}
-epreds_pid_new6 <- dat_M1i %>% 
-  data_grid(timebin, prime, pid = 7:12) %>%
-  add_epred_draws(model_M1i, 
-                  re_formula = NULL, 
-                  allow_new_levels=T,    
-                  sample_new_levels = "gaussian") %>% 
-  mutate(prime = factor(prime, 
-                        levels=c(1,2,3),
-                        labels=c("blank","congruent","incongruent"))) %>%
-  ungroup()
-```
-
-Summarize and plot predicted hazard values.
-
-```{r epred-new-pid-6-summary}
-epreds_pid_new6 %>% 
-  group_by(pid,timebin,prime) %>% 
-  median_qi(.width=.95)
-```
-
-```{r epred-new-pid-6-plot}
- ggplot(epreds_pid_new6, aes(x=timebin, y=.epred, 
-                       fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",
-                    .width=.95,
-                    alpha = 0.5) +
-    scale_fill_brewer(palette = "Dark2") +
-    scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-    theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted hazard") +
-    ggtitle("Six new hypothetical participants") +
-    facet_wrap(~pid, ncol=3)
-```
-
-```{r save-epred-pid-6-new, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new_6.png", width = 10, height = 8, dpi = 800)
+
+p2
+```
+
+Combine the plots.
+
+```{r save-epred-combined, eval=F}
+# combine 2 plots
+p1_theme <- p1 + theme(legend.position = "none")
+
+(p1_theme / p2) +
+    plot_annotation(tag_levels = 'A') +
+    plot_layout(guides = "collect",
+                axes = "collect_x") & 
+    theme(legend.position = "bottom")
+```
+
+Save plot.
+
+```{r save-combined-pred, eval=F}
+# save combined plot
+ggsave("Tutorial_2_Bayesian/figures/M1i_pred_combined.png", width = 10, height = 12, dpi = 800)
 ```
@@
 
-Summarize and plot.
+Summarize.
 
@@
 
-Save as table.
+Save summary as table.
 
@@
 ```{r epred-grand-diff-plot}
- ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
+ p11 <- ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
                        fill=contrast)) +
-  stat_lineribbon(alpha = 0.5, 
+  stat_lineribbon(alpha = 0.4, 
                   point_interval = "mean_qi",
@@
   labs(y = "difference in predicted hazard") +
-  ggtitle("Grand average marginal effect") +
+  #ggtitle("Grand average marginal effect") +
   facet_wrap(~contrast)
+
+p11
 ```
@@
 
+Show table.
+
+```{r}
+kable(table, caption = "Point (mean) and 95% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.")
+```
+
 Plot the contrasts in predicted hazard values.
@@
 ```{r epred-pid-diff-plot}
- ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
+ p22 <- ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
                               fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, 
+    stat_lineribbon(alpha = 0.4, 
                     point_interval = "mean_qi",
@@
     labs(y = "difference in predicted hazard") +
-    ggtitle("Subject-specific AMEs") +
+    #ggtitle("Subject-specific AMEs") +
     facet_wrap(~pid, ncol=3)
+
+p22
 ```
@@
 
-[[remove next section and 2 code chunks....]]
-And check whether an average of subject-specific AMEs is equal to the grand AME... NO !!!
-
-```{r}
-epreds_pid_diffs_av <- epreds_pid_diffs %>% group_by(timebin,contrast,.draw) %>%
-  mutate(average_condmeans = mean(diff))
-```
-
-```{r}
- ggplot(epreds_pid_diffs_av, aes(x=timebin, y=diff, 
-                       fill=contrast)) +
-  stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.8,.95)) +
- # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
-  geom_hline(yintercept = 0, color = "red", lty = 3) +
-  scale_y_continuous(limits=c(-0.5,0.3)) +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_color_brewer(palette = "Dark2") +
-  theme(axis.text.x = element_text(angle=90)) +
-  labs(y = "difference in predicted hazard") +
-  ggtitle("Average of subject-specific AMEs") +
-  facet_wrap(~contrast)
+Combine the plots.
+
+```{r save-ames-combined, eval=F}
+# combine 2 plots
+p11_theme <- p11 + theme(legend.position = "none")
+
+(p11_theme / p22) +
+    plot_annotation(tag_levels = 'A') +
+   # plot_layout(guides = "collect",
+  #              axes = "collect_x") & 
+    theme(legend.position = "bottom")
+```
+
+Save plot.
+
+```{r save-combined-ames, eval=F}
+# save combined plot
+ggsave("Tutorial_2_Bayesian/figures/M1i_ame_combined.png", width = 10, height = 12, dpi = 800)
 ```
@@
 
+Show table.
+
+```{r}
+kable(table_prop, caption = "Summarizing the posterior distributions of each contrast by their proportion below and above zero.
+prop_below = proportion below zero; prop_above = proportion above zero.")
+```
+
 Thus, the probability that the contrast "congruent minus blank" is larger than 0, is larger than .9 in bins 6 to 8.
@@
 ```
+
+# 8. Prior predictive checks
+
+## 8.1. Model M0i
+
+### Intermezzo: Prior Predictive Check for M0i.
+
+Following Gelman et al.
+(2020), we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
+First, sample the prior distributions using sample_prior="only".
+
+```{r M0i-prior, eval=F}
+plan(multicore)
+
+model_M0i_prior <-                  
+   brm(data = data_M0i,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + timebin + (0 + timebin | pid),
+       prior = priors_M0i,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0i_prior")
+```
+
+```{r inspect-M0i-prior}
+model_M0i_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0i_prior.rds")
+```
+
+Next, use them to predict prior data. But to better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.
+
+```{r ppc-manual, eval=F}
+# prior predictive check for 1 time bin
+# see http://bruno.nicenboim.me/bayescogsci/
+
+# cloglog and inverse-cloglog link functions
+inverse_cloglog <- function(x) {return(1-(exp(-1*exp(x))))}
+cloglog <- function(x) {return(log(-1*log(1-x)))}
+
+# Specify
+N_samples <- 2000 # the number of samples from the skew_normal
+N_obs <- 100      # the number of simulated observations per sample
+set.seed(1)
+
+# function to generate binary observations from samples
+cloglog_predictive_distribution <- function(samples, Nobs) {
+
+    # empty data frame with headers
+    df_pred <- tibble(trial = numeric(0),
+                      y_pred = numeric(0),
+                      iter = numeric(0))
+    # i iterates from 1 to the length of samples
+    for (i in seq_along(samples)) {
+      cloglog_haz <- samples[i]
+      df_pred <- bind_rows(df_pred,
+          tibble(trial = 1:N_obs,
+                 y_pred = rbinom(Nobs, 1, inverse_cloglog(cloglog_haz)),
+                 iter = i))
+    }
+    df_pred 
+ }
+
+# sample prior cloglog-hazard values for 1 bin
+cloglog_samples <- rskew_normal(N_samples,-1,1,-2)
+
+# apply the function
+prior_pred <- cloglog_predictive_distribution(cloglog_samples,
+                                              N_obs)
+
+# calculate hazard as the mean over binary observations
+prior_pred_haz <- prior_pred %>% 
+  group_by(iter) %>% 
+  summarise(pred_haz = mean(y_pred))
+
+# plot 
+ggplot(prior_pred_haz, aes(x=pred_haz)) +
+  geom_histogram(binwidth=.01) +
+  scale_x_continuous(limits = c(0,1)) +
+  labs(x = "simulated hazard",
+       title = "Prior predictive distribution",
+       subtitle = str_c(N_samples," samples; ",
+                        N_obs," observations per sample"))
+```
+
+```{r save-ppc, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc-1-bin, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png")
+```
+
+The shape of the prior predictive distribution for 1 bin looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
+
+Now use add_predicted_draws() to simulate prior-based data for our 10 bins and 6 subjects, while taking into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.
+
+```{r ppc-M0i-prior, eval=F}
+# Generate prior predictive hazard functions for 6 participants
+newdata_prior = tibble(timebin = 6:15) %>% 
+           expand_grid(pid = 1:6)
+
+# Specify
+N_obs = 100
+set.seed(2)
+
+# prepare columns of data set
+df_pred <- tibble(timebin = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# Call add_predicted_draws() for each simulated observation
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(model_M0i_prior, 
+                                    newdata=newdata_prior, 
+                                    summary=F, 
+                                    ndraws=NULL) %>% # all 2000
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across observations)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,timebin,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot 200 draws per participant
+ggplot(data = df_pred_haz %>% filter(.draw < 200), 
+       aes(x=timebin, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   geom_line(data=df_pred_haz %>% filter(.draw == 20), 
+             color="red", 
+             linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
+  labs(x = "time bin",
+       y = "simulated hazard",
+       title = "Prior predictive distributions",
+       subtitle = str_c("200 samples; ",N_obs," observations per sample")) +
+  facet_wrap(~pid)
+```
+
+```{r save-ppc-ten-bins-six-subjects, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0i_ppc.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc-M0i, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0i_ppc.png")
+```
+
+The red hazard function is one of the 200 functions plotted, for each of six participants. They can take on any shape, with values below .5 being more likely than values above .5.
+
+After our prior predictive check, we can fit model M0i to get the posterior distributions.
+
+## 8.2 Model M0c
+
+### Intermezzo. Prior predictive check for model M0c.
+
+Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs.
+
+To get an idea of how the intercept and both slopes control the shape of the hazard function, the following code allows to change the coefficients and generate (cloglog-)hazard functions.
+Once you have found the coefficients for the intercept, period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to determine a suited mean and standard deviation for each coefficient.
+
+```{r search-coef}
+dat <- tibble(x = -3:6,
+              x2 = x^2,
+              cloglog1 = -1.8   + 0.3*x    - 0.1*x2,  # black
+              cloglog2 = -1     + 0.41*x   - 0.04*x2,  # green
+              cloglog3 = -3     - 0.13*x   + 0.12*x2, # red
+              cloglog4 = -0.40  + 0.6*x    - 0.11*x2, # blue
+              haz1 = 1 - exp(-1*exp(cloglog1)),
+              haz2 = 1 - exp(-1*exp(cloglog2)),
+              haz3 = 1 - exp(-1*exp(cloglog3)),
+              haz4 = 1 - exp(-1*exp(cloglog4)))
+
+p1<-ggplot(dat, aes(x=x)) +  
+      geom_line(aes(y=cloglog1), color="black") +
+      geom_line(aes(y=cloglog2), color="green") +
+      geom_line(aes(y=cloglog3), color="red") +
+      geom_line(aes(y=cloglog4), color="blue") +
+      scale_y_continuous(limits=c(-6,2)) +
+      scale_x_continuous(breaks=c(-3:6)) +
+      labs(x = "timebin", y = "cloglog-hazard")
+
+p2<-ggplot(dat, aes(x=x))+
+      geom_line(aes(y=haz1), color="black") + 
+      geom_line(aes(y=haz2), color="green") +
+      geom_line(aes(y=haz3), color="red") +
+      geom_line(aes(y=haz4), color="blue") +
+      scale_y_continuous(limits=c(0,1)) +
+      scale_x_continuous(breaks=c(-3:6)) +
+      labs(x = "timebin", y = "hazard")
+
+p1|p2
+```
+
+For example, the coefficients for x vary from -0.13 to 0.6, and we might specify N(.3,.25) as a prior distribution, Those for x-squared vary from -0.11 to 0.12, and we might specify N(0,.06) as a prior distribution.
+
+Suppose we expect (cloglog-)hazard functions that increase almost linearly with time.
+We also set the standard deviation of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.
+
+```{r priors-M0c-ppc1}
+priors_M0c_ppc1 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.3,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .01)", class = "sd"), 
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+Sample the prior distributions.
+
+```{r fit-model-M0c-prior1, eval=F}
+plan(multicore)
+
+model_M0c_prior1 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc1,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior1")
+```
+
+```{r load-M0c-prior1}
+model_M0c_prior1 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior1.rds")
+```
+
+Check the prior distributions.
+
+```{r prior1-distr}
+# extract prior draws
+post <- as_draws_df(model_M0c_prior1) %>% 
+  select(-lp__) %>% 
+  as_tibble()
+
+tidy_prior <- post %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter))
+
+# plot
+ggplot(tidy_prior, aes(x = name, y = value, fill = name)) +  
+  stat_halfeye(alpha=0.7, 
+               point_interval = "median_qi",
+               .width=c(.8,.99), 
+               show.legend = F) +
+  labs(x = "parameter",
+       title = "Samples from the prior distributions")
+```
+
+Next, generate prior predictive check for each participant.
+
+Create a function.
+
+```{r function-plot-ppc, eval=F}
+plot_ppc <- function(fit_prior, N_samples, N_obs, new_data_prior){
+
+set.seed(1)
+
+# prepare columns of data set with predictions
+df_pred <- tibble(period_9 = integer(0),
+                  period_9_sq = integer(0),
+                  pid = integer(0),
+                  .row = integer(0),
+                  .chain = integer(0),
+                  .iteration = integer(0),
+                  .draw = integer(0),
+                  .prediction = integer(0),
+                  obs = integer(0))
+
+# extract predictions
+for(i in 1:N_obs){
+  prior_pred <- add_predicted_draws(fit_prior, 
+                                    newdata = new_data_prior, 
+                                    summary=F, 
+                                    ndraws=N_samples) %>% 
+                mutate(obs = i)
+  df_pred <- bind_rows(df_pred, prior_pred)
+}
+
+# calculate hazard per draw (average across obs)
+df_pred_haz <- df_pred %>% 
+  group_by(pid,period_9,.draw) %>% 
+  summarise(pred_haz = mean(.prediction))
+
+# plot N_samples draws per participant, highlighting three
+sel = sample(1:N_samples,3,replace=F)
+
+ggplot(df_pred_haz %>% 
+   filter(.draw <N_samples), aes(x=period_9, y=pred_haz, group=.draw)) +
+   geom_line(color="black") +
+   # highlight
+   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
+             color="red", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
+            color="green", linewidth=2) +
+   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
+          color="blue", linewidth=2) +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
+  facet_wrap(~pid)
+}
+```
+
+Apply the function.
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p1 <- plot_ppc(model_M0c_prior1,2000,100,newdata_prior)
+p1
+```
+
+For each of 2000 draws from the prior distributions, a black hazard function is plotted. Three randomly selected hazard functions are highlighted in red, green, and blue.
+
+```{r save-ppc1-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc1.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc1-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc1.png")
+```
+
+Or suppose we believe apriori that the hazard functions are flat.
+We specify our new priors:
+
+```{r priors-M0c-ppc2}
+priors_M0c_ppc2 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0,.001)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.001)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .001)", class = "sd"), 
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+Sample the prior distributions 2000 times:
+
+```{r fit-model-M0c-prior2, eval=F}
+plan(multicore)
+
+model_M0c_prior2 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc2,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior2")
+```
+
+```{r load-M0c-prior2}
+model_M0c_prior2 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior2.rds")
+```
+
+Apply the function plot_ppc().
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p2 <- plot_ppc(model_M0c_prior2,2000,100,newdata_prior)
+p2
+```
+
+Now the prior predictive hazard functions are all rather flat, with an intercept controlled by the prior skew-normal distribution, i.e., values below .5 are more likely than values above .5.
+
+```{r save-ppc2-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc2.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc2-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc2.png")
+```
+
+Finally, perform a ppc for the selected prior distributions for M0c (see lines 369-374), and increase the standard deviation for the normal prior for class "sd" somewhat.
+
+```{r priors-M0c-ppc3}
+priors_M0c_ppc3 <- c(
+  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
+  set_prior("normal(0.3,.25)", class = "b", coef = "period_9"), 
+  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
+  set_prior("normal(0, .1)", class = "sd"),  
+  set_prior("lkj(2)", class = "cor")  
+)
+```
+
+```{r fit-model-M0c-prior3, eval=F}
+plan(multicore)
+
+model_M0c_prior3 <-
+   brm(data = data_M0c,
+       family = bernoulli(link="cloglog"),
+       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
+                          (0 + Intercept + period_9 + period_9_sq | pid),
+       prior = priors_M0c_ppc3,
+       chains = 1, cores = 1, 
+       iter = 3000, warmup = 1000,
+       control = list(adapt_delta = 0.999, 
+                      step_size = 0.04, 
+                      max_treedepth = 12),
+       seed = 12, init = "0",
+       sample_prior = "only",
+       file = "Tutorial_2_Bayesian/models/model_M0c_prior3")
+```
+
+```{r load-M0c-prior3}
+model_M0c_prior3 <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior3.rds")
+```
+
+Apply the function plot_ppc().
+
+```{r eval=F}
+# make new data
+newdata_prior = tibble(period_9 = c(-3:6)) %>% 
+  mutate(period_9_sq = period_9^2) %>%
+  expand_grid(pid = 1:6)
+
+# make ppc plot
+p3 <- plot_ppc(model_M0c_prior3,2000,100,newdata_prior)
+p3
+```
+
+A wide range of differently shaped hazard functions can be observed, and early bins are unlikely to have high hazard values, consistent with our prior beliefs. 
+Note that some simulated hazard functions start off with a high hazard in bin -3.
+These reflect a combination of negative slopes and a positive coefficient for period_9_sq.
+These possibilities are required to encompass hazard functions that initially decrease and then increase, as observed in the empirical hazard functions for incongruent primes.
+
+```{r save-ppc3-M0c, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M0c_ppc3.png", width = 10, height = 8, dpi = 800)
+```
+
+```{r plot-ppc3-M0c, out.width='80%'}
+knitr::include_graphics("Tutorial_2_Bayesian/figures/M0c_ppc3.png")
+```
+
+After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions.

### 2024-12-09 Tutorial_2b.Rmd
update tuts

@@
 
+#
+```{r plot-post-M0i-ca-alt}
+tidy_fixed_ca <- post_ca %>% 
+  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
+  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
+  pivot_longer(-c(chain, draw, iter)) %>% 
+  mutate(timebin = str_sub(name,10,11),
+         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
+         condition = str_sub(name,17,18),
+         condition = factor(str_remove(condition,"e"),
+                            levels=c(1,2,3),
+                            labels=c("blank","congruent","incongruent")))
+head(tidy_fixed_ca)
+tail(tidy_fixed_ca)
+
+# plot
+p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
+  stat_halfeye(point_interval = "median_qi",
+               .width = c(.80,.95),
+               alpha=0.7,
+               position = pd) +
+  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
+       x = "time bin", 
+       y = "logit-ca") +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_y_continuous(limits=c(-10,10)) +
+  theme(axis.text.x = element_text(angle=90)) #+
+  #facet_wrap(~condition)
+p_tidy_fixed_ca
+```
+
+```{r save-post-alt, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr_alt.png", width = 10, height = 8, dpi = 800)
+```
+
+#
+
 ## Plot predicted conditional accuracy functions
@@
     scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted conditional accuracy") +
-    ggtitle("Average participant")
+    labs(y = "predicted conditional accuracy") #+
+    #ggtitle("Average participant")
+p1
 ```
@@
 ```{r plot-epred-pid}
- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
+p2 <- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                        fill=prime, color=prime)) +
@@
     labs(y = "predicted conditional accuracy") +
-    ggtitle("Each participant (N = 6)") +
+   # ggtitle("Each participant (N = 6)") +
     facet_wrap(~pid, ncol=3)
+p2
 ```
@@
 ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png", width = 10, height = 8, dpi = 800)
+```
+
+Combine the plots.
+
+```{r save-epred-combined, eval=F}
+# combine 2 plots
+p1_theme <- p1 + theme(legend.position = "none")
+
+(p1_theme / p2) +
+    plot_annotation(tag_levels = 'A') +
+    plot_layout(guides = "collect",
+                axes = "collect_x") & 
+    theme(legend.position = "bottom")
+```
+
+Save plot.
+
+```{r save-combined-pred, eval=F}
+# save combined plot
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png", width = 10, height = 12, dpi = 800)
 ```

### 2024-12-09 Tutorial_3a.Rmd
update tuts

@@
               effectname = names(fixef(model_M1i_f)),
-              position = 1:30) 
-est
+              position = rep(6:15,3),
+              prime = factor(rep(1:3, each=10),labels=c("blank","congruent","incongruent")) )
+est %>% print(n=30)
 ```
@@
 ```
+
+#
+```{r plot-estimates}
+# check hex codes for color paletter
+mypal <- brewer.pal(8,"Dark2")
+
+# specify colors and shapes
+cols    <- c( "c1" = mypal[1], "c2" = mypal[2])
+shapes  <- c("s1" = 16, "s2" = 17)
+
+# plot
+p1 <- ggplot(data = est, aes(x = position)) +
+ geom_point(aes( y = brm[,"Estimate"],  
+                 color = "c1", 
+                 shape = "s1"),
+                 size = 1.5) +
+ geom_point(aes( y = glmer$estimate, 
+                 color = "c2", 
+                 shape = "s2"),
+                 size = 1.5, 
+                 position = position_nudge(x = .2))  +
+ geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
+                    color = "c1"), 
+                show.legend = F) +
+ geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
+                    color = "c2"), 
+                show.legend = F, 
+                position = position_nudge(x = .2)) +
+# geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
+ scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +
+
+ labs( x = "bin endpoint", y = "cloglog-hazard" ) +
+ scale_color_manual(name = "Model", 
+                                breaks = c("c1", "c2"), 
+                                values = cols,
+                                labels = c("brm", "glmer")) +
+ scale_shape_manual(name = "Model", 
+                              breaks = c("s1", "s2"),
+                              values = shapes,
+                              labels = c("brm", "glmer")) +
+  theme(panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~prime)
+p1
+```
+
+```{r save-plot-comparison, eval=F}
+ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=10,dpi=800)
+```

### 2024-12-09 manuscript/manuscript.Rmd
update ms. and ms-notes

@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6), and for each participant in the data set (Figure 7).
-
-(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6A), and for each participant in the data set (Figure 6B).
+
+(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).
 
 ```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_grand.png")
-```
-
-(ref:descr-plot7-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.
-
-```{r plot-pred-pid-effects, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_pid.png")
-```
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 8; grand average marginal effect) and for each participant in the data set (Figure 9; subject-specific average marginal effect).
-
-(ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the average participant.
-
-```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_grand_AME.png")
-```
-
-(ref:descr-plot9-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for each participant.
-
-```{r plot-pid-ame-effects, fig.cap = "(ref:descr-plot9-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pid_AMEs.png")
-```
-
-Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_combined.png")
+```
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+
+(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the average participant (A), and for each participant (B).
+
+```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ame_combined.png")
+```
+
+<!-- Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.-->
 
@@
 
+The point estimates and quantile intervals can be reported in a table (see Tutorial_2a.Rmd for details).
+
 #### Example conclusions for M1i
@@
 
-The posterior distribution of each contrast can also be summarized by considering its proportion below or above some value, like zero.
-Table 5 shows the proportion of the posterior distribution below or above zero, for each time bin and contrast.
-
-(ref:table-props-caption) Summarizing the posterior distributions of each contrast by their proportion below and above zero.
-
-(ref:table-props-note-caption) prop_below = proportion below zero; prop_above = proportion above zero.
-
-```{r table-contrasts_props, eval=T}
- apa_table(
-   tab_contrast_props,
-   caption = "(ref:table-props-caption)",
-   note = "(ref:table-props-note-caption)",
-   placement = "H",
-#   font_size = "scriptsize",
-   longtable = TRUE,
-   repeat_header = TRUE,
-   format.args = list(
-        digits = c(0,0,2,2),
-        decimal.mark = ".", big.mark = "")
-)
-```
-
-Thus, the probability that the contrast "congruent minus blank" is larger than 0, is larger than .9 in bins 6 to 8.
-And the probability that the contrast "incongruent minus blank" is smaller than 0, is larger than .9 in bins 9 to 12.
-
 If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
@@
 
-To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 10, together with point (median) and interval estimates (80% and 95% credible intervals).
-
-(ref:descr-plot10-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
-
-```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
+To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 8, together with point (median) and interval estimates (80% and 95% credible intervals).
+
+(ref:descr-plot8-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
+
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 11), and for each participant in the data set (Figure 12).
-
-(ref:plot11-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.
-
-```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot11-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_grand.png")
-```
-
-(ref:plot12-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.
-
-```{r plot-pred-pid-effects-ca, fig.cap = "(ref:plot12-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png")
-```
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 13; grand average marginal effect).
-
-(ref:descr-plot13-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.
-
-```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot13-caption)", out.width='80%'}
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 9A), and for each participant in the data set (Figure 9B).
+
+(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).
+
+```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot9-caption)", out.width='80%'}
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png")
+```
+
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 10; grand average marginal effect).
+
+(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.
+
+```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png")
@@
 
-Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.
+<!-- Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.-->
 
@@
 
-```{r table-ca-grand-ame}
+```{r table-ca-grand-ame, eval=F}
  apa_table(
@@
 
-Based on Figure 13 and Table 6, we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+The point estimates and quantile intervals can be reported in a table (see Tutorial_2b.Rmd for details).
+
+Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
@@
 
-In Figure 14 we compare the parameter estimates of model M1i from brm() with those of model M1i_f from glmer().
-
-(ref:descr-plot14-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
-
-```{r plot-comparison, fig.cap = "(ref:descr-plot14-caption)", out.width='80%'}
+In Figure 11 we compare the parameter estimates of model M1i from brm() with those of model M1i_f from glmer().
+
+(ref:descr-plot11-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
+
+```{r plot-comparison, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png")
@@
 
-Figure 14 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
+Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
 However, model M1i_f did not converge and resulted in a singular fit.
@@
 We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
-The raw data and the simulated data are plotted in Figure 15 and show quite close correspondence, which is re-assuring.
+The raw data and the simulated data are plotted in Figure 12 and show quite close correspondence, which is re-assuring.
 But, this is only one dataset.
@@
 
-(ref:simdat-plot15-caption) Raw data from @panisWhatShapingRT2016 and simulated data from 10 participants.
-
-```{r simdat, fig.cap = "(ref:simdat-plot15-caption)", out.width='80%'}
+(ref:simdat-plot12-caption) Raw data from @panisWhatShapingRT2016 and simulated data from 10 participants.
+
+```{r simdat, fig.cap = "(ref:simdat-plot12-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_4_Planning/figures/raw_vs_d1.jpeg")
@@
 
-Summary results from Simulation 1 are shown in Figure 16A.
-Figure 16A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 13A.
+Figure 13A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
@@
 Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
-Summary results from Simulation 2 are shown in Figure 16B.
+Summary results from Simulation 2 are shown in Figure 13B.
 A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):
@@
 
-(ref:power-plot16-caption) Statistical power across data Simulation 1 (A) and Simulation 2 (B).
+(ref:power-plot13-caption) Statistical power across data Simulation 1 (A) and Simulation 2 (B).
 Power was calculated as the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
@@
 
-```{r power, fig.cap = "(ref:power-plot16-caption)", out.width='100%'}
+```{r power, fig.cap = "(ref:power-plot13-caption)", out.width='100%'}
 knitr::include_graphics("../Tutorial_4_Planning/figures/bin6_power_col.jpeg")

### 2024-12-09 manuscript/manuscript.tex
update ms. and ms-notes

@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6), and for each participant in the data set (Figure 7).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6A), and for each participant in the data set (Figure 6B).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_grand}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_combined}
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.}\label{fig:plot-pred-grand-effects}
+\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
 \end{figure}
 
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_pid}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ame_combined}
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.}\label{fig:plot-pred-pid-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
 \end{figure}
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 8; grand average marginal effect) and for each participant in the data set (Figure 9; subject-specific average marginal effect).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_grand_AME}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the average participant.}\label{fig:plot-grand-ame-effects}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pid_AMEs}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for each participant.}\label{fig:plot-pid-ame-effects}
-\end{figure}
-
-Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% credible interval, for the average participant.
+The point estimates and quantile intervals can be reported in a table (see Tutorial\_2a.Rmd for details).
 
@@
 
-The posterior distribution of each contrast can also be summarized by considering its proportion below or above some value, like zero.
-Table 5 shows the proportion of the posterior distribution below or above zero, for each time bin and contrast.
-
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{TableNotes}[para]
-\normalsize{\textit{Note.} prop\_below = proportion below zero; prop\_above = proportion above zero.}
-\end{TableNotes}
-
-\begin{longtable}{llll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table-contrasts_props}Summarizing the posterior distributions of each contrast by their proportion below and above zero.}\\
-\toprule
-timebin & \multicolumn{1}{c}{contrast} & \multicolumn{1}{c}{prop\_above} & \multicolumn{1}{c}{prop\_below}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:table-contrasts_props} continued}}\\
-\toprule
-timebin & \multicolumn{1}{c}{contrast} & \multicolumn{1}{c}{prop\_above} & \multicolumn{1}{c}{prop\_below}\\
-\midrule
-\endhead
-6 & congruent minus blank & 0.99 & 0.01\\
-7 & congruent minus blank & 0.99 & 0.01\\
-8 & congruent minus blank & 0.95 & 0.05\\
-9 & congruent minus blank & 0.79 & 0.21\\
-10 & congruent minus blank & 0.24 & 0.76\\
-11 & congruent minus blank & 0.15 & 0.85\\
-12 & congruent minus blank & 0.33 & 0.67\\
-13 & congruent minus blank & 0.28 & 0.72\\
-14 & congruent minus blank & 0.19 & 0.81\\
-15 & congruent minus blank & 0.47 & 0.53\\
-6 & incongruent minus blank & 0.98 & 0.02\\
-7 & incongruent minus blank & 0.92 & 0.08\\
-8 & incongruent minus blank & 0.12 & 0.88\\
-9 & incongruent minus blank & 0.00 & 1.00\\
-10 & incongruent minus blank & 0.01 & 0.99\\
-11 & incongruent minus blank & 0.02 & 0.98\\
-12 & incongruent minus blank & 0.04 & 0.96\\
-13 & incongruent minus blank & 0.20 & 0.80\\
-14 & incongruent minus blank & 0.27 & 0.73\\
-15 & incongruent minus blank & 0.58 & 0.42\\
-\bottomrule
-\addlinespace
-\insertTableNotes
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
-Thus, the probability that the contrast ``congruent minus blank'' is larger than 0, is larger than .9 in bins 6 to 8.
-And the probability that the contrast ``incongruent minus blank'' is smaller than 0, is larger than .9 in bins 9 to 12.
-
 If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
@@
 
-To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 10, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 8, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 11), and for each participant in the data set (Figure 12).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 9A), and for each participant in the data set (Figure 9B).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_grand}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined}
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant.}\label{fig:plot-pred-grand-effects-ca}
+\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
 \end{figure}
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_pid}
-
-}
-
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for each participant.}\label{fig:plot-pred-pid-effects-ca}
-\end{figure}
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 13; grand average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 10; grand average marginal effect).
 
@@
 
-Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95\% credible interval, for the average participant.
-
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{TableNotes}[para]
-\normalsize{\textit{Note.} diff = difference in predicted conditional accuracy.}
-\end{TableNotes}
-
-\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table-ca-grand-ame}Point (mean) and 95\% credible interval summary of estimated differences in conditional accuracy, for each time bin and contrast, in the average participant.}\\
-\toprule
-contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff\_ca} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:table-ca-grand-ame} continued}}\\
-\toprule
-contrast & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{diff\_ca} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper}\\
-\midrule
-\endhead
-congruent minus blank & 6 & 0.36 & 0.06 & 0.74\\
-congruent minus blank & 7 & 0.32 & 0.13 & 0.57\\
-congruent minus blank & 8 & 0.16 & 0.06 & 0.31\\
-congruent minus blank & 9 & 0.07 & 0.02 & 0.17\\
-congruent minus blank & 10 & 0.00 & -0.03 & 0.03\\
-congruent minus blank & 11 & -0.03 & -0.11 & 0.02\\
-congruent minus blank & 12 & -0.02 & -0.10 & 0.03\\
-congruent minus blank & 13 & -0.04 & -0.20 & 0.06\\
-congruent minus blank & 14 & -0.20 & -0.59 & 0.07\\
-congruent minus blank & 15 & -0.06 & -0.49 & 0.28\\
-incongruent minus blank & 6 & -0.61 & -0.91 & -0.22\\
-incongruent minus blank & 7 & -0.64 & -0.84 & -0.38\\
-incongruent minus blank & 8 & -0.70 & -0.86 & -0.40\\
-incongruent minus blank & 9 & -0.21 & -0.59 & 0.04\\
-incongruent minus blank & 10 & -0.03 & -0.15 & 0.02\\
-incongruent minus blank & 11 & -0.02 & -0.07 & 0.02\\
-incongruent minus blank & 12 & 0.00 & -0.04 & 0.04\\
-incongruent minus blank & 13 & 0.01 & -0.06 & 0.08\\
-incongruent minus blank & 14 & 0.02 & -0.09 & 0.18\\
-incongruent minus blank & 15 & -0.09 & -0.39 & 0.23\\
-\bottomrule
-\addlinespace
-\insertTableNotes
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
-Based on Figure 13 and Table 6, we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+The point estimates and quantile intervals can be reported in a table (see Tutorial\_2b.Rmd for details).
+
+Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
@@
 
-In Figure 14 we compare the parameter estimates of model M1i from brm() with those of model M1i\_f from glmer().
+In Figure 11 we compare the parameter estimates of model M1i from brm() with those of model M1i\_f from glmer().
 
@@
 
-Figure 14 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
+Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
 However, model M1i\_f did not converge and resulted in a singular fit.
@@
 
-Summary results from Simulation 1 are shown in Figure 16A.
-Figure 16A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 13A.
+Figure 13A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.

### 2024-12-10 manuscript/SupplementaryMaterial.Rmd
update ms. and ms-notes.

@@
 \noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 15.
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 14.
 
-(ref:descr-plot15-caption) The logit and cloglog link functions.
+(ref:descr-plot14-caption) The logit and cloglog link functions.
 
-```{r plot-link-functions, fig.cap = "(ref:descr-plot15-caption)", out.width='80%'}
+```{r plot-link-functions, fig.cap = "(ref:descr-plot14-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
@@
 
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X~1~) can be written as follows:
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X~1~) can be written as follows:
 
@@
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 16 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 15 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
 
-(ref:descr-plot16-caption) Prior distributions for the Intercept on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
+(ref:descr-plot15-caption) Prior distributions for the Intercept on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).
 
-```{r plot-priors, fig.cap = "(ref:descr-plot16-caption)", out.width='80%'}
+```{r plot-priors, fig.cap = "(ref:descr-plot15-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")

### 2024-12-10 manuscript/manuscript.Rmd
update ms. and ms-notes.

@@
 Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
-keywords          : "response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology"
-wordcount         : "X"
+keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
+wordcount         : "13574"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in Example 4 show that the effect starts between 100 and 200 ms (in timebin 2) and is gone when the waiting time reaches 500 ms or more.
-In contrast, in Example 5, the effect starts around 300 ms and is gone by 700 ms.
-And in the Example 6, the effect reverses between 500 and 600 ms.
+For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in timebin 2) and is gone when the waiting time reaches 500 ms or more.
+In contrast, in example 5, the effect starts around 300 ms and is gone by 700 ms.
+And in example 6, the effect reverses between 500 and 600 ms.
 What kind of theory or theories could account for such effects?
@@
 
-We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
@@
 
-In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
+In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
 We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
+In Tutorial 2b we fit Bayesian multilevel regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
 
@@
 We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
-[[more to come here, once we have written the tutorial]].
-
-In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+
+In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multilevel regression modelling steps.
 Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
@@
 Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in part D of the supplementary material.
+As such, we only supply regression equations in section D of the supplementary material.
 
@@
 Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
+Error bars represent +/- 1 standard error of the respective proportion.
 While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
@@
 We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly summarise list the benefits below, and refer the reader to section F of Supplementary Materials for more detailed coverage of the benefits.
-A summary of the benefits are as follows:
+Instead, we briefly list the benefits below, and refer the reader to section F of the supplementary material for more detailed coverage of the benefits.
+The benefits include:
 
@@
 EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see below).
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
 
@@
 Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Error bars represent +/- 1 standard error of the respective proportion.
 Bin width equals 100 ms.
@@
 
-Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
+Our lab adopts a estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
 We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
@@
 We used `r my_r_citation$r` for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see part B in the supplemental material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section B in the supplemental material for a list of the custom functions).
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the supplementary material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
@@
 Vertical dotted lines indicate the estimated median RTs.
-Error bars represent +/- 1 Standard Error of the respective proportion.
+Error bars represent +/- 1 standard error of the respective proportion.
 
@@
 
-So far in this paper, we have use a simple experimental design, which involved one condition with three levels.
+So far in this paper, we have used a simple experimental design, which involved one condition with three levels.
 But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than three levels.
@@
 
-In this third tutorial, we illustrate how to fit Bayesian multi-level regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see part C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 
-The middle column of Figure 16 in part E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 16 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Figure 15 in section E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 15 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6A), and for each participant in the data set (Figure 6B).
-
-(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the grand average (Figure 6A), and for each participant in the data set (Figure 6B).
+
+(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
-
-(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the average participant (A), and for each participant (B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the grand average (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+
+(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the grand average (A), and for each participant (B).
 
@@
 
-<!-- Table 4 shows the summaries of the estimated hazard differences for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.-->
-
-```{r load-tables, eval=T, echo=F}
-tab_grand_ame <- read_csv("../Tutorial_2_Bayesian/tables/grand_AMEs.csv")
-#tab_pid_ame <- read_csv("../Tutorial_2_Bayesian/tables/pid_AMEs.csv")
-tab_contrast_props <- read_csv("../Tutorial_2_Bayesian/tables/contrasts_props.csv")
-```
-
-(ref:table-grand-ame-caption) Point (mean) and 95% credible interval summary of estimated differences in hazard, for each time bin and contrast, in the average participant.
-
-(ref:table-grand-ame-note-caption) diff = difference in predicted hazard.
-
-```{r table-grand-ame, eval = F}
- apa_table(
-   tab_grand_ame,
-   caption = "(ref:table-grand-ame-caption)",
-   note = "(ref:table-grand-ame-note-caption)",
-   placement = "H",
-#   font_size = "scriptsize",
-   longtable = TRUE,
-   repeat_header = TRUE,
-   format.args = list(
-        digits = c(0,0,2,2,2),
-        decimal.mark = ".", big.mark = "")
-)
-```
-
 The point estimates and quantile intervals can be reported in a table (see Tutorial_2a.Rmd for details).
@@
 
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
-The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 7 (95% CrI = [0.04, 0.25]).
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
+The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
 For the other bins, the 95% credible interval contained zero.
-The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin 9 (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin 10 (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin 11 (95% CrI = [-0.40, -0.03]).
+The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin (320,360] (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin (360,400] (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin (400,440] (95% CrI = [-0.40, -0.03]).
 For the other bins, the 95% credible interval contained zero.
-Note that we could also have calculated hazard ratios instead of hazard differences.
 
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 7B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
@@
 
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 9).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 7B).
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
-In Tutorial_2a.Rmd we provide a number of model formula that should get you going.
+In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
 
@@
 
-In this fourth tutorial, we illustrate how to fit a Bayesian multi-level regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
+In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
 The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
@@
 
-(ref:descr-plot8-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
+(ref:descr-plot8-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca. ca = conditional accuracy.
 
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 9A), and for each participant in the data set (Figure 9B).
-
-(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the grand average (Figure 9A), and for each participant in the data set (Figure 9B).
+
+(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 10; grand average marginal effect).
-
-(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10; grand average marginal effect).
+
+(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average.
 
@@
 
-<!-- Table 6 shows the summaries of the estimated differences in conditional accuracy for both contrasts in terms of a point estimate (the mean) and the upper and lower bounds of the 95% credible interval, for the average participant.-->
-
-```{r load-tables-ca, eval=T, echo=F}
-tab_ca_grand_ame <- read_csv("../Tutorial_2_Bayesian/tables/ca_grand_AMEs.csv")
-```
-
-(ref:table-ca-grand-ame-caption) Point (mean) and 95% credible interval summary of estimated differences in conditional accuracy, for each time bin and contrast, in the average participant.
-
-(ref:table-ca-grand-ame-note-caption) diff = difference in predicted conditional accuracy.
-
-```{r table-ca-grand-ame, eval=F}
- apa_table(
-   tab_ca_grand_ame,
-   caption = "(ref:table-ca-grand-ame-caption)",
-   note = "(ref:table-ca-grand-ame-note-caption)",
-   placement = "H",
-#   font_size = "scriptsize",
-   longtable = TRUE,
-   repeat_header = TRUE,
-   format.args = list(
-        digits = c(0,0,2,2,2),
-        decimal.mark = ".", big.mark = "")
-)
-```
-
-The point estimates and quantile intervals can be reported in a table (see Tutorial_2b.Rmd for details).
-
-Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
@@
 From this simulation, we can see that somewhere between an effect size of a 50% and 75% reduction in hazard value, power increases to a range that most researchers would consider acceptable (i.e., >95% power).
-To probe this space a little further, we decided to run a second simulation, which varied different parameters
+To probe this space a little further, we decided to run a second simulation, which varied different parameters.
 
@@
 In Simulation 1, the effect size was varied between a 25%, 50% and 75% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
-In Simulation 2, the effect size was varied between a 50%, 60% and 70% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80% (red), 90% (black) and 95% (blue) power.
-Abbreviations: rep_n = the number of trials per experimental condition; sub_j = the number of participants per simulated experiment.
+In Simulation 2, the effect size was varied between a 50%, 60% and 70% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).
+The dashed lines represent 80% (red), 90% (black) and 95% (blue) power.
+Abbreviations: rep_n = the number of trials per experimental condition; subj_n = the number of participants per simulated experiment.
 
@@
 
-EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions. 
-For example, some of the models discussed in Tutorial 2a (M2) contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions. 
+For example, some of the models discussed in Tutorial 2a contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than 1 time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
+From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
@@
 For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
-These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 600 ms for simple detection tasks, 1000 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
-Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response), introducing a response deadline is recommended when designing RT experiments.
+These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 500 ms for simple detection tasks, 800 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
+Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response in the analysis time window), introducing a response deadline is recommended when designing RT experiments.
 Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard and conditional accuracy models.
@@
 One way to deal with this is to describe and interpret the different patterns.
-Another way is to run a clustering algorithm on the individual hazard estimates across all conditions.
+Another way is to run a clustering algorithm on the individual hazard estimates across all bins and conditions.
 The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
@@
 
-Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021], which is a from of Small-N approach (Smith & Little, 2018).
+Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021], which is a from of small-*N* approach (Smith & Little, 2018).
 This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
@@
 
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multi-level hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 
@@
 The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this event constitutes a single transition between an "idle" state and a "responded" state.
+For RT data, this button-press event constitutes a single transition between an "idle" state and a "responded" state.
 However, in certain situations, more than one event of interest might exist.

### 2024-12-10 manuscript/manuscript.tex
update ms. and ms-notes.

@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology\newline\indent Word count: X}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 13574}
 \usepackage{lineno}
@@
   pdflang={en-EN},
-  pdfkeywords={response times, event history analysis, Bayesian multi-level regression models, experimental psychology, cognitive psychology},
+  pdfkeywords={response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology},
   hidelinks,
@@
 
-We provide six different tutorials, which are written in the R programming language and publicly available on our Github and the Open Science Framework (OSF) pages, along with all of the other code and material associated with the project.
+We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
@@
 
-In Tutorial 2a, we illustrate how one can fit Bayesian multi-level regression models to RT data using the R package brms.
+In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
 We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
-In Tutorial 2b we fit Bayesian multi-level regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
+In Tutorial 2b we fit Bayesian multilevel regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
 
@@
 We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
-{[}{[}more to come here, once we have written the tutorial{]}{]}.
-
-In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
+
+In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multilevel regression modelling steps.
 Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
@@
   EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see below).
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
 \item
@@
 Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Error bars represent +/- 1 standard error of the respective proportion.
 Bin width equals 100 ms.}\label{fig:plot3}
@@
 
-Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
+Our lab adopts a estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
@@
 We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-The content of the tutorials, in terms of EHA and multi-level regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see part A of the supplementary material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the supplementary material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
@@
 Vertical dotted lines indicate the estimated median RTs.
-Error bars represent +/- 1 Standard Error of the respective proportion.}\label{fig:eha-plot}
+Error bars represent +/- 1 standard error of the respective proportion.}\label{fig:eha-plot}
 \end{figure}
@@
 
-In this third tutorial, we illustrate how to fit Bayesian multi-level regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data set used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see part C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant (Figure 6A), and for each participant in the data set (Figure 6B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the grand average (Figure 6A), and for each participant in the data set (Figure 6B).
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
+\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
 \end{figure}
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the average participant (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the grand average (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
 
@@
 
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
 \end{figure}
@@
 
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
-The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin 6 (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin 7 (95\% CrI = {[}0.04, 0.25{]}).
+If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
+The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
 For the other bins, the 95\% credible interval contained zero.
-The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin 6 (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin 9 (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin 10 (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin 11 (95\% CrI = {[}-0.40, -0.03{]}).
+The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin (320,360{]} (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin (360,400{]} (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin (400,440{]} (95\% CrI = {[}-0.40, -0.03{]}).
 For the other bins, the 95\% credible interval contained zero.
-Note that we could also have calculated hazard ratios instead of hazard differences.
 
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 9.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 7B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
@@
 
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 9).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 7B).
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
-In Tutorial\_2a.Rmd we provide a number of model formula that should get you going.
+In Tutorial\_2a.Rmd we provide a number of model formulae that should get you going.
 
@@
 
-\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca.}\label{fig:plot-ca-fixed-effects}
+\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca. ca = conditional accuracy.}\label{fig:plot-ca-fixed-effects}
 \end{figure}
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the average participant (Figure 9A), and for each participant in the data set (Figure 9B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the grand average (Figure 9A), and for each participant in the data set (Figure 9B).
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the average participant (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
+\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
 \end{figure}
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the average participant (Figure 10; grand average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10; grand average marginal effect).
 
@@
 
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the average participant.}\label{fig:plot-ca-grand-ame-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average.}\label{fig:plot-ca-grand-ame-effects}
 \end{figure}
 
-The point estimates and quantile intervals can be reported in a table (see Tutorial\_2b.Rmd for details).
-
-Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 14).
+Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
@@
 From this simulation, we can see that somewhere between an effect size of a 50\% and 75\% reduction in hazard value, power increases to a range that most researchers would consider acceptable (i.e., \textgreater95\% power).
-To probe this space a little further, we decided to run a second simulation, which varied different parameters
+To probe this space a little further, we decided to run a second simulation, which varied different parameters.
 
@@
 In Simulation 1, the effect size was varied between a 25\%, 50\% and 75\% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
-In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.
-Abbreviations: rep\_n = the number of trials per experimental condition; sub\_j = the number of participants per simulated experiment.}\label{fig:power}
+In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).
+The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.
+Abbreviations: rep\_n = the number of trials per experimental condition; subj\_n = the number of participants per simulated experiment.}\label{fig:power}
 \end{figure}
@@
 
-EHA can quickly become very complex when adding more than 1 time scale, due to the many possible higher-order interactions.
-For example, some of the models discussed in Tutorial 2a (M2) contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions.
+For example, some of the models discussed in Tutorial 2a contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than 1 time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
+From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
@@
 
-Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021), which is a from of Small-N approach (Smith \& Little, 2018).
+Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021), which is a from of small-\emph{N} approach (Smith \& Little, 2018).
 This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
@@
 The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this event constitutes a single transition between an ``idle'' state and a ``responded'' state.
+For RT data, this button-press event constitutes a single transition between an ``idle'' state and a ``responded'' state.
 However, in certain situations, more than one event of interest might exist.
@@
 \noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 15.
+Popular link functions include the logit link and the complementary log-log link, as shown in Figure 14.
 
@@
 
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
+An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
 
@@
 To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 16 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Figure 15 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).

### 2024-12-10 notes/manuscript_notes.Rmd
update ms. and ms-notes
update ms. notes
add note
update ms. and ms-notes.

@@
 
-->
+-> Done. Moved to tutorial only.
 
@@
 
--> cloglog-hazard is better I guess.
+-> cloglog-hazard is correct.
 
@@
 
-->
+-> Done. !!!!! delete one....
 
@@
 That saves loads of space and time scanning values when the figures are beautiful and standalone as they are.
+
+-> Done.
 
@@
 
-2. As (1) combine Figures 8 and 9.
-
-->
+-> Done.
+
+2. As (1) combine Figures 8 and 9. (ames)
+
+-> Done.
 
@@
 
-->
+-> Title removed before combining with Figure 7.
 
@@
 
-->
+-> AMEs was removed before combining with fig 8.
 
@@
 
+-> Done.
+
 Put Table 6 in the tutorial and remove it from the main text.
+
+-> Done.
 
@@
 
-# Sven dec 8
-
-I also moved prior predictive checks to last section of 2a.
-
-Mention 2b ppc...
-
-Section 4...
-
-Tutorial 2a ... M1d...
+-> Done.
+
+# Sven dec 10
+
+adjust small error in Figure 3 + check Figure 10 + add second panel to Figure 10
+
+change ms code in section 4.1.2 (make it same as in tut)
+
+Mention tut 2b priors...   todo
+
+Section 4... comments?
+
+appendices/separate Rmd file...
+
+clean up html of Rmd files

### 2024-12-10 notes/manuscript_notes.Rmd
update ms-notes.

@@
 
-- check legends of all Figures, and add more info if needed
+- check legends of all Figures, and add more info if needed -> Done
 
@@
 
-- add info on simulation and power analysis (section 4.4 + summary in section 4)
-
-- delete code in Tutorial_2a about brand new hypothetical subject?
-
-- use timebin or time bins?
-
-- use multi-level or multilevel ?
+- add info on simulation and power analysis (section 4.4 + summary in section 4). Done.
+
+- delete code in Tutorial_2a about brand new hypothetical subject? Done.
+
+- use timebin or time bins? Timebin.
+
+- use multi-level or multilevel ? Multilevel.
 
@@
 
--> Done. !!!!! delete one....
+-> Done.
 
@@
 
-# Sven dec 10
+# Sven dec 10 still to do:
 
@@
 
-Mention tut 2b priors...   todo
-
-Section 4... comments?
-
-appendices/separate Rmd file...
-
-clean up html of Rmd files
+appendix/separate Rmd file... + {var} in functions...?
+
+clean up html files (delete old figures, Mention tut 2b priors, ...)
+
+change Figure comparison for ca
+
+check for timebin, multilevel

### 2024-12-10 manuscript/manuscript.Rmd
RR edits

@@
 abstract: |
-  Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
+  Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
 However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
@@
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
-To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
-We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multilevel regression modelling.
-Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
+To help achieve broader uptake of EHA, in this paper we outline a set of tutorials that detail one distributional method known as discrete-time EHA.
+We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design, as well as how to manage inter-individual differences.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
-Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
+Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
@@
 survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
-We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
 Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
@@
 We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
@@
 
-For a comprehensive background context to EHA, we recommend several excellent textbooks [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010].
-Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks [@gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
+We recommend several excellent textbooks for a comprehensive background context to EHA [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010] and for a more general introduction to understanding regression equations [@gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
@@
 
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel).
 These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
@@
 We perform data wrangling following tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
-Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of descriptive EHA a lot quicker and more efficient.
+In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
+In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.
 
@@
 
-Our lab adopts a estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
+Our lab adopts an estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
 We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
-This means that wherever possible we include varying intercepts and slopes per participant
+This means that wherever possible we include varying intercepts and slopes per participant.
 To make inferences, we use two main approaches.
@@
 
-Tutorials 1a and 1b show how to calculate and plot the descriptive statistics of EHA/SAT when there are one and two independent variables, respectively.
+Tutorials 1a and 1b show how to calculate and plot the descriptive statistics of EHA/SAT when there are one or two independent variables, respectively.
 Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
@@
 In contrast, the 'person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the 'person-trial-bin' data set generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+Therefore, the 'person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
 This data set is used when fitting hazard models (Tutorials 2a and 3a).
@@
 
-(ref:descr-plot4-caption) Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6.
+(ref:descr-plot4-caption) Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
 Vertical dotted lines indicate the estimated median RTs.
@@
 
-These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
+These distributional results suggest that participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
-When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial_1a.Rmd).
+When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial_1a.Rmd).
 
@@
 So far in this paper, we have used a simple experimental design, which involved one condition with three levels.
-But psychological experiments are often more complex, with crossed factorial designs with more conditions and more than three levels.
+But psychological experiments are often more complex, with crossed factorial designs and/or conditions with more than three levels.
 The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
@@
 
-In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
@@
 Brm() will then construct k-1 indicator variables (see model M1d in Tutorial_2a.Rmd for an example).
-With index coding, one constructs an index variable which contains integers that correspond to different categories (see models M0i and M1i below).
+With index coding, one constructs an index variable that contains integers that correspond to different categories (see models M0i and M1i below).
 As explained by @mcelreathStatisticalRethinkingBayesian2020, the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.
@@
 
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
@@
 
-In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-To keep the tutorial short, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i_ca.
+In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors.
+To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i_ca.
 
@@
 
-In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data set used in Tutorial 1a.
+In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
 
-To keep this tutorial short, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
+Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-The resulting hazard model is called M1i_f.
+The resulting hazard model is called M1i_f with the appended "_f" denoting a frequentist model.
 
-In Figure 11 we compare the parameter estimates of model M1i from brm() with those of model M1i_f from glmer().
+In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i_f.
 
@@
 This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
-But the price you pay for being able to fit more complex random effects models in a Bayesian framework is computation time.
+But the price you pay for being able to fit models with more complex varying effects structures via a Bayesian framework is increased computation time.
 In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
@@
 
-In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-To keep it short, we only fitted the effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
+In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
+To be concise, we only fit effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-Again, the resulting conditional accuracy model M1i_ca_f did not converge and resulted in a singular fit.
+The resulting conditional accuracy model, which we labelled M1i_ca_f, did not converge and resulted in a singular fit.
+Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
 
@@
 
-The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it [@gelmanBayesianWorkflow2020].
-The basic structure and code follows the examples outlined by Solomon Kurz in his
-'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper [@debruineUnderstandingMixedEffectsModels2021].
+The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it [@gelmanBayesianWorkflow2020].
+The basic structure and code follows the examples outlined by Solomon Kurz in his 'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper [@debruineUnderstandingMixedEffectsModels2021].
 
@@
 
-1. Fit a regression model to an existing dataset.
-2. Use the regression model parameters to simulate one new dataset.
+1. Fit a regression model to existing data.
+2. Use the regression model parameters to simulate new data.
 3. Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
@@
 Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data.
-However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be computationally demanding for desktop machines.
+However, when each model takes several hours to build, and we may want to simulate many 1000s of datasets, it can be computationally demanding for desktop machines.
 So, for ease, here we just use the raw simulated datasets to guide future expectations.
@@
 We chose to focus on a subsample of the data to ease the computational burden.
-We also used a full varying effects structure, with the model formula, as follows:
+We also used a full varying effects structure, with the model formula as follows:
 
@@
 Now that we have summarised our simulated data, what planning decisions could we make about a future study?
-How many trials per condition should we collect and how many participants should we test? 
+More concretely, how many trials per condition should we collect and how many participants should we test? 
 Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources [@lakensSampleSizeJustification2022]. 
 There is no straigtforward and clear-cut answer.
-Some considerations might be...
+Some considerations might be as follows:
 
@@
   - Are you running multiple studies that have some form of replication built in?
-  - What resources do you have at your disposal, such as time, money and personnel?
+  - What level of resources do you have at your disposal, such as time, money and personnel?
   - How easy or difficult is it to obtain the specific type of sample?
@@
 If we were running this kind of study in our lab, what would we do?
-We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon [@panisWhatShapingRT2016].
-Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80% power mark.
-If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide >90% power for an effect size of 0.5.
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed previously [@panisWhatShapingRT2016].
+Then we might pick the corresponding combination of trial count per condition (e.g., 200) and participant sample size (e.g., N=10 or N=15) that takes you over the 80% power mark.
+If we wanted to maximise power based on these simulations, and we had the time and resources available, then we would test N=20 participants, which would provide >90% power for an effect size of 0.5.
 
@@
 EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions. 
-For example, some of the models discussed in Tutorial 2a contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+For example, some of the models discussed in Tutorial 2a, which we did not focus on in the main text, contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
@@
 Estimating the temporal distributions of RT and accuracy provide a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
-Statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
 We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing EHA and investigating the temporal profile of cognitive states.

### 2024-12-10 notes/manuscript_notes.Rmd
update ms-notes and tut-notes
update ms-notes
RR notes

@@
 
-clean up html files (delete old figures, Mention tut 2b priors, ...)
+clean up html files (delete old figures, Mention tut 2b priors line 101, ...)
 
@@
 
-check for timebin, multilevel
+check for timebin -> time bin
+
+# Rich Dec 10th
+
+## supplementary figures ##
+
+in the main text, change to Supplementary Figure X (as appropriate). 
+e.g., page 28, Figure 15 when talking about priors.
+
+## figure 11 ##
+
+very minor point. I would make figure smaller in height. There is a lot of white space and a shorter plot woudl still communicate the main message. 
+I would also fatten the point and the lines to make them easier to see.
+Add more dodge if necessary.
+
+## conclusion ##
+
+you say
+
+>Statistically controlling for the passage of time during data analysis is equally important as experimental control during the design of an experiment, to better understand human behavior in experimental paradigms.
+
+But I don't think we need it, and I don't understand the second half of the sentence. If we do need it, we need to make it clear what it means.
+
+For now, I deleted it.

### 2024-12-11 Tutorial_4.Rmd
RR updated tutorial 4.

@@
 pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
-         "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux")
+         "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux",
+         "rstan")
 
@@
         future.rng.onMisuse = "ignore") ## automatically set in RStudio
+
+rstan_options(auto_write = TRUE)
 
@@
 The vignettes provided by the faux package are particularly helpful and worth following before you take a deep dive in the below code. 
+There is also an associated [journal article](https://doi.org/10.1177/2515245920965119), which is worth reading.
 
@@
 
-1. Fit a regression model to an existing dataset.
-2. Use the regression model parameters to simulate one new dataset.
+1. Fit a regression model to existing data.
+2. Use the regression model parameters to simulate new data.
 3. Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
-4. Summarise the simulated data to estimate likely power or precision of the research design options.
+4. Summarise the simulated data to estimate likely power or precision of the various research designs.
 
@@
 The basic workflow above is a modified version of that recommended by Solomon and Lisa.
-In a typical workflow, once the datasets are simulated, a regression model would be fit to each datatset.
-Then, the parameters from each model would be summarised to make a judgment about the likely precision or power of the desing.
-However, in the case of Model_4 in this tutorial, which took ~8 hours to build, we are not able to easily build 1000 models per variation in sample size and/or trial count, as that would take an unreasonably long time.
+In a typical workflow, once the datasets are simulated, a regression model would be fit to each dataset.
+Then, the parameters from each model would be summarised to make a judgment about the likely precision or power of the design.
+However, in the case of some of the models in this tutorial, which can take several hours to build, we are not able to easily build 1000 models per variation in parameter setting, such as sample size and/or trial count, as that would take an unreasonably long time on a desktop machine.
 Instead, we choose to take a simpler approach and summarise the data per bin and condition, rather than the parameter estimates from the model.
 Although this is not as satisfying, we feel it can still be a useful guide to set expectations. 
-To further save time and computational resources, we also choose to use a much simpler model than Model_4, in order to outline the basic approach to planning via data simulation. 
-The details on this simpler model structure are provided below.
+To further save time and computational resources, for the purposes of this planning tutorial, we also choose to use a smaller dataset than was used in the earlier tutorials. 
+The details on this simpler dataset and modelling procedure are provided below.
 
@@
 
-Here, we again use data form Panis and Schmidt (2016), but to make things simpler, we focus on only 6 timebins (timebins 4-9) and two conditions (con and inc).
+Here, we again use data form Panis and Schmidt (2016), but to make things simpler, we focus on only 6 time bins (time bins 4-9) and two prime conditions (congruent and incongruent).
 
@@
 data <- read_csv("Tutorial_4_planning/data/tidyppp_Exp1_n6.csv") %>%
-  ## no mask condition only, neutral prime removed, timebins 4-9 only
+  ## no mask condition only, neutral prime removed, time bins 4-9 only
   filter(no_ma == 1, no_pr < 1, timebin %in% c(4:9)) %>% 
@@
   group_by(pid, timebin, prime) %>% 
-  summarise(outcome = mean(outcome))
+  summarise(outcome = mean(outcome, na.rm = TRUE))
 head(data_pid)
@@
   group_by(timebin, prime) %>% 
-  summarise(outcome = mean(outcome))
+  summarise(outcome = mean(outcome, na.rm = TRUE))
 head(data_group)
@@
 
-So, the approach here is to build an initial model using an index coding approach, but with a much simpler dataset than has been used in prior tutorials.
-
-Because we are building Bayesian regression models, which generate a posterior distribution per condition of interest in our index coding approach, we can simply calculate contrasts of interests and associated prevision or interval estimates (95% quantile intervals, for example) to address any questions we may have about comparisons between conditions, such as inc vs con in a particular timebin.
+The approach here is to build an initial model using an index coding approach, but with a subset of the data that was used in prior tutorials.
+
+Because we are building Bayesian regression models, which generate a posterior distribution per condition of interest in our index coding approach, we can simply calculate contrasts of interests and associated precision or interval estimates (95% quantile intervals, for example) to address any questions we may have about comparisons between conditions, such as incongruent vs congruent in a particular time bin.
 
@@
 
-For an index coding model, priors will be set differently to how they were set in the earlier tutorials, because the earlier tutorials did not use an index coding approach.
 Using an index coding approach, each condition has a prior for the mean of that condition.
@@
 (0,2) looks like it covers all of our likely values, even if it is too broad.
+
+For simplicity, we use these normal(0,2) priors for the mean per condition.
+However, if we were doing this for real, we would probably choose different priors that are similar to ones that were discussed in Tutorial_2a as well as in Supplmentary Materials.
 
@@
 
-- model checking by looking at the convergence of chains and model diagnostics. 
+- check the model by looking at the convergence of chains and other model diagnostics. 
 We'll skip this step for now, just to move on to the main focus of planning and data simulation.
@@
 - Summarise and visualise the parameter estimates and/or posterior predictions.
+
 Below, we quickly summarise fixed effects from the model in cloglog values, and we also convert them to hazard values, which tend to be far easier to interpret.
@@
   labs(title = "Model bi cloglog",
-       x = "timebin", y = "cloglog") +
+       x = "time bin", y = "cloglog") +
   scale_fill_brewer(palette = "Dark2") 
@@
 ggsave ("Tutorial_4_planning/figures/index_cloglog.jpeg",
-        width = 6, height = 4)
+        width = 6, height = 4, dpi = 800)
 ```
@@
   labs(title = "hazard estimates",
-       x = "timebin", y = "hazard") +
+       x = "time bin", y = "hazard") +
   scale_fill_brewer(palette = "Dark2") +
@@
 ggsave ("Tutorial_4_planning/figures/index_hazard.jpeg",
-        width = 6, height = 4)
+        width = 6, height = 4, dpi = 800)
 ```
@@
 
-Note: For multi-level data, data simulation can be quite a convoluted process, and we therefore strongly recommend reading the power blogs by Solomon Kurz and working through the faux package vignettes, before getting stuck into the below code.
+Note: For multi-level data, data simulation can be quite a convoluted process, and we therefore strongly recommend reading the power blogs by Solomon Kurz and working through the faux package [vignettes](https://debruine.github.io/faux/index.html), before getting stuck into the below code.
 
@@
 
-## b1 to b6 = timebin 1 to 6.
+## b1 to b6 = time bin 1 to 6.
 ## c = con
@@
   group_by(subj, condition, bin) %>% 
-  summarise(outcome = mean(outcomen)) %>% 
+  summarise(outcome = mean(outcomen, na.rm = TRUE)) %>% 
   mutate(prime = if_else(condition == "cond1", "con", "inc"),
@@
   group_by(condition, bin) %>% 
-  summarise(outcome = mean(outcomen)) %>% 
+  summarise(outcome = mean(outcomen, na.rm = TRUE)) %>% 
   mutate(prime = if_else(condition == "cond1", "con", "inc"),
@@
 
-this is the descriptive data per bin and condition for hazards
-probability of a response/event, given that it has not happened yet
+this is the descriptive data per bin and condition for the hazard - probability of a response/event, given that it has not happened yet
 
@@
 
-Given that the effect of interest was ~80% reduction in Sven's 2016 paper, let's play around with a range of effect sizes from e.g., 0.25, 0.5, 0.75.
+Given that the effect of interest was ~80% reduction in Sven's 2016 paper (Panis & Schmidt, 2016), let's play around with a range of effect sizes from e.g., 0.25, 0.5, 0.75.
 These would translate to a 75% reduction, 50% reduction and 25% reduction from con to inc.
@@
 create a function to simulate data
-These values are from the initial model and data from Sven's paper, but they can be modified when running the function, which we do below.
+These values are from the initial model and data from Sven's paper, which we created above in section 4, but they can be modified when running the function, which we do below.
 
@@
 
-vary trial count per condition and effect size, keep pid fixed at 10.
+vary trial count per condition and effect size, keep pid fixed at N=10.
 This will take some time, probably hours...go get a beer/coffee.
@@
   scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
-  labs(x="timebin",
+  labs(x="time bin",
        y="outcome") + 
@@
 
-basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound quantiles exclude zero?
+basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound 95% quantile intervals exclude zero?
 This of course is not a model-based estimate.
-But I think it should give a good guide to what we might expect.
+But we think it should give a reasonable guide to what we might expect.
 
@@
 
-- effects of con vs inc that are as large as previously reported by Panis et al., 2016 would be easy to detect with 100 trials and 10 pids. And probably less data than that also.
+- effects of con vs inc that are as large as previously reported by Panis & Schimdt (2016) would be easy to detect with 100 trials and 10 pids. And probably less data than that also.
 
@@
 
-# 11. Run some more follow-up sims #
-
-Let's try a 50%, 60% and 70% reduction across N=10,15 and 20.
+# 11. Run a follow-up simulation #
+
+Let's try a 50%, 60% and 70% reduction across N=10, 15 and 20.
 
@@
 vary effect size and pid.
-Again, as before, this will take some time, probably hours...go get a beer/coffee.
+Again, as before, this will take some time, probably hours...go get another beer/coffee.
 And again, eval is set to FALSE, to save time for the tutorial.
@@
 
-basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound quantiles exclude zero?
-This of course is not a model-based estimate.
-But I think it should give a good guide to what we might expect.
+Same basic idea as in Simulation 1.
 

### 2024-12-11 manuscript/manuscript.Rmd
update ms.

@@
 my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) 
-
 ```

### 2024-12-11 manuscript/manuscript.tex
update ms.

@@
 \abstract{%
-Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
+Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
 However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
@@
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
-To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA.
-We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multilevel regression modelling.
-Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences.
+To help achieve broader uptake of EHA, in this paper we outline a set of tutorials that detail one distributional method known as discrete-time EHA.
+We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design, as well as how to manage inter-individual differences.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
-Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
+Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
 }
@@
 survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
-We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
 Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
@@
 We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
@@
 
-For a comprehensive background context to EHA, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003).
-Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
+We recommend several excellent textbooks for a comprehensive background context to EHA (Allison, 2010; Singer \& Willett, 2003) and for a more general introduction to understanding regression equations (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
 Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
@@
 
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition.
+Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel).
 These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
@@
 We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
-Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of descriptive EHA a lot quicker and more efficient.
+In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
+In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.
 
@@
 
-Our lab adopts a estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
+Our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
 We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
-This means that wherever possible we include varying intercepts and slopes per participant
+This means that wherever possible we include varying intercepts and slopes per participant.
 To make inferences, we use two main approaches.
@@
 In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' data set generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
+Therefore, the `person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
 This data set is used when fitting hazard models (Tutorials 2a and 3a).
@@
 
-\caption{Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6.
+\caption{Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
 Vertical dotted lines indicate the estimated median RTs.
@@
 
-These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
+These distributional results suggest that participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
 Qualitatively similar results were obtained for the other five participants.
-When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).
+When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial\_1a.Rmd).
 
@@
 
-In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data set used in Tutorial 1a.
+In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data used in Tutorial 1a.
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
@@
 Brm() will then construct k-1 indicator variables (see model M1d in Tutorial\_2a.Rmd for an example).
-With index coding, one constructs an index variable which contains integers that correspond to different categories (see models M0i and M1i below).
+With index coding, one constructs an index variable that contains integers that correspond to different categories (see models M0i and M1i below).
 As explained by McElreath (2020), the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.
@@
 
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A.
 The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
@@
 
-In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors.
-To keep the tutorial short, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i\_ca.
+In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors.
+To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i\_ca.
 
@@
 
-In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data set used in Tutorial 1a.
+In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
 
-To keep this tutorial short, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
+Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting hazard model is called M1i\_f.
-
-In Figure 11 we compare the parameter estimates of model M1i from brm() with those of model M1i\_f from glmer().
+The resulting hazard model is called M1i\_f with the appended ``\_f'' denoting a frequentist model.
+
+In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i\_f.
 
@@
 
-In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a.
-To keep it short, we only fitted the effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
+In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
+To be concise, we only fit effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-Again, the resulting conditional accuracy model M1i\_ca\_f did not converge and resulted in a singular fit.
+The resulting conditional accuracy model, which we labelled M1i\_ca\_f, did not converge and resulted in a singular fit.
+Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
 
@@
 
-The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
-The basic structure and code follows the examples outlined by Solomon Kurz in his
-`power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
+The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
+The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
 
@@
 \item
-  Fit a regression model to an existing dataset.
-\item
-  Use the regression model parameters to simulate one new dataset.
+  Fit a regression model to existing data.
+\item
+  Use the regression model parameters to simulate new data.
 \item
@@
 Now that we have summarised our simulated data, what planning decisions could we make about a future study?
-How many trials per condition should we collect and how many participants should we test?
+More concretely, how many trials per condition should we collect and how many participants should we test?
 Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022).
 There is no straigtforward and clear-cut answer.
-Some considerations might be\ldots{}
+Some considerations might be as follows:
 
@@
 \item
-  What resources do you have at your disposal, such as time, money and personnel?
+  What level of resources do you have at your disposal, such as time, money and personnel?
 \item
@@
 If we were running this kind of study in our lab, what would we do?
-We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon (Panis \& Schmidt, 2016).
-Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80\% power mark.
-If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
+We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed previously (Panis \& Schmidt, 2016).
+Then we might pick the corresponding combination of trial count per condition (e.g., 200) and participant sample size (e.g., N=10 or N=15) that takes you over the 80\% power mark.
+If we wanted to maximise power based on these simulations, and we had the time and resources available, then we would test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
 
@@
 EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions.
-For example, some of the models discussed in Tutorial 2a contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
+For example, some of the models discussed in Tutorial 2a, which we did not focus on in the main text, contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).

### 2024-12-13 manuscript/manuscript.Rmd
update sims and ms

@@
 (ref:descr-plot3-caption) Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Discrete time-to-event and conditional accuracy data were simulated for 1000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
 Error bars represent +/- 1 standard error of the respective proportion.

### 2024-12-13 manuscript/manuscript.tex
update sims and ms

@@
 \caption{Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 2000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
+Discrete time-to-event and conditional accuracy data were simulated for 1000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
 Error bars represent +/- 1 standard error of the respective proportion.

### 2024-12-13 sims/sims.Rmd
update sims and ms

@@
 
-This file simulates the data used in Figure 1 of the manuscript and then creates some plots.
+This file simulates the data used in Figures 1, 2, and 3 of the manuscript and then creates some plots.
 
@@
 
-```{r}
+```{r figure-1}
 p_haz_inset_facet <- p8.1b +
@@
 
-# Section 11: Simulate hazard data for 1 subject, 1 condition, 200 trials, and plot empirical hazard and survivor functions with error bars (Figure 2)
+# Section 11 (Figure 2): Simulate hazard data for 1 subject, 1 condition, 200 trials, and plot empirical hazard and survivor functions with error bars
 
@@
 
-## hazard
-
-### generate data set
+## generate data set
 
@@
 ## apply function sim_haz
-data_single2 <- sim_haz(subj_n = 1, rep_n = 1000) %>%
+ntrials <- 1000
+
+data_single2 <- sim_haz(subj_n = 1, rep_n = ntrials) %>%
                mutate(period = bin,
@@
                mutate(acc = ifelse(period<6 & condition=="incongruent",0,1),
-                      acc = ifelse(period==6 & condition=="incongruent" & trial < 500,0,acc)) 
+                      acc = ifelse(period==6 & condition=="incongruent" & trial < ntrials/2,0,acc)) 
 head(data_single2, n=20)
 summary(data_single2)
-
-```
-
-### set up life table
+```
+
+## set up life table
 
@@
 
-## add ca(t)
+## get ca(t)
 
@@
 
-### plot hazard functions
+## plot hazard functions
 
@@
 
-## accuracy
-
-create some data
-
-```{r}
-acc <- tibble(
-  time = seq(1,10,1),
-  congruent = rep(1,length = 10),
-  incongruent = c(0, 0, 0, 0, 0, .5, 1, 1, 1, 1)
-) %>% 
-  pivot_longer(-time,
-               names_to = "condition",
-               values_to = "accuracy") %>% 
-  mutate(condition = factor(condition))
-head(acc,n=20)
-```
-
-line plot per bin
+## plot ca
 
@@
 
-Combine plots
+## Combine plots
 

### 2024-12-13 Tutorial_2b.Rmd
update 2b and ms.

@@
 ## Set up priors for M1i_ca.
+
+As we expect conditional accuracies ranging between (and including) 0 and 1, we use a flat prior on the logit scale for our intercepts (see section E in the supplemental material).
 
@@
 
-#
-```{r plot-post-M0i-ca-alt}
-tidy_fixed_ca <- post_ca %>% 
-  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
-  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
-  pivot_longer(-c(chain, draw, iter)) %>% 
-  mutate(timebin = str_sub(name,10,11),
-         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
-         condition = str_sub(name,17,18),
-         condition = factor(str_remove(condition,"e"),
-                            levels=c(1,2,3),
-                            labels=c("blank","congruent","incongruent")))
-head(tidy_fixed_ca)
-tail(tidy_fixed_ca)
-
-# plot
-p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
-  stat_halfeye(point_interval = "median_qi",
-               .width = c(.80,.95),
-               alpha=0.7,
-               position = pd) +
-  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
-       x = "time bin", 
-       y = "logit-ca") +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-  scale_y_continuous(limits=c(-10,10)) +
-  theme(axis.text.x = element_text(angle=90)) #+
-  #facet_wrap(~condition)
-p_tidy_fixed_ca
-```
-
-```{r save-post-alt, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr_alt.png", width = 10, height = 8, dpi = 800)
-```
-
-#
-
 ## Plot predicted conditional accuracy functions
@@
 
-Save plot.
+Save combined plot.
 
@@
 ```{r grand-ame-plot}
- ggplot(epreds_grand_diffs, aes(x=timebin, y=diff_ca, 
+p11 <- ggplot(epreds_grand_diffs, aes(x=timebin, y=diff_ca, 
                        fill=contrast)) +
@@
   labs(y = "difference in predicted ca") +
-  ggtitle("Grand average marginal effect") +
+#  ggtitle("Grand average marginal effect") +
   facet_wrap(~contrast)
+p11
 ```
@@
 ```{r epred-pid-ame-plot}
- ggplot(epreds_pid_diffs, aes(x=timebin, y=diff_ca, 
+p22 <- ggplot(epreds_pid_diffs, aes(x=timebin, y=diff_ca, 
                        fill=contrast, color=contrast)) +
@@
     labs(y = "difference in predicted ca") +
-    ggtitle("Subject-specific AMEs") +
+   # ggtitle("Subject-specific AMEs") +
     facet_wrap(~pid, ncol=3)
+p22
 ```
@@
 ```
+
+Combine the plots.
+
+```{r save-ames-combined-ca, eval=F}
+# combine 2 plots
+p11_theme <- p11 + theme(legend.position = "none")
+
+(p11_theme / p22) +
+    plot_annotation(tag_levels = 'A') +
+   # plot_layout(guides = "collect",
+  #              axes = "collect_x") & 
+    theme(legend.position = "bottom")
+```
+
+Save combined plot.
+
+```{r save-combined-ames-ca, eval=F}
+# save combined plot
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png", width = 10, height = 12, dpi = 800)
+```

### 2024-12-13 manuscript/manuscript.Rmd
update 2b and ms.

@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10; grand average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
 
-(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average.
+(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average (A), and for each participant (B).
 
 ```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png")
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png")
 ```
 
-Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], and (280,320], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10).
-Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
+Based on Figure 10A we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
 

### 2024-12-13 manuscript/manuscript.tex
update 2b and ms.

@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10; grand average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_grand_AME}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined}
 
@@
 
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average.}\label{fig:plot-ca-grand-ame-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-ca-grand-ame-effects}
 \end{figure}
 
-Based on Figure 10 we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, and (280,320{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10).
-Incongruent primes have a negative effect on the conditional accuracy of emitted responses in those time bins, relative to the estimates in the baseline condition.
+Based on Figure 10A we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
 

### 2024-12-13 Tutorial_1a.Rmd
update tut 1a and ms.

@@
 data_wr <- data_wr %>% 
-  rename(pid = vp, condition = prime_type, acc = respac, trial = TrialNr)
+  rename(pid = vp, condition = prime_type, acc = respac, trial = TrialNr) %>%
 
@@
 # create factor for condition (prime type)
-data_wr <- data_wr %>% 
   mutate(condition = condition + 1,

### 2024-12-13 manuscript/manuscript.Rmd
update tut 1a and ms.

@@
 data_wr<-read_csv("../Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
-colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
 data_wr <- data_wr %>% 
+  rename(pid = vp, condition = prime_type, acc = respac, trial = TrialNr) %>% 
   mutate(condition = condition + 1, # original levels were 0, 1, 2.

### 2024-12-13 manuscript/manuscript.tex
update tut 1a and ms.

@@
 \NormalTok{data\_wr}\OtherTok{\textless{}{-}}\FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/DataExp1\_6subjects\_wrangled.csv"}\NormalTok{)}
-\FunctionTok{colnames}\NormalTok{(data\_wr) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"pid"}\NormalTok{,}\StringTok{"bl"}\NormalTok{,}\StringTok{"tr"}\NormalTok{,}\StringTok{"condition"}\NormalTok{,}\StringTok{"resp"}\NormalTok{,}\StringTok{"acc"}\NormalTok{,}\StringTok{"rt"}\NormalTok{,}\StringTok{"trial"}\NormalTok{) }
 \NormalTok{data\_wr }\OtherTok{\textless{}{-}}\NormalTok{ data\_wr }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{pid =}\NormalTok{ vp, }\AttributeTok{condition =}\NormalTok{ prime\_type, }\AttributeTok{acc =}\NormalTok{ respac, }\AttributeTok{trial =}\NormalTok{ TrialNr) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =}\NormalTok{ condition }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, }\CommentTok{\# original levels were 0, 1, 2.}

### 2024-12-13 manuscript/Suppl_material/supplementary.Rmd
update ms. and new suppl-material folder

--- 2024-12-10 manuscript/SupplementaryMaterial.Rmd
@@
-# Supplementary material
+---
+title             : no
+shorttitle        : "A tutorial on hazard analysis"
+bibliography      : ["extrareferences2.bib"]
+floatsintext      : yes
+linenumbers       : yes
+draft             : no
+mask              : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+classoption       : "man"
+output            : papaja::apa6_pdf
+editor_options: 
+  chunk_output_type: console
+header-includes:
+  - \raggedbottom
+---
+
+```{r setup, include = FALSE}
+pkg <- c("papaja", "citr", "tidyverse", "RColorBrewer", "patchwork")
+
+lapply(pkg, library, character.only = TRUE)
+```
+
+```{r analysis-preferences}
+# Seed for random number generation
+set.seed(42)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+```
+
+```{r plot-settings}
+## theme settings for ggplot
+theme_set(
+  theme_bw() +
+    theme(text = element_text(size = 18, face = "bold"), 
+          title = element_text(size = 18, face = "bold"),
+          legend.position = "bottom")
+)
+
+## Set the amount of dodge in figures
+pd <- position_dodge(0.7)
+pd2 <- position_dodge(1)
+```
+
+```{r global-chunk-settings}
+## set the figure options
+knitr::opts_chunk$set(fig.pos='H', out.extra = '', out.height = "67%",
+                      fig.align = "center") # initial version
+```
+
+<!-- the below two lines directly change the default Figure and Table naming to Supplementary -->
+
+<!-- see here for more details - https://stackoverflow.com/questions/47173279/internationalization-r-knitr-figure-caption-label -->
+
+\renewcommand{\figurename}{Supplementary Figure}
+\renewcommand{\tablename}{Supplemetentary Table}
+
+# Supplemental Material
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 14.
+Popular link functions include the logit link and the complementary log-log link, as shown in Supplementary Figure 1.
 
@@
 ```{r plot-link-functions, fig.cap = "(ref:descr-plot14-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
+knitr::include_graphics("../../Tutorial_2_Bayesian/figures/linkfunctions.png")
 ```
@@
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 15 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Supplementary Figure 2 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
@@
 ```{r plot-priors, fig.cap = "(ref:descr-plot15-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")
+knitr::include_graphics("../../Tutorial_2_Bayesian/figures/plot_of_priors.png")
 ```
@@
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+
+# References

### 2024-12-13 manuscript/Suppl_material/supplementary.tex
update ms. and new suppl-material folder

--- 2024-12-13 manuscript/manuscript.tex
@@
 \documentclass[
-  man, donotrepeattitle,floatsintext]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 \usepackage{xcolor}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 13574}
 \usepackage{lineno}
@@
 \hypersetup{
-  pdftitle={Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows},
-  pdfauthor={Sven Panis1 \& Richard Ramsey1},
   pdflang={en-EN},
-  pdfkeywords={response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology},
   hidelinks,
@@
 
-\title{Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows}
-\author{Sven Panis\textsuperscript{1} \& Richard Ramsey\textsuperscript{1}}
+\author{\phantom{0}}
 \date{}
@@
 
-\authornote{
-
-Neural Control of Movement lab, Department of Health Sciences and Technology (D-HEST).
-Social Brain Sciences lab, Department of Humanities, Social and Political Sciences (D-GESS).
-
-The authors made the following contributions.
-Sven Panis: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Richard Ramsey: Conceptualization, Writing - Review \& Editing, Supervision.
-
-Correspondence concerning this article should be addressed to Sven Panis, ETH GLC, room G16.2, Gloriastrasse 37/39, 8006 Zürich.
-E-mail: \href{mailto:sven.panis@hest.ethz.ch}{\nolinkurl{sven.panis@hest.ethz.ch}}
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} ETH Zürich}
-
-\abstract{%
-Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
-However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
-The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states.
-One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.
-To help achieve broader uptake of EHA, in this paper we outline a set of tutorials that detail one distributional method known as discrete-time EHA.
-We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design, as well as how to manage inter-individual differences.
-We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
-Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
-}
+\affiliation{\phantom{0}}
 
 \begin{document}
-\maketitle
 
-\section{1. Introduction}\label{introduction}
+\renewcommand{\figurename}{Supplementary Figure}
+\renewcommand{\tablename}{Supplemetentary Table}
 
-\subsection{1.1 Motivation and background context: Comparing means versus distributional shapes}\label{motivation-and-background-context-comparing-means-versus-distributional-shapes}
-
-In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
-Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
-
-As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, right).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../sims/figures/haz_inset_facet}
-
-}
-
-\caption{Means versus distributional shapes for six different simulated data set examples.
-The first second after stimulus onset is divided in ten bins of 100 ms.
-Timebin indicates the bin rank.
-The first bin is (0,100{]}, the last bin is (900,1000{]}.
-For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
-Insets show mean response time per condition.}\label{fig:plot1}
-\end{figure}
-
-Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in timebin 2) and is gone when the waiting time reaches 500 ms or more.
-In contrast, in example 5, the effect starts around 300 ms and is gone by 700 ms.
-And in example 6, the effect reverses between 500 and 600 ms.
-What kind of theory or theories could account for such effects?
-Are there new auxiliary assumptions that theories need to adopt?
-And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
-As we show later using published examples, for many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-
-From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1693; Makeham, 1860).
-The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
-
-\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
-
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
-We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
-
-We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003).
-Indeed, our primary aim here is to introduce the set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
-
-We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
-Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
-Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.
-
-In Tutorial 1a, we illustrate how to process or ``wrangle'' a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
-The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
-
-In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
-We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
-In Tutorial 2b we fit Bayesian multilevel regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
-
-In Tutorial 3a, we shortly illustrate how to fit similar multilevel regression models for RT data in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to EHA.
-In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
-
-In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
-We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
-
-In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multilevel regression modelling steps.
-Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
-
-\section{2. A brief introduction to event history analysis}\label{a-brief-introduction-to-event-history-analysis}
-
-We recommend several excellent textbooks for a comprehensive background context to EHA (Allison, 2010; Singer \& Willett, 2003) and for a more general introduction to understanding regression equations (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
-Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
-Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in section D of the supplementary material.
-
-\subsection{2.1 Basic features of event history analysis}\label{basic-features-of-event-history-analysis}
-
-To apply EHA, one must be able to:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
-\item
-  define time point zero (e.g., target stimulus onset, fixation onset, etc.);
-\item
-  measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
-\end{enumerate}
-
-In EHA, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
-Since our focus here is on hazard models that use discrete time units, we describe that approach.
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600 ms).
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in part A of the supplementary material.
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_surv_single}
-
-}
-
-\caption{Discrete-time hazard and survivor functions.
-Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
-Error bars represent +/- 1 standard error of the respective proportion.
-While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t=5) = .44.
-Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.}\label{fig:plot2}
-\end{figure}
-
-\subsection{2.2 Benefits of event history analysis}\label{benefits-of-event-history-analysis}
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the supplementary material for more detailed coverage of the benefits.
-The benefits include:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
-\item
-  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
-\item
-  EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
-\item
-  Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
-\item
-  EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
-\end{enumerate}
-
-\subsection{2.3 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
-
-To make EHA more relevant to researchers studying cognitive psychology and cognitive neuroscience, in this section we provide a relevant worked example and consider implications that are relevant to that domain of research.
-
-\subsubsection{2.3.1 A worked example}\label{a-worked-example}
-
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a discrimination task.
-For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
-
-Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical choice RT example that is inspired by real data (Panis \& Schmidt, 2016), but simplified to make the main point clearer (Figure 3).
-In a standard priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
-The prime can then be congruent or incongruent with the target.
-
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel).
-These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
-For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
-
-This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT and accuracy across trials.
-Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials, and are indexed by each recorded response.
-But such a conclusion is not supported when the effects are explored over a timeline.
-Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_acc_single}
-
-}
-
-\caption{Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 1000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
-Error bars represent +/- 1 standard error of the respective proportion.
-Bin width equals 100 ms.}\label{fig:plot3}
-\end{figure}
-
-Unlocking the temporal states of cognitive processes can be revealing for theory development and the understanding of basic psychological processes.
-Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
-Do all participants show similar temporal states or are there individual differences?
-Do such individual differences extend to those individuals that have been diagnosed with some form of psychopathology?
-How do temporal states relate to brain-based mechanisms that might be studied using other methods from cognitive neuroscience?
-And how much of theory in cognitive psychology would be in need of revision if mean-average comparisons were supplemented with a temporal states approach?
-
-\subsubsection{2.3.2 Implications for designing experiments}\label{implications-for-designing-experiments}
-
-Performing EHA in experimental psychology has implications for how experiments are designed.
-Indeed, if trials are categorised as a function of when responses occur, then each timebin will only include a subset of the total number of trials.
-For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
-Those 100 trials must be distributed in some manner across the chosen number of bins.
-
-In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
-Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
-This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
-
-In contrast to the large-\emph{N} design that typically average across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
-This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
-
-\section{3. An overview of the general analytical workflow}\label{an-overview-of-the-general-analytical-workflow}
-
-Although the focus is on EHA/SAT, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
-
-\subsection{3.1 Data science workflow and descriptive statistics}\label{data-science-workflow-and-descriptive-statistics}
-
-We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
-In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.
-
-\subsection{3.2 Inferential statistical approach}\label{inferential-statistical-approach}
-
-Our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
-We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
-This means that wherever possible we include varying intercepts and slopes per participant.
-To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020).
-We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
-
-\subsection{3.3 Implementation}\label{implementation}
-
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-
-\section{4. Tutorials}\label{tutorials}
-
-Tutorials 1a and 1b show how to calculate and plot the descriptive statistics of EHA/SAT when there are one or two independent variables, respectively.
-Tutorials 2a and 2b illustrate how to use Bayesian multilevel modeling to fit hazard and conditional accuracy models, respectively.
-Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section B in the supplemental material for a list of the custom functions).
-
-Our list of tutorials is as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  1a. Wrangle raw data and calculate descriptive stats for one independent variable
-\item
-  1b. Wrangle raw data and calculate descriptive stats for two independent variables
-\item
-  2a. Bayesian multilevel modeling for h(t)
-\item
-  2b. Bayesian multilevel modeling for ca(t)
-\item
-  3a. Frequentist multilevel modeling for h(t)
-\item
-  3b. Frequentist multilevel modeling for ca(t)
-\item
-  \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \setcounter{enumi}{3}
-  \tightlist
-  \item
-    Simulation and power analysis for planning experiments
-  \end{enumerate}
-\end{itemize}
-
-\subsection{4.1 Tutorial 1a: Calculating descriptive statistics using a life table}\label{tutorial-1a-calculating-descriptive-statistics-using-a-life-table}
-
-\subsubsection{4.1.1 Data wrangling aims}\label{data-wrangling-aims}
-
-Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes, using a life table.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
-
-Second, we want to produce two different data sets that can each be submitted to different types of inferential modelling approaches.
-The two types of data structure we label as `person-trial' data and `person-trial-bin' data.
-The `person-trial' data (Table 1) will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment.
-This data set is used when fitting conditional accuracy models (Tutorials 2b and 3b).
-
-\begin{table}[H]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:ca-data-table}Data structure for `person-trial' data}
-
-\begin{tabular}{lllll}
-\toprule
-pid & \multicolumn{1}{c}{trial} & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{rt} & \multicolumn{1}{c}{accuracy}\\
-\midrule
-1 & 1 & congruent & 373.49 & 1\\
-1 & 2 & incongruent & 431.31 & 1\\
-1 & 3 & congruent & 455.43 & 0\\
-1 & 4 & incongruent & 622.41 & 1\\
-1 & 5 & incongruent & 535.98 & 1\\
-1 & 6 & incongruent & 540.08 & 1\\
-1 & 7 & congruent & 511.07 & 1\\
-1 & 8 & incongruent & 444.42 & 1\\
-1 & 9 & congruent & 678.69 & 1\\
-1 & 10 & congruent & 549.79 & 1\\
-\bottomrule
-\addlinespace
-\end{tabular}
-
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} The first 10 trials for participant 1 are shown. These data are simulated and for illustrative purposes only.}
-\end{tablenotes}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial.
-Therefore, the `person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-This data set is used when fitting hazard models (Tutorials 2a and 3a).
-It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
-Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
-
-\begin{table}[H]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:ha-data-table}Data structure for `person-trial-bin' data}
-
-\begin{tabular}{lllll}
-\toprule
-pid & \multicolumn{1}{c}{trial} & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{timebin} & \multicolumn{1}{c}{event}\\
-\midrule
-1 & 1 & congruent & 1 & 0\\
-1 & 1 & congruent & 2 & 0\\
-1 & 1 & congruent & 3 & 0\\
-1 & 1 & congruent & 4 & 1\\
-1 & 2 & incongruent & 1 & 0\\
-1 & 2 & incongruent & 2 & 0\\
-1 & 2 & incongruent & 3 & 0\\
-1 & 2 & incongruent & 4 & 0\\
-1 & 2 & incongruent & 5 & 1\\
-\bottomrule
-\addlinespace
-\end{tabular}
-
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} The first 2 trials for participant 1 from Table 1 are shown. The width of the time bins is 100 ms. These data are simulated and for illustrative purposes only.}
-\end{tablenotes}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsubsection{4.1.2 A real data wrangling example}\label{a-real-data-wrangling-example}
-
-To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016).
-In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
-Participants had to indicate the direction in which the double arrow pointed using their corresponding index finger, within 800 ms after target onset.
-Response time and accuracy were recorded on each trial.
-Prime type (blank, congruent, incongruent) and mask type were manipulated.
-Here we focus on the subset of trials in which no mask was presented.
-The 13-ms prime stimulus was a double arrow presented 187 ms before target onset in the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
-
-There are several data wrangling steps to be taken. First, we need to load the data before we (a) supply required column names, and (b) specify the factor condition with the correct levels and labels.
-
-The required column names are as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  ``pid'', indicating unique participant IDs;
-\item
-  ``trial'', indicating each unique trial per participant;
-\item
-  ``condition'', a factor indicating the levels of the independent variable (1, 2, \ldots) and the corresponding labels;
-\item
-  ``rt'', indicating the response times in ms;
-\item
-  ``acc'', indicating the accuracies (1/0).
-\end{itemize}
-
-In the code of Tutorial 1a, this is accomplished as follows.
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{data\_wr}\OtherTok{\textless{}{-}}\FunctionTok{read\_csv}\NormalTok{(}\StringTok{"../Tutorial\_1\_descriptive\_stats/data/DataExp1\_6subjects\_wrangled.csv"}\NormalTok{)}
-\NormalTok{data\_wr }\OtherTok{\textless{}{-}}\NormalTok{ data\_wr }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{rename}\NormalTok{(}\AttributeTok{pid =}\NormalTok{ vp, }\AttributeTok{condition =}\NormalTok{ prime\_type, }\AttributeTok{acc =}\NormalTok{ respac, }\AttributeTok{trial =}\NormalTok{ TrialNr) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condition =}\NormalTok{ condition }\SpecialCharTok{+} \DecValTok{1}\NormalTok{, }\CommentTok{\# original levels were 0, 1, 2.}
-         \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }
-                            \AttributeTok{levels=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{), }
-                            \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{,}\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the supplementary material for their definitions.
-To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{data\_nested }\OtherTok{\textless{}{-}}\NormalTok{ data\_wr }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{group\_nest}\NormalTok{(pid)}
-\NormalTok{data\_final }\OtherTok{\textless{}{-}}\NormalTok{ data\_nested }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\# ! user input: censoring time, and bin width}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{censored =} \FunctionTok{map}\NormalTok{(data, censor, }\DecValTok{600}\NormalTok{, }\DecValTok{40}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# create person{-}trial{-}bin data set}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}          
-  \CommentTok{\# create life tables without ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}   
-  \CommentTok{\# calculate ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condacc =} \FunctionTok{map}\NormalTok{(censored, calc\_ca)) }\SpecialCharTok{\%\textgreater{}\%}      
-  \CommentTok{\# create life tables with ca(t)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable\_ca =} \FunctionTok{map2}\NormalTok{(lifetable, condacc, join\_lt\_ca)) }\SpecialCharTok{\%\textgreater{}\%}    
-  \CommentTok{\# create plots }
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{plot =} \FunctionTok{map2}\NormalTok{(}\AttributeTok{.x =}\NormalTok{ lifetable\_ca, }\AttributeTok{.y =}\NormalTok{ pid, plot\_eha,}\DecValTok{1}\NormalTok{))  }
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-Note that the censoring time should be a multiple of the bin width (both in ms).
-The censoring time should be a time point after which no informative responses are expected anymore.
-In experiments that implement a response deadline in each trial the censoring time can equal that deadline time point.
-Trials with a RT larger than the censoring time, or trials in which no response is emitted during the data collection period, are treated as right-censored observations in EHA.
-In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
-Removing such trials before calculating the mean event time will result in underestimation of the true mean.
-
-The person-trial-bin oriented data set is created by our custom function ptb(), and it has one row for each time bin (of each trial) that is at risk for event occurrence.
-The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-
-The next step is to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha().
-When creating the plots, some warning messages will likely be generated, like these:
-
-\begin{itemize}
-\tightlist
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_line()}).
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_point()}).
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_segment()}).
-\end{itemize}
-
-The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
-One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
-In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
-
-Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
-
-\begin{table}[H]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:life-table}The life table for the blank prime condition of participant 6.}
-
-\begin{tabular}{lllllllll}
-\toprule
-bin & \multicolumn{1}{c}{risk\_set} & \multicolumn{1}{c}{events} & \multicolumn{1}{c}{hazard} & \multicolumn{1}{c}{se\_haz} & \multicolumn{1}{c}{survival} & \multicolumn{1}{c}{se\_surv} & \multicolumn{1}{c}{ca} & \multicolumn{1}{c}{se\_ca}\\
-\midrule
-0 & 220 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
-40 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-80 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-120 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-160 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-200 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-240 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-280 & 220 & 7 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
-320 & 213 & 13 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
-360 & 200 & 26 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
-400 & 174 & 40 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
-440 & 134 & 48 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
-480 & 86 & 37 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
-520 & 49 & 32 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00\\
-560 & 17 & 9 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00\\
-600 & 8 & 4 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00\\
-\bottomrule
-\addlinespace
-\end{tabular}
-
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} The column named ``bin'' indicates the endpoint of each time bin (in ms), and includes time point zero.
-For example the first bin is (0,40{]} with the starting point excluded and the endpoint included.
-At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
-se = standard error.
-ca = conditional accuracy.
-NA = undefined.}
-\end{tablenotes}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
-By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=21.33in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
-
-}
-
-\caption{Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
-Vertical dotted lines indicate the estimated median RTs.
-Error bars represent +/- 1 standard error of the respective proportion.}\label{fig:eha-plot}
-\end{figure}
-
-For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
-
-Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-
-However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.06 for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
-
-These distributional results suggest that participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus.
-Qualitatively similar results were obtained for the other five participants.
-When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial\_1a.Rmd).
-
-In general, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
-Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
-
-At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA/SAT.
-As we will show in later Tutorials, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
-But first we consider calculating the descriptive statistics for two independent variables.
-
-\subsection{4.2 Tutorial 1b: Generalising to a more complex design}\label{tutorial-1b-generalising-to-a-more-complex-design}
-
-So far in this paper, we have used a simple experimental design, which involved one condition with three levels.
-But psychological experiments are often more complex, with crossed factorial designs and/or conditions with more than three levels.
-The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
-We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
-
-To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial\_1b.Rmd for those that are interested.
-
-\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-discrete-time-to-event-data}
-
-In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data used in Tutorial 1a.
-Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).
-
-\subsubsection{4.3.1 Hazard model considerations}\label{hazard-model-considerations}
-
-There are several analytic decisions one has to make when fitting a discrete-time hazard model.
-First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material).
-The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor.
-And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding.
-With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level.
-Brm() will then construct k-1 indicator variables (see model M1d in Tutorial\_2a.Rmd for an example).
-With index coding, one constructs an index variable that contains integers that correspond to different categories (see models M0i and M1i below).
-As explained by McElreath (2020), the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.
-
-In the case of a large-\emph{N} design without repeated measurements, the parameters of a discrete-time hazard model can be estimated using standard logistic regression software after expanding the typical person-trial data set into a person-trial-bin data set (Allison, 2010).
-When there is clustering in the data, as in the case of a small-\emph{N} design with repeated measurements, the parameters of a discrete-time hazard model can be estimated using population-averaged methods (e.g., Generalized Estimating Equations), and Bayesian or frequentist generalized linear mixed models (Allison, 2010).
-
-In general, there are three assumptions one can make or relax when adding experimental predictor variables and other covariates: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).
-
-In tutorial\_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We decided to select the analysis time window (200,600{]} and the cloglog link.
-Below, we shortly discuss two of these models.
-The person-trial-bin data set is prepared as follows.
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# read in the file we saved in tutorial 1a}
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"Tutorial\_1\_descriptive\_stats/data/inputfile\_hazard\_modeling.csv"}\NormalTok{)}
-
-\NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} 
-\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
-\FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-       \CommentTok{\# define categorical predictor TIME as index variable named timebin}
-\FunctionTok{mutate}\NormalTok{(}\AttributeTok{timebin =} \FunctionTok{factor}\NormalTok{(period, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{15}\NormalTok{)),}
-       \CommentTok{\# factor "condition" using reference coding, with "blank" as the reference level}
-       \AttributeTok{condition =} \FunctionTok{factor}\NormalTok{(condition, }\AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"blank"}\NormalTok{, }\StringTok{"congruent"}\NormalTok{,}\StringTok{"incongruent"}\NormalTok{)),}
-       \CommentTok{\# categorical predictor "prime" with index coding}
-       \AttributeTok{prime =} \FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"blank"}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FunctionTok{ifelse}\NormalTok{(condition}\SpecialCharTok{==}\StringTok{"congruent"}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{)),}
-       \AttributeTok{prime =} \FunctionTok{factor}\NormalTok{(prime, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-\subsubsection{4.3.2 Prior distributions}\label{prior-distributions}
-
-To get the posterior distribution of each model parameter given the data, we need to specify prior distributions for the model parameters which reflect our prior beliefs.
-In Tutorial\_2a.Rmd we perform a few prior predictive checks to make sure our selected prior distributions reflect our prior beliefs (Gelman, Vehtari, et al., 2020).
-
-The middle column of Figure 15 in section E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 15 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
-
-\subsubsection{4.3.3 Model M0i: A null model with index coding}\label{model-m0i-a-null-model-with-index-coding}
-
-When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
-In this first model, we use a general specification of TIME using index coding, and do not include experimental predictors.
-We call this model ``M0i''.
-
-Before we fit model M0i, we select the necessary columns from the data, and specify our priors. In the code of Tutorial 2a, model M0i is specified as follows.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_M0i }\OtherTok{\textless{}{-}}                    
-   \FunctionTok{brm}\NormalTok{(}\AttributeTok{data =}\NormalTok{ data\_M0i,}
-       \AttributeTok{family =} \FunctionTok{bernoulli}\NormalTok{(}\AttributeTok{link=}\StringTok{"cloglog"}\NormalTok{),}
-       \AttributeTok{formula =}\NormalTok{ event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin }\SpecialCharTok{|}\NormalTok{ pid),}
-       \AttributeTok{prior =}\NormalTok{ priors\_M0i,}
-       \AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}\NormalTok{, }
-       \AttributeTok{iter =} \DecValTok{3000}\NormalTok{, }\AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-       \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.999}\NormalTok{, }
-                      \AttributeTok{step\_size =} \FloatTok{0.04}\NormalTok{, }
-                      \AttributeTok{max\_treedepth =} \DecValTok{12}\NormalTok{),}
-       \AttributeTok{seed =} \DecValTok{12}\NormalTok{, }\AttributeTok{init =} \StringTok{"0"}\NormalTok{,}
-       \AttributeTok{file =} \StringTok{"Tutorial\_2\_Bayesian/models/model\_M0i"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-After selecting the bernoulli family and the cloglog link, the model formula is specified.
-The specification ``0 + \ldots{}'' removes the default intercept in brm().
-The fixed effects include an intercept for each level of timebin.
-Each of these intercepts is allowed to vary across individuals (variable pid).
-We request 2000 samples from the posterior distribution for each of four chains.
-Estimating model M0i took about 30 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
-
-\subsubsection{4.3.4 Model M1i: Adding the effects of prime-target congruency}\label{model-m1i-adding-the-effects-of-prime-target-congruency}
-
-Previous research has shown that psychological effects typically change over time (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009).
-In the next model, therefore, we use index coding for both TIME (variable ``timebin'') and the categorical predictor prime-target-congruency (variable ``prime''), so that we get 30 grand intercepts, one for each combination of timebin level and prime level.
-Here is the model formula of this model that we call ``M1i''.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{|}\NormalTok{ pid)}
-\end{Highlighting}
-\end{Shaded}
-
-Estimating model M1i took about 124 minutes.
-
-\subsubsection{4.3.5 Compare the models.}\label{compare-the-models.}
-
-We can compare the two models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2020).
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M0i model_M1i 
-##         0         1
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## model_M0i model_M1i 
-##         0         1
-\end{verbatim}
-
-\normalsize
-
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other simpler model.
-
-\subsubsection{4.3.6 Evaluating parameter estimates in model M1i}\label{evaluating-parameter-estimates-in-model-m1i}
-
-To make inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80\% and 95\% credible intervals).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_postdistr}
-
-}
-
-\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i.}\label{fig:plot-fixed-effects}
-\end{figure}
-
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the grand average (Figure 6A), and for each participant in the data set (Figure 6B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_combined}
-
-}
-
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
-\end{figure}
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the grand average (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ame_combined}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
-\end{figure}
-
-The point estimates and quantile intervals can be reported in a table (see Tutorial\_2a.Rmd for details).
-
-\paragraph{Example conclusions for M1i}\label{example-conclusions-for-m1i}
-
-What can we conclude from model M1i about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
-In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of 187 ms)?
-
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A.
-The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
-For the other bins, the 95\% credible interval contained zero.
-The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin (320,360{]} (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin (360,400{]} (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin (400,440{]} (95\% CrI = {[}-0.40, -0.03{]}).
-For the other bins, the 95\% credible interval contained zero.
-
-There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
-In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240{]}.
-In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440{]}.
-The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
-
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 7B.
-Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
-
-Future studies could (a) increase the number of participants to estimate the proportion of ``dippers'' in the subject population, and/or (b) try to explain why this dip occurs.
-For example, Panis and Schmidt (2016) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
-Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to.
-To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after \textasciitilde360 ms, and towards the end of the experiment.
-This effect might be masked for incongruent primes by the response competition effect.
-
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 7B).
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
-In Tutorial\_2a.Rmd we provide a number of model formulae that should get you going.
-
-\subsection{4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}
-
-In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors.
-To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i\_ca.
-
-To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 8, together with point (median) and interval estimates (80\% and 95\% credible intervals).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_postdistr}
-
-}
-
-\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca. ca = conditional accuracy.}\label{fig:plot-ca-fixed-effects}
-\end{figure}
-
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the grand average (Figure 9A), and for each participant in the data set (Figure 9B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined}
-
-}
-
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
-\end{figure}
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-ca-grand-ame-effects}
-\end{figure}
-
-Based on Figure 10A we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
-Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
-
-\subsection{4.5 Tutorial 3a: Fitting Frequentist hazard models}\label{tutorial-3a-fitting-frequentist-hazard-models}
-
-In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
-The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-
-Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting hazard model is called M1i\_f with the appended ``\_f'' denoting a frequentist model.
-
-In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i\_f.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_3_Frequentist/comparison}
-
-}
-
-\caption{Parameter estimates for model M1i from brm() -- means and 95\% credible intervals -- and model M1i\_f from glmer() -- maximum likelihood estimates and 95\% confidence intervals.}\label{fig:plot-comparison}
-\end{figure}
-
-Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
-However, model M1i\_f did not converge and resulted in a singular fit.
-This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
-But the price you pay for being able to fit models with more complex varying effects structures via a Bayesian framework is increased computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
-
-\subsection{4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models}\label{tutorial-3b-fitting-frequentist-conditional-accuracy-models}
-
-In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
-To be concise, we only fit effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting conditional accuracy model, which we labelled M1i\_ca\_f, did not converge and resulted in a singular fit.
-Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
-
-\subsection{4.7 Tutorial 4: Planning}\label{tutorial-4-planning}
-
-In the final tutorial, we look at planning a future experiment, which uses EHA.
-
-\subsubsection{4.7.1 Background}\label{background}
-
-The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
-The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
-
-\subsubsection{4.7.2 Basic workflow}\label{basic-workflow}
-
-The basic workflow is as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Fit a regression model to existing data.
-\item
-  Use the regression model parameters to simulate new data.
-\item
-  Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
-\item
-  Summarise the simulated data to estimate likely power or precision of the research design options.
-\end{enumerate}
-
-Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data.
-However, when each model takes several hours to build, and we may want to simulate many 1000s of datasets, it can be computationally demanding for desktop machines.
-So, for ease, here we just use the raw simulated datasets to guide future expectations.
-
-In the below, we only provide a high-level summary of the process and let readers dive into the details within the tutorial should they feel so inclined.
-
-\subsubsection{4.7.3 Fit a regression model and simulate one dataset}\label{fit-a-regression-model-and-simulate-one-dataset}
-
-We again use the data from Panis and Schmidt (2016) to provide a worked example.
-We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
-We chose to focus on a subsample of the data to ease the computational burden.
-We also used a full varying effects structure, with the model formula as follows:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{event }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ timebin}\SpecialCharTok{:}\NormalTok{prime }\SpecialCharTok{|}\NormalTok{ pid)}
-\end{Highlighting}
-\end{Shaded}
-
-We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
-The raw data and the simulated data are plotted in Figure 12 and show quite close correspondence, which is re-assuring.
-But, this is only one dataset.
-What we really want to do is simulate many datasets and vary parameters of interest, which is what we turn to in the next section.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_4_Planning/figures/raw_vs_d1}
-
-}
-
-\caption{Raw data from Panis and Schmidt (2016) and simulated data from 10 participants.}\label{fig:simdat}
-\end{figure}
-
-\subsubsection{4.7.4 Simulate and summarise data across a range of parameter values}\label{simulate-and-summarise-data-across-a-range-of-parameter-values}
-
-Here we use the same data simulation process as used above, but instead of simulating one dataset, we simulate 1000 datasets per variation in parameter values.
-Specifically, in Simulation 1, we vary the number of trials per condition (100, 200, and 400), as well as the effect size in bin 6.
-We focus on bin 6 only, in terms of varying the effect size, just to make things simpler and easier to understand.
-The effect size observed in bin 6 in this subsample of data was a 79\% reduction in hazard value from the congruent prime (0.401 hazard value) to the incongruent prime condition (0.085 hazard value).
-In other words, a hazard ratio of 0.21 (e.g., 0.085/0.401 = 0.21).
-As a starting point, we chose three effect sizes, which covered a fairly broad range of hazard ratios (0.25, 0.5, 0.75), which correspond to a 75\%, 50\% and 25\% reduction in hazard value as a function of prime condition.
-
-Summary results from Simulation 1 are shown in Figure 13A.
-Figure 13A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
-In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
-We are aware that ``power'' is not part of a Bayesian analytical workflow, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
-
-The results of Simulation 1 show that if we were targeting an effect size similar to the one reported in the original study, then testing 10 participants and collecting 100 trials per condition would be enough to provide over 95\% power.
-However, we could not be as confident about smaller effects, such as a hazard ratio of 50\% or 25\%.
-From this simulation, we can see that somewhere between an effect size of a 50\% and 75\% reduction in hazard value, power increases to a range that most researchers would consider acceptable (i.e., \textgreater95\% power).
-To probe this space a little further, we decided to run a second simulation, which varied different parameters.
-
-In Simulation 2, we varied the effect size between a different range of values (0.5, 0.4, 0.3), which correspond to a 50\%, 60\% and 70\% reduction in hazard value as a function of prime condition.
-In addition, we varied the number of participants per experiment between 10, 15, and 20 participants.
-Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
-Summary results from Simulation 2 are shown in Figure 13B.
-A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):
-
-\begin{itemize}
-\tightlist
-\item
-  For a 70\% reduction (0.3 hazard ratio), N=10 would give nearly 100\% power.
-\item
-  For a 60\% reduction (0.4 hazard ratio), N=10 would give nearly 90\% power.
-\item
-  For a 50\% reduction (0.5 hazard ratio), N=15 would give over 80\% power.
-\end{itemize}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\linewidth,height=0.67\textheight,]{../Tutorial_4_Planning/figures/bin6_power_col}
-
-}
-
-\caption{Statistical power across data Simulation 1 (A) and Simulation 2 (B).
-Power was calculated as the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
-In Simulation 1, the effect size was varied between a 25\%, 50\% and 75\% reduction in hazard value, whereas the trial count was varied between 100, 200 and 400 trials per condition (the number of participants was fixed at N=10).
-In Simulation 2, the effect size was varied between a 50\%, 60\% and 70\% reduction in hazard value, whereas the number of participants was varied between N=10, 15 and 20 (the number of trials per condition was fixed at 200).
-The dashed lines represent 80\% (red), 90\% (black) and 95\% (blue) power.
-Abbreviations: rep\_n = the number of trials per experimental condition; subj\_n = the number of participants per simulated experiment.}\label{fig:power}
-\end{figure}
-
-\subsubsection{4.7.5 Planning decisions}\label{planning-decisions}
-
-Now that we have summarised our simulated data, what planning decisions could we make about a future study?
-More concretely, how many trials per condition should we collect and how many participants should we test?
-Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022).
-There is no straigtforward and clear-cut answer.
-Some considerations might be as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  How much power or precision are you looking to obtain in this particular study?
-\item
-  Are you running multiple studies that have some form of replication built in?
-\item
-  What level of resources do you have at your disposal, such as time, money and personnel?
-\item
-  How easy or difficult is it to obtain the specific type of sample?
-\end{itemize}
-
-If we were running this kind of study in our lab, what would we do?
-We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed previously (Panis \& Schmidt, 2016).
-Then we might pick the corresponding combination of trial count per condition (e.g., 200) and participant sample size (e.g., N=10 or N=15) that takes you over the 80\% power mark.
-If we wanted to maximise power based on these simulations, and we had the time and resources available, then we would test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
-
-\textbf{But}, and this is an important ``but'', unless there are unavoidable reasons, no matter what planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
-Instead, we would run a follow-up experiment that replicates and extends the initial result.
-By doing so, we would aim to avoid the Cult of the Isolated Single Study (Nelder, 1999; Tong, 2019), and thus reduce the reliance on any one type of planning tool, such as a power analysis.
-Then, we would look for common patterns across two or more experiments, rather than trying to make the case that a single study on its own has sufficient evidential value to hit some criterion mark.
-
-\section{5. Discussion}\label{discussion}
-
-This main motivation for writing this paper is the observation that EHA and SAT analysis remain under-used in psychological research.
-As a consequence, the field of psychological research is not taking full advantage of the many benefits EHA/SAT provides compared to more conventional analyses.
-By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to mean-average comparisons.
-While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
-As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss issues relating to model complexity and interpretability, individual differences, as well as limitations of the approach and future extensions.
-
-\subsection{5.1 What are the main use-cases of EHA for understanding cognition and brain function?}\label{what-are-the-main-use-cases-of-eha-for-understanding-cognition-and-brain-function}
-
-For those researchers, like ourselves, who are primarily interested in understanding human cognitive and brain systems, we consider two broadly-defined, main use-cases of EHA.
-First, as we hope to have made clear by this point, EHA is one way to investigating a ``temporal states'' approach to cognitive processes.
-EHA provides one way to uncover when cognitive states may start and stop, as well as what they may be tied to or interact with.
-Therefore, if your research questions concern \textbf{when} and \textbf{for how long} psychological states occur, our EHA tutorials could be useful tools for you to use.
-
-Second, even if you are not primarily interested in studying the temporal states of cognition, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on mean-average comparisons.
-Given that distinctly different inferences can be made from the same data based on whether one computes a mean-average across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement mean-average comparisons with EHA.
-One could envisage scenarios where the implicit assumption of an effect manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain aspects of time.
-
-\subsection{5.2 Model complexity versus interpretability}\label{model-complexity-versus-interpretability}
-
-EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions.
-For example, some of the models discussed in Tutorial 2a, which we did not focus on in the main text, contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
-However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
-From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
-As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
-This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4).
-
-\subsection{5.3 Individual differences}\label{individual-differences}
-
-One important issue is that of possible individual differences in the overall location of the distribution, and the time course of psychological effects.
-For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
-These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 500 ms for simple detection tasks, 800 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
-Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response in the analysis time window), introducing a response deadline is recommended when designing RT experiments.
-Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard and conditional accuracy models.
-
-But even when using a response deadline, participants can differ qualitatively in the effects they display (see Panis, 2020).
-One way to deal with this is to describe and interpret the different patterns.
-Another way is to run a clustering algorithm on the individual hazard estimates across all bins and conditions.
-The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
-One might then exclude the smaller sub-group of participants before fitting a hazard model or consider the possibility that different cognitive processes may be at play during task performance across the different sub-groups.
-
-Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021), which is a from of small-\emph{N} approach (Smith \& Little, 2018).
-This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
-This approach allows one to quantify how typical or uncommon an observed effect is in the population, and the uncertainty around this estimate.
-
-\subsection{5.4 Limitations}\label{limitations}
-
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
-
-Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
-Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
-Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
-For instance, if the maximum session length deemed reasonable is between 1 and 2 hours, what is the maximum number of trials per condition that you could reasonably collect?
-After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
-Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
-
-We did not discuss continuous-time EHA, nor continuous-time SAT analysis.
-As indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods.
-Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
-However, they require much more data to estimate the continuous-time hazard (rate) function well.
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.
-
-\subsection{5.5 Extensions}\label{extensions}
-
-The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this button-press event constitutes a single transition between an ``idle'' state and a ``responded'' state.
-However, in certain situations, more than one event of interest might exist.
-For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before being absorbed into a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize EHA to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
-Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (Allison, 2010).
-For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
-Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.
-
-\section{6. Conclusions}\label{conclusions}
-
-Estimating the temporal distributions of RT and accuracy provide a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
-We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing EHA and investigating the temporal profile of cognitive states.
-On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.
-
-\newpage
-
-\section{References}\label{references}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-allisonDiscreteTimeMethodsAnalysis1982}
-Allison, P. D. (1982). Discrete-{Time Methods} for the {Analysis} of {Event Histories}. \emph{Sociological Methodology}, \emph{13}, 61. \url{https://doi.org/10.2307/270718}
-
-\bibitem[\citeproctext]{ref-allisonSurvivalAnalysisUsing2010}
-Allison, P. D. (2010). \emph{Survival analysis using {SAS}: A practical guide} (2. ed). Cary, NC: SAS Press.
-
-\bibitem[\citeproctext]{ref-R-citr}
-Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://github.com/crsh/citr}
-
-\bibitem[\citeproctext]{ref-R-papaja}
-Aust, F., \& Barth, M. (2023). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\bibitem[\citeproctext]{ref-pap2024}
-Aust, F., \& Barth, M. (2024). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
-
-\bibitem[\citeproctext]{ref-bakerPowerContoursOptimising2021}
-Baker, D.
-H., Vilidaite, G., Lygo, F.
-A., Smith, A.
-K., Flack, T.
-R., Gouws, A.
-D., \& Andrews, T.
-J.
-(2021).
-Power contours: {Optimising} sample size and precision in experimental psychology and human neuroscience.
-\emph{Psychological Methods}, \emph{26}(3), 295--314.
-\url{https://doi.org/10.1037/met0000337}
-
-\bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
-Barr, D.
-J., Levy, R., Scheepers, C., \& Tily, H.
-J.
-(2013).
-Random effects structure for confirmatory hypothesis testing: {Keep} it maximal.
-\emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001.
-\url{https://doi.org/10.1016/j.jml.2012.11.001}
-
-\bibitem[\citeproctext]{ref-R-tinylabels}
-Barth, M. (2023). \emph{{tinylabels}: Lightweight variable labels}. Retrieved from \url{https://cran.r-project.org/package=tinylabels}
-
-\bibitem[\citeproctext]{ref-R-lme4}
-Bates, D., Mächler, M., Bolker, B., \& Walker, S.
-(2015).
-Fitting linear mixed-effects models using {lme4}.
-\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
-\url{https://doi.org/10.18637/jss.v067.i01}
-
-\bibitem[\citeproctext]{ref-R-Matrix}
-Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://CRAN.R-project.org/package=Matrix}
-
-\bibitem[\citeproctext]{ref-R-RJ-2021-048}
-Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
-
-\bibitem[\citeproctext]{ref-blossfeldTechniquesEventHistory2002}
-Blossfeld, H.-P., \& Rohwer, G. (2002). \emph{Techniques of event history modeling: {New} approaches to causal analysis, 2nd ed} (pp. x, 310). Mahwah, NJ, US: Lawrence Erlbaum Associates Publishers.
-
-\bibitem[\citeproctext]{ref-box-steffensmeierEventHistoryModeling2004}
-Box-Steffensmeier, J. M. (2004). Event history modeling: A guide for social scientists. Cambridge: University Press.
-
-\bibitem[\citeproctext]{ref-R-brms_a}
-Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
-
-\bibitem[\citeproctext]{ref-R-brms_b}
-Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411. \url{https://doi.org/10.32614/RJ-2018-017}
-
-\bibitem[\citeproctext]{ref-R-brms_c}
-Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with {brms} and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}
-
-\bibitem[\citeproctext]{ref-debruineUnderstandingMixedEffectsModels2021}
-DeBruine, L.
-M., \& Barr, D.
-J.
-(2021).
-Understanding {Mixed-Effects Models Through Data Simulation}.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920965119.
-\url{https://doi.org/10.1177/2515245920965119}
-
-\bibitem[\citeproctext]{ref-R-Rcpp_b}
-Eddelbuettel, D., \& Balamuta, J.
-J.
-(2018).
-{Extending {R} with {C++}: A Brief Introduction to {Rcpp}}.
-\emph{The American Statistician}, \emph{72}(1), 28--36.
-\url{https://doi.org/10.1080/00031305.2017.1375990}
-
-\bibitem[\citeproctext]{ref-R-Rcpp_a}
-Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. \url{https://doi.org/10.18637/jss.v040.i08}
-
-\bibitem[\citeproctext]{ref-R-cmdstanr}
-Gabry, J., Češnovar, R., Johnson, A., \& Bronder, S. (2024). \emph{Cmdstanr: R interface to 'CmdStan'}. Retrieved from \url{https://github.com/stan-dev/cmdstanr}
-
-\bibitem[\citeproctext]{ref-R-bayesplot}
-Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in bayesian workflow. \emph{J. R. Stat. Soc. A}, \emph{182}, 389--402. \url{https://doi.org/10.1111/rssa.12378}
-
-\bibitem[\citeproctext]{ref-gelmanRegressionOtherStories2020}
-Gelman, A., Hill, J., \& Vehtari, A.
-(2020).
-Regression and {Other Stories}.
-https://www.cambridge.org/highereducation/books/regression-and-other-stories/DD20DD6C9057118581076E54E40C372C; Cambridge University Press.
-\url{https://doi.org/10.1017/9781139161879}
-
-\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
-Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., \ldots{} Modrák, M. (2020). \emph{Bayesian {Workflow}}. arXiv. \url{https://doi.org/10.48550/arXiv.2011.01808}
-
-\bibitem[\citeproctext]{ref-R-standist}
-Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
-
-\bibitem[\citeproctext]{ref-R-lubridate}
-Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. Retrieved from \url{https://www.jstatsoft.org/v40/i03/}
-
-\bibitem[\citeproctext]{ref-Halley1693}
-Halley, E.
-(1693).
-VI.
-An estimate of the degrees of the mortality of mankind; drawn from curious tables of the births and funerals at the city of breslaw; with an attempt to ascertain the price of annuities upon lives.
-\emph{Philosophical Transactions of the Royal Society of London}, \emph{17}(196), 596--610.
-\url{https://doi.org/10.1098/rstl.1693.0007}
-
-\bibitem[\citeproctext]{ref-heiss2021}
-Heiss, A. (2021, November 10). A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel {Bayesian} Models. \url{https://doi.org/10.59350/wbn93-edb02}
-
-\bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
-Holden, J.
-G., Van Orden, G.
-C., \& Turvey, M.
-T.
-(2009).
-Dispersion of response times reveals cognitive dynamics.
-\emph{Psychological Review}, \emph{116}(2), 318--342.
-\url{https://doi.org/10.1037/a0014849}
-
-\bibitem[\citeproctext]{ref-hosmerAppliedSurvivalAnalysis2011}
-Hosmer, D. W., Lemeshow, S., \& May, S. (2011). \emph{Applied {Survival Analysis}: {Regression Modeling} of {Time} to {Event Data}} (2nd ed). Hoboken: John Wiley \& Sons.
-
-\bibitem[\citeproctext]{ref-inceBayesianInferencePopulation2021}
-Ince, R. A., Paton, A. T., Kay, J. W., \& Schyns, P. G. (2021). Bayesian inference of population prevalence. \emph{eLife}, \emph{10}, e62461. \url{https://doi.org/10.7554/eLife.62461}
-
-\bibitem[\citeproctext]{ref-kantowitzInterpretationReactionTime2021}
-Kantowitz, B.
-H., \& Pachella, R.
-G.
-(2021).
-The {Interpretation} of {Reaction Time} in {Information-Processing Research} 1.
-\emph{Human Information Processing}, 41--82.
-\url{https://doi.org/10.4324/9781003176688-2}
-
-\bibitem[\citeproctext]{ref-R-tidybayes}
-Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
-
-\bibitem[\citeproctext]{ref-kelsoOutlineGeneralTheory2013}
-Kelso, J.
-A.
-S., Dumas, G., \& Tognoli, E.
-(2013).
-Outline of a general theory of behavior and brain coordination.
-\emph{Neural Networks: The Official Journal of the International Neural Network Society}, \emph{37}, 120--131.
-\url{https://doi.org/10.1016/j.neunet.2012.09.003}
-
-\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
-Kruschke, J.
-K., \& Liddell, T.
-M.
-(2018).
-The {Bayesian New Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power analysis from a {Bayesian} perspective.
-\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 178--206.
-\url{https://doi.org/10.3758/s13423-016-1221-4}
-
-\bibitem[\citeproctext]{ref-kurzAppliedLongitudinalDataAnalysis2023}
-Kurz, A. S. (2023a). \emph{Applied longitudinal data analysis in brms and the tidyverse} (version 0.0.3). Retrieved from \url{https://bookdown.org/content/4253/}
-
-\bibitem[\citeproctext]{ref-kurzStatisticalRethinkingSecondEd2023}
-Kurz, A. S. (2023b). \emph{Statistical rethinking with brms, ggplot2, and the tidyverse: {Second} edition} (version 0.4.0). Retrieved from \url{https://bookdown.org/content/4857/}
-
-\bibitem[\citeproctext]{ref-lakensSampleSizeJustification2022}
-Lakens, D. (2022). Sample {Size Justification}. \emph{Collabra: Psychology}, \emph{8}(1), 33267. \url{https://doi.org/10.1525/collabra.33267}
-
-\bibitem[\citeproctext]{ref-landesIntroductionEventHistory2020}
-Landes, J., Engelhardt, S. C., \& Pelletier, F. (2020). An introduction to event history analyses for ecologists. \emph{Ecosphere}, \emph{11}(10), e03238. \url{https://doi.org/10.1002/ecs2.3238}
-
-\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
-Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
-
-\bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-Makeham, W. M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
-
-\bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
-McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
-
-\bibitem[\citeproctext]{ref-meyerModernMentalChronometry1988}
-Meyer, D. E., Osman, A. M., Irwin, D. E., \& Yantis, S. (1988). Modern mental chronometry. \emph{Biological Psychology}, \emph{26}(1-3), 3--67. \url{https://doi.org/10.1016/0301-0511(88)90013-0}
-
-\bibitem[\citeproctext]{ref-R-tibble}
-Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}
-
-\bibitem[\citeproctext]{ref-nelderStatisticsStatisticalScience1999}
-Nelder, J.
-A.
-(1999).
-From {Statistics} to {Statistical Science}.
-\emph{Journal of the Royal Statistical Society.
-Series D (The Statistician)}, \emph{48}(2), 257--269.
-Retrieved from \url{https://www.jstor.org/stable/2681191}
-
-\bibitem[\citeproctext]{ref-R-RColorBrewer}
-Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}
-
-\bibitem[\citeproctext]{ref-panisHowCanWe2020}
-Panis, S.
-(2020).
-How can we learn what attention is?
-{Response} gating via multiple direct routes kept in check by inhibitory control processes.
-\emph{Open Psychology}, \emph{2}(1), 238--279.
-\url{https://doi.org/10.1515/psych-2020-0107}
-
-\bibitem[\citeproctext]{ref-panisStudyingDynamicsVisual2020}
-Panis, S., Moran, R., Wolkersdorfer, M.
-P., \& Schmidt, T.
-(2020).
-Studying the dynamics of visual search behavior using {RT} hazard and micro-level speed--accuracy tradeoff functions: {A} role for recurrent object recognition and cognitive control processes.
-\emph{Attention, Perception, \& Psychophysics}, \emph{82}(2), 689--714.
-\url{https://doi.org/10.3758/s13414-019-01897-z}
-
-\bibitem[\citeproctext]{ref-panisAnalyzingResponseTimes2020}
-Panis, S., Schmidt, F., Wolkersdorfer, M.
-P., \& Schmidt, T.
-(2020).
-Analyzing {Response Times} and {Other Types} of {Time-to-Event Data Using Event History Analysis}: {A Tool} for {Mental Chronometry} and {Cognitive Psychophysiology}.
-\emph{I-Perception}, \emph{11}(6), 2041669520978673.
-\url{https://doi.org/10.1177/2041669520978673}
-
-\bibitem[\citeproctext]{ref-panisWhatShapingRT2016}
-Panis, S., \& Schmidt, T.
-(2016).
-What {Is Shaping RT} and {Accuracy Distributions}?
-{Active} and {Selective Response Inhibition Causes} the {Negative Compatibility Effect}.
-\emph{Journal of Cognitive Neuroscience}, \emph{28}(11), 1651--1671.
-\url{https://doi.org/10.1162/jocn_a_00998}
-
-\bibitem[\citeproctext]{ref-panisWhenDoesInhibition2022}
-Panis, S., \& Schmidt, T.
-(2022).
-When does {``inhibition of return''} occur in spatial cueing tasks?
-{Temporally} disentangling multiple cue-triggered effects using response history and conditional accuracy analyses.
-\emph{Open Psychology}, \emph{4}(1), 84--114.
-\url{https://doi.org/10.1515/psych-2022-0005}
-
-\bibitem[\citeproctext]{ref-panisNeuropsychologicalEvidenceTemporal2017}
-Panis, S., Torfs, K., Gillebert, C.
-R., Wagemans, J., \& Humphreys, G.
-W.
-(2017).
-Neuropsychological evidence for the temporal dynamics of category-specific naming.
-\emph{Visual Cognition}, \emph{25}(1-3), 79--99.
-\url{https://doi.org/10.1080/13506285.2017.1330790}
-
-\bibitem[\citeproctext]{ref-panisTimecourseContingenciesPerceptual2009}
-Panis, S., \& Wagemans, J.
-(2009).
-Time-course contingencies in perceptual organization and identification of fragmented object outlines.
-\emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{35}(3), 661--687.
-\url{https://doi.org/10.1037/a0013547}
-
-\bibitem[\citeproctext]{ref-R-patchwork}
-Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://CRAN.R-project.org/package=patchwork}
-
-\bibitem[\citeproctext]{ref-R-nlme}
-Pinheiro, J. C., \& Bates, D. M. (2000). \emph{Mixed-effects models in s and s-PLUS}. New York: Springer. \url{https://doi.org/10.1007/b98882}
-
-\bibitem[\citeproctext]{ref-R-base}
-R Core Team. (2024). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\bibitem[\citeproctext]{ref-ripleyMASSSupportFunctions2024}
-Ripley, B., Venables, B., Bates, D.
-M., ca 1998), K.
-H.
-(partial.
-port, ca 1998), A.
-G.
-(partial.
-port, \& polr), D.
-F.
-(support.
-functions for.
-(2024).
-\emph{{MASS}: {Support Functions} and {Datasets} for {Venables} and {Ripley}'s {MASS}}.
-
-\bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
-Singer, J. D., \& Willett, J. B. (2003). \emph{Applied {Longitudinal Data Analysis}: {Modeling Change} and {Event Occurrence}}. Oxford, New York: Oxford University Press.
-
-\bibitem[\citeproctext]{ref-smithSmallBeautifulDefense2018}
-Smith, P.
-L., \& Little, D.
-R.
-(2018).
-Small is beautiful: {In} defense of the small-{N} design.
-\emph{Psychonomic Bulletin \& Review}, \emph{25}(6), 2083--2101.
-\url{https://doi.org/10.3758/s13423-018-1451-8}
-
-\bibitem[\citeproctext]{ref-steeleGeneralMultilevelMultistate2004}
-Steele, F., Goldstein, H., \& Browne, W.
-(2004).
-A general multilevel multistate competing risks model for event history data, with an application to a study of contraceptive use dynamics.
-\emph{Statistical Modelling}, \emph{4}(2), 145--159.
-\url{https://doi.org/10.1191/1471082X04st069oa}
-
-\bibitem[\citeproctext]{ref-teachmanAnalyzingSocialProcesses1983}
-Teachman, J.
-D.
-(1983).
-Analyzing social processes: {Life} tables and proportional hazards models.
-\emph{Social Science Research}, \emph{12}(3), 263--301.
-\url{https://doi.org/10.1016/0049-089X(83)90015-7}
-
-\bibitem[\citeproctext]{ref-tongStatisticalInferenceEnables2019}
-Tong, C.
-(2019).
-Statistical {Inference Enables Bad Science}; {Statistical Thinking Enables Good Science}.
-\emph{The American Statistician}, \emph{73}(sup1), 246--261.
-\url{https://doi.org/10.1080/00031305.2018.1518264}
-
-\bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}
-Townsend, J.
-T.
-(1990).
-Truth and consequences of ordinal differences in statistical distributions: {Toward} a theory of hierarchical inference.
-\emph{Psychological Bulletin}, \emph{108}(3), 551--567.
-\url{https://doi.org/10.1037/0033-2909.108.3.551}
-
-\bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}
-Wickelgren, W. A. (1977). Speed-accuracy tradeoff and information processing dynamics. \emph{Acta Psychologica}, \emph{41}(1), 67--85. \url{https://doi.org/10.1016/0001-6918(77)90012-9}
-
-\bibitem[\citeproctext]{ref-R-ggplot2}
-Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-forcats}
-Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}
-
-\bibitem[\citeproctext]{ref-R-stringr}
-Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-tidyverse}
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., \ldots{} Yutani, H.
-(2019).
-Welcome to the {tidyverse}.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
-
-\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
-Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G.
-(2023).
-\emph{R for data science: Import, tidy, transform, visualize, and model data} (2nd edition).
-Beijing Boston Farnham Sebastopol Tokyo: O'Reilly.
-
-\bibitem[\citeproctext]{ref-R-dplyr}
-Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-purrr}
-Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
-
-\bibitem[\citeproctext]{ref-R-readr}
-Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-tidyr}
-Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-winterStatisticsLinguistsIntroduction2019}
-Winter, B. (2019). \emph{Statistics for {Linguists}: {An Introduction Using R}}. New York: Routledge. \url{https://doi.org/10.4324/9781315165547}
-
-\bibitem[\citeproctext]{ref-wolkersdorferTemporalDynamicsSequential2020}
-Wolkersdorfer, M.
-P., Panis, S., \& Schmidt, T.
-(2020).
-Temporal dynamics of sequential motor activation in a dual-prime paradigm: {Insights} from conditional accuracy and hazard functions.
-\emph{Attention, Perception, \& Psychophysics}, \emph{82}(5), 2581--2602.
-\url{https://doi.org/10.3758/s13414-020-02010-5}
-
-\end{CSLReferences}
-
-\newpage
-
-\section{Supplementary material}\label{supplementary-material}
+\section{Supplemental Material}\label{supplemental-material}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../../Tutorial_2_Bayesian/figures/linkfunctions}
 
@@
 
-To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 15 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
+To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
+Row C in Supplementary Figure 2 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+According to Kurz (2023), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../../Tutorial_2_Bayesian/figures/plot_of_priors}
 
@@
 First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density {[}and mass{]} functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
+Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are the preferred ways to describe the RT + accuracy data.
@@
 (Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer et al., 1988).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
 
@@
 EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-allisonSurvivalAnalysisUsing2010}
+Allison, P. D. (2010). \emph{Survival analysis using {SAS}: A practical guide} (2. ed). Cary, NC: SAS Press.
+
+\bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
+Holden, J.
+G., Van Orden, G.
+C., \& Turvey, M.
+T.
+(2009).
+Dispersion of response times reveals cognitive dynamics.
+\emph{Psychological Review}, \emph{116}(2), 318--342.
+\url{https://doi.org/10.1037/a0014849}
+
+\bibitem[\citeproctext]{ref-kantowitzInterpretationReactionTime2021}
+Kantowitz, B.
+H., \& Pachella, R.
+G.
+(2021).
+The {Interpretation} of {Reaction Time} in {Information-Processing Research} 1.
+\emph{Human Information Processing}, 41--82.
+\url{https://doi.org/10.4324/9781003176688-2}
+
+\bibitem[\citeproctext]{ref-kelsoOutlineGeneralTheory2013}
+Kelso, J.
+A.
+S., Dumas, G., \& Tognoli, E.
+(2013).
+Outline of a general theory of behavior and brain coordination.
+\emph{Neural Networks: The Official Journal of the International Neural Network Society}, \emph{37}, 120--131.
+\url{https://doi.org/10.1016/j.neunet.2012.09.003}
+
+\bibitem[\citeproctext]{ref-kurzAppliedLongitudinalDataAnalysis2023}
+Kurz, A. S. (2023). \emph{Applied longitudinal data analysis in brms and the tidyverse} (version 0.0.3). Retrieved from \url{https://bookdown.org/content/4253/}
+
+\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
+Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
+\bibitem[\citeproctext]{ref-meyerModernMentalChronometry1988}
+Meyer, D. E., Osman, A. M., Irwin, D. E., \& Yantis, S. (1988). Modern mental chronometry. \emph{Biological Psychology}, \emph{26}(1-3), 3--67. \url{https://doi.org/10.1016/0301-0511(88)90013-0}
+
+\bibitem[\citeproctext]{ref-panisStudyingDynamicsVisual2020}
+Panis, S., Moran, R., Wolkersdorfer, M.
+P., \& Schmidt, T.
+(2020).
+Studying the dynamics of visual search behavior using {RT} hazard and micro-level speed--accuracy tradeoff functions: {A} role for recurrent object recognition and cognitive control processes.
+\emph{Attention, Perception, \& Psychophysics}, \emph{82}(2), 689--714.
+\url{https://doi.org/10.3758/s13414-019-01897-z}
+
+\bibitem[\citeproctext]{ref-panisAnalyzingResponseTimes2020}
+Panis, S., Schmidt, F., Wolkersdorfer, M.
+P., \& Schmidt, T.
+(2020).
+Analyzing {Response Times} and {Other Types} of {Time-to-Event Data Using Event History Analysis}: {A Tool} for {Mental Chronometry} and {Cognitive Psychophysiology}.
+\emph{I-Perception}, \emph{11}(6), 2041669520978673.
+\url{https://doi.org/10.1177/2041669520978673}
+
+\bibitem[\citeproctext]{ref-panisWhatShapingRT2016}
+Panis, S., \& Schmidt, T.
+(2016).
+What {Is Shaping RT} and {Accuracy Distributions}?
+{Active} and {Selective Response Inhibition Causes} the {Negative Compatibility Effect}.
+\emph{Journal of Cognitive Neuroscience}, \emph{28}(11), 1651--1671.
+\url{https://doi.org/10.1162/jocn_a_00998}
+
+\bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}
+Townsend, J.
+T.
+(1990).
+Truth and consequences of ordinal differences in statistical distributions: {Toward} a theory of hierarchical inference.
+\emph{Psychological Bulletin}, \emph{108}(3), 551--567.
+\url{https://doi.org/10.1037/0033-2909.108.3.551}
+
+\bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}
+Wickelgren, W. A. (1977). Speed-accuracy tradeoff and information processing dynamics. \emph{Acta Psychologica}, \emph{41}(1), 67--85. \url{https://doi.org/10.1016/0001-6918(77)90012-9}
+
+\end{CSLReferences}
 

### 2024-12-13 manuscript/manuscript.Rmd
update ms. and new suppl-material folder

@@
 classoption       : "man, donotrepeattitle"
-appendix          : SupplementaryMaterial.Rmd
+#appendix          : SupplementaryMaterial.Rmd
 output            : papaja::apa6_pdf
@@
 Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in section D of the supplementary material.
+As such, we only supply regression equations in section D of the Supplemental Material.
 
@@
 Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in part A of the supplementary material.
+The equations that define both of these functions are reported in part A of the Supplemental Material.
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
@@
 We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the supplementary material for more detailed coverage of the benefits.
+Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
 The benefits include:
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section B in the supplemental material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section B in the Supplemental Material for a list of the custom functions).
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the supplementary material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the Supplemental Material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the Supplemental Material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
@@
 
-The middle column of Figure 15 in section E of the supplementary material shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B in Figure 15 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Supplementary Figure 2 (section E of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 2 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 :::
-
-\newpage
-
-```{r child = "SupplementaryMaterial.Rmd"}
-```

### 2024-12-13 manuscript/manuscript.tex
update ms. and new suppl-material folder

@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the supplementary material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the Supplemental Material for their definitions.
 To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
@@
 First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material).
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the Supplemental Material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
@@
 
-\bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
-Holden, J.
-G., Van Orden, G.
-C., \& Turvey, M.
-T.
-(2009).
-Dispersion of response times reveals cognitive dynamics.
-\emph{Psychological Review}, \emph{116}(2), 318--342.
-\url{https://doi.org/10.1037/a0014849}
-
 \bibitem[\citeproctext]{ref-hosmerAppliedSurvivalAnalysis2011}
@@
 
-\bibitem[\citeproctext]{ref-kelsoOutlineGeneralTheory2013}
-Kelso, J.
-A.
-S., Dumas, G., \& Tognoli, E.
-(2013).
-Outline of a general theory of behavior and brain coordination.
-\emph{Neural Networks: The Official Journal of the International Neural Network Society}, \emph{37}, 120--131.
-\url{https://doi.org/10.1016/j.neunet.2012.09.003}
-
 \bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
@@
 Landes, J., Engelhardt, S. C., \& Pelletier, F. (2020). An introduction to event history analyses for ecologists. \emph{Ecosphere}, \emph{11}(10), e03238. \url{https://doi.org/10.1002/ecs2.3238}
-
-\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
-Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
 
@@
 
-\bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}
-Townsend, J.
-T.
-(1990).
-Truth and consequences of ordinal differences in statistical distributions: {Toward} a theory of hierarchical inference.
-\emph{Psychological Bulletin}, \emph{108}(3), 551--567.
-\url{https://doi.org/10.1037/0033-2909.108.3.551}
-
 \bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}
@@
 
-\newpage
-
-\section{Supplementary material}\label{supplementary-material}
-
-\subsection{A.
-Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-probability-mass-and-conditional-accuracy-functions}
-
-The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991).
-After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-Because waiting times can only increase, discrete-time EHA focuses on the discrete-time hazard function
-
-\noindent h(t) = P(RT = t\textbar{} RT \(\geq\) t) \hfill  (1)
-
-\noindent and the discrete-time survivor function
-
-\noindent S(t) = P(RT \(>\) t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} \hfill  (2)
-
-\noindent and not on the probability mass function
-
-\noindent P(t) = P(RT = t) = h(t).S(t-1) \hfill  (3)
-
-\noindent nor the cumulative distribution function
-
-\noindent F(t) = P(RT \(\leq\) t) = 1-S(t) \hfill  (4)
-
-The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
-This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
-While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
-The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1.
-From equation 3 we find that hazard in bin t is equal to P(t)/S(t-1).
-
-For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
-
-\noindent ca(t) = P(correct \textbar{} RT = t) \hfill  (5)
-
-\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
-
-The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \(\geq\) t) tells you on how many percent of the trials the estimate h(t) = P(RT = t\textbar{} RT \(\geq\) t) is based.
-The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct \textbar{} RT = t) is based.
-
-While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y.
-If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
-Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).
-The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.
-
-\subsection{B. Custom functions for descriptive discrete-time hazard analysis}\label{b.-custom-functions-for-descriptive-discrete-time-hazard-analysis}
-
-We defined 12 custom functions that we list here.
-
-\begin{itemize}
-\tightlist
-\item
-  censor(df,timeout,bin\_width) : divide the time segment (0,timeout{]} in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
-\item
-  ptb(df) : transform the person-trial data set to the person-trial-bin data set
-\item
-  setup\_lt(ptb) : set up a life table for each level of 1 independent variable
-\item
-  setup\_lt\_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
-\item
-  calc\_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
-\item
-  calc\_ca\_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
-\item
-  join\_lt\_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
-\item
-  join\_lt\_ca\_2IV(df1, df2) : add the ca(t) estimates to the life tables (2 independent variables)
-\item
-  extract\_median(df) : estimate quantiles S(t)\textsubscript{.50} (1 independent variable)
-\item
-  extract\_median\_2IV(df) : estimate quantiles S(t)\textsubscript{.50} (2 independent variables)
-\item
-  plot\_eha(df, subj, haz\_yaxis=1, first\_bin\_shown=1, aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (1 independent variable), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
-\item
-  plot\_eha\_2IV(df, subj, haz\_yaxis=1, first\_bin\_shown=1, aggregated\_data=F, Nsubj=6) : create plots of the discrete-time functions (2 independent variables), and specify the upper limit of the y-axis in the hazard plot, with which bin to start plotting, whether the data is aggregated across participants, and across how many participants
-\end{itemize}
-
-When you want to analyse simple RT data from a detection experiment with one independent variable, the functions calc\_ca() and join\_lt\_ca() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot\_eha().
-When you want to analyse simple RT data from a detection experiment with two independent variables, the functions calc\_ca\_2IV() and join\_lt\_ca\_2IV() should not be used, and the code to plot the conditional accuracy functions should be removed from the function plot\_eha\_2IV().
-
-\subsection{C. Link functions}\label{c.-link-functions}
-
-Popular link functions include the logit link and the complementary log-log link, as shown in Figure 14.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/linkfunctions}
-
-}
-
-\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
-\end{figure}
-
-\subsection{D. Regression equations}\label{d.-regression-equations}
-
-An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
-
-\noindent cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\beta\)\textsubscript{0}ONE + \(\beta\)\textsubscript{1}(TIME-9) + \(\beta\)\textsubscript{2}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{3}X\textsubscript{1} + \(\beta\)\textsubscript{4}X\textsubscript{2} + \(\beta\)\textsubscript{5}X\textsubscript{2}(TIME-9){]} \hfill  (6)
-
-The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the parameters \(\beta\)\textsubscript{0} to \(\beta\)\textsubscript{2} multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
-The second set of terms (the beta parameters \(\beta\)\textsubscript{3} to \(\beta\)\textsubscript{5}) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
-Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{3} cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{4} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{4} + \(\beta\)\textsubscript{5} cloglog-hazard units (when TIME-9 = 1), and so forth.
-To interpret the effects of a predictor, its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
-When using the logit link, exponentiating a \(\beta\) parameter results in an odds ratio.
-
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
-
-\noindent cloglog{[}h(t){]} = {[}\(\beta\)\textsubscript{0}D1 + \(\beta\)\textsubscript{1}D2 + \(\beta\)\textsubscript{2}D3 + \(\beta\)\textsubscript{3}D4 + \(\beta\)\textsubscript{4}D5 + \(\beta\)\textsubscript{5}D6{]} + {[}\(\beta\)\textsubscript{6}X\textsubscript{1}{]} \hfill  (7)
-
-\subsection{E. Prior distributions}\label{e.-prior-distributions}
-
-To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023a) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Figure 15 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
-According to Kurz (2023a), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/plot_of_priors}
-
-}
-
-\caption{Prior distributions for the Intercept on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).}\label{fig:plot-priors}
-\end{figure}
-
-To gain a sense of what prior \emph{cloglog} values would approximate a uniform distribution on the hazard probability scale, we followed Kurz's approach and simulated a large number of draws from the Uniform(0,1) distribution, converted them to the cloglog metric, and fitted a skew-normal model (due to the asymmetry of the cloglog link function).
-Row E shows that using a skew-normal distribution with a mean of -0.59, a standard deviation of 1.26, and a skewness of -4.22 as a prior on the cloglog scale, approximates a uniform distribution on the probability scale.
-However, because hazard values below .5 are more likely in RT studies, using a skew-normal distribution with a mean of -1, a standard deviation of 1, and a skewness of -2 as a prior on the cloglog scale (row F), might be a good weakly informative prior for the intercept(s) in a cloglog-hazard model.
-
-\subsection{F. Advantages of hazard analysis}\label{f.-advantages-of-hazard-analysis}
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-First, as discussed by Holden, Van Orden, and Turvey (2009), ``probability density {[}and mass{]} functions can appear nearly identical, both statistically and to the naked eye, and yet are clearly different on the basis of their hazard functions (but not vice versa).
-Hazard functions are thus more diagnostic than density functions'' (p.~331) when one is interested in studying the detailed shape of a RT distribution (see also Figure 1 in Panis, Schmidt, et al., 2020).
-Therefore, when the goal is to study how psychological effects change over time, hazard and conditional accuracy functions are the preferred ways to describe the RT + accuracy data.
-
-Second, because RT distributions may differ from one another in multiple ways, Townsend (1990) developed a dominance hierarchy of statistical differences between two arbitrary distributions A and B.
-For example, if h\textsubscript{A}(t) \textgreater{} h\textsubscript{B}(t) for all t, then both hazard functions are said to show a complete ordering.
-Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
-For example, when mean A \textless{} mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t \textgreater{} 300 ms, or only for t \textless{} 500 ms), or they may cross each other one or more times.
-
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
-This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
-In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
-As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
-This orthodox procedure leads to underestimation of the true mean.
-By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
-
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer et al., 1988).
-
-Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, et al., 2020; Panis \& Schmidt, 2016).
-
 \end{document}

### 2024-12-13 notes/manuscript_notes.Rmd
update sims and ms
update 2b and ms.
update tut 1a and ms.
update ms. and new suppl-material folder

@@
 
-adjust small error in Figure 3 + check Figure 10 + add second panel to Figure 10
+adjust small error in Figure 3
+
+-> Done.
+
+check Figure 10 + add second panel to Figure 10
+
+-> Done.
 
@@
 
+-> Done.
+
 appendix/separate Rmd file... + {var} in functions...?
 
-clean up html files (delete old figures, Mention tut 2b priors line 101, ...)
+-> separate supplementary.Rmd created.
+
+clean up html files 2a (delete old figures,
+
+->
+
+Mention tut 2b priors line 101, ...)
+-> Done.
 
 change Figure comparison for ca
+->
 
 check for timebin -> time bin
+
+->
+
+Change grand average to population-level expectations... and participant-level expectations...
+
+->
 
@@
 e.g., page 28, Figure 15 when talking about priors.
+
+-> Done.
 
@@
 
+->
+
 ## conclusion ##
@@
 For now, I deleted it.
+
+-> I agree.

### 2024-12-13 manuscript/manuscript.Rmd
update ms-notes and ms.

@@
     email         : "sven.panis@hest.ethz.ch"
-    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
-      - "Conceptualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
+#    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
+#      - "Conceptualization"
+#      - "Writing - Original Draft Preparation"
+#      - "Writing - Review & Editing"
   - name          : "Richard Ramsey"
     affiliation   : "1"
-    role:
-      - "Conceptualization"
-      - "Writing - Review & Editing"
-      - "Supervision"
+#    role:
+#      - "Conceptualization"
+#      - "Writing - Review & Editing"
+#      - "Supervision"
 affiliation:

### 2024-12-13 Tutorial_2a.Rmd
update 2a and ms.

@@
 
-Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data.
+Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data (section 3).
 The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent).
-We perform some prior predictive checks to determine the prior distributions we will employ.
-We compare models using WAIC and LOO, and interpret parameter estimates for one model.
-We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale.
+We performed some prior predictive checks to determine the prior distributions we will employ (section 8).
+We compare models using WAIC and LOO (section 4), and interpret parameter estimates for one model (section 5).
+We also plot the logit and cloglog link functions (section 6), and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale (section 7).
 
@@
 The first model we fit is a "random intercepts" model, where we fit a single grand intercept for each timebin and add random intercepts that vary between participants.
-This constitutes a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial).
+This constitutes a general specification of TIME, and can be used if we do not want to make assumptions about how (cloglog-)hazard changes over time (within a trial).
 
@@
 Prepare the data file, and specify priors.
-The skew_normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that hazard values above .5 are less likely than values below .5.
+The skew-normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that higher hazard values (above .5) are less likely than lower values (below .5).
+This prior is plotted in row F of Supplementary Figure 2 (section E in the Supplemental Material).
 A prior predictive check is performed in section 8.1.
@@
 
-Prepare the data file, and specify priors. To find the priors, we performed a prior predictive check which is reported in section 8.2.
+Prepare the data file, and specify priors. To find the class b priors, we performed a prior predictive check which is reported in section 8.2.
 
@@
   set_prior("normal(0,.06)", class= "b", coef = "period_9_sq"),  
+  # A N(0,.4) prior is set for the effects of congruent and incongruent primes
   set_prior("normal(0,.4)", class = "b", coef = "conditioncongruent"),
@@
 Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition.
-Model_M1d took about 184 minutes to run.
-Note that duplicate terms in the model formula (e.g., condition) are ignored.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model_M1d took about 184 minutes to run.
 
@@
 
-First, we can categorize the predictor trial number by grouping trials in one of three "stages" of the experiment (stage 1 = trials 1 to 500; stage 2 = trials 501 to 1000; stage 3 = trials 1001 and later) and use index coding (variable "stage" in ptb_data).
+First, we can categorize the predictor trial number by grouping trials in one of a number of "stages" in the experiment (e.g., stage 1 = trials 1 to 500; stage 2 = trials 501 to 1000; stage 3 = trials 1001 and later) and use index coding (variable "stage" in ptb_data).
 
@@
 
-Compare all three models.
+Compare all five models.
 
@@
 
-#
-```{r plot-post-M1i-alt}
-tidy_fixed <- post %>% 
-  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
-  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
-  pivot_longer(-c(chain, draw, iter)) %>% 
-  mutate(timebin = str_sub(name,10,11),
-         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
-         condition = str_sub(name,17,18),
-         condition = factor(str_remove(condition,"e"),
-                            levels=c(1,2,3),
-                            labels=c("blank","congruent","incongruent")))
-head(tidy_fixed)
-tail(tidy_fixed)
-
-# plot
-p_tidy_fixed <- ggplot(tidy_fixed, aes(x = timebin, y = value, fill=condition)) +  
-  stat_halfeye(point_interval="median_qi", 
-               .width = c(0.80,0.95),
-               alpha=0.6,
-               position=pd) +
-  labs(title = 'Posterior distributions for population-level\neffects in Model M1i',
-       x = "time bin", y = "cloglog-hazard") +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_color_brewer(palette = "Dark2") +
-
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-   theme(axis.text.x = element_text(angle=90)) 
-#+
-#  facet_wrap(~condition)
-p_tidy_fixed
-```
-
-```{r save-post, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_postdistr_alt3.png", width = 10, height = 8, dpi = 800)
-```
-
-#
+Save the posterior distribution plot.
 
@@
 
-Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, and for each participant in the data set.
-
-First, for the average participant, using add_epred_draws().
+Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the "average" participant at the population level, and for each participant in the data set.
+
+First, at the population level, using add_epred_draws().
 
@@
   add_epred_draws(model_M1i, 
-                  re_formula = NA) %>% 
+                  re_formula = NA) %>% # ignore random effects
   mutate(prime = factor(prime, 
@@
 
-Again, for the average participant, but now using posterior draws.
+Again, for the "average" participant at the population level, but now using posterior draws.
 
@@
     scale_y_continuous(limits=c(0,.6)) +
-    labs(y = "predicted hazard") +
-    ggtitle("Average participant")
+    labs(y = "predicted hazard") 
 ```
@@
   add_epred_draws(model_M1i, 
-                  re_formula = NULL) %>% 
+                  re_formula = NULL) %>% # include random effects
   mutate(prime = factor(prime, 
@@
 
-Save plot.
+Save the combined plot.
 
@@
 p11
-```
-
-```{r save-grand-AMEs, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_grand_AME.png", width = 10, height = 8, dpi = 800)
 ```
@@
 
-```{r save-pid-AMEs, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pid_AMEs.png", width = 10, height = 8, dpi = 800)
-```
-
 Combine the plots.
@@
 
-### AMEs for a brand new hypothetical participant
-
-```{r epred-new-pid-diff}
-epreds_pid_new_diffs <- epreds_pid_new %>%
-  pivot_wider(id_cols = c(pid,timebin,.draw),
-              names_from = "prime",
-              values_from = ".epred") %>% 
-  mutate(`congruent minus blank` = congruent - blank,
-         `incongruent minus blank` = incongruent -  blank) %>% 
-  select(-c(blank,congruent,incongruent)) %>% 
-  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
-               names_to = "contrast",
-               values_to = "diff") 
-epreds_pid_new_diffs 
-```
-
-Summarize and plot the contrasts in predicted hazard values.
-
-```{r epred-new-pid-diff-summary}
-epreds_pid_new_diffs %>% 
-  group_by(pid,timebin,contrast) %>% 
-  median_qi(.width = .95) %>%
-  arrange(contrast,timebin,.width) %>%
-  select(-c(.width,.point,.interval))
-```
-
-```{r epred-new-pid-diff-plot}
- ggplot(epreds_pid_new_diffs, aes(x=timebin, y=diff, 
-                       fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, 
-                    point_interval = "mean_qi",
-                    .width=c(.95)) +
-    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
-    geom_hline(yintercept = 0, color = "red", lty = 3) +
-    scale_fill_brewer(palette = "Dark2") +
-    scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-    theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(-1,1)) +
-    labs(y = "difference in predicted hazard") +
-    ggtitle("AMEs for a brand new\nhypothetical participant") +
-    facet_wrap(~contrast)
-```
-
-```{r save-pid-new-AMEs, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_pid_new_AMEs.png", width = 10, height = 8, dpi = 800)
-```
-
 ## Example conclusions for model M1i.
@@
 
-If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 8 and Table 4.
+If we want to study the population-level effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 7 (95% CrI = [0.04, 0.25]).
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 9.
+If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 7B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
@@
 ## 8.1. Model M0i
-
-### Intermezzo: Prior Predictive Check for M0i.
 
@@
 
-After our prior predictive check, we can fit model M0i to get the posterior distributions.
+After our prior predictive check, we can fit model M0i to get the posterior distributions (see section 3).
 
@@
 
-### Intermezzo. Prior predictive check for model M0c.
-
 Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs. 
-
 To get an idea of how the intercept and both slopes control the shape of the hazard function, the following code allows to change the coefficients and generate (cloglog-)hazard functions.
@@
 
-After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions.
+After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions (see section 3).

### 2024-12-13 Tutorial_3a.Rmd
update comparison figures, and ms.

--- 2024-11-26 Tutorial_3a.Rmd
@@
 est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
-              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(3,7,8)],
+              glmer = tidy(model_M1i_f, 
+                           effects = c("fixed"), 
+                           conf.int=T, 
+                           conf.level = 0.95)[,c(3,7,8)],
               effectname = names(fixef(model_M1i_f)),
-              position = 1:30) 
-est
+              position = rep(6:15,3),
+              prime = factor(rep(1:3, each=10),
+                             labels=c("blank","congruent","incongruent")) )
+est %>% print(n=30)
 ```
+
+Plot.
 
@@
                  shape = "s1"),
-                 size = 1.5) +
+                 size = 2.4) +
  geom_point(aes( y = glmer$estimate, 
@@
                  shape = "s2"),
-                 size = 1.5, 
-                 position = position_nudge(x = .2))  +
+                 size = 2.4, 
+                 position = position_nudge(x = .4))  +
  geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                     color = "c1"), 
-                show.legend = F) +
+                show.legend = F,
+                linewidth = 1.2) +
  geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
@@
                 show.legend = F, 
-                position = position_nudge(x = .2)) +
- geom_hline(yintercept=0,colour = "grey", linetype = 2) +
- scale_x_continuous(breaks = 1:30,labels = est$effectname) +
+                linewidth = 1.2,
+                position = position_nudge(x = .4)) +
+# geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
  scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +
 
- labs( x = "Effects in model M1i", y = "cloglog-hazard" ) +
+ labs(x = "bin endpoint", y = "cloglog-hazard") +
  scale_color_manual(name = "Model", 
@@
                     labels = c("brm", "glmer")) +
-  coord_flip() 
+  theme(panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle=90)) +
+  facet_wrap(~prime)
 p1
@@
 ```{r save-plot-comparison, eval=F}
-ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=8,dpi=800)
+ggsave("Tutorial_3_Frequentist/Comparison.png", height=4, width=10,dpi=800)
 ```

### 2024-12-13 Tutorial_3b.Rmd
update comparison figures, and ms.

@@
               effectname = names(fixef(model_M1i_ca_f)),
-              position = 1:30)
+              position = rep(6:15,3),
+              prime = factor(rep(1:3, each=10),
+                             labels=c("blank","congruent","incongruent")) )
+
+est %>% print(n=30)      
 ```
@@
                  shape = "s2"),
-                 size = 1.5, 
-                 position = position_nudge(x = .2))  +
+                 size = 2.4, 
+                 position = position_nudge(x = .4))  +
  geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                     color = "c1"), 
+                linewidth = 1.2,
                 show.legend = F) +
@@
                 show.legend = F,
-                position = position_nudge(x = .2)) +
- geom_hline(yintercept=0,colour = "grey", linetype = 2) +
- scale_x_continuous(breaks = 1:30,labels = est$effectname) +
+                linewidth = 1.2,
+                position = position_nudge(x = .4)) +
+ #geom_hline(yintercept=0,colour = "grey", linetype = 2) +
+ scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
  scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +
 
- labs( x = "Effects in model M1i_ca", y = "logit-ca" ) +
+ labs(x = "bin endpoint", y = "logit-ca") +
  scale_color_manual(name = "Model", 
@@
                     labels = c("brm", "glmer")) +
-  coord_flip() 
+ theme(panel.grid.minor = element_blank(),
+       axis.text.x = element_text(angle=90)) +
+facet_wrap(~prime)
 p1
@@
 ```{r save-plot-comparison, eval=F}
-ggsave("Tutorial_3_Frequentist/Comparison_ca.png", height=8, width=8,dpi=800)
+ggsave("Tutorial_3_Frequentist/Comparison_ca.png", height=4, width=10,dpi=800)
 ```

### 2024-12-13 Tutorial_1a.Rmd
update 1a html, lock file, ms-notes

@@
 It uses functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
-To set up a life table, the user has to select a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (i.e., a cutoff time, or timeout).
+To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (i.e., a cutoff time, or timeout).
 The goal is to find the optimal bin width and censoring time (both in ms) to portray the effects of interest.
@@
 
-Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 134-136 in custom_functions.R) as above (lines 96-99).
+Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 119-122).
 

### 2024-12-13 Tutorial_2a.Rmd
update 2a and 2b html

@@
 
-The shape of the prior predictive distribution for 1 bin looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.
+The shape of the prior predictive distribution for 1 bin looks as expected (compare with Supplementary Figure 2 (row F) in the Supplemental Material): hazard values below .5 are more likely than hazard values above .5.
 

### 2024-12-13 Tutorial_4.Rmd
update 3a 3b 4.Rmd

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, eval=F)
 ```
@@
 p4.1
-
+```
+
+```{r eval=F}
 ggsave ("Tutorial_4_planning/figures/index_cloglog.jpeg",
@@
 p4.2
-
+```
+
+```{r eval=F}
 ggsave ("Tutorial_4_planning/figures/index_hazard.jpeg",
@@
 p5.4
-
+```
+
+```{r eval=F}
 ggsave ("Tutorial_4_planning/figures/raw_vs_d1.jpeg",
@@
 
-```{r load-sim1}
+```{r load-sim1, eval=F}
 sx1 <- read_csv("Tutorial_4_planning/data/sim1/sim1_data.csv") %>%
@@
 
-```{r exp-summary-sim1}
+```{r exp-summary-sim1, eval=F}
 sx1_exp <- sx1 %>% 
@@
 
-```{r sim-summary-sim1}
+```{r sim-summary-sim1, eval=F}
 sx1_sim <- sx1_exp %>% 
@@
 
-```{r sim1-violin}
+```{r sim1-violin, eval=F}
 p9.1 <- ggplot(sx1_sim, 

### 2024-12-14 manuscript/manuscript.Rmd
update ms. and ms-notes

@@
 In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
-We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
+We perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
 In Tutorial 2b we fit Bayesian multilevel regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
@@
 Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in part A of the Supplemental Material.
+The equations that define both of these functions are reported in section A of the Supplemental Material.
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the grand average (Figure 6A), and for each participant in the data set (Figure 6B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 6A), and for each participant in the data set (Figure 6B).
 
-(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).
+(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the grand average (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A; grand average marginal effect) and at the participant level (Figure 7B; subject-specific average marginal effect).
 
-(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin for the grand average (A), and for each participant (B).
+(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).
 
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the grand average (Figure 9A), and for each participant in the data set (Figure 9B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 9A), and for each participant in the data set (Figure 9B).
 
-(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).
+(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
 
-(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average (A), and for each participant (B).
+(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).
 
@@
 
-Based on Figure 10A we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Based on Figure 10A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.

### 2024-12-14 manuscript/manuscript.tex
update ms-notes and ms.
update ms. and ms-notes

@@
 
-The authors made the following contributions.
-Sven Panis: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Richard Ramsey: Conceptualization, Writing - Review \& Editing, Supervision.
-
 Correspondence concerning this article should be addressed to Sven Panis, ETH GLC, room G16.2, Gloriastrasse 37/39, 8006 Zürich.
@@
 In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
-We also perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
+We perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
 In Tutorial 2b we fit Bayesian multilevel regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- for the grand average (Figure 6A), and for each participant in the data set (Figure 6B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 6A), and for each participant in the data set (Figure 6B).
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
+\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
 \end{figure}
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both for the grand average (Figure 7A; grand average marginal effect) and for each participant in the data set (Figure 7B; subject-specific average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A; grand average marginal effect) and at the participant level (Figure 7B; subject-specific average marginal effect).
 
@@
 
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
 \end{figure}
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- for the grand average (Figure 9A), and for each participant in the data set (Figure 9B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 9A), and for each participant in the data set (Figure 9B).
 
@@
 
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
+\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
 \end{figure}
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects for the grand average (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
 
@@
 
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin for the grand average (A), and for each participant (B).}\label{fig:plot-ca-grand-ame-effects}
+\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-ca-grand-ame-effects}
 \end{figure}
 
-Based on Figure 10A we see that congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Based on Figure 10A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.

### 2024-12-14 manuscript/manuscript.Rmd
update ms. and ms-notes

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "13574"
+wordcount         : "11539 (body) + 1528 (references)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 The first second after stimulus onset is divided in ten bins of 100 ms.
-Timebin indicates the bin rank.
+Time bin indicates the bin rank.
 The first bin is (0,100], the last bin is (900,1000].
-For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
+For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
 Insets show mean response time per condition.
@@
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in timebin 2) and is gone when the waiting time reaches 500 ms or more.
+For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in time bin 2) and is gone when the waiting time reaches 500 ms or more.
 In contrast, in example 5, the effect starts around 300 ms and is gone by 700 ms.
@@
 Since our focus here is on hazard models that use discrete time units, we describe that approach.
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 timebins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in timebin 6 (any RTs from 501 ms to 600 ms).
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
 
@@
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
 
@@
 These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
-During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+During time bin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
 For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
@@
 Performing EHA in experimental psychology has implications for how experiments are designed.
-Indeed, if trials are categorised as a function of when responses occur, then each timebin will only include a subset of the total number of trials.
+Indeed, if trials are categorised as a function of when responses occur, then each time bin will only include a subset of the total number of trials.
 For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
@@
 We again use the data from @panisWhatShapingRT2016 to provide a worked example.
-We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
+We fit an index coding model on a subset of time bins (six time bins in total) and for two prime conditions (congruent and incongruent).
 We chose to focus on a subsample of the data to ease the computational burden.

### 2024-12-14 manuscript/manuscript.tex
update ms. and ms-notes

@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 13574}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11539 (body) + 1528 (references)}
 \usepackage{lineno}
@@
 The first second after stimulus onset is divided in ten bins of 100 ms.
-Timebin indicates the bin rank.
+Time bin indicates the bin rank.
 The first bin is (0,100{]}, the last bin is (900,1000{]}.
-For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that timebin, given that it has not yet occurred.
+For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
 Insets show mean response time per condition.}\label{fig:plot1}
@@
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
 
@@
 These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
+During time bin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
 For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
@@
 We again use the data from Panis and Schmidt (2016) to provide a worked example.
-We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent).
+We fit an index coding model on a subset of time bins (six time bins in total) and for two prime conditions (congruent and incongruent).
 We chose to focus on a subsample of the data to ease the computational burden.

### 2024-12-15 manuscript/manuscript.Rmd
update ms. and ms-notes

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "11539 (body) + 1528 (references)"
+wordcount         : "11643 (body) + 1528 (references)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 
+# Author contributions
+
+Conceptualization: S. Panis and R. Ramsey; Software: S. Panis and R. Ramsey; Writing - Original Draft Preparation: S. Panis; Writing - Review & Editing: S. Panis and R. Ramsey; Supervision: R. Ramsey.
+
+# Conflicts of Interest
+
+The author(s) declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+# Prior versions
+
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive:
+
+# Supplemental Material
+
+# Disclosures
+
+## Data, materials, and online resources
+
+Link to public archive: https://github.com/sven-panis/Tutorial_Event_History_Analysis
+
+Supplemental Material:
+
+## Ethical approval
+
+Ethical approval was not required for this tutorial in which we reanalyze existing data sets.
+
+\newpage
+
 # References

### 2024-12-15 manuscript/manuscript.tex
update ms. and ms-notes

@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11539 (body) + 1528 (references)}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11643 (body) + 1528 (references)}
 \usepackage{lineno}
@@
 
+\section{Author contributions}\label{author-contributions}
+
+Conceptualization: S.
+Panis and R.
+Ramsey; Software: S.
+Panis and R.
+Ramsey; Writing - Original Draft Preparation: S.
+Panis; Writing - Review \& Editing: S.
+Panis and R.
+Ramsey; Supervision: R.
+Ramsey.
+
+\section{Conflicts of Interest}\label{conflicts-of-interest}
+
+The author(s) declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+\section{Prior versions}\label{prior-versions}
+
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive:
+
+\section{Supplemental Material}\label{supplemental-material}
+
+\section{Disclosures}\label{disclosures}
+
+\subsection{Data, materials, and online resources}\label{data-materials-and-online-resources}
+
+Link to public archive: \url{https://github.com/sven-panis/Tutorial_Event_History_Analysis}
+
+Supplemental Material:
+
+\subsection{Ethical approval}\label{ethical-approval}
+
+Ethical approval was not required for this tutorial in which we reanalyze existing data sets.
+
+\newpage
+
 \section{References}\label{references}

### 2024-12-15 sims/sims.Rmd
update figures sims

@@
   scale_colour_brewer(palette = "Dark2") +
-  labs(x="timebin",
+  labs(x="time bin",
        y="probability")
@@
 
-ggsave("sims/figures/rt_haz.jpeg",
-       width = 12, height = 8, dpi=800)
+#ggsave("sims/figures/rt_haz.jpeg",
+#      width = 12, height = 8, dpi=800)
 ```
@@
 
-ggsave("sims/figures/haz_surv.jpeg",
-       width = 7, height = 5, dpi=800)
+#ggsave("sims/figures/haz_surv.jpeg",
+#       width = 7, height = 5, dpi=800)
 ```
@@
   ggtitle("hazard and survivor functions") +
-  xlab("timebin")
+  xlab("time bin")
 p11.1
@@
   ggtitle("hazard")  +
-  xlab("timebin") + ylab("probability") + guides(color=guide_legend(title="prime type"))
+  xlab("time bin") + ylab("probability") + guides(color=guide_legend(title="prime type"))
 p12.1
@@
   scale_y_continuous(limits = c(0,1)) +
-  xlab("timebin") +
+  xlab("time bin") +
   ylab("probability") +

### 2024-12-15 Tutorial_2a.Rmd
change x-axis labels 2a and 2b

@@
     scale_y_continuous(limits=c(0,.7)) +
-    labs(y = "predicted hazard")
+    labs(x = "time bin", y = "predicted hazard")
 
@@
     scale_y_continuous(limits=c(0,.6)) +
-    labs(y = "predicted hazard") 
+    labs(x = "time bin", y = "predicted hazard") 
 ```
@@
     scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted hazard") +
+    labs(x = "time bin", y = "predicted hazard") +
    # ggtitle("Each participant (N = 6)") +
@@
   theme(axis.text.x = element_text(angle=90)) +
-  labs(y = "difference in predicted hazard") +
+  labs(x = "time bin", y = "difference in predicted hazard") +
   #ggtitle("Grand average marginal effect") +
@@
     scale_y_continuous(limits=c(-0.6,0.6)) +
-    labs(y = "difference in predicted hazard") +
+    labs(x = "time bin", y = "difference in predicted hazard") +
     #ggtitle("Subject-specific AMEs") +

### 2024-12-15 Tutorial_2b.Rmd
change x-axis labels 2a and 2b

@@
     scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted conditional accuracy") #+
+    labs(x = "time bin", y = "predicted conditional accuracy") #+
     #ggtitle("Average participant")
@@
     scale_y_continuous(limits=c(0,1)) +
-    labs(y = "predicted conditional accuracy") +
+    labs(x = "time bin", y = "predicted conditional accuracy") +
    # ggtitle("Each participant (N = 6)") +
@@
   theme(axis.text.x = element_text(angle=90)) +
-  labs(y = "difference in predicted ca") +
+  labs(x = "time bin", y = "difference in predicted ca") +
 #  ggtitle("Grand average marginal effect") +
@@
     scale_y_continuous(limits=c(-1,1)) +
-    labs(y = "difference in predicted ca") +
+    labs(x = "time bin", y = "difference in predicted ca") +
    # ggtitle("Subject-specific AMEs") +

### 2024-12-15 manuscript/manuscript.Rmd
update abstract

@@
 abstract: |
-  Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
-However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+  Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widespread impact on our understanding of human cognition.
+However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
 The ability to reveal finer-grained, "temporal states" of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states.
+Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterize the time course of psychological states.
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.

### 2024-12-15 manuscript/manuscript.tex
update abstract

@@
 \abstract{%
-Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition.
-However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widespread impact on our understanding of human cognition.
+However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
 The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
-Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states.
+Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterize the time course of psychological states.
 One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches.

### 2024-12-15 manuscript/manuscript.Rmd
update ms pdf

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "11643 (body) + 1528 (references)"
+wordcount         : "11643 (body) + 1528 (references) + 2394 (supplemental material)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A; grand average marginal effect), and for each participant (Figure 10B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A) and for each participant (Figure 10B).
 

### 2024-12-15 manuscript/manuscript.tex
update ms pdf

@@
 {}{\PatchFailed}
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11643 (body) + 1528 (references)}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11643 (body) + 1528 (references) + 2394 (supplemental material)}
 \usepackage{lineno}

### 2024-12-16 Tutorial_4.Rmd
updated tutorial 4. added sim data loading from the OSF.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, eval=F)
+knitr::opts_chunk$set(echo = TRUE, eval=T)
 ```
@@
          "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux",
-         "rstan")
+         "rstan", "osfr")
 
@@
 For Lisa's faux package, see here: [https://debruine.github.io/faux/](https://debruine.github.io/faux/).
-The vignettes provided by the faux package are particularly helpful and worth following before you take a deep dive in the below code. 
+The vignettes provided by the faux package are particularly helpful and worth following before you take a deep dive into the below code. 
 There is also an associated [journal article](https://doi.org/10.1177/2515245920965119), which is worth reading.
@@
 For simplicity, we use these normal(0,2) priors for the mean per condition.
-However, if we were doing this for real, we would probably choose different priors that are similar to ones that were discussed in Tutorial_2a as well as in Supplmentary Materials.
+However, if we were doing this for real, we would probably choose different priors that are similar to ones that were discussed in Tutorial_2a, as well as in Supplementary Materials.
 
@@
 
-read in the file, if it is already computed
-
-```{r load-sim1, eval=F}
+read in the file, if it is already computed.
+
+If necessary, read it from the OSF to start with (because it is a large file). 
+This step is only required to be done once.
+After that, the file will be in your local directory and you can skip this chunk in the future.
+To do so, set eval=TRUE in the chunk below, then re-set to FALSE once the data have been downloaded.
+
+```{r load-osf-sim1, eval=F}
+## create a relevant folder
+dir.create("Tutorial_4_planning/data/sim1/")
+
+## read in the data from the OSF and store it in a relevant folder
+osf_sim1_data <-
+  # retrieve the OSF (main) project
+  osf_retrieve_node("3dbcs") %>% ## this is just one project picked at random
+  # list files in this repository - there must be a more efficient way than this...
+  osf_ls_files(pattern = "Tutorial_4_planning") %>%
+  osf_ls_files(pattern = "data") %>%
+  osf_ls_files(pattern = "sim1") %>%
+  osf_ls_files(pattern = "sim1_data") %>%
+  osf_download(path = "./Tutorial_4_planning/data/sim1/")
+```
+
+read the file from a local folder
+
+```{r load-sim1, eval=T}
 sx1 <- read_csv("Tutorial_4_planning/data/sim1/sim1_data.csv") %>%
@@
 
-```{r exp-summary-sim1, eval=F}
+```{r exp-summary-sim1, eval=T}
 sx1_exp <- sx1 %>% 
@@
 
-```{r sim-summary-sim1, eval=F}
+```{r sim-summary-sim1, eval=T}
 sx1_sim <- sx1_exp %>% 
@@
 
-```{r sim1-violin, eval=F}
+```{r sim1-violin, eval=T}
 p9.1 <- ggplot(sx1_sim, 
@@
 
-read in the file, if it is already computed
-
-```{r load-sim2}
+read in the file, if it is already computed.
+
+If necessary, read it from the OSF to start with (because it is a large file). 
+This step is only required to be done once.
+After that, the file will be in your local directory and you can skip this chunk in the future.
+To do so, set eval=TRUE in the chunk below, then re-set to FALSE once the data have been downloaded.
+
+```{r load-osf-sim2, eval=F}
+## create a relevant folder
+dir.create("Tutorial_4_planning/data/sim2/")
+
+## read in the data from the OSF and store it in a relevant folder
+osf_sim2_data <-
+  # retrieve the OSF (main) project
+  osf_retrieve_node("3dbcs") %>% ## this is just one project picked at random
+  # list files in this repository - there must be a more efficient way than this...
+  osf_ls_files(pattern = "Tutorial_4_planning") %>%
+  osf_ls_files(pattern = "data") %>%
+  osf_ls_files(pattern = "sim2") %>%
+  osf_ls_files(pattern = "sim2_data") %>%
+  osf_download(path = "./Tutorial_4_planning/data/sim2/")
+```
+
+read the file from a local folder
+
+```{r load-sim2, eval=T}
 sx2 <- read_csv("Tutorial_4_planning/data/sim2/sim2_data.csv") %>%
@@
 
-```{r}
+```{r sim1-sim2-power-plot}
 p12.1 <- (p10.4 | p11.6) +

### 2024-12-16 Tutorial_4.Rmd
removed unnecessary text

@@
   # retrieve the OSF (main) project
-  osf_retrieve_node("3dbcs") %>% ## this is just one project picked at random
+  osf_retrieve_node("3dbcs") %>% 
   # list files in this repository - there must be a more efficient way than this...

### 2024-12-16 Tutorial_4.Rmd
updated the way osf data is read in by checking if it exists first

@@
          "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux",
-         "rstan", "osfr")
+         "rstan", "osfr", "here")
 
@@
 After that, the file will be in your local directory and you can skip this chunk in the future.
-To do so, set eval=TRUE in the chunk below, then re-set to FALSE once the data have been downloaded.
-
-```{r load-osf-sim1, eval=F}
-## create a relevant folder
-dir.create("Tutorial_4_planning/data/sim1/")
-
-## read in the data from the OSF and store it in a relevant folder
-osf_sim1_data <-
-  # retrieve the OSF (main) project
+
+```{r load-osf-sim1, eval=T}
+## create a relevant folder, if it does not already exist
+sim1_output_dir <- here("Tutorial_4_planning/data/sim1/")
+
+if (!dir.exists(here(sim1_output_dir))) {
+  dir.create(here(sim1_output_dir))
+}
+
+## read in the data from the OSF and store it in a relevant folder, if it does not already exist
+sim1_file_name <- here("Tutorial_4_planning/data/sim1/sim1_data.csv")
+
+if (!file.exists(here(sim1_file_name))) {
   osf_retrieve_node("3dbcs") %>% 
-  # list files in this repository - there must be a more efficient way than this...
   osf_ls_files(pattern = "Tutorial_4_planning") %>%
@@
   osf_ls_files(pattern = "sim1_data") %>%
-  osf_download(path = "./Tutorial_4_planning/data/sim1/")
+  osf_download(path = sim1_output_dir)
+}
 ```
@@
 
-```{r load-osf-sim2, eval=F}
-## create a relevant folder
-dir.create("Tutorial_4_planning/data/sim2/")
-
-## read in the data from the OSF and store it in a relevant folder
-osf_sim2_data <-
-  # retrieve the OSF (main) project
+```{r load-osf-sim2, eval=T}
+## create a relevant folder, if it does not already exist
+sim2_output_dir <- here("Tutorial_4_planning/data/sim2/")
+
+if (!dir.exists(here(sim2_output_dir))) {
+  dir.create(here(sim2_output_dir))
+}
+
+## read in the data from the OSF and store it in a relevant folder, if it does not already exist
+sim2_file_name <- here("Tutorial_4_planning/data/sim2/sim2_data.csv")
+
+if (!file.exists(here(sim2_file_name))) {
   osf_retrieve_node("3dbcs") %>% 
-  # list files in this repository - there must be a more efficient way than this...
   osf_ls_files(pattern = "Tutorial_4_planning") %>%
@@
   osf_ls_files(pattern = "sim2_data") %>%
-  osf_download(path = "./Tutorial_4_planning/data/sim2/")
+  osf_download(path = sim2_output_dir)
+}
 ```

### 2024-12-16 Tutorial_1b.Rmd
update Rmd & html files

@@
 ---
-title: "Tutorial_1b"
+title: "Tutorial 1b"
 author: "Sven Panis"

### 2024-12-16 Tutorial_2a.Rmd
update Rmd & html files

@@
 ---
-title: "Tutorial_2a"
+title: "Tutorial 2a"
 author: "Sven Panis"
@@
 
-For example, the coefficients for x vary from -0.13 to 0.6, and we might specify N(.3,.25) as a prior distribution, Those for x-squared vary from -0.11 to 0.12, and we might specify N(0,.06) as a prior distribution.
+For example, the coefficients for x vary from -0.13 to 0.6, and we might specify N(.3,.25) as a prior distribution.
+Those for x-squared vary from -0.11 to 0.12, and we might specify N(0,.06) as a prior distribution.
 
@@
 
-Finally, perform a ppc for the selected prior distributions for M0c (see lines 369-374), and increase the standard deviation for the normal prior for class "sd" somewhat.
+Finally, perform a ppc for the selected prior distributions for M0c (see lines 220-224), and increase the standard deviation for the normal prior for class "sd" somewhat.
 

### 2024-12-16 Tutorial_2b.Rmd
update Rmd & html files

@@
 ---
-title: "Tutorial_2b"
+title: "Tutorial 2b"
 author: "sven panis"
@@
 
-As we expect conditional accuracies ranging between (and including) 0 and 1, we use a flat prior on the logit scale for our intercepts (see section E in the supplemental material).
+As we expect conditional accuracies ranging between (and including) 0 and 1, we use a flat prior on the logit scale for our intercepts (see Supplementary Figure 2 in section E of the Supplemental Material).
 
@@
 
-```{r save-epred-grand, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_grand.png", width = 10, height = 8, dpi = 800)
-```
-
 ### Subject-specific
@@
 p2
-```
-
-```{r save-epred-pid, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png", width = 10, height = 8, dpi = 800)
 ```
@@
 
-```{r save-grand-AMEs, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png", width = 10, height = 8, dpi = 800)
-```
-
 ### Subject-specific AMEs
@@
 
-```{r save-pid-AMEs, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pid_AMEs.png", width = 10, height = 8, dpi = 800)
-```
-
 Combine the plots.

### 2024-12-16 Tutorial_3a.Rmd
update 3a 3b 4.Rmd
update Rmd & html files

@@
 ---
-title: "Tutorial_3a"
+title: "Tutorial 3a"
 author: "Sven Panis"
@@
 ```{r load-pkg, results='hide'}
-pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")
+pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed", "rstan")
 
@@
 load("Tutorial_3_Frequentist/model_M1i_f.RData")
-summary(model_M1i_f)
+fixef(model_M1i_f)
 ```

### 2024-12-16 Tutorial_3b.Rmd
update 3a 3b 4.Rmd
update Rmd & html files

@@
 ---
-title: "Tutorial_3b"
+title: "Tutorial 3b"
 author: "sven panis"
@@
 load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
-summary(model_M1i_ca_f)
+fixef(model_M1i_ca_f)
 ```

### 2024-12-16 manuscript/manuscript.Rmd
update readme, ms., and refs

@@
 pkg2 <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
-         "brms", "tidybayes", "bayesplot", "future", "parallel")
+         "brms", "tidybayes", "bayesplot", "future", "parallel", "rstan")
 
@@
 The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it [@gelmanBayesianWorkflow2020].
-The basic structure and code follows the examples outlined by Solomon Kurz in his 'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as the related paper [@debruineUnderstandingMixedEffectsModels2021].
+The basic structure and code follows the examples outlined by Solomon Kurz in his 'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as these related papers [@debruineUnderstandingMixedEffectsModels2021; @pargentTutorialTailoredSimulationBased2024].
 

### 2024-12-16 manuscript/manuscript.tex
update readme, ms., and refs

@@
   {\xapptocmd\section
-    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
     {}{\InnerPatchFailed}%
@@
 {}{\PatchFailed}
+\makeatother
 \keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11643 (body) + 1528 (references) + 2394 (supplemental material)}
@@
 Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.
+Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024a), which makes it computationally reproducible, in terms of the underlying data and figures.
 
@@
 
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024b), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
@@
 The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
-The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).
+The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as these related papers (DeBruine \& Barr, 2021; Pargent, Koch, Kleine, Lermer, \& Gaube, 2024).
 
@@
 
+\bibitem[\citeproctext]{ref-pap2024}
+Aust, F., \& Barth, M. (2024a). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
+
 \bibitem[\citeproctext]{ref-R-papaja}
-Aust, F., \& Barth, M. (2023). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\bibitem[\citeproctext]{ref-pap2024}
-Aust, F., \& Barth, M. (2024). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
+Aust, F., \& Barth, M. (2024b). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
 
@@
 \bibitem[\citeproctext]{ref-R-Matrix}
-Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://CRAN.R-project.org/package=Matrix}
+Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://Matrix.R-forge.R-project.org}
 
@@
 \bibitem[\citeproctext]{ref-R-tidybayes}
-Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
+Kay, M. (2024). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
 
@@
 
+\bibitem[\citeproctext]{ref-pargentTutorialTailoredSimulationBased2024}
+Pargent, F., Koch, T.
+K., Kleine, A.-K., Lermer, E., \& Gaube, S.
+(2024).
+A {Tutorial} on {Tailored Simulation-Based Sample-Size Planning} for {Experimental Designs With Generalized Linear Mixed Models}.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{7}(4), 25152459241287132.
+\url{https://doi.org/10.1177/25152459241287132}
+
 \bibitem[\citeproctext]{ref-R-patchwork}
-Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://CRAN.R-project.org/package=patchwork}
+Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://patchwork.data-imaginist.com}
 
@@
 
+\bibitem[\citeproctext]{ref-R-StanHeaders}
+Stan Development Team. (2020). \emph{{StanHeaders}: Headers for the {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
+
+\bibitem[\citeproctext]{ref-R-rstan}
+Stan Development Team. (2024). \emph{{RStan}: The {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
+
 \bibitem[\citeproctext]{ref-steeleGeneralMultilevelMultistate2004}

### 2024-12-17 Tutorial_1a.Rmd
update tuts 1a and 3a, ms.

@@
 
-This Tutorial_1a file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to calculate descriptive statistics.
-It uses functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
+In Tutorial 1a we wrangle raw data from Experiment 1 of Panis and Schmidt (2016), and illustrate how to calculate descriptive statistics.
+We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
 To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (i.e., a cutoff time, or timeout).

### 2024-12-17 Tutorial_3a.Rmd
update tuts 1a and 3a, ms.

@@
 
-In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).
+In Tutorial 3a we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).
 

### 2024-12-17 manuscript/manuscript.Rmd
update tuts 1a and 3a, ms.

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "11643 (body) + 1528 (references) + 2394 (supplemental material)"
+wordcount         : "11664 (body) + 1593 (references) + 2394 (supplemental material)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 
-Supplemental Material:
+Supplemental Material: Panis_Ramsey_suppl_material.pdf
 

### 2024-12-17 manuscript/manuscript.tex
update tuts 1a and 3a, ms.

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11643 (body) + 1528 (references) + 2394 (supplemental material)}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11664 (body) + 1593 (references) + 2394 (supplemental material)}
 \usepackage{lineno}
@@
 
-Supplemental Material:
+Supplemental Material: Panis\_Ramsey\_suppl\_material.pdf
 

### 2025-06-14 manuscript/ResponseReview1.Rmd
add Response-letter (and createFigsReview) Rmd file

---
title: "Review 1: Author responses"
author: "Panis & Ramsey"
date: "12.06.2025"
output: 
  pdf_document:
      keep_tex: true
header-includes:
  - \usepackage{framed}
  - \usepackage{xcolor}
  - \let\oldquote\quote
  - \let\endoldquote\endquote
  - \renewenvironment{quote}{\begin{leftbar}}{\end{leftbar}}
---

```{r setup, include=FALSE}
# By default, the code chunks are hidden for brevity. Set the echo argument to TRUE for blocks you want to share the code in the output for. 
knitr::opts_chunk$set(echo = F)
```

```{r packages and data, include=F }

```

# Editor

> 15-Mar-2025
>
> Dear Dr. Panis:
>
> Thank you for submitting your Tutorial (AMPPS-24-0232) entitled "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows" to Advances in Methods and Practices in Psychological Science (AMPPS). I have now received input from three experts in the field, and you will find the reviews below this message and attached. As you will see, all three reviewers believe your paper can make a strong contribution to the literature. Prior to considering the reviewers' input, I read your paper to form an independent opinion. I share the positive sentiments about this work and invite you to revise and resubmit your manuscript for further consideration at AMPPS.

> The reviews are through and thoughtful, and I will not reiterate all the details here. For my part, I would like your revision to focus on accessibility. Your paper is a bit challenging-- it's long and a little tough to get through. What can make it easier for less familiar readers? Perhaps more of a basic set up earlier? A glossary defining specific terms (e.g., what is an event? how is it defines? censoring? hazards vs. survival curves, conditional probabilities, etc.)? Also, clarify when readers need to shift from frequentist to Bayesian model... and why? The main consideration here is that wish to make the paper much more accessible-- from start to finish-- for readers who are novices and entirely unfamiliar with this area. This can be done any number of ways, and R3 also includes some suggestions for making these reviews. 
I encourage you to have the revised paper read and pre-reviewed by colleagues who are novices in this area with the explicit goal of making the paper accessible to them-- this will help you hit the more for our broad readership.

> If you choose to submit a revision, please include include a letter detailing your point-by-point responses to each reviewer comment and indicating how you changed the manuscript to address them. The revised manuscript may undergo further peer review. We ask that you submit your revision within three months. Please let us know if you will not be able to meet this deadline.

> To submit your revision, log into http://mc.manuscriptcentral.com/ampps and enter your Author Center, where you will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.

> IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please delete any redundant or outdated files before completing the submission.

> Once again, thank you for submitting your manuscript to AMPPS and I look forward to receiving your revision.

> Sincerely,

> David A. Sbarra, Ph.D.
> Editor, Advances in Methods and Practices in Psychological Science

We tried to focus on accessibility in the revision of our Tutorial AMPPS-24-0232.

# Reviewer: 1

> Comments to the Author
> It was a pleasure to review this manuscript as someone who has spent years applying survival analysis methods to psychological data. This is an ambitious manuscript and I believe it achieves its goals. The key contributions of this manuscript from my perspective are the creation of custom functions for data wrangling, and the section on planning studies/simulations for power analysis (of course, in addition to the introduction to this category of methods and the practical tutorials provided. I have some comments on this work that I believe could strengthen this manuscript’s contribution.

> General comments:

> The Introduction is really nicely framed in terms of providing compelling information for why this class of statistical models would be useful in some specific areas of psychology (i.e., reaction time data). Given that there are already a couple of tutorial-types of articles introducing specific applications of survival analysis in developmental psychology (references below, both regarding the use of these methods for behavioural observations of emotional expressions in a multilevel framework for recurring events), it would be helpful to further frame the unique contribution that this work adds. The authors may also consider further emphasizing the specific theoretical contributions that using these methods may lead to in the areas of reaction time data etc.

Response: In the revised ms. We added the following section: In this tutorial we provide code to implement discrete-time hazard models in a Bayesian framework, with a focus on cognitive experimental psychology.

> Lougheed, J. P., Benson, L., Ram, N., & Cole, P. M. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53–65. https://doi.org/10.1037/dev0000619

> Stoolmiller, M. (2016). An introduction to using multivariate multilevel survival analysis to study coercive family process. In T. J. Dishion & J. J. Snyder (Eds.), The Oxford handbook of coercive relationship dynamics (pp. 363–378). Oxford University Press.

> Stoolmiller, M., & Snyder, J. (2006). Modeling heterogeneity in social interaction processes using multilevel survival analysis. Psychological Methods, 11(2), 164–177. https://doi.org/10.1037/1082-989X.11.2.164

Response: We added these references to the ms.

> Similarly, it would be helpful to make it clear that this manuscript pertains to estimating single-event occurrences rather than recurring events (which are described in the Lougheed and Stoolmiller tutorials mentioned above)

Response: On page x of the revised ms., we mention that we estimate single-event occurrences that are repeatedly measured in each participant.

> Specific comments:

> The authors use several different terms to refer to time-to-event models throughout the manuscript. The title refers to EHA, and the running head refers to Hazard analysis. Survival analysis is also a common term in Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller & Snyder, 2006). Disambiguating the various terms early in the manuscript and then constantly using one term throughout the manuscript would be helpful.

Response: We disambiguate the various term early in the revised ms. (page x).

> Page 5: it would be helpful to refer to the same time unit in reference to Figure 1, which shows time bins along the x-axis but the examples discussed in the manuscript refer to units of milliseconds.

Response: We refer to the same time unit in Figure 1 and the text.

> Page 7: The authors state that they are not aware of EHA tutorials for psychological time to event data. Previoulsy mentioned tutorial articles by Lougheed and Stoolmiller could be mentioned here— these are demonstrating applications to behavioural observation data within psychology.

Response: We added these 3 references to the revised ms.

> Page 8: I’m curious about the decision to focus on discrete time rather than continuous time models. It would help the reader to explain the decision either way, or refer the reader to Lougheed et al., 2019 who discuss this distinction and choices that researchers may make regarding how to handle decisions around time given data formats. I don’t work with reaction time data, but it seems to me that many experiments in this area would collect data in continuous time intervals, and it’s not clear to me what the benefit of discretizing into time bins is. Now that I’ve read the entire manuscript, I see that some information is presented on this on Page 50, but it would be helpful to make this information known to the reader earlier in the manuscript.

Response: We address the issue of treating time as discrete or continuous on page x of the revised ms.

> Page 8-9, excellent and accessible explanations of the survival function in the discrete time framework.

Response:

> Page 11: In my experience, researchers in psychology are generally not familiar with the concept of right censored data. This term is first used on page 11 with no explanation— a brief, even parenthetical, definition would be helpful. For example, the parenthetical definition on page 48 could be provided on page 11.

Response: We define right-censoring on page x of the revised ms.

> Taken together, this is an excellent manuscript with the potential to further bring survival analysis methods into Psychology, where time-to-event and right-censored cases abound, yet very few people seem to consider these issues.

# Reviewer: 2

> Comments to the Author
> Please see attached file for comments.

> Reviewer report for AMPPS-24-0232: “Event History Analysis for psychological time-to-
event data: A tutorial in R with examples in Bayesian and frequentist workflows”
Summary: The authors present a tutorial paper on the analysis of discrete time-to-event
outcomes. They motivate their tutorials with a description of psychological experiments
that produce time-to-event data and explain how simply summarizing event-time
outcomes using means results in a loss of potentially important information. They then
explain how more information can be gleaned from event-time data if they are
summarized using event history analysis methods. They focus specifically on the
analysis of discrete time-to-event outcomes. They provide code for conducting the
analyses in the tutorials on their Github. Overall, I like the idea of this paper and
strongly agree that using methods appropriate for analyzing time-to-event data, rather
than simply summarizing event-time outcomes using means, is very important and
currently underused. While I like the premise of the paper and think that such a paper
could be a valuable contribution to the literature, I do think the paper has room for
improvement in a number of areas (see comments below). The writing is approachable
(which is advantageous for a tutorial paper) but some text in the paper was repetitive
and could be a bit more polished; additional editing with a focus on clarity and brevity
would be beneficial.

Response: We removed repetitive text, and try to focus on clarity and brevity in the revised ms.

> Major comments:
1. Some discussion on the choice of bin width is needed. Since all of the example
time-to-event data is collected in continuous time, the analyst must discretize it
by defining bins. Defining bins is a key step in the analysis pipeline and so
should not be overlooked. It seems like results—and potentially conclusions
drawn—may depend on the choice of bin width. What factors should the analyst
consider when defining the width of the bins? Does the number of bins matter?
Or is the number of events per bin important? What about other factors? Should
analysts (or potentially these authors too) consider multiple bin widths when
conducting an analysis? An example comparing results when bin width varies
could be both interesting and informative.

Response: We discuss the issue of defining bins on page x of the revised ms.

> 2. In the discussion, the authors mention that they choose to present a tutorial on
discrete time analysis methods because they believe these are easier to
understand. My personal impression is that awareness of continuous time
analysis methods is slightly higher and so researchers may be more familiar with
methods such as Cox models, even if they have not used them themselves;
however, this could vary by specific sub-field. I agree with the authors that the
interpretation of the hazard is easier in the discrete time setting than in the
continuous time setting, but I do think that the distinction between the
interpretation of the hazard in continuous vs discrete time should be discussed in
the introduction. When I first read the introduction, I was confused as to why the
focus was on discrete time methods, when the outcome measures collected in
the motivating data were collected in continuous time. I would suggest either of
the following: (a) provide concrete examples of when truly discrete timeoutcomes are collected to strengthen the motivation of the paper and justification
for use of discrete time analysis methods or (b) keep the motivating examples as-
is and include a short comparison of discrete vs. continuous time analysis
methods to help explain your reasoning for choosing to focus on discrete time
analysis methods.

Response: We include a short comparison of both methods on page x of the revised ms.

> 3. Censoring is an important part of time-to-event data and so more description of
censoring—and how it is (not) accounted for in “orthodox” analysis methods vs.
discrete time-to-event analysis methods would be useful, beyond the already-
included mention of censoring in Section 5.3.

Response: We define (right-)censoring on page x of the revised ms.

> 4. Section 1.2 is a bit long. Please shorten this section to improve the clarity of the
text and reduce redundancy. There is also overlap between text in Section 1.2
and the start of Section 4.

Response: We shortened this section and removed the overlap.

> 5. Section 2.3.2. I like the idea of making a connection with experimental design,
but am having trouble seeing how the discussion in this section is specific to EHA
and how these comments do not apply also to the “orthodox method”. I think this
section could be improved by providing more specific distinctions between
experimental design implications for EHA vs. the “orthodox method”.

Response: We added the following section on page x of the revised ms.: …

> 6. Line 517: What are model weights? Please explain briefly for the reader,
including how the interpretation differs for loo and waic. Also, in the provided
reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
than WAIC, so could you please include a definition of or reference for dWAIC in
your context? Lastly, when reporting the model weights in the example in the
text, I’d suggest including more than 2 digits so that it is clear if the reported
weights are actually 0 and 1 or if rounding has just made them appear so.

Response: We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.

> 7. Tutorials 3a and 3b: Rather than simply stating that the frequentist models did
not converge and resulted in singular fits, it would be helpful to provide more
discussion. Do you have an idea why the models did not converge? Since this
is a tutorial paper, more explanation is needed. These frequentist examples
don’t add much to the paper in their current form, so I’d either suggest either (a)
providing some suggestions to the reader of alternative approaches or specific
strategies that could be used to define a model that will converge or (b) removing
these examples from the main text and directing the reader to the supplement
instead.

Response: We removed the frequentist examples from the main text into the supplementary material.

> Minor comments:
> 1. Is “mean-average comparison” standard terminology? This phrase seems
redundant to me, since means and averages are the same thing.

Response: We removed this term.

> 2. Figure 1: In the caption, please state what the small error bars in the inset plots
are showing.

Response:

> 3. Please consider updating all plots so that they are understandable when viewed
in black and white. In addition to using different colors, I’d suggest distinguishing
the conditions using shapes (e.g., circles vs. triangles) and different line type
(e.g., solid vs. dashed lines).

Response:

> 4. Figure 2 caption:a. Please provide a reference for the statement beginning, “Because the
survival function…”.

Response:

> b. I am confused by the sentence beginning: “For example, the high hazard
of …”. Could you please provide some additional explanation or a
reference for more details on this interpretation?

Response:

> 5. Lines 156-166: Generally, survival probabilities are described as probabilities of
remaining event-free up until a certain point, rather than the probability of an
event occurring in the future. Some clarity in wording could be used here.

Response:

> 6. Please confirm that you have defined all acronyms, abbreviations, and functions
at their first use in the text. E.g., please define WAIC, LOO, P(t), ca(t), SOA, CrI,
among others.

Response: We define all acronyms on their first mention in the revised ms.

> 7. Section 3.1 could be condensed in the main paper (additional details could be
included in the supplementary material if important). Also, please double check
your definition of functional programming.

Response:

> 8. Table 2: It could be helpful to show an example of a trial where an event never
occurs, to help the reader connect what is written in the text (line 325-327) to
what appears in the table.

Response:

> 9. Figure 4: Why is there no bin 40?

Response:

> 10. Line 423: The phrase “early responses” is confusing as it implies temporal
ordering of the responses (i.e., early vs. late trials). I’d suggest using “fast
response” or “rapid responses”, or some other phrase along those lines.

Response:

> 11. Is the discussion of reference and index coding (lines 457 – 464) needed? In the
introduction, the authors state that knowledge of regression is assumed so I am
not sure this is necessary to include, but I defer to the authors’ judgment.

Response: Index coding is not used a lot in experimental psychology so we included a short discussion.

> 12. Figure 6: I only see one width of credible interval for each line on this plot—
please clarify if you are plotting 80% or 95% credible intervals.

Response:

> 13. Line 536: To me, “subject-specific” and “marginal” mean the opposite things.
That is, subject-specific is an effect that is conditional on an individual while a
“marginal” effect has been marginalized/averaged over individuals. Could you
please clarify your use of these terms here?

Response:

> 14. Lines 584-587: Since this is a tutorial paper, I’d suggest including explanations
for why you make changes (a), (b), and (c).

Response:

> 15. Line 659: What is a “fully varying effects structure”? Please define/explain.

Response:

> 16. Line 832-833: Please provide a reference for the statement, “However, they
require much more data…”.

Response:

> 17. Github code: Instruction 2 of the README states “open the
reproducible_workflow.Rproj” – I can’t find this R project, should it be the
“Tutorial_EHA.Rproj” instead?

Response:

# Reviewer: 3

> Comments to the Author
This tutorial guides readers through how to wrangle cognitive reaction time data and address research questions pertaining to temporal dynamics of cognition using event history analysis (EHA) and speed/accuracy tradeoff analysis (SAT). Models are illustrated for hazard functions, and conditional accuracy functions. A nice feature of this study is that each model is presented in both the Bayesian and frequentist frameworks. Having both frameworks provides side-by-side comparisons such as when the Bayesian framework might be preferred (e.g., when models do not converge in the Frequentist framework, which is common when modeling multilevel data). The tutorial also includes a section on simulation and power analysis for planning experiments. I believe this tutorial could eventually be a useful contribution to fellow psychological scientists. However, in its current state, it was difficult to follow/ make my way through (despite existing expertise in EHA), and also difficult to understand what the main contributions were above and beyond existing works. Additionally, it was not clear why the discrete time framework was chosen. The temporal granularity of reaction time data seems like it would be better suited for a continuous time framework. Altogether, the tutorial covers important concepts, but it would benefit from greater accuracy and clearer definitions of key terms. Carefully removing redundancies, getting to the tutorial section much quicker (currently not until pg. 17), and being more precise with explanations of technical jargon would enhance its clarity and accuracy.

Response: We discuss the main contributions above and beyond existing works on page x of the revised ms. We address the issue of treating time as discrete or continuous on page x of the revised ms. We removed redundancies  and included more precise explanations of technical jargon.

> <b>Code for the tutorials:</b> I went to the github page and found a folder for each of the 4 tutorials, which included sub-folders for data, figures, and tables. It took a minute for me to realize that the actual R code for each tutorial was in the main folder, not within each sub-folder. This may also be confusing to readers. Additionally, each of the 4 tutorials, along with their sub-tutorials are provided in separate .Rmd files. There were several download steps needed to get all the information. To me, this made the user experience tedious, and made the tutorials seem disjointed from one another. The authors might consider a more comprehensive single Rmd file that is neatly organized with an interactive table of contents (in the knitted html file). To me, this would make the experience much more enjoyable and useful for readers. Given that this is a tutorial, it would also be nice if the the RMarkdowns were somewhat more standalone e.g., by including a brief description of what a life table is and how to interpret the plot.

Response: We explain the location of the files in the README file. We included a brief description of a life table and how to interpret the plot in the R Mardown tutorial files. A single download step is possible by either cloning the project, or downloading the ZIP file.

> <b>Contribution of this tutorial:</b>
It wasn’t clear to me as a reader what the unique contribution is for this tutorial and what gaps in the literature it fills beyond existing texts and tutorials. For example, is this a tutorial that is meant to specifically focus on experimental psychology cognitive task data – and that this type of data has special characteristics/associated research questions that necessitate their own stand-alone tutorial? If so, it may be helpful to give more detail on what these specific characteristics tend to be that may make it different from the many other types of psychology data. Or, is this a tutorial on discrete time survival analysis that is meant to be broadly applicable to psychological scientists, and the cognitive reaction time data is just for illustrative purposes. If so, then much less emphasis should be placed on the specific data type and more emphasis should be placed more broadly on types of psychological data/ research questions this method could be applied to. I think the distinction makes a big difference in the entire framing of the tutorial, and choosing one or the other will make the tutorial content much easier for readers to absorb.

Response: This is a tutorial specifically focused on Bayesian methods and cognitive experimental psychology data, such as response times and saccade latencies.
Analyzing such event latencies with EHA can give better answers about the time course of psychological processes compared to (a) ANOVA on means and (b) currently popular quantile-based distributional methods. 
Information about the time course of cognitive processes is present in the detailed shape of the hazard function which quantifies the moment-to-moment changes in behavioral performance, in terms of the risk of response occurrence, and the accuracy of responses emitted. EHA thus allows answering research questions about the onset of psychological effects, about time-locking of effects to stimuli, and about how the effect changes on multiple time scales, i.e., with the passage of time within trials, across trials, across blocks within a session, etc. 
Applying EHA also has implications for experimental design, such as response deadlines, timer on trial onset, etc.

> On page 7, it states that there are not tutorials aimed at psychological time-to-event data. Yet, there are at least a few existing empirical article examples with accompanying tutorials in psychology (listed below), and arguably some of the popular textbooks are also geared toward psychological scientists (e.g., Singer & Willett). Similarly, many of the textbooks contain code to accompany their text. Regardless, even if those are not counted, I still think it will be beneficial for readers if the authors can make a clear case for what is unique about psychological time-to-event data that necessitates a stand alone tutorial/ what gaps in the existing textbooks/empirical articles/tutorials the present tutorial serves to fill.

Response: We believe a stand-alone tutorial for cognitive experimental psychology is necessary because (a) psychologists are reluctant to abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as we know, none of the existing tutorials focuses on Bayesian methods.

> Lougheed, J. P., Benson, L., Cole, P. M., & Ram, N. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53.

> Elmer, T., van Duijn, M. A., Ram, N., & Bringmann, L. F. (2023). Modeling categorical time-to-event data: The example of social interaction dynamics captured with event-contingent experience sampling methods. Psychological Methods.

> Mills, M. (2011). The fundamentals of survival and event history analysis. Introducing Survival Analysis and Event History Analysis. London: SAGE Publications, 1-17.

Response: We included these references in the revised ms.

> <b>Organization of the tutorial:</b>The current version of the tutorial is much longer than a typical AMPPStutorial. There are a few possibilities for dealing with this. First, it may be that the tutorial could be split into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own. Another possibility is that the focus could be primarily on the Bayesian models, and then leave the frequentist models for either a supplement, or to only live in the RMarkdown. This possibility may also be beneficial due to the model convergence issues with at least one of the frequentist models (which is common with multilevel data and why many psychological scientists are moving to the Bayesian framework). A third possibility is that the overall tutorial text could be greatly streamlined, with major cuts to both the introduction and the discussion sections.

Response: We focus primarily on the Bayesian models in the revised ms.  We also made major cuts to both the introduction and the discussion sections.

> <b>Clarity and accuracy: </b>
-There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided. For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.

Response: We provide clear descriptions of censoring and risk sets on page x of the revised ms.

> Additionally, many key details, such as the EHA equations are relegated to supplementary materials. In my opinion, any key information that makes the tutorial more accessible and precise should not be in supplementary materials.

Response: We placed the equations for h(t), S(t), and ca(t) in the revised ms. on page x. The equations for P(t) are still in the supplementary material.

> There are many abbreviations that were difficult to follow. EHA is arguably standard, but SAT (for speed/accuracy trade-off) is less familiar, and may therefore benefit from not being an abbreviation. Similarly referring to hazard function as h(t), conditional accuracy function as c(t) was less clear than just saying hazard function and conditional accuracy function.

Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in main body of the revised ms.

> It would be helpful if the research questions that can be addressed by the methods presented are clearly presented in each section. I think that currently they are getting lost in the tutorial.

Response:

> There are several opportunities to describe the method more precisely. For example, the discrete time hazard function gives a ‘conditional’ probability (pg. 9).

Response:

> Additionally, it may be helpful to use more standard notations such as h(t_ij) to indicate the conditional probability that the event occurs in bin j for individual i, given that the individual had not already experienced the event in a previous time period (lines 159-160, pg. 9).

Response:

>  Similarly, the survival probability is the probability that the event occurs after bin j, *given that the individual lhas survived up until bin j*.

Response:

> Throughout the tutorial, reference to individuals in the notation is missing.

Response: We added reference to individuals in the notation.

> Another inaccuracy is that in the discrete time framework, it is not an instantaneous likelihood (line 161, pg. 9).

Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood.

>  <b>Introduction:</b>
-Section 1.1 – I had a difficult time understanding this section as the opening to the tutorial, and the last paragraph seemed out of place/abrupt. felt too vague to be understandable as the opening to the tutorial.

Response:

> Given that the speed-accuracy tradeoff analysis is described as central to this tutorial, it may benefit from being introduced earlier than pg. 7.

Response: We introduce SAT analysis already on page x in the revised ms.

> It seems that comparing and contrasting Bayesian and frequentist methods may be outside the scope of a tutorial (pg. 7)

Response: We agree and focus now on the Bayesian models in the revised ms.

> As mentioned above, tutorial 4 – while related – could potentially be its own stand-alone tutorial.

Response: We decided to keep tutorial 4.

> Censoring is a key component of survival analysis that is missing from the introduction

Response:

> It might be useful to indicate early on the distinctions between multilevel data, single events, and recurring events – and that this particular tutorial is only focused on single events.

Response: We included the following text on page x of the revised ms.: This tutorial deals with single-event analyses which focus on the time until a single event occurs in a trial (e.g., a button press), with repeated measures obtained for each individual. Recurrent event analysis deals with multiple events that can happen repeatedly to the same individual within a single observation period.

> It’s not clear why a discrete time framework is used given the granular temporal resolution of reaction time data – this would also mean that the data would not need to be separated into bins. Most textbooks (e.g., Allison’s) often indicate discrete time methods are common for temporal scales such as months, years, or decades (not milliseconds). Clearer justification for the discrete time framework and separating the data into bins would be helpful.

Response:

> Clearer description of the data would be helpful for readers (as well as only having this description in one place in the tutorial to be more concise). For example, on pg. 10 it is indicated that 200 trials of 1 experimental condition are simulated. Can the event occur at any time in those 200 trials? If the event occurs in trial 150, do the remaining 50 trials still occur, or does the time series end for that person? How are those extra trials after an event occurred handled in the data wrangling and analysis. Can events occur more than once?

Response:

> Section 2.2 (pg. 10, lines 168-169), if describing what ‘statisticians and mathematical psychologists say’, it would be helpful to provide supporting references.

Response: We provide supporting references on page x of the revised ms.

> The numbered list on pg. 11 contained jargon statements without explanations. Suggest to simplify and keep the focus on the data you are using and specific research questions you are using survival analysis as a tool to address.

Response: We removed the numbered list in the revised ms.

> In section 2.3, suggest to lead with the data and research questions of interest and then show how features of survival analysis make it a good method to examine the research questions. I also felt that up until now, the tutorial has been very confusing and hard to follow. It could be that starting the tutorial at section 2.3 (with a few extra details filled in), may be an alternative.

Response: We follow the suggestion of this reviewer to start the tutorial at section 2.3. That is, we immediately provide an example experiment/data set with a descriptive analysis that illustrates the benefits of applying EHA to psychological time-to-event data such as response times and saccade latencies. This new example data set is taken from Panis & Schmidt (2022).

> Again, section 2.3.1 (the majority of pg. 12) was unclear and difficult to follow for someone who already has knowledge of survival analysis, but not much familiarity with cognitive process/reaction time data.

Response:

> From what is written, I did not have a clear sense of how temporal states of cognitive processes and theory development could occur (pg. 14, lines 231-29). These ideas and the research questions are scattered in many places throughout the tutorial, and are not very clear/straightforward. It may be helpful to locate them in just one place in the tutorial and provide a clearer description to guide readers. Similarly, the statement on pg. 14 (lines 238-239) felt rather strong without any references, and also out of scope for a tutorial.

Response:

> It wasn’t clear how section 3.1 added value to the tutorial

Response:

> Section 3.2 – it would be helpful if there was clarification in the text on when it would not be possible to include varying intercepts and slopes (i.e., at least clarifying that random intercepts are a requirement, but sometimes random slopes are removed based on model convergence issues – however, this also begs the question of whether a model should even be presented if it does not converge when it includes random slopes…)

Response:

> Additionally in section 3.2, the text on pg. 16 is vague and does not give much detail to readers.

Response:

> Overall, the introduction spans across 16 pages, which feels long even for a standard empirical article, let alone a tutorial. As currently written, I think many readers will get lost before even getting to the tutorials section.

Response: We restructured and shortened the introduction.

> <b>Tutorials section:</b>
-Rather than only using h(t), and ca(t), suggest to use hazard function and conditional accuracy function along with their mathematical notation. Without having presented any equations, using terms such as h(t) and ca(t) seems odd.

Response: We try to limit the number of abbreviations used in the main body of the revised ms.

> Similarly, P(t) is never defined for readers.

Response: P(t)  is defined only in the supplementary material.

> Section 4.1.1. – it would be helpful to clarify that the risk set also pertains to individuals (not just bins)

Response: We clarify this on page x of the revised ms.

> The connection between trials and bins could be made clearer (this is in reference to the text on pg. 18, but I think the clarity would need to come earlier on in the tutorial when describing the data)

Response:

> Comprehension of the text on pg. 18 would be enhanced if both right and left censoring had been clearly introduced previously. Although left censoring is probably unlikely in the reaction time data used in the tutorial – if it is the case that this tutorial is meant for broader data types in psychology, left censoring would certainly be relevant to many other use cases.

Response: We introduce types of censoring on page x of the revised ms.

> Section 4.1.2 was a clearer data description than had been presented up until that point in the tutorial. Again, it is suggested that the manuscript/tutorial could be dramatically reduced in length if information was more concisely but descriptively written in a single place (e.g., data descriptions).

Response: We provide the information on data descriptions in a single place (page x of the revised ms.).

> Suggest to be more precise with the language on pg. 21 (line 355) – which data are nested within participants? The trial data? The binned data?

Response:

> pg. 22 (line 358) the way that censoring is described here is confusing.

Response:

> pg. 22 discusses warning messages and lets readers know these can be ignored but does not give details as to why. Since this is a didactic tutorial, it may be helpful to have at least some explanation as to why the warning can be ignored.

Response:

> pg. 22 – since this is a tutorial, it might also be helpful to unpack in more detail the section about identifying individuals who may be guessing etc – especially since these are very important considerations.

Response: We unpack in more detail how individuals can be identified that may be guessing, or that do not perform the task as intended.

> section 4.3.1 – (line 451) it would be helpful if more details is given on how to set the analysis time window. What counts as “enough data”? Given that this is a tutorial, these types of detail are critical for readers to be able to implement the method in their own work

Response:

> section 4.3.1 (line 453) – more detail is needed on the cloglog link function and how it differs from the more commonly known about logit link.

Response:

> section 4.3.1 (line 456) – what goes into the decision about whether to treat time as a categorical or continuous predictor?

Response:

> section 4.3.3. – it might be helpful to give a brief indicator up front to readers about the rationale/purpose of fitting both M0i and M1i, especially given that no interpretation is given for M0i. Similarly, it would be helpful to give a bit more detail on the purpose of the comparison of M0i and M1i (given that this is a didactic tutorial).

Response: We provide a brief description of the rationale of comparing models M0i and M1i on page x of the revised ms.

> section 4.3.6. – same thing here – since this is a tutorial, at least brief details on what the purpose of this section is would be helpful. It might also be helpful here to define what is meant by credible interval so that more novice readers don’t confuse it with a confidence interval.

Response:

> The term SOA is used on pg. 34, but had never previously been defined or described.

Response:

> The text on pg. 35 (lines 567-574) seems too in-depth for a tutorial

Response:

> pg. 36 – it would be helpful to give more detail on how to interpret the logit-ca scale

Response:

> section 4.7 – as previously mentioned, it is possible that tutorial 4 could be a stand alone tutorial. If it’s not though, more details are needed up front to tell readers what the purpose is and why this is useful.

Response:

> <b>Discussion section:</b> The Discussion section would benefit from being shortened (AMPPS guidelines indicate discussions should be a brief summary of their contents rather than a general discussion).

Response: We shortened the discussion section.

> is ‘how long’ a component of survival analysis – or is it more accurate to say ‘whether and when psychological states occur’?

Response: It is more accurate to say ‘whether and when’ psychological states occur. However, based on those results, we can deduce information about how long a psychological state lasts within a trial.

> pg. 47 (lines 753-761) – I had a difficult time understanding this section

Response: We rewrote this section.

> section 5.5 – unclear that the multi-state model is actually an example where a multi-state model would be relevant… (the conceptual example given seems like it would be more linked to either a hidden Markov model or recurring event survival model)

Response: We provide a better example in the revised ms.

>  <b>Figures:</b>
-Figure 1 is difficult to follow. It is unclear how the means are being calculated. For example, it seems like in examples 2 and 3, the means between conditions should be different. I was also confused in example 4 because the difference in probability between the two conditions in time bin 2 seems tiny. I was also confused because it still looked like there was a big difference in probability between the two conditions at bin 5. Was bin 6 by chance what was meant instead of bin 5? It also still looked like there were differences later on such as in bin 7. Example 5 also left me confused. It looked to me like the effect started at bin 4 and ended at bin 8. Clarification (and perhaps simplifying) this figure would be useful.

Response:

> The description for figure 2 is unclear, particularly with respect to the contrasting knowledge that can be gained from the survival function and hazard function

Response:

### 2025-06-14 manuscript/CreateFiguresReview.Rmd
add Response-letter (and createFigsReview) Rmd file

---
title: "CreateFiguresReview"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Load the libraries that we will be using.

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "patchwork","cowplot","ggtext","xfun")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Create Figure on experimental paradigms

create detection data

```{r}
max_RT = 500
data_detection_present <- tibble(RT = c(254,NA,234,max_RT,343,283),   
                                 trial = c(6:1))
data_detection_absent <- tibble(RT = c(max_RT,NA,max_RT,max_RT,256,max_RT),   
                                 trial = c(6:1))
```

create detection plot

```{r}
lines_p <- tibble(xstart = 0,
                  xend =   data_detection_present$RT,
                  ystart = data_detection_present$trial,
                  yend =   data_detection_present$trial)

lines_a <- tibble(xstart = 0,
                  xend =   data_detection_absent$RT,
                  ystart = data_detection_absent$trial,
                  yend =   data_detection_absent$trial)

p1 <- data_detection_present %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_p,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill ="black") + # 4 for x
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Target present")+
  ggtitle("Detection task - one button") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

p2 <- data_detection_absent %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(4,NA,4,4,21,4), size =3, fill ="black") +
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Target absent") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Detection task - single button"
p_detection <- (p1/p2) + 
  # plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_detection
```

create categorization data

```{r}
max_RT = 800
data_cat_A <- tibble(RT = c(354,NA,434,max_RT,443,383),   
                                 trial = c(6:1))
data_cat_B <- tibble(RT = c(656,NA,max_RT,437,476,543),   
                                 trial = c(6:1))
```

create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   data_cat_A$RT,
                  ystart = data_cat_A$trial,
                  yend =   data_cat_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   data_cat_B$RT,
                  ystart = data_cat_B$trial,
                  yend =   data_cat_B$trial)

p3 <- data_cat_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill =c("white",NA,"black","black","white","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Discrimination task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

p4 <- data_cat_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,4,21,21,21), size =3, fill =c("black",NA,"black","white","black","white")) +
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Discrimination task - two buttons"
p_discrimination <- (p3/p4) + 
 #  plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_discrimination
```

create unstable-perception data and plot

```{r}
max_RT = 120
data_percept_A <- tibble(RT = c(2,26,44,62,84,113,max_RT,
                                5,31,49,67,89,107,max_RT),   
                                 trial = c(1,1,1,1,1,1,1,
                                           2,2,2,2,2,2,2))
data_percept_B <- tibble(RT = c(4,36,54,75,88,112,max_RT,
                                5,31,45,68,89,114,max_RT),   
                                 trial = c(1,1,1,1,1,1,1,
                                           2,2,2,2,2,2,2))
```

create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   max_RT,
                  ystart = data_percept_A$trial,
                  yend =   data_percept_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   max_RT,
                  ystart = data_percept_B$trial,
                  yend =   data_percept_B$trial)

p7 <- data_percept_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), 
             size =3, 
             fill =c("black","white","black","white","black","white","black", "white","black","white","black","white","black","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT), breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1,2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Bistable perception task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.title = element_textbox_simple())

p8 <- data_percept_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), size =3, fill =c("white","black","white","black","white","black","black","black","white","black","white","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1:2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Discrimination task - two buttons"
p_percept <- (p7/p8) + 
 #  plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_percept
```

create SOA data and plot

```{r}
max_RT = 1200
data_soa_A <- tibble(RT = c(654,NA,734,max_RT,243,683),   
                               trial = c(6:1))
data_soa_B <- tibble(RT = c(956,NA,max_RT,225,776,843),   
                                 trial = c(6:1))
```

create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   data_soa_A$RT,
                  ystart = data_soa_A$trial,
                  yend =   data_soa_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   data_soa_B$RT,
                  ystart = data_soa_B$trial,
                  yend =   data_soa_B$trial)

p5 <- data_soa_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
    geom_rect(mapping=aes(xmin=400, xmax=500, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill =c("white",NA,"black","black","white","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,200,400,600,800,1000,1200), labels=c(-400,-200,0,200,400,600,800)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Discrimination task with SOA manipulation - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.margin = unit(c(1,1,4,1),"lines"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_textbox_simple()) +

  coord_cartesian(clip = "off")+
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(-400,-200,0,200,400,600,800),
           vjust = 2.8,
           size=5,
           fontface="bold") +
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(0,200,400,600,800,1000,1200),
           vjust = 4.2,
           size  = 5,
           fontface="bold")
#+
#  annotate(geom = "text",
#           x = 600,
#           y = 7,
#           label = "RT (ms)",
#           vjust = 4.2,
#           size=7,
#           fontface="bold")

p6 <- data_soa_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
    geom_rect(mapping=aes(xmin=400, xmax=500, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,4,21,21,21), size =3, fill =c("white",NA,"black","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,200,400,600,800,1000,1200), labels=c(-400,-200,0,200,400,600,800)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.margin = unit(c(1,1,4,1),"lines"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +

  coord_cartesian(clip = "off")+
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(-400,-200,0,200,400,600,800),
           vjust = 2.8,
           size=5,
            fontface="bold") +
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(0,200,400,600,800,1000,1200),
           vjust = 4.2,
           size  = 5,
           fontface="bold") +
  annotate(geom = "text",
           x = 600,
           y = 7,
           label = "RT (ms)",
           vjust = 4.2,
           size=7,
           fontface="bold")

#title="Discrimination task - two buttons"
p_discriminationSOA <- (p5/p6) 
 #  plot_annotation(title = title) +
  # plot_layout(axes = "collect_x")

p_discriminationSOA
```

create single Figure

```{r}
plot_grid(p_detection, p_discrimination,p_discriminationSOA,p_percept,
  nrow = 2,
  labels = c("A","B","C","D"),
  label_size = 22#,
  #align = "v"
)
```

save Figure

```{r}
ggsave("REVIEW1/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
```

### 2025-06-14 manuscript/ResponseReview1.tex
Add ResponseReview1.pdf

--- 2025-06-14 manuscript/ResponseReview1.Rmd
@@
----
-title: "Review 1: Author responses"
-author: "Panis & Ramsey"
-date: "12.06.2025"
-output: 
-  pdf_document:
-      keep_tex: true
-header-includes:
-  - \usepackage{framed}
-  - \usepackage{xcolor}
-  - \let\oldquote\quote
-  - \let\endoldquote\endquote
-  - \renewenvironment{quote}{\begin{leftbar}}{\end{leftbar}}
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
+% Set default figure placement to htbp
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{framed}
+\usepackage{xcolor}
+\let\oldquote\quote
+\let\endoldquote\endquote
+\renewenvironment{quote}{\begin{leftbar}}{\end{leftbar}}
+\usepackage{bookmark}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Review 1: Author responses},
+  pdfauthor={Panis \& Ramsey},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-```{r setup, include=FALSE}
-# By default, the code chunks are hidden for brevity. Set the echo argument to TRUE for blocks you want to share the code in the output for. 
-knitr::opts_chunk$set(echo = F)
-```
+\title{Review 1: Author responses}
+\author{Panis \& Ramsey}
+\date{12.06.2025}
 
-```{r packages and data, include=F }
+\begin{document}
+\maketitle
 
-```
+\section{Editor}\label{editor}
 
-# Editor
+\begin{quote}
+15-Mar-2025
 
-> 15-Mar-2025
->
-> Dear Dr. Panis:
->
-> Thank you for submitting your Tutorial (AMPPS-24-0232) entitled "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows" to Advances in Methods and Practices in Psychological Science (AMPPS).
-I have now received input from three experts in the field, and you will find the reviews below this message and attached.
-As you will see, all three reviewers believe your paper can make a strong contribution to the literature.
-Prior to considering the reviewers' input, I read your paper to form an independent opinion.
-I share the positive sentiments about this work and invite you to revise and resubmit your manuscript for further consideration at AMPPS.
+Dear Dr.~Panis:
 
-> The reviews are through and thoughtful, and I will not reiterate all the details here.
-For my part, I would like your revision to focus on accessibility.
-Your paper is a bit challenging-- it's long and a little tough to get through.
-What can make it easier for less familiar readers?
-Perhaps more of a basic set up earlier?
-A glossary defining specific terms (e.g., what is an event?
-how is it defines?
-censoring?
-hazards vs.
-survival curves, conditional probabilities, etc.)?
-Also, clarify when readers need to shift from frequentist to Bayesian model...
-and why?
-The main consideration here is that wish to make the paper much more accessible-- from start to finish-- for readers who are novices and entirely unfamiliar with this area.
-This can be done any number of ways, and R3 also includes some suggestions for making these reviews.
-I encourage you to have the revised paper read and pre-reviewed by colleagues who are novices in this area with the explicit goal of making the paper accessible to them-- this will help you hit the more for our broad readership.
+Thank you for submitting your Tutorial (AMPPS-24-0232) entitled ``Event
+History Analysis for psychological time-to-event data: A tutorial in R
+with examples in Bayesian and frequentist workflows'' to Advances in
+Methods and Practices in Psychological Science (AMPPS). I have now
+received input from three experts in the field, and you will find the
+reviews below this message and attached. As you will see, all three
+reviewers believe your paper can make a strong contribution to the
+literature. Prior to considering the reviewers' input, I read your paper
+to form an independent opinion. I share the positive sentiments about
+this work and invite you to revise and resubmit your manuscript for
+further consideration at AMPPS.
+\end{quote}
 
-> If you choose to submit a revision, please include include a letter detailing your point-by-point responses to each reviewer comment and indicating how you changed the manuscript to address them.
-The revised manuscript may undergo further peer review.
-We ask that you submit your revision within three months.
-Please let us know if you will not be able to meet this deadline.
+\begin{quote}
+The reviews are through and thoughtful, and I will not reiterate all the
+details here. For my part, I would like your revision to focus on
+accessibility. Your paper is a bit challenging-- it's long and a little
+tough to get through. What can make it easier for less familiar readers?
+Perhaps more of a basic set up earlier? A glossary defining specific
+terms (e.g., what is an event? how is it defines? censoring? hazards
+vs.~survival curves, conditional probabilities, etc.)? Also, clarify
+when readers need to shift from frequentist to Bayesian model\ldots{}
+and why? The main consideration here is that wish to make the paper much
+more accessible-- from start to finish-- for readers who are novices and
+entirely unfamiliar with this area. This can be done any number of ways,
+and R3 also includes some suggestions for making these reviews. I
+encourage you to have the revised paper read and pre-reviewed by
+colleagues who are novices in this area with the explicit goal of making
+the paper accessible to them-- this will help you hit the more for our
+broad readership.
+\end{quote}
 
-> To submit your revision, log into http://mc.manuscriptcentral.com/ampps and enter your Author Center, where you will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.
+\begin{quote}
+If you choose to submit a revision, please include include a letter
+detailing your point-by-point responses to each reviewer comment and
+indicating how you changed the manuscript to address them. The revised
+manuscript may undergo further peer review. We ask that you submit your
+revision within three months. Please let us know if you will not be able
+to meet this deadline.
+\end{quote}
 
-> IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please delete any redundant or outdated files before completing the submission.
+\begin{quote}
+To submit your revision, log into
+\url{http://mc.manuscriptcentral.com/ampps} and enter your Author
+Center, where you will find your manuscript title listed under
+``Manuscripts with Decisions.'' Under ``Actions,'' click on ``Create a
+Revision.'' Your manuscript number has been appended to denote a
+revision.
+\end{quote}
 
-> Once again, thank you for submitting your manuscript to AMPPS and I look forward to receiving your revision.
+\begin{quote}
+IMPORTANT: Your original files are available to you when you upload your
+revised manuscript. Please delete any redundant or outdated files before
+completing the submission.
+\end{quote}
 
-> Sincerely,
+\begin{quote}
+Once again, thank you for submitting your manuscript to AMPPS and I look
+forward to receiving your revision.
+\end{quote}
 
-> David A. Sbarra, Ph.D.
-> Editor, Advances in Methods and Practices in Psychological Science
+\begin{quote}
+Sincerely,
+\end{quote}
 
-We tried to focus on accessibility in the revision of our Tutorial AMPPS-24-0232.
+\begin{quote}
+David A. Sbarra, Ph.D. Editor, Advances in Methods and Practices in
+Psychological Science
+\end{quote}
 
-# Reviewer: 1
+We tried to focus on accessibility in the revision of our Tutorial
+AMPPS-24-0232.
 
-> Comments to the Author
-> It was a pleasure to review this manuscript as someone who has spent years applying survival analysis methods to psychological data.
-This is an ambitious manuscript and I believe it achieves its goals.
-The key contributions of this manuscript from my perspective are the creation of custom functions for data wrangling, and the section on planning studies/simulations for power analysis (of course, in addition to the introduction to this category of methods and the practical tutorials provided.
-I have some comments on this work that I believe could strengthen this manuscript’s contribution.
+\section{Reviewer: 1}\label{reviewer-1}
 
-> General comments:
+\begin{quote}
+Comments to the Author It was a pleasure to review this manuscript as
+someone who has spent years applying survival analysis methods to
+psychological data. This is an ambitious manuscript and I believe it
+achieves its goals. The key contributions of this manuscript from my
+perspective are the creation of custom functions for data wrangling, and
+the section on planning studies/simulations for power analysis (of
+course, in addition to the introduction to this category of methods and
+the practical tutorials provided. I have some comments on this work that
+I believe could strengthen this manuscript's contribution.
+\end{quote}
 
-> The Introduction is really nicely framed in terms of providing compelling information for why this class of statistical models would be useful in some specific areas of psychology (i.e., reaction time data).
-Given that there are already a couple of tutorial-types of articles introducing specific applications of survival analysis in developmental psychology (references below, both regarding the use of these methods for behavioural observations of emotional expressions in a multilevel framework for recurring events), it would be helpful to further frame the unique contribution that this work adds.
-The authors may also consider further emphasizing the specific theoretical contributions that using these methods may lead to in the areas of reaction time data etc.
+\begin{quote}
+General comments:
+\end{quote}
 
-Response: In the revised ms.
-We added the following section: In this tutorial we provide code to implement discrete-time hazard models in a Bayesian framework, with a focus on cognitive experimental psychology.
+\begin{quote}
+The Introduction is really nicely framed in terms of providing
+compelling information for why this class of statistical models would be
+useful in some specific areas of psychology (i.e., reaction time data).
+Given that there are already a couple of tutorial-types of articles
+introducing specific applications of survival analysis in developmental
+psychology (references below, both regarding the use of these methods
+for behavioural observations of emotional expressions in a multilevel
+framework for recurring events), it would be helpful to further frame
+the unique contribution that this work adds. The authors may also
+consider further emphasizing the specific theoretical contributions that
+using these methods may lead to in the areas of reaction time data etc.
+\end{quote}
 
-> Lougheed, J.
-P., Benson, L., Ram, N., & Cole, P.
-M.
-(2019).
-Multilevel survival analysis: Studying the timing of children’s recurring behaviors.
-Developmental Psychology, 55(1), 53–65.
-https://doi.org/10.1037/dev0000619
+Response: In the revised ms. We added the following section: In this
+tutorial we provide code to implement discrete-time hazard models in a
+Bayesian framework, with a focus on cognitive experimental psychology.
 
-> Stoolmiller, M.
-(2016).
-An introduction to using multivariate multilevel survival analysis to study coercive family process.
-In T.
-J.
-Dishion & J.
-J.
-Snyder (Eds.), The Oxford handbook of coercive relationship dynamics (pp.
-363–378).
-Oxford University Press.
+\begin{quote}
+Lougheed, J. P., Benson, L., Ram, N., \& Cole, P. M. (2019). Multilevel
+survival analysis: Studying the timing of children's recurring
+behaviors. Developmental Psychology, 55(1), 53--65.
+\url{https://doi.org/10.1037/dev0000619}
+\end{quote}
 
-> Stoolmiller, M., & Snyder, J.
-(2006).
-Modeling heterogeneity in social interaction processes using multilevel survival analysis.
-Psychological Methods, 11(2), 164–177.
-https://doi.org/10.1037/1082-989X.11.2.164
+\begin{quote}
+Stoolmiller, M. (2016). An introduction to using multivariate multilevel
+survival analysis to study coercive family process. In T. J. Dishion \&
+J. J. Snyder (Eds.), The Oxford handbook of coercive relationship
+dynamics (pp.~363--378). Oxford University Press.
+\end{quote}
+
+\begin{quote}
+Stoolmiller, M., \& Snyder, J. (2006). Modeling heterogeneity in social
+interaction processes using multilevel survival analysis. Psychological
+Methods, 11(2), 164--177.
+\url{https://doi.org/10.1037/1082-989X.11.2.164}
+\end{quote}
 
@@
 
-> Similarly, it would be helpful to make it clear that this manuscript pertains to estimating single-event occurrences rather than recurring events (which are described in the Lougheed and Stoolmiller tutorials mentioned above)
+\begin{quote}
+Similarly, it would be helpful to make it clear that this manuscript
+pertains to estimating single-event occurrences rather than recurring
+events (which are described in the Lougheed and Stoolmiller tutorials
+mentioned above)
+\end{quote}
 
-Response: On page x of the revised ms., we mention that we estimate single-event occurrences that are repeatedly measured in each participant.
+Response: On page x of the revised ms., we mention that we estimate
+single-event occurrences that are repeatedly measured in each
+participant.
 
-> Specific comments:
+\begin{quote}
+Specific comments:
+\end{quote}
 
-> The authors use several different terms to refer to time-to-event models throughout the manuscript.
-The title refers to EHA, and the running head refers to Hazard analysis.
-Survival analysis is also a common term in Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller & Snyder, 2006).
-Disambiguating the various terms early in the manuscript and then constantly using one term throughout the manuscript would be helpful.
+\begin{quote}
+The authors use several different terms to refer to time-to-event models
+throughout the manuscript. The title refers to EHA, and the running head
+refers to Hazard analysis. Survival analysis is also a common term in
+Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller \&
+Snyder, 2006). Disambiguating the various terms early in the manuscript
+and then constantly using one term throughout the manuscript would be
+helpful.
+\end{quote}
 
-Response: We disambiguate the various term early in the revised ms. (page x).
+Response: We disambiguate the various term early in the revised ms.
+(page x).
 
-> Page 5: it would be helpful to refer to the same time unit in reference to Figure 1, which shows time bins along the x-axis but the examples discussed in the manuscript refer to units of milliseconds.
+\begin{quote}
+Page 5: it would be helpful to refer to the same time unit in reference
+to Figure 1, which shows time bins along the x-axis but the examples
+discussed in the manuscript refer to units of milliseconds.
+\end{quote}
 
@@
 
-> Page 7: The authors state that they are not aware of EHA tutorials for psychological time to event data.
-Previoulsy mentioned tutorial articles by Lougheed and Stoolmiller could be mentioned here— these are demonstrating applications to behavioural observation data within psychology.
+\begin{quote}
+Page 7: The authors state that they are not aware of EHA tutorials for
+psychological time to event data. Previoulsy mentioned tutorial articles
+by Lougheed and Stoolmiller could be mentioned here--- these are
+demonstrating applications to behavioural observation data within
+psychology.
+\end{quote}
 
@@
 
-> Page 8: I’m curious about the decision to focus on discrete time rather than continuous time models.
-It would help the reader to explain the decision either way, or refer the reader to Lougheed et al., 2019 who discuss this distinction and choices that researchers may make regarding how to handle decisions around time given data formats.
-I don’t work with reaction time data, but it seems to me that many experiments in this area would collect data in continuous time intervals, and it’s not clear to me what the benefit of discretizing into time bins is.
-Now that I’ve read the entire manuscript, I see that some information is presented on this on Page 50, but it would be helpful to make this information known to the reader earlier in the manuscript.
+\begin{quote}
+Page 8: I'm curious about the decision to focus on discrete time rather
+than continuous time models. It would help the reader to explain the
+decision either way, or refer the reader to Lougheed et al., 2019 who
+discuss this distinction and choices that researchers may make regarding
+how to handle decisions around time given data formats. I don't work
+with reaction time data, but it seems to me that many experiments in
+this area would collect data in continuous time intervals, and it's not
+clear to me what the benefit of discretizing into time bins is. Now that
+I've read the entire manuscript, I see that some information is
+presented on this on Page 50, but it would be helpful to make this
+information known to the reader earlier in the manuscript.
+\end{quote}
 
-Response: We address the issue of treating time as discrete or continuous on page x of the revised ms.
+Response: We address the issue of treating time as discrete or
+continuous on page x of the revised ms.
 
-> Page 8-9, excellent and accessible explanations of the survival function in the discrete time framework.
+\begin{quote}
+Page 8-9, excellent and accessible explanations of the survival function
+in the discrete time framework.
+\end{quote}
 
@@
 
-> Page 11: In my experience, researchers in psychology are generally not familiar with the concept of right censored data.
-This term is first used on page 11 with no explanation— a brief, even parenthetical, definition would be helpful.
-For example, the parenthetical definition on page 48 could be provided on page 11.
+\begin{quote}
+Page 11: In my experience, researchers in psychology are generally not
+familiar with the concept of right censored data. This term is first
+used on page 11 with no explanation--- a brief, even parenthetical,
+definition would be helpful. For example, the parenthetical definition
+on page 48 could be provided on page 11.
+\end{quote}
 
@@
 
-> Taken together, this is an excellent manuscript with the potential to further bring survival analysis methods into Psychology, where time-to-event and right-censored cases abound, yet very few people seem to consider these issues.
+\begin{quote}
+Taken together, this is an excellent manuscript with the potential to
+further bring survival analysis methods into Psychology, where
+time-to-event and right-censored cases abound, yet very few people seem
+to consider these issues.
+\end{quote}
 
-# Reviewer: 2
+\section{Reviewer: 2}\label{reviewer-2}
 
-> Comments to the Author
-> Please see attached file for comments.
+\begin{quote}
+Comments to the Author Please see attached file for comments.
+\end{quote}
 
-> Reviewer report for AMPPS-24-0232: “Event History Analysis for psychological time-to-
-event data: A tutorial in R with examples in Bayesian and frequentist workflows”
-Summary: The authors present a tutorial paper on the analysis of discrete time-to-event
-outcomes. They motivate their tutorials with a description of psychological experiments
-that produce time-to-event data and explain how simply summarizing event-time
-outcomes using means results in a loss of potentially important information. They then
-explain how more information can be gleaned from event-time data if they are
-summarized using event history analysis methods. They focus specifically on the
-analysis of discrete time-to-event outcomes. They provide code for conducting the
-analyses in the tutorials on their Github. Overall, I like the idea of this paper and
-strongly agree that using methods appropriate for analyzing time-to-event data, rather
-than simply summarizing event-time outcomes using means, is very important and
-currently underused. While I like the premise of the paper and think that such a paper
-could be a valuable contribution to the literature, I do think the paper has room for
-improvement in a number of areas (see comments below). The writing is approachable
-(which is advantageous for a tutorial paper) but some text in the paper was repetitive
-and could be a bit more polished; additional editing with a focus on clarity and brevity
-would be beneficial.
+\begin{quote}
+Reviewer report for AMPPS-24-0232: ``Event History Analysis for
+psychological time-to- event data: A tutorial in R with examples in
+Bayesian and frequentist workflows'' Summary: The authors present a
+tutorial paper on the analysis of discrete time-to-event outcomes. They
+motivate their tutorials with a description of psychological experiments
+that produce time-to-event data and explain how simply summarizing
+event-time outcomes using means results in a loss of potentially
+important information. They then explain how more information can be
+gleaned from event-time data if they are summarized using event history
+analysis methods. They focus specifically on the analysis of discrete
+time-to-event outcomes. They provide code for conducting the analyses in
+the tutorials on their Github. Overall, I like the idea of this paper
+and strongly agree that using methods appropriate for analyzing
+time-to-event data, rather than simply summarizing event-time outcomes
+using means, is very important and currently underused. While I like the
+premise of the paper and think that such a paper could be a valuable
+contribution to the literature, I do think the paper has room for
+improvement in a number of areas (see comments below). The writing is
+approachable (which is advantageous for a tutorial paper) but some text
+in the paper was repetitive and could be a bit more polished; additional
+editing with a focus on clarity and brevity would be beneficial.
+\end{quote}
 
-Response: We removed repetitive text, and try to focus on clarity and brevity in the revised ms.
+Response: We removed repetitive text, and try to focus on clarity and
+brevity in the revised ms.
 
-> Major comments:
-1. Some discussion on the choice of bin width is needed. Since all of the example
-time-to-event data is collected in continuous time, the analyst must discretize it
-by defining bins. Defining bins is a key step in the analysis pipeline and so
-should not be overlooked. It seems like results—and potentially conclusions
-drawn—may depend on the choice of bin width. What factors should the analyst
-consider when defining the width of the bins? Does the number of bins matter?
-Or is the number of events per bin important? What about other factors? Should
-analysts (or potentially these authors too) consider multiple bin widths when
-conducting an analysis? An example comparing results when bin width varies
-could be both interesting and informative.
+\begin{quote}
+Major comments: 1. Some discussion on the choice of bin width is needed.
+Since all of the example time-to-event data is collected in continuous
+time, the analyst must discretize it by defining bins. Defining bins is
+a key step in the analysis pipeline and so should not be overlooked. It
+seems like results---and potentially conclusions drawn---may depend on
+the choice of bin width. What factors should the analyst consider when
+defining the width of the bins? Does the number of bins matter? Or is
+the number of events per bin important? What about other factors? Should
+analysts (or potentially these authors too) consider multiple bin widths
+when conducting an analysis? An example comparing results when bin width
+varies could be both interesting and informative.
+\end{quote}
 
-Response: We discuss the issue of defining bins on page x of the revised ms.
+Response: We discuss the issue of defining bins on page x of the revised
+ms.
 
-> 2. In the discussion, the authors mention that they choose to present a tutorial on
-discrete time analysis methods because they believe these are easier to
-understand. My personal impression is that awareness of continuous time
-analysis methods is slightly higher and so researchers may be more familiar with
-methods such as Cox models, even if they have not used them themselves;
-however, this could vary by specific sub-field. I agree with the authors that the
-interpretation of the hazard is easier in the discrete time setting than in the
-continuous time setting, but I do think that the distinction between the
-interpretation of the hazard in continuous vs discrete time should be discussed in
-the introduction. When I first read the introduction, I was confused as to why the
-focus was on discrete time methods, when the outcome measures collected in
-the motivating data were collected in continuous time. I would suggest either of
-the following: (a) provide concrete examples of when truly discrete timeoutcomes are collected to strengthen the motivation of the paper and justification
-for use of discrete time analysis methods or (b) keep the motivating examples as-
-is and include a short comparison of discrete vs. continuous time analysis
-methods to help explain your reasoning for choosing to focus on discrete time
-analysis methods.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  In the discussion, the authors mention that they choose to present a
+  tutorial on discrete time analysis methods because they believe these
+  are easier to understand. My personal impression is that awareness of
+  continuous time analysis methods is slightly higher and so researchers
+  may be more familiar with methods such as Cox models, even if they
+  have not used them themselves; however, this could vary by specific
+  sub-field. I agree with the authors that the interpretation of the
+  hazard is easier in the discrete time setting than in the continuous
+  time setting, but I do think that the distinction between the
+  interpretation of the hazard in continuous vs discrete time should be
+  discussed in the introduction. When I first read the introduction, I
+  was confused as to why the focus was on discrete time methods, when
+  the outcome measures collected in the motivating data were collected
+  in continuous time. I would suggest either of the following: (a)
+  provide concrete examples of when truly discrete timeoutcomes are
+  collected to strengthen the motivation of the paper and justification
+  for use of discrete time analysis methods or (b) keep the motivating
+  examples as- is and include a short comparison of discrete
+  vs.~continuous time analysis methods to help explain your reasoning
+  for choosing to focus on discrete time analysis methods.
+\end{enumerate}
+\end{quote}
 
-Response: We include a short comparison of both methods on page x of the revised ms.
+Response: We include a short comparison of both methods on page x of the
+revised ms.
 
-> 3. Censoring is an important part of time-to-event data and so more description of
-censoring—and how it is (not) accounted for in “orthodox” analysis methods vs.
-discrete time-to-event analysis methods would be useful, beyond the already-
-included mention of censoring in Section 5.3.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Censoring is an important part of time-to-event data and so more
+  description of censoring---and how it is (not) accounted for in
+  ``orthodox'' analysis methods vs. discrete time-to-event analysis
+  methods would be useful, beyond the already- included mention of
+  censoring in Section 5.3.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 4. Section 1.2 is a bit long. Please shorten this section to improve the clarity of the
-text and reduce redundancy. There is also overlap between text in Section 1.2
-and the start of Section 4.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Section 1.2 is a bit long. Please shorten this section to improve the
+  clarity of the text and reduce redundancy. There is also overlap
+  between text in Section 1.2 and the start of Section 4.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 5. Section 2.3.2. I like the idea of making a connection with experimental design,
-but am having trouble seeing how the discussion in this section is specific to EHA
-and how these comments do not apply also to the “orthodox method”. I think this
-section could be improved by providing more specific distinctions between
-experimental design implications for EHA vs. the “orthodox method”.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Section 2.3.2. I like the idea of making a connection with
+  experimental design, but am having trouble seeing how the discussion
+  in this section is specific to EHA and how these comments do not apply
+  also to the ``orthodox method''. I think this section could be
+  improved by providing more specific distinctions between experimental
+  design implications for EHA vs.~the ``orthodox method''.
+\end{enumerate}
+\end{quote}
 
-Response: We added the following section on page x of the revised ms.: …
+Response: We added the following section on page x of the revised ms.:
+\ldots{}
 
-> 6. Line 517: What are model weights? Please explain briefly for the reader,
-including how the interpretation differs for loo and waic. Also, in the provided
-reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
-than WAIC, so could you please include a definition of or reference for dWAIC in
-your context? Lastly, when reporting the model weights in the example in the
-text, I’d suggest including more than 2 digits so that it is clear if the reported
-weights are actually 0 and 1 or if rounding has just made them appear so.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{5}
+\tightlist
+\item
+  Line 517: What are model weights? Please explain briefly for the
+  reader, including how the interpretation differs for loo and waic.
+  Also, in the provided reference (Kruz 2023a), I was only able to find
+  a discussion of dWAIC, rather than WAIC, so could you please include a
+  definition of or reference for dWAIC in your context? Lastly, when
+  reporting the model weights in the example in the text, I'd suggest
+  including more than 2 digits so that it is clear if the reported
+  weights are actually 0 and 1 or if rounding has just made them appear
+  so.
+\end{enumerate}
+\end{quote}
 
-Response: We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.
+Response: We explain model weights on page x, include a reference for
+dWAIC, and report the weights with 3 digits.
 
-> 7. Tutorials 3a and 3b: Rather than simply stating that the frequentist models did
-not converge and resulted in singular fits, it would be helpful to provide more
-discussion. Do you have an idea why the models did not converge? Since this
-is a tutorial paper, more explanation is needed. These frequentist examples
-don’t add much to the paper in their current form, so I’d either suggest either (a)
-providing some suggestions to the reader of alternative approaches or specific
-strategies that could be used to define a model that will converge or (b) removing
-these examples from the main text and directing the reader to the supplement
-instead.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{6}
+\tightlist
+\item
+  Tutorials 3a and 3b: Rather than simply stating that the frequentist
+  models did not converge and resulted in singular fits, it would be
+  helpful to provide more discussion. Do you have an idea why the models
+  did not converge? Since this is a tutorial paper, more explanation is
+  needed. These frequentist examples don't add much to the paper in
+  their current form, so I'd either suggest either (a) providing some
+  suggestions to the reader of alternative approaches or specific
+  strategies that could be used to define a model that will converge or
+  (b) removing these examples from the main text and directing the
+  reader to the supplement instead.
+\end{enumerate}
+\end{quote}
 
-Response: We removed the frequentist examples from the main text into the supplementary material.
+Response: We removed the frequentist examples from the main text into
+the supplementary material.
 
-> Minor comments:
-> 1. Is “mean-average comparison” standard terminology? This phrase seems
-redundant to me, since means and averages are the same thing.
+\begin{quote}
+Minor comments: 1. Is ``mean-average comparison'' standard terminology?
+This phrase seems redundant to me, since means and averages are the same
+thing.
+\end{quote}
 
@@
 
-> 2. Figure 1: In the caption, please state what the small error bars in the inset plots
-are showing.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Figure 1: In the caption, please state what the small error bars in
+  the inset plots are showing.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 3. Please consider updating all plots so that they are understandable when viewed
-in black and white. In addition to using different colors, I’d suggest distinguishing
-the conditions using shapes (e.g., circles vs. triangles) and different line type
-(e.g., solid vs. dashed lines).
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Please consider updating all plots so that they are understandable
+  when viewed in black and white. In addition to using different colors,
+  I'd suggest distinguishing the conditions using shapes (e.g., circles
+  vs.~triangles) and different line type (e.g., solid vs.~dashed lines).
+\end{enumerate}
+\end{quote}
 
@@
 
-> 4. Figure 2 caption:a. Please provide a reference for the statement beginning, “Because the
-survival function…”.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Figure 2 caption:a. Please provide a reference for the statement
+  beginning, ``Because the survival function\ldots{}''.
+\end{enumerate}
+\end{quote}
 
@@
 
-> b. I am confused by the sentence beginning: “For example, the high hazard
-of …”. Could you please provide some additional explanation or a
-reference for more details on this interpretation?
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\alph{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  I am confused by the sentence beginning: ``For example, the high
+  hazard of \ldots{}''. Could you please provide some additional
+  explanation or a reference for more details on this interpretation?
+\end{enumerate}
+\end{quote}
 
@@
 
-> 5. Lines 156-166: Generally, survival probabilities are described as probabilities of
-remaining event-free up until a certain point, rather than the probability of an
-event occurring in the future. Some clarity in wording could be used here.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Lines 156-166: Generally, survival probabilities are described as
+  probabilities of remaining event-free up until a certain point, rather
+  than the probability of an event occurring in the future. Some clarity
+  in wording could be used here.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 6. Please confirm that you have defined all acronyms, abbreviations, and functions
-at their first use in the text. E.g., please define WAIC, LOO, P(t), ca(t), SOA, CrI,
-among others.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{5}
+\tightlist
+\item
+  Please confirm that you have defined all acronyms, abbreviations, and
+  functions at their first use in the text. E.g., please define WAIC,
+  LOO, P(t), ca(t), SOA, CrI, among others.
+\end{enumerate}
+\end{quote}
 
-Response: We define all acronyms on their first mention in the revised ms.
+Response: We define all acronyms on their first mention in the revised
+ms.
 
-> 7. Section 3.1 could be condensed in the main paper (additional details could be
-included in the supplementary material if important). Also, please double check
-your definition of functional programming.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{6}
+\tightlist
+\item
+  Section 3.1 could be condensed in the main paper (additional details
+  could be included in the supplementary material if important). Also,
+  please double check your definition of functional programming.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 8. Table 2: It could be helpful to show an example of a trial where an event never
-occurs, to help the reader connect what is written in the text (line 325-327) to
-what appears in the table.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{7}
+\tightlist
+\item
+  Table 2: It could be helpful to show an example of a trial where an
+  event never occurs, to help the reader connect what is written in the
+  text (line 325-327) to what appears in the table.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 9. Figure 4: Why is there no bin 40?
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{8}
+\tightlist
+\item
+  Figure 4: Why is there no bin 40?
+\end{enumerate}
+\end{quote}
 
@@
 
-> 10. Line 423: The phrase “early responses” is confusing as it implies temporal
-ordering of the responses (i.e., early vs. late trials). I’d suggest using “fast
-response” or “rapid responses”, or some other phrase along those lines.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{9}
+\tightlist
+\item
+  Line 423: The phrase ``early responses'' is confusing as it implies
+  temporal ordering of the responses (i.e., early vs.~late trials). I'd
+  suggest using ``fast response'' or ``rapid responses'', or some other
+  phrase along those lines.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 11. Is the discussion of reference and index coding (lines 457 – 464) needed? In the
-introduction, the authors state that knowledge of regression is assumed so I am
-not sure this is necessary to include, but I defer to the authors’ judgment.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{10}
+\tightlist
+\item
+  Is the discussion of reference and index coding (lines 457 -- 464)
+  needed? In the introduction, the authors state that knowledge of
+  regression is assumed so I am not sure this is necessary to include,
+  but I defer to the authors' judgment.
+\end{enumerate}
+\end{quote}
 
-Response: Index coding is not used a lot in experimental psychology so we included a short discussion.
+Response: Index coding is not used a lot in experimental psychology so
+we included a short discussion.
 
-> 12. Figure 6: I only see one width of credible interval for each line on this plot—
-please clarify if you are plotting 80% or 95% credible intervals.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{11}
+\tightlist
+\item
+  Figure 6: I only see one width of credible interval for each line on
+  this plot--- please clarify if you are plotting 80\% or 95\% credible
+  intervals.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 13. Line 536: To me, “subject-specific” and “marginal” mean the opposite things.
-That is, subject-specific is an effect that is conditional on an individual while a
-“marginal” effect has been marginalized/averaged over individuals. Could you
-please clarify your use of these terms here?
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{12}
+\tightlist
+\item
+  Line 536: To me, ``subject-specific'' and ``marginal'' mean the
+  opposite things. That is, subject-specific is an effect that is
+  conditional on an individual while a ``marginal'' effect has been
+  marginalized/averaged over individuals. Could you please clarify your
+  use of these terms here?
+\end{enumerate}
+\end{quote}
 
@@
 
-> 14. Lines 584-587: Since this is a tutorial paper, I’d suggest including explanations
-for why you make changes (a), (b), and (c).
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{13}
+\tightlist
+\item
+  Lines 584-587: Since this is a tutorial paper, I'd suggest including
+  explanations for why you make changes (a), (b), and (c).
+\end{enumerate}
+\end{quote}
 
@@
 
-> 15. Line 659: What is a “fully varying effects structure”? Please define/explain.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{14}
+\tightlist
+\item
+  Line 659: What is a ``fully varying effects structure''? Please
+  define/explain.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 16. Line 832-833: Please provide a reference for the statement, “However, they
-require much more data…”.
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{15}
+\tightlist
+\item
+  Line 832-833: Please provide a reference for the statement, ``However,
+  they require much more data\ldots{}''.
+\end{enumerate}
+\end{quote}
 
@@
 
-> 17. Github code: Instruction 2 of the README states “open the
-reproducible_workflow.Rproj” – I can’t find this R project, should it be the
-“Tutorial_EHA.Rproj” instead?
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{16}
+\tightlist
+\item
+  Github code: Instruction 2 of the README states ``open the
+  reproducible\_workflow.Rproj'' -- I can't find this R project, should
+  it be the ``Tutorial\_EHA.Rproj'' instead?
+\end{enumerate}
+\end{quote}
 
@@
 
-# Reviewer: 3
+\section{Reviewer: 3}\label{reviewer-3}
 
-> Comments to the Author
-This tutorial guides readers through how to wrangle cognitive reaction time data and address research questions pertaining to temporal dynamics of cognition using event history analysis (EHA) and speed/accuracy tradeoff analysis (SAT).
-Models are illustrated for hazard functions, and conditional accuracy functions.
-A nice feature of this study is that each model is presented in both the Bayesian and frequentist frameworks.
-Having both frameworks provides side-by-side comparisons such as when the Bayesian framework might be preferred (e.g., when models do not converge in the Frequentist framework, which is common when modeling multilevel data).
-The tutorial also includes a section on simulation and power analysis for planning experiments.
-I believe this tutorial could eventually be a useful contribution to fellow psychological scientists.
-However, in its current state, it was difficult to follow/ make my way through (despite existing expertise in EHA), and also difficult to understand what the main contributions were above and beyond existing works.
-Additionally, it was not clear why the discrete time framework was chosen.
-The temporal granularity of reaction time data seems like it would be better suited for a continuous time framework.
-Altogether, the tutorial covers important concepts, but it would benefit from greater accuracy and clearer definitions of key terms.
-Carefully removing redundancies, getting to the tutorial section much quicker (currently not until pg.
-17), and being more precise with explanations of technical jargon would enhance its clarity and accuracy.
+\begin{quote}
+Comments to the Author This tutorial guides readers through how to
+wrangle cognitive reaction time data and address research questions
+pertaining to temporal dynamics of cognition using event history
+analysis (EHA) and speed/accuracy tradeoff analysis (SAT). Models are
+illustrated for hazard functions, and conditional accuracy functions. A
+nice feature of this study is that each model is presented in both the
+Bayesian and frequentist frameworks. Having both frameworks provides
+side-by-side comparisons such as when the Bayesian framework might be
+preferred (e.g., when models do not converge in the Frequentist
+framework, which is common when modeling multilevel data). The tutorial
+also includes a section on simulation and power analysis for planning
+experiments. I believe this tutorial could eventually be a useful
+contribution to fellow psychological scientists. However, in its current
+state, it was difficult to follow/ make my way through (despite existing
+expertise in EHA), and also difficult to understand what the main
+contributions were above and beyond existing works. Additionally, it was
+not clear why the discrete time framework was chosen. The temporal
+granularity of reaction time data seems like it would be better suited
+for a continuous time framework. Altogether, the tutorial covers
+important concepts, but it would benefit from greater accuracy and
+clearer definitions of key terms. Carefully removing redundancies,
+getting to the tutorial section much quicker (currently not until pg.
+17), and being more precise with explanations of technical jargon would
+enhance its clarity and accuracy.
+\end{quote}
 
-Response: We discuss the main contributions above and beyond existing works on page x of the revised ms.
-We address the issue of treating time as discrete or continuous on page x of the revised ms.
-We removed redundancies  and included more precise explanations of technical jargon.
+Response: We discuss the main contributions above and beyond existing
+works on page x of the revised ms. We address the issue of treating time
+as discrete or continuous on page x of the revised ms. We removed
+redundancies and included more precise explanations of technical jargon.
 
-> <b>Code for the tutorials:</b> I went to the github page and found a folder for each of the 4 tutorials, which included sub-folders for data, figures, and tables.
-It took a minute for me to realize that the actual R code for each tutorial was in the main folder, not within each sub-folder.
-This may also be confusing to readers.
-Additionally, each of the 4 tutorials, along with their sub-tutorials are provided in separate .Rmd files.
-There were several download steps needed to get all the information.
-To me, this made the user experience tedious, and made the tutorials seem disjointed from one another.
-The authors might consider a more comprehensive single Rmd file that is neatly organized with an interactive table of contents (in the knitted html file).
-To me, this would make the experience much more enjoyable and useful for readers.
-Given that this is a tutorial, it would also be nice if the the RMarkdowns were somewhat more standalone e.g., by including a brief description of what a life table is and how to interpret the plot.
+\begin{quote}
+Code for the tutorials: I went to the github page and found a folder for
+each of the 4 tutorials, which included sub-folders for data, figures,
+and tables. It took a minute for me to realize that the actual R code
+for each tutorial was in the main folder, not within each sub-folder.
+This may also be confusing to readers. Additionally, each of the 4
+tutorials, along with their sub-tutorials are provided in separate .Rmd
+files. There were several download steps needed to get all the
+information. To me, this made the user experience tedious, and made the
+tutorials seem disjointed from one another. The authors might consider a
+more comprehensive single Rmd file that is neatly organized with an
+interactive table of contents (in the knitted html file). To me, this
+would make the experience much more enjoyable and useful for readers.
+Given that this is a tutorial, it would also be nice if the the
+RMarkdowns were somewhat more standalone e.g., by including a brief
+description of what a life table is and how to interpret the plot.
+\end{quote}
 
-Response: We explain the location of the files in the README file.
-We included a brief description of a life table and how to interpret the plot in the R Mardown tutorial files.
-A single download step is possible by either cloning the project, or downloading the ZIP file.
+Response: We explain the location of the files in the README file. We
+included a brief description of a life table and how to interpret the
+plot in the R Mardown tutorial files. A single download step is possible
+by either cloning the project, or downloading the ZIP file.
 
-> <b>Contribution of this tutorial:</b>
-It wasn’t clear to me as a reader what the unique contribution is for this tutorial and what gaps in the literature it fills beyond existing texts and tutorials.
-For example, is this a tutorial that is meant to specifically focus on experimental psychology cognitive task data – and that this type of data has special characteristics/associated research questions that necessitate their own stand-alone tutorial?
-If so, it may be helpful to give more detail on what these specific characteristics tend to be that may make it different from the many other types of psychology data.
-Or, is this a tutorial on discrete time survival analysis that is meant to be broadly applicable to psychological scientists, and the cognitive reaction time data is just for illustrative purposes.
-If so, then much less emphasis should be placed on the specific data type and more emphasis should be placed more broadly on types of psychological data/ research questions this method could be applied to.
-I think the distinction makes a big difference in the entire framing of the tutorial, and choosing one or the other will make the tutorial content much easier for readers to absorb.
+\begin{quote}
+Contribution of this tutorial: It wasn't clear to me as a reader what
+the unique contribution is for this tutorial and what gaps in the
+literature it fills beyond existing texts and tutorials. For example, is
+this a tutorial that is meant to specifically focus on experimental
+psychology cognitive task data -- and that this type of data has special
+characteristics/associated research questions that necessitate their own
+stand-alone tutorial? If so, it may be helpful to give more detail on
+what these specific characteristics tend to be that may make it
+different from the many other types of psychology data. Or, is this a
+tutorial on discrete time survival analysis that is meant to be broadly
+applicable to psychological scientists, and the cognitive reaction time
+data is just for illustrative purposes. If so, then much less emphasis
+should be placed on the specific data type and more emphasis should be
+placed more broadly on types of psychological data/ research questions
+this method could be applied to. I think the distinction makes a big
+difference in the entire framing of the tutorial, and choosing one or
+the other will make the tutorial content much easier for readers to
+absorb.
+\end{quote}
 
-Response: This is a tutorial specifically focused on Bayesian methods and cognitive experimental psychology data, such as response times and saccade latencies.
-Analyzing such event latencies with EHA can give better answers about the time course of psychological processes compared to (a) ANOVA on means and (b) currently popular quantile-based distributional methods.
-Information about the time course of cognitive processes is present in the detailed shape of the hazard function which quantifies the moment-to-moment changes in behavioral performance, in terms of the risk of response occurrence, and the accuracy of responses emitted.
-EHA thus allows answering research questions about the onset of psychological effects, about time-locking of effects to stimuli, and about how the effect changes on multiple time scales, i.e., with the passage of time within trials, across trials, across blocks within a session, etc.
-Applying EHA also has implications for experimental design, such as response deadlines, timer on trial onset, etc.
+Response: This is a tutorial specifically focused on Bayesian methods
+and cognitive experimental psychology data, such as response times and
+saccade latencies. Analyzing such event latencies with EHA can give
+better answers about the time course of psychological processes compared
+to (a) ANOVA on means and (b) currently popular quantile-based
+distributional methods. Information about the time course of cognitive
+processes is present in the detailed shape of the hazard function which
+quantifies the moment-to-moment changes in behavioral performance, in
+terms of the risk of response occurrence, and the accuracy of responses
+emitted. EHA thus allows answering research questions about the onset of
+psychological effects, about time-locking of effects to stimuli, and
+about how the effect changes on multiple time scales, i.e., with the
+passage of time within trials, across trials, across blocks within a
+session, etc. Applying EHA also has implications for experimental
+design, such as response deadlines, timer on trial onset, etc.
 
-> On page 7, it states that there are not tutorials aimed at psychological time-to-event data.
-Yet, there are at least a few existing empirical article examples with accompanying tutorials in psychology (listed below), and arguably some of the popular textbooks are also geared toward psychological scientists (e.g., Singer & Willett).
-Similarly, many of the textbooks contain code to accompany their text.
-Regardless, even if those are not counted, I still think it will be beneficial for readers if the authors can make a clear case for what is unique about psychological time-to-event data that necessitates a stand alone tutorial/ what gaps in the existing textbooks/empirical articles/tutorials the present tutorial serves to fill.
+\begin{quote}
+On page 7, it states that there are not tutorials aimed at psychological
+time-to-event data. Yet, there are at least a few existing empirical
+article examples with accompanying tutorials in psychology (listed
+below), and arguably some of the popular textbooks are also geared
+toward psychological scientists (e.g., Singer \& Willett). Similarly,
+many of the textbooks contain code to accompany their text. Regardless,
+even if those are not counted, I still think it will be beneficial for
+readers if the authors can make a clear case for what is unique about
+psychological time-to-event data that necessitates a stand alone
+tutorial/ what gaps in the existing textbooks/empirical
+articles/tutorials the present tutorial serves to fill.
+\end{quote}
 
-Response: We believe a stand-alone tutorial for cognitive experimental psychology is necessary because (a) psychologists are reluctant to abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as we know, none of the existing tutorials focuses on Bayesian methods.
+Response: We believe a stand-alone tutorial for cognitive experimental
+psychology is necessary because (a) psychologists are reluctant to
+abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as
+we know, none of the existing tutorials focuses on Bayesian methods.
 
-> Lougheed, J. P., Benson, L., Cole, P. M., & Ram, N. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53.
+\begin{quote}
+Lougheed, J. P., Benson, L., Cole, P. M., \& Ram, N. (2019). Multilevel
+survival analysis: Studying the timing of children's recurring
+behaviors. Developmental Psychology, 55(1), 53.
+\end{quote}
 
-> Elmer, T., van Duijn, M.
-A., Ram, N., & Bringmann, L.
-F.
-(2023).
-Modeling categorical time-to-event data: The example of social interaction dynamics captured with event-contingent experience sampling methods.
-Psychological Methods.
+\begin{quote}
+Elmer, T., van Duijn, M. A., Ram, N., \& Bringmann, L. F. (2023).
+Modeling categorical time-to-event data: The example of social
+interaction dynamics captured with event-contingent experience sampling
+methods. Psychological Methods.
+\end{quote}
 
-> Mills, M. (2011). The fundamentals of survival and event history analysis. Introducing Survival Analysis and Event History Analysis. London: SAGE Publications, 1-17.
+\begin{quote}
+Mills, M. (2011). The fundamentals of survival and event history
+analysis. Introducing Survival Analysis and Event History Analysis.
+London: SAGE Publications, 1-17.
+\end{quote}
 
@@
 
-> <b>Organization of the tutorial:</b>The current version of the tutorial is much longer than a typical AMPPStutorial.
-There are a few possibilities for dealing with this.
-First, it may be that the tutorial could be split into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own.
-Another possibility is that the focus could be primarily on the Bayesian models, and then leave the frequentist models for either a supplement, or to only live in the RMarkdown.
-This possibility may also be beneficial due to the model convergence issues with at least one of the frequentist models (which is common with multilevel data and why many psychological scientists are moving to the Bayesian framework).
-A third possibility is that the overall tutorial text could be greatly streamlined, with major cuts to both the introduction and the discussion sections.
+\begin{quote}
+Organization of the tutorial:The current version of the tutorial is much
+longer than a typical AMPPStutorial. There are a few possibilities for
+dealing with this. First, it may be that the tutorial could be split
+into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own.
+Another possibility is that the focus could be primarily on the Bayesian
+models, and then leave the frequentist models for either a supplement,
+or to only live in the RMarkdown. This possibility may also be
+beneficial due to the model convergence issues with at least one of the
+frequentist models (which is common with multilevel data and why many
+psychological scientists are moving to the Bayesian framework). A third
+possibility is that the overall tutorial text could be greatly
+streamlined, with major cuts to both the introduction and the discussion
+sections.
+\end{quote}
 
-Response: We focus primarily on the Bayesian models in the revised ms.  We also made major cuts to both the introduction and the discussion sections.
+Response: We focus primarily on the Bayesian models in the revised ms.
+We also made major cuts to both the introduction and the discussion
+sections.
 
-> <b>Clarity and accuracy: </b>
--There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided.
-For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.
+\begin{quote}
+Clarity and accuracy: -There are several instances where jargon terms
+and key concepts pertinent to survival analysis need to be defined
+and/or described in slightly more detail than what is currently
+provided. For example, clearer descriptions of censoring and risk sets
+(plus what it means to be at risk) should be given in the introduction.
+\end{quote}
 
-Response: We provide clear descriptions of censoring and risk sets on page x of the revised ms.
+Response: We provide clear descriptions of censoring and risk sets on
+page x of the revised ms.
 
-> Additionally, many key details, such as the EHA equations are relegated to supplementary materials.
-In my opinion, any key information that makes the tutorial more accessible and precise should not be in supplementary materials.
+\begin{quote}
+Additionally, many key details, such as the EHA equations are relegated
+to supplementary materials. In my opinion, any key information that
+makes the tutorial more accessible and precise should not be in
+supplementary materials.
+\end{quote}
 
-Response: We placed the equations for h(t), S(t), and ca(t) in the revised ms. on page x. The equations for P(t) are still in the supplementary material.
+Response: We placed the equations for h(t), S(t), and ca(t) in the
+revised ms. on page x. The equations for P(t) are still in the
+supplementary material.
 
-> There are many abbreviations that were difficult to follow.
-EHA is arguably standard, but SAT (for speed/accuracy trade-off) is less familiar, and may therefore benefit from not being an abbreviation.
-Similarly referring to hazard function as h(t), conditional accuracy function as c(t) was less clear than just saying hazard function and conditional accuracy function.
+\begin{quote}
+There are many abbreviations that were difficult to follow. EHA is
+arguably standard, but SAT (for speed/accuracy trade-off) is less
+familiar, and may therefore benefit from not being an abbreviation.
+Similarly referring to hazard function as h(t), conditional accuracy
+function as c(t) was less clear than just saying hazard function and
+conditional accuracy function.
+\end{quote}
 
-Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in main body of the revised ms.
+Response: Although SAT is rather standard in cognitive experimental
+psychology, we tried to limit the use of abbreviations in main body of
+the revised ms.
 
-> It would be helpful if the research questions that can be addressed by the methods presented are clearly presented in each section. I think that currently they are getting lost in the tutorial.
+\begin{quote}
+It would be helpful if the research questions that can be addressed by
+the methods presented are clearly presented in each section. I think
+that currently they are getting lost in the tutorial.
+\end{quote}
 
@@
 
-> There are several opportunities to describe the method more precisely. For example, the discrete time hazard function gives a ‘conditional’ probability (pg. 9).
+\begin{quote}
+There are several opportunities to describe the method more precisely.
+For example, the discrete time hazard function gives a `conditional'
+probability (pg. 9).
+\end{quote}
 
@@
 
-> Additionally, it may be helpful to use more standard notations such as h(t_ij) to indicate the conditional probability that the event occurs in bin j for individual i, given that the individual had not already experienced the event in a previous time period (lines 159-160, pg.
-9).
+\begin{quote}
+Additionally, it may be helpful to use more standard notations such as
+h(t\_ij) to indicate the conditional probability that the event occurs
+in bin j for individual i, given that the individual had not already
+experienced the event in a previous time period (lines 159-160, pg. 9).
+\end{quote}
 
@@
 
->  Similarly, the survival probability is the probability that the event occurs after bin j, *given that the individual lhas survived up until bin j*.
+\begin{quote}
+Similarly, the survival probability is the probability that the event
+occurs after bin j, \emph{given that the individual lhas survived up
+until bin j}.
+\end{quote}
 
@@
 
-> Throughout the tutorial, reference to individuals in the notation is missing.
+\begin{quote}
+Throughout the tutorial, reference to individuals in the notation is
+missing.
+\end{quote}
 
@@
 
-> Another inaccuracy is that in the discrete time framework, it is not an instantaneous likelihood (line 161, pg. 9).
+\begin{quote}
+Another inaccuracy is that in the discrete time framework, it is not an
+instantaneous likelihood (line 161, pg. 9).
+\end{quote}
 
-Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood.
+Response: Indeed, only continuous-time hazard can be called an
+instantaneous likelihood.
 
->  <b>Introduction:</b>
--Section 1.1 – I had a difficult time understanding this section as the opening to the tutorial, and the last paragraph seemed out of place/abrupt.
-felt too vague to be understandable as the opening to the tutorial.
+\begin{quote}
+Introduction: -Section 1.1 -- I had a difficult time understanding this
+section as the opening to the tutorial, and the last paragraph seemed
+out of place/abrupt. felt too vague to be understandable as the opening
+to the tutorial.
+\end{quote}
 
@@
 
-> Given that the speed-accuracy tradeoff analysis is described as central to this tutorial, it may benefit from being introduced earlier than pg. 7.
+\begin{quote}
+Given that the speed-accuracy tradeoff analysis is described as central
+to this tutorial, it may benefit from being introduced earlier than pg.
+7.
+\end{quote}
 
@@
 
-> It seems that comparing and contrasting Bayesian and frequentist methods may be outside the scope of a tutorial (pg. 7)
+\begin{quote}
+It seems that comparing and contrasting Bayesian and frequentist methods
+may be outside the scope of a tutorial (pg. 7)
+\end{quote}
 
-Response: We agree and focus now on the Bayesian models in the revised ms.
+Response: We agree and focus now on the Bayesian models in the revised
+ms.
 
-> As mentioned above, tutorial 4 – while related – could potentially be its own stand-alone tutorial.
+\begin{quote}
+As mentioned above, tutorial 4 -- while related -- could potentially be
+its own stand-alone tutorial.
+\end{quote}
 
@@
 
-> Censoring is a key component of survival analysis that is missing from the introduction
+\begin{quote}
+Censoring is a key component of survival analysis that is missing from
+the introduction
+\end{quote}
 
@@
 
-> It might be useful to indicate early on the distinctions between multilevel data, single events, and recurring events – and that this particular tutorial is only focused on single events.
+\begin{quote}
+It might be useful to indicate early on the distinctions between
+multilevel data, single events, and recurring events -- and that this
+particular tutorial is only focused on single events.
+\end{quote}
 
-Response: We included the following text on page x of the revised ms.: This tutorial deals with single-event analyses which focus on the time until a single event occurs in a trial (e.g., a button press), with repeated measures obtained for each individual.
-Recurrent event analysis deals with multiple events that can happen repeatedly to the same individual within a single observation period.
+Response: We included the following text on page x of the revised ms.:
+This tutorial deals with single-event analyses which focus on the time
+until a single event occurs in a trial (e.g., a button press), with
+repeated measures obtained for each individual. Recurrent event analysis
+deals with multiple events that can happen repeatedly to the same
+individual within a single observation period.
 
-> It’s not clear why a discrete time framework is used given the granular temporal resolution of reaction time data – this would also mean that the data would not need to be separated into bins.
-Most textbooks (e.g., Allison’s) often indicate discrete time methods are common for temporal scales such as months, years, or decades (not milliseconds).
-Clearer justification for the discrete time framework and separating the data into bins would be helpful.
+\begin{quote}
+It's not clear why a discrete time framework is used given the granular
+temporal resolution of reaction time data -- this would also mean that
+the data would not need to be separated into bins. Most textbooks (e.g.,
+Allison's) often indicate discrete time methods are common for temporal
+scales such as months, years, or decades (not milliseconds). Clearer
+justification for the discrete time framework and separating the data
+into bins would be helpful.
+\end{quote}
 
@@
 
-> Clearer description of the data would be helpful for readers (as well as only having this description in one place in the tutorial to be more concise).
-For example, on pg.
-10 it is indicated that 200 trials of 1 experimental condition are simulated.
-Can the event occur at any time in those 200 trials?
-If the event occurs in trial 150, do the remaining 50 trials still occur, or does the time series end for that person?
-How are those extra trials after an event occurred handled in the data wrangling and analysis.
-Can events occur more than once?
+\begin{quote}
+Clearer description of the data would be helpful for readers (as well as
+only having this description in one place in the tutorial to be more
+concise). For example, on pg. 10 it is indicated that 200 trials of 1
+experimental condition are simulated. Can the event occur at any time in
+those 200 trials? If the event occurs in trial 150, do the remaining 50
+trials still occur, or does the time series end for that person? How are
+those extra trials after an event occurred handled in the data wrangling
+and analysis. Can events occur more than once?
+\end{quote}
 
@@
 
-> Section 2.2 (pg. 10, lines 168-169), if describing what ‘statisticians and mathematical psychologists say’, it would be helpful to provide supporting references.
+\begin{quote}
+Section 2.2 (pg. 10, lines 168-169), if describing what `statisticians
+and mathematical psychologists say', it would be helpful to provide
+supporting references.
+\end{quote}
 
@@
 
-> The numbered list on pg.
-11 contained jargon statements without explanations.
-Suggest to simplify and keep the focus on the data you are using and specific research questions you are using survival analysis as a tool to address.
+\begin{quote}
+The numbered list on pg. 11 contained jargon statements without
+explanations. Suggest to simplify and keep the focus on the data you are
+using and specific research questions you are using survival analysis as
+a tool to address.
+\end{quote}
 
@@
 
-> In section 2.3, suggest to lead with the data and research questions of interest and then show how features of survival analysis make it a good method to examine the research questions.
-I also felt that up until now, the tutorial has been very confusing and hard to follow.
-It could be that starting the tutorial at section 2.3 (with a few extra details filled in), may be an alternative.
+\begin{quote}
+In section 2.3, suggest to lead with the data and research questions of
+interest and then show how features of survival analysis make it a good
+method to examine the research questions. I also felt that up until now,
+the tutorial has been very confusing and hard to follow. It could be
+that starting the tutorial at section 2.3 (with a few extra details
+filled in), may be an alternative.
+\end{quote}
 
-Response: We follow the suggestion of this reviewer to start the tutorial at section 2.3.
-That is, we immediately provide an example experiment/data set with a descriptive analysis that illustrates the benefits of applying EHA to psychological time-to-event data such as response times and saccade latencies.
-This new example data set is taken from Panis & Schmidt (2022).
+Response: We follow the suggestion of this reviewer to start the
+tutorial at section 2.3. That is, we immediately provide an example
+experiment/data set with a descriptive analysis that illustrates the
+benefits of applying EHA to psychological time-to-event data such as
+response times and saccade latencies. This new example data set is taken
+from Panis \& Schmidt (2022).
 
-> Again, section 2.3.1 (the majority of pg.
-12) was unclear and difficult to follow for someone who already has knowledge of survival analysis, but not much familiarity with cognitive process/reaction time data.
+\begin{quote}
+Again, section 2.3.1 (the majority of pg. 12) was unclear and difficult
+to follow for someone who already has knowledge of survival analysis,
+but not much familiarity with cognitive process/reaction time data.
+\end{quote}
 
@@
 
-> From what is written, I did not have a clear sense of how temporal states of cognitive processes and theory development could occur (pg.
-14, lines 231-29).
-These ideas and the research questions are scattered in many places throughout the tutorial, and are not very clear/straightforward.
-It may be helpful to locate them in just one place in the tutorial and provide a clearer description to guide readers.
-Similarly, the statement on pg.
-14 (lines 238-239) felt rather strong without any references, and also out of scope for a tutorial.
+\begin{quote}
+From what is written, I did not have a clear sense of how temporal
+states of cognitive processes and theory development could occur (pg.
+14, lines 231-29). These ideas and the research questions are scattered
+in many places throughout the tutorial, and are not very
+clear/straightforward. It may be helpful to locate them in just one
+place in the tutorial and provide a clearer description to guide
+readers. Similarly, the statement on pg. 14 (lines 238-239) felt rather
+strong without any references, and also out of scope for a tutorial.
+\end{quote}
 
@@
 
-> It wasn’t clear how section 3.1 added value to the tutorial
+\begin{quote}
+It wasn't clear how section 3.1 added value to the tutorial
+\end{quote}
 
@@
 
-> Section 3.2 – it would be helpful if there was clarification in the text on when it would not be possible to include varying intercepts and slopes (i.e., at least clarifying that random intercepts are a requirement, but sometimes random slopes are removed based on model convergence issues – however, this also begs the question of whether a model should even be presented if it does not converge when it includes random slopes…)
+\begin{quote}
+Section 3.2 -- it would be helpful if there was clarification in the
+text on when it would not be possible to include varying intercepts and
+slopes (i.e., at least clarifying that random intercepts are a
+requirement, but sometimes random slopes are removed based on model
+convergence issues -- however, this also begs the question of whether a
+model should even be presented if it does not converge when it includes
+random slopes\ldots)
+\end{quote}
 
@@
 
-> Additionally in section 3.2, the text on pg. 16 is vague and does not give much detail to readers.
+\begin{quote}
+Additionally in section 3.2, the text on pg. 16 is vague and does not
+give much detail to readers.
+\end{quote}
 
@@
 
-> Overall, the introduction spans across 16 pages, which feels long even for a standard empirical article, let alone a tutorial.
-As currently written, I think many readers will get lost before even getting to the tutorials section.
+\begin{quote}
+Overall, the introduction spans across 16 pages, which feels long even
+for a standard empirical article, let alone a tutorial. As currently
+written, I think many readers will get lost before even getting to the
+tutorials section.
+\end{quote}
 
@@
 
-> <b>Tutorials section:</b>
--Rather than only using h(t), and ca(t), suggest to use hazard function and conditional accuracy function along with their mathematical notation.
-Without having presented any equations, using terms such as h(t) and ca(t) seems odd.
+\begin{quote}
+Tutorials section: -Rather than only using h(t), and ca(t), suggest to
+use hazard function and conditional accuracy function along with their
+mathematical notation. Without having presented any equations, using
+terms such as h(t) and ca(t) seems odd.
+\end{quote}
 
-Response: We try to limit the number of abbreviations used in the main body of the revised ms.
+Response: We try to limit the number of abbreviations used in the main
+body of the revised ms.
 
-> Similarly, P(t) is never defined for readers.
+\begin{quote}
+Similarly, P(t) is never defined for readers.
+\end{quote}
 
@@
 
-> Section 4.1.1. – it would be helpful to clarify that the risk set also pertains to individuals (not just bins)
+\begin{quote}
+Section 4.1.1. -- it would be helpful to clarify that the risk set also
+pertains to individuals (not just bins)
+\end{quote}
 
@@
 
-> The connection between trials and bins could be made clearer (this is in reference to the text on pg.
-18, but I think the clarity would need to come earlier on in the tutorial when describing the data)
+\begin{quote}
+The connection between trials and bins could be made clearer (this is in
+reference to the text on pg. 18, but I think the clarity would need to
+come earlier on in the tutorial when describing the data)
+\end{quote}
 
@@
 
-> Comprehension of the text on pg.
-18 would be enhanced if both right and left censoring had been clearly introduced previously.
-Although left censoring is probably unlikely in the reaction time data used in the tutorial – if it is the case that this tutorial is meant for broader data types in psychology, left censoring would certainly be relevant to many other use cases.
+\begin{quote}
+Comprehension of the text on pg. 18 would be enhanced if both right and
+left censoring had been clearly introduced previously. Although left
+censoring is probably unlikely in the reaction time data used in the
+tutorial -- if it is the case that this tutorial is meant for broader
+data types in psychology, left censoring would certainly be relevant to
+many other use cases.
+\end{quote}
 
@@
 
-> Section 4.1.2 was a clearer data description than had been presented up until that point in the tutorial.
-Again, it is suggested that the manuscript/tutorial could be dramatically reduced in length if information was more concisely but descriptively written in a single place (e.g., data descriptions).
+\begin{quote}
+Section 4.1.2 was a clearer data description than had been presented up
+until that point in the tutorial. Again, it is suggested that the
+manuscript/tutorial could be dramatically reduced in length if
+information was more concisely but descriptively written in a single
+place (e.g., data descriptions).
+\end{quote}
 
-Response: We provide the information on data descriptions in a single place (page x of the revised ms.).
+Response: We provide the information on data descriptions in a single
+place (page x of the revised ms.).
 
-> Suggest to be more precise with the language on pg. 21 (line 355) – which data are nested within participants? The trial data? The binned data?
+\begin{quote}
+Suggest to be more precise with the language on pg. 21 (line 355) --
+which data are nested within participants? The trial data? The binned
+data?
+\end{quote}
 
@@
 
-> pg. 22 (line 358) the way that censoring is described here is confusing.
+\begin{quote}
+pg. 22 (line 358) the way that censoring is described here is confusing.
+\end{quote}
 
@@
 
-> pg.
-22 discusses warning messages and lets readers know these can be ignored but does not give details as to why.
-Since this is a didactic tutorial, it may be helpful to have at least some explanation as to why the warning can be ignored.
+\begin{quote}
+pg. 22 discusses warning messages and lets readers know these can be
+ignored but does not give details as to why. Since this is a didactic
+tutorial, it may be helpful to have at least some explanation as to why
+the warning can be ignored.
+\end{quote}
 
@@
 
-> pg.
-22 – since this is a tutorial, it might also be helpful to unpack in more detail the section about identifying individuals who may be guessing etc – especially since these are very important considerations.
+\begin{quote}
+pg. 22 -- since this is a tutorial, it might also be helpful to unpack
+in more detail the section about identifying individuals who may be
+guessing etc -- especially since these are very important
+considerations.
+\end{quote}
 
-Response: We unpack in more detail how individuals can be identified that may be guessing, or that do not perform the task as intended.
+Response: We unpack in more detail how individuals can be identified
+that may be guessing, or that do not perform the task as intended.
 
-> section 4.3.1 – (line 451) it would be helpful if more details is given on how to set the analysis time window.
-What counts as “enough data”?
-Given that this is a tutorial, these types of detail are critical for readers to be able to implement the method in their own work
+\begin{quote}
+section 4.3.1 -- (line 451) it would be helpful if more details is given
+on how to set the analysis time window. What counts as ``enough data''?
+Given that this is a tutorial, these types of detail are critical for
+readers to be able to implement the method in their own work
+\end{quote}
 
@@
 
-> section 4.3.1 (line 453) – more detail is needed on the cloglog link function and how it differs from the more commonly known about logit link.
+\begin{quote}
+section 4.3.1 (line 453) -- more detail is needed on the cloglog link
+function and how it differs from the more commonly known about logit
+link.
+\end{quote}
 
@@
 
-> section 4.3.1 (line 456) – what goes into the decision about whether to treat time as a categorical or continuous predictor?
+\begin{quote}
+section 4.3.1 (line 456) -- what goes into the decision about whether to
+treat time as a categorical or continuous predictor?
+\end{quote}
 
@@
 
-> section 4.3.3.
-– it might be helpful to give a brief indicator up front to readers about the rationale/purpose of fitting both M0i and M1i, especially given that no interpretation is given for M0i.
-Similarly, it would be helpful to give a bit more detail on the purpose of the comparison of M0i and M1i (given that this is a didactic tutorial).
+\begin{quote}
+section 4.3.3. -- it might be helpful to give a brief indicator up front
+to readers about the rationale/purpose of fitting both M0i and M1i,
+especially given that no interpretation is given for M0i. Similarly, it
+would be helpful to give a bit more detail on the purpose of the
+comparison of M0i and M1i (given that this is a didactic tutorial).
+\end{quote}
 
-Response: We provide a brief description of the rationale of comparing models M0i and M1i on page x of the revised ms.
+Response: We provide a brief description of the rationale of comparing
+models M0i and M1i on page x of the revised ms.
 
-> section 4.3.6.
-– same thing here – since this is a tutorial, at least brief details on what the purpose of this section is would be helpful.
-It might also be helpful here to define what is meant by credible interval so that more novice readers don’t confuse it with a confidence interval.
+\begin{quote}
+section 4.3.6. -- same thing here -- since this is a tutorial, at least
+brief details on what the purpose of this section is would be helpful.
+It might also be helpful here to define what is meant by credible
+interval so that more novice readers don't confuse it with a confidence
+interval.
+\end{quote}
 
@@
 
-> The term SOA is used on pg. 34, but had never previously been defined or described.
+\begin{quote}
+The term SOA is used on pg. 34, but had never previously been defined or
+described.
+\end{quote}
 
@@
 
-> The text on pg. 35 (lines 567-574) seems too in-depth for a tutorial
+\begin{quote}
+The text on pg. 35 (lines 567-574) seems too in-depth for a tutorial
+\end{quote}
 
@@
 
-> pg. 36 – it would be helpful to give more detail on how to interpret the logit-ca scale
+\begin{quote}
+pg. 36 -- it would be helpful to give more detail on how to interpret
+the logit-ca scale
+\end{quote}
 
@@
 
-> section 4.7 – as previously mentioned, it is possible that tutorial 4 could be a stand alone tutorial.
-If it’s not though, more details are needed up front to tell readers what the purpose is and why this is useful.
+\begin{quote}
+section 4.7 -- as previously mentioned, it is possible that tutorial 4
+could be a stand alone tutorial. If it's not though, more details are
+needed up front to tell readers what the purpose is and why this is
+useful.
+\end{quote}
 
@@
 
-> <b>Discussion section:</b> The Discussion section would benefit from being shortened (AMPPS guidelines indicate discussions should be a brief summary of their contents rather than a general discussion).
+\begin{quote}
+Discussion section: The Discussion section would benefit from being
+shortened (AMPPS guidelines indicate discussions should be a brief
+summary of their contents rather than a general discussion).
+\end{quote}
 
@@
 
-> is ‘how long’ a component of survival analysis – or is it more accurate to say ‘whether and when psychological states occur’?
+\begin{quote}
+is `how long' a component of survival analysis -- or is it more accurate
+to say `whether and when psychological states occur'?
+\end{quote}
 
-Response: It is more accurate to say ‘whether and when’ psychological states occur.
-However, based on those results, we can deduce information about how long a psychological state lasts within a trial.
+Response: It is more accurate to say `whether and when' psychological
+states occur. However, based on those results, we can deduce information
+about how long a psychological state lasts within a trial.
 
-> pg. 47 (lines 753-761) – I had a difficult time understanding this section
+\begin{quote}
+pg. 47 (lines 753-761) -- I had a difficult time understanding this
+section
+\end{quote}
 
@@
 
-> section 5.5 – unclear that the multi-state model is actually an example where a multi-state model would be relevant… (the conceptual example given seems like it would be more linked to either a hidden Markov model or recurring event survival model)
+\begin{quote}
+section 5.5 -- unclear that the multi-state model is actually an example
+where a multi-state model would be relevant\ldots{} (the conceptual
+example given seems like it would be more linked to either a hidden
+Markov model or recurring event survival model)
+\end{quote}
 
@@
 
->  <b>Figures:</b>
--Figure 1 is difficult to follow.
-It is unclear how the means are being calculated.
-For example, it seems like in examples 2 and 3, the means between conditions should be different.
-I was also confused in example 4 because the difference in probability between the two conditions in time bin 2 seems tiny.
-I was also confused because it still looked like there was a big difference in probability between the two conditions at bin 5.
-Was bin 6 by chance what was meant instead of bin 5?
-It also still looked like there were differences later on such as in bin 7.
-Example 5 also left me confused.
-It looked to me like the effect started at bin 4 and ended at bin 8.
-Clarification (and perhaps simplifying) this figure would be useful.
+\begin{quote}
+Figures: -Figure 1 is difficult to follow. It is unclear how the means
+are being calculated. For example, it seems like in examples 2 and 3,
+the means between conditions should be different. I was also confused in
+example 4 because the difference in probability between the two
+conditions in time bin 2 seems tiny. I was also confused because it
+still looked like there was a big difference in probability between the
+two conditions at bin 5. Was bin 6 by chance what was meant instead of
+bin 5? It also still looked like there were differences later on such as
+in bin 7. Example 5 also left me confused. It looked to me like the
+effect started at bin 4 and ended at bin 8. Clarification (and perhaps
+simplifying) this figure would be useful.
+\end{quote}
 
@@
 
-> The description for figure 2 is unclear, particularly with respect to the contrasting knowledge that can be gained from the survival function and hazard function
+\begin{quote}
+The description for figure 2 is unclear, particularly with respect to
+the contrasting knowledge that can be gained from the survival function
+and hazard function
+\end{quote}
 
 Response:
+
+\end{document}

### 2025-06-14 sims/sims.Rmd
Add code for Figure 1 in revision1

@@
 data_n10 <- sim_haz(subj_n = 10, rep_n = 200)
+head(data_n10)
+
 data_n20 <- sim_haz(subj_n = 20, rep_n = 200)
@@
 
-ggsave("sims/figures/haz_acc_single.jpeg",
-       width = 7, height = 6, dpi=800)
-
-```
+#ggsave("sims/figures/haz_acc_single.jpeg",
+#       width = 7, height = 6, dpi=800)
+
+```
+
+# section 13: REVISION 1: Create New Figure 1
+
+In Figure 1 of the revision, we want to display for 1 subject the hazard and ca functions for 2 conditions, and the corresponding four means.
+
+Steps:
+
+## 1. Generate discrete time-to-event data for 1 subject, 2 conditions, 100 trials, and 10 time bins;
+
+Adjust function:
+
+```{r}
+sim_haz <- function(subj_n = 10, rep_n = 200, bin_n = 10, cond_n = 2,  # these can be changed when calling the f
+               hazard_cond1 = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
+               hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, .5, 0.3, 0.7, 1, 1,1),
+               hazard = c(hazard_cond1, hazard_cond2),   # 
+                ... # helps the function work with pmap() below
+                ) {
+
+  # set up data structure
+  data <- add_random(subj = subj_n, rep = rep_n) %>% 
+  add_within("subj", condition = c("cond1", "cond2")) %>%
+  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
+  add_within("subj", bin = 1:bin_n) %>% 
+  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
+         event = rbinom(n(), 1, hazf)) %>% 
+  group_by(subj, rep, condition) %>%
+  mutate(cumsum = cumsum(event),
+         outcomel = event==1 & cumsum(event) < 2,
+         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
+                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
+  drop_na(outcomen) %>% 
+  ungroup()
+}
+```
+
+Generate data:
+
+```{r}
+## make it reproducible
+set.seed(7)
+
+n_trials=200
+## apply function sim_haz
+data_fig1_hazard <- sim_haz(subj_n = 1, rep_n = n_trials) %>%
+               mutate(subj = factor(subj),
+                      trial = as.integer(str_sub(rep, start=4, end=6)),
+                      acc = ifelse(bin<5 & condition=="cond2",0,1),
+                      acc = ifelse(bin==5 & trial < 100,0,acc)) %>%
+               select(-c("contrast","cumsum","outcomel","outcomen","rep"))
+
+head(data_fig1_hazard, n=20)
+summary(data_fig1_hazard)
+```
+
+Set up life table for generated data:
+
+```{r}
+lifetable_fig1 <- data_fig1_hazard %>%
+    mutate(event = str_c("event", event)) %>%
+    group_by(condition,bin) %>% 
+    count(event) %>% 
+    ungroup() %>% 
+    pivot_wider(names_from = event,
+                values_from = n) %>% 
+    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
+           event1 = ifelse(is.na(event1),0,event1),
+           risk_set = event0 + event1) %>% # define the risk set
+    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
+    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
+           pmass = event1/max(risk_set),
+           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
+    group_by(condition) %>%
+    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
+           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
+           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
+    ungroup() 
+```
+
+Calaculate conditional accuracy:
+
+```{r}
+ca <- data_fig1_hazard %>% 
+    filter(event == 1) %>%
+    group_by(condition,bin) %>%
+    summarize(ca = mean(acc) %>% round(digits = 2),
+              n = n(),
+              .groups = 'drop') %>%
+    ungroup() %>%
+    mutate(se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) 
+ca
+```
+
+## 2. Plot hazard and conditional accuracy
+
+```{r}
+p13.1 <-ggplot(lifetable_fig1, aes(x = bin, y = hazard,
+                         colour = condition)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10, limits = c(1,10),labels = c(1:10)*100) +
+  scale_y_continuous(limits = c(0,1)) +
+ # ggtitle("hazard and conditional accuracy functions") +
+  xlab("time bin endpoint") +
+  theme(axis.text.x = element_text(angle = 90))
+p13.1
+
+#ggsave("sims/figures/haz_surv_single.jpeg",
+#       width = 7, height = 5, dpi=800)
+```
+
+```{r}
+p13.2 <-ggplot(ca, aes(x = bin, y = ca,
+                         colour = condition)) +
+  geom_line(linewidth = 1) +
+  geom_point(size=3) +
+  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
+  scale_colour_brewer(palette = "Dark2") + 
+  scale_x_continuous(breaks = 1:10, limits=c(1,10),labels = c(1:10)*100) +
+  scale_y_continuous(limits = c(0,1)) +
+  xlab("time bin endpoint") +
+    ylab("conditional accuracy") +
+   theme(axis.text.x = element_text(angle = 90))
+
+ # ggtitle("conditional accuracy functions") 
+#+ 
+#  guides(color=guide_legend(title="prime type"))
+p13.2
+```
+
+Combine plots
+
+```{r}
+p13.3 <- (p13.1 / p13.2) +
+  plot_layout(guides = 'collect',
+              axes = 'collect')
+p13.3
+```
+
+## 3. Create RT data (assuming uniform distribution within a bin) and calculate means
+
+use a uniform distribution to provide upper and lower bounds
+
+```{r}
+rt_dat_fig1 <- data_fig1_hazard %>%
+  filter(event==1) %>% 
+  mutate(u = bin*100,
+         l = u-100,
+         m = (l + u)/2) %>% 
+  group_by(subj, trial, condition, bin) %>% 
+  # mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
+  mutate(rt = runif(n=1, min=l, max=u))
+
+head(rt_dat_fig1)
+summary(rt_dat_fig1)
+```
+
+create summary RT data per condition at the pid level for plotting mean comparisons
+
+```{r}
+rt_dat_fig1_pid <- rt_dat_fig1 %>%
+  group_by(condition) %>% 
+  summarise(mean = mean(rt),
+            sd = sd(rt),
+            n = n(), # 
+            sem = (sd/sqrt(n)))
+rt_dat_fig1_pid
+```
+
+create summary ACC data per condition at the pid level for plotting mean comparisons
+
+```{r}
+acc_dat_fig1_pid <- rt_dat_fig1 %>%
+  group_by(condition) %>% 
+  summarise(mean = mean(acc),
+            sd = sd(acc),
+            n = n(), # 
+            sem = (sd/sqrt(n)))
+acc_dat_fig1_pid
+```
+
+## 4. plot rt + accuracy
+
+```{r}
+p13.4 <- ggplot(rt_dat_fig1_pid, aes(x = condition, y = mean, fill=condition, color=condition)) +
+  geom_col(alpha = 0.9, width = 0.6) +
+  geom_errorbar(
+    aes(ymin = mean - sem, ymax = mean + sem),
+    width = 0.2,
+    linewidth = 0.8,
+    color = "black"
+  ) +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+ #  ggtitle("mean RT by condition") +
+   scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, 100)) +
+   ylab("mean RT (ms)")
+p13.4
+```
+
+```{r}
+p13.5 <- ggplot(acc_dat_fig1_pid, aes(x = condition, y = mean, fill=condition, color=condition)) +
+  geom_col(alpha = 0.9, width = 0.6) +
+  geom_errorbar(
+    aes(ymin = mean - sem, ymax = mean + sem),
+    width = 0.2,
+    linewidth = 0.8,
+    color = "black"
+  ) +
+   scale_fill_brewer(palette = "Dark2") +
+   scale_colour_brewer(palette = "Dark2") +
+ #  ggtitle("mean accuracy by condition") +
+   scale_y_continuous(limits = c(0, 1)) +
+   ylab("accuracy")
+p13.5
+```
+
+plot together
+
+```{r}
+p13.6 <- (p13.4 / p13.5) +
+  plot_layout(guides = 'collect',
+              axes = 'collect')
+p13.6
+
+plot_grid(p13.6, p13.3,
+  nrow = 1,
+  labels = c("A","B"),
+  label_size = 22
+)
+
+```
+
+```{r}
+ggsave("sims/figures/Fig1_revision1.jpeg",
+       width = 10, height = 10, dpi=800)
+
+```

### 2025-06-14 sims/sims.Rmd
Improve Figure 1 details

@@
  # ggtitle("hazard and conditional accuracy functions") +
-  xlab("time bin endpoint") +
-  theme(axis.text.x = element_text(angle = 90))
+  xlab("time bin endpoint (ms)") +
+  ylab("hazard") +
+  theme(axis.text.x = element_text(angle = 90),
+        panel.grid.minor.x = element_blank()) +
+  annotate("label", x = 2, y = .95,
+           label = "h(t) = P(RT = t | RT ≥ t)",
+           fill = "white",           # White background
+           color = "black", 
+           size = 7, 
+           hjust = 0)
 p13.1
@@
   scale_y_continuous(limits = c(0,1)) +
-  xlab("time bin endpoint") +
+  xlab("time bin endpoint (ms)") +
     ylab("conditional accuracy") +
-   theme(axis.text.x = element_text(angle = 90))
+   theme(axis.text.x = element_text(angle = 90),
+         panel.grid.minor.x = element_blank()) +
+   annotate("label", x = 2, y = .25,
+           label = "ca(t) = P(correct | RT = t)",
+           fill = "white",           # White background
+           color = "black", 
+           size = 7, 
+           hjust = 0)
 
@@
    scale_y_continuous(limits = c(0, 1)) +
-   ylab("accuracy")
+   ylab("mean accuracy (%)")
 p13.5
@@
 p13.6
-
+```
+
+Plot mean and distributional results:
+
+```{r}
 plot_grid(p13.6, p13.3,
@@
 )
-
 ```

### 2025-06-16 sims/sims.Rmd
small changes to figure1 for revision

@@
 ```{r load-pkg}
-pkg <- c("tidyverse", "RColorBrewer", "patchwork", "truncnorm", "faux", "ggpmisc")
+pkg <- c("tidyverse", "RColorBrewer", "patchwork", "truncnorm", "faux", "ggpmisc","cowplot")
 
@@
         panel.grid.minor.x = element_blank()) +
-  annotate("label", x = 2, y = .95,
+  annotate("label", x = 1.7, y = .95,
            label = "h(t) = P(RT = t | RT ≥ t)",
@@
          panel.grid.minor.x = element_blank()) +
-   annotate("label", x = 2, y = .25,
+   annotate("label", x = 1.5, y = .25,
            label = "ca(t) = P(correct | RT = t)",

### 2025-06-16 manuscript/Suppl_material/supplementary.Rmd
Initial revision of 1. Introduction

@@
 
+## B. Benefits of event history analysis
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
+Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
+The benefits include:
+
+1. Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
+
+2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
+
+3.
+EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
+
+4.
+Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+[@allisonSurvivalAnalysisUsing2010].
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
+
+5.
+EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
+
 ## B. Custom functions for descriptive discrete-time hazard analysis

### 2025-06-16 manuscript/manuscript.Rmd
added link to psycharxiv
Initial revision of 1. Introduction

@@
 In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
+Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
 Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 
-As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
+As a simple illustration, Figure 1 shows how comparing means between two conditions conceal the shapes of the underlying RT and accuracy distributions.
+We simulated a RT + accuracy data set for a single subject who performed 200 trials (i.e., repeated measurements) in each of two conditions.
+For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500] or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
+Similarly, while this subject makes less errors in condition 1 (86% accuracy) compared to condition 2 (64% accuracy), the conditional accuracy functions show that (a) erroneous responses in condition 1 are confined to a single time bin, and (b) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+
+<!-- For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
 In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, right).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
-
-(ref:descr-fig1-caption) Means versus distributional shapes for six different simulated data set examples.
-The first second after stimulus onset is divided in ten bins of 100 ms.
-Time bin indicates the bin rank.
+Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.-->
+
+(ref:descr-fig1-caption) Mean performance versus distributional analyses.
+(A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
+(B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
+The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms.
 The first bin is (0,100], the last bin is (900,1000].
-For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
-Insets show mean response time per condition.
+Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
+The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
+Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
+
+<!-- For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
+Insets show mean response time per condition.-->
 
 ```{r plot1, fig.cap = "(ref:descr-fig1-caption)", out.width="90%"}
-knitr::include_graphics("../sims/figures/haz_inset_facet.jpeg")
+knitr::include_graphics("../sims/figures/Fig1_revision1.jpeg")
 ```
@@
 Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in time bin 2) and is gone when the waiting time reaches 500 ms or more.
-In contrast, in example 5, the effect starts around 300 ms and is gone by 700 ms.
-And in example 6, the effect reverses between 500 and 600 ms.
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+
+For many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
+
+<!-- For example, the distributions in Figure 1 show that the effect starts between 400 and 500 ms (in time bin 5) and is gone when the waiting time reaches 700 ms or more.
+
 What kind of theory or theories could account for such effects?
@@
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
-As we show later using published examples, for many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-
+-->
+
+## 1.2 Aims and structure of the paper
+
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
+Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
+
+<!-- We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
+Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.-->
+
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
 Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @Halley1693; @williammatthewmakehamLawMortalityConstruction1860].
-The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020].
-
-## 1.2 Aims and structure of the paper
-
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
-We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020; XXXXX].
 
@@
 Indeed, our primary aim here is to introduce the set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
-
-We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
+In section A of the Supplemental Material we explain how the data from typical RT tasks (detection, discrimination or categorization, bistable perception) are treated in EHA.
+
+We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
@@
 
-In Tutorial 1a, we illustrate how to process or "wrangle" a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
+<!--In Tutorial 1a, we illustrate how to process or "wrangle" a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
 The descriptive statistics are plotted, and we comment on their interpretation.
@@
 In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
-We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
-
-In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020] on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multilevel regression modelling steps.
-Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.-->.
 
@@
 In EHA, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+
 Since our focus here is on hazard models that use discrete time units, we describe that approach.
@@
 The equations that define both of these functions are reported in section A of the Supplemental Material.
+
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
+In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
@@
 
-## 2.2 Benefits of event history analysis
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
-The benefits include:
-
-1. Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
-
-2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
-
-3.
-EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
-
-4.
-Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-[@allisonSurvivalAnalysisUsing2010].
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
-
-5.
-EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
-
 ## 2.3 Event history analysis in the context of experimental psychology
@@
 ```{r plot3, fig.cap = "(ref:descr-plot3-caption)", out.width="80%"}
-knitr::include_graphics("../sims/figures/haz_acc_single.jpeg")
+#knitr::include_graphics("../sims/figures/haz_acc_single.jpeg")
 ```
@@
 
-All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive:
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: https://doi.org/10.31234/osf.io/57bh6
 

### 2025-06-16 manuscript/manuscript.tex
added link to psycharxiv
Initial revision of 1. Introduction

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \def\languageshorthands#1{}
+\ifLuaTeX
+  \usepackage[english]{selnolig} % disable illegal ligatures
+\fi
 % Manuscript styling
@@
 \raggedbottom
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 In experimental psychology, it is standard practice to analyse response times (RTs), saccade latencies, and fixation durations by calculating average performance across a series of trials.
-Such mean-average comparisons have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how long it lasts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
+Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
 Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 
-As a simple illustration, Figure 1 shows the results of several simulated RT data sets, which show how mean-average comparisons between two conditions can conceal the shape of the underlying RT distributions.
-For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, right).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.
+As a simple illustration, Figure 1 shows how comparing means between two conditions conceal the shapes of the underlying RT and accuracy distributions.
+We simulated a RT + accuracy data set for a single subject who performed 200 trials (i.e., repeated measurements) in each of two conditions.
+For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500{]} or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
+Similarly, while this subject makes less errors in condition 1 (86\% accuracy) compared to condition 2 (64\% accuracy), the conditional accuracy functions show that (a) erroneous responses in condition 1 are confined to a single time bin, and (b) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../sims/figures/haz_inset_facet}
-
-}
-
-\caption{Means versus distributional shapes for six different simulated data set examples.
-The first second after stimulus onset is divided in ten bins of 100 ms.
-Time bin indicates the bin rank.
+{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../sims/figures/Fig1_revision1}
+
+}
+
+\caption{Mean performance versus distributional analyses.
+(A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
+(B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
+The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms.
 The first bin is (0,100{]}, the last bin is (900,1000{]}.
-For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
-Insets show mean response time per condition.}\label{fig:plot1}
+Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
+The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
+Error bars represent +/- 1 standard error of the mean (A) or proportion (B).}\label{fig:plot1}
 \end{figure}
@@
 Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between different theoretical accounts of psychological and/or brain-based processes.
-For example, the distributions in example 4 show that the effect starts between 100 and 200 ms (in time bin 2) and is gone when the waiting time reaches 500 ms or more.
-In contrast, in example 5, the effect starts around 300 ms and is gone by 700 ms.
-And in example 6, the effect reverses between 500 and 600 ms.
-What kind of theory or theories could account for such effects?
-Are there new auxiliary assumptions that theories need to adopt?
-And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
-As we show later using published examples, for many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the early emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+
+For many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
+
+\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
+
+In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
+survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
+Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
+
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
 From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
 Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1693; Makeham, 1860).
-The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis, Schmidt, et al., 2020).
-
-\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
-
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
-We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.
+The same logic has been applied to psychological time-to-event data, as previously demonstrated {[}Panis, Schmidt, et al. (2020); XXXXX{]}.
 
@@
 Indeed, our primary aim here is to introduce the set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
-
-We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
+In section A of the Supplemental Material we explain how the data from typical RT tasks (detection, discrimination or categorization, bistable perception) are treated in EHA.
+
+We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
 The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
@@
 
-In Tutorial 1a, we illustrate how to process or ``wrangle'' a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
-The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
-
-In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
-We perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
-In Tutorial 2b we fit Bayesian multilevel regression models to \emph{timed} accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
-
-In Tutorial 3a, we shortly illustrate how to fit similar multilevel regression models for RT data in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to EHA.
-In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
-
-In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
-We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.
-
-In summary, even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multilevel regression modelling steps.
-Therefore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.
+.
 
@@
 The equations that define both of these functions are reported in section A of the Supplemental Material.
+
 The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
+In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
 In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
@@
 
-\subsection{2.2 Benefits of event history analysis}\label{benefits-of-event-history-analysis}
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
-The benefits include:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
-\item
-  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
-\item
-  EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
-\item
-  Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
-\item
-  EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
-\end{enumerate}
-
 \subsection{2.3 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
@@
 Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_acc_single}
-
-}
-
-\caption{Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 1000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
-Error bars represent +/- 1 standard error of the respective proportion.
-Bin width equals 100 ms.}\label{fig:plot3}
-\end{figure}
 
@@
 This data set is used when fitting conditional accuracy models (Tutorials 2b and 3b).
+
+\begin{verbatim}
+## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
+## Use 'xfun::attr2()' instead.
+## See help("Deprecated")
+\end{verbatim}
 
@@
 Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
+
+\begin{verbatim}
+## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
+## Use 'xfun::attr2()' instead.
+## See help("Deprecated")
+\end{verbatim}
 
@@
 Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
+
+\begin{verbatim}
+## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
+## Use 'xfun::attr2()' instead.
+## See help("Deprecated")
+\end{verbatim}
 
@@
 
-All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive:
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: \url{https://doi.org/10.31234/osf.io/57bh6}
 
@@
 McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
-
-\bibitem[\citeproctext]{ref-meyerModernMentalChronometry1988}
-Meyer, D. E., Osman, A. M., Irwin, D. E., \& Yantis, S. (1988). Modern mental chronometry. \emph{Biological Psychology}, \emph{26}(1-3), 3--67. \url{https://doi.org/10.1016/0301-0511(88)90013-0}
 

### 2025-06-18 manuscript/manuscript.Rmd
add section 2.1

@@
 For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500] or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
-Similarly, while this subject makes less errors in condition 1 (86% accuracy) compared to condition 2 (64% accuracy), the conditional accuracy functions show that (a) erroneous responses in condition 1 are confined to a single time bin, and (b) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+Similarly, while this subject makes less errors in condition 1 (86% accuracy) compared to condition 2 (64% accuracy), the conditional accuracy functions show that (a) the effect is present only for responses emitted before 400 ms, (b) erroneous responses in condition 1 are confined to a single time bin, and (c) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+
+Why does this matter for research in psychology?
+Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+
+For many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
 
@@
 Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
+Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
+For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (< 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (>500 ms) are controlled by the target stimulus.
 
@@
 ```
-
-Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-
-For many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
 
@@
 
+To apply EHA, one must be able to:
+
+1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+
+2. define time point zero (e.g., target stimulus onset, fixation onset, etc.);
+
+3. measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+
+## 2.1 Single, repeatable, and recurrent events
+
+While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
+For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus that (s)he has to detect by pressing a button within a certain time window (e.g., 1 second from target onset).
+The single time-to-detection is thus repeatedly measured across multiple trials.
+In the parlance of EHA, in each trial the participant is "at risk" for response occurrence as long as the response has not occurred yet.
+
+In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window.
+This is known as a "competing risks" situation, because in each trial, the participant can transition from an idle state to either a correct response state or an incorrect response state.
+
+In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
+In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period (2 minutes).
+We do not analyse recurrent events in this tutorial.
+More information about recurrent events analysis can be found in REF and REF...
+
+In secion A of the Supplementary Material we visualize the types of time-to-event data that can be obtained in these RT tasks.
+
+## 2.2 Right censoring vs data trimming
+
+## 2.3 Discrete vs continuous time units
+
+All measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window, 
+- interest in the shape of the hazard function <-> Cox
+- non-smooth functions <-> parametric
+- definition hazard prob and logistic regression
+
+## 2.4 Discrete-time Hazard functions and conditional accuracy fun
+
+def hazard
+
+def survival
+
+def cumulative distribution
+
+def prob mass function
+
+def conditional accuracz function
+
+## 2.5 Number of time bins, repeated measures, and samples
+
+## 2.6 Bayesian vs. frequentist approach
+
+-------------------
+
 We recommend several excellent textbooks for a comprehensive background context to EHA [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010] and for a more general introduction to understanding regression equations [@gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
@@
 ## 2.1 Basic features of event history analysis
-
-To apply EHA, one must be able to:
-
-1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
-
-2. define time point zero (e.g., target stimulus onset, fixation onset, etc.);
-
-3. measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
 

### 2025-06-18 manuscript/manuscript.tex
add section 2.1

@@
 
+Why does this matter for research in psychology?
+Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+
+For many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
+
 \begin{figure}[H]
@@
 
-Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the early emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-
-For many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, as well as adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
-
 \subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
@@
 
+To apply EHA, one must be able to:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+\item
+  define time point zero (e.g., target stimulus onset, fixation onset, etc.);
+\item
+  measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+\end{enumerate}
+
+\subsection{2.1 Single, repeatable, and recurrent events}\label{single-repeatable-and-recurrent-events}
+
+While people can die only once, in experimental RT tasks the events of interest - button presses, saccade onsets, etc.
+- are typically repeatable.
+For example, in a one-button detection task the participant is presented in each trial with a faint target stimulus that (s)he has to detect by pressing a button within a certain time (e.g., 1 second).
+
+Detection: single event, repeatable: hazard function + clustering
+Discrimination: Two events, repeatable: competing risks
+Bistable perception task: Two events, repeatable, and recurrent
+
+\subsection{2.2 Right censoring vs data trimming}\label{right-censoring-vs-data-trimming}
+
+\subsection{2.3 Discrete vs continuous time units}\label{discrete-vs-continuous-time-units}
+
+All measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window,
+- interest in the shape of the hazard function \textless-\textgreater{} Cox
+- non-smooth functions \textless-\textgreater{} parametric
+- definition hazard prob and logistic regression
+
+\subsection{2.4 Discrete-time Hazard functions vs cumulative distribution functions}\label{discrete-time-hazard-functions-vs-cumulative-distribution-functions}
+
+def hazard
+
+def survival
+
+def cumulative distribution
+
+def prob mass function
+
+def conditional accuracz function
+
+\subsection{2.5 Number of time bins, repeated measures, and samples}\label{number-of-time-bins-repeated-measures-and-samples}
+
+\subsection{2.6 Bayesian vs.~frequentist approach}\label{bayesian-vs.-frequentist-approach}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
 We recommend several excellent textbooks for a comprehensive background context to EHA (Allison, 2010; Singer \& Willett, 2003) and for a more general introduction to understanding regression equations (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
@@
 \subsection{2.1 Basic features of event history analysis}\label{basic-features-of-event-history-analysis}
-
-To apply EHA, one must be able to:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
-\item
-  define time point zero (e.g., target stimulus onset, fixation onset, etc.);
-\item
-  measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
-\end{enumerate}
 

### 2025-06-18 manuscript/manuscript.Rmd
update 2.2 and 2.3

@@
 In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
+survival analysis, hazard analysis, duration analysis, failure time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
 Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
@@
 Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @Halley1693; @williammatthewmakehamLawMortalityConstruction1860].
-The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020; XXXXX].
+The same logic can be applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020; XXXXX].
 
@@
 Indeed, our primary aim here is to introduce the set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
-In section A of the Supplemental Material we explain how the data from typical RT tasks (detection, discrimination or categorization, bistable perception) are treated in EHA.
 
 We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data measured in RT tasks.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
@@
 
-To apply EHA, one must be able to:
-
-1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+EHA is a class of statistical methods to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010). To apply EHA, one must be able to:
+
+1.
+define an event of interest that represents a qualitative change - a transition from one discrete state to another - that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
 
@@
 While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
-For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus that (s)he has to detect by pressing a button within a certain time window (e.g., 1 second from target onset).
-The single time-to-detection is thus repeatedly measured across multiple trials.
-In the parlance of EHA, in each trial the participant is "at risk" for response occurrence as long as the response has not occurred yet.
+For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset).
+In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 1 second long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, and the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs.
 
 In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window.
-This is known as a "competing risks" situation, because in each trial, the participant can transition from an idle state to either a correct response state or an incorrect response state.
+In the world of EHA, this is known as a "competing risks" situation, because in each trial the participant can transition from an idle state to either a "correct response" state or an "incorrect response" state.
 
 In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
-In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period (2 minutes).
-We do not analyse recurrent events in this tutorial.
+In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.
+
+In section A of the Supplemental Material we visualize the types of time-to-event data that are obtained in these typical RT tasks (detection, discrimination or categorization, bistable perception).
+Note that we do not analyse recurrent events in this tutorial.
 More information about recurrent events analysis can be found in REF and REF...
 
-In secion A of the Supplementary Material we visualize the types of time-to-event data that can be obtained in these RT tasks.
-
-## 2.2 Right censoring vs data trimming
+## 2.2 Right censoring versus data trimming
+
+What do you do with trials in which no response occurs during the observation period?
+EHA treats such trials as *right-censored* observations on the variable RT, because all we know is that RT is greater than some value.
+Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
+For example, in the one-button detection task example from above, all trials have a *censoring time* of 1 second, but some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+
+EHA can deal in a straight-forward fashion with right-censored time-to-event data.
+In contrast, experimental psychologists are used to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Discarding data can introduce biases, however.
 
@@
 
-All measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window, 
-- interest in the shape of the hazard function <-> Cox
-- non-smooth functions <-> parametric
-- definition hazard prob and logistic regression
-
-## 2.4 Discrete-time Hazard functions and conditional accuracy fun
+All man-made measurements of duration are discrete in nature.
+However, when the temporal resolution is high relative to the duration of the observation window, researchers typically treat time as continuous.
+RT data can thus be analysed using continuous-time EHA methods which use the exact event times, including parametric models (e.g., an exponential hazard model, a Weibull hazard model, a lognormal hazard model) and the popular Cox regression model ().
+
+However, in this tutorial we focus on discrete-time methods for three reasons:
+First, we are interested in studying the shape of the hazard function (Cox regression ignores this and only tests the effects of covariates); 
+Second, empirical hazard and conditional accuracy functions from certain RT tasks (e.g., interference tasks; Figure 1B) can show abrupt changes in their shape (parametric methods assume smooth distributions), and the shape of the hazard function in many experimental tasks is still unknown (parametric methods assume well-defined probability distributions);
+Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are already familiar.
+
+In sum, due to their simplicity and flexibility, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA and learn to apply EHA, even though continuous-time methods might be more suited in certain situations.
+
+## 2.4 Discrete-time hazard functions and conditional accuracy functions
+
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B), let RT be a discrete random variable denoting the *rank* of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the detection response in trial 1 might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+
+Interval-censored... +
+
+Discrete-time EHA focuses on the discrete-time hazard function of event occurrence
+
+and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in section A of the Supplemental Material.
 
@@
 
-## 2.5 Number of time bins, repeated measures, and samples
+The discrete-time hazard function for individual i in trial j of condition k gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+
+In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
+
+ life table...
+
+## 2.5 Number of samples, repeated measures, time bins
+
+In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
+
+Power
+
+Number of time bins?
+
+bin width
 
 ## 2.6 Bayesian vs. frequentist approach
+
+To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy.
+Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
+Note that time-varying covariates are not covered in this tutorial.
+
+Heterogeneity
+
+fitting problems
 
@@
 Since our focus here is on hazard models that use discrete time units, we describe that approach.
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in section A of the Supplemental Material.
-
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
 

### 2025-06-18 manuscript/manuscript.tex
update 2.2 and 2.3

@@
 For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500{]} or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
-Similarly, while this subject makes less errors in condition 1 (86\% accuracy) compared to condition 2 (64\% accuracy), the conditional accuracy functions show that (a) erroneous responses in condition 1 are confined to a single time bin, and (b) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+Similarly, while this subject makes less errors in condition 1 (86\% accuracy) compared to condition 2 (64\% accuracy), the conditional accuracy functions show that (a) the effect is present only for responses emitted before 400 ms, (b) erroneous responses in condition 1 are confined to a single time bin, and (c) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
 
@@
 The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
-Error bars represent +/- 1 standard error of the mean (A) or proportion (B).}\label{fig:plot1}
+Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
+Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
+For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (\textless{} 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (\textgreater500 ms) are controlled by the target stimulus.}\label{fig:plot1}
 \end{figure}
@@
 In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
+survival analysis, hazard analysis, duration analysis, failure time analysis, and transition analysis (Singer \& Willett, 2003).
 Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
@@
 Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1693; Makeham, 1860).
-The same logic has been applied to psychological time-to-event data, as previously demonstrated {[}Panis, Schmidt, et al. (2020); XXXXX{]}.
+The same logic can be applied to psychological time-to-event data, as previously demonstrated {[}Panis, Schmidt, et al. (2020); XXXXX{]}.
 
@@
 Indeed, our primary aim here is to introduce the set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
-In section A of the Supplemental Material we explain how the data from typical RT tasks (detection, discrimination or categorization, bistable perception) are treated in EHA.
 
 We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data measured in RT tasks.
 Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
@@
 
-To apply EHA, one must be able to:
+EHA is a class of statistical methods to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010). To apply EHA, one must be able to:
 
@@
 \item
-  define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+  define an event of interest that represents a qualitative change - a transition from one discrete state to another - that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
 \item
@@
 
-While people can die only once, in experimental RT tasks the events of interest - button presses, saccade onsets, etc.
-- are typically repeatable.
-For example, in a one-button detection task the participant is presented in each trial with a faint target stimulus that (s)he has to detect by pressing a button within a certain time (e.g., 1 second).
-
-Detection: single event, repeatable: hazard function + clustering
-Discrimination: Two events, repeatable: competing risks
-Bistable perception task: Two events, repeatable, and recurrent
-
-\subsection{2.2 Right censoring vs data trimming}\label{right-censoring-vs-data-trimming}
+While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
+For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset).
+In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 1 second long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, and the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs.
+
+In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window.
+In the world of EHA, this is known as a ``competing risks'' situation, because in each trial the participant can transition from an idle state to either a ``correct response'' state or an ``incorrect response'' state.
+
+In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
+In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.
+
+In section A of the Supplemental Material we visualize the types of time-to-event data that are obtained in these typical RT tasks (detection, discrimination or categorization, bistable perception).
+Note that we do not analyse recurrent events in this tutorial.
+More information about recurrent events analysis can be found in REF and REF\ldots{}
+
+\subsection{2.2 Right censoring versus data trimming}\label{right-censoring-versus-data-trimming}
+
+What do you do with trials in which no response occurs during the observation period?
+EHA treats such trials as \emph{right-censored} observations on the variable RT, because all we know is that RT is greater than some value.
+Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
+For example, in the one-button detection task example from above, all trials have a \emph{censoring time} of 1 second, but some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+
+EHA can deal in a straight-forward fashion with right-censored time-to-event data.
+In contrast, experimental psychologists are used to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Discarding data can introduce biases, however.
 
@@
 
-All measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window,
-- interest in the shape of the hazard function \textless-\textgreater{} Cox
-- non-smooth functions \textless-\textgreater{} parametric
-- definition hazard prob and logistic regression
-
-\subsection{2.4 Discrete-time Hazard functions vs cumulative distribution functions}\label{discrete-time-hazard-functions-vs-cumulative-distribution-functions}
+All man-made measurements of duration are discrete in nature.
+However, when the temporal resolution is high relative to the duration of the observation window, researchers typically treat time as continuous.
+RT data can thus be analysed using continuous-time EHA methods which use the exact event times, including parametric models (e.g., an exponential hazard model, a Weibull hazard model, a lognormal hazard model) and the popular Cox regression model ().
+
+However, in this tutorial we focus on discrete-time methods for three reasons:
+First, we are interested in studying the shape of the hazard function (Cox regression ignores this and only tests the effects of covariates);
+Second, empirical hazard and conditional accuracy functions from certain RT tasks (e.g., interference tasks; Figure 1B) can show abrupt changes in their shape (parametric methods assume smooth distributions), and the shape of the hazard function in many experimental tasks is still unknown (parametric methods assume well-defined probability distributions);
+Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are familiar.
+
+Thus, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA, even though continuous-time methods might be preferred in some situations.
+
+\subsection{2.4 Discrete-time hazard functions and conditional accuracy functions}\label{discrete-time-hazard-functions-and-conditional-accuracy-functions}
+
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+
+Interval-censored\ldots{} +
+
+Discrete-time EHA focuses on the discrete-time hazard function of event occurrence
+
+and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in section A of the Supplemental Material.
 
@@
 
-\subsection{2.5 Number of time bins, repeated measures, and samples}\label{number-of-time-bins-repeated-measures-and-samples}
+The discrete-time hazard function for individual i in trial j of condition k gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+
+In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
+In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
+
+life table\ldots{}
+
+\subsection{2.5 Number of samples, repeated measures, time bins}\label{number-of-samples-repeated-measures-time-bins}
+
+In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
+
+Power
+
+Number of time bins?
 
 \subsection{2.6 Bayesian vs.~frequentist approach}\label{bayesian-vs.-frequentist-approach}
+
+To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy.
+Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
+Note that time-varying covariates are not used in this tutorial.
+
+Heterogeneity
+
+fitting problems
 
@@
 Since our focus here is on hazard models that use discrete time units, we describe that approach.
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in section A of the Supplemental Material.
-
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
 

### 2025-06-18 manuscript/manuscript.Rmd
add section 2.4

@@
 
+Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
+For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (< 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (>500 ms) are controlled by the target stimulus.
+
 For many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
@@
 Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
-Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
-For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (< 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (>500 ms) are controlled by the target stimulus.
 
@@
 For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset).
-In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 1 second long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, and the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs.
+In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 1 second long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, and the individual always starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs.
 
@@
 For example, the detection response in trial 1 might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-
-Interval-censored... +
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence
-
-and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in section A of the Supplemental Material.
-
-def hazard
-
-def survival
-
-def cumulative distribution
-
-def prob mass function
-
-def conditional accuracz function
-
-The discrete-time hazard function for individual i in trial j of condition k gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-
-In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
-
- life table...
-
-## 2.5 Number of samples, repeated measures, time bins
-
-In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
-
-Power
-
-Number of time bins?
-
-bin width
-
-## 2.6 Bayesian vs. frequentist approach
-
-To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy.
+Thus, the RT data are interval-censored, because we only use the information that a<RT<=b when the response occurs in time bin (a,b].
+
+While experimental psychologists are familiar with the cumulative distribution function or F(t) = P(RT <= t) and the probability mass function or P(t) = P(RT = t), discrete-time EHA focuses on the discrete-time hazard function of event occurrence:
+
+\noindent h(t) = P(RT = t| RT $\geq$ t)   \hfill  (1)
+
+\noindent and the discrete-time survivor function:
+
+\noindent S(t) = P(RT $>$ t) = 1 - F(t) =  [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)]  \hfill  (2)
+
+The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. 
+In other words, it reflects the instantaneous risk that the response occurs in bin t, given that it has not yet occurred in one of the prior bins.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event does not occur before the endpoint of bin t.
+As a result, only the hazard function conveys the risk of event occurrence associated with each bin, and ...
+suited for online tracking of performance..
+cfr mouse cursor movements....
+
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
+
+\noindent ca(t) = P(correct | RT = t)   \hfill  (5)
+
+\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
+
+As we will illustrate in Tutorials 1a and 1b, performing a descriptive EHA/SAT analysis by calculating the sample-based estimates of h(t), S(t) and ca(t) for each combination of participant and condition requires setting up a *life table*.
+Definition life table...
+
+## 2.5 Bayesian vs. frequentist approaches to regression
+
+To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy, i.e., perform inferential EHA/SAT analysis.
 Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
@@
 
-Heterogeneity
-
-fitting problems
-
+Heterogeneity -> Multilevel survival analysis: Methods, Models and Applications
+Austin 2017 !!
+
+fitting problems -> Bayesian
+
+## 2.6 Number of samples, repeated measures, time bins
+
+In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
+To test process models of cognition, .. advises  to use small-N designs, ... eACH SUBJECT REPLICATION UNIT
+
+Power IS A COMPLEX FUNCTION OF .... (REF A, REF B ON POWER WITH EXPONENTIAL)
+
+Number of time bins?
+
+bin width
 -------------------
@@
 
-## 2.1 Basic features of event history analysis
+<!--## 2.1 Basic features of event history analysis
 
@@
 For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
 
@@
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
-
-# 3. An overview of the general analytical workflow
+-->
+
+<!--# 3. An overview of the general analytical workflow
 
@@
 
-## 3.3 Implementation
+## 3.3 Implementation-->
+
+# 4. Tutorials
 
@@
 `r my_r_citation$pkgs` <!-- create footnote -->
-
-# 4. Tutorials
 

### 2025-06-18 manuscript/manuscript.tex
add section 2.4

@@
 
+Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
+For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (\textless{} 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (\textgreater500 ms) are controlled by the target stimulus.
+
 For many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
@@
 The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
-Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
-Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
-For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (\textless{} 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (\textgreater500 ms) are controlled by the target stimulus.}\label{fig:plot1}
+Error bars represent +/- 1 standard error of the mean (A) or proportion (B).}\label{fig:plot1}
 \end{figure}
@@
 For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset).
-In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 1 second long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, and the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs.
+In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 1 second long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, and the individual always starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs.
 
@@
 Second, empirical hazard and conditional accuracy functions from certain RT tasks (e.g., interference tasks; Figure 1B) can show abrupt changes in their shape (parametric methods assume smooth distributions), and the shape of the hazard function in many experimental tasks is still unknown (parametric methods assume well-defined probability distributions);
-Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are familiar.
-
-Thus, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA, even though continuous-time methods might be preferred in some situations.
+Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are already familiar.
+
+In sum, due to their simplicity and flexibility, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA and learn to apply EHA, even though continuous-time methods might be more suited in certain situations.
 
@@
 
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-
-Interval-censored\ldots{} +
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence
-
-and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in section A of the Supplemental Material.
-
-def hazard
-
-def survival
-
-def cumulative distribution
-
-def prob mass function
-
-def conditional accuracz function
-
-The discrete-time hazard function for individual i in trial j of condition k gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-
-In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
-
-life table\ldots{}
-
-\subsection{2.5 Number of samples, repeated measures, time bins}\label{number-of-samples-repeated-measures-time-bins}
-
-In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
-
-Power
-
-Number of time bins?
-
-\subsection{2.6 Bayesian vs.~frequentist approach}\label{bayesian-vs.-frequentist-approach}
+After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B), let RT be a discrete random variable denoting the \emph{rank} of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, the detection response in trial 1 might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+Thus, the RT data are interval-censored, because we only use the information that a\textless RT\textless=b when the response occurs in time bin (a,b{]}.
+
+While experimental psychologists are familiar with the cumulative distribution function or F(t) = P(RT \textless= t) and the probability mass function or P(t) = P(RT = t), discrete-time EHA focuses on the discrete-time hazard function of event occurrence:
+
+\noindent h(t) = P(RT = t\textbar{} RT \(\geq\) t) \hfill  (1)
+
+\noindent and the discrete-time survivor function:
+
+\noindent S(t) = P(RT \(>\) t) = 1 - F(t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} \hfill  (2)
+
+The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
+In other words, it reflects the instantaneous risk that the response occurs in bin t, given that it has not yet occurred in one of the prior bins.
+In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event does not occur before the endpoint of bin t.
+As a result, only the hazard function conveys the risk of event occurrence associated with each bin, and \ldots{} suited for online tracking of performance..
+cfr mouse cursor movements\ldots.
+
+For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
+
+\noindent ca(t) = P(correct \textbar{} RT = t) \hfill  (5)
+
+\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
+
+As we will illustrate in Tutorials 1a and 1b, performing a descriptive EHA/SAT analysis by calculating the sample-based estimates of h(t), S(t) and ca(t) for each combination of participant and condition requires setting up a \emph{life table}.
+Definition life table\ldots{}
+
+\subsection{2.5 Bayesian vs.~frequentist approaches to regression}\label{bayesian-vs.-frequentist-approaches-to-regression}
 
@@
 Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
-Note that time-varying covariates are not used in this tutorial.
+Note that time-varying covariates are not covered in this tutorial.
 
@@
 
+Multilevel survival analysis: Methods, Models and Applications
+Austin 2017 !!
+
 fitting problems
 
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+\subsection{2.6 Number of samples, repeated measures, time bins}\label{number-of-samples-repeated-measures-time-bins}
+
+In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
+To test process models of cognition, researchers tend to use small-N designs, \ldots{}
+
+Power
+
+Number of time bins?
+
+\subsection{bin width}\label{bin-width}
 
@@
 As such, we only supply regression equations in section D of the Supplemental Material.
-
-\subsection{2.1 Basic features of event history analysis}\label{basic-features-of-event-history-analysis}
-
-In EHA, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
-
-Since our focus here is on hazard models that use discrete time units, we describe that approach.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../sims/figures/haz_surv_single}
-
-}
-
-\caption{Discrete-time hazard and survivor functions.
-Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
-Error bars represent +/- 1 standard error of the respective proportion.
-While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t=5) = .44.
-Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.}\label{fig:plot2}
-\end{figure}
-
-\subsection{2.3 Event history analysis in the context of experimental psychology}\label{event-history-analysis-in-the-context-of-experimental-psychology}
-
-To make EHA more relevant to researchers studying cognitive psychology and cognitive neuroscience, in this section we provide a relevant worked example and consider implications that are relevant to that domain of research.
-
-\subsubsection{2.3.1 A worked example}\label{a-worked-example}
-
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a discrimination task.
-For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
-
-Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical choice RT example that is inspired by real data (Panis \& Schmidt, 2016), but simplified to make the main point clearer (Figure 3).
-In a standard priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
-The prime can then be congruent or incongruent with the target.
-
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel).
-These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed).
-During time bin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
-For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
-
-This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT and accuracy across trials.
-Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials, and are indexed by each recorded response.
-But such a conclusion is not supported when the effects are explored over a timeline.
-Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
-
-Unlocking the temporal states of cognitive processes can be revealing for theory development and the understanding of basic psychological processes.
-Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
-Do all participants show similar temporal states or are there individual differences?
-Do such individual differences extend to those individuals that have been diagnosed with some form of psychopathology?
-How do temporal states relate to brain-based mechanisms that might be studied using other methods from cognitive neuroscience?
-And how much of theory in cognitive psychology would be in need of revision if mean-average comparisons were supplemented with a temporal states approach?
-
-\subsubsection{2.3.2 Implications for designing experiments}\label{implications-for-designing-experiments}
-
-Performing EHA in experimental psychology has implications for how experiments are designed.
-Indeed, if trials are categorised as a function of when responses occur, then each time bin will only include a subset of the total number of trials.
-For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
-Those 100 trials must be distributed in some manner across the chosen number of bins.
-
-In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
-Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
-This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
-
-In contrast to the large-\emph{N} design that typically average across many participants without being able to scrutinize individual data patterns, small-\emph{N} designs retain crucial information about the data patterns of individual observers.
-This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
-Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
-
-\section{3. An overview of the general analytical workflow}\label{an-overview-of-the-general-analytical-workflow}
-
-Although the focus is on EHA/SAT, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
-
-\subsection{3.1 Data science workflow and descriptive statistics}\label{data-science-workflow-and-descriptive-statistics}
-
-We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
-In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.
-
-\subsection{3.2 Inferential statistical approach}\label{inferential-statistical-approach}
-
-Our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
-We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
-This means that wherever possible we include varying intercepts and slopes per participant.
-To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020).
-We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
-
-\subsection{3.3 Implementation}\label{implementation}
 
@@
 
-\bibitem[\citeproctext]{ref-bakerPowerContoursOptimising2021}
-Baker, D.
-H., Vilidaite, G., Lygo, F.
-A., Smith, A.
-K., Flack, T.
-R., Gouws, A.
-D., \& Andrews, T.
-J.
-(2021).
-Power contours: {Optimising} sample size and precision in experimental psychology and human neuroscience.
-\emph{Psychological Methods}, \emph{26}(3), 295--314.
-\url{https://doi.org/10.1037/met0000337}
-
-\bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
-Barr, D.
-J., Levy, R., Scheepers, C., \& Tily, H.
-J.
-(2013).
-Random effects structure for confirmatory hypothesis testing: {Keep} it maximal.
-\emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001.
-\url{https://doi.org/10.1016/j.jml.2012.11.001}
-
 \bibitem[\citeproctext]{ref-R-tinylabels}
@@
 Kay, M. (2024). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
-
-\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
-Kruschke, J.
-K., \& Liddell, T.
-M.
-(2018).
-The {Bayesian New Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power analysis from a {Bayesian} perspective.
-\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 178--206.
-\url{https://doi.org/10.3758/s13423-016-1221-4}
 
@@
 
-\bibitem[\citeproctext]{ref-smithSmallBeautifulDefense2018}
-Smith, P.
-L., \& Little, D.
-R.
-(2018).
-Small is beautiful: {In} defense of the small-{N} design.
-\emph{Psychonomic Bulletin \& Review}, \emph{25}(6), 2083--2101.
-\url{https://doi.org/10.3758/s13423-018-1451-8}
-
 \bibitem[\citeproctext]{ref-R-StanHeaders}
@@
 
-\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
-Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G.
-(2023).
-\emph{R for data science: Import, tidy, transform, visualize, and model data} (2nd edition).
-Beijing Boston Farnham Sebastopol Tokyo: O'Reilly.
-
 \bibitem[\citeproctext]{ref-R-dplyr}

### 2025-06-18 manuscript/Suppl_material/supplementary.Rmd
update 2.2 and 2.3
update supps

@@
 This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
+
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+
 The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1.
@@
 EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived "dips" that could signal the presence of active top-down respone inhibition processes [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
 

### 2025-06-19 manuscript/manuscript.tex
small update

@@
 
-To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy.
+To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy, i.e., perform inferential EHA/SAT analysis.
 Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
@@
 
-Heterogeneity
-
-Multilevel survival analysis: Methods, Models and Applications
+Heterogeneity -\textgreater{} Multilevel survival analysis: Methods, Models and Applications
 Austin 2017 !!
 
-fitting problems
+fitting problems -\textgreater{} Bayesian
 
@@
 In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
-To test process models of cognition, researchers tend to use small-N designs, \ldots{}
-
-Power
+To test process models of cognition, .. advises to use small-N designs, \ldots{} eACH SUBJECT REPLICATION UNIT
+
+Power IS A COMPLEX FUNCTION OF \ldots. (REF A, REF B ON POWER WITH EXPONENTIAL)
 
@@
 
+\section{4. Tutorials}\label{tutorials}
+
 We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024b), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-
-\section{4. Tutorials}\label{tutorials}
 

### 2025-06-20 manuscript/reviews/ResponseReview1.Rmd
updated packages and re-organised folders

--- 2025-06-14 manuscript/ResponseReview1.Rmd
@@
 >
-> Thank you for submitting your Tutorial (AMPPS-24-0232) entitled "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows" to Advances in Methods and Practices in Psychological Science (AMPPS).
+> Thank you for submitting your Tutorial  entitled "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows" .
 I have now received input from three experts in the field, and you will find the reviews below this message and attached.
@@
 Prior to considering the reviewers' input, I read your paper to form an independent opinion.
-I share the positive sentiments about this work and invite you to revise and resubmit your manuscript for further consideration at AMPPS.
+I share the positive sentiments about this work and invite you to revise and resubmit your manuscript for further consideration.
 
@@
 
-> To submit your revision, log into http://mc.manuscriptcentral.com/ampps and enter your Author Center, where you will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.
-
-> IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please delete any redundant or outdated files before completing the submission.
-
-> Once again, thank you for submitting your manuscript to AMPPS and I look forward to receiving your revision.
-
 > Sincerely,
 
-> David A. Sbarra, Ph.D.
-> Editor, Advances in Methods and Practices in Psychological Science
-
-We tried to focus on accessibility in the revision of our Tutorial AMPPS-24-0232.
+We tried to focus on accessibility in the revision of our Tutorial.
 

### 2025-06-20 manuscript/manuscript.Rmd
Feedback and edits on the updated manuscript.

--- 2025-06-18 manuscript/manuscript.Rmd
@@
 # R references
-r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
-my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) 
+# r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
+# my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)
 ```
@@
 
-As a simple illustration, Figure 1 shows how comparing means between two conditions conceal the shapes of the underlying RT and accuracy distributions.
-We simulated a RT + accuracy data set for a single subject who performed 200 trials (i.e., repeated measurements) in each of two conditions.
-For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500] or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
-Similarly, while this subject makes less errors in condition 1 (86% accuracy) compared to condition 2 (64% accuracy), the conditional accuracy functions show that (a) the effect is present only for responses emitted before 400 ms, (b) erroneous responses in condition 1 are confined to a single time bin, and (c) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+As a simple illustration, Figure 1 summarises simulated data that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
+Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
 
 Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-
-For many psychological questions, such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+For many psychological questions, the estimation of such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes and by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
 Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
 
-<!-- For instance, in examples 1-3, mean RT is always comparable between two conditions, while the distributions differ (Figure 1, left).
-In contrast, in examples 4-6, mean RT is lower in condition 2 compared to condition 1, but the RT distributions differ in each case (Figure 1, right).
-Therefore, a comparison of means would lead to a similar conclusion in examples 1-3, as well as examples 4-6, whereas a comparison of the distributions would lead to a different conclusion in every case.-->
-
-(ref:descr-fig1-caption) Mean performance versus distributional analyses.
+(ref:descr-fig1-caption) Simulated data showing mean performance versus distributional analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
-Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
-For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (< 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (>500 ms) are controlled by the target stimulus.
-
-<!-- For our purposes here, it is enough to know that the distributions plotted represent the probability of an event occurring in that time bin, given that it has not yet occurred.
-Insets show mean response time per condition.-->
 
@@
 
-<!-- For example, the distributions in Figure 1 show that the effect starts between 400 and 500 ms (in time bin 5) and is gone when the waiting time reaches 700 ms or more.
+## 1.2 Aims
 
-What kind of theory or theories could account for such effects?
-Are there new auxiliary assumptions that theories need to adopt?
-And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
--->
+Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
 
-## 1.2 Aims and structure of the paper
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+Set within this context, our overall aim is to introduce a set of tutorials, which explain **how** to do such analyses in the context of experimental psychology, rather than repeat in any detail **why** you may do them.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
 
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
-Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
+## 1.3 Structure
 
-<!-- We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience.
-Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.-->
+In what follows, the paper is organised in three main sections.
+In Section 2, we provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper and why such an approach might be relevant for research in experimental psychology.
+In Section 3, we outline a series of tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
+In Section 4, we discuss the strengths and weaknesses of the approach for researchers in experimental psychology.
 
-Even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
-From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died [e.g., @Halley1693; @williammatthewmakehamLawMortalityConstruction1860].
-The same logic has been applied to psychological time-to-event data, as previously demonstrated [@panisAnalyzingResponseTimes2020; XXXXX].
+# 2. What is event history analaysis and why is it relevant to research in experimental psychology?
 
-We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @allisonSurvivalAnalysisUsing2010].
-Indeed, our primary aim here is to introduce the set of tutorials, which explain **how** to do such analyses, rather than repeat in any detail **why** you may do them.
-In section A of the Supplemental Material we explain how the data from typical RT tasks (detection, discrimination or categorization, bistable perception) are treated in EHA.
+## 2.1 A brief introduction to event history analysis
 
-We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/    Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data.
-Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
-Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja package [@pap2024], which makes it computationally reproducible, in terms of the underlying data and figures.
+EHA is a statistical approach to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
+In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
+For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
+We also supply all relevant regression equations in section D of the Supplemental Material.
 
-<!--In Tutorial 1a, we illustrate how to process or "wrangle" a previously published RT + accuracy data set to calculate descriptive statistics when there is one independent variable.
-The descriptive statistics are plotted, and we comment on their interpretation.
-In Tutorial 1b we provide a generalisation of this approach to illustrate how one can calculate the descriptive statistics when using a more complex design, such as when there are two independent variables.
+### 2.1.1 Terminolgy and minimum requirements for EHA
 
-In Tutorial 2a, we illustrate how one can fit Bayesian multilevel regression models to RT data using the R package brms.
-We perform prior predictive checks, compare models, and interpret the plots of the predicted hazard functions for the selected model, and the posterior distributions of our contrasts of interest.
-In Tutorial 2b we fit Bayesian multilevel regression models to *timed* accuracy data to perform a micro-level speed-accuracy tradeoff (SAT) analysis, which complements the EHA of RT data for choice RT data.
+To avoid possible confusion in terminology used, it is worth noting that EHA is frequently known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer & Willett, 2003).
+In this paper, we choose to use the term EHA throughout.
 
-In Tutorial 3a, we shortly illustrate how to fit similar multilevel regression models for RT data in a frequentist framework using the R package lme4.
-We then briefly compare and contrast these inferential frameworks when applied to EHA.
-In Tutorial 3b, we illustrate how to perform the SAT analysis in a frequentist framework.
+In terms of minimum requirements to apply EHA, one must be able to:
 
-In tutorial 4, we illustrate one approach to planning how much data to collect in an experiment using EHA.
-We use data simulation techniques to vary sample size and trial count per condition until a certain degree of statistical power or precision is reached.-->.
-
-# 2. A brief introduction to event history analysis
-
-To apply EHA, one must be able to:
-
-1. define an event of interest that represents a qualitative change that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
+1.
+define an event of interest that represents a qualitative change - a transition from one discrete state to another - that can be situated in time (e.g., a button press, a saccade onset, a fixation offset, etc.);
 
@@
 
-## 2.1 Single, repeatable, and recurrent events
+### 2.1.2 Types of EHA
 
-While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
-For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus that (s)he has to detect by pressing a button within a certain time window (e.g., 1 second from target onset).
-The single time-to-detection is thus repeatedly measured across multiple trials.
-In the parlance of EHA, in each trial the participant is "at risk" for response occurrence as long as the response has not occurred yet.
+There are different types of EHA.
+For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. 
+As a lab, and mainly for practical reasons, we have much more experience using discrete time EHA, and that is the approach that we describe and focus on in this paper. 
+This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable. 
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well. 
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than ~200 trials per condition per experiment.
+Moreover, as indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
 
-In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window.
-This is known as a "competing risks" situation, because in each trial, the participant can transition from an idle state to either a correct response state or an incorrect response state.
+To apply discrete time EHA, one divides time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+Then let t be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular experimental condition. 
+For example, a response on a single trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+One then calculates the discrete-time hazard function of event occurrence.
+The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in
+bin t, given that the event does not occur in previous bins. 
+In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet
+occurred in the past, i.e., in one of the prior bins.
 
-In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
-In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period (2 minutes).
-We do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in REF and REF...
-
-In secion A of the Supplementary Material we visualize the types of time-to-event data that can be obtained in these RT tasks.
-
-## 2.2 Right censoring vs data trimming
-
-## 2.3 Discrete vs continuous time units
-
-All measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window, 
-- interest in the shape of the hazard function <-> Cox
-- non-smooth functions <-> parametric
-- definition hazard prob and logistic regression
-
-## 2.4 Discrete-time Hazard functions and conditional accuracy fun
-
-def hazard
-
-def survival
-
-def cumulative distribution
-
-def prob mass function
-
-def conditional accuracz function
-
-## 2.5 Number of time bins, repeated measures, and samples
-
-## 2.6 Bayesian vs. frequentist approach
-
--------------------
-
-We recommend several excellent textbooks for a comprehensive background context to EHA [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010] and for a more general introduction to understanding regression equations [@gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
-Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
-Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in section D of the Supplemental Material.
-
-## 2.1 Basic features of event history analysis
-
-In EHA, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
-
-Since our focus here is on hazard models that use discrete time units, we describe that approach.
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins), let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the first response might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-
-Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2).
-The equations that define both of these functions are reported in section A of the Supplemental Material.
-
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous risk that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event occurs after bin t.
-In other words, the survivor function gives you for each time bin the likelihood that the event occurs in the future, i.e., in one of the subsequent time bins.
-
-(ref:descr-plot2-caption) Discrete-time hazard and survivor functions.
-Discrete time-to-event data were simulated for 200 trials of 1 experimental condition.
-Error bars represent +/- 1 standard error of the respective proportion.
-While the hazard function is the vehicle for inferring the time course of cognitive processes, the survival probability S(t-1) can help to qualify or provide context to the interpretation of the hazard probability h(t).
-For example, the high hazard of .60 = h(t=6) is experienced only by 44 percent of the trials, as S(t=5) = .44.
-Because the survivor function is a decreasing function of time, the error bars in later parts of the hazard function will always be wider and less precise compared to earlier parts.
-
-```{r plot2, fig.cap = "(ref:descr-plot2-caption)", out.width="80%"}
-knitr::include_graphics("../sims/figures/haz_surv_single.jpeg")
-```
-
-## 2.3 Event history analysis in the context of experimental psychology
-
-To make EHA more relevant to researchers studying cognitive psychology and cognitive neuroscience, in this section we provide a relevant worked example and consider implications that are relevant to that domain of research.
-
-### 2.3.1 A worked example
-
-In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a discrimination task.
-For example, a task may involve choosing between two response options with only one of them being correct.
-For such two-choice RT data, the discrete-time EHA of the RT data (hazard and survivor functions) can be extended with a discrete-time SAT analysis of the timed accuracy data.
-Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function, which gives you the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
+In those cases, one can also calculate the conditional accuracy function (Figure B).
+The conditional accuracy function gives you for each time bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
 
-Integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes.
-To illustrate, we consider a hypothetical choice RT example that is inspired by real data [@panisWhatShapingRT2016], but simplified to make the main point clearer (Figure 3).
-In a standard priming paradigm, there is a prime stimulus (e.g., an arrow pointing left or right) followed by a target stimulus (another arrow pointing left or right).
-The prime can then be congruent or incongruent with the target.
+## 2.2 Benefits of event history analysis for research in experimental psychology
 
-Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel).
-These results show that for short waiting times (< bin 6), responses always follow the prime (and not the target, as instructed).
-During time bin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition.
-For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here). 
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (section F of the Supplemental Material). 
+Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
 
-This joint pattern of results is interesting because it can provide meaningfully different conclusions about psychological processes compared to conventional analyses, such as computing mean-average RT and accuracy across trials.
-Mean-average RT would only represent the overall ability of cognition to overcome interference, on average, across trials.
-For instance, if mean-average RT was higher in incongruent than congruent trials, one may conclude that cognitive mechanisms that support interference control are working as expected across trials, and are indexed by each recorded response.
-But such a conclusion is not supported when the effects are explored over a timeline.
-Instead, the psychological conclusion is much more nuanced and suggests that multiple states start, stop and possibly interact over a particular temporal window.
+We highlight three benefits that we think are relevant to the domain of experimental psychology. 
+First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about cognition and theoretical development.
+As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
+For example, what kind of theory or theories could account for the effects reported in Figure 1B? 
+Are there new auxiliary assumptions that theories need to adopt? 
+And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
 
-(ref:descr-plot3-caption) Discrete-time hazard and conditional accuracy functions.
-Discrete time-to-event and conditional accuracy data were simulated for 1000 trials for each of two priming conditions (congruent and incongruent prime stimuli).
-Error bars represent +/- 1 standard error of the respective proportion.
-Bin width equals 100 ms.
+Second, compared to more conventional analytical approaches, EHA uses more of the data and deals with msising data differently.
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Discarding data can introduce biases, however.
+Rather than treat non-responses as missing data, EHA treats such trials as *right-censored* observations on the variable RT, because all we know is that RT is greater than some value. 
+Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data. 
+For example, if the censor time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+The use of right-censoring in EHA, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology.
 
-```{r plot3, fig.cap = "(ref:descr-plot3-caption)", out.width="80%"}
-#knitr::include_graphics("../sims/figures/haz_acc_single.jpeg")
-```
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks (Supp Fig X). 
+EHA can also be used with a range of common experimental manipualtions, such as stimulus-onset-asynchrony (Supp Fig X). 
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology.
 
-Unlocking the temporal states of cognitive processes can be revealing for theory development and the understanding of basic psychological processes.
-Possibly more importantly, however, is that it simultaneously opens the door to address many new and previously unanswered questions.
-Do all participants show similar temporal states or are there individual differences?
-Do such individual differences extend to those individuals that have been diagnosed with some form of psychopathology?
-How do temporal states relate to brain-based mechanisms that might be studied using other methods from cognitive neuroscience?
-And how much of theory in cognitive psychology would be in need of revision if mean-average comparisons were supplemented with a temporal states approach?
-
-### 2.3.2 Implications for designing experiments
+## 2.3 Implications for research design in experimental psychology
 
 Performing EHA in experimental psychology has implications for how experiments are designed. 
-Indeed, if trials are categorised as a function of when responses occur, then each time bin will only include a subset of the total number of trials.
-For example, let's consider an experiment where each participant performs 2 conditions and there are 100 trial repetitions per condition.
-Those 100 trials must be distributed in some manner across the chosen number of bins.
+More specifically, we consider two implications that researchers will need to consider when using discrete time EHA.
 
-In such experimental designs, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+First, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
 Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-*N* designs).
 This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
-
-In contrast to the large-*N* design that typically average across many participants without being able to scrutinize individual data patterns,  small-*N*  designs  retain  crucial  information  about  the  data  patterns of individual observers.
-This can be advantageous whenever participants differ systematically in their strategies or in the time courses of their effects, so that averaging them would lead to misleading data patterns.
 Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-*N* designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall [@bakerPowerContoursOptimising2021; @smithSmallBeautifulDefense2018].
 
-# 3. An overview of the general analytical workflow
-
-Although the focus is on EHA/SAT, we also want to briefly comment on broader aspects of our general analytical workflow, which relate more to data science and data analysis workflows.
-
-## 3.1 Data science workflow and descriptive statistics
-
-We perform data wrangling following tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
-In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
-In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.
-
-## 3.2 Inferential statistical approach
-
-Our lab adopts an estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
-We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
-This means that wherever possible we include varying intercepts and slopes per participant.
-To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2020].
-We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
-
-## 3.3 Implementation
-
-We used `r my_r_citation$r` for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
-
-`r my_r_citation$pkgs` <!-- create footnote -->
+Second, the width of each time bin will need to be determined. 
+For instance, in Figure 1B we chose 100ms in an arbitrary manner.
+In reality, however, bin width will need to be set by considering a number of factors simultabeously.
+For example, [[Sven - add a few sentences here]]
 
 # 4. Tutorials
+
+We used r my_r_citation$r for all reported analyses.
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 The next step is to set up the life table using our custom function setup_lt(), calculate the conditional accuracies using our custom function calc_ca(), add the ca(t) estimates to the life table using our custom function join_lt_ca(), and then plot the descriptive statistics using our custom function plot_eha().
-When creating the plots, some warning messages will likely be generated, like these:
-
-* Removed 2 rows containing missing values or values outside the scale range (`geom_line()`). 
-* Removed 2 rows containing missing values or values outside the scale range (`geom_point()`). 
-* Removed 2 rows containing missing values or values outside the scale range (`geom_segment()`).
-
-The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant. 
@@
 Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors [@singerAppliedLongitudinalData2003].
+
+In general, when fitting regression models, our lab adopts an estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
+We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
+This means that wherever possible we include varying intercepts and slopes per participant.
+To make inferences, we use two main approaches.
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2020].
+We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
 
@@
 
-## 4.5 Tutorial 3a: Fitting Frequentist hazard models
+<!-- ## 4.5 Tutorial 3a: Fitting Frequentist hazard models -->
 
-In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
+<!-- In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
 The general process is similar to that in Tutorial 2a, except that there are no priors to set.
+-->
 
-Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
+<!-- Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
 Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
 The resulting hazard model is called M1i_f with the appended "_f" denoting a frequentist model.
+-->
 
-In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i_f.
+<!-- In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i_f. -->
 
-(ref:descr-plot11-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
+<!-- (ref:descr-plot11-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
+-->
 
-```{r plot-comparison, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png")
-```
+<!-- ```{r plot-comparison, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'} -->
+<!-- knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png") -->
+<!-- ``` -->
 
-Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
+<!-- Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
 However, model M1i_f did not converge and resulted in a singular fit.
@@
 In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
+-->
 
-## 4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models
+<!-- ## 4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models -->
 
-In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
+<!-- In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
 To be concise, we only fit effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
@@
 Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
+-->
 
@@
 
-# 5. Discussion
+# 4. Discussion
 
@@
 
-## 5.1 What are the main use-cases of EHA for understanding cognition and brain function?
+## 4.1 What are the main use-cases of EHA for understanding cognition and brain function?
 
@@
 
-## 5.2 Model complexity versus interpretability
+## 4.2 Model complexity versus interpretability
 
@@
 
-## 5.3 Individual differences
-
-One important issue is that of possible individual differences in the overall location of the distribution, and the time course of psychological effects.
-For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
-These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 500 ms for simple detection tasks, 800 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
-Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response in the analysis time window), introducing a response deadline is recommended when designing RT experiments.
-Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard and conditional accuracy models.
-
-But even when using a response deadline, participants can differ qualitatively in the effects they display [see @panisHowCanWe2020].
-One way to deal with this is to describe and interpret the different patterns.
-Another way is to run a clustering algorithm on the individual hazard estimates across all bins and conditions.
-The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
-One might then exclude the smaller sub-group of participants before fitting a hazard model or consider the possibility that different cognitive processes may be at play during task performance across the different sub-groups.
-
-Another approach to deal with individual differences is Bayesian prevalence [@inceBayesianInferencePopulation2021], which is a from of small-*N* approach (Smith & Little, 2018).
-This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
-This approach allows one to quantify how typical or uncommon an observed effect is in the population, and the uncertainty around this estimate.
-
-## 5.4 Limitations
+## 4.3 Limitations
 
@@
 
-We did not discuss continuous-time EHA, nor continuous-time SAT analysis.
-As indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods.
-Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
-However, they require much more data to estimate the continuous-time hazard (rate) function well.
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than ~200 trials per condition per experiment.
-
-## 5.5 Extensions
-
-The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this button-press event constitutes a single transition between an "idle" state and a "responded" state.
-However, in certain situations, more than one event of interest might exist.
-For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before being absorbed into a final "death" state.
-When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize EHA to transitions between three or more states [@steeleGeneralMultilevelMultistate2004].
-Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@allisonSurvivalAnalysisUsing2010].
-For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
-Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.
-
-# 6. Conclusions
+# 5. Conclusions
 

### 2025-06-20 manuscript/manuscript.tex
Feedback and edits on the updated manuscript.

--- 2025-06-18 manuscript/manuscript.tex
@@
 
-As a simple illustration, Figure 1 shows how comparing means between two conditions conceal the shapes of the underlying RT and accuracy distributions.
-We simulated a RT + accuracy data set for a single subject who performed 200 trials (i.e., repeated measurements) in each of two conditions.
-For example, while this subject is 71 ms faster on average in condition 1 (481 ms) compared to condition 2 (552 ms), the corresponding hazard functions of response occurrence show that the effect starts in time period (400,500{]} or bin t=5, and is present in three consecutive time bins (i.e., for 300 ms).
-Similarly, while this subject makes less errors in condition 1 (86\% accuracy) compared to condition 2 (64\% accuracy), the conditional accuracy functions show that (a) the effect is present only for responses emitted before 400 ms, (b) erroneous responses in condition 1 are confined to a single time bin, and (c) the observed conditional accuracies (0, 1, 0.51, 0.48) are never even close to the mean accuracies.
+As a simple illustration, Figure 1 summarises simulated data that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
+Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
 
 Why does this matter for research in psychology?
-Compared to the aggregation of data across trials, a distributional approach offers the possibility to reveal the time course of psychological states.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higer in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-
-Note that the distributional shapes are inspired by published results from interference paradigms such as priming and cueing tasks (refs).
-For example, if the target stimulus at time zero is preceded by a prime stimulus that can be congruent (condition 1) or incongruent (condition 2) to the target, then the distributions show that (a) the fastest responses (\textless{} 400 ms) are triggered exclusively by the prime, (b) between 400 and 500 ms, the prime-triggered response tendency is actively inhibited (leading to conditional accuracies close to .5), and (c) the slowest responses (\textgreater500 ms) are controlled by the target stimulus.
-
-For many psychological questions, such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes, by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
+For many psychological questions, the estimation of such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes and by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
 Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
@@
 
-\caption{Mean performance versus distributional analyses.
+\caption{Simulated data showing mean performance versus distributional analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 
-\subsection{1.2 Aims and structure of the paper}\label{aims-and-structure-of-the-paper}
-
-In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a.
-survival analysis, hazard analysis, duration analysis, failure time analysis, and transition analysis (Singer \& Willett, 2003).
-Our ultimate goal is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
-
-Even though EHA is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
-From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
-Indeed, hundreds of years ago, analytical methods were developed to predict the duration of time until people died (e.g., Halley, 1693; Makeham, 1860).
-The same logic can be applied to psychological time-to-event data, as previously demonstrated {[}Panis, Schmidt, et al. (2020); XXXXX{]}.
-
-We first provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper.
-However, this will remain relatively short, as this has been covered in detail before (Allison, 1982, 2010; Singer \& Willett, 2003).
-Indeed, our primary aim here is to introduce the set of tutorials, which explain \textbf{how} to do such analyses, rather than repeat in any detail \textbf{why} you may do them.
-
-We then provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data measured in RT tasks.
-Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes.
-Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R.
-Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024a), which makes it computationally reproducible, in terms of the underlying data and figures.
-
-.
-
-\section{2. A brief introduction to event history analysis}\label{a-brief-introduction-to-event-history-analysis}
-
-EHA is a class of statistical methods to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010). To apply EHA, one must be able to:
+\subsection{1.2 Aims}\label{aims}
+
+Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
+
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews (Allison, 1982; Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; e.g., Singer \& Willett, 2003; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+Set within this context, our overall aim is to introduce a set of tutorials, which explain \textbf{how} to do such analyses in the context of experimental psychology, rather than repeat in any detail \textbf{why} you may do them.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
+
+\subsection{1.3 Structure}\label{structure}
+
+In what follows, the paper is organised in three main sections.
+In Section 2, we provide a brief overview of EHA to orient the reader to the basic concepts that we will use throughout the paper and why such an approach might be relevant for research in experimental psychology.
+In Section 3, we outline a series of tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/Tutorial_Event_History_Analysis}), along with all of the other code and material associated with the project.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
+In Section 4, we discuss the strengths and weaknesses of the approach for researchers in experimental psychology.
+
+\section{2.
+What is event history analaysis and why is it relevant to research in experimental psychology?}\label{what-is-event-history-analaysis-and-why-is-it-relevant-to-research-in-experimental-psychology}
+
+\subsection{2.1 A brief introduction to event history analysis}\label{a-brief-introduction-to-event-history-analysis}
+
+EHA is a statistical approach to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
+In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
+For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Singer \& Willett, 2003; Winter, 2019).
+We also supply all relevant regression equations in section D of the Supplemental Material.
+
+\subsubsection{2.1.1 Terminolgy and minimum requirements for EHA}\label{terminolgy-and-minimum-requirements-for-eha}
+
+To avoid possible confusion in terminology used, it is worth noting that EHA is frequently known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
+In this paper, we choose to use the term EHA throughout.
+
+In terms of minimum requirements to apply EHA, one must be able to:
 
@@
 
-\subsection{2.1 Single, repeatable, and recurrent events}\label{single-repeatable-and-recurrent-events}
-
-While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
-For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset).
-In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 1 second long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, and the individual always starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs.
-
-In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window.
-In the world of EHA, this is known as a ``competing risks'' situation, because in each trial the participant can transition from an idle state to either a ``correct response'' state or an ``incorrect response'' state.
-
-In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
-In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.
-
-In section A of the Supplemental Material we visualize the types of time-to-event data that are obtained in these typical RT tasks (detection, discrimination or categorization, bistable perception).
-Note that we do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in REF and REF\ldots{}
-
-\subsection{2.2 Right censoring versus data trimming}\label{right-censoring-versus-data-trimming}
-
-What do you do with trials in which no response occurs during the observation period?
-EHA treats such trials as \emph{right-censored} observations on the variable RT, because all we know is that RT is greater than some value.
-Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
-For example, in the one-button detection task example from above, all trials have a \emph{censoring time} of 1 second, but some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
-
-EHA can deal in a straight-forward fashion with right-censored time-to-event data.
-In contrast, experimental psychologists are used to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
-Discarding data can introduce biases, however.
-
-\subsection{2.3 Discrete vs continuous time units}\label{discrete-vs-continuous-time-units}
-
-All man-made measurements of duration are discrete in nature.
-However, when the temporal resolution is high relative to the duration of the observation window, researchers typically treat time as continuous.
-RT data can thus be analysed using continuous-time EHA methods which use the exact event times, including parametric models (e.g., an exponential hazard model, a Weibull hazard model, a lognormal hazard model) and the popular Cox regression model ().
-
-However, in this tutorial we focus on discrete-time methods for three reasons:
-First, we are interested in studying the shape of the hazard function (Cox regression ignores this and only tests the effects of covariates);
-Second, empirical hazard and conditional accuracy functions from certain RT tasks (e.g., interference tasks; Figure 1B) can show abrupt changes in their shape (parametric methods assume smooth distributions), and the shape of the hazard function in many experimental tasks is still unknown (parametric methods assume well-defined probability distributions);
-Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are already familiar.
-
-In sum, due to their simplicity and flexibility, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA and learn to apply EHA, even though continuous-time methods might be more suited in certain situations.
-
-\subsection{2.4 Discrete-time hazard functions and conditional accuracy functions}\label{discrete-time-hazard-functions-and-conditional-accuracy-functions}
-
-After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B), let RT be a discrete random variable denoting the \emph{rank} of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, the detection response in trial 1 might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-Thus, the RT data are interval-censored, because we only use the information that a\textless RT\textless=b when the response occurs in time bin (a,b{]}.
-
-While experimental psychologists are familiar with the cumulative distribution function or F(t) = P(RT \textless= t) and the probability mass function or P(t) = P(RT = t), discrete-time EHA focuses on the discrete-time hazard function of event occurrence:
-
-\noindent h(t) = P(RT = t\textbar{} RT \(\geq\) t) \hfill  (1)
-
-\noindent and the discrete-time survivor function:
-
-\noindent S(t) = P(RT \(>\) t) = 1 - F(t) = {[}1-h(t){]}.{[}1-h(t-1){]}.{[}1-h(t-2){]}\ldots{[}1-h(1){]} \hfill  (2)
-
-The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous risk that the response occurs in bin t, given that it has not yet occurred in one of the prior bins.
-In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the survival probability, the probability that the event does not occur before the endpoint of bin t.
-As a result, only the hazard function conveys the risk of event occurrence associated with each bin, and \ldots{} suited for online tracking of performance..
-cfr mouse cursor movements\ldots.
-
-For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
-
-\noindent ca(t) = P(correct \textbar{} RT = t) \hfill  (5)
-
-\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+\subsubsection{2.1.2 Types of EHA}\label{types-of-eha}
+
+There are different types of EHA.
+For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
+As a lab, and mainly for practical reasons, we have much more experience using discrete time EHA, and that is the approach that we describe and focus on in this paper.
+This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable.
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well.
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.
+Moreover, as indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
+
+To apply discrete time EHA, one divides time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+Then let t be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+For example, a response on a single trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+One then calculates the discrete-time hazard function of event occurrence.
+The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in
+bin t, given that the event does not occur in previous bins.
+In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet
+occurred in the past, i.e., in one of the prior bins.
+
+In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
+In those cases, one can also calculate the conditional accuracy function (Figure B).
+The conditional accuracy function gives you for each time bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 
-As we will illustrate in Tutorials 1a and 1b, performing a descriptive EHA/SAT analysis by calculating the sample-based estimates of h(t), S(t) and ca(t) for each combination of participant and condition requires setting up a \emph{life table}.
-Definition life table\ldots{}
-
-\subsection{2.5 Bayesian vs.~frequentist approaches to regression}\label{bayesian-vs.-frequentist-approaches-to-regression}
-
-To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy.
-Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation).
-Note that time-varying covariates are not covered in this tutorial.
-
-Heterogeneity
-
-Multilevel survival analysis: Methods, Models and Applications
-Austin 2017 !!
-
-fitting problems
-
-\subsection{2.6 Number of samples, repeated measures, time bins}\label{number-of-samples-repeated-measures-time-bins}
-
-In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition.
-To test process models of cognition, researchers tend to use small-N designs, \ldots{}
-
-Power
-
-Number of time bins?
-
-\subsection{bin width}\label{bin-width}
-
-We recommend several excellent textbooks for a comprehensive background context to EHA (Allison, 2010; Singer \& Willett, 2003) and for a more general introduction to understanding regression equations (Gelman, Hill, \& Vehtari, 2020; Winter, 2019).
-Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before.
-Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-As such, we only supply regression equations in section D of the Supplemental Material.
-
-We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024b), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses.
+\subsection{2.2 Benefits of event history analysis for research in experimental psychology}\label{benefits-of-event-history-analysis-for-research-in-experimental-psychology}
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here).
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (section F of the Supplemental Material).
+Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
+
+We highlight three benefits that we think are relevant to the domain of experimental psychology.
+First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about cognition and theoretical development.
+As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
+For example, what kind of theory or theories could account for the effects reported in Figure 1B?
+Are there new auxiliary assumptions that theories need to adopt?
+And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
+
+Second, compared to more conventional analytical approaches, EHA uses more of the data and deals with msising data differently.
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Discarding data can introduce biases, however.
+Rather than treat non-responses as missing data, EHA treats such trials as \emph{right-censored} observations on the variable RT, because all we know is that RT is greater than some value.
+Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
+For example, if the censor time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+The use of right-censoring in EHA, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology.
+
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks (Supp Fig X).
+EHA can also be used with a range of common experimental manipualtions, such as stimulus-onset-asynchrony (Supp Fig X).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology.
+
+\subsection{2.3 Implications for research design in experimental psychology}\label{implications-for-research-design-in-experimental-psychology}
+
+Performing EHA in experimental psychology has implications for how experiments are designed.
+More specifically, we consider two implications that researchers will need to consider when using discrete time EHA.
+
+First, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
+This approach emphasizes the precision and reproducibility of data patterns at the individual participant level to increase the inferential validity of the design (Baker et al., 2021; Smith \& Little, 2018).
+Note that because statistical power derives both from the number of participants and from the number of repeated measures per participant and condition, small-\emph{N} designs can still achieve what are generally considered acceptable levels of statistical power, if they have a sufficient amount of data overall (Baker et al., 2021; Smith \& Little, 2018).
+
+Second, the width of each time bin will need to be determined.
+For instance, in Figure 1B we chose 100ms in an arbitrary manner.
+In reality, however, bin width will need to be set by considering a number of factors simultabeously.
+For example, {[}{[}Sven - add a few sentences here{]}{]}
+
+\section{4. Tutorials}\label{tutorials}
+
+We used r my\_r\_citation\$r for all reported analyses.
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-
-\section{4. Tutorials}\label{tutorials}
 
@@
 This data set is used when fitting conditional accuracy models (Tutorials 2b and 3b).
-
-\begin{verbatim}
-## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
-## Use 'xfun::attr2()' instead.
-## See help("Deprecated")
-\end{verbatim}
 
@@
 Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
-
-\begin{verbatim}
-## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
-## Use 'xfun::attr2()' instead.
-## See help("Deprecated")
-\end{verbatim}
 
@@
 The next step is to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha().
-When creating the plots, some warning messages will likely be generated, like these:
-
-\begin{itemize}
-\tightlist
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_line()}).
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_point()}).
-\item
-  Removed 2 rows containing missing values or values outside the scale range (\texttt{geom\_segment()}).
-\end{itemize}
-
-The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
@@
 Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
-
-\begin{verbatim}
-## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
-## Use 'xfun::attr2()' instead.
-## See help("Deprecated")
-\end{verbatim}
 
@@
 
+In general, when fitting regression models, our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
+We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
+This means that wherever possible we include varying intercepts and slopes per participant.
+To make inferences, we use two main approaches.
+We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020).
+We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
+
 \subsubsection{4.3.1 Hazard model considerations}\label{hazard-model-considerations}
@@
 
-\subsection{4.5 Tutorial 3a: Fitting Frequentist hazard models}\label{tutorial-3a-fitting-frequentist-hazard-models}
-
-In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
-The general process is similar to that in Tutorial 2a, except that there are no priors to set.
-
-Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting hazard model is called M1i\_f with the appended ``\_f'' denoting a frequentist model.
-
-In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i\_f.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_3_Frequentist/comparison}
-
-}
-
-\caption{Parameter estimates for model M1i from brm() -- means and 95\% credible intervals -- and model M1i\_f from glmer() -- maximum likelihood estimates and 95\% confidence intervals.}\label{fig:plot-comparison}
-\end{figure}
-
-Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
-However, model M1i\_f did not converge and resulted in a singular fit.
-This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
-But the price you pay for being able to fit models with more complex varying effects structures via a Bayesian framework is increased computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
-
-\subsection{4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models}\label{tutorial-3b-fitting-frequentist-conditional-accuracy-models}
-
-In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
-To be concise, we only fit effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024).
-The resulting conditional accuracy model, which we labelled M1i\_ca\_f, did not converge and resulted in a singular fit.
-Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
-
 \subsection{4.7 Tutorial 4: Planning}\label{tutorial-4-planning}
@@
 
-\section{5. Discussion}\label{discussion}
+\section{4. Discussion}\label{discussion}
 
@@
 
-\subsection{5.1 What are the main use-cases of EHA for understanding cognition and brain function?}\label{what-are-the-main-use-cases-of-eha-for-understanding-cognition-and-brain-function}
+\subsection{4.1 What are the main use-cases of EHA for understanding cognition and brain function?}\label{what-are-the-main-use-cases-of-eha-for-understanding-cognition-and-brain-function}
 
@@
 
-\subsection{5.2 Model complexity versus interpretability}\label{model-complexity-versus-interpretability}
+\subsection{4.2 Model complexity versus interpretability}\label{model-complexity-versus-interpretability}
 
@@
 
-\subsection{5.3 Individual differences}\label{individual-differences}
-
-One important issue is that of possible individual differences in the overall location of the distribution, and the time course of psychological effects.
-For example, when you wait for a response of the participant on each trial, you allow the participant to have control over the trial duration, and some participants might respond only when they are confident that their emitted response will be correct.
-These issues can be avoided by introducing a (relatively short) response deadline in each trial, e.g., 500 ms for simple detection tasks, 800 ms for more difficult discrimination tasks, or 2 s for tasks requiring extended high-level processing.
-Because EHA can deal in a straightforward fashion with right-censored observations (i.e., trials without an observed response in the analysis time window), introducing a response deadline is recommended when designing RT experiments.
-Furthermore, introducing a response deadline and asking participants to respond before the deadline as much as possible, will also lead to individual distributions that overlap in time, which is important when selecting a common analysis time window when fitting hazard and conditional accuracy models.
-
-But even when using a response deadline, participants can differ qualitatively in the effects they display (see Panis, 2020).
-One way to deal with this is to describe and interpret the different patterns.
-Another way is to run a clustering algorithm on the individual hazard estimates across all bins and conditions.
-The obtained dendrogram can then be used to identify a (hopefully big) cluster of participants that behave similarly, and to identify a (hopefully small) cluster of participants with different behavioral patterns.
-One might then exclude the smaller sub-group of participants before fitting a hazard model or consider the possibility that different cognitive processes may be at play during task performance across the different sub-groups.
-
-Another approach to deal with individual differences is Bayesian prevalence (Ince, Paton, Kay, \& Schyns, 2021), which is a from of small-\emph{N} approach (Smith \& Little, 2018).
-This method looks at effects within each individual in the study and asks how likely it would be to see the same result if the experiment was repeated with a new person chosen from the wider population at random.
-This approach allows one to quantify how typical or uncommon an observed effect is in the population, and the uncertainty around this estimate.
-
-\subsection{5.4 Limitations}\label{limitations}
+\subsection{4.3 Limitations}\label{limitations}
 
@@
 
-We did not discuss continuous-time EHA, nor continuous-time SAT analysis.
-As indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods.
-Given that RT is typically treated as a continuous variable, it is possible that continuous-time methods will ultimately prevail.
-However, they require much more data to estimate the continuous-time hazard (rate) function well.
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.
-
-\subsection{5.5 Extensions}\label{extensions}
-
-The hazard models in this tutorial assume that there is one event of interest.
-For RT data, this button-press event constitutes a single transition between an ``idle'' state and a ``responded'' state.
-However, in certain situations, more than one event of interest might exist.
-For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before being absorbed into a final ``death'' state.
-When you have data on the timing of these transitions, one can apply multi-state hazard models, which generalize EHA to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
-Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial.
-Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (Allison, 2010).
-For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence.
-Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.
-
-\section{6. Conclusions}\label{conclusions}
+\section{5. Conclusions}\label{conclusions}
 
@@
 
-\bibitem[\citeproctext]{ref-R-citr}
-Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://github.com/crsh/citr}
-
-\bibitem[\citeproctext]{ref-pap2024}
-Aust, F., \& Barth, M. (2024a). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
-
-\bibitem[\citeproctext]{ref-R-papaja}
-Aust, F., \& Barth, M. (2024b). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
-
-\bibitem[\citeproctext]{ref-R-tinylabels}
-Barth, M. (2023). \emph{{tinylabels}: Lightweight variable labels}. Retrieved from \url{https://cran.r-project.org/package=tinylabels}
-
-\bibitem[\citeproctext]{ref-R-lme4}
-Bates, D., Mächler, M., Bolker, B., \& Walker, S.
-(2015).
-Fitting linear mixed-effects models using {lme4}.
-\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
-\url{https://doi.org/10.18637/jss.v067.i01}
-
-\bibitem[\citeproctext]{ref-R-Matrix}
-Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://Matrix.R-forge.R-project.org}
-
-\bibitem[\citeproctext]{ref-R-RJ-2021-048}
-Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+\bibitem[\citeproctext]{ref-bakerPowerContoursOptimising2021}
+Baker, D.
+H., Vilidaite, G., Lygo, F.
+A., Smith, A.
+K., Flack, T.
+R., Gouws, A.
+D., \& Andrews, T.
+J.
+(2021).
+Power contours: {Optimising} sample size and precision in experimental psychology and human neuroscience.
+\emph{Psychological Methods}, \emph{26}(3), 295--314.
+\url{https://doi.org/10.1037/met0000337}
+
+\bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
+Barr, D.
+J., Levy, R., Scheepers, C., \& Tily, H.
+J.
+(2013).
+Random effects structure for confirmatory hypothesis testing: {Keep} it maximal.
+\emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001.
+\url{https://doi.org/10.1016/j.jml.2012.11.001}
 
@@
 Box-Steffensmeier, J. M. (2004). Event history modeling: A guide for social scientists. Cambridge: University Press.
-
-\bibitem[\citeproctext]{ref-R-brms_a}
-Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
-
-\bibitem[\citeproctext]{ref-R-brms_b}
-Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411. \url{https://doi.org/10.32614/RJ-2018-017}
-
-\bibitem[\citeproctext]{ref-R-brms_c}
-Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with {brms} and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}
 
@@
 
-\bibitem[\citeproctext]{ref-R-Rcpp_b}
-Eddelbuettel, D., \& Balamuta, J.
-J.
-(2018).
-{Extending {R} with {C++}: A Brief Introduction to {Rcpp}}.
-\emph{The American Statistician}, \emph{72}(1), 28--36.
-\url{https://doi.org/10.1080/00031305.2017.1375990}
-
-\bibitem[\citeproctext]{ref-R-Rcpp_a}
-Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. \url{https://doi.org/10.18637/jss.v040.i08}
-
-\bibitem[\citeproctext]{ref-R-cmdstanr}
-Gabry, J., Češnovar, R., Johnson, A., \& Bronder, S. (2024). \emph{Cmdstanr: R interface to 'CmdStan'}. Retrieved from \url{https://github.com/stan-dev/cmdstanr}
-
-\bibitem[\citeproctext]{ref-R-bayesplot}
-Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in bayesian workflow. \emph{J. R. Stat. Soc. A}, \emph{182}, 389--402. \url{https://doi.org/10.1111/rssa.12378}
-
 \bibitem[\citeproctext]{ref-gelmanRegressionOtherStories2020}
@@
 
-\bibitem[\citeproctext]{ref-R-standist}
-Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
-
-\bibitem[\citeproctext]{ref-R-lubridate}
-Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. Retrieved from \url{https://www.jstatsoft.org/v40/i03/}
-
-\bibitem[\citeproctext]{ref-Halley1693}
-Halley, E.
-(1693).
-VI.
-An estimate of the degrees of the mortality of mankind; drawn from curious tables of the births and funerals at the city of breslaw; with an attempt to ascertain the price of annuities upon lives.
-\emph{Philosophical Transactions of the Royal Society of London}, \emph{17}(196), 596--610.
-\url{https://doi.org/10.1098/rstl.1693.0007}
-
 \bibitem[\citeproctext]{ref-heiss2021}
@@
 Hosmer, D. W., Lemeshow, S., \& May, S. (2011). \emph{Applied {Survival Analysis}: {Regression Modeling} of {Time} to {Event Data}} (2nd ed). Hoboken: John Wiley \& Sons.
-
-\bibitem[\citeproctext]{ref-inceBayesianInferencePopulation2021}
-Ince, R. A., Paton, A. T., Kay, J. W., \& Schyns, P. G. (2021). Bayesian inference of population prevalence. \emph{eLife}, \emph{10}, e62461. \url{https://doi.org/10.7554/eLife.62461}
 
@@
 
-\bibitem[\citeproctext]{ref-R-tidybayes}
-Kay, M. (2024). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
+\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
+Kruschke, J.
+K., \& Liddell, T.
+M.
+(2018).
+The {Bayesian New Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power analysis from a {Bayesian} perspective.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 178--206.
+\url{https://doi.org/10.3758/s13423-016-1221-4}
 
@@
 
-\bibitem[\citeproctext]{ref-williammatthewmakehamLawMortalityConstruction1860}
-Makeham, W. M. (1860). \emph{On the {Law} of {Mortality} and the {Construction} of {Annuity Tables}}. {The Assurance Magazine, and Journal of the Institute of Actuaries}.
-
 \bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
 McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
-
-\bibitem[\citeproctext]{ref-R-tibble}
-Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}
 
@@
 
-\bibitem[\citeproctext]{ref-R-RColorBrewer}
-Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}
-
 \bibitem[\citeproctext]{ref-panisHowCanWe2020}
@@
 
-\bibitem[\citeproctext]{ref-R-patchwork}
-Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://patchwork.data-imaginist.com}
-
-\bibitem[\citeproctext]{ref-R-nlme}
-Pinheiro, J. C., \& Bates, D. M. (2000). \emph{Mixed-effects models in s and s-PLUS}. New York: Springer. \url{https://doi.org/10.1007/b98882}
-
-\bibitem[\citeproctext]{ref-R-base}
-R Core Team. (2024). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\bibitem[\citeproctext]{ref-ripleyMASSSupportFunctions2024}
-Ripley, B., Venables, B., Bates, D.
-M., ca 1998), K.
-H.
-(partial.
-port, ca 1998), A.
-G.
-(partial.
-port, \& polr), D.
-F.
-(support.
-functions for.
-(2024).
-\emph{{MASS}: {Support Functions} and {Datasets} for {Venables} and {Ripley}'s {MASS}}.
-
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
@@
 
-\bibitem[\citeproctext]{ref-R-StanHeaders}
-Stan Development Team. (2020). \emph{{StanHeaders}: Headers for the {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
-
-\bibitem[\citeproctext]{ref-R-rstan}
-Stan Development Team. (2024). \emph{{RStan}: The {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
-
-\bibitem[\citeproctext]{ref-steeleGeneralMultilevelMultistate2004}
-Steele, F., Goldstein, H., \& Browne, W.
-(2004).
-A general multilevel multistate competing risks model for event history data, with an application to a study of contraceptive use dynamics.
-\emph{Statistical Modelling}, \emph{4}(2), 145--159.
-\url{https://doi.org/10.1191/1471082X04st069oa}
+\bibitem[\citeproctext]{ref-smithSmallBeautifulDefense2018}
+Smith, P.
+L., \& Little, D.
+R.
+(2018).
+Small is beautiful: {In} defense of the small-{N} design.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(6), 2083--2101.
+\url{https://doi.org/10.3758/s13423-018-1451-8}
 
@@
 
-\bibitem[\citeproctext]{ref-R-ggplot2}
-Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-forcats}
-Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}
-
-\bibitem[\citeproctext]{ref-R-stringr}
-Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-tidyverse}
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., \ldots{} Yutani, H.
-(2019).
-Welcome to the {tidyverse}.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
-
-\bibitem[\citeproctext]{ref-R-dplyr}
-Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-purrr}
-Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
-
-\bibitem[\citeproctext]{ref-R-readr}
-Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
-
-\bibitem[\citeproctext]{ref-R-tidyr}
-Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
-
 \bibitem[\citeproctext]{ref-winterStatisticsLinguistsIntroduction2019}

### 2025-06-20 manuscript/notes/notes.Rmd
notes on the updated manuscript

---
title: "notes"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# June 20th 2025 #

## package loading ##

loading packages at the top of the file is a bit of a mess at the minute and appears rather complicated and it needs cleaning up.
there are several repeat library loads e.g., tidyverse and patchwork

## r-references ##

this bib r-references r-ref thing file prevented the manuscript from building, so I had to comment it out.
This needs fixing.

## intro ##

- I made the intro shorter. We still make the same points, just in far fewer words.

## section 2 ##

I took out the following and changed it to be far shorter. 
If you want to include any of it in supps, then let's discuss it.
Instead, I chose to focus on picking out the benefits of EHA for research in experimental psychology.
That way, we are being explicit about what we are aiming to do in this section.
And we are therefore making it clear that we are NOT providing a detailed description of EHA.
There is simply not enough space.
And we look confused and contradictory because we don't actually do what we say.

You will notice that I have mentioned many of the below points, just not in detail.
And again, that's becuase that is not the point of the paper.

- leave the survivor function in supps because it just seems unnecessary, especially since we do not plot it anymore.

- Sven - can you add a relevant reference for the following statement:
Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF??).

- in section 2.3 can you add a line or two about bin width please?

<!-- ## 2.1 Single, repeatable, and recurrent events -->

<!-- While people can die only once, in experimental RT tasks the events of interest are typically repeatable. For example, in the target-present condition of a one-button detection task the participant is presented in each trial with a faint target stimulus whose presence (s)he has to detect by pressing a button within a certain time window (e.g., the first second after target onset). In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 1 second long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, and the individual always starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs. -->

<!-- In a two-button discrimination task, the participant is presented in each trial with a target stimulus that (s)he has to categorize by pressing one of two buttons within a certain time window. In the world of EHA, this is known as a "competing risks" situation, because in each trial the participant can transition from an idle state to either a "correct response" state or an "incorrect response" state. -->

<!-- In a bistable perception task, the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation. In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.  -->

<!-- In section A of the Supplemental Material we visualize the types of time-to-event data that are obtained in these typical RT tasks (detection, discrimination or categorization, bistable perception). Note that we do not analyse recurrent events in this tutorial. More information about recurrent events analysis can be found in REF and REF... -->

<!-- ## 2.2 Right censoring versus data trimming -->

<!-- What do you do with trials in which no response occurs during the observation period? EHA treats such trials as *right-censored* observations on the variable RT, because all we know is that RT is greater than some value. Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data. For example, in the one-button detection task example from above, all trials have a *censoring time* of 1 second, but some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.  -->

<!-- EHA can deal in a straight-forward fashion with right-censored time-to-event data. In contrast, experimental psychologists are used to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF). Discarding data can introduce biases, however.  -->

<!-- ## 2.3 Discrete vs continuous time units -->

<!-- All man-made measurements of duration are discrete in nature. However, when the temporal resolution is high relative to the duration of the observation window, researchers typically treat time as continuous. RT data can thus be analysed using continuous-time EHA methods which use the exact event times, including parametric models (e.g., an exponential hazard model, a Weibull hazard model, a lognormal hazard model) and the popular Cox regression model (). -->

<!-- However, in this tutorial we focus on discrete-time methods for three reasons: -->
<!-- First, we are interested in studying the shape of the hazard function (Cox regression ignores this and only tests the effects of covariates);  -->
<!-- Second, empirical hazard and conditional accuracy functions from certain RT tasks (e.g., interference tasks; Figure 1B) can show abrupt changes in their shape (parametric methods assume smooth distributions), and the shape of the hazard function in many experimental tasks is still unknown (parametric methods assume well-defined probability distributions); -->
<!-- Third, in discrete time, hazard is simply defined as a conditional probability (see 2.4) and we can apply logistic regression modeling with which most experimental psychologists are already familiar. -->

<!-- In sum, due to their simplicity and flexibility, we believe that discrete-time methods are a good starting point for experimental psychologists that want to abandon ANOVA and learn to apply EHA, even though continuous-time methods might be more suited in certain situations. -->

<!-- ## 2.4 Discrete-time hazard functions and conditional accuracy functions -->

<!-- After dividing time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B), let RT be a discrete random variable denoting the *rank* of the time bin in which a particular person's response occurs in a particular experimental condition. For example, the detection response in trial 1 might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms). Thus, the RT data are interval-censored, because we only use the information that a<RT<=b when the response occurs in time bin (a,b]. -->

<!-- While experimental psychologists are familiar with the cumulative distribution function or F(t) = P(RT <= t) and the probability mass function or P(t) = P(RT = t), discrete-time EHA focuses on the discrete-time hazard function of event occurrence: -->

<!-- \noindent h(t) = P(RT = t| RT $\geq$ t)   \hfill  (1) -->

<!-- \noindent and the discrete-time survivor function:  -->

<!-- \noindent S(t) = P(RT $>$ t) = 1 - F(t) =  [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)]  \hfill  (2) -->

<!-- The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins.  -->
<!-- In other words, it reflects the instantaneous risk that the response occurs in bin t, given that it has not yet occurred in one of the prior bins. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the survival probability, the probability that the event does not occur before the endpoint of bin t. As a result, only the hazard function conveys the risk of event occurrence associated with each bin, and ... suited for online tracking of performance.. cfr mouse cursor movements.... -->

<!-- For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function  -->

<!-- \noindent ca(t) = P(correct | RT = t)   \hfill  (5) -->

<!-- \noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010]. The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function. -->
<!-- We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.  -->

<!-- As we will illustrate in Tutorials 1a and 1b, performing a descriptive EHA/SAT analysis by calculating the sample-based estimates of h(t), S(t) and ca(t) for each combination of participant and condition requires setting up a *life table*. -->
<!-- Definition life table... -->

<!-- ## 2.5 Bayesian vs. frequentist approaches to regression -->

<!-- To study how the risk of a response, and the accuracy of an emitted response, depends on covariates (i.e., explanatory predictor variables) we can estimate regression models for hazard and for conditional accuracy, i.e., perform inferential EHA/SAT analysis. Such covariates can be constant over within-trial time (e.g., gender, race, trial number, block number) or vary with within-trial time (e.g., heart rate, eye gaze position, eye pupil dilation). Note that time-varying covariates are not covered in this tutorial. -->

<!-- Heterogeneity -> Multilevel survival analysis: Methods, Models and Applications -->
<!-- Austin 2017 !! -->

<!-- fitting problems -> Bayesian -->

<!-- ## 2.6 Number of samples, repeated measures, time bins -->

<!-- In a typical RT data set from a within-subject design, there are N individuals and M repeated measures or trials per experimental condition. -->
<!-- To test process models of cognition, .. advises  to use small-N designs, ... eACH SUBJECT REPLICATION UNIT -->

<!-- Power IS A COMPLEX FUNCTION OF .... (REF A, REF B ON POWER WITH EXPONENTIAL) -->

<!-- Number of time bins? -->

<!-- bin width -->
<!-- ------------------- -->

## tutorial section ##

- the frequentist tutorial still needs removing and placing in supps. 
then just refer to it in the main text.
I commented it out for now.

## 4.1.2 ##

please move the below text to the script or tutorial itself.
e.g., inlcude a commented out note that explains this can be ignored.
there is no need for this in the paper itself.

<!-- When creating the plots, some warning messages will likely be generated, like these: -->

<!-- * Removed 2 rows containing missing values or values outside the scale range (`geom_line()`).  -->
<!-- * Removed 2 rows containing missing values or values outside the scale range (`geom_point()`).  -->
<!-- * Removed 2 rows containing missing values or values outside the scale range (`geom_segment()`). -->

<!-- The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored. -->

page 19. Figure 2. The image warped / wrong size and needs fixing.

## combining figures ##

we currently have a lot of space dedicated to results figures e.g., Figures 3,4,5,6,7,8.

I suggest an alternative, which is more efficient and compact.

Let's at least try it and see how we think it compares.

I would reduce 6 figures into 2.

One hazard figure.

One ca figure.

Each figure has 3 panels across 3 rows.
panel a - parameter estimates (current figure 3 e.g.)
panel b - predicted hazard group level (current figure 4a)
panel c - condition differences in hazard (current figure 5a).

Then do the same for CA in a separate figure.

And then report individual differences in Supplementary and refer to the figure in the main text.

What I like about this approach is that:

A) we save a tremendous amount of space.
B) we still make the same points.
C) we do not really focus in any material way on individual differences, so why both with the plots in the main text.
D) folks still have easy access to the code and the plots in Supps should they feel they need them.
E) we are really responding well to reviewers in terms of saving space and being more compact.

## discussion ##

I removed the unnecessary sections that we discussed previously.

### 2025-06-23 manuscript/Suppl_material/supplementary.Rmd
update Supplementary Material section A

@@
 
-## A. Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions
+## A. RT tasks in the context of EHA: right-censoring versus random censoring
+
+While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
+In Supplementary Figure 1A, we present a hypothetical single-subject RT data set from a detection experiment with one response button and two experimental conditions: target present and target absent.
+For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which the target stimulus is presented or not depending on the experimental condition the trials belongs to, and finally a blank screen until an overt response is detected or 500 ms have passed.
+Supplementary Figure 1A shows the typical pattern: the participants presses the button sometime before 500 ms in most trials in the target-present condition (black dots), while (s)he does not press the button in most trials of the target-absent condition (black crosses).
+
+In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 500 ms long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
+
+Right-censoring occurs when we only know that the event time is larger than some value, i.e., the censoring time.
+The most common type of right-censoring is “singly Type I censoring” which applies when the experiment uses a fixed response deadline for all trials.
+“Type I” means that the censoring time is fixed and under the control of the experimenter; “Singly” refers to the fact that all observations have the same censoring time (Allison, 2010).
+Clearly, discarding right-censored observations before calculating the median or mean RT will result in biased estimates, especially in the target-absent condition.
+
+(ref:descr-plot-RT-tasks-caption) RT tasks and EHA concepts.
+The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line).
+In each trial of each condition of each task, an overt response time (RT) is indicated by a circle.
+(A) Single-button detection data requires a single-event analysis.
+(B) Two-button discrimination data requires a competing risks analysis.
+(C) Competing risk data in combination with an SOA manipulation might require increasing the observation period in each trial.
+(D) Two-button bistable perception data requires a recurrent event / multistate analysis.
+
+```{r plot-RT-tasks, fig.cap = "(ref:descr-plot-RT-tasks-caption)", out.width="80%"}
+knitr::include_graphics("../../manuscript/Suppl_material/Plot_paradigms.png")
+```
+
+Importantly, when estimating the shape of the RT distribution in each experimental condition, EHA takes into account right-censored observations.
+As a result, EHA has a wider scope than current methods, as it can also be applied to RT data from experimental paradigms in which RTs are typically not measured, such as masking experiments and studies on the attentional blink.
+Also, EHA allows to increase experimental design efficiency, because we can introduce a response deadline instead of waiting until a response is emitted in each trial, so that we do not have to wait for very slow responses which would be trimmed anyway.
+
+There are different types of censoring.
+Left censoring occurs when all that is known about an observation on a variable T is that it is less than some value.
+Interval censoring combines right and left censoring so that all you know about T is that a < T < b, for some values of a and b (Allison, 2010).
+Random censoring occurs when observations are terminated for reasons that are not under the control of the experimenter.
+
+For example, in Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B.
+For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which a female or male face is presented, and finally a blank screen until an overt response is detected or 800 ms have passed.
+Participants are required to press the left button when identifying a male, and the right button when identifying a female.
+In Supplementary Figure 1B, correct responses are indicated with black dots, error responses with white dots, and the few right-censored observations with crosses.
+
+In EHA parlance this is known as a *competing risks* situation, because from target display screen onset (time zero) onwards, both correct and incorrect responses are competing with another in each trial until one occurs which determines the event type and latency that are recorded.
+In other words, in each trial the participant can transition from an idle state to either a "correct response" state or an "incorrect response" state.
+
+There are two ways to deal with a competing-risks situation.
+One approach is to model the hazard function of correct response occurrence and treat incorrect responses as right-censored observations, and vice versa.
+When estimating the hazard of correct response occurrence, then the unpredictable error responses will introduce random censoring (and vice versa).
+Importantly, all statistical methods for time-to-event data require that random censoring be noninformative: a trial that is censored at time c should be representative of all those trials with the same values of the explanatory variables that survive to c (Allison, 2010).
+However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another (Burle et al., 2004; Eriksen et al., 1985; Praamstra & Seiss, 2005).
+Thus, another approach is to model the hazard of (any) response occurrence and extend it with a speed-accuracy trade-off (SAT) analysis (plotting accuracy as a function of latency).
+This is the approach taken here (see Figure 1), and many experimental psychologists are familiar with SAT.
+
+Studying the effect of SOA on mean RT is still a popular approach to study the time course of cognitive processes (ref).
+For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue).
+However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis & Schmidt, 2016).
+For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels).
+To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset.
+Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these "negative" RTs in the analysis (lower x-axis labels).
+In general, all RTs must be positive for EHA to work.
+
+Finally, in a bistable perception task (Supplementary Figure 1D), the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes in each trial, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
+In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.
+Note that we do not analyse recurrent events in this tutorial.
+More information about recurrent events analysis can be found in REF and REF...
+
+## B. Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions
 

### 2025-06-23 manuscript/Suppl_material/supplementary.tex
update Supplementary Material section A

@@
 {}{\PatchFailed}
+\makeatother
 \usepackage{lineno}
@@
 
-\subsection{A.
-Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions}\label{a.-definitions-of-discrete-time-hazard-surivor-probability-mass-and-conditional-accuracy-functions}
+\subsection{A. RT tasks in the context of EHA: right-censoring versus random censoring}\label{a.-rt-tasks-in-the-context-of-eha-right-censoring-versus-random-censoring}
+
+While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
+In Supplementary Figure 1A, we present a hypothetical single-subject RT data set from a detection experiment with one response button and two experimental conditions: target present and target absent.
+For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which the target stimulus is presented or not depending on the experimental condition the trials belongs to, and finally a blank screen until an overt response is detected or 500 ms have passed.
+Supplementary Figure 1A shows the typical pattern: the participants presses the button sometime before 500 ms in most trials in the target-present condition (black dots), while (s)he does not press the button in most trials of the target-absent condition (black crosses).
+
+In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 500 ms long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
+
+Right-censoring occurs when we only know that the event time is larger than some value, i.e., the censoring time.
+The most common type of right-censoring is ``singly Type I censoring'' which applies when the experiment uses a fixed response deadline for all trials.
+``Type I'' means that the censoring time is fixed and under the control of the experimenter; ``Singly'' refers to the fact that all observations have the same censoring time (Allison, 2010).
+Clearly, discarding right-censored observations before calculating the median or mean RT will result in biased estimates, especially in the target-absent condition.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../../manuscript/Suppl_material/Plot_paradigms}
+
+}
+
+\caption{RT tasks and EHA concepts.
+The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line).
+In each trial of each condition of each task, an overt response time (RT) is indicated by a circle.
+(A) Single-button detection data requires a single-event analysis.
+(B) Two-button discrimination data requires a competing risks analysis.
+(C) Competing risk data in combination with an SOA manipulation might require increasing the observation period in each trial.
+(D) Two-button bistable perception data requires a recurrent event / multistate analysis.}\label{fig:plot-RT-tasks}
+\end{figure}
+
+Importantly, when estimating the shape of the RT distribution in each experimental condition, EHA takes into account right-censored observations.
+As a result, EHA has a wider scope than current methods, as it can also be applied to RT data from experimental paradigms in which RTs are typically not measured, such as masking experiments and studies on the attentional blink.
+Also, EHA allows to increase experimental design efficiency, because we can introduce a response deadline instead of waiting until a response is emitted in each trial, so that we do not have to wait for very slow responses which would be trimmed anyway.
+
+There are different types of censoring.
+Left censoring occurs when all that is known about an observation on a variable T is that it is less than some value.
+Interval censoring combines right and left censoring so that all you know about T is that a \textless{} T \textless{} b, for some values of a and b (Allison, 2010).
+Random censoring occurs when observations are terminated for reasons that are not under the control of the experimenter.
+
+For example, in Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B.
+For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which a female or male face is presented, and finally a blank screen until an overt response is detected or 800 ms have passed.
+Participants are required to press the left button when identifying a male, and the right button when identifying a female.
+In Supplementary Figure 1B, correct responses are indicated with black dots, error responses with white dots, and the few right-censored observations with crosses.
+
+In EHA parlance this is known as a \emph{competing risks} situation, because from target display screen onset (time zero) onwards, both correct and incorrect responses are competing with another in each trial until one occurs which determines the event type and latency that are recorded.
+In other words, in each trial the participant can transition from an idle state to either a ``correct response'' state or an ``incorrect response'' state.
+
+There are two ways to deal with a competing-risks situation.
+One approach is to model the hazard function of correct response occurrence and treat incorrect responses as right-censored observations, and vice versa.
+When estimating the hazard of correct response occurrence, then the unpredictable error responses will introduce random censoring (and vice versa).
+Importantly, all statistical methods for time-to-event data require that random censoring be noninformative: a trial that is censored at time c should be representative of all those trials with the same values of the explanatory variables that survive to c (Allison, 2010).
+However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another (Burle et al., 2004; Eriksen et al., 1985; Praamstra \& Seiss, 2005).
+Thus, another approach is to model the hazard of (any) response occurrence and extend it with a speed-accuracy trade-off (SAT) analysis (plotting accuracy as a function of latency).
+This is the approach taken here (see Figure 1), and many experimental psychologists are familiar with SAT.
+
+Studying the effect of SOA on mean RT is still a popular approach to study the time course of cognitive processes (ref).
+For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue).
+However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis \& Schmidt, 2016).
+For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels).
+To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset.
+Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these ``negative'' RTs in the analysis (lower x-axis labels).
+In general, all RTs must be positive for EHA to work.
+
+Finally, in a bistable perception task (Supplementary Figure 1D), the participant is looking at an ambiguous stimulus (e.g., the duck-rabbit illusion, the Necker cube) for two minutes in each trial, for example, and asked to press a button each time when her/his perception switches from one possibe interpretation to the other possible interpretation.
+In this task, there are two events (percept A switches to percept B, percept B switches to percept A) that can recur within the same observation period of two minutes, so that the individual transitions back and forth between two states.
+Note that we do not analyse recurrent events in this tutorial.
+More information about recurrent events analysis can be found in REF and REF\ldots{}
+
+\subsection{B.
+Definitions of discrete-time hazard, surivor, probability mass, and conditional accuracy functions}\label{b.-definitions-of-discrete-time-hazard-surivor-probability-mass-and-conditional-accuracy-functions}
 
@@
 This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
+
 While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
+
 The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1.
@@
 The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.
+
+\subsection{B. Benefits of event history analysis}\label{b.-benefits-of-event-history-analysis}
+
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
+Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
+The benefits include:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
+\item
+  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
+\item
+  EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
+Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
+\item
+  Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+(Allison, 2010).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+\item
+  EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
+\end{enumerate}
 
@@
 (Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer et al., 1988).
 
@@
 EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived ``dips'' that could signal the presence of active top-down respone inhibition processes (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
 

### 2025-06-23 manuscript/manuscript.Rmd
update Supplementary Material section A

@@
 
+In section A of the Supplemental Material we visualize the types of time-to-event data that are obtained in these typical RT tasks (detection, discrimination or categorization, bistable perception).
+
 ### 2.1.1 Terminolgy and minimum requirements for EHA
@@
 For example, [[Sven - add a few sentences here]]
+
+The optimal bin width will depend on many factors, including (a) the length of the observation period in each trial, (b) the rarity of event occurrence, (c) the number of trials per condition per participant, (d) the shape of the hazard function, etc.
+
+Finding an appropriate bin width in a given user case will require testing a number of options, when calculating and plotting the descriptive statistics (see below).
+The goal is to find the smallest bin width that is supported by the amount of data available.
+Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
+Too small bin widths will result in irregular hazard functions as many bins will have no events, and thus hazard estimates of zero.
+Interestingly, the time bins do not need to have the same width.
+For example, Panis (2020) used larger bins towards the end of the observation period, as fewer events occurred then.
 

### 2025-06-24 manuscript/manuscript.Rmd
update ms.

--- 2025-06-20 manuscript/manuscript.Rmd
@@
 title             : "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows"
-shorttitle        : "A tutorial on hazard analysis"
+shorttitle        : "A tutorial on event history analysis"
 author: 
@@
   Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widespread impact on our understanding of human cognition.
-However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and how they evolve with increasing waiting time.
 The ability to reveal finer-grained, "temporal states" of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
@@
 To help achieve broader uptake of EHA, in this paper we outline a set of tutorials that detail one distributional method known as discrete-time EHA.
-We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design, as well as how to manage inter-individual differences.
+We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design.
 We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
@@
 
-pkg2 <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
+pkg2 <- c("cmdstanr", "standist", 
          "brms", "tidybayes", "bayesplot", "future", "parallel", "rstan")
@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger et al., 2019).
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
@@
 
-As a simple illustration, Figure 1 summarises simulated data that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
+As a simple illustration, Figure 1 summarises simulated single-subject data (200 trials per condition) that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
 
-Why does this matter for research in psychology?
-For many psychological questions, the estimation of such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes and by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
-
-(ref:descr-fig1-caption) Simulated data showing mean performance versus distributional analyses.
+(ref:descr-fig1-caption) Simulated data showing mean performance versus distributional (EHA/SAT) analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 [see also @panisWhatShapingRT2016].
+
+Why does this matter for research in psychology?
+For many psychological questions, the estimation of such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for better temporal methods (ref1, ref2).
+
 ## 1.2 Aims
@@
 Set within this context, our overall aim is to introduce a set of tutorials, which explain **how** to do such analyses in the context of experimental psychology, rather than repeat in any detail **why** you may do them.
-Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHAf in the future.
 
@@
 
-EHA is a statistical approach to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
+EHA is a class of statistical approaches to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
 In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
-We also supply all relevant regression equations in section D of the Supplemental Material.
+We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks in section A of the Supplemental Material, and supply relevant regression equations in section E of the Supplemental Material.
 
@@
 
-To avoid possible confusion in terminology used, it is worth noting that EHA is frequently known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer & Willett, 2003).
+To avoid possible confusion in terminology used, it is worth noting that EHA is known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer & Willett, 2003).
 In this paper, we choose to use the term EHA throughout.
 
-In terms of minimum requirements to apply EHA, one must be able to:
+In terms of minimum requirements to apply a single-event EHA, one must be able to:
 
@@
 
-2. define time point zero (e.g., target stimulus onset, fixation onset, etc.);
+2. define time point zero in each trial (e.g., target stimulus onset, fixation onset, etc.);
 
-3. measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+3. measure the passage of time between time point zero and event occurrence in discrete or continuous time units in each trial.
 
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable. 
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well. 
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than ~200 trials per condition per experiment.
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (REFS). 
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per experiment (REF PANIS).
 Moreover, as indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
 
-To apply discrete time EHA, one divides time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
-Then let t be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular experimental condition. 
-For example, a response on a single trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-One then calculates the discrete-time hazard function of event occurrence.
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular trial (i.e., repeated measure). 
+For example, a response in one trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+One then calculates the sample-based estimate of the discrete-time hazard function of event occurrence for each experimental condition (Figure 1B top).
+The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in
 bin t, given that the event does not occur in previous bins. 
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet
-occurred in the past, i.e., in one of the prior bins.
+In other words, it reflects the instantaneous risk that the event occurs in the current bin t, given that it has not yet
+occurred in the past, i.e., in one of the prior bins (t-1, t-2, ..., 1).
 
 In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
-In those cases, one can also calculate the conditional accuracy function (Figure B).
-The conditional accuracy function gives you for each time bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+In those cases, one can also calculate the conditional accuracy function (Figure 1B bottom).
+The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
+
+The definitions of these and other discrete-time functions is given in section B of the Supplemental Material.
 
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here). 
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (section F of the Supplemental Material). 
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material). 
 Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
@@
 We highlight three benefits that we think are relevant to the domain of experimental psychology. 
-First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about cognition and theoretical development.
+First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard functions and their associated conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about the microgenesis and temporal organization of cognition and theoretical development.
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
-For example, what kind of theory or theories could account for the effects reported in Figure 1B? 
-Are there new auxiliary assumptions that theories need to adopt? 
+For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B? 
+Are there new auxiliary assumptions that computational models need to adopt (ref search)? 
+Will the temporal effect patterns align nicely with EEG findings (ref IOR)?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
 
-Second, compared to more conventional analytical approaches, EHA uses more of the data and deals with msising data differently.
-It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Second, compared to more conventional analytical approaches, EHA uses more of the data because it deals with missing data differently.
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT (REF).
 Discarding data can introduce biases, however.
@@
 Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
-For example, if the censor time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
-The use of right-censoring in EHA, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology.
+For example, if the censoring time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+The fact that EHA can deal with right-censoring, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology (ANOVA, linear regression, delta plots).
 
-Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks (Supp Fig X). 
-EHA can also be used with a range of common experimental manipualtions, such as stimulus-onset-asynchrony (Supp Fig X). 
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology.
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (REFS).
 
@@
 Performing EHA in experimental psychology has implications for how experiments are designed. 
-More specifically, we consider two implications that researchers will need to consider when using discrete time EHA.
+More specifically, we consider three implications that researchers will need to consider when using discrete-time EHA.
 
-First, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+First, one can use a response deadline in each trial because EHA deals with right-censored observations.
+
+Second, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
 Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-*N* designs).
@@
 
-Second, the width of each time bin will need to be determined. 
+Third, the width of each time bin will need to be determined. 
 For instance, in Figure 1B we chose 100ms in an arbitrary manner.
-In reality, however, bin width will need to be set by considering a number of factors simultabeously.
-For example, [[Sven - add a few sentences here]]
+In reality, however, bin width will need to be set by considering a number of factors simultaneously.
+The optimal bin width will depend on (a) the length of the observation period in each trial, (b) the rarity of event occurrence, (c) the number of repeated measures (or trials) per condition per participant, and (d) the shape of the hazard function.
+Finding an appropriate bin width in a given user case before fitting models will require testing a number of options, when calculating and plotting the descriptive statistics (see section 3.1).
+The goal is to find the smallest bin width that is supported by the amount of data available.
+Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
+Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
+Interestingly, the time bins do not need to have the same width.
+For example, Panis (2020) used larger bins towards the end of the observation period, as fewer events occurred there.
 
-# 4. Tutorials
+# 3. Tutorials
 
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section B in the Supplemental Material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C in the Supplemental Material for a list of the custom functions).
 
@@
 
-## 4.1 Tutorial 1a: Calculating descriptive statistics using a life table
+## 3.1 Tutorial 1a: Calculating descriptive statistics using a life table
 
-### 4.1.1 Data wrangling aims
+### 3.1.1 Data wrangling aims
 
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes, using a life table.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+First, we want to summarise and visualise descriptive statistics using a life table.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), possibly the conditional accuracy function ca(t), and their estimated standard errors (se).
+The definitions of these functions are provided in section A of the Supplemental Material.
 
@@
 
-### 4.1.2 A real data wrangling example
+### 3.1.2 A real data wrangling example
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the Supplemental Material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t).
 To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
@@
 
-Note that the censoring time should be a multiple of the bin width (both in ms).
+Note that the censoring time (here: 600 ms) should be a multiple of the bin width (here: 40 ms).
 The censoring time should be a time point after which no informative responses are expected anymore.
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
+
 In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+Also, to select a bin width for fitting models, one should test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
+Too small bin widths will result in erratic hazard functions because many bins will have estimates equal to zero.
 
@@
 
-```{r eha-plot, fig.cap = "(ref:descr-plot4-caption)"}
+```{r eha-plot, fig.cap = "(ref:descr-plot4-caption)", out.width="90%"}
 knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt.png")
@@
 
-## 4.2 Tutorial 1b: Generalising to a more complex design
+## 3.2 Tutorial 1b: Generalising to a more complex design
 
@@
 
-## 4.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data
+## 3.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data
 
@@
 
-### 4.3.1 Hazard model considerations
+### 3.3.1 Hazard model considerations
 
@@
 
-### 4.3.2 Prior distributions
+### 3.3.2 Prior distributions
 
@@
 
-### 4.3.3 Model M0i: A null model with index coding
+### 3.3.3 Model M0i: A null model with index coding
 
@@
 
-### 4.3.4 Model M1i: Adding the effects of prime-target congruency
+### 3.3.4 Model M1i: Adding the effects of prime-target congruency
 
@@
 
-### 4.3.5 Compare the models.
+### 3.3.5 Compare the models.
 
@@
 
-### 4.3.6 Evaluating parameter estimates in model M1i
+### 3.3.6 Evaluating parameter estimates in model M1i
 
@@
 What can we conclude from model M1i about our research question, i.e., the temporal dynamics of the effect of prime-target congruency on RT?
-In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target SOA of 187 ms)?
+In other words, in which of the 40-ms time bins between 200 and 600 ms after target onset does changing the prime from blank to congruent or incongruent affect the hazard of response occurrence (for a prime-target stimulus-onset-asynchrony of 187 ms)?
 
@@
 
-## 4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
+## 3.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
 
@@
 
-## 4.7 Tutorial 4: Planning
+## 3.5 Tutorial 4: Planning
 
@@
 
-### 4.7.1 Background
+### 3.5.1 Background
 
@@
 
-### 4.7.2 Basic workflow
+### 3.5.2 Basic workflow
 
@@
 
-### 4.7.3 Fit a regression model and simulate one dataset
+### 3.5.3 Fit a regression model and simulate one dataset
 
@@
 
-### 4.7.4 Simulate and summarise data across a range of parameter values
+### 3.5.4 Simulate and summarise data across a range of parameter values
 
@@
 
-### 4.7.5 Planning decisions
+### 3.5.5 Planning decisions
 
@@
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss issues relating to model complexity and interpretability, individual differences, as well as limitations of the approach and future extensions.
+In the following, we discuss the main use-cases, issues relating to model complexity and interpretability, as well as limitations of the approach and future extensions.
 
@@
 First, as we hope to have made clear by this point, EHA is one way to investigating a "temporal states" approach to cognitive processes.
-EHA provides one way to uncover when cognitive states may start and stop, as well as what they may be tied to or interact with.
-Therefore, if your research questions concern **when** and **for how long** psychological states occur, our EHA tutorials could be useful tools for you to use.
+EHA provides one way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
+Therefore, if your research questions concern **when psychological states occur, and how they are temporally organized**, our EHA tutorials could be useful tools for you to use.
 
-Second, even if you are not primarily interested in studying the temporal states of cognition, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on mean-average comparisons.
-Given that distinctly different inferences can be made from the same data based on whether one computes a mean-average across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement mean-average comparisons with EHA.
-One could envisage scenarios where the implicit assumption of an effect manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain aspects of time.
+Second, even if you are not primarily interested in studying the temporal organization of cognitive states, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on comparisons between means.
+Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
+One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
+Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.
 
@@
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general.
+From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (REF DFT).
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
@@
 
-Compared to the orthodox method -- comparing mean-averages between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 
 Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
-Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the hazard function.
+Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the discete-time hazard function.
 Therefore, we recommend researchers to collect as many trials as possible per experimental condition, given the available resources and considering the participant experience (e.g., fatigue and boredom).
@@
 Estimating the temporal distributions of RT and accuracy provide a rich source of information on the time course of cognitive processing, which have been largely undervalued in the history of experimental psychology and cognitive neuroscience.
-We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing EHA and investigating the temporal profile of cognitive states.
+We hope that by providing a set of hands-on, step-by-step tutorials, which come with custom-built and freely available code, researchers will feel more comfortable embracing EHA and investigating the shape of empirical hazard functions and the temporal profile of cognitive states.
 On a broader level, we think that wider adoption of such approaches will have a meaningful impact on the inferences drawn from data, as well as the development of theories regarding the structure of cognition.

### 2025-06-24 manuscript/manuscript.tex
update ms.

@@
 
-\shorttitle{A tutorial on hazard analysis}
+\shorttitle{A tutorial on event history analysis}
 
@@
 Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widespread impact on our understanding of human cognition.
-However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration.
+However, the orthodox method for analyzing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and how they evolve with increasing waiting time.
 The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data.
@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
+Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger et al., 2019).
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
@@
 
-As a simple illustration, Figure 1 summarises simulated data that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
+As a simple illustration, Figure 1 summarises simulated single-subject data (200 trials per condition) that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-
-Why does this matter for research in psychology?
-For many psychological questions, the estimation of such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes and by adding a relatively under-used dimension -- the passage of time -- to the theory building toolkit.
-Thus, a distributional approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to better discriminate between different theoretical accounts of psychological and/or brain-based processes.
 
@@
 
-\caption{Simulated data showing mean performance versus distributional analyses.
+\caption{Simulated data showing mean performance versus distributional (EHA/SAT) analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 (see also Panis \& Schmidt, 2016).
+
+Why does this matter for research in psychology?
+For many psychological questions, the estimation of such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for better temporal methods (ref1, ref2).
+
 \subsection{1.2 Aims}\label{aims}
@@
 Set within this context, our overall aim is to introduce a set of tutorials, which explain \textbf{how} to do such analyses in the context of experimental psychology, rather than repeat in any detail \textbf{why} you may do them.
-Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHAf in the future.
 
@@
 
-EHA is a statistical approach to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
+EHA is a class of statistical approaches to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
 In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Singer \& Willett, 2003; Winter, 2019).
-We also supply all relevant regression equations in section D of the Supplemental Material.
+We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks in section A of the Supplemental Material, and supply relevant regression equations in section E of the Supplemental Material.
 
@@
 
-To avoid possible confusion in terminology used, it is worth noting that EHA is frequently known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
+To avoid possible confusion in terminology used, it is worth noting that EHA is known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003).
 In this paper, we choose to use the term EHA throughout.
 
-In terms of minimum requirements to apply EHA, one must be able to:
+In terms of minimum requirements to apply a single-event EHA, one must be able to:
 
@@
 \item
-  define time point zero (e.g., target stimulus onset, fixation onset, etc.);
-\item
-  measure the passage of time between time point zero and event occurrence in discrete or continuous time units.
+  define time point zero in each trial (e.g., target stimulus onset, fixation onset, etc.);
+\item
+  measure the passage of time between time point zero and event occurrence in discrete or continuous time units in each trial.
 \end{enumerate}
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable.
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well.
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological time-to-event data sets for which there are less than \textasciitilde200 trials per condition per experiment.
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (REFS).
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per experiment (REF PANIS).
 Moreover, as indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
 
-To apply discrete time EHA, one divides time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
-Then let t be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
-For example, a response on a single trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-One then calculates the discrete-time hazard function of event occurrence.
-The discrete-time hazard function gives you, for each time bin, the probability that the event occurs (sometime) in
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial (i.e., repeated measure).
+For example, a response in one trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
+One then calculates the sample-based estimate of the discrete-time hazard function of event occurrence for each experimental condition (Figure 1B top).
+The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in
 bin t, given that the event does not occur in previous bins.
-In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet
-occurred in the past, i.e., in one of the prior bins.
+In other words, it reflects the instantaneous risk that the event occurs in the current bin t, given that it has not yet
+occurred in the past, i.e., in one of the prior bins (t-1, t-2, \ldots, 1).
 
 In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
-In those cases, one can also calculate the conditional accuracy function (Figure B).
-The conditional accuracy function gives you for each time bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+In those cases, one can also calculate the conditional accuracy function (Figure 1B bottom).
+The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 
+The definitions of these and other discrete-time functions is given in section B of the Supplemental Material.
+
 \subsection{2.2 Benefits of event history analysis for research in experimental psychology}\label{benefits-of-event-history-analysis-for-research-in-experimental-psychology}
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here).
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (section F of the Supplemental Material).
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material).
 Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
@@
 We highlight three benefits that we think are relevant to the domain of experimental psychology.
-First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard and conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about cognition and theoretical development.
+First, as illustrated in Figure 1, compared to averaging data across trials, integrating results between hazard functions and their associated conditional accuracy functions for choice RT data can be informative for understanding psychological processes, in terms of inferences about the microgenesis and temporal organization of cognition and theoretical development.
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
-For example, what kind of theory or theories could account for the effects reported in Figure 1B?
-Are there new auxiliary assumptions that theories need to adopt?
+For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B?
+Are there new auxiliary assumptions that computational models need to adopt (ref search)?
+Will the temporal effect patterns align nicely with EEG findings (ref IOR)?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
 
-Second, compared to more conventional analytical approaches, EHA uses more of the data and deals with msising data differently.
-It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs before calculating a mean RT (REF).
+Second, compared to more conventional analytical approaches, EHA uses more of the data because it deals with missing data differently.
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT (REF).
 Discarding data can introduce biases, however.
@@
 Right-censoring is a type of missing data problem and a nearly universal feature of survival data including RT data.
-For example, if the censor time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
-The use of right-censoring in EHA, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology.
-
-Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks (Supp Fig X).
-EHA can also be used with a range of common experimental manipualtions, such as stimulus-onset-asynchrony (Supp Fig X).
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology.
+For example, if the censoring time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
+The fact that EHA can deal with right-censoring, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology (ANOVA, linear regression, delta plots).
+
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (REFS).
 
@@
 Performing EHA in experimental psychology has implications for how experiments are designed.
-More specifically, we consider two implications that researchers will need to consider when using discrete time EHA.
-
-First, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
+More specifically, we consider three implications that researchers will need to consider when using discrete-time EHA.
+
+First, one can use a response deadline in each trial because EHA deals with right-censored observations.
+
+Second, since the number of trials per condition are spread across bins, it is important to have a relatively large number of trial repetitions per participant and per condition.
 Accordingly, experimental designs using this approach typically focus on factorial, within-subject designs, in which a large number of observations are made on a relatively small number of participants (so-called small-\emph{N} designs).
@@
 
-Second, the width of each time bin will need to be determined.
+Third, the width of each time bin will need to be determined.
 For instance, in Figure 1B we chose 100ms in an arbitrary manner.
-In reality, however, bin width will need to be set by considering a number of factors simultabeously.
-For example, {[}{[}Sven - add a few sentences here{]}{]}
-
-\section{4. Tutorials}\label{tutorials}
+In reality, however, bin width will need to be set by considering a number of factors simultaneously.
+The optimal bin width will depend on (a) the length of the observation period in each trial, (b) the rarity of event occurrence, (c) the number of repeated measures (or trials) per condition per participant, and (d) the shape of the hazard function.
+Finding an appropriate bin width in a given user case before fitting models will require testing a number of options, when calculating and plotting the descriptive statistics (see section 3.1).
+The goal is to find the smallest bin width that is supported by the amount of data available.
+Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
+Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
+Interestingly, the time bins do not need to have the same width.
+For example, Panis (2020) used larger bins towards the end of the observation period, as fewer events occurred there.
+
+\section{3. Tutorials}\label{tutorials}
 
@@
 
-\subsection{4.1 Tutorial 1a: Calculating descriptive statistics using a life table}\label{tutorial-1a-calculating-descriptive-statistics-using-a-life-table}
-
-\subsubsection{4.1.1 Data wrangling aims}\label{data-wrangling-aims}
+\subsection{3.1 Tutorial 1a: Calculating descriptive statistics using a life table}\label{tutorial-1a-calculating-descriptive-statistics-using-a-life-table}
+
+\subsubsection{3.1.1 Data wrangling aims}\label{data-wrangling-aims}
 
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics that relate to our main research questions about the time course of psychological processes, using a life table.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), P(t), possibly ca(t), and their estimated standard errors (se).
+First, we want to summarise and visualise descriptive statistics using a life table.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), possibly the conditional accuracy function ca(t), and their estimated standard errors (se).
+The definitions of these functions are provided in section A of the Supplemental Material.
 
@@
 
+\begin{verbatim}
+## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
+## Use 'xfun::attr2()' instead.
+## See help("Deprecated")
+\end{verbatim}
+
 \begin{table}[H]
@@
 
-\subsubsection{4.1.2 A real data wrangling example}\label{a-real-data-wrangling-example}
+\subsubsection{3.1.2 A real data wrangling example}\label{a-real-data-wrangling-example}
 
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t) -- see section A of the Supplemental Material for their definitions.
+Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t).
 To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
@@
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
+
 In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+Also, to select a bin width for fitting models, one should test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
+Too small bin widths will result in erratic hazard functions because many bins will have estimates equal to zero.
 
 Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6.
+
+\begin{verbatim}
+## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
+## Use 'xfun::attr2()' instead.
+## See help("Deprecated")
+\end{verbatim}
 
@@
 
-{\centering \includegraphics[width=21.33in,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
+{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt}
 
@@
 
-\subsection{4.2 Tutorial 1b: Generalising to a more complex design}\label{tutorial-1b-generalising-to-a-more-complex-design}
+\subsection{3.2 Tutorial 1b: Generalising to a more complex design}\label{tutorial-1b-generalising-to-a-more-complex-design}
 
@@
 
-\subsection{4.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-discrete-time-to-event-data}
+\subsection{3.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-discrete-time-to-event-data}
 
@@
 
-\subsubsection{4.3.1 Hazard model considerations}\label{hazard-model-considerations}
+\subsubsection{3.3.1 Hazard model considerations}\label{hazard-model-considerations}
 
@@
 
-\subsubsection{4.3.2 Prior distributions}\label{prior-distributions}
+\subsubsection{3.3.2 Prior distributions}\label{prior-distributions}
 
@@
 
-\subsubsection{4.3.3 Model M0i: A null model with index coding}\label{model-m0i-a-null-model-with-index-coding}
+\subsubsection{3.3.3 Model M0i: A null model with index coding}\label{model-m0i-a-null-model-with-index-coding}
 
@@
 
-\subsubsection{4.3.4 Model M1i: Adding the effects of prime-target congruency}\label{model-m1i-adding-the-effects-of-prime-target-congruency}
+\subsubsection{3.3.4 Model M1i: Adding the effects of prime-target congruency}\label{model-m1i-adding-the-effects-of-prime-target-congruency}
 
@@
 
-\subsubsection{4.3.5 Compare the models.}\label{compare-the-models.}
+\subsubsection{3.3.5 Compare the models.}\label{compare-the-models.}
 
@@
 
-\subsubsection{4.3.6 Evaluating parameter estimates in model M1i}\label{evaluating-parameter-estimates-in-model-m1i}
+\subsubsection{3.3.6 Evaluating parameter estimates in model M1i}\label{evaluating-parameter-estimates-in-model-m1i}
 
@@
 
-\subsection{4.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}
+\subsection{3.4 Tutorial 2b: Fitting Bayesian conditional accuracy models}\label{tutorial-2b-fitting-bayesian-conditional-accuracy-models}
 
@@
 
-\subsection{4.7 Tutorial 4: Planning}\label{tutorial-4-planning}
+\subsection{3.5 Tutorial 4: Planning}\label{tutorial-4-planning}
 
@@
 
-\subsubsection{4.7.1 Background}\label{background}
+\subsubsection{3.5.1 Background}\label{background}
 
@@
 
-\subsubsection{4.7.2 Basic workflow}\label{basic-workflow}
+\subsubsection{3.5.2 Basic workflow}\label{basic-workflow}
 
@@
 
-\subsubsection{4.7.3 Fit a regression model and simulate one dataset}\label{fit-a-regression-model-and-simulate-one-dataset}
+\subsubsection{3.5.3 Fit a regression model and simulate one dataset}\label{fit-a-regression-model-and-simulate-one-dataset}
 
@@
 
-\subsubsection{4.7.4 Simulate and summarise data across a range of parameter values}\label{simulate-and-summarise-data-across-a-range-of-parameter-values}
+\subsubsection{3.5.4 Simulate and summarise data across a range of parameter values}\label{simulate-and-summarise-data-across-a-range-of-parameter-values}
 
@@
 
-\subsubsection{4.7.5 Planning decisions}\label{planning-decisions}
+\subsubsection{3.5.5 Planning decisions}\label{planning-decisions}
 
@@
 First, as we hope to have made clear by this point, EHA is one way to investigating a ``temporal states'' approach to cognitive processes.
-EHA provides one way to uncover when cognitive states may start and stop, as well as what they may be tied to or interact with.
-Therefore, if your research questions concern \textbf{when} and \textbf{for how long} psychological states occur, our EHA tutorials could be useful tools for you to use.
-
-Second, even if you are not primarily interested in studying the temporal states of cognition, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on mean-average comparisons.
-Given that distinctly different inferences can be made from the same data based on whether one computes a mean-average across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement mean-average comparisons with EHA.
-One could envisage scenarios where the implicit assumption of an effect manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain aspects of time.
+EHA provides one way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
+Therefore, if your research questions concern \textbf{when psychological states occur, and how they are temporally organized}, our EHA tutorials could be useful tools for you to use.
+
+Second, even if you are not primarily interested in studying the temporal organization of cognitive states, EHA could still be a useful tool to consider using, in order to qualify inferences that are being made based on comparisons between means.
+Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
+One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
+Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.
 

### 2025-06-24 manuscript/Suppl_material/supplementary.Rmd
update Suppl. material

@@
 title             : no
-shorttitle        : "A tutorial on hazard analysis"
+shorttitle        : "A tutorial on event history analysis"
 bibliography      : ["extrareferences2.bib"]
@@
 (B) Two-button discrimination data requires a competing risks analysis.
-(C) Competing risk data in combination with an SOA manipulation might require increasing the observation period in each trial.
+(C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial.
 (D) Two-button bistable perception data requires a recurrent event / multistate analysis.
@@
 
-Studying the effect of SOA on mean RT is still a popular approach to study the time course of cognitive processes (ref).
+Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes (ref).
 For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue).
@@
 
-## B. Benefits of event history analysis
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
-The benefits include:
-
-1. Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
-
-2. RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
-
-3.
-EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
-
-4.
-Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-[@allisonSurvivalAnalysisUsing2010].
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
-
-5.
-EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
-
-## B. Custom functions for descriptive discrete-time hazard analysis
+## C. Custom functions for descriptive discrete-time hazard analysis
 
@@
 
-## C. Link functions
+## D. Link functions
 
@@
 
-## D. Regression equations
+## E. Regression equations
 
@@
 
-## E. Prior distributions
+## F. Prior distributions
 
@@
 
-## F. Advantages of hazard analysis
+## G. Advantages of hazard analysis
 
@@
 
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
+Fourth, hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, pupil dilation, etc.
+[@allisonSurvivalAnalysisUsing2010].
 This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology [@meyerModernMentalChronometry1988].
@@
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived "dips" that could signal the presence of active top-down respone inhibition processes [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
 

### 2025-06-24 manuscript/Suppl_material/supplementary.tex
update Suppl. material

@@
 (B) Two-button discrimination data requires a competing risks analysis.
-(C) Competing risk data in combination with an SOA manipulation might require increasing the observation period in each trial.
+(C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial.
 (D) Two-button bistable perception data requires a recurrent event / multistate analysis.}\label{fig:plot-RT-tasks}
@@
 
-\subsection{B. Benefits of event history analysis}\label{b.-benefits-of-event-history-analysis}
-
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons.
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks.
-Instead, we briefly list the benefits below, and refer the reader to section F of the Supplemental Material for more detailed coverage of the benefits.
-The benefits include:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Hazard functions are more diagnostic than density functions when one is interested in studying the detailed shape of a RT distribution (Holden et al., 2009).
-\item
-  RT distributions may differ from each other in multiple ways, and hazard functions allow one to capture these differences which mean-average comparisons may conceal (Townsend, 1990).
-\item
-  EHA takes account of more of the data collected in a typical speeded response experiment, by virtue of not discarding right-censored observations.
-Trials with very long RTs are not discarded, but instead contribute to the risk set in each time bin (see section 4.1.2 below).
-\item
-  Hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
-(Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
-\item
-  EHA can help to solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
-\end{enumerate}
-
-\subsection{B. Custom functions for descriptive discrete-time hazard analysis}\label{b.-custom-functions-for-descriptive-discrete-time-hazard-analysis}
+\subsection{C. Custom functions for descriptive discrete-time hazard analysis}\label{c.-custom-functions-for-descriptive-discrete-time-hazard-analysis}
 
@@
 
-\subsection{C. Link functions}\label{c.-link-functions}
+\subsection{D. Link functions}\label{d.-link-functions}
 
@@
 
-\subsection{D. Regression equations}\label{d.-regression-equations}
+\subsection{E. Regression equations}\label{e.-regression-equations}
 
@@
 
-\subsection{E. Prior distributions}\label{e.-prior-distributions}
+\subsection{F. Prior distributions}\label{f.-prior-distributions}
 
@@
 
-\subsection{F. Advantages of hazard analysis}\label{f.-advantages-of-hazard-analysis}
+\subsection{G. Advantages of hazard analysis}\label{g.-advantages-of-hazard-analysis}
 
@@
 
-Fourth, hazard modeling allows incorporating time-varying explanatory covariates such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, etc.
+Fourth, hazard modeling allows one to incorporate time-varying explanatory covariates, such as heart rate, electroencephalogram (EEG) signal amplitude, gaze location, pupil dilation, etc.
 (Allison, 2010).
-This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer et al., 1988).
+This is useful for linking physiological effects to behavioral effects when performing cognitive psychophysiology (Meyer, Osman, Irwin, \& Yantis, 1988).
 
@@
 Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived ``dips'' that could signal the presence of active top-down respone inhibition processes (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
+As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
 

### 2025-06-24 manuscript/reviews/ResponseReview1.Rmd
update Response to Reviewers

@@
 author: "Panis & Ramsey"
-date: "12.06.2025"
+date: "24.06.2025"
 output: 
@@
 
-We tried to focus on accessibility in the revision of our Tutorial.
+Response: Thank you for allowing us to submit a revision of our Tutorial (AMPPS-24-0232).
+
+We tried to focus on accessibility in the revision of our Tutorial in several ways. 
+First, we replaced Figure 1 with a simpler Figure that directly conveys our message, and deleted Figures 2 and 3.
+Second, we shortened the Introduction and Discussion sections.
+Third, we added a section in the supplementary material (section A) in which we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception).
+We also mention the unique contribution of our work, we removed redundancies, and tried to improve the clarity of the paper.
 
@@
 
-Response: In the revised ms.
-We added the following section: In this tutorial we provide code to implement discrete-time hazard models in a Bayesian framework, with a focus on cognitive experimental psychology.
+Response: In the revised ms. we frame the unique contribution that this works adds on pages 5-6:
+"...we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which
+provide worked examples of the key data processing and Bayesian multilevel regression modelling steps."
 
@@
 
-Response: On page x of the revised ms., we mention that we estimate single-event occurrences that are repeatedly measured in each participant.
+Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.
 
@@
 
-Response: We disambiguate the various term early in the revised ms. (page x).
+Response: We disambiguate the various term in the revised ms. on page 7, and use one term throughout.
 
@@
 
-Response: We refer to the same time unit in Figure 1 and the text.
+Response: Figure 1 has been replaced.
 
@@
 
-Response: We address the issue of treating time as discrete or continuous on page x of the revised ms.
+Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:
+
+"...the definition of hazard and the type of models employed depend on whether one is using
+continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
+more experience using discrete-time EHA, and that is the approach that we describe and
+focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
+treated as a continuous variable. However, continuous forms of EHA require much more
+data to estimate the continuous-time hazard (rate) function well. Thus, by trading
+a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
+for dealing with typical psychological RT data sets for which there are less than ~200 trials
+per condition per experiment. Moreover, as indicated by Allison (2010),
+learning discrete-time EHA methods first will help in learning continuous-time methods, so
+it seems like a good starting point."
 
@@
 
-Response:
+Response: We moved the explanation of the survival function to the Supplementary Material in the revised ms.
 
@@
 
-Response: We define right-censoring on page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of the revised ms.
 
@@
 
-Response: We discuss the issue of defining bins on page x of the revised ms.
+Response: We discuss the issue of defining bins on page 11 of the revised ms.:
+
+"Third, the width of each time bin will need to be determined. For instance, in Figure
+1B we chose 100ms in an arbitrary manner. In reality, however, bin width will need to be
+set by considering a number of factors simultaneously. The optimal bin width will depend
+on (a) the length of the observation period in each trial, (b) the rarity of event occurrence,
+(c) the number of repeated measures (or trials) per condition per participant, and (d) the
+shape of the hazard function. Finding an appropriate bin width in a given user case before
+fitting models will require testing a number of options, when calculating and plotting the
+descriptive statistics (see section 3.1). The goal is to find the smallest bin width that is
+supported by the amount of data available. Based on our experience, a bin width of 50 ms
+is a good starting value when the number of repeated measures is 100 or less. Too small
+bin widths will result in erratic hazard functions as many bins will have no events, and
+thus hazard estimates of zero. Interestingly, the time bins do not need to have the same
+width. For example, Panis (2020) used larger bins towards the end of the observation
+period, as fewer events occurred there."
 
@@
 
-Response: We include a short comparison of both methods on page x of the revised ms.
+Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:
+
+"...the definition of hazard and the type of models employed depend on whether one is using
+continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
+more experience using discrete-time EHA, and that is the approach that we describe and
+focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
+treated as a continuous variable. However, continuous forms of EHA require much more
+data to estimate the continuous-time hazard (rate) function well. Thus, by trading
+a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
+for dealing with typical psychological RT data sets for which there are less than ~200 trials
+per condition per experiment. Moreover, as indicated by Allison (2010),
+learning discrete-time EHA methods first will help in learning continuous-time methods, so
+it seems like a good starting point."
 
@@
 
-Response: We define (right-)censoring on page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of the revised ms.
 
@@
 
-Response: We added the following section on page x of the revised ms.: …
+Response: We created a new Figure 1 that makes the intended point, and removed section 2.3.2.
 
@@
 
-Response: We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.
+Response: TODO...
+
+We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.
 
@@
 
-Response: We removed the frequentist examples from the main text into the supplementary material.
+Response: We removed the frequentist examples from the main text.
 
@@
 
-Response:
+Response: The old Figure 1 has been replaced.
 
@@
 
-Response:
+Response: TODO
 
@@
 
-Response:
+Response: Figure 2 has been removed. We only discuss the survival function in the Supplemental Material (section B) of the revised ms.
 
@@
 
-Response:
+Response: We removed this section.
 
@@
 
-Response:
+Response: We follow this definition in section B of the Supplemental Material.
 
@@
 
-Response:
+Response: Section 3 has been removed in the revised ms.
 
@@
 
-Response:
+Response: We added data visualizations in section A of the Supplemental Material.
 
@@
 
-Response:
+Response: TODO -> correction needed.
 
@@
 
-Response:
+Response: TODO.
 
@@
 
-Response:
+Response: Yes this has been corrected now.
 
@@
 
-Response: We discuss the main contributions above and beyond existing works on page x of the revised ms.
-We address the issue of treating time as discrete or continuous on page x of the revised ms.
-We removed redundancies  and included more precise explanations of technical jargon.
+Response: We discuss the main contributions above and beyond existing works on pages 5-6 of the revised ms.
+We address the issue of treating time as discrete or continuous on pages 7-8 of the revised ms.
+We removed redundancies, get to the tutorial section quicker (page 11) and included more precise explanations of technical jargon.
 
@@
 
-Response: We explain the location of the files in the README file.
-We included a brief description of a life table and how to interpret the plot in the R Mardown tutorial files.
-A single download step is possible by either cloning the project, or downloading the ZIP file.
+Response: We explain the location of the files in the README file. A single download step is possible by either cloning the project, or downloading the ZIP file.
+We included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
 
@@
 
-Response: This is a tutorial specifically focused on Bayesian methods and cognitive experimental psychology data, such as response times and saccade latencies.
-Analyzing such event latencies with EHA can give better answers about the time course of psychological processes compared to (a) ANOVA on means and (b) currently popular quantile-based distributional methods.
-Information about the time course of cognitive processes is present in the detailed shape of the hazard function which quantifies the moment-to-moment changes in behavioral performance, in terms of the risk of response occurrence, and the accuracy of responses emitted.
-EHA thus allows answering research questions about the onset of psychological effects, about time-locking of effects to stimuli, and about how the effect changes on multiple time scales, i.e., with the passage of time within trials, across trials, across blocks within a session, etc.
-Applying EHA also has implications for experimental design, such as response deadlines, timer on trial onset, etc.
+Response: This tutorial is meant to specifically focus on experimental psychology RT data and this type of data has special characteristics that necessitate their own stand-alone tutorial.
+Therefore, we added visualizations of the types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception) in section A of the Supplemental Material.
 
@@
 
-Response: We believe a stand-alone tutorial for cognitive experimental psychology is necessary because (a) psychologists are reluctant to abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as we know, none of the existing tutorials focuses on Bayesian methods.
+Response: In the revised ms. we frame the unique contribution that this works adds on pages 5-6:
+"...we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which
+provide worked examples of the key data processing and Bayesian multilevel regression modelling steps."
 
@@
 
-Response: We focus primarily on the Bayesian models in the revised ms.  We also made major cuts to both the introduction and the discussion sections.
+Response: We focus primarily on the Bayesian models in the revised ms., and keep Tutorial 4.  We also made major cuts to both the introduction and the discussion sections to shorten the ms.
 
@@
 
-Response: We provide clear descriptions of censoring and risk sets on page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.
 
@@
 
-Response: We placed the equations for h(t), S(t), and ca(t) in the revised ms. on page x. The equations for P(t) are still in the supplementary material.
+Response: We placed the equations for h(t) and ca(t) in Figure 1 of the revised ms.
 
@@
 
-Response:
+Response: Indeed, this was a big error on our part. Thank you for pointing this out.
 
@@
 
-Response:
+Response: TODO. We add indexes for individual and trial.
 
@@
 
-Response: We added reference to individuals in the notation.
+Response: TODO. We added reference to individuals and trials (repeated measures) in the notation.
 
@@
 
-Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood.
+Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood. We replaced the term "likelihood" with "risk".
 
@@
 
-Response:
+Response: We rewrote this section.
 
@@
 
-Response: We introduce SAT analysis already on page x in the revised ms.
+Response: We introduce conditional accuracy functions already in Figure 1 of the revised ms.
 
@@
 
-Response:
+Response: We introduce the definition of right-censoring on page 10 of the revised ms.
 
@@
 
-Response: We included the following text on page x of the revised ms.: This tutorial deals with single-event analyses which focus on the time until a single event occurs in a trial (e.g., a button press), with repeated measures obtained for each individual.
-Recurrent event analysis deals with multiple events that can happen repeatedly to the same individual within a single observation period.
+Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.
+In section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events.
 
@@
 
-Response:
+Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:
+
+"...the definition of hazard and the type of models employed depend on whether one is using
+continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
+more experience using discrete-time EHA, and that is the approach that we describe and
+focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
+treated as a continuous variable. However, continuous forms of EHA require much more
+data to estimate the continuous-time hazard (rate) function well. Thus, by trading
+a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
+for dealing with typical psychological RT data sets for which there are less than ~200 trials
+per condition per experiment. Moreover, as indicated by Allison (2010),
+learning discrete-time EHA methods first will help in learning continuous-time methods, so
+it seems like a good starting point."
 
@@
 
-Response:
+Response: We provided a clearer description of the data.
 
@@
 
-Response: We provide supporting references on page x of the revised ms.
+Response: TODO. We provide supporting references on page x of the revised ms.
 
@@
 
-Response: We follow the suggestion of this reviewer to start the tutorial at section 2.3.
-That is, we immediately provide an example experiment/data set with a descriptive analysis that illustrates the benefits of applying EHA to psychological time-to-event data such as response times and saccade latencies.
-This new example data set is taken from Panis & Schmidt (2022).
+Response: Section 2.3 has been rewritten. The new Figure 1 in the tutorial directly focuses  on the comparison between analysing means vs. event occurrence.
 
@@
 
-Response:
+Response: Section 2.3.1 has been removed.
 
@@
 
-Response:
+Response: We explain the idea of temporal states of cognitive processes now on page 5 of the revised ms. and illustrate it in Figure 1.
 
@@
 
-Response:
+Response: Section 3 has been removed.
 
@@
 
-Response: We restructured and shortened the introduction.
+Response: We agree, and restructured and shortened the introduction.
 
@@
 
-Response: P(t)  is defined only in the supplementary material.
+Response: P(t) was and still is defined only in the Supplementary Material (section B).
 
@@
 
-Response: We clarify this on page x of the revised ms.
+Response: TODO.
 
@@
 
-Response:
+Response: This connection should be clear now based on section A of the Supplemental Material, and based on the text on pages 7 and 8.
 
@@
 
-Response: We introduce types of censoring on page x of the revised ms.
+Response: We introduce various types of censoring in section A of the Supplemental Material.
 
@@
 
-Response: We provide the information on data descriptions in a single place (page x of the revised ms.).
+Response: We provide the information on data descriptions in a single place in the revised ms.
 
@@
 
-Response:
+Response: TODO. We agree and adjusted the description.
 
@@
 
-Response:
+Response: On line 362 it was stated that "the warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored".
+
+We removed these messages from the paper and mention them now only in the tutorial itself.
 
@@
 
-Response: We unpack in more detail how individuals can be identified that may be guessing, or that do not perform the task as intended.
+Response: TODO.
 
@@
 
-Response:
+Response: A comparison between the cloglog and logit link functions is provided in section C of the Supplemental Material.
 
@@
 
-Response: We provide a brief description of the rationale of comparing models M0i and M1i on page x of the revised ms.
+Response: TODO.
 
@@
 
-Response:
+Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in experimental psychology.
 
@@
 
-Response:
+Response: We decided to keep this example, as it can help readers with interpreting effects in hazard and conditional accuracy functions.
 
@@
 
-Response:
+Response: We decided not to include more detail because logistic regression is very familiar to experimental psychologists.
 
@@
 
-Response:
+Response: We decided to keep tutorial 4, and discuss its usefulness on page 45 of the revised ms.
 
@@
 
+Nevertheless, we replaced this with "how effects evolve with increasing waiting time".
+
 > pg. 47 (lines 753-761) – I had a difficult time understanding this section
 
-Response: We rewrote this section.
+Response: TODO.
 
@@
 
-Response: We provide a better example in the revised ms.
+Response: We removed section 5.5.
 
@@
 
-Response:
+Response: We created a new Figure 1.
 
@@
 
-Response:
+Response: Figure 2 has been removed.

### 2025-06-24 manuscript/reviews/ResponseReview1.tex
update Response to Reviewers

--- 2025-06-14 manuscript/ResponseReview1.tex
@@
 \author{Panis \& Ramsey}
-\date{12.06.2025}
+\date{24.06.2025}
 
@@
 
-Thank you for submitting your Tutorial (AMPPS-24-0232) entitled ``Event
-History Analysis for psychological time-to-event data: A tutorial in R
-with examples in Bayesian and frequentist workflows'' to Advances in
-Methods and Practices in Psychological Science (AMPPS). I have now
-received input from three experts in the field, and you will find the
-reviews below this message and attached. As you will see, all three
-reviewers believe your paper can make a strong contribution to the
-literature. Prior to considering the reviewers' input, I read your paper
-to form an independent opinion. I share the positive sentiments about
-this work and invite you to revise and resubmit your manuscript for
-further consideration at AMPPS.
+Thank you for submitting your Tutorial entitled ``Event History Analysis
+for psychological time-to-event data: A tutorial in R with examples in
+Bayesian and frequentist workflows'' . I have now received input from
+three experts in the field, and you will find the reviews below this
+message and attached. As you will see, all three reviewers believe your
+paper can make a strong contribution to the literature. Prior to
+considering the reviewers' input, I read your paper to form an
+independent opinion. I share the positive sentiments about this work and
+invite you to revise and resubmit your manuscript for further
+consideration.
 \end{quote}
@@
 \begin{quote}
-To submit your revision, log into
-\url{http://mc.manuscriptcentral.com/ampps} and enter your Author
-Center, where you will find your manuscript title listed under
-``Manuscripts with Decisions.'' Under ``Actions,'' click on ``Create a
-Revision.'' Your manuscript number has been appended to denote a
-revision.
-\end{quote}
-
-\begin{quote}
-IMPORTANT: Your original files are available to you when you upload your
-revised manuscript. Please delete any redundant or outdated files before
-completing the submission.
-\end{quote}
-
-\begin{quote}
-Once again, thank you for submitting your manuscript to AMPPS and I look
-forward to receiving your revision.
-\end{quote}
-
-\begin{quote}
 Sincerely,
@@
 
-\begin{quote}
-David A. Sbarra, Ph.D. Editor, Advances in Methods and Practices in
-Psychological Science
-\end{quote}
-
-We tried to focus on accessibility in the revision of our Tutorial
-AMPPS-24-0232.
+Response: Thank you for allowing us to submit a revision of our Tutorial
+(AMPPS-24-0232).
+
+We tried to focus on accessibility in the revision of our Tutorial in
+several ways. First, we replaced Figure 1 with a simpler Figure that
+directly conveys our message, and deleted Figures 2 and 3. Second, we
+shortened the Introduction and Discussion sections. Third, we added a
+section in the supplementary material (section A) in which we visualize
+the different types of time-to-event data that are obtained in typical
+RT tasks (detection, discrimination, bistable perception). We also
+mention the unique contribution of our work, we removed redundancies,
+and tried to improve the clarity of the paper.
 
@@
 
-Response: In the revised ms. We added the following section: In this
-tutorial we provide code to implement discrete-time hazard models in a
-Bayesian framework, with a focus on cognitive experimental psychology.
+Response: In the revised ms. we frame the unique contribution that this
+works adds on pages 5-6: ``\ldots we are not aware of any tutorials that
+are aimed specifically at psychological RT (+ accuracy) data, and which
+provide worked examples of the key data processing and Bayesian
+multilevel regression modelling steps.''
 
@@
 
-Response: On page x of the revised ms., we mention that we estimate
+Response: On page 7 of the revised ms., we mention that we measure
 single-event occurrences that are repeatedly measured in each
@@
 
-Response: We disambiguate the various term early in the revised ms.
-(page x).
+Response: We disambiguate the various term in the revised ms. on page 7,
+and use one term throughout.
 
@@
 
-Response: We address the issue of treating time as discrete or
-continuous on page x of the revised ms.
+Response: We address our choice for discrete-time methods on pages 7-8
+of the revised ms.:
+
+``\ldots the definition of hazard and the type of models employed depend
+on whether one is using continuous or discrete time units. As a lab, and
+mainly for practical reasons, we have much more experience using
+discrete-time EHA, and that is the approach that we describe and focus
+on in this paper. This choice may seem counter-intuitive, given that RT
+is typically treated as a continuous variable. However, continuous forms
+of EHA require much more data to estimate the continuous-time hazard
+(rate) function well. Thus, by trading a bit of temporal resolution for
+a lower number of trials, discrete-time methods seem ideal for dealing
+with typical psychological RT data sets for which there are less than
+\textasciitilde200 trials per condition per experiment. Moreover, as
+indicated by Allison (2010), learning discrete-time EHA methods first
+will help in learning continuous-time methods, so it seems like a good
+starting point.''
 
@@
 
-Response:
+Response: We moved the explanation of the survival function to the
+Supplementary Material in the revised ms.
 
@@
 
-Response: We define right-censoring on page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of
+the revised ms.
 
@@
 
-Response: We discuss the issue of defining bins on page x of the revised
-ms.
+Response: We discuss the issue of defining bins on page 11 of the
+revised ms.:
+
+``Third, the width of each time bin will need to be determined. For
+instance, in Figure 1B we chose 100ms in an arbitrary manner. In
+reality, however, bin width will need to be set by considering a number
+of factors simultaneously. The optimal bin width will depend on (a) the
+length of the observation period in each trial, (b) the rarity of event
+occurrence, (c) the number of repeated measures (or trials) per
+condition per participant, and (d) the shape of the hazard function.
+Finding an appropriate bin width in a given user case before fitting
+models will require testing a number of options, when calculating and
+plotting the descriptive statistics (see section 3.1). The goal is to
+find the smallest bin width that is supported by the amount of data
+available. Based on our experience, a bin width of 50 ms is a good
+starting value when the number of repeated measures is 100 or less. Too
+small bin widths will result in erratic hazard functions as many bins
+will have no events, and thus hazard estimates of zero. Interestingly,
+the time bins do not need to have the same width. For example, Panis
+(2020) used larger bins towards the end of the observation period, as
+fewer events occurred there.''
 
@@
 
-Response: We include a short comparison of both methods on page x of the
-revised ms.
+Response: We address our choice for discrete-time methods on pages 7-8
+of the revised ms.:
+
+``\ldots the definition of hazard and the type of models employed depend
+on whether one is using continuous or discrete time units. As a lab, and
+mainly for practical reasons, we have much more experience using
+discrete-time EHA, and that is the approach that we describe and focus
+on in this paper. This choice may seem counter-intuitive, given that RT
+is typically treated as a continuous variable. However, continuous forms
+of EHA require much more data to estimate the continuous-time hazard
+(rate) function well. Thus, by trading a bit of temporal resolution for
+a lower number of trials, discrete-time methods seem ideal for dealing
+with typical psychological RT data sets for which there are less than
+\textasciitilde200 trials per condition per experiment. Moreover, as
+indicated by Allison (2010), learning discrete-time EHA methods first
+will help in learning continuous-time methods, so it seems like a good
+starting point.''
 
@@
 
-Response: We define (right-)censoring on page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of
+the revised ms.
 
@@
 
-Response: We added the following section on page x of the revised ms.:
-\ldots{}
+Response: We created a new Figure 1 that makes the intended point, and
+removed section 2.3.2.
 
@@
 
-Response: We explain model weights on page x, include a reference for
-dWAIC, and report the weights with 3 digits.
+Response: TODO\ldots{}
+
+We explain model weights on page x, include a reference for dWAIC, and
+report the weights with 3 digits.
 
@@
 
-Response: We removed the frequentist examples from the main text into
-the supplementary material.
+Response: We removed the frequentist examples from the main text.
 
@@
 
-Response:
+Response: Figure 2 has been removed. We only discuss the survival
+function in the Supplemental Material (section B) of the revised ms.
 
@@
 
-Response:
+Response: We follow this definition in section B of the Supplemental
+Material.
 
@@
 
-Response:
+Response: We added data visualizations in section A of the Supplemental
+Material.
 
@@
 
-Response:
+Response: TODO -\textgreater{} correction needed.
 
@@
 Response: We discuss the main contributions above and beyond existing
-works on page x of the revised ms. We address the issue of treating time
-as discrete or continuous on page x of the revised ms. We removed
-redundancies and included more precise explanations of technical jargon.
+works on pages 5-6 of the revised ms. We address the issue of treating
+time as discrete or continuous on pages 7-8 of the revised ms. We
+removed redundancies, get to the tutorial section quicker (page 11) and
+included more precise explanations of technical jargon.
 
@@
 
-Response: We explain the location of the files in the README file. We
-included a brief description of a life table and how to interpret the
-plot in the R Mardown tutorial files. A single download step is possible
-by either cloning the project, or downloading the ZIP file.
+Response: We explain the location of the files in the README file. A
+single download step is possible by either cloning the project, or
+downloading the ZIP file. We included a brief description of a life
+table and how to interpret the plot in the R Markdown tutorial files.
 
@@
 
-Response: This is a tutorial specifically focused on Bayesian methods
-and cognitive experimental psychology data, such as response times and
-saccade latencies. Analyzing such event latencies with EHA can give
-better answers about the time course of psychological processes compared
-to (a) ANOVA on means and (b) currently popular quantile-based
-distributional methods. Information about the time course of cognitive
-processes is present in the detailed shape of the hazard function which
-quantifies the moment-to-moment changes in behavioral performance, in
-terms of the risk of response occurrence, and the accuracy of responses
-emitted. EHA thus allows answering research questions about the onset of
-psychological effects, about time-locking of effects to stimuli, and
-about how the effect changes on multiple time scales, i.e., with the
-passage of time within trials, across trials, across blocks within a
-session, etc. Applying EHA also has implications for experimental
-design, such as response deadlines, timer on trial onset, etc.
+Response: This tutorial is meant to specifically focus on experimental
+psychology RT data and this type of data has special characteristics
+that necessitate their own stand-alone tutorial. Therefore, we added
+visualizations of the types of time-to-event data that are obtained in
+typical RT tasks (detection, discrimination, bistable perception) in
+section A of the Supplemental Material.
 
@@
 
-Response: We believe a stand-alone tutorial for cognitive experimental
-psychology is necessary because (a) psychologists are reluctant to
-abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as
-we know, none of the existing tutorials focuses on Bayesian methods.
+Response: In the revised ms. we frame the unique contribution that this
+works adds on pages 5-6: ``\ldots we are not aware of any tutorials that
+are aimed specifically at psychological RT (+ accuracy) data, and which
+provide worked examples of the key data processing and Bayesian
+multilevel regression modelling steps.''
 
@@
 
-Response: We focus primarily on the Bayesian models in the revised ms.
-We also made major cuts to both the introduction and the discussion
-sections.
+Response: We focus primarily on the Bayesian models in the revised ms.,
+and keep Tutorial 4. We also made major cuts to both the introduction
+and the discussion sections to shorten the ms.
 
@@
 
-Response: We provide clear descriptions of censoring and risk sets on
-page x of the revised ms.
+Response: We introduce the definition of right-censoring on page 10 of
+the revised ms. We discuss what it means to be at risk in section A of
+the Supplemental Material.
 
@@
 
-Response: We placed the equations for h(t), S(t), and ca(t) in the
-revised ms. on page x. The equations for P(t) are still in the
-supplementary material.
+Response: We placed the equations for h(t) and ca(t) in Figure 1 of the
+revised ms.
 
@@
 
-Response:
+Response: Indeed, this was a big error on our part. Thank you for
+pointing this out.
 
@@
 
-Response: We added reference to individuals in the notation.
+Response: TODO. We added reference to individuals and trials (repeated
+measures) in the notation.
 
@@
 Response: Indeed, only continuous-time hazard can be called an
-instantaneous likelihood.
+instantaneous likelihood. We replaced the term ``likelihood'' with
+``risk''.
 
@@
 
-Response: We introduce SAT analysis already on page x in the revised ms.
+Response: We introduce conditional accuracy functions already in Figure
+1 of the revised ms.
 
@@
 
-Response:
+Response: We introduce the definition of right-censoring on page 10 of
+the revised ms.
 
@@
 
-Response: We included the following text on page x of the revised ms.:
-This tutorial deals with single-event analyses which focus on the time
-until a single event occurs in a trial (e.g., a button press), with
-repeated measures obtained for each individual. Recurrent event analysis
-deals with multiple events that can happen repeatedly to the same
-individual within a single observation period.
+Response: On page 7 of the revised ms., we mention that we measure
+single-event occurrences that are repeatedly measured in each
+participant. In section A of the Supplemental Material we mention that
+this Tutorial does not cover recurrent events.
 
@@
 
-Response:
+Response: We address our choice for discrete-time methods on pages 7-8
+of the revised ms.:
+
+``\ldots the definition of hazard and the type of models employed depend
+on whether one is using continuous or discrete time units. As a lab, and
+mainly for practical reasons, we have much more experience using
+discrete-time EHA, and that is the approach that we describe and focus
+on in this paper. This choice may seem counter-intuitive, given that RT
+is typically treated as a continuous variable. However, continuous forms
+of EHA require much more data to estimate the continuous-time hazard
+(rate) function well. Thus, by trading a bit of temporal resolution for
+a lower number of trials, discrete-time methods seem ideal for dealing
+with typical psychological RT data sets for which there are less than
+\textasciitilde200 trials per condition per experiment. Moreover, as
+indicated by Allison (2010), learning discrete-time EHA methods first
+will help in learning continuous-time methods, so it seems like a good
+starting point.''
 
@@
 
-Response: We provide supporting references on page x of the revised ms.
+Response: TODO. We provide supporting references on page x of the
+revised ms.
 
@@
 
-Response: We follow the suggestion of this reviewer to start the
-tutorial at section 2.3. That is, we immediately provide an example
-experiment/data set with a descriptive analysis that illustrates the
-benefits of applying EHA to psychological time-to-event data such as
-response times and saccade latencies. This new example data set is taken
-from Panis \& Schmidt (2022).
+Response: Section 2.3 has been rewritten. The new Figure 1 in the
+tutorial directly focuses on the comparison between analysing means
+vs.~event occurrence.
 
@@
 
-Response:
+Response: We explain the idea of temporal states of cognitive processes
+now on page 5 of the revised ms. and illustrate it in Figure 1.
 
@@
 
-Response: P(t) is defined only in the supplementary material.
+Response: P(t) was and still is defined only in the Supplementary
+Material (section B).
 
@@
 
-Response:
+Response: This connection should be clear now based on section A of the
+Supplemental Material, and based on the text on pages 7 and 8.
 
@@
 
-Response: We introduce types of censoring on page x of the revised ms.
+Response: We introduce various types of censoring in section A of the
+Supplemental Material.
 
@@
 Response: We provide the information on data descriptions in a single
-place (page x of the revised ms.).
+place in the revised ms.
 
@@
 
-Response:
+Response: On line 362 it was stated that ``the warning messages are
+generated because some bins have no hazard and ca(t) estimates, and no
+error bars. They can thus safely be ignored''.
+
+We removed these messages from the paper and mention them now only in
+the tutorial itself.
 
@@
 
-Response: We unpack in more detail how individuals can be identified
-that may be guessing, or that do not perform the task as intended.
+Response: TODO.
 
@@
 
-Response:
+Response: A comparison between the cloglog and logit link functions is
+provided in section C of the Supplemental Material.
 
@@
 
-Response: We provide a brief description of the rationale of comparing
-models M0i and M1i on page x of the revised ms.
+Response: TODO.
 
@@
 
-Response:
+Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in
+experimental psychology.
 
@@
 
-Response:
+Response: We decided to keep this example, as it can help readers with
+interpreting effects in hazard and conditional accuracy functions.
 
@@
 
-Response:
+Response: We decided not to include more detail because logistic
+regression is very familiar to experimental psychologists.
 
@@
 
-Response:
+Response: We decided to keep tutorial 4, and discuss its usefulness on
+page 45 of the revised ms.
 
@@
 
+Nevertheless, we replaced this with ``how effects evolve with increasing
+waiting time''.
+
 \begin{quote}
@@
 
-Response: We rewrote this section.
+Response: TODO.
 
@@
 
-Response:
+Response: Figure 2 has been removed.
 

### 2025-06-24 sims/sims.Rmd
update Figure 1

@@
  # ggtitle("hazard and conditional accuracy functions") +
-  xlab("time bin endpoint (ms)") +
+  xlab("time bin t's endpoint (ms)") +
   ylab("hazard") +
@@
   scale_y_continuous(limits = c(0,1)) +
-  xlab("time bin endpoint (ms)") +
+  xlab("time bin t's endpoint (ms)") +
     ylab("conditional accuracy") +

### 2025-06-25 manuscript/manuscript.Rmd
update ms. + add references

@@
 # R references
-# r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
-# my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)
+ r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation/ creates a .bib-reference file
+ my_r_citation <- cite_r(file = "r-references.bib", footnote = FALSE)
 ```
@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger et al., 2019).
+Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology [@bolgerCausalProcessesPsychology2019].
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
@@
 For many psychological questions, the estimation of such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
-Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for better temporal methods (ref1, ref2).
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for a temporal science of behavior [@ABNEY2025].
 
@@
 
-Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data.
+Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
 
@@
 Set within this context, our overall aim is to introduce a set of tutorials, which explain **how** to do such analyses in the context of experimental psychology, rather than repeat in any detail **why** you may do them.
-Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHAf in the future.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
 
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
-We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks in section A of the Supplemental Material, and supply relevant regression equations in section E of the Supplemental Material.
+We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks and supply relevant regression equations in sections A and E of the Supplemental Material, respectively.
 
@@
 For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. 
-As a lab, and mainly for practical reasons, we have much more experience using discrete time EHA, and that is the approach that we describe and focus on in this paper. 
+As a lab, and mainly for practical reasons, we have much more experience using discrete-time EHA, and that is the approach that we describe and focus on in this paper. 
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable. 
 However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (REFS). 
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per experiment (REF PANIS).
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per experiment [@panisAnalyzingResponseTimes2020].
 Moreover, as indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
@@
 We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
-
-The definitions of these and other discrete-time functions is given in section B of the Supplemental Material.
+The definitions of these and other discrete-time functions are given in section B of the Supplemental Material.
 
@@
 
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here). 
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons [@holdenDispersionResponseTimes2009; @townsendTruthConsequencesOrdinal1990; @luceResponseTimesTheir1991].
 We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material). 
@@
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
-For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B? 
-Are there new auxiliary assumptions that computational models need to adopt (ref search)? 
-Will the temporal effect patterns align nicely with EEG findings (ref IOR)?
+For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B [@panisWhatShapingRT2016]? 
+Are there new auxiliary assumptions that computational models need to adopt [@panisStudyingDynamicsVisual2020]? 
+Will the temporal effect patterns align nicely with EEG findings [@panisWhenDoesInhibition2022]?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
@@
 Second, compared to more conventional analytical approaches, EHA uses more of the data because it deals with missing data differently.
-It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT (REF).
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT [@bergerComparisonDifferentResponse2021].
 Discarding data can introduce biases, however.
@@
 Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (REFS).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition [@barackTwoViewsCognitive2021].
 
@@
 Interestingly, the time bins do not need to have the same width.
-For example, Panis (2020) used larger bins towards the end of the observation period, as fewer events occurred there.
+For example, @panisHowCanWe2020 used larger bins towards the end of the observation period, as fewer events occurred there.
 
 # 3. Tutorials
-
-We used r my_r_citation$r for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 Tutorials 3a and 3b show how to implement, respectively, multilevel models for hazard and conditional accuracy in the frequentist framework.
+Tutorial 4 shows how to use simulation and power analysis for planning experiments.
 Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C in the Supplemental Material for a list of the custom functions).
@@
 * 4.  Simulation and power analysis for planning experiments
+
+We used `r my_r_citation` for all reported analyses.
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 

### 2025-06-25 manuscript/manuscript.tex
update ms. + add references

@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger et al., 2019).
+Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger, Zee, Rossignac-Milon, \& Hassin, 2019).
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
@@
 For many psychological questions, the estimation of such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
-Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for better temporal methods (ref1, ref2).
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for a temporal science of behavior (Abney, Fausey, Suarez-Rivera, \& Tamis-LeMonda, 2025).
 
@@
 Set within this context, our overall aim is to introduce a set of tutorials, which explain \textbf{how} to do such analyses in the context of experimental psychology, rather than repeat in any detail \textbf{why} you may do them.
-Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHAf in the future.
+Therefore, we hope that our tutorials will provide a pathway for research avenues in experimental psychology that have the potential to benefit from using EHA in the future.
 
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Singer \& Willett, 2003; Winter, 2019).
-We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks in section A of the Supplemental Material, and supply relevant regression equations in section E of the Supplemental Material.
+We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks and supply relevant regression equations in sections A and E of the Supplemental Material, respectively.
 
@@
 For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
-As a lab, and mainly for practical reasons, we have much more experience using discrete time EHA, and that is the approach that we describe and focus on in this paper.
+As a lab, and mainly for practical reasons, we have much more experience using discrete-time EHA, and that is the approach that we describe and focus on in this paper.
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable.
@@
 
-Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (REF?? - Sven please cite a relevant REF here).
+Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (Holden, Van Orden, \& Turvey, 2009; Luce, 1991; Townsend, 1990).
 We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material).
@@
 As such, the approach permits different kinds of questions to be asked, different inferences to be made, and it holds the potential to discriminate between theoretical accounts of psychological and/or brain-based processes.
-For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B?
-Are there new auxiliary assumptions that computational models need to adopt (ref search)?
-Will the temporal effect patterns align nicely with EEG findings (ref IOR)?
+For example, what kind of theory or set of mechanisms could account for the shape of the functions and the temporally localized effects reported in Figure 1B (Panis \& Schmidt, 2016)?
+Are there new auxiliary assumptions that computational models need to adopt (Panis, Moran, et al., 2020)?
+Will the temporal effect patterns align nicely with EEG findings (Panis \& Schmidt, 2022)?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
@@
 Second, compared to more conventional analytical approaches, EHA uses more of the data because it deals with missing data differently.
-It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT (REF).
+It is conventional with RT data to either (a) use a response deadline and discard all trials without a response, or (b) wait in each trial until a response occurs and then apply data trimming techniques, i.e., discarding too short or too long RTs (and perhaps also erroneous responses) before calculating a mean RT (Berger \& Kiefer, 2021).
 Discarding data can introduce biases, however.
@@
 Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (REFS).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (Barack \& Krakauer, 2021).
 
@@
 
-We used r my\_r\_citation\$r for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-
 Tutorials 1a and 1b show how to calculate and plot the descriptive statistics of EHA/SAT when there are one or two independent variables, respectively.
@@
 \end{itemize}
+
+We used R (Version 4.4.0; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023) for all reported analyses.
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-ABNEY2025}
+Abney, D.
+H., Fausey, C.
+M., Suarez-Rivera, C., \& Tamis-LeMonda, C.
+S.
+(2025).
+Advancing a temporal science of behavior.
+\emph{Trends in Cognitive Sciences}.
+\url{https://doi.org/10.1016/j.tics.2025.05.010}
+
 \bibitem[\citeproctext]{ref-allisonDiscreteTimeMethodsAnalysis1982}
@@
 Allison, P. D. (2010). \emph{Survival analysis using {SAS}: A practical guide} (2. ed). Cary, NC: SAS Press.
+
+\bibitem[\citeproctext]{ref-R-citr}
+Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://github.com/crsh/citr}
+
+\bibitem[\citeproctext]{ref-R-papaja}
+Aust, F., \& Barth, M. (2024). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://doi.org/10.32614/CRAN.package.papaja}
 
@@
 
+\bibitem[\citeproctext]{ref-barackTwoViewsCognitive2021}
+Barack, D. L., \& Krakauer, J. W. (2021). Two views on the cognitive brain. \emph{Nature Reviews Neuroscience}, \emph{22}(6), 359--371. \url{https://doi.org/10.1038/s41583-021-00448-6}
+
 \bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
@@
 
+\bibitem[\citeproctext]{ref-R-tinylabels}
+Barth, M. (2023). \emph{{tinylabels}: Lightweight variable labels}. Retrieved from \url{https://cran.r-project.org/package=tinylabels}
+
+\bibitem[\citeproctext]{ref-R-lme4}
+Bates, D., Mächler, M., Bolker, B., \& Walker, S.
+(2015).
+Fitting linear mixed-effects models using {lme4}.
+\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
+\url{https://doi.org/10.18637/jss.v067.i01}
+
+\bibitem[\citeproctext]{ref-R-Matrix}
+Bates, D., Maechler, M., \& Jagan, M. (2024). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://Matrix.R-forge.R-project.org}
+
+\bibitem[\citeproctext]{ref-R-RJ-2021-048}
+Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+
+\bibitem[\citeproctext]{ref-bergerComparisonDifferentResponse2021}
+Berger, A., \& Kiefer, M.
+(2021).
+Comparison of {Different Response Time Outlier Exclusion Methods}: {A Simulation Study}.
+\emph{Frontiers in Psychology}, \emph{12}, 675558.
+\url{https://doi.org/10.3389/fpsyg.2021.675558}
+
 \bibitem[\citeproctext]{ref-blossfeldTechniquesEventHistory2002}
@@
 
+\bibitem[\citeproctext]{ref-bolgerCausalProcessesPsychology2019}
+Bolger, N., Zee, K.
+S., Rossignac-Milon, M., \& Hassin, R.
+R.
+(2019).
+Causal processes in psychology are heterogeneous.
+\emph{Journal of Experimental Psychology: General}, \emph{148}(4), 601--618.
+\url{https://doi.org/10.1037/xge0000558}
+
 \bibitem[\citeproctext]{ref-box-steffensmeierEventHistoryModeling2004}
 Box-Steffensmeier, J. M. (2004). Event history modeling: A guide for social scientists. Cambridge: University Press.
+
+\bibitem[\citeproctext]{ref-R-brms_a}
+Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\bibitem[\citeproctext]{ref-R-brms_b}
+Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411. \url{https://doi.org/10.32614/RJ-2018-017}
+
+\bibitem[\citeproctext]{ref-R-brms_c}
+Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with {brms} and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}
 
@@
 
+\bibitem[\citeproctext]{ref-R-Rcpp_b}
+Eddelbuettel, D., \& Balamuta, J.
+J.
+(2018).
+{Extending {R} with {C++}: A Brief Introduction to {Rcpp}}.
+\emph{The American Statistician}, \emph{72}(1), 28--36.
+\url{https://doi.org/10.1080/00031305.2017.1375990}
+
+\bibitem[\citeproctext]{ref-R-Rcpp_a}
+Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. \url{https://doi.org/10.18637/jss.v040.i08}
+
+\bibitem[\citeproctext]{ref-R-cmdstanr}
+Gabry, J., Češnovar, R., Johnson, A., \& Bronder, S. (2024). \emph{Cmdstanr: R interface to 'CmdStan'}. Retrieved from \url{https://github.com/stan-dev/cmdstanr}
+
+\bibitem[\citeproctext]{ref-R-bayesplot}
+Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in bayesian workflow. \emph{J. R. Stat. Soc. A}, \emph{182}, 389--402. \url{https://doi.org/10.1111/rssa.12378}
+
 \bibitem[\citeproctext]{ref-gelmanRegressionOtherStories2020}
@@
 
+\bibitem[\citeproctext]{ref-R-standist}
+Girard, J. (2024). \emph{Standist: What the package does (one line, title case)}. Retrieved from \url{https://github.com/jmgirard/standist}
+
+\bibitem[\citeproctext]{ref-R-lubridate}
+Grolemund, G., \& Wickham, H. (2011). Dates and times made easy with {lubridate}. \emph{Journal of Statistical Software}, \emph{40}(3), 1--25. Retrieved from \url{https://www.jstatsoft.org/v40/i03/}
+
 \bibitem[\citeproctext]{ref-heiss2021}
 Heiss, A. (2021, November 10). A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel {Bayesian} Models. \url{https://doi.org/10.59350/wbn93-edb02}
+
+\bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
+Holden, J.
+G., Van Orden, G.
+C., \& Turvey, M.
+T.
+(2009).
+Dispersion of response times reveals cognitive dynamics.
+\emph{Psychological Review}, \emph{116}(2), 318--342.
+\url{https://doi.org/10.1037/a0014849}
 
@@
 
+\bibitem[\citeproctext]{ref-R-tidybayes}
+Kay, M. (2024). \emph{{tidybayes}: Tidy data and geoms for {Bayesian} models}. \url{https://doi.org/10.5281/zenodo.1308151}
+
 \bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
@@
 
+\bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
+Luce, R. D. (1991). \emph{Response times: Their role in inferring elementary mental organization} (1. issued as paperback). Oxford: Univ. Press.
+
 \bibitem[\citeproctext]{ref-mcelreathStatisticalRethinkingBayesian2020}
 McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
+
+\bibitem[\citeproctext]{ref-R-tibble}
+Müller, K., \& Wickham, H. (2023). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}
 
@@
 
+\bibitem[\citeproctext]{ref-R-RColorBrewer}
+Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. Retrieved from \url{https://CRAN.R-project.org/package=RColorBrewer}
+
 \bibitem[\citeproctext]{ref-panisHowCanWe2020}
@@
 
+\bibitem[\citeproctext]{ref-R-patchwork}
+Pedersen, T. L. (2024). \emph{Patchwork: The composer of plots}. Retrieved from \url{https://patchwork.data-imaginist.com}
+
+\bibitem[\citeproctext]{ref-R-nlme}
+Pinheiro, J. C., \& Bates, D. M. (2000). \emph{Mixed-effects models in s and s-PLUS}. New York: Springer. \url{https://doi.org/10.1007/b98882}
+
+\bibitem[\citeproctext]{ref-R-base}
+R Core Team. (2024). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
@@
 
+\bibitem[\citeproctext]{ref-R-StanHeaders}
+Stan Development Team. (2020). \emph{{StanHeaders}: Headers for the {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
+
+\bibitem[\citeproctext]{ref-R-rstan}
+Stan Development Team. (2024). \emph{{RStan}: The {R} interface to {Stan}}. Retrieved from \url{https://mc-stan.org/}
+
 \bibitem[\citeproctext]{ref-teachmanAnalyzingSocialProcesses1983}
@@
 
+\bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}
+Townsend, J.
+T.
+(1990).
+Truth and consequences of ordinal differences in statistical distributions: {Toward} a theory of hierarchical inference.
+\emph{Psychological Bulletin}, \emph{108}(3), 551--567.
+\url{https://doi.org/10.1037/0033-2909.108.3.551}
+
 \bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}
 Wickelgren, W. A. (1977). Speed-accuracy tradeoff and information processing dynamics. \emph{Acta Psychologica}, \emph{41}(1), 67--85. \url{https://doi.org/10.1016/0001-6918(77)90012-9}
+
+\bibitem[\citeproctext]{ref-R-ggplot2}
+Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}
+
+\bibitem[\citeproctext]{ref-R-forcats}
+Wickham, H. (2023a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://forcats.tidyverse.org/}
+
+\bibitem[\citeproctext]{ref-R-stringr}
+Wickham, H. (2023b). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://stringr.tidyverse.org}
+
+\bibitem[\citeproctext]{ref-R-tidyverse}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the {tidyverse}.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\bibitem[\citeproctext]{ref-R-dplyr}
+Wickham, H., François, R., Henry, L., Müller, K., \& Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://dplyr.tidyverse.org}
+
+\bibitem[\citeproctext]{ref-R-purrr}
+Wickham, H., \& Henry, L. (2023). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://purrr.tidyverse.org/}
+
+\bibitem[\citeproctext]{ref-R-readr}
+Wickham, H., Hester, J., \& Bryan, J. (2024). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://readr.tidyverse.org}
+
+\bibitem[\citeproctext]{ref-R-tidyr}
+Wickham, H., Vaughan, D., \& Girlich, M. (2024). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://tidyr.tidyverse.org}
 

### 2025-06-25 manuscript/manuscript.Rmd
update ms. and references

@@
 
-Even though EHA is a widely used statistical tool and there already exist many excellent reviews [e.g., @singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004] and tutorials [e.g., @allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004; @millsIntroducingSurvivalEvent2011] and tutorials [@allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020; @elmerModelingCategoricalTimetoevent2023; @lougheedMultilevelSurvivalAnalysis2019; @stoolmillerIntroductionUsingMultivariate2015; @stoolmillerModelingHeterogeneitySocial2006], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
 Set within this context, our overall aim is to introduce a set of tutorials, which explain **how** to do such analyses in the context of experimental psychology, rather than repeat in any detail **why** you may do them.
@@
 
-EHA is a class of statistical approaches to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations (Allison, 2010).
+EHA is a class of statistical approaches to study the occurrence and timing of events, such as disease onset, marriages, arrests, and job terminations [@allisonSurvivalAnalysisUsing2010].
 In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
@@
 
-To avoid possible confusion in terminology used, it is worth noting that EHA is known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer & Willett, 2003).
+To avoid possible confusion in terminology used, it is worth noting that EHA is known by various labels, such as survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis [@singerAppliedLongitudinalData2003].
 In this paper, we choose to use the term EHA throughout.
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable. 
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (REFS). 
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well [@bloxomEstimatingResponseTime1984; @luceResponseTimesTheir1991; @vanzandtHowFitResponse2000].
 Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per experiment [@panisAnalyzingResponseTimes2020].

### 2025-06-25 manuscript/manuscript.tex
update ms. and references

@@
 
-Even though EHA is a widely used statistical tool and there already exist many excellent reviews (Allison, 1982; Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; e.g., Singer \& Willett, 2003; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+Even though EHA is a widely used statistical tool and there already exist many excellent reviews (Allison, 1982; Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Mills, 2011; Singer \& Willett, 2003; Teachman, 1983) and tutorials (Allison, 2010; Elmer, Van Duijn, Ram, \& Bringmann, 2023; Landes, Engelhardt, \& Pelletier, 2020; Lougheed, Benson, Cole, \& Ram, 2019; Stoolmiller, 2015; Stoolmiller \& Snyder, 2006), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
 Set within this context, our overall aim is to introduce a set of tutorials, which explain \textbf{how} to do such analyses in the context of experimental psychology, rather than repeat in any detail \textbf{why} you may do them.
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable.
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (REFS).
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per experiment (REF PANIS).
+However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991; Van Zandt, 2000).
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per experiment (Panis, Schmidt, et al., 2020).
 Moreover, as indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
@@
 
+\bibitem[\citeproctext]{ref-bloxomEstimatingResponseTime1984}
+Bloxom, B.
+(1984).
+Estimating response time hazard functions: {An} exposition and extension.
+\emph{Journal of Mathematical Psychology}, \emph{28}(4), 401--420.
+\url{https://doi.org/10.1016/0022-2496(84)90008-7}
+
 \bibitem[\citeproctext]{ref-bolgerCausalProcessesPsychology2019}
@@
 Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. \url{https://doi.org/10.18637/jss.v040.i08}
+
+\bibitem[\citeproctext]{ref-elmerModelingCategoricalTimetoevent2023}
+Elmer, T., Van Duijn, M.
+A.
+J., Ram, N., \& Bringmann, L.
+F.
+(2023).
+Modeling categorical time-to-event data: {The} example of social interaction dynamics captured with event-contingent experience sampling methods.
+\emph{Psychological Methods}.
+\url{https://doi.org/10.1037/met0000598}
 
@@
 
+\bibitem[\citeproctext]{ref-lougheedMultilevelSurvivalAnalysis2019}
+Lougheed, J.
+P., Benson, L., Cole, P.
+M., \& Ram, N.
+(2019).
+Multilevel survival analysis: {Studying} the timing of children's recurring behaviors.
+\emph{Developmental Psychology}, \emph{55}(1), 53--65.
+\url{https://doi.org/10.1037/dev0000619}
+
 \bibitem[\citeproctext]{ref-luceResponseTimesTheir1991}
@@
 McElreath, R. (2020). \emph{Statistical {Rethinking}: {A Bayesian Course} with {Examples} in {R} and {STAN}} (2nd ed.). New York: {Chapman and Hall/CRC}. \url{https://doi.org/10.1201/9780429029608}
+
+\bibitem[\citeproctext]{ref-millsIntroducingSurvivalEvent2011}
+Mills, M.
+(2011).
+\emph{Introducing {Survival} and {Event History Analysis}}.
+1 Oliver's Yard,~55 City Road,~London~EC1Y 1SP~United Kingdom: SAGE Publications Ltd.
+\url{https://doi.org/10.4135/9781446268360}
 
@@
 
+\bibitem[\citeproctext]{ref-stoolmillerIntroductionUsingMultivariate2015}
+Stoolmiller, M.
+(2015).
+\emph{An {Introduction} to {Using Multivariate Multilevel Survival Analysis} to {Study Coercive Family Process}} (Vol.
+1; T.
+J.
+Dishion \& J.
+Snyder, Eds.).
+Oxford University Press.
+\url{https://doi.org/10.1093/oxfordhb/9780199324552.013.27}
+
+\bibitem[\citeproctext]{ref-stoolmillerModelingHeterogeneitySocial2006}
+Stoolmiller, M., \& Snyder, J.
+(2006).
+Modeling heterogeneity in social interaction processes using multilevel survival analysis.
+\emph{Psychological Methods}, \emph{11}(2), 164--177.
+\url{https://doi.org/10.1037/1082-989X.11.2.164}
+
 \bibitem[\citeproctext]{ref-teachmanAnalyzingSocialProcesses1983}
@@
 
+\bibitem[\citeproctext]{ref-vanzandtHowFitResponse2000}
+Van Zandt, T. (2000). How to fit a response time distribution. \emph{Psychonomic Bulletin \& Review}, \emph{7}(3), 424--465. \url{https://doi.org/10.3758/BF03214357}
+
 \bibitem[\citeproctext]{ref-wickelgrenSpeedaccuracyTradeoffInformation1977}

### 2025-06-30 manuscript/reviews/ResponseReview1.Rmd
update ResponseReviewers
commments on the response.

@@
 
-Response: Thank you for allowing us to submit a revision of our Tutorial (AMPPS-24-0232).
+Response: Thank you for providing constructive feedback and allowing us to submit a revision of our manuscript (AMPPS-24-0232).
 
@@
 We also mention the unique contribution of our work, we removed redundancies, and tried to improve the clarity of the paper.
+As such, we have substantially revised the manuscript in accordance with feedback from you as well as the reviewers.
 
@@
 
-Response: We disambiguate the various term in the revised ms. on page 7, and use one term throughout.
+Response: We disambiguate the various terms in the revised ms. on page 7, and we now only use one term throughout.
 
@@
 
-Response: We moved the explanation of the survival function to the Supplementary Material in the revised ms.
+Response: Thank you for the positive feedback. 
+To make the manuscript more accessible and shorter and to respond to some other comments from the editor and other reviewers, we moved the explanation of the survival function to the Supplementary Material in the revised ms.
 
@@
 > Taken together, this is an excellent manuscript with the potential to further bring survival analysis methods into Psychology, where time-to-event and right-censored cases abound, yet very few people seem to consider these issues.
+
+Thank you very much for your helpful and positive feedback.
 
@@
 
-Response: We removed repetitive text, and try to focus on clarity and brevity in the revised ms.
+Response: Thanks for your constructive and encouraging feedback.
+As we also noted in our response to the editor, we have made substantial revisions to the manuscript that include removing repetitive text, and placing a stronger focus on clarity and brevity in the revised ms.
 
@@
 
-Response: We explain the location of the files in the README file. A single download step is possible by either cloning the project, or downloading the ZIP file.
-We included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
+Response: We explain the location of the files and the folder structure in the README file.
+In addition we make it clear that s single download step is possible by either cloning the project, or downloading the ZIP file.
+We also included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
+At this point, we also prefer a modular approach to the tutorials rather than combining them. 
+We appreciate that not everyone might share this preference, but we think there are many benefits to a modular workflow in this instance.
 
@@
 Response: We placed the equations for h(t) and ca(t) in Figure 1 of the revised ms. 
+Beyond Figure 1, we still choose to keep any further reference of equations to supplementary materials. 
+This is because we feel that one barrier to uptake of these methods is an aversion to equations.
+And by placing them in supplementary materials we feel that we strike a nice balance between non-technical accessibility and a fullness of mathemathical explanation for key ideas and concepts that remain easily accessible to interested readers.
 
@@
 
-Response: We decided to keep tutorial 4.
+Response: We decided to keep tutorial 4, as we feel that planning future studies is an important and difficult step to take in these kinds of studies, which we wanted to keep close to the rest of the material for ease of understanding a reasonably complete workflow.
 

### 2025-06-30 manuscript/notes/notes.Rmd
update Notes
commments on the response.

@@
 
+-> I removed redundant packages.
+
 ## r-references ##
@@
 
+-> I got it working but packages are (for now) mentioned in text instead of in footnote...
+
 ## intro ##
@@
 - I made the intro shorter. We still make the same points, just in far fewer words.
+
+-> Well done!
 
@@
 If you want to include any of it in supps, then let's discuss it.
+
+-> Section A of the revised Suppl. Material now contains the Figure of RT tasks and discussion of the types of time-to-event data obtained in typical RT tasks.
+
 Instead, I chose to focus on picking out the benefits of EHA for research in experimental psychology.
@@
 
+-> I agree.
+
 - Sven - can you add a relevant reference for the following statement:
@@
 
+-> Done.
+
 - in section 2.3 can you add a line or two about bin width please?
+
+-> Done.
 
@@
 
+-> TODO.
+
 ## 4.1.2 ##
@@
 
+-> TODO.
+
 page 19. Figure 2. The image warped / wrong size and needs fixing.
+
+-> Done.
 
@@
 
+-> TODO.
+
 ## discussion ##
@@
 I removed the unnecessary sections that we discussed previously.
+
+-> ok.
+
+# June 30th 2025 #
+
+## section 1.1 ##
+
+- re-combine paragraphs 2 and 3. That way, you don't start a paragraph with "For example ..."
+
+## section 2.3 ##
+
+- para 2. 1 sentence is not a paragraph. is this not finished yet? I don't understand hopw that sentence fits.
+
+## response to reviewers ##
+
+- the formatting really helps distinguish between editor/reviewer comments and our responses. Nice one. it is so much easier to read.
+
+- the below quote is not contradictory, but it sure sounds it.
+Can we rework it?
+Do you see what I mean?
+single events that repeat...
+Maybe just an extra sentence or two to make the relevant distinction clear.
+
+>Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.
+
+on line 510 of the response, you say:
+
+>Response: We provided a clearer description of the data.
+
+I think it is essential to always provide page, para and/or line numbers for reviewers so that they can easily check things.
+This same logic should be applied to all of our responses where necessary.
+Once the page numbers change, it can be a nightmare to find and verufy changes as a reviewer.

### 2025-07-04 manuscript/manuscript.Rmd
small updates across ms.

@@
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
-Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
+Such absolute timing information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection [@panisAnalyzingResponseTimes2020].
 
@@
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 [see also @panisWhatShapingRT2016].
 
@@
 ```
-
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 [see also @panisWhatShapingRT2016].
 
@@
 In Section 3, we outline a series of tutorials, which are written in the R programming language and publicly available on our Github page (https://github.com/sven-panis/Tutorial_Event_History_Analysis), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, plotting descriptive statistics, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
 In Section 4, we discuss the strengths and weaknesses of the approach for researchers in experimental psychology.
@@
 
-Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable perception tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
 The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition [@barackTwoViewsCognitive2021].
@@
 Third, the width of each time bin will need to be determined. 
-For instance, in Figure 1B we chose 100ms in an arbitrary manner.
+For instance, in Figure 1B we chose 100 ms in an arbitrary manner.
 In reality, however, bin width will need to be set by considering a number of factors simultaneously.
@@
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to mean-average comparisons.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to standard approaches.
 While we have focused here on within-subject, factorial, small-*N* designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-*N* designs with only one measurement per subject, between-subject designs, etc.).
@@
 For those researchers, like ourselves, who are primarily interested in understanding human cognitive and brain systems, we consider two broadly-defined, main use-cases of EHA.
-First, as we hope to have made clear by this point, EHA is one way to investigating a "temporal states" approach to cognitive processes.
-EHA provides one way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
+First, as we hope to have made clear by this point, EHA is one way to investigating a "temporal states" approach to cognitive processes, by tracking behavior as a function of step-wise increases in absolute waiting time.
+EHA thus provides a way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
 Therefore, if your research questions concern **when psychological states occur, and how they are temporally organized**, our EHA tutorials could be useful tools for you to use.
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.
+Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice (REF why you should use EHA...).
+
+<!--One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
+Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.-->
 
@@
 
-EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions. 
+Hazard and conditional accuracy models can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions. 
 For example, some of the models discussed in Tutorial 2a, which we did not focus on in the main text, contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
-However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises a number of sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
 From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (REF DFT).
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
-As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
-This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4).
+As one example, if there are several times cales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
+This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4 and section 2.3).
 
@@
 
-Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods (ref ??).
 
-Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
+Another issue is that you need a relatively large number of trials per condition to estimate the discrete-time hazard function with relatively high temporal resolution (e.g., 20 ms), which is required when testing predictions of process models of cognition.
 Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the discete-time hazard function.
@@
 After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
-Finally, there is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
+There is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
+Finally, if you have a lot of repeated measurements per condition per participant, you can of course also try continuous-time methods (ref).
 

### 2025-07-04 manuscript/manuscript.tex
small updates across ms.

@@
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
-Such information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
+Such absolute timing information is useful not only for the interpretation of experimental effects under investigation, but also for cognitive psychophysiology and computational model selection (Panis, Schmidt, Wolkersdorfer, \& Schmidt, 2020).
 
@@
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 (see also Panis \& Schmidt, 2016).
 
@@
 
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 (see also Panis \& Schmidt, 2016).
-
 Why does this matter for research in psychology?
@@
 In Section 3, we outline a series of tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/Tutorial_Event_History_Analysis}), along with all of the other code and material associated with the project.
-The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
+The tutorials provide hands-on, concrete examples of key parts of the analytical process, such as data wrangling, plotting descriptive statistics, model fitting and planning future studies, so that others can apply EHA to their own time-to-event data measured in RT tasks.
 In Section 4, we discuss the strengths and weaknesses of the approach for researchers in experimental psychology.
@@
 
-Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable percpetion tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
+Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable perception tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
 The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (Barack \& Krakauer, 2021).
@@
 Third, the width of each time bin will need to be determined.
-For instance, in Figure 1B we chose 100ms in an arbitrary manner.
+For instance, in Figure 1B we chose 100 ms in an arbitrary manner.
 In reality, however, bin width will need to be set by considering a number of factors simultaneously.
@@
 By providing a freely available set of tutorials, which provide step-by-step guidelines and ready-to-use R code, we hope that researchers will feel more comfortable using EHA/SAT in the future.
-Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to mean-average comparisons.
+Indeed, we hope that our tutorials may help to overcome a barrier to entry with EHA/SAT, which is that such approaches require more analytical complexity compared to standard approaches.
 While we have focused here on within-subject, factorial, small-\emph{N} designs, it is important to realize that EHA/SAT can be applied to other designs as well (large-\emph{N} designs with only one measurement per subject, between-subject designs, etc.).
@@
 For those researchers, like ourselves, who are primarily interested in understanding human cognitive and brain systems, we consider two broadly-defined, main use-cases of EHA.
-First, as we hope to have made clear by this point, EHA is one way to investigating a ``temporal states'' approach to cognitive processes.
-EHA provides one way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
+First, as we hope to have made clear by this point, EHA is one way to investigating a ``temporal states'' approach to cognitive processes, by tracking behavior as a function of step-wise increases in absolute waiting time.
+EHA thus provides a way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
 Therefore, if your research questions concern \textbf{when psychological states occur, and how they are temporally organized}, our EHA tutorials could be useful tools for you to use.
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.
+Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice (REF why you should use EHA\ldots).
 
@@
 
-EHA can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions.
+Hazard and conditional accuracy models can quickly become very complex when adding more than one time scale, due to the many possible higher-order interactions.
 For example, some of the models discussed in Tutorial 2a, which we did not focus on in the main text, contain two time scales as covariates: the passage of time on the within-trial time scale, and the passage of time on the across-trial (or within-experiment) time scale.
-However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises three sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
+However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises a number of sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
 From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (REF DFT).
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
-As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
-This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4).
+As one example, if there are several times cales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
+This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4 and section 2.3).
 
@@
 
-Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
-
-Another issue is that you need a relatively large number of trials per condition to estimate the hazard function with high temporal resolution, which is required when testing predictions of process models of cognition.
+Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods (ref ??).
+
+Another issue is that you need a relatively large number of trials per condition to estimate the discrete-time hazard function with relatively high temporal resolution (e.g., 20 ms), which is required when testing predictions of process models of cognition.
 Indeed, in general, there is a trade-off between the number of trials per condition and the temporal resolution (i.e., bin width) of the discete-time hazard function.

### 2025-07-05 manuscript/manuscript.Rmd
changes to ms.

@@
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
-The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
+The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.1.2.
 Error bars represent +/- 1 standard error of the mean (A) or proportion (B).
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
-We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks and supply relevant regression equations in sections A and E of the Supplemental Material, respectively.
+We supply a few regression equations in section E of the Supplemental Material.
 
@@
 
-In terms of minimum requirements to apply a single-event EHA, one must be able to:
+In terms of minimum requirements to apply EHA, one must be able to:
 
@@
 
+These minimal requirements are fulfilled by the RT data obtained in single-button detection tasks, where the time-to-response is repeatedly measured in different trials in the same individual.
+In section A of the Supplemental Material we visualize this and other types of time-to-event data which are typically obtained in discrimination and bistable perception tasks.
+
 ### 2.1.2 Types of EHA
 
-There are different types of EHA.
+There are different types of modeling approaches in EHA.
 For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. 
@@
 However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well [@bloxomEstimatingResponseTime1984; @luceResponseTimesTheir1991; @vanzandtHowFitResponse2000].
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per experiment [@panisAnalyzingResponseTimes2020].
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per participant [@panisAnalyzingResponseTimes2020].
 Moreover, as indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons [@holdenDispersionResponseTimes2009; @townsendTruthConsequencesOrdinal1990; @luceResponseTimesTheir1991].
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material). 
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section G of the Supplemental Material). 
 Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
@@
 More specifically, we consider three implications that researchers will need to consider when using discrete-time EHA.
-
 First, one can use a response deadline in each trial because EHA deals with right-censored observations.
@@
 Tutorial 4 shows how to use simulation and power analysis for planning experiments.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C in the Supplemental Material for a list of the custom functions).
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C of the Supplemental Material for a list of the custom functions).
 
-Our list of tutorials is as follows:
+<!-- Our list of tutorials is as follows:
 
@@
 * 3b. Frequentist multilevel modeling for ca(t)
-* 4.  Simulation and power analysis for planning experiments
+* 4.  Simulation and power analysis for planning experiments-->
 
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 We used `r my_r_citation` for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on @allisonSurvivalAnalysisUsing2010, @singerAppliedLongitudinalData2003, @mcelreathStatisticalRethinkingBayesian2020, @heiss2021,  @kurzAppliedLongitudinalDataAnalysis2023, and @kurzStatisticalRethinkingSecondEd2023.
 
@@
 First, we want to summarise and visualise descriptive statistics using a life table.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), possibly the conditional accuracy function ca(t), and their estimated standard errors (se).
-The definitions of these functions are provided in section A of the Supplemental Material.
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
+The definitions of these quantities are provided in section B of the Supplemental Material.
 
@@
 
-```{r ca-data-table}
+```{r ca-data-table, warning=F}
 apa_table(
@@
 Therefore, the 'person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-This data set is used when fitting hazard models (Tutorials 2a and 3a).
+This data set is used when fitting discrete-time hazard models (Tutorials 2a and 3a).
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
@@
 
-```{r ha-data-table}
+```{r ha-data-table, warning=F}
 apa_table(
@@
 
-To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016.
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from @panisWhatShapingRT2016, who were interested in the temporal dynamics of the effect of prime-target congruency in RT and accuracy data.
 In their first experiment, @panisWhatShapingRT2016 presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 Response time and accuracy were recorded on each trial.
-Prime type (blank, congruent, incongruent) and mask type were manipulated.
-Here we focus on the subset of trials in which no mask was presented.
+Prime type (blank, congruent, incongruent) and mask type were manipulated across trials (i.e., repeated measures of time-to-response).
+Here we focus for each participant on the subset of 220 trials in which no mask was presented.
 The 13-ms prime stimulus was a double arrow presented 187 ms before target onset in the congruent (same direction as target) and incongruent (opposite direction as target) prime conditions.
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t).
+Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
 To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
@@
 Note that the censoring time (here: 600 ms) should be a multiple of the bin width (here: 40 ms).
-The censoring time should be a time point after which no informative responses are expected anymore.
+The censoring time should be a time point after which no informative responses are expected anymore, in case one waits for a response in each trial.
 In experiments that implement a response deadline in each trial the censoring time can equal that deadline time point.
-Trials with a RT larger than the censoring time, or trials in which no response is emitted during the data collection period, are treated as right-censored observations in EHA.
+Trials with a RT larger than the censoring time, or trials in which no response is emitted during the data observation period, are treated as right-censored observations in EHA.
 In other words, these trials are not discarded, because they contain the information that the event did not occur before the censoring time.
-Removing such trials before calculating the mean event time will result in underestimation of the true mean.
+Removing such trials before calculating the mean event time would result in underestimation of the true mean.
 
@@
 The variable "event" in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-
-The next step is to set up the life table using our custom function setup_lt(), calculate the conditional accuracies using our custom function calc_ca(), add the ca(t) estimates to the life table using our custom function join_lt_ca(), and then plot the descriptive statistics using our custom function plot_eha().
+The next steps are to set up the life table using our custom function setup_lt(), calculate the conditional accuracies using our custom function calc_ca(), add the ca(t) estimates to the life table using our custom function join_lt_ca(), and then plot the descriptive statistics using our custom function plot_eha().
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
 
-In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
-Also, to select a bin width for fitting models, one should test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
+In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may not be following task instructions (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+Also, to select a suited bin width for model fitting, one can test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
 Too small bin widths will result in erratic hazard functions because many bins will have estimates equal to zero.
@@
 
-```{r life-table}
+```{r life-table, warning=F}
 life_tab <- read_csv("../Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
@@
 Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
-By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any data set.
+By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any RT data set.
 
@@
 When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial_1a.Rmd).
-
-In general, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
+More generally, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
 Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
-At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA/SAT.
-As we will show in later Tutorials, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
-But first we consider calculating the descriptive statistics for two independent variables.
+At this point, we have calculated and plotted the descriptive statistics for each type of prime stimulus in EHA/SAT.
+As we will show in later Tutorials, statistical models for hazard and conditional accuracy functions can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
+But first we consider calculating the descriptive statistics for within-subject designs with two independent variables.
 
@@
 The purpose of Tutorial 1b, therefore, is to provide a generalisation of the basic approach, which extends to a more complicated design.
-We felt that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
+We feel that this might be useful for researchers in experimental psychology that typically use crossed factorial designs.
 
@@
 
-## 3.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data
+## 3.3 Tutorial 2a: Fitting Bayesian hazard models to interval-censored RT data
 
@@
 There are several analytic decisions one has to make when fitting a discrete-time hazard model.
-First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the Supplemental Material).
+First, because the first few bins typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant.
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section D of the Supplemental Material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data [@singerAppliedLongitudinalData2003].
-Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor.
-And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding.
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
+For example, when you want to know if cloglog-hazard is changing linearly or quadracticly over time, you should treat TIME as a continuous predictor.
+When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding.
 With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level.
@@
 
-The middle column of Supplementary Figure 2 (section E of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 2 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Supplementary Figure 3 (section F of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 3 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
-In this first model, we use a general specification of TIME using index coding, and do not include experimental predictors.
+In this first baseline or reference model, we use a general specification of TIME using index coding, and do not include experimental predictors.
 We call this model "M0i".
+The other model (see section 3.3.4) extends model M0i by including our experimental predictor prime type.
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A; grand average marginal effect) and at the participant level (Figure 7B; subject-specific average marginal effect).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A) and at the participant level (Figure 7B).
 
@@
 
-The point estimates and quantile intervals can be reported in a table (see Tutorial_2a.Rmd for details).
+The point estimates and quantile intervals can also be reported in a table (see Tutorial_2a.Rmd for details).
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 4A.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 7B.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 4B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
 
-Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
+<!-- Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
 For example, @panisWhatShapingRT2016 concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
@@
 To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
-This effect might be masked for incongruent primes by the response competition effect.
+This effect might be masked for incongruent primes by the response competition effect.-->
 
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 7B).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 4B).
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
@@
 -->
+
+Before we move to our Tutorial on planning experiments, we also provide code to fit hazard and conditional accuracy models in the frequentist framework (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
+However, because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here.
 
@@
 As such, the general workflow and associated code can be modified and applied more broadly to other contexts and research questions.
-In the following, we discuss the main use-cases, issues relating to model complexity and interpretability, as well as limitations of the approach and future extensions.
+In the following, we discuss the main use-cases, issues relating to model complexity and interpretability, as well as limitations of the approach.
 
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice (REF why you should use EHA...).
+Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
 
@@
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises a number of sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (REF DFT).
+From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general [@schonerDynamicThinkingPrimer2016].
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
@@
 
-Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods (ref ??).
+Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
 
@@
 After consideration, it might be worth conducting multiple testing sessions per participant and/or reducing the number of experimental conditions.
-There is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
-Finally, if you have a lot of repeated measurements per condition per participant, you can of course also try continuous-time methods (ref).
+There is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process [@bakerPowerContoursOptimising2021].
+Finally, if you have a lot of repeated measurements per condition per participant, you can of course also try continuous-time methods [@allisonSurvivalAnalysisUsing2010; @elmerModelingCategoricalTimetoevent2023].
 

### 2025-07-05 manuscript/manuscript.tex
changes to ms.

@@
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
-The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.
+The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.1.2.
 Error bars represent +/- 1 standard error of the mean (A) or proportion (B).}\label{fig:plot1}
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Singer \& Willett, 2003; Winter, 2019).
-We also visualize and discuss the types of time-to-event data that are obtained in typical RT tasks and supply relevant regression equations in sections A and E of the Supplemental Material, respectively.
+We supply a few regression equations in section E of the Supplemental Material.
 
@@
 
+These minimal requirements are fulfilled by the RT data obtained in single-button detection tasks, where the time-to-response is repeatedly measured in different trials in the same individual.
+In section A of the Supplemental Material we visualize this and other types of time-to-event data which are typically obtained in discrimination and bistable perception tasks.
+
 \subsubsection{2.1.2 Types of EHA}\label{types-of-eha}
 
-There are different types of EHA.
+There are different types of modeling approaches in EHA.
 For example, the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units.
@@
 However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991; Van Zandt, 2000).
-Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per experiment (Panis, Schmidt, et al., 2020).
+Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per participant (Panis, Schmidt, et al., 2020).
 Moreover, as indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
@@
 Statisticians and mathematical psychologists recommend focusing on the hazard function when analyzing time-to-event data for various reasons (Holden, Van Orden, \& Turvey, 2009; Luce, 1991; Townsend, 1990).
-We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section F of the Supplemental Material).
+We do not cover these benefits in detail here, as these are more general topics that have been covered elsewhere in textbooks (see also section G of the Supplemental Material).
 Instead, here we focus on the benefits as we see them for common research programmes in experimental psychology.
@@
 Tutorial 4 shows how to use simulation and power analysis for planning experiments.
-Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C in the Supplemental Material for a list of the custom functions).
-
-Our list of tutorials is as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  1a. Wrangle raw data and calculate descriptive stats for one independent variable
-\item
-  1b. Wrangle raw data and calculate descriptive stats for two independent variables
-\item
-  2a. Bayesian multilevel modeling for h(t)
-\item
-  2b. Bayesian multilevel modeling for ca(t)
-\item
-  3a. Frequentist multilevel modeling for h(t)
-\item
-  3b. Frequentist multilevel modeling for ca(t)
-\item
-  \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \setcounter{enumi}{3}
-  \tightlist
-  \item
-    Simulation and power analysis for planning experiments
-  \end{enumerate}
-\end{itemize}
-
+Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C of the Supplemental Material for a list of the custom functions).
+
+The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 We used R (Version 4.4.0; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023) for all reported analyses.
-The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
 
@@
 Therefore, the `person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin.
-This data set is used when fitting hazard models (Tutorials 2a and 3a).
+This data set is used when fitting discrete-time hazard models (Tutorials 2a and 3a).
 It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest.
 Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.
-
-\begin{verbatim}
-## Warning in attr(x, "align"): 'xfun::attr()' is deprecated.
-## Use 'xfun::attr2()' instead.
-## See help("Deprecated")
-\end{verbatim}
 
@@
 
-To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016).
+To illustrate how to quickly set up life tables for calculating the descriptive statistics (functions of discrete time), we use a published data set on masked response priming from Panis and Schmidt (2016), who were interested in the temporal dynamics of the effect of prime-target congruency in RT and accuracy data.
 In their first experiment, Panis and Schmidt (2016) presented a double arrow for 94 ms that pointed left or right as the target stimulus with an onset at time point zero in each trial.
@@
 
-Next, we can set up the life tables and plots of the discrete-time functions h(t), S(t), ca(t), and P(t).
+Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
 To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
@@
 The variable ``event'' in the person-trial-bin oriented data set indicates whether a response occurs (1) or not (0) for each bin.
-
-The next step is to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha().
+The next steps are to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha().
 One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.
 
-In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may be guessing (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
-Also, to select a bin width for fitting models, one should test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
+In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may not be following task instructions (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.
+Also, to select a suited bin width for model fitting, one can test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
 Too small bin widths will result in erratic hazard functions because many bins will have estimates equal to zero.
@@
 When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial\_1a.Rmd).
-
-In general, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
+More generally, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
 Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
 
-At this point, we have calculated, summarised and plotted descriptive statistics for the key variables in EHA/SAT.
-As we will show in later Tutorials, statistical models for h(t) and ca(t) can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
-But first we consider calculating the descriptive statistics for two independent variables.
+At this point, we have calculated and plotted the descriptive statistics for each type of prime stimulus in EHA/SAT.
+As we will show in later Tutorials, statistical models for hazard and conditional accuracy functions can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
+But first we consider calculating the descriptive statistics for within-subject designs with two independent variables.
 
@@
 
-\subsection{3.3 Tutorial 2a: Fitting Bayesian hazard models to discrete time-to-event data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-discrete-time-to-event-data}
+\subsection{3.3 Tutorial 2a: Fitting Bayesian hazard models to interval-censored RT data}\label{tutorial-2a-fitting-bayesian-hazard-models-to-interval-censored-rt-data}
 
@@
 There are several analytic decisions one has to make when fitting a discrete-time hazard model.
-First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant.
-Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the Supplemental Material).
+First, because the first few bins typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant.
+Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section D of the Supplemental Material).
 The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003).
-Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor.
-And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding.
+Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
+For example, when you want to know if cloglog-hazard is changing linearly or quadracticly over time, you should treat TIME as a continuous predictor.
+When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding.
 With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level.
@@
 When you do not want to make assumptions about the shape of the hazard function, or its shape is not smooth but irregular, then you can use a general specification of TIME, i.e., fit one grand intercept per time bin.
-In this first model, we use a general specification of TIME using index coding, and do not include experimental predictors.
+In this first baseline or reference model, we use a general specification of TIME using index coding, and do not include experimental predictors.
 We call this model ``M0i''.
+The other model (see section 3.3.4) extends model M0i by including our experimental predictor prime type.
 
@@
 
-The point estimates and quantile intervals can be reported in a table (see Tutorial\_2a.Rmd for details).
+The point estimates and quantile intervals can also be reported in a table (see Tutorial\_2a.Rmd for details).
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 4A.
 The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 7B.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 4B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
 
-Future studies could (a) increase the number of participants to estimate the proportion of ``dippers'' in the subject population, and/or (b) try to explain why this dip occurs.
-For example, Panis and Schmidt (2016) concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
-Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to.
-To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after \textasciitilde360 ms, and towards the end of the experiment.
-This effect might be masked for incongruent primes by the response competition effect.
-
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 7B).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 4B).
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
@@
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
+
+Before we move to our Tutorial on planning experiments, we also provide code to fit hazard and conditional accuracy models in the frequentist framework (see Tutorial\_3a.Rmd and Tutorial\_3b.Rmd).
+However, because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here.
 
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice (REF why you should use EHA\ldots).
+Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
 
@@
 However, when trials are presented in blocks, and blocks of trials within sessions, and when the experiment comprises a number of sessions, then four time scales can be defined (within-trial, within-block, within-session, and within-experiment).
-From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (REF DFT).
+From a theoretical perspective, adding more than one time scale -- and their interactions -- can be important to capture plasticity and other learning effects that may play out on such longer time scales, and that are probably present in each experiment in general (Schöner \& Spencer, 2016).
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
@@
 There is a user-friendly online tool for calculating statistical power as a function of the number of trials as well as the number of participants, and this might be worth consulting to guide the research design process (Baker et al., 2021).
-Finally, if you have a lot of repeated measurements per condition per participant, you can of course also try continuous-time methods (ref).
+Finally, if you have a lot of repeated measurements per condition per participant, you can of course also try continuous-time methods (Allison, 2010; Elmer et al., 2023).
 
@@
 
+\bibitem[\citeproctext]{ref-schonerDynamicThinkingPrimer2016}
+Schöner, G., \& Spencer, J. P. (2016). \emph{Dynamic thinking: A primer on dynamic field theory}. New York, NY: Oxford University Press. \url{https://doi.org/10.1093/acprof:oso/9780199300563.001.0001}
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}

### 2025-07-05 manuscript/Suppl_material/supplementary.Rmd
changes to suppl. material

@@
 
-Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes (ref).
+Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes.
 For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue).
@@
 Note that we do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in REF and REF...
+More information about recurrent events analysis can be found in @stoolmillerModelingHeterogeneitySocial2006 and @millsIntroducingSurvivalEvent2011.
 
@@
 The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].
-After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial.
 Because waiting times can only increase, discrete-time EHA focuses on the discrete-time hazard function
@@
 The hazard rate function gives you the instantaneous *rate* of event occurrence at time point t, given that the event has not occurred yet.
+Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models [@singerAppliedLongitudinalData2003].
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Supplementary Figure 1.
+Popular link functions include the logit link and the complementary log-log link, as shown in Supplementary Figure 2.
 
-(ref:descr-plot14-caption) The logit and cloglog link functions.
+(ref:descr-plot14-caption) The logit and complementary log-log (cloglog) link functions.
 
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
+An example (single-level) discrete-time hazard model for individual i in bin t of trial j with three predictors (TIME, X~1~, X~2~), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
 
-\noindent cloglog[h(t)] = ln(-ln[1-h(t)]) =  [$\beta$~0~ONE + $\beta$~1~(TIME-9) + $\beta$~2~(TIME-9)$^2$] +                                    [$\beta$~3~X~1~ + $\beta$~4~X~2~ + $\beta$~5~X~2~(TIME-9)]  \hfill  (6)
+\noindent cloglog[h~ij~(t)] = ln(-ln[1-h~ij~(t)]) =  [$\beta$~0~ONE + $\beta$~1~(TIME-9) + $\beta$~2~(TIME-9)$^2$] +                                    [$\beta$~3~X~1ij~ + $\beta$~4~X~2ij~ + $\beta$~5~X~2ij~(TIME-9)]  \hfill  (6)
 
 The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the parameters $\beta$~0~ to $\beta$~2~ multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~i~ take on a value of zero).
+The first set of terms within brackets, the parameters $\beta$~0~ to $\beta$~2~ multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X~ij~ take on a value of zero).
 The second set of terms (the beta parameters $\beta$~3~ to $\beta$~5~) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X~1~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~3~ cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X~2~ in the example), then the effect of a 1 unit increase in X~2~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~4~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~4~ + $\beta$~5~ cloglog-hazard units (when TIME-9 = 1), and so forth.
+For example, the effect of a 1 unit increase in X~1ij~ is to vertically shift the whole baseline cloglog-hazard function by $\beta$~3~ cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X~2ij~ in the example), then the effect of a 1 unit increase in X~2ij~ is to vertically shift the predicted cloglog-hazard in bin 9 by $\beta$~4~ cloglog-hazard units (when TIME-9 = 0), in bin 10 by $\beta$~4~ + $\beta$~5~ cloglog-hazard units (when TIME-9 = 1), and so forth.
 To interpret the effects of a predictor, its $\beta$ parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
@@
 
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X~1~) can be written as follows:
+An example (single-level) discrete-time hazard model for individual i in bin t of trial j with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X~1ij~) can be written as follows:
 
-\noindent cloglog[h(t)] =  [$\beta$~0~D1 + $\beta$~1~D2 + $\beta$~2~D3 + $\beta$~3~D4 + $\beta$~4~D5 + $\beta$~5~D6] + [$\beta$~6~X~1~]  \hfill  (7)
+\noindent cloglog[h~ij~(t)] =  [$\beta$~0~D1 + $\beta$~1~D2 + $\beta$~2~D3 + $\beta$~3~D4 + $\beta$~4~D5 + $\beta$~5~D6] + [$\beta$~6~X~1ij~]  \hfill  (7)
 
@@
 To gain a sense of what prior *logit* values would approximate a uniform distribution on the probability scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Supplementary Figure 2 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Supplementary Figure 3 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to @kurzAppliedLongitudinalDataAnalysis2023, such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
@@
 
-## G. Advantages of hazard analysis
+## G. Advantages of EHA
 

### 2025-07-05 manuscript/Suppl_material/supplementary.tex
changes to suppl. material

@@
 Note that we do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in REF and REF\ldots{}
+More information about recurrent events analysis can be found in Stoolmiller and Snyder (2006) and Mills (2011).
 
@@
 The shape of a distribution of waiting times can be described in multiple ways (Luce, 1991).
-After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition.
+After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial.
 Because waiting times can only increase, discrete-time EHA focuses on the discrete-time hazard function
@@
 The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.
+Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models (Singer \& Willett, 2003).
 
@@
 
-Popular link functions include the logit link and the complementary log-log link, as shown in Supplementary Figure 1.
+Popular link functions include the logit link and the complementary log-log link, as shown in Supplementary Figure 2.
 
@@
 
-\caption{The logit and cloglog link functions.}\label{fig:plot-link-functions}
+\caption{The logit and complementary log-log (cloglog) link functions.}\label{fig:plot-link-functions}
 \end{figure}
@@
 
-An example (single-level) discrete-time hazard model with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
+An example (single-level) discrete-time hazard model for individual i in bin t of trial j with three predictors (TIME, X\textsubscript{1}, X\textsubscript{2}), the cloglog link function, and a second-order polynomial specification for TIME can be written as follows:
 
-\noindent cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\beta\)\textsubscript{0}ONE + \(\beta\)\textsubscript{1}(TIME-9) + \(\beta\)\textsubscript{2}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{3}X\textsubscript{1} + \(\beta\)\textsubscript{4}X\textsubscript{2} + \(\beta\)\textsubscript{5}X\textsubscript{2}(TIME-9){]} \hfill  (6)
+\noindent cloglog{[}h\textsubscript{ij}(t){]} = ln(-ln{[}1-h\textsubscript{ij}(t){]}) = {[}\(\beta\)\textsubscript{0}ONE + \(\beta\)\textsubscript{1}(TIME-9) + \(\beta\)\textsubscript{2}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{3}X\textsubscript{1ij} + \(\beta\)\textsubscript{4}X\textsubscript{2ij} + \(\beta\)\textsubscript{5}X\textsubscript{2ij}(TIME-9){]} \hfill  (6)
 
 The main predictor variable TIME is the time bin index t that is centered on value 9 in this example.
-The first set of terms within brackets, the parameters \(\beta\)\textsubscript{0} to \(\beta\)\textsubscript{2} multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{i} take on a value of zero).
+The first set of terms within brackets, the parameters \(\beta\)\textsubscript{0} to \(\beta\)\textsubscript{2} multiplied by their polynomial specifications of (centered) time, represents the shape of the baseline cloglog-hazard function (i.e., when all predictors X\textsubscript{ij} take on a value of zero).
 The second set of terms (the beta parameters \(\beta\)\textsubscript{3} to \(\beta\)\textsubscript{5}) represents the vertical shift in the baseline cloglog-hazard for a 1 unit increase in the respective predictor variable.
 Predictors can be discrete, continuous, and time-varying or time-invariant.
-For example, the effect of a 1 unit increase in X\textsubscript{1} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{3} cloglog-hazard units.
-However, if the predictor interacts linearly with TIME (see X\textsubscript{2} in the example), then the effect of a 1 unit increase in X\textsubscript{2} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{4} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{4} + \(\beta\)\textsubscript{5} cloglog-hazard units (when TIME-9 = 1), and so forth.
+For example, the effect of a 1 unit increase in X\textsubscript{1ij} is to vertically shift the whole baseline cloglog-hazard function by \(\beta\)\textsubscript{3} cloglog-hazard units.
+However, if the predictor interacts linearly with TIME (see X\textsubscript{2ij} in the example), then the effect of a 1 unit increase in X\textsubscript{2ij} is to vertically shift the predicted cloglog-hazard in bin 9 by \(\beta\)\textsubscript{4} cloglog-hazard units (when TIME-9 = 0), in bin 10 by \(\beta\)\textsubscript{4} + \(\beta\)\textsubscript{5} cloglog-hazard units (when TIME-9 = 1), and so forth.
 To interpret the effects of a predictor, its \(\beta\) parameter is exponentiated, resulting in a hazard ratio (due to the use of the cloglog link).
@@
 
-An example (single-level) discrete-time hazard model with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X\textsubscript{1}) can be written as follows:
+An example (single-level) discrete-time hazard model for individual i in bin t of trial j with a general specification for TIME (separate intercepts for each of six bins, where D1 to D6 are binary indicator variables identifying each bin) and a single predictor (X\textsubscript{1ij}) can be written as follows:
 
-\noindent cloglog{[}h(t){]} = {[}\(\beta\)\textsubscript{0}D1 + \(\beta\)\textsubscript{1}D2 + \(\beta\)\textsubscript{2}D3 + \(\beta\)\textsubscript{3}D4 + \(\beta\)\textsubscript{4}D5 + \(\beta\)\textsubscript{5}D6{]} + {[}\(\beta\)\textsubscript{6}X\textsubscript{1}{]} \hfill  (7)
+\noindent cloglog{[}h\textsubscript{ij}(t){]} = {[}\(\beta\)\textsubscript{0}D1 + \(\beta\)\textsubscript{1}D2 + \(\beta\)\textsubscript{2}D3 + \(\beta\)\textsubscript{3}D4 + \(\beta\)\textsubscript{4}D5 + \(\beta\)\textsubscript{5}D6{]} + {[}\(\beta\)\textsubscript{6}X\textsubscript{1ij}{]} \hfill  (7)
 
@@
 To gain a sense of what prior \emph{logit} values would approximate a uniform distribution on the probability scale, Kurz (2023) simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t distribution.
-Row C in Supplementary Figure 2 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
+Row C in Supplementary Figure 3 shows that using a t-distribution with 7.61 degrees of freedom and a scale parameter of 1.57 as a prior on the logit scale, approximates a uniform distribution on the probability scale.
 According to Kurz (2023), such a prior might be a good prior for the intercept(s) in a logit-hazard model, while the N(0,1) prior in row D might be a good prior for the non-intercept parameters in a logit-hazard model, as it gently regularizes p towards .5 (i.e., a zero effect on the logit scale).
@@
 
-\subsection{G. Advantages of hazard analysis}\label{g.-advantages-of-hazard-analysis}
+\subsection{G. Advantages of EHA}\label{g.-advantages-of-eha}
 
@@
 
+\bibitem[\citeproctext]{ref-millsIntroducingSurvivalEvent2011}
+Mills, M.
+(2011).
+\emph{Introducing {Survival} and {Event History Analysis}}.
+1 Oliver's Yard,~55 City Road,~London~EC1Y 1SP~United Kingdom: SAGE Publications Ltd.
+\url{https://doi.org/10.4135/9781446268360}
+
 \bibitem[\citeproctext]{ref-panisStudyingDynamicsVisual2020}
@@
 
+\bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}
+Singer, J. D., \& Willett, J. B. (2003). \emph{Applied {Longitudinal Data Analysis}: {Modeling Change} and {Event Occurrence}}. Oxford, New York: Oxford University Press.
+
+\bibitem[\citeproctext]{ref-stoolmillerModelingHeterogeneitySocial2006}
+Stoolmiller, M., \& Snyder, J.
+(2006).
+Modeling heterogeneity in social interaction processes using multilevel survival analysis.
+\emph{Psychological Methods}, \emph{11}(2), 164--177.
+\url{https://doi.org/10.1037/1082-989X.11.2.164}
+
 \bibitem[\citeproctext]{ref-townsendTruthConsequencesOrdinal1990}

### 2025-07-05 manuscript/notes/notes.Rmd
changes to notes.

@@
 
--> TODO.
+-> Done.
 
@@
 
+-> Done. I separated them before to avoid an almost blank page in the ms...
+
 ## section 2.3 ##
@@
 
+-> Changed.
+
 ## response to reviewers ##
@@
 - the formatting really helps distinguish between editor/reviewer comments and our responses. Nice one. it is so much easier to read.
+
+-> I agree.
 
@@
 
+-> This sentence has been removed.
+I added this sentence: "These minimal requirements are fulfilled by the RT data obtained in single-button detection tasks, where the time-to-response is repeatedly measured in different trials in the same individual.
+In section A of the Supplemental Material we visualize this and other types of time-to-event data which are typically obtained in discrimination and bistable perception tasks."
+Thus, based on section A in the Supplemental Material, the distinction between single event and repeated events should be clear now.
+
 on line 510 of the response, you say:
@@
 Once the page numbers change, it can be a nightmare to find and verufy changes as a reviewer.
+
+-> I agree.

### 2025-07-06 manuscript/manuscript.Rmd
changes to ms.

@@
 The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
-The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+The conditional accuracy function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
@@
 The goal is to find the smallest bin width that is supported by the amount of data available.
-Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
+Based on our experience, a bin width of 40 ms is a good starting value when the number of repeated measures is 100 or less.
 Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
@@
 Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
-To do so using a functional programming approach, one has to nest the data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
+To do so using a functional programming approach, one has to nest the person-trial data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.
 
@@
 More generally, these results go against the (often implicit) assumption in research on priming that all observed responses are primed responses to the target stimulus.
-Instead, the distributional data show that early responses are triggered exclusively by the prime stimulus, while only later responses reflect primed responses to the target stimulus.
+Instead, the distributional data show that fast responses are triggered exclusively by the prime stimulus, while only the slower responses reflect primed responses to the target stimulus.
 
@@
 In general, when fitting regression models, our lab adopts an estimation approach to multilevel regression [@kruschkeBayesianNewStatistics2018; @winterStatisticsLinguistsIntroduction2019], which is heavily influenced by the Bayesian framework as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2020; @kurzStatisticalRethinkingSecondEd2023].
-We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013].
+We also use a "keep it maximal" approach by specifying a full varying (or random) effects structure [@barrRandomEffectsStructure2013].
 This means that wherever possible we include varying intercepts and slopes per participant.
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2020].
+We compare models of different complexity using information criteria and cross-validation, to evaluate out-of-sample predictive accuracy [@mcelreathStatisticalRethinkingBayesian2020].
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 
-We can compare the two models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria [@mcelreathStatisticalRethinkingBayesian2020; @kurzAppliedLongitudinalDataAnalysis2023].
+There are two popular strategies to evaluate how well models will perform in predicting new data on average: Leave-One-Out (LOO) cross-validation and the Widely Applicable Information Criterion or WAIC [@mcelreathStatisticalRethinkingBayesian2020].
+LOO-weights represent the optimal linear combination of models for predictive performance, with higher weights for models with better out-of-sample predictive performance.
+WAIC-weights represent the relative evidence for each model, with higher weights for models with a better fit while accounting for model complexity [@mcelreathStatisticalRethinkingBayesian2020; @kurzAppliedLongitudinalDataAnalysis2023].
 
@@
 ```{r show-weights, echo=T, eval=T}
-model_weights(model_M0i, model_M1i, weights = "loo") %>% round(digits = 2)
-model_weights(model_M0i, model_M1i, weights = "waic") %>% round(digits = 2) 
+model_weights(model_M0i, model_M1i, weights = "loo") %>% round(digits = 2) %>% format(nsmall=2)
+model_weights(model_M0i, model_M1i, weights = "waic") %>% round(digits = 1) %>% format(nsmall=2)
 ```
@@
 
-Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to the other simpler model.
+Clearly, both the loo and waic weighting schemes assign a weight of 1 to model M1i, and a weight of 0 to model M0i.
 
@@
 
-To make inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80% and 95% credible intervals).
+A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 4A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 4A.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
@@
 In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely).
 To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i_ca.

### 2025-07-06 manuscript/manuscript.tex
changes to ms.

@@
 The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
-The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
+The conditional accuracy function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
 We refer to this extended (hazard + conditional accuracy) analysis for choice RT data as EHA/SAT.
@@
 Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
-To do so using a functional programming approach, one has to nest the data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
+To do so using a functional programming approach, one has to nest the person-trial data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.
 
@@
 In general, when fitting regression models, our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020).
-We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013).
+We also use a ``keep it maximal'' approach by specifying a full varying (or random) effects structure (Barr, Levy, Scheepers, \& Tily, 2013).
 This means that wherever possible we include varying intercepts and slopes per participant.
 To make inferences, we use two main approaches.
-We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020).
+We compare models of different complexity using information criteria and cross-validation, to evaluate out-of-sample predictive accuracy (McElreath, 2020).
 We also take the most complex model and evaluate key parameters of interest using point and interval estimates.
@@
 
-We can compare the two models using the Widely Applicable Information Criterion (WAIC) and Leave-One-Out (LOO) cross-validation, and look at model weights for both criteria (Kurz, 2023a; McElreath, 2020).
+There are two popular strategies to evaluate how well models will perform in predicting new data on average: Leave-One-Out (LOO) cross-validation and the Widely Applicable Information Criterion or WAIC (McElreath, 2020).
+LOO-weights represent the optimal linear combination of models for predictive performance, with higher weights for models with better out-of-sample predictive performance.
+WAIC-weights represent the relative evidence for each model, with higher weights for models with a better fit while accounting for model complexity (Kurz, 2023a; McElreath, 2020).
 
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"loo"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{format}\NormalTok{(}\AttributeTok{nsmall=}\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 ## model_M0i model_M1i 
-##         0         1
+##    "0.00"    "1.00"
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{) }
+\FunctionTok{model\_weights}\NormalTok{(model\_M0i, model\_M1i, }\AttributeTok{weights =} \StringTok{"waic"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{round}\NormalTok{(}\AttributeTok{digits =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{format}\NormalTok{(}\AttributeTok{nsmall=}\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 
-To make inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 4A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 4A.
 The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
@@
 In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a.
-The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors.
+The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely).
 To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i\_ca.

### 2025-07-06 manuscript/Suppl_material/supplementary.Rmd
changes to suppl. material

@@
 Note that we do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in @stoolmillerModelingHeterogeneitySocial2006 and @millsIntroducingSurvivalEvent2011.
+More information about recurrent events analysis can be found in @stoolmillerModelingHeterogeneitySocial2006, @millsIntroducingSurvivalEvent2011, and @lougheedMultilevelSurvivalAnalysis2019.
 
@@
 
-While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event does not occur before the endpoint of bin t.
 

### 2025-07-06 manuscript/Suppl_material/supplementary.tex
changes to suppl. material

@@
 Note that we do not analyse recurrent events in this tutorial.
-More information about recurrent events analysis can be found in Stoolmiller and Snyder (2006) and Mills (2011).
+More information about recurrent events analysis can be found in Stoolmiller and Snyder (2006), Mills (2011), and Lougheed, Benson, Cole, and Ram (2019).
 
@@
 
-While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event does not occur before the endpoint of bin t.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.Rmd
changes to response to reviewers

@@
 
-Response: In the revised ms. we frame the unique contribution that this works adds on pages 5-6:
+Response: In the revised ms. we frame the unique contribution that this works adds on page 6:
 "...we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which
@@
 
-Response: We added these references to the ms.
+Response: We added these references to the revised ms. on page 6.
 
@@
 
-Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.
+Response: We added a section in the Supplemental Material (section A in the revised ms.) which explains the types of time-to-event data typically obtained in RT tasks (detection, discrimination, and bistable perception tasks).
+This section makes it clear that our tutorials pertain only to estimating single-event occurrences.
+It also visualizes recurrent events and includes the Lougheed and Stoolmiller tutorials.
 
@@
 
-Response: We disambiguate the various terms in the revised ms. on page 7, and we now only use one term throughout.
+Response: We disambiguate the various terms in the revised ms. on pages 7-8, and we now only use one term throughout.
 
@@
 
-Response: We added these 3 references to the revised ms.
+Response: We added these 3 references to the revised ms. on page 6.
 
@@
 
-Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:
+Response: We address our choice for discrete-time methods on pages 8-9 of the revised ms.:
 
-"...the definition of hazard and the type of models employed depend on whether one is using
-continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
-more experience using discrete-time EHA, and that is the approach that we describe and
-focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
-treated as a continuous variable. However, continuous forms of EHA require much more
-data to estimate the continuous-time hazard (rate) function well. Thus, by trading
-a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
-for dealing with typical psychological RT data sets for which there are less than ~200 trials
-per condition per experiment. Moreover, as indicated by Allison (2010),
+"... the definition of hazard and the type of models employed depend on whether one is using continuous or discrete time units. As a lab, and mainly for practical reasons,
+we have much more experience using discrete-time EHA, and that is the approach that we
+describe and focus on in this paper. This choice may seem counter-intuitive, given that RT
+is typically treated as a continuous variable. However, continuous forms of EHA require
+much more data to estimate the continuous-time hazard (rate) function well (Bloxom,
+1984; Luce, 1991; Van Zandt, 2000). Thus, by trading a bit of temporal resolution for a
+lower number of trials, discrete-time methods seem ideal for dealing with typical
+psychological RT data sets for which there are less than ~200 trials per condition per
+participant (Panis, Schmidt, et al., 2020). Moreover, as indicated by Allison (2010),
 learning discrete-time EHA methods first will help in learning continuous-time methods, so
@@
 
-Response: We discuss the issue of defining bins on page 11 of the revised ms.:
+Response: We discuss the issue of defining bins on page 12 of the revised ms.:
 
 "Third, the width of each time bin will need to be determined. For instance, in Figure
-1B we chose 100ms in an arbitrary manner. In reality, however, bin width will need to be
+1B we chose 100 ms in an arbitrary manner. In reality, however, bin width will need to be
 set by considering a number of factors simultaneously. The optimal bin width will depend
@@
 descriptive statistics (see section 3.1). The goal is to find the smallest bin width that is
-supported by the amount of data available. Based on our experience, a bin width of 50 ms
+supported by the amount of data available. Based on our experience, a bin width of 40 ms
 is a good starting value when the number of repeated measures is 100 or less. Too small
@@
 > 6. Line 517: What are model weights? Please explain briefly for the reader,
-including how the interpretation differs for loo and waic. Also, in the provided
+including how the interpretation differs for loo and waic. 
+Also, in the provided
 reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
 than WAIC, so could you please include a definition of or reference for dWAIC in
-your context? Lastly, when reporting the model weights in the example in the
+your context? 
+Lastly, when reporting the model weights in the example in the
 text, I’d suggest including more than 2 digits so that it is clear if the reported
@@
 
-Response: TODO...
-
-We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.
+Response: We briefly explain model weights for loo and waic on page 28 of the revised ms.
+A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we report more digits when reporting the model weights.
 
@@
 
-Response: We removed the frequentist examples from the main text.
+Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 37 of the revised ms.
+We have no specific idea why the models did not converge, so we simply mention on page 37:
+"... because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here."
 
@@
 
-Response: TODO
+Response:TODO.  To improve the Figure (figure number 2 in the revised ms.). However, we added bin 40 in the revised ms.
 
@@
 
-Response: TODO
+Response: We replace early with fast on page 22 of the revised ms.
 
@@
 
-Response: TODO.
+Response: We erroneously used the term marginal. We removed this term in the revised ms.
 
@@
 
-Response: TODO.
+Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 33 of the revised ms.
+The logit is typically used for logistic regression by experimental psychologists, so we feel that no extra explanation is needed.
 
@@
 
-Response: TODO.
+Response: We explain this on page 24 of the revised ms.:
+"We also use a "keep it maximal" approach by specifying a full varying (or random) effects structure. This means that wherever possible we include varying intercepts and slopes per participant."
 
@@
 
-Response: TODO.
+Response: References are provided on page 8 of the revised ms.
 
@@
 
-Response: We discuss the main contributions above and beyond existing works on pages 5-6 of the revised ms.
-We address the issue of treating time as discrete or continuous on pages 7-8 of the revised ms.
-We removed redundancies, get to the tutorial section quicker (page 11) and included more precise explanations of technical jargon.
+Response: We discuss the main contributions above and beyond existing works on page 6 of the revised ms.
+We address the issue of treating time as discrete or continuous on pages 8-9 of the revised ms.
+We removed redundancies, get to the tutorial section quicker (page 12) and included more precise explanations of technical jargon.
 
@@
 Response: We explain the location of the files and the folder structure in the README file.
-In addition we make it clear that s single download step is possible by either cloning the project, or downloading the ZIP file.
+In addition we make it clear that a single download step is possible by either cloning the project, or downloading the ZIP file.
 We also included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
@@
 
-Response: We included these references in the revised ms.
+Response: We included these references in the revised ms. on page 6.
 
@@
 Response: We placed the equations for h(t) and ca(t) in Figure 1 of the revised ms. 
-Beyond Figure 1, we still choose to keep any further reference of equations to supplementary materials. 
+Beyond Figure 1, we still choose to keep any further reference of equations to the supplemental material (section B). 
 This is because we feel that one barrier to uptake of these methods is an aversion to equations.
-And by placing them in supplementary materials we feel that we strike a nice balance between non-technical accessibility and a fullness of mathemathical explanation for key ideas and concepts that remain easily accessible to interested readers.
+And by placing them in the supplemental material we feel that we strike a nice balance between non-technical accessibility and a fullness of mathemathical explanation for key ideas and concepts that remain easily accessible to interested readers.
 
@@
 
-Response: TODO.
+Response: We discuss general research questions on page 10 of the revised ms.
+
+And we added research questions in each section of Tutorial 2a.
 
@@
 
-Response: Indeed, this was a big error on our part. Thank you for pointing this out.
+Response: Indeed, this was a big error on our part, which we corrected in the revised ms. Thank you for pointing this out.
 
@@
 
-Response: TODO. We add indexes for individual and trial.
+Response: We added indexes for individual and trial in the regression model equations in section E of the supplemental material.
+Furthermore, our new description on pages 8 and 9, together with the new section A in the supplemental material, should make this issue clear for the reader.
 
@@
 
-Response: TODO.
+Response: We define S(t) = P(RT > t) now as "the probability that the event does not occur before the endpoint of bin t" in section B of the Supplemental Material.
 
@@
 
-Response: TODO. We added reference to individuals and trials (repeated measures) in the notation.
+Response: We added reference to individuals and trials (repeated measures) in the notation of the model equations in section E of the Supplemental Material.
+Furthermore, we feel that the information provided by our new description on pages 8 and 9, together with the new section A in the supplemental material, is sufficient for understanding what we mean without including a reference to individuals (and trials) in the main text.
 
@@
 
-Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.
-In section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events.
+Response: In section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events, and we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception tasks).
+This should make it clear to the reader that we are dealing with single events that are repeatedly measured within the same individual.
 
@@
 
-Response: We provided a clearer description of the data.
+Response: We think that there is a misunderstanding of what "trial" means.
+We now provide a clear description of the meaning of "trial" and of the various types of time-to-data obtained in typical RT tasks in section A of the Supplemental Material.
 
@@
 
-Response: TODO. We provide supporting references on page x of the revised ms.
+Response: We provide supporting references on pages 9-10 of the revised ms.
 
@@
 
-Response: Section 2.3 has been rewritten. The new Figure 1 in the tutorial directly focuses  on the comparison between analysing means vs. event occurrence.
+Response: Section 2.3 has been rewritten.
+The new Figure 1 in the tutorial directly focuses  on the comparison between analysing means vs.
+distributional shapes.
+We also discuss research questions on page 10 of the revised ms.
 
@@
 
-Response: We explain the idea of temporal states of cognitive processes now on page 5 of the revised ms. and illustrate it in Figure 1.
+Response: We explain the idea of temporal states of cognitive processes now on pages 3-4 of the revised ms. and illustrate it in Figure 1.
 
@@
 
-Response: TODO.
+Response: We do not discuss frequentist approaches in the main text anymore (to save space) and only mention that the reader can consult Tutorials 3a and 3b.
+We agree that a model should not be presented if it does not converge when it includes random slopes.
+However, to save space, and because this is an active area of reserch, we do not attempt to clarify when it would not be possible to include varying intercepts and slopes.
 
@@
 
-Response: TODO.
+Response: We clarify this in section A of the Supplemental Material.
 
@@
 
-Response: We introduce various types of censoring in section A of the Supplemental Material.
+Response: We discuss various types of censoring (right, left, random informative, random uninformative) in section A of the Supplemental Material.
 
@@
 
-Response: TODO.
+Response: We specify "person-trial data" in the revised ms.
 
@@
 
-Response: TODO. We agree and adjusted the description.
+Response: We agree and adjusted the description (see pages 10-11 of the revised ms.).
 
@@
 
-We removed these messages from the paper and mention them now only in the tutorial itself.
+We now removed these messages from the paper and mention them now only in the tutorial itself.
 
@@
 
-Response: TODO.
+Response: We know of no more detail to report.
 
@@
 
-Response: TODO.
+Response: We changed this sentence into: "... because the first few bins
+typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant." on page 24 of the revised ms.
 
@@
 
-Response: A comparison between the cloglog and logit link functions is provided in section C of the Supplemental Material.
+Response: A visual comparison between the cloglog and logit link functions is provided in section C of the Supplemental Material.
 
@@
 
-Response: TODO.
+Response: We added the following information on page 24 of the revised ms.:
+"Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
+For example, when you want to know if cloglog-hazard is changing linearly or quadracticly over time, you should treat TIME as a continuous predictor.
+When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding."
 
@@
 
-Response: TODO.
+Response: We mention now that model M0i is a baseline or reference model, to be compared with model M01 (see page 26 of the revised ms.).
 
@@
 
-Response: TODO.
+Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 29 of the revised ms.
 
@@
 
-Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in experimental psychology.
+Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in experimental psychology. We now only use the term SOA in the Supplemental Material.
 
@@
 However, based on those results, we can deduce information about how long a psychological state lasts within a trial.
-
 Nevertheless, we replaced this with "how effects evolve with increasing waiting time".
@@
 
-Response: TODO.
+Response: We rewrote this section.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.tex
changes to response to reviewers

@@
 
-Response: Thank you for allowing us to submit a revision of our Tutorial
-(AMPPS-24-0232).
+Response: Thank you for providing constructive feedback and allowing us
+to submit a revision of our manuscript (AMPPS-24-0232).
 
@@
 mention the unique contribution of our work, we removed redundancies,
-and tried to improve the clarity of the paper.
+and tried to improve the clarity of the paper. As such, we have
+substantially revised the manuscript in accordance with feedback from
+you as well as the reviewers.
 
@@
 Response: In the revised ms. we frame the unique contribution that this
-works adds on pages 5-6: ``\ldots we are not aware of any tutorials that
+works adds on page 6: ``\ldots we are not aware of any tutorials that
 are aimed specifically at psychological RT (+ accuracy) data, and which
@@
 
-Response: On page 7 of the revised ms., we mention that we measure
-single-event occurrences that are repeatedly measured in each
-participant.
+Response: We added a section in the Supplemental Material (section A in
+the revised ms.) which explains the types of time-to-event data
+typically obtained in RT tasks (detection, discrimination, and bistable
+perception tasks). This section makes it clear that our tutorials
+pertain only to estimating single-event occurrences. It also visualizes
+recurrent events and includes the Lougheed and Stoolmiller tutorials.
 
@@
 
-Response: We disambiguate the various term in the revised ms. on page 7,
-and use one term throughout.
+Response: We disambiguate the various terms in the revised ms. on pages
+7-8, and we now only use one term throughout.
 
@@
 
-Response: We address our choice for discrete-time methods on pages 7-8
+Response: We address our choice for discrete-time methods on pages 8-9
 of the revised ms.:
 
-``\ldots the definition of hazard and the type of models employed depend
-on whether one is using continuous or discrete time units. As a lab, and
-mainly for practical reasons, we have much more experience using
-discrete-time EHA, and that is the approach that we describe and focus
-on in this paper. This choice may seem counter-intuitive, given that RT
-is typically treated as a continuous variable. However, continuous forms
-of EHA require much more data to estimate the continuous-time hazard
-(rate) function well. Thus, by trading a bit of temporal resolution for
-a lower number of trials, discrete-time methods seem ideal for dealing
+``\ldots{} the definition of hazard and the type of models employed
+depend on whether one is using continuous or discrete time units. As a
+lab, and mainly for practical reasons, we have much more experience
+using discrete-time EHA, and that is the approach that we describe and
+focus on in this paper. This choice may seem counter-intuitive, given
+that RT is typically treated as a continuous variable. However,
+continuous forms of EHA require much more data to estimate the
+continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991;
+Van Zandt, 2000). Thus, by trading a bit of temporal resolution for a
+lower number of trials, discrete-time methods seem ideal for dealing
 with typical psychological RT data sets for which there are less than
-\textasciitilde200 trials per condition per experiment. Moreover, as
-indicated by Allison (2010), learning discrete-time EHA methods first
-will help in learning continuous-time methods, so it seems like a good
-starting point.''
+\textasciitilde200 trials per condition per participant (Panis, Schmidt,
+et al., 2020). Moreover, as indicated by Allison (2010), learning
+discrete-time EHA methods first will help in learning continuous-time
+methods, so it seems like a good starting point.''
 
@@
 
-Response: We moved the explanation of the survival function to the
-Supplementary Material in the revised ms.
+Response: Thank you for the positive feedback. To make the manuscript
+more accessible and shorter and to respond to some other comments from
+the editor and other reviewers, we moved the explanation of the survival
+function to the Supplementary Material in the revised ms.
 
@@
 \end{quote}
+
+Thank you very much for your helpful and positive feedback.
 
@@
 
-Response: We removed repetitive text, and try to focus on clarity and
-brevity in the revised ms.
+Response: Thanks for your constructive and encouraging feedback. As we
+also noted in our response to the editor, we have made substantial
+revisions to the manuscript that include removing repetitive text, and
+placing a stronger focus on clarity and brevity in the revised ms.
 
@@
 
-Response: We discuss the issue of defining bins on page 11 of the
+Response: We discuss the issue of defining bins on page 12 of the
 revised ms.:
@@
 ``Third, the width of each time bin will need to be determined. For
-instance, in Figure 1B we chose 100ms in an arbitrary manner. In
+instance, in Figure 1B we chose 100 ms in an arbitrary manner. In
 reality, however, bin width will need to be set by considering a number
@@
 find the smallest bin width that is supported by the amount of data
-available. Based on our experience, a bin width of 50 ms is a good
+available. Based on our experience, a bin width of 40 ms is a good
 starting value when the number of repeated measures is 100 or less. Too
@@
 
-Response: TODO\ldots{}
-
-We explain model weights on page x, include a reference for dWAIC, and
-report the weights with 3 digits.
+Response: We briefly explain model weights for loo and waic on page 28
+of the revised ms. A discussion of WAIC in Kurz (2023a) can be found in
+section 4.6.4., so we did not include a reference for dWAIC in the
+revised ms. Lastly, we report more digits when reporting the model
+weights.
 
@@
 
-Response: We removed the frequentist examples from the main text.
+Response: We removed the frequentist examples from the main text and
+direct the reader to the Tutorials on page 37 of the revised ms. We have
+no specific idea why the models did not converge, so we simply mention
+on page 37: ``\ldots{} because multilevel generalized linear regression
+models often do not converge with complex random-coefficient structures,
+we do not discuss them here.''
 
@@
 
-Response: TODO
+Response:TODO. To improve the Figure (figure number 2 in the revised
+ms.). However, we added bin 40 in the revised ms.
 
@@
 
-Response: TODO.
+Response: We erroneously used the term marginal. We removed this term in
+the revised ms.
 
@@
 
-Response: TODO.
+Response: We now write ``The general process is similar to Tutorial 2a,
+except that (a) we use the person-trial data, (b) we use the symmetric
+logit link function, and (c) we change the priors (our prior belief is
+that conditional accuracy values between 0 and 1 are equally likely).''
+on page 33 of the revised ms. The logit is typically used for logistic
+regression by experimental psychologists, so we feel that no extra
+explanation is needed.
 
@@
 
-Response: TODO.
+Response: We explain this on page 24 of the revised ms.: ``We also use
+a''keep it maximal'' approach by specifying a full varying (or random)
+effects structure. This means that wherever possible we include varying
+intercepts and slopes per participant.''
 
@@
 Response: We discuss the main contributions above and beyond existing
-works on pages 5-6 of the revised ms. We address the issue of treating
-time as discrete or continuous on pages 7-8 of the revised ms. We
-removed redundancies, get to the tutorial section quicker (page 11) and
-included more precise explanations of technical jargon.
+works on page 6 of the revised ms. We address the issue of treating time
+as discrete or continuous on pages 8-9 of the revised ms. We removed
+redundancies, get to the tutorial section quicker (page 12) and included
+more precise explanations of technical jargon.
 
@@
 
-Response: We explain the location of the files in the README file. A
-single download step is possible by either cloning the project, or
-downloading the ZIP file. We included a brief description of a life
-table and how to interpret the plot in the R Markdown tutorial files.
+Response: We explain the location of the files and the folder structure
+in the README file. In addition we make it clear that a single download
+step is possible by either cloning the project, or downloading the ZIP
+file. We also included a brief description of a life table and how to
+interpret the plot in the R Markdown tutorial files. At this point, we
+also prefer a modular approach to the tutorials rather than combining
+them. We appreciate that not everyone might share this preference, but
+we think there are many benefits to a modular workflow in this instance.
 
@@
 Response: We placed the equations for h(t) and ca(t) in Figure 1 of the
-revised ms.
+revised ms. Beyond Figure 1, we still choose to keep any further
+reference of equations to the supplemental material (section B). This is
+because we feel that one barrier to uptake of these methods is an
+aversion to equations. And by placing them in the supplemental material
+we feel that we strike a nice balance between non-technical
+accessibility and a fullness of mathemathical explanation for key ideas
+and concepts that remain easily accessible to interested readers.
 
@@
 
-Response: TODO.
+Response: We discuss general research questions on page 10 of the
+revised ms.
+
+And we added research questions in each section of Tutorial 2a.
 
@@
 
-Response: Indeed, this was a big error on our part. Thank you for
-pointing this out.
+Response: Indeed, this was a big error on our part, which we corrected
+in the revised ms. Thank you for pointing this out.
 
@@
 
-Response: TODO. We add indexes for individual and trial.
+Response: We added indexes for individual and trial in the regression
+model equations in section E of the supplemental material. Furthermore,
+our new description on pages 8 and 9, together with the new section A in
+the supplemental material, should make this issue clear for the reader.
 
@@
 
-Response: TODO.
+Response: We define S(t) = P(RT \textgreater{} t) now as ``the
+probability that the event does not occur before the endpoint of bin t''
+in section B of the Supplemental Material.
 
@@
 
-Response: TODO. We added reference to individuals and trials (repeated
-measures) in the notation.
+Response: We added reference to individuals and trials (repeated
+measures) in the notation of the model equations in section E of the
+Supplemental Material. Furthermore, we feel that the information
+provided by our new description on pages 8 and 9, together with the new
+section A in the supplemental material, is sufficient for understanding
+what we mean without including a reference to individuals (and trials)
+in the main text.
 
@@
 
-Response: We decided to keep tutorial 4.
+Response: We decided to keep tutorial 4, as we feel that planning future
+studies is an important and difficult step to take in these kinds of
+studies, which we wanted to keep close to the rest of the material for
+ease of understanding a reasonably complete workflow.
 
@@
 
-Response: On page 7 of the revised ms., we mention that we measure
-single-event occurrences that are repeatedly measured in each
-participant. In section A of the Supplemental Material we mention that
-this Tutorial does not cover recurrent events.
+Response: In section A of the Supplemental Material we mention that this
+Tutorial does not cover recurrent events, and we visualize the different
+types of time-to-event data that are obtained in typical RT tasks
+(detection, discrimination, bistable perception tasks). This should make
+it clear to the reader that we are dealing with single events that are
+repeatedly measured within the same individual.
 
@@
 
-Response: We provided a clearer description of the data.
+Response: We think that there is a misunderstanding of what ``trial''
+means. We now provide a clear description of the meaning of ``trial''
+and of the various types of time-to-data obtained in typical RT tasks in
+section A of the Supplemental Material.
 
@@
 
-Response: TODO. We provide supporting references on page x of the
-revised ms.
+Response: We provide supporting references on pages 9-10 of the revised
+ms.
 
@@
 tutorial directly focuses on the comparison between analysing means
-vs.~event occurrence.
+vs.~distributional shapes. We also discuss research questions on page 10
+of the revised ms.
 
@@
 Response: We explain the idea of temporal states of cognitive processes
-now on page 5 of the revised ms. and illustrate it in Figure 1.
+now on pages 3-4 of the revised ms. and illustrate it in Figure 1.
 
@@
 
-Response: TODO.
+Response: We do not discuss frequentist approaches in the main text
+anymore (to save space) and only mention that the reader can consult
+Tutorials 3a and 3b. We agree that a model should not be presented if it
+does not converge when it includes random slopes. However, to save
+space, and because this is an active area of reserch, we do not attempt
+to clarify when it would not be possible to include varying intercepts
+and slopes.
 
@@
 
-Response: We introduce various types of censoring in section A of the
-Supplemental Material.
+Response: We discuss various types of censoring (right, left, random
+informative, random uninformative) in section A of the Supplemental
+Material.
 
@@
 
-Response: TODO.
+Response: We specify ``person-trial data'' in the revised ms.
 
@@
 
-Response: TODO. We agree and adjusted the description.
+Response: We agree and adjusted the description (see pages 10-11 of the
+revised ms.).
 
@@
 
-We removed these messages from the paper and mention them now only in
-the tutorial itself.
+We now removed these messages from the paper and mention them now only
+in the tutorial itself.
 
@@
 
-Response: TODO.
+Response: We changed this sentence into: ``\ldots{} because the first
+few bins typically contain no responses, one has to select an analysis
+time window, i.e., a contiguous set of bins for which there is data for
+each participant.'' on page 24 of the revised ms.
 
@@
 
-Response: A comparison between the cloglog and logit link functions is
-provided in section C of the Supplemental Material.
+Response: A visual comparison between the cloglog and logit link
+functions is provided in section C of the Supplemental Material.
 
@@
 
-Response: TODO.
+Response: We added the following information on page 24 of the revised
+ms.: ``Third, one has to choose whether to treat TIME (i.e., the time
+bin index t) as a categorical or continuous predictor (see also section
+E of the Supplemental Material). For example, when you want to know if
+cloglog-hazard is changing linearly or quadracticly over time, you
+should treat TIME as a continuous predictor. When you are only
+interested in the effect of covariates on hazard, you can treat TIME as
+a categorical predictor (i.e., fit an intercept for each bin), in which
+case you can choose between reference coding and index coding.''
 
@@
 
-Response: TODO.
+Response: We mention now that model M0i is a baseline or reference
+model, to be compared with model M01 (see page 26 of the revised ms.).
 
@@
 
-Response: TODO.
+Response: We added the purpose (``To make causal inferences\ldots{}'')
+and define a credible interval on page 29 of the revised ms.
 
@@
 Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in
-experimental psychology.
+experimental psychology. We now only use the term SOA in the
+Supplemental Material.
 
@@
 states occur. However, based on those results, we can deduce information
-about how long a psychological state lasts within a trial.
-
-Nevertheless, we replaced this with ``how effects evolve with increasing
-waiting time''.
+about how long a psychological state lasts within a trial. Nevertheless,
+we replaced this with ``how effects evolve with increasing waiting
+time''.
 

### 2025-07-06 Tutorial_1a.Rmd
update tutorial 1a

@@
 
-In Tutorial 1a we wrangle raw data from Experiment 1 of Panis and Schmidt (2016), and illustrate how to calculate descriptive statistics.
+In Tutorial 1a we wrangle raw data from Experiment 1 of Panis and Schmidt (2016), and illustrate how to calculate descriptive statistics (functions of discrete time).
 We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
-To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (i.e., a cutoff time, or timeout).
+
+A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t) = P(RT = t | RT >= t), survival probability S(t) = P(Rt > t), probability mass P(t) = P(RT = t), possibly the conditional accuracy ca(t) = P(correct | RT = t), and their estimated standard errors (se).
+
+To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (a cutoff time, or timeout; the response deadline).
 The goal is to find the optimal bin width and censoring time (both in ms) to portray the effects of interest.
@@
     # create plots, and set haz_yaxis to 1, and first_bin_shown to 2
-    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha, 1, 2))  
+    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha, 1, 1))  
     # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
@@
 
-Save the plots.
+Note that warning messages are generated when plucking a plot, like this:
+"1: Removed 2 rows containing missing values or values outside the scale range".
+These warning messages are generated because some bins have no hazard and ca(t) estimates, and also no error bars. They can thus safely be ignored.
+
+Now save the plots.
 
@@
 
-# extract the lifetable (column 7 in data_final) for condition 1 (neutral) for subject 6
+# extract the lifetable (column 7 in data_final) for condition 1 (blank) for subject 6
 lifetable_blank_s6 <- data_final[[7]][[6]] %>% 

### 2025-07-06 manuscript/manuscript.Rmd
change ms. + suppl. mat.

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "11664 (body) + 1593 (references) + 2394 (supplemental material)"
+wordcount         : "10131 (body) + 1709 (references) + 3473 ( body supplemental material) + 393 (refs suppl. mat.)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]

### 2025-07-06 manuscript/manuscript.tex
change ms. + suppl. mat.

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 11664 (body) + 1593 (references) + 2394 (supplemental material)}
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10131 (body) + 1709 (references) + 3473 ( body supplemental material) + 393 (refs suppl.
+mat.)}
 \usepackage{lineno}

### 2025-07-06 Tutorial_2a.Rmd
update tutorial 2a (change figure)

@@
                           fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",.width=.95,alpha = 0.5) +
+    stat_lineribbon(point_interval="median_qi",.width=c(.8,.95),alpha = 0.5) +
     scale_fill_brewer(palette = "Dark2") +
@@
     stat_lineribbon(point_interval="median_qi",
-                    .width=.95,
+                    .width=c(.8,.95),
                     alpha = 0.5) +

### 2025-07-06 manuscript/reviews/ResponseReview1.Rmd
update response to reviewers

@@
 
-Response:TODO.  To improve the Figure (figure number 2 in the revised ms.). However, we added bin 40 in the revised ms.
+Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin in Figure 2 (previously 4) of the revised ms.
 
@@
 
-Response: TODO -> correction needed.
+Response: We corrected this (Figure 4 in the revised ms.).
 
@@
 In addition we make it clear that a single download step is possible by either cloning the project, or downloading the ZIP file.
-We also included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.
+We also included a brief description of a life table and the definitions of the discrete-time functions in the R Markdown file called Tutorial_1a.Rmd.
 At this point, we also prefer a modular approach to the tutorials rather than combining them. 
@@
 distributional shapes.
-We also discuss research questions on page 10 of the revised ms.
+We also discuss general research questions on page 10 of the revised ms.
 
@@
 We agree that a model should not be presented if it does not converge when it includes random slopes.
-However, to save space, and because this is an active area of reserch, we do not attempt to clarify when it would not be possible to include varying intercepts and slopes.
+However, to save space, and because this is an active area of reserch, we do not attempt to clarify when it would not be possible to include varying intercepts and/or slopes.
 
@@
 Response: On line 362 it was stated that "the warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored".
-
 We now removed these messages from the paper and mention them now only in the tutorial itself.
@@
 
-Response: We know of no more detail to report.
+Response: We have no more detail to share on this issue.
+On page 19 of the revised ms.
+we now write "In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may not be following task instructions (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.".
 

### 2025-07-06 manuscript/reviews/ResponseReview1.tex
update response to reviewers

@@
 
-Response:TODO. To improve the Figure (figure number 2 in the revised
-ms.). However, we added bin 40 in the revised ms.
+Response: Because no responses occurred in the first bin with endpoint
+40 ms. However, we added this bin in Figure 2 (previously 4) of the
+revised ms.
 
@@
 
-Response: TODO -\textgreater{} correction needed.
+Response: We corrected this (Figure 4 in the revised ms.).
 
@@
 step is possible by either cloning the project, or downloading the ZIP
-file. We also included a brief description of a life table and how to
-interpret the plot in the R Markdown tutorial files. At this point, we
-also prefer a modular approach to the tutorials rather than combining
-them. We appreciate that not everyone might share this preference, but
-we think there are many benefits to a modular workflow in this instance.
+file. We also included a brief description of a life table and the
+definitions of the discrete-time functions in the R Markdown file called
+Tutorial\_1a.Rmd. At this point, we also prefer a modular approach to
+the tutorials rather than combining them. We appreciate that not
+everyone might share this preference, but we think there are many
+benefits to a modular workflow in this instance.
 
@@
 tutorial directly focuses on the comparison between analysing means
-vs.~distributional shapes. We also discuss research questions on page 10
-of the revised ms.
+vs.~distributional shapes. We also discuss general research questions on
+page 10 of the revised ms.
 
@@
 to clarify when it would not be possible to include varying intercepts
-and slopes.
+and/or slopes.
 
@@
 generated because some bins have no hazard and ca(t) estimates, and no
-error bars. They can thus safely be ignored''.
-
-We now removed these messages from the paper and mention them now only
-in the tutorial itself.
+error bars. They can thus safely be ignored''. We now removed these
+messages from the paper and mention them now only in the tutorial
+itself.
 
@@
 
-Response: We know of no more detail to report.
+Response: We have no more detail to share on this issue. On page 19 of
+the revised ms. we now write ``In general, it is important to visually
+inspect the functions first for each participant, in order to identify
+individuals that may not be following task instructions (e.g., a flat
+conditional accuracy function at .5 indicates that someone is just
+guessing), outlying individuals, and/or different groups with
+qualitatively different behavior.''.
 

### 2025-07-06 manuscript/manuscript.Rmd
recreate pdfs after updating R

@@
 # R references
- r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation/ creates a .bib-reference file
+ r_refs(file = "r-references.bib") #, append=F) # append=F prevents multiple copies of the same citation/ creates a .bib-reference file
  my_r_citation <- cite_r(file = "r-references.bib", footnote = FALSE)

### 2025-07-06 manuscript/manuscript.tex
recreate pdfs after updating R

@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.4.0; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 

### 2025-07-06 Tutorial_2a.Rmd
update Tutorial 2a

@@
 
-Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the "average" participant at the population level, and for each participant in the data set.
+Following Heiss (2021) -https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/#average-marginal-effects-, we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the "average" participant at the population level, and for each participant in the data set.
 
@@
 
-Following Heiss (2021), we are actually interested in the difference in predicted hazard between congruent and blank primes on the one hand, and between incongruent and blank primes on the other hand, for each time bin.
+Following Heiss (2021) - https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/#average-marginal-effects -, we are actually interested in the difference in predicted hazard between congruent and blank primes on the one hand, and between incongruent and blank primes on the other hand, for each time bin.
 When based on the grand mean, this is known as the grand average marginal effect (AME).

### 2025-07-06 manuscript/reviews/ResponseReview1.Rmd
update Response to reviewers.

@@
 
-Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin in Figure 2 (previously 4) of the revised ms.
+Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin back in Figure 2 (previously 4) of the revised ms.
 
@@
 
-Response: We replace early with fast on page 22 of the revised ms.
+Response: We replaced early with fast on page 22 of the revised ms.
 
@@
 
-Response: Index coding is not used a lot in experimental psychology so we included a short discussion.
+Response: Index coding is not used a lot in experimental psychology so we included a short description.
 
@@
 
-Response: We erroneously used the term marginal. We removed this term in the revised ms.
+Response: We feel there is no real consensus in the literature about the use of the term marginal.
+We thus removed this term in the revised ms.
+But we kept the term in tutorials 2a and 2b where we base our use of the terms on this cited reference: Heiss (2021) - https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/#average-marginal-effects.
 
@@
 
-Response: Yes this has been corrected now.
+Response: Yes, this has been corrected now.
 
@@
 
-Response: We discuss general research questions on page 10 of the revised ms.
+Response: We now discuss general research questions on page 10 of the revised ms.
 
-And we added research questions in each section of Tutorial 2a.
+We did not add more information about possible research questions in each section of Tutorial 2a, because we feel that (a) research questions are highly dependent on the field of study (perception, behavioral control, memory), (b) this would make the tutorial even longer, and (c) the knowledge of experimental psychologists on regression modeling should be large enough for them to link the model structure to possible research questions, once they have read the paper and worked through the code examples.
 
@@
 We agree that a model should not be presented if it does not converge when it includes random slopes.
-However, to save space, and because this is an active area of reserch, we do not attempt to clarify when it would not be possible to include varying intercepts and/or slopes.
+However, to save space, and because this is an active area of research, we do not attempt to clarify when it would be possible to include varying intercepts and/or slopes.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.tex
update Response to reviewers.

@@
 Response: Because no responses occurred in the first bin with endpoint
-40 ms. However, we added this bin in Figure 2 (previously 4) of the
+40 ms. However, we added this bin back in Figure 2 (previously 4) of the
 revised ms.
@@
 Response: Index coding is not used a lot in experimental psychology so
-we included a short discussion.
+we included a short description.
 
@@
 
-Response: We erroneously used the term marginal. We removed this term in
-the revised ms.
+Response: We feel there is no real consensus in the literature about the
+use of the term marginal. We thus removed this term in the revised ms.
+But we kept the term in tutorials 2a and 2b where we base our use of the
+terms on this cited reference: Heiss (2021) -
+\url{https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/\#average-marginal-effects}.
 
@@
 
-Response: We discuss general research questions on page 10 of the
+Response: We now discuss general research questions on page 10 of the
 revised ms.
 
-And we added research questions in each section of Tutorial 2a.
+We did not add more information about possible research questions in
+each section of Tutorial 2a, because we feel that (a) research questions
+are highly dependent on the field of study (perception, behavioral
+control, memory), (b) this would make the tutorial even longer, and (c)
+the knowledge of experimental psychologists on regression modeling
+should be large enough for them to link the model structure to possible
+research questions, once they have read the paper and worked through the
+code examples.
 
@@
 does not converge when it includes random slopes. However, to save
-space, and because this is an active area of reserch, we do not attempt
-to clarify when it would not be possible to include varying intercepts
+space, and because this is an active area of research, we do not attempt
+to clarify when it would be possible to include varying intercepts
 and/or slopes.

### 2025-07-06 manuscript/manuscript.Rmd
update ms.

@@
 We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design.
-We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations of this work.
 Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
@@
 
-(ref:descr-fig1-caption) Simulated data showing mean performance versus distributional (EHA/SAT) analyses.
+(ref:descr-fig1-caption) Simulated single-subject data showing mean performance versus distributional (EHA/SAT) analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 The first bin is (0,100], the last bin is (900,1000].
+Two hundred trials were simulated in each condition.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019].
+For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019; @millsIntroducingSurvivalEvent2011].
 We supply a few regression equations in section E of the Supplemental Material.
@@
 The goal is to find the smallest bin width that is supported by the amount of data available.
-Based on our experience, a bin width of 40 ms is a good starting value when the number of repeated measures is 100 or less.
+Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
 Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
@@
 Our data wrangling procedures serve two related purposes.
-First, we want to summarise and visualise descriptive statistics using a life table.
+First, we want to calculate descriptive statistics for each condition in each individual using a life table.
 A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
@@
 
-Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
+Figure 2 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
 By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any RT data set.
 
-(ref:descr-plot4-caption) Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
+(ref:descr-plot2-caption) Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
 Vertical dotted lines indicate the estimated median RTs.
@@
 
-```{r eha-plot, fig.cap = "(ref:descr-plot4-caption)", out.width="90%"}
+```{r eha-plot, fig.cap = "(ref:descr-plot2-caption)", out.width="90%"}
 knitr::include_graphics("../Tutorial_1_descriptive_stats/figures/Plot_for_subject6_PanisSchmidt.png")
@@
 
-At this point, we have calculated and plotted the descriptive statistics for each type of prime stimulus in EHA/SAT.
+At this point, we have calculated and plotted the descriptive statistics for each type of prime stimulus.
 As we will show in later Tutorials, statistical models for hazard and conditional accuracy functions can be implemented as generalized linear mixed regression models predicting event occurrence (1/0) and conditional accuracy (1/0) in each bin of a selected time window for analysis.
@@
 
-The middle column of Supplementary Figure 3 (section F of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 3 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Supplementary Figure 2 (section D of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 2 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 
@@
 
-To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80% and 95% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
 
-(ref:descr-plot5-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
+(ref:descr-plot3-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
 
-```{r plot-fixed-effects, fig.cap = "(ref:descr-plot5-caption)", out.width='80%'}
+```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_postdistr.png")
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 6A), and for each participant in the data set (Figure 6B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 4A), and for each participant in the data set (Figure 4B).
 
-(ref:descr-plot6-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
+(ref:descr-plot4-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
 
-```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
+```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot4-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_combined.png")
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A) and at the participant level (Figure 7B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 5A) and at the participant level (Figure 5B).
 
-(ref:descr-plot7-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).
+(ref:descr-plot5-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).
 
-```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot7-caption)", out.width='80%'}
+```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot5-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ame_combined.png")
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 4A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 5A.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 4B.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
+Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
+In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
 
@@
 
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 4B).
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
-In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
-
 ## 3.4 Tutorial 2b: Fitting Bayesian conditional accuracy models
@@
 
-To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 8, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 6, together with point (median) and interval estimates (80% and 95% credible intervals).
 
-(ref:descr-plot8-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca. ca = conditional accuracy.
+(ref:descr-plot6-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca. ca = conditional accuracy.
 
-```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 9A), and for each participant in the data set (Figure 9B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 7A), and for each participant in the data set (Figure 7B).
 
-(ref:plot9-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
+(ref:plot7-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
 
-```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot9-caption)", out.width='80%'}
+```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot7-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png")
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A) and for each participant (Figure 10B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 8A) and for each participant (Figure 8B).
 
-(ref:descr-plot10-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).
+(ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).
 
-```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot10-caption)", out.width='80%'}
+```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png")
@@
 
-Based on Figure 10A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Based on Figure 8A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 8A).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
@@
 We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
-The raw data and the simulated data are plotted in Figure 12 and show quite close correspondence, which is re-assuring.
+The raw data and the simulated data are plotted in Figure 9 and show quite close correspondence, which is re-assuring.
 But, this is only one dataset.
@@
 
-Summary results from Simulation 1 are shown in Figure 13A.
-Figure 13A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 10A.
+Figure 10A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
@@
 Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
-Summary results from Simulation 2 are shown in Figure 13B.
+Summary results from Simulation 2 are shown in Figure 10B.
 A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):

### 2025-07-06 manuscript/manuscript.tex
update ms.

--- 2025-07-06 manuscript/manuscript.tex
@@
 
-\caption{Simulated data showing mean performance versus distributional (EHA/SAT) analyses.
+\caption{Simulated single-subject data showing mean performance versus distributional (EHA/SAT) analyses.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 The first bin is (0,100{]}, the last bin is (900,1000{]}.
+Two hundred trials were simulated in each condition.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 In this section, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology.
-For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Singer \& Willett, 2003; Winter, 2019).
+For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Mills, 2011; Singer \& Willett, 2003; Winter, 2019).
 We supply a few regression equations in section E of the Supplemental Material.
@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.4.0; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.1; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.13.1; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{rstan} (Version 2.32.6; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.4; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 
@@
 
-Figure 4 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
+Figure 2 displays the discrete-time hazard, survivor, conditional accuracy, and probability mass functions for each prime condition for participant 6.
 By using discrete-time hazard functions of event occurrence -- in combination with conditional accuracy functions for two-choice tasks -- one can provide an unbiased, time-varying, and probabilistic description of the latency and accuracy of responses based on all trials of any RT data set.
@@
 
-To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 5, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
@@
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 6A), and for each participant in the data set (Figure 6B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 4A), and for each participant in the data set (Figure 4B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 7A) and at the participant level (Figure 7B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 5A) and at the participant level (Figure 5B).
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 4A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 5A.
 The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 4B.
+If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
-
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time (Figure 4B).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time.
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
@@
 
-To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 8, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 6, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 
@@
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 9A), and for each participant in the data set (Figure 9B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 7A), and for each participant in the data set (Figure 7B).
 
@@
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 10A) and for each participant (Figure 10B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 8A) and for each participant (Figure 8B).
 
@@
 
-Based on Figure 10A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 10A).
+Based on Figure 8A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 8A).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
@@
 
-Summary results from Simulation 1 are shown in Figure 13A.
-Figure 13A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 10A.
+Figure 10A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.

### 2025-07-06 manuscript/reviews/ResponseReview1.Rmd
update responses to reviewers

@@
 descriptive statistics (see section 3.1). The goal is to find the smallest bin width that is
-supported by the amount of data available. Based on our experience, a bin width of 40 ms
+supported by the amount of data available. Based on our experience, a bin width of 50 ms
 is a good starting value when the number of repeated measures is 100 or less. Too small
@@
 
-We did not add more information about possible research questions in each section of Tutorial 2a, because we feel that (a) research questions are highly dependent on the field of study (perception, behavioral control, memory), (b) this would make the tutorial even longer, and (c) the knowledge of experimental psychologists on regression modeling should be large enough for them to link the model structure to possible research questions, once they have read the paper and worked through the code examples.
+We did not add more information about possible research questions in each section of Tutorial 2a, because we feel that (a) research questions are highly dependent on the field of study (perception, behavioral control, memory), (b) this would make the tutorial even longer, and (c) we mention in the paper that we do not discuss **why** you might perform EHA.
+In other words, we think that the knowledge of experimental psychologists on regression modeling should be large enough for them to link the model structure to possible research questions, once they have read the paper and worked through the code examples.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.tex
update responses to reviewers

@@
 find the smallest bin width that is supported by the amount of data
-available. Based on our experience, a bin width of 40 ms is a good
+available. Based on our experience, a bin width of 50 ms is a good
 starting value when the number of repeated measures is 100 or less. Too
@@
 control, memory), (b) this would make the tutorial even longer, and (c)
-the knowledge of experimental psychologists on regression modeling
-should be large enough for them to link the model structure to possible
-research questions, once they have read the paper and worked through the
-code examples.
+we mention in the paper that we do not discuss \textbf{why} you might
+perform EHA. In other words, we think that the knowledge of experimental
+psychologists on regression modeling should be large enough for them to
+link the model structure to possible research questions, once they have
+read the paper and worked through the code examples.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.Rmd
update ms. and Response to reviewers

@@
 Response: We briefly explain model weights for loo and waic on page 28 of the revised ms.
-A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we report more digits when reporting the model weights.
+A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we print more digits when reporting the model weights.
 
@@
 
-Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin back in Figure 2 (previously 4) of the revised ms.
+Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin back in Figure 2 (previously 4) of the revised ms to avoid confusion.
 
@@
 
-Response: P(t) was and still is defined only in the Supplementary Material (section B).
+Response: P(t) was and still is defined only in the Supplementary Material (section B in the revised version).
 
@@
 "Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
-For example, when you want to know if cloglog-hazard is changing linearly or quadracticly over time, you should treat TIME as a continuous predictor.
+For example, when you want to know if cloglog-hazard is changing linearly or quadraticly over time, you should treat TIME as a continuous predictor.
 When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding."
@@
 Response: We mention now that model M0i is a baseline or reference model, to be compared with model M01 (see page 26 of the revised ms.).
+We also mention that we only report 2 models in the paper from a larger set of models that are fitted in the file Tutorial_2a.Rmd.
 
@@
 
-Response: We decided to keep this example, as it can help readers with interpreting effects in hazard and conditional accuracy functions.
+Response: We removed this text.
 
@@
 
-Response: We decided to keep tutorial 4, and discuss its usefulness on page 45 of the revised ms.
+Response: We decided to keep tutorial 4, and discuss its usefulness on pages 45-46 of the revised ms.
 
@@
 
-Response: We shortened the discussion section.
+Response: We shortened the Discussion section.
 

### 2025-07-06 manuscript/reviews/ResponseReview1.tex
update ms. and Response to reviewers

@@
 section 4.6.4., so we did not include a reference for dWAIC in the
-revised ms. Lastly, we report more digits when reporting the model
+revised ms. Lastly, we print more digits when reporting the model
 weights.
@@
 40 ms. However, we added this bin back in Figure 2 (previously 4) of the
-revised ms.
+revised ms to avoid confusion.
 
@@
 Response: P(t) was and still is defined only in the Supplementary
-Material (section B).
+Material (section B in the revised version).
 
@@
 E of the Supplemental Material). For example, when you want to know if
-cloglog-hazard is changing linearly or quadracticly over time, you
-should treat TIME as a continuous predictor. When you are only
-interested in the effect of covariates on hazard, you can treat TIME as
-a categorical predictor (i.e., fit an intercept for each bin), in which
-case you can choose between reference coding and index coding.''
+cloglog-hazard is changing linearly or quadraticly over time, you should
+treat TIME as a continuous predictor. When you are only interested in
+the effect of covariates on hazard, you can treat TIME as a categorical
+predictor (i.e., fit an intercept for each bin), in which case you can
+choose between reference coding and index coding.''
 
@@
 model, to be compared with model M01 (see page 26 of the revised ms.).
+We also mention that we only report 2 models in the paper from a larger
+set of models that are fitted in the file Tutorial\_2a.Rmd.
 
@@
 
-Response: We decided to keep this example, as it can help readers with
-interpreting effects in hazard and conditional accuracy functions.
+Response: We removed this text.
 
@@
 Response: We decided to keep tutorial 4, and discuss its usefulness on
-page 45 of the revised ms.
+pages 45-46 of the revised ms.
 

### 2025-07-06 manuscript/manuscript.Rmd
update ms. and Response to reviewers

@@
 Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
-For example, when you want to know if cloglog-hazard is changing linearly or quadracticly over time, you should treat TIME as a continuous predictor.
+For example, when you want to know if cloglog-hazard is changing linearly or quadraticly over time, you should treat TIME as a continuous predictor.
 When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding.
@@
 
-The middle column of Supplementary Figure 2 (section D of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
-While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 2 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+The middle column of Supplementary Figure 3 (section F of the Supplemental Material) shows six examples of prior distributions for an intercept on the logit and/or cloglog scales.
+While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 3 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
 

### 2025-07-06 manuscript/manuscript.tex
update ms. and Response to reviewers

@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 

### 2025-07-07 manuscript/manuscript.Rmd
update ms.

@@
 We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design.
-We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations of this work.
+We finish the article by considering the benefits of the approach for understanding psychological states, as well as its limitations.
 Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms.
+The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
 The first bin is (0,100], the last bin is (900,1000].
@@
 
-Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
+Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
 
@@
 
-To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10 time bins; Figure 1B).
 Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular trial (i.e., repeated measure). 
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-Therefore, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
+In any case, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
 

### 2025-07-07 manuscript/manuscript.tex
update ms.

@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms.
+The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
 The first bin is (0,100{]}, the last bin is (900,1000{]}.
@@
 
-To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1:10 time bins; Figure 1B).
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10 time bins; Figure 1B).
 Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial (i.e., repeated measure).
@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 

### 2025-07-07 manuscript/reviews/ResponseReview1.Rmd
update date in Response to Reviewers

@@
 author: "Panis & Ramsey"
-date: "24.06.2025"
+date: "07.07.2025"
 output: 

### 2025-07-07 manuscript/manuscript.Rmd
edits on the revision

@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology [@bolgerCausalProcessesPsychology2019].
+Indeed, the view that mean values represent truth and variations around the mean are error is deeply ingrained in experimental psychology [@bolgerCausalProcessesPsychology2019].
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events [@panisHowCanWe2020; @panisNeuropsychologicalEvidenceTemporal2017; @panisStudyingDynamicsVisual2020; @panisTimecourseContingenciesPerceptual2009; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022; @wolkersdorferTemporalDynamicsSequential2020].
@@
 
-As a simple illustration, Figure 1 summarises simulated single-subject data (200 trials per condition) that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
+As a simple illustration, Figure 1 summarises simulated data for one subject that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
@@
 In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 [see also @panisWhatShapingRT2016].
+In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature [@panisWhatShapingRT2016].
 
@@
 For many psychological questions, the estimation of such "temporal states" information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
-Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for a temporal science of behavior [@ABNEY2025].
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible response to the recent call for a temporal science of behavior [@ABNEY2025].
 
@@
 
-Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
+Our ultimate aim in this paper is twofold.
+First, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data.
+Second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
 
 Even though EHA is a widely used statistical tool and there already exist many excellent reviews [@singerAppliedLongitudinalData2003; @allisonDiscreteTimeMethodsAnalysis1982; @blossfeldTechniquesEventHistory2002; @hosmerAppliedSurvivalAnalysis2011; @teachmanAnalyzingSocialProcesses1983; @box-steffensmeierEventHistoryModeling2004; @millsIntroducingSurvivalEvent2011] and tutorials [@allisonSurvivalAnalysisUsing2010; @landesIntroductionEventHistory2020; @elmerModelingCategoricalTimetoevent2023; @lougheedMultilevelSurvivalAnalysis2019; @stoolmillerIntroductionUsingMultivariate2015; @stoolmillerModelingHeterogeneitySocial2006], we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+
 Set within this context, our overall aim is to introduce a set of tutorials, which explain **how** to do such analyses in the context of experimental psychology, rather than repeat in any detail **why** you may do them.
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics [@singerAppliedLongitudinalData2003; @allisonSurvivalAnalysisUsing2010; @gelmanRegressionOtherStories2020; @winterStatisticsLinguistsIntroduction2019; @millsIntroducingSurvivalEvent2011].
-We supply a few regression equations in section E of the Supplemental Material.
+We also supply relevant regression equations in section E of the Supplemental Material.
 
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable. 
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well [@bloxomEstimatingResponseTime1984; @luceResponseTimesTheir1991; @vanzandtHowFitResponse2000].
+However, continuous forms of EHA require much more data to reliably estimate the continuous-time hazard (rate) function [@bloxomEstimatingResponseTime1984; @luceResponseTimesTheir1991; @vanzandtHowFitResponse2000].
 Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per participant [@panisAnalyzingResponseTimes2020].
@@
 For example, a response in one trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
-One then calculates the sample-based estimate of the discrete-time hazard function of event occurrence for each experimental condition (Figure 1B top).
+One then calculates the sample-based estimate of the discrete-time hazard function of event occurrence for each experimental condition (Figure 1B upper panel).
 The discrete-time hazard function gives you, for each time bin, the conditional probability that the event occurs (sometime) in
@@
 In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
-In those cases, one can also calculate the conditional accuracy function (Figure 1B bottom).
+In those cases, one can also calculate the conditional accuracy function (Figure 1B lower panel).
 The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
@@
 For example, if the censoring time was 1 second, then some trials result in observed event times (those with a RT below 1 second), while the other trials result in response times that are right-censored at 1 second.
-The fact that EHA can deal with right-censoring, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology (ANOVA, linear regression, delta plots).
+The fact that EHA can deal with right-censoring, therefore, presents a analytical strength of the approach compared to many common approaches in experimental psychology (e.g., ANOVA, linear regression, delta plots).
 
@@
 Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
-Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
-Interestingly, the time bins do not need to have the same width.
+Overly small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
+Of note, however, is that time bins do not need to have the same width.
 For example, @panisHowCanWe2020 used larger bins towards the end of the observation period, as fewer events occurred there.
@@
 From a practical perspective, therefore, some choices need to be made to balance the amount of data that is being collected per participant, condition and across the varying timescales.
-As one example, if there are several times cales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
+As one example, if there are several timescales of relevance, then it might be prudent for interpretational purposes to limit the number of experimental predictor variables (conditions).
 This is of course where planning and data simulation efforts would be important to provide a guide to experimental design choices (see Tutorial 4 and section 2.3).
@@
 Compared to the orthodox method -- comparing means between conditions -- the most important limitation of multilevel hazard and conditional accuracy modeling is that it might take a long time to estimate the parameters using Bayesian methods or the model might have to be simplified significantly to use frequentist methods.
+Relatedly, as these models can be quite complex in terms of the number of possible parameters, more thought is required at the model specification and model building stages.
 

### 2025-07-07 manuscript/notes/notes.Rmd
added notes

@@
 -> I agree.
+
+# July 7th 2025 #
+
+## repsponse document ##
+
+-page 2. we say:
+>It also visualizes recurrent events and includes the Lougheed and Stoolmiller tutorials.
+
+Do we mean "includes reference to the Lougheed and Stoolmiller tutorials"? If not, it is a little odd to read.
+
+-page 6. we say this, which is a direct quote from the manuscript:
+
+>“... because multilevel generalized linear regression models often do not converge with complex
+random-coefficient structures, we do not discuss them here.”
+
+I also read this in the main text for context. 
+It doesn't make sense as it is written in the main text, so let's move it to supps.
+Instead, in the main text, let's just say this.
+
+> Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
+
+If we have anything more elaborate to say about model fitting etc., then let's say it in the tutorial/s.
+
+And in the response to the reviewer, let's NOT say "We have no idea why the models did not converge ..."
+Instead, let's say it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure. (We can also say this in the tutorial or supps, as necessary).
+And then can we cite the lme4 tutorial paper or something else to provide a resource for folks to use?
+
+The idea would be that (1) all frequentist stuff is in supps and (2) we provide a reference for others to follow.
+
+-page 16 of the review. Page 28 of the main manuscript.
+
+>To make causal inferences...
+
+Was this language in the original submission? e.g., did we say causal in the original?
+
+If we want to keep this language, can we cite the following textbook (see below)? 
+It has a really nice section on the link between experimental designs and causal inferences. 
+We don't need to say anything else, just insert the reference.
+
+https://experimentology.io/
+
+Frank, M.
+C., Braginsky, M., Cachia, J., Coles, N.
+A., Hardwicke, T.
+E., Hawkins, R.
+D., Mathur, M.
+B., & Williams, R.
+2025.
+Experimentology: An Open Science Approach to Experimental Psychology Methods.
+Stanford University.
+https://doi.org/10.25936/3JP6-5M50.
+(Also published by MIT Press, ISBN 978-0-262-55256-1)
+
+BibTex:
+
+@book{experimentology2025,
+  author    = {Frank, Michael C. and Braginsky, Mika and Cachia, Julie and Coles, Nicholas A. and Hardwicke, Tom E. and Hawkins, Robert D. and Mathur, Maya B. and Williams, Rondeline},
+  title     = {{Experimentology: An Open Science Approach to Experimental Psychology Methods}},
+  year      = {2025},
+  publisher = {Stanford University},
+  doi       = {10.25936/3JP6-5M50},
+  note      = {Also published by MIT Press, ISBN 978-0-262-55256-1}
+}
+
+## manuscript ##
+
+-page 11. The below sentence:
+
+>First, one can use a response deadline in each trial because EHA deals with right-censored observations.
+
+This still needs explaining/contextualising with an extra sentence or two. It does not stand alone as it is currently written.

### 2025-07-07 manuscript/manuscript.tex
create pdf of ms after review by Rich

@@
 Such comparisons between means have been the workhorse of experimental psychology over the last century, and have had a substantial impact on theory development as well as our understanding of the structure of cognition and brain function.
-Indeed, the view that mean values are truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger, Zee, Rossignac-Milon, \& Hassin, 2019).
+Indeed, the view that mean values represent truth and variations around the mean are error is deeply ingrained in experimental psychology (Bolger, Zee, Rossignac-Milon, \& Hassin, 2019).
 However, differences in mean RT conceal important pieces of information, such as when an experimental effect starts, how it evolves with increasing waiting time, and whether its onset is time-locked to other events (Panis, 2020; Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022; Panis, Torfs, Gillebert, Wagemans, \& Humphreys, 2017; Panis \& Wagemans, 2009; Wolkersdorfer, Panis, \& Schmidt, 2020).
@@
 In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
-In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1 (see also Panis \& Schmidt, 2016).
+In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature (Panis \& Schmidt, 2016).
 
@@
 For many psychological questions, the estimation of such ``temporal states'' information can be theoretically meaningful by leading to more fine-grained understanding of psychological processes.
-Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible answer to the recent call for a temporal science of behavior (Abney, Fausey, Suarez-Rivera, \& Tamis-LeMonda, 2025).
+Because EHA adds a relatively under-used but ever-present dimension -- the passage of time -- to the theory building toolkit, it provides one possible response to the recent call for a temporal science of behavior (Abney, Fausey, Suarez-Rivera, \& Tamis-LeMonda, 2025).
 
@@
 
-Our ultimate aim in this paper is twofold: first, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
+Our ultimate aim in this paper is twofold.
+First, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data.
+Second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
 
 Even though EHA is a widely used statistical tool and there already exist many excellent reviews (Allison, 1982; Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Mills, 2011; Singer \& Willett, 2003; Teachman, 1983) and tutorials (Allison, 2010; Elmer, Van Duijn, Ram, \& Bringmann, 2023; Landes, Engelhardt, \& Pelletier, 2020; Lougheed, Benson, Cole, \& Ram, 2019; Stoolmiller, 2015; Stoolmiller \& Snyder, 2006), we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which provide worked examples of the key data processing and Bayesian multilevel regression modelling steps.
+
 Set within this context, our overall aim is to introduce a set of tutorials, which explain \textbf{how} to do such analyses in the context of experimental psychology, rather than repeat in any detail \textbf{why} you may do them.
@@
 For those who want more detailed treatment of EHA and/or regression equations, we refer the reader to several excellent textbooks on these topics (Allison, 2010; Gelman, Hill, \& Vehtari, 2020; Mills, 2011; Singer \& Willett, 2003; Winter, 2019).
-We supply a few regression equations in section E of the Supplemental Material.
+We also supply relevant regression equations in section E of the Supplemental Material.
 
@@
 This choice may seem counter-intuitive, given that RT is typically treated as a continuous variable.
-However, continuous forms of EHA require much more data to estimate the continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991; Van Zandt, 2000).
+However, continuous forms of EHA require much more data to reliably estimate the continuous-time hazard (rate) function (Bloxom, 1984; Luce, 1991; Van Zandt, 2000).
 Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per participant (Panis, Schmidt, et al., 2020).
@@
 In the context of experimental psychology, it is often (but not always), the case that responses can be classified as correct or incorrect.
-In those cases, one can also calculate the conditional accuracy function (Figure 1B bottom).
+In those cases, one can also calculate the conditional accuracy function (Figure 1B lower panel).
 The conditional accuracy function gives you for each time bin the conditional probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
@@
 Based on our experience, a bin width of 50 ms is a good starting value when the number of repeated measures is 100 or less.
-Too small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
-Interestingly, the time bins do not need to have the same width.
+Overly small bin widths will result in erratic hazard functions as many bins will have no events, and thus hazard estimates of zero.
+Of note, however, is that time bins do not need to have the same width.
 For example, Panis (2020) used larger bins towards the end of the observation period, as fewer events occurred there.
@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 

### 2025-07-08 manuscript/manuscript.Rmd
update ms.

@@
 More specifically, we consider three implications that researchers will need to consider when using discrete-time EHA.
-First, one can use a response deadline in each trial because EHA deals with right-censored observations.
+First, because EHA deals with right-censored observations, one can use a fixed response deadline in each trial.
+This will increase design efficiency as one does not need to wait for very long RTs that would be trimmed anyway.
 
@@
 
-Before we move to our Tutorial on planning experiments, we also provide code to fit hazard and conditional accuracy models in the frequentist framework (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
-However, because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here.
+Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
 

### 2025-07-08 manuscript/manuscript.tex
update ms.

@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021, 2021), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{futures} (Bengtsson, 2021b), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021a), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
 
@@
 
-Before we move to our Tutorial on planning experiments, we also provide code to fit hazard and conditional accuracy models in the frequentist framework (see Tutorial\_3a.Rmd and Tutorial\_3b.Rmd).
-However, because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here.
+Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial\_3a.Rmd and Tutorial\_3b.Rmd).
 
@@
 \bibitem[\citeproctext]{ref-R-RJ-2021-048}
-Bengtsson, H. (2021). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+Bengtsson, H. (2021a). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+
+\bibitem[\citeproctext]{ref-R-futures}
+Bengtsson, H.
+(2021b).
+{futures}: A unifying framework for parallel and distributed processing in r using futures.
+\emph{The R Journal}, \emph{13}(2), 208--227.
+\url{https://doi.org/10.32614/RJ-2021-048}
 

### 2025-07-08 manuscript/reviews/ResponseReview1.Rmd
update Notes and ResponseToReviewers

@@
 This section makes it clear that our tutorials pertain only to estimating single-event occurrences.
-It also visualizes recurrent events and includes the Lougheed and Stoolmiller tutorials.
+It also visualizes recurrent events and includes references to the Lougheed and Stoolmiller tutorials.
 
@@
 Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 37 of the revised ms.
-We have no specific idea why the models did not converge, so we simply mention on page 37:
-"... because multilevel generalized linear regression models often do not converge with complex random-coefficient structures, we do not discuss them here."
+
+We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure.
 

### 2025-07-08 manuscript/reviews/ResponseReview1.tex
update date in Response to Reviewers
update Notes and ResponseToReviewers

@@
 \author{Panis \& Ramsey}
-\date{24.06.2025}
+\date{07.07.2025}
 
@@
 pertain only to estimating single-event occurrences. It also visualizes
-recurrent events and includes the Lougheed and Stoolmiller tutorials.
+recurrent events and includes references to the Lougheed and Stoolmiller
+tutorials.
 
@@
 Response: We removed the frequentist examples from the main text and
-direct the reader to the Tutorials on page 37 of the revised ms. We have
-no specific idea why the models did not converge, so we simply mention
-on page 37: ``\ldots{} because multilevel generalized linear regression
-models often do not converge with complex random-coefficient structures,
-we do not discuss them here.''
+direct the reader to the Tutorials on page 37 of the revised ms.
+
+We mention in Tutorials 3a and 3b that it is common for models to fail
+to converge in lme4 when they have a reasonably complex random effects
+structure.
 

### 2025-07-08 Tutorial_1a.Rmd
update tutorials with bibliography

@@
 date: "`r Sys.Date()`"
+bibliography: "refs_tutorials.bib"
 output:
@@
 
-In Tutorial 1a we wrangle raw data from Experiment 1 of Panis and Schmidt (2016), and illustrate how to calculate descriptive statistics (functions of discrete time).
+In Tutorial 1a we wrangle raw data from Experiment 1 of @panisWhatShapingRT2016, and illustrate how to calculate descriptive statistics (functions of discrete time).
 We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.
@@
 
-Read in the raw data file from Panis and Schmidt (2016) - can be skipped.
+Read in the raw data file from @panisWhatShapingRT2016 - can be skipped. This file can be found on ResearchGate.
 
@@
 
-# 5. Optional: set up life tables and make plots for data aggregated across participants
+# 5. Optional: set up life tables and make one plot for data aggregated across participants
 
@@
   #            first3 bins in the plots).
-  # aggregated_data: For a dataset with data aggregated across individuals, set to TRUE. 
+  # aggregated_data: When plotting data aggregated across individuals, set to TRUE. 
   #            This only controls the title of the plots.
@@
 ```
+
+# 6. References.

### 2025-07-08 Tutorial_1b.Rmd
update tutorials with bibliography

@@
 
-Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).
+Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of @panisWhatShapingRT2016 when there are TWO independent variables (here: mask type and prime type).
 
@@
 
-Read in the raw data from Panis and Schmidt (2016).
+Read in the raw data from @panisWhatShapingRT2016. This file can be found on ResearchGate.
 
@@
 ```
+
+# 5. References.

### 2025-07-08 Tutorial_2a.Rmd
update tutorials with bibliography

--- 2025-07-06 Tutorial_2a.Rmd
@@
 date: "`r Sys.Date()`"
+bibliography: "refs_tutorials.bib"
 output: 
@@
 Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data (section 3).
-The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent).
+The data are taken from the first experiment of @panisWhatShapingRT2016, but using only the no-mask trials (with factor prime = blank, congruent, or incongruent).
 We performed some prior predictive checks to determine the prior distributions we will employ (section 8).
@@
 
-The predictive accuracy of a set of models can be compared using WAIC and LOO.
+The predictive accuracy of a set of models can be compared using WAIC and LOO [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].
 
@@
 ```{r model-weights}
-model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "loo") %>% round(digits = 3)
-
-model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "waic") %>% round(digits = 3)
+model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "loo") %>% round(digits = 3) %>% format(nsmall=2)
+
+model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "waic") %>% round(digits = 3) %>% format(nsmall=2)
 ```
@@
 
-To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
+To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
 Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.
@@
 
-Following Gelman et al.
-(2020), we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
+Following @gelmanBayesianWorkflow2020, we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs.
 First, sample the prior distributions using sample_prior="only".
@@
 After these prior predictive checks, we can increase the standard deviation for class "sd" to 1 to cover many possible inter-individual differences, and fit model M0c to get the posterior distributions (see section 3).
+
+# 9. References.
+
+Heiss, Andrew (2021). “A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel Bayesian Models.” November 10, 2021. https://doi.org/10.59350/wbn93-edb02.

### 2025-07-08 Tutorial_2b.Rmd
update tutorials 2b, 3a, 3b
update tutorials with bibliography

@@
 author: "sven panis"
-date: "2024-10-18"
+date: "`r Sys.Date()`"
+bibliography: "refs_tutorials.bib"
 output: 
@@
 
-In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data.
+In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data, using the input file created in Tutorial 1a.
 
@@
 
+See Heiss (2021).
+
 ### Grand mean AMEs
@@
 ```
+
+# 5. References.
+
+Heiss, Andrew (2021). “A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel Bayesian Models.” November 10, 2021. https://doi.org/10.59350/wbn93-edb02.

### 2025-07-08 Tutorial_3a.Rmd
update Tutorials 3a and 3b with 1 sentence
update tutorials with bibliography

--- 2024-12-16 Tutorial_3a.Rmd
@@
 
-In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).
+In Tutorial 3a we build a Frequentist hazard regression model for the first experiment of @panisWhatShapingRT2016 using only the no-mask trials (prime = blank, congruent, or incongruent).
+Note that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure [@eager2017mixedeffectsmodelsterrible].
 
@@
 ```
+
+# 5. References

### 2025-07-08 Tutorial_3b.Rmd
update Tutorials 3a and 3b with 1 sentence
update tutorials with bibliography

@@
 author: "sven panis"
-date: "2024-10-18"
+date: "`r Sys.Date()`"
+bibliography: "refs_tutorials.bib"
 output:
@@
 
-In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).
+In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of @panisWhatShapingRT2016 using only the no-mask trials (prime = blank, congruent, or incongruent).
+Note that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure [@eager2017mixedeffectsmodelsterrible].
 

### 2025-07-08 manuscript/manuscript.Rmd
add reference to Frank et al., 2025

@@
 
-To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make causal inferences from the parameter estimates in model M1i [@experimentology2025], we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80% and 95% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.

### 2025-07-08 manuscript/manuscript.tex
add reference to Frank et al., 2025

@@
 
-To make causal inferences from the parameter estimates in model M1i, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make causal inferences from the parameter estimates in model M1i (Frank et al., 2025), we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
@@
 
+\bibitem[\citeproctext]{ref-experimentology2025}
+Frank, M.
+C., Braginsky, M., Cachia, J., Coles, N.
+A., Hardwicke, T.
+E., Hawkins, R.
+D., \ldots{} Williams, R.
+(2025).
+\emph{{Experimentology: An Open Science Approach to Experimental Psychology Methods}}.
+Stanford University.
+\url{https://doi.org/10.25936/3JP6-5M50}
+
 \bibitem[\citeproctext]{ref-R-cmdstanr}

### 2025-07-08 manuscript/Suppl_material/supplementary.Rmd
adjust Suppl. Figure 1

@@
 
-In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 500 ms long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
+In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation or follow-up period* is 500 ms long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
 
@@
 (D) Two-button bistable perception data requires a recurrent event / multistate analysis.
+Black bars indicate stimlus presentation.
 
@@
 
-For example, in Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B.
+In Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B.
 For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which a female or male face is presented, and finally a blank screen until an overt response is detected or 800 ms have passed.
@@
 To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset.
-Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these "negative" RTs in the analysis (lower x-axis labels).
+Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these prime-triggered "negative" RTs in the analysis (lower x-axis labels).
 In general, all RTs must be positive for EHA to work.
@@
 
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT > t-1) = P(RT $\geq$ t) tells you on how many percent of the trials the estimate h(t) = P(RT = t| RT $\geq$ t) is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct | RT = t) is based.
+
 For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
@@
 
-The survivor function provides a context for the hazard function, as S(t-1) = P(RT > t-1) = P(RT $\geq$ t) tells you on how many percent of the trials the estimate h(t) = P(RT = t| RT $\geq$ t) is based.
-The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct | RT = t) is based.
-
 While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y]: x < RT $\leq$ y.
-If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial.
 Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t).
@@
 
-Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
+Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the observation period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline).
 This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth.
 In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times.
-As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean.
-This orthodox procedure leads to underestimation of the true mean.
-By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.
+Ignoring trials without a response leads to underestimation of the true mean.
+By introducing a fixed censoring time for all trials, trials with long RTs are not discarded but contribute to the risk set of each bin.
 

### 2025-07-08 manuscript/Suppl_material/supplementary.tex
adjust Suppl. Figure 1

@@
 
-In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 500 ms long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
+In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation or follow-up period} is 500 ms long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.
 
@@
 (C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial.
-(D) Two-button bistable perception data requires a recurrent event / multistate analysis.}\label{fig:plot-RT-tasks}
+(D) Two-button bistable perception data requires a recurrent event / multistate analysis.
+Black bars indicate stimlus presentation.}\label{fig:plot-RT-tasks}
 \end{figure}
@@
 To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset.
-Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these ``negative'' RTs in the analysis (lower x-axis labels).
+Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these prime-triggered ``negative'' RTs in the analysis (lower x-axis labels).
 In general, all RTs must be positive for EHA to work.
@@
 
+The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \(\geq\) t) tells you on how many percent of the trials the estimate h(t) = P(RT = t\textbar{} RT \(\geq\) t) is based.
+The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct \textbar{} RT = t) is based.
+
 For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function
@@
 
-The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \(\geq\) t) tells you on how many percent of the trials the estimate h(t) = P(RT = t\textbar{} RT \(\geq\) t) is based.
-The probability mass function provides a context for the conditional accuracy function, as P(t) = P(RT = t) tells you on how many percent of the trials the estimate ca(t) = P(correct \textbar{} RT = t) is based.
-
 While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y.
-If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition.
+If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial.
 Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t).

### 2025-07-08 manuscript/reviews/CreateFiguresReview.Rmd
updated packages and re-organised folders
changes to response to reviewers
adjust Suppl. Figure 1

--- 2025-06-14 manuscript/CreateFiguresReview.Rmd
@@
 
-# 2. Create Figure on experimental paradigms
+# 2. Create Supplementary Figure 1 on experimental paradigms
 
@@
 
-create categorization plot
+create bistable perception plot
 
@@
   geom_vline(xintercept=max_RT,linetype = 3) +
-#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
+  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
   geom_segment(data = lines_a,
@@
   geom_vline(xintercept=0) +
-#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
+  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
   geom_segment(data = lines_b,
@@
 
-create categorization plot
+create categorization + SOA plot
 
@@
 ```{r}
-ggsave("REVIEW1/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
+ggsave("manuscript/Suppl_material/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
 ```

### 2025-07-08 sims/sims.Rmd
adjust Figure 1

--- 2025-06-16 sims/sims.Rmd
@@
 
-This file simulates the data used in Figures 1, 2, and 3 of the manuscript and then creates some plots.
+This file simulates the data used in old Figures 1, 2, and 3 of the manuscript and then creates some plots.
+
+Figure 1 of revision 1 can be found in section 13.
 
@@
 p13.1 <-ggplot(lifetable_fig1, aes(x = bin, y = hazard,
-                         colour = condition)) +
+                         colour = condition, linetype=condition)) +
   geom_line(linewidth = 1) +
@@
 p13.2 <-ggplot(ca, aes(x = bin, y = ca,
-                         colour = condition)) +
+                         colour = condition, linetype=condition)) +
   geom_line(linewidth = 1) +

### 2025-07-08 Tutorial_1a.Rmd
update Figure 2 (differently shaped points)

@@
 
-To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (a cutoff time, or timeout; the response deadline).
-The goal is to find the optimal bin width and censoring time (both in ms) to portray the effects of interest.
+To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (a cutoff time, or timeout; the response deadline if one is used).
+The goal is to find a good bin width and censoring time (both in ms) to portray the effects of interest.
 Note that all response times must be positive.

### 2025-07-08 custom_functions.R

@@ -132,30 +132,30 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
   
   # plot the hazard functions
   p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
-    geom_line(aes(y=hazard)) +
-    geom_point(aes(y=hazard), 
-               size=1) + 
+    geom_line(aes(y=hazard), linewidth=1) +
+    geom_point(aes(y=hazard,shape=condition), 
+               size=2.5) + 
     geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), 
-                   show.legend = F) +
-    scale_color_brewer(palette = "Dark2") +
+                   show.legend = F, linewidth = 0.6) +
+    scale_color_brewer(palette = "Dark2", name="Condition") +
+  scale_shape_discrete(name ="Condition", solid=T) + 
     scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,haz_yaxis)) +
     labs(x = "bin endpoints (ms)", 
-         y = "h(t)",
-         color = "Condition") +
+         y = "h(t)") +
     theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
   
   # plot the survivor functions
   p2 <-df %>%
     ggplot(aes(x=period, color=condition, group=condition)) +
-    geom_line(aes(y=survival)) +
-    geom_point(aes(y=survival), 
-               size=1) +
+    geom_line(aes(y=survival), linewidth=1) +
+    geom_point(aes(y=survival,shape=condition), 
+               size=2.5) +
     geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), 
-                   show.legend = F) +
+                   show.legend = F, linewidth = 0.6) +
     
     # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). 
     # Make sure you apply the same levels and labels for the factor condition as in Tutorial 1a. 
@@ -166,54 +166,56 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                                labels = c("blank","congruent","incongruent"))),
[…]

### 2025-07-08 Tutorial_2a.Rmd
update Figures 4 and 5 (add linetypes)

@@
                           fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",.width=c(.8,.95),alpha = 0.5) +
+    stat_lineribbon(aes(linetype=prime),
+                    point_interval="median_qi",
+                    .width=c(.8,.95),
+                    alpha = 0.5) +
     scale_fill_brewer(palette = "Dark2") +
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), 
+                     breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
-    scale_y_continuous(limits=c(0,.7)) +
+    scale_y_continuous(limits=c(0,.6)) +
     labs(x = "time bin", y = "predicted hazard")
@@
            fill=condition, color=condition)) +
-    stat_lineribbon(point_interval="median_qi",
+    stat_lineribbon(aes(linetype=condition),
+                    point_interval="median_qi",
                     .width=.95,
@@
                        fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",
+    stat_lineribbon(aes(linetype=prime),
+                    point_interval="median_qi",
                     .width=c(.8,.95),
@@
                               fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.4, 
+    stat_lineribbon(aes(linetype=contrast),
+                    alpha = 0.4, 
                     point_interval = "mean_qi",

### 2025-07-08 sims/sims.Rmd
update legend Figure 1

@@
 p13.1 <-ggplot(lifetable_fig1, aes(x = bin, y = hazard,
-                         colour = condition, linetype=condition)) +
-  geom_line(linewidth = 1) +
+                         colour = condition)) +
+  geom_line(aes(linetype=condition),
+            linewidth = 1) +
   geom_point(size=3) +
-  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
+  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), 
+                 show.legend = F) +
   scale_colour_brewer(palette = "Dark2") + 
-  scale_x_continuous(breaks = 1:10, limits = c(1,10),labels = c(1:10)*100) +
+  scale_x_continuous(breaks = 1:10, 
+                     limits = c(1,10),
+                     labels = c(1:10)*100) +
   scale_y_continuous(limits = c(0,1)) +
- # ggtitle("hazard and conditional accuracy functions") +
   xlab("time bin t's endpoint (ms)") +
   ylab("hazard") +
-  theme(axis.text.x = element_text(angle = 90),
-        panel.grid.minor.x = element_blank()) +
+
   annotate("label", x = 1.7, y = .95,
@@
            size = 7, 
-           hjust = 0)
+           hjust = 0) +
+    theme(axis.text.x = element_text(angle = 90),
+        panel.grid.minor.x = element_blank(),
+        legend.key.size = unit(1.3, 'cm'))
+
 p13.1
@@
 p13.2 <-ggplot(ca, aes(x = bin, y = ca,
-                         colour = condition, linetype=condition)) +
-  geom_line(linewidth = 1) +
+                         colour = condition)) +
+  geom_line(aes(linetype=condition),
+            linewidth = 1) +
   geom_point(size=3) +
-  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
+  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), 
+                 show.legend = F) +
   scale_colour_brewer(palette = "Dark2") + 
-  scale_x_continuous(breaks = 1:10, limits=c(1,10),labels = c(1:10)*100) +
+  scale_x_continuous(breaks = 1:10, 
+                     limits=c(1,10),
+                     labels = c(1:10)*100) +
   scale_y_continuous(limits = c(0,1)) +
@@
     ylab("conditional accuracy") +
-   theme(axis.text.x = element_text(angle = 90),
-         panel.grid.minor.x = element_blank()) +
    annotate("label", x = 1.5, y = .25,
@@
            size = 7, 
-           hjust = 0)
+           hjust = 0) +
+    theme(axis.text.x = element_text(angle = 90),
+        panel.grid.minor.x = element_blank(),
+        legend.key.size = unit(1.3, 'cm'))
 

### 2025-07-08 Tutorial_2b.Rmd
update Figures 7 and 8 (different linetypes)

@@
                           fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",
+    stat_lineribbon(aes(linetype=prime),
+                    point_interval="median_qi",
                     alpha = 0.5, 
@@
                        fill=prime, color=prime)) +
-    stat_lineribbon(point_interval="median_qi",
+    stat_lineribbon(aes(linetype=prime),
+                    point_interval="median_qi",
                     alpha = 0.5, 
@@
                        fill=contrast, color=contrast)) +
-    stat_lineribbon(alpha = 0.5, 
+    stat_lineribbon(aes(linetype=contrast),
+                    alpha = 0.5, 
                     point_interval = "mean_qi",

### 2025-07-08 Tutorial_4.Rmd
update Figure 9 (different linetypes)

@@
 p3.1 <- ggplot(data_group, aes(x=timebin, y = outcome, colour = prime)) +
-   geom_line(aes(group = prime)) + geom_point() +
+   geom_line(aes(group = prime, linetype = prime)) + geom_point() +
    geom_jitter(data=data_pid, alpha = 0.5, width = 0.1, height = 0) +
@@
 p5.2 <- ggplot(d1_group, aes(x=bin, y = event, colour = condition)) +
-   geom_line(aes(group = condition)) + 
+   geom_line(aes(group = condition, linetype=condition)) + 
    geom_point() +
@@
 p5.3 <- ggplot(d1_out_group, aes(x = bin, y = outcome, colour = prime)) +
-  geom_line(aes(group = prime)) + 
+  geom_line(aes(group = prime, linetype = prime)) + 
   geom_point() +

### 2025-07-08 manuscript/reviews/ResponseReview1.Rmd
update ResponseToReviewers

@@
 Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 37 of the revised ms.
-
-We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure.
+We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.
 
@@
 
-Response: TODO
+Response: We updated all plots by distinguishing the conditions using different shapes and/or line types, so that they are understandable when viewed in black and white.
 
@@
 We thus removed this term in the revised ms.
-But we kept the term in tutorials 2a and 2b where we base our use of the terms on this cited reference: Heiss (2021) - https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/#average-marginal-effects.
+But we kept the term in tutorials 2a and 2b where we base our use of the terms on this cited reference: Heiss (2021).
 

### 2025-07-08 manuscript/reviews/ResponseReview1.tex
update ResponseToReviewers

@@
 Response: We removed the frequentist examples from the main text and
-direct the reader to the Tutorials on page 37 of the revised ms.
-
-We mention in Tutorials 3a and 3b that it is common for models to fail
-to converge in lme4 when they have a reasonably complex random effects
-structure.
+direct the reader to the Tutorials on page 37 of the revised ms. We
+mention in Tutorials 3a and 3b that it is common for models to fail to
+converge in lme4 when they have a reasonably complex random effects
+structure, and we provide a reference.
 
@@
 
-Response: TODO
+Response: We updated all plots by distinguishing the conditions using
+different shapes and/or line types, so that they are understandable when
+viewed in black and white.
 
@@
 But we kept the term in tutorials 2a and 2b where we base our use of the
-terms on this cited reference: Heiss (2021) -
-\url{https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide/\#average-marginal-effects}.
+terms on this cited reference: Heiss (2021).
 

### 2025-07-09 Tutorial_2a.Rmd
Create 2 new Figures (hazard vs ca results)

@@
 
+## Create Figure for revised ms.
+
+```{r fig-revision, eval=F}
+# combine 3 plots
+p11_theme <- p11 + theme(legend.position = "none")
+
+(p_tidy_fixed / p1 / p11) +
+    plot_annotation(tag_levels = 'A') +
+   # plot_layout(guides = "collect",
+  #              axes = "collect_x") & 
+    theme(legend.position = "bottom")
+
+```
+
+```{r save-fig-revision, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/Revision_results_hazard.png", width = 10, height = 22, dpi = 800)
+```
+
 ## Example conclusions for model M1i.

### 2025-07-09 Tutorial_2b.Rmd
Create 2 new Figures (hazard vs ca results)

@@
 
+## Create Figure for revised ms.
+
+```{r fig-revision, eval=F}
+# combine 3 plots
+p11_theme <- p11 + theme(legend.position = "none")
+
+(p_tidy_fixed_ca / p1 / p11) +
+    plot_annotation(tag_levels = 'A') +
+   # plot_layout(guides = "collect",
+  #              axes = "collect_x") & 
+    theme(legend.position = "bottom")
+
+```
+
+```{r save-fig-revision, eval=F}
+ggsave("Tutorial_2_Bayesian/figures/Revision_results_ca.png", width = 10, height = 22, dpi = 800)
+```
+
 # 5. References.

### 2025-07-09 manuscript/manuscript.Rmd
Create 2 new Figures (hazard vs ca results)

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10131 (body) + 1709 (references) + 3473 ( body supplemental material) + 393 (refs suppl. mat.)"
+wordcount         : "10211 (body) + 1764 (references) + 3473 ( body supplemental material) + 393 (refs suppl. mat.)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]

### 2025-07-09 manuscript/manuscript.tex
Create 2 new Figures (hazard vs ca results)

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10131 (body) + 1709 (references) + 3473 ( body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10211 (body) + 1764 (references) + 3473 ( body supplemental material) + 393 (refs suppl.
 mat.)}

### 2025-07-09 Tutorial_2b.Rmd
change model results and sizes figs 3 and 4

@@
 # save combined plot
-ggsave("Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png", width = 10, height = 12, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png", width = 14, height = 12, dpi = 800)
 ```

### 2025-07-09 manuscript/manuscript.Rmd
change model results and sizes figs 3 and 4

@@
 
-To make causal inferences from the parameter estimates in model M1i [@experimentology2025], we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make causal inferences from the parameter estimates in model M1i [@experimentology2025], we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3A, together with point (median) and interval estimates (80% and 95% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
 
-(ref:descr-plot3-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
+(ref:descr-plot3-caption) Discrete-time hazard modeling results.
+(A) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
+(B) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(C) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level.
 
-```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_postdistr.png")
+```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.width='90%'}
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_postdistr.png")
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_hazard.png")
 ```
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 4A), and for each participant in the data set (Figure 4B).
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 3B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 3C).
+The point estimates and quantile intervals can also be reported in a table (see Tutorial_2a.Rmd for details).
 
-(ref:descr-plot4-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
+<!-- , and for each participant in the data set (Figure 4B).-->
+
+<!-- (ref:descr-plot4-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).-->
 
 ```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot4-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_combined.png")
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_combined.png")
 ```
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 5A) and at the participant level (Figure 5B).
+<!-- As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 3C).-->
 
-(ref:descr-plot5-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).
+<!-- and at the participant level (Figure 5B). -->
+
+<!-- (ref:descr-plot5-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).-->
 
 ```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot5-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ame_combined.png")
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ame_combined.png")
 ```
-
-The point estimates and quantile intervals can also be reported in a table (see Tutorial_2a.Rmd for details).
 
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 5A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 3C.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
@@
 The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
+Future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
+In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
 
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
+<!--If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
 Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
 Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
-In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
+Therefore,-->
 
@@
 
-To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 6, together with point (median) and interval estimates (80% and 95% credible intervals).
+To make inferences from the parameter estimates in model M1i_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 4A, together with point (median) and interval estimates (80% and 95% credible intervals).
 
-(ref:descr-plot6-caption) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca. ca = conditional accuracy.
+(ref:descr-plot6-caption) Conditional accuracy modeling results.
+(A) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
+(B) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(C) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.
+ca = conditional accuracy.
 
 ```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
+knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_ca.png")
 ```
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 7A), and for each participant in the data set (Figure 7B).
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 4B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 4C).
+Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first three time bins, relative to blank primes.
 
-(ref:plot7-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).
+<!--, and for each participant in the data set (Figure 7B).-->
+
+<!-- (ref:plot7-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).-->
 
 ```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot7-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png")
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png")
 ```
 
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 8A) and for each participant (Figure 8B).
+<!-- and for each participant (Figure 8B).-->
 
-(ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).
+<!-- (ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).-->
 
 ```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
-knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png")
+#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png")
 ```
-
-Based on Figure 8A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 8A).
-Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
 
@@
 We then took parameters from this model and used them to create a single dataset with 200 trials per condition for 10 individual participants.
-The raw data and the simulated data are plotted in Figure 9 and show quite close correspondence, which is re-assuring.
+The raw data and the simulated data are plotted in Figure 5 and show quite close correspondence, which is re-assuring.
 But, this is only one dataset.
@@
 
-Summary results from Simulation 1 are shown in Figure 10A.
-Figure 10A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 6A.
+Figure 6A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
@@
 Given that trial count per condition made little difference to power in Simulation 1, we fixed trial count at 200 trials per condition in Simulation 2.
-Summary results from Simulation 2 are shown in Figure 10B.
+Summary results from Simulation 2 are shown in Figure 6B.
 A summary of these power calculations might be as follows (trial count = 200 per condition in all cases):

### 2025-07-09 manuscript/manuscript.tex
change model results and sizes figs 3 and 4

@@
 
-To make causal inferences from the parameter estimates in model M1i (Frank et al., 2025), we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make causal inferences from the parameter estimates in model M1i (Frank et al., 2025), we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 3A, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 A credible interval is a range of values that contains a parameter's true value with a specified probability, given the observed data and model.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_postdistr}
+{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_hazard}
 
@@
 
-\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i.}\label{fig:plot-fixed-effects}
+\caption{Discrete-time hazard modeling results.
+(A) Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i.
+(B) Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin at the population level.}\label{fig:plot-fixed-effects}
 \end{figure}
 
-Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 4A), and for each participant in the data set (Figure 4B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_pred_combined}
-
-}
-
-\caption{Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-pred-grand-effects}
-\end{figure}
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects, both at the population level (Figure 5A) and at the participant level (Figure 5B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ame_combined}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-grand-ame-effects}
-\end{figure}
-
+Because the parameter estimates are on the cloglog-hazard scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted hazard values -- at the population level (Figure 3B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 3C).
 The point estimates and quantile intervals can also be reported in a table (see Tutorial\_2a.Rmd for details).
@@
 
-If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 5A.
+If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 3C.
 The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
@@
 The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
-
-If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
-Note that three participants (1, 2, and 3) show a negative difference for the contrast ``congruent minus incongruent'' in bin (360,400{]} -- subject 2 also in bin (320,360{]}.
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to ``dip'' around that time.
-Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
+Future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.
 In Tutorial\_2a.Rmd we provide a number of model formulae that should get you going.
@@
 
-To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 6, together with point (median) and interval estimates (80\% and 95\% credible intervals).
+To make inferences from the parameter estimates in model M1i\_ca, we first plot the densities of the draws from the posterior distributions of its population-level parameters in Figure 4A, together with point (median) and interval estimates (80\% and 95\% credible intervals).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_postdistr}
+{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
 
@@
 
-\caption{Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca. ca = conditional accuracy.}\label{fig:plot-ca-fixed-effects}
+\caption{Conditional accuracy modeling results.
+(A) Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca.
+(B) Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.
+ca = conditional accuracy.}\label{fig:plot-ca-fixed-effects}
 \end{figure}
 
-Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 7A), and for each participant in the data set (Figure 7B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined}
-
-}
-
-\caption{Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-pred-grand-effects-ca}
-\end{figure}
-
-As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 8A) and for each participant (Figure 8B).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined}
-
-}
-
-\caption{Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).}\label{fig:plot-ca-grand-ame-effects}
-\end{figure}
-
-Based on Figure 8A we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 8A).
-Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first time bins, relative to the estimates in the baseline condition.
+Because the parameter estimates are on the logit-ca scale, we can ease our interpretation by plotting the expected value of the posterior predictive distribution -- the predicted conditional accurcies -- at the population level (Figure 4B).
+As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 4C).
+Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
+Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first three time bins, relative to blank primes.
 
@@
 
-Summary results from Simulation 1 are shown in Figure 10A.
-Figure 10A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
+Summary results from Simulation 1 are shown in Figure 6A.
+Figure 6A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
 In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.

### 2025-07-09 manuscript/manuscript.Rmd
update wordcount before review by Rich

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10211 (body) + 1764 (references) + 3473 ( body supplemental material) + 393 (refs suppl. mat.)"
+wordcount         : "10094 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl. mat.)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]

### 2025-07-09 manuscript/manuscript.tex
update wordcount before review by Rich

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10211 (body) + 1764 (references) + 3473 ( body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10094 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl.
 mat.)}

### 2025-07-09 manuscript/reviews/ResponseReview1.Rmd
Small corrections

@@
 
-Response: We added a section in the Supplemental Material (section A in the revised ms.) which explains the types of time-to-event data typically obtained in RT tasks (detection, discrimination, and bistable perception tasks).
+Response: We explicitly mention this on page 6 of the revised ms.
+Also, we added a section in the Supplemental Material (section A in the revised ms.) which explains the types of time-to-event data typically obtained in RT tasks (detection, discrimination, and bistable perception tasks).
 This section makes it clear that our tutorials pertain only to estimating single-event occurrences.

### 2025-07-09 manuscript/reviews/ResponseReview1.tex
Small corrections

@@
 
-Response: We added a section in the Supplemental Material (section A in
-the revised ms.) which explains the types of time-to-event data
-typically obtained in RT tasks (detection, discrimination, and bistable
-perception tasks). This section makes it clear that our tutorials
-pertain only to estimating single-event occurrences. It also visualizes
-recurrent events and includes references to the Lougheed and Stoolmiller
-tutorials.
+Response: We explicitly mention this on page 6 of the revised ms. Also,
+we added a section in the Supplemental Material (section A in the
+revised ms.) which explains the types of time-to-event data typically
+obtained in RT tasks (detection, discrimination, and bistable perception
+tasks). This section makes it clear that our tutorials pertain only to
+estimating single-event occurrences. It also visualizes recurrent events
+and includes references to the Lougheed and Stoolmiller tutorials.
 

### 2025-07-09 manuscript/manuscript.Rmd
Small corrections

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10094 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl. mat.)"
+wordcount         : "10082 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl. mat.)"
 bibliography      : ["extrareferences2.bib", "r-references.bib"]
@@
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
+Two hundred trials are simulated in each condition.
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
@@
 The first bin is (0,100], the last bin is (900,1000].
-Two hundred trials were simulated in each condition.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 
-To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10 time bins; Figure 1B).
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10; Figure 1B).
 Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular trial (i.e., repeated measure). 
@@
 Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable perception tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition [@barackTwoViewsCognitive2021].
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the RT domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition [@barackTwoViewsCognitive2021].
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of @panisWhatShapingRT2016, which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial_1b.Rmd for those that are interested.
+As we use the same functional programming approach as in Tutorial 1a, we simply refer the reader to Tutorial_1b.Rmd.
+
+<!--present the sample-based functions for each participant as part of Tutorial_1b.Rmd for those that are interested.-->
 
@@
 There are several analytic decisions one has to make when fitting a discrete-time hazard model.
-First, because the first few bins typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant.
+First, because the first few bins often contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant.
 Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section D of the Supplemental Material).
@@
 
-```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='80%'}
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='90%'}
 #knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")

### 2025-07-09 manuscript/manuscript.tex
Small corrections

@@
 The first bin is (0,100{]}, the last bin is (900,1000{]}.
-Two hundred trials were simulated in each condition.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 
-To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10 time bins; Figure 1B).
+To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10; Figure 1B).
 Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial (i.e., repeated measure).
@@
 Third, the approach is generalisable and applicable to many tasks that are commonly used in experimental psychology, such as detection, discrimination and bistable perception tasks, and to a range of common experimental manipulations, such as stimulus-onset-asynchrony (see section A of the Supplemental Material).
-The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (Barack \& Krakauer, 2021).
+The upshot is that one general analytical approach, which holds several potential advantages, is widely applicable to many substantive use-cases in the RT domain of experimental psychology, irrespective of the analyst's current view on the nature of cognition (Barack \& Krakauer, 2021).
 
@@
 To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type.
-As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for each participant as part of Tutorial\_1b.Rmd for those that are interested.
+As we use the same functional programming approach as in Tutorial 1a, we simply refer the reader to Tutorial\_1b.Rmd.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
+{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
 

### 2025-07-09 manuscript/reviews/CreateFiguresReview.Rmd
improve Suppl. Fig 1

@@
   geom_vline(xintercept=max_RT,linetype = 3) +
-  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
+  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.84, ymax=2.9), color="black", alpha=0.5) +
   geom_segment(data = lines_a,
@@
   geom_vline(xintercept=0) +
-  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
+  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.84, ymax=2.9), color="black", alpha=0.5) +
   geom_segment(data = lines_b,

### 2025-07-09 manuscript/notes/notes.Rmd
update Notes and ResponseToReviewers
add reference to Frank et al., 2025
more notes on the revision

@@
 
+-> corrected.
+
 -page 6. we say this, which is a direct quote from the manuscript:
@@
 Instead, let's say it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure. (We can also say this in the tutorial or supps, as necessary).
+
+-> Done.
+
 And then can we cite the lme4 tutorial paper or something else to provide a resource for folks to use?
@@
 The idea would be that (1) all frequentist stuff is in supps and (2) we provide a reference for others to follow.
+
+-> I added a reference.
 
@@
 Was this language in the original submission? e.g., did we say causal in the original?
+
+-> No we did not. I changed this to make a distinction between model prediction (for the model comparison section) and model inference.
 
@@
 
+-> I added this reference to the ms.
+
 ## manuscript ##
@@
 This still needs explaining/contextualising with an extra sentence or two. It does not stand alone as it is currently written.
+
+-> Done.
+
+# July 9th #
+
+## manuscript ##
+
+- figure 1 legend.
+
+write out EHA/SAT in full.
+At this stage, these terms have not been defined, so using abbreviations is meaningless.
+Or just remove the labels completely and say " ...
+versus a distributional analysis".
+
+the brackets are wrong in the Figure 1 legend. (0,100]. This should be [0,100], right? This happens twice.
+
+- page 6.
+first para.
+EHA is not defined in the main text yet, but we use the abbreviation.
+We cannot do this.
+And int he response, don't we say that we are not doing this?
+This becomes very confusing and incoherent.
+We have to define any abbreviations we use at their first instance in the main text.
+
+One option would be to add a line on page 3 para 2, like this:
+
+...compared to the aggregation of data across trials (Figure 1A), a distributional approach known as event history analysis (EHA) offers ...
+
+If not here, we need it somewhere else. This is really basic and needs to done consistently.
+
+- page 13.
+tutorials.
+I would get rid of all of the package references to ease the flow of reading.
+Can they go in supplementary materials?
+Or somewhere else?
+At the moment, it isn't working with them piled into a massive paragraph.
+
+- page 19. delete the sentence below, as we already say that earlier in the manuscript:
+
+>Too small bin widths...
+
+- Figure 3.
+Can you make the figure longer?
+It looks squashed and is hard to see/understand.
+If it takes up the whole page and the legend goes off the page, then that's ok.
+Same goes for Figure 4 also.
+
+- Figures 3 and 4.
+The colours are wrong in panel C.
+This is really confusing.
+We need to set the colours to be consistent.
+e.g., if congruent is orange on A and B, it should be orange in C.
+To do so, in the code for panel C, you need to use scale_fill_manual(values=c("#D95F02FF", "#7570B3FF")).
+ANd then choose the relevant hex codes e.g., https://emilhvitfeldt.github.io/r-color-palettes/discrete/RColorBrewer/Dark2/index.html.
+
+Does this make sense? It is really important. Please check the colour codes are correct.
+
+- page 33. para 2. We need to cite Bates et al., 2015 for the lme4 package.
+
+- in the manuscript.Rmd file, can we delete all text that we are not keeping?
+It seems really untidy/annoying to have loads of text commented out.
+Does this make sense?
+Just delete it and commit the change.
+I already deleted some stuff, but there is more to do maybe?

### 2025-07-09 manuscript/manuscript.Rmd
edits to the revised manuscript

@@
 As a simple illustration, Figure 1 summarises simulated data for one subject that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
-Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach known as event history analysis (EHA) offers the possibility to reveal the time course of psychological states (Figure 1B).
 For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
@@
 
-(ref:descr-fig1-caption) Simulated single-subject data showing mean performance versus distributional (EHA/SAT) analyses.
+(ref:descr-fig1-caption) Simulated single-subject data showing mean performance versus a distributional analysis.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
-The first bin is (0,100], the last bin is (900,1000].
+The first bin is [0,100], the last bin is [900,1000].
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10; Figure 1B).
-Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular trial (i.e., repeated measure). 
+Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person’s response occurs in a particular trial across a repeated measures design. 
 For example, a response in one trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
@@
 Additionally, to further simplify the process for other users, the first two tutorials rely on a set of our own custom functions that make sub-processes easier to automate, such as data wrangling and plotting functions (see section C of the Supplemental Material for a list of the custom functions).
-
-<!-- Our list of tutorials is as follows:
-
-* 1a. Wrangle raw data and calculate descriptive stats for one independent variable
-* 1b. Wrangle raw data and calculate descriptive stats for two independent variables
-* 2a. Bayesian multilevel modeling for h(t)
-* 2b. Bayesian multilevel modeling for ca(t)
-* 3a. Frequentist multilevel modeling for h(t)
-* 3b. Frequentist multilevel modeling for ca(t)
-* 4.  Simulation and power analysis for planning experiments-->
 
@@
 Also, to select a suited bin width for model fitting, one can test and compare various bin widths in the censor function, and select the smallest one that is supported by the data.
-Too small bin widths will result in erratic hazard functions because many bins will have estimates equal to zero.
 
@@
 While a normal distribution with relatively large variance is often used as a weakly informative prior for continuous dependent variables, rows A and B of Supplementary Figure 3 show that specifying such distributions on the logit and cloglog scales actually leads to rather informative distributions on the original probability scale, as most mass is pushed to probabilities of 0 and 1.
+As such, we modify the prior formulation in order to make sure that it remains consistent with a weakly informative approach (see section F of the Supplemental Material).
 
@@
 
-Estimating model M1i took about 124 minutes.
+Estimating model M1i took about 124 minutes using the same MacBook Pro.
 
@@
 
-<!-- , and for each participant in the data set (Figure 4B).-->
-
-<!-- (ref:descr-plot4-caption) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).-->
-
-```{r plot-pred-grand-effects, fig.cap = "(ref:descr-plot4-caption)", out.width='80%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_pred_combined.png")
-```
-
-<!-- As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 3C).-->
-
-<!-- and at the participant level (Figure 5B). -->
-
-<!-- (ref:descr-plot5-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level (A), and for each participant (B).-->
-
-```{r plot-grand-ame-effects, fig.cap = "(ref:descr-plot5-caption)", out.width='80%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ame_combined.png")
-```
-
 #### Example conclusions for M1i
@@
 In Tutorial_2a.Rmd we provide a number of model formulae that should get you going.
-
-<!--If we want to focus more on inter-individual differences, we can study the subject-specific hazard functions in Figure 5B.
-Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
-Therefore,-->
-
-<!-- Future studies could (a) increase the number of participants to estimate the proportion of "dippers" in the subject population, and/or (b) try to explain why this dip occurs.
-For example, @panisWhatShapingRT2016 concluded that active, top-down, task-guided response inhibition effects emerge around 360 ms after the onset of the stimulus following the prime (here: the target stimulus).
-Such a top-down inhibitory effect might exist in our priming data set, because after some time participants will learn that the first stimulus is not the one they have to respond to.
-To prevent a premature overt response to the prime they thus might gradually increase a global response threshold during the remainder of the experiment, which could result in a lower hazard in congruent trials compared to blank trials, for bins after ~360 ms, and towards the end of the experiment.
-This effect might be masked for incongruent primes by the response competition effect.-->
 
@@
 
-<!--, and for each participant in the data set (Figure 7B).-->
-
-<!-- (ref:plot7-caption) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level (A), and for each participant (B).-->
-
-```{r plot-pred-grand-effects-ca, fig.cap = "(ref:plot7-caption)", out.width='80%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png")
-```
-
-<!-- and for each participant (Figure 8B).-->
-
-<!-- (ref:descr-plot8-caption) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level (A), and for each participant (B).-->
-
-```{r plot-ca-grand-ame-effects, fig.cap = "(ref:descr-plot8-caption)", out.width='80%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_ame_combined.png")
-```
-
-<!-- ## 4.5 Tutorial 3a: Fitting Frequentist hazard models -->
-
-<!-- In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a.
-The general process is similar to that in Tutorial 2a, except that there are no priors to set.
--->
-
-<!-- Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-The resulting hazard model is called M1i_f with the appended "_f" denoting a frequentist model.
--->
-
-<!-- In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i_f. -->
-
-<!-- (ref:descr-plot11-caption) Parameter estimates for model M1i from brm() -- means and 95% credible intervals -- and model M1i_f from glmer() -- maximum likelihood estimates and 95% confidence intervals.
--->
-
-<!-- ```{r plot-comparison, fig.cap = "(ref:descr-plot11-caption)", out.width='80%'} -->
-<!-- knitr::include_graphics("../Tutorial_3_Frequentist/comparison.png") -->
-<!-- ``` -->
-
-<!-- Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure.
-However, model M1i_f did not converge and resulted in a singular fit.
-This is of course one of the reasons why Bayesian modeling has become so popular in recent years.
-But the price you pay for being able to fit models with more complex varying effects structures via a Bayesian framework is increased computation time.
-In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.
--->
-
-<!-- ## 4.6 Tutorial 3b: Fitting Frequentist conditional accuracy models -->
-
-<!-- In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a.
-To be concise, we only fit effects from model M1i_ca (see Tutorial 2b) using the function glmer() from the R package lme4.
-Alternatively, one could also use the function glmmPQL() from the R package MASS [@ripleyMASSSupportFunctions2024].
-The resulting conditional accuracy model, which we labelled M1i_ca_f, did not converge and resulted in a singular fit.
-Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.
--->
-
 Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
@@
 The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it [@gelmanBayesianWorkflow2020].
-The basic structure and code follows the examples outlined by Solomon Kurz in his 'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/) as well as these related papers [@debruineUnderstandingMixedEffectsModels2021; @pargentTutorialTailoredSimulationBased2024].
+The basic structure and code follows the examples outlined by Solomon Kurz in his 'power' blog posts (https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/) and Lisa Debruine's R package faux{} (https://debruine.github.io/faux/), as well as these related papers [@debruineUnderstandingMixedEffectsModels2021; @pargentTutorialTailoredSimulationBased2024].
 
@@
 Figure 6A depicts statistical "power" as calculated by the percentage of lower-bound 95% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
-In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
+In other words, we calculate the fraction of simulated datasets that generated an effect of prime that excludes the criterion mark of zero.
 We are aware that "power" is not part of a Bayesian analytical workflow, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
@@
 
-**But**, and this is an important "but", unless there are unavoidable reasons, no matter what planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
+But, and this is an important caveat, unless there are unavoidable reasons, no matter what kind of planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
 Instead, we would run a follow-up experiment that replicates and extends the initial result. 
@@
 EHA thus provides a way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
-Therefore, if your research questions concern **when psychological states occur, and how they are temporally organized**, our EHA tutorials could be useful tools for you to use.
+Therefore, if your research questions concern **when psychological states occur, and how they are temporally organized**, our EHA tutorials could be useful tools to use for basic knowledge development, as well as theory building.
 
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-In any case, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
-
-<!--One could envisage scenarios where the implicit assumption of an effect in mean RTs manifesting across all of the time bins measured would not be supported by EHA.
-Therefore, the conclusion of interest would not apply to all responses, but instead it would be restricted to certain periods of within-trial time.-->
+For instance, EHA might reveal that the conclusion of interest based on averaging across trials does not apply to all responses, but is instead restricted to certain periods of within-trial time.
 

### 2025-07-09 Tutorial_2a.Rmd
change model results and sizes figs 3 and 4
increase size Figures 3 and 4

@@
   scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_color_brewer(palette = "Dark2") +
+  scale_fill_manual(values=c("#D95F02FF", "#7570B3FF")) +
   theme(axis.text.x = element_text(angle=90)) +
@@
 ```{r save-fig-revision, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/Revision_results_hazard.png", width = 10, height = 22, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/Revision_results_hazard.png", width = 14, height = 20, dpi = 800)
 ```

### 2025-07-09 Tutorial_2b.Rmd
Small corrections
increase size Figures 3 and 4

@@
   scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_color_brewer(palette = "Dark2") +
+  scale_fill_manual(values=c("#D95F02FF", "#7570B3FF")) +
   theme(axis.text.x = element_text(angle=90)) +
   labs(x = "time bin", y = "difference in predicted ca") +
-#  ggtitle("Grand average marginal effect") +
   facet_wrap(~contrast)
@@
 ```{r save-fig-revision, eval=F}
-ggsave("Tutorial_2_Bayesian/figures/Revision_results_ca.png", width = 10, height = 22, dpi = 800)
+ggsave("Tutorial_2_Bayesian/figures/Revision_results_ca.png", width = 14, height = 20, dpi = 800)
 ```

### 2025-07-09 manuscript/manuscript.Rmd
update ms, notes, and suppl. mat.

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10082 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl. mat.)"
-bibliography      : ["extrareferences2.bib", "r-references.bib"]
+wordcount         : "10115 (body) + 1764 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
+bibliography      : ["extrareferences.bib", "extrareferences2.bib", "r-references.bib"]
 floatsintext      : yes
@@
 # R references
- r_refs(file = "r-references.bib") #, append=F) # append=F prevents multiple copies of the same citation/ creates a .bib-reference file
- my_r_citation <- cite_r(file = "r-references.bib", footnote = FALSE)
+ r_refs(file = "r-references.bib") # creates a .bib-reference file
+
+ my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) # use in main text to cite R
+
 ```
@@
 As a simple illustration, Figure 1 summarises simulated data for one subject that shows how comparing means between two conditions can conceal the shapes of the underlying RT and accuracy distributions.
-Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach known as event history analysis (EHA) offers the possibility to reveal the time course of psychological states (Figure 1B).
+Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+Here we apply a distributional method known as event history analysis (EHA) extended with speed-accuracy tradeoff (SAT) analysis.
 For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
@@
 The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
-The first bin is [0,100], the last bin is [900,1000].
+The first bin is (0,100], the last bin is (900,1000].
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 
-<!--present the sample-based functions for each participant as part of Tutorial_1b.Rmd for those that are interested.-->
-
 ## 3.3 Tutorial 2a: Fitting Bayesian hazard models to interval-censored RT data
@@
 
-```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.width='90%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_postdistr.png")
+```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.height='87%', out.width='90%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_hazard.png")
@@
 (A) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
-(B) Point (median) and 80/95% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(B) Point (median) and 80/95% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
 (C) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.
-ca = conditional accuracy.
 
-```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.width='90%'}
-#knitr::include_graphics("../Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png")
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.height='87%', out.width='90%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_ca.png")
@@
 
-Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial_3a.Rmd and Tutorial_3b.Rmd).
+Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework in Tutorial_3a.Rmd and Tutorial_3b.Rmd, using the R package lme4() [@R-lme4].
 

### 2025-07-09 manuscript/manuscript.tex
update ms, notes, and suppl. mat.

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10094 (body) + 1764 (references) + 3442 ( body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10115 (body) + 1764 (references) + 3442 (body supplemental material) + 393 (refs suppl.
 mat.)}
@@
 Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach offers the possibility to reveal the time course of psychological states (Figure 1B).
+Here we apply a distributional method known as event history analysis (EHA) extended with speed-accuracy tradeoff (SAT) analysis.
 For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
@@
 
-\caption{Simulated single-subject data showing mean performance versus distributional (EHA/SAT) analyses.
+\caption{Simulated single-subject data showing mean performance versus a distributional analysis.
 (A) The mean RT (top) and overall accuracy (bottom) for two conditions are plotted.
@@
 To apply discrete-time EHA, one divides the within-trial time in discrete, contiguous time bins indexed by t (e.g., t = 1 to 10; Figure 1B).
-Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial (i.e., repeated measure).
+Then let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular trial across a repeated measures design.
 For example, a response in one trial might occur at 546 ms and it would be in time bin 6 (any RTs from 501 ms to 600 ms).
@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024) and the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{futures} (Bengtsson, 2021b), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021a), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023) for all reported analyses.
+We used R (Version 4.5.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{futures} (Bengtsson, 2021b), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021a), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023).},
+
+for all reported analyses.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_hazard}
+{\centering \includegraphics[width=0.9\linewidth,height=0.87\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_hazard}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
+{\centering \includegraphics[width=0.9\linewidth,height=0.87\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
 
@@
 (A) Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca.
-(B) Point (median) and 80/95\% credible interval summaries of the conditional accuracy estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
-(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.
-ca = conditional accuracy.}\label{fig:plot-ca-fixed-effects}
+(B) Point (median) and 80/95\% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
+(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.}\label{fig:plot-ca-fixed-effects}
 \end{figure}
@@
 
-Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework using the R package lme4() (see Tutorial\_3a.Rmd and Tutorial\_3b.Rmd).
+Finally, because many researchers will be more familiar with frequentist statistics, we also provide code to fit hazard and conditional accuracy models in the frequentist framework in Tutorial\_3a.Rmd and Tutorial\_3b.Rmd, using the R package lme4() (Bates et al., 2015).
 
@@
 The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
-The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as these related papers (DeBruine \& Barr, 2021; Pargent, Koch, Kleine, Lermer, \& Gaube, 2024).
+The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}), as well as these related papers (DeBruine \& Barr, 2021; Pargent, Koch, Kleine, Lermer, \& Gaube, 2024).
 
@@
 Figure 6A depicts statistical ``power'' as calculated by the percentage of lower-bound 95\% confidence intervals that exclude zero when the difference between prime condition is calculated (congruent - incongruent).
-In other words, what fraction of the simulated datasets generated an effect of prime that excludes the criterion mark of zero.
+In other words, we calculate the fraction of simulated datasets that generated an effect of prime that excludes the criterion mark of zero.
 We are aware that ``power'' is not part of a Bayesian analytical workflow, but we choose to include it here, as it is familiar to most researchers in experimental psychology.
@@
 
-\textbf{But}, and this is an important ``but'', unless there are unavoidable reasons, no matter what planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
+But, and this is an important caveat, unless there are unavoidable reasons, no matter what kind of planning choices we made based on these data simulations, we would not solely rely on data collected from one single study.
 Instead, we would run a follow-up experiment that replicates and extends the initial result.
@@
 EHA thus provides a way to uncover the microgenesis of cognitive effects, by revealing when cognitive states may start and stop, how states are replaced with others, as well as what they may be tied to or interact with.
-Therefore, if your research questions concern \textbf{when psychological states occur, and how they are temporally organized}, our EHA tutorials could be useful tools for you to use.
+Therefore, if your research questions concern \textbf{when psychological states occur, and how they are temporally organized}, our EHA tutorials could be useful tools to use for basic knowledge development, as well as theory building.
 
@@
 Given that distinctly different inferences can be made from the same data based on whether one computes a mean across trials or a RT distribution of events (Figure 1), it may be important for researchers to supplement comparisons between means with EHA.
-In any case, if you have a lot of right-censored observations in your RT data set, and/or your research question concerns whether and when responses occur, and whether and when experimental manipulations affect the instantaneous risk of response occurrence, then EHA should be your method of choice.
+For instance, EHA might reveal that the conclusion of interest based on averaging across trials does not apply to all responses, but is instead restricted to certain periods of within-trial time.
 

### 2025-07-09 manuscript/notes/notes.Rmd
update ms, notes, and suppl. mat.

@@
 
+-> EHA/SAT has been removed.
+
 the brackets are wrong in the Figure 1 legend. (0,100]. This should be [0,100], right? This happens twice.
+
+-> No this was correct. I changed it again.
 
@@
 
+-> Because previous reviewers saw SAT as a separate analysis than EHA, I am reluctant to say "Indeed, compared to the aggregation of data across trials (Figure 1A), a distributional approach known as event history analysis (EHA)
+offers the possibility to reveal the time course of psychological states (Figure 1B)" as this implies that Figure 1B is an EHA.
+Instead, I added this sentence: "Here we apply a distributional method known as event history analysis (EHA) extended with speed-accuracy tradeoff (SAT) analysis."
+
 - page 13.
@@
 
+-> I put them back in a footnote. I like it this way. Let me know if you want them in an extra section in the Suppl. Material...
+
 - page 19. delete the sentence below, as we already say that earlier in the manuscript:
@@
 >Too small bin widths...
+
+-> I cannot find this sentence...??
 
@@
 Same goes for Figure 4 also.
+
+-> Done. When the figure gets too large the caption goes off the page without appearing on the next page... I have to see how this can be solved.
 
@@
 
+-> I changed the colors.
+
 - page 33. para 2. We need to cite Bates et al., 2015 for the lme4 package.
+
+-> Done.
 
@@
 I already deleted some stuff, but there is more to do maybe?
+
+->  Done.
+
+- Just quickly, one thing I noticed, which seems strange, is that we have 6 reference files (.bib files).
+Some of them are repeats - is this intentional?
+It seems very strange and a recipe for confusion and difficulty reproducing our workflow.
+In some cases, I can see the need - e.g., supps or tutorial refs that are separate.
+But then why do we have things like extrarefernces and extrareferneces2?
+I really think we should clean this up, so that we have one reference file per document.
+If we need a separate r-reference file for the manuscript then that’s fine.
+Can you explain the need or make them more efficient?
+
+-> I explain their needs in README: extrareferences are manually selected references (not in Zotero); extrareferences2.bib is created by exporting a library from Zotero.

### 2025-07-10 manuscript/reviews/ResponseReview1.Rmd
add info on reducing 6 to 2 figures
update ms. and ResponseReview

@@
 author: "Panis & Ramsey"
-date: "07.07.2025"
+date: "`r Sys.Date()`"
 output: 
@@
 
-Response: We introduce the definition of right-censoring on page 10 of the revised ms.
+Response: We introduce the concept of right-censoring on page 11 of the revised ms.
 
@@
 As we also noted in our response to the editor, we have made substantial revisions to the manuscript that include removing repetitive text, and placing a stronger focus on clarity and brevity in the revised ms.
+For example, we describe the modeling results in the paper using only two instead of six figures.
 
@@
 
-Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 37 of the revised ms.
+Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 34 of the revised ms.
 We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.
@@
 
-Response: The old Figure 1 has been replaced.
+Response: The old Figure 1 has been replaced, and we state what error bars are showing in the new Figure 1.
 
@@
 
-Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 33 of the revised ms.
+Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 32 of the revised ms.
 The logit is typically used for logistic regression by experimental psychologists, so we feel that no extra explanation is needed.
@@
 We address the issue of treating time as discrete or continuous on pages 8-9 of the revised ms.
-We removed redundancies, get to the tutorial section quicker (page 12) and included more precise explanations of technical jargon.
+We removed redundancies, get to the tutorial section quicker (page 13) and included more precise explanations of technical jargon.
 
@@
 
-Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in main body of the revised ms.
+Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in the main body of the revised ms.
 
@@
 
-Response: We provide supporting references on pages 9-10 of the revised ms.
+Response: We provide supporting references on page 10 of the revised ms.
 
@@
 However, to save space, and because this is an active area of research, we do not attempt to clarify when it would be possible to include varying intercepts and/or slopes.
+We do mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.
 
@@
 
-Response: We specify "person-trial data" in the revised ms.
+Response: We specify "person-trial data" in the revised ms on page 18.
 
@@
 
-Response: A visual comparison between the cloglog and logit link functions is provided in section C of the Supplemental Material.
+Response: A visual comparison between the cloglog and logit link functions is provided in section D of the Supplemental Material.
 
@@
 
-Response: We mention now that model M0i is a baseline or reference model, to be compared with model M01 (see page 26 of the revised ms.).
+Response: We mention now that model M0i is a baseline or reference model, to be compared with model M1i (see page 26 of the revised ms.).
 We also mention that we only report 2 models in the paper from a larger set of models that are fitted in the file Tutorial_2a.Rmd.
@@
 
-Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 29 of the revised ms.
+Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 28 of the revised ms.
 
@@
 
-Response: We decided to keep tutorial 4, and discuss its usefulness on pages 45-46 of the revised ms.
+Response: We decided to keep tutorial 4, and discuss its usefulness on pages 42-43 of the revised ms.
 
@@
 
-Response: We created a new Figure 1.
+Response: We created a new, simpler Figure 1.
 

### 2025-07-10 manuscript/reviews/ResponseReview1.tex
add info on reducing 6 to 2 figures
update ms. and ResponseReview

@@
 \author{Panis \& Ramsey}
-\date{07.07.2025}
+\date{2025-07-10}
 
@@
 
-Response: We introduce the definition of right-censoring on page 10 of
-the revised ms.
+Response: We introduce the concept of right-censoring on page 11 of the
+revised ms.
 
@@
 revisions to the manuscript that include removing repetitive text, and
-placing a stronger focus on clarity and brevity in the revised ms.
+placing a stronger focus on clarity and brevity in the revised ms. For
+example, we describe the modeling results in the paper using only two
+instead of six figures.
 
@@
 Response: We removed the frequentist examples from the main text and
-direct the reader to the Tutorials on page 37 of the revised ms. We
+direct the reader to the Tutorials on page 34 of the revised ms. We
 mention in Tutorials 3a and 3b that it is common for models to fail to
@@
 
-Response: The old Figure 1 has been replaced.
+Response: The old Figure 1 has been replaced, and we state what error
+bars are showing in the new Figure 1.
 
@@
 that conditional accuracy values between 0 and 1 are equally likely).''
-on page 33 of the revised ms. The logit is typically used for logistic
+on page 32 of the revised ms. The logit is typically used for logistic
 regression by experimental psychologists, so we feel that no extra
@@
 as discrete or continuous on pages 8-9 of the revised ms. We removed
-redundancies, get to the tutorial section quicker (page 12) and included
+redundancies, get to the tutorial section quicker (page 13) and included
 more precise explanations of technical jargon.
@@
 Response: Although SAT is rather standard in cognitive experimental
-psychology, we tried to limit the use of abbreviations in main body of
-the revised ms.
+psychology, we tried to limit the use of abbreviations in the main body
+of the revised ms.
 
@@
 
-Response: We provide supporting references on pages 9-10 of the revised
-ms.
+Response: We provide supporting references on page 10 of the revised ms.
 
@@
 to clarify when it would be possible to include varying intercepts
-and/or slopes.
+and/or slopes. We do mention in Tutorials 3a and 3b that it is common
+for models to fail to converge in lme4 when they have a reasonably
+complex random effects structure, and we provide a reference.
 
@@
 
-Response: We specify ``person-trial data'' in the revised ms.
+Response: We specify ``person-trial data'' in the revised ms on page 18.
 
@@
 Response: A visual comparison between the cloglog and logit link
-functions is provided in section C of the Supplemental Material.
+functions is provided in section D of the Supplemental Material.
 
@@
 Response: We mention now that model M0i is a baseline or reference
-model, to be compared with model M01 (see page 26 of the revised ms.).
+model, to be compared with model M1i (see page 26 of the revised ms.).
 We also mention that we only report 2 models in the paper from a larger
@@
 Response: We added the purpose (``To make causal inferences\ldots{}'')
-and define a credible interval on page 29 of the revised ms.
+and define a credible interval on page 28 of the revised ms.
 
@@
 Response: We decided to keep tutorial 4, and discuss its usefulness on
-pages 45-46 of the revised ms.
+pages 42-43 of the revised ms.
 

### 2025-07-10 manuscript/manuscript.Rmd
update ms. and ResponseReview

@@
 First, we want to convince readers of the many benefits of using EHA when dealing with psychological RT data.
-Second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time speed-accuracy tradeoff (SAT) analysis on timed accuracy data in case of choice RT data (Figure 1B).
+Second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a (single event) discrete-time EHA on RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data in case of choice RT data (Figure 1B).
 
@@
 
-(ref:descr-plot3-caption) Discrete-time hazard modeling results.
+(ref:descr-plot3-caption) Discrete-time hazard modeling results at the population level.
 (A) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i.
-(B) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
-(C) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin at the population level.
+(B) Point (median) and 80/95% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin.
+(C) Point (mean) and 80/95% credible interval summaries of estimated differences in hazard in each time bin.
 
@@
 
-(ref:descr-plot6-caption) Conditional accuracy modeling results.
+(ref:descr-plot6-caption) Conditional accuracy modeling results at the population level.
 (A) Medians and 80/95% credible intervals of the posterior distributions of the population-level parameters of model M1i_ca.
-(B) Point (median) and 80/95% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
-(C) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.
+(B) Point (median) and 80/95% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin.
+(C) Point (mean) and 80/95% credible interval summaries of estimated differences in conditional accuracy in each time bin.
 

### 2025-07-10 manuscript/manuscript.tex
update ms. and ResponseReview

@@
 
-\caption{Discrete-time hazard modeling results.
+\caption{Discrete-time hazard modeling results at the population level.
 (A) Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i.
-(B) Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
-(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin at the population level.}\label{fig:plot-fixed-effects}
+(B) Point (median) and 80/95\% credible interval summaries of the hazard estimates (expected values of the draws from the posterior predictive distributions) in each time bin.
+(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in hazard in each time bin.}\label{fig:plot-fixed-effects}
 \end{figure}
@@
 
-\caption{Conditional accuracy modeling results.
+\caption{Conditional accuracy modeling results at the population level.
 (A) Medians and 80/95\% credible intervals of the posterior distributions of the population-level parameters of model M1i\_ca.
-(B) Point (median) and 80/95\% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin at the population level.
-(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin at the population level.}\label{fig:plot-ca-fixed-effects}
+(B) Point (median) and 80/95\% credible interval summaries of the conditional accuracy (ca) estimates (expected values of the draws from the posterior predictive distributions) in each time bin.
+(C) Point (mean) and 80/95\% credible interval summaries of estimated differences in conditional accuracy in each time bin.}\label{fig:plot-ca-fixed-effects}
 \end{figure}

### 2025-07-10 manuscript/notes/notes.Rmd
update Notes
update Notes.
extra notes/comments

@@
 
--> No this was correct. I changed it again.
+-> No this was correct. I changed it again, because time point zero is not part of bin 1.
 
@@
 
--> I cannot find this sentence...??
+-> Done.
 
@@
 
--> Done. When the figure gets too large the caption goes off the page without appearing on the next page... I have to see how this can be solved.
+-> Done. When I make the figures even larger then the caption goes off the page without appearing on the next page... I have to see how this can be solved.
 
@@
 -> I explain their needs in README: extrareferences are manually selected references (not in Zotero); extrareferences2.bib is created by exporting a library from Zotero.
+
+# July 10th #
+
+- the readme needs a lot of work.
+ Take a look here or elsewhere and update it: https://unlimited.ethz.ch/spaces/RC/pages/194119720/How+to+write+a+Readme+file
+
+This is really important.
+
+Also, some stuff is just wrong/out of date as far as I can tell. e.g., why are Figures 1,2 and 3 in the sims folder?
+
+This is the first contact point that folks have with the project so the layout and formatting needs to be crystal clear.
+
+Take a look at this prior project as an example and try to copy the use of bold and section headings.
+https://github.com/rich-ramsey/knot_dmc
+
+- The reference files still need fixing.
+We can't be having loads of extrareferences files.
+it just makes no sense computationally.
+Just have one file called references.bib and combine the extrareferences files into it.
+Then have a separate file called r-references as you already do.
+It is easy to do as it is just text.
+
+And is there supposed to be a duplicated r-references file in the root folder and the /manuscript/ folder?
+If so, why?
+Can this be cleared up.
+You make no mention of it in the readme.
+So it seems very odd.
+
+- r-references in the footnote is fine.
+
+-  Like I said before, the brackets cannot stay like that in Figure 1.
+It just makes no sense.
+Instead, just use words and no brackets.
+e.g., The first bin is for responses up to 100ms (or whatever) and the last bin is for responses between 901 and 1000.
+Or just do [1,100] and [901,1000], if that's easier.
+
+- Figures 3 and 4.
+Ok, so, if it is difficult to get things on one page, then just make them as big as possible.
+Otherwise, the only other option I can see if splitting the figures into two.
+Parameter estimates in one figure and posterior preds in another.
+However, I kind of like them all in one place at the moment, because it directly connects model parameters with predictions and it is efficient in terms of space.
+WHat do you think?
+Maybe we just accept the limits of visualisation?
+Or make the legend text single space, but I'm not sure how easy that is in papaja(), as I've never tried it.

### 2025-07-10 manuscript/manuscript.tex
update ms. and notes

@@
 The content of the tutorials, in terms of EHA and multilevel regression modelling, is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2020), Heiss (2021), Kurz (2023a), and Kurz (2023b).
-We used R (Version 4.5.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{futures} (Bengtsson, 2021b), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021a), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023).},
+We used R (Version 4.5.1; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.13.0; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.22.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.9.0.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{futures} (Bengtsson, 2021), \emph{ggplot2} (Version 3.5.2; Wickham, 2016), \emph{lme4} (Version 1.1.37; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.4; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.3; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.168; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.3; Aust \& Barth, 2024), \emph{patchwork} (Version 1.3.0; Pedersen, 2024), \emph{purrr} (Version 1.0.4; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.14; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{rstan} (Version 2.32.7; Stan Development Team, 2024), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{StanHeaders} (Version 2.32.10; Stan Development Team, 2020), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.3.0; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.7; Kay, 2024), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019) and \emph{tinylabels} (Version 0.2.5; Barth, 2023).},
 
@@
 
-\bibitem[\citeproctext]{ref-R-RJ-2021-048}
-Bengtsson, H. (2021a). A unifying framework for parallel and distributed processing in r using futures. \emph{The R Journal}, \emph{13}(2), 208--227. \url{https://doi.org/10.32614/RJ-2021-048}
-
 \bibitem[\citeproctext]{ref-R-futures}
 Bengtsson, H.
-(2021b).
+(2021).
 {futures}: A unifying framework for parallel and distributed processing in r using futures.

### 2025-07-10 manuscript/Suppl_material/supplementary.Rmd
change bib files

@@
 shorttitle        : "A tutorial on event history analysis"
-bibliography      : ["extrareferences2.bib"]
+bibliography      : ["refs_suppl_material.bib"]
 floatsintext      : yes

### 2025-07-10 manuscript/manuscript.Rmd
update ms. and notes
change bib files

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10115 (body) + 1764 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
-bibliography      : ["extrareferences.bib", "extrareferences2.bib", "r-references.bib"]
+wordcount         : "10102 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
+bibliography      : ["refs_manuscript.bib", "r-references.bib"]
 floatsintext      : yes
@@
 # R references
- r_refs(file = "r-references.bib") # creates a .bib-reference file
+ #r_refs(file = "r-references.bib") # creates a .bib-reference file
 

### 2025-07-10 manuscript/manuscript.Rmd
update Notes, ms. and README

@@
   - \raggedbottom
+  - \usepackage{setspace}
+  - \captionsetup[figure]{font={stretch=1,scriptsize}}
 ---
@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
-The first bin is (0,100], the last bin is (900,1000].
+The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (t = 1 to 10).
+Each time bin is a half-open interval; for example, the first bin is the interval (0,100], with 0 excluded and 100 included.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 
-```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.height='87%', out.width='90%'}
+```{r plot-fixed-effects, fig.cap = "(ref:descr-plot3-caption)", out.height='95%', out.width='90%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_hazard.png")
@@
 
-```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.height='87%', out.width='90%'}
+```{r plot-ca-fixed-effects, fig.cap = "(ref:descr-plot6-caption)", out.height='95%', out.width='90%'}
 knitr::include_graphics("../Tutorial_2_Bayesian/figures/Revision_results_ca.png")

### 2025-07-10 manuscript/manuscript.tex
update Notes, ms. and README

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10115 (body) + 1764 (references) + 3442 (body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10102 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl.
 mat.)}
@@
 \raggedbottom
+\usepackage{setspace}
+\captionsetup[figure]{font={stretch=1,scriptsize}}
 \usepackage{bookmark}
@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten bins of 100 ms (t = 1 to 10).
-The first bin is (0,100{]}, the last bin is (900,1000{]}.
+The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (t = 1 to 10).
+Each time bin is a half-open interval; for example, the first bin is the interval (0,100{]}, with 0 excluded and 100 included.
 Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.87\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_hazard}
+{\centering \includegraphics[width=0.9\linewidth,height=0.95\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_hazard}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth,height=0.87\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
+{\centering \includegraphics[width=0.9\linewidth,height=0.95\textheight,]{../Tutorial_2_Bayesian/figures/Revision_results_ca}
 

### 2025-07-10 manuscript/notes/notes.Rmd
update ms. and notes
update Notes, ms. and README

@@
 
+-> I copied the use of bold and section headings, and updated the README file.
+
 - The reference files still need fixing.
@@
 
+-> I created one refs_manuscript.bib file, and one refs_suppl_material.bib file.
+
 - r-references in the footnote is fine.
+
+-> ok.
 
@@
 e.g., The first bin is for responses up to 100ms (or whatever) and the last bin is for responses between 901 and 1000.
-Or just do [1,100] and [901,1000], if that's easier.
+Or just do ]1,100] and [901,1000], if that's easier.
+
+-> I use words now, and keep 1 example.
 
@@
 Or make the legend text single space, but I'm not sure how easy that is in papaja(), as I've never tried it.
+
+-> I increased the size of both figures and made the figure (and table) captions smaller using
+header-includes:
+  - \usepackage{setspace}
+  - \captionsetup[figure]{font={stretch=1,scriptsize}}
+
+This also decreased Table sizes, and footnote size. I think this is fine. But I have to check references to pages again in ResponseReview1.pdf...

### 2025-07-11 manuscript/manuscript.Rmd
update ms.
use normalsize in header

@@
   - \usepackage{setspace}
-  - \captionsetup[figure]{font={stretch=1,scriptsize}}
+  - \captionsetup[figure]{font={stretch=1,normalsize}} 
 ---

### 2025-07-11 manuscript/reviews/ResponseReview1.Rmd
update ms, notes, and suppl. material

@@
 
-Response: We briefly explain model weights for loo and waic on page 28 of the revised ms.
+Response: We briefly explain model weights for loo and waic on page 27 of the revised ms.
 A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we print more digits when reporting the model weights.
@@
 
-Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 34 of the revised ms.
+Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 32 of the revised ms.
 We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.
@@
 
-Response: We replaced early with fast on page 22 of the revised ms.
+Response: We replaced early with fast on page 21 of the revised ms.
 
@@
 
-Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 32 of the revised ms.
+Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 30 of the revised ms.
 The logit is typically used for logistic regression by experimental psychologists, so we feel that no extra explanation is needed.
@@
 
-Response: We explain this on page 24 of the revised ms.:
+Response: We explain this on pages 22-23 of the revised ms.:
 "We also use a "keep it maximal" approach by specifying a full varying (or random) effects structure. This means that wherever possible we include varying intercepts and slopes per participant."
@@
 
-Response: We introduce the definition of right-censoring on page 10 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.
+Response: We introduce the definition of right-censoring on page 11 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.
 
@@
 
-Response: In section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events, and we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception tasks).
+Response: We explicitly mention "single event" on page 6 of the revised ms.
+Also, in section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events, and we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception tasks).
 This should make it clear to the reader that we are dealing with single events that are repeatedly measured within the same individual.
@@
 We agree that a model should not be presented if it does not converge when it includes random slopes.
-However, to save space, and because this is an active area of research, we do not attempt to clarify when it would be possible to include varying intercepts and/or slopes.
+However, to save space, and because this is an active area of research, we do not attempt to clarify when it would not be possible to include varying intercepts and/or slopes.
 We do mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.
@@
 
-Response: This connection should be clear now based on section A of the Supplemental Material, and based on the text on pages 7 and 8.
+Response: This connection should be clear now based on section A of the Supplemental Material, and based on the text on pages 8 and 9.
 
@@
 
-Response: We specify "person-trial data" in the revised ms on page 18.
+Response: We specify "person-trial data" in the revised ms on page 17.
 
@@
 Response: We have no more detail to share on this issue.
-On page 19 of the revised ms.
+On page 18 of the revised ms.
 we now write "In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may not be following task instructions (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.".
@@
 Response: We changed this sentence into: "... because the first few bins
-typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant." on page 24 of the revised ms.
+typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant." on page 23 of the revised ms.
 
@@
 
-Response: We added the following information on page 24 of the revised ms.:
+Response: We added the following information on page 23 of the revised ms.:
 "Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material).
@@
 
-Response: We mention now that model M0i is a baseline or reference model, to be compared with model M1i (see page 26 of the revised ms.).
+Response: We mention now that model M0i is a baseline or reference model, to be compared with model M1i (see page 25 of the revised ms.).
 We also mention that we only report 2 models in the paper from a larger set of models that are fitted in the file Tutorial_2a.Rmd.
@@
 
-Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 28 of the revised ms.
+Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 27 of the revised ms.
 
@@
 
-Response: We decided to keep tutorial 4, and discuss its usefulness on pages 42-43 of the revised ms.
+Response: We decided to keep tutorial 4, and discuss its usefulness on pages 40-41 of the revised ms.
 

### 2025-07-11 manuscript/reviews/ResponseReview1.tex
update ms, notes, and suppl. material

@@
 \author{Panis \& Ramsey}
-\date{2025-07-10}
+\date{2025-07-11}
 
@@
 
-Response: We briefly explain model weights for loo and waic on page 28
+Response: We briefly explain model weights for loo and waic on page 27
 of the revised ms. A discussion of WAIC in Kurz (2023a) can be found in
@@
 Response: We removed the frequentist examples from the main text and
-direct the reader to the Tutorials on page 34 of the revised ms. We
+direct the reader to the Tutorials on page 32 of the revised ms. We
 mention in Tutorials 3a and 3b that it is common for models to fail to
@@
 that conditional accuracy values between 0 and 1 are equally likely).''
-on page 32 of the revised ms. The logit is typically used for logistic
+on page 30 of the revised ms. The logit is typically used for logistic
 regression by experimental psychologists, so we feel that no extra
@@
 
-Response: We explain this on page 24 of the revised ms.: ``We also use
-a''keep it maximal'' approach by specifying a full varying (or random)
-effects structure. This means that wherever possible we include varying
-intercepts and slopes per participant.''
+Response: We explain this on pages 22-23 of the revised ms.: ``We also
+use a''keep it maximal'' approach by specifying a full varying (or
+random) effects structure. This means that wherever possible we include
+varying intercepts and slopes per participant.''
 
@@
 
-Response: We introduce the definition of right-censoring on page 10 of
+Response: We introduce the definition of right-censoring on page 11 of
 the revised ms. We discuss what it means to be at risk in section A of
@@
 
-Response: In section A of the Supplemental Material we mention that this
-Tutorial does not cover recurrent events, and we visualize the different
-types of time-to-event data that are obtained in typical RT tasks
-(detection, discrimination, bistable perception tasks). This should make
-it clear to the reader that we are dealing with single events that are
-repeatedly measured within the same individual.
+Response: We explicitly mention ``single event'' on page 6 of the
+revised ms. Also, in section A of the Supplemental Material we mention
+that this Tutorial does not cover recurrent events, and we visualize the
+different types of time-to-event data that are obtained in typical RT
+tasks (detection, discrimination, bistable perception tasks). This
+should make it clear to the reader that we are dealing with single
+events that are repeatedly measured within the same individual.
 
@@
 space, and because this is an active area of research, we do not attempt
-to clarify when it would be possible to include varying intercepts
+to clarify when it would not be possible to include varying intercepts
 and/or slopes. We do mention in Tutorials 3a and 3b that it is common
@@
 Response: This connection should be clear now based on section A of the
-Supplemental Material, and based on the text on pages 7 and 8.
+Supplemental Material, and based on the text on pages 8 and 9.
 
@@
 
-Response: We specify ``person-trial data'' in the revised ms on page 18.
+Response: We specify ``person-trial data'' in the revised ms on page 17.
 
@@
 
-Response: We have no more detail to share on this issue. On page 19 of
+Response: We have no more detail to share on this issue. On page 18 of
 the revised ms. we now write ``In general, it is important to visually
@@
 time window, i.e., a contiguous set of bins for which there is data for
-each participant.'' on page 24 of the revised ms.
+each participant.'' on page 23 of the revised ms.
 
@@
 
-Response: We added the following information on page 24 of the revised
+Response: We added the following information on page 23 of the revised
 ms.: ``Third, one has to choose whether to treat TIME (i.e., the time
@@
 Response: We mention now that model M0i is a baseline or reference
-model, to be compared with model M1i (see page 26 of the revised ms.).
+model, to be compared with model M1i (see page 25 of the revised ms.).
 We also mention that we only report 2 models in the paper from a larger
@@
 Response: We added the purpose (``To make causal inferences\ldots{}'')
-and define a credible interval on page 28 of the revised ms.
+and define a credible interval on page 27 of the revised ms.
 
@@
 Response: We decided to keep tutorial 4, and discuss its usefulness on
-pages 42-43 of the revised ms.
+pages 40-41 of the revised ms.
 

### 2025-07-11 manuscript/Suppl_material/supplementary.Rmd
update ms, notes, and suppl. material

--- 2025-07-08 manuscript/Suppl_material/supplementary.Rmd
@@
   - \raggedbottom
+  - \usepackage{setspace}
+  - \captionsetup[figure]{font={stretch=1,normalsize}} 
 ---

### 2025-07-11 manuscript/notes/notes.Rmd
update ms, notes, and suppl. material

--- 2025-07-10 manuscript/notes/notes.Rmd
@@
 e.g., The first bin is for responses up to 100ms (or whatever) and the last bin is for responses between 901 and 1000.
-Or just do [1,100] and [901,1000], if that's easier.
+Or just do ]1,100] and [901,1000], if that's easier.
+
+-> I use words now, and keep 1 example for illustration.
 
@@
 Or make the legend text single space, but I'm not sure how easy that is in papaja(), as I've never tried it.
+
+-> I increased the size of both figures and made the figure (and table) captions smaller using
+header-includes:
+  - \usepackage{setspace}
+  - \captionsetup[figure]{font={stretch=1,normalsize}}
+
+This also decreased Table sizes, and footnote size. I think this is actually fine.

### 2025-07-11 Tutorial_2b.Rmd
small corrections

@@
 
-As we expect conditional accuracies ranging between (and including) 0 and 1, we use a flat prior on the logit scale for our intercepts (see Supplementary Figure 2 in section E of the Supplemental Material).
+As we expect conditional accuracies ranging between (and including) 0 and 1, we use a flat prior on the logit scale for our intercepts (see Supplementary Figure 3 in section F of the Supplemental Material).
 

### 2025-07-11 manuscript/manuscript.Rmd
small corrections

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10102 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
+wordcount         : "10115 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
 bibliography      : ["refs_manuscript.bib", "r-references.bib"]

### 2025-07-11 manuscript/manuscript.tex
use normalsize in header
small corrections

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10102 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10115 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl.
 mat.)}
@@
 \usepackage{setspace}
-\captionsetup[figure]{font={stretch=1,scriptsize}}
+\captionsetup[figure]{font={stretch=1,normalsize}}
 \usepackage{bookmark}

### 2025-07-13 Tutorial_1a.Rmd
remove bin brackets and readme.html; update Figs 1,2 and Table 3

@@
 
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t) = P(RT = t | RT >= t), survival probability S(t) = P(Rt > t), probability mass P(t) = P(RT = t), possibly the conditional accuracy ca(t) = P(correct | RT = t), and their estimated standard errors (se).
+A life table includes for each time bin indexed by t, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t) = P(RT = t | RT >= t), survival probability S(t) = P(Rt > t), probability mass P(t) = P(RT = t), possibly the conditional accuracy ca(t) = P(correct | RT = t), and their estimated standard errors (se).
 
@@
   mutate(bin = period*binsize,
+         t = period,
          events = event1) %>%
-  select(bin,risk_set,events,hazard,se_haz,
-         survival,se_surv,ca,se_ca)
+  select(bin,t,risk_set,events,hazard,se_haz,
+         survival,se_surv,ca,se_ca,pmass,se_pmass)
 
@@
 # create a row for time point 0
-row0 <- c(bin=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,
-          survival=1,se_surv=0,ca=NA,se_ca=NA)
+row0 <- c(bin=0,t=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,
+          survival=1,se_surv=0,ca=NA,se_ca=NA,pmass=0,se_pmass=0)
 
@@
   arrange(bin)
+
+# change names
+names(lifetable_blank_s6_complete) <- c("bin","index t", 
+                                        "RS","#events", "h(t)", "se[h(t)])", "S(t)", "se[S(t)]", "ca(t)", "se[ca(t)]", "P(t)", "se[P(t)]")
+
 lifetable_blank_s6_complete
@@
 ```{r save-single-lifetable, eval=F}
-write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
+write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_blank_s6.csv")
 ```

### 2025-07-13 manuscript/manuscript.Rmd
remove bin brackets and readme.html; update Figs 1,2 and Table 3

@@
 #r_refs(file = "r-references.bib") # creates a .bib-reference file
-
- my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) # use in main text to cite R
-
+my_r_citation <- cite_r(file = "r-references.bib", 
+                        footnote = TRUE) # use in main text to cite R 
 ```
@@
 Here we apply a distributional method known as event history analysis (EHA) extended with speed-accuracy tradeoff (SAT) analysis.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in the hazard of response occurrence is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), the hazard of response occurrence is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
 In a third state (>500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature [@panisWhatShapingRT2016].
+Note that we will always refer to a time bin by its upper bound.
+For example, time bin "500" in Figure 1B refers to the time interval running from 400 ms to 500 ms, with the lower bound of 400 ms excluded, and the upper bound of 500 ms included.
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature [@panisWhenDoesInhibition2022].
 
@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (t = 1 to 10).
-Each time bin is a half-open interval; for example, the first bin is the interval (0,100], with 0 excluded and 100 included.
-Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
+The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (indexed by t = 1 to 10).
+The hazard and conditional accuracy estimates are plotted at the upper bound of each time bin.
 The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.1.2.
@@
 
-# 2. What is event history analaysis and why is it relevant to research in experimental psychology?
+# 2. What is event history analysis and why is it relevant to research in experimental psychology?
 
@@
 First, we want to calculate descriptive statistics for each condition in each individual using a life table.
-A life table includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
+A life table (see Table 3) includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
 The definitions of these quantities are provided in section B of the Supplemental Material.
@@
 data_final <- data_nested %>% 
-  # ! user input: censoring time, and bin width
+  # ! user input: censoring time, and bin width in milliseconds
   mutate(censored = map(data, censor, 600, 40)) %>%   
@@
   mutate(ptb_data = map(censored, ptb)) %>%          
-  # create life tables without ca(t)
+  # create life tables without conditional accuracies
   mutate(lifetable = map(ptb_data, setup_lt)) %>%   
-  # calculate ca(t)
+  # calculate conditional accuracies
   mutate(condacc = map(censored, calc_ca)) %>%      
-  # create life tables with ca(t)
+  # create life tables with conditional accuracies
   mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    
@@
 
-(ref:life-note-caption) The column named "bin" indicates the endpoint of each time bin (in ms), and includes time point zero.
-For example the first bin is (0,40] with the starting point excluded and the endpoint included.
-At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
-se = standard error.
-ca = conditional accuracy.
-NA = undefined.
+(ref:life-note-caption) The column named "bin" indicates the upper bound of each time bin (in ms), and includes time point zero.
+At time point zero, no events can occur and therefore both the discrete-time hazard h(t=0) and the conditional accuracy ca(t=0) are undefined.
+RS = risk set; se = standard error; NA = undefined.
 
 ```{r life-table, warning=F}
-life_tab <- read_csv("../Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
+life_tab <- read_csv("../Tutorial_1_descriptive_stats/tables/lifetable_blank_s6.csv")
 
@@
   format.args = list(
-    digits = c(0,0,0,2,2,2,2,2,2),
+    digits = c(0,0,0,0,2,2,2,2,2,2,2,2),
     decimal.mark = ".", big.mark = "")
@@
 
-(ref:descr-plot2-caption) Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
+(ref:descr-plot2-caption) Estimated sample-based discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
 Vertical dotted lines indicate the estimated median RTs.
@@
 
-For example, for participant 6, the estimated hazard values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+For example, for participant 6, the estimated hazard values in bin 280 are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 In other words, when the waiting time has increased until *240 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
 
-Furthermore, the estimated conditional accuracy values in bin (240,280] are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+Furthermore, the estimated conditional accuracy values in bin 280 are `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin 280, then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(7) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(7) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(7) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 
 However, when the waiting time has increased until *400 ms* after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(hazard) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(hazard) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(hazard) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440], then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin 440, then the probability that it is correct is estimated to be `r data_final[[7]][[6]] %>% filter(condition=="blank") %>% select(ca) %>% slice(11) %>% pull()`, `r data_final[[7]][[6]] %>% filter(condition=="congruent") %>% select(ca) %>% slice(11) %>% pull()`, and `r data_final[[7]][[6]] %>% filter(condition=="incongruent") %>% select(ca) %>% slice(11) %>% pull()` for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 In tutorial_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We decided to select the analysis time window (200,600] and the cloglog link.
+We decided to select the 200-600 ms analysis time window, and the cloglog link.
 Below, we shortly discuss two of these models.
@@
 ptb_data <- ptb_data %>% 
-# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
+# select analysis time range: 200-600 ms, with 10 bins (time bin ranks 6 to 15)
 filter(period > 5) %>%
@@
 model_M1i <- readRDS("../Tutorial_2_Bayesian/models/model_M1i.rds")
-#model_M1d <- readRDS("../Tutorial_2_Bayesian/models/model_M1d.rds")
 
@@
 model_M1i <- add_criterion(model_M1i, c("loo", "waic"))
-#model_M1d <- add_criterion(model_M1d, c("loo", "waic"))
 ```
@@
 If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 3C.
-The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin (240,280]) (95% CrI = [0.04, 0.25]).
+The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 240 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 280 (95% CrI = [0.04, 0.25]).
 For the other bins, the 95% credible interval contained zero.
-The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin (200,240] (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin (320,360] (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin (360,400] (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin (400,440] (95% CrI = [-0.40, -0.03]).
+The contrast "incongruent minus blank" was estimated to be 0.09 hazard units in bin 240 (95% CrI = [0.01, 0.21]), -0.19 hazard units in bin 360 (95% CrI = [-0.31, -0.06]), -0.27 hazard units in bin 400 (95% CrI = [-0.45, -0.04]), and -0.23 hazard units in bin 440 (95% CrI = [-0.40, -0.03]).
 For the other bins, the 95% credible interval contained zero.
@@
 There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
-In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240].
-In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in time bin 240.
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period 320-440 ms.
 The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
@@
 As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 4C).
-Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240], (240,280], (280,320], and (320,360], relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
+Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins 240, 280, 320, and 360, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first three time bins, relative to blank primes.

### 2025-07-13 manuscript/manuscript.tex
remove bin brackets and readme.html; update Figs 1,2 and Table 3

@@
 Here we apply a distributional method known as event history analysis (EHA) extended with speed-accuracy tradeoff (SAT) analysis.
-For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in hazard is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
-In a second state (400 to 500 ms), hazard is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
+For example, Figure 1B shows a first state (up to 400 ms after target onset) for which the early upswing in the hazard of response occurrence is equal for both conditions, and the emitted responses are always correct in condition 1 and always incorrect in condition 2.
+In a second state (400 to 500 ms), the hazard of response occurrence is higher in condition 1, and conditional accuracies are close to .5 in both conditions.
 In a third state (\textgreater500 ms), the effect disappears in hazard, and all conditional accuracies are equal to 1.
-Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature (Panis \& Schmidt, 2016).
+Note that we will always refer to a time bin by its upper bound.
+For example, time bin ``500'' in Figure 1B refers to the time interval running from 400 ms to 500 ms, with the lower bound of 400 ms excluded, and the upper bound of 500 ms included.
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature (Panis \& Schmidt, 2022).
 
@@
 (B) The discrete-time hazard functions (top) and conditional accuracy functions (bottom) are plotted for the same data.
-The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (t = 1 to 10).
-Each time bin is a half-open interval; for example, the first bin is the interval (0,100{]}, with 0 excluded and 100 included.
-Note that the hazard and conditional accuracy estimates are plotted at the endpoint of each time bin.
+The first second after target stimulus onset (time zero) is divided in ten time bins of 100 ms (indexed by t = 1 to 10).
+The hazard and conditional accuracy estimates are plotted at the upper bound of each time bin.
 The definitions of discrete-time hazard and conditional accuracy are further explained in section 2.1.2.
@@
 \section{2.
-What is event history analaysis and why is it relevant to research in experimental psychology?}\label{what-is-event-history-analaysis-and-why-is-it-relevant-to-research-in-experimental-psychology}
+What is event history analysis and why is it relevant to research in experimental psychology?}\label{what-is-event-history-analysis-and-why-is-it-relevant-to-research-in-experimental-psychology}
 
@@
 \NormalTok{data\_final }\OtherTok{\textless{}{-}}\NormalTok{ data\_nested }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\# ! user input: censoring time, and bin width}
+  \CommentTok{\# ! user input: censoring time, and bin width in milliseconds}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{censored =} \FunctionTok{map}\NormalTok{(data, censor, }\DecValTok{600}\NormalTok{, }\DecValTok{40}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ptb\_data =} \FunctionTok{map}\NormalTok{(censored, ptb)) }\SpecialCharTok{\%\textgreater{}\%}          
-  \CommentTok{\# create life tables without ca(t)}
+  \CommentTok{\# create life tables without conditional accuracies}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable =} \FunctionTok{map}\NormalTok{(ptb\_data, setup\_lt)) }\SpecialCharTok{\%\textgreater{}\%}   
-  \CommentTok{\# calculate ca(t)}
+  \CommentTok{\# calculate conditional accuracies}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{condacc =} \FunctionTok{map}\NormalTok{(censored, calc\_ca)) }\SpecialCharTok{\%\textgreater{}\%}      
-  \CommentTok{\# create life tables with ca(t)}
+  \CommentTok{\# create life tables with conditional accuracies}
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lifetable\_ca =} \FunctionTok{map2}\NormalTok{(lifetable, condacc, join\_lt\_ca)) }\SpecialCharTok{\%\textgreater{}\%}    
@@
 
-\begin{tabular}{lllllllll}
+\begin{tabular}{llllllllllll}
 \toprule
-bin & \multicolumn{1}{c}{risk\_set} & \multicolumn{1}{c}{events} & \multicolumn{1}{c}{hazard} & \multicolumn{1}{c}{se\_haz} & \multicolumn{1}{c}{survival} & \multicolumn{1}{c}{se\_surv} & \multicolumn{1}{c}{ca} & \multicolumn{1}{c}{se\_ca}\\
+bin & \multicolumn{1}{c}{index t} & \multicolumn{1}{c}{RS} & \multicolumn{1}{c}{\#events} & \multicolumn{1}{c}{h(t)} & \multicolumn{1}{c}{se[h(t)])} & \multicolumn{1}{c}{S(t)} & \multicolumn{1}{c}{se[S(t)]} & \multicolumn{1}{c}{ca(t)} & \multicolumn{1}{c}{se[ca(t)]} & \multicolumn{1}{c}{P(t)} & \multicolumn{1}{c}{se[P(t)]}\\
 \midrule
-0 & 220 & NA & NA & NA & 1.00 & 0.00 & NA & NA\\
-40 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-80 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-120 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-160 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-200 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-240 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA\\
-280 & 220 & 7 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17\\
-320 & 213 & 13 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12\\
-360 & 200 & 26 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05\\
-400 & 174 & 40 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00\\
-440 & 134 & 48 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02\\
-480 & 86 & 37 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00\\
-520 & 49 & 32 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00\\
-560 & 17 & 9 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00\\
-600 & 8 & 4 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00\\
+0 & 0 & 220 & NA & NA & NA & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+40 & 1 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+80 & 2 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+120 & 3 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+160 & 4 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+200 & 5 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+240 & 6 & 220 & 0 & 0.00 & 0.00 & 1.00 & 0.00 & NA & NA & 0.00 & 0.00\\
+280 & 7 & 220 & 7 & 0.03 & 0.01 & 0.97 & 0.01 & 0.29 & 0.17 & 0.03 & 0.01\\
+320 & 8 & 213 & 13 & 0.06 & 0.02 & 0.91 & 0.02 & 0.77 & 0.12 & 0.06 & 0.02\\
+360 & 9 & 200 & 26 & 0.13 & 0.02 & 0.79 & 0.03 & 0.92 & 0.05 & 0.12 & 0.02\\
+400 & 10 & 174 & 40 & 0.23 & 0.03 & 0.61 & 0.03 & 1.00 & 0.00 & 0.18 & 0.03\\
+440 & 11 & 134 & 48 & 0.36 & 0.04 & 0.39 & 0.03 & 0.98 & 0.02 & 0.22 & 0.03\\
+480 & 12 & 86 & 37 & 0.43 & 0.05 & 0.22 & 0.03 & 1.00 & 0.00 & 0.17 & 0.03\\
+520 & 13 & 49 & 32 & 0.65 & 0.07 & 0.08 & 0.02 & 1.00 & 0.00 & 0.15 & 0.02\\
+560 & 14 & 17 & 9 & 0.53 & 0.12 & 0.04 & 0.01 & 1.00 & 0.00 & 0.04 & 0.01\\
+600 & 15 & 8 & 4 & 0.50 & 0.18 & 0.02 & 0.01 & 1.00 & 0.00 & 0.02 & 0.01\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} The column named ``bin'' indicates the endpoint of each time bin (in ms), and includes time point zero.
-For example the first bin is (0,40{]} with the starting point excluded and the endpoint included.
-At time point zero, no events can occur and therefore h(t=0) and ca(t=0) are undefined.
-se = standard error.
-ca = conditional accuracy.
-NA = undefined.}
+\normalsize{\textit{Note.} The column named ``bin'' indicates the upper bound of each time bin (in ms), and includes time point zero.
+At time point zero, no events can occur and therefore both the discrete-time hazard h(t=0) and the conditional accuracy ca(t=0) are undefined.
+RS = risk set; se = standard error; NA = undefined.}
 \end{tablenotes}
@@
 
-\caption{Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
+\caption{Estimated sample-based discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6.
 Vertical dotted lines indicate the estimated median RTs.
@@
 
-For example, for participant 6, the estimated hazard values in bin (240,280{]} are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
+For example, for participant 6, the estimated hazard values in bin 280 are 0.03, 0.17, and 0.20 for the blank, congruent, and incongruent prime conditions, respectively.
 In other words, when the waiting time has increased until \emph{240 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is more than five times larger for both prime-present conditions, compared to the blank prime condition.
 
-Furthermore, the estimated conditional accuracy values in bin (240,280{]} are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
-In other words, if a response is emitted in bin (240,280{]}, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+Furthermore, the estimated conditional accuracy values in bin 280 are 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
+In other words, if a response is emitted in bin 280, then the probability that it is correct is estimated to be 0.29, 1, and 0 for the blank, congruent, and incongruent prime conditions, respectively.
 
 However, when the waiting time has increased until \emph{400 ms} after target onset, then the conditional probability of response occurrence in the next 40 ms is estimated to be 0.36, 0.25, and 0.06 for the blank, congruent, and incongruent prime conditions, respectively.
-And when a response does occur in bin (400,440{]}, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
+And when a response does occur in bin 440, then the probability that it is correct is estimated to be 0.98, 1, and 1 for the blank, congruent, and incongruent prime conditions, respectively.
 
@@
 In tutorial\_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We decided to select the analysis time window (200,600{]} and the cloglog link.
+We decided to select the 200-600 ms analysis time window, and the cloglog link.
 Below, we shortly discuss two of these models.
@@
 \NormalTok{ptb\_data }\OtherTok{\textless{}{-}}\NormalTok{ ptb\_data }\SpecialCharTok{\%\textgreater{}\%} 
-\CommentTok{\# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)}
+\CommentTok{\# select analysis time range: 200{-}600 ms, with 10 bins (time bin ranks 6 to 15)}
 \FunctionTok{filter}\NormalTok{(period }\SpecialCharTok{\textgreater{}} \DecValTok{5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
@@
 If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on the credible Intervals (CrIs) in Figure 3C.
-The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}).
+The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin 240 (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin 280 (95\% CrI = {[}0.04, 0.25{]}).
 For the other bins, the 95\% credible interval contained zero.
-The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin (320,360{]} (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin (360,400{]} (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin (400,440{]} (95\% CrI = {[}-0.40, -0.03{]}).
+The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin 240 (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin 360 (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin 400 (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin 440 (95\% CrI = {[}-0.40, -0.03{]}).
 For the other bins, the 95\% credible interval contained zero.
@@
 There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
-In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240{]}.
-In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440{]}.
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in time bin 240.
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period 320-440 ms.
 The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
@@
 As we are actually interested in the effects of congruent and incongruent primes, relative to the blank prime condition, we can construct two contrasts (congruent-blank, incongruent-blank), and plot the posterior distributions of these contrast effects at the population level (Figure 4C).
-Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins (200,240{]}, (240,280{]}, (280,320{]}, and (320,360{]}, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
+Based on Figure 4C we see that on the population level congruent primes have a positive effect on the conditional accuracy of emitted responses in time bins 240, 280, 320, and 360, relative to the estimates in the baseline condition (blank prime; red dashed lines in Figure 4C).
 Incongruent primes have a negative effect on the conditional accuracy of emitted responses in the first three time bins, relative to blank primes.

### 2025-07-13 sims/sims.Rmd
remove bin brackets and readme.html; update Figs 1,2 and Table 3

@@
 
-This file simulates the data used in old Figures 1, 2, and 3 of the manuscript and then creates some plots.
-
-Figure 1 of revision 1 can be found in section 13.
+This file simulates the data used in Figures 1, 2, and 3 of the first version of the manuscript and then creates some plots.
+
+Figure 1 of revised manuscript can be found in section 13.
 
@@
   theme_bw() +
-    theme(text = element_text(size = 22, face = "bold"), 
-          title = element_text(size = 22, face = "bold"),
+    theme(text = element_text(size = 20, face = "bold"), 
+          title = element_text(size = 20, face = "bold"),
           legend.position = "bottom")
@@
   scale_y_continuous(limits = c(0,1)) +
-  xlab("time bin t's endpoint (ms)") +
+  xlab("time bin (upper bound in ms)") +
   ylab("hazard") +
@@
   scale_y_continuous(limits = c(0,1)) +
-  xlab("time bin t's endpoint (ms)") +
+  xlab("time bin (upper bound in ms)") +
     ylab("conditional accuracy") +

### 2025-07-13 custom_functions.R

@@ -138,12 +138,12 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
     geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), 
                    show.legend = F, linewidth = 0.6) +
     scale_color_brewer(palette = "Dark2", name="Condition") +
-  scale_shape_discrete(name ="Condition", solid=T) + 
+    scale_shape_discrete(name ="Condition", solid=T) + 
     scale_x_continuous(breaks = c(first_bin_shown:(cutoff/binsize)), 
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,haz_yaxis)) +
-    labs(x = "bin endpoints (ms)", 
+    labs(x = "time bin (upper bound in ms)", 
          y = "h(t)") +
     theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
@@ -174,7 +174,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,1)) +
-    labs(x = "bin endpoints (ms)", 
+    labs(x = "time bin (upper bound in ms)", 
          y = "S(t)") +
     theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
@@ -194,7 +194,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                        limits = c(first_bin_shown,cutoff/binsize)) +
 
     scale_y_continuous(limits = c(0,1)) +
-    labs(x = "bin endpoints (ms)", 
+    labs(x = "time bin (upper bound in ms)", 
          y = "ca(t)") +
     theme(panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle=90))
@@ -214,7 +214,7 @@ plot_eha <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F,Nsu
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
[…]

### 2025-07-13 Tutorial_2a.Rmd
small corrections
update Figures 3 and 4

@@
 We compare models using WAIC and LOO (section 4), and interpret parameter estimates for one model (section 5).
-We also plot the logit and cloglog link functions (section 6), and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale (section 7).
+We also plot the logit and cloglog link functions (section 6), and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., discrete-time hazard) scale (section 7).
 
@@
 The skew-normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that higher hazard values (above .5) are less likely than lower values (below .5).
-This prior is plotted in row F of Supplementary Figure 2 (section E in the Supplemental Material).
+This prior is plotted in row F of Supplementary Figure 3 (section F in the Supplemental Material).
 A prior predictive check is performed in section 8.1.
@@
   labs(title = 'Posterior distributions for population-level\neffects in Model M1i',
-       x = "time bin", y = "cloglog-hazard") +
+       x = "time bin (upper bound in ms)", y = "cloglog-hazard") +
   scale_fill_brewer(palette = "Dark2") +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
   theme(axis.text.x = element_text(angle=90)) +
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), 
+    scale_x_discrete(labels = c(6:15)*40, 
                      breaks = 6:15) +
@@
     scale_y_continuous(limits=c(0,.6)) +
-    labs(x = "time bin", y = "predicted hazard")
+    labs(x = "time bin (upper bound in ms)", y = "predicted hazard")
 
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(0,.6)) +
-    labs(x = "time bin", y = "predicted hazard") 
+    labs(x = "time bin (upper bound in ms)", y = "predicted hazard") 
 ```
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(0,1)) +
-    labs(x = "time bin", y = "predicted hazard") +
-   # ggtitle("Each participant (N = 6)") +
+    labs(x = "time bin (upper bound in ms)", y = "predicted hazard") +
     facet_wrap(~pid, ncol=3)
@@
   scale_y_continuous(limits=c(-0.5,0.3)) +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
   scale_fill_manual(values=c("#D95F02FF", "#7570B3FF")) +
   theme(axis.text.x = element_text(angle=90)) +
-  labs(x = "time bin", y = "difference in predicted hazard") +
-  #ggtitle("Grand average marginal effect") +
+  labs(x = "time bin (upper bound in ms)", y = "difference in predicted hazard") +
   facet_wrap(~contrast)
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(-0.6,0.6)) +
-    labs(x = "time bin", y = "difference in predicted hazard") +
-    #ggtitle("Subject-specific AMEs") +
+    labs(x = "time bin (upper bound in ms)", y = "difference in predicted hazard") +
     facet_wrap(~pid, ncol=3)
@@
     plot_annotation(tag_levels = 'A') +
-   # plot_layout(guides = "collect",
-  #              axes = "collect_x") & 
     theme(legend.position = "bottom")
@@
     plot_annotation(tag_levels = 'A') +
-   # plot_layout(guides = "collect",
-  #              axes = "collect_x") & 
     theme(legend.position = "bottom")
-
 ```
@@
 
-If we want to study the population-level effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A.
+If we want to study the population-level effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on the population-level contrasts.
 The contrast "congruent minus blank" was estimated to be 0.09 hazard units in bin 6 (95% CrI = [0.02, 0.17]), and 0.14 hazard units in bin 7 (95% CrI = [0.04, 0.25]).
@@
 There are thus two phases of performance for the average person between 200 and 600 ms after target onset.
-In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period (200, 240].
-In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period (320,440].
+In the first phase, the addition of a congruent or incongruent prime stimulus increases the hazard of response occurrence compared to blank prime trials in the time period 200-240 ms.
+In the second phase, only the incongruent prime decreases the hazard of response occurrence compared to blank primes, in the time period 320-440 ms.
 The sign of the effect of incongruent primes on the hazard of response occurrrence thus depends on how much waiting time has passed since target onset.
@@
 
-If we want to focus more on inter-individual differences, we can study the subject-specific differences in hazard in Figure 7B.
-Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin (360,400] -- subject 2 also in bin (320,360].
+If we want to focus more on inter-individual differences, we can study the subject-specific contrasts.
+Note that three participants (1, 2, and 3) show a negative difference for the contrast "congruent minus incongruent" in bin 400 -- subject 2 also in bin 360.
 
@@
 
-Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time (Figure 9).
+Interestingly, all subjects show a tendency in their mean difference (congruent minus blank) to "dip" around that time.
 Therefore, future modeling efforts could incorporate the trial number into the model formula, in order to also study how the effects of prime type on hazard change on the long experiment-wide time scale, next to the short trial-wide time scale.

### 2025-07-13 Tutorial_2b.Rmd
update Figures 3 and 4

@@
   labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
-       x = "time bin", 
+       x = "time bin (upper bound in ms)", 
        y = "logit-ca") +
   scale_fill_brewer(palette = "Dark2") +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
   scale_y_continuous(limits=c(-10,10)) +
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(0,1)) +
-    labs(x = "time bin", y = "predicted conditional accuracy") #+
-    #ggtitle("Average participant")
+    labs(x = "time bin (upper bound in ms)", 
+         y = "predicted conditional accuracy")
+
 p1
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(0,1)) +
-    labs(x = "time bin", y = "predicted conditional accuracy") +
-   # ggtitle("Each participant (N = 6)") +
+    labs(x = "time bin (upper bound in ms)", 
+         y = "predicted conditional accuracy") +
     facet_wrap(~pid, ncol=3)
@@
   scale_y_continuous(limits=c(-1,1)) +
-  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+  scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
   scale_fill_manual(values=c("#D95F02FF", "#7570B3FF")) +
   theme(axis.text.x = element_text(angle=90)) +
-  labs(x = "time bin", y = "difference in predicted ca") +
+  labs(x = "time bin (upper bounds in ms)", 
+       y = "difference in predicted ca") +
   facet_wrap(~contrast)
+
 p11
@@
                values_to = "diff_ca")
-epreds_pid_diffs # 160 000 rows 
+
+epreds_pid_diffs  
 ```
@@
                     .width=c(.8,.95)) +
-    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
     geom_hline(yintercept = 0, color = "red", lty = 3) +
@@
     scale_color_brewer(palette = "Dark2") +
-    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
+    scale_x_discrete(labels = c(6:15)*40, breaks = 6:15) +
     theme(axis.text.x = element_text(angle=90)) +
     scale_y_continuous(limits=c(-1,1)) +
-    labs(x = "time bin", y = "difference in predicted ca") +
-   # ggtitle("Subject-specific AMEs") +
+    labs(x = "time bin (upper bound in ms)", 
+         y = "difference in predicted ca") +
     facet_wrap(~pid, ncol=3)

### 2025-07-13 Tutorial_1b.Rmd
update Figures Tutorial 1b

@@
 
-Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of @panisWhatShapingRT2016 when there are TWO independent variables (here: mask type and prime type).
+Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of @panisWhatShapingRT2016 when there are two independent variables (here: mask type and prime type).
 

### 2025-07-13 custom_functions.R

@@ -350,7 +350,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
                       labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                       limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,haz_yaxis)) +
-    labs(x = "bin endpoints (ms)",
+    labs(x = "time bin (upper bound in ms)",
         y = "h(t)",
         color = "Prime") +
     theme(panel.grid.minor = element_blank(),
@@ -377,7 +377,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,1)) +
-    labs(x = "bin endpoints (ms)",
+    labs(x = "time bin (upper bound in ms)",
          y = "S(t)",
          color = "Prime") +
     theme(panel.grid.minor = element_blank(),
@@ -399,7 +399,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,1)) +
-    labs(x = "bin endpoints (ms)",
+    labs(x = "time bin (upper bound in ms)",
          y = "ca(t)",
          color = "Prime") +
     theme(panel.grid.minor = element_blank(),
@@ -421,7 +421,7 @@ plot_eha_2IV <- function(df,subj,haz_yaxis=1,first_bin_shown=1,aggregated_data=F
                        labels = c(first_bin_shown:(cutoff/binsize)*binsize),
                        limits = c(first_bin_shown,cutoff/binsize)) +
     scale_y_continuous(limits = c(0,.5)) +
-    labs(x = "bin endpoints (ms)",
+    labs(x = "time bin (upper bound in ms)",
          y = "P(t)",
          color = "Prime") +
     theme(panel.grid.minor = element_blank(),

### 2025-07-13 manuscript/Suppl_material/supplementary.Rmd
update suppl. material

@@
 
-While people can die only once, in experimental RT tasks the events of interest are typically repeatable.
+While people can die only once, in experimental response time (RT) tasks the events of interest are typically repeatable.
 In Supplementary Figure 1A, we present a hypothetical single-subject RT data set from a detection experiment with one response button and two experimental conditions: target present and target absent.
@@
 
-(ref:descr-plot-RT-tasks-caption) RT tasks and EHA concepts.
+(ref:descr-plot-RT-tasks-caption) Response time (RT) tasks and EHA concepts.
 The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line).
-In each trial of each condition of each task, an overt response time (RT) is indicated by a circle.
+In each trial of each condition of each task, an overt response is indicated by a circle, and right-censored observations by a cross.
 (A) Single-button detection data requires a single-event analysis.
@@
 Thus, another approach is to model the hazard of (any) response occurrence and extend it with a speed-accuracy trade-off (SAT) analysis (plotting accuracy as a function of latency).
-This is the approach taken here (see Figure 1), and many experimental psychologists are familiar with SAT.
+This is the approach taken here (see Figure 1B), and many experimental psychologists are familiar with SAT.
 
@@
 
-While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y]: x < RT $\leq$ y.
+While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time.
 If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial.
@@
 Ignoring trials without a response leads to underestimation of the true mean.
-By introducing a fixed censoring time for all trials, trials with long RTs are not discarded but contribute to the risk set of each bin.
+By introducing a fixed censoring time for all trials, trials with long RTs are not discarded but contribute to the risk set of each bin (see Table 3).
 
@@
 Finally, as explained by @kelsoOutlineGeneralTheory2013, it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived "dips" that could signal the presence of active top-down respone inhibition processes [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016].
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical hazard (and possibly conditional accuracy) distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived "dips" that could signal the presence of active top-down respone inhibition processes [@panisStudyingDynamicsVisual2020; @panisWhatShapingRT2016; @panisWhenDoesInhibition2022].
 As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.

### 2025-07-13 manuscript/Suppl_material/supplementary.tex
update ms, notes, and suppl. material
update suppl. material

@@
 \raggedbottom
+\usepackage{setspace}
+\captionsetup[figure]{font={stretch=1,normalsize}}
 \usepackage{bookmark}
@@
 
-\caption{RT tasks and EHA concepts.
+\caption{Response time (RT) tasks and EHA concepts.
 The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line).
-In each trial of each condition of each task, an overt response time (RT) is indicated by a circle.
+In each trial of each condition of each task, an overt response is indicated by a circle, and right-censored observations by a cross.
 (A) Single-button detection data requires a single-event analysis.
@@
 
-While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y.
+While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time.
 If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial.
@@
 Finally, as explained by Kelso, Dumas, and Tognoli (2013), it is crucial to first have a precise description of the macroscopic behavior of a system (here: h(t) and possibly ca(t) functions) in order to know what to derive on the microscopic level.
-EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical RT hazard distributions.
-Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived ``dips'' that could signal the presence of active top-down respone inhibition processes (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016).
+EHA can thus solve the problem of model mimicry, i.e., the fact that different computational models can often predict the same mean RTs as observed in the empirical data, but not necessarily the detailed shapes of the empirical hazard (and possibly conditional accuracy) distributions.
+Also, fitting parametric functions or computational models to data without studying the shape of the empirical discrete-time h(t) and ca(t) functions can miss important features in the data, such as short-lived ``dips'' that could signal the presence of active top-down respone inhibition processes (Panis, Moran, Wolkersdorfer, \& Schmidt, 2020; Panis \& Schmidt, 2016, 2022).
 As such, EHA can be a tool to help distinguish between competing theories of cognition and brain function.
@@
 
+\bibitem[\citeproctext]{ref-panisWhenDoesInhibition2022}
+Panis, S., \& Schmidt, T.
+(2022).
+When does {``inhibition of return''} occur in spatial cueing tasks?
+{Temporally} disentangling multiple cue-triggered effects using response history and conditional accuracy analyses.
+\emph{Open Psychology}, \emph{4}(1), 84--114.
+\url{https://doi.org/10.1515/psych-2022-0005}
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}

### 2025-07-13 manuscript/manuscript.Rmd
update word count

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10115 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl. mat.)"
+wordcount         : "10154 (body) + 1742 (references) + 3453 (body supplemental material) + 428 (refs suppl. mat.)"
 bibliography      : ["refs_manuscript.bib", "r-references.bib"]

### 2025-07-13 manuscript/manuscript.tex
update word count

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10115 (body) + 1742 (references) + 3442 (body supplemental material) + 393 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10154 (body) + 1742 (references) + 3453 (body supplemental material) + 428 (refs suppl.
 mat.)}

### 2025-07-14 manuscript/reviews/ResponseReview1.Rmd
add bibtex entries to supps. Update Response Letter.

@@
 knitr::opts_chunk$set(echo = F)
-```
-
-```{r packages and data, include=F }
-
 ```
@@
 
-Response: Thank you for providing constructive feedback and allowing us to submit a revision of our manuscript (AMPPS-24-0232).
+Response: Dear Editor,
+
+thank you for providing constructive feedback and allowing us to submit a revision of our manuscript (AMPPS-24-0232).
 
 We tried to focus on accessibility in the revision of our Tutorial in several ways. 
-First, we replaced Figure 1 with a simpler Figure that directly conveys our message, and deleted Figures 2 and 3.
+First, we reduced the number of Figures: we replaced Figure 1 with a simpler Figure that directly conveys our message and deleted Figures 2 and 3, and we replaced the 6 old figures on modeling results with 2 new figures.
 Second, we shortened the Introduction and Discussion sections.
@@
 Response: Thank you for the positive feedback. 
-To make the manuscript more accessible and shorter and to respond to some other comments from the editor and other reviewers, we moved the explanation of the survival function to the Supplementary Material in the revised ms.
+To make the manuscript more accessible and shorter and to respond to some other comments from the editor and other reviewers, we moved the explanation of the survival function to the Supplementary Material in the revised ms (section B).
 
@@
 
-Response: We introduce the concept of right-censoring on page 11 of the revised ms.
+Response: We introduce the concept of right-censoring on page 11 of the revised ms, and the conventional way of dealing with missing data (or very long RTs) on page 10.
 
@@
 
-Response: We explain this on pages 22-23 of the revised ms.:
+Response: We explain this on pages 22 of the revised ms.:
 "We also use a "keep it maximal" approach by specifying a full varying (or random) effects structure. This means that wherever possible we include varying intercepts and slopes per participant."
@@
 
-Response: We discuss the main contributions above and beyond existing works on page 6 of the revised ms.
+Response: Thank you for your constructive feedback.
+We discuss the main contributions above and beyond existing works on page 6 of the revised ms.
 We address the issue of treating time as discrete or continuous on pages 8-9 of the revised ms.

### 2025-07-14 manuscript/reviews/ResponseReview1.tex
add bibtex entries to supps. Update Response Letter.

@@
 \author{Panis \& Ramsey}
-\date{2025-07-11}
+\date{2025-07-14}
 
@@
 
-Response: Thank you for providing constructive feedback and allowing us
-to submit a revision of our manuscript (AMPPS-24-0232).
+Response: Dear Editor,
+
+thank you for providing constructive feedback and allowing us to submit
+a revision of our manuscript (AMPPS-24-0232).
 
 We tried to focus on accessibility in the revision of our Tutorial in
-several ways. First, we replaced Figure 1 with a simpler Figure that
-directly conveys our message, and deleted Figures 2 and 3. Second, we
-shortened the Introduction and Discussion sections. Third, we added a
-section in the supplementary material (section A) in which we visualize
-the different types of time-to-event data that are obtained in typical
-RT tasks (detection, discrimination, bistable perception). We also
-mention the unique contribution of our work, we removed redundancies,
-and tried to improve the clarity of the paper. As such, we have
-substantially revised the manuscript in accordance with feedback from
-you as well as the reviewers.
+several ways. First, we reduced the number of Figures: we replaced
+Figure 1 with a simpler Figure that directly conveys our message and
+deleted Figures 2 and 3, and we replaced the 6 old figures on modeling
+results with 2 new figures. Second, we shortened the Introduction and
+Discussion sections. Third, we added a section in the supplementary
+material (section A) in which we visualize the different types of
+time-to-event data that are obtained in typical RT tasks (detection,
+discrimination, bistable perception). We also mention the unique
+contribution of our work, we removed redundancies, and tried to improve
+the clarity of the paper. As such, we have substantially revised the
+manuscript in accordance with feedback from you as well as the
+reviewers.
 
@@
 
-Response: We discuss the main contributions above and beyond existing
-works on page 6 of the revised ms. We address the issue of treating time
-as discrete or continuous on pages 8-9 of the revised ms. We removed
-redundancies, get to the tutorial section quicker (page 13) and included
-more precise explanations of technical jargon.
+Response: Thank you for your constructive feedback. We discuss the main
+contributions above and beyond existing works on page 6 of the revised
+ms. We address the issue of treating time as discrete or continuous on
+pages 8-9 of the revised ms. We removed redundancies, get to the
+tutorial section quicker (page 13) and included more precise
+explanations of technical jargon.
 

### 2025-07-14 manuscript/Suppl_material/supplementary.Rmd
add bibtex entries to supps. Update Response Letter.

@@
 The most common type of right-censoring is “singly Type I censoring” which applies when the experiment uses a fixed response deadline for all trials.
-“Type I” means that the censoring time is fixed and under the control of the experimenter; “Singly” refers to the fact that all observations have the same censoring time (Allison, 2010).
+“Type I” means that the censoring time is fixed and under the control of the experimenter; “Singly” refers to the fact that all observations have the same censoring time [@allisonSurvivalAnalysisUsing2010].
 Clearly, discarding right-censored observations before calculating the median or mean RT will result in biased estimates, especially in the target-absent condition.
@@
 
-There are different types of censoring.
+There are different types of censoring [@allisonSurvivalAnalysisUsing2010].
 Left censoring occurs when all that is known about an observation on a variable T is that it is less than some value.
-Interval censoring combines right and left censoring so that all you know about T is that a < T < b, for some values of a and b (Allison, 2010).
+Interval censoring combines right and left censoring so that all you know about T is that a < T < b, for some values of a and b.
 Random censoring occurs when observations are terminated for reasons that are not under the control of the experimenter.
@@
 When estimating the hazard of correct response occurrence, then the unpredictable error responses will introduce random censoring (and vice versa).
-Importantly, all statistical methods for time-to-event data require that random censoring be noninformative: a trial that is censored at time c should be representative of all those trials with the same values of the explanatory variables that survive to c (Allison, 2010).
-However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another (Burle et al., 2004; Eriksen et al., 1985; Praamstra & Seiss, 2005).
+Importantly, all statistical methods for time-to-event data require that random censoring be noninformative: a trial that is censored at time c should be representative of all those trials with the same values of the explanatory variables that survive to c [@allisonSurvivalAnalysisUsing2010].
+However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another [@BURLE2004153; @praamstra9513; @eriksenElectromyographicExaminationResponse1985].
 Thus, another approach is to model the hazard of (any) response occurrence and extend it with a speed-accuracy trade-off (SAT) analysis (plotting accuracy as a function of latency).
@@
 For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue).
-However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis & Schmidt, 2016).
+However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus [@panisHowCanWe2020; @panisWhatShapingRT2016].
 For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels).
@@
 
-The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
+The discrete-time hazard function of event occurrence gives you for each bin the conditional probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
 This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 
-While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event does not occur before the endpoint of bin t.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event does not occur before the upper bound of bin t.
 
@@
 
-\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
+\noindent which gives you for each bin the conditional probability that a response is correct given that it is emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010].
 The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 For example, if h~A~(t) > h~B~(t) for all t, then both hazard functions are said to show a complete ordering.
-Townsend (1990) concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
+@townsendTruthConsequencesOrdinal1990 concluded that stronger conclusions can be drawn from data when comparing the hazard functions using EHA.
 For example, when mean A < mean B, the hazard functions might show a complete ordering (i.e., for all t), a partial ordering (e.g., only for t > 300 ms, or only for t < 500 ms), or they may cross each other one or more times.

### 2025-07-14 manuscript/Suppl_material/supplementary.tex
add bibtex entries to supps. Update Response Letter.

@@
 
-There are different types of censoring.
+There are different types of censoring (Allison, 2010).
 Left censoring occurs when all that is known about an observation on a variable T is that it is less than some value.
-Interval censoring combines right and left censoring so that all you know about T is that a \textless{} T \textless{} b, for some values of a and b (Allison, 2010).
+Interval censoring combines right and left censoring so that all you know about T is that a \textless{} T \textless{} b, for some values of a and b.
 Random censoring occurs when observations are terminated for reasons that are not under the control of the experimenter.
@@
 Importantly, all statistical methods for time-to-event data require that random censoring be noninformative: a trial that is censored at time c should be representative of all those trials with the same values of the explanatory variables that survive to c (Allison, 2010).
-However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another (Burle et al., 2004; Eriksen et al., 1985; Praamstra \& Seiss, 2005).
+However, the random censoring by errors when estimating the hazard function of correct responses (and vice versa) is very likely informative, because response channels are known to compete with one another (Burle, Vidal, Tandonnet, \& Hasbroucq, 2004; Eriksen, Coles, Morris, \& O'hara, 1985; Praamstra \& Seiss, 2005).
 Thus, another approach is to model the hazard of (any) response occurrence and extend it with a speed-accuracy trade-off (SAT) analysis (plotting accuracy as a function of latency).
@@
 
-The discrete-time hazard function of event occurrence gives you for each bin the probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
+The discrete-time hazard function of event occurrence gives you for each bin the conditional probability that the event occurs (sometime) in that bin, given that the event has not occurred yet in previous bins.
 This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before.
 
-While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event does not occur before the endpoint of bin t.
+While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event does not occur before the upper bound of bin t.
 
@@
 
-\noindent which gives you for each bin the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
+\noindent which gives you for each bin the conditional probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).
 The ca(t) function is also known as the micro-level speed-accuracy tradeoff (SAT) function.
@@
 
+\bibitem[\citeproctext]{ref-BURLE2004153}
+Burle, B., Vidal, F., Tandonnet, C., \& Hasbroucq, T.
+(2004).
+Physiological evidence for response inhibition in choice reaction time tasks.
+\emph{Brain and Cognition}, \emph{56}(2), 153--164.
+\url{https://doi.org/10.1016/j.bandc.2004.06.004}
+
+\bibitem[\citeproctext]{ref-eriksenElectromyographicExaminationResponse1985}
+Eriksen, C.
+W., Coles, M.
+G.
+H., Morris, L.
+R., \& O'hara, W.
+P.
+(1985).
+An electromyographic examination of response competition.
+\emph{Bulletin of the Psychonomic Society}, \emph{23}(3), 165--168.
+\url{https://doi.org/10.3758/BF03329816}
+
 \bibitem[\citeproctext]{ref-holdenDispersionResponseTimes2009}
@@
 
+\bibitem[\citeproctext]{ref-panisHowCanWe2020}
+Panis, S.
+(2020).
+How can we learn what attention is?
+{Response} gating via multiple direct routes kept in check by inhibitory control processes.
+\emph{Open Psychology}, \emph{2}(1), 238--279.
+\url{https://doi.org/10.1515/psych-2020-0107}
+
 \bibitem[\citeproctext]{ref-panisStudyingDynamicsVisual2020}
@@
 
+\bibitem[\citeproctext]{ref-praamstra9513}
+Praamstra, P., \& Seiss, E.
+(2005).
+The neurophysiology of response competition: {Motor} cortex activation and inhibition following subliminal response priming.
+\emph{Journal of Cognitive Neuroscience}, \emph{17}(3), 483--493.
+\url{https://doi.org/10.1162/0898929053279513}
+
 \bibitem[\citeproctext]{ref-singerAppliedLongitudinalData2003}

### 2025-07-14 manuscript/manuscript.Rmd
small updates to ms. and Suppl. Material

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10154 (body) + 1742 (references) + 3453 (body supplemental material) + 428 (refs suppl. mat.)"
+wordcount         : "10165 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl. mat.)"
 bibliography      : ["refs_manuscript.bib", "r-references.bib"]
@@
 For example, time bin "500" in Figure 1B refers to the time interval running from 400 ms to 500 ms, with the lower bound of 400 ms excluded, and the upper bound of 500 ms included.
-Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature [@panisWhenDoesInhibition2022].
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature [@panisWhenDoesInhibition2022; @panisWhatShapingRT2016; @panisHowCanWe2020].
 
@@
 Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than ~200 trials per condition per participant [@panisAnalyzingResponseTimes2020].
-Moreover, as indicated by @allisonSurvivalAnalysisUsing2010, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
+Moreover, as indicated by @singerAppliedLongitudinalData2003, learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
 
@@
 Are there new auxiliary assumptions that computational models need to adopt [@panisStudyingDynamicsVisual2020]? 
-Will the temporal effect patterns align nicely with EEG findings [@panisWhenDoesInhibition2022]?
+Will the temporal effect patterns align nicely with electroencephalography (EEG) findings [@panisWhenDoesInhibition2022]?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
@@
 First, we want to calculate descriptive statistics for each condition in each individual using a life table.
-A life table (see Table 3) includes for each time bin, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
+A life table (see Table 3) includes for each time bin indexed by t, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t), survival probability S(t), probability mass P(t), possibly the conditional accuracy ca(t), and their estimated standard errors (se).
 The definitions of these quantities are provided in section B of the Supplemental Material.
@@
 
-Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
+Next, we can set up the life tables and plot for each participant and condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
 To do so using a functional programming approach, one has to nest the person-trial data within participants using the group_nest() function, and supply a user-defined censoring time and bin width to our custom function "censor()", as follows.

### 2025-07-14 manuscript/manuscript.tex
small updates to ms. and Suppl. Material

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10154 (body) + 1742 (references) + 3453 (body supplemental material) + 428 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10165 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl.
 mat.)}
@@
 For example, time bin ``500'' in Figure 1B refers to the time interval running from 400 ms to 500 ms, with the lower bound of 400 ms excluded, and the upper bound of 500 ms included.
-Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature (Panis \& Schmidt, 2022).
+Importantly from a face-validity perspective, this pattern of simulated data can be seen in the experimental psychology literature (Panis, 2020; Panis \& Schmidt, 2016, 2022).
 
@@
 Thus, by trading a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal for dealing with typical psychological RT data sets for which there are less than \textasciitilde200 trials per condition per participant (Panis, Schmidt, et al., 2020).
-Moreover, as indicated by Allison (2010), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
+Moreover, as indicated by Singer and Willett (2003), learning discrete-time EHA methods first will help in learning continuous-time methods, so it seems like a good starting point.
 
@@
 Are there new auxiliary assumptions that computational models need to adopt (Panis, Moran, et al., 2020)?
-Will the temporal effect patterns align nicely with EEG findings (Panis \& Schmidt, 2022)?
+Will the temporal effect patterns align nicely with electroencephalography (EEG) findings (Panis \& Schmidt, 2022)?
 And are there new experiments that need to be performed to test the novel predictions that follow from these analyses?
@@
 
-Next, we can set up the life tables and plot for each condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
+Next, we can set up the life tables and plot for each participant and condition the discrete-time hazard function h(t), survivor function S(t), probability mass function P(t), and conditional accuracy function ca(t).
 To do so using a functional programming approach, one has to nest the person-trial data within participants using the group\_nest() function, and supply a user-defined censoring time and bin width to our custom function ``censor()'', as follows.

### 2025-07-14 Tutorial_1a.Rmd
update Tutorials and ms.

@@
 
-Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 119-122).
+Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 124-126).
 

### 2025-07-14 Tutorial_2a.Rmd
update Tutorials and ms.

@@
 
-M0d took about 77 minutes.
+M0d took about 77 minutes using the same MacBook Pro.
 
@@
 
-Model_M0c took about 19 minutes to run.
+Model_M0c took about 19 minutes to run using the same MacBook Pro.
 
@@
 
-Model M1i took about 124 minutes to fit.
+Model M1i took about 124 minutes to fit using the same MacBook Pro.
 
@@
 
-Second, when the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
+Second, when the shape of the hazard function is rather smooth, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME.
 Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").
@@
 Note that duplicate terms in the model formula (e.g., condition) are ignored.
-Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model_M1d took about 184 minutes to run.
+Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model_M1d took about 184 minutes to run using the same MacBook Pro.
 
@@
     labs(y = "difference in predicted hazard", x = "timebin") +
-    ggtitle("Grand average marginal effect") +
     facet_wrap(~contrast)
@@
 
-The shape of the prior predictive distribution for 1 bin looks as expected (compare with Supplementary Figure 2 (row F) in the Supplemental Material): hazard values below .5 are more likely than hazard values above .5.
+The shape of the prior predictive distribution for 1 bin looks as expected (compare with Supplementary Figure 3 (row F) in section F of the Supplemental Material): hazard values below .5 are more likely than hazard values above .5.
 

### 2025-07-14 Tutorial_3a.Rmd
update Tutorials and ms.

@@
 
-Model_M1i_f took about 3 hours.
+Model_M1i_f took about 3 hours on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
@@
                 position = position_nudge(x = .4)) +
-# geom_hline(yintercept=0,colour = "grey", linetype = 2) +
  scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
  scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +
-
- labs(x = "bin endpoint", y = "cloglog-hazard") +
+ labs(x = "time bin (upper bound in ms)", y = "cloglog-hazard") +
  scale_color_manual(name = "Model", 

### 2025-07-14 Tutorial_3b.Rmd
update Tutorials and ms.

@@
 
-Model_M1i_ca_f took about 90 minutes to fit.
+Model_M1i_ca_f took about 90 minutes to fit on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).
 
@@
                 position = position_nudge(x = .4)) +
- #geom_hline(yintercept=0,colour = "grey", linetype = 2) +
  scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
  scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +
-
- labs(x = "bin endpoint", y = "logit-ca") +
+ labs(x = "time bin (upper bound in ms)", y = "logit-ca") +
  scale_color_manual(name = "Model", 

### 2025-07-14 manuscript/manuscript.Rmd
update word count

@@
 keywords          : "response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology"
-wordcount         : "10165 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl. mat.)"
+wordcount         : "10167 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl. mat.)"
 bibliography      : ["refs_manuscript.bib", "r-references.bib"]
@@
 In tutorial_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We decided to select the 200-600 ms analysis time window, and the cloglog link.
+We decided to select the 200-600 ms time window for inferential analysis, and the cloglog link.
 Below, we shortly discuss two of these models.

### 2025-07-14 manuscript/manuscript.tex
update word count

@@
 \makeatother
-\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10165 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl.
+\keywords{response times, event history analysis, Bayesian multilevel regression models, experimental psychology, cognitive psychology\newline\indent Word count: 10167 (body) + 1742 (references) + 3458 (body supplemental material) + 539 (refs suppl.
 mat.)}
@@
 In tutorial\_2a.Rmd we fit several Bayesian multilevel models (i.e., generalized linear mixed models) that differ in complexity to the person-trial-bin oriented data set that we created in Tutorial 1a.
-We decided to select the 200-600 ms analysis time window, and the cloglog link.
+We decided to select the 200-600 ms time window for inferential analysis, and the cloglog link.
 Below, we shortly discuss two of these models.

### 2025-07-14 manuscript/manuscript.Rmd
update filename disclosures
update doi

@@
 
-All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: https://doi.org/10.31234/osf.io/57bh6
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: https://doi.org/10.31234/osf.io/57bh6_v2
 
@@
 
-Supplemental Material: Panis_Ramsey_suppl_material.pdf
+Supplemental Material: Panis_Ramsey_suppl_material_R1.pdf
 

### 2025-07-14 manuscript/manuscript.tex
update filename disclosures
update doi

@@
 
-All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: \url{https://doi.org/10.31234/osf.io/57bh6}
+All of the submitted manuscript and Supplemental Material was previously posted to a preprint archive: \url{https://doi.org/10.31234/osf.io/57bh6_v2}
 
@@
 
-Supplemental Material: Panis\_Ramsey\_suppl\_material.pdf
+Supplemental Material: Panis\_Ramsey\_suppl\_material\_R1.pdf
 
